### 2016-05-16 admin/CURIS2016/curis.tex
init!

\documentclass[12pt]{article}
\usepackage{balance,times,graphics,setspace}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}

\setlength{\parskip}{.4em}
\setlength{\parindent}{0cm}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}

\begin{document}
\section*{Project Name}

Alia: Building a worker--owned gig market

\section*{Description}

Working with the National Domestic Workers Alliance (NDWA),
we're building a cooperative labor market for house cleaners;
This labor market will allow workers to
manage and run both
their individual businesses \&
the platform on which they work.
We expect to expand significantly over the summer,
with communities coming ``online'' in
the SF Bay Area,
New York City,
and Seattle
(we're open to substantial growth,
but the purpose of this project isn't to scale).

This research is about exploring how a digitally--mediated cooperative labor market
would operate; in other words, what would a worker--owned version of
Handy,
or Uber,
or Amazon's Mechanical Turk,
look like?
The broader question at stake is how to enable and facilitate
collective governance,
something we see to varying extents in online communities like
Reddit, Wikipedia, etc\dots,
but which we don't seem to understand well enough to make routine,
or to replicate consistently.

More tangibly,
over the summer we're going to be working with
groups of workers that are on (more or less) the same platform.
We'll be building features to facilitate work
(one example:
communicating availability and negotiating booking a time)
as well as tools to facilitate decision--making
(one possible example:
voting mechanisms to hold a referendum).

\section*{Recommended Background}

Ideally, you have experience in
Human--Computer Interaction
(think CS 147, 247).
If your background is especially unorthodox,
you might have some experience in ethnographic methods
(think Anthropology, Sociology),
and that would be interesting.
Much of the day--to--day work will involve programming for the web
(think CS 142),
so regardless of your research goals you should be able to code
(and be reasonably competent with existing libraries, frameworks, and APIs).

It might be worth pointing out that
I don't expect you to be an expert in these things right now.
You should be competent enough to start working on day 1, but
this is a chance to hone your skills and maybe learn something completely new.

If your background doesn't fit with anything above,
please tell us the most ambitious thing you've undertaken,
academic or not!
(Hard work can overcome lack of preparation!)

\section*{Number of students allowed}

1--2

\section*{Prerequisite/Preparation}

Your preparation would be to try and backfill some of
the stuff in the ``Recommended Background'' section
that you don't already have.
That'll be unique to your circumstances and background,
so make this judgment call on your own;
if we end up working together,
I can put together a more actionable list of things to do between now and this summer.

\bibliographystyle{acm}
\bibliography{../references.bib}
\end{document}

### 2016-05-16 admin/coopTimeline/assignment1.tex
init!

\documentclass[11pt]{article}
\usepackage{balance,graphics,times,setspace}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue,anchorcolor=blue]{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}

% \hypersetup{ colorlinks=true,linkcolor=blue}
% \setlength{\parskip}{.4em}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
\def\labelitemi{--}

% \newlist{inlinelist}{enumerate*}{1}
% \setlist*[inlinelist,1]{%
%   label=\arabic*),
% }

\begin{document}
  \begin{center}
  \Large{\textbf{Timeline}}
  \end{center}
  % \textit{A person's community of origin is an irrelevant factor in emerging modes of digitally enabled labor.}

% \begin{enumerate}[label=Week \arabic*:]
\section*{November: The Forum}
  \begin{itemize}
    \item Recruiting interested workers (10--20 people)

    \textit{We don't need to be ``Internet--scale'' right off the bat; the big challenge will be figuring out how to make interactions work in a digital medium; we'll tackle the ``interactions at scale'' issue later.}
    \item Hashing out norms ---
    bylaws, policies, etc\dots
    \item Figuring out system specifications ---
    Defining how an automatic system should behave \& what interactions it should make available
    (e.g. how much information about workers customers should be able to see)

    \textit{Much of this has been worked out with Casa Latina, but I think it's important that every work group arrives at a conclusion about this sort of stuff actively, rather than have it put upon them.}
    \item Start narrowing down a developer.

    \textit{Finding a good developer will be absolutely crucial.
    I (Ali) still want to be very involved so if something happens to the site/platform I have a vague sense of how to fix it,
    but this person should be primarily committed to the project.
    We can work out how that will generally work in detail after we've narrowed our candidates list.}
  \end{itemize}
\section*{December: Prototype}
  \begin{itemize}
    \item Immediately: Ali making it work via ``\href{https://en.wikipedia.org/wiki/Wizard_of_Oz_experiment}{Wizard of Oz}''

    \textit{This is part of why we don't want lots of workers --- so that I can handle the amount of work figuring out work assignments and whatnot. It'll also inform whichever person is building the system.}
  \end{itemize}
\section*{December--January: Software}
  \begin{itemize}
    \item Hire a developer, make code open--source for collaborators.

    \textit{As before, this developer will be implementing major components of the platform, although the code itself should remain open--source, especially if we hope to encourage contributors to participate.}
    \item Essentially, start working on software to replace all of the things Ali's doing in the prototype run.
  \end{itemize}
\section*{January: Ramp--up}
  \begin{itemize}
    \item At this point we should be transitioning to the automated/software system,
    so Ali won't be doing dispatch anymore.

    \textit{
      This is where we'll identify issues that maybe Ali was intuitively working around without documenting,
      or handling any bugs or issues that might cause problems for workers absent a human handling dispatch.}
  \end{itemize}
\section*{February--April: 3--month run}
  \begin{itemize}
    \item Handling everything via the software--based system (i.e. Ali won't be running dispatch, making reviews work, etc\dots)

    \textit{The goal of this 3--month phase is to establish a running track of the cooperative working (both in the sense of getting work done and functioning properly).}
  \end{itemize}
\section*{May: Write}
  \begin{itemize}
    \item I'll want to write about the findings of this
    (hopefully) successfully--running cooperative work group.
    With the deadline for \href{http://cscw.acm.org/2016/index.php}{Computer--Supported Cooperative--Work}
    (A Human--Computer Interaction conference) in late May,
    I'll want to have ample time to write up my notes.
  \end{itemize}
\section*{Later: Scale}

% \end{enumerate}
% \bibliographystyle{acm}
% \bibliography{../references.bib}
\end{document}

### 2016-05-16 admin/coopTimeline/timeline.tex
init!

--- 2016-05-16 admin/coopTimeline/assignment1.tex
@@
     Defining how an automatic system should behave \& what interactions it should make available
-    (e.g. how much information about workers customers should be able to see)
+    (e.g. how much information customers should be able to see)
 
@@
 \section*{Later: Scale}
+  \begin{itemize}
+    \item First to around 50 workers, then to around 100, then to around 150 or 200.
+    \item \textit{Each milestone represents a different \textbf{kind} of group --- a group of 10 operates differently from a group of 50; a 50--person group operates differently from a 100--person group; etc\dots}
+  \end{itemize}
 

### 2016-05-16 chi/papers/chi2016labor/self-review.tex
init!

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}

\setlength{\parindent}{0em}
\setlength{\parskip}{0em}
\renewcommand{\baselinestretch}{0}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}
\begin{document}
\begin{itemize}
\item
Everything is so great. 5/5
\item
Nominate for best paper. much wow. such research.
\end{itemize}

\section*{Introduction}
\begin{itemize}
\item
``Gig work'' is not a new thing; piecework has been around since the 19th century, and a lot of that research and work can inform this.
\item why has the notion of ``on the side'' work changed? This seems too important to dismiss out of hand.
\item The questions the paper claims to answer are the following:
\begin{enumerate}
\item
What components are inextricable from successful markets?
\item
what can be disentangled?
\item
how would worker--run groups operate themselves?
\end{enumerate}
Let's see if this gets answered...
\end{itemize}

\section*{Background}
\begin{itemize}
\item \textbf{Collective Action} sounds unrelated to the topic introduced earlier; worker--centric labor markets are different from worker--run labor markets, and it sounds like this is conflating the two; did I miss that connection? The only thing I see that really makes a strong connection in the introduction is the vague reference to ``giving workers a sense of locus'', but that's not entirely the same as worker--run markets\ldots
\end{itemize}

\section*{Background}
Seems fine, not sure how these really

\end{document}

### 2016-05-16 classes/comm220/assignment1/assignment1.tex
init!

\documentclass[11pt]{article}
\usepackage{balance,graphics,times,setspace}
\usepackage{hyperref}
% \usepackage{fontspec}
% \usepackage{xltxtra}
% \usepackage{xunicode}
\usepackage[margin=1in]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}

% \doublespacing
\usepackage[tiny,compact]{titlesec}

% \usepackage{xesearch}
% \newcounter{words}
% \newenvironment{counted}{%
%   \setcounter{words}{0}
%   \SearchList!{wordcount}{\stepcounter{words}}
%     {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
%     n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
%   \UndoBoundary{'}
%   \SearchOrder{p;}}{%
%   \StopSearching}

% \usepackage[style=mla,backend=biber]{biblatex}
% \addbibresource{references.bib} % you would use your own bib file here

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}

\begin{document}
  \begin{center}
  \large{\bf Response Paper 1} --- 
  Ali Alkhatib
  \end{center}
  \textit{A person's community of origin is an irrelevant factor in emerging modes of digitally enabled labor.}

% \begin{counted}
% The study of Human--Computer Interaction (HCI) has come to the realization
% --- arguably several times ---
% that the emergent cultures of digital communities are deeply complex
% and substantively influential,
% contributing largely to the way communities of people perceive and react to the world
% \cite{dynamo,turkopticon}.
% Having realized this,
% substantial effort has studied the practice of designing and fostering online communities and cultures
% in an effort to better--understand and perhaps ultimately manage this factor
% \cite{successfulOnlineCommunities,earl2011digitally}.

% The body of research inquiring into online cultures and the study of shaping digital communities continues to grow,
% but the space describing the interplay between what we might simplistically refer to as ``offline culture'' and ``online culture'' remains relatively untrodden.
% Some research explores the digital embodiment of offline culture
% \cite{boellstorff2008coming},
% while others have thought more deeply about the adoption of technology given cultural contexts
% \cite{toyama2015geek,postcolonialComputing},
% but few in the realm of HCI seem to pay adequate attention to the dramatic influence that one's offline identity has on online participation.

% Instead,

In \textit{Geek Heresy}, Kentaro Toyama identifies three major world views describing the potential impact of technology on society:
\begin{inlinelist}
  \item ``utopians'',
  \item ``skeptics'', and
  \item ``contextualists''
\end{inlinelist}
\cite{toyama2015geek}.
% The thesis of his text drives the argument that both extremes
% --- the utopians and the skeptics ---
% make reductive assumptions about the importance of human factors in technological interventions.
From the perspectives of utopians, the promise goes
that technology will universally and unwaveringly improve our lives, and thus implicitly
assumes that people will adopt technological interventions in fundamentally the same ways that designers and engineers adopted their own creations.
Naturally, failures of programs like these, such as One Laptop Per Child (OLPC), embolden skeptics, who point out that
if anything, the positive outcomes of technological interventions are exceptions to the rule.
Skeptics then argue,
perhaps making similar assumptions that cultural and human factors are homogeneous,
that technology will unwaveringly marginalize and disempower the very people on whose behalf technologists ostensibly intervened.

Toyama argues a more cautious outlook, which he calls the ``contextualist'' perspective.
That is, the outcomes of technological interventions depend on myriad ``human factors''.
While Toyama's case studies very narrowly discuss humanitarian interventions
such as providing children in under--served schools with devices such as tablet computers,
and global access to the Internet through projects like \texttt{Internet.org},
his underlying thesis provides an important insight into other aspects of Human--Computer Interaction (HCI) and Computer--Supported Cooperative Work (CSCW).
Research in these fields is slowly making the realization that
an individual's culture is deeply intertwined in that individual's participation in online communities.
Perhaps nowhere is this more apparent than in the ways culture influences digitally enabled labor.

This paper will attempt to show that
culture deeply influences the emerging modes of digitally enabled labor that we call ``piecework'' by both enabling and challenging this kind of work.
Indeed, I'll show that attempts to ignore and abstract these factors by market--operators are causing, or at least exacerbating,
frustration,
mistrust,
and dissatisfaction
among workers in these nascent markets.
This discussion will consider a number of cases in some depth, namely
\begin{inlinelist}
  \item Amazon Mechanical Turk (AMT) and other information work markets,
  \item Uber, Lyft, and other embodied piecework markets, and finally
  \item we'll see that the culture of piecework effectively reflects the developing culture of transient participation
\end{inlinelist}.

A substantial body of research in the Human--Computer Interaction space has focused on information work on markets like Amazon Mechanical Turk (AMT), UpWork, and CrowdFlower
\cite{taskSearch,turkopticon,crowdworkFuture,foundry}.
% However, some researchers have explored embodied piecework in markets such as Uber
% and
% research generally in crowdsourcing might inform this inquiry as well
% \cite{uberAlgorithm,redballoon}.
Our discussion will start here, with information work,
in particular looking at the turmoil ``Turkers'' endure and the challenges they face,
made more intransigent by attempts by market--operators like Amazon to stifle open discussion of norms.
% leading to piecework through marketplaces such as Uber, TaskRabbit, and others.

Brabham discusses a number of issues which permeate digital labor such as
ethical concerns,
questions surrounding intellectual property,
and difficulties managing culturally diverse groups of workers
\cite{brabham2013crowdsourcing}.
These all illustrate the relevancy of offline norms and culture on digitally enabled labor ---
as legal scholars struggle to apply existing laws to emerging forms of work,
we see the statuses of workers and even the nature of digitally enabled work itself change over time
\cite{fedsUber}.
But these issues illustrate that digitally enabled labor can be a setting upon which offline issues are projected.
I would assert that these problems can be disentangled by negotiation with legal entities, and are arguably solvable given the systems that exist.
% But these examples don't speak to the core thesis.
% In fact, they seem to suggest that the model of projecting offline culture onto online settings is the correct one.

I argue that a deeper problem affects information labor markets, most visibly evident on Amazon Mechanical Turk (AMT).
If we look to the internal turmoil among Turkers,
and the struggles they face as they wrestle with notions of collective identity,
we see that the design of AMT itself stymies communication among workers, and even between workers and requesters
\cite{crowdworkFuture,Ross}.
These studies further illustrate a contentious and changing sense of self among Turkers,
attempting to reconcile differing views on propriety, tolerating mistreatment, and ways of responding to egregious abuses.
Differing views on all of these issues, and especially the perception that some workers are not as invested as others, foments a sense of mistrust among workers.

It's worth taking a moment to discuss the negative repercussions that follow when a culture of mistrust forms;
Sennett's (2007) work proves especially informative in this context.
In particular, Sennett describes three social deficits, one of which being fundamentally related to trust.
This issue of a lack of trust in an organizational setting can be caused in part by radically changing personnel, a characteristic of piecework often framed as a feature along the lines of the interchangeability of workers
\cite{sennett2007culture}.
Issues such as mistrust make it difficult for participants in these markets to rely on one another,
as shared investment in the quality of the market isn't felt,
precipitating a ``market for lemons'' and ultimately a depreciation of the value of labor
\cite{akerlof1970market}.

Efforts have been made to benefit workers by facilitating communication among workers internally (e.g. Turkopticon) and between workers and requesters (e.g. \texttt{WeAreDynamo.org}, or ``Dynamo''), and these efforts are worth some consideration \cite{turkopticon,dynamo}.
For example, in the drafting of commons--written ethical guidelines for academic requesters,
Turkers themselves note that workers from different backgrounds will bring different expectations regarding issues such as pay, communicativeness, and responsibility to worker well--being,
and that considerate requesters should make appropriate decisions depending on the workers they solicit
\cite{dynamo}.

This finding is important for a number of reasons,
but within the context of the overarching argument it's critical that
the problems Turkers face are baked into AMT, seemingly by design.
Resolving the tensions researchers found between Turkers from the U.S. and Turkers from other countries
(where costs of living differ dramatically)
would require a substantial redesign of the marketplace to allow workers to collectively agree on 
guidelines,
norms,
and ultimately standards of practice.
Amazon's response to these complicating factors has increasingly been to limit who may use AMT either as a requester or as a worker, at one point only allowing those with American tax records to register to use the service at all.

So it was empirically true that
a set of norms could be agreed upon by workers from various backgrounds with differing cultural views,
but it was an endeavor not without its challenges.
More troubling, the issues of differing norms and expectations from one worker to another were never unambiguously resolved;
the guidelines recommend even now that requesters consider the unique circumstances of individual workers, and behave accordingly.
In this way, the guidelines set in place by Dynamo corroborate the thesis of this paper
--- that cultural factors stemming from an individual's ``community of origin'' strongly influence the individual's participation in digitally enabled labor.
In this case, it proved too difficult a challenge to disentangle even given a system designed in part to mediate these contentious issues.

We turn now to digitally enabled ``embodied'' piecework
--- consisting of labor markets like Uber, TaskRabbit, etc\dots ---
where we find yet more culturally entangled issues problematizing work.
While the nature of these issues differs in some ways from the problems Turkers and other information workers face,
I argue they share a similar source: failure to consider the human factors
--- the cultures of participants ---
in designing and later operating these markets.

Shirky (2008) discusses a phenomenon in \textit{Here Comes Everybody};
``homophily'', or the grouping of like with like,
and how it greatly increases the likelihood of similar things being together (in this case in particular the grouping of people)
\cite{shirky2008here}.
Goldsmith and Wu (2006) talk of the same phenomenon without calling it by name,
in particular describing how the Internet allows us to be \textit{more} sensitive to the contexts of those with whom we interact,
rather than \textit{less}.
More concretely, they illustrate the emergence of hyper--local markets enabled by the Internet
--- the delivery of perishable goods
and the connection of very niche local interest groups\footnote{Shirky discusses niche local groups as well (especially in Chapter 8 --- ``Solving Social Dilemmas''), however Goldsmith and Wu's discussion proves more germane to digitally enabled labor.}
\cite{goldsmith2006controls}.

Given these perspectives, it might seem obvious in retrospect that digitally enabled labor markets
like Uber, Handy, and TaskRabbit
would eventually emerge.
However, even in these local marketplaces, we find friction and dissatisfaction among workers on a global scale.
Researchers have documented worker resistance of the systems imposed on them
\cite{uberAlgorithm}.
In at least one case, conflict between conventional and new--style workers has led to violence \cite{uberRiots}.

% \textit{talk about cultural differences between gig workers}

Miller (2011) relays the promise that technologists made,
that the Internet would abstract issues like geography away from us,
and allow us to interact meaningfully with people around the world
\cite{miller2011understanding}.
Instead, we find that, at least in some cases,
the Internet allows us to be more tightly--knit than before
(a finding corroborated by several studies of social network graphs
\cite{mcpherson2001birds,takhteyev2012geography}\footnote{Briefly, these papers find that many of the relationships found in social networks are geographically proximal, and not unbounded, providing yet more empirical evidence for the phenomenon of ``homophily'' that Shirky discusses.}).
In a literal sense,
an individual's community (in this case a geographic community) influences and perhaps even determines the networks they form socially and professionally.

% Having considered a

Having looked at the troubling issues entangled in digitally enabled labor, especially through the lenses of piecework both offline and online, one might leave with the impression that this kind of work is diametrically at odds with the interests of those who participate.
I would not argue that this is the case.
In fact, I will make the claim that
--- for all the problematizing characteristics to which researchers have pointed ---
its redeeming qualities have gone underreported, arguably detracting from our understanding of the culture of contemporary piecework.

One might observe that piecework in its contemporary instantiation is informed to some extent by the type of work done by those imagining and operating these markets.
Computer Scientists, trained to abstract problems and decompose them into smaller parts, have similarly imagined systems that abstract the human work implicitly and sometimes explicitly as inanimate parts.
On AMT, the portal for requesters to sign in illustrates the system of workers as a series of gears and cogs, belying the human workers under the metaphorical table.
In this way, the individual whose community of origin arguably affects the digitally enabled labor we discuss here is not the worker \textit{per se}, but another stakeholder rarely mentioned in discussions about this nascent market.

I would take this argument further, and say not only that it conforms to a familiar paradigm to system--designers, but that it also appeals to workers consciously looking to trade security (in this case, of employment) for flexibility.
Brabham doesn't make the point that crowdsourcing be capable of particularly rapid results, but this feature seems to emerge as a central feature of the idea
\cite{foundry}.
This aspect of crowdsourcing, allowing workers to come and go freely, parallels Shirky's (2008) thoughts on ``flash mobs'', wherein he describes the physical embodiment of what seems online like an abrupt, almost arbitrary appearance of hundreds or thousands of people
\cite{shirky2008here}.
But where Shirky writes about the scale of collective action enabled by digital media,
I find it more compelling that participants in these rapidly--formed groups can join and depart without that group suffering a substantial loss.

Shirky gestures to this feature of commons--based production in his writing on Wikipedia;
the fact that participants can come and go as needed allows them each to contribute a small part of the necessary whole
--- perhaps by making a single timely or worthwhile post ---
without the community relying unduly on any single contributor
\cite{shirky2008here}.
Brabham is careful to place Wikipedia and similar commons--based peer production systems outside of the scope of what he calls ``crowdsourcing'';
for the purposes of this overarching discussion, I can accept that exclusion, except to say that this characteristic remains a shared feature of crowdsourcing
--- that people can contribute as much or as little as they want without penalty for only superficial participation ---
is a crucial, shared element of piecework
\cite{brabham2013crowdsourcing}.
It is this aspect of piecework
--- a characteristic that is both increasingly familiar and amenable to people whose culture is similarly increasingly defined by their familiarity with the Internet ---
that makes it so tempting to young adults to participate.

This paper makes the argument that piecework is both enabled and problematized by the cultural forces at work in the marketplaces that have emerged thus far.
Namely,
\begin{inlinelist}
  \item we saw that information work like Turking endures turmoil over the conflicting paradigms of the heterogeneously composed workers who make up these workforces,
  \item we considered the conflicts that emerge in embodied piecework through markets like Uber, and
  \item we found that the cultures of system--designers as well as users informs and even determines the outcomes both of these systems as well as users' participation in these systems
\end{inlinelist}.
Perhaps what remains is the question of how to wrestle with the conflicting cultural factors that affect piecework,
and how we can enable piecework labor markets that attempt to reconcile the cultural issues complicating this kind of work without abandoning the features which appeal to workers in existing markets.
% and how we might enable work that appeals to workers for the reasons that piecework is so appealing --- offering features such as flexibility --- without making these marketplaces as disempowering, frustrating, and otherwise marginalizing as existing markets have been shown to be.
% Namely, while piecework challenges participants in this nascent labor market to find ways of surviving and thriving,

% Indeed, the very features which define piecework appeal to younger workers by offering transient, low--barrier entries to engage in much the same ways that online services do.
% These conflicting forces ultimately do little to affect the broader thesis --- that the community from which one originates dramatically affects one's participation in digitally enabled labor markets, whether that community is substantively informed by offline experiences or an intermingling of offline and digitally mediated experiences.
%

% But this argument belies a more complex issue that we will identify and with which we'll have to engage.
% That issue being that culture is not a strictly offline artifact projecting itself online,
% but something that shapes and changes through participation in technologically enabled communities.
% In this way, a person's community of ``origin'' and the community that they form through digital labor
% --- whether that labor is information work like ``Turking'', or the kind of ``gig work'' represented by Uber, Lyft, and TaskRabbit ---
% arguably get entangled together, and certainly become relevant to each other.
% Ultimately, this paper will attempt to demonstrate that culture is a core component of digitally enabled labor,
% and that offline cultures have informed and continue to engage with this emerging form of labor.% digitally enabled labor.

% We can explore this thesis through several examples,
% seeing that cultural influences are not only relevant, but critical factors in emerging digital labor.

% Various examples in e--commerce show us that technology and community can be deeply connected.
% Goldsmith and Wu point out that
% e--commerce sites almost immediately ask users their location
% (or attempt to determine that information automatically)
% to identify what unique laws, norms, and practices to acknowledge;
% but more than this, they illustrate a unique feature of the Internet:

% That geographic constraints affect digitally enabled commerce is arguably a given.
% Among the various modes of exchange made possible by technology,
% other forms of digital labor prove more surprising,
% and certainly more vexing, to researchers.
% To illustrate this, we turn to digitally enabled information work such as ``Turking'',
% predominantly online labor conducted on marketplaces like Mechanical Turk by companies like Amazon.
% Here, we arrive at the notion of crowdsourcing Brabham defines in his book on the subject
% \cite{brabham2013crowdsourcing}.

%

% The case of click--work is a fascinating one; Amazon describes Mechanical Turk as ``\textit{artificial} artificial intelligence'',
% and suggests structuring ``micro--work'' in such a way that the desired output given a certain input would be unambiguous.
% A cynical view might see this as attempting to minimize the potential influence of

% Just as Brabham carefully defines ``crowdsourcing''
% to rule out things like commons--based peer production and open source project contributions,
% we need to carefully outline what we mean by ``labor'' and even ``digitally enabled labor''
% \cite{brabham2013crowdsourcing}.

% Using our lens, scoped very narrowly to see the influence that culture has on work, we'll see that digital information labor

% Before we dive into these issues,
% it will prove worthwhile to frame some of the relevant terms and ideas.
% Specifically, since this discussion will largely focus on digital enabled labor and culture,
% we should carefully scope what we mean by ``labor'' and ``work'',
% and similarly frame ``culture'' accordingly.

% As Turner discussed in lectures,
% attempting to define work and labor can become a
% labyrinthian search for characteristics that intuitively delineate what is work or labor from what is not.
% To use his example, it's at least debatable whether a researcher's free time spent thinking represents work;
% certainly, the product of that thought will contribute to his or her career output
% in the form of publications,
% but this substantively describes the ``cognitive surplus'' Shirky describes at length
% \cite{shirky2010cognitive}.
% What good fundamental argument is to be made to differentiate this work from that of InnoCentive

% In the study of Human--Computer Interaction, researchers have discovered time and time again that
% the influence of culture in digitally--enabled communities is far--reaching
% \cite{postcolonialComputing,toyama2015geek}.
% This realization has been corroborated by researchers in the social sciences and humanities
% \cite{boellstorff2008coming,ito2005technosocial}. % ito2010mobilizing,ito2009hanging,boellstorff2012ethnography}.

% In \textit{Geek Heresy}, Kentaro Toyama identifies three major world views describing the potential impact of technology on society:
% \begin{inlinelist}
%   \item ``utopians'',
%   \item ``skeptics'', and
%   \item ``contextualists''
% \end{inlinelist}
% \cite{toyama2015geek}.
% Toyama's thesis, that technological intervention can benefit communities,
% but that this intervention must be conscious of human factors,
% carefully dismantles the unbridled optimism widely held and implicitly spread by ``utopians'',
% who advocate strongly for free access to the Internet and more ubiquitous computing devices
% (see: \texttt{Internet.org}, One Laptop Per Child).
% Without addressing the underlying

% In his book, Toyama seems to avoid an underlying insight:
% that these idealists hope that technology can be applied with similar effects across cultures.

% % Toyama avoids discussing the sources of what one might call a ``utopian'' world view,
% % but in doing so he may have overlooked a crucial insight:

% the cultures we bring with us as we build and engage with technologically--mediated systems vastly informs our participation in those systems.
% Nowhere is this clearer than in the culturally interwoven fabric of economies,
% as we will see in the microcosm of digitally enabled labor.

% \end{counted}
% \footnotesize{word count: \thewords}
\bibliographystyle{acm}
\bibliography{../../references.bib}
\end{document}

### 2016-05-16 classes/comm220/assignment2/assignment2.tex
init!

%!TEX program = xelatex
\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,times}
\usepackage{hyperref}
% \usepackage{fontspec}
% \usepackage{xltxtra}
% \usepackage{xunicode}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}

% \doublespacing
\usepackage[tiny,compact]{titlesec}

\usepackage{xesearch}
\newcounter{words}
\newenvironment{counted}{%
  \setcounter{words}{0}
  \SearchList!{wordcount}{\stepcounter{words}}
    {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
    n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
  \UndoBoundary{'}
  \SearchOrder{p;}}{%
  \StopSearching}

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}
\def\labelitemi{\ }
\newcommand*\elide{\textup{[\,\dots]}}
\newcommand{\sectitle}[1]{\textbf{#1}\\}

\begin{document}

  \begin{center}
  \large{\bf Response Paper 2} --- 
  Ali Alkhatib
  \end{center}

\textit{Software has more power than law to shape the use of digital media.}

\begin{counted}

\sectitle{Introduction}
In 1996, John Perry Barlow published ``A Declaration of the Independence of Cyberspace'',
drawing a line between what he described as two worlds:
one, predicated on property, identity, and context;
and the other, predicated on ethics, thoughts, and egalitarianism
\cite{barlow2009declaration}.
Barlow's missive reflects a cultural sense deeply ingrained in cyber culture
--- both then and now ---
that people on the Internet are free from the shackles of offline social, legal, and cultural institutions.
% Barlow's emphatic claim that his is a culture disentangled from the hooks of physical embodiment is somewhat undermined by the geographic reference that his letter was written in Davos, Switzerland.
Indeed, his claim has resonated on the Internet, but it remains relatively untested;
are the patterns of use of digital media shaped primarily by software as Barlow seems to suggest,
or are they in fact primarily influenced by laws like the ones from which he so emphatically attempts to distance himself?
% do laws substantively influence and shape the use of digital media or are they, as Barlow hopes to

In this paper, I argue that the roles of software and legal institutions in shaping the use of digital media are more nuanced than a simple ranking of one over the other.
I will attempt to show that software potently articulates potential uses of digital media;
the law, in turn, outlines the constraints on how people may use digital media.
While these forces appear inherently to compete with one another,
I will posit that the bottom--up influence of software
and the top--down influence of law
help provide important guidance in shaping the ways people adopt and use digital media.

To substantiate this thesis,
we will look to three points:
\begin{inlinelist}
\item We will consider a handful of cases illustrating the enabling potential of software systems, including digitally--enabled work and the emergence of ``big data'' as a way of approaching previously overwhelming problems;
\item we will look to legal cases demonstrating the state's ability to restrict behavior on these systems, such as the application of existing and new laws to enforce laws prohibiting stalking, harassment, and abuse;
\item finally, we will attempt to reconcile these opposing forces with an analytic framework.
\end{inlinelist}
% To substantiate this thesis, we will see cases exercising the extremes of
% software enabling use,
% law limiting it,
% and finally we will consider a case of law and software in use together to control use absolutely.
% Specifically,
% This relationship is strongly evidenced in a number of domains:
% \begin{inlinelist}
% \item in digitally--enabled labor markets and the conflicts we see between these nascent systems and protections guaranteed by legal institutions; and
% \item in the tension between system--users and system--designers over privacy and the use of users' data.
% Given these discouraging examples of the potential interactions between legal systems and software, we will finally conclude with a look at
% % as revelations related to privacy break mental models informed largely by cultural (including legal) worldviews;
% \item open--source, ``free'', software development and specifically
% how these cases prove that software and the law can operate in a collaborative, rather than an adversarial, nature.
% \end{inlinelist}

\sectitle{The empowering force of software}
Digitally--enabled labor markets,
in particular those articulating ``crowdsourcing'' as described by Jeff Howe and others
\cite{brabham2013crowdsourcing,howe2008crowdsourcing},
have enabled such radical change in the formulations of labor that one might think it calls for Dalton's comparisons of primitive economies and market industrialism, which he points out are
``different\dots\ not in degree but in kind''
\cite{dalton1961economic}.
In other words, ``crowdsourcing''
--- at least as Brabham and Howe define it ---
is not just the scaling of work in one dimension, but changing the axis of work itself.
Brabham attempts to distance crowdsourcing from conventional labor
--- and even from other digitally--enabled labor such as ``production of the commons'' (e.g. Wikipedia) ---
though deeper consideration of Riis's turn--of--the--century exposition on piecework suggests important parallels 
\cite{brabham2013crowdsourcing,riis2004other}.

Nevertheless, the changes to labor seem more reminiscent of a ``great transformation'', 
on a scale of which Polanyi had written decades prior
\cite{polanyi1944great}.
In fact, through this lens, an argument can be made that some form of substantive transformation
--- the commoditization of labor as discrete,
modular pieces and the emergence of jobs rather than the notion of a ``career'' as such ---
is taking place after what might be regarded as a ``false start''
documented by Riis and later fought by early labor unions
\cite{ebbinghaus1999institutions,riis2004other}.

We can think about the emergence of piecework,
at the turn of the twentieth century and then again in the last decade,
as the introduction of systems
--- social in the first case, but predicated on software in the contemporary case.
Both advocate, in a bottom--up fashion,
unconventional paradigms of work. %, hoping for norms to emerge slowly over time.
In the first case,
the number of worker--advocacy groups in the United States skyrocketed between the late nineteenth and early twentieth centuries 
(until reaching what Hannan and Freemon hypothesized was a ``carrying capacity'' of labor unions)
and pressure from these unions pressed legal institutions to formalize many of the protections for which many of these organizations had originally fought
\cite{hannan1987ecology}.

% In this sense, the law formalized a norm advocated by collective--action groups,
% a phenomenon I argue parallels the events we are seeing today;
Systems make it possible for workers to communicate, rally, and coordinate to improve their circumstances
\cite{turkopticon,dynamo,uberAlgorithm}.
We see that software allows people to communicate and coordinate in ways that were previously unheard--of and otherwise considered impossible
\cite{shirky2008here,miller2011understanding,costanza2014out}.

The enabling influence of software is not limited to labor markets.
In \textit{Numbersense}, Fung (2013) makes an argument for the potential to use massive amounts of data
--- about the self and the other ---
to make crucial insights informing decisions
\cite{fung2013numbersense}.
His perspective and optimism is not unique;
Mayer-Sch{\"o}nberger \& Cukier (2013) describe countless uses of ``Big Data'' to inform and improve the lives of those who participate in the collection of data at scale
\cite{mayer2013big}.
While Lanier (2014) claims that ``no one in science thinks of big data as a \ldots\ silver bullet''
\cite{lanier2014owns},
the overwhelming narrative among scientists suggests an attempt to find almost god--like applications of ``big data''
\cite{carrel2012quantified,li2011understanding,wolf2010data,bell2009total}.
% In this sense,
Indeed, data is widely regarded as capable of solving any problem,
limited only by the talent of those who wield it\footnote{
Big data is not universally perceived as a cure for all problems;
Bowker (2000) takes a critical look at the limitations of categorization
% data as such,
and the quantification, measurement,
and logging of everything that the aforementioned research advocates,
and how this over--reliance may problematize our understanding of the world
\cite{bowker2000sorting}.}.

I argue that this is an inherent affordance of software,
to empower and enable those with access to these systems and tools to behave in new ways,
and that software has a potentially significant role in influencing people's use of digital media.

% I would assert that we should expect robust bottom--up efforts to protect and empower workers,
% and that the cultural shift among ``gig workers'' will precipitate laws codifying many of the protections workers ensured via these mechanisms.
% With that being said,
% for the bottom--up efforts I predict to be successful,
% they must provide workers with alternatives to existing markets,
% or at the very least the means to coordinate and effect change on those markets,
% in order to influence how markets operate.
% In this

\sectitle{The constraining influence of law}
While legal systems don't yet adequately protect workers in contemporary piecework markets,
colloquially known as the ``gig economy'',
recent events suggest that their institutions are beginning to act to formalize protections determined to be necessary by workers in these new markets
\cite{fedsUber,uberSuit}.
Further paralleling the historical example,
the argument has been made that more rigorous government involvement is necessary to protect workers
by articulating limits on how people may use digital media in the context of crowdsourcing and this mode of labor
\cite{fixingChaos}.
This line of advocacy fundamentally assumes that the role of law is to rein in potentially dangerous forces, such as digitally--enabled labor markets predicated on software.

% Mismatches in users' expectations regarding privacy proves a fertile ground for exploring how software enables certain uses, while law often limits it.

It should also come as little surprise that
laws have emerged outlining reasonable uses of data
and limiting how data can be used to harm users.
As Zittrain (2008) points out, privacy policies and content licenses
--- legally binding terms describing the relationships between parties of various natures ---
establish limits on the behavior of all parties involved
\cite{zittrain2008future}.
It is worth noting that these limitations thus constitute guarantees upon which others can rely,
which Zittrain argues is an important feature of the Internet,
especially well--illustrated in his discussion of Net Neutrality.

% Lanier (2014) more overtly argues for limitations on the ways technology and ``Big Data'' about people can be used
% \cite{lanier2014owns}.
% While some of the propositions he makes are novel\footnote{Lanier's suggestion that people be compensated by system--builders for the value of their data is particularly inventive},
% Lanier's arguments fundamentally rest on the principle that
% legal institutions ought to limit how users
% (e.g. advertisers and government intelligence agencies)
% use data.
% While in some sense this argument is prescriptive rather than descriptive,
% --- Lanier is advocating for policies that are not the status quo ---
Lanier (2014) and Bowker (2000) take a critical stance of the empowering nature of personal data at scale,
but these critiques crystallize the \textit{geist}
of the sense of empowerment that is widely felt in software.
Lanier's assumption appears to be that this is the default role of laws
--- to outline how people can behave
(in this case in the context of using digital media)
by drawing boundaries beyond which represent social and legal transgressions ---
and that the nature of any debate regarding
legislative intervention is fundamentally about whether it should draw an otherwise
artificial boundary on people's behavior with regard to digital media.
% In some sense, Lanier contributes to the thesis of this paper by operating according to the otherwise--invisible influence that is culture itself.

Privacy is but one narrow facet exposing law's influence of the use of digital media.
Laws regarding behavior on the Internet expose the attempts to limit the ways people use digital media,
as Citron (2014) and Anderson (2013) show.
Technology and software affords people the ability to run a website or service over the Internet;
Anderson (2013) describes the awkward attempts of ``offline police'' to become ``Internet police''
as they attempt to follow spammers, child pornographers, and drug dealers online;
in particular,
Anderson points to the applications of offline laws
which codify the limitations on acceptable behavior online
\cite{anderson2013internet}.

Citron presents a more focused persuasive case that the law should %later argues law should
--- and increasingly does ---
outline limits on the behavior that occurs on websites and through online systems;
transgressions such as cyber stalking and ``revenge porn'' \cite{citron2014hate}.
While, as Citron points out in his examples of cyber harassment,
there is ample need and motivation to revise and update laws,
% there is ample room to improve and revise laws to better--reflect 
% In cases like these,
it's hard to downplay the role of law in shaping peoples' use of digital media.
Moreover, here we see again that the primary force defining the boundaries on people's behavior
stems not from the affordances and influences of software, but the directives of law.

Law further shapes the use of digital media in increasingly challenging ways.
Mayer-Sch{\"o}nberger points out that ``policy makers are compelling \ldots\ data collectors to perfect the digital memory of all of us''
\cite{mayer2011delete},
but this isn't the whole story;
the ``right to be forgotten'' articulated by the European Union illustrates a seemingly conflicting force,
prompting legal scholars to attempt to reconcile the push to exploit the affordances of software systems 
while concurrently addressing ``an urgent problem in the digital age''
\cite{rosen2012right}.
Even in cases as contradictory as this one,
it's clear that law carries with it an inherent power to limit and bound behavior
--- in this case, the behavior of system--designers ---
thus overwhelmingly shaping the use of digital media.

% Here, we see a more contentious relationship between software
% --- which arguably enables analysis and insights previously unavailable ---
% and law
% --- which draws strict boundaries on the uses of data and software--systems ---

% We can see law and software influencing behavior in other aspects of our lives as well;
% the mismatch of users' expectations regarding privacy compared to
% the reality they experience illustrates how systems can afford for certain interactions
% which are intuitively predicated on cultural norms such as laws.
% More concisely, people's use of digital media is informed
% --- sometimes erroneously --- by expectations codified

% In this case, I'll show that people use systems according primarily to their expectations of privacy
% --- largely informed and affirmed by conventional systems such as laws ---
% and that those expectations are routinely violated by software--based systems.
% It is the mismatch between expectation and reality that surprises and frustrates users of systems.

% \textit{Insert section about privacy revelations; argument that problems exist because software is at odds with culture, formalized by law.}

\sectitle{Reconciling conflicting powers} % --- the good, the bad, and the ugly}
% It can be difficult to reconcile the power of software and the power of law as determining factors in the use of digital media, but 
This imposing power of governments and laws may be conceptually related to ``bio--power'', a term which Foucault uses to reference to how ``the human species became the object of a political strategy''
\cite{foucault2009security}.
I suggest that the power laws employ, to threaten to restrict human bodies, describes the paradigm on which all laws operate
--- and that this paradigm is fundamentally the reason that laws curtail and limit.

Foucault's insight arguably emerges when we look critically at the myriad censorship tools employed in China in the forms of ``The Great Firewall'', preventing access to certain sites and services, and the hundreds of thousands of censors who tirelessly flag and remove content deemed inappropriate.
King, Pan, \& Roberts (2013) illustrate this confluence of law and software, and how the Chinese government effectively ``clips'' the social ties of collective action efforts to stifle organizers' efforts
\cite{king2014reverse}.
By tightly constraining the enabling force of software systems,
and invoking the ethos of the police and law through references to ``Chacha'' --- the Internet police,
China expertly shapes its citizens' use of digital media
\cite{king2013censorship}.

\sectitle{Conclusion}
We've explored the ways that software and law influence and shape the use of digital media,
coming away with a complex and conflicted understanding of the relationship that these two forces have
with one another.
We saw that software and digital systems have the power to enable behavior of various qualities
--- both to empower and to disenfranchise people.
We also dove into myriad ways that laws curtail and limit behavior and the use of digital media
--- again to empower and, sometimes, to disenfranchise people.
From these cases, we can reasonably conclude that software, ``cyberspace'', the embodied world, and law are far more entangled and interrelated than Barlow (1996) would have liked to admit in his ``declaration of independence''.

It may be tempting to ascribe greater value to law or software over the other
as we weigh the potency of these forces to influence digital media use,
but this would be a mistake.
I would argue that it would be more accurate to say that both factors must interact,
and that the ways in which people use digital media are enabled and determined by software and constrained by laws.

% It may be more accurate to describe the relationships between law, software, and the ways in which people use the myriad digital media that emerge as a confluence of necessary and sufficient conditions.
% In the same way that a certain predicates may be necessary for a certain conclusion to emerge
% Zittrain discusses another issue which highlights the potential for law to limit uses,
% in this case pointing to ``patent [and copyright] thickets'' \cite{zittrain2008future}.
% Examples like these typify a conflict between software and 
% Here we see the counter--example to the risk of software being too powerful
% --- as we saw in the case of digitally--enabled labor markets ---
% in that laws governing patents, copyrights, and other intellectual property
% potentially stymie the use of digital media.
% The aforementioned cases might lead one to think
% --- and argue, as Gray does \cite{fixingChaos} ---
% a narrative that generally places software in an adversarial position with relation to law.
% That is, that technology and government are inherently at odds with one another,
% and that such is the natural relationship in which users of digital media find themselves participating.

% I hope to illustrate instead that law and software sometimes cooperate,
% more persuasively influencing people's behavior
% by working together than either force could have accomplished alone. This will be evident

% Arguably the most salient case of law and technology working together can be found in China,
% where ``The Great Firewall of China'' prevents Internet--users from accessing various services,
% and countless censors work endlessly to flag and remove certain data \cite{king2013censorship}.

% We will find,
% independent of the morality of the policies and systems themselves,
% that this synergy of technology and law overwhelmingly influences how people use digital media.

% I argue instead that law and software must interact constructively,
% fostering the formation of norms via affordances made available in software
% and affirmed by legal systems codifying those norms.
% We can see examples of positive relationships in a number of settings,
% but perhaps the best--documented example is that of open--source software development.
% In this space, we will find robust 
% \textit{Insert awesome optimistic case here: Maybe something about how FSF uses legal frameworks to make it possible to formalize a grassroots movement for free software. Gonna be supes awesome.}

% In both cases,
% conflicts with the law, a top--down power , % s (such as legal frameworks)
% make the importance of collaboration between software and the law crucial.

% \begin{inlinelist}
% \item worker--advocacy groups such as labor unions,
% \item laws designed to protect workers, and
% \item blorp
% \end{inlinelist}

% \begin{inlinelist}
%   \item the conflicts between digitally--enabled labor markets and legal frameworks protecting workers illustrate the turmoil we should expect to see when worldviews represented by systems built from software and those expressed by laws are not in alignment;
%   \item privacy of individuals in the landscape of digitally--mediated environments has become a contentious topic, due in part to the difficulties in reconciling how we perceive privacy with existing laws, both in the contexts of the \textit{spirit} and the \textit{letter} of laws;
%   \item final
%   \end{inlinelist}

\end{counted}
\footnotesize{word count: \thewords}

\bibliographystyle{acm} % this will cause an error if there are no citations
\bibliography{../../references.bib}
\end{document}

### 2016-05-16 classes/comm220/final/final.tex
init!

% % !TEX program = xelatex
\documentclass[11pt,titlepage]{article}
\usepackage{balance,graphics,setspace,parskip,times}
% \usepackage{fontspec}
% \usepackage{xltxtra}
% \usepackage{xunicode}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}

\doublespacing
\usepackage[tiny,compact]{titlesec}

% \usepackage{xesearch}
% \newcounter{words}
% \newenvironment{counted}{%
%   \setcounter{words}{0}
%   \SearchList!{wordcount}{\stepcounter{words}}
%     {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
%     n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
%   \UndoBoundary{'}
%   \SearchOrder{p;}}{%
%   \StopSearching}

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}
\def\labelitemi{\ }
\newcommand*\elide{\textup{[\,\dots]}}
\newcommand{\sectitle}[1]{\textbf{#1}\\}

\author{
  Ali Alkhatib\\
  ali.alkhatib@cs.stanford.edu\\
  \textit{Couldn't find an envelope; will coordinate to meet next quarter (if acceptable)}
}

\title{Final Paper\\
\normalsize{
  \textit{Digital media are an egalitarian force in American society}\\
  Communication 220
  }
}

\begin{document}
\maketitle

\sectitle{Introduction}
Since --- and indeed before ---
the 90s, digital media have promised to unburden participants of the baggage of conventional cultural mores
\cite{barlow2009declaration};
Barlow in particular describes a very crystallized notion of
``cyberspace'' and its ``independence of the tyrannies \dots impose[d] on'' its denizens,
independence from conflicts inherent to embodied existence,
and finally a space to form a culture of universal human equality ---
to foster an egalitarian culture that transcends geographic space,
as well as the holdovers of culture from which he and others like him wished to break.

But the track records of digital media have been mixed, at best;
even in cases where we see what appears,
at first glance,
to be the emergent characteristics of egalitarianism,
closer inspection reveals sexist, racist, and variously exclusive
communities that are ultimately prohibitively difficult for outsiders to penetrate.
More troubling,
deeper consideration reveals worrying parallels with the counterculture movement of the 60s,
namely the trend of experimenting of with nominally egalitarian communes,
which ultimately proved to be sites of intense sexually \& racially charged practices of exclusion.
% with some facets of the broader counterculture movement 
% and generally informed the sensibilities and milieu of the Internet.

At another level, we see that digitally--mediated sites of emergent culture
are being substantively designed and shaped by certain groups of people and,
importantly,
occupied and used by other groups of people, leading to a worrying stark hierarchy ---
% a homogeneous group of people
% --- whereas these sites are largely occupied by others ---
% leading to hierarchical structure almost reminiscent of pre--industrialism ---
that is, people largely divided according to whether
they control digital media or are controlled by them.

I will attempt to argue that
digital media and the sites that emerge therein are
at least potentially
similar to the communes of the counterculture era,
and that the evidence suggests that
many of these sites have similarly paralleled the egalitarian communities from decades past.
I will go further, however, to offer a more optimistic outlook:
that the sites of interaction mediated by digital media can more easily be informed by one another,
and that the failures of one community to empower its members equally
need not necessarily repeat themselves across communities,
as so many communes often did.

To make this argument,
we need first to make the case that
digital media have been egalitarian forces in at least some facets of society,
which we will attempt to illustrate through
\begin{inlinelist}
  % \item the WELL, a community whose values precipitated the broader ethos of the Web, which later emerged,
  \item peer production work
      such as the Free Open Source Software (FOSS) movement \& Wikis, and
  \item efforts toward collective action and solidarity in online communities
\end{inlinelist}.
These cases will make a very simple argument at face value
--- that digital media are monotonically egalitarian forces ---
but we will explore these cases in depth and identify problems these systems have faced,
and sometimes caused, for participants.

We will also explore the much broader case of
systems themselves as tools used to separate users from builders.
Namely, we will look to
\begin{inlinelist}
  \item the oppressive nature of ``crowdwork'' and the broader ``gig labor''
  --- novel names for a well--established practice of
  speculative work which started at least as early as the turn of the 20th century
  (but which has found new purchase in the last 10 years), and
  \item the use of ``big data'' to track and analyze users, and
  the power system designers wield over their users by tracking people as data
  % \item the cultural, gendered, and other factors which emerge even among developers and engineers
\end{inlinelist}.

\sectitle{Peer Production: FOSS and Wikipedia}
Numerous examples come to mind in making the case that
digital media generate egalitarian communities;
perhaps the most compelling examples are those of peer production and
in particular
the decentralized organization that typifies many Free Open Source (FOSS) projects
as well as the ethos of Wikipedia and the many ``Wikis'' which have emerged since.

Suffice it to say that many have pointed to Linux and other open source projects,
in Raymond's case
identifying these as illustrative of an alternative to the ``cathedral'' organizational paradigm by
offering access to a sometimes hectic, ad--hoc ``bazaar'' instead
\cite{raymond1999cathedral,weber2004success,crowston2006core}.
The influence of open source software has been well--documented,
and certainly it represents a compelling counter--argument to
the proprietary ownership paradigm which has otherwise dominated software
\cite{kelty2008two,weber2004success}.

It seems, however, that open source projects often have complicated levels of involvement
--- if not explicitly named, then certainly implied by varying levels of engagement and interaction
\cite{crowston2006core,mockus2002two}.
Kelty points out that, in the community of developers he studied,
most developers were male, into heavy metal, and culturally relatively homogeneous
--- an issue that will resurface later in our analysis of big data,
but is worth pointing to here in our consideration of whether and to what extent FOSS is in fact egalitarian
\cite{kelty2008two}.
Another issue, that many of these open--source projects are so sparsely engaged that
it may be difficult to assess whether these \textit{would} be sites of egalitarianism,
is difficult to speak to, but worth noting
\cite{barcellini2006users}.

Wikipedia, by contrast, deliberately affords participants varying levels of commitment to take on,
and importantly almost any person with access to the Internet and an ability and willingness to write
can participate in the community.
This peer production model,
described by Howe and others as similar but subtly different from crowdsourcing,
theoretically allows all to come,
make edits and suggestions,
and debate edits on their merits rather than the status of the proponent
\cite{howe2008crowdsourcing,atlee2008collective,benkler2013peer}.
Here then we see at least a somewhat viable argument that
digital media facilitate the formation of egalitarian communities.

Shirky makes a similar point toward crowdsourcing in his discussion on the observation that
``everyone is a media outlet'' in this new setting
\cite{shirky2008here}.
His outlook is less optimistic
--- that the \textit{professions} involved in journalism and media have changed
in their mixing with ``amateurs'' ---
but if nothing else this contributes to a broader argument that
digital media have democratized the ability to publish content%
\footnote{%
Mackinnon (2012) discusses this effect to some extent through her analysis of WordPress as well,
but her discussion is largely set in China; while her findings are arguably still relevant
to American society, Shirky's general discussion above suffices
\cite{mackinnon2012consent}.}.

Unfortunately, again we find that reality differs from the promise;
researchers have discovered over the years that significant,
often implicit barriers to participation exist
without sufficient explanation for potential new editors
\cite{halfaker2012rise,zittrain2008future}.
Keegan \& Gergle in particular highlight the hypocrisy of
Wikipedia's stated ``egalitarianism'' in contrast to the reality that new editors experience
\cite{keegan2010egalitarians}.
This backlash that users often face has been documented and,
used to conclude that ``Wikipedians are \textbf{born}, not made'' (emphasis added)
\cite{wikipediansBornNotMade}.

\sectitle{Collective action}
Given the problematic nature of FOSS projects
and Wikipedia's enduring challenges to encourage new users to participate as editors,
one might conclude that
digital media unequivocally results in these problematized instances of egalitarianism.
Instead, I suggest we look further for evidence that
digital media represent forces toward egalitarianism, if not egalitarianism itself.

Collective action coordinated by digital media has,
Shirky and others argue,
at least the potential for democratized organization and planning
\cite{shirky2008here,costanza2014out}.
Here then, in some limited (and even stretched) sense,
digital media as democratized tools to facilitate collective action become
egalitarian influences in their own rights.
Miller in particular illustrates this notion through the use of the ``rhizome''
as a way of describing networked societies
\cite{miller2011understanding}.

Further, the tools needed to expose mistreatment of pieceworkers at the turn of the 20th century
--- namely, the camera Jacob Riis brought into the homes of workers ---
were made more accessible through the propagation of digital media and the tools needed to make them,
to say nothing of distributing that media,
made dramatically easier by blogging such as WordPress and other systems like it
\cite{riis2004other,facesOfMechanicalTurk}.
Further,
communities like Turkopticon \& Dynamo afford members some sense of equality among one another,
structured by guidelines on behavior to ensure equal opportunity to present views and
% some sense of anonymity and
separation from risk of controversial or contentious views affecting them
outside of the context of that forum
\cite{turkopticon,dynamo}.

\sectitle{Crowdwork, gig labor, and ties to piecework}
Having briefly mentioned communities of crowdworkers,
% (``Turkopticon'' itself is a play on the words ``panopticon'' and ``[Amazon Mechanical] Turk''),
we should acknowledge the unique role that crowdwork and gig labor play in this argument,
and in particular use this to segue from the notion of digitally--situated communities of egalitarianism
(or, as I have argued, \textit{not}) toward a much more general critique of
these communities as the boundary itself between the user and the builder.

Nominally, the gigwork enabled by Amazon's Mechanical Turk platform
(and other systems such as Upwork, Elance, etc\dots) make it possible for workers
--- who otherwise could not participate in the economy ---
to engage meaningfully, even earning a living.
In practice, however, these workers struggle to make ends meet
\cite{Ross}.
These markets for ``information work'' are, it turns out, far from unique;
markets for drivers for hire, cleaners, movers, and other various tasks
prove stifling and challenging for workers to benefit from.

Cynically,
one might argue that
the case for egalitarianism is still being made within the context of this workforce;
that is, everyone is struggling roughly equally compared to one another\footnote{This,
it turns out, is not true;
Irani et al. found that payment stratification benefits more senior Turkers,
a finding that has been corroborated incidentally by others
\cite{turkopticon}.
Still, we could cede this argument entirely in favor of the larger one to be made,
that egalitarianism must take into account the steep inequalities between various stakeholders.}.
This argument is specious, but it might be pernicious enough to merit addressing;
simply, the notion of a community achieving egalitarian ideals must be met entirely.
Amazon Mechanical Turk
--- and indeed \textit{all} of the marketplaces which promote piecework ---
relies on gross information asymmetry between the worker and the employer.
\cite{turkopticon,silberman2010ethics}.
That information asymmetry prevents participants from interacting as equals,
fundamentally preventing this and any market like it from being a site of egalitarianism.
Indeed, even communities \textit{connected to} these markets potentially suffer the hazards of
preferential statuses in these markets, especially illustrated by heterarchical relationships,
such as seeking support in one context given high status in another
\cite{dynamo}.

Further, the challenges these workers face sometimes come from the system itself,
but crucially from the finite designs of engineers,
causing its own form of oppression that differs from the kinds of coercive forces that offline markets
--- often the results of innumerable influencing factors ---
in that we can trace and even design responses to these issues
\cite{uberAlgorithm,turkopticon}.

\sectitle{Big Data}
In the previous section, we considered the effect that human--made systems such as
Uber, Lyft, etc\dots
have on workers, and the potential to track the design choices of system--engineers compared to the
confluence of broader marketplaces.
In this section,
we'll explore another aspect of human--engineered systems that separates and devalues
some classes of people from others --- the use of ``big data'' as a lens to consider the implications of
how imbalanced the power is between users of a digitally mediated system and the designers of that system.

Two issues come to light in this discussion prompted by big data.
The first is the general issue that system--designers have control over
the architecture of a digitally--mediated environment in only vaguely similar ways to
a conventional architect's control over a physical, embodied space.
Lessig (2006) describes a useful mental model %for conceptualizing the ways various forces influence a body
wherein four forces act on a body:
\begin{inlinelist}
  \item architecture,
  \item market,
  \item norms, and finally
  \item laws
\end{inlinelist}
\cite{lessig2006code}.
As we will soon come to realize,
these forces are largely consolidated into the agendas of system--builders.

It's important to acknowledge that
developers of digital systems have enormous control over the \textbf{architecture} of that system
(e.g. designing a user's interaction with a space
in such a way that it encourages or discourages certain behavior,
going as far as influencing ``emotional contagion'' \cite{kramer2014experimental}).
Additionally, markets such as Uber, Lyft, Amazon, and others
demonstrably have almost --- and perhaps effectively ---
unilateral, complete control over the \textbf{markets} which they expose through tools such as
data lock--in and threats of banishment for violating their own terms.
Norms, it would seem, are more difficult for a system--designer to influence;
deeper consideration reveals, however, that system--designers take insights from myriad fields such as
psychology, anthropology, and sociology, to inform the architecture of their system and 
encourage the emergence of \textbf{norms} preferred by \textit{them}
\cite{anderson2013steering,successfulOnlineCommunities}.

We return to Kelty's observation that developers have been a relatively homogeneous community of people;
notable exceptions aside, most developers in contemporary Silicon Valley have been,
and continue to be, male and relatively culturally homogeneous
\cite{ullman2012close,kelty2008two}.
It would be challenging even for culturally--minded people to appreciate the differing world--views
expressed around the world; for people whose fields of study range engineering and computer science
--- and especially those with minimal to no training in the study of, or exposure to various cultures ---
deep,
\textit{emic} understanding of the cultures of the people who use their systems
may be impossible for the engineers who build these major sites of social interaction.

Given that cold reality,
to say nothing of the interpersonal relationships that take place on digital media,
it's perhaps unreasonable to claim that the designers of systems and the users of systems are ever
part of a single egalitarian society.
Those that build systems are fundamentally more empowered than those that use them,
and with little to no understanding, and limited interaction between these two groups,
it seems unlikely that the characteristics of egalitarianism
--- for instance, mutual respect and understanding ---
will emerge.
I will continue to argue that this power imbalance is further skewed
in favor of those that design, implement, and control systems,
due to their necessary access to (and \textit{de facto} ownership of) users' data.

The second issue that emerges is the core topic of ``big data''
--- that users' collections data often become hostage to system--designers,
and that this fundamentally disrupts any endeavor toward an egalitarian society,
especially in the case to which Barlow attempts to speak (that of cyberspace)
\cite{barlow2009declaration}.
While Bush (1945) predicts a future where
the transmission of information might be reduced to a triviality in cost,
he seemed not to consider the circumstances in which we find ourselves ---
that the information itself might be worth too much to trivialize its free transmission
\cite{bush1979we}.

Indeed,
information has become a cornerstone of many of the companies which drive digital media spaces
\cite{sennett2007culture,mayer2013big}.
It should not, then, be surprising that companies as diverse as
driver--for--hire companies like Uber \& Lyft,
social networking sites such as Facebook \& Twitter,
and providers of services throughout (Google, Apple, \& Microsoft)
all generally maintain as secretly as possible both
the nature and content of what they track, even from the subjects of said tracking.
Lanier (2014) argues interestingly that perhaps
companies should be obligated to compensate people for the value extracted from data collected on them,
but this proposition
--- while interesting ---
is both economically infeasible and technically challenging
\cite{lanier2014owns}.

% \sectitle{Gender, race, \& developers}

\sectitle{Potential to be different}
Thus far, I've described an almost dystopian outlook on
the prospects of digital media being egalitarian forces in society at all,
let alone American society%
  \footnote{I've avoided discussing this issue until now,
  except here to say that efforts toward global egalitarianism are qualitatively
  --- not quantitatively ---
  more complicated.
  That is, the myriad political regimes attempting to moderate and legislate the use of technology
  (discussed to varying degrees by Goldsmith \& Wu, King et al., and others)
  are in themselves challenges similar in complexity to those we discuss here
  \cite{goldsmith2006controls,king2013censorship,king2014reverse}.
  Finding ways to relate these political and cultural contexts holistically,
  emerging with a comprehensive world--view on digital media's ability to affect egalitarian influence
  on various cultures and global culture as a whole, therefore,
  would fall far outside the scope of this paper.}.
%
There is however, reason to be hopeful, if not optimistic.
To say nothing of the efforts described earlier to enable discussion among people as equals
\cite{turkopticon,dynamo},
the broader field of Human--Computer Interaction benefits from insights both in design
--- leading realizations of the benefits of ``participatory design'',
wherein users and builders of systems work collaboratively ---
as well various fields of social and cultural inquiry
\cite{bodker1991cooperative,successfulOnlineCommunities}.

Where Turner points out how the counterculture revolution
and communes in particular
invariably became sites of implicitly defined prejudices,
I argue that our ability to learn from other communities rapidly and freely makes it possible
to avoid sliding into the same causes of failure which almost invariably struck communes
\cite{turner2005counterculture}.
Dynamo learned from Turkopticon and Wikipedia
\cite{dynamo,turkopticon};
bodies of literature are largely informed by existing communities,
the results of this literature often becoming the basis for new endeavors
\cite{bernstein20114chan,horowitz2010anatomy,Nov:2007:MW:1297797.1297798,wikipediansBornNotMade}.

All of this is to say that,
when we
--- in particular, researchers, but also system--designers and \textit{people} in general ---
recognize subtle, pervasive barriers discouraging people from participating in editing Wikipedia,
we can leverage mountains of research into encouraging participation,
leading to design interventions that precipitate and facilitate more inclusive environments.
Indeed, we can point to research on various fronts which corroborates this potential,
the most salient of which being the efforts to change Wikipedia's administrative practices
--- as well as the design of the system itself ---
to ``steer'' more constructive moderation of edits
\cite{ciampaglia2015moodbar,kraut2010dealing,pal2011early}.

\sectitle{Closing thoughts}
I've attempted to articulate a complicated perspective on,
and in fact relationship with,
digital media through their ability to be an egalitarian influence on American society.
Ultimately, my arguments attempted to illustrate that
numerous examples ostensibly settling this debate exist,
but that each of those examples is problematized by deeper,
underlying complications to the relationships between stakeholders.

I chose to end, however, with a call toward optimism to suggest that
it is not built \textit{into} digital media that egalitarianism is,
at best,
a false promise; that instead,
it may be possible to achieve the egalitarianism Barlow wishes to see in his declaration,
but that it will
--- as always ---
take our best efforts to avoid making the same mistakes we've made in the past,
marginalizing people and imposing our own world--views in an attempt to
homogenize cyber--culture as it forms.
I believe, with care and consideration of the past
--- a skill admittedly all too rare among those who need it most ---
system--designers can help construct a cyberspace that welcomes all as Barlow wrote in 1996.

% \end{counted}
% \footnotesize{word count: \thewords}

\bibliographystyle{acm} % this will cause an error if there are no citations!
\bibliography{../../../references.bib}
\end{document}

### 2016-05-16 classes/mse330/final/appendix.tex
init!

%!TEX root = final.tex
\newpage
\appendix
\section{Appendix} \label{appendixA}
Below we provide further details including confusion matrices and
reliability plots on some of the models used.

\begin{figure}[h]
  \centering
    \begin{subfigure}[p!]{\textwidth}
            \caption{Non--Quality of Life Cases}
            \begin{subfigure}[p!]{0.49\textwidth}
              \includegraphics[width=\textwidth]{figures/appa.png}
              \label{fig:AppA}
            \end{subfigure}
            ~
            \begin{subfigure}[p!]{0.49\textwidth}
              \includegraphics[width=\textwidth]{figures/appb.png}
              \label{fig:AppB}
            \end{subfigure}
    \end{subfigure}

    \begin{subfigure}[p!]{\textwidth}
                \caption{Quality of Life Cases}
                \begin{subfigure}[p!]{0.49\textwidth}

                  \includegraphics[width=\textwidth]{figures/appc.png}
                  \label{fig:AppC}
                \end{subfigure}
                ~
                \begin{subfigure}[p!]{0.49\textwidth}

                  \includegraphics[width=\textwidth]{figures/appd.png}
                  \label{fig:AppD}
                \end{subfigure}
      \end{subfigure}
      \caption{Random Forest Model}
\end{figure}
\newpage

\begin{figure}[h]
  \centering
    \begin{subfigure}[p!]{\textwidth}
    \caption{ROR Modeling using GLM}
        \begin{subfigure}[p1]{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/glmplots/ror_fc.png}
        \end{subfigure}
        ~
        \begin{subfigure}[p1]{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/glmplots/rorq_fc.png}
        \end{subfigure}
    \end{subfigure}

    \begin{subfigure}[p!]{\textwidth}
    \caption{Plea Rates Modeled using GLM}
        \begin{subfigure}[p1]{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/glmplots/plea_fc.png}

        \end{subfigure}
        ~
        \begin{subfigure}[p1]{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/glmplots/pleaq_fc.png}

        \end{subfigure}
    \end{subfigure}

\end{figure}

\subsection{Validating GLM}
With a half and half test/train split,
the ROR model had acceptable performance,
while the plea model's performance was rather unremarkable.
Though their accuracy rates are not abysmal,
it is important to keep in mind that
both outcomes being predicted have relatively high frequency in the data
(ROR rate is close to 65 percent and plea rate is around 50 percent).

### 2016-05-16 classes/mse330/final/conclusion.tex
init!

%!TEX root = final.tex
Bail--setting remains an ill--explored area of research;
this is particularly the case given its critical role in the criminal justice system.
Moreover, the nature of research in this space makes it challenging to
run analyses on substantial quantities of data, as we were able to in this case.
By adopting a number of modeling approaches and methods of analyses,
we were able to verify the effect that pretrial incarceration has on
a defendant's willingness to plead guilty.
Perhaps unsurprisingly,
we found evidence that offers credibility to the fear that bail--setting carries
an undue impact on the process.

Our first attempts to quantify and measure
a dimension that was previously under--explored
suggests that people are experiencing worse outcomes than others
on the basis of their inability to post bail.
While these findings suggest that the courts are not providing the fairest possible
experience to all those who pass through the justice system,
we see these findings as a silver lining.
Indeed, this research begins the discussion of
quantifying and
measuring
the extent to which pretrial incarceration is affecting defendants' plea rates.
This tentative step of measuring, and perhaps tracking, the effects of
pretrial incarceration may yield insights to ensure that defendants are receiving
just and due process.

### 2016-05-16 classes/mse330/final/discussion.tex
init!

%!TEX root = final.tex
As described in the results above,
we find that our first hypothesis,
that a more burdensome bail will increase likelihood for a defendant to plead out,
is confirmed.
Our secondary hypothesis is also confirmed,
though less strongly.

More specifically,
we observe that in non--QoL cases,
there is a substantive distinction between the first two burden levels,
which do not involve jail time,
and the latter two burden levels,
which do.
Jailing a defendant
increases the predicted probability of pleading
compared to burdens not involving jail time.
We observe a similar,
though slightly smaller effect in QoL cases.
With regards to our second hypothesis,
we find that despite differences in the cases received by judges
(which are unfortunately not randomized),
some variance in the handling of cases by judges
can be attributed to the specific judge handling the case.

These two findings together underscore the importance of an objective,
randomized,
and well--informed justice system in order to achieve fair outcomes.
The degree to which a judge is informed or able to estimate
a defendant's means and ability to pay bail can
make a significant difference in outcomes for that defendant.
This suggests that work aimed at estimating
accurate and
straightforward predictive heuristics for judges to use in setting bail is
a promising strategy towards facilitating fairness for all defendants.
Likewise,
individual differences in setting bail provide the opportunity for
further statistical analyses to discern the effect of
different bail--setting norms on defendant outcomes,
and again suggest that helping judges adjudicate consistently over time and
in comparison to each other can
help make the justice system more fair towards defendants.

### 2016-05-16 classes/mse330/final/final.tex
init!

% !TEX program = xelatex
\input{../../../starter.tex}
\setmainfont[
Ligatures=TeX,
Extension=.otf,
UprightFont= *-regular,
BoldFont=*-bold,
ItalicFont=*-italic,
BoldItalicFont=*-bolditalic]{texgyretermes}

\title{The Effect of Bail on Guilty Pleas\\
\large{\textit{MS\&E 330}}}
\author{
Ali Alkhatib, Nikhil Garg, Martin Lindsey, Danaë Metaxa--Kakavouli
\\
\{\href{mailto:al2@stanford.edu}{al2},
  \href{mailto:nkgarg@stanford.edu}{nkgarg},
  \href{mailto:mlindsey@stanford.edu}{mlindsey},
  \href{mailto:metaxa@stanford.edu}{metaxa}\}@stanford.edu}

\bibliographystyle{apalike}

\begin{document}

\begin{titlingpage}
  \maketitle
\end{titlingpage}

\begin{abstract}
Research on the subject of
bail,
the burden it places on defendants, and
the persuasive and perhaps coercive influence it bears on the justice system
has been explored to some extent ethnographically, but thus far
researchers have not had the data necessary to conduct rigorous quantitative studies on
bail--setting and
the factors which influence pleading behavior. % and the influencing factors therein have not had sufficient data
With a unique insight into the demographics and characteristics of defendants and their outcomes,
we explore two questions:
\begin{inlinelist}
\item Does a greater burden, as measured by bail, cause people to plead guilty? And
\item Are judges handling cases uniquely, or uniformly?
\end{inlinelist}
Our findings reveal that heightened burdens do seem to cause defendants to plead guilty at a higher rate,
and that there is some evidence that variance in bail--setting is unique to judges.
\end{abstract}

\section{Introduction}
\input{intro}

\section{Literature Review}
\input{related_literature}

\section{Method}
\input{method}

\section{Results}
\input{results}

\section{Discussion}
\input{discussion}

\section{Conclusion}
\input{conclusion}

\input{appendix}

\newpage
\bibliography{../../../references.bib}

\end{document}

### 2016-05-16 classes/mse330/final/intro.tex
init!

%!TEX root = final.tex
Maintaining an efficient justice system is in the interest of both the professionals managing and the citizens interacting with it.
Computational methods including large--scale data analysis and machine learning excel at quantitatively identifying points of inefficiency.
We propose to use these techniques to build models predicting points of inefficiency in the pretrial processes,
in order to provide recommendations for improving efficiency to judges and other court officials.

Notably,
a large majority of cases are never sent to trial; rather,
charges are dropped or settled out of court.
Pretrial procedures,
then,
such as the setting of bail or decision to pursue a case,
are especially fruitful sites for consideration with regards to efficiency.
A more insightful understanding of how to approach these pretrial procedures will benefit court officials in the time and money spent processing cases needlessly.
For defendants,
these insights could lead them to spend less time in prison awaiting trial if they are unable to pay bail,
and fewer resources towards a case that will never be brought to trial.

With this potential for intervention in mind,
we propose to examine the Gotham dataset,
using case features available to the judge,
and in particular bail,
which is set by the judge,
to predict whether a defendant will plead or not later in court proceedings.
The goal of this procedure is to analyze one attribute of a defendant's process through the justice system from the perspective of a judge,
including information the judge knows about the case and,
most importantly,
the decision a judge can make: setting bail.

Specifically,
we conceive of bail setting as imposing a type of (intentional) burden upon a defendant.
This burden can occur at several different levels,
from release on a defendant's own recognizance (ROR) --- a nonexistent burden,
to refusal to set bail at all in order to keep a defendant in custody --- a very high burden.
In between these two levels are bails set which the defendant is able to pay (or find a bondsman to put up bond),
and bails which the defendant is not able to pay.
These two categories are irrespective of the bail amount,
and instead are specific to the means of the defendant who is either able or not to pay bail.
Similarly,
it is worth considering whether different judges set bail in substantially different ways given similar cases,
as a starting point for different analyses.

We make \nextitemizecount{} hypotheses:
\begin{enumerate}
\item Effect of burden on pleading:
      that placing a larger bail burden on a defendant will correlate with a
      higher likelihood to plead later in the case.
      We posit a causal psychological mechanism by which
      defendants who have been required to put up substantial financial assets or,
      even further,
      had to spend time in jail awaiting trial are placed in a vulnerable position
      by the justice system and
      therefore are less likely to prolong their case defend themselves forcefully against charges
      --- in other words, more likely to plead guilty.
\item Our secondary hypothesis predicts that
      different judges will set substantially different bail on similar cases,
      since we anticipate that there is significant room for human error or
      difference this aspect of the justice system.
\end{enumerate}

The underlying goal of this work is
to provide predictions which aid judges in setting effective but
not excessive bail by
better understanding one way in which the process of setting bail influences
defendant experiences with and attitudes toward the justice system.

### 2016-05-16 classes/mse330/final/method.tex
init!

%!TEX root = final.tex
In order to analyze our two research questions,
we first quantify bail burden and preprocess the data,
and subsequently perform statistical analyses on the data.
The methods used in these steps are detailed below.

\subsection{Bail Burden}
We conceive of the possible burden placed on a defendant by bail as fitting into one of
\nextitemizecount{} categories:
\begin{inlinelist}
  \item When a defendant is Released on his or her Own Recognizance (ROR),
  the burden is negligible.
  Defendants are required neither to put up money nor to spend time incarcerated; 
  \item when bail is set and a defendant is able to pay it immediately,
  the defendant is lightly burdened; 
  \item when bail is set and a defendant is able to pay it but with a delay,
    the defendant is significantly burdened both
    financially and by
    needing to spend time incarcerated prior to meeting bail; 
  \item when a defendant either has no option to be released on bail,
  or is unable to meet bail,
    he or she is \textit{severely} burdened,
    having to spend the entire time until trial incarcerated.
\end{inlinelist}

\subsection{Preprocessing}
Several preprocessing steps were taken to ensure the validity of our analyses.
Due to some inconsistencies in volume of cases prior to 2013,
only cases from after (and including) 2013 were analyzed.
Additionally,
cases involving violent offenses were removed since
there are extra considerations and complexities in bail--setting for violent offenders,
who may be considered safety risks to the community and evaluated differently for bail.
We also only consider cases that made it past arraignment,
since cases where charges were dropped at arraignment are ineligible for the analyses we perform.
Finally,
for most analyses we consider Quality of Life (QoL) and non--Quality of Life (non--QoL) offenses separately.
Quality of life offenses make up a category of illegal activity which is connected with the broken windows theory of policing
--- these are misdemeanor offenses such as
  disorderly conduct,
  vagrancy,
  or
  loitering.
QoL and non--QoL offenses may differ substantially in the populations being charged and in their handling in the justice system,
so we perform the same analyses in parallel across these two types of data.

\subsection{Analyzing the Effect of Bail on the Probability to Plea}
We analyze the effect of bail burden on the probability of pleading through
two main techniques and
using both random forest models and GLM.
First,
we train models to predict the probability of pleading using all case characteristics through arraignment,
including the burden level set.
Then we create test sets composed of different burden levels
(through two different techniques,
described next),
and analyze the test set probability of pleading conditional on each burden level.
We present the results both as a CDF of the test sets' probability of pleading and as means of the distribution.

To verify the effectiveness of our models,
we produce calibration plots and include them in the appendix.
For each row in the original test set,
we predict the probability of pleading and ROR,
respectively,
for our two models.
We bin these probabilities and observe the true probability of observing the dependent variable for each of the bins.

\subsubsection{``All Else Equal''}
For our preliminary analysis,
we create the test sets for each burden level through an ``all else equal'' assumption
--- that we can change the burden level of a given case while holding all other case characteristics the same.
We duplicate the test set into four sets.
In each data set,
we set the burden level to
  ``ROR'',
  ``Trivial'',
  ``Trivial with Delay'',
  and
  ``Jail'',
respectively.
We then send these test sets to the model and view the resulting predicted probability of pleading.
We note that the ``all else equal'' assumption is not true in general.
In a murder case,
for example,
it would not make sense to change the burden level from ``Jail'' to ``ROR''.
However,
because we analyze only non--violent cases,
the assumption holds approximately.

\subsubsection{Propensity Score Matching}
To further support our analysis,
we relax the assumption through initial propensity score matching.
For this method,
we split the burden into only two levels: ROR and Non--ROR.
We train a model to predict the probability of ROR.
Two test sets are then matched on the probability of ROR through \texttt{matchit}
\citep{matchit},
and the burden bits are set to ROR and Non--ROR,
respectively.
They are then sent to the probability of pleading models.
To do true propensity score matching and make sure our assumption holds,
we should have filtered out cases with either very low or very high probability of ROR.
Otherwise,
we are still using cases where it would not make sense to change the ROR value.

\subsection{Analyzing the Effect of Judge on Bail}
\subsubsection{Analysis --- Random Forest}
For an initial analysis to observe judge effects,
we predict the Probability of ROR being granted for each case,
without using \texttt{JudgeID} in our prediction model.
This analysis uses the same Random Forest ROR model trained in the Propensity Matching analysis.
This model gives a baseline for judge independent granting of ROR.
For each judge with over 100 cases in the test set,
we determine the judge's mean predicted probability of ROR using the model.
This value is the percentage of a given judge's cases that the ``average'' judge would ROR.
We compare this value to the true percentage of cases in which each judge actually granted ROR.

\subsubsection{Analysis --- Generalized Linear Model}
An ``all else equal'' approach was taken in the GLM analysis
--- the general strategy was to create a copy of the data with a test variable set to a given value,
input these modified datasets into the model for each value the test variable can assume,
and then obtain summary statistics for each possible value
to then compare with the summary statistics associated with the other values;
by doing this,
any predictive power this model might have can be used
to check for the relative influences of certain factors on the outcome that this model predicts.

The ROR model was used to check if certain arraignment judges differed substantially in how they treated ROR.
The plea model was used to check if certain levels of burden had a higher predicted rate reaching a plea disposition.
As a cursory additional check,
matching edge cases were identified using a software package
(i.e. cases that can be considered similar on model inputs but differed on observed outcomes),
and in each case,
the same analysis was performed on a reduced dataset of these edge cases to similar results.

Within this reduced set of edge cases,
the assumption of ``all else equal'' has greater validity,
as the matching software is trusted to compute a quantitative measure of this
``equality'' between cases and identify precisely those which are ``sufficiently equal''.
The effects of judge on ROR are summarized in
figures~\ref{fig:ROR_QoL_GLM} and
        \ref{fig:ROR_non--QoL_GLM}.
For non--QoL cases,
judges seem to have more similar ROR rates compared to QoL cases
(one of the most prolific judges is also an outlier in the QoL dataset).
The effects of burden on the rate of plea dispositions are summarized in
figures~\ref{fig:QoL_plead_GLM} and
        \ref{fig:non--QoL_plead_GLM}.

For both QoL and non--QoL cases,
the lower levels of burden had close distributions,
but the distribution for ``nontrivial'',
the highest level of burden had its mass shifted noticeably toward higher plea rates.
Additionally,
the predicted plea distributions on the QoL dataset had mass shifted more to the right compared to the non--QoL distributions.
Analogous plots for the same analysis performed with the ROR model on the matched non--QoL dataset are found in
figures~\ref{fig:mean_ROR_non--QoL_GLM} and
        \ref{fig:mean_ROR_QoL_GLM};
the other combinations of QoL or non--QoL and ROR or plea model yielded matched datasets that were too small
(i.e. on the order of tens of cases) to use.

### 2016-05-16 classes/mse330/final/related_literature.tex
init!

%!TEX root = final.tex
Broadly, we identify \nextitemizecount{} areas of research which provide some insight into this topic:
\begin{inlinelist}
  \item The use of
    \textit{\lnameref{subsec:bail}},
    and its stated and empirical purposes,
  \item the study of the effect that bail has on
    \textit{\lnameref{subsec:burden}},
    and
  \item the role of
    \textit{\lnameref{subsec:plea}},
    both for the courts and defendants.
\end{inlinelist}
We will discuss the intersection of these bodies of work as they relate to our research question.

\subsection{\nmu Bail}\label{subsec:bail}
In this section, we will explore the murky waters of the use of bail,
but ultimately settle on the conclusion that the courts have made:
that bail is fundamentally a necessity in a judicial system,
despite the philosophical questions and issues it inherently raises.
Despite the need for bail as a robust safeguard against undue imprisonment,
we will illustrate that the practice of bail--setting is
--- perhaps decidedly --- difficult to discuss in broad terms,
often leaving researchers primarily with
    qualitative,
    ethnographic, and sometimes
    anecdotal
evidence upon which to work.

Bail represents an attempt to resolve a challenging cognitive dissonance.
As \citet{foote1959bail} points out, the imprisonment of a defendant who is nominally considered
``innocent until found guilty''
arouses challenges of principle;
if the courts are to assume that a defendant is innocent (and thus treat them as such),
then holding them in custody until the trial begins is patently wrong, as it deprives a
\textit{presumably} innocent person of freedom.

Nevertheless, the United States Constitution's 8th amendment
--- as well as a series of laws attempting to reform bail
(one in 1966, another in 1984) ---
ostensibly protect defendants from unreasonable and excessive bail amounts,
otherwise to be set at the discretion of the presiding judge
\citep{berg1985bail}.
As \citeauthor{berg1985bail} notes, however,
the most recent revisions to the Bail Reform Act
--- those passed in 1984 ---
make it explicit that judges may set or even deny bail on the basis of
the judge's assessment of the defendant's danger to the community.
In \citet{1984schall},
the court ruled that
``there was nothing inherently unconstitutional in preventive detention''
and that
``the restraint on [the defendant]'s liberty
did not amount to punishment because
that was not the express purpose of the pretrial detention''.

The precedents established in the cases of
\citet{1984schall} and
\citet{1987united} broadly supports the judge's prerogative to set bail
--- and importantly, to \textit{not} set it
(in other words, to remand the defendant with no option to pay bail),
if the judge determines that safety to the community
(or the defendant's appearance upon a later court date)
could not be assured by any bail amount.
This latitude becomes an important component that we attempt to explore
in the hopes of determining whether if
--- at least within a single jurisdiction ---
judge bail--setting variance is high or not.
The implication of this question may be that
some judges are more merciful,
and some decidedly harsher,
than their peers in their assessments of defendants.

\subsection{\nmu Burden}\label{subsec:burden}
The 1984 Bail Reform Act thus
afforded the courts a great deal of freedom in setting bail,
bringing us naturally to the question of what effect this has on defendants awaiting trial.
In this area, we find myriad researchers have contributed;
we will attempt to highlight some of the work that is
especially illuminating and
representative of the body of work.

\citeauthor{kellough2002remand} identify qualitatively that the
``moral assessments'' of defendants appear to influence the burden judges assign in the form of bail--setting.
This, they hypothesize, may explain
(at least to some extent)
the disparity in race between
those that are let free in the interim before their court date, and
those that are held in custody on remand
\citep{kellough2002remand}.

\citeauthor{zacharias1997justice} further discusses the burden that bail may impose on defendants,
in this case highlighting the leverage that prosecutors may have over defendants,
for instance in the form of the unfamiliar and unwelcome environment
that temporary incarceration represents, and the promise to expedite one's interaction with the justice system
\citep{zacharias1997justice}.
This, finally, leads us to the outcome variable which we will ultimately investigate.

\subsection{\nmu Plea \nmu Bargaining}\label{subsec:plea}
Most tangibly, the persuasive
(perhaps \textit{coercive})
force that prosecutors can exercise in tandem with judges is
the pressure to encourage defendants to plead guilty in the pretrial phase.
This aspect of the justice system is especially unique, both because
plea--bargaining defies the adversarial model of the court and trial process and because
these dynamics are so ill--understood at broad levels
\citep{zacharias1997justice}.

The research on the \lnameref{subsec:burden} that bail represents on defendants
--- especially defendants with limited financial means and similarly limited access to cash ---
has \textit{qualitatively} found that
trial outcomes may be affected by the burden of pretrial incarceration.
To illustrate this, we turn to
\citet{grossman1983plea} and \citet{alschuler1981changing};
\citeauthor{grossman1983plea} estimates that as many as
``90 percent of convictions \dots result from a negotiated plea of guilty''
\citep[see also][with similar findings]{mccoy1980plea,kaplan1977american,newman1966conviction}.

Most of the existing research in plea bargaining,
especially \citeauthor{grossman1983plea}'s work,
focuses on the economic role of the plea bargain as a
``screening device'';
\citeauthor{alschuler1981changing}
contributes to the argument that pretrial bargaining may serve the defendant's best interests,
observing that neither the courts nor the public defenders who often represent defendants
have sufficient time to see all cases through open court
\citep{alschuler1981changing}.
\citeauthor{alschuler1981changing}
makes the noteworthy observation that we lack sufficient information to determine whether 
  the pressure of incarceration,
  the prosecutor's persuasive offer to settle, and
  the defending attorney's desire to expedite the closure of outstanding cases
is in fact causing defendants to settle more frequently,
primarily due to a dearth of broad--based data, and certainly
--- as \citeauthor{kellough2002remand} describe it ---
a simple ``paucity of research [in plea bargaining]''
\citep{kellough2002remand}.

### 2016-05-16 classes/mse330/final/results.tex
init!

%!TEX root = final.tex
\subsection{Analyzing the Effect of Bail on the Probability to Plea}
\subsubsection{All Else Equal --- Random Forest}
We first carry out the All Else Equal analysis through a Random Forest model.
Appendix \ref{appendixA} shows the calibration information for the Random Forest models for both
Non--Quality of Life and
Quality of Life cases.
We observe that the models are accurate and serviceable,
though the Non--QoL model consistently over--predicts the probability of pleading in the unaltered test set.

Tables~\ref{table:non--QoL_plead} and
       \ref{table:QoL_plead},
respectively,
include the mean predicted probability of pleading at different burden levels for
Non--Quality of Life and
Quality of Life cases, respectively.
Figures~\ref{fig:non--QoL_plead} and
        \ref{fig:QoL_plead}, show the CDFs of same values determined through the model.
Applying the GLM to the dataset reveals roughly similar predictions, as illustrated by
Figures~\ref{fig:non--QoL_plead_GLM} and
        \ref{fig:QoL_plead_GLM}.

We observe that,
especially for Non--Quality of Life cases,
jailing someone substantively increases the predicted probability of pleading when compared other burden levels.
This effect can be observed through both the higher means and the right--shift on the CDF plots.
The effect is similar, though much less pronounced, for Quality of Life cases.

\begin{table}
\centering
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
  \hline
\textbf{Burden Level} & \textbf{Mean Predicted Probability of Pleading} \\ \hline
    ROR & 0.3631537 \\ \hline
    Trivial & 0.3834242 \\ \hline
    Delay & 0.3744825 \\ \hline
    Jail & 698843  \\ \hline
  \end{tabular}
  \caption{Mean Probability of Pleading for Non--Quality of Life offenses through All Else Equal Analysis using Random Forest model}
  \label{table:non--QoL_plead}
\end{table}

\begin{table}
\centering
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
  \hline
\textbf{Burden Level} & \textbf{Mean Predicted Probability of Pleading} \\ \hline
    ROR & 0.6072936 \\ \hline
    Trivial & 0.6102411 \\ \hline
    Delay & 0.5999771 \\ \hline
    Jail & 0.648984 \\ \hline
  \end{tabular}
  \caption{Mean Probability of Pleading for Quality of Life offenses through All Else Equal Analysis using Random Forest model}
  \label{table:QoL_plead}
\end{table}

\begin{figure}[t!]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figx.png}
    \label{fig:non--QoL_plead}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figy.png}
    \label{fig:QoL_plead}
  \end{subfigure}
  \caption{CDFs of predicted probability of pleading guilty at different burden levels
           determined through the All Else Equal Random Forest model}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/plea_cdf.png}
    \label{fig:non--QoL_plead_GLM}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/pleaq_cdf.png}
    \label{fig:QoL_plead_GLM}
  \end{subfigure}
  \caption{CDFs of predicted probability of pleading guilty at different burden levels
           determined through the Generalized Linear Model}
\end{figure}

\subsubsection{Propensity Score Matching --- Random Forest}
Finally, we carry out the Propensity Score analysis through a Random Forest model.
Appendix \ref{appendixA}
also includes the calibration information for the Random Forest models to predict ROR for each type of case.
We observe that the ROR prediction models are accurate and unbiased.

Tables \ref{table:ROR_non--QoL_propensity_score} and
       \ref{table:ROR_QoL_propensity_score},
respectively, include the mean predicted probability of pleading for
  ROR and Non--ROR, for
  Non--Quality of Life and Quality of Life cases,
  respectively.
Figures~\ref{fig:ROR_non--QoL_propensity_score} and
        \ref{fig:ROR_QoL_propensity_score} show
the CDFs of same values determined through the model
(also see findings through GLM approach in
Figures~\ref{fig:ROR_non--QoL_GLM} and
        \ref{fig:ROR_QoL_GLM}).

Our results match the All Else Equal analysis.
Especially for Non--Quality of Life cases, a Non--ROR burden level increases the predicted probability of pleading.
Future work should repeat this analysis using
Jail and
Non--Jail burden levels,
as those levels are where the substantive difference occurs in the All Else Equal analysis.

\begin{table}
\centering
  \begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
    \hline
    \textbf{Burden Level} & \textbf{Mean Predicted Probability of Pleading} \\ \hline
    ROR & 0.3706164 \\ \hline
    Non--ROR & 0.412566 \\ \hline
  \end{tabular}
  \caption{Mean Probability of Pleading for Non--Quality of Life offenses through Random Forest Modeling}% Propensity Score Matching}
  \label{table:ROR_non--QoL_propensity_score}
\end{table}

\begin{table}
\centering
  \begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
    \hline
    \textbf{Burden Level} & \textbf{Mean Predicted Probability of Pleading} \\ \hline
    ROR & 0.6127946 \\ \hline
    Non--ROR & 0.6363191 \\ \hline
  \end{tabular}
  \caption{Mean Probability of Pleading for Quality of Life offenses through Random Forest Modeling}% Propensity Score Matching}
  \label{table:ROR_QoL_propensity_score}
\end{table}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figxx.png}
    \label{fig:ROR_non--QoL_propensity_score}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figyy.png}
    \label{fig:ROR_QoL_propensity_score}
  \end{subfigure}
  \caption{CDFs for the mean predicted probability of pleading for ROR and non--ROR through Random Forest Modeling}% Propensity Score Matching}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/ror_cdf.png}
    \label{fig:ROR_non--QoL_GLM}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/rorq_cdf.png}
    \label{fig:ROR_QoL_GLM}
  \end{subfigure}
  \caption{CDFs for the mean predicted probability of pleading for ROR and non--ROR through Generalized Linear Model}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figaa.png}
    \label{fig:mean_ROR_non--QoL}
  \end{subfigure}
  ~
  \begin{subfigure}{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/figbb.png}
    \label{fig:mean_ROR_QoL}
  \end{subfigure}
  \caption{mean probability of ROR for each judge, vs.
  actual percentage of ROR modeled by Random Forest}

  \begin{subfigure}{0.49\textwidth}
    \caption{Non--Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/ror_scatter.png}
    \label{fig:mean_ROR_non--QoL_GLM}
  \end{subfigure}
  ~
  \begin{subfigure}{0.49\textwidth}
    \caption{Quality of Life cases}
    \includegraphics[width=\textwidth]{figures/glmplots/rorq_scatter.png}
    \label{fig:mean_ROR_QoL_GLM}
  \end{subfigure}
  \caption{mean probability of ROR for each judge, vs.
  actual percentage of ROR modeled by GLM}
\end{figure}

\subsection{Analyzing the Effect of Judge on Bail}
Figures~\ref{fig:mean_ROR_non--QoL} and
        \ref{fig:mean_ROR_QoL} show
        the mean predicted probability of ROR
        for each judge versus the judge's actual percentage of ROR,
        for Non--Quality of Life and Quality of Life cases,
respectively.
The plots show that some of the variation in
the judges' percentage of cases that are granted ROR is due to
non--randomness in the cases that they adjudicate.
However,
we observe that for a given mean predicted probability of ROR (on the Y axis),
there is significant judge variance on the true percentage of cases that are granted ROR.
Thus,
this initial analysis shows that different judges do utilize ROR differently for the same types of cases.
As before,
figures~\ref{fig:mean_ROR_non--QoL_GLM} and
        \ref{fig:mean_ROR_QoL_GLM} illustrate similar patterns using a different modeling approach (GLM).

### 2016-05-16 cscw/papers/2017/labor/trackingWork.tex
init!

\documentclass{sigchi}

\usepackage{todonotes,txfonts,balance,graphics,color}
\usepackage{booktabs,textcomp,microtype,ccicons}
% \usepackage{babel}
% \usepackage{csquotes}
\usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
\usepackage{enumitem}
\usepackage[T1]{fontenc}
\usepackage[super]{nth}
\usepackage[pdftex,pdfpagelabels=false]{hyperref}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}

\def\plaintitle{Naming Things is Hard: Real Title Following Colon}
\def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
\def\emptyauthor{}
\def\plainkeywords{Please don't make me pick keywords.
This is like asking a teacher to give the bullet points of
what a student missed in lecture.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}
\bibliography{../../../../references}

% End of preamble. Here it comes the document.
\begin{document}
\balance{}
\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
}

\maketitle

\begin{abstract}
  With growing attention on gig work
  --- ranging from the ``sharing economy'' to microtasks ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  --- driven in part by the discretization,
  routenization,
  and external management of said work ---
  not only suggests, but in fact generates what we have observed:
  members of this transient workforce increasingly feeling
  disempowered,
  marginalized,
  and frustrated by the systems and platforms on which they work.

  After evaluating this framing
  through a series of case studies
  in various industries falling broadly under the ``gig work'' category,
  we turn our theoretical lens to look to the future,
  to identify worthwhile questions and
  points of inquiry that researchers in social computing should consider
  as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section{Introduction}
This section should outline the trend of contemporary gig work in the last 10 years,
in particular tracking the changes since
\citeauthor{crowdworkFuture} discussed
\citetitle{crowdworkFuture}
in \citeyear{crowdworkFuture}
\cite{crowdworkFuture}.
This would review the literature since then,
then point to the movements toward worker unionization, advocacy, etc\dots
that have happened in the last year or two.

In the reflection on the literature published in the last 5 years since
\citetitle{crowdworkFuture},
we notice a broader change in the form of work that's being done;
motivated by this realization,
we look at the emergent phenomena in piecework both as interconnected.
By looking at the movements toward
discretization,
routenization,
and management
as necessarily linked --- even causally --- to one another, 
we give ourselves a framing of contemporary piecework that explains
and arguably predicts
what we have seen thus far.

\subsection{Piecework as a lens to understand gig work}
It's not new that researchers have discussed gig work as an extension or resurgence
of piecework, but it is sufficiently important for the framing of this argument that
it merits a more direct treatment and some explanation here.
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
a term almost lost to history, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking
\cite{scott1975women}.
Given the scope, however,
we will have to leave this complex topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\begin{inlinelist}
\item the work is done in the home;
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}

% The work involved often meant sewing together pieces of denim,
% was largely done by women, and turned

Crowd work,
they argued in \citeyear{crowdworkFuture},
represents a threat to the future of work inasmuch as it marginalized
and perhaps even harmed
workers;
their driving question, then, was
whether the future of
(crowd, but arguably all discretized)
work might be one in which they wish their children to participate someday;
it seems strikingly like one that might have been asked
as patches of denim were first being delivered to the homes of early pieceworkers.

This paper will argue that the similarity is more than striking;
it was an inevitable milestone in the maturation of the type of work that is reemerging today.
\textit{How to continue???}

% we engaged in ethnographic fieldwork
% to attempt to understand why gig work appeals to those that engage in it,
% and to gain a better understanding of how they relate to the marketplaces in which they work.

% \textit{I'd also like to bring up piecework here,
% and briefly make the case that
% the kind of work we're seeing should be thought of as a continuation
% or resurgence
% of piecework, with some important distinctions that differentiate it.}

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
starting with the discretization of tasks,
followed by routenization and the rise of workflows,
and finally the external management of workers.
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.

We explore each topic
--- discretization, routenization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The past decade of research has shed light on crowd work from various perspectives,
and revealed 
\subsection{Discretization}
\textit{I want to argue that modern gig work is discretized,
and that this parallels some of the piecework that we're historically familiar with}
Work has become discretized in the last 10 years, particularly in the information work sector.
Breakthroughs like ``Find--fix--verify'', ``Cascade'', etc\dots
make it easier to assign smaller components of work to myriad workers and
recompose the constituent parts into something more complex
\cite{bernstein2015soylent,chilton2013cascade}.
Some work has even turned to ``break it down'',
inasmuch as decomposing what we might call ``macro'' tasks into microtasks
\cite{cheng2015break}
(\textit{there's a lot more to talk critique from this paper; sorry Justin.}).

The parallels between this kind of work and
historical piecework of the turn of the \nth{20} century
is by no means new
[\textit{citation needed --- who said this first? most notably? Turner?}].
\textit{Elaborate on gig work, piecework, etc\dots}

\subsection{Routenization}
The \textbf{discretization} and decomposition of work certainly facilitates
breaking tasks down,
parallelizing work,
and getting more done across a broader array of people;
but the key advantage in turning macro tasks 
into many micro tasks is that we can make that work \textit{routine}.

By making larger bodies of work less contextually situated,
we can define processes that make that work
and the instructions thereof
useful for virtually anyone with shared cultural or intellectual background.

Routenization can be seen in the contemporary in workflows
[\textit{There are various papers that talk about creating a workflow;
maybe this is where FFV belongs? Foundry?}].

\textit{Harken back to piecework
and the instructions that denim workers had to make clothing;
the instructions were clear and scoped enough that you could repeat the task ad infinitum.}

\subsection{External Management}
\textbf{Discretization}, enabling \textbf{routenization}, thus allows us to
evaluate work in abstraction.
This means that we can

\textit{I want to cite Justin's work on crowdsourcing effort to highlight that
we can evaluate workers for the work they do with increasingly finite amounts of time;
also, Ranjay's work that embraces failure because
we can essentially algorithmically make sense of failure (let's call it noise)
and re--capture signal.}

\subsection{Resistance}
So it's not surprising, then, that workers are frustrated:
Turkopticon, Dynamo, and many other pieces of work point this out
\cite{turkopticon,dynamo,uberAlgorithm}
[\textit{also want to cite Brian's work from CHI this year}].
In popular culture, reporters have inquired about
what ``the future of labor unions'' will look like
[\textit{Imagine I had cited an online article and
it wasn't totally inappropriate for a conference}].

\textit{The parallels here would be difficult to ignore;
Lily and others have pointed this out,
but it's worth thinking about the broader trend of
resisting management,
driven by routenization,
enabled by discretization.
Do that here.}

\subsection{Looking forward}
If we agree that
discretization,
routenization,
management,
and even resistance
necessarily follow one another according to this theoretical lens,
then we have to use it to attempt to envision
what comes next.

\textit{I've been going back and forth regarding how to frame this paper;
the approach that tries to look at
things that are similar vs things that are different
doesn't seem to work, but
I've left the thoughts here because it's not all completely bad.}

\section{Things stay the same}
How is gig work the same as it's been historically?

I think this section would be compelling to draw parallels between
the narratives drivers gave about the flexibility, autonomy, etc\dots
and that which we might have seen among pieceworkers
(predominantly women, who benefited from being able to work from home).

\subsection{Flexibility}
Are there cultural differences between the people that did piecework and
the people that do gig work now?
I'm not sure there are significant differences that have affected the outcomes so far.

Many of the workers to whom piecework appealed were mothers, wives, etc\dots
who mostly stayed at home for various reasons
(certainly largely it was cultural
--- women weren't afforded equal access to labor opportunities,
making in--home job opportunities not only compelling, but also one of few available options).

Gig workers are in some senses similarly constrained:
workers on Amazon Mechanical Turk
--- those that use it as a primary source of income, at least ---
report being homebound for various reasons (e.g. medical, parenting, etc\dots).
Society and circumstance have made it difficult
or impossible
to join the contemporary, conventional workforce;
gig work re--opens that door.

But there are differences; during our research over the summer,
we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
Many of them told us about their home lives
--- about children, spouses, and other commitments ---
to which they wanted to dedicate more time.
One driver
(let's call him Ra\'{u}l)
told me about how he drove for Lyft
after working as an inventory manager at a hospital for more than nine years.

I asked him why he quit that job and forewent
the benefits,
predictability,
and career growth opportunity
that his old job offered.
He told me that when his daughter was born,
he was overcome with a desire to spend more time with her.
No longer satisfied with work
where he often left before his daughter would wake up and return after she fell asleep,
Ra\'{u}l decided to start driving for Lyft,
because he could drive in the evenings when his daughter was asleep.

Other drivers reported similar benefits;
gig work affords its workers flexibility that conventional careers don't allow.

\subsection{Routenization of work}
This leads into the next section, but I want to bring up the process of
making work about mass--manufacturing, at least inasmuch as
the instructions are the same for everyone.

The Internet arguably has made it much easier to broadcast those instructions,
but it's had this deeper effect of enabling some amount of back--and--forth
between the worker and the (algorithmic) manager.

\subsection{Taylorism}
The routenization of work makes it possible to measure that process,
optimize it for certain characteristics, and ultimately
lead to Taylorism and scientific management.
This is not new;
researchers have studied and written about the slow creep of
algorithmic management and 
discretization \& routenization of work tasks.

\textit{We hope to take a step back from the context in which this work is often applied, and
look for its place in the larger trends and theories
to make sense of the trends of gig work at large.}

Industrialization and the automobile assembly line makes this famous,
but piecework functioned on the principle that everyone was making similar
or identical
garments and other products.

Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
sometimes bootstrapped,
as in Ranjay's talk on 
``Embracing Error to Enable Rapid Crowdsourcing'',
but more conventionally in work flows like ``Find--Fix--Verify''.

\textit{This might be an opportunity to reflect on
how pieceworkers internalized the work they were doing,
responded to the stress of the uncertainty of potentially rejected work,
etc\dots but I'm not familiar with research in that space.}

\subsection{The emergence of decentralized workplaces}
The practice of in--home piecework was consumed by
the centralization of factories
(the effect of which we'll talk about later, since
we can talk about how this made unions more practical),
but for a time many pieceworkers
at the turn of the \nth{20} century
worked out of their homes.
Strikingly, many of the gig workers
\citeauthor{crowdworkFuture}
discussed in
\citeyear{crowdworkFuture},
and indeed many more continue to do what we might call ``information work''
--- that is, work that predominantly demands human computation ---
but increasingly we're seeing the movement toward transient work that largely requires embodied presence
\cite{uberAlgorithm} (and others).

\subsection{What was this about?}
I was going to check in quickly and take a photo of your whiteboard
(hoping that you'd kept our conversation notes around)
but your room was \textit{literally} full of people lol.

\section{Things are changing}
The medium on which this work is being done
--- and to an extent the medium used to manage workers ---
has dramatically changed things as well, however;
workers are distributed around the world,
working out of their cars in the cases of livery services
(notably, never returning to a base of operations)
across and between cities as well as nations,
or in their homes
(paralleling the trend of piecework even more closely).

Trying to understand how gig work has differed from piecework should at least
start with looking at the different characteristics of the work involved.
After that, we should think about how
the demography and culture of the people engaging in this kind of work
have changed versus that of the pieceworkers.

\subsection{Differences in the work itself}
Gig work has all of the above similarities with piecework,
but there are key differences.

Piecework emerged at the turn of the \nth{20} century,
right at the time that telecommunications began to boom in the United States. % too far a reach?
\textit{Did telecom enable remote management?}

Gig work in its contemporary formation is largely mediated by
ubiquitously accessible digital media
(the Internet, telephony, etc\dots)
and importantly has relied on this technology to facilitate the remote management of workers
\cite{uberAlgorithm}.

\textit{Has the work fundamentally changed,
or are we just being managed remotely in different ways?
I'm not entirely sure.}

How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
How does this work differ from the work in which truckers, taxi drivers, and other
independent contractors have been participating for decades?
We argue that the substantive difference in these markets is the speed of the market itself,
motivated by the technology which mediates it.
Because workers can be sourced and dispatched virtually instantly,
businesses that engage in this kind of work
(e.g. Uber, Amazon Mechanical Turk, etc\dots)
have taken to removing other bottlenecks,
like vetting workers upfront

\section{More deliberate work}
% \subsection{Differences in the people that do the work}
% ???

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-05-16 cscw/papers/2017/theoryPaper/designGuidance/vulnerableWorkers.tex
init!

%!TEX root = ../proceedings.tex
% \subsection{Vulnerable workers, and how to protect them}
\textit{Many workers in the gig economy are vulnerable to exploitation;
careless or malicious systems can endanger workers.}

At organization E, workers expressed concern when we discussed the ability to show a profile picture to customers in advance of their arrival.
These concerns stemmed from fears of their privacy being violated,
and discrimination over their race, age, gender, and other characteristics.
In particular, workers at organization E were afraid that customers would reject them for being too old.

organization E addressed this issue by emphasizing the qualifications of workers,
abstracting their names and details for first-time jobs.
Workers fundamentally have final say over whether to be matched to a customer,
based on general data about the location,
an indication of whether they've worked together before, and a rough estimate of the amount of work solicited
(In most cases, a precise estimate of how long a job will take is difficult to make without someone familiar with the work on-site to make an informed estimate).

If a customer has hired through organization E before and wants to hire a previously hired worker,
a rudimentary system allows volunteer dispatch staff to identify those workers for matching.
This allows workers to gradually develop a body of clients over time;
eventually, they no longer rely on organization E to refer customers to them at all.
Organization A utilized a similar approach, allowing customers to request a worker they've worked with in the past,
but otherwise not revealing information about workers to customers.

Many of the existing marketplaces for gig work don't expose such affordances to request a previously hired worker.
% making it difficult or impossible to grow a regular client base without leaving the system entirely.
This may be deliberate, as it effectively prevents workers from fostering a community of clients.
By preventing workers and customers from becoming familiar,
workers become dependent on the market created by the platform to provide leads on customers.
Whether a system promotes such dependence or not is up to the system's designers,
but we articulate a possible approach which makes repeat work possible.

As we illustrate in Figure ~\ref{fig:previousHire},
a simple interface exposing basic information about a previously hired worker can remind a customer whom they want to select.
Further, a technological solution such as this can ease the negotiation of scheduling;
in our case, we suggest prompting the worker for their availability.
Once several times are suggested, the customer can select and confirm a time.

\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/recurrent}
    \caption{By exposing some information about previously matched workers,
    a system can facilitate workers developing a reliable customer base.}~\label{fig:previousHire}
\end{figure}

Working with organization B, nurses we spoke to expressed concern over disrespectful or uncooperative patients.
Nurses described various safety concerns including
violent neighborhoods, dangerous pets such as dogs, and patients in declining mental health.
Nurses told us that these were inherent risks associated with in-home care.
Some of the nurses that we spoke to clarified that, in cases where they felt a particular danger,
they may decline to meet a patient and communicate that concern later.

\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/startwork}
    \caption{A simple interface for workers,
    allowing them to contact customers,
    facilitates the resolution of potential issues before work begins,
    such as difficulty finding a location, ambiguity in instructions, etc\dots}~\label{fig:startwork}
\end{figure}

Technological systems can assist workers in cases like these;
by exposing customer contact information to workers (and potentially vice versa),
workers can contact customers before they begin work to clarify or resolve any issues.

### 2016-05-16 proposals/CyberProposal/paper.tex
init!

\documentclass[11pt]{article}
\usepackage{balance,graphics,times,setspace}
\usepackage{hyperref,graphicx,sidecap}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
\newcommand{\sectitle}[1]{\textit{#1}}

\begin{document}

\section*{Collective Action and Governance in an Online Piecework Economy}

\begin{itemize}[leftmargin=0em,label={}]
  \item Michael Bernstein, Computer Science, Stanford University
  \item Margaret Levi, Political Science and CASBS, Stanford University
\end{itemize}

\section*{Abstract:} 
The digital gig economy has led to a resurgence of piecework.
Without shared factories and water coolers,
how do digital pieceworkers coordinate,
build solidarity,
and take collective action?
We will engage in fieldwork with digital pieceworkers who work in data entry,
domestic services,
and on--demand driving to understand how they counter algorithmic systems and engage in collective behavior.
We will then design and launch a new collectively--governed platform for domestic workers,
in collaboration with Fair Care Labs at the NDWA
(National Domestic Workers Alliance).
Our goal is to both understand the factors that drive collective behavior in supposedly--distributed gig platforms,
and to introduce new designs and infrastructure that enable the growth of worker collective action in the digital piecework
(gig)
economy.
Our results shed light on the strengths and weaknesses of  digital piecework cooperativism,
as well as the policies necessary for it to succeed.

\section*{Relationship to previously funded work:}
PIs Bernstein,
Levi,
Johari and Valentine have an ongoing funded Cyber project titled ``Cyber Work: The Future of Networked Labor".
This project outlined major goals:
\begin{inlinelist}
\item ethnographic studies of online work practice,
\item the creation of crowd organizations,
\item collective action,
and
\item new labor platforms
\end{inlinelist}.
The project was scoped too ambitiously for the funding available last year,
requesting \$1.2 million over three years.
Thus, we have split out part of
(1) and all of
(3) into this separate proposal.
(2) and (4) are making rapid progress under the existing funding,
and currently have significant publications in preparation and under review,
as well as an alpha version of the platform launched.
If the committee wishes to fund this proposal,
we are comfortable with either pitching it as a separate proposal or folding it into the existing Cyber Work proposal as part of a renewal request.
The timing of the funding start date is flexible.

\begin{itemize}[leftmargin=0em,label={}]
\item Proposed budget: \$100,000 per year
\item Collaborators: Ali Alkhatib, Niloufar Salehi
\end{itemize}
\thispagestyle{empty}
\pagebreak
\clearpage
\setcounter{page}{1}

Piecework has historically disempowered workers and empowered employers.
Subdividing work and paying per task drives employees to work faster and longer, affording them little control over working conditions.
For most of its history, piecework offered no benefits, occupational health and safety, or social insurance.
When piecework combined with job routinization, workers became interchangeable and deindividuated.
This combination of piecework and routinization produced the dominant payment system during the industrial era,
starting with agriculture and home production but quickly moving into factories.
However, by the turn of the 20th century in the United States, campaigns for workers' rights yielded regulations on conditions, and by the late 20th century, piecework was associated primarily with migrant labor and sweatshops.
This contracted piecework economy has remained relatively stable since.

Today, with the growth of the digital gig economy, piecework is reemerging in new forms.
Networked computational infrastructure is now mediating each unit of work and worker.
Uber's taxi drivers are paid per ride and assigned jobs by an algorithm
\cite{uberAlgorithm,hall2015analysis}.
Information workers, such as those on Amazon Mechanical Turk and Upwork, are paid per task, performing vast quantities of data and administrative work
\cite{martin2014being}.
% [Martin et al., 2014].
Increasingly numerous workers make money through one--off tasks like housecleaning and food delivery, all mediated by online platforms.
Current legal and policy frameworks classify these workers as independent contractors, precluding them from accessing employee benefits and job security associated with employment.
Complicating matters, today's workers often despise unfair employers but prize the flexibility and autonomy afforded by their platforms
\cite{martin2014being}.
% [Martin et al., 2014].
Many view themselves as masters of their own fate, free to enter and leave the workforce at will and without repercussion.
Despite this, they still seek collective representation, as Uber drivers have done repeatedly to combat falling rates.

How do workers react to, and create collective counterbalances to, the digital piecework economy?
On the one hand, studies of online communities predict that workers online will face greater challenges, because trust is more difficult to build in digitally mediated environments
\cite{successfulOnlineCommunities,kollock2005managing,cook2005cooperation}.
% [Kraut et al., 2012, Kollock and Smith, 2005, Cook et al., 2005].
Without factory floors, water coolers, and shared neighborhoods
\cite{waterCooler},
% [Small 2015],
how do workers coordinate and build solidarity?
On the other hand,
online forums may make it easier for workers to share information and strategies,
and successful collaborative efforts such as Wikipedia suggest that the online environment may support forms of cooperation that could not have succeeded as readily offline.

As yet,
we have limited knowledge of how online pieceworkers could adapt to contemporary labor markets.
We do not know how workers are engaging in coordinated and cooperative actions,
nor what form they should take in an era of digital work.
Are their efforts similar to historical behaviors for offline pieceworkers,
such as unions and worker cooperatives?
Or do the affordances of the online world result in different outcomes?
Furthermore,
could technological affordances grant pieceworkers the ability to organize more effectively,
and even to run their own marketplaces?

We propose paired grounded ethnographic fieldwork and system development to 1) understand the forces shaping informal and formal worker reactions to the online piecework economy,
and 2) create a technical and design foundation for collective action,
ownership and governance in online piecework.

First, we will analyze worker behavior in the current sociotechnical ecosystem through a year of fieldwork on major Mechanical Turk
(data entry)
platforms and forums,
and on gig economy platforms such as Uber
(driving)
and Handy
(domestic work).
Through this fieldwork, we will investigate the techniques that workers use to manage their work environments and job opportunities.
We will seek to understand the challenges they face when balancing an identity of being self--employed with an identity of being a pieceworker for a privately--owned labor platform, and the techniques that workers and employers use to circumvent each other's restrictions, barriers, and behaviors.

Second, we will create and launch a cooperatively owned ``gig" marketplace for domestic workers in collaboration with the National Domestic Workers Alliance
(NDWA).
Our own prior work established a collective action platform for crowd workers
\cite{dynamo}.
We now seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
The NDWA's Fair Care Labs are providing access to interested domestic workers
(house cleaners), and we are designing and implementing the platform with them as a pilot group.
The goal is to understand how to design for a set of distributed workers to achieve consensus on issues such as management, service fees, and arbitration.

Our team is uniquely qualified to pursue this opportunity.
Levi has years of experience studying labor, collectives, and unions.
She is pairing with Bernstein, a computer scientist who studies online labor platforms and collective action.
Bernstein, alongside PhD students Salehi and Alkhatib, created the first platform for collective action efforts amongst digital pieceworkers, focused on the Amazon Mechanical Turk platform.
This effort has galvanized hundreds of workers and resulted in outcomes including public media campaigns and the creation of ethical work guidelines for Mechanical Turk employers.

This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.
What affordances and protections do workers need in order to make their voices heard and to have collective power?
What protections should platforms be required to provide to their workers?
What support is necessary for people who work on multiple platforms, frequently moving between different employers?
Are unions the answer, or some other form of worker organization?

\section{Ethnographic analysis of current worker coping and counter--strategies}
The first study will consist of ethnographic fieldwork studying workers in modern piecework labor markets 
--- e.g., Mechanical Turk, Uber, and Handy ---
to understand how social coping strategies emerge among workers.
How do workers air their grievances?
How do they come together to push for change?
How have they adapted to the working conditions of these digital hiring halls?

Lee et al.
and others have identified a number of ways that workers subvert the intent of system--designers
\cite{uberAlgorithm}.
These patterns of behavior elude algorithmic tracking and measurement because they deliberately avoid the a priori assumptions made by the designers of systems,
who attempt to structure these sites of work in ways to incentivize preferred behavior.
Identifying and understanding the details of this behavior thus begins with a qualitative,
ethnographic endeavor.

We will perform one year of fieldwork with digital pieceworkers.
To understand their breadth of experience,
we will sample roughly fifty workers from information work
(for example,
Mechanical Turk and CrowdFlower),
and another fifty workers in the ``sharing economy''
(including Uber,
Lyft,
and AirBnB),
and fifty in gig work
(namely Handy and TaskRabbit).
Given existing relationships and reputations as credible researchers in communities of Turkers,
we can recruit participants through known existing channels.
In the cases of platforms such as Uber,
AirBnB,
and Handy,
we can recruit workers for nominally legitimate work
(for example,
requesting that a driver drives for 15 minutes to a previously selected destination)
and invite them to participate in our interview.

Through a combination of semi--structured interviews,
surveys and grounded analysis of worker forums,
we will seek to better understand workers' relationships with the externally--managed platforms,
with their clients,
and with each other.
Our questions will include contextualizing questions asking how long the worker has worked in that platform and what sort of work they engaged in before they worked on that platform.
Excluding questions and topics of discussion that emerge from these prompts,
we will ask workers what parts of their platforms and styles of work they enjoy and what parts they find frustrating,
which will segue into further questions about whether and how people work around the system when it fails to do what they wish.

Further,
we will investigate how workers rely on each other despite being distributed and never meeting face--to--face,
and how they collectively act to counter the directives and desires of their employers.
We seek to uncover how norms emerge, how workers discover one another, and how techniques for subverting systems are discovered and communicated across these geographically dispersed, and often seemingly disconnected, communities of workers.

These insights will inform the broader question of how collective action among digitally mediated communities works, and in particular expose how members of networked communities 
--- typically considered too decentralized and diffuse to allow such phenomena to occur ---
successfully engage in collective action.
We will identify workers' behaviors that expose and capitalize on flaws' in employers' algorithms and policies 
--- for example, how Lyft drivers will rapidly toggle their availability on and off in order to contravene the Lyft availability algorithm, which would otherwise automatically grow the geographic radius in which a driver must pick up a ride if they have not found a ride recently.

\section{Designing future platforms}

Armed with a better understanding of workers' needs and collective behaviors, we next seek to create a platform that can become a template for online work that better represents workers' needs.
However, solutions that exist for traditional, non-gig workers
--- especially labor unions ---
face resistance from workers in the gig economy
\cite{martin2014being,dynamo}.
% [Martin et al., 2014; Salehi et al., 2015].
This distaste may be due to a general decline in public opinion of unions.
However, the traditional mechanics of a labor union are also a poor match: challenging issues include managing the benefits and proportional representation of workers whose contracts may last only minutes or hours, enforcing collective decisions on a decentralized network of workers, and establishing trust between workers who may never meet face--to--face.
In our early fieldwork, for example, a worker on Amazon Mechanical Turk stated:
``If by `union' you mean a `labor union', I would not feel comfortable taking part.
It runs against my grain because I am an individualist.
[\dots] I consider myself self--employed\dots\  not working for anyone in particular.''

Given these constraints, what alternative model would give workers more power in an online gig--economy platform while still enabling a market to emerge and function? To explore one such alternative, we will collaborate with the National Domestic Workers Alliance's innovation arm, the Fair Care Labs, to create tools for online pieceworkers to collectively manage and govern their own labor markets.
While the literature in social computing design and crowdsourcing already features many tools for successful decentralized collaborations (e.g., Wikipedia), and collective action movements (e.g., change.org), templates and guidelines for sustained collective action and governance online in high--stakes situations that affect workers' livelihoods are rare.
Such efforts tend to swell quickly, but soon find themselves embroiled as they struggle to satisfy myriad stakeholders .
Our goal is to introduce design mechanisms to enable collectives of gig workers to form, debate policy, and implement their decisions over a long period of time.

Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction. The web and mobile application, Alia, enables the basic affordances of ``gig'' work for domestic work
--- rapid scheduling of any available worker in the network (Figure 1).

% Armed with a better understanding of workers' needs and collective behaviors, we next seek to create a platform that can become a template for online work that better represents workers' needs.
% However, solutions that exist for traditional, non--gig workers 
% --- especially labor unions 
% --- face resistance from workers in the gig economy
% \cite{martin2014being,dynamo}.
% This distaste may be due to a general decline in public opinion of unions.
% However, the traditional mechanics of a labor union also are a poor match: challenging issues include managing the benefits and proportional representation of workers whose contracts may last only minutes or hours, enforcing collective decisions on a decentralized network of workers, and establishing trust between workers who may never meet face--to--face.
% In our early fieldwork, for example, a worker on Amazon Mechanical Turk stated:
% ``If by `union' you mean a `labor union', I would not feel comfortable taking part.
% It runs against my grain because I am an individualist.
% [\dots] I consider myself self--employed\dots\space not working for anyone in particular.''

% Given these constraints, what alternative model would give workers more power in an online gig--economy platform while still enabling a market to emerge and function?
% To explore one such alternative, we will collaborate with the National Domestic Workers Alliance's innovation arm, the Fair Care Labs, to create tools for online pieceworkers to collectively manage and govern their own labor markets.
% While the literature in social computing design and crowdsourcing already features many tools for successful decentralized collaborations
% (e.g., Wikipedia), and collective action movements
% (e.g., change.org), sustained collective action and governance online is rare.
% Such efforts tend to swell quickly and then fizzle in interest.
% Our goal is to introduce design mechanisms to enable collectives of gig workers to form, debate policy, and implement their decisions over a long period of time.

% Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers
% (house cleaners).
% This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
% However, our interest
% (and Fair Care Labs')
% is to generalize its insights to other workers and industries.
% While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
% In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction.
% The web and mobile application, Alia, enables the basic affordances of ``gig" work for domestic work 
% --- rapid scheduling of any available worker in the network
% (Figure 1)

\begin{SCfigure}
  \caption{\textit{Alia, our prototype cooperative platform for collective governance in a gig marketplace.
  Alia is targeted as a marketplace for domestic workers to do rapid online bookings.
  Its main research goal is to examine affordances for collective decisionmaking and governance in a cooperative marketplace.}}
    \includegraphics[width=0.24\textwidth]{figures/1.png}
    \includegraphics[width=0.24\textwidth]{figures/2.png}
    \includegraphics[width=0.24\textwidth]{figures/3.png}
\end{SCfigure}
% Figure 1.
% Alia, our prototype cooperative platform for collective governance in a gig marketplace.
% Alia is targeted as a marketplace for domestic workers to do rapid online bookings.
% Its main research goal is to examine affordances for collective decisionmaking and governance in a cooperative marketplace.

The research question that this system asks is: how can distributed, loosely--affiliated workers make effective decisions to govern a cooperative marketplace?
Traditional co--ops operate through scheduled elections to a board and maintain in--group unity via regular or mandatory meetings.
However, worker cooperatives at national and international scales could become unwieldy to govern; empirically, worker cooperatives become proportionally rarer as one looks to larger and larger companies, perhaps in part because the gubernatorial challenges either prevent companies from growing to
(or surviving at)
that scale.

Instead of attempting to scale a single cooperative instance, we propose mechanisms through which local cohorts of workers can `fork' a cooperative's guidelines to create an editable copy, then launch their own co--op.
This new co--op would still operate within the global Alia infrastructure, allowing the proliferation of many local co--ops that benefit from the single central infrastructure.
Examples of projects like this one already exist 
--- Wikipedia's source code, maintained by the Wikimedia Foundation, is available for others to fork and run independently.
Indeed, instantiations of Wikipedia operating under different administrative policies exist: Conservapedia, the Wikia platform, and others illustrate the potential to fork this originating source code and run parallel communities.

We believe that these local groups will be more competent at making decisions rapidly and responsively, while still benefiting from both the network effects of a larger cooperative organization as well as the focus of engineering effort on a single codebase.
If successful, online cooperatives might then be able to scale in ways rarely seen in their offline counterparts.

As Alia begins attracting workers, we will again engage in semi--structured interviews of workers on the platform, surveys and grounded analysis of the platform's forums.
We seek to understand: what issues are most contentious for workers to agree on?

How distributed or centralized do they want the platform's governance to be?
When concerns arise similarly to the externally--owned marketplace, how does the cooperative market deal with them?
We hypothesize that workers will want a loose confederation.
Because workers hold strong self--images as freelancers, they may be wary of systems tightly regulating their behavior.

\section*{Logistics and timeline}
Alia will be open source on GitHub via the MIT license.
The platform will be hosted on Heroku, and data will be secured and privately backed up using Postgres.
The effort will proceed over two years.
In the first year, we will perform the qualitative research with workers in order to understand their platform policy and governance behaviors.
Simultaneously, we will design and implement the first version of the Alia platform.
In the second year, we will launch Alia, grow its use locally in the Bay Area, and perform fieldwork with Alia workers to understand their relationship with the cooperative.
We are thus requesting two years of funding for the pair of graduate students who are leading the project.
They, and the PIs, bring together years of experience in qualitative and historical fieldwork, design and system development.
We are also requesting one month of summer salary funding for each of the PIs to support their work on the project.
If the project is successful in its goals, we may seek a third year of funding to grow the platform nationwide and continue our instrumentation and research on it.

\section*{Policy implications and dissemination strategy}
Our policy and design goal is to understand the kinds of worker collectives that can succeed in an online piecework economy, and to ensure that workers have access to them.
Our research targets will be top--tier venues in human--computer interaction
(e.g., ACM CHI, ACM CSCW)
and political science and sociology
(e.g., APSR, ASR, AJS, AJPS, POP).
One main outlet for broader dissemination will be as part of the Future of Work project at CASBS 
--- which recently published a series of essays in Pacific Standard and placed op eds in important newspapers and online sources; the relevant editors remain keen for new and additional material, and various other magazines
(print and online), such as American Prospect and Boston Review are also interested in possible publication of articles from this project.
A second set of outlets will be through the NDWA's relationships with labor advocates and the press.

\bibliographystyle{apalike}
\bibliography{../references.bib}
\end{document}

### 2016-05-16 proposals/RSFproposal/paper.tex
init!

% !TEX program = xelatex
\documentclass[11pt]{article}
\usepackage{balance,graphics,fontspec,setspace,titling,fmtcount,graphicx}
\usepackage{natbib,sidecap,subcaption,hyperref,setspace}

\usepackage[page,toc]{appendix}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}

\setcitestyle{authoryear, open={(},close={)}}

% \linespread{2}

% \doublespace
\setmainfont[
Ligatures=TeX,
Extension=.otf,
UprightFont= *-regular,
BoldFont=*-bold,
ItalicFont=*-italic,
BoldItalicFont=*-bolditalic]{texgyretermes}
\setmonofont{Source Code Pro}

\setlength{\parskip}{.4em}
\setlength{\parindent}{0em}
\newcounter{countitems}
\newcounter{nextitemizecount}
\newcommand{\setupcountitems}{%
  \stepcounter{nextitemizecount}%
  \setcounter{countitems}{0}%
  \preto\item{\stepcounter{countitems}}%
}
\makeatletter
\newcommand{\computecountitems}{%
  \edef\@currentlabel{\number\c@countitems}%
  \label{countitems@\number\numexpr\value{nextitemizecount}-1\relax}%
}
\newcommand{\nextitemizecount}{%
  \getrefnumber{countitems@\number\c@nextitemizecount}%
}
\makeatother
\AtBeginEnvironment{inlinelist}{\setupcountitems}
\AtEndEnvironment{inlinelist}{\computecountitems}
\AtBeginEnvironment{enumerate}{\setupcountitems}
\AtEndEnvironment{enumerate}{\computecountitems}
\AtBeginEnvironment{itemize}{\setupcountitems}
\AtEndEnvironment{itemize}{\computecountitems}

\titlespacing*{\section}
{0pt}{0pt}{0pt}
\titleformat*{\subsection}{\it}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
\newcommand{\sectitle}[1]{\textit{#1}}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}

\title{Worker Experience and Collective Action in an Online Piecework Economy}
\author{
Michael Bernstein\\
Computer Science, Stanford University
\and
Margaret Levi\\
Political Science and CASBS, Stanford University
}
\begin{document}
\begin{titlingpage}
  \maketitle
\end{titlingpage}

\pagebreak
\begin{abstract}
The digital gig economy has led to a resurgence of piecework.
Without shared factories and water coolers,
how do digital pieceworkers coordinate,
build solidarity,
and take collective action?
We will engage in fieldwork to understand
how digital pieceworkers engage in collective behavior
to better their work conditions.
Ethnographic fieldwork will take place on worker forums,
and in a cooperatively-owned labor market.
Our goal is to understand the drivers and inhibitors of collective behavior in online work.
Results will impact the design and infrastructure of the growing gig economy,
and the policies necessary for collective action by workers to succeed online.
\end{abstract}

\section{Introduction}
Generally, piecework disempowers workers and empowers employers.
Piecework drives employees to work faster and longer,
affording them little control over their working conditions.
For most of its history, piecework offered no
  benefits,
  occupational health and safety,
  or social insurance.
When piecework combines with the routinization of jobs,
workers become interchangeable and deindividuated,
another form of disempowerment.
Piecework was the dominant form of payment system during the industrial era,
starting with agriculture and home production but quickly moving into factories.
At the turn of the 20th century in the U.S.,
campaigns for workers' rights had yielded regulations on conditions.
By the late 20th century, however, piecework was associated primarily with migrant labor and sweatshops.

Today, with the growth of the digital gig economy, piecework is reemerging in new forms.
Networked computational infrastructure is now mediating each unit of work and worker.
Uber's taxi drivers are paid per ride and assigned jobs by an algorithm
\citep{uberAlgorithm,hall2015analysis}.
Information workers
--- such as those on Amazon Mechanical Turk (AMT) and Upwork ---
are paid per task, performing vast quantities of data and administrative work
\citep{martin2014being}.
Increasing numbers of workers make money through one--off tasks like
  housecleaning
  and food delivery,
all mediated by online platforms.
Current legal and policy frameworks classify these workers as
\textit{independent contractors},
precluding them from accessing employee benefits and job security associated with employment.

Digital piecework reframes many of the historical dynamics.
In one sense, it is a return to piecework in the home,
where the worker controlled the hours worked.
Seeing few alternatives,
home pieceworkers found themselves deciding whether to work in dangerous conditions or to starve.
However, in another sense,
online pieceworkers make this scenario quite different,
prizing the flexibility and autonomy afforded by their platforms
\citep{martin2014being}.
Many view themselves as masters of their own fate, free to enter and leave the workforce at will and without repercussion.

These trends in labor are not new,
but have accelerated with the growth in digital mediation.
With relatively scant research on the magnitude of this digitally mediated,
transient labor force,
and lack of consensus regarding the exact boundaries describing this mode of labor,
estimating the true scale of the gig economy has proven difficult.
Nevertheless,
TIME's most recent poll of so--called
``sharing economy'' workers led them to conclude that
as many as forty five million Americans have offered goods or services through these marketplaces
\citep{sharingEconomyPoll}.
With nearly 20 percent of jobs in the United States that could be potentially sent
``down the wire''
\citep{blinder2006offshoring}
via services such as this,
the gig economy's rapid growth and large potential workforce necessitates a focused research and policy agenda.

Especially important,
we as yet have limited knowledge of how
collectives of online pieceworkers are adapting to contemporary labor markets.
Furthermore, we do not know how workers are engaging in coordinated and cooperative reactions.
Are their efforts similar to historical efforts for offline pieceworkers,
such as unions and worker cooperatives?
Or do the affordances of the online world result in different outcomes?

How do workers react, and create collective counterbalances,
to the digital piecework economy?
On one hand, studies of online communities predict that
workers online will face greater challenges because
trust is more difficult to build in digitally mediated environments
\citep{successfulOnlineCommunities,kollock2005managing,cook2005cooperation}.
Without
  factory floors,
  water coolers, and
  shared neighborhoods
\citep{waterCooler},
how do workers coordinate and build solidarity?
On the other hand,
online forums may make it easier for workers to
  share information and
  coordinate strategies with relative anonymity,
and successful collective efforts
such as Wikipedia
suggest that online environments may support collectives
that could not have succeeded as readily offline.

We propose grounded ethnographic fieldwork to understand the forces shaping informal and formal worker reactions to the online piecework economy.
We will perform a coupled analysis of worker behavior in
\begin{inlinelist}
  \item the current sociotechnical ecosystem and in 
  \item emerging platforms for worker cooperativism and collective action
\end{inlinelist}
First, through a year of fieldwork on
  major Mechanical Turk (data entry) platforms and forums, and
  on gig economy platforms such as Uber (driving) and
  Handy (domestic work),
we will investigate the techniques that workers utilize
to manage their work environments and job opportunities.
We will seek to understand the challenges they face when
balancing an identity of being self--employed with
          an identity of being a pieceworker for a privately--owned labor platform.
Second, we will engage with workers using
recently developed platforms for collective action on Mechanical Turk
for forming worker cooperatives in the gig economy
to understand how formal efforts succeed and fail.

Our team is uniquely qualified to pursue this opportunity.
Levi has years of experience studying
  labor,
  collectives, and
  unions.
She is pairing with Bernstein,
a computer scientist who studies online labor platforms and collective action.
Bernstein, alongside Ph.D.
students Salehi and Alkhatib,
created the first platform for collective action efforts amongst digital pieceworkers,
focused on the Amazon Mechanical Turk platform.
This effort has galvanized hundreds of workers and
resulted in outcomes including public media campaigns and
the creation of ethical work guidelines for Mechanical Turk employers.

This work carries policy implications for the emerging,
and as yet largely unregulated, gig economy.
What affordances and protections do workers need in order to unionize,
or gather, effectively?
What protections should platforms be required to provide to their workers?
What support is necessary for people who work on multiple platforms,
frequently moving between different employers?

\section{Related Work}

The form of work that we propose to study has ties to previously--studied communities,
but various features make the context and framing of this field site unique.
Myriad lenses can be used to appreciate the qualities of gig work; we will explore
\nextitemizecount{}
of them here:
\begin{inlinelist}
  \item Taylorism and the rise of the routinization of work, and
  \item piecework, and the parallels between historical piecework and its contemporary gig work.
\end{inlinelist}

The parallels between Taylorism and gig work on digitally mediated labor markets
--- especially markets for information work such as AMT ---
are striking, perhaps for unsurprising reasons.
Whereas these markets advertise the work as
``\textit{artificial} artificial intelligence'',
the reality, as
\citeauthor{turkopticon}
point out,
belies the true nature of the platform and dehumanizes the workers behind the scenes
\citep{turkopticon}.
This distancing from the human features of AMT promotes a framing of work that is
\begin{inlinelist}
  \item compartmentalizable (wherein tasks can be broken into smaller components),
  \item decontextualizable (describing the alienation of work about which
  \citeauthor{marx2012economic}
  writes
  \citep{marx2012economic}), and
  \item readily testable (that is, provably true or untrue).
\end{inlinelist}

The relationship between piecework and gig work is important to appreciate, both
to frame our understanding of emergent trends in labor
as well as
to inform our theories for worker empowerment later.
Through this lens, we find the work of
\citeauthor{riis2004other}
and others informative
\citep{riis2004other,gringeri1994getting,herzog1980hand}.
The study of piecework represents not only a historical insight,
but a global one as well;
\citeauthor{shaheed1983invisible}
describe the role of piecework among women in Lahore,
and
\citeauthor{hahn1996feminization}
explores the gendered nature of piecework in India
\citep{shaheed1983invisible,hahn1996feminization}.

Considering the empowerment of workers in broader contexts,
we find a substantial body of literature which informs our theoretical framing as well.
In framing this body of research,
we identify \nextitemizecount{} themes:
\begin{inlinelist}
  \item the collaborative nature of gig work, namely on Amazon Mechanical Turk (AMT),
  \item efforts toward collective action and the challenges that stymie these groups, and
  \item more historical perspectives on collective action and the challenges that problematize it.
\end{inlinelist}

To understand sustained collective action,
\citet{russell1982collective}
calls for an
``anthropological investigation of minute interrelationships''
and the institutions in which those relationships occur
\citep[see also
\citet{hirschman1970exit,ostrom1990governing,polletta2002freedom}
]{russell1982collective}.
% [Hardin,
% 1982,
% also see Hirschman [1970]; Ostrom [1990]; Polletta [2002]].
Digital ethnographic fieldwork has detailed how workers gather in virtual spaces
--- such as forums and chat rooms ---
to
  discuss their work,
  share best practices, and
  manage their relationship with employers
\citep{martin2014being}.
The crowd of independent,
dispersed workers is in fact a collaborative network
\citep{crowdcollab}.
While this research has established that
  workers share information and build social connections,
it is not clear how they may engage in more complicated forms of collective action.

Research on online collective action explores
the social dynamics involved when groups aggregate their efforts online to push a cause.
In a setting where people can cheaply communicate with like--minded individuals,
collectives can overcome many of the challenges that stymie offline groups
\citep{earl2011digitally}.
But they still face free rider problems
\citep{olsonlogic},
the choice of exit or voice
\citep{hirschman1970exit},
and newcomers perceptions of favoritism towards veterans
\citep{polletta2002freedom}.

For example,
digitally mediated communities
such as Wikipedia
confront accusations of elite ``gatekeepers'',
paralleling the perceptions of oligarchism within allegedly egalitarian and
participatory governance systems
\citep{keegan2010egalitarians}.
We can learn from the experiments attempting to make sites of peer production
like Wikipedia
more inclusive and transparent.
Further,
a wealth of literature identifies both
  worthwhile practices and
  pitfalls that might endanger similar organizations
\citep{weber2004success,mockus2002two}.

Work to date addresses
--- but only to a limited extent ---
the limits and possibilities of cooperation among online workers.
Our research enables us to go into far greater depth on these questions
than previous studies have been able to do.
The combination of field work,
comparisons of different kinds of jobs and work organization,
and experiments with platforms will allow us to identify more specific
  mechanisms,
  processes, and
  contexts
that facilitate (or block) varieties of collective action.

\section{Prior work by the PIs}

Michael Bernstein is an Assistant Professor at Stanford and a computer scientist who
  designs,
  engineers, and
  studies
online work.
He is the recipient of
  a Sloan Research Fellowship,
  NSF CAREER grant, and
  George M. Sprowls Award for best doctoral thesis in Computer Science at MIT.
His group's research has received Best Paper awards and nominations at premier venues in
human--computer interaction.
Recently, his group, led by Ph.D. students
  Niloufar Salehi and
  Ali Alkhatib,
created the We Are Dynamo platform,
which enabled workers on Amazon Mechanical Turk to take collective action.
Hundreds of workers have participated in the site,
leading to two major campaigns:
\begin{inlinelist}
  \item The creation of worker--authored ethical research guidelines
  for research on Amazon Mechanical Turk at \url{http://guidelines.wearedynamo.org};
  \item A public media campaign of open letters to
  Amazon CEO Jeff Bezos
  to advocate for workers' perspectives on labor issues with Mechanical Turk.
\end{inlinelist}
Subsequent to this effort, several requests
(e.g., reinstatement of some workers accounts)
were implemented by Amazon.

Margaret Levi is a comparative political economist,
whose work focuses on
  government and governance,
  cooperation and trust,
  citizenship, and
  labor.
From 1994--2004,
she was the co--director with
  Russell Hardin and
  Karen Cook
of the RSF project on trust,
and she remains co--general editor of the RSF Trust series.
She co--authored one book in the series and edited or co--edited several others.
Levi has also written three books that explore the relationships of
  workers with their
    governments,
    employers, and
    unions.
In particular,
she has investigated the conditions under which
\begin{inlinelist}
  \item workers overcome collective action problems to organize unions
    in the United States and
    around the world,
  \item gain recognition for their unions and
    collective bargaining rights, and
  \item develop an expanded community of fate
  that takes into account distant others and
  on whose behalf the workers are willing to take costly actions.
\end{inlinelist}

\section{Research Proposal}
This research calls for cross--cultural analysis,
first exploring the communities and support structures which emerge in piecework markets run by external managers,
and subsequently their counterpart organizations and structures which form in cooperatively--run piecework labor markets.
Our work will explore two fundamentally different types of labor markets: labor markets run by external managers
(e.g. Uber, Lyft, Amazon Mechanical Turk, etc\dots),
and labor markets run cooperatively
(i.e. owned and managed by the system's workers themselves).

\subsection{Investigating Current Coping Strategies}\label{ethnography}
The first study will consist of ethnographic fieldwork studying workers in modern piecework labor markets
--- e.g., Mechanical Turk, Uber, and Handy ---
to understand how social coping strategies emerge among workers.
How do workers air their grievances?
How do they come together to push for change?
How have they adapted to the working conditions of these digital hiring halls?

\citet{uberAlgorithm}
and others have identified a number of ways that workers
  navigate and even
  subvert the intent of system--designers.
These patterns of behavior elude
  algorithmic tracking and
  measurement
because they deliberately avoid the \textit{a priori} assumptions made by the designers of systems,
who attempt to structure these sites of work in ways to incentivize preferred behavior.
Identifying and understanding the details of this behavior thus begins with a
  qualitative,
  ethnographic
endeavor.

We will perform fieldwork with digital pieceworkers.
To understand their breadth of experience,
we will sample roughly fifty workers from information work
(for example, Mechanical Turk and CrowdFlower),
and another
fifty workers in the ``sharing economy''
(including
  Uber,
  Lyft, and
  AirBnB),
and fifty in gig work
(namely Handy and TaskRabbit).
Given
  existing relationships and
  reputations
as credible researchers in communities of \textit{Turkers},
we can recruit participants through known existing channels.
In the cases of platforms such as
  Uber,
  AirBnB, and
  Handy,
we can recruit workers for nominally legitimate work
(for example, requesting that a driver drives for 15 minutes to a previously selected destination)
and invite them to participate in our interview.

Through a combination of
  semi--structured interviews,
  surveys, and
  grounded analysis of worker forums,
we will seek to better understand workers' relationships
  with the externally--managed platforms,
  with their clients, and
  with each other.
Our questions will include
contextualizing questions asking
  how long the worker has worked in that platform and
  what sort of work they engaged in before they worked on that platform.
Excluding questions and topics of discussion that emerge from these prompts,
we will ask workers
  what parts of their platforms and styles of work they enjoy and
  what parts they find frustrating,
which will segue into further questions about
whether and how people work around the system when it fails to do what they wish.

Further,
we will investigate how workers rely on each other
despite being distributed and never meeting face--to--face,
and how they collectively act to counter the directives and desires of their employers.
We seek to uncover
  how norms emerge,
  how workers discover one another, and
  how techniques for subverting systems are
discovered and
communicated across these geographically dispersed,
and often seemingly disconnected,
communities of workers.

These insights will inform the broader question of
how collective action among digitally mediated communities works,
and in particular expose how members of networked communities
--- typically considered too decentralized and diffuse to allow such phenomena to occur ---
successfully engage in collective behavior.
These behaviors often work
to expose and
capitalize on unexpected flaws in employers' algorithms and policies
--- for example, how Uber drivers will rapidly toggle their availability on and off
in order to contravene the Uber availability algorithm,
which would otherwise automatically grow the geographic radius in which a driver must pick up a ride if they have not found a ride recently
\citep{uberAlgorithm}.

\subsection{Understanding Emerging Cooperativism}\label{Alia}
Armed with a better understanding of workers' needs and collective behaviors,
we next seek to study worker behavior in a platform that better represents workers' needs.
However, solutions that exist for traditional, non--gig workers
--- especially labor unions ---
face resistance from workers in the gig economy
\citep{martin2014being,dynamo}.
This distaste may be due to a general decline in public opinion of unions.
However,
the traditional mechanics of a labor union are also a poor match: challenging issues include managing the benefits and proportional representation of workers whose contracts may last only minutes or hours,
enforcing collective decisions on a decentralized network of workers,
and establishing trust between workers who may never meet face--to--face.
In our early fieldwork, for example,
a worker on Amazon Mechanical Turk stated:
``If by `union' you mean a `labor union',
I would not feel comfortable taking part.
It runs against my grain because I am an individualist.
[\dots] I consider myself self--employed\dots~not working for anyone in particular''
\citep{dynamo}.

Given these constraints,
we seek to understand:
what model would give workers more power in an online gig--economy platform
while still enabling a market to emerge and function?
To explore one such alternative,
we are collaborating with the National Domestic Workers Alliance's innovation arm,
the Fair Care Labs,
on tools for online pieceworkers to collectively manage and govern their own labor markets.
While the literature in
social computing design and
crowdsourcing
already features many tools for successful decentralized collaborations
(e.g., Wikipedia),
and collective action movements
(e.g., change.org),
templates and guidelines for
sustained collective action and
governance online in high--stakes situations
that affect workers' livelihoods are rare.
Such efforts tend to swell quickly,
but soon find themselves embroiled as they struggle to satisfy myriad stakeholders.
Our goal is to introduce design mechanisms
to enable collectives of gig workers to form,
debate policy,
and implement their decisions over a long period of time.

In collaboration with
the National Domestic Workers Alliance's innovation arm
--- the Fair Care Labs ---
we are designing and launching a cooperative gig labor market for
domestic workers (house cleaners).
This cooperative labor market, ``Alia'',
is designed for domestic workers in the California Bay Area.
However, our interest
--- and the Fair Care Labs' ---
is to generalize its insights to other workers and industries.
While the driving domain is already crowded with services such as Lyft and Uber,
domestic work is a more ideal area to launch a research prototype.
In addition,
the NDWA Fair Care Labs' support and network will help us gain rapid traction.
The web and mobile application,
Alia, enables the basic affordances of ``gig'' work for domestic work
--- rapid scheduling of any available worker in the network
(see figure~\ref{fig:Alia} on page~\pageref{fig:Alia}).

\begin{figure}
\centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/1.png}
        % \caption{A gull}
        \label{fig:sched}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/2.png}
        % \caption{A tiger}
        \label{fig:confirm}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/3.png}
        % \caption{A mouse}
        \label{fig:profile}
    \end{subfigure}
    \caption{Alia, our prototype cooperative platform for collective governance in a gig marketplace.
             Alia is targeted as a marketplace for domestic workers to do rapid online bookings.
             Its main research goal is to examine affordances for
             collective decision--making and
             governance in a cooperative marketplace.}\label{fig:Alia}
\end{figure}
% \begin{SCfigure}
% \centering
% \caption{}
% \includegraphics[width=0.7\textwidth]{figures/main.png}
% \end{SCfigure}

The research question that this system asks is:
  how can distributed,
  loosely--affiliated workers
  make effective decisions to govern a cooperative marketplace?
Traditional co--ops operate through scheduled elections to a board and maintain
in--group unity via regular or mandatory meetings.
However,
worker cooperatives at national and international scales could become unwieldy to govern;
empirically,
worker cooperatives become proportionally rarer as one looks to larger and larger companies,
perhaps in part because the governmental challenges either prevent companies from growing to
(or surviving at)
that scale.

Instead of attempting to scale a single cooperative instance,
we propose mechanisms through which local cohorts of workers can `fork' a co--op's guidelines
to create an editable copy, then launch their own co--op.
This new co--op would still operate within the global Alia infrastructure,
allowing the proliferation of many local co--ops that benefit from the single central infrastructure.
Examples of projects like this one already exist
--- Wikipedia’s source code,
maintained by the Wikimedia Foundation,
is available for others to fork and run independently.
Indeed,
instantiations of Wikipedia operating under different administrative policies exist:
Conservapedia,
the Wikia platform, and
others illustrate the potential to fork this originating source code and run parallel communities.

Our second study leverages our access to this cooperatively owned marketplace,
attempting to understand similar dynamics but
focused on those that arise when online pieceworkers
manage and
govern their own labor market.
We seek to understand:
what issues are most contentious for workers to agree on?
How distributed or centralized do they want the platform's governance to be?
When concerns arise similarly to the externally--owned marketplace,
how does the cooperative market deal with them?
We hypothesize that workers will want a loose confederation.
Because workers hold strong self--images as freelancers,
they may be wary of systems tightly regulating their behavior.

Given our privileged access to the platform,
we will engage in a systematic sampling of fifty workers for semi--structured interviews.
We will sample workers randomly from active users of the system,
stratifying at the level of local co--op to ensure that
we capture local variations even if the co--ops in the network are different sizes.
Separately, we will sample workers who were previously active on the platform but have dropped out.
This second sample will be similarly stratified by local co--op within the Alia system.
The first group
--- active co--op workers ---
will give us access to experiences from workers
who have invested time and effort into the future of distributed worker cooperativism.
However, such workers might be self--selecting to be predisposed to such a model.
Thus, we are separately sampling the second group
--- inactive co--op workers ---
to gain insight into those who decided the model did not sufficiently support their needs.

The semi--structured interviews will begin with
a critical incident analysis that asks the worker to recall
a time that their own efforts were supported by the cooperative's decisions.
Example incidents might include when the collective identity helped the worker acquire more work,
or when it helped them engage in arbitration with an unhappy client.
Second, we will seek critical incidents when the worker felt unsupported by the cooperative's decisions.
Example incidents in this case might include
decisions about taxing all cooperative transactions
in order to pay for administrative and server support for the co--op.
We will then ask workers to describe their level of engagement with
the cooperative's governance,
their levels of trust and familiarity with others in the co--op,
and and the kinds of policy issues that they wish the cooperative would engage with.

The interviews are one thrust of a three--pronged mixed methods analysis.
The second thrust is surveys given to platform users, based on the same sampling techniques.
One of the core challenges of gig work is the distributed nature of its workers,
and trust amongst distributed pseudonymous individuals can be low
\citep{rains2007impact}.
Thus, we will utilize survey instruments intended to study internet--specific social capital
\citep{ellison2007benefits}.
In particular, the survey instrument measures
\begin{inlinelist}
  \item bridging social capital,
  which is social capital accumulated between groups of weaker ties and associated with
    risk-taking,
    information diffusion, and
    participation in a group context; and
  \item bonding social capital,
  which is social capital accumulated between stronger ties and
  associated with favors exchanged between close relationships.
\end{inlinelist}
These measures of social capital will help us understand
the levels of trust in the cooperative generally
(bridging social capital) and
with specific individuals in the cooperative
(bonding social capital).

Next, we will engage in a grounded analysis of the platform's forums.
Alia will conduct most of its cooperative decision--making on an online forum.
We will perform an ethnomethodological analysis of forum posts
on Alia to better understand its process for decision making.
We seek to answer:
what techniques do workers use for coalition-building around an online issue,
and how are conflicts resolved?
What percentage of discussion on the forum is dedicated to policy debate,
versus water cooler chatter or mentorship?
What percentage of co-op workers engage in policy debates,
what percentage lurk (read but do not post), and
what percentage do not even log in to follow debates?

Finally, we will perform
a cross--cultural analysis between the current platforms in
Section \ref{ethnography}
and the cooperative platform in
Section \ref{Alia},
comparing the social dynamics that arise in both platforms.
We expect to find that the social mechanisms people adopt to cope with
--- and to navigate ---
cooperatively--designed and
managed marketplaces differ substantially.
Our policy and design goal is
to understand the kinds of worker collectives that can succeed in an online piecework economy, and
to ensure that workers have access to them.

\section{Work Plan and Timeline}
The effort will proceed over two years.
In the first year,
we will perform the qualitative research with workers in order to understand
their platform policy and governance behaviors.
This includes the 100 interviews with information workers on
  Amazon Mechanical Turk and ``gig workers'' on
  Uber,
  Lyft, and
  Handy.
We will analyze the data using a grounded theory approach.
Simultaneously, in the first year,
we design and implement the Alia platform.

In the second year, we will launch Alia,
grow its use locally in the California Bay Area,
and perform fieldwork with Alia workers to understand their relationship with the cooperative.
This includes interviews with fifty workers,
surveys,
and ethnomethodological analysis of the co--op online forums.

A summary of the work plan is below:

% \begin{center}
  \begin{tabular}{|p{0.15\textwidth}|p{0.8\textwidth}|}
    \hline
    % \thead{<Header 1>} & \thead{<Header 2>} \\ %& \thead{<Header 3>} & \thead{<Header 4>} \\ \hline
\textbf{Period} & \textbf{Milestones} \\
\hline
Year one &
\begin{inlinelist}
  \item Fifty interviews with Mechanical Turk workers;
  \item fifty interviews with Uber/Lyft/Handy workers;
  \item implementation of Alia
\end{inlinelist}
\\
\hline
Year two &
\begin{inlinelist}
  \item Launch and grow Alia;
  \item fifty interviews with Alia workers;
  \item survey of Alia workers;
  \item ethnomethodological analysis of Alia forum
\end{inlinelist}
\\
\hline
  \end{tabular}
% \end{center}

\section{Budget Justification}
Our budget asks for support for a graduate student to lead the
  interviews,
  analysis, and
  platform implementation
over the two years.
This student,
  Ali Alkhatib,
holds undergraduate degrees in
  Anthropology and
  Informatics,
and is currently a Computer Science Ph.D. student,
so he has the training required to engage in both activities.
A second graduate student will aid for one quarter each year,
especially with managing the Alia platform.
PIs Levi and Bernstein will support the effort with one month of summer salary each year.

\section{Current and Pending Support}
Bernstein has received an NSF CAREER award (~\$500k) for research on crowdsourcing,
but none of the currently proposed projects were described in that award.
Levi is the PI for CASBS of a grant from the Rockefeller Foundation (\$150k)
to consider the future of work and workers.
This grant supports a working group and essay solicitation/editing but
no research assistants or other support for Levi.
Bernstein and Levi have received internal grant funding from
Stanford's Cyber Initiative (\$100k)
to engage in a social scientific and engineering effort for the future of online work.
That grant is focused on the design and evaluation of
new forms of teams and organizations for crowd workers,
as well as the design and evaluation of new microwork platforms that grant workers stronger reputation signals they can utilize to get hired.
Bernstein and Levi have applied for additional funding from
Stanford's Cyber Initiative for work described in this proposal (\$100k);
review is pending.

\section{Dissemination Strategy}
Our policy and design goal is to understand
the kinds of worker collectives that can succeed
in an online piecework economy,
and to ensure that workers have access to them.
Our research targets will be top--tier venues in human--computer interaction
(e.g., ACM CHI, ACM CSCW)
and political science and sociology
(e.g., APSR, ASR, AJS, AJPS, POP).

One main outlet for broader dissemination will be as part of
the Future of Work project at CASBS,
which recently published a series of essays in
Pacific Standard and placed op eds in important newspapers and online sources.
The relevant editors remain keen for new and additional material,
and various other magazines
(print and online),
such as American Prospect and Boston Review,
are also interested in possible publication of articles from this project.

A second set of outlets will be through
the NDWA's relationships with labor advocates and the press.
Our collaborators within the NDWA will manage this effort.
They are eager to position the NDWA and
the NDWA's Fair Care Labs
as an organization working on behalf of the growing class of gig workers.

\section{Conclusion}
An increasing number of workers are re--entering a piecework economy
via online data work and ``gig'' work.
This new online piecework is highly distributed and
individualistic in character, yet
workers still face the same challenges as ever in
representation and
collective behavior.
How are workers reacting to their situation,
and how are they empowered and inhibited in creating collectives
to air grievances and make decisions?
We seek to understand whether the set of online tools that have led to major collective efforts such as Wikipedia and social networks will likewise produce a new set of collective opportunities for workers.

Our work will broadly inform the ongoing policy debates about how to regulate the gig economy.
Gig workers are debated regularly in the pages of major newspapers,
however little literature has deeply investigated their position within the labor force,
how they weave together an income via many small piecework jobs,
and how they organize themselves.
Our goal is to aid this debate by contributing strong empirical evidence of workers' needs and lives under this new piecework economy.

\pagebreak
\begin{appendices}

\section{Social Capital Survey Scale}
These scales are adapted from \citet*{ellison2007benefits}.
Each item is posed as a 7--point Likert scale agree/disagree statement.

\subsection*{Bridging Social Capital Scale}
\begin{itemize}
  \item I feel I am part of the Alia community
  \item I am interested in what goes on at Alia
  \item Alia is a good place to be
  \item I would be willing to contribute money to Alia
  \item Interacting with people at Alia makes me want to try new things
  \item Interacting with people at Alia makes me feel like a part of a larger community
  \item I am willing to spend time to support general Alia activities
  \item In Alia, I come into contact with new people all the time
  \item Interacting with people at Alia reminds me that everyone in the world is connected
\end{itemize}

\subsection*{Bonding Social Capital Scale}
\begin{itemize}
  \item There are several people in Alia I trust to solve my problems
  \item If I needed an emergency loan of \$100, I know someone in Alia I can turn to
  \item There is someone in Alia I can turn to for advice about making very important decisions
  \item If I needed an emergency loan of \$100, I know someone in Alia I can turn to
  \item The people I interact with in Alia would be good job references for me
  \item I do not know people in Alia well enough to get them to do anything important (reversed)
\end{itemize}

\end{appendices}

\pagebreak
\bibliographystyle{plainnat}
\bibliography{../../references}
\end{document}

### 2016-05-16 proposals/nsfGRFP/personal_statement.tex
init!

\documentclass[12pt]{article}
% Load basic packages
\usepackage{balance,graphics,times,parskip,setspace}
\usepackage[pdftex]{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\singlespacing
\usepackage[tiny,compact]{titlesec}
\setlength{\parskip}{.4em}

\newcommand{\sectitle}[1]{\textbf{\MakeUppercase{#1}}}

% \setlist[enumerate,1]{%
%   label=\arabic*.,
% }

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}
\begin{document}

\begin{center}
\textsc{\MakeUppercase{Peer--managed Crowdsource Labor Markets}} --- Ali Alkhatib
\end{center}

% \textit{Roadmap:}
% \begin{itemize} \itemsep0pt \parskip0pt
%   \item Worked on Dynamo; learned about collective action
%   \item Continued thinking about issues, driven to explore technologically enabled work in general
%   \item Found myriad issues, began thinking more deeply about this subject:
%   \begin{itemize}
%     \item What are the common problems?
%     \item Who are the actors in these problems?
%     \item How can these actors contribute to solutions?
%   \end{itemize}
%   \item Went to MSR to study ``gig workers'' and design a worker-centric system
%   \begin{itemize}
%     \item Learned about cooperative labor groups
%     \item Designed interactions that place the benefits of workers first
%     \item Formed alliances with key stakeholders
%   \end{itemize}
%   \item We discovered characteristics that allow for the emergence of cooperative groups
%   \begin{itemize}
%     \item Some barriers prevent these approaches from scaling:
%     \begin{itemize}
%       \item Trusting members when membership is enormous
%       \item Sense of stakeholder status at vanishingly small individual stake
%       \item ???
%     \end{itemize}
%   \end{itemize}
%   \item Finding solutions to these problems is critical.
%   \begin{itemize}
%     \item Work is increasingly trending toward ``gig work''
%     \item The law is catching up, but markets continue to emerge finding new ways to shirk laws
%     \item Fundamentally, the best way to guarantee healthy market is to prove one can exist
%   \end{itemize}
%   \item I'm the best person to do this research:
%   \begin{itemize}
%     \item I come from a background in Anthropology; enables deep ethnographic research
%     \item I understand the technical domain better than any standard Anthropologist would
%     \item I have research experience in this space \& existing goodwill and trust with workers
%   \end{itemize}
% \end{itemize}

% Last year while working on Dynamo,
% a platform for collective action among workers on Amazon Mechanical Turk (AMT),
% I noticed a trend.
% It was subtle, but workers in the relatively nascent ``gig economy'' were beginning
% to speak out against the working conditions and treatment they were enduring until now.
% Among other things, workers began to demand fairer, more transparent treatment,
% more rigorous vetting of malicious requesters,
% and in some cases, liaisons between workers and the managers who operate these markets.

% In particular, they were airing their complaints to the operators of these markets,
% exposing a complicated legal maneuver these operators had contrived to minimize their legal exposure.

% Turkopticon and Dynamo empowered information workers in the gig market, but
% as I soon discovered, these platforms were stopgap solutions to deeply flawed social arrangements.
% To be specific, these markets assume an antagonistic role with relation to workers.
% In balanced markets, the deleterious effects of antagonism are mitigated by the competing forces.
% On markets like AMT and other ``gig work'' markets, however,

% Last year I began collaborating with workers on Amazon Mechanical Turk (AMT),
% known amongst one another as ``Turkers'',
% to study technologically enabled collective action.
% Together we built ``Dynamo'',
% a platform for organizing collective action which enabled workers to coordinate their joint protest of mistreatment;
% since its founding, Dynamo has fostered numerous campaigns designed to elevate these workers' statuses in the eyes of job requesters and market operators.

% I continued thinking about the issues that emerged while we worked on Dynamo,
% and eventually I found myself studying technologically enabled work in general,
% looking for undercurrents between these fields.
% I found parallels

% In these last 5 years I've gone from archaeological digs in Belize to designing crowd labor markets at Microsoft Research.
% In the time between then and now, I've

Relatively few Anthropologists can comfortably code in R or Python,
and similarly few Computer Scientists readily engage with Foucault's writing on biopower and biopolitics.
This type of dichotomy tends to prevent Computer Scientists from using their tools to explore socio--cultural topics deeply,
and social scientists from making use of powerful quantitative methods and techniques.
Vanishingly few researchers can claim to have engaged deeply in both fields.
As an Anthropologist, I have helped run archaeological field schools and led ethnographic research in field sites including Central and South America;
as a Computer Scientist, I have designed and helped build online systems such as platforms for collective action and labor markets for gig workers.

 % Microsoft Research.
% In this statement,
% I hope to illustrate that I have precisely the skill--set and motivation necessary to

% I have engaged so deeply in Anthropology and Archeology as staffing field schools in Central and South America,
% and demonstrated my competence as a Computer Scientist in my design and implementation of crowd labor markets at Microsoft Research.

% With my background in Anthropology and my skills as a Computer Scientist,
% I am uniquely capable of exploring deeply interdisciplinary, socio--technical questions.

% To be sure, the path I took to get to where I am was far from direct;
% I spent years at a community college with little idea of what I wanted to do with my life.
% My academic performance reflected my lack of motivation, and even after several quarters

I began my academic career at a community college with no idea of what I wanted to do with my life.
This ennui had plagued me through high school,
and my performance suffered so greatly that I nearly didn't graduate;
I completed several courses at a community college,
applying those credits to high school courses I hadn't successfully completed during my high school career,
and earned my diploma months later than intended.
When I arrived at community college,
I knew that it was a patch for a much deeper problem:
I didn't know what I cared enough about to pursue with earnest.

I had a vague sense that I wanted to study people and I was interested in technology,
but I had neither the motivation nor awareness of a field such as Human--Computer Interaction (HCI) to spark and focus that motivation.
I took courses in countless subjects, and
while all of these subjects interested me,
the passion that I saw in others was absent for me,
so I continued to search.

It was an Archeology course which sparked my passion for learning.
My then--professor extolled Anthropology's myriad uses,
and as I learned about archaeological sites and the material records they left,
I realized that the methodologies described in Archeology were the ones I wanted to learn and apply to living cultures.
I later discovered Cultural Anthropology
and found that while, methodologically, the quantitative methods of Archeology were at odds with contemporary Cultural Anthropology,
I could nevertheless find a niche and safely pursue my passion there.

Over the next 4 years I pursued anthropological and archaeological research,
pursuing ethnographic fieldwork in Belize and leading field school exercises in Ecuador between integrating and managing a database of tens of thousands of artifact finds from over a decade of past excavations.
In Belize and later in Ecuador,
I studied lithic shards and ceramic bits,
finding parallels in the silicon bits and array shards of digital cultures.
I not only developed skills that would enable me to study living cultures in novel ways,
but I acquired a passion and ambition for the role and status of the social sciences in an increasingly data--oriented world.

After transferring to UC Irvine I pursued an intensive curriculum of up to twice as many units per quarter as are typically allowed,
engaging in student government and multiple research initiatives through the Undergraduate Research Opportunities Program (UROP).
While engaged in the Multidisciplinary Design Program (MDP),
sponsored by UROP and advised by Dr.
Geoffrey Bowker and Dr.
Judith Gregory,
I found that Informatics offered the technical background I would need to pursue graduate--level research of digital communities and cultures.
In its diverse research area utilizing qualitative and quantitative methods I saw an opportunity to round myself out;
I consequently decided to pursue a degree in Informatics,
with an emphasis in Human--Computer Interaction,
in concurrence with my degree in Anthropology.
To accomplish this,
I enrolled in as many as 33 units per quarter while working and pursuing independent research.

Eager to gain knowledge,
tools,
and skills I had thus far been lacking,
I found myself taking so many more courses
--- and so many of which without many of the prerequisites associated ---
that my academic performance suffered.
That point is evidenced by several of the grades on my transcripts,
when I took advanced engineering courses in the implementation of programming languages,
emphasizing non-imperative languages like Prolog and Haskell,
for which I was woefully unprepared.
Keen to learn these languages,
and perhaps headstrong,
I enrolled in the course anyway and earned the lowest grade I received as an undergraduate.
Still,
the experience illuminated new modes of thought,
illustrated through the logic of these programming languages,
which I maintain I would not have had the opportunity to experience had I not taken the course at that time.

% That decisive point in my academic career
% --- the course in Archeology ---
% opened a door to a career in research that I had never realized existed,
% let alone that I could be a part of.
% More than that,
% when I realized that I could carve a path for myself,
% I found the internal drive I needed to make it possible;
% the goal I needed.
% My two years at UC Irvine
% --- in which I earned two separate degrees in disparate subjects,
% earned membership in the Phi Beta Kappa honors society,
% engaged in numerous research projects (one of which resulted in an undergraduate thesis),
% and managed to maintain a job ---
% illustrate that once I have identified and placed the full weight of my energy and focus on my goals,
% achieving them becomes easy.

My experience as a researcher while affiliated with a community college,
in Belize and Ecuador,
taught me both flexibility and perseverance.
In the highlands of Ecuador
--- with low oxygen from high altitudes, near--freezing climate,
and scorching sun owing to the reduced atmospheric coverage ---
we had to be both patient and determined in our fieldwork.
In Belize,
the grueling heat,
swarms of potentially malarial mosquitos,
and the frustrating opposition we experienced from locals suspicious of our intentions drove many to give up.
Those of us that pushed onward gained an appreciation for the rigor that goes into personally directed,
previously unexplored research questions.
Research as an archaeologist and in my fieldwork in other countries instilled in me a work ethic that has benefited me in every endeavor I have approached since then.

My research experience at UC Irvine further emphasizes both the interdisciplinary nature of my background as well as my ability to pursue research of the scale and ambition described in my research proposal.
In my first year at UC Irvine,
I pursued thoroughly interdisciplinary research through the previously mentioned MDP crossing the Donald Bren School of Information and Computer Science with the School of Social Ecology.
In this research,
I leveraged observational and ethnographic fieldwork I had learned over the years of fieldwork I had done as a pure anthropologist in Ecuador and Belize.
I also tapped into design skills I was beginning to learn as a student in Informatics and Human--Computer Interaction,
weaving my skills into a single skillset that spans disciplines.
Here,
I honed the crucial skill of incorporating and synthesizing wisdom from disparate fields,
a skill I have since found to be profoundly valuable.

In my second and final year at UC Irvine,
I engaged in independent research for my honors thesis on Quantified Self,
advised by Dr.
Tom Boellstorff and supported by the Undergraduate Research Opportunities Program at UC Irvine.
Every aspect,
from navigating ethical considerations with the Institutional Review Board,
to balancing the quantitative and qualitative research methods I had learned and was continuing to learn,
was both enormously difficult and profoundly rewarding.
Studying quantitative data, as well as critiquing the process of reflecting on my automatically--logged data,
% and digital paradigms alongside perhaps more conventional socio--cultural theories and frameworks,
helped concretize my sense of the intellectual space that I wanted to explore further.

After graduating from UCI,
I moved to Stanford to begin work as an independent researcher under Michael Bernstein, where
I studied collective action and how crowd workers deal with marginalizing, frustrating, and sometimes antagonistic labor markets.
Through that research,
I gained a deeper appreciation for the needs gig workers faced,
and the inherent challenges they encountered trying to mitigate these issues.
Our research led to ``Dynamo'',
a platform used by hundreds of Turkers to coordinate collective action to improve their circumstances and relationships between themselves and requesters.
Our research has produced a peer--reviewed publication,
a still--operating platform for Turkers,
and a set of ethical guidelines for academic requesters
\cite{dynamo}.

As successful as it was, Dynamo's launch
left me with a pervading sense that
our ability to positively affect worker outcomes in this emerging type of labor was fundamentally limited as long as we dealt in tools to \textit{ameliorate} the grievances of existing systems.
I felt increasingly that the most outsize impact we could have on markets like AMT, Uber, and the like would be to mitigate their exploitation and marginalization of workers;
the only way that we could illustrate a gig labor market that benefits workers and customers would be to articulate the system ourselves.
% It became increasingly apparent that solutions to the problems gig workers faced would come not from patches to existing systems,
% but from systems designed at the outset to empower and benefit workers.

I began working with Microsoft Research and the FUSE Labs last summer to explore the viability of worker--centric labor markets.
Working with the Service Employees International Union (SEIU), the International Brotherhood of Electrical Workers (IBEW), and the National Domestic Workers Alliance (NDWA) and their innovation arm the Fair Care Labs,
I conducted ethnographic fieldwork of day--laborers in and around Seattle, designed, and ultimately implemented a front--end system illustrating the interactions and affordances that a worker--centric labor market would offer.

My body of research has given me a strong background of research ranging myriad fields,
each leaving an indelible mark on the way and the rigor with which I pursue research.
I believe that these experiences,
and those that I will acquire over the next several years at Stanford,
will help me establish myself as a leading thinker, bridging the social sciences and computer science.

Robert A.
Heinlein coined a term well known today in programming culture: ``grok''.
It means to understand an idea so personally that it becomes a part of who you are.
Anthropology has a similar term for something that is understood personally and intuitively
--- ``emic'' knowledge.
Two distinct,
almost fundamentally different fields of study share a notion of the intimate familiarity of an idea,
demonstrating the potential to bridge these two areas that has yet to be explored more fully.
At its core,
this notion describes the artificial distinction and the incredible convergent evolution of hypotheses,
theories,
and models that the giants of our respective fields have invented.
These mirrored ideas,
and others like them,
may represent decades of research that might have been able to benefit from and synthesize with the findings of other fields,
if only someone adept enough in both fields could retrieve them and communicate them back.

With my ability to code--switch between computer science and the social sciences,
I believe that I am in a unique position to accomplish research that few others are properly equipped to do
--- the kind of research that deftly wields complex ideas from across the intellectual landscape.
Ultimately,
I hope to champion the view that conventional social sciences and what we call computational social sciences can
--- and indeed must ---
greatly influence one another;
that the social sciences can inform Human--Computer Interaction,
and that HCI and CSCW can inform the social sciences as well.
To do so will require the most adept translators of concepts and theories that we can train,
but to fail in this endeavor would be to separate
--- artificially and unnecessarily ---
two groups of thinkers who otherwise would benefit profoundly from one another's wisdom.

\renewcommand\refname{\sectitle{references}}
% % REFERENCES FORMAT
% % References must be the same font size as other body text.
% \bibliographystyle{SIGCHI-Reference-Format}
\bibliographystyle{acm}
{\footnotesize
\bibliography{../references.bib}
}

\end{document}

% TEMPLATE CREATED BY ANDREW FOR 2015 NSF GRFP APPLICATION. IN 
% NO WAY DOES HE GUARANTEE THIS TEMPLATE COMPLIES WITH
% FORMATTING RESTRICTIONS FOUND IN THE GRFP SOLICITATION. YOU
% SHOULD CHECK THE FORMATTING YOURSELF.
%
% ANDREW MAY BE CONTACTED VIA:
%      HIS WEBSITE: www.math.uga.edu/~amaurer
%      EMAIL:       andrew.b.maurer@gmail.com

### 2016-05-16 proposals/nsfGRFP/proceedings.tex
init!

\documentclass[12pt]{article}
% Load basic packages
\usepackage{balance,graphics,times,parskip,setspace}
\usepackage[pdftex]{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\singlespacing
\usepackage[tiny,compact]{titlesec}
\setlength{\parskip}{.4em}

\newcommand{\sectitle}[1]{\textbf{\MakeUppercase{#1}}}

% \setlist[enumerate,1]{%
%   label=\arabic*.,
% }

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}

\begin{document}
  \begin{center}
  \large{\bf Crowd--Operated Gig Markets and Crowdsourcing}
  % Ali Alkhatib
  \end{center}

\sectitle{introduction}\\
In the past several years, researchers have watched new forms of labor,
ranging from information work like Amazon Mechanical Turk (AMT)
to offline gigs like Uber and TaskRabbit emerge and balloon into industries now worth billions.
While workers in these markets point out that the transient ``gig work'' these markets enable
allow them to participate in the market when they might have been unable to do so before,
for instance due to medical conditions preventing them from making long--term commitments,
or competing obligations such as parenthood or caregiving.

On the other hand, this ``gig economy'' has been criticized for commoditizing
and ultimately marginalizing workers
\cite{uberAlgorithm}.
Efforts have attempted to make these markets more rewarding, or at least more bearable,
for workers
\cite{turkopticon,dynamo},
but we still see no fundamentally differentiated labor markets for workers;
virtually all of the existing markets trade one manager for another, maintaining an adversarial role between laborers and their managers.

Workers have faced issues like these for more than a century,
but they bear relevance to the HCI community broadly for two major reasons:
\begin{inlinelist}
  \item the features of digital systems allow researchers to have an outsize impacts on workers,
  whereas historically markets have been fractured \& untenable,
  and
  % --- or emulate their characteristics ---
  \item as researchers of technical systems and social interactions,
  researchers have a prerogative to attempt to steer systems toward more beneficial outcomes.
  \textit{researchers, broad impact, etc\dots}
\end{inlinelist}

\sectitle{hypothesis}\\
I propose to investigate the viability of a technologically--mediated,
worker--run market as an alternative to existing systems.
I hypothesize that
such a market can compete effectively,
and that workers and consumers would benefit more, % than under the status quo,
for a number of reasons.
% as a result of community--determined norms and rules.
% and that it would benefit workers \& consumers due to the equitable relationships that it inherently fosters.
For workers,
\begin{inlinelist} % \itemsep 0pt \parskip 0pt
  \item communally determined rules and norms are perceived as, and perhaps are, fairer, and
  \item a protocol for labor markets would allow markets to communicate things like worker reputation.
\end{inlinelist}
Customers, for their parts, would be able
% \begin{inlinelist} % \itemsep 0pt \parskip 0pt
  % \item to discriminate between markets on various criteria, and
  to leverage market competition, minimizing predatory pricing models.
% \end{inlinelist}

\sectitle{background}\\
The most immediate question is why technologically--mediated
worker--run markets are so rare.
One may have predicted that worker--run markets would emerge inevitably
as online systems parallel (and sometimes model) offline settings.
Furthermore, the structure of the Internet
--- a ``rhizome'', as Miller and numerous others describe \cite{miller2011understanding} ---
suggests that communities should be able to form naturally,
unencumbered by barriers like geography.
Futurists predicted that the Internet would democratize communication,
enabling grassroots organization at unprecedented scales.

The empirical research suggests that the reality of this promise is more qualified
\cite{dynamo,toyama2015geek}.
Indeed, the factors which stymie community--run digital markets are not technical.
Researchers have built on existing markets to experiment with the organization of work;
others have implemented labor markets entirely from scratch
\cite{Alek2011,foundry}.
The components of gig labor market
--- such as scheduling, dispatch, \& payments ---
are sufficiently explored that we should consider them ``solved problems''.

The challenge of a worker--run market seems to be sociocultural.
The emergence, communication, and enforcement of norms
add a dimension of complexity to the individual cases of collective action represented by the study of individual events.
Engineering instances of collective action is a qualitatively different challenge from
fostering a community of ongoing collective action
\cite{russell1982collective}.
% (e.g. \cite{catalyst,dynamo}).

\sectitle{Methodology}\\
I intend to develop a worker--run labor market,
which the National Domestic Workers Alliance (NDWA)
will operate through their innovation arm, the Fair Care Labs.
I hope to demonstrate that cooperatively operated
digital labor markets can be successful, and that technologically--mediated
cooperative labor markets are viable alternatives to markets operated as firms.

The most immediate goal is to implement a system consisting of
the technical requirements of gig labor markets described before
(factors such as scheduling, dispatch, \& payments).
This system would resemble a mobile application,
similar in appearance to existing labor platforms,
but importantly differentiated by affordances for community governance:
\begin{inlinelist}
  \item forums for discourse,
  \item voting systems for referendums, and
  \item bylaws written and approved by the community approval.
\end{inlinelist}

The Fair Care Labs plans to deploy this system in two locales
--- the San Francisco Bay Area and the Seattle area ---
in order to allow me to provide high--level guidance
and to make course--corrections as necessary.
This should ameliorate concerns both about funding the system
and sourcing workers;
my involvement is the extent to which I will be able to contribute as a researcher.

% Past the instantiation of the marketplace and the recruitment of workers,
% the questions that remain relate to
% the ongoing relationships that form,
% the norms that emerge, and the methods and practices Computer Scientists might employ to foster productive relationships,
% informed in large parts by insights made through Anthropological inquiry.

% As early as 1982 it had been argued that
Using Hardin's perspective of collective action,
distinguishing between ``one--shot'' and ``ongoing'' campaigns,
we realize it's necessary to consider how groups sustain collective action over time.
% \cite{russell1982collective}.
Thus, designing ongoing collective action requires skills in 
system--building,
as well as a ``nearly anthropological investigation'' to engage in the sociocultural questions herein
\cite{russell1982collective}.
This describes the type of evaluation involved:
a reflection on the ways in which people work and relate with one another,
and the techniques we can use as engineers of social systems to guide relationships.

% In this area, I am relatively uniquely qualified.
The ethnographic and design work I've done at Microsoft Research (MSR),
my work at Stanford University (Dynamo
\cite{dynamo}),
and my background and training as an Anthropologist
make me ideally skilled to pursue this area.
I have demonstrated that I can build systems,
and crucially I have proven experience in the qualitative,
nebulous fieldwork necessary for this research.

% With the NDWA funding and operating this market,
% Continuing the work I began with the NDWA while working with MSR,
% and having secured their commitment to fund and operate this market,
% I'm confident that a functionally viable labor market will be given a real chance to succeed.
% While the NDWA are committed to more worker--centric labor markets,
% my ability to contribute depends in part on my ability to continue my career.

% \dots
% boom microphone drop give money please

\renewcommand\refname{\sectitle{references}}
% % REFERENCES FORMAT
% % References must be the same font size as other body text.
% \bibliographystyle{SIGCHI-Reference-Format}
\bibliographystyle{acm}
{\footnotesize
\bibliography{references}
}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-05-16 proposals/nsfGRFP/research_statement.tex
init!

--- 2016-05-16 proposals/nsfGRFP/proceedings.tex
@@
   \begin{center}
-  \large{\bf Crowd--Operated Gig Markets and Crowdsourcing}
+  \large{\bf Worker--Managed Crowdsource Labor Markets}
   % Ali Alkhatib
@@
 \sectitle{introduction}\\
-In the past several years, researchers have watched new forms of labor,
-ranging from information work like Amazon Mechanical Turk (AMT)
-to offline gigs like Uber and TaskRabbit emerge and balloon into industries now worth billions.
-While workers in these markets point out that the transient ``gig work'' these markets enable
-allow them to participate in the market when they might have been unable to do so before,
-for instance due to medical conditions preventing them from making long--term commitments,
-or competing obligations such as parenthood or caregiving.
+In the past several years, researchers have watched new forms of labor
+--- exemplified by labor markets like Amazon Mechanical Turk (AMT), Uber, and TaskRabbit ---
+emerge and balloon into billion--dollar industries.
+%
+While these markets allow people to engage in the workforce in novel ways,
+this ``gig economy'' has been criticized for commoditizing
+and marginalizing workers
+\cite{uberAlgorithm}.
 
-On the other hand, this ``gig economy'' has been criticized for commoditizing
-and ultimately marginalizing workers
-\cite{uberAlgorithm}.
-Efforts have attempted to make these markets more rewarding, or at least more bearable,
-for workers
-\cite{turkopticon,dynamo},
-but we still see no fundamentally differentiated labor markets for workers;
-virtually all of the existing markets trade one manager for another, maintaining an adversarial role between laborers and their managers.
+On Turkopticon \cite{turkopticon} and in my prior work on Dynamo \cite{dynamo},
+we attempted to make these markets more rewarding for workers,
+but no labor markets to date place workers first;
+existing markets trade one manager for another, maintaining an adversarial role between laborers and their managers.
 
@@
   \item as researchers of technical systems and social interactions,
-  researchers have a prerogative to attempt to steer systems toward more beneficial outcomes.
-  \textit{researchers, broad impact, etc\dots}
+  researchers have a prerogative to attempt to steer systems toward more beneficial outcomes,
+  especially given the growth of this form of labor and the potential implications for the future of work.
 \end{inlinelist}
@@
 Customers, for their parts, would be able
-% \begin{inlinelist} % \itemsep 0pt \parskip 0pt
-  % \item to discriminate between markets on various criteria, and
 to leverage market competition, minimizing predatory pricing models.
-% \end{inlinelist}
 
@@
 
-The challenge of a worker--run market seems to be sociocultural.
-The emergence, communication, and enforcement of norms
-add a dimension of complexity to the individual cases of collective action represented by the study of individual events.
+No; we're finding that the challenges to worker--centric markets are social.
+Research has shown that crowds can be directed
+and can make tactical decisions on an ad hoc basis,
+but collective governance and policy--making,
+especially in competitive economic settings, remains an open question.
+% The emergence, communication, and enforcement of norms
+The coalescence of culture
+adds a dimension of complexity to the individual cases of collective action represented by the study of individual events.
 Engineering instances of collective action is a qualitatively different challenge from
@@
 
-The most immediate goal is to implement a system consisting of
-the technical requirements of gig labor markets described before
-(factors such as scheduling, dispatch, \& payments).
-This system would resemble a mobile application,
-similar in appearance to existing labor platforms,
-but importantly differentiated by affordances for community governance:
-\begin{inlinelist}
-  \item forums for discourse,
-  \item voting systems for referendums, and
-  \item bylaws written and approved by the community approval.
-\end{inlinelist}
+My immediate task is to create a labor market which illustrates the characteristics described
+--- technical features like scheduling, dispatch, \& payments, as well as
+the necessary affordances for collective governance:
+% The most immediate goal is to implement a system consisting of
+% the technical requirements of gig labor markets described before
+% (factors such as scheduling, dispatch, \& payments).
+% I will develop a mobile application,
+% similar in appearance to existing labor platforms,
+% but importantly differentiated by affordances for community governance:
+% \begin{inlinelist}
+  % \item
+  forums for discourse,
+  % \item
+  voting systems for referendums, and
+  % \item
+  communally written bylaws. %  written and approved by the community.
+% \end{inlinelist}
+These tools will be the starting point for a worker--run crowd labor market,
+upon which more affordances will be designed and implemented as needed.
 
-The Fair Care Labs plans to deploy this system in two locales
---- the San Francisco Bay Area and the Seattle area ---
+The Fair Care Labs plans to continue the relationship that began over the summer while I was affiliated with Microsoft Research,
+deploying this system in the San Francisco Bay Area and the Seattle area
 in order to allow me to provide high--level guidance
-and to make course--corrections as necessary.
+and to make any necessary course--corrections.
 This should ameliorate concerns both about funding the system
@@
 % In this area, I am relatively uniquely qualified.
-The ethnographic and design work I've done at Microsoft Research (MSR),
-my work at Stanford University (Dynamo
-\cite{dynamo}),
+The ethnographic and design work I've done with various labor advocacy groups,
+my prior work on Dynamo
+\cite{dynamo},
 and my background and training as an Anthropologist
-make me ideally skilled to pursue this area.
+make me ideally skilled to pursue this line of research.
 I have demonstrated that I can build systems,
 and crucially I have proven experience in the qualitative,
-nebulous fieldwork necessary for this research.
+sometimes nebulous fieldwork necessary for this research.
 
@@
 {\footnotesize
-\bibliography{references}
+\bibliography{../references.bib}
 }

### 2016-05-17 cscw/papers/2017/labor/trackingWork.tex
incidental updates

@@
 Routenization can be seen in the contemporary in workflows
+\cite{foundry,bernstein2015soylent}
 [\textit{There are various papers that talk about creating a workflow;

### 2016-05-19 chi/papers/2016/groundwork/background.tex
major revisions to intro and generally framing need to update the rest of the paper to reflect the change in topics/categories that we look through

%!TEX root = proceedings.tex
\section{Background \& Theory}

The existing knowledge from which we draw in this research can primarily be traced to two pools of research;
the first comes from the extensive body of research surrounding collective action,
and the work that has gone into describing, categorizing, predicting, and even designing to foster campaigns of collective action.
The second major source of our knowledge comes from the study of on-demand or gig markets.
By synthesizing the discoveries from these sources of research, we identify \textit{a priori} guidance at this intersection of two subjects.

\subsection{The Long History of Collective Action}
Grassroots, community-led organizations are not new;
a substantial body of literature illustrates myriad approaches to guiding communities and assisting in collective action.
Specifically, when we consider the role of insight into collective action, we refer to what Hardin describes as ``directed at an ongoing problem"
\cite{russell1982collective}.
The implication here, he argues, is that the guidance on this form of collective action is dramatically more nuanced than ``one-shot" collective action, demanding an ``anthropological investigation of minute interrelationships".
We might call this ``ongoing" collective action.
Economist Mancur Olson proposes, in part, that collective action depends on some large, generally inactive group in order to succeed
\cite{olsonlogic}.

After decades of observations, Hardin posits that collective action is too commonplace for Olson's thesis to hold;
He suggests that the requirements for collective action which Olson theorizes may have changed
--- specifically, lowering the threshold ---
as a result of myriad factors outside of the scope of this research, except to point out that recent work in online collective action prompts further scrutiny of Olson's thesis and the critiques later researchers have levied.
We suggest an alternative consideration: that much of the research in collective action in the space of HCI in fact corroborates the latent community requirement Olson recommends.
Myriad collective action endeavors seem to succeed in part because they precipitate a collective of latent, willing participants in some form of community action
\cite{catalyst,dynamo,foundry}.

% Given this interpretation of successful cases of collective action, we infer that designing a system specifically with the precipitation of a crowd as future participants may be a requisite factor in successful collective action.

\subsection{On-Demand Markets and Their Workers}

A robust and growing body of research exploring collective action and movements enabled by the Internet adds to a body of knowledge previously uninformed by the tools the Internet affords.
Where collective action research coordinated and executed offline describes challenges symptomatic of social structures,
researchers of online communities can and do offer design guidance for the structure of online communities
\cite{Hirsch:2009:FLA:1516016.1516024}.
Substantial contributions deeply investigating online communities, such as Wikipedia, have lent system designers guidance in designing communities geared toward some ongoing collective action online
\cite{Nov:2007:MW:1297797.1297798,wikipediansBornNotMade,whyWikipedians}.
In this last case, studies of Wikipedia and its users
--- known colloquially as ``Wikipedians" ---
is especially instructive, as it addresses the distinction Olson makes regarding ``one-shot" collective action and what we will call ``ongoing" collective action.

Researchers have described the potential of crowd work, and perhaps more importantly outlined various problems with said markets
\cite{crowdworkFuture}.
More recently, researchers explored the impact of the sharing economy among socioeconomically disadvantaged communities,
and studied norms and behaviors among the workers of individual markets like Uber
\cite{DillahuntPromise,uberAlgorithm}.
The directives offered by Kittur et al. provide substantial guidance in the conceptual design of a ``crowd market",
but the direct application of this research to a communally led organization
--- a worker-run labor market --- merits further consideration.
% but the broader transformative effects of the sharing economy passed scrutiny relatively quietly.

The focus of technologically enabled peer economies is well-explored;
researchers have pointedly identified the efforts of both for and non-profit online peer markets and the challenges workers and market operators face
\cite{uberAlgorithm,Bellotti:2014:TCS:2611247.2557061,Thebault-Spieker:2015:ASS:2675133.2675278}.
Some research has implemented independent micro-work markets for their own purposes
\cite{Alt:2010:LCE:1868914.1868921},
though the purpose of this research was not to explore the idea of a worker-centric or worker-led market.

Computational social scientists have documented workers' efforts to circumvent the systems imposed on them by market operators
\cite{uberAlgorithm},
and more directly researchers have observed the continuing effort to resist and critique markets for ``gig work", in these cases in the context of online labor, where micro-work on Amazon Mechanical Turk (AMT) predates offline gig work companies such as Uber
\cite{turkopticon,dynamo}.

Nevertheless, frustrations with these marketplaces persist,
and the trends among emerging marketplaces seem to commoditize workers more and more aggressively.
These ``patches" of existing markets appear to have only marginal effects on the qualities of these markets;
Uber drivers continue to resist the algorithmic matching while walking a fine line to avoid retribution from management, and some of the most frustrating requesters on AMT continue to antagonize Turkers.

\subsection{Juxtaposed, but not Synthesized}
Broadly, these bodies of research have overlapped in limited cases
\cite{dynamo,turkopticon}.
The intersection of ``community-driven action as a mode of designing" \& ``design of technologically-enabled peer markets" represents a field site ready for the application of existing knowledge, and perhaps the production of new.
We begin to bridge this research by applying our learnings regarding collective action (both offline and online) to what we have learned from research surrounding the ``gig economy" in its various names.
Drawing on the spirit of critical theory research and the grievances found among gig workers \cite{turkopticon,uberAlgorithm}, we began our fieldwork thinking about the viability of a self-directed crowd, informed perhaps by the

### 2016-05-19 chi/papers/2016/groundwork/designGuidance/assigningWork.tex
major revisions to intro and generally framing need to update the rest of the paper to reflect the change in topics/categories that we look through

%!TEX root = ../proceedings.tex
% \subsection{Assigning Work}
\textit{Who gets new work opportunities first is a contentious issue which should be negotiated by people, enforced by technology.}

When new work becomes available, who should have first access to claim it proves to be a contentious topic.
Depending on the intended velocity of the market, various suggestions have emerged.

In existing technologically enabled markets, we find some of these decisions have been made carefully.
Some of these approaches are worth considering, and can be described thusly:
\begin{itemize} \itemsep0pt \parskip0pt
  \item On AMT, Turkers find themselves in constant competition with one another to claim HITs;
  the effects of this task searching behavior have been explored in some detail \cite{taskSearch}.
  \item Airbnb, Craigslist, and others invite workers to describe the services or products they offer, and leave the work of matching to customers.
  The complex nature of finding a suitable place to live may make this approach preferable over automated matching.
  \item Uber, Lyft, and others use algorithmic matching to determine who should be offered work;
  it is unclear to workers and consumers who is given a job offer first,
  but drivers we spoke to assumed that the nearest drivers were given first offer of a new job in markets such as Uber \& Lyft.
\end{itemize}

Among each of these approaches designers can make choices that dramatically affect the experiences for workers and customers.
The first option, used by Amazon Mechanical Turk, Craigslist, etc\dots makes it possible for job turnover to be extremely high, but the cost is that workers may feel stressed by the uncertainty over whether a job they're considering has already been claimed.
The second approach mentioned, which Airbnb uses, gives customers the choice to contact ``workers", but the process of finding a provider makes the overall matching process time-consuming.

When we spoke with members of organization B and organization A,
the consensus we initially found was that seniority within a reasonable proximity to the work would be the fairest solution to the problem we described.
After some discussion, it emerged that this position was not universally shared among members of either group.
Perhaps understandably, younger members of organization B felt that workers who had been working longer that day without having been offered any jobs,
or workers who were significantly closer, should eventually be given higher priority for work.

organization E chose yet another option, primarily using a random selection process at the beginning of each hiring day by pulling the names of currently present workers out of a jug;
every eligible worker has a reasonable chance at being selected first, just as they have a chance of coming up last.
Workers could short-circuit the selection process by volunteering to do work beneficial to the group
--- for example, making coffee, or sweeping the common room's floors.
By volunteering to do this work, they're guaranteed first priority for incoming jobs the following day,
This ensures that they'll be given an offer for work quickly on the following day.

The best option for most markets might be the model Uber, Lyft, and others have adopted,
where customers are automatically routed to workers,
who are given a time window where they have exclusive claim to a job that's been posted.
If they reject a job offer, the job moves to the next eligible worker.

\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/scheduling}
    \caption{In the client's perspective, a lightweight scheduling system can suffice to communicate availability.}~\label{fig:clientSchedule}
\end{figure}
\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/jobtimesoffer}
    \caption{In the worker's perspective, scanning suggested start times can make scheduling easier than entering a series of windows of availability each week or month.}~\label{fig:workerSchedule}
\end{figure}

The task of finding eligible, willing workers can be simplified by considering availability,
narrowing the list of candidate workers to those who are interested in working during a given time frame.
This can, however, be a particularly frustrating challenge for on-demand workers.
One of the most common reasons we heard from gig workers who preferred such work was the flexibility that mode of work offered;
in other words, trying to get workers to commit to windows of availability goes against the nature of that work --- the freedom --- which made it appealing to those we consulted.

This challenge was ameliorated in various ways across field sites; At organization E, workers were assigned each day even if work was solicited days or weeks in advance (unless, of course, a specific worker was requested for repeat work). This process neatly handles the potentially ambiguous question of availability for a number of workers, but it's limiting in two ways:

\begin{enumerate} \itemsep0pt \parskip0pt
  \item organization E's scale is ultimately limited by their ability to process workers on a daily basis (this was particularly salient as a volunteer, processing workers as they arrived to check at 7 A.M.)
  \item Workers who wished to be available for leads had to travel to the organization E offices; sometimes workers would travel north as many as 30 minutes, only to get a job 30 minutes south again.
\end{enumerate}

Instead of setting availability in advance, we suggest optimizing the process of scheduling so that workers and customers spend a minimal amount of time negotiating one another's availability.

This leaves open the question of how that list of workers is determined;
who is the first choice, and who's second if the first worker doesn't claim the job?
Who ends up getting last pick?

We can only offer that this factor is determined in large part by the organization running the system.
A template for a worker-run marketplace should therefore expose these options,
and more importantly surface the debate at the core of these discussions,
while abstracting away the technical details such as implementation.
These cases also illustrate that systems can be designed in ways that meaningfully and fairly reward workers for doing necessary work for the benefit of the community as a whole.

### 2016-05-19 chi/papers/2016/groundwork/discussion.tex
major revisions to intro and generally framing need to update the rest of the paper to reflect the change in topics/categories that we look through

%!TEX root = proceedings.tex
\section{Discussion}
We spent several months working alongside workers in various industries,
iterating on mockups and prototypes,
and getting feedback from members and administrators of worker-advocacy groups like organization C, organization B, and organization D.
With their insights, we identified a number of aspects of gig work that potentially marginalize workers,
and in doing so we begin to articulate ways to avoid such an outcome.
These potentially marginalizing aspects in the peer economy consist of the following:

\begin{enumerate}\itemsep0pt \parskip0pt
  \item Constructive feedback
  \item Assigning work fairly
  \item Managing customer expectations
  \item Protecting vulnerable workers
  \item Reconciling worker identities
  \item Assessing worker qualifications
  \item Communicating worker quality
  % \item providing workers with substantive constructive feedback,
  % \item fairly assigning work in groups
  % \item managing customer expectations
  % \item protecting vulnerable workers
  % \item respecting worker identity
  % \item assessing worker qualifications
  % \item communicating worker quality
\end{enumerate}

We discussed the importance of such a system given the existing trends in technologically enabled labor markets,
and suggest an alternative market design wherein workers might operate their marketplace collectively.
Finally, informed by ethnographic fieldwork and interviews, we distill what we believe conscientious system-builders would need to create a worker-centric peer economy.

It's worth noting that while we worked with workers from various industries,
seeking to find common threads between each community,
we can't speak to the applicability of these findings across all conceivable industries and modes of work.
Information workers such as those on Amazon Mechanical Turk may differ in significant,
fundamental ways compared to drivers-for-hire such as Uber drivers.

Above all, we hope to convey the potential insight one can gain by working with stakeholders,
and the importance of collaboratively designing labor markets with the workers themselves.

As Kittur et al. point out, as researchers, computer scientists, and participants in the technological community we ask whether
\textit{``\dots we foresee a future crowd workplace in which we would want our children to participate''} \cite{crowdworkFuture}.
Indeed, we must answer this question, by articulating an economy that empowers and respects workers, rather than marginalizes and exploits them.

### 2016-05-19 chi/papers/2017/referrals/referrals.tex
major revisions to intro and generally framing need to update the rest of the paper to reflect the change in topics/categories that we look through

--- 2016-05-17 cscw/papers/2017/labor/trackingWork.tex
@@
 \section{Introduction}
+In the past decade
+microtasks, the ``sharing economy'', and other instantiations of on--demand contract labor
+have grown in popularity both in the academic community as well as the collective consciousness
+\cite{Ross,CrowdsourcingUserStudies,uberAlgorithm}.
+
 This section should outline the trend of contemporary gig work in the last 10 years,
@@
 \subsection{Piecework as a lens to understand gig work}
-It's not new that researchers have discussed gig work as an extension or resurgence
-of piecework, but it is sufficiently important for the framing of this argument that
-it merits a more direct treatment and some explanation here.
+It's not new to discuss gig work as an extension or even a resurgence of piecework
+\cite{crowdworkFuture,fixingChaos}.
+Nevertheless, it's important enough for the framing of this argument that
+a more direct treatment and some explanation is worthwhile.
 \citeauthor{crowdworkFuture}
@@
 it was an inevitable milestone in the maturation of the type of work that is reemerging today.
-\textit{How to continue???}
+``But'',
+as \citeauthor{scholz2012digital} points out,
+``it would be wrong to conclude that
+in the realm of digital labor there is nothing new under the sun''
+\cite{scholz2012digital}.
+Indeed, we will attempt to reason about more than the larger--scale progression from
+the discretization of work toward scientific management and future steps;
+we will attempt to make sense of how the modern processes in digital labor,
+in particular the use of the Internet as a medium to manage as well as \textit{do} work,
+affect these unfolding events.
 

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
fleshing out 3 of the case studies

@@
   --- driven in part by the discretization,
-  routenization,
+  routinization,
   and external management of said work ---
@@
 \section{Introduction}
-This section should outline the trend of contemporary gig work in the last 10 years,
-in particular tracking the changes since
-\citeauthor{crowdworkFuture} discussed
-\citetitle{crowdworkFuture}
-in \citeyear{crowdworkFuture}
-\cite{crowdworkFuture}.
-This would review the literature since then,
-then point to the movements toward worker unionization, advocacy, etc\dots
-that have happened in the last year or two.
+
+% \itshape
+% What am I trying to say here?
+% I want to get across that researchers have looked at gig work in various ways,
+% but not come up with very persuasive overarching explanations for this whole process.
+% That is, there's no broader theory making sense of why all of this is happening.
+
+% Rather than make the argument that
+% these things are all independently similar to the piecework
+% (something to which researchers have already made the connection),
+% I want to frame this as a huge process, involving:
+% \begin{enumerate}
+%   \item distribution of work,
+%   \textbf{which leads to\dots}
+%   \item discretization and routinization of work,
+%   \textbf{which allows\dots}
+%   \item management (scientific \& external),
+%   \textbf{which causes workers to feel\dots}
+%   \item alienation from work (Marx's idea) and resistance to management (also Marx probably but a more general labor advocacy theory too)
+% \end{enumerate}
+% okay, let's try this:
+% \upshape
+
+The past decade has seen
+microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
+grow to occupy the fascination of both academic circles and our culture as a whole
+\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
+The research community has identified myriad connections between this emergent form of work
+and the historically situated ``piecework'',
+as well as a number of observations relating to
+the frustration workers feel stemming from the management of this work
+\cite{turkopticon,dynamo,uberAlgorithm}.
+This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.
+
+While much of this work appears to gesture toward
+the parallels between contemporary on--demand work and piecework,
+it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
+This paper will attempt to do so by arguing that
+the topics social computing researchers have investigated are
+not just parallel to historical piecework and
+the process of factorization that took place in the early \nth{20} century;
+indeed, these phenomena are directly related, and that they precipitate
+the social outcomes we have observed,
+especially with regard to worker frustration, coordination, and resistance.
+
+% This section should outline the trend of contemporary gig work in the last 10 years,
+% in particular tracking the changes since
+% \citeauthor{crowdworkFuture} discussed
+% \citetitle{crowdworkFuture}
+% in \citeyear{crowdworkFuture}
+% \cite{crowdworkFuture},.
+% In particular, we point to \citeauthor{crowdworkFuture}'s discussion of worthwhile
+% questions to ask regarding \citetitle{crowdworkFuture}
+% \cite{crowdworkFuture}.
+% This would review the literature since then,
+% then point to the movements toward worker unionization, advocacy, etc\dots
+% that have happened in the last year or two.
 
 In the reflection on the literature published in the last 5 years since
+\citeauthor{crowdworkFuture}'s 
 \citetitle{crowdworkFuture},
-we notice a broader change in the form of work that's being done;
-motivated by this realization,
-we look at the emergent phenomena in piecework both as interconnected.
+we notice a broader trend describing the change in work that's being done
+% motivated by this realization,
+% we look at the emergent phenomena in piecework both as interconnected
+\cite{crowdworkFuture}.
+Crowd--powered work is not just distributed across homes,
+but dispersed among drivers, cleaners and other on--demand workers.
 By looking at the movements toward
-discretization,
-routenization,
-and management
-as necessarily linked --- even causally --- to one another, 
-we give ourselves a framing of contemporary piecework that explains
-and arguably predicts
-what we have seen thus far.
-
-\subsection{Piecework as a lens to understand gig work}
-It's not new that researchers have discussed gig work as an extension or resurgence
-of piecework, but it is sufficiently important for the framing of this argument that
-it merits a more direct treatment and some explanation here.
-\citeauthor{crowdworkFuture}
-investigate the future of crowd work by
-situating and interrogating it through ``piecework'',
-a term almost lost to history, but which for a time described
-work done in the home,
-in manageable tasks,
-often involving clear instructions
-and payment only for work completed, not work done
-(the differentiation, here, being that
-one would be paid for the \textit{output} of the work,
-not the \textit{duration})
-\cite{crowdworkFuture}.
-This work was largely in textiles and was predominantly done by women,
-which by itself is a topic worthy of unpacking
-\cite{scott1975women}.
-Given the scope, however,
-we will have to leave this complex topic with the superficial takeaway that
-piecework and microtasks share some structural similarities:
-\begin{inlinelist}
-\item the work is done in the home;
-\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
-\item the worker's status
-(not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy.
-\end{inlinelist}
+distributing work,
+routinizing and breaking down tasks,
+and externally managing workers
+% discretization,
+% routinization,
+% and management
+as linked to one another, 
+we give ourselves a framing of contemporary piecework that explains,
+and arguably predicts,
+what we have seen thus far
+--- and perhaps what we should expect to see going forward.
 
@@
 
-This paper will argue that the similarity is more than striking;
+This paper will argue that the familiarity is more than passing;
 it was an inevitable milestone in the maturation of the type of work that is reemerging today.
-\textit{How to continue???}
+``But'',
+as \citeauthor{scholz2012digital} points out,
+``it would be wrong to conclude that
+in the realm of digital labor there is nothing new under the sun''
+\cite{scholz2012digital}.
+Indeed, we will attempt to reason about more than the larger--scale progression from
+the discretization of work toward scientific management and future steps;
+we will attempt to make sense of how the modern processes in digital labor,
+in particular the use of the Internet as a medium to manage as well as \textit{do} work,
+affect these unfolding events.
 
@@
 
+\subsection{Piecework as a lens to understand gig work}
+% It's not new to discuss gig work as an extension or even a resurgence of piecework
+% \cite{crowdworkFuture,fixingChaos}.
+% Nevertheless,
+% it's sufficiently core to the argument here that a review is worthwhile.
+\citeauthor{crowdworkFuture}
+investigate the future of crowd work by
+situating and interrogating it through ``piecework'',
+a term almost lost to history, but which for a time described
+work done in the home,
+in manageable tasks,
+often involving clear instructions
+and payment only for work completed, not work done
+(the differentiation, here, being that
+one would be paid for the \textit{output} of the work,
+not the \textit{duration})
+\cite{crowdworkFuture}.
+This work was largely in textiles and was predominantly done by women,
+which by itself is a topic worthy of unpacking
+\cite{scott1975women}.
+Given the scope, 
+we'll leave this topic with the superficial takeaway that
+piecework and microtasks share some structural similarities:
+\begin{inlinelist}
+\item the work is done in the home;
+\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
+\item the worker's status
+(not only socially, but also economically)
+is ambiguous, or at least the subject of some controversy.
+\end{inlinelist}
+It may be said that
+on--demand work shares, if not inspiration, then many of the same outcomes as piecework.
+
 \subsection{Context}
@@
 starting with the discretization of tasks,
-followed by routenization and the rise of workflows,
+followed by routinization and the rise of workflows,
 and finally the external management of workers.
@@
 We explore each topic
---- discretization, routenization, \& management ---
+--- discretization, routinization, \& management ---
 by looking at case studies in social computing.
@@
 \section{Case Studies}
-The past decade of research has shed light on crowd work from various perspectives,
-and revealed 
-\subsection{Discretization}
+The existing body of research has shed light on on--demand from various perspectives,
+and revealed a number of topics that,
+through our framing, are clearly situated together.
+Those topics, at a very high level, are
+\begin{enumerate}
+\item the \textbf{distribution} of work,
+\item the \textbf{routinization} and \textbf{discretization} of work, and finally
+\item the \textbf{relationships} between the workers and the product of their work,
+      as well as workers and managers.
+\end{enumerate}
+
+We will follow these topics, using them as prompts for case studies in
+the emergence and
+development of contemporary piecework.
+
+\subsection{Distributing Work}
+In both the historical and contemporary cases,
+distributed, speculative work
+% seemed a reasonable way
+has promised
+to diffuse the costs of capital--intensive resources
+(whether they were sewing machines or laboratory equipment).
+
+\textit{There's a reference to something that's \textbf{not} 99designs,
+that describes a call for scientific solutions/designs to solve some problem
+that the requesters have.
+Honestly it's killing me that I can't remember the name,
+and I need to find this before I do anything productive ever again.}
+
+The HCI community is perhaps most familiar with examples such as
+99designs,
+Uber, Lyft,
+and Amazon Mechanical Turk (AMT),
+which all allow requesters in various forms to tap into
+resources such as
+cars,
+computers, and above all
+``cognitive surplus''
+with relative ease
+\cite{DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
+This insight, that
+workers can be geographically distributed,
+% (especially given careful central management of those workers, discussed later),
+has proven remarkably compelling
+\cite{sensitiveTasks}.
+
+The insight here seems to have been that
+a substantial amount of work does not necessarily need to be done in collocated space,
+but can be done from one's own home or other space arranged by the worker.
+Moreover, by shifting the workplace into the worker's domain,
+the cost of capital involved in that work can more easily be placed on workers' backs.
+While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
+(inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
+these new systems go further,
+by framing ownership and responsibility for costs
+(e.g. automotive maintenance)
+on drivers.
+In our fieldwork, we discovered many longtime ``on--demand'' drivers
+(affiliated with platforms such as Uber and Lyft)
+who readily voiced frustration with the costs of
+frequent oil changes and the unexpected replacement of expensive parts.
+
+Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
+textile workers were given the source material with which to work,
+but were generally expected to use their own sewing materials and work in their own homes
+\cite{hapke2004sweatshop}.
+
+% the change seemed to begin with
+% the redistribution of work into the homes and onto the diffuse capital of workers.
+% 99designs may represent one of the first illustrative cases of
+% speculative, distributed labor in the contemporary form,
+% but speculative labor and even on--demand design work is not necessarily new;
+% \cite{99designs}.
+
+% distributed; uber, etc... ->
+% routinization; crowdwork - > optimization
+% - > accommodation and resistance, alienation?
+% tom malone
+
+% This framing of labor worked for some time in the late \nth{19} and early \nth{20} centuries,
+
+% This work is only made to seem new by framing the units of work as the output of ``task--completion APIs'',
+% making it possible
+% (and indeed advantageous)
+% to allocate work on the fly to people in different places
+% \cite{storiesIraniSilberman}.
+
+\subsection{Discretization and Routinization}
+
+% I never talk about factorization it's too soon ahhhh
+With piecework and factorization making mass manufacturing more viable,
+businesses could produce at scales that were previously unimaginable.
+Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
+One of the issues that challenged mass production at the turn of the \nth{20} century,
+as well as today in crowd work and elsewhere,
+has been that of ensuring that the quality of the work is sufficiently good.
+One of the
+
+% hey so i need to bring in citations on piecework for this stuff because I didn't exactly live through this time
+
+% is ANY of this good?
+
+% The approach to piecework described above hinged in some sense on
+% the shared understanding of the work to be done;
+% workers generally understood what was required of sewing together pieces of fabric.
+% Still, the work was speculative in that inadequately sewn clothing would not earn payment.
+% This represented a risk for pieceworkers,
+% who might spend time working on 10 garments only to be paid for 2 or 3.
+% It also represented a potential cost for management,
+% who would need to salvage the material
+% (if it was even possible),
+% and give that material to another pieceworker to do it correctly.
+
+% i need to find the source for this, because it's honestly only a vague quote in my head,
+% so... maybe i imagined it?
+
+% One of the solutions that
+
+% Broadly speaking, the solution to this problem is to make instructions clearer and easier to follow.
+% Further, by establishing a shared understanding
+% --- a ``collective conscious'' or even a corporate culture regarding work ---
+
 \textit{I want to argue that modern gig work is discretized,
 and that this parallels some of the piecework that we're historically familiar with}
+
 Work has become discretized in the last 10 years, particularly in the information work sector.
-Breakthroughs like ``Find--fix--verify'', ``Cascade'', etc\dots
+Advances such as
+``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})
+and
+``Cascade'' % (described in \citetitle{chilton2013cascade})
 make it easier to assign smaller components of work to myriad workers and
@@
 \cite{bernstein2015soylent,chilton2013cascade}.
-Some work has even turned to ``break it down'',
-inasmuch as decomposing what we might call ``macro'' tasks into microtasks
-\cite{cheng2015break}
-(\textit{there's a lot more to talk critique from this paper; sorry Justin.}).
+Some work has taken special interest in the process of breaking tasks down into more manageable parts
+% Some work has even turned to break it down'',
+inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
+both for the purposes of crowd--sourcing in the sense of out--sourcing work
+as well as ``selfsourcing''
+\cite{cheng2015break,teevan2014selfsourcing}.
 
@@
 
-\subsection{Routenization}
+% \subsection{routinization}
 The \textbf{discretization} and decomposition of work certainly facilitates
@@
 
-Routenization can be seen in the contemporary in workflows
+routinization can be seen in the contemporary in workflows
 \cite{foundry,bernstein2015soylent}
@@
 \subsection{External Management}
-\textbf{Discretization}, enabling \textbf{routenization}, thus allows us to
-evaluate work in abstraction.
-This means that we can
+\textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
+manage work more meaningfully.
+Routinization in particular made it possible
+(and indeed preferable) to assume that people
+% as \citeauthor{hu1961parallel} describes the premise of this sort of work,
+``\dots are of equal ability and \dots can do any of the $n$ jobs''
+\cite{hu1961parallel}.
+\citeauthor{storiesIraniSilberman}
+find similarly that
+this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
+% making it possible
+% (and indeed advantageous)
+% to allocate work on the fly to people in different places
+\cite{storiesIraniSilberman}.
+
+Crucially,
+\citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
+which in \citeyear{hu1961parallel}
+meant that workers would at least have to be trained in person, if not instructed as well.
+Throughout the industrial revolution, workers flocked to centralized locales of production,
+and physically situated factories made it both
+difficult to ignore the general pattern of mistreatment and
+relatively easy to organize.
+
+% construction and manufacturing in various forms
+% (from textiles to automobiles)
+% encourage the consolidation of capital--intensive resources like factories.
+% The prohibitively expensive nature of factory equipment may have been a key factor in 
+\citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
+
+% evaluate work in abstraction.
+% This means that we can
 
@@
 resisting management,
-driven by routenization,
+driven by routinization,
 enabled by discretization.
@@
 discretization,
-routenization,
+routinization,
 management,
@@
 
-\subsection{Routenization of work}
+\subsection{routinization of work}
 This leads into the next section, but I want to bring up the process of
@@
 \subsection{Taylorism}
-The routenization of work makes it possible to measure that process,
+The routinization of work makes it possible to measure that process,
 optimize it for certain characteristics, and ultimately
@@
 algorithmic management and 
-discretization \& routenization of work tasks.
+discretization \& routinization of work tasks.
 

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
discretization content

@@
 
+Distributing work in method described above may be thought of as vertically slicing work such that 
+each person is responsible for the whole task
+--- making a whole garment or being responsible for driving in a single neighborhood.
+Broken down in this way, work could grow to unprecedented scales,
+but the quality of the work would remain relatively variable.
+Textile work being a salient example,
+some seamstresses might be better than others at the same task,
+making this framing of the work initially problematic.
+
+A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
+manageable routines that could be taught relatively easily,
+and whose work output could be evaluated in abstraction from the rest of the work.
+In Ford's assembly line, this meant that workers were not responsible for building a whole car,
+but a single very narrowly defined action that needed to be done on every car.
+
+This approach paralleled what would be known as ``Taylorism'',
+and its influences can be seen today in crowd work and micro--tasks.
+\textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
+horizontal cuts instead of vertical}
+\cite{cheng2015break,writingMicroTasks}.
+% and might generally be regarded collectively as ``scientific management''
+% The compelling solution that emerged in piecework became what we now know as Taylorism;
+% rather than
+
 % I never talk about factorization it's too soon ahhhh
-With piecework and factorization making mass manufacturing more viable,
-businesses could produce at scales that were previously unimaginable.
-Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
-One of the issues that challenged mass production at the turn of the \nth{20} century,
-as well as today in crowd work and elsewhere,
-has been that of ensuring that the quality of the work is sufficiently good.
-One of the
+% With piecework and factorization making mass manufacturing more viable,
+% businesses could produce at scales that were previously unimaginable.
+% Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
+% One of the issues that challenged mass production at the turn of the \nth{20} century,
+% as well as today in crowd work and elsewhere,
+% has been that of ensuring that the quality of the work is sufficiently good.
 
@@
 
-\textit{Harken back to piecework
-and the instructions that denim workers had to make clothing;
-the instructions were clear and scoped enough that you could repeat the task ad infinitum.}
+% \textit{Harken back to piecework
+% and the instructions that denim workers had to make clothing;
+% the instructions were clear and scoped enough that you could repeat the task ad infinitum.}
 

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
making tasks discrete, giving all of the same type to a single worker makes them faster (according to lasecki et al). todo: expand on that finding!

@@
 distributed, speculative work
-% seemed a reasonable way
 has promised
@@
 frequent oil changes and the unexpected replacement of expensive parts.
-
 Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
@@
 \cite{cheng2015break,writingMicroTasks}.
+By breaking tasks into their constituent parts,
+and especially as the nature of the work could be aligned with other, similar tasks,
+workers could become ``experts'' in a small aspect of the work that they did,
+speeding their work dramatically
+\cite{delayAndOrderLasecki}.
+
+Taylorism enabled
+--- and would later become synonymous with --- 
+scientific management, which we will discuss later
 % and might generally be regarded collectively as ``scientific management''

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
cleanup

@@
 \cite{cheng2015break,writingMicroTasks}.
-By breaking tasks into their constituent parts,
-and especially as the nature of the work could be aligned with other, similar tasks,
+By decomposing and recomposing tasks,
+and in particular by assigning similarly natured work to the same workers,
+% and especially as the nature of the work could be aligned with other, similar tasks,
 workers could become ``experts'' in a small aspect of the work that they did,
@@
 \cite{delayAndOrderLasecki}.
+Perhaps more important, however, was that
+the breaking down of work into tasks has made it more practical to evaluate work at each stage
+[\textit{citation needed}]
 

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
comments

@@
 \setlength{\pdfpageheight}{\pprh}
+
+\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
 
@@
 and the historically situated ``piecework'',
+\msb{Did the research community really do this?
+Or is \textbf{your} contribution?
+I want to understand whether the literature threw around the word as a metaphor, or if they really meant it in the full historical context of the term.}
 as well as a number of observations relating to
@@
 the process of factorization that took place in the early \nth{20} century;
-indeed, these phenomena are directly related, and that they precipitate
+indeed, these phenomena are directly related \msb{how?
+what's the difference between parallel and directly related?}, and that they \msb{`they' means `the topics social computing researchers have investigated'?
+I'm not sure I understand which topics these are, or how they precipitate outcomes} precipitate
 the social outcomes we have observed,
 especially with regard to worker frustration, coordination, and resistance.
+\msb{At this point I'm not sure what the argument is.
+I see several possible arguments being made: 1) People already talked about piecework, and I'm going to talk about it more and argue its `direct' relation as opposed to `parallel' relation.
+2) People have talked about the \textit{effects} of gig work, and I'm going to tie them all together under one \textit{cause}---piecework---and show how that cause led to them historically and will now too.
+3) There are topics that social computing researchers investigated (not sure which) and they are directly related to these outcomes.
+I sort of assumed it was going to be number two, but that's not what I understood here.}
 
@@
 Crowd--powered work is not just distributed across homes,
-but dispersed among drivers, cleaners and other on--demand workers.
+but dispersed among drivers, cleaners and other on--demand workers \msb{I don't understand the distinction.
+Was Kittur et al's vision about homes somehow?
+Or are you saying that it was about information work but in fact is occuring elsewhere?}.
 By looking at the movements toward
@@
 --- and perhaps what we should expect to see going forward.
+\msb{this paragraph sounded like half review of previous literature, and then the second half switched into ``my contribution'' mode.
+Either a clearer transition or a more singular purpose would help.}
 
@@
 as patches of denim were first being delivered to the homes of early pieceworkers.
+\msb{this seems speculative---I'm not sure whether they knew what was coming.
+and wasn't income distribution different then?
+Who was the `they' in this case?
+Industrialists?}
 
 This paper will argue that the familiarity is more than passing;
-it was an inevitable milestone in the maturation of the type of work that is reemerging today.
+it \msb{what is `it'?} was an inevitable milestone in the maturation of the type of work that is reemerging today.
+\msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable.
+Why was it inevitable then, and why is it inevitable now?
+What is it about piecework that generates this outcome?
+In two sentences or less...}
 ``But'',
@@
 affect these unfolding events.
+\msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say.
+I think it needs to connect the dots more visibly for the reader.
+Is it that we're using histories and theories of piecework to reason about that progression as it applies to today?}
+\msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}
 
@@
 situating and interrogating it through ``piecework'',
-a term almost lost to history, but which for a time described
+\msb{I would not say that piecework was exactly the centerpiece of that paper. It was used as a point of contrast.}
+a term almost lost to history \msb{a bit exaggerated}, but which for a time described
 work done in the home,
@@
 not the \textit{duration})
-\cite{crowdworkFuture}.
+\cite{crowdworkFuture} \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
 This work was largely in textiles and was predominantly done by women,
-which by itself is a topic worthy of unpacking
+which by itself is a topic worthy of unpacking \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
 \cite{scott1975women}.
@@
 piecework and microtasks share some structural similarities:
+\msb{may be just me, but that felt like it was belittling the reader, like their puny brains wouldn't be able to understand more than a superficial takeaway}
 \begin{inlinelist}
-\item the work is done in the home;
+\item the work is done in the home \msb{true? isn't it still piecework if it's in a factory?};
 \item the worker is paid for each discrete piece of work done, regardless of time or effort; and
@@
 (not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy.
+is ambiguous, or at least the subject of some controversy \msb{my first reaction was ``this isn't definitional''.
+But now I look back and see that you're just saying there are similarities.
+It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
 \end{inlinelist}
 It may be said that
-on--demand work shares, if not inspiration, then many of the same outcomes as piecework.
+on--demand work shares, if not inspiration \msb{I doubt it was inspiration --- `hey, let's create something like factory floors but for information work!'}, then many of the same outcomes as piecework.
+\msb{what are outcomes?}
 
@@
 This paper attempts to make sense of the broader research on this piecework,
-or ``gig work'', by framing this as one of several steps in the marginalization of workers,
+or ``gig work'', by framing this as one of several steps in the marginalization of workers, \msb{this feels out of order to me.
+I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
 starting with the discretization of tasks,
@@
 we can both make sense of past events and perhaps reasonably predict next steps.
+\msb{I don't think we've offered much proof that this is the case yet.
+Why must they follow necessarily?
+Can you offer a cause and effect relationship?}
 
@@
 and I need to find this before I do anything productive ever again.}
+\msb{netflix prize?}
 
@@
 Moreover, by shifting the workplace into the worker's domain,
-the cost of capital involved in that work can more easily be placed on workers' backs.
+the cost of capital involved in that work can more easily be placed on workers' backs \msb{I don't think that was the original motivation for AMT or Uber --- it was about reaching people who could not otherwise join the labor force for Amazon's data cleaning or for taxi-ing.}.
 While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
@@
 is by no means new
-[\textit{citation needed --- who said this first? most notably? Turner?}].
+[\textit{citation needed --- who said this first?
+most notably?
+Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
 \textit{Elaborate on gig work, piecework, etc\dots}

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
.

--- 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
@@
 the breaking down of work into tasks has made it more practical to evaluate work at each stage
-[\textit{citation needed}]
+[\textit{citation needed}].
 
-Taylorism enabled
---- and would later become synonymous with --- 
-scientific management, which we will discuss later
+% Taylorism enabled
+% --- and would later become synonymous with --- 
+% scientific management, which we will discuss later
 % and might generally be regarded collectively as ``scientific management''
@@
 
-\textit{I want to argue that modern gig work is discretized,
-and that this parallels some of the piecework that we're historically familiar with}
+% \textit{I want to argue that modern gig work is discretized,
+% and that this parallels some of the piecework that we're historically familiar with}
 
-Work has become discretized in the last 10 years, particularly in the information work sector.
-Advances such as
+Opening the door to more discrete modules of work known as tasks 
+% Work has become discretized in the last 10 years, particularly in the information work sector.
+Contributions such as
 ``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})

### 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
fix PineGreen definition
any diffs?

@@
 \setlength{\pdfpageheight}{\pprh}
-
+\definecolor{PineGreen}{HTML}{008800}
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
@@
 % \newcommand\tabhead[1]{\small\textbf{#1}}
-\bibliography{../../../../references}
+\jbibliography{../../../../references}
 
@@
 \cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
-The research community has identified myriad connections between this emergent form of work
+The research community has made connections between this emergent form of work
 and the historically situated ``piecework'',
-\msb{Did the research community really do this?
+% \msb{Did the research community really do this?
 Or is \textbf{your} contribution?
@@
 \cite{turkopticon,dynamo,uberAlgorithm}.
-This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.
+With that being said, the connections drawn between
+historical piecework and
+contemporary crowd work
+have largely been used as a high--level tool to contextualize
+what otherwise seemed to be a difficult 
+% This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.
 

### 2016-05-23 chi/papers/2017/referrals/referrals.tex
some revisions on core argument; cleaned up referrals.tex

--- 2016-05-16 cscw/papers/2017/labor/trackingWork.tex
@@
 \begin{abstract}
-  With growing attention on gig work
-  --- ranging from the ``sharing economy'' to microtasks ---
-  scholars have made connections to frameworks like Taylorism,
-  and mechanisms such as worker advocacy and empowerment,
-  to make sense of our observations of
-  on--demand work and the workers that power this movement.
-  We argue that our the underlying trend towards ``piecework''
-  --- driven in part by the discretization,
-  routenization,
-  and external management of said work ---
-  not only suggests, but in fact generates what we have observed:
-  members of this transient workforce increasingly feeling
-  disempowered,
-  marginalized,
-  and frustrated by the systems and platforms on which they work.
+  This paper is going to be about giving workers an opportunity to refer other workers.
+  Some of the old literature on job referring suggests that
+  it's a good way for networked workers to do stuff.
+  \citeauthor{crowdcollab} found that \citetitle{crowdcollab},
+  so why not take advantage of that?
 
-  After evaluating this framing
-  through a series of case studies
-  in various industries falling broadly under the ``gig work'' category,
-  we turn our theoretical lens to look to the future,
-  to identify worthwhile questions and
-  points of inquiry that researchers in social computing should consider
-  as we attempt to anticipate and perhaps shape the future of work.
+  We'll try various incentive structures
+  (no penalty/no rejection, task approval, referral commission, and referral bonus)
 \end{abstract}
@@
 
+Hi there.
 \section{Introduction}
-This section should outline the trend of contemporary gig work in the last 10 years,
-in particular tracking the changes since
-\citeauthor{crowdworkFuture} discussed
-\citetitle{crowdworkFuture}
-in \citeyear{crowdworkFuture}
-\cite{crowdworkFuture}.
-This would review the literature since then,
-then point to the movements toward worker unionization, advocacy, etc\dots
-that have happened in the last year or two.
+The practice of referring work to other skilled workers has been well--studied in economics and more broadly in the social sciences for decades
+[\textit{citation needed}].
+But researchers in crowd work haven't sufficiently explored the potential
+to employ this characteristic of socially connected workers
+to give professionals opportunities to better route tasks.
+One possible reason for this may have been the assumption that workers are not necessarily well--connected.
+This would be an intuitively reasonable assumption
+--- with no in--built forums or communities for workers on various gig labor markets
+(e.g. Uber, Amazon Mechanical Turk (AMT), etc\dots),
+one might reasonably conclude that these workers are not collaborative.
 
-In the reflection on the literature published in the last 5 years since
-\citetitle{crowdworkFuture},
-we notice a broader change in the form of work that's being done;
-motivated by this realization,
-we look at the emergent phenomena in piecework both as interconnected.
-By looking at the movements toward
-discretization,
-routenization,
-and management
-as necessarily linked --- even causally --- to one another, 
-we give ourselves a framing of contemporary piecework that explains
-and arguably predicts
-what we have seen thus far.
-
-\subsection{Piecework as a lens to understand gig work}
-It's not new that researchers have discussed gig work as an extension or resurgence
-of piecework, but it is sufficiently important for the framing of this argument that
-it merits a more direct treatment and some explanation here.
-\citeauthor{crowdworkFuture}
-investigate the future of crowd work by
-situating and interrogating it through ``piecework'',
-a term almost lost to history, but which for a time described
-work done in the home,
-in manageable tasks,
-often involving clear instructions
-and payment only for work completed, not work done
-(the differentiation, here, being that
-one would be paid for the \textit{output} of the work,
-not the \textit{duration})
-\cite{crowdworkFuture}.
-This work was largely in textiles and was predominantly done by women,
-which by itself is a topic worthy of unpacking
-\cite{scott1975women}.
-Given the scope, however,
-we will have to leave this complex topic with the superficial takeaway that
-piecework and microtasks share some structural similarities:
-\begin{inlinelist}
-\item the work is done in the home;
-\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
-\item the worker's status
-(not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy.
-\end{inlinelist}
-
-% The work involved often meant sewing together pieces of denim,
-% was largely done by women, and turned
-
-Crowd work,
-they argued in \citeyear{crowdworkFuture},
-represents a threat to the future of work inasmuch as it marginalized
-and perhaps even harmed
-workers;
-their driving question, then, was
-whether the future of
-(crowd, but arguably all discretized)
-work might be one in which they wish their children to participate someday;
-it seems strikingly like one that might have been asked
-as patches of denim were first being delivered to the homes of early pieceworkers.
-
-This paper will argue that the similarity is more than striking;
-it was an inevitable milestone in the maturation of the type of work that is reemerging today.
-\textit{How to continue???}
-
-% we engaged in ethnographic fieldwork
-% to attempt to understand why gig work appeals to those that engage in it,
-% and to gain a better understanding of how they relate to the marketplaces in which they work.
-
-% \textit{I'd also like to bring up piecework here,
-% and briefly make the case that
-% the kind of work we're seeing should be thought of as a continuation
-% or resurgence
-% of piecework, with some important distinctions that differentiate it.}
-
-\subsection{Context}
-Importantly, since the reemergence of piecework we've seen substantial frustration
-and resistance among the workers in this area
-\cite{uberAlgorithm,turkopticon,dynamo}.
-This paper attempts to make sense of the broader research on this piecework,
-or ``gig work'', by framing this as one of several steps in the marginalization of workers,
-starting with the discretization of tasks,
-followed by routenization and the rise of workflows,
-and finally the external management of workers.
-All of this is to say that these milestones follow sequentially,
-not coincidentally but necessarily,
-and that by tracing this path using the corpus of scholarship on labor and workers
-we can both make sense of past events and perhaps reasonably predict next steps.
-
-We explore each topic
---- discretization, routenization, \& management ---
-by looking at case studies in social computing.
-Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
-we turn to look ahead, envisioning future areas that researchers in social computing
---- and particularly digitally mediated work ---
-should explore.
-
-\section{Case Studies}
-The past decade of research has shed light on crowd work from various perspectives,
-and revealed 
-\subsection{Discretization}
-\textit{I want to argue that modern gig work is discretized,
-and that this parallels some of the piecework that we're historically familiar with}
-Work has become discretized in the last 10 years, particularly in the information work sector.
-Breakthroughs like ``Find--fix--verify'', ``Cascade'', etc\dots
-make it easier to assign smaller components of work to myriad workers and
-recompose the constituent parts into something more complex
-\cite{bernstein2015soylent,chilton2013cascade}.
-Some work has even turned to ``break it down'',
-inasmuch as decomposing what we might call ``macro'' tasks into microtasks
-\cite{cheng2015break}
-(\textit{there's a lot more to talk critique from this paper; sorry Justin.}).
-
-The parallels between this kind of work and
-historical piecework of the turn of the \nth{20} century
-is by no means new
-[\textit{citation needed --- who said this first? most notably? Turner?}].
-\textit{Elaborate on gig work, piecework, etc\dots}
-
-\subsection{Routenization}
-The \textbf{discretization} and decomposition of work certainly facilitates
-breaking tasks down,
-parallelizing work,
-and getting more done across a broader array of people;
-but the key advantage in turning macro tasks 
-into many micro tasks is that we can make that work \textit{routine}.
-
-By making larger bodies of work less contextually situated,
-we can define processes that make that work
-and the instructions thereof
-useful for virtually anyone with shared cultural or intellectual background.
-
-Routenization can be seen in the contemporary in workflows
-[\textit{There are various papers that talk about creating a workflow;
-maybe this is where FFV belongs? Foundry?}].
-
-\textit{Harken back to piecework
-and the instructions that denim workers had to make clothing;
-the instructions were clear and scoped enough that you could repeat the task ad infinitum.}
-
-\subsection{External Management}
-\textbf{Discretization}, enabling \textbf{routenization}, thus allows us to
-evaluate work in abstraction.
-This means that we can
-
-\textit{I want to cite Justin's work on crowdsourcing effort to highlight that
-we can evaluate workers for the work they do with increasingly finite amounts of time;
-also, Ranjay's work that embraces failure because
-we can essentially algorithmically make sense of failure (let's call it noise)
-and re--capture signal.}
-
-\subsection{Resistance}
-So it's not surprising, then, that workers are frustrated:
-Turkopticon, Dynamo, and many other pieces of work point this out
-\cite{turkopticon,dynamo,uberAlgorithm}
-[\textit{also want to cite Brian's work from CHI this year}].
-In popular culture, reporters have inquired about
-what ``the future of labor unions'' will look like
-[\textit{Imagine I had cited an online article and
-it wasn't totally inappropriate for a conference}].
-
-\textit{The parallels here would be difficult to ignore;
-Lily and others have pointed this out,
-but it's worth thinking about the broader trend of
-resisting management,
-driven by routenization,
-enabled by discretization.
-Do that here.}
-
-\subsection{Looking forward}
-If we agree that
-discretization,
-routenization,
-management,
-and even resistance
-necessarily follow one another according to this theoretical lens,
-then we have to use it to attempt to envision
-what comes next.
-
-\textit{I've been going back and forth regarding how to frame this paper;
-the approach that tries to look at
-things that are similar vs things that are different
-doesn't seem to work, but
-I've left the thoughts here because it's not all completely bad.}
-
-\section{Things stay the same}
-How is gig work the same as it's been historically?
-
-I think this section would be compelling to draw parallels between
-the narratives drivers gave about the flexibility, autonomy, etc\dots
-and that which we might have seen among pieceworkers
-(predominantly women, who benefited from being able to work from home).
-
-\subsection{Flexibility}
-Are there cultural differences between the people that did piecework and
-the people that do gig work now?
-I'm not sure there are significant differences that have affected the outcomes so far.
-
-Many of the workers to whom piecework appealed were mothers, wives, etc\dots
-who mostly stayed at home for various reasons
-(certainly largely it was cultural
---- women weren't afforded equal access to labor opportunities,
-making in--home job opportunities not only compelling, but also one of few available options).
-
-Gig workers are in some senses similarly constrained:
-workers on Amazon Mechanical Turk
---- those that use it as a primary source of income, at least ---
-report being homebound for various reasons (e.g. medical, parenting, etc\dots).
-Society and circumstance have made it difficult
-or impossible
-to join the contemporary, conventional workforce;
-gig work re--opens that door.
-
-But there are differences; during our research over the summer,
-we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
-Many of them told us about their home lives
---- about children, spouses, and other commitments ---
-to which they wanted to dedicate more time.
-One driver
-(let's call him Ra\'{u}l)
-told me about how he drove for Lyft
-after working as an inventory manager at a hospital for more than nine years.
-
-I asked him why he quit that job and forewent
-the benefits,
-predictability,
-and career growth opportunity
-that his old job offered.
-He told me that when his daughter was born,
-he was overcome with a desire to spend more time with her.
-No longer satisfied with work
-where he often left before his daughter would wake up and return after she fell asleep,
-Ra\'{u}l decided to start driving for Lyft,
-because he could drive in the evenings when his daughter was asleep.
-
-Other drivers reported similar benefits;
-gig work affords its workers flexibility that conventional careers don't allow.
-
-\subsection{Routenization of work}
-This leads into the next section, but I want to bring up the process of
-making work about mass--manufacturing, at least inasmuch as
-the instructions are the same for everyone.
-
-The Internet arguably has made it much easier to broadcast those instructions,
-but it's had this deeper effect of enabling some amount of back--and--forth
-between the worker and the (algorithmic) manager.
-
-\subsection{Taylorism}
-The routenization of work makes it possible to measure that process,
-optimize it for certain characteristics, and ultimately
-lead to Taylorism and scientific management.
-This is not new;
-researchers have studied and written about the slow creep of
-algorithmic management and 
-discretization \& routenization of work tasks.
-
-\textit{We hope to take a step back from the context in which this work is often applied, and
-look for its place in the larger trends and theories
-to make sense of the trends of gig work at large.}
-
-Industrialization and the automobile assembly line makes this famous,
-but piecework functioned on the principle that everyone was making similar
-or identical
-garments and other products.
-
-Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
-sometimes bootstrapped,
-as in Ranjay's talk on 
-``Embracing Error to Enable Rapid Crowdsourcing'',
-but more conventionally in work flows like ``Find--Fix--Verify''.
-
-\textit{This might be an opportunity to reflect on
-how pieceworkers internalized the work they were doing,
-responded to the stress of the uncertainty of potentially rejected work,
-etc\dots but I'm not familiar with research in that space.}
-
-\subsection{The emergence of decentralized workplaces}
-The practice of in--home piecework was consumed by
-the centralization of factories
-(the effect of which we'll talk about later, since
-we can talk about how this made unions more practical),
-but for a time many pieceworkers
-at the turn of the \nth{20} century
-worked out of their homes.
-Strikingly, many of the gig workers
-\citeauthor{crowdworkFuture}
-discussed in
-\citeyear{crowdworkFuture},
-and indeed many more continue to do what we might call ``information work''
---- that is, work that predominantly demands human computation ---
-but increasingly we're seeing the movement toward transient work that largely requires embodied presence
-\cite{uberAlgorithm} (and others).
-
-\subsection{What was this about?}
-I was going to check in quickly and take a photo of your whiteboard
-(hoping that you'd kept our conversation notes around)
-but your room was \textit{literally} full of people lol.
-
-\section{Things are changing}
-The medium on which this work is being done
---- and to an extent the medium used to manage workers ---
-has dramatically changed things as well, however;
-workers are distributed around the world,
-working out of their cars in the cases of livery services
-(notably, never returning to a base of operations)
-across and between cities as well as nations,
-or in their homes
-(paralleling the trend of piecework even more closely).
-
-Trying to understand how gig work has differed from piecework should at least
-start with looking at the different characteristics of the work involved.
-After that, we should think about how
-the demography and culture of the people engaging in this kind of work
-have changed versus that of the pieceworkers.
-
-\subsection{Differences in the work itself}
-Gig work has all of the above similarities with piecework,
-but there are key differences.
-
-Piecework emerged at the turn of the \nth{20} century,
-right at the time that telecommunications began to boom in the United States. % too far a reach?
-\textit{Did telecom enable remote management?}
-
-Gig work in its contemporary formation is largely mediated by
-ubiquitously accessible digital media
-(the Internet, telephony, etc\dots)
-and importantly has relied on this technology to facilitate the remote management of workers
-\cite{uberAlgorithm}.
-
-\textit{Has the work fundamentally changed,
-or are we just being managed remotely in different ways?
-I'm not entirely sure.}
-
-How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
-How does this work differ from the work in which truckers, taxi drivers, and other
-independent contractors have been participating for decades?
-We argue that the substantive difference in these markets is the speed of the market itself,
-motivated by the technology which mediates it.
-Because workers can be sourced and dispatched virtually instantly,
-businesses that engage in this kind of work
-(e.g. Uber, Amazon Mechanical Turk, etc\dots)
-have taken to removing other bottlenecks,
-like vetting workers upfront
-
-\section{More deliberate work}
-% \subsection{Differences in the people that do the work}
-% ???
+But, as
+\citeauthor{crowdcollab} identifies,
+crowd workers are substantially more collaborative than we originally may have inferred
+\cite{crowdcollab}.
+Given this realization, we pose the question:
+can requesters operationalize this finding to get work done more effectively than conventional methods would allow?
+Perhaps more importantly, can employing this finding allow workers
+to contribute meaningfully to tasks
+in ways that do not necessarily fall under the existing paradigms
+(e.g. ``find--fix--verify'')
+\cite{bernstein2015soylent}?
 

### 2016-05-23 cscw/papers/2017/labor/trackingWork.tex
some revisions on core argument; cleaned up referrals.tex

--- 2016-05-20 cscw/papers/2017/labor/trackingWork.tex
@@
 \begin{document}
-\balance{}
 \title{\plaintitle}
@@
 \cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
-The research community has identified myriad connections between this emergent form of work
+The research community has made connections between this emergent form of work
 and the historically situated ``piecework'',
-\msb{Did the research community really do this?
-Or is \textbf{your} contribution?
-I want to understand whether the literature threw around the word as a metaphor, or if they really meant it in the full historical context of the term.}
 as well as a number of observations relating to
@@
 \cite{turkopticon,dynamo,uberAlgorithm}.
-This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.
+But the connections between
+historical piecework and
+contemporary crowd work
+haven't been deeply made as yet.
+% um, note to self: say more.
 
@@
 the process of factorization that took place in the early \nth{20} century;
-indeed, these phenomena are directly related \msb{how?
-what's the difference between parallel and directly related?}, and that they \msb{`they' means `the topics social computing researchers have investigated'?
-I'm not sure I understand which topics these are, or how they precipitate outcomes} precipitate
-the social outcomes we have observed,
-especially with regard to worker frustration, coordination, and resistance.
-\msb{At this point I'm not sure what the argument is.
-I see several possible arguments being made: 1) People already talked about piecework, and I'm going to talk about it more and argue its `direct' relation as opposed to `parallel' relation.
-2) People have talked about the \textit{effects} of gig work, and I'm going to tie them all together under one \textit{cause}---piecework---and show how that cause led to them historically and will now too.
-3) There are topics that social computing researchers investigated (not sure which) and they are directly related to these outcomes.
-I sort of assumed it was going to be number two, but that's not what I understood here.}
-
-% This section should outline the trend of contemporary gig work in the last 10 years,
-% in particular tracking the changes since
-% \citeauthor{crowdworkFuture} discussed
-% \citetitle{crowdworkFuture}
-% in \citeyear{crowdworkFuture}
-% \cite{crowdworkFuture},.
-% In particular, we point to \citeauthor{crowdworkFuture}'s discussion of worthwhile
-% questions to ask regarding \citetitle{crowdworkFuture}
-% \cite{crowdworkFuture}.
-% This would review the literature since then,
-% then point to the movements toward worker unionization, advocacy, etc\dots
-% that have happened in the last year or two.
+indeed, these phenomena precipitate and reinforce each other.
 
@@
 \cite{crowdworkFuture}.
-Crowd--powered work is not just distributed across homes,
-but dispersed among drivers, cleaners and other on--demand workers \msb{I don't understand the distinction.
-Was Kittur et al's vision about homes somehow?
-Or are you saying that it was about information work but in fact is occuring elsewhere?}.
-By looking at the movements toward
+Following the improvements in
+breaking down tasks,
+delegating work, and
+managing workers more broadly,
+Crowd--powered has continued to parallel historical piecework's trajectory
+by outgrowing ad--hoc worker groups and coalescing into more formal working groups;
+we might call this the beginning of factorization.
+% "Factorization" is not the right word but I honestly have no idea what the right term is.
+% The process of formalizing work into factories? Should I tweet/post/ask about this somewhere?
+
+% Crowd work,
+% they argued in \citeyear{crowdworkFuture},
+% represents a threat to the future of work inasmuch as it marginalized
+% and perhaps even harmed
+% workers;
+% their driving question, then, was
+% whether the future of
+% (crowd, but arguably all discretized)
+% work might be one in which they wish their children to participate someday;
+% it seems strikingly like one that might have been asked
+% as patches of denim were first being delivered to the homes of early pieceworkers.
+% % \msb{this seems speculative---I'm not sure whether they knew what was coming. and wasn't income distribution different then? Who was the `they' in this case? Industrialists?}
+% agreed; i'm commenting it out for now to see how I feel without making this argument at all. If I really feel like it's missing, I should find evidence that contemporaries discussed it...
+
+Making sense of this 
+by looking at the movements toward
 distributing work,
@@
 and externally managing workers
-% discretization,
-% routinization,
-% and management
 as linked to one another, 
@@
 --- and perhaps what we should expect to see going forward.
-\msb{this paragraph sounded like half review of previous literature, and then the second half switched into ``my contribution'' mode.
-Either a clearer transition or a more singular purpose would help.}
 
-% The work involved often meant sewing together pieces of denim,
-% was largely done by women, and turned
+This paper will attempt to demonstrate that the familiarity is more than passing;
+that the phenomena researchers have observed in on--demand digitally mediated labor markets
+were inevitable milestones in the birth and life of piecework.
+We will set out to show that piecework itself
+--- and the features inherent to it (such as decomposition of tasks) --- % and others?
+precipitated the negative outcomes that historically led workers to collective organization and action in the early \nth{20} century;
+that aspects of piecework like 
+% Indeed, the model of work distribution and management we know as piecework
+% --- including the methods and work--flows that facilitate this kind of work ---
+% ultimately cause what we recognize broadly as
+% frustration,
+% alienation, and
+% resistance
+% among gig workers.
 
-Crowd work,
-they argued in \citeyear{crowdworkFuture},
-represents a threat to the future of work inasmuch as it marginalized
-and perhaps even harmed
-workers;
-their driving question, then, was
-whether the future of
-(crowd, but arguably all discretized)
-work might be one in which they wish their children to participate someday;
-it seems strikingly like one that might have been asked
-as patches of denim were first being delivered to the homes of early pieceworkers.
-\msb{this seems speculative---I'm not sure whether they knew what was coming.
-and wasn't income distribution different then?
-Who was the `they' in this case?
-Industrialists?}
-
-This paper will argue that the familiarity is more than passing;
-it \msb{what is `it'?} was an inevitable milestone in the maturation of the type of work that is reemerging today.
-\msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable.
+% \msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable.
 Why was it inevitable then, and why is it inevitable now?
@@
 \cite{scholz2012digital}.
-Indeed, we will attempt to reason about more than the larger--scale progression from
-the discretization of work toward scientific management and future steps;
-we will attempt to make sense of how the modern processes in digital labor,
-in particular the use of the Internet as a medium to manage as well as \textit{do} work,
-affect these unfolding events.
-\msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say.
+For one, information work
+(e.g. AMT \& Upwork)
+and digitally managed work in general make it substantially easier
+to keep workers geographically dispersed
+compared to the factory workers that followed home--based piecework.
+This aspect of work renders workers virtually invisible to the task solicitors
+--- or ``requesters'' ---
+despite being significantly more connected than historical pieceworkers were % this is a weak point...
+\cite{turkopticon}.
+Secondly, while
+factory foremen and other middle--men
+traditionally mediated the relationships between workers and managers,
+today the visible agents are \textit{systems};
+rather than employing individuals who ``personified the functions of management''
+and can thus negotiate workers' needs,
+socio--technical systems mediate these interactions % often with unrewarding outcomes
+\cite{wray1949marginal}
+% \cite{storiesIraniSilberman}.
+Finally, this characteristic of modern piecework makes
+worker coordination for collective advocacy and action significantly more difficult,
+necessitating special consideration to maintain the inertia of collective will
+while focusing that energy productively
+\cite{dynamo}.
+% here are undoubtedly more examples of differences.
+% How do I not stake a claim that this is comprehensive?
+% Or should I really flesh this out and make it (comprehensive)?
+
+% Indeed, we will attempt to reason about more than the larger--scale progression from
+% the discretization of work toward scientific management and future steps;
+% we will attempt to make sense of how the modern processes in digital labor,
+% in particular the use of the Internet as a medium to manage as well as \textit{do} work,
+% affect these unfolding events.
+% \msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say.
 I think it needs to connect the dots more visibly for the reader.
 Is it that we're using histories and theories of piecework to reason about that progression as it applies to today?}
-\msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}
-
-% we engaged in ethnographic fieldwork
-% to attempt to understand why gig work appeals to those that engage in it,
-% and to gain a better understanding of how they relate to the marketplaces in which they work.
-
-% \textit{I'd also like to bring up piecework here,
-% and briefly make the case that
-% the kind of work we're seeing should be thought of as a continuation
-% or resurgence
-% of piecework, with some important distinctions that differentiate it.}
+% \msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}
 
 \subsection{Piecework as a lens to understand gig work}
-% It's not new to discuss gig work as an extension or even a resurgence of piecework
-% \cite{crowdworkFuture,fixingChaos}.
-% Nevertheless,
-% it's sufficiently core to the argument here that a review is worthwhile.
 \citeauthor{crowdworkFuture}
@@
 
-\textit{There's a reference to something that's \textbf{not} 99designs,
-that describes a call for scientific solutions/designs to solve some problem
-that the requesters have.
-Honestly it's killing me that I can't remember the name,
-and I need to find this before I do anything productive ever again.}
-\msb{netflix prize?}
+% \textit{There's a reference to something that's \textbf{not} 99designs,
+% that describes a call for scientific solutions/designs to solve some problem
+% that the requesters have.
+% Honestly it's killing me that I can't remember the name,
+% and I need to find this before I do anything productive ever again.}
+% \msb{netflix prize?}
+% that's an interesting example in the same vein as what i'm thinking about, but this is a whole business of it
+% \textbf{INNOCENTIVE} oh my gosh it just clicked my life is complete right now at 9:54am.
+% I don't even need to submit anymore.
+% I'm satisfied having pulled this strand of a memory out of my brain.
 
 The HCI community is perhaps most familiar with examples such as
-99designs,
+Innocentive,
 Uber, Lyft,
@@
 with relative ease
-\cite{DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
+\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
 This insight, that
-workers can be geographically distributed,
+workers can be geographically distributed and tasks decomposed,
 % (especially given careful central management of those workers, discussed later),
 has proven remarkably compelling
-\cite{sensitiveTasks}.
+and an effective fulcrum for leveraging the Internet for highly scalable work
+\cite{sensitiveTasks,embracingErrorKrishna}.
 
@@
 
+\balance{}
 \printbibliography

### 2016-05-25 cscw/papers/2017/labor/trackingWork.tex
first section seems mostly done; some stuff still commented out needs to be brought back in (click here to read more: the second bullet point will SHOCK you! - section 2 is getting fleshed out, and some citations are getting thrown in, but i also have this paper from ~1963 that's eerily prescient about digitally mediated gig work - section 3 is not really done, but we know we're going to make this argument about factories emerging based on the practice of establishing worker pools (something AMT affords for by letting you make hidden HITs, for example), so i think we can flesh that out soon...

@@
 \definecolor{PineGreen}{HTML}{008800}
+\definecolor{BrickRed}{HTML}{FF0000}
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
+\newcommand{\ali}[1]{{\color{BrickRed}[\textit{al2: #1}]}}
 
@@
 \section{Introduction}
-
-% \itshape
-% What am I trying to say here?
-% I want to get across that researchers have looked at gig work in various ways,
-% but not come up with very persuasive overarching explanations for this whole process.
-% That is, there's no broader theory making sense of why all of this is happening.
-
-% Rather than make the argument that
-% these things are all independently similar to the piecework
-% (something to which researchers have already made the connection),
-% I want to frame this as a huge process, involving:
-% \begin{enumerate}
-%   \item distribution of work,
-%   \textbf{which leads to\dots}
-%   \item discretization and routinization of work,
-%   \textbf{which allows\dots}
-%   \item management (scientific \& external),
-%   \textbf{which causes workers to feel\dots}
-%   \item alienation from work (Marx's idea) and resistance to management (also Marx probably but a more general labor advocacy theory too)
-% \end{enumerate}
-% okay, let's try this:
-% \upshape
-
 The past decade has seen
@@
 we notice a broader trend describing the change in work that's being done
-% motivated by this realization,
-% we look at the emergent phenomena in piecework both as interconnected
 \cite{crowdworkFuture}.
@@
 % as patches of denim were first being delivered to the homes of early pieceworkers.
-% % \msb{this seems speculative---I'm not sure whether they knew what was coming. and wasn't income distribution different then? Who was the `they' in this case? Industrialists?}
 % agreed; i'm commenting it out for now to see how I feel without making this argument at all. If I really feel like it's missing, I should find evidence that contemporaries discussed it...
@@
 were inevitable milestones in the birth and life of piecework.
-We will set out to show that piecework itself
---- and the features inherent to it (such as decomposition of tasks) --- % and others?
-precipitated the negative outcomes that historically led workers to collective organization and action in the early \nth{20} century;
-that aspects of piecework like 
-% Indeed, the model of work distribution and management we know as piecework
-% --- including the methods and work--flows that facilitate this kind of work ---
-% ultimately cause what we recognize broadly as
-% frustration,
-% alienation, and
-% resistance
-% among gig workers.
+We will set out to show that this perspective of crowd work as an embodiment of piecework
+predicts the myriad outcomes of contemporary crowd work, including
+our developments of work--flows further abstracting work,
+the troubling effects of those developments on factors such as pay and work quality, and
+many if not all of the frustrations researchers have discovered among workers
+in the study of crowd and gig work.
 
-% \msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable.
-Why was it inevitable then, and why is it inevitable now?
-What is it about piecework that generates this outcome?
-In two sentences or less...}
 ``But'',
@@
 socio--technical systems mediate these interactions % often with unrewarding outcomes
-\cite{wray1949marginal}
+\cite{wray1949marginal}.
 % \cite{storiesIraniSilberman}.
@@
 \cite{dynamo}.
+\ali{I really don't like this paragraph\dots}
 % here are undoubtedly more examples of differences.
@@
 % Or should I really flesh this out and make it (comprehensive)?
-
-% Indeed, we will attempt to reason about more than the larger--scale progression from
-% the discretization of work toward scientific management and future steps;
-% we will attempt to make sense of how the modern processes in digital labor,
-% in particular the use of the Internet as a medium to manage as well as \textit{do} work,
-% affect these unfolding events.
-% \msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say.
-I think it needs to connect the dots more visibly for the reader.
-Is it that we're using histories and theories of piecework to reason about that progression as it applies to today?}
-% \msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}
 
@@
 situating and interrogating it through ``piecework'',
-\msb{I would not say that piecework was exactly the centerpiece of that paper. It was used as a point of contrast.}
-a term almost lost to history \msb{a bit exaggerated}, but which for a time described
-work done in the home,
+which historically described work done in the home,
 in manageable tasks,
@@
 not the \textit{duration})
-\cite{crowdworkFuture} \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
-This work was largely in textiles and was predominantly done by women,
-which by itself is a topic worthy of unpacking \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
-\cite{scott1975women}.
+\cite{crowdworkFuture}.
+% \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
+% This work was largely in textiles and was predominantly done by women,
+% which by itself is a topic worthy of unpacking
+% \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
+% \ali{let's just leave this in the graveyard for now; i feel like someone who knows what they're talking about with regard to feminist and gender studies
+% (e.g. the Bardzells, Irani) would really do justice here, but i'd be getting in over my head}
+% \cite{scott1975women}.
 Given the scope,
-we'll leave this topic with the superficial takeaway that
-piecework and microtasks share some structural similarities:
-\msb{may be just me, but that felt like it was belittling the reader, like their puny brains wouldn't be able to understand more than a superficial takeaway}
+we can frame piecework and gig work as sharing these important similarities:
 \begin{inlinelist}
-\item the work is done in the home \msb{true? isn't it still piecework if it's in a factory?};
+\item this form of work began in the home
 \item the worker is paid for each discrete piece of work done, regardless of time or effort; and
@@
 (not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy \msb{my first reaction was ``this isn't definitional''.
+is ambiguous, or at least the subject of some controversy.
+% \msb{my first reaction was ``this isn't definitional''.
 But now I look back and see that you're just saying there are similarities.
 It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
-\end{inlinelist}
-It may be said that
-on--demand work shares, if not inspiration \msb{I doubt it was inspiration --- `hey, let's create something like factory floors but for information work!'}, then many of the same outcomes as piecework.
-\msb{what are outcomes?}
+\end{inlinelist}.
 
@@
 This paper attempts to make sense of the broader research on this piecework,
-or ``gig work'', by framing this as one of several steps in the marginalization of workers, \msb{this feels out of order to me.
-I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
-starting with the discretization of tasks,
-followed by routinization and the rise of workflows,
-and finally the external management of workers.
+or ``gig work'', by framing this as one of several steps in the marginalization of workers,
+% \msb{this feels out of order to me. I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
+%
+% this needs to be re-written to reflect the new split
+%
+%
+% starting with the discretization of tasks,
+% followed by routinization and the rise of work--flows,
+% and finally the external management of workers.
+% 
 All of this is to say that these milestones follow sequentially,
@@
 we can both make sense of past events and perhaps reasonably predict next steps.
-\msb{I don't think we've offered much proof that this is the case yet.
-Why must they follow necessarily?
-Can you offer a cause and effect relationship?}
+% \msb{I don't think we've offered much proof that this is the case yet. Why must they follow necessarily? Can you offer a cause and effect relationship?}
 
@@
 through our framing, are clearly situated together.
-Those topics, at a very high level, are
+Those topics are, at a high level, as follows:
 \begin{enumerate}
-\item the \textbf{distribution} of work,
-\item the \textbf{routinization} and \textbf{discretization} of work, and finally
-\item the \textbf{relationships} between the workers and the product of their work,
-      as well as workers and managers.
+\item the processes involved in making work into tasks, or \textbf{discretization},
+in ``\nameref{sec:MakingGigsWork}'';
+\item the outcomes of that discretization, both on the work itself as well as the workers,
+in ``\nameref{sec:Fallout}'';
+and finally
+\item the results of these forces; namely the turning of work toward \textbf{factories},
+which we will discuss in ``\dots''.
 \end{enumerate}
@@
 
-\subsection{Distributing Work}
+\subsection{Making Gig(s) Work}\label{sec:MakingGigsWork}
+% this is phase 1: the decomposition & abstraction of work;
+% all of the research that has gone into work--flows and
+% all the things we can do with crowd work.
+
+In this section, we'll consider several components of contemporary piecework
+that relate to its historical parallel.
+In particular, we'll look at the ``decomposition'' of complex skill--intensive tasks
+into smaller, more abstracted work, as well as the routinization that this allows.
+
 In both the historical and contemporary cases,
@@
 
-The insight here seems to have been that
-a substantial amount of work does not necessarily need to be done in collocated space,
-but can be done from one's own home or other space arranged by the worker.
-Moreover, by shifting the workplace into the worker's domain,
-the cost of capital involved in that work can more easily be placed on workers' backs \msb{I don't think that was the original motivation for AMT or Uber --- it was about reaching people who could not otherwise join the labor force for Amazon's data cleaning or for taxi-ing.}.
+% The insight here seems to have been that
+% a substantial amount of work does not necessarily need to be done in collocated space,
+% but can be done from one's own home or other space arranged by the worker.
+% Moreover, by shifting the workplace into the worker's domain,
+% the cost of capital involved in that work can more easily be placed on workers' backs
+% \msb{I don't think that was the original motivation for AMT or Uber --- it was about reaching people who could not otherwise join the labor force for Amazon's data cleaning or for taxi-ing.}.
+% I didn't mean that this was some insight that Uber or Lyft had, but that this was some part of what makes Uber work;
+% i would argue categorically that these companies wouldn't functionally work if they weren't able to obviate the 
+% maintenance costs that a factory owner would associate with their capital (e.g. auto costs)
 While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
 (inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
-these new systems go further,
+these new systems go further
 by framing ownership and responsibility for costs
@@
 
-% the change seemed to begin with
-% the redistribution of work into the homes and onto the diffuse capital of workers.
-% 99designs may represent one of the first illustrative cases of
-% speculative, distributed labor in the contemporary form,
-% but speculative labor and even on--demand design work is not necessarily new;
-% \cite{99designs}.
-
-% distributed; uber, etc... ->
-% routinization; crowdwork - > optimization
-% - > accommodation and resistance, alienation?
-% tom malone
-
-% This framing of labor worked for some time in the late \nth{19} and early \nth{20} centuries,
-
-% This work is only made to seem new by framing the units of work as the output of ``task--completion APIs'',
-% making it possible
-% (and indeed advantageous)
-% to allocate work on the fly to people in different places
-% \cite{storiesIraniSilberman}.
-
-\subsection{Discretization and Routinization}
+% \subsection{Discretization and Routinization}
 
@@
 historical piecework of the turn of the \nth{20} century
-is by no means new
-[\textit{citation needed --- who said this first?
+is by no means new.
+% [\textit{citation needed --- who said this first?
 most notably?
 Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
-\textit{Elaborate on gig work, piecework, etc\dots}
+% \textit{Elaborate on gig work, piecework, etc\dots}
 
@@
 
-routinization can be seen in the contemporary in workflows
-\cite{foundry,bernstein2015soylent}
-[\textit{There are various papers that talk about creating a workflow;
-maybe this is where FFV belongs? Foundry?}].
+routinization can be seen in the contemporary in work--flows
+\cite{foundry,bernstein2015soylent}.
+% [\textit{There are various papers that talk about creating a work--flow;
+% maybe this is where FFV belongs? Foundry?}].
 
@@
 
-\subsection{External Management}
-\textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
-manage work more meaningfully.
-Routinization in particular made it possible
-(and indeed preferable) to assume that people
-% as \citeauthor{hu1961parallel} describes the premise of this sort of work,
-``\dots are of equal ability and \dots can do any of the $n$ jobs''
-\cite{hu1961parallel}.
-\citeauthor{storiesIraniSilberman}
-find similarly that
-this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
-% making it possible
-% (and indeed advantageous)
-% to allocate work on the fly to people in different places
-\cite{storiesIraniSilberman}.
+\subsection{The Fallout of Gig Work}\label{sec:Fallout}
+If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
+the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
+\citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
+(see also \citeauthor{dynamo}'s work
+continuing in the spirit of this observation to generate collective action to improve worker conditions)
+\cite{turkopticon,dynamo}.
+\citeauthor{uberAlgorithm}
+find similarly that workers on gig work platforms are frustrated by the systems on which they work,
+to say little of the policies which these systems enforce
+\cite{uberAlgorithm}.
 
-Crucially,
-\citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
-which in \citeyear{hu1961parallel}
-meant that workers would at least have to be trained in person, if not instructed as well.
-Throughout the industrial revolution, workers flocked to centralized locales of production,
-and physically situated factories made it both
-difficult to ignore the general pattern of mistreatment and
-relatively easy to organize.
+\itshape
+sections here should include:
+\begin{enumerate}
+\item ``flexibility'' \cite{friedman2014workers,DesigningForLabourUber}
+\item leads to ``low pay'' \cite{turkopticon,dynamo}
+\item leads to ``variable quality work''
+\end{enumerate}
 
-% construction and manufacturing in various forms
-% (from textiles to automobiles)
-% encourage the consolidation of capital--intensive resources like factories.
-% The prohibitively expensive nature of factory equipment may have been a key factor in 
-\citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
+\upshape
 
-% evaluate work in abstraction.
-% This means that we can
+\subsection{Factories}
 
-\textit{I want to cite Justin's work on crowdsourcing effort to highlight that
-we can evaluate workers for the work they do with increasingly finite amounts of time;
-also, Ranjay's work that embraces failure because
-we can essentially algorithmically make sense of failure (let's call it noise)
-and re--capture signal.}
+% I commented everything going forward out not because it's all trash (it's not!)
+% but because we pretty significantly reframed the cases to be studying and
+% i didn't want the structure from before to influence the rewrite;
+% we should be bringing content from below back above. just, judiciously...
 
-\subsection{Resistance}
-So it's not surprising, then, that workers are frustrated:
-Turkopticon, Dynamo, and many other pieces of work point this out
-\cite{turkopticon,dynamo,uberAlgorithm}
-[\textit{also want to cite Brian's work from CHI this year}].
-In popular culture, reporters have inquired about
-what ``the future of labor unions'' will look like
-[\textit{Imagine I had cited an online article and
-it wasn't totally inappropriate for a conference}].
+% \subsection{Discretization and Routinization}
 
-\textit{The parallels here would be difficult to ignore;
-Lily and others have pointed this out,
-but it's worth thinking about the broader trend of
-resisting management,
-driven by routinization,
-enabled by discretization.
-Do that here.}
+% Distributing work in method described above may be thought of as vertically slicing work such that 
+% each person is responsible for the whole task
+% --- making a whole garment or being responsible for driving in a single neighborhood.
+% Broken down in this way, work could grow to unprecedented scales,
+% but the quality of the work would remain relatively variable.
+% Textile work being a salient example,
+% some seamstresses might be better than others at the same task,
+% making this framing of the work initially problematic.
+
+% A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
+% manageable routines that could be taught relatively easily,
+% and whose work output could be evaluated in abstraction from the rest of the work.
+% In Ford's assembly line, this meant that workers were not responsible for building a whole car,
+% but a single very narrowly defined action that needed to be done on every car.
+
+% This approach paralleled what would be known as ``Taylorism'',
+% and its influences can be seen today in crowd work and micro--tasks.
+% \textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
+% horizontal cuts instead of vertical}
+% \cite{cheng2015break,writingMicroTasks}.
+% By decomposing and recomposing tasks,
+% and in particular by assigning similarly natured work to the same workers,
+% workers could become ``experts'' in a small aspect of the work that they did,
+% speeding their work dramatically
+% \cite{delayAndOrderLasecki}.
+% Perhaps more important, however, was that
+% the breaking down of work into tasks has made it more practical to evaluate work at each stage
+% [\textit{citation needed}].
+
+% Opening the door to more discrete modules of work known as tasks 
+% Contributions such as
+% ``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})
+% and
+% ``Cascade'' % (described in \citetitle{chilton2013cascade})
+% make it easier to assign smaller components of work to myriad workers and
+% recompose the constituent parts into something more complex
+% \cite{bernstein2015soylent,chilton2013cascade}.
+% Some work has taken special interest in the process of breaking tasks down into more manageable parts
+% inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
+% both for the purposes of crowd--sourcing in the sense of out--sourcing work
+% as well as ``selfsourcing''
+% \cite{cheng2015break,teevan2014selfsourcing}.
+
+% The parallels between this kind of work and
+% historical piecework of the turn of the \nth{20} century
+% is by no means new
+% [\textit{citation needed --- who said this first?
+most notably?
+Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
+% \textit{Elaborate on gig work, piecework, etc\dots}
+
+% The \textbf{discretization} and decomposition of work certainly facilitates
+% breaking tasks down,
+% parallelizing work,
+% and getting more done across a broader array of people;
+% but the key advantage in turning macro tasks 
+% into many micro tasks is that we can make that work \textit{routine}.
+
+% By making larger bodies of work less contextually situated,
+% we can define processes that make that work
+% and the instructions thereof
+% useful for virtually anyone with shared cultural or intellectual background.
+
+% routinization can be seen in the contemporary in work--flows
+% \cite{foundry,bernstein2015soylent}
+% [\textit{There are various papers that talk about creating a work--flow;
+% maybe this is where FFV belongs? Foundry?}].
+
+% \subsection{External Management}
+% \textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
+% manage work more meaningfully.
+% Routinization in particular made it possible
+% (and indeed preferable) to assume that people
+% ``\dots are of equal ability and \dots can do any of the $n$ jobs''
+% \cite{hu1961parallel}.
+% \citeauthor{storiesIraniSilberman}
+% find similarly that
+% this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
+% \cite{storiesIraniSilberman}.
+
+% Crucially,
+% \citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
+% which in \citeyear{hu1961parallel}
+% meant that workers would at least have to be trained in person, if not instructed as well.
+% Throughout the industrial revolution, workers flocked to centralized locales of production,
+% and physically situated factories made it both
+% difficult to ignore the general pattern of mistreatment and
+% relatively easy to organize.
+
+% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
+
+% \textit{I want to cite Justin's work on crowdsourcing effort to highlight that
+% we can evaluate workers for the work they do with increasingly finite amounts of time;
+% also, Ranjay's work that embraces failure because
+% we can essentially algorithmically make sense of failure (let's call it noise)
+% and re--capture signal.}
+
+% \subsection{Resistance}
+% So it's not surprising, then, that workers are frustrated:
+% Turkopticon, Dynamo, and many other pieces of work point this out
+% \cite{turkopticon,dynamo,uberAlgorithm}
+% [\textit{also want to cite Brian's work from CHI this year}].
+% In popular culture, reporters have inquired about
+% what ``the future of labor unions'' will look like
+% [\textit{Imagine I had cited an online article and
+% it wasn't totally inappropriate for a conference}].
+
+% \textit{The parallels here would be difficult to ignore;
+% Lily and others have pointed this out,
+% but it's worth thinking about the broader trend of
+% resisting management,
+% driven by routinization,
+% enabled by discretization.
+% Do that here.}
 

### 2016-05-25 cscw/papers/2017/labor/trackingWork.tex
i rerererewrote some stuff

@@
 
-In this section, we'll consider several components of contemporary piecework
-that relate to its historical parallel.
-In particular, we'll look at the ``decomposition'' of complex skill--intensive tasks
-into smaller, more abstracted work, as well as the routinization that this allows.
+This section will largely discuss the processes at work that make distributed,
+digitally managed work both possible and indeed preferable for ``requesters''
+(in other words, the employers who solicit workers).
+This body of research spans a broad field
+within the CSCW and broader HCI community.
+In this context,
+we'll look at the body of research through the lens of
+highlighting contributions which expand what we
+(as ``requesters'' of work) can do by managing workers in novel ways.
+This work broadly consists of three areas:
+\begin{inlinelist}
+\item \nameref{sec:decomposition},
+\item \nameref{sec:workAbstraction}, and
+\item \nameref{sec:flexibility}.
+\end{inlinelist}
+
+\subsubsection{Decomposition}\label{sec:decomposition}
+Scholarship describing and exploring
+the decomposition of tasks is perhaps the most established of the above areas among HCI researchers;
+\citeauthor{bernsteinSoylent} made early contributions to this area with \textit{Soylent}
+(by coining and employing the ``find--fix--verify'' approach)
+\cite{bernsteinSoylent}.
+% \citeauthor{horowitz2010anatomy} found that optimally directing questions to relevant answerers
+% on various domains could yield relevant answers at enormous scales
+% \cite{horowitz2010anatomy}.
+\citeauthor{cheng2015break} found that microtasks
+--- though not necessarily \textit{faster} than ``macrotasks'' ---
+yield higher quality work,
+particularly when that work might be readily interrupted
+% (we will, however, discuss the implications of these findings in further depth later)
+\cite{cheng2015break}.
+\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work
+by exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+
+\subsubsection{Work Abstraction}\label{sec:workAbstraction}
+Decomposition allows requesters to assign work tasks without concern for the broader context.
+\citeauthor{chilton2013cascade} perhaps best illustrates this,
+demonstrating with \textit{Cascade} that it's possible to
+break certain classes of tasks apart
+% (recursively, if necessary)
+in such a way that they yield hierarchical categories
+previously thought to be safely within the domain of expert workers
+with top--down awareness of the context of the work as a whole
+\cite{chilton2013cascade}.
+
+Much of the research in the space of designing crowd work
+sought to illustrate the potential to take highly creative or skilled work
+and generate high--quality work.
+Perhaps the most salient case study here can be found in
+\citeauthor{foundry}'s \textit{Foundry}, which employed
+``flash teams'' to achieve expert--level outcomes via thoughtful
+decomposition of work as ``modular tasks''
+\cite{foundry}.
+
+\subsubsection{Flexibility}\label{sec:flexibility}
+Earlier we discussed \citeauthor{cheng2015break}'s work
+identifying the impact that interruption has on worker performance.
+This work both points to and embodies a broader sentiment in
+both the study and practice
+of crowd work that microtasks should be designed flexibly,
+fully exploiting the abstracted nature of each piece of work
+\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
+While \citeauthor{cheng2015break} identified costs with breaking tasks into smaller components
+in the form of higher cumulative time to complete
+(albeit much shorter real time to complete, owing to parallelization),
+\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+similar tasks together.
+
+% This work on decomposition is not strictly new;
+``Decomposition'' generally illustrates the same concepts of work that ``Taylorism''
+and scientific management sought to embody
+---
+\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
+\citeyear{silberman2010ethics}
+\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
 
@@
 % \subsection{Discretization and Routinization}
-
+% \subsubsection{Decomposition}\label{sec:decomposition}
 Distributing work in method described above may be thought of as vertically slicing work such that 
@@
 Contributions such as
-``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})
+``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
 and
@@
 recompose the constituent parts into something more complex
-\cite{bernstein2015soylent,chilton2013cascade}.
+\cite{bernsteinSoylent,chilton2013cascade}.
 Some work has taken special interest in the process of breaking tasks down into more manageable parts
@@
 is by no means new.
-% [\textit{citation needed --- who said this first?
-most notably?
-Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
-% \textit{Elaborate on gig work, piecework, etc\dots}
 
-% \subsection{routinization}
 The \textbf{discretization} and decomposition of work certainly facilitates
@@
 routinization can be seen in the contemporary in work--flows
-\cite{foundry,bernstein2015soylent}.
+\cite{foundry,bernsteinSoylent}.
 % [\textit{There are various papers that talk about creating a work--flow;
@@
 \item leads to ``low pay'' \cite{turkopticon,dynamo}
-\item leads to ``variable quality work''
+\item leads to ``variable quality work'' \cite{MaliciousCrowdworkersGadiraju,AccuracyTradeoffsLasecki}.
 \end{enumerate}
-
 \upshape
 
-\subsection{Factories}
+\subsection{Factories}\label{sec:Factorization}
 
@@
 % Contributions such as
-% ``find--fix--verify'' % (described in \citetitle{bernstein2015soylent})
+% ``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
 % and
@@
 % recompose the constituent parts into something more complex
-% \cite{bernstein2015soylent,chilton2013cascade}.
+% \cite{bernsteinSoylent,chilton2013cascade}.
 % Some work has taken special interest in the process of breaking tasks down into more manageable parts
@@
 % routinization can be seen in the contemporary in work--flows
-% \cite{foundry,bernstein2015soylent}
+% \cite{foundry,bernsteinSoylent}
 % [\textit{There are various papers that talk about creating a work--flow;

### 2016-05-26 cscw/papers/2017/labor/trackingWork.tex
subsection 2 is more or less fleshed out; ss3 is still not really done

@@
 
-\usepackage{todonotes,txfonts,balance,graphics,color}
+\usepackage{todonotes,txfonts,balance,graphics,color,alphalph}
 \usepackage{booktabs,textcomp,microtype,ccicons}
@@
 \usepackage{enumitem}
+\usepackage{moreenum}
 \usepackage[T1]{fontenc}
-\usepackage[super]{nth}
+\usepackage{nth}
 \usepackage[pdftex,pdfpagelabels=false]{hyperref}
@@
 \item the results of these forces; namely the turning of work toward \textbf{factories},
-which we will discuss in ``\dots''.
+which we will discuss in ``\nameref{sec:Factorization}''.
 \end{enumerate}
@@
 \item \nameref{sec:workAbstraction}, and
-\item \nameref{sec:flexibility}.
-\end{inlinelist}
+\item \nameref{sec:flexibility}
+\end{inlinelist}.
+As we explore this work, we'll
+attempt to relate the advances in the design of crowd work to
+the research contributions made in the research of assembly line manufacturing,
+Taylorism, and
+scientific management 
+during the \nth{20} century
+\cite{hu1961parallel}.
 
@@
 \cite{bernsteinSoylent}.
-% \citeauthor{horowitz2010anatomy} found that optimally directing questions to relevant answerers
-% on various domains could yield relevant answers at enormous scales
-% \cite{horowitz2010anatomy}.
 \citeauthor{cheng2015break} found that microtasks
@@
 particularly when that work might be readily interrupted
-% (we will, however, discuss the implications of these findings in further depth later)
 \cite{cheng2015break}.
@@
 \cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+While this work doesn't strictly fall under ``crowdsourcing'',
+the major contributions here
+seem uncontroversially to be inspired by the design of crowd work.
 
-\subsubsection{Work Abstraction}\label{sec:workAbstraction}
-Decomposition allows requesters to assign work tasks without concern for the broader context.
-\citeauthor{chilton2013cascade} perhaps best illustrates this,
-demonstrating with \textit{Cascade} that it's possible to
-break certain classes of tasks apart
-% (recursively, if necessary)
-in such a way that they yield hierarchical categories
-previously thought to be safely within the domain of expert workers
-with top--down awareness of the context of the work as a whole
-\cite{chilton2013cascade}.
-
-Much of the research in the space of designing crowd work
+Much of the research in the space of designing crowd work has
 sought to illustrate the potential to take highly creative or skilled work
-and generate high--quality work.
-Perhaps the most salient case study here can be found in
+and generate high--quality results.
+Perhaps the most notable case study here can be found in
 \citeauthor{foundry}'s \textit{Foundry}, which employed
@@
 
+Work decomposition is far from new;
+``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
+and scientific management sought to embody
+---
+\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
+\citeyear{silberman2010ethics}
+\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
+In both the historical and contemporary cases of decomposed work,
+work was,
+at least initially,
+distributed in the form of tasks to the homes of workers;
+\citeauthor{riisOtherSideLives} captured this in his documentary work
+\citetitle{riisOtherSideLives}
+in \citeyear{riisOtherSideLives}
+\cite{riisOtherSideLives}.
+
+% In both the historical and contemporary cases,
+% distributed, speculative work
+% has promised
+% to diffuse the costs of capital--intensive resources
+% (whether they were sewing machines or laboratory equipment).
+
+\subsubsection{Work Abstraction}\label{sec:workAbstraction}
+Decomposition allows requesters to assign tasks without concern for the broader context.
+While we'll discuss this aspect of crowd work more critically later,
+it's worth pointing out that discrete blocks of work containing all the relevant context for a worker
+allows workers to engage with virtually any component of work without worrying that their lack of 
+higher--level awareness of the goals of the requester might negatively affect their work.
+
+\citeauthor{chilton2013cascade} perhaps best illustrated this with
+\textit{Cascade} by demonstrating that it's possible to
+break certain classes of tasks apart
+in such a way that they yield taxonomies of various subjects,
+a task previously thought to be safely within the domain of expert workers
+with top--down awareness of the context of the work as a whole
+\cite{chilton2013cascade}.
+\citeauthor{verroios2014context} further illustrate this potential by
+forming a task one might consider highly contextually dependent
+--- summarizing the contents of a movie ---
+in such a way that crowd workers could contribute small pieces of work without
+needing to know the content of the rest of the project
+\cite{verroios2014context}.
+
+Here, \citeauthor{hu1961parallel}'s work,
+saying of assembly line work that
+``It is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
+parallels the approach that dominated early research into crowd work
+--- namely, using non--expert crowds for complex work
+\cite{hu1961parallel}.
+This mindset in \citeauthor{hu1961parallel}'s analysis,
+and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
+substantively owes its existence to scientific management
+and the rigorous decomposition of work into tasks, discussed earlier,
+and persists to this day as it colors
+researchers' goals and objectives in the study and design of crowd work.
+
+Piecework's influence on the abstraction of work into tasks,
+described above, is more than just caused by the decomposition of work;
+work abstraction itself makes it possible for workers to come and go flexibly,
+prompting work requesters to consider ways to design these now discrete tasks in ways that
+maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
+\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
+It's to this area that we now turn our attention.
+
 \subsubsection{Flexibility}\label{sec:flexibility}
 Earlier we discussed \citeauthor{cheng2015break}'s work
-identifying the impact that interruption has on worker performance.
+measuring the impact that interruption has on worker performance.
 This work both points to and embodies a broader sentiment in
@@
 
-% This work on decomposition is not strictly new;
-``Decomposition'' generally illustrates the same concepts of work that ``Taylorism''
-and scientific management sought to embody
----
-\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
-\citeyear{silberman2010ethics}
-\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
+Given the importance of consistent work results, one might intuit that
+requesters would prefer high--quality workers who can be relied upon to be available
+(even for contextually independent tasks),
+which would appear to contradict the benefits of flexibility already discussed;
+requesters have thus made significant headway toward
+``embracing error'' to allow requesters to maximize the benefits of a flexible,
+even transient,
+workforce.
 
-In both the historical and contemporary cases,
-distributed, speculative work
-has promised
-to diffuse the costs of capital--intensive resources
-(whether they were sewing machines or laboratory equipment).
+\citeauthor{embracingErrorKrishna} offer orders--of--magnitude improvements
+in various binary classification tasks
+% --- a typical application crowd work ---
+on the principle that myriad diverse workers complete these tasks
+in order to accurately inform the model.
+And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
+some researchers have designed work predicated on the expectation of this phenomenon:
+\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
+crowd workers would never be given enough information to piece together sensitive information about
+any single topic
+\cite{sensitiveTasks}.
+
+\subsection{The Fallout of Crowd Work}\label{sec:Fallout}
+% If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
+% the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
+\citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
+(see also \citeauthor{dynamo}'s work
+continuing in the spirit of this observation to generate collective action to improve worker conditions)
+\cite{turkopticon,dynamo}.
+\citeauthor{uberAlgorithm}
+find similarly that workers on gig work platforms are frustrated by the systems on which they work,
+to say little of the policies which these systems enforce
+\cite{uberAlgorithm}.
+
+We discussed the benefits of flexibility
+(both in the sense of having arbitrary workers perform tasks and
+in the sense that we can design tasks to be more resilient to poor work)
+in the previous section.
+It's from that point in the literature that we turn our attention to
+the perhaps unintended effects of crowd work
+and the affordances for transience that we build into this mode of work.
+We'll address two major areas of work under this subject:
+\begin{inlinelist}
+\item \nameref{sec:lowPay}; and
+\item \nameref{sec:varQualWork}.
+\end{inlinelist}
+
+\subsubsection{Low Pay}\label{sec:lowPay}
+\citeauthor{laborEconomicsOfCrowdsourcingHorton}
+identified problems with crowd work wages relatively early on,
+attempting to address this imbalance from a behavioral economic perspective ---
+that is, identifying and presenting a model that describes a worker's
+``\textit{reservation wage}''
+\cite{laborEconomicsOfCrowdsourcingHorton}.
+This work has largely informed much of the research into and practice of estimating crowd work compensation
+\cite{incentivesShaw,paolacci2010running}.
+More cynically,
+we would describe this work as identifying optimal levels to activate race conditions among piece workers.
+
+But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}'',
+a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
+\cite{turkopticon}.
+Of particular relevance here,
+\citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
+wage theft and
+pay rates that translate to well under minimum wage.
+Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},
+we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
+was \$1.38, while the mean was \$3.63
+\cite{laborEconomicsOfCrowdsourcingHorton}.
+
+Understanding workers' motivations given these conditions has thus become a goal for some researchers
+\cite{whyWouldAnyoneBrewer}.
+\citeauthor{Sun20111033} conclude that
+``\dots~solvers participate in online tasks
+not only for money
+but also for enjoyment
+or the sense of self--worth''
+\cite{Sun20111033}.
+This might have rung true in \citeyear{Sun20111033},
+and certainly corroborates \citeauthor{Ross}'s findings after investigating
+``who are the crowdworkers'',
+but as \citeauthor{whoareNOTtheTurkers} points out
+``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
+\cite{Ross,whoareNOTtheTurkers}.
+
+Now, \citeauthor{turkopticon}
+and later
+\citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
+via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s work
+\cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
+
+\subsubsection{Variable quality work}\label{sec:varQualWork}
+Meanwhile,
+researchers have struggled with what we might call ``variable quality work''
+along two dimensions:
+First, to use the characterization of one of these contributions,
+``understanding malicious behavior''
+\cite{MaliciousCrowdworkersGadiraju}.
+While some work has cast workers as ``malicious'' agents,
+the evidence thus far suggests that
+workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
+(a topic of discussion to which we'll return later)
+\cite{uberAlgorithm}.
+The second dimension of research in this space generally attempts
+to eke out the highest quality work possible from workers
+given the apparent difficulty in predicting work outcomes
+
+% Low pay yields variable quality work for a number of reasons,
+% but before we discuss the causes of this effect, we should discuss the
+
+% Variance in the quali
+
+The effect that poor wages have had on piece work and factory workers is hardly unknown or novel;
+\citeauthor{gantt1913work} discuss this exact mechanism in his book on
+\citetitle{gantt1913work}, pointing out that
+``\dots~where there is no union,
+the class wage is practically gauged by the wages the poor workman will accept,
+and the good workman soon becomes discouraged and \textit{sets his pace by that of his less efficient neighbor},
+with the result that the general tone of the shop is lowered'' (emphasis added)
+\cite{gantt1913work}.
+\ali{Do I need to follow up or does this basically speak for itself?}
+
+This research is similar to, but subtly different from, the notion of the ``market for `lemons'''
+which \citeauthor{fort2011amazon} discuss;
+specifically, \citeauthor{akerlof1970market}'s writing of a ``market for `lemons'''
+describes a marketplace where the quality of the product or service is unknown to the buyer
+\cite{fort2011amazon,akerlof1970market}.
+The effect of this \textit{perceived} uncertainty is that
+the \textit{actual} trustworthiness drops precipitously
+as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,
+leaving only the ones who cannot or will not establish their trustworthiness.
+
+Suffice it to say, then, that poor pay and poor work are linked,
+and that we should not be surprised to find this relationship play out online as strongly as it does offline.
+Indeed,
+
+% \itshape
+% sections here should include:
+% \begin{enumerate}
+% \item ``flexibility'' \cite{friedman2014workers,DesigningForLabourUber}
+% \item leads to ``low pay'' \cite{turkopticon,dynamo}
+% \item leads to ``variable quality work'' \cite{MaliciousCrowdworkersGadiraju,AccuracyTradeoffsLasecki}.
+% \end{enumerate}
+% \upshape
+
+\subsection{Factories}\label{sec:Factorization}
+
+\rule{\linewidth}{1pt}
 
@@
 as well as ``selfsourcing''
-\cite{cheng2015break,teevan2014selfsourcing}.
+\cite{cheng2015break,selfsourcingTeevan2014}.
 
@@
 % the instructions were clear and scoped enough that you could repeat the task ad infinitum.}
-
-\subsection{The Fallout of Gig Work}\label{sec:Fallout}
-If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
-the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
-\citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
-(see also \citeauthor{dynamo}'s work
-continuing in the spirit of this observation to generate collective action to improve worker conditions)
-\cite{turkopticon,dynamo}.
-\citeauthor{uberAlgorithm}
-find similarly that workers on gig work platforms are frustrated by the systems on which they work,
-to say little of the policies which these systems enforce
-\cite{uberAlgorithm}.
-
-\itshape
-sections here should include:
-\begin{enumerate}
-\item ``flexibility'' \cite{friedman2014workers,DesigningForLabourUber}
-\item leads to ``low pay'' \cite{turkopticon,dynamo}
-\item leads to ``variable quality work'' \cite{MaliciousCrowdworkersGadiraju,AccuracyTradeoffsLasecki}.
-\end{enumerate}
-\upshape
-
-\subsection{Factories}\label{sec:Factorization}
 
@@
 % as well as ``selfsourcing''
-% \cite{cheng2015break,teevan2014selfsourcing}.
+% \cite{cheng2015break,selfsourcingTeevan2014}.
 

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
cleanup

@@
 \begin{enumerate}
-\item the processes involved in making work into tasks, or \textbf{discretization},
-in ``\nameref{sec:MakingGigsWork}'';
-\item the outcomes of that discretization, both on the work itself as well as the workers,
-in ``\nameref{sec:Fallout}'';
+\item the processes involved in making work into tasks, or \textbf{discretization};
+% in ``\nameref{sec:MakingGigsWork}'';
+\item the outcomes of that discretization, both on the work itself as well as the workers;
+% in ``\nameref{sec:Fallout}'';
 and finally
-\item the results of these forces; namely the turning of work toward \textbf{factories},
-which we will discuss in ``\nameref{sec:Factorization}''.
+\item the results of these forces; namely the turning of work toward \textbf{factories}.
+% which we will discuss in ``\nameref{sec:Factorization}''.
 \end{enumerate}
@@
 % all the things we can do with crowd work.
+
+The HCI community is perhaps most familiar with examples of task--based work such as
+99designs,
+Amazon Mechanical Turk (AMT),
+and increasingly
+Uber \& Lyft,
+which all allow requesters in various forms to tap into
+resources such as
+cars,
+computers, and above all
+``cognitive surplus''
+with relative ease
+\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
+This insight, that
+workers can be geographically distributed and tasks decomposed,
+% (especially given careful central management of those workers, discussed later),
+has proven remarkably compelling
+and an effective fulcrum for leveraging the Internet for highly scalable work
+\cite{sensitiveTasks,embracingErrorKrishna}.
 
@@
 \subsubsection{Decomposition}\label{sec:decomposition}
+Piecework may be thought of as vertically slicing work such that 
+each person is responsible for the whole task
+--- making a whole garment, in this case.
+Broken down in this way, work could grow to unprecedented scales,
+but the quality of the work would remain relatively variable.
+Textile work being a salient example,
+it took time for workers to acquire sufficient skill
+to do every aspect of the piece work so that the garment would be accepted.
+% some seamstresses might be better than others at the same task,
+% making this framing of the work initially problematic.
+
+A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
+manageable routines that could be taught relatively easily,
+and whose work output could be evaluated in abstraction from the rest of the work.
+In Ford's assembly line, this meant that workers were not responsible for building a whole car,
+but a single very narrowly defined action that needed to be done on every car.
+
+This approach paralleled what would be known as ``Taylorism'',
+and its influences can be seen today in crowd work and micro--tasks.
+\textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
+horizontal cuts instead of vertical}
+\cite{cheng2015break,writingMicroTasks}.
+By decomposing and recomposing tasks,
+and in particular by assigning similarly natured work to the same workers,
+% and especially as the nature of the work could be aligned with other, similar tasks,
+workers could become ``experts'' in a small aspect of the work that they did,
+speeding their work dramatically
+\cite{delayAndOrderLasecki}.
+Perhaps more important, however, was that
+the breaking down of work into tasks has made it more practical to evaluate work at each stage
+[\textit{citation needed}].
+
 Scholarship describing and exploring
@@
 
-Work decomposition is far from new;
+Work decomposition, then, is far from new;
 ``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
@@
 saying of assembly line work that
-``It is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
+``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
 parallels the approach that dominated early research into crowd work
@@
 
+\itshape
+So work output is variable, and that's bad. What are we doing about it?
+Historically, this is what led to factories; you would pay some experts who understood what they needed to do
+and they would form a cohort into which you invest more.
+
+Some research already looks at stuff like investing in workers
+\cite{shepherdingDow}.
+What else is there that kind of goes in this direction?
+Informally, we know that this happens a lot
+\cite{jonBrelig}.
+In fact, the formalization of factories 
+% But \citeauthor{shepherdingDow}'s contribution operates at an individual scale;
+% giving feedback --- even automated --- to individual workers replicates many of the same features
+% (and limitations) of conventional offline work
+\upshape
+
+\section{Bleak Futures}
+We've traced a path from piecework itself through
+the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
+from there, we followed the trail from decomposition to work abstraction,
+and the potential to break work away from workers themselves,
+which in turn facilitates flexibility for requesters as well as workers.
+\ali{oh, forgot to mention that
+I want to cite something from CHI 2016 that mentioned
+the flexibility that Uber/Lyft give drivers}
+
 \rule{\linewidth}{1pt}
-
-% \textit{There's a reference to something that's \textbf{not} 99designs,
-% that describes a call for scientific solutions/designs to solve some problem
-% that the requesters have.
-% Honestly it's killing me that I can't remember the name,
-% and I need to find this before I do anything productive ever again.}
-% \msb{netflix prize?}
-% that's an interesting example in the same vein as what i'm thinking about, but this is a whole business of it
-% \textbf{INNOCENTIVE} oh my gosh it just clicked my life is complete right now at 9:54am.
-% I don't even need to submit anymore.
-% I'm satisfied having pulled this strand of a memory out of my brain.
-
-The HCI community is perhaps most familiar with examples such as
-Innocentive,
-Uber, Lyft,
-and Amazon Mechanical Turk (AMT),
-which all allow requesters in various forms to tap into
-resources such as
-cars,
-computers, and above all
-``cognitive surplus''
-with relative ease
-\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
-This insight, that
-workers can be geographically distributed and tasks decomposed,
-% (especially given careful central management of those workers, discussed later),
-has proven remarkably compelling
-and an effective fulcrum for leveraging the Internet for highly scalable work
-\cite{sensitiveTasks,embracingErrorKrishna}.
 
@@
 % maintenance costs that a factory owner would associate with their capital (e.g. auto costs)
-While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
-(inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
-these new systems go further
-by framing ownership and responsibility for costs
-(e.g. automotive maintenance)
-on drivers.
-In our fieldwork, we discovered many longtime ``on--demand'' drivers
-(affiliated with platforms such as Uber and Lyft)
-who readily voiced frustration with the costs of
-frequent oil changes and the unexpected replacement of expensive parts.
-Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
-textile workers were given the source material with which to work,
-but were generally expected to use their own sewing materials and work in their own homes
-\cite{hapke2004sweatshop}.
+
+% While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
+% (inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
+% these new systems go further
+% by framing ownership and responsibility for costs
+% (e.g. automotive maintenance)
+% on drivers.
+% In our fieldwork, we discovered many longtime ``on--demand'' drivers
+% (affiliated with platforms such as Uber and Lyft)
+% who readily voiced frustration with the costs of
+% frequent oil changes and the unexpected replacement of expensive parts.
+% Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
+% textile workers were given the source material with which to work,
+% but were generally expected to use their own sewing materials and work in their own homes
+% \cite{hapke2004sweatshop}.
 
@@
 % \subsubsection{Decomposition}\label{sec:decomposition}
-Distributing work in method described above may be thought of as vertically slicing work such that 
-each person is responsible for the whole task
---- making a whole garment or being responsible for driving in a single neighborhood.
-Broken down in this way, work could grow to unprecedented scales,
-but the quality of the work would remain relatively variable.
-Textile work being a salient example,
-some seamstresses might be better than others at the same task,
-making this framing of the work initially problematic.
-
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
-manageable routines that could be taught relatively easily,
-and whose work output could be evaluated in abstraction from the rest of the work.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-but a single very narrowly defined action that needed to be done on every car.
-
-This approach paralleled what would be known as ``Taylorism'',
-and its influences can be seen today in crowd work and micro--tasks.
-\textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
-horizontal cuts instead of vertical}
-\cite{cheng2015break,writingMicroTasks}.
-By decomposing and recomposing tasks,
-and in particular by assigning similarly natured work to the same workers,
-% and especially as the nature of the work could be aligned with other, similar tasks,
-workers could become ``experts'' in a small aspect of the work that they did,
-speeding their work dramatically
-\cite{delayAndOrderLasecki}.
-Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage
-[\textit{citation needed}].
 
@@
 
-Opening the door to more discrete modules of work known as tasks 
-% Work has become discretized in the last 10 years, particularly in the information work sector.
-Contributions such as
-``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
-and
-``Cascade'' % (described in \citetitle{chilton2013cascade})
-make it easier to assign smaller components of work to myriad workers and
-recompose the constituent parts into something more complex
-\cite{bernsteinSoylent,chilton2013cascade}.
-Some work has taken special interest in the process of breaking tasks down into more manageable parts
-% Some work has even turned to break it down'',
-inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
-both for the purposes of crowd--sourcing in the sense of out--sourcing work
-as well as ``selfsourcing''
-\cite{cheng2015break,selfsourcingTeevan2014}.
-
-The parallels between this kind of work and
-historical piecework of the turn of the \nth{20} century
-is by no means new.
-
-The \textbf{discretization} and decomposition of work certainly facilitates
-breaking tasks down,
-parallelizing work,
-and getting more done across a broader array of people;
-but the key advantage in turning macro tasks 
-into many micro tasks is that we can make that work \textit{routine}.
-
-By making larger bodies of work less contextually situated,
-we can define processes that make that work
-and the instructions thereof
-useful for virtually anyone with shared cultural or intellectual background.
-
-routinization can be seen in the contemporary in work--flows
-\cite{foundry,bernsteinSoylent}.
+% Opening the door to more discrete modules of work known as tasks 
+% % Work has become discretized in the last 10 years, particularly in the information work sector.
+% Contributions such as
+% ``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
+% and
+% ``Cascade'' % (described in \citetitle{chilton2013cascade})
+% make it easier to assign smaller components of work to myriad workers and
+% recompose the constituent parts into something more complex
+% \cite{bernsteinSoylent,chilton2013cascade}.
+% Some work has taken special interest in the process of breaking tasks down into more manageable parts
+% % Some work has even turned to break it down'',
+% inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
+% both for the purposes of crowd--sourcing in the sense of out--sourcing work
+% as well as ``selfsourcing''
+% \cite{cheng2015break,selfsourcingTeevan2014}.
+
+% The parallels between this kind of work and
+% historical piecework of the turn of the \nth{20} century
+% is by no means new.
+% 
+% The \textbf{discretization} and decomposition of work certainly facilitates
+% breaking tasks down,
+% parallelizing work,
+% and getting more done across a broader array of people;
+% but the key advantage in turning macro tasks 
+% into many micro tasks is that we can make that work \textit{routine}.
+
+% By making larger bodies of work less contextually situated,
+% we can define processes that make that work
+% and the instructions thereof
+% useful for virtually anyone with shared cultural or intellectual background.
+
+% routinization can be seen in the contemporary in work--flows
+% \cite{foundry,bernsteinSoylent}.
 % [\textit{There are various papers that talk about creating a work--flow;
@@
 
-Piecework emerged at the turn of the \nth{20} century,
-right at the time that telecommunications began to boom in the United States. % too far a reach?
-\textit{Did telecom enable remote management?}
+% Piecework emerged at the turn of the \nth{20} century,
+% right at the time that telecommunications began to boom in the United States. % too far a reach?
+% \textit{Did telecom enable remote management?}
 

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
refining intro

@@
 \cite{dynamo}.
-\ali{I really don't like this paragraph\dots}
+% \ali{I really don't like this paragraph\dots}
 % here are undoubtedly more examples of differences.
@@
 \subsection{Piecework as a lens to understand gig work}
+
 \citeauthor{crowdworkFuture}
-investigate the future of crowd work by
-situating and interrogating it through ``piecework'',
-which historically described work done in the home,
+write of the future of crowd work in an effort to investigate
+whether crowd work will be a worthwhile form of work or another instantiation of ``piecework'',
+a term historically used to describe work done in the home,
 in manageable tasks,
@@
 is ambiguous, or at least the subject of some controversy.
-% \msb{my first reaction was ``this isn't definitional''.
-But now I look back and see that you're just saying there are similarities.
-It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
 \end{inlinelist}.
 
-\subsection{Context}
-Importantly, since the reemergence of piecework we've seen substantial frustration
-and resistance among the workers in this area
+% \subsection{Context}
+In the past decade, researchers have observed frustration
+grow among on--demand workers,
+with expression of this frustration spanning a wide range of tactics
 \cite{uberAlgorithm,turkopticon,dynamo}.
-This paper attempts to make sense of the broader research on this piecework,
-or ``gig work'', by framing this as one of several steps in the marginalization of workers,
-% \msb{this feels out of order to me. I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
-%
-% this needs to be re-written to reflect the new split
-%
-%
-% starting with the discretization of tasks,
-% followed by routinization and the rise of work--flows,
-% and finally the external management of workers.
-% 
-All of this is to say that these milestones follow sequentially,
-not coincidentally but necessarily,
-and that by tracing this path using the corpus of scholarship on labor and workers
-we can both make sense of past events and perhaps reasonably predict next steps.
-% \msb{I don't think we've offered much proof that this is the case yet. Why must they follow necessarily? Can you offer a cause and effect relationship?}
-
-We explore each topic
---- discretization, routinization, \& management ---
-by looking at case studies in social computing.
-Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
-we turn to look ahead, envisioning future areas that researchers in social computing
---- and particularly digitally mediated work ---
-should explore.
+Attempting to make sense of these case studies has been challenging
+in part because
+a wholly encompassing framework for understanding this form of work
+has thus proven difficult to capture.
+
+This paper attempts to make sense of the broader research on this new form of labor,
+or ``gig work'', by evaluating this work through a more familiar lens: piecework.
+More concretely, by looking at task--based or ``gig'' work as
+an instantiation and continuation of piecework,
+and by looking for patterns of behavior that the corresponding literature predicts
+on this basis, we can do the following:
+\begin{inlinelist}
+  \item we can make sense of the phenomena so far as part of a much larger series of interrelated events;
+  \item we can bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
+  \item we can offer predictions of what social computing researchers, and workers themselves, should expect to see on the horizon of on--demand work.
+\end{inlinelist}
+
+We'll look at a broad range of cases under a number of major themes
+we propose as broadly describing the types of research being done in crowd work
+and more generally in what we argue is contemporary piecework.
+After validating this lens as a way of reasoning about on--demand labor,
+we'll attempt to use this perspective to suggest areas of research worth anticipating,
+and developments we should expect to see in the maturation of digitally mediated work.
 
@@
 \item the processes involved in making work into tasks, or \textbf{discretization};
-% in ``\nameref{sec:MakingGigsWork}'';
 \item the outcomes of that discretization, both on the work itself as well as the workers;
-% in ``\nameref{sec:Fallout}'';
 and finally
 \item the results of these forces; namely the turning of work toward \textbf{factories}.
-% which we will discuss in ``\nameref{sec:Factorization}''.
 \end{enumerate}
 
-We will follow these topics, using them as prompts for case studies in
-the emergence and
-development of contemporary piecework.
+% We will follow these topics, using them as prompts for case studies in
+% the emergence and
+% development of contemporary piecework.
 
@@
 it took time for workers to acquire sufficient skill
-to do every aspect of the piece work so that the garment would be accepted.
+to do every aspect of the work so that the garment would be accepted by the company soliciting that work.
 % some seamstresses might be better than others at the same task,

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
notes from meeting
adding implications section

@@
 \usepackage{booktabs,textcomp,microtype,ccicons}
+\usepackage[normalem]{ulem}
 % \usepackage{babel}
@@
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
-\newcommand{\ali}[1]{{\color{BrickRed}[\textit{al2: #1}]}}
+\newcommand{\ali}[1]{{\color{BrickRed}[\itshape al2: #1\upshape]}}
 
@@
 \begin{enumerate}
-\item the processes involved in making work into tasks, or \textbf{discretization};
-\item the outcomes of that discretization, both on the work itself as well as the workers;
+\item the \textbf{processes} involved in making work into tasks, or discretization;
+\item the outcomes (and indeed the \textbf{fallout}) of that discretization,
+both on the work itself as well as the workers;
 and finally
-\item the results of these forces; namely the turning of work toward \textbf{factories}.
+\item the \textbf{relationships} between workers and requesters of the work
+--- both \textit{cooperative} and \textit{adversarial} cases.
 \end{enumerate}
@@
 
-\subsection{Making Gig(s) Work}\label{sec:MakingGigsWork}
+\subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
 % this is phase 1: the decomposition & abstraction of work;
@@
 
-% \itshape
-% sections here should include:
-% \begin{enumerate}
-% \item ``flexibility'' \cite{friedman2014workers,DesigningForLabourUber}
-% \item leads to ``low pay'' \cite{turkopticon,dynamo}
-% \item leads to ``variable quality work'' \cite{MaliciousCrowdworkersGadiraju,AccuracyTradeoffsLasecki}.
-% \end{enumerate}
-% \upshape
-
-\subsection{Factories}\label{sec:Factorization}
-
+\subsection{Relationships Between Workers and Managers}\label{sec:relationships}
+
+\subsubsection{External Management}
 \itshape
-So work output is variable, and that's bad. What are we doing about it?
-Historically, this is what led to factories; you would pay some experts who understood what they needed to do
-and they would form a cohort into which you invest more.
-
-Some research already looks at stuff like investing in workers
-\cite{shepherdingDow}.
-What else is there that kind of goes in this direction?
-Informally, we know that this happens a lot
-\cite{jonBrelig}.
-In fact, the formalization of factories 
-% But \citeauthor{shepherdingDow}'s contribution operates at an individual scale;
-% giving feedback --- even automated --- to individual workers replicates many of the same features
-% (and limitations) of conventional offline work
+
+Scientific management in other words; should I just call it that and get right into it?
+I kind of want the focus to be on the adversarial nature of the management,
+and ``external'' seems to connote that more strongly than ``scientific'',
+which has a tone of being more\dots~Measured? Objective? Better?
+
+There's a fair amount to say here, and lots to cite, so I'm not super concerned.
+I do want to bring up \citeauthor{storiesIraniSilberman}'s
+\citetitle{storiesIraniSilberman} \cite{storiesIraniSilberman}
+as a way of pointing out that even in our effort to study piecework ethnographically,
+and despite our attempts to be reflexive and conscientious of
+the power dynamics that we bring into any discussion
+(as engineers, as people with megaphones, etc\dots)
+we still manage to influence and perhaps steer.
+
 \upshape
-
-\section{Bleak Futures}
-We've traced a path from piecework itself through
-the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
-from there, we followed the trail from decomposition to work abstraction,
-and the potential to break work away from workers themselves,
-which in turn facilitates flexibility for requesters as well as workers.
-\ali{oh, forgot to mention that
-I want to cite something from CHI 2016 that mentioned
-the flexibility that Uber/Lyft give drivers}
-
-\rule{\linewidth}{1pt}
-
-% The insight here seems to have been that
-% a substantial amount of work does not necessarily need to be done in collocated space,
-% but can be done from one's own home or other space arranged by the worker.
-% Moreover, by shifting the workplace into the worker's domain,
-% the cost of capital involved in that work can more easily be placed on workers' backs
-% \msb{I don't think that was the original motivation for AMT or Uber --- it was about reaching people who could not otherwise join the labor force for Amazon's data cleaning or for taxi-ing.}.
-% I didn't mean that this was some insight that Uber or Lyft had, but that this was some part of what makes Uber work;
-% i would argue categorically that these companies wouldn't functionally work if they weren't able to obviate the 
-% maintenance costs that a factory owner would associate with their capital (e.g. auto costs)
-
-% While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
-% (inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
-% these new systems go further
-% by framing ownership and responsibility for costs
-% (e.g. automotive maintenance)
-% on drivers.
-% In our fieldwork, we discovered many longtime ``on--demand'' drivers
-% (affiliated with platforms such as Uber and Lyft)
-% who readily voiced frustration with the costs of
-% frequent oil changes and the unexpected replacement of expensive parts.
-% Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
-% textile workers were given the source material with which to work,
-% but were generally expected to use their own sewing materials and work in their own homes
-% \cite{hapke2004sweatshop}.
-
-% \subsection{Discretization and Routinization}
-% \subsubsection{Decomposition}\label{sec:decomposition}
 
@@
 
-% I never talk about factorization it's too soon ahhhh
-% With piecework and factorization making mass manufacturing more viable,
-% businesses could produce at scales that were previously unimaginable.
-% Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
-% One of the issues that challenged mass production at the turn of the \nth{20} century,
-% as well as today in crowd work and elsewhere,
-% has been that of ensuring that the quality of the work is sufficiently good.
-
-% hey so i need to bring in citations on piecework for this stuff because I didn't exactly live through this time
-
-% is ANY of this good?
-
-% The approach to piecework described above hinged in some sense on
-% the shared understanding of the work to be done;
-% workers generally understood what was required of sewing together pieces of fabric.
-% Still, the work was speculative in that inadequately sewn clothing would not earn payment.
-% This represented a risk for pieceworkers,
-% who might spend time working on 10 garments only to be paid for 2 or 3.
-% It also represented a potential cost for management,
-% who would need to salvage the material
-% (if it was even possible),
-% and give that material to another pieceworker to do it correctly.
-
-% i need to find the source for this, because it's honestly only a vague quote in my head,
-% so... maybe i imagined it?
-
-% One of the solutions that
-
-% Broadly speaking, the solution to this problem is to make instructions clearer and easier to follow.
-% Further, by establishing a shared understanding
-% --- a ``collective conscious'' or even a corporate culture regarding work ---
-
-% \textit{I want to argue that modern gig work is discretized,
-% and that this parallels some of the piecework that we're historically familiar with}
-
-% Opening the door to more discrete modules of work known as tasks 
-% % Work has become discretized in the last 10 years, particularly in the information work sector.
-% Contributions such as
-% ``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
-% and
-% ``Cascade'' % (described in \citetitle{chilton2013cascade})
-% make it easier to assign smaller components of work to myriad workers and
-% recompose the constituent parts into something more complex
-% \cite{bernsteinSoylent,chilton2013cascade}.
-% Some work has taken special interest in the process of breaking tasks down into more manageable parts
-% % Some work has even turned to break it down'',
-% inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
-% both for the purposes of crowd--sourcing in the sense of out--sourcing work
-% as well as ``selfsourcing''
-% \cite{cheng2015break,selfsourcingTeevan2014}.
-
-% The parallels between this kind of work and
-% historical piecework of the turn of the \nth{20} century
-% is by no means new.
-% 
-% The \textbf{discretization} and decomposition of work certainly facilitates
-% breaking tasks down,
-% parallelizing work,
-% and getting more done across a broader array of people;
-% but the key advantage in turning macro tasks 
-% into many micro tasks is that we can make that work \textit{routine}.
-
-% By making larger bodies of work less contextually situated,
-% we can define processes that make that work
-% and the instructions thereof
-% useful for virtually anyone with shared cultural or intellectual background.
-
-% routinization can be seen in the contemporary in work--flows
-% \cite{foundry,bernsteinSoylent}.
-% [\textit{There are various papers that talk about creating a work--flow;
-% maybe this is where FFV belongs? Foundry?}].
-
-% \textit{Harken back to piecework
-% and the instructions that denim workers had to make clothing;
-% the instructions were clear and scoped enough that you could repeat the task ad infinitum.}
-
-% I commented everything going forward out not because it's all trash (it's not!)
-% but because we pretty significantly reframed the cases to be studying and
-% i didn't want the structure from before to influence the rewrite;
-% we should be bringing content from below back above. just, judiciously...
-
-% \subsection{Discretization and Routinization}
-
-% Distributing work in method described above may be thought of as vertically slicing work such that 
-% each person is responsible for the whole task
-% --- making a whole garment or being responsible for driving in a single neighborhood.
-% Broken down in this way, work could grow to unprecedented scales,
-% but the quality of the work would remain relatively variable.
-% Textile work being a salient example,
-% some seamstresses might be better than others at the same task,
-% making this framing of the work initially problematic.
-
-% A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
-% manageable routines that could be taught relatively easily,
-% and whose work output could be evaluated in abstraction from the rest of the work.
-% In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-% but a single very narrowly defined action that needed to be done on every car.
-
-% This approach paralleled what would be known as ``Taylorism'',
-% and its influences can be seen today in crowd work and micro--tasks.
-% \textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
-% horizontal cuts instead of vertical}
-% \cite{cheng2015break,writingMicroTasks}.
-% By decomposing and recomposing tasks,
-% and in particular by assigning similarly natured work to the same workers,
-% workers could become ``experts'' in a small aspect of the work that they did,
-% speeding their work dramatically
-% \cite{delayAndOrderLasecki}.
-% Perhaps more important, however, was that
-% the breaking down of work into tasks has made it more practical to evaluate work at each stage
-% [\textit{citation needed}].
-
-% Opening the door to more discrete modules of work known as tasks 
-% Contributions such as
-% ``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
-% and
-% ``Cascade'' % (described in \citetitle{chilton2013cascade})
-% make it easier to assign smaller components of work to myriad workers and
-% recompose the constituent parts into something more complex
-% \cite{bernsteinSoylent,chilton2013cascade}.
-% Some work has taken special interest in the process of breaking tasks down into more manageable parts
-% inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
-% both for the purposes of crowd--sourcing in the sense of out--sourcing work
-% as well as ``selfsourcing''
-% \cite{cheng2015break,selfsourcingTeevan2014}.
-
-% The parallels between this kind of work and
-% historical piecework of the turn of the \nth{20} century
-% is by no means new
-% [\textit{citation needed --- who said this first?
-most notably?
-Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
-% \textit{Elaborate on gig work, piecework, etc\dots}
-
-% The \textbf{discretization} and decomposition of work certainly facilitates
-% breaking tasks down,
-% parallelizing work,
-% and getting more done across a broader array of people;
-% but the key advantage in turning macro tasks 
-% into many micro tasks is that we can make that work \textit{routine}.
-
-% By making larger bodies of work less contextually situated,
-% we can define processes that make that work
-% and the instructions thereof
-% useful for virtually anyone with shared cultural or intellectual background.
-
-% routinization can be seen in the contemporary in work--flows
-% \cite{foundry,bernsteinSoylent}
-% [\textit{There are various papers that talk about creating a work--flow;
-% maybe this is where FFV belongs? Foundry?}].
-
-% \subsection{External Management}
-% \textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
-% manage work more meaningfully.
-% Routinization in particular made it possible
-% (and indeed preferable) to assume that people
-% ``\dots are of equal ability and \dots can do any of the $n$ jobs''
-% \cite{hu1961parallel}.
-% \citeauthor{storiesIraniSilberman}
-% find similarly that
-% this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
-% \cite{storiesIraniSilberman}.
-
-% Crucially,
-% \citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
-% which in \citeyear{hu1961parallel}
-% meant that workers would at least have to be trained in person, if not instructed as well.
-% Throughout the industrial revolution, workers flocked to centralized locales of production,
-% and physically situated factories made it both
-% difficult to ignore the general pattern of mistreatment and
-% relatively easy to organize.
-
-% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
-
-% \textit{I want to cite Justin's work on crowdsourcing effort to highlight that
-% we can evaluate workers for the work they do with increasingly finite amounts of time;
-% also, Ranjay's work that embraces failure because
-% we can essentially algorithmically make sense of failure (let's call it noise)
-% and re--capture signal.}
+\subsubsection{Alienated Workers}
+\itshape
+
+Marx's big entrance. Lily has lots of work pointing to this, and
+lots of that work points to yet more work gesturing in this direction.
+I want to bring a lot of that work
+that hasn't made it into the discussion yet
+into the conversation here.
+
+Note to self: examples of citations would help make it clearer for MSB which ones I'm overlooking,
+if nothing else
+
+\upshape
+
+\subsubsection{Resistance}
+
+\citeauthor{uberAlgorithm}'s paper on Uber drivers
+\cite{uberAlgorithm}.
+
+\citeauthor{dynamo} on \textit{Dynamo};
+\citeauthor{turkopticon} on \textit{Turkopticon}
+\cite{turkopticon,dynamo}.
 
@@
 
-\subsection{Looking forward}
-If we agree that
-discretization,
-routinization,
-management,
-and even resistance
-necessarily follow one another according to this theoretical lens,
-then we have to use it to attempt to envision
-what comes next.
-
-\textit{I've been going back and forth regarding how to frame this paper;
-the approach that tries to look at
-things that are similar vs things that are different
-doesn't seem to work, but
-I've left the thoughts here because it's not all completely bad.}
-
-\section{Things stay the same}
-How is gig work the same as it's been historically?
-
-I think this section would be compelling to draw parallels between
-the narratives drivers gave about the flexibility, autonomy, etc\dots
-and that which we might have seen among pieceworkers
-(predominantly women, who benefited from being able to work from home).
-
-\subsection{Flexibility}
-Are there cultural differences between the people that did piecework and
-the people that do gig work now?
-I'm not sure there are significant differences that have affected the outcomes so far.
-
-Many of the workers to whom piecework appealed were mothers, wives, etc\dots
-who mostly stayed at home for various reasons
-(certainly largely it was cultural
---- women weren't afforded equal access to labor opportunities,
-making in--home job opportunities not only compelling, but also one of few available options).
-
-Gig workers are in some senses similarly constrained:
-workers on Amazon Mechanical Turk
---- those that use it as a primary source of income, at least ---
-report being homebound for various reasons (e.g. medical, parenting, etc\dots).
-Society and circumstance have made it difficult
-or impossible
-to join the contemporary, conventional workforce;
-gig work re--opens that door.
-
-But there are differences; during our research over the summer,
-we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
-Many of them told us about their home lives
---- about children, spouses, and other commitments ---
-to which they wanted to dedicate more time.
-One driver
-(let's call him Ra\'{u}l)
-told me about how he drove for Lyft
-after working as an inventory manager at a hospital for more than nine years.
-
-I asked him why he quit that job and forewent
-the benefits,
-predictability,
-and career growth opportunity
-that his old job offered.
-He told me that when his daughter was born,
-he was overcome with a desire to spend more time with her.
-No longer satisfied with work
-where he often left before his daughter would wake up and return after she fell asleep,
-Ra\'{u}l decided to start driving for Lyft,
-because he could drive in the evenings when his daughter was asleep.
-
-Other drivers reported similar benefits;
-gig work affords its workers flexibility that conventional careers don't allow.
-
-\subsection{routinization of work}
-This leads into the next section, but I want to bring up the process of
-making work about mass--manufacturing, at least inasmuch as
-the instructions are the same for everyone.
-
-The Internet arguably has made it much easier to broadcast those instructions,
-but it's had this deeper effect of enabling some amount of back--and--forth
-between the worker and the (algorithmic) manager.
-
-\subsection{Taylorism}
-The routinization of work makes it possible to measure that process,
-optimize it for certain characteristics, and ultimately
-lead to Taylorism and scientific management.
-This is not new;
-researchers have studied and written about the slow creep of
-algorithmic management and 
-discretization \& routinization of work tasks.
-
-\textit{We hope to take a step back from the context in which this work is often applied, and
-look for its place in the larger trends and theories
-to make sense of the trends of gig work at large.}
-
-Industrialization and the automobile assembly line makes this famous,
-but piecework functioned on the principle that everyone was making similar
-or identical
-garments and other products.
-
-Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
-sometimes bootstrapped,
-as in Ranjay's talk on 
-``Embracing Error to Enable Rapid Crowdsourcing'',
-but more conventionally in work flows like ``Find--Fix--Verify''.
-
-\textit{This might be an opportunity to reflect on
-how pieceworkers internalized the work they were doing,
-responded to the stress of the uncertainty of potentially rejected work,
-etc\dots but I'm not familiar with research in that space.}
-
-\subsection{The emergence of decentralized workplaces}
-The practice of in--home piecework was consumed by
-the centralization of factories
-(the effect of which we'll talk about later, since
-we can talk about how this made unions more practical),
-but for a time many pieceworkers
-at the turn of the \nth{20} century
-worked out of their homes.
-Strikingly, many of the gig workers
-\citeauthor{crowdworkFuture}
-discussed in
-\citeyear{crowdworkFuture},
-and indeed many more continue to do what we might call ``information work''
---- that is, work that predominantly demands human computation ---
-but increasingly we're seeing the movement toward transient work that largely requires embodied presence
-\cite{uberAlgorithm} (and others).
-
-\subsection{What was this about?}
-I was going to check in quickly and take a photo of your whiteboard
-(hoping that you'd kept our conversation notes around)
-but your room was \textit{literally} full of people lol.
-
-\section{Things are changing}
-The medium on which this work is being done
---- and to an extent the medium used to manage workers ---
-has dramatically changed things as well, however;
-workers are distributed around the world,
-working out of their cars in the cases of livery services
-(notably, never returning to a base of operations)
-across and between cities as well as nations,
-or in their homes
-(paralleling the trend of piecework even more closely).
-
-Trying to understand how gig work has differed from piecework should at least
-start with looking at the different characteristics of the work involved.
-After that, we should think about how
-the demography and culture of the people engaging in this kind of work
-have changed versus that of the pieceworkers.
-
-\subsection{Differences in the work itself}
-Gig work has all of the above similarities with piecework,
-but there are key differences.
-
-% Piecework emerged at the turn of the \nth{20} century,
-% right at the time that telecommunications began to boom in the United States. % too far a reach?
-% \textit{Did telecom enable remote management?}
-
-Gig work in its contemporary formation is largely mediated by
-ubiquitously accessible digital media
-(the Internet, telephony, etc\dots)
-and importantly has relied on this technology to facilitate the remote management of workers
-\cite{uberAlgorithm}.
-
-\textit{Has the work fundamentally changed,
-or are we just being managed remotely in different ways?
-I'm not entirely sure.}
-
-How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
-How does this work differ from the work in which truckers, taxi drivers, and other
-independent contractors have been participating for decades?
-We argue that the substantive difference in these markets is the speed of the market itself,
-motivated by the technology which mediates it.
-Because workers can be sourced and dispatched virtually instantly,
-businesses that engage in this kind of work
-(e.g. Uber, Amazon Mechanical Turk, etc\dots)
-have taken to removing other bottlenecks,
-like vetting workers upfront
-
-\section{More deliberate work}
+\section{Bleak Futures}
+We've traced a path from piecework itself through
+the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
+from there, we followed the trail from decomposition to work abstraction,
+and the potential to break work away from workers themselves,
+which in turn facilitates flexibility for requesters as well as workers
+
+\subsection{Factories}\label{sec:Factorization}
+
+\itshape
+So work output is variable, and that's bad. What are we doing about it?
+Historically, this is what led to factories; you would pay some experts who understood what they needed to do
+and they would form a cohort into which you invest more.
+
+Some research already looks at stuff like investing in workers
+\cite{shepherdingDow}.
+What else is there that kind of goes in this direction?
+Informally, we know that this happens a lot
+\cite{jonBrelig}.
+AMT, meanwhile, offers requesters the ability to create tasks which are
+not just hidden from unqualified workers by default, but completely.
+Requesters have taken to using lists of worker IDs which reference
+workers who have proven their reliability,
+representing a sort of proto--organization of loosely connected workers.
+\upshape
+
+\section{Implications for Design}
+If it's agreed that
+the major topics we've discussed thus far are related and
+--- at least to \textit{some} extent ---
+precipitated in the fashion we argue,
+then we have a rare opportunity as researchers,
+and as agents of change in the communities we study,
+to affect change on the dynamics of crowd and on--demand work
+as they continue to develop.
+
+Without claiming to have easy, cut--and--dry solutions to these problems,
+we can nevertheless bring to attention a number of critical opportunities to
+learn from historical parallels in piecework and factory labor,
+and make informed decisions regarding whether
+(or indeed how)
+we may want to influence outcomes.
+The challenges we bring to attention here are as follows:
+\begin{inlinelist}
+  \item codifying investment toward collective goods into the designs of systems;
+  \item (re--)decentralizing the internet (hey, it's lowercase now!); and
+  \item enabling reputation transferral.
+\end{inlinelist}
+
+\subsection{Codify the common good}
+\subsection{Decentralize the internet (again)}
+\subsection{Deal with reputation}
+
+\rule{\linewidth}{1pt}
+
+% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
+
+% \textit{I've been going back and forth regarding how to frame this paper;
+% the approach that tries to look at
+% things that are similar vs things that are different
+% doesn't seem to work, but
+% I've left the thoughts here because it's not all completely bad.}
+
+% \section{Things stay the same}
+% How is gig work the same as it's been historically?
+
+% I think this section would be compelling to draw parallels between
+% the narratives drivers gave about the flexibility, autonomy, etc\dots
+% and that which we might have seen among pieceworkers
+% (predominantly women, who benefited from being able to work from home).
+
+% \subsection{Flexibility}
+% Are there cultural differences between the people that did piecework and
+% the people that do gig work now?
+% I'm not sure there are significant differences that have affected the outcomes so far.
+
+% Many of the workers to whom piecework appealed were mothers, wives, etc\dots
+% who mostly stayed at home for various reasons
+% (certainly largely it was cultural
+% --- women weren't afforded equal access to labor opportunities,
+% making in--home job opportunities not only compelling, but also one of few available options).
+
+% Gig workers are in some senses similarly constrained:
+% workers on Amazon Mechanical Turk
+% --- those that use it as a primary source of income, at least ---
+% report being homebound for various reasons (e.g. medical, parenting, etc\dots).
+% Society and circumstance have made it difficult
+% or impossible
+% to join the contemporary, conventional workforce;
+% gig work re--opens that door.
+
+% But there are differences; during our research over the summer,
+% we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
+% Many of them told us about their home lives
+% --- about children, spouses, and other commitments ---
+% to which they wanted to dedicate more time.
+% One driver
+% (let's call him Ra\'{u}l)
+% told me about how he drove for Lyft
+% after working as an inventory manager at a hospital for more than nine years.
+
+% I asked him why he quit that job and forewent
+% the benefits,
+% predictability,
+% and career growth opportunity
+% that his old job offered.
+% He told me that when his daughter was born,
+% he was overcome with a desire to spend more time with her.
+% No longer satisfied with work
+% where he often left before his daughter would wake up and return after she fell asleep,
+% Ra\'{u}l decided to start driving for Lyft,
+% because he could drive in the evenings when his daughter was asleep.
+
+% Other drivers reported similar benefits;
+% gig work affords its workers flexibility that conventional careers don't allow.
+
+% \subsection{routinization of work}
+% This leads into the next section, but I want to bring up the process of
+% making work about mass--manufacturing, at least inasmuch as
+% the instructions are the same for everyone.
+
+% The Internet arguably has made it much easier to broadcast those instructions,
+% but it's had this deeper effect of enabling some amount of back--and--forth
+% between the worker and the (algorithmic) manager.
+
+% \subsection{Taylorism}
+% The routinization of work makes it possible to measure that process,
+% optimize it for certain characteristics, and ultimately
+% lead to Taylorism and scientific management.
+% This is not new;
+% researchers have studied and written about the slow creep of
+% algorithmic management and 
+% discretization \& routinization of work tasks.
+
+% \textit{We hope to take a step back from the context in which this work is often applied, and
+% look for its place in the larger trends and theories
+% to make sense of the trends of gig work at large.}
+
+% Industrialization and the automobile assembly line makes this famous,
+% but piecework functioned on the principle that everyone was making similar
+% or identical
+% garments and other products.
+
+% Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
+% sometimes bootstrapped,
+% as in Ranjay's talk on 
+% ``Embracing Error to Enable Rapid Crowdsourcing'',
+% but more conventionally in work flows like ``Find--Fix--Verify''.
+
+% \textit{This might be an opportunity to reflect on
+% how pieceworkers internalized the work they were doing,
+% responded to the stress of the uncertainty of potentially rejected work,
+% etc\dots but I'm not familiar with research in that space.}
+
+% \subsection{The emergence of decentralized workplaces}
+% The practice of in--home piecework was consumed by
+% the centralization of factories
+% (the effect of which we'll talk about later, since
+% we can talk about how this made unions more practical),
+% but for a time many pieceworkers
+% at the turn of the \nth{20} century
+% worked out of their homes.
+% Strikingly, many of the gig workers
+% \citeauthor{crowdworkFuture}
+% discussed in
+% \citeyear{crowdworkFuture},
+% and indeed many more continue to do what we might call ``information work''
+% --- that is, work that predominantly demands human computation ---
+% but increasingly we're seeing the movement toward transient work that largely requires embodied presence
+% \cite{uberAlgorithm} (and others).
+
+% \subsection{What was this about?}
+% I was going to check in quickly and take a photo of your whiteboard
+% (hoping that you'd kept our conversation notes around)
+% but your room was \textit{literally} full of people lol.
+
+% \section{Things are changing}
+% The medium on which this work is being done
+% --- and to an extent the medium used to manage workers ---
+% has dramatically changed things as well, however;
+% workers are distributed around the world,
+% working out of their cars in the cases of livery services
+% (notably, never returning to a base of operations)
+% across and between cities as well as nations,
+% or in their homes
+% (paralleling the trend of piecework even more closely).
+
+% Trying to understand how gig work has differed from piecework should at least
+% start with looking at the different characteristics of the work involved.
+% After that, we should think about how
+% the demography and culture of the people engaging in this kind of work
+% have changed versus that of the pieceworkers.
+
+% \subsection{Differences in the work itself}
+% Gig work has all of the above similarities with piecework,
+% but there are key differences.
+
+% % Piecework emerged at the turn of the \nth{20} century,
+% % right at the time that telecommunications began to boom in the United States. % too far a reach?
+% % \textit{Did telecom enable remote management?}
+
+% Gig work in its contemporary formation is largely mediated by
+% ubiquitously accessible digital media
+% (the Internet, telephony, etc\dots)
+% and importantly has relied on this technology to facilitate the remote management of workers
+% \cite{uberAlgorithm}.
+
+% \textit{Has the work fundamentally changed,
+% or are we just being managed remotely in different ways?
+% I'm not entirely sure.}
+
+% How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
+% How does this work differ from the work in which truckers, taxi drivers, and other
+% independent contractors have been participating for decades?
+% We argue that the substantive difference in these markets is the speed of the market itself,
+% motivated by the technology which mediates it.
+% Because workers can be sourced and dispatched virtually instantly,
+% businesses that engage in this kind of work
+% (e.g. Uber, Amazon Mechanical Turk, etc\dots)
+% have taken to removing other bottlenecks,
+% like vetting workers upfront
+
+% \section{More deliberate work}
 % \subsection{Differences in the people that do the work}

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
comments

@@
 \section{Introduction}
+\msb{I think a contribution of this paper can be, and should be, unifiying crowd work, sharing economy, and gig economy all under one theoretical banner: piecework.
+Come out and say that super visibly somewhere in this intro as a major contribution.}
+
 The past decade has seen
-microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
+paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
 grow to occupy the fascination of both academic circles and our culture as a whole
@@
 \cite{turkopticon,dynamo,uberAlgorithm}.
-But the connections between
+\msb{A weak differentiating sentence: } But the connections between
 historical piecework and
@@
 % um, note to self: say more.
+\msb{This paragraph doesn't set up the tension and release effectively.
+I suggest focusing this paragraph on a brief summary of the things that people have pointed out (low pay, Taylorism, difficult collective action).
+Basically paint them as a bunch of disconnected observations --- which is what they have been.
+Then at the start of the next paragraph, come out with a clear and visible thesis: say that all of these things fall under the banner of piecework.
+(and now you can finally mention that while piecework has been brought up, nobody has made the historical connections yet) Then you can move on to the next paragraph below, which is about how it's basically a necessary outcome.}
 
@@
 the process of factorization that took place in the early \nth{20} century;
-indeed, these phenomena precipitate and reinforce each other.
+indeed, these phenomena \textit{precipitate and reinforce each other}.
 
@@
 managing workers more broadly,
-Crowd--powered has continued to parallel historical piecework's trajectory
-by outgrowing ad--hoc worker groups and coalescing into more formal working groups;
+Crowd--power has continued to parallel historical piecework's trajectory
+by outgrowing ad--hoc worker groups and coalescing into more formal working groups \msb{give an example, since the factorization claim is really gutsy right now otherwise};
 we might call this the beginning of factorization.
@@
 
-Making sense of this 
+Making sense of this \msb{of what?} 
 by looking at the movements toward
@@
 
-This paper will attempt to demonstrate that the familiarity is more than passing;
+This paper will attempt to demonstrate that the familiarity \msb{what familiarity?} is more than passing;
 that the phenomena researchers have observed in on--demand digitally mediated labor markets
@@
 many if not all of the frustrations researchers have discovered among workers
-in the study of crowd and gig work.
+in the study of crowd and gig work. \msb{nice paragraph}
 
@@
 For one, information work
-(e.g. AMT \& Upwork)
+(e.g. Amazon Mechanical Turk (AMT) \& Upwork)
 and digitally managed work in general make it substantially easier
@@
 not the \textit{duration})
-\cite{crowdworkFuture}.
+\cite{crowdworkFuture}. \msb{split up that sentence. One is the def'n of piecework, the other is the comparison you're drawing}
 % \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
@@
 Given the scope,
-we can frame piecework and gig work as sharing these important similarities:
+we can frame piecework and gig work \msb{wait, are you using gig work as the umbrella term? I think you need a category term that covers all of them} as sharing these important similarities:
 \begin{inlinelist}
@@
 This paper attempts to make sense of the broader research on this new form of labor,
-or ``gig work'', by evaluating this work through a more familiar lens: piecework.
+or ``gig work'' \msb{so is ``gig work'' your umbrella term? we need a definition of this at some point}, by evaluating this work through a more familiar lens: piecework.
 More concretely, by looking at task--based or ``gig'' work as
@@
 \section{Case Studies}
-The existing body of research has shed light on on--demand from various perspectives,
+The existing body of research has shed light on on--demand \msb{missing noun?} from various perspectives,
 and revealed a number of topics that,
@@
 
+\msb{the past two paragraphs feel to me like they should be riddled with citations to the historical literature. They feel kind of bald without them.}
+
 This approach paralleled what would be known as ``Taylorism'',
@@
 \cite{bernsteinSoylent}.
+\msb{given future focus on Foundry, this is too bernstein-centric.
+instead, talk about crowdforge's map-reduce framework and Little's Human Computation Algorithms HCOMP 2009 with its iterative algorithms}
 \citeauthor{cheng2015break} found that microtasks
@@
 \subsubsection{Flexibility}\label{sec:flexibility}
+\msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
+
 Earlier we discussed \citeauthor{cheng2015break}'s work
@@
 both the study and practice
-of crowd work that microtasks should be designed flexibly,
+of crowd work that microtasks should be designed flexibly \msb{not sure what flexibly means here, and the rest of the sentence isn't helping me unpack it.
+flexibility for requester?
+worker?
+flexible algorthms?},
 fully exploiting the abstracted nature of each piece of work
@@
 
+\msb{I still dont' have a clear conceptual picture of what flexibility is, and what about it sets this section apart. Ranjay's stuff is about speed...? I'm confused.}
+
 \subsection{The Fallout of Crowd Work}\label{sec:Fallout}
@@
 \subsubsection{Low Pay}\label{sec:lowPay}
+\msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
+
 \citeauthor{laborEconomicsOfCrowdsourcingHorton}
@@
 More cynically,
-we would describe this work as identifying optimal levels to activate race conditions among piece workers.
+we would describe this work as identifying optimal levels to activate race conditions \msb{???} among piece workers.
 
@@
 \cite{gantt1913work}.
-\ali{Do I need to follow up or does this basically speak for itself?}
+\ali{Do I need to follow up or does this basically speak for itself?} \msb{looks good}
 
@@
 and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-Indeed,
+Indeed, \msb{finished sentence?}
 

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
comments that i should've pushed 6 hours ago yay for falling asleep

--- 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
@@
 \subsection{Codify the common good}
+As \citeauthor{lessig2006code} points out in his book,
+digital media give designers the opportunity to design and build into the systems
+policies and practices to contribute to the collective benefit of the people therein
+\cite{lessig2006code}.
+Historically, the confluence of forces \citeauthor{lessig2006code} describes
+would ultimately result in outcomes such as benefits for workers,
+funds for sick leave and vacation, and other conveniences.
+The transient nature of on--demand work would seem to problematize this arrangement,
+but we can discuss and explore the viability of building into systems the mechanisms necessary
+to save a portion of payment from every gig,
+record taxable income, or
+myriad other generally administrative tasks automatically.
+
 \subsection{Decentralize the internet (again)}
+Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
+forcing workers either to choose one or juggle participation in these markets with great difficulty.
+Drivers
+
 \subsection{Deal with reputation}
 
-\rule{\linewidth}{1pt}
+% \rule{\linewidth}{1pt}
 

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex

@@
 \section{Introduction}
+% \msb{I think a contribution of this paper can be, and should be, unifiying crowd work, sharing economy, and gig economy all under one theoretical banner: piecework.
+Come out and say that super visibly somewhere in this intro as a major contribution.}
+
 The past decade has seen
-microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
+paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
 grow to occupy the fascination of both academic circles and our culture as a whole
@@
 as well as a number of observations relating to
-the frustration workers feel stemming from the management of this work
+the frustration workers feel stemming from the management of this work,
+but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
+or crowd work itself,
+has proven difficult;
+% as a result,
+the connection between crowd work and a historical parallel contextualizing its ongoing developments 
+hasn't been deeply made yet
 \cite{turkopticon,dynamo,uberAlgorithm}.
-But the connections between
-historical piecework and
-contemporary crowd work
-haven't been deeply made as yet.
+% \msb{A weak differentiating sentence: }
+% But the connections between
+% historical piecework and
+% contemporary crowd work
+% haven't been deeply made as yet.
 % um, note to self: say more.
+% \msb{This paragraph doesn't set up the tension and release effectively.
+I suggest focusing this paragraph on a brief summary of the things that people have pointed out (low pay, Taylorism, difficult collective action).
+Basically paint them as a bunch of disconnected observations --- which is what they have been.
+Then at the start of the next paragraph, come out with a clear and visible thesis: say that all of these things fall under the banner of piecework.
+(and now you can finally mention that while piecework has been brought up, nobody has made the historical connections yet) Then you can move on to the next paragraph below, which is about how it's basically a necessary outcome.}
 
@@
 managing workers more broadly,
-Crowd--powered has continued to parallel historical piecework's trajectory
-by outgrowing ad--hoc worker groups and coalescing into more formal working groups;
-we might call this the beginning of factorization.
-% "Factorization" is not the right word but I honestly have no idea what the right term is.
-% The process of formalizing work into factories? Should I tweet/post/ask about this somewhere?
+crowd--powered work has continued to parallel historical piecework's trajectory
+by outgrowing ad--hoc worker groups and coalescing into more formal working groups,
+for example through the use of a ``white list'' of familiar workers.
+We might call this the regularization of factory work.
 
-% Crowd work,
-% they argued in \citeyear{crowdworkFuture},
-% represents a threat to the future of work inasmuch as it marginalized
-% and perhaps even harmed
-% workers;
-% their driving question, then, was
-% whether the future of
-% (crowd, but arguably all discretized)
-% work might be one in which they wish their children to participate someday;
-% it seems strikingly like one that might have been asked
-% as patches of denim were first being delivered to the homes of early pieceworkers.
-% agreed; i'm commenting it out for now to see how I feel without making this argument at all. If I really feel like it's missing, I should find evidence that contemporaries discussed it...
-
-Making sense of this 
+Making sense of the broader field of crowd work
 by looking at the movements toward
@@
 
-This paper will attempt to demonstrate that the familiarity is more than passing;
+This paper will attempt to demonstrate that the familiarity
+that we incidentally observe between what we generally call crowd work \& gig work
+and the historical practice of piecework and subsequent factory work
+is more than passing;
 that the phenomena researchers have observed in on--demand digitally mediated labor markets
@@
 write of the future of crowd work in an effort to investigate
-whether crowd work will be a worthwhile form of work or another instantiation of ``piecework'',
+whether crowd work will be a worthwhile form of work or another instantiation of ``piecework''
+\cite{crowdworkFuture}.
+``Piecework'' here makes use of
 a term historically used to describe work done in the home,
@@
 one would be paid for the \textit{output} of the work,
-not the \textit{duration})
-\cite{crowdworkFuture}.
-% \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
-% This work was largely in textiles and was predominantly done by women,
-% which by itself is a topic worthy of unpacking
-% \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
-% \ali{let's just leave this in the graveyard for now; i feel like someone who knows what they're talking about with regard to feminist and gender studies
-% (e.g. the Bardzells, Irani) would really do justice here, but i'd be getting in over my head}
-% \cite{scott1975women}.
+not the \textit{duration}).
+% \msb{split up that sentence. One is the def'n of piecework, the other is the comparison you're drawing}
 Given the scope,
-we can frame piecework and gig work as sharing these important similarities:
+we can frame piecework and on--demand labor
+% \msb{
+% \sout{wait, are you using gig work as the umbrella term? I think you need a category term that covers all of them}
+% \ali{let's use ``on--demand labor'', since
+% that seems general enough without accidentally including historical piecework}}
+as sharing these important similarities:
 \begin{inlinelist}
@@
 
-This paper attempts to make sense of the broader research on this new form of labor,
-or ``gig work'', by evaluating this work through a more familiar lens: piecework.
+This paper attempts to make sense of the broader research on this new form of work,
+\textit{on--demand labor}, by evaluating this work through a more familiar lens: piecework.
 More concretely, by looking at task--based or ``gig'' work as
@@
 \section{Case Studies}
-The existing body of research has shed light on on--demand from various perspectives,
+The existing body of research has shed light on on--demand labor from various perspectives,
 and revealed a number of topics that,
@@
 Broken down in this way, work could grow to unprecedented scales,
-but the quality of the work would remain relatively variable.
+but the quality of the work would remain relatively variable
+\cite{murray1983decentralisation}.
 Textile work being a salient example,
 it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work.
-% some seamstresses might be better than others at the same task,
-% making this framing of the work initially problematic.
+to do every aspect of the work so that the garment would be accepted by the company soliciting that work
+\cite{vezina1992light}.
 
@@
 manageable routines that could be taught relatively easily,
-and whose work output could be evaluated in abstraction from the rest of the work.
+and whose work output could be evaluated in abstraction from the rest of the work
+\cite{restructuringPieceworkBaker}.
 In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-but a single very narrowly defined action that needed to be done on every car.
+but a single very narrowly defined action that needed to be done on every car
+\cite{towardsGlobalFordism}.
+By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,
+``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
+\cite{schoenberger1988fordism}.
 
-This approach paralleled what would be known as ``Taylorism'',
-and its influences can be seen today in crowd work and micro--tasks.
-\textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
-horizontal cuts instead of vertical}
-\cite{cheng2015break,writingMicroTasks}.
+This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),
+can be seen today in crowd work and on--demand labor through the application of micro--tasks.
+\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,
+facilitating evaluation and parallelization
+\cite{writingMicroTasks}.
 By decomposing and recomposing tasks,
 and in particular by assigning similarly natured work to the same workers,
-% and especially as the nature of the work could be aligned with other, similar tasks,
 workers could become ``experts'' in a small aspect of the work that they did,
@@
 the breaking down of work into tasks has made it more practical to evaluate work at each stage
-[\textit{citation needed}].
+\cite{rogstadius2011assessment}.
 
@@
 the decomposition of tasks is perhaps the most established of the above areas among HCI researchers;
-\citeauthor{bernsteinSoylent} made early contributions to this area with \textit{Soylent}
-(by coining and employing the ``find--fix--verify'' approach)
-\cite{bernsteinSoylent}.
+\citeauthor{crowdForgeKittur} specifically drive at this goal by addressing the possibility of
+``crowdsourcing complex work''
+\cite{crowdForgeKittur}.
 \citeauthor{cheng2015break} found that microtasks
@@
 \cite{riisOtherSideLives}.
-
-% In both the historical and contemporary cases,
-% distributed, speculative work
-% has promised
-% to diffuse the costs of capital--intensive resources
-% (whether they were sewing machines or laboratory equipment).
 
@@
 \subsubsection{Flexibility}\label{sec:flexibility}
+% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
+
 Earlier we discussed \citeauthor{cheng2015break}'s work
@@
 both the study and practice
-of crowd work that microtasks should be designed flexibly,
+of crowd work that microtasks should be designed resiliently against the variability of workers,
+% \msb{not sure what flexibly means here, and the rest of the sentence isn't helping me unpack it. flexibility for requester? worker? flexible algorthms?},
 fully exploiting the abstracted nature of each piece of work
 \cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
+That is to say, micro--tasks should be designed such that a single worker's poor performance,
+or a good worker's sudden departure,
+would not significantly impact the agenda of the work as a whole.
 While \citeauthor{cheng2015break} identified costs with breaking tasks into smaller components
@@
 in various binary classification tasks
-% --- a typical application crowd work ---
-on the principle that myriad diverse workers complete these tasks
-in order to accurately inform the model.
+on the principle that diverse workers complete these tasks
+in order to accurately inform the model on the variety of delays in response times.
 And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
@@
 \cite{sensitiveTasks}.
+
+Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
+\citeauthor{tolliday1986between}'s treatment describing
+turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
+Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
+[and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
+piece rates~\dots~rather than a day wage system''
+\cite{tolliday1986between}.
+
+% \msb{I still dont' have a clear conceptual picture of what flexibility is, and what about it sets this section apart. Ranjay's stuff is about speed...? I'm confused.}
 
@@
 \subsubsection{Low Pay}\label{sec:lowPay}
+% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
+
 \citeauthor{laborEconomicsOfCrowdsourcingHorton}
@@
 
+Turmoil over low --- and declining --- pay was among the chief grievances among then nascent
+British labor unions in the early \nth{20} century
+\cite{turner1952trade}.
+This, \citeauthor{ebbinghaus1999institutions} argued,
+fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
+(to which we'll return later when we discuss \citeauthor{levi2009union}'s reexamination of labor advocacy organizations)
+\cite{ebbinghaus1999institutions}.
+This realization has similarly fueled a body of research into
+the various incentive structures available to piecework employers
+\cite{roy1953work}.
+
+The parallels between the complaints of low pay among crowd workers and other on--demand workers
+and the pieceworkers and later factory workers in the \nth{20} century
+are inescapable.
+We argue further that the \textit{causes} here
+--- work decomposition,
+work abstraction, and
+flexibility ---
+lead inexorably to low and declining pay for workers.
+
 \subsubsection{Variable quality work}\label{sec:varQualWork}
-Meanwhile,
-researchers have struggled with what we might call ``variable quality work''
+Researchers have struggled with what we might generously call work of ``variable quality''
 along two dimensions:
@@
 to eke out the highest quality work possible from workers
-given the apparent difficulty in predicting work outcomes
+given the apparent difficulty in predicting work outcomes.
 
@@
 and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-Indeed,
+Indeed, \msb{finished sentence? \ali{Indeed.}}
 

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
cleaned out comments? feedback as i push through the italicized stuff?

@@
 \cite{incentivesShaw,paolacci2010running}.
-More cynically,
-we would describe this work as identifying optimal levels to activate race conditions \msb{???} among piece workers.
 
@@
 \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
-via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s work
+via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
 \cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
 
-Turmoil over low --- and declining --- pay was among the chief grievances among then nascent
+On--demand workers were not the first to be exploited along the dimension of low pay rates.
+Frustration over low (and declining) pay was one of the chief grievances among then nascent
 British labor unions in the early \nth{20} century
@@
 fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
-(to which we'll return later when we discuss \citeauthor{levi2009union}'s reexamination of labor advocacy organizations)
-\cite{ebbinghaus1999institutions}.
+(to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
+\cite{ebbinghaus1999institutions,levi2009union}.
 This realization has similarly fueled a body of research into
@@
 lead inexorably to low and declining pay for workers.
+Moreover, we point out that low pay leads to other negative outcomes both
+in on--demand work
+as well as
+in piecework and on assembly lines.
 
@@
 Researchers have struggled with what we might generously call work of ``variable quality''
-along two dimensions:
-First, to use the characterization of one of these contributions,
+along two dimensions.
+The first, to use the characterization of one of these contributions, we can call
 ``understanding malicious behavior''
 \cite{MaliciousCrowdworkersGadiraju}.
-While some work has cast workers as ``malicious'' agents,
+While some work has cast workers as ``malicious'' or at least adversarial parties,
 the evidence thus far suggests that
@@
 to eke out the highest quality work possible from workers
-given the apparent difficulty in predicting work outcomes.
+given the apparent difficulty in predicting work outcomes
+\cite{embracingErrorKrishna}.
 
@@
 
-The effect that poor wages have had on piece work and factory workers is hardly unknown or novel;
+The effect low wages have had on piece work and factory workers is well--known;
 \citeauthor{gantt1913work} discuss this exact mechanism in his book on
@@
 \cite{gantt1913work}.
-\ali{Do I need to follow up or does this basically speak for itself?} \msb{looks good}
 
@@
 and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-Indeed, \msb{finished sentence? \ali{Indeed.}}
 

### 2016-05-28 cscw/papers/2017/labor/trackingWork.tex
final?

@@
 
-\def\plaintitle{Naming Things is Hard: Real Title Following Colon}
+\def\plaintitle{Reexamining Crowd Work: A Historical Framing of On--Demand Labor as Piecework}
 \def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
 \def\emptyauthor{}
-\def\plainkeywords{Please don't make me pick keywords.
-This is like asking a teacher to give the bullet points of
-what a student missed in lecture.}
+\def\plainkeywords{Crowdsourcing}
 \def\plaingeneralterms{Documentation, Standardization}
@@
 \begin{abstract}
-  With growing attention on gig work
-  --- ranging from the ``sharing economy'' to microtasks ---
+  With growing attention in on--demand labor
+  --- ranging from the ``sharing economy'' to information work ---
   scholars have made connections to frameworks like Taylorism,
@@
   We argue that our the underlying trend towards ``piecework''
-  --- driven in part by the discretization,
-  routinization,
-  and external management of said work ---
-  not only suggests, but in fact generates what we have observed:
-  members of this transient workforce increasingly feeling
-  disempowered,
-  marginalized,
-  and frustrated by the systems and platforms on which they work.
+  predicts and indeed precipitates both the contributions
+  that have been made toward the development of on--demand labor and crowd work
+  as well as the fallout among workers and researchers with regard to the disillusionment and alienation of work.
+  % not only suggests, but in fact generates what we have observed:
+  % members of this transient workforce increasingly feeling
+  % disempowered,
+  % marginalized,
+  % and frustrated by the systems and platforms on which they work.
 
   After evaluating this framing
-  through a series of case studies
-  in various industries falling broadly under the ``gig work'' category,
-  we turn our theoretical lens to look to the future,
+  through a series of case studies,
+  % in various on--demand industries,
+  we turn our lens to look to the future,
   to identify worthwhile questions and
@@
 
-Suffice it to say, then, that poor pay and poor work are linked,
+\subsection{Relationships Between Workers and Managers}\label{sec:relationships}
+Suffice it to say that poor pay and poor work are linked,
 and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-
-\subsection{Relationships Between Workers and Managers}\label{sec:relationships}
+But the poor treatment of workers by managers
+--- both human and algorithmic ---
+do more than affect the economic relationships between workers and employers.
+Here, then, we turn to examine this facet of on--demand work
+and how these dynamics strikingly replicate the relationships
+researchers in labor advocacy encountered in the study of piecework and factory work.
+
+This topic can be condensed into three major areas:
+\begin{inlinelist}
+\item external (scientific) management, and the evaluation of workers as functional modules; and
+% \item the alienation of workers under this form of evaluation; and
+\item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
+\end{inlinelist}
 
 \subsubsection{External Management}
-\itshape
-
-Scientific management in other words; should I just call it that and get right into it?
-I kind of want the focus to be on the adversarial nature of the management,
-and ``external'' seems to connote that more strongly than ``scientific'',
-which has a tone of being more\dots~Measured? Objective? Better?
-
-There's a fair amount to say here, and lots to cite, so I'm not super concerned.
-I do want to bring up \citeauthor{storiesIraniSilberman}'s
-\citetitle{storiesIraniSilberman} \cite{storiesIraniSilberman}
-as a way of pointing out that even in our effort to study piecework ethnographically,
-and despite our attempts to be reflexive and conscientious of
-the power dynamics that we bring into any discussion
-(as engineers, as people with megaphones, etc\dots)
-we still manage to influence and perhaps steer.
-
-\upshape
-
-% Taylorism enabled
-% --- and would later become synonymous with --- 
-% scientific management, which we will discuss later
-% and might generally be regarded collectively as ``scientific management''
-% The compelling solution that emerged in piecework became what we now know as Taylorism;
-% rather than
-
-\subsubsection{Alienated Workers}
-\itshape
-
-Marx's big entrance. Lily has lots of work pointing to this, and
-lots of that work points to yet more work gesturing in this direction.
-I want to bring a lot of that work
-that hasn't made it into the discussion yet
-into the conversation here.
-
-Note to self: examples of citations would help make it clearer for MSB which ones I'm overlooking,
-if nothing else
-
-\upshape
+We discussed Fordism and Taylorism earlier in our discussions of
+\nameref{sec:decomposition} and \nameref{sec:flexibility},
+but here the core of these paradigmatic views
+--- the scientific management of work ---
+becomes relevant.
+We use ``external'' here instead of ``scientific'', however,
+to more broadly capture the disconnect between managers and workers.
+By describing it as thus,
+we can touch on the relationship that workers have with \textit{researchers}, as well,
+even though that work is not strictly
+--- or just not exclusively ---
+of the same nature as the management and experience as when interacting with requesters.
+
+First, intuitively, the variable--quality work we discussed previously has led to
+a large and growing body of research attempting to evaluate workers' performance and error rates
+across numerous dimensions;
+for example, \citeauthor{measuringCrowdsourcingCheng} explore the error rates of workers by
+operating on a sliding scale giving workers varying amounts of time to accomplish micro--tasks
+\cite{measuringCrowdsourcingCheng}.
+\citeauthor{storiesIraniSilberman} describe the treatment of workers
+as sorts of ``human APIs'' that can, importantly, be rigorously evaluated
+\cite{storiesIraniSilberman}.
+\citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
+cognitively demanding tasks on workers,
+informing crowdsourcing research by suggesting the use of cognitive load assessments such as
+NASA Task Load Index surveys to evaluate workers pre and post--tasks
+\cite{embracingErrorKrishna,measuringCrowdsourcingCheng}.
+
+External management comes in other forms than scientific, as previously mentioned.
+Researchers in particular have noticed that their relationships with on--demand workers are,
+at the least, complex.
+\citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
+specifically, their interactions with field sites in which they work
+as designers and mediators of change influence the relationships they have with Turkers
+\cite{storiesIraniSilberman}.
+
+The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
+and even physiological study of pieceworkers closely resembles the
+research into cognitive loads and stress levels that we discussed among on--demand crowd workers
+\cite{pieceworkBiologicalHarm,hu1961parallel}.
+Even the complicated relationships between observers and workers themselves are not necessarily new
+\cite{riisOtherSideLives}.
+\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined among Turkers
+\cite{facesOfMechanicalTurk}.
+We argue that these investigations naturally follow the flourishing and fallout of what is,
+fundamentally, piecework mediated in new ways.
+
+% \subsubsection{Alienated Workers}
+% Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
+% T
+
+% \itshape
+
+% Marx's big entrance. Lily has lots of work pointing to this, and
+% lots of that work points to yet more work gesturing in this direction.
+% I want to bring a lot of that work
+% that hasn't made it into the discussion yet
+% into the conversation here.
+
+% Note to self: examples of citations would help make it clearer for MSB which ones I'm overlooking,
+% if nothing else
+
+% \upshape
 
 \subsubsection{Resistance}
-
-\citeauthor{uberAlgorithm}'s paper on Uber drivers
+It shouldn't surprise us, then, that workers have resisted the management imposed on them both by people and by systems,
+often without recourse or room for feedback.
+Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
+many toggle their availability to avoid being dispatched to more distant locations,
+resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
 \cite{uberAlgorithm}.
 
-\citeauthor{dynamo} on \textit{Dynamo};
-\citeauthor{turkopticon} on \textit{Turkopticon}
+Resistance has sometimes been more coordinated, as well;
+we see this in \citeauthor{turkopticon}'s coverage on \textit{Turkopticon} as workers collectively accumulated information about requesters, and
+in \citeauthor{dynamo}'s work on \textit{Dynamo}, which generated ``Guidelines for Academic Requesters'' written by crowd workers
 \cite{turkopticon,dynamo}.
+
+Resistance against managers in piecework and factory labor settings are deeply well--explored,
+but perhaps the most relevant case study to draw on here is to be found in
+\citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',
+where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
+\cite{waldinger1996helots}.
+The achievements of labor advocacy groups such as labor unions as resistant,
+even adversarial organizations counter--balancing the management
+is somewhat well--understood
+\cite{russell1982collective,craig1992behavior}.
+We argue that these threads of resistance against management in various forms
+are in fact one.
 
@@
 
-\section{Bleak Futures}
+\section{The Bleak Future of Crowd Work}
 We've traced a path from piecework itself through
 the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
-from there, we followed the trail from decomposition to work abstraction,
-and the potential to break work away from workers themselves,
-which in turn facilitates flexibility for requesters as well as workers
-
-\subsection{Factories}\label{sec:Factorization}
-
-\itshape
-So work output is variable, and that's bad. What are we doing about it?
-Historically, this is what led to factories; you would pay some experts who understood what they needed to do
-and they would form a cohort into which you invest more.
-
-Some research already looks at stuff like investing in workers
-\cite{shepherdingDow}.
-What else is there that kind of goes in this direction?
-Informally, we know that this happens a lot
-\cite{jonBrelig}.
+in tracing this process, we touched on the relationships between decomposition,
+work \& worker abstraction,
+flexibility, and followed through both the general fallout of crowd work in the research community
+as well as the fallout between workers and the managers and other external parties
+--- including researchers.
+
+Throughout these case studies, we have pointed out the parallels between
+the contemporary research in on--demand labor and
+the much larger body of research constituting our understanding of topics such as
+piecework, factory work, and laborer relations.
+If we agree that this framing is useful and informative, then
+several topics emerge as relatively open questions in the study of crowd work and on--demand labor.
+Two of the most pressing questions are
+\begin{inlinelist}
+\item the beginnings of factories, and 
+\item the decline of relevance of worker advocacy organizations.
+\end{inlinelist}
+We will discuss those questions here.
+
+\subsection{The beginnings of factories}\label{sec:Factorization}
+We established earlier that abstracted work and low wages tend to result in variable outcomes,
+which presents problems for employers.
+% So work output is variable, and that's bad. What are we doing about it?
+Historically, this is what led to factories; by employing a cohort of known workers,
+we can be reasonably assured that the quality of the work will be better than random.
+Furthermore, we can invest more resources in training workers and
+get workers to do more complex work with more context.
+% you would pay some experts who understood what they needed to do
+% and they would form a cohort into which you invest more.
+
+Some research already looks at research such as investing in workers, and
+informally, we know that this happens among industry requesters
+\cite{jonBrelig,shepherdingDow}.
 AMT, meanwhile, offers requesters the ability to create tasks which are
@@
 representing a sort of proto--organization of loosely connected workers.
-\upshape
+
+This, then, suggests that the beginning of the regularization of workforces
+--- a sort of coalescence of factories ---
+is already happening.
+If our framing of on--demand labor is accurately describing an underlying relationship with piecework,
+then we should watch for the emergence and popularization of persistent teams of workers.
+% While this field site will likely prove exceedingly difficult to 
+% Social computing researchers should seek to understand this process more
+% fully, as we a
+
+\subsection{The decline of advocacy organizations}
+The rise of labor unions in the \nth{20} century seems to have been precipitated by
+severely unjust conditions imposed on workers in factories and elsewhere
+\cite{ebbinghaus1999institutions}.
+Incidents broadly describing this dynamic can be found in research on AMT
+\cite{turkopticon,dynamo}.
+If these are prototypical labor advocacy organizations of contemporary on--demand work,
+the next question we should look to is if
+--- and indeed \textit{how} ---
+these institutions might face challenges in the future.
+
+For insight on this, we return to \citeauthor{levi2009union}'s study of labor unions,
+and identify that
+``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
+\cite{levi2009union}.
+This perception already appears to be emerging in digitally mediated peer--governed organizations
+\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
+If these organizations and others are to avoid the same fate that labor unions faced,
+they should take care to study this phenomenon and attempt to avoid it.
 
@@
 
-\subsection{Decentralize the internet (again)}
+\subsection{Decentralize the internet --- again}
 Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
 forcing workers either to choose one or juggle participation in these markets with great difficulty.
-Drivers
+An ``API'' for on--demand labor markets could make it possible for any person
+or organization to instantiate their own marketplace and inter--operate with.
+This can be changed, and indeed must, if we are to realize the hopes of early researchers who advocated the
+democratizing nature and power of the internet
+\cite{barlow2009declaration,lanier2014owns}.
 
 \subsection{Deal with reputation}
+Reputation systems in on--demand labor markets are fundamentally broken.
+To say nothing of the fact that information workers (such as those on AMT) can't transfer their reputations to qualitatively different forms of labor
+like driving--for--hire (e.g. Uber), even within the same industry
+it's currently not feasible for workers to transfer their reputations or other information from one place to another.
+This affects more than the reputation and trustworthiness of workers;
+accounting for things such as taxes, benefits, etc\dots is all but left to the individual workers, who struggle with myriad bureaucratic obstacles.
+We can design systems that facilitate the aggregation and, more importantly, the transferral of reputation, income, and other features of work.
+
+\section{Discussion}
+We've discussed a number of aspects of on--demand work that
+offer parallels with historical piecework.
+Perhaps more importantly, we've hopefully demonstrated that the dynamics we observe in on--demand work
+are interrelated and follow from one another just as necessarily as they did in
+the development and maturation of piecework and factory work through the \nth{20} century.
+This framing on on--demand work should, we hope, provide us with the necessary historical context
+to make better--informed design decisions about how we want ``the future of crowd work'' to look.
+
+\section{Conclusion}
+\citeauthor{crowdworkFuture} discussed many of the challenges and problems in crowd work in \citeyear{crowdworkFuture},
+but didn't necessarily situate the notion of crowd work in a broader context.
+This paper attempts to fill that gap, and in doing so hopes
+to give the research community theoretical grounding to work with and within on--demand labor more successfully.
+But more than that, we hope to have addressed important questions to inform how we actually might make crowd work
+a career in which we want our children to work.
 

### 2016-05-28 cscw/papers/2017/labor/trackingWork.tex
inexplicably, we can keep submitting, so i added some citations i had really wanted to add

@@
 \begin{abstract}
-  With growing attention in on--demand labor
+  With growing attention toward on--demand labor
   --- ranging from the ``sharing economy'' to information work ---
-  scholars have made connections to frameworks like Taylorism,
-  and mechanisms such as worker advocacy and empowerment,
+  scholars have made connections to various frameworks 
+  and mechanisms such as worker advocacy, empowerment, and Taylorism,
   to make sense of our observations of
   on--demand work and the workers that power this movement.
-  We argue that our the underlying trend towards ``piecework''
-  predicts and indeed precipitates both the contributions
+  We argue that the literature surrounding ``piecework''
+  informs and even predicts both the contributions
   that have been made toward the development of on--demand labor and crowd work
-  as well as the fallout among workers and researchers with regard to the disillusionment and alienation of work.
-  % not only suggests, but in fact generates what we have observed:
-  % members of this transient workforce increasingly feeling
-  % disempowered,
-  % marginalized,
-  % and frustrated by the systems and platforms on which they work.
+  as well as the fallout among workers and researchers
+  with regard to the disillusionment and alienation of work.
 
@@
   through a series of case studies,
-  % in various on--demand industries,
-  we turn our lens to look to the future,
+  we look to the future
   to identify worthwhile questions and
-  points of inquiry that researchers in social computing should consider
+  points of inquiry, such as the movement toward factories,
+  that researchers in social computing should consider
   as we attempt to anticipate and perhaps shape the future of work.
@@
 \section{Introduction}
-% \msb{I think a contribution of this paper can be, and should be, unifiying crowd work, sharing economy, and gig economy all under one theoretical banner: piecework.
-Come out and say that super visibly somewhere in this intro as a major contribution.}
-
 The past decade has seen
@@
 has proven difficult;
-% as a result,
 the connection between crowd work and a historical parallel contextualizing its ongoing developments 
@@
 \cite{turkopticon,dynamo,uberAlgorithm}.
-% \msb{A weak differentiating sentence: }
-% But the connections between
-% historical piecework and
-% contemporary crowd work
-% haven't been deeply made as yet.
-% um, note to self: say more.
-% \msb{This paragraph doesn't set up the tension and release effectively.
-I suggest focusing this paragraph on a brief summary of the things that people have pointed out (low pay, Taylorism, difficult collective action).
-Basically paint them as a bunch of disconnected observations --- which is what they have been.
-Then at the start of the next paragraph, come out with a clear and visible thesis: say that all of these things fall under the banner of piecework.
-(and now you can finally mention that while piecework has been brought up, nobody has made the historical connections yet) Then you can move on to the next paragraph below, which is about how it's basically a necessary outcome.}
 
@@
 \subsubsection{Low Pay}\label{sec:lowPay}
-% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
 
@@
 
-This topic can be condensed into three major areas:
+This topic can be condensed into two major areas:
 \begin{inlinelist}
@@
 \cite{pieceworkBiologicalHarm,hu1961parallel}.
-Even the complicated relationships between observers and workers themselves are not necessarily new
-\cite{riisOtherSideLives}.
-\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined among Turkers
-\cite{facesOfMechanicalTurk}.
-We argue that these investigations naturally follow the flourishing and fallout of what is,
-fundamentally, piecework mediated in new ways.
+Even the complicated relationships between observers and workers themselves are not necessarily new;
+\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
+to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
+\cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.
+
+Similarly, \citeauthor{pollard1963factory}'s words on the punishment factory workers faced ---
+for example, that ``unsatisfactory work was punished \dots~by fines or by dismissal'' ---
+seems especially relevant given the fears we now know to be ubiquitous on platforms such as AMT, Uber, and other on--demand markets
+\cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.
 
 % \subsubsection{Alienated Workers}
+% The dissatisfaction on crowd work systems is not as well--explored as their actual responses to that dissatisfaction
+% (in various forms of resistance),
+% but it's nevertheless worthwhile to point out some of the observations that have been made here,
+% as external management, discussed earlier, and the discussion on worker resistance, to be discussed next,
+% are bridged by this quality of on--demand labor markets.
+
 % Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
@@
 % \itshape
-
 % Marx's big entrance. Lily has lots of work pointing to this, and
@@
 \subsubsection{Resistance}
-It shouldn't surprise us, then, that workers have resisted the management imposed on them both by people and by systems,
-often without recourse or room for feedback.
+It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,
+often without recourse or opportunity for feedback, let alone substantive input.
 Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
@@
 \cite{levi2009union}.
-This perception already appears to be emerging in digitally mediated peer--governed organizations
+% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
+% unable to satisfy their needs or otherwise more invested in the 
+This perception already appears to be emerging in digitally mediated peer--governed organizations,
+as \citeauthor{keegan2010egalitarians} and others have illustratively documented
 \cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
@@
   \item codifying investment toward collective goods into the designs of systems;
-  \item (re--)decentralizing the internet (hey, it's lowercase now!); and
+  \item (re--)decentralizing the internet; and
   \item enabling reputation transferral.
@@
 
-% \rule{\linewidth}{1pt}
-
-% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case
-
-% \textit{I've been going back and forth regarding how to frame this paper;
-% the approach that tries to look at
-% things that are similar vs things that are different
-% doesn't seem to work, but
-% I've left the thoughts here because it's not all completely bad.}
-
-% \section{Things stay the same}
-% How is gig work the same as it's been historically?
-
-% I think this section would be compelling to draw parallels between
-% the narratives drivers gave about the flexibility, autonomy, etc\dots
-% and that which we might have seen among pieceworkers
-% (predominantly women, who benefited from being able to work from home).
-
-% \subsection{Flexibility}
-% Are there cultural differences between the people that did piecework and
-% the people that do gig work now?
-% I'm not sure there are significant differences that have affected the outcomes so far.
-
-% Many of the workers to whom piecework appealed were mothers, wives, etc\dots
-% who mostly stayed at home for various reasons
-% (certainly largely it was cultural
-% --- women weren't afforded equal access to labor opportunities,
-% making in--home job opportunities not only compelling, but also one of few available options).
-
-% Gig workers are in some senses similarly constrained:
-% workers on Amazon Mechanical Turk
-% --- those that use it as a primary source of income, at least ---
-% report being homebound for various reasons (e.g. medical, parenting, etc\dots).
-% Society and circumstance have made it difficult
-% or impossible
-% to join the contemporary, conventional workforce;
-% gig work re--opens that door.
-
-% But there are differences; during our research over the summer,
-% we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
-% Many of them told us about their home lives
-% --- about children, spouses, and other commitments ---
-% to which they wanted to dedicate more time.
-% One driver
-% (let's call him Ra\'{u}l)
-% told me about how he drove for Lyft
-% after working as an inventory manager at a hospital for more than nine years.
-
-% I asked him why he quit that job and forewent
-% the benefits,
-% predictability,
-% and career growth opportunity
-% that his old job offered.
-% He told me that when his daughter was born,
-% he was overcome with a desire to spend more time with her.
-% No longer satisfied with work
-% where he often left before his daughter would wake up and return after she fell asleep,
-% Ra\'{u}l decided to start driving for Lyft,
-% because he could drive in the evenings when his daughter was asleep.
-
-% Other drivers reported similar benefits;
-% gig work affords its workers flexibility that conventional careers don't allow.
-
-% \subsection{routinization of work}
-% This leads into the next section, but I want to bring up the process of
-% making work about mass--manufacturing, at least inasmuch as
-% the instructions are the same for everyone.
-
-% The Internet arguably has made it much easier to broadcast those instructions,
-% but it's had this deeper effect of enabling some amount of back--and--forth
-% between the worker and the (algorithmic) manager.
-
-% \subsection{Taylorism}
-% The routinization of work makes it possible to measure that process,
-% optimize it for certain characteristics, and ultimately
-% lead to Taylorism and scientific management.
-% This is not new;
-% researchers have studied and written about the slow creep of
-% algorithmic management and 
-% discretization \& routinization of work tasks.
-
-% \textit{We hope to take a step back from the context in which this work is often applied, and
-% look for its place in the larger trends and theories
-% to make sense of the trends of gig work at large.}
-
-% Industrialization and the automobile assembly line makes this famous,
-% but piecework functioned on the principle that everyone was making similar
-% or identical
-% garments and other products.
-
-% Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
-% sometimes bootstrapped,
-% as in Ranjay's talk on 
-% ``Embracing Error to Enable Rapid Crowdsourcing'',
-% but more conventionally in work flows like ``Find--Fix--Verify''.
-
-% \textit{This might be an opportunity to reflect on
-% how pieceworkers internalized the work they were doing,
-% responded to the stress of the uncertainty of potentially rejected work,
-% etc\dots but I'm not familiar with research in that space.}
-
-% \subsection{The emergence of decentralized workplaces}
-% The practice of in--home piecework was consumed by
-% the centralization of factories
-% (the effect of which we'll talk about later, since
-% we can talk about how this made unions more practical),
-% but for a time many pieceworkers
-% at the turn of the \nth{20} century
-% worked out of their homes.
-% Strikingly, many of the gig workers
-% \citeauthor{crowdworkFuture}
-% discussed in
-% \citeyear{crowdworkFuture},
-% and indeed many more continue to do what we might call ``information work''
-% --- that is, work that predominantly demands human computation ---
-% but increasingly we're seeing the movement toward transient work that largely requires embodied presence
-% \cite{uberAlgorithm} (and others).
-
-% \subsection{What was this about?}
-% I was going to check in quickly and take a photo of your whiteboard
-% (hoping that you'd kept our conversation notes around)
-% but your room was \textit{literally} full of people lol.
-
-% \section{Things are changing}
-% The medium on which this work is being done
-% --- and to an extent the medium used to manage workers ---
-% has dramatically changed things as well, however;
-% workers are distributed around the world,
-% working out of their cars in the cases of livery services
-% (notably, never returning to a base of operations)
-% across and between cities as well as nations,
-% or in their homes
-% (paralleling the trend of piecework even more closely).
-
-% Trying to understand how gig work has differed from piecework should at least
-% start with looking at the different characteristics of the work involved.
-% After that, we should think about how
-% the demography and culture of the people engaging in this kind of work
-% have changed versus that of the pieceworkers.
-
-% \subsection{Differences in the work itself}
-% Gig work has all of the above similarities with piecework,
-% but there are key differences.
-
-% % Piecework emerged at the turn of the \nth{20} century,
-% % right at the time that telecommunications began to boom in the United States. % too far a reach?
-% % \textit{Did telecom enable remote management?}
-
-% Gig work in its contemporary formation is largely mediated by
-% ubiquitously accessible digital media
-% (the Internet, telephony, etc\dots)
-% and importantly has relied on this technology to facilitate the remote management of workers
-% \cite{uberAlgorithm}.
-
-% \textit{Has the work fundamentally changed,
-% or are we just being managed remotely in different ways?
-% I'm not entirely sure.}
-
-% How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
-% How does this work differ from the work in which truckers, taxi drivers, and other
-% independent contractors have been participating for decades?
-% We argue that the substantive difference in these markets is the speed of the market itself,
-% motivated by the technology which mediates it.
-% Because workers can be sourced and dispatched virtually instantly,
-% businesses that engage in this kind of work
-% (e.g. Uber, Amazon Mechanical Turk, etc\dots)
-% have taken to removing other bottlenecks,
-% like vetting workers upfront
-
-% \section{More deliberate work}
-% \subsection{Differences in the people that do the work}
-% ???
-
 \balance{}

### 2016-06-15 chi/papers/2017/referrals/referrals.tex
total rewrite of the tracking labor intro

@@
 
-Hi there.
+\itshape
+Sample intro. Does this sell the idea?
+\upshape
+
 \section{Introduction}
-The practice of referring work to other skilled workers has been well--studied in economics and more broadly in the social sciences for decades
-[\textit{citation needed}].
-But researchers in crowd work haven't sufficiently explored the potential
-to employ this characteristic of socially connected workers
-to give professionals opportunities to better route tasks.
-One possible reason for this may have been the assumption that workers are not necessarily well--connected.
-This would be an intuitively reasonable assumption
---- with no in--built forums or communities for workers on various gig labor markets
-(e.g. Uber, Amazon Mechanical Turk (AMT), etc\dots),
-one might reasonably conclude that these workers are not collaborative.
-
-But, as
-\citeauthor{crowdcollab} identifies,
-crowd workers are substantially more collaborative than we originally may have inferred
-\cite{crowdcollab}.
-Given this realization, we pose the question:
-can requesters operationalize this finding to get work done more effectively than conventional methods would allow?
-Perhaps more importantly, can employing this finding allow workers
-to contribute meaningfully to tasks
-in ways that do not necessarily fall under the existing paradigms
-(e.g. ``find--fix--verify'')
-\cite{bernstein2015soylent}?
 

### 2016-06-15 cscw/papers/2017/labor/trackingWork.tex
total rewrite of the tracking labor intro

@@
 \section{Introduction}
-The past decade has seen
-paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
-grow to occupy the fascination of both academic circles and our culture as a whole
-\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
-The research community has made connections between this emergent form of work
-and the historically situated ``piecework'',
-as well as a number of observations relating to
-the frustration workers feel stemming from the management of this work,
-but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
-or crowd work itself,
-has proven difficult;
-the connection between crowd work and a historical parallel contextualizing its ongoing developments 
-hasn't been deeply made yet
-\cite{turkopticon,dynamo,uberAlgorithm}.
+% \itshape 
+\begin{itemize}
+  \item on--demand labor has consumed an increasing proportion of the labor market in the past ten years [\textit{Pew research on on--demand labor}]
+  \item the work surrounding this topic has ranged from Turk work processes to resisting Uber \& other gig labor platforms \cite{bernsteinSoylent,uberAlgorithm}
+  \item some researchers have tackled understanding this new instantiation of work, but come up short.
+  \item we offer a framing on this topic that situates the research so far on a timeline of the maturation of another form of work that emerged approximately 150 years ago
+  --- that is, piecework --- which convincingly suggests that on--demand labor in the form of Turking, driving for Uber, etc\dots are in fact little more than piecework re--surging today.
+  \item this paper will trace the body of research describing microwork --- and later gig work --- and place it in the context of historical piecework and the industrial revolution as a whole.
+  \item informed by this new lens, we'll then turn our attention to the future of this re--emergent form of work, and suggest ways that researchers and users of on--demand labor might influence the outcomes we predict in this paper.
+\end{itemize}
 
-While much of this work appears to gesture toward
-the parallels between contemporary on--demand work and piecework,
-it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
-This paper will attempt to do so by arguing that
-the topics social computing researchers have investigated are
-not just parallel to historical piecework and
-the process of factorization that took place in the early \nth{20} century;
-indeed, these phenomena \textit{precipitate and reinforce each other}.
+% \upshape
 
-In the reflection on the literature published in the last 5 years since
-\citeauthor{crowdworkFuture}'s 
-\citetitle{crowdworkFuture},
-we notice a broader trend describing the change in work that's being done
-\cite{crowdworkFuture}.
-Following the improvements in
-breaking down tasks,
-delegating work, and
-managing workers more broadly,
-crowd--powered work has continued to parallel historical piecework's trajectory
-by outgrowing ad--hoc worker groups and coalescing into more formal working groups,
-for example through the use of a ``white list'' of familiar workers.
-We might call this the regularization of factory work.
+% The past decade has seen
+% paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
+% grow to occupy the fascination of both academic circles and our culture as a whole
+% \cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
+% The research community has made connections between this emergent form of work
+% and the historically situated ``piecework'',
+% as well as a number of observations relating to
+% the frustration workers feel stemming from the management of this work,
+% but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
+% or crowd work itself,
+% has proven difficult;
+% the connection between crowd work and a historical parallel contextualizing its ongoing developments 
+% hasn't been deeply made yet
+% \cite{turkopticon,dynamo,uberAlgorithm}.
 
-Making sense of the broader field of crowd work
-by looking at the movements toward
-distributing work,
-routinizing and breaking down tasks,
-and externally managing workers
-as linked to one another, 
-we give ourselves a framing of contemporary piecework that explains,
-and arguably predicts,
-what we have seen thus far
---- and perhaps what we should expect to see going forward.
+% While much of this work appears to gesture toward
+% the parallels between contemporary on--demand work and piecework,
+% it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
+% This paper will attempt to do so by arguing that
+% the topics social computing researchers have investigated are
+% not just parallel to historical piecework and
+% the process of factorization that took place in the early \nth{20} century;
+% indeed, these phenomena \textit{precipitate and reinforce each other}.
 
-This paper will attempt to demonstrate that the familiarity
-that we incidentally observe between what we generally call crowd work \& gig work
-and the historical practice of piecework and subsequent factory work
-is more than passing;
-that the phenomena researchers have observed in on--demand digitally mediated labor markets
-were inevitable milestones in the birth and life of piecework.
-We will set out to show that this perspective of crowd work as an embodiment of piecework
-predicts the myriad outcomes of contemporary crowd work, including
-our developments of work--flows further abstracting work,
-the troubling effects of those developments on factors such as pay and work quality, and
-many if not all of the frustrations researchers have discovered among workers
-in the study of crowd and gig work.
+% In the reflection on the literature published in the last 5 years since
+% \citeauthor{crowdworkFuture}'s 
+% \citetitle{crowdworkFuture},
+% we notice a broader trend describing the change in work that's being done
+% \cite{crowdworkFuture}.
+% Following the improvements in
+% breaking down tasks,
+% delegating work, and
+% managing workers more broadly,
+% crowd--powered work has continued to parallel historical piecework's trajectory
+% by outgrowing ad--hoc worker groups and coalescing into more formal working groups,
+% for example through the use of a ``white list'' of familiar workers.
+% We might call this the regularization of factory work.
 
-``But'',
-as \citeauthor{scholz2012digital} points out,
-``it would be wrong to conclude that
-in the realm of digital labor there is nothing new under the sun''
-\cite{scholz2012digital}.
-For one, information work
-(e.g. Amazon Mechanical Turk (AMT) \& Upwork)
-and digitally managed work in general make it substantially easier
-to keep workers geographically dispersed
-compared to the factory workers that followed home--based piecework.
-This aspect of work renders workers virtually invisible to the task solicitors
---- or ``requesters'' ---
-despite being significantly more connected than historical pieceworkers were % this is a weak point...
-\cite{turkopticon}.
-Secondly, while
-factory foremen and other middle--men
-traditionally mediated the relationships between workers and managers,
-today the visible agents are \textit{systems};
-rather than employing individuals who ``personified the functions of management''
-and can thus negotiate workers' needs,
-socio--technical systems mediate these interactions % often with unrewarding outcomes
-\cite{wray1949marginal}.
-% \cite{storiesIraniSilberman}.
-Finally, this characteristic of modern piecework makes
-worker coordination for collective advocacy and action significantly more difficult,
-necessitating special consideration to maintain the inertia of collective will
-while focusing that energy productively
-\cite{dynamo}.
-% \ali{I really don't like this paragraph\dots}
-% here are undoubtedly more examples of differences.
-% How do I not stake a claim that this is comprehensive?
-% Or should I really flesh this out and make it (comprehensive)?
+% Making sense of the broader field of crowd work
+% by looking at the movements toward
+% distributing work,
+% routinizing and breaking down tasks,
+% and externally managing workers
+% as linked to one another, 
+% we give ourselves a framing of contemporary piecework that explains,
+% and arguably predicts,
+% what we have seen thus far
+% --- and perhaps what we should expect to see going forward.
+
+% This paper will attempt to demonstrate that the familiarity
+% that we incidentally observe between what we generally call crowd work \& gig work
+% and the historical practice of piecework and subsequent factory work
+% is more than passing;
+% that the phenomena researchers have observed in on--demand digitally mediated labor markets
+% were inevitable milestones in the birth and life of piecework.
+% We will set out to show that this perspective of crowd work as an embodiment of piecework
+% predicts the myriad outcomes of contemporary crowd work, including
+% our developments of work--flows further abstracting work,
+% the troubling effects of those developments on factors such as pay and work quality, and
+% many if not all of the frustrations researchers have discovered among workers
+% in the study of crowd and gig work.
+
+% ``But'',
+% as \citeauthor{scholz2012digital} points out,
+% ``it would be wrong to conclude that
+% in the realm of digital labor there is nothing new under the sun''
+% \cite{scholz2012digital}.
+% For one, information work
+% (e.g. Amazon Mechanical Turk (AMT) \& Upwork)
+% and digitally managed work in general make it substantially easier
+% to keep workers geographically dispersed
+% compared to the factory workers that followed home--based piecework.
+% This aspect of work renders workers virtually invisible to the task solicitors
+% --- or ``requesters'' ---
+% despite being significantly more connected than historical pieceworkers were % this is a weak point...
+% \cite{turkopticon}.
+% Secondly, while
+% factory foremen and other middle--men
+% traditionally mediated the relationships between workers and managers,
+% today the visible agents are \textit{systems};
+% rather than employing individuals who ``personified the functions of management''
+% and can thus negotiate workers' needs,
+% socio--technical systems mediate these interactions % often with unrewarding outcomes
+% \cite{wray1949marginal}.
+% % \cite{storiesIraniSilberman}.
+% Finally, this characteristic of modern piecework makes
+% worker coordination for collective advocacy and action significantly more difficult,
+% necessitating special consideration to maintain the inertia of collective will
+% while focusing that energy productively
+% \cite{dynamo}.
+% % \ali{I really don't like this paragraph\dots}
+% % here are undoubtedly more examples of differences.
+% % How do I not stake a claim that this is comprehensive?
+% % Or should I really flesh this out and make it (comprehensive)?
 

### 2016-07-12 chi/papers/2017/trust/trust.tex
cscw revisions; intro spin-off

--- 2016-06-15 chi/papers/2017/referrals/referrals.tex
@@
 \def\plaintitle{Naming Things is Hard: Real Title Following Colon}
-\def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
+\def\plainauthor{All the people (al2, msb)}
 \def\emptyauthor{}
@@
 \begin{abstract}
-  This paper is going to be about giving workers an opportunity to refer other workers.
-  Some of the old literature on job referring suggests that
-  it's a good way for networked workers to do stuff.
-  \citeauthor{crowdcollab} found that \citetitle{crowdcollab},
-  so why not take advantage of that?
-
-  We'll try various incentive structures
-  (no penalty/no rejection, task approval, referral commission, and referral bonus)
+  todo: this
 \end{abstract}
@@
 
-\itshape
-Sample intro. Does this sell the idea?
-\upshape
+\section{Introduction}
+Trust precipitates many, if not all, social interactions
+(wow, much depth).
 
-\section{Introduction}
+\begin{enumerate}
+\item In the initial relationship, to the extent the situation is novel and ambiguous, faith in humanity will lead to trusting beliefs.
+\item In the initial relationship, trusting stance will lead to trusting intention. The effects of trusting stance on trusting intention will not be mediated by trusting beliefs.
+\item In the initial relationship, situational normality belief will lead to trusting intention.
+\item In the initial relationship, structural assurance belief will lead to trusting intention.
+\item In the initial relationship, the trusting beliefs will be a function of structural assurance belief and situational normality belief.
+\item In the initial relationship, trusting intention will be a function of benevolence belief, competence belief, honesty belief, and predictability belief.
+\item In the initial relationship, faith in humanity and trusting stance will lead to structural assurance belief.
+\item In the initial relationship, categorization processes that place the other person in a positive grouping will tend to produce high levels of trusting beliefs.
+\item In the initial relationship, token control efforts will strengthen the tendency of categorization processes, faith in humanity, and structural assurance belief to produce high levels of trusting beliefs.
+\item In the initial relationship, high trusting intention is likely to be very fragile when it is supported
+by only one or two antecedents, when it relies almost exclusively on assumptions, and when perceived risk is high.
+\item In the initial relationship, high trusting intention is likely to be robust when (1) a combination of several of its antecedents encourages the trustor to ignore, rationalize, or absorb the negative actions of the other and (2) continued success or low perceived risk of failure consequences cause little critical attention to be paid to the other's behavior.
+Subjects with high disposition to trust or institution-based trust levels are likely to pay even less critical attention.
+\item In the initial relationship, high trusting intention is likely to be robust when (1) the parties interact face to face, frequently in positive ways, or (2) the trusted party has built a widely known good reputation.
+Social interaction affects trusting intention by its positive effects on trusting beliefs and institution-based trust
+\cite{mcknight1998initial}.
+\end{enumerate}
 
@@
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2016-07-12 cscw/papers/2017/labor/introduction.tex
cscw revisions; intro spin-off

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Introduction}
% what would a cynical reader make bullet points out of?
% each sentence feels like a thought that's unrelated to the next one
The past decade has seen a proliferation of transient labor.
% beginning with ``information work'' such as on
% Amazon Mechanical Turk (AMT).
% ??? ->
Stemming from the recognition that
complex tasks might be achievable
--- and indeed trivial ---
through the effective management of the ``crowd'',
researchers have rapidly begun exploring
the space for opportunities to facilitate new (and revisit familiar) forms of work
\cite{bernsteinSoylent,redballoon}.
%  ??? -->
Broadly, this work has imagined the various applications of ``gig''--based work,
and making it work for ``requesters''
--- academics, industries, and to some extent end--users
who might benefit from the product of crowd--powered work
\cite{hong2015group,jonBrelig,paolacci2010running}.
% ??? -->
Since then,
the application of digitally mediated environments acting as rapid ``hiring halls''
for on--demand tasks
has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
% parentheticals giving examples?
have garnered the interest of both industry and research.

Researchers have made efforts to understand the \textit{people} that have gravitated toward on--demand work
since its emergence and popularization,
but as the form of work has grown and changed, so too have the demographics of workers
\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the realization of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody
\cite{turkopticon,dynamo}.
Other work has focused on the outcomes of work,
reflecting on the resistance workers express against digitally mediated labor markets
\cite{uberAlgorithm}.
Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.

\subsection{Piecework as a lens to understand gig work}
These bodies of literature have conversed,
and indeed begun to identify that their theories are interrelated in some way,
but thus far a convincing model for making sense of
--- let alone predicting ---
the phenomena in on--demand labor has proven elusive.
For instance, 
making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
is understandable in its own context, but
the relationship between the work itself (``turking'')
and the myriad social outcomes has only briefly been explored
\cite{crowdcollab}.

We offer a framing for \textit{all} on--demand work as a modern instantiation of ``piecework'',
referenced briefly by
\citeauthor{crowdworkFuture}
in
\citeyear{crowdworkFuture}
as a loose analogy to the form of work emerging at the time,
but not given as much weight as it perhaps could have been
\cite{crowdworkFuture}.
``Piecework'' here makes use of
a term historically used to describe work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation then being that
one would be paid for the \textit{output} of the work,
not the \textit{duration}).

This framing prompts us to note several immediately apparent similarities:
% Given the scope,
% we can frame piecework and on--demand labor
% as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}

% In the past decade, researchers have observed frustration
% grow among on--demand workers,
% with expression of this frustration spanning a wide range of tactics
% \cite{uberAlgorithm,turkopticon,dynamo}.
% Attempting to make sense of these case studies has been challenging
% in part because
% a wholly encompassing framework for understanding this form of work
% has thus proven difficult to capture.

But more than this,
the framing of on--demand labor as piecework (re--instantiated)
allows us to attempt to make sense of the broader research on this new form of work
by evaluating this work through a much more refined lens.
More concretely, by looking at task--based or ``gig'' work as
an instantiation (or even a continuation) of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can
\begin{inlinelist}
  \item make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
  \item offer predictions of what social computing researchers,
        and workers themselves,
        should expect to see on the horizon of on--demand work.
\end{inlinelist}

We'll look at a broad range of cases under a number of major themes
we propose as broadly describing the types of research being done in crowd work
and more generally in what we argue is contemporary piecework.
After validating this lens as a way of reasoning about on--demand labor,
we'll attempt to use this perspective to suggest areas of research worth anticipating,
and developments we should expect to see in the maturation of digitally mediated work.
Finally, we will offer design implications based on this research\dots \ali{???}

\onlyinsubfile{
\itshape
\subsection{notes/what i want to get across\dots}
\begin{itemize}
  \item on--demand labor has consumed an increasing proportion of the labor market in the past ten years [\textit{Pew research on on--demand labor}];
  the work  has ranged from Turk work processes
  \item some of it has been about resisting Uber \& other gig labor platforms \cite{bernsteinSoylent,uberAlgorithm};
  some researchers have tackled understanding this new instantiation of work, but come up short (\textit{how}).

  \item (what binds this together)
  \item we offer a framing on this topic that situates the research so far on a timeline of the maturation of another form of work that emerged approximately 150 years ago
  --- that is, piecework --- which convincingly suggests that on--demand labor in the form of Turking, driving for Uber, etc\dots are in fact little more than piecework re--surging today.

  \item this paper will trace the body of research describing microwork --- and later gig work --- and place it in the context of historical piecework and the industrial revolution as a whole.
  \item informed by this new lens, we'll then turn our attention to the future of this re--emergent form of work, and suggest ways that researchers and users of on--demand labor might influence the outcomes we predict in this paper.
\end{itemize}
\upshape
}

\end{document}

### 2016-07-12 random/final.tex
cscw revisions; intro spin-off

% % !TEX program = xelatex
\documentclass[10pt]{article}
\usepackage{balance,graphics,setspace,parskip,times}
% \usepackage{fontspec}
% \usepackage{xltxtra}
% \usepackage{xunicode}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}

\usepackage[tiny,compact]{titlesec}

% \usepackage{xesearch}
% \newcounter{words}
% \newenvironment{counted}{%
%   \setcounter{words}{0}
%   \SearchList!{wordcount}{\stepcounter{words}}
%     {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
%     n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
%   \UndoBoundary{'}
%   \SearchOrder{p;}}{%
%   \StopSearching}

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}
\def\labelitemi{\ }
\newcommand*\elide{\textup{[\,\dots]}}
\newcommand{\sectitle}[1]{\textbf{#1}\\}

% \author{
%   Ali Alkhatib\\
%   ali.alkhatib@cs.stanford.edu\\
%   \textit{Couldn't find an envelope; will coordinate to meet next quarter (if acceptable)}
% }

% \title{Final Paper\\
% \normalsize{
%   \textit{Digital media are an egalitarian force in American society}\\
%   }
% }

\begin{document}
The future of work may resemble a hunter--gatherer mode of work;
foraging for work throughout the day to sustain ourselves
rather than allowing us the stability of careers that we know today.

Alex wakes up in the morning to his alarm and
checks his phone for notifications of non--urgent tasks for which he's qualified
--- for instance, doing a second or third round of revisions for an essay for which he's already given feedback ---
while eating breakfast.
He flips from microtask to microtask over cereal and juice,
indicating whether the people in the pictures he's flipping through are having an argument or not.
He has a vague sense that he's providing training data to some artificial intelligence project, but the details escape him.
Or he never sought after them.
He doesn't care.

Occasionally Alex gets a prompt asking if he can set aside time later in the week to do some tasks
--- in this case, committing to training someone to do better in technical interviews by doing mock interviews over Skype ---
adding a few more blocks of appointments to his week.
His calendar is mostly empty, but
he tries to consolidate appointments as much as possible so he doesn't have to wake up at all hours of the day.
% After agreeing to some of those tasks, populating more of his calendar over the next several days,
% he packs up his laptop and phone and heads out.

Alex finishes his breakfast, wraps up his last task, and starts packing his messenger bag for the day.
As he starts to head out, he taps through an app interface to see if he can get a cleaner to come by his place today.

Michelle? No, she's never on time\dots~\\
Eric? No, seemed a little shady\dots~\\
\textbf{Lupe} is available, and she was reliable.

Alex pings Lupe to see if she can swing by today and do the laundry that's been piling up in the hamper.

Alex's phone steadily buzzes every few seconds with notifications that people are looking for workers.
Some are looking for a ride to the coffee shop he's already heading to;
others are looking for a ride downtown;
still others are looking for a ride to the airport.
% Knowing that his intent is to go to Philz Coffee every morning and seeing that Alex has recently woken up,
% Uber starts suggesting to him people in his building or neighborhood who are interested in going in the same direction.
Today Alex wants to get some work and later meet with some friends, so he declines them all.
He doesn't like driving people around anyway;
passengers are rude and the pay isn't very good.

He makes his way to Philz ignoring the now streaming solicitations on his phone:
Pick up someone's medicine;
tutor someone's kid in math;
babysit someone's kid for a few hours on an abrupt need
(for \textit{4 times} the normal value because of the urgency of the task request);
eventually Alex gets frustrated, stops, and toggles his availability to ``off''.
The phone rests quietly for the rest of the drive.

As Alex parks his car, Lupe's response comes in:
she'll be by in 30 minutes to start the laundry and will do wash \& dry.
Alex asks if Lupe can fold the clothing and, in the meantime while the laundry is running,
clean up his apartment. She'll later agree for an added fee.

Alex gets to Philz, opens his laptop, and navigates to his work site.
He's just received a notification: he's been granted the qualification to squash Google bugs,
so his eligible work list has expanded from around 10,000 tasks to 30,000.
This is excellent; most of the new tasks Alex has seeing lately were either too far out of his wheelhouse
or it was too unclear what the problem was; he's wary of taking on work where the input might be unclear;
once he worked on an issue that ended up taking almost a week to sort out and only paid \$10.

If Alex puts more time into squashing bugs on Facebook,
he might be promoted to a higher role that can take on more interesting,
lucrative tasks,
but the low--hanging fruit at Google and Apple are more plentiful,
so Alex opts to continue ``playing the field'' where the tasks are easier,
and more importantly more likely to pay off.

A few hours pass, and Alex has closed 40 bugs, making \$60;
he messages his friends to see where they are:

\begin{itemize}
\item Stephen is on his way, just dropping someone off before going off the clock
(\textit{Stephen drives for a living,
usually working nights and weekends in the city but occasionally picking up airport runs, like now}).
He should be by in around 15 minutes.
\item Erin messages that she has to cancel.
She's been getting jobs rejected lately
(\textit{Erin is a designer for more firms than she can count});
she needs to pick up some more gigs to make up the money she's lost from those rejections
(and, if possible, to recoup her reputation after a string of rejections).
\end{itemize}

Stephen catches up with Alex at Philz;
like Erin, Stephen has been having trouble with customers lately, but his work
requires him to go out and drive, putting himself ``in harm's way,'' as it were,
% Stephen's been having a rough go of things just as Erin has, but his work
% means that he has to get out there to do work,
whereas Erin's work is all at home.
In that sense, her work is a little more desirable;
Erin \textit{could}, she jokes, do all her work from Bermuda.
She's never been, but she's saving for it. Maybe next year.

Stephen tells Alex about some of the terrible passengers he had over the weekend:
two were absolutely trashed, both making a mess in his car;
a third passenger seemed to make it his mission to harass Stephen the whole way,
interrogating him for the route he took (even though his app navigates for him),
and generally treating him terribly.

Alex sighs and tells Stephen to just keep at it; 
Stephen sighs and says that he can put up with it as long as he doesn't get ``fat--fingered''
--- a term referencing when passengers accidentally (or deliberately) give their drivers one--star reviews.
Another of those, and Stephen might get suspended from the platform.
And it changes every day; today Palo Alto's threshold is 4.6; yesterday it was 4.3; tomorrow it might be 4.7
--- and if you wake up one morning and your rating is too low for the are you live in,
then you'll just have to do work somewhere else or stop working.

Alex glances at his phone.
The screen keeps flashing subtly as new jobs stream in.
One from Twitter.
Another, this time from Apple.
This one from Microsoft pays (almost suspiciously) well.
These are all well--paying jobs
--- a few dollars for closing a simple ticket, up to as much as a C--note for something more involved.
There are other problems associated with this work,
like maintaining confidentiality and keeping everything mentally sandboxed, 
but none of it puts him in physical danger the way on--demand driving does with Stephen;
and his work is more cut--and--dry than Erin's,
so he generally knows when he submits work whether it'll be approved or rejected.

Stephen's phone buzzes this time;
someone's looking for an SUV driver for a long trip to SoCal.
It's too good an offer to pass up.
Stephen bids farewell to Alex and quickly makes his way to the door.
Alex pulls his laptop back out and gets back to work.

% \end{counted}
% \footnotesize{word count: \thewords}

\bibliographystyle{acm} % this will cause an error if there are no citations!
\bibliography{../../../references.bib}
\end{document}

### 2016-07-12 trustList/final.tex
cscw revisions; intro spin-off

--- 2016-07-12 chi/papers/2017/trust/trust.tex
@@
-\documentclass{sigchi}
+% % !TEX program = xelatex
+\documentclass[10pt]{article}
+\usepackage{balance,graphics,setspace,parskip,times,hyperref}
+\usepackage[margin=1in]{geometry}
+\usepackage[inline]{enumitem}
+\usepackage[tiny,compact]{titlesec}
 
-\usepackage{todonotes,txfonts,balance,graphics,color}
-\usepackage{booktabs,textcomp,microtype,ccicons}
-% \usepackage{babel}
-% \usepackage{csquotes}
-\usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
-\usepackage{enumitem}
-\usepackage[T1]{fontenc}
-\usepackage[super]{nth}
-\usepackage[pdftex,pdfpagelabels=false]{hyperref}
-\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
-\usepackage[utf8]{inputenc} % for a UTF8 editor only
-
-% Paper metadata (use plain text, for PDF inclusion and later
-% re-using, if desired).  Use \emtpyauthor when submitting for review
-% so you remain anonymous.
-
+\setlength{\parskip}{.4em}
 \newlist{inlinelist}{enumerate*}{1}
-\setlist*[inlinelist,1]{%
+\setlist*[inlinelist,1]{
   label=\arabic*),
@@
 
-\def\plaintitle{Naming Things is Hard: Real Title Following Colon}
-\def\plainauthor{All the people (al2, msb)}
-\def\emptyauthor{}
-\def\plainkeywords{Please don't make me pick keywords.
-This is like asking a teacher to give the bullet points of
-what a student missed in lecture.}
-\def\plaingeneralterms{Documentation, Standardization}
+\def\labelitemi{\ }
+\newcommand*\elide{\textup{[\,\dots]}}
+\newcommand{\sectitle}[1]{\textbf{#1}\\}
 
-% llt: Define a global style for URLs, rather that the default one
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{
-    \def\UrlFont{\sf}
-  }{
-    \def\UrlFont{\small\bf\ttfamily}
-  }}
-\makeatother
-\urlstyle{leo}
-
-% To make various LaTeX processors do the right thing with page size.
-\def\pprw{8.5in}
-\def\pprh{11in}
-\special{papersize=\pprw,\pprh}
-\setlength{\paperwidth}{\pprw}
-\setlength{\paperheight}{\pprh}
-\setlength{\pdfpagewidth}{\pprw}
-\setlength{\pdfpageheight}{\pprh}
-
-% Make sure hyperref comes last of your loaded packages, to give it a
-% fighting chance of not being over-written, since its job is to
-% redefine many LaTeX commands.
-\definecolor{linkColor}{RGB}{6,125,233}
-\hypersetup{%
-  pdftitle={\plaintitle},
-% Use \plainauthor for final version.
-%  pdfauthor={\plainauthor},
-  pdfauthor={\emptyauthor},
-  pdfkeywords={\plainkeywords},
-  bookmarksnumbered,
-  pdfstartview={FitH},
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
-  urlcolor=linkColor,
-  breaklinks=true,
-  hypertexnames=false
-}
-
-% create a shortcut to typeset table headings
-% \newcommand\tabhead[1]{\small\textbf{#1}}
-\bibliography{../../../../references}
-
-% End of preamble. Here it comes the document.
 \begin{document}
-\balance{}
-\title{\plaintitle}
-
-\numberofauthors{3}
-\author{%
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-}
-
-\maketitle
-
-\begin{abstract}
-  todo: this
-\end{abstract}
-
-\category{H.5.m.}{Information Interfaces and Presentation
-  (e.g. HCI)}{Miscellaneous} \category{See
-  \url{http://acm.org/about/class/1998/} for the full list of ACM
-  classifiers. This section is required.}{}{}
-
-\keywords{\plainkeywords}
-
-\section{Introduction}
-Trust precipitates many, if not all, social interactions
-(wow, much depth).
-
+\section*{propositions of trust}
 \begin{enumerate}
 \item In the initial relationship, to the extent the situation is novel and ambiguous, faith in humanity will lead to trusting beliefs.
+
+\textit{not sure how to discuss this except maybe to engage with the general worldview about humanity perhaps shared among drug users/sellers, but idk}
+
 \item In the initial relationship, trusting stance will lead to trusting intention. The effects of trusting stance on trusting intention will not be mediated by trusting beliefs.
+
 \item In the initial relationship, situational normality belief will lead to trusting intention.
@@
 \item In the initial relationship, high trusting intention is likely to be robust when (1) the parties interact face to face, frequently in positive ways, or (2) the trusted party has built a widely known good reputation.
-Social interaction affects trusting intention by its positive effects on trusting beliefs and institution-based trust
-\cite{mcknight1998initial}.
+Social interaction affects trusting intention by its positive effects on trusting beliefs and institution-based trust.
+
 \end{enumerate}
+\cite{mcknight1998initial}
 
-\printbibliography
-
+\balance{}
+\bibliographystyle{acm}
+\bibliography{../references}
 \end{document}

### 2016-07-15 drafts/groundwork/designGuidance/workerQuality.tex
reorg to be less hopeful :P

%!TEX root = ../proceedings.tex
% \subsection{Worker Quality \& Ratings}
\textit{Groups use social pressure to encourage good actors or discourage bad actors \& ensure high quality, but rarely both.}

% \textit{there should be a whole subsection where I talk about how stressful quantitative ratings are, and I can even reference some research on eBay seller profiles and how useless they are}

Another issue that arose during fieldwork surrounded the verification that a worker was of high quality.
To reference existing platforms again and using Uber and Lyft in this case, the quality of a worker can be illustrated simplistically by the quantitative ratings that drivers often have.
The distinction between qualification and quality is important;
to use driving as an example once again, a driver may be qualified (typically, licensed to drive \& insured) but not high quality
--- for example, struggling to navigate in tricky roads, unfamiliar with the area, etc\dots

We found two primary ways of communicating worker quality:
\begin{itemize} \itemsep0pt \parskip0pt
  \item guaranteeing outcomes
  \item vouching for good work
\end{itemize}

At organizations with high barriers to membership, work can be guaranteed by the organization of workers.
This guarantee gives customers relative confidence that any issues will be rectified at a cost fully absorbed by the organization, rather than the customer.
To cover these costs, work groups generally tax workers for each job to form a ``rainy day fund" in anticipation of such an expense.

In the case of organization A, if work is not done satisfactorily well, organization A's guarantee involves paying for a new worker to do the job correctly at the organization's expense.
Workers whose work requires fixing are punished in various ways depending on the nature of the error and whether the worker has done inadequate work in the recent past;
workers essentially face increasingly severe punishments as they repeatedly make mistakes (or make mistakes of greater cost).

It's important to note that these organizations can make this guarantee for two reasons.
The first reason is that membership exposes workers to work opportunities that make attempting to gain membership worthwhile.
The second reason is that membership eligibility is non-trivial to achieve, and transgressing community norms and risking expulsion carries significant consequences, especially given the work opportunities that they would be jeopardizing.

Organization E addressed this challenge using community pressure fostered by shared emotional investment in the organization.
Through democratic administrative systems and frequent meetings mandatory for all active workers, a sense of communal buy-in seemed to take form among workers who felt that their failure to do good work would let down the rest of their peers.
Organization E further stoked this sense of community investment by reading feedback from customers, made anonymous by organization E, to the group during all-hands meetings.
Workers reportedly felt a sense of shared success when good feedback arrived, and similarly felt a shared failure when feedback was critical.

We discovered that quantitative ratings weigh heavily on workers,
causing anxiety over arbitrary and opaque feedback negatively affecting their eligibility as workers.
Platforms like Uber and Lyft warn their workers that a user rating below a certain threshold
(for example, rolling average under 4.8 on a 1-5 scale) may result in disciplinary action.
Drivers, for instance, told us of courses they had to take if the average of their ratings fell below 4.6/5.

We can --- and will --- discuss ways to make quantitative rating systems less damaging in practice,
but we find that numeric ratings overwhelmingly tend to harm workers;
feedback --- rather than ratings --- proves a more effective way to improve workers
\cite{numericCritique},
and customers' responses to negative ratings seem to be exaggerated
\cite{ebayRatings}.
Given these apparent weaknesses in numeric ratings, we propose that efforts to communicate reputation should attempt to maximize qualitative data, rather than minimize it as many systems (e.g. Uber, Lyft, etc\dots) do.

This sense of shared reputation and the desire among individuals to uphold it was reportedly very strong.
We discovered that workers, fearing their refusal to do certain kinds of work would reflect poorly on organization E, would agree to do work even when it was inappropriate to do so.
Examples of impropriety include using unfamiliar mechanical equipment and working in hazardous settings with inadequate protection.
In light of this, the organization has had to remind workers to take a firm stance when customers make unreasonable requests of them.

Whether worker motivation stemmed genuinely from investment in their community
or from a more practical desire not to lose the job they've gotten that day is unclear.
However, considering the high demand we found during our fieldwork
we suspect that this practical desire was not significantly influential.
In other words,
workers could turn down work and confidently expect to be offered another worthwhile job,
if they were only concerned with earning money.

### 2016-07-15 drafts/groundwork/fieldwork_advice.tex
reorg to be less hopeful :P

%!TEX root = proceedings.tex
\section{Method \& Position}
We approach this research with some amount of reflexivity, as described by Geertz and elaborated by others
\cite{geertz1988works,marcus1999anthropology},
to acknowledge the role we play as participants in the culture we study.
Reflexivity affords other benefits, however;
it allows us to avoid an undue and perhaps hopeless struggle to separate ourselves and objectify the communities we studied, and engage with participants more transparently, rather than attempting to abandon our preconceptions and existing mental models.
% I want to make a brief reference to reflexivity as an asset for social scientists, but if the argument is too opaque here then the answer is to make that point more fully so it's clearer, but that's not really a good option, so...

% We learned several constructive lessons in our fieldwork, especially relevant to a researcher whose intent is to design and implement a system for the members of a given community.
% do I list the subsections here? I feel like that would be a lot clearer, but I'm thinking maybe I should just dive into it

\subsection{Finding Participants, Making Partners}
Our process began by finding organizations with demonstrated success in areas such as worker advocacy and collective action, hoping that their experience in offline work would inform online working groups.
We found a variety of organizations ranging small worker cooperatives to national-scale labor unions, and began to make contact with several.
After several critical introductions by mutual contacts, we had a diverse range of partners ranging in size and spanning several industries.

We worked with 5 organizations: Organization A, organization B, organization C, organization D, and organization E.
Organization A is a labor union that spans the nation, although we specifically worked with a union chapter local to our area;
workers in this group are regulated by state laws describing qualifications and eligibility to work.
This represented one of the more formalized, organized groups that we worked with.
Organization B represents more than one million workers across various industries.
The specific vertical of work we studied involved health care, and the regulations involved with this group was even greater than with organization A.
Organization C is a national labor union and worker advocacy group which supports tens of thousands of workers, many of whom women and of minority groups.
We worked with a local group that organization C supports named organization E, which helps domestic workers and house cleaners find work our approximate geographic area.
Finally, we worked briefly with organization D,
a labor union that broadly advocates on behalf of many organizations of workers and expressed interest in representing existing workers in the ``gig economy'',
such as Uber and Lyft drivers.

% We spent more than three months making initial contact, introducing our backgrounds and motivations, and working with these groups.
With each group, we explained our interest and motivation as researchers exploring the idea of a worker-centric labor market;
we also attempted to describe what we hoped to contribute to this space.
At the time, we imagined a ``worker cooperative"
--- a worker-owned organization that often makes decisions democratically ---
and we found surprising (at the time) resistance from some worker-led organizations with which we spoke.
But this presented us with opportunities to learn about the perceived shortcomings of various organization structures.

As our interviews and discussions progressed, we mocked up designs based on what we learned.
By bringing these mock-ups to subsequent meetings, and iterating quickly on the feedback workers provided us,
we were able to demonstrate in our own ways as computer scientists and designers that we were both listening to and learning from their input.
This process, time-consuming and labor-intensive though it was, made important strides to prove the overarching claim that we were invested in these groups and respected the guidance they provided as equal partners.

Mock-ups offered another benefit that may be uniquely accessible to designers.
As we found ourselves struggling to explain concepts such as privacy and trustworthiness in tangible ways, and thus failing to get substantive answers to questions around these issues, we realized that we could illustrate various paradigms, extrapolate on them, and learn how workers felt about design choices through their preference and especially from their suggestions for revision.

We ultimately spent more than three months making contact with, and learning from, these groups.
For several weeks (admittedly fleeting, in the context of ethnographic fieldwork),
we learned through participant-observation:
answering phone dispatch requests and triaging the worker assignment system when interactions mediated by existing technologies broke down.
We gained significantly deeper insights about workers and customers as a result of this time,
but engaging in the uncomfortable, time-intensive, and tiring work of liaising with customers and triaging worker dispatch proved our level of commitment as greater in scope than a short-lived research project.

### 2016-07-15 drafts/labor/trackingWork.tex
reorg to be less hopeful :P

--- 2016-05-28 cscw/papers/2017/labor/trackingWork.tex
@@
 
-\usepackage{todonotes,txfonts,balance,graphics,color,alphalph}
-\usepackage{booktabs,textcomp,microtype,ccicons}
+\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth}
 \usepackage[normalem]{ulem}
-% \usepackage{babel}
-% \usepackage{csquotes}
 \usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
-\usepackage{enumitem}
-\usepackage{moreenum}
 \usepackage[T1]{fontenc}
-\usepackage{nth}
 \usepackage[pdftex,pdfpagelabels=false]{hyperref}
@@
 \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+\makeatletter
+\def\blx@maxline{77}
+\makeatother
 
@@
 
-% End of preamble. Here it comes the document.
+\newcommand{\onlyinsubfile}[1]{#1}
+\newcommand{\notinsubfile}[1]{}
 \begin{document}
+\renewcommand{\onlyinsubfile}[1]{}
+\renewcommand{\notinsubfile}[1]{#1}
 \title{\plaintitle}
@@
 \keywords{\plainkeywords}
-
-\section{Introduction}
-The past decade has seen
-paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
-grow to occupy the fascination of both academic circles and our culture as a whole
-\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
-The research community has made connections between this emergent form of work
-and the historically situated ``piecework'',
-as well as a number of observations relating to
-the frustration workers feel stemming from the management of this work,
-but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
-or crowd work itself,
-has proven difficult;
-the connection between crowd work and a historical parallel contextualizing its ongoing developments 
-hasn't been deeply made yet
-\cite{turkopticon,dynamo,uberAlgorithm}.
-
-While much of this work appears to gesture toward
-the parallels between contemporary on--demand work and piecework,
-it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
-This paper will attempt to do so by arguing that
-the topics social computing researchers have investigated are
-not just parallel to historical piecework and
-the process of factorization that took place in the early \nth{20} century;
-indeed, these phenomena \textit{precipitate and reinforce each other}.
-
-In the reflection on the literature published in the last 5 years since
-\citeauthor{crowdworkFuture}'s 
-\citetitle{crowdworkFuture},
-we notice a broader trend describing the change in work that's being done
-\cite{crowdworkFuture}.
-Following the improvements in
-breaking down tasks,
-delegating work, and
-managing workers more broadly,
-crowd--powered work has continued to parallel historical piecework's trajectory
-by outgrowing ad--hoc worker groups and coalescing into more formal working groups,
-for example through the use of a ``white list'' of familiar workers.
-We might call this the regularization of factory work.
-
-Making sense of the broader field of crowd work
-by looking at the movements toward
-distributing work,
-routinizing and breaking down tasks,
-and externally managing workers
-as linked to one another, 
-we give ourselves a framing of contemporary piecework that explains,
-and arguably predicts,
-what we have seen thus far
---- and perhaps what we should expect to see going forward.
-
-This paper will attempt to demonstrate that the familiarity
-that we incidentally observe between what we generally call crowd work \& gig work
-and the historical practice of piecework and subsequent factory work
-is more than passing;
-that the phenomena researchers have observed in on--demand digitally mediated labor markets
-were inevitable milestones in the birth and life of piecework.
-We will set out to show that this perspective of crowd work as an embodiment of piecework
-predicts the myriad outcomes of contemporary crowd work, including
-our developments of work--flows further abstracting work,
-the troubling effects of those developments on factors such as pay and work quality, and
-many if not all of the frustrations researchers have discovered among workers
-in the study of crowd and gig work.
-
-``But'',
-as \citeauthor{scholz2012digital} points out,
-``it would be wrong to conclude that
-in the realm of digital labor there is nothing new under the sun''
-\cite{scholz2012digital}.
-For one, information work
-(e.g. Amazon Mechanical Turk (AMT) \& Upwork)
-and digitally managed work in general make it substantially easier
-to keep workers geographically dispersed
-compared to the factory workers that followed home--based piecework.
-This aspect of work renders workers virtually invisible to the task solicitors
---- or ``requesters'' ---
-despite being significantly more connected than historical pieceworkers were % this is a weak point...
-\cite{turkopticon}.
-Secondly, while
-factory foremen and other middle--men
-traditionally mediated the relationships between workers and managers,
-today the visible agents are \textit{systems};
-rather than employing individuals who ``personified the functions of management''
-and can thus negotiate workers' needs,
-socio--technical systems mediate these interactions % often with unrewarding outcomes
-\cite{wray1949marginal}.
-% \cite{storiesIraniSilberman}.
-Finally, this characteristic of modern piecework makes
-worker coordination for collective advocacy and action significantly more difficult,
-necessitating special consideration to maintain the inertia of collective will
-while focusing that energy productively
-\cite{dynamo}.
-% \ali{I really don't like this paragraph\dots}
-% here are undoubtedly more examples of differences.
-% How do I not stake a claim that this is comprehensive?
-% Or should I really flesh this out and make it (comprehensive)?
-
-\subsection{Piecework as a lens to understand gig work}
-
-\citeauthor{crowdworkFuture}
-write of the future of crowd work in an effort to investigate
-whether crowd work will be a worthwhile form of work or another instantiation of ``piecework''
-\cite{crowdworkFuture}.
-``Piecework'' here makes use of
-a term historically used to describe work done in the home,
-in manageable tasks,
-often involving clear instructions
-and payment only for work completed, not work done
-(the differentiation, here, being that
-one would be paid for the \textit{output} of the work,
-not the \textit{duration}).
-% \msb{split up that sentence. One is the def'n of piecework, the other is the comparison you're drawing}
-Given the scope,
-we can frame piecework and on--demand labor
-% \msb{
-% \sout{wait, are you using gig work as the umbrella term? I think you need a category term that covers all of them}
-% \ali{let's use ``on--demand labor'', since
-% that seems general enough without accidentally including historical piecework}}
-as sharing these important similarities:
-\begin{inlinelist}
-\item this form of work began in the home
-\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
-\item the worker's status
-(not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy.
-\end{inlinelist}.
-
-% \subsection{Context}
-In the past decade, researchers have observed frustration
-grow among on--demand workers,
-with expression of this frustration spanning a wide range of tactics
-\cite{uberAlgorithm,turkopticon,dynamo}.
-Attempting to make sense of these case studies has been challenging
-in part because
-a wholly encompassing framework for understanding this form of work
-has thus proven difficult to capture.
-
-This paper attempts to make sense of the broader research on this new form of work,
-\textit{on--demand labor}, by evaluating this work through a more familiar lens: piecework.
-More concretely, by looking at task--based or ``gig'' work as
-an instantiation and continuation of piecework,
-and by looking for patterns of behavior that the corresponding literature predicts
-on this basis, we can do the following:
-\begin{inlinelist}
-  \item we can make sense of the phenomena so far as part of a much larger series of interrelated events;
-  \item we can bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
-  \item we can offer predictions of what social computing researchers, and workers themselves, should expect to see on the horizon of on--demand work.
-\end{inlinelist}
-
-We'll look at a broad range of cases under a number of major themes
-we propose as broadly describing the types of research being done in crowd work
-and more generally in what we argue is contemporary piecework.
-After validating this lens as a way of reasoning about on--demand labor,
-we'll attempt to use this perspective to suggest areas of research worth anticipating,
-and developments we should expect to see in the maturation of digitally mediated work.
+\subfile{introduction.tex}
 

### 2016-07-15 drafts/theoryPaper/designGuidance/feedback.tex
reorg to be less hopeful :P

%!TEX root = ../proceedings.tex
% \subsection{Feedback}
\textit{Ratings cause anxiety, whereas feedback can satisfy the same administrative needs without causing workers undue stress.}

At organization E, we learned that workers interpreted negative feedback very personally,
making it difficult for them to internalize that feedback constructively and act on the suggestions customers made.
Multiple issues may have been at play here;
cultural differences might explain a mismatch in how feedback is offered and how it is received,
but so too would the power imbalance between undocumented domestic workers and their customers.

organization E didn't seem to investigate the cause of this effect,
but their solution circumvented this problem entirely.
Rather than giving workers unfiltered, raw feedback as it streamed in,
organization E intercepted reviews from customers and distilled them into constructive, actionable feedback.
Both praise and criticism would sometimes be read aloud to the entire group,
with identifying information removed.

\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/feedback}
    \caption{Given quantitative prompts and qualitative follow-up questions,
  research suggests that users are more likely to write more,
  which can better-inform.}~\label{fig:workerFeedback}
\end{figure}

The purpose of reading positive feedback publicly was to reaffirm the group's sense of worth in a shared sense of success;
the purpose of reading the negative feedback, meanwhile, was to prompt workers to reflect on what went ``wrong'', and how to avoid such an outcome in the future.
A shared sense of failure also seemed to affect workers in these cases;
whether this is an outcome of public readings of reviews, or more complex relationships between workers and organization E, is unclear.

Speaking with Uber and Lyft drivers, we found an alternative approach to providing workers with feedback;
drivers are made aware of an aggregated rating
--- generally a moving average of the previous \textit{n} ratings ---
but they are not exposed to qualitative feedback in any form, even when customers provide it.

A driver's current aggregated rating, it turns out, is extremely important:
a driver's rating determines the worker's eligibility to do work, and falling below various thresholds carries various consequences. Reports suggesting that these thresholds are finely tuned as well as closely guarded secrets make these markets particularly emotionally taxing for workers \cite{leakedUber}.

When we spoke to drivers, they relayed stories of apparent obligations to take expensive remedial courses if their overall rating dropped below a certain threshold,
and widely held fears of suspension and deactivation for falling below an unknown level weigh heavily on workers' minds.
Frustration directed at drunk, clumsy, or naive passengers accidentally or deliberately giving drivers a rating of 4 (out of 5) stars further exacerbates stress.

Markets which aim to empower ``gig workers'' should use quantitative rating systems sparingly,
or to prompt more detailed qualitative feedback, as we illustrate in Figure
\ref{fig:workerFeedback}.
Quantitative ratings in the form of Likert scales followed by qualitative prompts for feedback seem to precipitate more detailed feedback
\cite{numericCritique}.

% While quantitative metrics
% --- like approval rates and moving average ratings ---
% make large-scale evaluation seemingly manageable and generalizable,
% the professional stress these
Quantitative metrics --- especially those that are opaquely evaluated --- do not benefit workers, nor does it seem they even inform appropriate customer behavior \cite{ebayRatings}.
Meanwhile, constructive feedback suggesting improvements may provide workers with the necessary information to improve as professionals without causing undue stress over issues such as worker eligibility..

### 2016-07-15 drafts/theoryPaper/designGuidance/workerIdentity.tex
reorg to be less hopeful :P

%!TEX root = ../proceedings.tex
% \subsection{Worker identity versus ``gig workers"}
\textit{The very features that make gig work appealing also stymie attempts to coalesce stable worker advocacy groups.}

At organization E, organization A, organization C, and organization B we observed a subtle but important characteristic;
everyone at these organizations identified primarily by the work that they did.
Members of organization E self-identified as house cleaners and day laborers;
People at organization A were workers, first and foremost.
The same fundamental sentiment was shared by nurses affiliated with organization B.

We found that this carried an important effect;
at organization E the effect was most salient that workers identified together and recognized the importance of maintaining organization E's positive reputation among customers.
Every week, organization E administrative staff would read reviews of workers,
made anonymous beforehand,
allowing the workers to reflect on these frequent successes and occasional failures.

Perhaps most surprising was a shared sense of success and failure when workers listened to this feedback;
people we spoke to felt that they had upheld or let down their peers and the organization of which they were part,
even if the review was not of their own work.
This strong sense of group identity seemed to drive workers to work hard to maintain the already good reputation that organization E enjoyed.

When we spoke to drivers in the gig economy, some told us that they had previously worked for competing yellow cab companies.
They explained that they gave up on conventional cab companies was because requirements to drive for yellow cabs felt onerous.
Drivers worked for Uber and Lyft because those markets allowed them to drive as much or, importantly, as little as they wanted.
At some yellow cab companies, a monthly fee was required for access to the cabs themselves.

This presents a potential challenge for those hoping to build communities fostering collective decision-making and action.
As Kraut et al. point out, a crucial component in designing successful online communities is identification with the group
\cite{successfulOnlineCommunities}.
This is already a difficult goal to achieve, now exacerbated by the nature of the community involved.
Features that makes gig work appealing
--- the freedom to come and go as one pleases, for instance ---
make it that much more difficult to form a shared sense of community based on shared identities.

Indeed, workers in the gig economy tended not to think of themselves as identified by the work that they do ---
many drivers, for instance, volunteered that they were musicians, security guards, and even elementary school teachers.
Their work as drivers, then, was not substantively important to who they are.
Workers at organization E, meanwhile, identified as cleaners, day laborers, etc\dots without qualification.

Among many of the ``gig workers" that we spoke to, the sense of freedom and independence turned out to be an important feature of their identity.
The relative freedom over when and indeed whether to work was, it seemed, powerfully appealing.

This represents a difficult dilemma for labor unions and other conventional worker advocacy groups,
which rely on the shared sense of identity of workers,
who are now increasingly detaching their identities from the work that they do.

We assume that a worker-centric market needs emotional investment from its participants in order to succeed,
and the evidence seems to corroborate this sense \cite{dynamo,olsonlogic,russell1982collective}.
Given this premise, it seems that workers in a system must appreciate their reliance on their group's collective success.
Balancing this need with the desire that workers have expressed to remain unburdened by various requirements might prove to be a difficult act, but a necessary one.

We have little concrete guidance for this challenge
except to to say that system designers might strive to find ways to make it palatable for gig workers to identify collectively
and to recognize that their individual success is collectively determined by the quality of all of their work.
By forming a sense of shared identity and investment in that community,
a constructive sort of self-policing might emerge, as briefly discussed by Lave \& Wenger \cite{lave1998communities}.

### 2016-07-15 drafts/theoryPaper/intro.tex
reorg to be less hopeful :P

%!TEX root = proceedings.tex
\section{Introduction}

% Karl Polyani's \textit{The Great Transformation} describes the processes that led Western Europe from Feudalism to what he calls ``market society"
% \cite{GreatTransformation}.
% His insights gave birth to ``Substantivism", whose subscribers think of economies broadly as ways of navigating the world
% --- socially and otherwise ---
% rather than more formally as ways of maximizing the utilization of scarce resources.
% In the Substantivist school of thought the negotiations found in the ``sharing" or ``peer economy" between and among workers, customers, and the market operators themselves collectively represent a broader economy than the money-for-(rides, housing, etc\dots) exchange that formalists would explore.

In the past several years, a new type of work has emerged where customers can hire someone to complete an individual task, or ``gig''.
This task might involve delivering a package, or driving someone downtown.
Such a model principally requires the sharing of capital-intensive goods, like access to a car or home,
which led to the popularization of its familiar name, the ``sharing economy''.
People could ``share'' their homes (Airbnb, Couchsurfing \cite{airbnbOfficial,couchsurfingOfficial}), cars (Uber, Lyft, and others \cite{uberOfficial,lyftOfficial}), and increasingly one's own time (TaskRabbit, Zaarly, and many others \cite{taskrabbitOfficial,zaarlyOfficial}).

Workers' demographics have changed dramatically in as little as half a decade largely discreetly.
Where workers in the sharing economy were once car and home-owners who had free time to spare,
many now think of these companies and the markets they expose as their primary source of income.
In the last several years, workers for Uber, Homejoy,
and other market operators have initiated
--- and in some cases won \cite{homejoySuit,uberSuit} ---
suits describing mistreatment and misclassification of these workers as
``independent contractors" rather than ``employees" (protected and regulated by labor laws in the United States) \cite{fedsUber}.

In the sharing economy's nascent years, companies enticed workers to join for the potential to do work ``on the side": offering rides to others when they had free time, or renting out their apartment when they were out of town for a weekend.
For various reasons, the culture has since changed, and the notion of doing work in one's free time has largely disappeared.
Instead, drivers report primarily working as drivers, and that their primary sources of income consist of the aggregated sum of passengers they pick up through ride-sharing markets.
Even in the hotelier industry, providers have purchased apartments or re-purposed their own homes primarily to serve guest occupants, rather than to rent out incidentally when they have a spare room or are out of town.

In these ways, workers are neither ``peers", nor are they ``sharing" resources that would otherwise go underutilized.
But they're not conventional workers, either.
They acquire capital
--- sometimes co-signing on leases with the companies that run these markets ---
under their own names, run their businesses relatively independently, and make the majority or even totality of their income based on each individual job, or ``gig", cumulatively summed up.
With their careers described as a series of individual jobs, each self-contained and relatively independent of the others, people have renamed it the ``gig economy", more fairly referencing the differentiating nature of this work.

The widespread nature of these changes suggests that this is part of a larger trend in what may become the future of work;
far from the hopeful but cautious predictions offered of information workers and ``crowd work'' \cite{crowdworkFuture}.
Workers, increasingly find themselves objectified, marginalized, and frustrated by oppressive systems.

We considered, then, how one might design a worker-centric peer market;
how can system designers create technologically enabled markets as successful as existing markets like Uber, Lyft, and others, while also:
\begin{itemize} \itemsep0pt \parskip0pt
  \item giving workers a sense of locus
  \item benefiting workers as well as consumers
  \item facilitating worker organization and communication
  \item enabling collective decision \& action among workers
\end{itemize}

To answer these questions, we engaged in extensive fieldwork alongside workers and labor organizations in a number of industries.
Building in part from backgrounds in the social sciences and as trained computer scientists,
we learned about workers and the industries in which they work from workers themselves.
We report on the processes of making contact with variously formally organized worker advocacy groups,
illustrate some of the ways that we can learn from informants most effectively given our own skills,
and finally describe some of the findings we made as a result of our own use of these methods.

Informed by the input of dozens of workers from numerous industries ranging from highly regulated to informal,
we identify a number of aspects of on-demand work which system-designers should consider in the creation of a worker-centric labor market.
We offer guidance on these design considerations,
and in some cases illustrate the suggested approaches we generated in tandem with these partner organizations and workers.

Specifically, we offer contributions to the following questions
\begin{itemize}\itemsep0pt \parskip0pt
  \item What components of existing markets are inextricable from the features which make these markets successful?
  \item What can be disentangled and abstracted away?
  \item How would these groups be operated?
\end{itemize}

### 2016-07-15 drafts/theoryPaper/proceedings.tex
reorg to be less hopeful :P

\documentclass{sigchi}
\usepackage{graphics,balance,txfonts,mathptmx,color,booktabs,color,booktabs,textcomp,microtype,ccicons}
\usepackage[pdftex]{hyperref}
\usepackage[utf8]{inputenc}
\def\plaintitle{Laying Groundwork For a Worker--Centric Peer Economy}
\def\plainauthor{Ali Alkhatib, Justin Cranshaw, Andrés Monroy-Hernandez}
\def\emptyauthor{Anonymous for Review}
\def\plainkeywords{Peer economy; Collaborative markets}
\def\plaingeneralterms{Documentation, Standardization}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},       pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords}, bookmarksnumbered,
  pdfstartview={FitH},          colorlinks,
  citecolor=black,              filecolor=black,          linkcolor=black,
  urlcolor=linkColor,           breaklinks=true,
}

\begin{document}

\title{\plaintitle}

\numberofauthors{4}
\author{
  \alignauthor{Ali Alkhatib                 \affaddr{Stanford University}   \affaddr{Stanford, USA}
        \email{ali.alkhatib@cs.stanford.edu}}
  \alignauthor{Justin Cranshaw              \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
        \email{justincr@microsoft.com}}
  \alignauthor{Andrés Monroy-Hernandez      \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
        \email{amh@microsoft.com}}
  \alignauthor{Michael Bernstein      \affaddr{Stanford University}    \affaddr{Stanford, USA}
        \email{msb@cs.stanford.edu}}
}
\numberofauthors{1} \author{ \alignauthor{\emptyauthor} }
\maketitle

\begin{abstract}
The ``gig economy''
--- where work is measured through discrete tasks rather than through interconnected (and often long--running) careers ---
has become a site of significant frustration among workers who have come to rely on these systems and this form of work.
This paper explores some of the sources of that stress,
and attempts to relate this seemingly nascent work style to more familiar forms of labor
--- namely, ``piece work'',
a form of work which emerged at the dawn of the 20th century.
While we find and identify parallels between ``gig work'' and the more theoretically familiar ``piece work'',
we emphasize divergences which make the conflation of these ideas problematic.

We further explore opportunities and tactics to empower workers,
informed by ethnographic fieldwork including participant--observation,
semi--structured interviews,
and participatory design.
Our observation and interviews lead us to conclude that the most intuitive mechanisms for empowering workers
--- namely, labor unions advocating collectively and counter--balancing the manager's outsize leverage ---
may both structurally and culturally fail to capture the needs of ``gig workers''.
We argue that design interventions to empower ``gig workers'' must build on the theoretically familiar,
much as ``gig work'' itself draws inspiration from ``piece work'',
but that it is crucial that interventions are designed to support,
rather than reject,
the unique cultural and system qualities of gig workers.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section*{Dealing with Monopsonies}
\section*{Collective Bargaining}

% \input{intro.tex}
% \input{background.tex}
% \input{fieldwork_advice.tex}
% \input{design_guidance.tex}
% \input{discussion.tex}

% % REFERENCES FORMAT
% % References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{../references}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-07-16 drafts/labor/trackingWork.tex
various intro revs

@@
 \usepackage[normalem]{ulem}
-\usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
+% \usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
+\usepackage{biblatex}
 \usepackage[T1]{fontenc}
@@
 % \newcommand\tabhead[1]{\small\textbf{#1}}
-\bibliography{../../../../references}
+\bibliography{references}
 
@@
 
+\cite{olson2014ways}
 The HCI community is perhaps most familiar with examples of task--based work such as

### 2016-07-16 cscw/papers/2017/labor/introduction.tex
cleanup
cleanup?
some other intro revisions..

@@
+% %!TEX root = trackingWork.tex
 \documentclass[trackingWork]{subfiles}
@@
 \section{Introduction}
-% what would a cynical reader make bullet points out of?
 % each sentence feels like a thought that's unrelated to the next one
-The past decade has seen a proliferation of transient labor.
-% beginning with ``information work'' such as on
-% Amazon Mechanical Turk (AMT).
-% ??? ->
+In the past 10 years, TEST TEST TEST
+Amazon Mechanical Turk (AMT) has been the site of a flourishing economy for ``information work''
+\cite{paolacci2010running}.
+
+The past decade has seen a proliferation of a new type of labor measured in discrete, atomic tasks and on--off gigs;
+the workers that do these tasks float into and out of the labor market so freely that it has been called ``transient''
+by some
+\onlyinsubfile{\ali{who talks about this? definitely \cite{solinger1995chinese} but not the right context! (although, super interesting)}}.
 Stemming from the recognition that

### 2016-07-16 cscw/papers/2017/labor/trackingWork.tex
some other intro revisions..

--- 2016-05-28 cscw/papers/2017/labor/trackingWork.tex
@@
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
-\newcommand{\ali}[1]{{\color{BrickRed}[\itshape al2: #1\upshape]}}
+\newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}
 

### 2016-07-16 drafts/groundwork/designGuidance/workerQualifications.tex

%!TEX root = ../proceedings.tex
% \subsection{Worker Qualifications}
\textit{Workers in variably regulated markets similarly seek to prove their qualification;
they accomplish this in different ways.}

In interviewing workers across industries we found that workers with formalized qualifications wished to emphasize the value of the formalized qualifications that they offer;
in the case of electrical workers, strictly regulated by the state in which we were conducting research, workers feared that unlicensed, unregulated workers potentially tarnished the reputations of workers with legitimate credentials.

In the driver-for-hire market, we generally trust that drivers for Uber, Lyft, and even yellow cab companies are legally qualified to drive a car.
While passengers rarely ask to see proof of license and it is rarely shown (except, sometimes, in the cases of yellow cabs), it is widely assumed that a worker given access to the market and its consumers has sufficiently proven to the market operators (Uber, Lyft, etc\dots) that the driver is qualified.
The perceived risk of being caught without a valid driver's license seems sufficiently discouraging that fears of abuse and deception are not widely held in the United States, where this research was conducted.

As we interviewed members of organization A, we discovered a complex web of laws describing the ratios of variously qualified workers and other workers on construction and other work sites.
In short, a work site would be deemed in violation of construction regulations if too many inexperienced workers are working without sufficient more senior workers are not on the premises to oversee the work.
This requirement proves challenging to follow under the status quo as an worker calling in sick abruptly one morning may cause the work site to fall out of regulation.

Computer Scientists may observe that this problem simple enough to identify, at least to augment a human-driven process of finding another worker with sufficient credentials to take the absent worker's place for the duration of the absence.
We consider this one of the more formalized qualifications processes, and is roughly in line with the qualifications concerns we encountered when interviewing home health care workers through organization B.

We found less formally regulated, often homegrown, modes of qualifying workers at other organizations.
At organization E and through organization C, we found that organizations would determine worker qualifications using highly specialized criteria.
For instance, at organization E, we found that day laborers kept track not just of things like whether they were proficient English speakers, but also kept track of a wide array of qualifications describing whether they were eligible to take jobs involving yard work, heavy lifting, painting in various forms (e.g.
with a brush versus with a spray canister), and numerous other factors.

This system generally worked, with a notable exception.
One day, while volunteering for the phone dispatch system at organization E, we received a call asking for someone who could paint with a spray canister.
We correctly matched a worker to that call, but when the worker arrived to pick the worker up (the cheapest of three options for getting the worker to the work site), it came to light that the worker did not want to work for a contractor
--- which this customer was.
Our solution was to hastily find another worker who would be able to paint, which we managed to do expeditiously.
Later, we discovered that we had neglected to verify that the worker was able to paint with a spray canister, which he was not.
Ultimately, the worker was sent home and the work was not completed that day.

It should go without saying that this was a failure in attempting to match a worker and a customer.
The individual points of failure, however, can inform the design of an automated system to match workers and customers.
The first in the series of otherwise avoidable mistakes occurred as a result of the worker not being able to make an informed decision about the customer requesting him.
When the problem emerged, we sourced a worker according to a different procedure from the method used during the rest of the day
--- specifically, hastily ---
and as a result we overlooked details that otherwise would have prevented the second worker from claiming (or even wanting to claim) the work for which he was not qualified.
These considerations would be simple, but not necessarily intuitive, to implement in a system.

### 2016-07-16 drafts/groundwork/design_guidance.tex

%!TEX root = proceedings.tex
\section{Emergent Themes}
We identified several aspects of the on-demand markets we studied that seem to apply in generalizable ways.
These characteristics and issues extend beyond the more narrowly scoped issues of scheduling workers and handling payments which might seem more at the focal point of designing a labor market, although those issues too must be addressed;
we focus specifically on the social negotiations that a technological system must broker.
These guidelines are illustrated using mock-ups and a mobile application front-end developed in tandem with organization C, organization B, \& organization E, and consist of the following considerations:

\begin{enumerate} \itemsep0pt \parskip0pt
  \item Constructive feedback
  \item Assigning work fairly
  \item Managing customer expectations
  \item Protecting vulnerable workers
  \item Reconciling worker identities
  \item Assessing worker qualifications
  \item Communicating worker quality
% \item Feedback
% \item Assigning Work
% \item Customer Expectations
% \item Vulnerable Workers
% \item Worker Identity
% \item Worker Qualifications
% \item Worker Quality \& Ratings
\end{enumerate}

\subsection{1.        Constructive Feedback}
\input{designGuidance/feedback.tex}
\subsection{2.        Assigning Work Fairly}
\input{designGuidance/assigningWork.tex}
\subsection{3.        Managing Customer Expectations}
\input{designGuidance/customerExpectations.tex}
\subsection{4.        Protecting Vulnerable Workers}
\input{designGuidance/vulnerableWorkers.tex}
\subsection{5.        Reconciling Worker Identities}
\input{designGuidance/workerIdentity.tex}
\subsection{6.        Assessing Worker Qualifications}
\input{designGuidance/workerQualifications.tex}
\subsection{7.        Communicating Worker Quality}
\input{designGuidance/workerQuality.tex}

### 2016-07-16 drafts/labor/introduction.tex
reorg to be less hopeful :P

--- 2016-07-12 cscw/papers/2017/labor/introduction.tex
@@
+% %!TEX root = ./trackingWork.tex
 \documentclass[trackingWork]{subfiles}
@@
 \section{Introduction}
-% what would a cynical reader make bullet points out of?
 % each sentence feels like a thought that's unrelated to the next one
-The past decade has seen a proliferation of transient labor.
+The past decade has seen a proliferation in on--demand work
+where workers' statuses are so fleeting that
+they have been described as ``transient''.
+After identifying that myriad tasks
+--- from the mundane to the complex ---
+can be accomplished
 % beginning with ``information work'' such as on
@@
 % ??? ->
-Stemming from the recognition that
-complex tasks might be achievable
---- and indeed trivial ---
-through the effective management of the ``crowd'',
-researchers have rapidly begun exploring
-the space for opportunities to facilitate new (and revisit familiar) forms of work
-\cite{bernsteinSoylent,redballoon}.
-%  ??? -->
+% Stemming from the recognition that
+% complex tasks might be achievable
+% --- and indeed trivial ---
+by directing and managing the ``crowd'',
+the research and and industry communities have flocked to sites of labor
+like Amazon's Mechanical Turk (AMT)
+to solicit workers.
+Researchers in particular have taken to the space in earnest,
+finding opportunities to enable new forms of work
+as well as using these workers (``Turkers'')
+as representative populations of the public
+\cite{bernsteinSoylent,redballoon,paolacci2010running}.
+
+This approach to managing work (and in particular workers)
+can be differentiated from other formulations by a constellation of traits:
+\begin{inlinelist}
+\item distributing the workforce,
+\item distancing workers from the context of the work, and
+\item compensating work on a per--task basis rather than for time
+\end{inlinelist}.
+These features can be found elsewhere
+--- for instance,
+distributed workforces
+are a well--known feature of a global economy,
+%
+and abstracting workers from the products of their labor
+is well--associated with
+\citeauthor{marx2012economic} ---
+but these individual components have been explored to some extent
+(for instance, \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
+and we focus on the newfound interest in
+this particular paradigm of work
+% While contract work is familiar enough,
+% much of this work focuses on comprehensive work
+% (for example, creative or professional work like landscaping, design, etc\dots)
+\cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.
+
+On--demand work under this formulation
+has sparked interest across industries ranging from
+livery (driving for hire, for example Uber),
+house--cleaning (Handy),
+and various other services (TaskRabbit)
+\cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+Today, a rapidly growing transient workforce is forming,
+itself assembling piece--by--piece as industries and researchers find more and more unexpected ways
+to benefit from a latent pool of variably skilled workers.
+
 Broadly, this work has imagined the various applications of ``gig''--based work,

### 2016-07-16 drafts/theoryPaper/designGuidance/customerExpectations.tex

%!TEX root = ../proceedings.tex
% \subsection{Customer Expectations}
\textit{Clear expectations of all participants in trade benefit all, but introducing those expectations can be difficult for all parties.}

Perhaps the most common source of dissatisfaction with gig workers comes from a miscommunication or simply a mismatch in the expectations of the customer and the worker.
The effects of this mismatch appear to have been discussed when discussing Mechanical Turk, leading to norms where ``good" requesters know to provide examples of correct work and highlight common mistakes, explicating the intent of the micro work.
On other platforms, clear expectations are established ahead of time
(on ride-sharing markets, the expected driving path, estimated time, and estimated cost is displayed to customers).

People we spoke with at organization B, organization E, and organization C felt that formalizing the expectations of workers was the most appropriate solution to this problem.
By making it clear to customers what workers are expected to do
(for instance, what specific tasks go into cleaning a kitchen),
expectations of customers are set reasonably, workers can be trained properly,
and disputes about whether the worker did the job properly or sufficiently become less ambiguous.

\begin{figure}[t]
\centering
  \includegraphics[width=1\columnwidth]{figures/checklist}
    \caption{Explicit checklists make it easier for workers to verify that they've completed all expected work;
    it also makes it apparent to customers that they need to add or clarify certain tasks.}~\label{fig:checklist}
\end{figure}

We tend to agree with this approach.
In our fieldwork we found that paper checklists made it clearer to customers what to expect to be done
--- and importantly, what to expect not to be done.
If customers wanted additional work completed, they could specify that and agree to the extra time it would take to do that job, which they sometimes did.

This ``contract'' between workers and customers can be written collaboratively by both parties,
outlined by the workers themselves, or directed principally by customers as they describe the work they need completed;
the exact process is less important than the shared understanding of the constituent tasks.

### 2016-07-19 drafts/labor/introduction.tex
notes and revisions for the theory paper

@@
 \begin{document}
+
+\newcommand{\note}[1]{\onlyinsubfile{\textit{#1}}}
+
 \section{Introduction}
@@
 
+\note{Is this different from other contract work? yes. here's why:}
+
 This approach to managing work (and in particular workers)
@@
 \begin{inlinelist}
-\item distributing the workforce,
+\item distributing the workforce, % gray
 \item distancing workers from the context of the work, and
 \item compensating work on a per--task basis rather than for time
+% citations! future of crowdwork or mary's or lily's work / quin and bederson
 \end{inlinelist}.
@@
 distributed workforces
-are a well--known feature of a global economy,
-%
+are a feature of global labor markets,
 and abstracting workers from the products of their labor
@@
 \citeauthor{marx2012economic} ---
-but these individual components have been explored to some extent
-(for instance, \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
+but these components have been explored only individually
+(for instance, by \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
 and we focus on the newfound interest in
-this particular paradigm of work
+work arranging these components in what holistically seems to be a new mode of work
 % While contract work is familiar enough,
@@
 \cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.
+
+\note{why is this important? because it's bigger than AMT}
 
@@
 to benefit from a latent pool of variably skilled workers.
+
+\note{this is getting into more general ``how has this moved forward'' kind of work. save for later?}
 
@@
 \cite{uberAlgorithm}.
+
 Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
@@
 and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.
+
+\note{this probably doesn't belong here but i need to put it somewhere!}
+
+From various perspectives and with an eye toward myriad stakeholders,
+this body of work has ostensibly sought to answer one overarching question:
+What does the future hold for work and those that do it? % expand on ...?
+% make it clearer
+% inevitable/predictable outcomes?
+% aspirational -> how can we make work better for people
+Whether that research has attempted to articulate the potential to achieve difficult and creative tasks,
+or wrestled with the frustration that workers face on a near--constant basis,
+these roads of inquiry have, it seems, looked to the horizon to try to predict where we are heading.
+
+Rather than looking necessarily to the creation of systems
+(both as genuine agents and as provocateurs (e.g. Soylent and Turkopticon)),
+we turn to the \textit{past}, uncovering ``piecework'' as
+a useful model for understanding the mode of work that has captured the imagination of entrepreneurs and researchers.
 

### 2016-07-25 drafts/labor/introduction.tex
more intro revisions

@@
-% %!TEX root = ./trackingWork.tex
 \documentclass[trackingWork]{subfiles}
@@
 
-\note{Is this different from other contract work? yes. here's why:}
+The many sites of work replicating
+the general style of labor popularized by AMT
+have predominantly involved work done on a computer
+or involving the transfer of human--processed information,
+leading many to refer to AMT as a site of ``information work"
+\cite{turkopticon,professionalCrowdworkEthics,IraniFromCriticalDesign,OlsonMakeDistanceWorkWork}.
+Still, with distributed work emerging with enough variation to problematize this description,
+researchers have tried
+to articulate taxonomies for the sorts of work done by many distributed workers
+\cite{yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
 
+Attempting to triangulate the many sources of research on this area,
+we offer the following takeaway:
 This approach to managing work (and in particular workers)
@@
 \cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.
+
+% \note{Is this different from other contract work? yes. here's why:}
+% \note{\citeauthor{quinnbedersonTaxonomy} offer a taxonomy of crowd work,
+% but at a high level my takeaway consists of the following dimensions:
+% This is work involving\dots
+% \begin{itemize}
+% \item some sort of \textbf{motivation}
+% \item some sort of \textbf{quality control}
+% \item some approach to \textbf{work aggregation} % this is actually kind of useful, but the wrong framing
+% \item some amount of \textbf{human skill}
+% \item some \textbf{process order involving a requester, a worker, and maybe a computer at various points}
+% \item some notion of \textbf{task/request ``cardinality''}
+% \end{itemize}
+% The authors get into the dimensions that these features can take on,
+% but it nevertheless seems so vague that almost any form of  work can be plotted on this 6--dimensional graph.
+% }
 

### 2016-07-26 drafts/labor/introduction.tex
mondays

@@
 % each sentence feels like a thought that's unrelated to the next one
-The past decade has seen a proliferation in on--demand work
-where workers' statuses are so fleeting that 
-they have been described as ``transient''.
-After identifying that myriad tasks 
+The past decade has seen a flourishing of on--demand work
+where the statuses of workers have become so fleeting that 
+these workers
+(known colloquially as ``Turkers'')
+have been described as ``transient''
+\cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
+% After identifying
+The realization
+that myriad tasks
 --- from the mundane to the complex ---
 can be accomplished
-% beginning with ``information work'' such as on
-% Amazon Mechanical Turk (AMT).
-% ??? ->
-% Stemming from the recognition that
-% complex tasks might be achievable
-% --- and indeed trivial ---
-by directing and managing the ``crowd'',
-the research and and industry communities have flocked to sites of labor
+by directing and managing this crowd of workers,
+has spurred the research and and industry communities to flock to sites of labor
 like Amazon's Mechanical Turk (AMT)
-to solicit workers.
+to explore the limits of this distributed, anonymous labor force.
 Researchers in particular have taken to the space in earnest,
 finding opportunities to enable new forms of work
-as well as using these workers (``Turkers'')
+as well as using Turkers
 as representative populations of the public

### 2016-07-28 drafts/labor/introduction.tex
major revisions based on flash meeting feedback

--- 2016-07-19 drafts/labor/introduction.tex
@@
 
-\newcommand{\note}[1]{\onlyinsubfile{\textit{#1}}}
-
+\newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
 \section{Introduction}
-% each sentence feels like a thought that's unrelated to the next one
-The past decade has seen a proliferation in on--demand work
-where workers' statuses are so fleeting that
-they have been described as ``transient''.
-After identifying that myriad tasks
---- from the mundane to the complex ---
-can be accomplished
-% beginning with ``information work'' such as on
-% Amazon Mechanical Turk (AMT).
-% ??? ->
-% Stemming from the recognition that
-% complex tasks might be achievable
-% --- and indeed trivial ---
-by directing and managing the ``crowd'',
-the research and and industry communities have flocked to sites of labor
+The past decade has seen a flourishing of on--demand work
+where the statuses of workers have become so fleeting that 
+these workers
+(known colloquially as ``Turkers'')
+have been described as ``transient''
+\cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
+The realization
+that tasks can be accomplished by directing and managing this crowd of workers
+has spurred the research and and industry communities to flock to sites of labor
 like Amazon's Mechanical Turk (AMT)
-to solicit workers.
+to explore the limits of this distributed, seemingly ephemeral labor force.
 Researchers in particular have taken to the space in earnest,
 finding opportunities to enable new forms of work
-as well as using these workers (``Turkers'')
+as well as using Turkers
 as representative populations of the public
@@
 
-\note{Is this different from other contract work? yes. here's why:}
+The many sites of work replicating and extending on
+the general style of labor popularized by AMT
+have predominantly involved work done on a computer
+or involving the human processing of data,
+leading many to call this ``information work"
+\cite{turkopticon,professionalCrowdworkEthics,IraniFromCriticalDesign,OlsonMakeDistanceWorkWork}.
+\citeauthor{howe2008crowdsourcing} defined ``crowdsourcing'' in \citeyear{howe2008crowdsourcing}
+as
+``taking a job traditionally performed by a designated agent
+(usually an employee)
+and outsourcing it to an undefined,
+generally large group of people in the form of an open call''
+\cite{howe2008crowdsourcing}.
 
-This approach to managing work (and in particular workers)
-can be differentiated from other formulations by a constellation of traits:
-\begin{inlinelist}
-\item distributing the workforce, % gray
-\item distancing workers from the context of the work, and
-\item compensating work on a per--task basis rather than for time
-% citations! future of crowdwork or mary's or lily's work / quin and bederson
-\end{inlinelist}.
-These features can be found elsewhere
---- for instance,
-distributed workforces
-are a feature of global labor markets,
-and abstracting workers from the products of their labor
-is well--associated with
-\citeauthor{marx2012economic} ---
-but these components have been explored only individually
-(for instance, by \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
-and we focus on the newfound interest in
-work arranging these components in what holistically seems to be a new mode of work
-% While contract work is familiar enough,
-% much of this work focuses on comprehensive work
-% (for example, creative or professional work like landscaping, design, etc\dots)
-\cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.
+In the years since,
+scholars have generated taxonomies for the work done by many distributed workers
+in an attempt to better categorize and reason about the many forms of work done 
+on information work platforms such as AMT, oDesk, etc\dots
+\cite{yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
+We add that,
+under \citeauthor{howe2008crowdsourcing}'s constraints,
+even \textit{more} new forms of work fall squarely under
+the metaphorical umbrella we collectively call ``crowdsourcing''.
 
-\note{why is this important? because it's bigger than AMT}
-
-On--demand work under this formulation
-has sparked interest across industries ranging from
+Indeed, this on--demand workforce has sparked interest
+across industries ranging
 livery (driving for hire, for example Uber),
 house--cleaning (Handy),
-and various other services (TaskRabbit)
+and generalized services (TaskRabbit)
 \cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
 Today, a rapidly growing transient workforce is forming,
-itself assembling piece--by--piece as industries and researchers find more and more unexpected ways
-to benefit from a latent pool of variably skilled workers.
+itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
+to benefit from a latent pool of previously vetted workers
+\cite{pewSharing}.
 
-\note{this is getting into more general ``how has this moved forward'' kind of work. save for later?}
+% Broadly, this work has imagined the various applications of ``gig''--based work,
+% and making it work for ``requesters''
+% --- academics, industries, and to some extent end--users
+% who might benefit from the product of crowd--powered work
+% \cite{hong2015group,jonBrelig,paolacci2010running}.
+% % ??? -->
+% Since then,
+% the application of digitally mediated environments acting as rapid ``hiring halls''
+% for on--demand tasks
+% has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
+% livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
+% % parentheticals giving examples?
+% have garnered the interest of both industry and research.
 
-Broadly, this work has imagined the various applications of ``gig''--based work,
-and making it work for ``requesters''
---- academics, industries, and to some extent end--users
-who might benefit from the product of crowd--powered work
-\cite{hong2015group,jonBrelig,paolacci2010running}.
-% ??? -->
-Since then,
-the application of digitally mediated environments acting as rapid ``hiring halls''
-for on--demand tasks
-has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
-livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
-% parentheticals giving examples?
-have garnered the interest of both industry and research.
-
-Researchers have made efforts to understand the \textit{people} that have gravitated toward on--demand work
+Researchers have made efforts to understand the people
+that have gravitated toward crowdsourcing platforms
 since its emergence and popularization,
@@
 
-Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
-workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
-and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.
+% Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
+% workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
+% and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.
 
-\note{this probably doesn't belong here but i need to put it somewhere!}
+% \note{this probably doesn't belong here but i need to put it somewhere!}
 
@@
 What does the future hold for work and those that do it? % expand on ...?
-% make it clearer
-% inevitable/predictable outcomes?
-% aspirational -> how can we make work better for people
-Whether that research has attempted to articulate the potential to achieve difficult and creative tasks,
-or wrestled with the frustration that workers face on a near--constant basis,
-these roads of inquiry have, it seems, looked to the horizon to try to predict where we are heading.
+To give several key examples,
+the driving motivation behind much of the research in this space has led us to ask,
+at a high level, three important questions:
+\begin{enumerate}
+  \item What are the (creative, cognitive, emotional) limits of crowdsourcing?
+    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
+  \item What forms of worker management and arrangement are viable (and moreover \textit{effective})?
+    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
+  \item What will work and the place of work look like for the workers?
+    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}.
+    \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
+\end{enumerate}
 
-Rather than looking necessarily to the creation of systems
-(both as genuine agents and as provocateurs (e.g. Soylent and Turkopticon)),
-we turn to the \textit{past}, uncovering ``piecework'' as
-a useful model for understanding the mode of work that has captured the imagination of entrepreneurs and researchers.
+\subsection{Piecework as a lens to understand crowdsourcing}
+This large and growing body of research has conversed
+to varying degrees with labor scholarship,
+but has not offered a persuasive framing for holistically explaining
+the developments in worker processes that researchers have developer, or
+the phenomena in social environments we have observed;
+nor has any research, to our knowledge,
+gone as far as predict future developments.
 
-\subsection{Piecework as a lens to understand gig work}
-These bodies of literature have conversed,
-and indeed begun to identify that their theories are interrelated in some way,
-but thus far a convincing model for making sense of
---- let alone predicting ---
-the phenomena in on--demand labor has proven elusive.
-For instance, 
-making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
-is understandable in its own context, but
-the relationship between the work itself (``turking'')
-and the myriad social outcomes has only briefly been explored
-\cite{crowdcollab}.
+% These bodies of literature have conversed,
+% and indeed begun to identify that their theories are interrelated in some way,
+% but thus far a convincing model for making sense of
+% --- let alone predicting ---
+% the phenomena in on--demand labor has proven elusive.
+% For instance, 
+% making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
+% is understandable in its own context, but
+% the relationship between the work itself (``turking'')
+% and the myriad social outcomes has only briefly been explored
+% \cite{crowdcollab}.
 
-We offer a framing for \textit{all} on--demand work as a modern instantiation of ``piecework'',
-referenced briefly by
+We offer a framing for crowd work spanning the aforementioned industries
+collectively as a contemporary instantiation of ``piecework''.
+Piecework as a metaphor for the type of work at hand is not new
+(indeed,
 \citeauthor{crowdworkFuture}
@@
 \citeyear{crowdworkFuture}
-as a loose analogy to the form of work emerging at the time,
-but not given as much weight as it perhaps could have been
+referenced crowd work as ``piecework'' briefly
+as a loose analogy to the form of work emerging at the time),
+
+% but not given as much weight as it perhaps could have been
 \cite{crowdworkFuture}.
-``Piecework'' here makes use of
-a term historically used to describe work done in the home,
-in manageable tasks,
-often involving clear instructions
-and payment only for work completed, not work done
-(the differentiation then being that
-one would be paid for the \textit{output} of the work,
-not the \textit{duration}).
+% ``Piecework'' here makes use of
+% a term historically used to describe work done in the home,
+% in manageable tasks,
+% often involving clear instructions
+% and payment only for work completed, not work done
+% (the differentiation then being that
+% one would be paid for the \textit{output} of the work,
+% not the \textit{duration}).
 
-This framing prompts us to note several immediately apparent similarities:
-% Given the scope,
-% we can frame piecework and on--demand labor
-% as sharing these important similarities:
-\begin{inlinelist}
-\item this form of work began in the home
-\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
-\item the worker's status
-(not only socially, but also economically)
-is ambiguous, or at least the subject of some controversy.
-\end{inlinelist}
-
-% In the past decade, researchers have observed frustration
-% grow among on--demand workers,
-% with expression of this frustration spanning a wide range of tactics
-% \cite{uberAlgorithm,turkopticon,dynamo}.
-% Attempting to make sense of these case studies has been challenging
-% in part because
-% a wholly encompassing framework for understanding this form of work
-% has thus proven difficult to capture.
+% This framing prompts us to note several immediately apparent similarities:
+% \begin{inlinelist}
+%   \item this form of work began in the home
+%   \item the worker is paid for each discrete piece of work done, regardless of time or effort; and
+%   \item the worker's status
+%   (not only socially, but also economically)
+%   is ambiguous, or at least the subject of some controversy.
+% \end{inlinelist}
 

### 2016-07-28 drafts/labor/trackingWork.tex
more intro revisions
major revisions based on flash meeting feedback

@@
 }
+
+\setlist{nolistsep}
 
@@
   We argue that the literature surrounding ``piecework''
-  informs and even predicts both the contributions
+  informs and even predicts both the contributions \cite{}
   that have been made toward the development of on--demand labor and crowd work
@@
 ---
-\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
-\citeyear{silberman2010ethics}
-\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
+\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
+\citeyear{professionalCrowdworkEthics}
+\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
 In both the historical and contemporary cases of decomposed work,
@@
 research into cognitive loads and stress levels that we discussed among on--demand crowd workers
-\cite{pieceworkBiologicalHarm,hu1961parallel}.
+\cite{hu1961parallel,pieceworkBiologicalHarm}.
 Even the complicated relationships between observers and workers themselves are not necessarily new;

### 2016-07-28 drafts/labor/introduction.tex
bibs and stuff
notes from meeting

@@
 \newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
+
 \section{Introduction}
@@
   \item What are the (creative, cognitive, emotional) limits of crowdsourcing?
+(what are the \textbf{complexity limits}) \note{(hospitals at high level?) -- team scaffolds} \note{why is there a high limit?
+is that different with crowds?}
     \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
   \item What forms of worker management and arrangement are viable (and moreover \textit{effective})?
+  \note{time in motion studies; job design; }
     \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
@@
 \end{enumerate}
+
+\note{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
@@
 collectively as a contemporary instantiation of ``piecework''.
-Piecework as a metaphor for the type of work at hand is not new
-(indeed,
+Piecework as a metaphor for the type of work at hand is not new.
+Indeed,
 \citeauthor{crowdworkFuture}
@@
 referenced crowd work as ``piecework'' briefly
-as a loose analogy to the form of work emerging at the time),
-
+as a loose analogy to the form of work emerging at the time
 % but not given as much weight as it perhaps could have been
@@
 and developments we should expect to see in the maturation of digitally mediated work.
-Finally, we will offer design implications based on this research\dots \ali{???}
+Finally, we will offer design implications based on this research\dots \note{???}
 
@@
 
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+}
+
 \end{document}

### 2016-08-01 drafts/labor/introduction.tex
home changes

@@
 
-From various perspectives and with an eye toward myriad stakeholders,
-this body of work has ostensibly sought to answer one overarching question:
+% From a wide range of academic and industrial perspectives
+% and increasingly identifying the many stakeholders in digitally mediated work,
+The extant body of work has ostensibly sought to answer one underlying question:
 What does the future hold for work and those that do it? % expand on ...?
-To give several key examples,
-the driving motivation behind much of the research in this space has led us to ask,
-at a high level, three important questions:
+Researchers have offered their input on this open question along three major threads of scholarship:
+% the driving motivation behind much of the research in this space has led us to ask,
+% at a high level, three important questions:
 \begin{enumerate}
-  \item What are the (creative, cognitive, emotional) limits of crowdsourcing?
-(what are the \textbf{complexity limits}) \note{(hospitals at high level?) -- team scaffolds} \note{why is there a high limit?
-is that different with crowds?}
-    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
-  \item What forms of worker management and arrangement are viable (and moreover \textit{effective})?
-  \note{time in motion studies; job design; }
-    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
+  \item What are the limits of crowdsourcing?
+  Perhaps more tightly constrained,
+  what can and cannot be done by crowd workers?
+  % (what are the \textbf{complexity limits}) \note{(hospitals at high level?) -- team scaffolds} \note{why is there a high limit? is that different with crowds?}
+    \cite[for example, see][]{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
+  \item What forms of work design, and worker management and arrangement, are viable?
+  % \note{time in motion studies; job design; } or is this more about managers, coops, etc...
+    \cite[for example, see][]{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
   \item What will work and the place of work look like for the workers?
-    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}.
-    \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
+    \cite[for example, see][]{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}
+    % \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
 \end{enumerate}
-
-\note{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
@@
 and developments we should expect to see in the maturation of digitally mediated work.
-Finally, we will offer design implications based on this research\dots \note{???}
+Finally, we will offer design implications based on this research\dots
+% \note{???}
 
-\onlyinsubfile{
-\itshape
-\subsection{notes/what i want to get across\dots}
+\note{
+\textbf{notes/what i want to get across\dots}
+
 \begin{itemize}
@@
 \end{itemize}
-\upshape
-}
+\upshape}
 

### 2016-08-01 drafts/labor/trackingWork.tex
home changes

@@
 \subfile{introduction.tex}
-
-\section{Case Studies}
-The existing body of research has shed light on on--demand labor from various perspectives,
-and revealed a number of topics that,
-through our framing, are clearly situated together.
-Those topics are, at a high level, as follows:
-\begin{enumerate}
-\item the \textbf{processes} involved in making work into tasks, or discretization;
-\item the outcomes (and indeed the \textbf{fallout}) of that discretization,
-both on the work itself as well as the workers;
-and finally
-\item the \textbf{relationships} between workers and requesters of the work
---- both \textit{cooperative} and \textit{adversarial} cases.
-\end{enumerate}
-
-% We will follow these topics, using them as prompts for case studies in
-% the emergence and
-% development of contemporary piecework.
-
-\subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
-% this is phase 1: the decomposition & abstraction of work;
-% all of the research that has gone into work--flows and
-% all the things we can do with crowd work.
-
-\cite{olson2014ways}
-The HCI community is perhaps most familiar with examples of task--based work such as
-99designs,
-Amazon Mechanical Turk (AMT),
-and increasingly
-Uber \& Lyft,
-which all allow requesters in various forms to tap into
-resources such as
-cars,
-computers, and above all
-``cognitive surplus''
-with relative ease
-\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
-This insight, that
-workers can be geographically distributed and tasks decomposed,
-% (especially given careful central management of those workers, discussed later),
-has proven remarkably compelling
-and an effective fulcrum for leveraging the Internet for highly scalable work
-\cite{sensitiveTasks,embracingErrorKrishna}.
-
-This section will largely discuss the processes at work that make distributed,
-digitally managed work both possible and indeed preferable for ``requesters''
-(in other words, the employers who solicit workers).
-This body of research spans a broad field
-within the CSCW and broader HCI community.
-In this context,
-we'll look at the body of research through the lens of
-highlighting contributions which expand what we
-(as ``requesters'' of work) can do by managing workers in novel ways.
-This work broadly consists of three areas:
-\begin{inlinelist}
-\item \nameref{sec:decomposition},
-\item \nameref{sec:workAbstraction}, and
-\item \nameref{sec:flexibility}
-\end{inlinelist}.
-As we explore this work, we'll
-attempt to relate the advances in the design of crowd work to
-the research contributions made in the research of assembly line manufacturing,
-Taylorism, and
-scientific management 
-during the \nth{20} century
-\cite{hu1961parallel}.
-
-\subsubsection{Decomposition}\label{sec:decomposition}
-Piecework may be thought of as vertically slicing work such that 
-each person is responsible for the whole task
---- making a whole garment, in this case.
-Broken down in this way, work could grow to unprecedented scales,
-but the quality of the work would remain relatively variable
-\cite{murray1983decentralisation}.
-Textile work being a salient example,
-it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work
-\cite{vezina1992light}.
-
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
-manageable routines that could be taught relatively easily,
-and whose work output could be evaluated in abstraction from the rest of the work
-\cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-but a single very narrowly defined action that needed to be done on every car
-\cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,
-``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
-\cite{schoenberger1988fordism}.
-
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),
-can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,
-facilitating evaluation and parallelization
-\cite{writingMicroTasks}.
-By decomposing and recomposing tasks,
-and in particular by assigning similarly natured work to the same workers,
-workers could become ``experts'' in a small aspect of the work that they did,
-speeding their work dramatically
-\cite{delayAndOrderLasecki}.
-Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage
-\cite{rogstadius2011assessment}.
-
-Scholarship describing and exploring
-the decomposition of tasks is perhaps the most established of the above areas among HCI researchers;
-\citeauthor{crowdForgeKittur} specifically drive at this goal by addressing the possibility of
-``crowdsourcing complex work''
-\cite{crowdForgeKittur}.
-\citeauthor{cheng2015break} found that microtasks
---- though not necessarily \textit{faster} than ``macrotasks'' ---
-yield higher quality work,
-particularly when that work might be readily interrupted
-\cite{cheng2015break}.
-\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work
-by exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-While this work doesn't strictly fall under ``crowdsourcing'',
-the major contributions here
-seem uncontroversially to be inspired by the design of crowd work.
-
-Much of the research in the space of designing crowd work has
-sought to illustrate the potential to take highly creative or skilled work
-and generate high--quality results.
-Perhaps the most notable case study here can be found in
-\citeauthor{foundry}'s \textit{Foundry}, which employed
-``flash teams'' to achieve expert--level outcomes via thoughtful
-decomposition of work as ``modular tasks''
-\cite{foundry}.
-
-Work decomposition, then, is far from new;
-``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
-and scientific management sought to embody
----
-\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
-\citeyear{professionalCrowdworkEthics}
-\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
-In both the historical and contemporary cases of decomposed work,
-work was,
-at least initially,
-distributed in the form of tasks to the homes of workers;
-\citeauthor{riisOtherSideLives} captured this in his documentary work
-\citetitle{riisOtherSideLives}
-in \citeyear{riisOtherSideLives}
-\cite{riisOtherSideLives}.
-
-\subsubsection{Work Abstraction}\label{sec:workAbstraction}
-Decomposition allows requesters to assign tasks without concern for the broader context.
-While we'll discuss this aspect of crowd work more critically later,
-it's worth pointing out that discrete blocks of work containing all the relevant context for a worker
-allows workers to engage with virtually any component of work without worrying that their lack of 
-higher--level awareness of the goals of the requester might negatively affect their work.
-
-\citeauthor{chilton2013cascade} perhaps best illustrated this with
-\textit{Cascade} by demonstrating that it's possible to
-break certain classes of tasks apart
-in such a way that they yield taxonomies of various subjects,
-a task previously thought to be safely within the domain of expert workers
-with top--down awareness of the context of the work as a whole
-\cite{chilton2013cascade}.
-\citeauthor{verroios2014context} further illustrate this potential by
-forming a task one might consider highly contextually dependent
---- summarizing the contents of a movie ---
-in such a way that crowd workers could contribute small pieces of work without
-needing to know the content of the rest of the project
-\cite{verroios2014context}.
-
-Here, \citeauthor{hu1961parallel}'s work,
-saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
-parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work
-\cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,
-and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
-substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,
-and persists to this day as it colors
-researchers' goals and objectives in the study and design of crowd work.
-
-Piecework's influence on the abstraction of work into tasks,
-described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,
-prompting work requesters to consider ways to design these now discrete tasks in ways that
-maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
-\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
-It's to this area that we now turn our attention.
-
-\subsubsection{Flexibility}\label{sec:flexibility}
-% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
-
-Earlier we discussed \citeauthor{cheng2015break}'s work
-measuring the impact that interruption has on worker performance.
-This work both points to and embodies a broader sentiment in
-both the study and practice
-of crowd work that microtasks should be designed resiliently against the variability of workers,
-% \msb{not sure what flexibly means here, and the rest of the sentence isn't helping me unpack it. flexibility for requester? worker? flexible algorthms?},
-fully exploiting the abstracted nature of each piece of work
-\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-That is to say, micro--tasks should be designed such that a single worker's poor performance,
-or a good worker's sudden departure,
-would not significantly impact the agenda of the work as a whole.
-While \citeauthor{cheng2015break} identified costs with breaking tasks into smaller components
-in the form of higher cumulative time to complete
-(albeit much shorter real time to complete, owing to parallelization),
-\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
-similar tasks together.
-
-Given the importance of consistent work results, one might intuit that
-requesters would prefer high--quality workers who can be relied upon to be available
-(even for contextually independent tasks),
-which would appear to contradict the benefits of flexibility already discussed;
-requesters have thus made significant headway toward
-``embracing error'' to allow requesters to maximize the benefits of a flexible,
-even transient,
-workforce.
-
-\citeauthor{embracingErrorKrishna} offer orders--of--magnitude improvements
-in various binary classification tasks
-on the principle that diverse workers complete these tasks
-in order to accurately inform the model on the variety of delays in response times.
-And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-some researchers have designed work predicated on the expectation of this phenomenon:
-\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
-crowd workers would never be given enough information to piece together sensitive information about
-any single topic
-\cite{sensitiveTasks}.
-
-Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
-\citeauthor{tolliday1986between}'s treatment describing
-turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
-Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
-[and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
-piece rates~\dots~rather than a day wage system''
-\cite{tolliday1986between}.
-
-% \msb{I still dont' have a clear conceptual picture of what flexibility is, and what about it sets this section apart. Ranjay's stuff is about speed...? I'm confused.}
-
-\subsection{The Fallout of Crowd Work}\label{sec:Fallout}
-% If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
-% the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
-\citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
-(see also \citeauthor{dynamo}'s work
-continuing in the spirit of this observation to generate collective action to improve worker conditions)
-\cite{turkopticon,dynamo}.
-\citeauthor{uberAlgorithm}
-find similarly that workers on gig work platforms are frustrated by the systems on which they work,
-to say little of the policies which these systems enforce
-\cite{uberAlgorithm}.
-
-We discussed the benefits of flexibility
-(both in the sense of having arbitrary workers perform tasks and
-in the sense that we can design tasks to be more resilient to poor work)
-in the previous section.
-It's from that point in the literature that we turn our attention to
-the perhaps unintended effects of crowd work
-and the affordances for transience that we build into this mode of work.
-We'll address two major areas of work under this subject:
-\begin{inlinelist}
-\item \nameref{sec:lowPay}; and
-\item \nameref{sec:varQualWork}.
-\end{inlinelist}
-
-\subsubsection{Low Pay}\label{sec:lowPay}
-
-\citeauthor{laborEconomicsOfCrowdsourcingHorton}
-identified problems with crowd work wages relatively early on,
-attempting to address this imbalance from a behavioral economic perspective ---
-that is, identifying and presenting a model that describes a worker's
-``\textit{reservation wage}''
-\cite{laborEconomicsOfCrowdsourcingHorton}.
-This work has largely informed much of the research into and practice of estimating crowd work compensation
-\cite{incentivesShaw,paolacci2010running}.
-
-But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}'',
-a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
-\cite{turkopticon}.
-Of particular relevance here,
-\citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
-wage theft and
-pay rates that translate to well under minimum wage.
-Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},
-we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
-was \$1.38, while the mean was \$3.63
-\cite{laborEconomicsOfCrowdsourcingHorton}.
-
-Understanding workers' motivations given these conditions has thus become a goal for some researchers
-\cite{whyWouldAnyoneBrewer}.
-\citeauthor{Sun20111033} conclude that
-``\dots~solvers participate in online tasks
-not only for money
-but also for enjoyment
-or the sense of self--worth''
-\cite{Sun20111033}.
-This might have rung true in \citeyear{Sun20111033},
-and certainly corroborates \citeauthor{Ross}'s findings after investigating
-``who are the crowdworkers'',
-but as \citeauthor{whoareNOTtheTurkers} points out
-``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
-\cite{Ross,whoareNOTtheTurkers}.
-
-Now, \citeauthor{turkopticon}
-and later
-\citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
-via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
-\cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
-
-On--demand workers were not the first to be exploited along the dimension of low pay rates.
-Frustration over low (and declining) pay was one of the chief grievances among then nascent
-British labor unions in the early \nth{20} century
-\cite{turner1952trade}.
-This, \citeauthor{ebbinghaus1999institutions} argued,
-fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
-(to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
-\cite{ebbinghaus1999institutions,levi2009union}.
-This realization has similarly fueled a body of research into
-the various incentive structures available to piecework employers
-\cite{roy1953work}.
-
-The parallels between the complaints of low pay among crowd workers and other on--demand workers
-and the pieceworkers and later factory workers in the \nth{20} century
-are inescapable.
-We argue further that the \textit{causes} here
---- work decomposition,
-work abstraction, and
-flexibility ---
-lead inexorably to low and declining pay for workers.
-Moreover, we point out that low pay leads to other negative outcomes both
-in on--demand work
-as well as
-in piecework and on assembly lines.
-
-\subsubsection{Variable quality work}\label{sec:varQualWork}
-Researchers have struggled with what we might generously call work of ``variable quality''
-along two dimensions.
-The first, to use the characterization of one of these contributions, we can call
-``understanding malicious behavior''
-\cite{MaliciousCrowdworkersGadiraju}.
-While some work has cast workers as ``malicious'' or at least adversarial parties,
-the evidence thus far suggests that
-workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
-(a topic of discussion to which we'll return later)
-\cite{uberAlgorithm}.
-The second dimension of research in this space generally attempts
-to eke out the highest quality work possible from workers
-given the apparent difficulty in predicting work outcomes
-\cite{embracingErrorKrishna}.
-
-% Low pay yields variable quality work for a number of reasons,
-% but before we discuss the causes of this effect, we should discuss the
-
-% Variance in the quali
-
-The effect low wages have had on piece work and factory workers is well--known;
-\citeauthor{gantt1913work} discuss this exact mechanism in his book on
-\citetitle{gantt1913work}, pointing out that
-``\dots~where there is no union,
-the class wage is practically gauged by the wages the poor workman will accept,
-and the good workman soon becomes discouraged and \textit{sets his pace by that of his less efficient neighbor},
-with the result that the general tone of the shop is lowered'' (emphasis added)
-\cite{gantt1913work}.
-
-This research is similar to, but subtly different from, the notion of the ``market for `lemons'''
-which \citeauthor{fort2011amazon} discuss;
-specifically, \citeauthor{akerlof1970market}'s writing of a ``market for `lemons'''
-describes a marketplace where the quality of the product or service is unknown to the buyer
-\cite{fort2011amazon,akerlof1970market}.
-The effect of this \textit{perceived} uncertainty is that
-the \textit{actual} trustworthiness drops precipitously
-as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,
-leaving only the ones who cannot or will not establish their trustworthiness.
-
-\subsection{Relationships Between Workers and Managers}\label{sec:relationships}
-Suffice it to say that poor pay and poor work are linked,
-and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-But the poor treatment of workers by managers
---- both human and algorithmic ---
-do more than affect the economic relationships between workers and employers.
-Here, then, we turn to examine this facet of on--demand work
-and how these dynamics strikingly replicate the relationships
-researchers in labor advocacy encountered in the study of piecework and factory work.
-
-This topic can be condensed into two major areas:
-\begin{inlinelist}
-\item external (scientific) management, and the evaluation of workers as functional modules; and
-% \item the alienation of workers under this form of evaluation; and
-\item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
-\end{inlinelist}
-
-\subsubsection{External Management}
-We discussed Fordism and Taylorism earlier in our discussions of
-\nameref{sec:decomposition} and \nameref{sec:flexibility},
-but here the core of these paradigmatic views
---- the scientific management of work ---
-becomes relevant.
-We use ``external'' here instead of ``scientific'', however,
-to more broadly capture the disconnect between managers and workers.
-By describing it as thus,
-we can touch on the relationship that workers have with \textit{researchers}, as well,
-even though that work is not strictly
---- or just not exclusively ---
-of the same nature as the management and experience as when interacting with requesters.
-
-First, intuitively, the variable--quality work we discussed previously has led to
-a large and growing body of research attempting to evaluate workers' performance and error rates
-across numerous dimensions;
-for example, \citeauthor{measuringCrowdsourcingCheng} explore the error rates of workers by
-operating on a sliding scale giving workers varying amounts of time to accomplish micro--tasks
-\cite{measuringCrowdsourcingCheng}.
-\citeauthor{storiesIraniSilberman} describe the treatment of workers
-as sorts of ``human APIs'' that can, importantly, be rigorously evaluated
-\cite{storiesIraniSilberman}.
-\citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
-cognitively demanding tasks on workers,
-informing crowdsourcing research by suggesting the use of cognitive load assessments such as
-NASA Task Load Index surveys to evaluate workers pre and post--tasks
-\cite{embracingErrorKrishna,measuringCrowdsourcingCheng}.
-
-External management comes in other forms than scientific, as previously mentioned.
-Researchers in particular have noticed that their relationships with on--demand workers are,
-at the least, complex.
-\citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
-specifically, their interactions with field sites in which they work
-as designers and mediators of change influence the relationships they have with Turkers
-\cite{storiesIraniSilberman}.
-
-The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
-and even physiological study of pieceworkers closely resembles the
-research into cognitive loads and stress levels that we discussed among on--demand crowd workers
-\cite{hu1961parallel,pieceworkBiologicalHarm}.
-Even the complicated relationships between observers and workers themselves are not necessarily new;
-\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
-to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
-\cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.
-
-Similarly, \citeauthor{pollard1963factory}'s words on the punishment factory workers faced ---
-for example, that ``unsatisfactory work was punished \dots~by fines or by dismissal'' ---
-seems especially relevant given the fears we now know to be ubiquitous on platforms such as AMT, Uber, and other on--demand markets
-\cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.
-
-% \subsubsection{Alienated Workers}
-% The dissatisfaction on crowd work systems is not as well--explored as their actual responses to that dissatisfaction
-% (in various forms of resistance),
-% but it's nevertheless worthwhile to point out some of the observations that have been made here,
-% as external management, discussed earlier, and the discussion on worker resistance, to be discussed next,
-% are bridged by this quality of on--demand labor markets.
-
-% Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
-% T
-
-% \itshape
-% Marx's big entrance. Lily has lots of work pointing to this, and
-% lots of that work points to yet more work gesturing in this direction.
-% I want to bring a lot of that work
-% that hasn't made it into the discussion yet
-% into the conversation here.
-
-% Note to self: examples of citations would help make it clearer for MSB which ones I'm overlooking,
-% if nothing else
-
-% \upshape
-
-\subsubsection{Resistance}
-It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,
-often without recourse or opportunity for feedback, let alone substantive input.
-Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
-many toggle their availability to avoid being dispatched to more distant locations,
-resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
-\cite{uberAlgorithm}.
-
-Resistance has sometimes been more coordinated, as well;
-we see this in \citeauthor{turkopticon}'s coverage on \textit{Turkopticon} as workers collectively accumulated information about requesters, and
-in \citeauthor{dynamo}'s work on \textit{Dynamo}, which generated ``Guidelines for Academic Requesters'' written by crowd workers
-\cite{turkopticon,dynamo}.
-
-Resistance against managers in piecework and factory labor settings are deeply well--explored,
-but perhaps the most relevant case study to draw on here is to be found in
-\citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',
-where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
-\cite{waldinger1996helots}.
-The achievements of labor advocacy groups such as labor unions as resistant,
-even adversarial organizations counter--balancing the management
-is somewhat well--understood
-\cite{russell1982collective,craig1992behavior}.
-We argue that these threads of resistance against management in various forms
-are in fact one.
-
-% \subsection{Resistance}
-% So it's not surprising, then, that workers are frustrated:
-% Turkopticon, Dynamo, and many other pieces of work point this out
-% \cite{turkopticon,dynamo,uberAlgorithm}
-% [\textit{also want to cite Brian's work from CHI this year}].
-% In popular culture, reporters have inquired about
-% what ``the future of labor unions'' will look like
-% [\textit{Imagine I had cited an online article and
-% it wasn't totally inappropriate for a conference}].
-
-% \textit{The parallels here would be difficult to ignore;
-% Lily and others have pointed this out,
-% but it's worth thinking about the broader trend of
-% resisting management,
-% driven by routinization,
-% enabled by discretization.
-% Do that here.}
+\subfile{cases.tex}
 

### 2016-08-01 drafts/labor/cases.tex
home changes

--- 2016-07-16 cscw/papers/2017/labor/trackingWork.tex
@@
-\documentclass{sigchi}
-
-\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth}
-\usepackage[normalem]{ulem}
-\usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
-\usepackage[T1]{fontenc}
-\usepackage[pdftex,pdfpagelabels=false]{hyperref}
-\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
-\usepackage[utf8]{inputenc} % for a UTF8 editor only
-
+\documentclass[trackingWork]{subfiles}
 \makeatletter
@@
 \makeatother
+\begin{document}
+\newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
 
-% Paper metadata (use plain text, for PDF inclusion and later
-% re-using, if desired).  Use \emtpyauthor when submitting for review
-% so you remain anonymous.
+\section{Major Research Questions}
+\note{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
-\newlist{inlinelist}{enumerate*}{1}
-\setlist*[inlinelist,1]{%
-  label=\arabic*),
-}
-
-\def\plaintitle{Reexamining Crowd Work: A Historical Framing of On--Demand Labor as Piecework}
-\def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
-\def\emptyauthor{}
-\def\plainkeywords{Crowdsourcing}
-\def\plaingeneralterms{Documentation, Standardization}
-
-% llt: Define a global style for URLs, rather that the default one
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{
-    \def\UrlFont{\sf}
-  }{
-    \def\UrlFont{\small\bf\ttfamily}
-  }}
-\makeatother
-\urlstyle{leo}
-
-% To make various LaTeX processors do the right thing with page size.
-\def\pprw{8.5in}
-\def\pprh{11in}
-\special{papersize=\pprw,\pprh}
-\setlength{\paperwidth}{\pprw}
-\setlength{\paperheight}{\pprh}
-\setlength{\pdfpagewidth}{\pprw}
-\setlength{\pdfpageheight}{\pprh}
-\definecolor{PineGreen}{HTML}{008800}
-\definecolor{BrickRed}{HTML}{FF0000}
-\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
-\newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}
-
-% Make sure hyperref comes last of your loaded packages, to give it a
-% fighting chance of not being over-written, since its job is to
-% redefine many LaTeX commands.
-\definecolor{linkColor}{RGB}{6,125,233}
-\hypersetup{%
-  pdftitle={\plaintitle},
-% Use \plainauthor for final version.
-%  pdfauthor={\plainauthor},
-  pdfauthor={\emptyauthor},
-  pdfkeywords={\plainkeywords},
-  bookmarksnumbered,
-  pdfstartview={FitH},
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
-  urlcolor=linkColor,
-  breaklinks=true,
-  hypertexnames=false
-}
-
-% create a shortcut to typeset table headings
-% \newcommand\tabhead[1]{\small\textbf{#1}}
-\bibliography{../../../../references}
-
-\newcommand{\onlyinsubfile}[1]{#1}
-\newcommand{\notinsubfile}[1]{}
-\begin{document}
-\renewcommand{\onlyinsubfile}[1]{}
-\renewcommand{\notinsubfile}[1]{#1}
-\title{\plaintitle}
-
-\numberofauthors{3}
-\author{%
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-}
-
-\maketitle
-
-\begin{abstract}
-  With growing attention toward on--demand labor
-  --- ranging from the ``sharing economy'' to information work ---
-  scholars have made connections to various frameworks 
-  and mechanisms such as worker advocacy, empowerment, and Taylorism,
-  to make sense of our observations of
-  on--demand work and the workers that power this movement.
-  We argue that the literature surrounding ``piecework''
-  informs and even predicts both the contributions
-  that have been made toward the development of on--demand labor and crowd work
-  as well as the fallout among workers and researchers
-  with regard to the disillusionment and alienation of work.
-
-  After evaluating this framing
-  through a series of case studies,
-  we look to the future
-  to identify worthwhile questions and
-  points of inquiry, such as the movement toward factories,
-  that researchers in social computing should consider
-  as we attempt to anticipate and perhaps shape the future of work.
-\end{abstract}
-
-\category{H.5.m.}{Information Interfaces and Presentation
-  (e.g. HCI)}{Miscellaneous} \category{See
-  \url{http://acm.org/about/class/1998/} for the full list of ACM
-  classifiers. This section is required.}{}{}
-
-\keywords{\plainkeywords}
-\subfile{introduction.tex}
-
-\section{Case Studies}
+\onlyinsubfile{
+\itshape
+% \section{Case Studies}
 The existing body of research has shed light on on--demand labor from various perspectives,
@@
 
-% We will follow these topics, using them as prompts for case studies in
-% the emergence and
-% development of contemporary piecework.
+\subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
 
-\subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
-% this is phase 1: the decomposition & abstraction of work;
-% all of the research that has gone into work--flows and
-% all the things we can do with crowd work.
-
+\cite{olson2014ways}
 The HCI community is perhaps most familiar with examples of task--based work such as
@@
 \cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.
-
-% \subsubsection{Alienated Workers}
-% The dissatisfaction on crowd work systems is not as well--explored as their actual responses to that dissatisfaction
-% (in various forms of resistance),
-% but it's nevertheless worthwhile to point out some of the observations that have been made here,
-% as external management, discussed earlier, and the discussion on worker resistance, to be discussed next,
-% are bridged by this quality of on--demand labor markets.
-
-% Unsurprisingly, workers have expressed a disassociation from the platforms on which they work and the work they do;
-% T
-
-% \itshape
-% Marx's big entrance. Lily has lots of work pointing to this, and
-% lots of that work points to yet more work gesturing in this direction.
-% I want to bring a lot of that work
-% that hasn't made it into the discussion yet
-% into the conversation here.
-
-% Note to self: examples of citations would help make it clearer for MSB which ones I'm overlooking,
-% if nothing else
-
-% \upshape
 
@@
 
-% \subsection{Resistance}
-% So it's not surprising, then, that workers are frustrated:
-% Turkopticon, Dynamo, and many other pieces of work point this out
-% \cite{turkopticon,dynamo,uberAlgorithm}
-% [\textit{also want to cite Brian's work from CHI this year}].
-% In popular culture, reporters have inquired about
-% what ``the future of labor unions'' will look like
-% [\textit{Imagine I had cited an online article and
-% it wasn't totally inappropriate for a conference}].
-
-% \textit{The parallels here would be difficult to ignore;
-% Lily and others have pointed this out,
-% but it's worth thinking about the broader trend of
-% resisting management,
-% driven by routinization,
-% enabled by discretization.
-% Do that here.}
-
-\section{The Bleak Future of Crowd Work}
-We've traced a path from piecework itself through
-the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
-in tracing this process, we touched on the relationships between decomposition,
-work \& worker abstraction,
-flexibility, and followed through both the general fallout of crowd work in the research community
-as well as the fallout between workers and the managers and other external parties
---- including researchers.
-
-Throughout these case studies, we have pointed out the parallels between
-the contemporary research in on--demand labor and
-the much larger body of research constituting our understanding of topics such as
-piecework, factory work, and laborer relations.
-If we agree that this framing is useful and informative, then
-several topics emerge as relatively open questions in the study of crowd work and on--demand labor.
-Two of the most pressing questions are
-\begin{inlinelist}
-\item the beginnings of factories, and 
-\item the decline of relevance of worker advocacy organizations.
-\end{inlinelist}
-We will discuss those questions here.
-
-\subsection{The beginnings of factories}\label{sec:Factorization}
-We established earlier that abstracted work and low wages tend to result in variable outcomes,
-which presents problems for employers.
-% So work output is variable, and that's bad. What are we doing about it?
-Historically, this is what led to factories; by employing a cohort of known workers,
-we can be reasonably assured that the quality of the work will be better than random.
-Furthermore, we can invest more resources in training workers and
-get workers to do more complex work with more context.
-% you would pay some experts who understood what they needed to do
-% and they would form a cohort into which you invest more.
-
-Some research already looks at research such as investing in workers, and
-informally, we know that this happens among industry requesters
-\cite{jonBrelig,shepherdingDow}.
-AMT, meanwhile, offers requesters the ability to create tasks which are
-not just hidden from unqualified workers by default, but completely.
-Requesters have taken to using lists of worker IDs which reference
-workers who have proven their reliability,
-representing a sort of proto--organization of loosely connected workers.
-
-This, then, suggests that the beginning of the regularization of workforces
---- a sort of coalescence of factories ---
-is already happening.
-If our framing of on--demand labor is accurately describing an underlying relationship with piecework,
-then we should watch for the emergence and popularization of persistent teams of workers.
-% While this field site will likely prove exceedingly difficult to 
-% Social computing researchers should seek to understand this process more
-% fully, as we a
-
-\subsection{The decline of advocacy organizations}
-The rise of labor unions in the \nth{20} century seems to have been precipitated by
-severely unjust conditions imposed on workers in factories and elsewhere
-\cite{ebbinghaus1999institutions}.
-Incidents broadly describing this dynamic can be found in research on AMT
-\cite{turkopticon,dynamo}.
-If these are prototypical labor advocacy organizations of contemporary on--demand work,
-the next question we should look to is if
---- and indeed \textit{how} ---
-these institutions might face challenges in the future.
-
-For insight on this, we return to \citeauthor{levi2009union}'s study of labor unions,
-and identify that
-``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
-\cite{levi2009union}.
-% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
-% unable to satisfy their needs or otherwise more invested in the 
-This perception already appears to be emerging in digitally mediated peer--governed organizations,
-as \citeauthor{keegan2010egalitarians} and others have illustratively documented
-\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
-If these organizations and others are to avoid the same fate that labor unions faced,
-they should take care to study this phenomenon and attempt to avoid it.
-
-\section{Implications for Design}
-If it's agreed that
-the major topics we've discussed thus far are related and
---- at least to \textit{some} extent ---
-precipitated in the fashion we argue,
-then we have a rare opportunity as researchers,
-and as agents of change in the communities we study,
-to affect change on the dynamics of crowd and on--demand work
-as they continue to develop.
-
-Without claiming to have easy, cut--and--dry solutions to these problems,
-we can nevertheless bring to attention a number of critical opportunities to
-learn from historical parallels in piecework and factory labor,
-and make informed decisions regarding whether
-(or indeed how)
-we may want to influence outcomes.
-The challenges we bring to attention here are as follows:
-\begin{inlinelist}
-  \item codifying investment toward collective goods into the designs of systems;
-  \item (re--)decentralizing the internet; and
-  \item enabling reputation transferral.
-\end{inlinelist}
-
-\subsection{Codify the common good}
-As \citeauthor{lessig2006code} points out in his book,
-digital media give designers the opportunity to design and build into the systems
-policies and practices to contribute to the collective benefit of the people therein
-\cite{lessig2006code}.
-Historically, the confluence of forces \citeauthor{lessig2006code} describes
-would ultimately result in outcomes such as benefits for workers,
-funds for sick leave and vacation, and other conveniences.
-The transient nature of on--demand work would seem to problematize this arrangement,
-but we can discuss and explore the viability of building into systems the mechanisms necessary
-to save a portion of payment from every gig,
-record taxable income, or
-myriad other generally administrative tasks automatically.
-
-\subsection{Decentralize the internet --- again}
-Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
-forcing workers either to choose one or juggle participation in these markets with great difficulty.
-An ``API'' for on--demand labor markets could make it possible for any person
-or organization to instantiate their own marketplace and inter--operate with.
-This can be changed, and indeed must, if we are to realize the hopes of early researchers who advocated the
-democratizing nature and power of the internet
-\cite{barlow2009declaration,lanier2014owns}.
-
-\subsection{Deal with reputation}
-Reputation systems in on--demand labor markets are fundamentally broken.
-To say nothing of the fact that information workers (such as those on AMT) can't transfer their reputations to qualitatively different forms of labor
-like driving--for--hire (e.g. Uber), even within the same industry
-it's currently not feasible for workers to transfer their reputations or other information from one place to another.
-This affects more than the reputation and trustworthiness of workers;
-accounting for things such as taxes, benefits, etc\dots is all but left to the individual workers, who struggle with myriad bureaucratic obstacles.
-We can design systems that facilitate the aggregation and, more importantly, the transferral of reputation, income, and other features of work.
-
-\section{Discussion}
-We've discussed a number of aspects of on--demand work that
-offer parallels with historical piecework.
-Perhaps more importantly, we've hopefully demonstrated that the dynamics we observe in on--demand work
-are interrelated and follow from one another just as necessarily as they did in
-the development and maturation of piecework and factory work through the \nth{20} century.
-This framing on on--demand work should, we hope, provide us with the necessary historical context
-to make better--informed design decisions about how we want ``the future of crowd work'' to look.
-
-\section{Conclusion}
-\citeauthor{crowdworkFuture} discussed many of the challenges and problems in crowd work in \citeyear{crowdworkFuture},
-but didn't necessarily situate the notion of crowd work in a broader context.
-This paper attempts to fill that gap, and in doing so hopes
-to give the research community theoretical grounding to work with and within on--demand labor more successfully.
-But more than that, we hope to have addressed important questions to inform how we actually might make crowd work
-a career in which we want our children to work.
-
-\balance{}
-\printbibliography
-
+\upshape
+}
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2016-08-02 drafts/labor/cases.tex
monday

@@
 \note{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
+
+\subsection{What are the limits of crowdsourcing?}
+For more than a decade
+research in crowd work has explored ways of expanding the work that crowdsourcing can do by
+  identifying challenges to this form of labor,
+  overcoming them through novel designs of work--flows and processes,
+and repeating the process
+\cite[etc\dots]{bernsteinSoylent}.
+The question that has emerged
+among these researchers and
+through the work that they have produced then
+has been whether there are limits to crowdsourcing
+(and, if so, what bounds them).
+
+In the field of piecework,
+the research covering this topic has both explored
+a breadth of tasks that might be rendered doable by piecemeal workers
+\textit{as well as} longitudinally documented the success of these approaches.
+
+Points to make:
+  \begin{itemize}
+    \item Blossoming of piecework
+    \begin{itemize}
+      \item high point: consultants?
+      \item most cases: auto workers/etc\dots
+      \item worst case: sweatshops (especially in developing nations)
+    \end{itemize}
+    \item what led to these outcomes?
+    \begin{itemize}
+      \item ``consultant'' work came out well because the work was complex;
+      this made it difficult to turn into a mass market commodity.
+      We see consultants ranging skill levels like
+      oDesk (implementing modules) to
+      Accenture (on--demand teams of consultants).
+      \item auto workers,
+      working in settings where capital couldn't be moved as easily
+      found themselves in the same workspace as a direct
+      --- if multi--stepped ---
+      result to the benefits of putting people in the same place to consolidate resources.
+      Moreover,
+      workers had leverage over factory owners as a result of that consolidated capital;
+      operating equipment required training.
+      \item sweatshop workers fared the worst,
+      for reasons that may seem obvious with hindsight.
+      Source materials for textiles are easier to ship than mechanical components such as engines,
+      making it easier for factories to relocate to developing nations
+      (where cost--of--living, and consequently wages, would be lower).
+      As wages, CoLAs, and QoL rise, workers begin asking for
+      (and later demanding)
+      higher wages and better conditions.
+      But where the auto workers have leverage,
+      textile workers find only a precarious economic balance now tipped by their collective action,
+      spurring manufacturers to move to a new locale
+
+    \end{itemize}
+  \end{itemize}
+
+\subsection{What forms of work design and worker management are viable?}
+\subsection{What will work and the place of work look like for workers?}
 
@@
 \subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
-
-\cite{olson2014ways}
 The HCI community is perhaps most familiar with examples of task--based work such as

### 2016-08-03 drafts/labor/cases.tex
home stuff

@@
 \begin{document}
-\newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
+% \newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
 
 \section{Major Research Questions}
-\note{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
+\ali{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
@@
       spurring manufacturers to move to a new locale
-
     \end{itemize}
@@
 \subsection{What forms of work design and worker management are viable?}
+
 \subsection{What will work and the place of work look like for workers?}
+The metaphorical mechanics of these dynamics are still at play;
+workers and managers continue to interact in adversarial manners,
+despite substantive work into aligning the motivations of workers and requesters
 
 \onlyinsubfile{
-\itshape
+\clearpage
 % \section{Case Studies}
@@
 are in fact one.
-
-\upshape
 }

### 2016-08-04 drafts/labor/cases.tex
wed

@@
 \subsection{What forms of work design and worker management are viable?}
+\begin{itemize}
+  \item researchers have looked at how to increase worker productivity
+  (e.g. finding the maximal speed at which gig workers can be expected to work before making errors)
+  \cite{measuringCrowdsourcingCheng}.
+  \item we've also seen people ``embrace error''
+  \cite{embracingErrorKrishna}.
+  \item still other research has looked into ways to sandbox workers from the context of their work 
+  \item but scholarship looking into the design and management of work and workers isn't new;
+  lots of research into getting pieceworkers to do work more quickly
+  \cite{seymour1954manual}.
+  \item Researchers have even asked the age old question of \textit{what motivates} pieceworkers
+  (echoing similar research on Wikipedia and Mechanical Turk)
+  \cite{roy1953work,Nov:2007:MW:1297797.1297798,kaufmann2011more}
+\end{itemize}
 

### 2016-08-09 drafts/labor/cases.tex
debugging a bunch of unnecessary stuff to get biblatex working...

@@
 and repeating the process
-\cite[etc\dots]{bernsteinSoylent}.
+\cite{bernsteinSoylent}.
 The question that has emerged
@@
 \citeauthor{riisOtherSideLives} captured this in his documentary work
-\citetitle{riisOtherSideLives}
+% \citetitle{riisOtherSideLives}
 in \citeyear{riisOtherSideLives}
@@
 \citeauthor{gantt1913work} discuss this exact mechanism in his book on
-\citetitle{gantt1913work}, pointing out that
+% \citetitle{gantt1913work}, pointing out that
 ``\dots~where there is no union,
@@
 }
+
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+}
+
 \end{document}

### 2016-08-09 drafts/labor/trackingWork.tex
debugging a bunch of unnecessary stuff to get biblatex working...

@@
 \usepackage[normalem]{ulem}
-% \usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
-\usepackage{biblatex}
+% \usepackage{url}
+% \def\UrlOrds{\do\*\do\-\do\~\do\'\do\"\do\-}%
+% \usepackage{multibib}
+\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
+% \usepackage{biblatex}
+% \usepackage[square,numbers]{jnatbib}
 \usepackage[T1]{fontenc}
-\usepackage[pdftex,pdfpagelabels=false]{hyperref}
+\usepackage[pdflang={en-US},pdftex]{hyperref}
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
 \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+% \bibliographystyle{SIGCHI-Reference-Format}
 
@@
 \bibliography{references}
+% \addbibresource{references}
 
@@
   We argue that the literature surrounding ``piecework''
-  informs and even predicts both the contributions \cite{}
+  informs and even predicts both the contributions
   that have been made toward the development of on--demand labor and crowd work
@@
 
-For insight on this, we return to \citeauthor{levi2009union}'s study of labor unions,
+For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
 and identify that
@@
 \cite{levi2009union}.
-% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
+Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
 % unable to satisfy their needs or otherwise more invested in the 

### 2016-08-09 cleanrefs.py

@@ -1,0 +1,38 @@
+import re,itertools
+with open('references.bib') as references:
+  refs=references.read()
+
+refObjects=re.findall("(@(?:.*?)\})\n", refs,re.MULTILINE|re.DOTALL)
+print("Found {} references...".format(len(refObjects)))
+
+bibliography=list()
+longest_field=0
+
+fieldName=re.compile("^\s*(\S+)\s*=\s*.*")
+
+fields=list(map(lambda n:n.split("\n"),refObjects))
+# print(refObjects[-1])
+# quit()
+# print(fields[0])
+
+results=set()
+for obj in fields:
+  temp=set(itertools.chain(*filter(None,map(fieldName.findall,obj))))
+  results=results.union(temp)
+# print(results)
+longest_field=len(max(results,key=lambda n:len(n)))
+
+# print(longest_field)
+# quit()
+for ref in refObjects:
+# for line in refObjects[0].split("\n"):
+  for line in ref.split("\n"):
+    output=list()
+    result=re.findall("^\s*(\S+)\s*=\s*\{(.*)\}(?:,|\})",line)
+    output.append("  {field:{width}} = {{{entry}}},".format(field=result[0][0],width=longest_field,entry=result[0][1]) if result else line or "}")
+    bibliography.append(output[0])
+    # print("\n".join(output))
+
+print("\n".join(bibliography))
+# for n in bibliography:
+#   print(n)

### 2016-08-11 drafts/labor/cases.tex
wednesday edits

@@
 
+% lol oh my gosh i just discovered figlet and it's changing my life
+% it's like i'm distributing warez
+
 \section{Major Research Questions}
@@
 
+%  _ _           _ _                __ 
+% | (_)_ __ ___ (_) |_ ___    ___  / _|
+% | | | '_ ` _ \| | __/ __|  / _ \| |_ 
+% | | | | | | | | | |_\__ \ | (_) |  _|
+% |_|_|_| |_| |_|_|\__|___/  \___/|_|
+
+%                            _                          _             
+%   ___ _ __ _____      ____| |___  ___  _   _ _ __ ___(_)_ __   __ _ 
+%  / __| '__/ _ \ \ /\ / / _` / __|/ _ \| | | | '__/ __| | '_ \ / _` |
+% | (__| | | (_) \ V  V / (_| \__ \ (_) | |_| | | | (__| | | | | (_| |
+%  \___|_|  \___/ \_/\_/ \__,_|___/\___/ \__,_|_|  \___|_|_| |_|\__, |
+%                                                               |___/
+
 \subsection{What are the limits of crowdsourcing?}
-For more than a decade
-research in crowd work has explored ways of expanding the work that crowdsourcing can do by
+Research in crowdsourcing has spent the better part of a decade
+exploring ways of growing the limits of crowdsourcing and
+finding the boundaries of crowd work and microtasks;
   identifying challenges to this form of labor,
@@
 through the work that they have produced then
-has been whether there are limits to crowdsourcing
-(and, if so, what bounds them).
+has been driving at \textit{whether} there are limits to crowdsourcing
+(and, if so, what factors determine those limits).
+Through this lens, we can point to
+a number of contributions to the field that have extended the boundaries of crowd work.
 
-In the field of piecework,
-the research covering this topic has both explored
-a breadth of tasks that might be rendered doable by piecemeal workers
-\textit{as well as} longitudinally documented the success of these approaches.
-
-Points to make:
-  \begin{itemize}
-    \item Blossoming of piecework
-    \begin{itemize}
-      \item high point: consultants?
-      \item most cases: auto workers/etc\dots
-      \item worst case: sweatshops (especially in developing nations)
-    \end{itemize}
-    \item what led to these outcomes?
-    \begin{itemize}
-      \item ``consultant'' work came out well because the work was complex;
-      this made it difficult to turn into a mass market commodity.
-      We see consultants ranging skill levels like
-      oDesk (implementing modules) to
-      Accenture (on--demand teams of consultants).
-      \item auto workers,
-      working in settings where capital couldn't be moved as easily
-      found themselves in the same workspace as a direct
-      --- if multi--stepped ---
-      result to the benefits of putting people in the same place to consolidate resources.
-      Moreover,
-      workers had leverage over factory owners as a result of that consolidated capital;
-      operating equipment required training.
-      \item sweatshop workers fared the worst,
-      for reasons that may seem obvious with hindsight.
-      Source materials for textiles are easier to ship than mechanical components such as engines,
-      making it easier for factories to relocate to developing nations
-      (where cost--of--living, and consequently wages, would be lower).
-      As wages, CoLAs, and QoL rise, workers begin asking for
-      (and later demanding)
-      higher wages and better conditions.
-      But where the auto workers have leverage,
-      textile workers find only a precarious economic balance now tipped by their collective action,
-      spurring manufacturers to move to a new locale
-    \end{itemize}
-  \end{itemize}
-
-\subsection{What forms of work design and worker management are viable?}
-\begin{itemize}
-  \item researchers have looked at how to increase worker productivity
-  (e.g. finding the maximal speed at which gig workers can be expected to work before making errors)
-  \cite{measuringCrowdsourcingCheng}.
-  \item we've also seen people ``embrace error''
-  \cite{embracingErrorKrishna}.
-  \item still other research has looked into ways to sandbox workers from the context of their work 
-  \item but scholarship looking into the design and management of work and workers isn't new;
-  lots of research into getting pieceworkers to do work more quickly
-  \cite{seymour1954manual}.
-  \item Researchers have even asked the age old question of \textit{what motivates} pieceworkers
-  (echoing similar research on Wikipedia and Mechanical Turk)
-  \cite{roy1953work,Nov:2007:MW:1297797.1297798,kaufmann2011more}
-\end{itemize}
-
-\subsection{What will work and the place of work look like for workers?}
-The metaphorical mechanics of these dynamics are still at play;
-workers and managers continue to interact in adversarial manners,
-despite substantive work into aligning the motivations of workers and requesters
-
-\onlyinsubfile{
-\clearpage
-% \section{Case Studies}
-The existing body of research has shed light on on--demand labor from various perspectives,
-and revealed a number of topics that,
-through our framing, are clearly situated together.
-Those topics are, at a high level, as follows:
-\begin{enumerate}
-\item the \textbf{processes} involved in making work into tasks, or discretization;
-\item the outcomes (and indeed the \textbf{fallout}) of that discretization,
-both on the work itself as well as the workers;
-and finally
-\item the \textbf{relationships} between workers and requesters of the work
---- both \textit{cooperative} and \textit{adversarial} cases.
-\end{enumerate}
-
-\subsection{The Processes of Making Gig(s) Work}\label{sec:MakingGigsWork}
-The HCI community is perhaps most familiar with examples of task--based work such as
-99designs,
-Amazon Mechanical Turk (AMT),
-and increasingly
-Uber \& Lyft,
-which all allow requesters in various forms to tap into
-resources such as
-cars,
-computers, and above all
-``cognitive surplus''
-with relative ease
-\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
-This insight, that
-workers can be geographically distributed and tasks decomposed,
-% (especially given careful central management of those workers, discussed later),
-has proven remarkably compelling
-and an effective fulcrum for leveraging the Internet for highly scalable work
-\cite{sensitiveTasks,embracingErrorKrishna}.
-
-This section will largely discuss the processes at work that make distributed,
-digitally managed work both possible and indeed preferable for ``requesters''
-(in other words, the employers who solicit workers).
-This body of research spans a broad field
-within the CSCW and broader HCI community.
-In this context,
-we'll look at the body of research through the lens of
-highlighting contributions which expand what we
-(as ``requesters'' of work) can do by managing workers in novel ways.
-This work broadly consists of three areas:
+The exploration of crowdsourcing's potential and limits has principally looked at
+manipulating and extending along three dimensions:
 \begin{inlinelist}
@@
 \end{inlinelist}.
-As we explore this work, we'll
-attempt to relate the advances in the design of crowd work to
-the research contributions made in the research of assembly line manufacturing,
-Taylorism, and
-scientific management 
-during the \nth{20} century
-\cite{hu1961parallel}.
+% We will further discuss the ways that various contributions drive along these dimensions.
+As we'll discuss through these dimensions,
+exploring the extents to which work can be
+decomposed,
+contextually abstracted, and
+made more resilient to attrition,
+we'll also point to corresponding piecework literature addressing these aspects;
+finally, we'll discuss how these elements will serve
+to constrain the upper and lower bounds of crowdsourcing as it relates
+to the question of the furthest limits of crowdsourcing.
+
+\ali{At this point I'm trying to retain as much of the old content as possible,
+but if the next few sections seem like I'm trying too hard to be lazy,
+say so and I'll do a rewrite rather than refactoring.}
+
+%      _                                          _ _   _             
+%   __| | ___  ___ ___  _ __ ___  _ __   ___  ___(_) |_(_) ___  _ __  
+%  / _` |/ _ \/ __/ _ \| '_ ` _ \| '_ \ / _ \/ __| | __| |/ _ \| '_ \ 
+% | (_| |  __/ (_| (_) | | | | | | |_) | (_) \__ \ | |_| | (_) | | | |
+%  \__,_|\___|\___\___/|_| |_| |_| .__/ \___/|___/_|\__|_|\___/|_| |_|
+%                                |_|                                  
+% lol i can't believe this worked i'm going to do this with every section.
 
 \subsubsection{Decomposition}\label{sec:decomposition}
-Piecework may be thought of as vertically slicing work such that 
-each person is responsible for the whole task
---- making a whole garment, in this case.
+Scholarship describing and exploring
+the decomposition of tasks is perhaps the most established
+of the previously mentioned three areas
+to HCI researchers;
+\citeauthor{crowdForgeKittur} specifically drive at this goal by addressing the possibility of
+crowdsourcing complex work
+\cite{crowdForgeKittur}.
+\citeauthor{cheng2015break} found that microtasks
+--- though not necessarily \textit{faster} than ``macrotasks'' ---
+yield higher quality work,
+particularly when that work is susceptible to frequent interruptions
+\cite{cheng2015break}.
+
+Much of the research in the space of designing crowd work has
+sought to illustrate the potential to take highly creative or skilled work
+and generate high--quality results.
+Perhaps the most notable case study here can be found in
+\citeauthor{foundry}'s \textit{Foundry}, which employed
+``flash teams'' to achieve expert--level outcomes via thoughtful
+decomposition of work as ``modular tasks''
+\cite{foundry}.
+
+Research exploring the decomposition of work more generally
+--- that is, without the constraint of an employer/requester directing workers ---
+has since emerged as well.
+\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work,
+exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+While some of this work doesn't strictly fall under ``crowdsourcing'',
+the topicality of this broader body of work will become more apparent
+as we trace the parallels of crowdsourcing with piecework.
+
+The work described thus far has illustrated myriad ways that
+we can manage workers for the purpose of accomplishing complex tasks with
+rapidly sourced workers from across the world.
+Critically, the work's defining characteristic involves
+the use of the Internet as a medium both 
+to coordinate workers, as well as
+to do the work itself.
+But research into the ``decomposition'' of work more generally
+illustrates the same concepts of work that ``Taylorism''
+and scientific management sought to embody
+---
+\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
+\citeyear{professionalCrowdworkEthics}
+\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
+In both the historical and contemporary cases of decomposed work,
+work was,
+at least initially,
+distributed in the form of tasks to the homes of workers;
+\citeauthor{riisOtherSideLives} captured this in
+in \citeyear{riisOtherSideLives}
+\cite{riisOtherSideLives}.
+
+Following this thread of work, it may be said that
+much of the work in crowdsourcing has found ways of ``vertically slicing'' work
+such that each person is responsible for as small a task as can be made reasonable
+--- taking a minute of audio, for instance, and transcribing just that.
+The aforementioned research has found many novel ways to slice work,
+communicate different amounts of context of work from one worker to the next,
+etc\dots
+but fundamentally they follow similar patterns.
+It's through this lens that we see the echoes of piecework.
+
+%                        _ _      _ 
+%  _ __   __ _ _ __ __ _| | | ___| |
+% | '_ \ / _` | '__/ _` | | |/ _ \ |
+% | |_) | (_| | | | (_| | | |  __/ |
+% | .__/ \__,_|_|  \__,_|_|_|\___|_|
+% |_|
+
+Piecework Piecework Piecework Piecework Piecework Piecework Piecework
+
+% Piecework's growing popularity was best documented at the turn of the \nth{20}
+% by \citeauthor{riisOtherSideLives} % in his work on \citetitle{riisOtherSideLives}
+% \cite{riisOtherSideLives}.
+
 Broken down in this way, work could grow to unprecedented scales,
@@
 
-Scholarship describing and exploring
-the decomposition of tasks is perhaps the most established of the above areas among HCI researchers;
-\citeauthor{crowdForgeKittur} specifically drive at this goal by addressing the possibility of
-``crowdsourcing complex work''
-\cite{crowdForgeKittur}.
-\citeauthor{cheng2015break} found that microtasks
---- though not necessarily \textit{faster} than ``macrotasks'' ---
-yield higher quality work,
-particularly when that work might be readily interrupted
-\cite{cheng2015break}.
-\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work
-by exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-While this work doesn't strictly fall under ``crowdsourcing'',
-the major contributions here
-seem uncontroversially to be inspired by the design of crowd work.
-
-Much of the research in the space of designing crowd work has
-sought to illustrate the potential to take highly creative or skilled work
-and generate high--quality results.
-Perhaps the most notable case study here can be found in
-\citeauthor{foundry}'s \textit{Foundry}, which employed
-``flash teams'' to achieve expert--level outcomes via thoughtful
-decomposition of work as ``modular tasks''
-\cite{foundry}.
-
-Work decomposition, then, is far from new;
-``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
-and scientific management sought to embody
----
-\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
-\citeyear{professionalCrowdworkEthics}
-\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
-In both the historical and contemporary cases of decomposed work,
-work was,
-at least initially,
-distributed in the form of tasks to the homes of workers;
-\citeauthor{riisOtherSideLives} captured this in his documentary work
-% \citetitle{riisOtherSideLives}
-in \citeyear{riisOtherSideLives}
-\cite{riisOtherSideLives}.
+%                     _            _         _                  _   _             
+% __      _____  _ __| | __   __ _| |__  ___| |_ _ __ __ _  ___| |_(_) ___  _ __  
+% \ \ /\ / / _ \| '__| |/ /  / _` | '_ \/ __| __| '__/ _` |/ __| __| |/ _ \| '_ \ 
+%  \ V  V / (_) | |  |   <  | (_| | |_) \__ \ |_| | | (_| | (__| |_| | (_) | | | |
+%   \_/\_/ \___/|_|  |_|\_\  \__,_|_.__/|___/\__|_|  \__,_|\___|\__|_|\___/|_| |_|
+% lollllllllllllllll
 
@@
 \cite{tolliday1986between}.
+
+In the field of piecework,
+the research covering this topic has both explored
+a breadth of tasks that might be rendered doable by piecemeal workers
+\textit{as well as} longitudinally documented the success of these approaches.
+Here, we
+
+Points to make:
+  \begin{itemize}
+    \item Blossoming of piecework
+    \begin{itemize}
+      \item high point: consultants?
+      \item most cases: auto workers/etc\dots
+      \item worst case: sweatshops (especially in developing nations)
+    \end{itemize}
+    \item what led to these outcomes?
+    \begin{itemize}
+      \item ``consultant'' work came out well because the work was complex;
+      this made it difficult to turn into a mass market commodity.
+      We see consultants ranging skill levels like
+      oDesk (implementing modules) to
+      Accenture (on--demand teams of consultants).
+      \item auto workers,
+      working in settings where capital couldn't be moved as easily
+      found themselves in the same workspace as a direct
+      --- if multi--stepped ---
+      result to the benefits of putting people in the same place to consolidate resources.
+      Moreover,
+      workers had leverage over factory owners as a result of that consolidated capital;
+      operating equipment required training.
+      \item sweatshop workers fared the worst,
+      for reasons that may seem obvious with hindsight.
+      Source materials for textiles are easier to ship than mechanical components such as engines,
+      making it easier for factories to relocate to developing nations
+      (where cost--of--living, and consequently wages, would be lower).
+      As wages, CoLAs, and QoL rise, workers begin asking for
+      (and later demanding)
+      higher wages and better conditions.
+      But where the auto workers have leverage,
+      textile workers find only a precarious economic balance now tipped by their collective action,
+      spurring manufacturers to move to a new locale
+    \end{itemize}
+  \end{itemize}
+
+\subsection{What forms of work design and worker management are viable?}
+\begin{itemize}
+  \item researchers have looked at how to increase worker productivity
+  (e.g. finding the maximal speed at which gig workers can be expected to work before making errors)
+  \cite{measuringCrowdsourcingCheng}.
+  \item we've also seen people ``embrace error''
+  \cite{embracingErrorKrishna}.
+  \item still other research has looked into ways to sandbox workers from the context of their work 
+  \item but scholarship looking into the design and management of work and workers isn't new;
+  lots of research into getting pieceworkers to do work more quickly
+  \cite{seymour1954manual}.
+  \item Researchers have even asked the age old question of \textit{what motivates} pieceworkers
+  (echoing similar research on Wikipedia and Mechanical Turk)
+  \cite{roy1953work,Nov:2007:MW:1297797.1297798,kaufmann2011more}
+\end{itemize}
+
+\subsection{What will work and the place of work look like for workers?}
+The metaphorical mechanics of these dynamics are still at play;
+workers and managers continue to interact in adversarial manners,
+despite substantive work into aligning the motivations of workers and requesters
+
+\onlyinsubfile{
+\clearpage
+% \section{Case Studies}
+The existing body of research has shed light on on--demand labor from various perspectives,
+and revealed a number of topics that,
+through our framing, are clearly situated together.
+Those topics are, at a high level, as follows:
+\begin{enumerate}
+\item the \textbf{processes} involved in making work into tasks, or discretization;
+\item the outcomes (and indeed the \textbf{fallout}) of that discretization,
+both on the work itself as well as the workers;
+and finally
+\item the \textbf{relationships} between workers and requesters of the work
+--- both \textit{cooperative} and \textit{adversarial} cases.
+\end{enumerate}
 

### 2016-08-11 drafts/labor/cases.tex
notes from meeting

@@
 % We will further discuss the ways that various contributions drive along these dimensions.
-As we'll discuss through these dimensions,
+As we'll discuss these dimensions,
 exploring the extents to which work can be
@@
 contextually abstracted, and
-made more resilient to attrition,
-we'll also point to corresponding piecework literature addressing these aspects;
-finally, we'll discuss how these elements will serve
+made more resilient to attrition of various forms,
+we'll also point to corresponding piecework literature addressing these aspects.
+Finally, we'll discuss how these elements will serve
 to constrain the upper and lower bounds of crowdsourcing as it relates
@@
 %                                |_|                                  
-% lol i can't believe this worked i'm going to do this with every section.
+% lol i can't believe this worked i'm going to do this with every (sub)*section.
 
@@
 to HCI researchers;
-\citeauthor{crowdForgeKittur} specifically drive at this goal by addressing the possibility of
+\citeauthor{crowdForgeKittur} specifically
+drive at this goal by addressing the possibility of
 crowdsourcing complex work
@@
 
-Piecework Piecework Piecework Piecework Piecework Piecework Piecework
+Piecework has seen work along this dimension spanning decades;
+\citeauthor{thompson1913time} investigate some of the ways that
+construction can benefit from the principles of scientific management.
+\citeauthor{thompson1913time}'s thesis asserts that
+task work is predicated on the accurate scientific management of work,
+including the ``miscellaneous tasks''.
+\citeauthor{thompson1913time} argues
+--- as early as \citeyear{thompson1913time} ---
+that ``\dots one may be challenged to find any class of work
+involving labor either indoors or out--of--doors
+where tasks cannot be fixed by proper time--study''
+\cite{thompson1913time}.
 
@@
 \cite{rogstadius2011assessment}.
+
+\subsubsection{So how does this affect crowdwork?}
+The work we've seen so far 
+\begin{itemize}
+  \item worst case: assembling iPhones (extant)
+  \item average case: railroad workers and assembly lines
+  \item high (complexity) case: 
+\end{itemize}
+
+% what can be decomposed? what can be evaluated?
+% technology/turuing completeness gets us a win
+% predefined expertise? you can find someone almost instantly;
+% ease of doing this makes it qualitatively different due to Channel Factors (diverting a channel/river)
+% huddler paper cites some decomposition stuff workflows are good/bad at things
+% 3-way slack w/msb and mav
+% academy of management 
+% van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
 

### 2016-08-11 drafts/labor/introduction.tex
notes from meeting

@@
 \makeatother
+
+% \usepackage[style=alphabetic]{biblatex}
 \begin{document}
@@
   % (what are the \textbf{complexity limits}) \note{(hospitals at high level?) -- team scaffolds} \note{why is there a high limit? is that different with crowds?}
-    \cite[for example, see][]{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
+    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
   \item What forms of work design, and worker management and arrangement, are viable?
   % \note{time in motion studies; job design; } or is this more about managers, coops, etc...
-    \cite[for example, see][]{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
+    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
   \item What will work and the place of work look like for the workers?
-    \cite[for example, see][]{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}
+    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}
     % \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
@@
 gone as far as predict future developments.
+
+% so CW?
+% mechanism/limitation in pw?
+% differences in cw? tech/access to intelligence?
+% predicted outcome then? -> decompose effort/expertise access
+% possible quirks/diffs?
 
@@
 in
-\citeyear{crowdworkFuture}
+\citeauthor{crowdworkFuture}
 referenced crowd work as ``piecework'' briefly
@@
   \balance{}
+  % \bibliography{references}
   \printbibliography

### 2016-08-11 drafts/labor/introduction.tex
notes were in the wrong file :x
various additional notes and a fix with some colors and todonotes or something

@@
 
-% \usepackage[style=alphabetic]{biblatex}
 \begin{document}
-
-\newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
 
@@
 
-% Broadly, this work has imagined the various applications of ``gig''--based work,
-% and making it work for ``requesters''
-% --- academics, industries, and to some extent end--users
-% who might benefit from the product of crowd--powered work
-% \cite{hong2015group,jonBrelig,paolacci2010running}.
-% % ??? -->
-% Since then,
-% the application of digitally mediated environments acting as rapid ``hiring halls''
-% for on--demand tasks
-% has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
-% livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
-% % parentheticals giving examples?
-% have garnered the interest of both industry and research.
-
 Researchers have made efforts to understand the people
@@
 
-% Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
-% workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
-% and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.
-
-% \note{this probably doesn't belong here but i need to put it somewhere!}
-
-% From a wide range of academic and industrial perspectives
-% and increasingly identifying the many stakeholders in digitally mediated work,
 The extant body of work has ostensibly sought to answer one underlying question:
@@
 Researchers have offered their input on this open question along three major threads of scholarship:
-% the driving motivation behind much of the research in this space has led us to ask,
-% at a high level, three important questions:
 \begin{enumerate}
@@
   what can and cannot be done by crowd workers?
-  % (what are the \textbf{complexity limits}) \note{(hospitals at high level?) -- team scaffolds} \note{why is there a high limit? is that different with crowds?}
     \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
   \item What forms of work design, and worker management and arrangement, are viable?
-  % \note{time in motion studies; job design; } or is this more about managers, coops, etc...
     \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
@@
     \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}
-    % \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
 \end{enumerate}
@@
 
-% so CW?
-% mechanism/limitation in pw?
-% differences in cw? tech/access to intelligence?
-% predicted outcome then? -> decompose effort/expertise access
-% possible quirks/diffs?
-
-% These bodies of literature have conversed,
-% and indeed begun to identify that their theories are interrelated in some way,
-% but thus far a convincing model for making sense of
-% --- let alone predicting ---
-% the phenomena in on--demand labor has proven elusive.
-% For instance, 
-% making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
-% is understandable in its own context, but
-% the relationship between the work itself (``turking'')
-% and the myriad social outcomes has only briefly been explored
-% \cite{crowdcollab}.
-
 We offer a framing for crowd work spanning the aforementioned industries
@@
 Indeed,
-\citeauthor{crowdworkFuture}
-in
-\citeauthor{crowdworkFuture}
+\citeauthor{crowdworkFuture} in \citeauthor{crowdworkFuture}
 referenced crowd work as ``piecework'' briefly
 as a loose analogy to the form of work emerging at the time
-% but not given as much weight as it perhaps could have been
 \cite{crowdworkFuture}.
-% ``Piecework'' here makes use of
-% a term historically used to describe work done in the home,
-% in manageable tasks,
-% often involving clear instructions
-% and payment only for work completed, not work done
-% (the differentiation then being that
-% one would be paid for the \textit{output} of the work,
-% not the \textit{duration}).
-
-% This framing prompts us to note several immediately apparent similarities:
-% \begin{inlinelist}
-%   \item this form of work began in the home
-%   \item the worker is paid for each discrete piece of work done, regardless of time or effort; and
-%   \item the worker's status
-%   (not only socially, but also economically)
-%   is ambiguous, or at least the subject of some controversy.
-% \end{inlinelist}
 
@@
 Finally, we will offer design implications based on this research\dots
-% \note{???}
 
-\note{
+\endnote{
 \textbf{notes/what i want to get across\dots}
@@
   \balance{}
-  % \bibliography{references}
   \printbibliography
+  \theendnotes
 }

### 2016-08-14 drafts/labor/trackingWork.tex
various additional notes and a fix with some colors and todonotes or something
moving things around

@@
-\documentclass{sigchi}
+\documentclass[chi_draft]{sigchi}
 
-\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth}
+\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes}
 \usepackage[normalem]{ulem}
@@
 \newcommand{\notinsubfile}[1]{}
+\newcommand{\note}[1]{\footnote{#1}}
 \begin{document}
@@
 \renewcommand{\notinsubfile}[1]{#1}
+\renewcommand{\note}[1]{\onlyinsubfile{\note{#1}}}
 \title{\plaintitle}
@@
 \maketitle
-
 \begin{abstract}

### 2016-08-16 drafts/labor/cases.tex
notes were in the wrong file :x
various additional notes and a fix with some colors and todonotes or something
mondays

@@
 \begin{document}
-% \newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
-
-% lol oh my gosh i just discovered figlet and it's changing my life
-% it's like i'm distributing warez
 
 \section{Major Research Questions}
-\ali{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
+% \ali{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
@@
   identifying challenges to this form of labor,
-  overcoming them through novel designs of work--flows and processes,
-and repeating the process
+  overcoming them through novel designs of work--flows and processes, and
+  repeating the process
 \cite{bernsteinSoylent}.
@@
 \end{inlinelist}.
-% We will further discuss the ways that various contributions drive along these dimensions.
-As we'll discuss these dimensions,
-exploring the extents to which work can be
+We'll explore these aspects of crowdsourcing,
+% As we'll discuss these dimensions,
+discussing the extents to which work can be
 decomposed,
 contextually abstracted, and
-made more resilient to attrition of various forms,
-we'll also point to corresponding piecework literature addressing these aspects.
+made more resilient to attrition of various forms.
+We'll also point to corresponding piecework literature addressing these aspects.
 Finally, we'll discuss how these elements will serve
@@
 to the question of the furthest limits of crowdsourcing.
-
-\ali{At this point I'm trying to retain as much of the old content as possible,
-but if the next few sections seem like I'm trying too hard to be lazy,
-say so and I'll do a rewrite rather than refactoring.}
 
@@
 %                                |_|                                  
-% lol i can't believe this worked i'm going to do this with every (sub)*section.
+%
 
-\subsubsection{Decomposition}\label{sec:decomposition}
-Scholarship describing and exploring
-the decomposition of tasks is perhaps the most established
-of the previously mentioned three areas
-to HCI researchers;
+% stronger topic sentences
+% boosterist
+
+\subsubsection{Decomposition (how high can we stack the salami?)}\label{sec:decomposition}
+Crowdsourcing has explored the decomposition of work for the better part of a decade;
 \citeauthor{crowdForgeKittur} specifically
@@
 \citeauthor{foundry}'s \textit{Foundry}, which employed
-``flash teams'' to achieve expert--level outcomes via thoughtful
-decomposition of work as ``modular tasks''
+``flash teams'' to accomplish work typically thought of as only within reach of experts
 \cite{foundry}.
@@
 While some of this work doesn't strictly fall under ``crowdsourcing'',
-the topicality of this broader body of work will become more apparent
-as we trace the parallels of crowdsourcing with piecework.
+the [scientific] management of the self as a worker
+--- of sorts ---
+will prove relevant as we trace the literature surrounding piecework.
 
-The work described thus far has illustrated myriad ways that
+The work described thus far has explored along myriad paths the ways that
 we can manage workers for the purpose of accomplishing complex tasks with
-rapidly sourced workers from across the world.
+rapidly sourced workers from all over the world.
 Critically, the work's defining characteristic involves
 the use of the Internet as a medium both 
-to coordinate workers, as well as
-to do the work itself.
+to coordinate workers, as well as to do the work itself.
 But research into the ``decomposition'' of work more generally
@@
 ---
-\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
-\citeyear{professionalCrowdworkEthics}
+\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger
+and warned of it in, but they were by no means alone in their warnings
 \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
-In both the historical and contemporary cases of decomposed work,
-work was,
-at least initially,
-distributed in the form of tasks to the homes of workers;
-\citeauthor{riisOtherSideLives} captured this in
-in \citeyear{riisOtherSideLives}
-\cite{riisOtherSideLives}.
 
@@
 such that each person is responsible for as small a task as can be made reasonable
---- taking a minute of audio, for instance, and transcribing just that.
-The aforementioned research has found many novel ways to slice work,
-communicate different amounts of context of work from one worker to the next,
-etc\dots
-but fundamentally they follow similar patterns.
-It's through this lens that we see the echoes of piecework.
+--- taking a minute of audio, for instance, and transcribing that.
+The findings we take away from the crowdsourcing literature might be summed up as
+a number of ways
+to break tasks down \cite[see][]{selfsourcingTeevan2014,foundry,crowdForgeKittur} and
+measuring the costs as they relate to work completion \&
+to a lesser extent to workers
+\cite[see][]{measuringCrowdsourcingCheng,cheng2015break,crowdworkFuture,professionalCrowdworkEthics}.
+
+Research into piecework has considered the decomposition of work;
+\citeauthor{10.2307/23702539} describes piecework in \citeyear{10.2307/23702539} as
+``\dots~based on examination of various shop jobs,
+which included calculation of the standard time and compensation for each task.
+Most piecework plans required additional clerical and supervisory people
+(such as piecework clerks, inspectors, and ``experts'')\dots'',
+allowing for what the \textit{American Machinist} would later call ``paper efficiency'',
+describing savings on paper that are in fact passed on to workers
+\cite{10.2307/23702539}.
+In this context,
+the introspection of tasks by experts and the careful work assignment \& management of workers
+we see in historical piecework seem to parallel crowdsourcing.
+
+% \citeauthor{Brown01041990} draws more decisive boundaries on piece rate approaches to remuneration,
+identifying in analysis of data from
+the U.S. Bureau of Labor and Statistics that
+``larger establishment are less likely to use piece rates and predominantly female establishments more likely to do so; unionized establishments make significantly greater use of standard--rate pay than do nonunion establishments; \textbf{incentive pay is less likely in jobs with a variety of duties than in jobs with a narrow set of routinized duties}; and the ease of monitoring work quality is correlated with the use of incentive pay''
+% \cite{Brown01041990}.
+
+What's different today? We can turn to \citeauthor{10.2307/23702539}'s reflection on railroad piecework
+for a hint of what might have enabled the future of crowdwork in which we now live:
+``\dots~only the largest and most wealthy railroads stood to gain substantially from scientific management,
+because only they had the resources to eliminate union resistance and
+\textit{pay the overhead involved in installing work reorganization}''
+(emphasis added)
+\cite{10.2307/23702539}.
+
+We argue that crowdsourcing is more widespread today,
+and has the potential to become even larger in scale,
+because the underlying architecture of crowdwork is digital;
+reformulating workflows to decompose tasks and institute piecework compensation programs
+is easier today than it was for railroad companies which had to move physical capital to do so.
+
+\ali{\citeauthor{10.2307/23702539} later argues that a significant obstacle to the introduction of piecework in railroad shops was the resistance of workers
+(the other being the resistance of management (for different and varying reasons)).
+In the cases of online platforms like AMT and digitally mediated platforms like Uber,
+where in both cases workers rarely if ever interact face--to--face and opportunities for coordination and collective action are severely limited,
+we should expect to see the challenges stymieing collective action that we see in \citeauthor{dynamo} and elsewhere
+\cite{10.2307/23702539,dynamo}.
+}
+
+%                     _            _         _                  _   _             
+% __      _____  _ __| | __   __ _| |__  ___| |_ _ __ __ _  ___| |_(_) ___  _ __  
+% \ \ /\ / / _ \| '__| |/ /  / _` | '_ \/ __| __| '__/ _` |/ __| __| |/ _ \| '_ \ 
+%  \ V  V / (_) | |  |   <  | (_| | |_) \__ \ |_| | | (_| | (__| |_| | (_) | | | |
+%   \_/\_/ \___/|_|  |_|\_\  \__,_|_.__/|___/\__|_|  \__,_|\___|\__|_|\___/|_| |_|
+%
+
+\subsubsection{Work Abstraction}\label{sec:workAbstraction}
+Decomposition allows requesters to assign parallelized tasks
+without undue concern for the broader context of the work.
+The earlier example of transcribed audio is admittedly a carefully curated one
+--- used to illustrate the ability
+to parallelize work across as many workers as there are tasks ---
+but even in this case problems arise;
+breaking an hour--long interview into
+3,600 1--second--long tasks would seem to make it impossible to do the tasks
+for lack of \textit{context},
+but the other extreme (one hour--long block)
+offers no performance improvement over conventionally sourced work.
+
+The question, then, becomes about how to break tasks down as much as possible
+(as discussed in the previous section)
+without losing the necessary context to do a task.
+Stated another way, how can we design work so that as little context as possible is provided,
+without leaving out context or information necessary to complete the task?
+As \citeauthor{verroios2014context} frame it,
+``the crowd must be able to act with
+global understanding when each contributor only has access to local views''
+\cite{verroios2014context}.
+
+While some research has attempted to increase workers' awareness of the work they do in
+apparent efforts to yield higher quality work,
+others have turned this constraint of crowdwork into a feature ---
+the abstraction from the work itself
+appears to allow requesters to engage workers in work with ---
+for instance ---
+sensitive, confidential or otherwise personally identifying information.
+We will discuss these and other avenues of research here.
+
+\textit{Cascade} demonstrated that it's possible to
+break certain classes of tasks apart
+in such a way that they yield taxonomies of various subjects,
+a task generally thought to be sufficiently complex that only expert workers
+with top--down awareness of the task
+--- in this case, with awareness of all the constituent colors ---
+could complete the task
+\cite{chilton2013cascade}.
+\citeauthor{verroios2014context} further illustrate this potential by
+forming a task one might consider highly contextually dependent
+--- summarizing the contents of a movie ---
+in such a way that crowd workers could contribute small pieces of work without
+needing to know the content of the rest of the project
+\cite{verroios2014context}.
+
+%   __ _           _ _     _ _ _ _         
+%  / _| | _____  _(_) |__ (_) (_) |_ _   _ 
+% | |_| |/ _ \ \/ / | '_ \| | | | __| | | |
+% |  _| |  __/>  <| | |_) | | | | |_| |_| |
+% |_| |_|\___/_/\_\_|_.__/|_|_|_|\__|\__, |
+%                                    |___/
+
+% how much volatility and quality can we deal with?
+% how to measure, get rid of people, recover from bad workers
+%mason's on getting workers to be better
+\subsubsection{Volatile quality and availability? How much can you deal with?}\label{sec:flexibility}
+A number of researchers have identified
+worker attrition,
+variability of worker performance,
+and uncertainty about good versus bad--faith actors as open questions of crowdwork
+\cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
+\ali{We can and should discuss the distinction between presumably ``bad faith'' workers
+\& workers who are merely responding in kind to bad requesters
+--- and the broader questions surrounding
+the roles that requesters as well as workers should play ---
+but let it suffice to say that
+requesters have been trying to understand and manage what appears to them as inconsistent work.
+Their ways of responding to that variance in work quality has largely involved
+making the work more flexible and resilient to work
+(although some work has gone into investigating the causes, rather than treating the symptoms)}
+
+Earlier we discussed \citeauthor{cheng2015break}'s work
+measuring the impact that interruption has on worker performance
+\cite{cheng2015break}.
+This work illustrates a broader sentiment in
+both the study and practice
+of crowdwork, that microtasks should be designed resiliently against the variability of workers,
+fully exploiting the abstracted nature of each piece of work
+\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
+That is to say, micro--tasks should be designed such that a single worker's poor performance,
+or a good worker's sudden departure,
+does not significantly impact the agenda of the work as a whole.
+While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
+in the form of higher cumulative time to complete
+(albeit much shorter real time to complete, owing to parallelization),
+\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+similar tasks together
+\cite[respectively]{cheng2015break,delayAndOrderLasecki}.
+
+\citeauthor{embracingErrorKrishna} take a different approach;
+by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds,
+the authors offer orders--of--magnitude improvements
+in various binary classification tasks
+% on the principle that diverse workers complete these tasks
+% in order to accurately inform the model on the variety of delays in response times.
+\cite{embracingErrorKrishna}.
+And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
+some researchers have designed work predicated on the expectation of it instead:
+\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
+crowd workers would never be given enough information to piece together sensitive information about
+any single topic
+\cite{sensitiveTasks}.
+
+The work thus far seems to attempt to maximize the quality of work among workers through various means:
+\begin{inlinelist}
+  \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju},
+  \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
+  \item Expecting certain levels of attrition and incorrectness and using that variability to their advantage \cite{embracingErrorKrishna}
+\end{inlinelist}.
+
+% \subsubsection{Piecework's role here}
+% We address the question of the limits of crowdsourcing in three parts:
+% \begin{inlinelist}
+% \item decomposition
+% \item work abstraction
+% and \item flexibility
+% \end{inlinelist}.
+% The literature on piecework has offered insights into each of these topics, which we will explore now.
+
+Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
+\citeauthor{tolliday1986between}'s treatment describing
+turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
+Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
+[and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
+piece rates~\dots~rather than a day wage system''
+\cite{tolliday1986between}.
+
+In the field of piecework,
+the research covering this topic has both explored
+a breadth of tasks that might be rendered doable by piecemeal workers
+\textit{as well as} longitudinally documented the success of these approaches.
+Here, we
+
+Here, \citeauthor{hu1961parallel}'s work,
+saying of assembly line work that
+``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
+parallels the approach that dominated early research into crowd work
+--- namely, using non--expert crowds for complex work
+\cite{hu1961parallel}.
+This mindset in \citeauthor{hu1961parallel}'s analysis,
+and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
+substantively owes its existence to scientific management
+and the rigorous decomposition of work into tasks, discussed earlier,
+and persists to this day as it colors
+researchers' goals and objectives in the study and design of crowd work.
+
+Piecework's influence on the abstraction of work into tasks,
+described above, is more than just caused by the decomposition of work;
+work abstraction itself makes it possible for workers to come and go flexibly,
+prompting work requesters to consider ways to design these now discrete tasks in ways that
+maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
+\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
+It's to this area that we now turn our attention.
 
@@
 \cite{thompson1913time}.
-
-% Piecework's growing popularity was best documented at the turn of the \nth{20}
-% by \citeauthor{riisOtherSideLives} % in his work on \citetitle{riisOtherSideLives}
-% \cite{riisOtherSideLives}.
 
@@
 
+% so CW?
+% mechanism/limitation in pw?
+% differences in cw? tech/access to intelligence?
+% predicted outcome then? -> decompose effort/expertise access
+% possible quirks/diffs?
+
 % what can be decomposed? what can be evaluated?
@@
 
-%                     _            _         _                  _   _             
-% __      _____  _ __| | __   __ _| |__  ___| |_ _ __ __ _  ___| |_(_) ___  _ __  
-% \ \ /\ / / _ \| '__| |/ /  / _` | '_ \/ __| __| '__/ _` |/ __| __| |/ _ \| '_ \ 
-%  \ V  V / (_) | |  |   <  | (_| | |_) \__ \ |_| | | (_| | (__| |_| | (_) | | | |
-%   \_/\_/ \___/|_|  |_|\_\  \__,_|_.__/|___/\__|_|  \__,_|\___|\__|_|\___/|_| |_|
-% lollllllllllllllll
-
-\subsubsection{Work Abstraction}\label{sec:workAbstraction}
-Decomposition allows requesters to assign tasks without concern for the broader context.
-While we'll discuss this aspect of crowd work more critically later,
-it's worth pointing out that discrete blocks of work containing all the relevant context for a worker
-allows workers to engage with virtually any component of work without worrying that their lack of 
-higher--level awareness of the goals of the requester might negatively affect their work.
-
-\citeauthor{chilton2013cascade} perhaps best illustrated this with
-\textit{Cascade} by demonstrating that it's possible to
-break certain classes of tasks apart
-in such a way that they yield taxonomies of various subjects,
-a task previously thought to be safely within the domain of expert workers
-with top--down awareness of the context of the work as a whole
-\cite{chilton2013cascade}.
-\citeauthor{verroios2014context} further illustrate this potential by
-forming a task one might consider highly contextually dependent
---- summarizing the contents of a movie ---
-in such a way that crowd workers could contribute small pieces of work without
-needing to know the content of the rest of the project
-\cite{verroios2014context}.
-
-Here, \citeauthor{hu1961parallel}'s work,
-saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
-parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work
-\cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,
-and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
-substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,
-and persists to this day as it colors
-researchers' goals and objectives in the study and design of crowd work.
-
-Piecework's influence on the abstraction of work into tasks,
-described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,
-prompting work requesters to consider ways to design these now discrete tasks in ways that
-maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
-\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
-It's to this area that we now turn our attention.
-
-\subsubsection{Flexibility}\label{sec:flexibility}
-% \msb{nothing in this subsubsection connects to the piecework literature. bring it back!}
-
-Earlier we discussed \citeauthor{cheng2015break}'s work
-measuring the impact that interruption has on worker performance.
-This work both points to and embodies a broader sentiment in
-both the study and practice
-of crowd work that microtasks should be designed resiliently against the variability of workers,
-% \msb{not sure what flexibly means here, and the rest of the sentence isn't helping me unpack it. flexibility for requester? worker? flexible algorthms?},
-fully exploiting the abstracted nature of each piece of work
-\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-That is to say, micro--tasks should be designed such that a single worker's poor performance,
-or a good worker's sudden departure,
-would not significantly impact the agenda of the work as a whole.
-While \citeauthor{cheng2015break} identified costs with breaking tasks into smaller components
-in the form of higher cumulative time to complete
-(albeit much shorter real time to complete, owing to parallelization),
-\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
-similar tasks together.
-
-Given the importance of consistent work results, one might intuit that
-requesters would prefer high--quality workers who can be relied upon to be available
-(even for contextually independent tasks),
-which would appear to contradict the benefits of flexibility already discussed;
-requesters have thus made significant headway toward
-``embracing error'' to allow requesters to maximize the benefits of a flexible,
-even transient,
-workforce.
-
-\citeauthor{embracingErrorKrishna} offer orders--of--magnitude improvements
-in various binary classification tasks
-on the principle that diverse workers complete these tasks
-in order to accurately inform the model on the variety of delays in response times.
-And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-some researchers have designed work predicated on the expectation of this phenomenon:
-\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
-crowd workers would never be given enough information to piece together sensitive information about
-any single topic
-\cite{sensitiveTasks}.
-
-Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
-\citeauthor{tolliday1986between}'s treatment describing
-turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
-Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
-[and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
-piece rates~\dots~rather than a day wage system''
-\cite{tolliday1986between}.
-
-In the field of piecework,
-the research covering this topic has both explored
-a breadth of tasks that might be rendered doable by piecemeal workers
-\textit{as well as} longitudinally documented the success of these approaches.
-Here, we
-
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+  % \nobalance{}
+  \clearpage
+  \nobalance{}
+  \begin{appendices}
   Points to make:
@@
 
-\onlyinsubfile{
-\clearpage
     % \section{Case Studies}
@@
     \end{enumerate}
-
-% \msb{I still dont' have a clear conceptual picture of what flexibility is, and what about it sets this section apart. Ranjay's stuff is about speed...? I'm confused.}
 
@@
     are in fact one.
-}
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
+  \end{appendices}
   }

### 2016-08-16 drafts/labor/cases.tex
small notes

@@
 
+% let's do topic sentences!
 \subsection{What are the limits of crowdsourcing?}
 Research in crowdsourcing has spent the better part of a decade
-exploring ways of growing the limits of crowdsourcing and
-finding the boundaries of crowd work and microtasks;
+exploring how to grow the limits of crowdsourcing and
+find the boundaries of crowd work and microtasks;
+this has largely involved
   identifying challenges to this form of labor,
@@
 \begin{inlinelist}
-  \item \nameref{sec:decomposition},
+  % \item \cref*{sec:decomposition},
+  \item \lowercase{\nameref{sec:decomposition}},
   \item \nameref{sec:workAbstraction}, and
@@
 
-\subsubsection{Decomposition (how high can we stack the salami?)}\label{sec:decomposition}
+\subsubsection{Achieving complexity}\label{sec:decomposition}
 Crowdsourcing has explored the decomposition of work for the better part of a decade;

### 2016-08-17 drafts/labor/cases.tex
more topic sentences in cases.tex - see bold in cases.pdf

@@
 \begin{document}
-
 \section{Major Research Questions}
-% \ali{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
+% \al¬i{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
@@
 \begin{inlinelist}
-  % \item \cref*{sec:decomposition},
-  \item \lowercase{\nameref{sec:decomposition}},
-  \item \nameref{sec:workAbstraction}, and
+  % \item \cref*{sec:Complexity},
+  \item \nameref{sec:Complexity},
+  \item \nameref{sec:Slicing}, and
   \item \nameref{sec:flexibility}
@@
 We'll explore these aspects of crowdsourcing,
-% As we'll discuss these dimensions,
 discussing the extents to which work can be
@@
 %
+% stronger topic sentences
+% boosterist (???)
 
-% stronger topic sentences
-% boosterist
-
-\subsubsection{Achieving complexity}\label{sec:decomposition}
-Crowdsourcing has explored the decomposition of work for the better part of a decade;
-\citeauthor{crowdForgeKittur} specifically
-drive at this goal by addressing the possibility of
+\subsubsection{Achieving greater complexity}\label{sec:Complexity}
+\topic{Crowdsourcing research has spent the better part of a decade attempting
+to prove the viability of crowdsourcing in increasingly complex work.}  % this is my topic sentence.
+\citeauthor{crowdForgeKittur}
+map the discussion toward this goal in their work on
 crowdsourcing complex work
 \cite{crowdForgeKittur}.
-\citeauthor{cheng2015break} found that microtasks
---- though not necessarily \textit{faster} than ``macrotasks'' ---
-yield higher quality work,
-particularly when that work is susceptible to frequent interruptions
-\cite{cheng2015break}.
+\citeauthor{Lasecki:2013:CCC:2501988.2502057} tackle an area of computationally assisted interaction
+--- conversational artificial intelligence ---
+and find ways to arrange workers and formulate a ``crowd--powered conversational assistant''
+\cite{Lasecki:2013:CCC:2501988.2502057}.
+Other work has followed a similar pattern,
+identifying target milestones in Computer Science
+that have presented significant challenges for researchers,
+and demonstrating that ``crowd--powered'' systems can tackle many of these challenges
+with generally greater success than purely computational attempts
+\cite[][and others]{Lasecki:2013:RCL:2441776.2441912,foundry}.
 
-Much of the research in the space of designing crowd work has
-sought to illustrate the potential to take highly creative or skilled work
-and generate high--quality results.
-Perhaps the most notable case study here can be found in
-\citeauthor{foundry}'s \textit{Foundry}, which employed
-``flash teams'' to accomplish work typically thought of as only within reach of experts
-\cite{foundry}.
+% NEEDS TO GO ELSEWHERE
+% \citeauthor{cheng2015break} found that microtasks
+% --- though not necessarily \textit{faster} than ``macrotasks'' ---
+% yield higher quality work,
+% particularly when that work is susceptible to frequent interruptions
+% \cite{cheng2015break}.
 
-Research exploring the decomposition of work more generally
---- that is, without the constraint of an employer/requester directing workers ---
-has since emerged as well.
-\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work,
+\topic{While much of the above work has focused on
+clever arrangements of systems to manage collectives of workers,
+another body of work has turned inward toward managing the self.}
+\citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
 exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
@@
 the [scientific] management of the self as a worker
---- of sorts ---
+(of sorts)
 will prove relevant as we trace the literature surrounding piecework.
 
-The work described thus far has explored along myriad paths the ways that
-we can manage workers for the purpose of accomplishing complex tasks with
-rapidly sourced workers from all over the world.
-Critically, the work's defining characteristic involves
-the use of the Internet as a medium both 
-to coordinate workers, as well as to do the work itself.
-But research into the ``decomposition'' of work more generally
-illustrates the same concepts of work that ``Taylorism''
-and scientific management sought to embody
----
-\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger
-and warned of it in, but they were by no means alone in their warnings
-\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
+\topic{A smaller but growing body of work has reflected on these bodies of work
+and brought to our attention a number of ethical questions surrounding the
+increasing complexity of crowdwork and the hazards that increasingly arise.}
+\citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
+--- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
+for instance ---
+but these and other dangers range an enormous landscape % too informal?
+\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
 
-Following this thread of work, it may be said that
-much of the work in crowdsourcing has found ways of ``vertically slicing'' work
-such that each person is responsible for as small a task as can be made reasonable
---- taking a minute of audio, for instance, and transcribing that.
-The findings we take away from the crowdsourcing literature might be summed up as
-a number of ways
-to break tasks down \cite[see][]{selfsourcingTeevan2014,foundry,crowdForgeKittur} and
-measuring the costs as they relate to work completion \&
-to a lesser extent to workers
-\cite[see][]{measuringCrowdsourcingCheng,cheng2015break,crowdworkFuture,professionalCrowdworkEthics}.
+\topic{We summarize the work discussed so far in this way:
+% crowdsourcing researchers have attempted to find the characteristics of crowdwork
+researchers of crowdsourcing and on--demand labor in general
+have attempted to find the characteristics which
+enable and stymie successful applications of this form of work,
+and how to get past or around these boundaries.}
+Whether this work has focused on the self (as in the case of ``selfsourcing'')
+or on others (as in \textit{CrowdForge} and \textit{Foundry}),
+crowdsourcing literature has generally studied the constituent work
+and attempted either to work around or indeed with the limitations of crowdwork to accomplish a difficult task
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
+% Following this thread of work, it may be said that
+% much of the work in crowdsourcing has found ways of ``vertically slicing'' work
+% such that each person is responsible for as small a task as can be made reasonable
+% --- taking a minute of audio, for instance, and transcribing that.
+% The findings we take away from the crowdsourcing literature might be summed up as
+% a number of ways
+% to break tasks down \cite[see][]{selfsourcingTeevan2014,foundry,crowdForgeKittur} and
+% measuring the costs as they relate to work completion \&
+% to a lesser extent to workers
+% \cite[see][]{measuringCrowdsourcingCheng,cheng2015break,crowdworkFuture,professionalCrowdworkEthics}.
 
-Research into piecework has considered the decomposition of work;
-\citeauthor{10.2307/23702539} describes piecework in \citeyear{10.2307/23702539} as
+\topic{Research into piecework has explored similar limits,
+with particularly striking parallels in
+the language used to describe factors necessary for piecework to thrive.}
+\citeauthor{10.2307/23702539} describe piecework in \citeyear{10.2307/23702539} as
 ``\dots~based on examination of various shop jobs,
@@
 
-% \citeauthor{Brown01041990} draws more decisive boundaries on piece rate approaches to remuneration,
+\citeauthor{Brown01041990} draws more decisive boundaries on piece rate approaches to remuneration,
 identifying in analysis of data from
 the U.S. Bureau of Labor and Statistics that
-``larger establishment are less likely to use piece rates and predominantly female establishments more likely to do so; unionized establishments make significantly greater use of standard--rate pay than do nonunion establishments; \textbf{incentive pay is less likely in jobs with a variety of duties than in jobs with a narrow set of routinized duties}; and the ease of monitoring work quality is correlated with the use of incentive pay''
-% \cite{Brown01041990}.
+``larger establishment are less likely to use piece rates and predominantly female establishments more likely to do so; unionized establishments make significantly greater use of standard--rate pay than do nonunion establishments; \textit{incentive pay is less likely in jobs with a variety of duties than in jobs with a narrow set of routinized duties}; and the ease of monitoring work quality is correlated with the use of incentive pay''
+\cite{Brown01041990}.
 
@@
 
-\subsubsection{Work Abstraction}\label{sec:workAbstraction}
+\subsubsection{Work Abstraction}\label{sec:Slicing}
 Decomposition allows requesters to assign parallelized tasks
@@
     We discussed Fordism and Taylorism earlier in our discussions of
-    \nameref{sec:decomposition} and \nameref{sec:flexibility},
+    \nameref{sec:Complexity} and \nameref{sec:flexibility},
     but here the core of these paradigmatic views

### 2016-08-17 drafts/labor/trackingWork.tex
small notes
more topic sentences in cases.tex - see bold in cases.pdf

@@
 \usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref}
 \usepackage[normalem]{ulem}
@@
 \setlist{nolistsep}
-
 \def\plaintitle{Reexamining Crowd Work: A Historical Framing of On--Demand Labor as Piecework}
@@
 \newcommand{\note}[1]{\footnote{#1}}
+\newcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
+\onlyinsubfile{\renewcommand{\topic}[1]{\textbf{#1}}} % if you're in a subfile, bold the \topic sentences
+
 \begin{document}

### 2016-08-18 drafts/labor/cases.tex
cleaned up some comments
revs

@@
 \section{Major Research Questions}
-% \al¬i{4 paragraphs on what happened in the above contexts; 2 paragraphs on best/worst outcomes \& why; ??? paragraphs on predictions for crowdsourcing}
 
-%  _ _           _ _                __ 
-% | (_)_ __ ___ (_) |_ ___    ___  / _|
-% | | | '_ ` _ \| | __/ __|  / _ \| |_ 
-% | | | | | | | | | |_\__ \ | (_) |  _|
-% |_|_|_| |_| |_|_|\__|___/  \___/|_|
-
-%                            _                          _             
-%   ___ _ __ _____      ____| |___  ___  _   _ _ __ ___(_)_ __   __ _ 
-%  / __| '__/ _ \ \ /\ / / _` / __|/ _ \| | | | '__/ __| | '_ \ / _` |
-% | (__| | | (_) \ V  V / (_| \__ \ (_) | |_| | | | (__| | | | | (_| |
-%  \___|_|  \___/ \_/\_/ \__,_|___/\___/ \__,_|_|  \___|_|_| |_|\__, |
-%                                                               |___/
-
-% let's do topic sentences!
 \subsection{What are the limits of crowdsourcing?}
@@
 \begin{inlinelist}
-  % \item \cref*{sec:Complexity},
   \item \nameref{sec:Complexity},
@@
 to the question of the furthest limits of crowdsourcing.
-
-%      _                                          _ _   _             
-%   __| | ___  ___ ___  _ __ ___  _ __   ___  ___(_) |_(_) ___  _ __  
-%  / _` |/ _ \/ __/ _ \| '_ ` _ \| '_ \ / _ \/ __| | __| |/ _ \| '_ \ 
-% | (_| |  __/ (_| (_) | | | | | | |_) | (_) \__ \ | |_| | (_) | | | |
-%  \__,_|\___|\___\___/|_| |_| |_| .__/ \___/|___/_|\__|_|\___/|_| |_|
-%                                |_|                                  
-%
-% stronger topic sentences
-% boosterist (???)
 
@@
 
-% NEEDS TO GO ELSEWHERE
-% \citeauthor{cheng2015break} found that microtasks
-% --- though not necessarily \textit{faster} than ``macrotasks'' ---
-% yield higher quality work,
-% particularly when that work is susceptible to frequent interruptions
-% \cite{cheng2015break}.
-
 \topic{While much of the above work has focused on
@@
 
-\topic{A smaller but growing body of work has reflected on these bodies of work
+\topic{A smaller but growing body of work has reflected on this
 and brought to our attention a number of ethical questions surrounding the
@@
 \topic{We summarize the work discussed so far in this way:
-% crowdsourcing researchers have attempted to find the characteristics of crowdwork
 researchers of crowdsourcing and on--demand labor in general
@@
 \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
-% Following this thread of work, it may be said that
-% much of the work in crowdsourcing has found ways of ``vertically slicing'' work
-% such that each person is responsible for as small a task as can be made reasonable
-% --- taking a minute of audio, for instance, and transcribing that.
-% The findings we take away from the crowdsourcing literature might be summed up as
-% a number of ways
-% to break tasks down \cite[see][]{selfsourcingTeevan2014,foundry,crowdForgeKittur} and
-% measuring the costs as they relate to work completion \&
-% to a lesser extent to workers
-% \cite[see][]{measuringCrowdsourcingCheng,cheng2015break,crowdworkFuture,professionalCrowdworkEthics}.
 
-\topic{Research into piecework has explored similar limits,
-with particularly striking parallels in
-the language used to describe factors necessary for piecework to thrive.}
-\citeauthor{10.2307/23702539} describe piecework in \citeyear{10.2307/23702539} as
+% \topic{Research into piecework has also explored the limits of piecework,
+% with some findings.}
+
+\ali{
+
+Talked to someone about this section and got some feedback\dots
+\begin{enumerate}
+  \item Explain what ``clever arrangement of work'' in the first paragraph means
+        (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
+  \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
+        (maybe even to justify its placement here); and
+  \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
+        to make the historical stuff on piecework
+        directly speak to the three previous paragraphs.
+  \item (Maybe?) summarize the piecework stuff
+        (the last paragraph approaches that, but more could be done).
+\end{enumerate}
+Thoughts on this?
+
+}
+
+% the language used to describe the factors considered for piecework to thrive.}
+First, we look to
+\citeauthor{10.2307/23702539},
+who describes piecework as
     ``\dots~based on examination of various shop jobs,
-which included calculation of the standard time and compensation for each task.
-Most piecework plans required additional clerical and supervisory people
-(such as piecework clerks, inspectors, and ``experts'')\dots'',
-allowing for what the \textit{American Machinist} would later call ``paper efficiency'',
-describing savings on paper that are in fact passed on to workers
+    which included calculation of the standard time and compensation for each task''
 \cite{10.2307/23702539}.
-In this context,
-the introspection of tasks by experts and the careful work assignment \& management of workers
-we see in historical piecework seem to parallel crowdsourcing.
-
-\citeauthor{Brown01041990} draws more decisive boundaries on piece rate approaches to remuneration,
-identifying in analysis of data from
-the U.S. Bureau of Labor and Statistics that
-``larger establishment are less likely to use piece rates and predominantly female establishments more likely to do so; unionized establishments make significantly greater use of standard--rate pay than do nonunion establishments; \textit{incentive pay is less likely in jobs with a variety of duties than in jobs with a narrow set of routinized duties}; and the ease of monitoring work quality is correlated with the use of incentive pay''
+\citeauthor{10.2307/23702539} later enumerates some of the roles required
+to facilitate piecework in the early \nth{20} century:
+    ``\dots~piecework clerks, inspectors, and ``experts''\dots''
+\cite{10.2307/23702539}.
+\citeauthor{Brown01041990} makes a more assertive
+claim about the factors necessary for piecework to thrive,
+specifically arguing that
+    ``\dots~incentive pay is less likely in jobs with
+    a variety of duties than in jobs with a narrow set of routinized duties''
 \cite{Brown01041990}.
-
-What's different today? We can turn to \citeauthor{10.2307/23702539}'s reflection on railroad piecework
-for a hint of what might have enabled the future of crowdwork in which we now live:
-``\dots~only the largest and most wealthy railroads stood to gain substantially from scientific management,
-because only they had the resources to eliminate union resistance and
-\textit{pay the overhead involved in installing work reorganization}''
-(emphasis added)
+\citeauthor{10.2307/23702539} makes another claim that
+successful cases of piecework owed themselves in part to the fact that
+    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
+    pay the overhead involved in installing work reorganization''
 \cite{10.2307/23702539}.
 
-We argue that crowdsourcing is more widespread today,
-and has the potential to become even larger in scale,
-because the underlying architecture of crowdwork is digital;
-reformulating workflows to decompose tasks and institute piecework compensation programs
-is easier today than it was for railroad companies which had to move physical capital to do so.
+\topic{Digital media have expanded the scope of viable piecework by pushing drastically on the limits
+cited by piecework researchers.}
+The research on piecework tells us that
+we should expect piecework to thrive in industries where
+the nature of the work is limited in complexity.
+Given the flourishing of on--demand labor platforms such as
+Uber, Amazon Mechanical Turk, and others, we ask ourselves
+what --- if anything --- has changed.
+We argue that the Internet has trivialized the costs and challenges of the earlier limiting factors for two reasons:
+\begin{inlinelist}
+  \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
+  \item The Internet allows us to leverage the benefits of ``economies of scale'' at very little cost
+        to the system--designer \cite{lessig2006code}
+\end{inlinelist}.
 
-\ali{\citeauthor{10.2307/23702539} later argues that a significant obstacle to the introduction of piecework in railroad shops was the resistance of workers
+\ali{\citeauthor{10.2307/23702539} also argues that
+a significant obstacle to the introduction of piecework in railroad shops was
+the resistance of workers
 (the other being the resistance of management (for different and varying reasons)).
 In the cases of online platforms like AMT and digitally mediated platforms like Uber,
-where in both cases workers rarely if ever interact face--to--face and opportunities for coordination and collective action are severely limited,
-we should expect to see the challenges stymieing collective action that we see in \citeauthor{dynamo} and elsewhere
-\cite{10.2307/23702539,dynamo}.
-}
+where in both cases workers rarely
+if ever
+interact face--to--face and
+opportunities for coordination and collective action are severely limited,
+we should expect to see
+the challenges stymieing collective action
+that we see in \citeauthor{dynamo} and elsewhere
+\cite{10.2307/23702539,dynamo}.}
 
-%                     _            _         _                  _   _             
-% __      _____  _ __| | __   __ _| |__  ___| |_ _ __ __ _  ___| |_(_) ___  _ __  
-% \ \ /\ / / _ \| '__| |/ /  / _` | '_ \/ __| __| '__/ _` |/ __| __| |/ _ \| '_ \ 
-%  \ V  V / (_) | |  |   <  | (_| | |_) \__ \ |_| | | (_| | (__| |_| | (_) | | | |
-%   \_/\_/ \___/|_|  |_|\_\  \__,_|_.__/|___/\__|_|  \__,_|\___|\__|_|\___/|_| |_|
-%
+\subsubsection{Slicing work smaller
+\joke{ --- or ``putting the {\tiny `micro'} in microwork''}
+}\label{sec:Slicing}
 
-\subsubsection{Work Abstraction}\label{sec:Slicing}
 Decomposition allows requesters to assign parallelized tasks
@@
   \begin{appendices}
-  Points to make:
-  \begin{itemize}
-    \item Blossoming of piecework
-    \begin{itemize}
-      \item high point: consultants?
-      \item most cases: auto workers/etc\dots
-      \item worst case: sweatshops (especially in developing nations)
-    \end{itemize}
-    \item what led to these outcomes?
-    \begin{itemize}
-      \item ``consultant'' work came out well because the work was complex;
-      this made it difficult to turn into a mass market commodity.
-      We see consultants ranging skill levels like
-      oDesk (implementing modules) to
-      Accenture (on--demand teams of consultants).
-      \item auto workers,
-      working in settings where capital couldn't be moved as easily
-      found themselves in the same workspace as a direct
-      --- if multi--stepped ---
-      result to the benefits of putting people in the same place to consolidate resources.
-      Moreover,
-      workers had leverage over factory owners as a result of that consolidated capital;
-      operating equipment required training.
-      \item sweatshop workers fared the worst,
-      for reasons that may seem obvious with hindsight.
-      Source materials for textiles are easier to ship than mechanical components such as engines,
-      making it easier for factories to relocate to developing nations
-      (where cost--of--living, and consequently wages, would be lower).
-      As wages, CoLAs, and QoL rise, workers begin asking for
-      (and later demanding)
-      higher wages and better conditions.
-      But where the auto workers have leverage,
-      textile workers find only a precarious economic balance now tipped by their collective action,
-      spurring manufacturers to move to a new locale
-    \end{itemize}
-  \end{itemize}
+\section{Graveyard}
+
+\citeauthor{cheng2015break} found that microtasks
+--- though not necessarily \textit{faster} than ``macrotasks'' ---
+yield higher quality work,
+particularly when that work is susceptible to frequent interruptions
+\cite{cheng2015break}.
 

### 2016-08-18 drafts/labor/trackingWork.tex
revs

--- 2016-08-14 drafts/labor/trackingWork.tex
@@
 \newcommand{\note}[1]{\footnote{#1}}
+\newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences
+\newcommand{\joke}[1]{#1} % if you're in a subfile, bold the \topic sentences
+
 \begin{document}
+\renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
+\renewcommand{\joke}[1]{} % don't do anything with \jokes! Just throw them away!
 \renewcommand{\onlyinsubfile}[1]{}

### 2016-08-19 drafts/labor/cases.tex
notes from meeting

@@
 \cite{crowdForgeKittur}.
+\ali{Types of things that people are trying to do that are complex: e.g. 1, e.g. 2, etc\dots}
 \citeauthor{Lasecki:2013:CCC:2501988.2502057} tackle an area of computationally assisted interaction
@@
 \cite[][and others]{Lasecki:2013:RCL:2441776.2441912,foundry}.
+
+\ali{the thing that has made all of this possible has been decomposition and the application of computational processes (e.g. MapReduce) to make it all possible}
 
@@
 
+\ali{
+Nuke this:
 \topic{We summarize the work discussed so far in this way:
@@
 
+Crowdforge, turkomatic, others talk about limiting factors; bring them up
+}
 % \topic{Research into piecework has also explored the limits of piecework,

### 2016-08-20 drafts/labor/cases.tex
friday edits, mostly rearranging, some additions

--- 2016-08-18 drafts/labor/cases.tex
@@
 
-\subsubsection{Achieving greater complexity}\label{sec:Complexity}
-\topic{Crowdsourcing research has spent the better part of a decade attempting
-to prove the viability of crowdsourcing in increasingly complex work.}  % this is my topic sentence.
-\citeauthor{crowdForgeKittur}
-map the discussion toward this goal in their work on
-crowdsourcing complex work
-\cite{crowdForgeKittur}.
-\citeauthor{Lasecki:2013:CCC:2501988.2502057} tackle an area of computationally assisted interaction
---- conversational artificial intelligence ---
-and find ways to arrange workers and formulate a ``crowd--powered conversational assistant''
-\cite{Lasecki:2013:CCC:2501988.2502057}.
-Other work has followed a similar pattern,
-identifying target milestones in Computer Science
-that have presented significant challenges for researchers,
-and demonstrating that ``crowd--powered'' systems can tackle many of these challenges
-with generally greater success than purely computational attempts
-\cite[][and others]{Lasecki:2013:RCL:2441776.2441912,foundry}.
-
-\topic{While much of the above work has focused on
-clever arrangements of systems to manage collectives of workers,
-another body of work has turned inward toward managing the self.}
-\citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
-exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-While some of this work doesn't strictly fall under ``crowdsourcing'',
-the [scientific] management of the self as a worker
-(of sorts)
-will prove relevant as we trace the literature surrounding piecework.
-
-\topic{A smaller but growing body of work has reflected on this
-and brought to our attention a number of ethical questions surrounding the
-increasing complexity of crowdwork and the hazards that increasingly arise.}
-\citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
---- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
-for instance ---
-but these and other dangers range an enormous landscape % too informal?
-\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
-
-\topic{We summarize the work discussed so far in this way:
-researchers of crowdsourcing and on--demand labor in general
-have attempted to find the characteristics which
-enable and stymie successful applications of this form of work,
-and how to get past or around these boundaries.}
-Whether this work has focused on the self (as in the case of ``selfsourcing'')
-or on others (as in \textit{CrowdForge} and \textit{Foundry}),
-crowdsourcing literature has generally studied the constituent work
-and attempted either to work around or indeed with the limitations of crowdwork to accomplish a difficult task
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
-
-% \topic{Research into piecework has also explored the limits of piecework,
-% with some findings.}
-
-\ali{
-
-Talked to someone about this section and got some feedback\dots
-\begin{enumerate}
-  \item Explain what ``clever arrangement of work'' in the first paragraph means
-        (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
-  \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
-        (maybe even to justify its placement here); and
-  \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
-        to make the historical stuff on piecework
-        directly speak to the three previous paragraphs.
-  \item (Maybe?) summarize the piecework stuff
-        (the last paragraph approaches that, but more could be done).
-\end{enumerate}
-Thoughts on this?
-
-}
-
-% the language used to describe the factors considered for piecework to thrive.}
-First, we look to
-\citeauthor{10.2307/23702539},
-who describes piecework as
-    ``\dots~based on examination of various shop jobs,
-    which included calculation of the standard time and compensation for each task''
-\cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} later enumerates some of the roles required
-to facilitate piecework in the early \nth{20} century:
-    ``\dots~piecework clerks, inspectors, and ``experts''\dots''
-\cite{10.2307/23702539}.
-\citeauthor{Brown01041990} makes a more assertive
-claim about the factors necessary for piecework to thrive,
-specifically arguing that
-    ``\dots~incentive pay is less likely in jobs with
-    a variety of duties than in jobs with a narrow set of routinized duties''
-\cite{Brown01041990}.
-\citeauthor{10.2307/23702539} makes another claim that
-successful cases of piecework owed themselves in part to the fact that
-    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
-    pay the overhead involved in installing work reorganization''
-\cite{10.2307/23702539}.
-
-\topic{Digital media have expanded the scope of viable piecework by pushing drastically on the limits
-cited by piecework researchers.}
-The research on piecework tells us that
-we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity.
-Given the flourishing of on--demand labor platforms such as
-Uber, Amazon Mechanical Turk, and others, we ask ourselves
-what --- if anything --- has changed.
-We argue that the Internet has trivialized the costs and challenges of the earlier limiting factors for two reasons:
-\begin{inlinelist}
-  \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
-  \item The Internet allows us to leverage the benefits of ``economies of scale'' at very little cost
-        to the system--designer \cite{lessig2006code}
-\end{inlinelist}.
-
-\ali{\citeauthor{10.2307/23702539} also argues that
-a significant obstacle to the introduction of piecework in railroad shops was
-the resistance of workers
-(the other being the resistance of management (for different and varying reasons)).
-In the cases of online platforms like AMT and digitally mediated platforms like Uber,
-where in both cases workers rarely
-if ever
-interact face--to--face and
-opportunities for coordination and collective action are severely limited,
-we should expect to see
-the challenges stymieing collective action
-that we see in \citeauthor{dynamo} and elsewhere
-\cite{10.2307/23702539,dynamo}.}
+\subfile{complexity}
 
@@
 in various binary classification tasks
-% on the principle that diverse workers complete these tasks
-% in order to accurately inform the model on the variety of delays in response times.
 \cite{embracingErrorKrishna}.
@@
 \end{inlinelist}.
-
-% \subsubsection{Piecework's role here}
-% We address the question of the limits of crowdsourcing in three parts:
-% \begin{inlinelist}
-% \item decomposition
-% \item work abstraction
-% and \item flexibility
-% \end{inlinelist}.
-% The literature on piecework has offered insights into each of these topics, which we will explore now.
 
@@
 
+\endnotetext{
     \citeauthor{cheng2015break} found that microtasks
@@
     \cite{cheng2015break}.
+}
 
@@
 
-    % Low pay yields variable quality work for a number of reasons,
-    % but before we discuss the causes of this effect, we should discuss the
-
-    % Variance in the quali
-
     The effect low wages have had on piece work and factory workers is well--known;
@@
 
+  \theendnotes
   \end{appendices}

### 2016-08-20 drafts/labor/complexity.tex
friday edits, mostly rearranging, some additions

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\subsubsection{Achieving greater complexity
\joke{ --- or ``making crowd--work \textit{yuge}''}
% i think it says a lot about my confidence in my latex skillz that i'm willing to put these in the subfile
}\label{sec:Complexity}

\topic{Crowdsourcing research has spent the better part of a decade attempting
to prove the viability of crowdsourcing in increasingly complex work.}  % this is my topic sentence.
\citeauthor{crowdForgeKittur}
map the discussion toward this goal in their work on
crowdsourcing complex work
\cite{crowdForgeKittur}.
The broader body of work has varied significantly in type
--- providing conversational assistants,
interpreting medical data,
telling coherent and compelling stories
\cite{Lasecki:2013:CCC:2501988.2502057,mavandadi2012distributed,KimStoria}.

\topic{This diverse body of research has involved similar approaches to problems,
namely consisting of insights made by the Computer Science community.}
The crowdwork literature typically identifies target milestones in CS
that have presented significant challenges for researchers,
leverages some of the approaches and insights that Computer Science researchers have already made
(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge}),
and arranges humans as computational black boxes within those approaches and processes
\cite[][and others]{foundry,crowdForgeKittur}.

% \ali{the thing that has made all of this possible has been decomposition and the application of computational processes (e.g. MapReduce) to make it all possible}

\endnotetext{
  This is also kind of off-topic, isn't it? where to take it?

  \topic{A smaller but growing body of work has reflected on this
  and brought to our attention a number of ethical questions surrounding the
  increasing complexity of crowdwork and the hazards that increasingly arise.}
  \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
  --- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
  for instance ---
  but these and other dangers range an enormous landscape % too informal?
  \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
% }

% \endnotetext{
  Nuke this?

  \topic{We summarize the work discussed so far in this way:
  researchers of crowdsourcing and on--demand labor in general
  have attempted to find the characteristics which
  enable and stymie successful applications of this form of work,
  and how to get past or around these boundaries.}
  Whether this work has focused on the self (as in the case of ``selfsourcing'')
  or on others (as in \textit{CrowdForge} and \textit{Foundry}),
  crowdsourcing literature has generally studied the constituent work
  and attempted either to work around
  or indeed with
  the limitations of crowdwork to accomplish a difficult task
  \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.

  Crowdforge, turkomatic, others talk about limiting factors; bring them up
}

\endnotetext{
    Talked to someone about this section and got some feedback\dots
    \begin{enumerate}
      \item Explain what ``clever arrangement of work'' in the first paragraph means
            (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
      \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
            (maybe even to justify its placement here); and
      \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
            to make the historical stuff on piecework
            directly speak to the three previous paragraphs.
      \item (Maybe?) summarize the piecework stuff
            (the last paragraph approaches that, but more could be done).
    \end{enumerate}
    Thoughts on this?
}

\topic{The research into piecework has identified a number of
major constraints that we should expect to find when
applying piecework approaches to labor markets.}
% To illustrate this, we look to 
% First, we look to
\citeauthor{10.2307/23702539},
who describes piecework as
    ``\dots~based on examination of various shop jobs,
    which included calculation of the standard time and compensation for each task''
\cite{10.2307/23702539}.
\citeauthor{10.2307/23702539} later enumerates some of the roles required
to facilitate piecework in the early \nth{20} century:
    ``\dots~piecework clerks, inspectors, and ``experts''\dots''
\cite{10.2307/23702539}.
\citeauthor{Brown01041990} makes a more assertive
claim about the factors necessary for piecework to thrive,
specifically arguing that
    ``\dots~incentive pay is less likely in jobs with
    a variety of duties than in jobs with a narrow set of routinized duties''
\cite{Brown01041990}.
\citeauthor{10.2307/23702539} makes another claim that
successful cases of piecework owed themselves in part to the fact that
    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
    pay the overhead involved in installing work reorganization''
\cite{10.2307/23702539}.

\topic{Digital media have expanded the scope of viable piecework by pushing drastically on the limits
cited by piecework researchers.}
The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity.
Given the flourishing of on--demand labor platforms such as
Uber, Amazon Mechanical Turk, and others, we ask ourselves
what --- if anything --- has changed.
We argue that the Internet has trivialized the costs and challenges of the earlier limiting factors for two reasons:
\begin{inlinelist}
  \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
  \item The Internet allows us to leverage the benefits of ``economies of scale'' at very little cost
        to the system--designer \cite{lessig2006code}
\end{inlinelist}.

\ali{\citeauthor{10.2307/23702539} also argues that
a significant obstacle to the introduction of piecework in railroad shops was
the resistance of workers
(the other being the resistance of management (for different and varying reasons)).
In the cases of online platforms like AMT and digitally mediated platforms like Uber,
where in both cases workers rarely
if ever
interact face--to--face and
opportunities for coordination and collective action are severely limited,
we should expect to see
the challenges stymieing collective action
that we see in \citeauthor{dynamo} and elsewhere
\cite{10.2307/23702539,dynamo}.}

\topic{}
% \topic{While much of the above work has focused on
% clever arrangements of systems to manage collectives of workers,
% another body of work has turned inward toward managing the self.}
\citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
While some of this work doesn't strictly fall under ``crowdsourcing'',
the [scientific] management of the self as a worker
(of sorts)
will prove relevant as we trace the literature surrounding piecework.

\end{document}

### 2016-08-25 drafts/labor/complexity.tex
moving complexity.tex stuff around

@@
 --- providing conversational assistants,
-interpreting medical data,
-telling coherent and compelling stories
+interpreting medical data, and
+telling coherent and compelling stories,
+to name a few examples
 \cite{Lasecki:2013:CCC:2501988.2502057,mavandadi2012distributed,KimStoria}.
 
-\topic{This diverse body of research has involved similar approaches to problems,
-namely consisting of insights made by the Computer Science community.}
+\topic{This body of research has involved similar approaches to problems,
+often involving insights made in Computer Science and applied to human work--flows.}
 The crowdwork literature typically identifies target milestones in CS
@@
 and arranges humans as computational black boxes within those approaches and processes
-\cite[][and others]{foundry,crowdForgeKittur}.
+\cite[][and others]{crowdForgeKittur,foundry}.
 
-% \ali{the thing that has made all of this possible has been decomposition and the application of computational processes (e.g. MapReduce) to make it all possible}
-
-\endnotetext{
-  This is also kind of off-topic, isn't it? where to take it?
-
-  \topic{A smaller but growing body of work has reflected on this
-  and brought to our attention a number of ethical questions surrounding the
-  increasing complexity of crowdwork and the hazards that increasingly arise.}
-  \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
-  --- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
-  for instance ---
-  but these and other dangers range an enormous landscape % too informal?
-  \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
-% }
-
-% \endnotetext{
-  Nuke this?
-
-  \topic{We summarize the work discussed so far in this way:
-  researchers of crowdsourcing and on--demand labor in general
-  have attempted to find the characteristics which
-  enable and stymie successful applications of this form of work,
-  and how to get past or around these boundaries.}
-  Whether this work has focused on the self (as in the case of ``selfsourcing'')
-  or on others (as in \textit{CrowdForge} and \textit{Foundry}),
-  crowdsourcing literature has generally studied the constituent work
-  and attempted either to work around
-  or indeed with
-  the limitations of crowdwork to accomplish a difficult task
-  \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
-
-  Crowdforge, turkomatic, others talk about limiting factors; bring them up
-}
-
-\endnotetext{
-    Talked to someone about this section and got some feedback\dots
-    \begin{enumerate}
-      \item Explain what ``clever arrangement of work'' in the first paragraph means
-            (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
-      \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
-            (maybe even to justify its placement here); and
-      \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
-            to make the historical stuff on piecework
-            directly speak to the three previous paragraphs.
-      \item (Maybe?) summarize the piecework stuff
-            (the last paragraph approaches that, but more could be done).
-    \end{enumerate}
-    Thoughts on this?
-}
+\topic{This approach has proven a compelling one because
+it leverages the in--built advantages that technology and digital media afford.}
+\textit{Foundry}'s tools for managing and arranging expert groups into a cohort
+allow researchers to convincingly argue that expert teams can be rapidly formed,
+just like non--expert teams
+\cite{foundry}.
 
@@
 major constraints that we should expect to find when
-applying piecework approaches to labor markets.}
-% To illustrate this, we look to 
-% First, we look to
+applying piecework approaches to work.}
 \citeauthor{10.2307/23702539},
@@
 
-\ali{\citeauthor{10.2307/23702539} also argues that
+\topic{\joke{great} topic sentence here.}
+% \topic{While much of the above work has focused on
+% clever arrangements of systems to manage collectives of workers,
+% another body of work has turned inward toward managing the self.}
+\citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
+exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+While some of this work doesn't strictly fall under ``crowdsourcing'',
+the [scientific] management of the self as a worker
+(of sorts)
+will prove relevant as we trace the literature surrounding piecework.
+
+\endnotetext{
+  This is also kind of off-topic, isn't it? where to take it?
+
+  \topic{A smaller but growing body of work has reflected on this
+  and brought to our attention a number of ethical questions surrounding the
+  increasing complexity of crowdwork and the hazards that increasingly arise.}
+  \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
+  --- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
+  for instance ---
+  but these and other dangers range an enormous landscape % too informal?
+  \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
+
+  kill this?
+
+  \topic{We summarize the work discussed so far in this way:
+  researchers of crowdsourcing and on--demand labor in general
+  have attempted to find the characteristics which
+  enable and stymie successful applications of this form of work,
+  and how to get past or around these boundaries.}
+  Whether this work has focused on the self (as in the case of ``selfsourcing'')
+  or on others (as in \textit{CrowdForge} and \textit{Foundry}),
+  crowdsourcing literature has generally studied the constituent work
+  and attempted either to work around
+  or indeed with
+  the limitations of crowdwork to accomplish a difficult task
+  \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
+
+  Crowdforge, turkomatic, others talk about limiting factors; bring them up
+
+    Talked to someone about this section and got some feedback\dots
+    \begin{enumerate}
+      \item Explain what ``clever arrangement of work'' in the first paragraph means
+            (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
+      \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
+            (maybe even to justify its placement here); and
+      \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
+            to make the historical stuff on piecework
+            directly speak to the three previous paragraphs.
+      \item (Maybe?) summarize the piecework stuff
+            (the last paragraph approaches that, but more could be done).
+    \end{enumerate}
+    Thoughts on this?
+
+\citeauthor{10.2307/23702539} also argues that
 a significant obstacle to the introduction of piecework in railroad shops was
@@
 that we see in \citeauthor{dynamo} and elsewhere
-\cite{10.2307/23702539,dynamo}.}
+\cite{10.2307/23702539,dynamo}.
 
-\topic{}
-% \topic{While much of the above work has focused on
-% clever arrangements of systems to manage collectives of workers,
-% another body of work has turned inward toward managing the self.}
-\citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
-exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-While some of this work doesn't strictly fall under ``crowdsourcing'',
-the [scientific] management of the self as a worker
-(of sorts)
-will prove relevant as we trace the literature surrounding piecework.
+}
+
+\onlyinsubfile{
+  % \balance{}
+  \printbibliography
+  \clearpage
+  \nobalance{}
+  \begin{appendices}
+  \theendnotes
+  \end{appendices}
+}
 

### 2016-08-25 drafts/labor/trackingWork.tex
friday edits, mostly rearranging, some additions
moving complexity.tex stuff around

@@
 
-\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix}
+\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
 \usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref}
 \usepackage[normalem]{ulem}
-% \usepackage{url}
-% \def\UrlOrds{\do\*\do\-\do\~\do\'\do\"\do\-}%
-% \usepackage{multibib}
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-% \usepackage{biblatex}
-% \usepackage[square,numbers]{jnatbib}
 \usepackage[T1]{fontenc}
@@
 % \bibliographystyle{SIGCHI-Reference-Format}
-
+\renewcommand*{\notesname}{Graveyard of old paragraphs}
 \makeatletter
@@
 \definecolor{BrickRed}{HTML}{FF0000}
+\definecolor{JokeGreen}{HTML}{00C953}
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
@@
 \newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences
-\newcommand{\joke}[1]{#1} % if you're in a subfile, bold the \topic sentences
+\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, bold the \topic sentences
 

### 2016-08-26 drafts/labor/cases.tex
thursday edits and wb feedback

@@
 
+We look at lots of papers that use the term ``crowdsourcing'' or ``crowdwork''
+in the abstracts and titles of their papers,
+especially the papers that we cited earlier, to try
+to answer the questions that we posed in the introduction.
+Then we look to the piecework literature to see whether and to what extent
+piecework answers the questions we, crowdsourcing researchers, have asked.
+We then see whether and how crowdsourcing as we know it has differed from piecework,
+and how that affects the predictions and conclusions made in the piecework literature.
+% look at review papers
+
+% what's the methodology?
+% make piecework literature much clearer about argument
+% make headers clearer to signal what i'm doing here.
+
 \subsection{What are the limits of crowdsourcing?}
-Research in crowdsourcing has spent the better part of a decade
+\topic{Research in crowdsourcing has spent the better part of a decade
 exploring how to grow the limits of crowdsourcing and
-find the boundaries of crowd work and microtasks;
-this has largely involved
+find the boundaries of crowd work and microtasks.}
+This has largely involved
   identifying challenges to this form of labor,
@@
   repeating the process
-\cite{bernsteinSoylent}.
+\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
 The question that has emerged

### 2016-08-26 drafts/labor/complexity.tex
thursday edits and wb feedback

@@
 }\label{sec:Complexity}
-
+% crowd weaver, aditya and someone else's book
 \topic{Crowdsourcing research has spent the better part of a decade attempting
@@
 \cite[][and others]{crowdForgeKittur,foundry}.
-
-\topic{This approach has proven a compelling one because
-it leverages the in--built advantages that technology and digital media afford.}
+This approach has proven a compelling one because
+it leverages the in--built advantages that technology and digital media afford.
 \textit{Foundry}'s tools for managing and arranging expert groups into a cohort
@@
 
-\topic{The research into piecework has identified a number of
-major constraints that we should expect to find when
-applying piecework approaches to work.}
+\topic{The research into piecework makes the case that
+piecework has been limited principally by the challenges of human management and oversight.}
 \citeauthor{10.2307/23702539},
@@
     ``\dots~based on examination of various shop jobs,
-    which included calculation of the standard time and compensation for each task''
-\cite{10.2307/23702539}.
+    which included calculation of the standard time and compensation for each task'',
+argues that piecework must be rigorously evaluated at a time that demands
+\textit{other people} perform the evaluation.
+% \cite{10.2307/23702539}.
 \citeauthor{10.2307/23702539} later enumerates some of the roles required
@@
 \cite{10.2307/23702539}.
-\citeauthor{Brown01041990} makes a more assertive
-claim about the factors necessary for piecework to thrive,
-specifically arguing that
+This criterion strictly limits the extent to which piecework can grow in complexity;
+it must, for instance, be quickly evaluable by another person.
+
+\topic{Piecework researchers also make claims regarding
+the organizations that benefit from piecework in the first place.}
+\citeauthor{Brown01041990} discusses the factors necessary for piecework to thrive:
     ``\dots~incentive pay is less likely in jobs with
@@
 \cite{Brown01041990}.
-\citeauthor{10.2307/23702539} makes another claim that
+\citeauthor{10.2307/23702539} adds further, that
 successful cases of piecework owed themselves in part to the fact that
@@
 \cite{10.2307/23702539}.
+Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
+make a persuasive argument that piecework is limited in complexity by
+managerial overhead and the fixed cost of adopting a piecework payment regime.
 

### 2016-08-26 drafts/labor/introduction.tex
moving things around
thursday edits and wb feedback

@@
 
+% what am i trying to say? that crowd work is an instantiation of piecework
+% framing/bullet points
+
 \section{Introduction}
-The past decade has seen a flourishing of on--demand work
-where the statuses of workers have become so fleeting that 
-these workers
+\topic{The past decade has seen a flourishing of on--demand work,
+motivated by the framing of work as
+contextually abstract and therefore
+interchangeable and modular.}
+The statuses of many of these workers
 (known colloquially as ``Turkers'')
-have been described as ``transient''
+have become so fleeting that 
+they have been described as ``transient''
 \cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
@@
 
+\topic{This matters because\dots}
 Indeed, this on--demand workforce has sparked interest
 across industries ranging
-livery (driving for hire, for example Uber),
+driving for hire (for example Uber),
 house--cleaning (Handy),
@@
   what can and cannot be done by crowd workers?
-    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
+    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,
+          Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
   \item What forms of work design, and worker management and arrangement, are viable?
-    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
+    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,
+          Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
+          Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
   \item What will work and the place of work look like for the workers?
-    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}
+    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,
+          takingAHITMcInnis}
 \end{enumerate}
@@
 Indeed,
-\citeauthor{crowdworkFuture} in \citeauthor{crowdworkFuture}
+\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
 referenced crowd work as ``piecework'' briefly
@@
 \cite{crowdworkFuture}.
+But more than this,
+the framing of on--demand labor as a re--instantiation of piecework
+gives us more material to make sense of the broader research on this new form of work
+by evaluating this work through a much more refined theoretical lens,
+informed by decades of rigorous, empirically based research.
 
-But more than this,
-the framing of on--demand labor as piecework (re--instantiated)
-allows us to attempt to make sense of the broader research on this new form of work
-by evaluating this work through a much more refined lens.
 More concretely, by looking at task--based or ``gig'' work as
@@
   \item make sense of the phenomena so far as part of a much larger series of interrelated events;
-  \item bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
+  \item bring into focus the ongoing work among
+        workers,
+        system--designers, and
+        researchers in this space;
+        and finally,
   \item offer predictions of what social computing researchers,
@@
 and developments we should expect to see in the maturation of digitally mediated work.
-Finally, we will offer design implications based on this research\dots
+Finally, we will offer design implications based on this research.
 
-\endnote{
-\textbf{notes/what i want to get across\dots}
-
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+  % \theendnotes
+\nobalance{}
+\clearpage
+\begin{appendices}
+\section{Sleeping kittens}
 \begin{itemize}
@@
 \end{itemize}
-\upshape}
+\end{appendices}
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-  \theendnotes
 }

### 2016-08-29 drafts/labor/introduction.tex
friday changes committed sunday

@@
 \documentclass[trackingWork]{subfiles}
+\onlyinsubfile{
+  \hypersetup{
+    citecolor  = blue,
+    linkcolor  = JokeGreen,
+  }
+}
+
 \makeatletter
@@
 \makeatother
-
 \begin{document}
-
-% what am i trying to say? that crowd work is an instantiation of piecework
-% framing/bullet points
-
 \section{Introduction}
 \topic{The past decade has seen a flourishing of on--demand work,
-motivated by the framing of work as
-contextually abstract and therefore
-interchangeable and modular.}
-The statuses of many of these workers
-(known colloquially as ``Turkers'')
-have become so fleeting that 
-they have been described as ``transient''
+largely driven by the reformulation of work as
+the constituent parts of larger tasks.}
+This framing of work into abstract blocks
+has allowed people to engage in work despite
+limited time,
+little to no awareness of the broader context of the work, and 
+(often) fleeting identities and associations
 \cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
 The realization
-that tasks can be accomplished by directing and managing this crowd of workers
+that complex tasks can be accomplished by directing and managing crowds of workers
 has spurred the research and and industry communities to flock to sites of labor
 like Amazon's Mechanical Turk (AMT)
-to explore the limits of this distributed, seemingly ephemeral labor force.
+to explore the limits of this distributed, fleeting workforce.
 Researchers in particular have taken to the space in earnest,
 finding opportunities to enable new forms of work
-as well as using Turkers
-as representative populations of the public
+using this population of ``Turkers''
+\note{as test subjects in a sort of at--scale laboratory for economic and other behavioral studies}
 \cite{bernsteinSoylent,redballoon,paolacci2010running}.
 
-The many sites of work replicating and extending on
-the general style of labor popularized by AMT
-have predominantly involved work done on a computer
-or involving the human processing of data,
-leading many to call this ``information work"
-\cite{turkopticon,professionalCrowdworkEthics,IraniFromCriticalDesign,OlsonMakeDistanceWorkWork}.
-\citeauthor{howe2008crowdsourcing} defined ``crowdsourcing'' in \citeyear{howe2008crowdsourcing}
-as
-``taking a job traditionally performed by a designated agent
-(usually an employee)
-and outsourcing it to an undefined,
-generally large group of people in the form of an open call''
-\cite{howe2008crowdsourcing}.
+\topic{This form of work has grown considerably in size,
+far beyond the domain of ``information work'' from which it first sprang.}
+While \citeauthor{howe2008crowdsourcing} described crowdsourcing as
+``outsourcing [work] to an undefined, generally large group of people in the form of an open call'',
+for years the instantiation of this work was limited to the utilization of
+human intelligence to process data and act on information
+\cite{CrowdsourcingUserStudies,movieSummarizationWu,
+      yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
+More recently, crowdsourcing of embodied work
+--- driving, cleaning, for instance ---
+has become a focus of on--demand labor markets
+\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+Today, on--demand work promises to become a \joke{yuge industry}
+\joke{[citation needed]}.
 
-In the years since,
-scholars have generated taxonomies for the work done by many distributed workers
-in an attempt to better categorize and reason about the many forms of work done 
-on information work platforms such as AMT, oDesk, etc\dots
-\cite{yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
-We add that,
-under \citeauthor{howe2008crowdsourcing}'s constraints,
-even \textit{more} new forms of work fall squarely under
-the metaphorical umbrella we collectively call ``crowdsourcing''.
+% \topic{This matters because\dots WHYYYYYYYYYYYYYYY\\}
+% Indeed, this on--demand workforce has sparked interest
+% across industries ranging
+% driving for hire (for example Uber),
+% house--cleaning (Handy),
+% and generalized services (TaskRabbit)
+% \cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+% Today, a rapidly growing transient workforce is forming,
+% itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
+% to benefit from a latent pool of previously vetted workers
+% \cite{pewSharing}.
 
-\topic{This matters because\dots}
-Indeed, this on--demand workforce has sparked interest
-across industries ranging
-driving for hire (for example Uber),
-house--cleaning (Handy),
-and generalized services (TaskRabbit)
-\cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-Today, a rapidly growing transient workforce is forming,
-itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
-to benefit from a latent pool of previously vetted workers
-\cite{pewSharing}.
-
+\topic{For all the growth we've observed in this labor market,
+we have also seen a complicated and conflicted culture emerge among its constituent workers.}
 Researchers have made efforts to understand the people
@@
 \cite{Ross,whoareNOTtheTurkers}.
-Some of this research has been motivated by the realization of the sociality of gig work,
+Some of this research has been motivated by the identification of the sociality of gig work,
 and the frustration and disenfranchisement that these systems embody
 \cite{turkopticon,dynamo}.
-Other work has focused on the outcomes of work,
+Other work has focused on the \textit{outcomes} of this frustration,
 reflecting on the resistance workers express against digitally mediated labor markets
@@
 
-The extant body of work has ostensibly sought to answer one underlying question:
-What does the future hold for work and those that do it? % expand on ...?
-Researchers have offered their input on this open question along three major threads of scholarship:
+\topic{The extant body of work has ostensibly sought to answer one underlying question:
+What does the future hold for work and those that do it?}
+Researchers have offered their input on this open question along three major threads:
 \begin{enumerate}
-  \item What are the limits of crowdsourcing?
-  Perhaps more tightly constrained,
-  what can and cannot be done by crowd workers?
+  \item What are the limits of crowdsourcing
         \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,
-          Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
+              Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}?
+        \note{Perhaps to be more precise, we ask two questions here:
+                \begin{enumerate}
+                  \item how far will crowdsourcing reach into the everyday lives of people, and
+                  \item how complex can crowdwork become?
+                \end{enumerate}}
   \item What forms of work design, and worker management and arrangement, are viable?
@@
   \printbibliography
-  % \theendnotes
   \nobalance{}
@@
   \section{Sleeping kittens}
-\begin{itemize}
-  \item on--demand labor has consumed an increasing proportion of the labor market in the past ten years [\textit{Pew research on on--demand labor}];
-  the work  has ranged from Turk work processes
-  \item some of it has been about resisting Uber \& other gig labor platforms \cite{bernsteinSoylent,uberAlgorithm};
-  some researchers have tackled understanding this new instantiation of work, but come up short (\textit{how}).
-
-  \item (what binds this together)
-  \item we offer a framing on this topic that situates the research so far on a timeline of the maturation of another form of work that emerged approximately 150 years ago
-  --- that is, piecework --- which convincingly suggests that on--demand labor in the form of Turking, driving for Uber, etc\dots are in fact little more than piecework re--surging today.
-
-  \item this paper will trace the body of research describing microwork --- and later gig work --- and place it in the context of historical piecework and the industrial revolution as a whole.
-  \item informed by this new lens, we'll then turn our attention to the future of this re--emergent form of work, and suggest ways that researchers and users of on--demand labor might influence the outcomes we predict in this paper.
-\end{itemize}
+  \theendnotes
   \end{appendices}

### 2016-08-29 drafts/labor/trackingWork.tex
friday changes committed sunday

@@
 \usepackage[pdflang={en-US},pdftex]{hyperref}
+
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
@@
 \newcommand{\note}[1]{\footnote{#1}}
-\newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences
-\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, bold the \topic sentences
-
+\newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
+\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!
 \begin{document}
 \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
-\renewcommand{\joke}[1]{} % don't do anything with \jokes! Just throw them away!
+\renewcommand{\joke}[1]{} % don't do anything with \jokes! Just throw them away! (Better to drop it than accidentally include it)
 \renewcommand{\onlyinsubfile}[1]{}
@@
 \renewcommand{\note}[1]{\onlyinsubfile{\note{#1}}}
+
 \title{\plaintitle}

### 2016-08-29 drafts/labor/cases.tex
slicing and flexibility spin-off

@@
 \subfile{complexity}
-
-\subsubsection{Slicing work smaller
-\joke{ --- or ``putting the {\tiny `micro'} in microwork''}
-}\label{sec:Slicing}
-
-Decomposition allows requesters to assign parallelized tasks
-without undue concern for the broader context of the work.
-The earlier example of transcribed audio is admittedly a carefully curated one
---- used to illustrate the ability
-to parallelize work across as many workers as there are tasks ---
-but even in this case problems arise;
-breaking an hour--long interview into
-3,600 1--second--long tasks would seem to make it impossible to do the tasks
-for lack of \textit{context},
-but the other extreme (one hour--long block)
-offers no performance improvement over conventionally sourced work.
-
-The question, then, becomes about how to break tasks down as much as possible
-(as discussed in the previous section)
-without losing the necessary context to do a task.
-Stated another way, how can we design work so that as little context as possible is provided,
-without leaving out context or information necessary to complete the task?
-As \citeauthor{verroios2014context} frame it,
-``the crowd must be able to act with
-global understanding when each contributor only has access to local views''
-\cite{verroios2014context}.
-
-While some research has attempted to increase workers' awareness of the work they do in
-apparent efforts to yield higher quality work,
-others have turned this constraint of crowdwork into a feature ---
-the abstraction from the work itself
-appears to allow requesters to engage workers in work with ---
-for instance ---
-sensitive, confidential or otherwise personally identifying information.
-We will discuss these and other avenues of research here.
-
-\textit{Cascade} demonstrated that it's possible to
-break certain classes of tasks apart
-in such a way that they yield taxonomies of various subjects,
-a task generally thought to be sufficiently complex that only expert workers
-with top--down awareness of the task
---- in this case, with awareness of all the constituent colors ---
-could complete the task
-\cite{chilton2013cascade}.
-\citeauthor{verroios2014context} further illustrate this potential by
-forming a task one might consider highly contextually dependent
---- summarizing the contents of a movie ---
-in such a way that crowd workers could contribute small pieces of work without
-needing to know the content of the rest of the project
-\cite{verroios2014context}.
+\subfile{slicing}
 

### 2016-08-29 drafts/labor/complexity.tex
slicing and flexibility spin-off

@@
 % crowd weaver, aditya and someone else's book
+
+\subsubsubsection{Crowdwork's perspective.}
 \topic{Crowdsourcing research has spent the better part of a decade attempting
@@
 
+\subsubsubsection{Piecework's perspective.}
 \topic{The research into piecework makes the case that
@@
 
+\subsubsubsection{What's changed.}
 \topic{Digital media have expanded the scope of viable piecework by pushing drastically on the limits
@@
 
-\topic{\joke{great} topic sentence here.}
-% \topic{While much of the above work has focused on
-% clever arrangements of systems to manage collectives of workers,
-% another body of work has turned inward toward managing the self.}
+\note{\topic{\joke{great} topic sentence here.}
 \citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
@@
 (of sorts)
-will prove relevant as we trace the literature surrounding piecework.
+will prove relevant as we trace the literature surrounding piecework.}
 
@@
 \onlyinsubfile{
-  % \balance{}
+  \balance{}
   \printbibliography
-  \clearpage
-  \nobalance{}
-  \begin{appendices}
-  \theendnotes
-  \end{appendices}
 }

### 2016-08-29 drafts/labor/flexibility.tex
slicing and flexibility spin-off

--- 2016-08-29 drafts/labor/cases.tex
@@
 \begin{document}
-\section{Major Research Questions}
 
-We look at lots of papers that use the term ``crowdsourcing'' or ``crowdwork''
-in the abstracts and titles of their papers,
-especially the papers that we cited earlier, to try
-to answer the questions that we posed in the introduction.
-Then we look to the piecework literature to see whether and to what extent
-piecework answers the questions we, crowdsourcing researchers, have asked.
-We then see whether and how crowdsourcing as we know it has differed from piecework,
-and how that affects the predictions and conclusions made in the piecework literature.
-% look at review papers
-
-% what's the methodology?
-% make piecework literature much clearer about argument
-% make headers clearer to signal what i'm doing here.
-
-\subsection{What are the limits of crowdsourcing?}
-\topic{Research in crowdsourcing has spent the better part of a decade
-exploring how to grow the limits of crowdsourcing and
-find the boundaries of crowd work and microtasks.}
-This has largely involved
-  identifying challenges to this form of labor,
-  overcoming them through novel designs of work--flows and processes, and
-  repeating the process
-\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
-The question that has emerged
-among these researchers and
-through the work that they have produced then
-has been driving at \textit{whether} there are limits to crowdsourcing
-(and, if so, what factors determine those limits).
-Through this lens, we can point to
-a number of contributions to the field that have extended the boundaries of crowd work.
-
-The exploration of crowdsourcing's potential and limits has principally looked at
-manipulating and extending along three dimensions:
-\begin{inlinelist}
-  \item \nameref{sec:Complexity},
-  \item \nameref{sec:Slicing}, and
-  \item \nameref{sec:flexibility}
-\end{inlinelist}.
-We'll explore these aspects of crowdsourcing,
-discussing the extents to which work can be
-decomposed,
-contextually abstracted, and
-made more resilient to attrition of various forms.
-We'll also point to corresponding piecework literature addressing these aspects.
-Finally, we'll discuss how these elements will serve
-to constrain the upper and lower bounds of crowdsourcing as it relates
-to the question of the furthest limits of crowdsourcing.
-
-\subfile{complexity}
-\subfile{slicing}
-
-%   __ _           _ _     _ _ _ _         
-%  / _| | _____  _(_) |__ (_) (_) |_ _   _ 
-% | |_| |/ _ \ \/ / | '_ \| | | | __| | | |
-% |  _| |  __/>  <| | |_) | | | | |_| |_| |
-% |_| |_|\___/_/\_\_|_.__/|_|_|_|\__|\__, |
-%                                    |___/
-
-% how much volatility and quality can we deal with?
-% how to measure, get rid of people, recover from bad workers
-%mason's on getting workers to be better
 \subsubsection{Volatile quality and availability? How much can you deal with?}\label{sec:flexibility}
@@
 
-Here, \citeauthor{hu1961parallel}'s work,
-saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
-parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work
-\cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,
-and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
-substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,
-and persists to this day as it colors
-researchers' goals and objectives in the study and design of crowd work.
-
-Piecework's influence on the abstraction of work into tasks,
-described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,
-prompting work requesters to consider ways to design these now discrete tasks in ways that
-maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
-\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
-It's to this area that we now turn our attention.
-
-%                        _ _      _ 
-%  _ __   __ _ _ __ __ _| | | ___| |
-% | '_ \ / _` | '__/ _` | | |/ _ \ |
-% | |_) | (_| | | | (_| | | |  __/ |
-% | .__/ \__,_|_|  \__,_|_|_|\___|_|
-% |_|
-
-Piecework has seen work along this dimension spanning decades;
-\citeauthor{thompson1913time} investigate some of the ways that
-construction can benefit from the principles of scientific management.
-\citeauthor{thompson1913time}'s thesis asserts that
-task work is predicated on the accurate scientific management of work,
-including the ``miscellaneous tasks''.
-\citeauthor{thompson1913time} argues
---- as early as \citeyear{thompson1913time} ---
-that ``\dots one may be challenged to find any class of work
-involving labor either indoors or out--of--doors
-where tasks cannot be fixed by proper time--study''
-\cite{thompson1913time}.
-
-Broken down in this way, work could grow to unprecedented scales,
-but the quality of the work would remain relatively variable
-\cite{murray1983decentralisation}.
-Textile work being a salient example,
-it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work
-\cite{vezina1992light}.
-
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
-manageable routines that could be taught relatively easily,
-and whose work output could be evaluated in abstraction from the rest of the work
-\cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-but a single very narrowly defined action that needed to be done on every car
-\cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,
-``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
-\cite{schoenberger1988fordism}.
-
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),
-can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,
-facilitating evaluation and parallelization
-\cite{writingMicroTasks}.
-By decomposing and recomposing tasks,
-and in particular by assigning similarly natured work to the same workers,
-workers could become ``experts'' in a small aspect of the work that they did,
-speeding their work dramatically
-\cite{delayAndOrderLasecki}.
-Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage
-\cite{rogstadius2011assessment}.
-
-\subsubsection{So how does this affect crowdwork?}
-The work we've seen so far 
-\begin{itemize}
-  \item worst case: assembling iPhones (extant)
-  \item average case: railroad workers and assembly lines
-  \item high (complexity) case: 
-\end{itemize}
-
-% so CW?
-% mechanism/limitation in pw?
-% differences in cw? tech/access to intelligence?
-% predicted outcome then? -> decompose effort/expertise access
-% possible quirks/diffs?
-
-% what can be decomposed? what can be evaluated?
-% technology/turuing completeness gets us a win
-% predefined expertise? you can find someone almost instantly;
-% ease of doing this makes it qualitatively different due to Channel Factors (diverting a channel/river)
-% huddler paper cites some decomposition stuff workflows are good/bad at things
-% 3-way slack w/msb and mav
-% academy of management 
-% van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
-
 \onlyinsubfile{
-  \balance{}
+  % \balance{}
   \printbibliography
-  % \nobalance{}
-  \clearpage
-  \nobalance{}
-  \begin{appendices}
-\section{Graveyard}
-
-\endnotetext{
-    \citeauthor{cheng2015break} found that microtasks
-    --- though not necessarily \textit{faster} than ``macrotasks'' ---
-    yield higher quality work,
-    particularly when that work is susceptible to frequent interruptions
-    \cite{cheng2015break}.
-}
-
-\subsection{What forms of work design and worker management are viable?}
-\begin{itemize}
-  \item researchers have looked at how to increase worker productivity
-  (e.g. finding the maximal speed at which gig workers can be expected to work before making errors)
-  \cite{measuringCrowdsourcingCheng}.
-  \item we've also seen people ``embrace error''
-  \cite{embracingErrorKrishna}.
-  \item still other research has looked into ways to sandbox workers from the context of their work 
-  \item but scholarship looking into the design and management of work and workers isn't new;
-  lots of research into getting pieceworkers to do work more quickly
-  \cite{seymour1954manual}.
-  \item Researchers have even asked the age old question of \textit{what motivates} pieceworkers
-  (echoing similar research on Wikipedia and Mechanical Turk)
-  \cite{roy1953work,Nov:2007:MW:1297797.1297798,kaufmann2011more}
-\end{itemize}
-
-\subsection{What will work and the place of work look like for workers?}
-The metaphorical mechanics of these dynamics are still at play;
-workers and managers continue to interact in adversarial manners,
-despite substantive work into aligning the motivations of workers and requesters
-
-    % \section{Case Studies}
-    The existing body of research has shed light on on--demand labor from various perspectives,
-    and revealed a number of topics that,
-    through our framing, are clearly situated together.
-    Those topics are, at a high level, as follows:
-    \begin{enumerate}
-    \item the \textbf{processes} involved in making work into tasks, or discretization;
-    \item the outcomes (and indeed the \textbf{fallout}) of that discretization,
-    both on the work itself as well as the workers;
-    and finally
-    \item the \textbf{relationships} between workers and requesters of the work
-    --- both \textit{cooperative} and \textit{adversarial} cases.
-    \end{enumerate}
-
-    \subsection{The Fallout of Crowd Work}\label{sec:Fallout}
-    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
-    % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
-    \citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
-    (see also \citeauthor{dynamo}'s work
-    continuing in the spirit of this observation to generate collective action to improve worker conditions)
-    \cite{turkopticon,dynamo}.
-    \citeauthor{uberAlgorithm}
-    find similarly that workers on gig work platforms are frustrated by the systems on which they work,
-    to say little of the policies which these systems enforce
-    \cite{uberAlgorithm}.
-
-    We discussed the benefits of flexibility
-    (both in the sense of having arbitrary workers perform tasks and
-    in the sense that we can design tasks to be more resilient to poor work)
-    in the previous section.
-    It's from that point in the literature that we turn our attention to
-    the perhaps unintended effects of crowd work
-    and the affordances for transience that we build into this mode of work.
-    We'll address two major areas of work under this subject:
-    \begin{inlinelist}
-    \item \nameref{sec:lowPay}; and
-    \item \nameref{sec:varQualWork}.
-    \end{inlinelist}
-
-    \subsubsection{Low Pay}\label{sec:lowPay}
-
-    \citeauthor{laborEconomicsOfCrowdsourcingHorton}
-    identified problems with crowd work wages relatively early on,
-    attempting to address this imbalance from a behavioral economic perspective ---
-    that is, identifying and presenting a model that describes a worker's
-    ``\textit{reservation wage}''
-    \cite{laborEconomicsOfCrowdsourcingHorton}.
-    This work has largely informed much of the research into and practice of estimating crowd work compensation
-    \cite{incentivesShaw,paolacci2010running}.
-
-    But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}'',
-    a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
-    \cite{turkopticon}.
-    Of particular relevance here,
-    \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
-    wage theft and
-    pay rates that translate to well under minimum wage.
-    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},
-    we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
-    was \$1.38, while the mean was \$3.63
-    \cite{laborEconomicsOfCrowdsourcingHorton}.
-
-    Understanding workers' motivations given these conditions has thus become a goal for some researchers
-    \cite{whyWouldAnyoneBrewer}.
-    \citeauthor{Sun20111033} conclude that
-    ``\dots~solvers participate in online tasks
-    not only for money
-    but also for enjoyment
-    or the sense of self--worth''
-    \cite{Sun20111033}.
-    This might have rung true in \citeyear{Sun20111033},
-    and certainly corroborates \citeauthor{Ross}'s findings after investigating
-    ``who are the crowdworkers'',
-    but as \citeauthor{whoareNOTtheTurkers} points out
-    ``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
-    \cite{Ross,whoareNOTtheTurkers}.
-
-    Now, \citeauthor{turkopticon}
-    and later
-    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
-    via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
-    \cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
-
-    On--demand workers were not the first to be exploited along the dimension of low pay rates.
-    Frustration over low (and declining) pay was one of the chief grievances among then nascent
-    British labor unions in the early \nth{20} century
-    \cite{turner1952trade}.
-    This, \citeauthor{ebbinghaus1999institutions} argued,
-    fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
-    (to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
-    \cite{ebbinghaus1999institutions,levi2009union}.
-    This realization has similarly fueled a body of research into
-    the various incentive structures available to piecework employers
-    \cite{roy1953work}.
-
-    The parallels between the complaints of low pay among crowd workers and other on--demand workers
-    and the pieceworkers and later factory workers in the \nth{20} century
-    are inescapable.
-    We argue further that the \textit{causes} here
-    --- work decomposition,
-    work abstraction, and
-    flexibility ---
-    lead inexorably to low and declining pay for workers.
-    Moreover, we point out that low pay leads to other negative outcomes both
-    in on--demand work
-    as well as
-    in piecework and on assembly lines.
-
-    \subsubsection{Variable quality work}\label{sec:varQualWork}
-    Researchers have struggled with what we might generously call work of ``variable quality''
-    along two dimensions.
-    The first, to use the characterization of one of these contributions, we can call
-    ``understanding malicious behavior''
-    \cite{MaliciousCrowdworkersGadiraju}.
-    While some work has cast workers as ``malicious'' or at least adversarial parties,
-    the evidence thus far suggests that
-    workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
-    (a topic of discussion to which we'll return later)
-    \cite{uberAlgorithm}.
-    The second dimension of research in this space generally attempts
-    to eke out the highest quality work possible from workers
-    given the apparent difficulty in predicting work outcomes
-    \cite{embracingErrorKrishna}.
-
-    The effect low wages have had on piece work and factory workers is well--known;
-    \citeauthor{gantt1913work} discuss this exact mechanism in his book on
-    % \citetitle{gantt1913work}, pointing out that
-    ``\dots~where there is no union,
-    the class wage is practically gauged by the wages the poor workman will accept,
-    and the good workman soon becomes discouraged and \textit{sets his pace by that of his less efficient neighbor},
-    with the result that the general tone of the shop is lowered'' (emphasis added)
-    \cite{gantt1913work}.
-
-    This research is similar to, but subtly different from, the notion of the ``market for `lemons'''
-    which \citeauthor{fort2011amazon} discuss;
-    specifically, \citeauthor{akerlof1970market}'s writing of a ``market for `lemons'''
-    describes a marketplace where the quality of the product or service is unknown to the buyer
-    \cite{fort2011amazon,akerlof1970market}.
-    The effect of this \textit{perceived} uncertainty is that
-    the \textit{actual} trustworthiness drops precipitously
-    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,
-    leaving only the ones who cannot or will not establish their trustworthiness.
-
-    \subsection{Relationships Between Workers and Managers}\label{sec:relationships}
-    Suffice it to say that poor pay and poor work are linked,
-    and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-    But the poor treatment of workers by managers
-    --- both human and algorithmic ---
-    do more than affect the economic relationships between workers and employers.
-    Here, then, we turn to examine this facet of on--demand work
-    and how these dynamics strikingly replicate the relationships
-    researchers in labor advocacy encountered in the study of piecework and factory work.
-
-    This topic can be condensed into two major areas:
-    \begin{inlinelist}
-    \item external (scientific) management, and the evaluation of workers as functional modules; and
-    % \item the alienation of workers under this form of evaluation; and
-    \item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
-    \end{inlinelist}
-
-    \subsubsection{External Management}
-    We discussed Fordism and Taylorism earlier in our discussions of
-    \nameref{sec:Complexity} and \nameref{sec:flexibility},
-    but here the core of these paradigmatic views
-    --- the scientific management of work ---
-    becomes relevant.
-    We use ``external'' here instead of ``scientific'', however,
-    to more broadly capture the disconnect between managers and workers.
-    By describing it as thus,
-    we can touch on the relationship that workers have with \textit{researchers}, as well,
-    even though that work is not strictly
-    --- or just not exclusively ---
-    of the same nature as the management and experience as when interacting with requesters.
-
-    First, intuitively, the variable--quality work we discussed previously has led to
-    a large and growing body of research attempting to evaluate workers' performance and error rates
-    across numerous dimensions;
-    for example, \citeauthor{measuringCrowdsourcingCheng} explore the error rates of workers by
-    operating on a sliding scale giving workers varying amounts of time to accomplish micro--tasks
-    \cite{measuringCrowdsourcingCheng}.
-    \citeauthor{storiesIraniSilberman} describe the treatment of workers
-    as sorts of ``human APIs'' that can, importantly, be rigorously evaluated
-    \cite{storiesIraniSilberman}.
-    \citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
-    cognitively demanding tasks on workers,
-    informing crowdsourcing research by suggesting the use of cognitive load assessments such as
-    NASA Task Load Index surveys to evaluate workers pre and post--tasks
-    \cite{embracingErrorKrishna,measuringCrowdsourcingCheng}.
-
-    External management comes in other forms than scientific, as previously mentioned.
-    Researchers in particular have noticed that their relationships with on--demand workers are,
-    at the least, complex.
-    \citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
-    specifically, their interactions with field sites in which they work
-    as designers and mediators of change influence the relationships they have with Turkers
-    \cite{storiesIraniSilberman}.
-
-    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
-    and even physiological study of pieceworkers closely resembles the
-    research into cognitive loads and stress levels that we discussed among on--demand crowd workers
-    \cite{hu1961parallel,pieceworkBiologicalHarm}.
-    Even the complicated relationships between observers and workers themselves are not necessarily new;
-    \citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
-    to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
-    \cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.
-
-    Similarly, \citeauthor{pollard1963factory}'s words on the punishment factory workers faced ---
-    for example, that ``unsatisfactory work was punished \dots~by fines or by dismissal'' ---
-    seems especially relevant given the fears we now know to be ubiquitous on platforms such as AMT, Uber, and other on--demand markets
-    \cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.
-
-    \subsubsection{Resistance}
-    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,
-    often without recourse or opportunity for feedback, let alone substantive input.
-    Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
-    many toggle their availability to avoid being dispatched to more distant locations,
-    resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
-    \cite{uberAlgorithm}.
-
-    Resistance has sometimes been more coordinated, as well;
-    we see this in \citeauthor{turkopticon}'s coverage on \textit{Turkopticon} as workers collectively accumulated information about requesters, and
-    in \citeauthor{dynamo}'s work on \textit{Dynamo}, which generated ``Guidelines for Academic Requesters'' written by crowd workers
-    \cite{turkopticon,dynamo}.
-
-    Resistance against managers in piecework and factory labor settings are deeply well--explored,
-    but perhaps the most relevant case study to draw on here is to be found in
-    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',
-    where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
-    \cite{waldinger1996helots}.
-    The achievements of labor advocacy groups such as labor unions as resistant,
-    even adversarial organizations counter--balancing the management
-    is somewhat well--understood
-    \cite{russell1982collective,craig1992behavior}.
-    We argue that these threads of resistance against management in various forms
-    are in fact one.
-
-  \theendnotes
-  \end{appendices}
 }

### 2016-08-29 drafts/labor/slicing.tex
slicing and flexibility spin-off

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\subsubsection{Slicing work smaller
\joke{ --- or ``putting the {\tiny `micro'} in microwork''}
}\label{sec:Slicing}

Concurrent with the body of work on the complexity of crowdsourcing
has been a thread of work exploring the decomposition of work.
In this section we'll discuss some of the approaches and the results 
the crowdsourcing community have taken from their work.
Then, as previously,
we'll discuss some of the findings that piecework research
has yielded.
Finally, we'll discuss similarities and differences between crowdwork and piecework,
and how crowdwork's limitations and potential differ from piecework's.

\subsubsubsection{Crowdwork's perspective.}
\topic{The research on crowdwork and decomposition has generally driven toward finding new ways
to strip unnecessary context from work,
allowing tasks to be broken down into smaller parts.}
As \citeauthor{verroios2014context} frame it,
``the crowd must be able to act with
global understanding when each contributor only has access to local views''
\cite{verroios2014context}.
% the decomposition of work allows requesters to assign tasks to workers in parallel,
% theoretically scaling to as many workers as there are tasks.
% As the nature of the work has grown in complexity,
% so too has the challenge of parallelizing and decontextualizing work;
Researchers have thus investigated the role of
distributed, contextually independent work
in the contexts of collaborative writing of various forms,
software development,
classification, and myriad other purposes
\cite{Kittur:2009:CCI:1518701.1518928,Baecker:1993:UID:169059.169312,
      10.2307/1562247,karnin2010crowdsourcing,bragg2013crowdsourcing,
      chilton2013cascade}.

% \topic{i was going to say something here about breaking work apart to the point that nobody can piece things together,
% but i think this would be more interesting as a commentary on trusting workers or something.}
% Given that much of the focus in crowdsourcing has focused on information work,
% \citeauthor{sensitiveTasks}, for instance,
% illustrate one approach to ensuring that workers never become aware of
% enough information to piece together sensitive data
% \cite{sensitiveTasks}.

\topic{One of the emergent properties of micro--tasks has been the relative cost of
\textit{finding} worthwhile tasks.}
The research community has documented and to some extent attempted to intervene in
the discovery of worthwhile tasks
\cite{taskSearch}.
\citeauthor{Cosley:2007:SUI:1216295.1216309}
attempts to address this by
directing workers to tasks through
``intelligent task routing''
\cite{Cosley:2007:SUI:1216295.1216309}.
Much of this work and the work at the periphery of this space, then,
has focused on
minimizing the amount of time that people need to spend doing
anything other than the work for which they are paid.

{
\subsubsubsection{Piecework's perspective.}
My goal for this section is to make this all about the assembly line and
scientific management, and how piecework literature tried to measure everything,
but found it untenable given the extra equipment that was necessary
(but generally which didn't exist) to track every movement and action that workers took.

\subsubsubsection{What's changed.}
The finding here is that
crowdwork can be more carefully micro--managed than piecework could be,
and that this is a double--edged sword:
we can effectively give feedback to workers on everything they do,
but this is emboldening us to try to
over--manage workers just as piecework tried to do.}

% The question, then, becomes about how to break tasks down as much as possible
% (as discussed in the previous section)
% without losing the necessary context to do a task.
% ow can we design work so that as little context as possible is provided,
% without leaving out context or information necessary to complete the task?

% While some research has attempted to increase workers' awareness of the work they do in
% apparent efforts to yield higher quality work,
% others have turned this constraint of crowdwork into a feature ---
% the abstraction from the work itself
% appears to allow requesters to engage workers in work with ---
% for instance ---
% sensitive, confidential or otherwise personally identifying information.
% We will discuss these and other avenues of research here.

% \textit{Cascade} demonstrated that it's possible to
% break certain classes of tasks apart
% in such a way that they yield taxonomies of various subjects,
% a task generally thought to be sufficiently complex that only expert workers
% with top--down awareness of the task
% --- in this case, with awareness of all the constituent colors ---
% could complete the task
% \cite{chilton2013cascade}.
% \citeauthor{verroios2014context} further illustrate this potential by
% forming a task one might consider highly contextually dependent
% --- summarizing the contents of a movie ---
% in such a way that crowd workers could contribute small pieces of work without
% needing to know the content of the rest of the project
% \cite{verroios2014context}.

\onlyinsubfile{
  % \balance{}
  \printbibliography
}

\end{document}

### 2016-09-05 drafts/labor/cases.tex
complexity and some slicing stuff

@@
 \subfile{slicing}
+\subfile{flexibility}
 
-%   __ _           _ _     _ _ _ _         
-%  / _| | _____  _(_) |__ (_) (_) |_ _   _ 
-% | |_| |/ _ \ \/ / | '_ \| | | | __| | | |
-% |  _| |  __/>  <| | |_) | | | | |_| |_| |
-% |_| |_|\___/_/\_\_|_.__/|_|_|_|\__|\__, |
-%                                    |___/
-
-% how much volatility and quality can we deal with?
-% how to measure, get rid of people, recover from bad workers
-%mason's on getting workers to be better
-\subsubsection{Volatile quality and availability? How much can you deal with?}\label{sec:flexibility}
-A number of researchers have identified
-worker attrition,
-variability of worker performance,
-and uncertainty about good versus bad--faith actors as open questions of crowdwork
-\cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
-\ali{We can and should discuss the distinction between presumably ``bad faith'' workers
-\& workers who are merely responding in kind to bad requesters
---- and the broader questions surrounding
-the roles that requesters as well as workers should play ---
-but let it suffice to say that
-requesters have been trying to understand and manage what appears to them as inconsistent work.
-Their ways of responding to that variance in work quality has largely involved
-making the work more flexible and resilient to work
-(although some work has gone into investigating the causes, rather than treating the symptoms)}
-
-Earlier we discussed \citeauthor{cheng2015break}'s work
-measuring the impact that interruption has on worker performance
-\cite{cheng2015break}.
-This work illustrates a broader sentiment in
-both the study and practice
-of crowdwork, that microtasks should be designed resiliently against the variability of workers,
-fully exploiting the abstracted nature of each piece of work
-\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-That is to say, micro--tasks should be designed such that a single worker's poor performance,
-or a good worker's sudden departure,
-does not significantly impact the agenda of the work as a whole.
-While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
-in the form of higher cumulative time to complete
-(albeit much shorter real time to complete, owing to parallelization),
-\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
-similar tasks together
-\cite[respectively]{cheng2015break,delayAndOrderLasecki}.
-
-\citeauthor{embracingErrorKrishna} take a different approach;
-by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds,
-the authors offer orders--of--magnitude improvements
-in various binary classification tasks
-\cite{embracingErrorKrishna}.
-And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-some researchers have designed work predicated on the expectation of it instead:
-\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
-crowd workers would never be given enough information to piece together sensitive information about
-any single topic
-\cite{sensitiveTasks}.
-
-The work thus far seems to attempt to maximize the quality of work among workers through various means:
-\begin{inlinelist}
-  \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju},
-  \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
-  \item Expecting certain levels of attrition and incorrectness and using that variability to their advantage \cite{embracingErrorKrishna}
-\end{inlinelist}.
-
-Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
-\citeauthor{tolliday1986between}'s treatment describing
-turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
-Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
-[and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
-piece rates~\dots~rather than a day wage system''
-\cite{tolliday1986between}.
-
-In the field of piecework,
-the research covering this topic has both explored
-a breadth of tasks that might be rendered doable by piecemeal workers
-\textit{as well as} longitudinally documented the success of these approaches.
-Here, we
-
+\endnotetext{
 Here, \citeauthor{hu1961parallel}'s work,
@@
 % van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
+}
 
@@
   \begin{appendices}
-\section{Graveyard}
+% \section{Graveyard}
 
@@
 }
-
+\endnotetext{
 \subsection{What forms of work design and worker management are viable?}
@@
     are in fact one.
-
+}
   \theendnotes

### 2016-09-05 drafts/labor/complexity.tex
complexity and some slicing stuff

--- 2016-08-26 drafts/labor/complexity.tex
@@
 
-\subsubsection{Achieving greater complexity
+\subsubsection[finding crowdwork's limits]{Achieving greater complexity
 \joke{ --- or ``making crowd--work \textit{yuge}''}
-% i think it says a lot about my confidence in my latex skillz that i'm willing to put these in the subfile
 }\label{sec:Complexity}
-% crowd weaver, aditya and someone else's book
+% crowd weaver, aditya's book
+
+\subsubsubsection{\crowdworkpers}
 \topic{Crowdsourcing research has spent the better part of a decade attempting
@@
 
+\subsubsubsection{\pieceworkpers}
 \topic{The research into piecework makes the case that
@@
 
-\topic{Digital media have expanded the scope of viable piecework by pushing drastically on the limits
+\subsubsubsection{\whatchanged}
+\topic{Digital media have expanded the scope of
+viable piecework by pushing drastically on the limits
 cited by piecework researchers.}
@@
 what --- if anything --- has changed.
-We argue that the Internet has trivialized the costs and challenges of the earlier limiting factors for two reasons:
+We argue that
+the Internet has trivialized
+the costs and challenges of the earlier limiting factors for two reasons:
 \begin{inlinelist}
   \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
-  \item The Internet allows us to leverage the benefits of ``economies of scale'' at very little cost
+  \item The Internet allows us to leverage the benefits of
+        ``economies of scale'' at very little cost
         to the system--designer \cite{lessig2006code}
@@
 
+\topic{Technology has made it possible
+for non--experts to do work that was once considered
+within the domain of experts.}
+\citeauthor{yuanAlmost} builds on the work of others
+(\textit{Voyant} and, more relevantly, \textit{CrowdCrit})
+to design workflows that yield ``expert--level feedback''
+\cite{yuanAlmost,Xu:2014:VGS:2531602.2531604,Luther:2014:CCA:2556420.2556788}.
+This body of work identifies ways to transform a variety of duties comprising complex tasks
+and distills them into ``a narrow set of routinized duties'',
+informed in part by researchers --- acting as inspectors --- and experts
+\cite[quotations from][]{10.2307/23702539}
+Where \citeauthor{10.2307/23702539} would call additionally for the identification of
+crowdsourcing's version of ``piecework clerks'', we point out that
+today algorithms manage workers as pieceworkers once did
+\cite{uberAlgorithm,10.2307/23702539}.
+
+\topic{Technology more directly facilitates the subversion of expertise requirements
+by giving non--experts access to information that would otherwise be unavailable.}
+Taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge''
+--- a demonstration of the driver's comprehensive familiarity.
+Researchers have identified significant growth of
+the hippocampal regions of the brains in veteran drivers,
+generally understood to be responsible for spatial functions such as navigation
+\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
+      skok2000managing,Woollett1407,woollett2011acquiring}.
+Services such as Google Maps \& Waze make it possible for
+people entirely unfamiliar with a city
+to know more about a city even than experts through
+the collective data generated by other users
+ranging topics such as police activity, congestion, construction, etc\dots
+\cite{silva2013traffic,hind2014outsmarting}.
+
+\topic{Crowdsourcing falters when the routinization of complex work proves difficult.}
+\ali{Something about complexity being difficult to make routine when
+there are lots of little tasks a la \citeauthor{Brown01041990}}
+
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+
+  \clearpage
+  \nobalance{}
+  \begin{appendices}
+  \section{Paragraph graveyard}
     \topic{\joke{great} topic sentence here.}
-% \topic{While much of the above work has focused on
-% clever arrangements of systems to manage collectives of workers,
-% another body of work has turned inward toward managing the self.}
     \citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
@@
 
-\endnotetext{
     This is also kind of off-topic, isn't it? where to take it?
@@
     \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
-  --- working for increasing amounts of time on tasks of growing complexity, only to discover that requesters are not willing to pay,
+    --- working for increasing amounts of time on tasks of growing complexity,
+    only to discover that requesters are not willing to pay,
     for instance ---
-  but these and other dangers range an enormous landscape % too informal?
+    but these and other dangers range an enormous landscape
     \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
-
-  kill this?
 
@@
   \cite{10.2307/23702539,dynamo}.
+  \end{appendices}
 
@@
 
-\onlyinsubfile{
-  % \balance{}
-  \printbibliography
-  \clearpage
-  \nobalance{}
-  \begin{appendices}
-  \theendnotes
-  \end{appendices}
-}
-
 \end{document}

### 2016-09-05 drafts/labor/flexibility.tex
complexity and some slicing stuff

@@
 
-\subsubsection{Volatile quality and availability? How much can you deal with?}\label{sec:flexibility}
+\subsubsection[worker volatility]{Volatile quality and availability? How much can you deal with?}
+\label{sec:flexibility}
+
+\ali{TODO:
+\begin{enumerate}
+  \item segue
+  \item \crowdworkpers
+  \item \pieceworkpers
+  \item \whatchanged
+\end{enumerate}
+}
+
 A number of researchers have identified
@@
   % \balance{}
-  \printbibliography
+  % \printbibliography
 }

### 2016-09-05 drafts/labor/slicing.tex
complexity and some slicing stuff

@@
 
-\subsubsection{Slicing work smaller
+\subsubsection[slicing work into smaller parts]{Slicing work smaller
 \joke{ --- or ``putting the {\tiny `micro'} in microwork''}
@@
 
-\subsubsubsection{Crowdwork's perspective.}
+\subsubsubsection{\crowdworkpers}
 \topic{The research on crowdwork and decomposition has generally driven toward finding new ways
@@
 \cite{verroios2014context}.
-% the decomposition of work allows requesters to assign tasks to workers in parallel,
-% theoretically scaling to as many workers as there are tasks.
-% As the nature of the work has grown in complexity,
-% so too has the challenge of parallelizing and decontextualizing work;
 Researchers have thus investigated the role of
@@
       chilton2013cascade}.
-
-% \topic{i was going to say something here about breaking work apart to the point that nobody can piece things together,
-% but i think this would be more interesting as a commentary on trusting workers or something.}
-% Given that much of the focus in crowdsourcing has focused on information work,
-% \citeauthor{sensitiveTasks}, for instance,
-% illustrate one approach to ensuring that workers never become aware of
-% enough information to piece together sensitive data
-% \cite{sensitiveTasks}.
 
@@
 
-{
-\subsubsubsection{Piecework's perspective.}
-My goal for this section is to make this all about the assembly line and
-scientific management, and how piecework literature tried to measure everything,
-but found it untenable given the extra equipment that was necessary
-(but generally which didn't exist) to track every movement and action that workers took.
+\subsubsubsection{\pieceworkpers}
+The beginnings of systematized task decomposition stretch back as far as the \nth{17} century,
+when Airy employed young boys at the Greenwich Observatory who
+``possessed the basic skills of mathematics, including
+`Arithmetic, the use of Logarithms, and Elementary Algebra'~''
+to \textit{compute} astronomical phenomena
+\cite{grier2013computers}.
+Airy's tasks were unique at the time for several reasons:
+\begin{enumerate}
+  \item each task was quickly verifiable by a qualified [human] computer,
+  \item tasks were discrete --- independent from one another, and
+  \item knowledge of the full scope of the project --- indeed, knowledge of anything more than the problem set at hand ---
+  was wholly unnecessary.
+\end{enumerate}
 
-\subsubsubsection{What's changed.}
+\subsubsubsection{\whatchanged}
+\ali{todo}
+
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+
+  \clearpage
+  \nobalance{}
+  \begin{appendices}
+    \section{Paragraph graveyard}
     The finding here is that
@@
     but this is emboldening us to try to
-over--manage workers just as piecework tried to do.}
+    over--manage workers just as piecework tried to do.
 
-% The question, then, becomes about how to break tasks down as much as possible
-% (as discussed in the previous section)
-% without losing the necessary context to do a task.
-% ow can we design work so that as little context as possible is provided,
-% without leaving out context or information necessary to complete the task?
-
-% While some research has attempted to increase workers' awareness of the work they do in
-% apparent efforts to yield higher quality work,
-% others have turned this constraint of crowdwork into a feature ---
-% the abstraction from the work itself
-% appears to allow requesters to engage workers in work with ---
-% for instance ---
-% sensitive, confidential or otherwise personally identifying information.
-% We will discuss these and other avenues of research here.
-
-% \textit{Cascade} demonstrated that it's possible to
-% break certain classes of tasks apart
-% in such a way that they yield taxonomies of various subjects,
-% a task generally thought to be sufficiently complex that only expert workers
-% with top--down awareness of the task
-% --- in this case, with awareness of all the constituent colors ---
-% could complete the task
-% \cite{chilton2013cascade}.
-% \citeauthor{verroios2014context} further illustrate this potential by
-% forming a task one might consider highly contextually dependent
-% --- summarizing the contents of a movie ---
-% in such a way that crowd workers could contribute small pieces of work without
-% needing to know the content of the rest of the project
-% \cite{verroios2014context}.
-
-\onlyinsubfile{
-  % \balance{}
-  \printbibliography
+    My goal for this section is to make two points:
+    \begin{enumerate}
+      \item show how this is related to the assembly line and scientific management,
+            and how piecework literature tried to measure everything,
+            but found it untenable given the extra equipment that was necessary
+            (but generally which didn't exist)
+            to track every movement and action that workers took.
+      \item show how this work was enabled by the ``verifiability'' of work output(?)
+    \end{enumerate}
+  \end{appendices}
 }

### 2016-09-06 drafts/labor/slicing.tex
more changes

@@
 has been a thread of work exploring the decomposition of work.
-In this section we'll discuss some of the approaches and the results 
-the crowdsourcing community have taken from their work.
-Then, as previously,
-we'll discuss some of the findings that piecework research
-has yielded.
+In this section we'll discuss some of the work that
+the crowdsourcing community have made toward making tasks more finite and discreet.
+Then, as in the previous section,
+we'll discuss some of the insights that piecework research has offered.
 Finally, we'll discuss similarities and differences between crowdwork and piecework,
@@
 \subsubsubsection{\crowdworkpers}
-\topic{The research on crowdwork and decomposition has generally driven toward finding new ways
-to strip unnecessary context from work,
-allowing tasks to be broken down into smaller parts.}
-As \citeauthor{verroios2014context} frame it,
+\topic{The crowdsourcing research into slicing work into smaller parts
+has largely centered around removing extraneous context from the worker's view.}
+
+While \citeauthor{Kinnaird:2012:WTM:2389176.2389219}
+advocated that
+greater context of the task would foster more reliably high--quality work
+(especially, according to their findings, when applied to more veteran crowd workers)
+% \cite{Kinnaird:2012:WTM:2389176.2389219}.
+% Nevertheless,
+the consensus that emerged
+seems best--represented by \citeauthor{verroios2014context}'s framing of
+the tension still present in the design of crowd work, that
 ``the crowd must be able to act with
 global understanding when each contributor only has access to local views''
-\cite{verroios2014context}.
-Researchers have thus investigated the role of
-distributed, contextually independent work
-in the contexts of collaborative writing of various forms,
-software development,
-classification, and myriad other purposes
-\cite{Kittur:2009:CCI:1518701.1518928,Baecker:1993:UID:169059.169312,
-      10.2307/1562247,karnin2010crowdsourcing,bragg2013crowdsourcing,
-      chilton2013cascade}.
+\cite{Kinnaird:2012:WTM:2389176.2389219,verroios2014context}.
+While \citeauthor{Kinnaird:2012:WTM:2389176.2389219} showed that
+it's possible to achieve higher quality work by
+providing people with more information about the work they're doing,
+an alternative approach to dealing with crowd workers
+--- one which treats micro task workers as
+``modular, protocol--defined computational services'' ---
+ultimately prevailed
+\cite{Kinnaird:2012:WTM:2389176.2389219,irani2015cultural}.
+
+% Researchers have thus investigated the role of
+% distributed, contextually independent work
+% in the contexts of collaborative writing of various forms,
+% software development,
+% classification, and myriad other purposes
+% \cite{Kittur:2009:CCI:1518701.1518928,Baecker:1993:UID:169059.169312,
+%       10.2307/1562247,karnin2010crowdsourcing,bragg2013crowdsourcing,
+%       chilton2013cascade}.
 
@@
 \begin{enumerate}
-  \item each task was quickly verifiable by a qualified [human] computer,
-  \item tasks were discrete --- independent from one another, and
-  \item knowledge of the full scope of the project --- indeed, knowledge of anything more than the problem set at hand ---
-  was wholly unnecessary.
+  \item Each task was quickly verifiable by a qualified (human) computer;
+  \item tasks were discrete --- that is, independent from one another; and
+  \item knowledge of the full scope of the project
+  --- indeed, knowledge of anything more than the problem set at hand ---
+  was unnecessary.
 \end{enumerate}

### 2016-09-06 drafts/labor/complexity.tex
various additions to slicing.tex; still adding to piecework perspective

@@
 we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity.
+the nature of the work is limited in complexity
+\cite{Brown01041990}.
 Given the flourishing of on--demand labor platforms such as
-Uber, Amazon Mechanical Turk, and others, we ask ourselves
+Uber, AMT, and others, we ask ourselves
 what --- if anything --- has changed.
@@
         ``economies of scale'' at very little cost
-        to the system--designer \cite{lessig2006code}
+        to the system--designer \cite{lessig2006code,miller2011understanding}
 \end{inlinelist}.
@@
 \ali{Something about complexity being difficult to make routine when
-there are lots of little tasks a la \citeauthor{Brown01041990}}
+there are lots of little, varied, unpredictable tasks (nod to \citeauthor{Brown01041990})}
 

### 2016-09-06 drafts/labor/introduction.tex
slicing and flexibility spin-off
various additions to slicing.tex; still adding to piecework perspective

@@
 \begin{enumerate}
-  \item What are the limits of crowdsourcing
-        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,
-              Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}?
-        \note{Perhaps to be more precise, we ask two questions here:
-                \begin{enumerate}
-                  \item how far will crowdsourcing reach into the everyday lives of people, and
-                  \item how complex can crowdwork become?
-                \end{enumerate}}
-  \item What forms of work design, and worker management and arrangement, are viable?
-        \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,
+  \item What are the limits of crowdsourcing?
+        Specifically,
+        \begin{inlinelist}
+          \item how complex can crowd work get? and
+          \item how far will crowd work reach into the everyday lives of people?
+        \end{inlinelist}
+        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
+              Nebeling:2016:WCW:2858036.2858169,
+              Hahn:2016:KAB:2858036.2858364};
+  \item To what extent can we decompose work (especially into micro--tasks)?
+        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
+              LykourentzouPersonalityMatters,KucherbaevReLauncher,
               Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
@@
   \item What will work and the place of work look like for the workers?
-        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,
-              takingAHITMcInnis}
+        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
+              whyWouldAnyoneBrewer,takingAHITMcInnis}
 \end{enumerate}
@@
   \nobalance{}
-  \clearpage
-  \begin{appendices}
-  \section{Sleeping kittens}
-  \theendnotes
-  \end{appendices}
+  % \clearpage
+  % \begin{appendices}
+  % \section{Sleeping kittens}
+  % \end{appendices}
 

### 2016-09-06 drafts/labor/slicing.tex
various additions to slicing.tex; still adding to piecework perspective

--- 2016-09-05 drafts/labor/slicing.tex
@@
 
-Concurrent with the body of work on the complexity of crowdsourcing
-has been a thread of work exploring the decomposition of work.
-In this section we'll discuss some of the approaches and the results 
-the crowdsourcing community have taken from their work.
-Then, as previously,
-we'll discuss some of the findings that piecework research
-has yielded.
-Finally, we'll discuss similarities and differences between crowdwork and piecework,
-and how crowdwork's limitations and potential differ from piecework's.
-
 \subsubsubsection{\crowdworkpers}
-\topic{The research on crowdwork and decomposition has generally driven toward finding new ways
-to strip unnecessary context from work,
-allowing tasks to be broken down into smaller parts.}
-As \citeauthor{verroios2014context} frame it,
-``the crowd must be able to act with
+% \ali{Does this belong here at all? The complexity argument just dives in.}
+\topic{The crowdsourcing research into work decomposition
+has largely focused on minimizing the additional context necessary to do tasks,
+and making it easier to do tasks with less time.}
+This first thread is
+perhaps best described by \citeauthor{verroios2014context} as
+making crowd workers ``\dots able to act with
 global understanding when each contributor only has access to local views''
 \cite{verroios2014context}.
-Researchers have thus investigated the role of
-distributed, contextually independent work
-in the contexts of collaborative writing of various forms,
-software development,
-classification, and myriad other purposes
-\cite{Kittur:2009:CCI:1518701.1518928,Baecker:1993:UID:169059.169312,
-      10.2307/1562247,karnin2010crowdsourcing,bragg2013crowdsourcing,
-      chilton2013cascade}.
+With the exception of a few cases
+(specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
+which finds that greater work context fosters more reliably high--quality work),
+the micro task paradigm has emerged as the overwhelming favorite
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016,
+      cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
 
-\topic{One of the emergent properties of micro--tasks has been the relative cost of
-\textit{finding} worthwhile tasks.}
-The research community has documented and to some extent attempted to intervene in
-the discovery of worthwhile tasks
-\cite{taskSearch}.
-\citeauthor{Cosley:2007:SUI:1216295.1216309}
-attempts to address this by
-directing workers to tasks through
-``intelligent task routing''
-\cite{Cosley:2007:SUI:1216295.1216309}.
-Much of this work and the work at the periphery of this space, then,
-has focused on
-minimizing the amount of time that people need to spend doing
-anything other than the work for which they are paid.
+\topic{As the additional context necessary to complete a task diminishes,
+the marginal cost of finding and
+\textit{doing} tasks has increasingly become the focus of research.}
+\citeauthor{taskSearch} illustrate the challenges on AMT,
+and some work has gone into ameliorating the problems specific to this work site
+(\textit{ReLauncher}), % does anyone care for project names?
+while other work designs tasks around gap time
+(\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning})
+\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,
+      Cai:2015:WLW:2702123.2702267}.
+Yet more work looks at the general framing of tasks,
+chaining and arranging them to maximally exploit
+the attention and stress threshold % ???
+of workers
+\cite{Cai:2016:CRI:2858036.2858237}.
+Rather than attempt to minimize the error rates in micro--tasks,
+as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested,
+we as a community have leaned \textit{into} the peril of
+low--context work,
+``embracing error'' in crowdsourcing
+\cite{embracingErrorKrishna}.
+
+Not all of the work toward optimizing crowd work--flows has gone toward
+minimizing the creative input of crowd workers;
+a thriving body of literature adopts
+practices such as pipelining to allow experts to participate in crowd work
+\cite{foundry}.
 
@@
     \end{enumerate}
+
+    As a result, the prevalent mindset of designing work for crowd workers
+    --- one which treats micro task workers as
+    ``modular, protocol--defined computational services'' ---
+    has inexorably alienated workers from the greater context of their work
+    \cite{irani2015cultural}.
+
+    \ali{i don't like this section \textit{here}, but i like it in general. what do?}
+    \topic{One of the emergent properties of micro--tasks has been the relative cost of
+    \textit{finding} worthwhile tasks.}
+    The research community has documented and to some extent attempted to intervene in
+    the discovery of worthwhile tasks
+    \cite{taskSearch}.
+    \citeauthor{Cosley:2007:SUI:1216295.1216309}
+    attempts to address this by
+    directing workers to tasks through
+    ``intelligent task routing''
+    \cite{Cosley:2007:SUI:1216295.1216309}.
+    Much of this work and the work at the periphery of this space, then,
+    has focused on
+    minimizing the amount of time that people need to spend doing
+    anything other than the work for which they are paid.
+
+    What we take away from this and the previous set of work is that
+    the value of adopting crowdsourcing for any particular task
+    seems to be mediated by two questions:
+    \begin{inlinelist}
+      \item How long does it take to train workers to do the work in question? and
+      \item How long does it take for the worker to do the work?
+    \end{inlinelist}
+    Minimizing these criteria has become
+    the overarching motivation of the crowdsourcing work design community
+    \cite{cheng2015break,Newell:2016:OMA:2858036.2858490}
+
   \end{appendices}

### 2016-09-08 drafts/labor/flexibility.tex
'closing' slicing.tex for now, moving on to flexibility

@@
 \begin{enumerate}
-  \item segue
   \item \crowdworkpers
@@
 
+\onlyinsubfile{
+  % \balance{}
+  \printbibliography
+  \clearpage
+  % \nobalance{}
+
+  \begin{appendices}
   A number of researchers have identified
@@
 
-\onlyinsubfile{
-  % \balance{}
-  % \printbibliography
+\end{appendices}
 }

### 2016-09-08 drafts/labor/slicing.tex
'closing' slicing.tex for now, moving on to flexibility

@@
 \subsubsubsection{\pieceworkpers}
-The beginnings of systematized task decomposition stretch back as far as the \nth{17} century,
+\topic{The research community relating to
+piecework and labor
+has been wrestling with the decomposition of work for centuries.}
+The beginnings of
+systematic task decomposition
+stretch back as far as the \nth{17} century,
 when Airy employed young boys at the Greenwich Observatory who
@@
 `Arithmetic, the use of Logarithms, and Elementary Algebra'~''
-to \textit{compute} astronomical phenomena
+to compute, by hand, astronomical phenomena
 \cite{grier2013computers}.
-Airy's tasks were unique at the time for several reasons:
-\begin{enumerate}
-  \item Each task was quickly verifiable by a qualified (human) computer;
-  \item tasks were discrete --- that is, independent from one another; and
-  \item knowledge of the full scope of the project
+These workers became the first \textit{computers}.
+
+The work Airy solicited was interesting for several reasons.
+First, work output was quickly verifiable;
+Airy could assign variably skilled workers to compute values,
+and have other workers check their work.
+\ali{I could point out that the opportunity to check work and
+repeat the task is a little like find--fix--verify, but is that jumping the gun?}
+Second, tasks were discrete --- that is, independent from one another.
+Finally, knowledge of the full scope of the project
 --- indeed, knowledge of anything more than the problem set at hand ---
 was unnecessary.
-\end{enumerate}
+
+The insight of breaking tasks down into smaller components didn't find its audience until
+the early \nth{20} century,
+with the rise of Fordism and scientific management (or Taylorism).
+From scientific management, we found that
+we could measure work at unprecedented resolution and precision.
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers, but certainly
+in Ford and Taylor's time --- and certainly in Airy's time ---
+highly instrumented, automatic measurement of workers was all but impossible.
+As a result,
+the distillation of work into smaller chunks
+ultimately reached a limit of usefulness.
+
+\ali{\citeauthor{marx2012economic} in here? Alienated from the context of the work?
+Critiques about piecework marginalizing workers?
+\citeauthor{american1921problem} wrote about this as well
+(albeit they were advocating from workers' perspectives, but who isn't?)}
 
 \subsubsubsection{\whatchanged}
-\ali{todo}
+\begin{inlinelist}
+  \item Computers make it possible to switch from one task to another
+  unlike any arbitrary manufacturing factory possibly could;
+  \item we've sliced works to such small sizes that the marginal costs
+  --- things like task--finding, cognitive load switching, etc\dots ---
+  have become relatively large;
+  \item instrumentation has become so advanced that
+  the curve of diminishing value on measuring and tracking workers has shifted significantly
+  (but not been obliterated);
+  \ali{some companies have suggested self--tracking through
+  programs that give workers fitbits and whatnot --- I could make the argument that
+  this is just an illustration of that, but
+  it's not really about work \textit{per se} unless
+  you think about it as the general management of workers. Thoughts?}
+\end{inlinelist}
 
 \onlyinsubfile{
-  \balance{}
+  % \balance{}
   \printbibliography

### 2016-09-08 drafts/theoryPaper/proceedings.tex
various intro revs
'closing' slicing.tex for now, moving on to flexibility

@@
 \usepackage[utf8]{inputenc}
-\def\plaintitle{Laying Groundwork For a Worker--Centric Peer Economy}
-\def\plainauthor{Ali Alkhatib, Justin Cranshaw, Andrés Monroy-Hernandez}
+\def\plaintitle{Here's a title!}
+\def\plainauthor{People here!}
 \def\emptyauthor{Anonymous for Review}
-\def\plainkeywords{Peer economy; Collaborative markets}
-\def\plaingeneralterms{Documentation, Standardization}
+\def\plainkeywords{???????????}
+\def\plaingeneralterms{????}
 
@@
 
-\numberofauthors{4}
+\numberofauthors{1}
 \author{
-  \alignauthor{Ali Alkhatib                 \affaddr{Stanford University}   \affaddr{Stanford, USA}
-        \email{ali.alkhatib@cs.stanford.edu}}
-  \alignauthor{Justin Cranshaw              \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
-        \email{justincr@microsoft.com}}
-  \alignauthor{Andrés Monroy-Hernandez      \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
-        \email{amh@microsoft.com}}
-  \alignauthor{Michael Bernstein      \affaddr{Stanford University}    \affaddr{Stanford, USA}
-        \email{msb@cs.stanford.edu}}
+    \alignauthor{Kenji Hata, Ranjay Krishna, Li Fei--Fei, Michael S. Bernstein}\\
+    \affaddr{Stanford University}\\
+    \email{\{kenjihata, ranjaykrishna, feifeili, msb\}@cs.stanford.edu}
   }
-\numberofauthors{1} \author{ \alignauthor{\emptyauthor} }
 \maketitle
-
-\begin{abstract}
-The ``gig economy''
---- where work is measured through discrete tasks rather than through interconnected (and often long--running) careers ---
-has become a site of significant frustration among workers who have come to rely on these systems and this form of work.
-This paper explores some of the sources of that stress,
-and attempts to relate this seemingly nascent work style to more familiar forms of labor
---- namely, ``piece work'',
-a form of work which emerged at the dawn of the 20th century.
-While we find and identify parallels between ``gig work'' and the more theoretically familiar ``piece work'',
-we emphasize divergences which make the conflation of these ideas problematic.
-
-We further explore opportunities and tactics to empower workers,
-informed by ethnographic fieldwork including participant--observation,
-semi--structured interviews,
-and participatory design.
-Our observation and interviews lead us to conclude that the most intuitive mechanisms for empowering workers
---- namely, labor unions advocating collectively and counter--balancing the manager's outsize leverage ---
-may both structurally and culturally fail to capture the needs of ``gig workers''.
-We argue that design interventions to empower ``gig workers'' must build on the theoretically familiar,
-much as ``gig work'' itself draws inspiration from ``piece work'',
-but that it is crucial that interventions are designed to support,
-rather than reject,
-the unique cultural and system qualities of gig workers.
-\end{abstract}
-
-\category{H.5.m.}{Information Interfaces and Presentation
-  (e.g. HCI)}{Miscellaneous} \category{See
-  \url{http://acm.org/about/class/1998/} for the full list of ACM
-  classifiers. This section is required.}{}{}
-
-\keywords{\plainkeywords}
-
-\section*{Dealing with Monopsonies}
-\section*{Collective Bargaining}
-
-% \input{intro.tex}
-% \input{background.tex}
-% \input{fieldwork_advice.tex}
-% \input{design_guidance.tex}
-% \input{discussion.tex}
-
-% % REFERENCES FORMAT
-% % References must be the same font size as other body text.
-\bibliographystyle{SIGCHI-Reference-Format}
-\bibliography{../references}
-
 \end{document}
 
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:
+% \begin{abstract}
+% The ``gig economy''
+% --- where work is measured through discrete tasks rather than through interconnected (and often long--running) careers ---
+% has become a site of significant frustration among workers who have come to rely on these systems and this form of work.
+% This paper explores some of the sources of that stress,
+% and attempts to relate this seemingly nascent work style to more familiar forms of labor
+% --- namely, ``piece work'',
+% a form of work which emerged at the dawn of the 20th century.
+% While we find and identify parallels between ``gig work'' and the more theoretically familiar ``piece work'',
+% we emphasize divergences which make the conflation of these ideas problematic.
+
+% We further explore opportunities and tactics to empower workers,
+% informed by ethnographic fieldwork including participant--observation,
+% semi--structured interviews,
+% and participatory design.
+% Our observation and interviews lead us to conclude that the most intuitive mechanisms for empowering workers
+% --- namely, labor unions advocating collectively and counter--balancing the manager's outsize leverage ---
+% may both structurally and culturally fail to capture the needs of ``gig workers''.
+% We argue that design interventions to empower ``gig workers'' must build on the theoretically familiar,
+% much as ``gig work'' itself draws inspiration from ``piece work'',
+% but that it is crucial that interventions are designed to support,
+% rather than reject,
+% the unique cultural and system qualities of gig workers.
+% \end{abstract}
+
+% \category{H.5.m.}{Information Interfaces and Presentation
+%   (e.g. HCI)}{Miscellaneous} \category{See
+%   \url{http://acm.org/about/class/1998/} for the full list of ACM
+%   classifiers. This section is required.}{}{}
+
+% \keywords{\plainkeywords}
+
+% \section*{Dealing with Monopsonies}
+% \section*{Collective Bargaining}
+
+% % \input{intro.tex}
+% % \input{background.tex}
+% % \input{fieldwork_advice.tex}
+% % \input{design_guidance.tex}
+% % \input{discussion.tex}
+
+% % % REFERENCES FORMAT
+% % % References must be the same font size as other body text.
+% \bibliographystyle{SIGCHI-Reference-Format}
+% \bibliography{references}
+
+% \end{document}
+
+% %%% Local Variables:
+% %%% mode: latex
+% %%% TeX-master: t
+% %%% End:

### 2016-09-12 drafts/labor/abstract.tex
renamed files, moved stuff around, adding implications sections

\documentclass[trackingWork]{subfiles}

\makeatletter
\def\blx@maxline{77}
\makeatother
% \zexternaldocument*{complexity}

\begin{document}

\begin{abstract}
Networked computation is enabling the rise of
crowdwork,
gig work, and
other forms of on--demand labor.
A large and growing body of scholarship has sought to predict
the socio--technical outcomes of this shift, especially
\nameref{sec:complexity} how complex and interdependent crowdwork's tasks can be,
\nameref{sec:decomposition} how thinly crowdwork can be sliced and modularized, and
\nameref{sec:relationships} what the collective outcomes will be for crowdworkers.
In this paper, we look to the historical scholarship on piecework
--- a strikingly similar trend of
work decomposition,
distribution, and
payment that was popular at the turn of the \nth{20} century ---
to understand how these questions might play out with modern crowdwork.
To do so, we identify the mechanisms that limited piecework historically, and
identify whether crowdwork faces the same mechanism limits or might differentiate itself.

% We consider the aforementioned open questions in the body of research on crowdwork
% under the following framework:
% \begin{inlinelist}
% \item we look to the familiar literature in human computation;
%       then
% \item we investigate the scholarship on piecework;
% \item we identify what (if anything) has changed in crowdwork to differentiate it from piecework;
%       and finally
% \item we attempt to provide a concise prediction on these open questions,
%       informed by piecework's insights.
% \end{inlinelist}

% With growing attention toward on--demand labor
% --- ranging from the ``sharing economy'' to information work ---
% scholars have made connections to various frameworks 
% and mechanisms such as worker advocacy, empowerment, and Taylorism,
% to make sense of our observations of
% on--demand work and the workers that power this movement.
% We argue that the literature surrounding ``piecework''
% informs and even predicts both the contributions
% that have been made toward the development of on--demand labor and crowd work
% as well as the fallout among workers and researchers
% with regard to the disillusionment and alienation of work.

% After evaluating this framing
% through a series of case studies,
% we look to the future
% to identify worthwhile questions and
% points of inquiry, such as the movement toward factories,
% that researchers in social computing should consider
% as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

% \onlyinsubfile{
% \clearpage
% \subfile{cases}
% }

\end{document}

### 2016-09-12 drafts/labor/cases.tex
renamed files, moved stuff around, adding implications sections

@@
 \begin{inlinelist}
-  \item \nameref{sec:Complexity},
-  \item \nameref{sec:Slicing}, and
-  \item \nameref{sec:flexibility}
+  \item \nameref{sec:complexity},
+  \item \nameref{sec:decomposition}, and
+  \item \nameref{sec:relationships}
 \end{inlinelist}.
@@
   \nobalance{}
-  \begin{appendices}
+  }
+  % \begin{appendices}
 % \section{Graveyard}
-
 \endnotetext{
+\section{Cases notes}
     \citeauthor{cheng2015break} found that microtasks
@@
     \cite{cheng2015break}.
-}
-\endnotetext{
+  % }
+  % \endnotetext{
   \subsection{What forms of work design and worker management are viable?}
@@
 
-    \subsection{The Fallout of Crowd Work}\label{sec:Fallout}
+    \subsection{The Fallout of Crowd Work}
     % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
@@
 
-    \subsubsection{Low Pay}\label{sec:lowPay}
+    \subsubsection{Low Pay}
 
@@
 
-    \subsubsection{Variable quality work}\label{sec:varQualWork}
+    \subsubsection{Variable quality work}
     Researchers have struggled with what we might generously call work of ``variable quality''
@@
 
-    \subsection{Relationships Between Workers and Managers}\label{sec:relationships}
+    \subsection{Relationships Between Workers and Managers}
     Suffice it to say that poor pay and poor work are linked,
@@
   }
-  \theendnotes
-  \end{appendices}
-  }
+  % \theendnotes
+  % \end{appendices}
 

### 2016-09-12 drafts/labor/complexity.tex
renamed files, moved stuff around, adding implications sections

@@
 
-\subsubsection[finding crowdwork's limits]{Achieving greater complexity
-\joke{ --- or ``making crowd--work \textit{yuge}''}
-}\label{sec:Complexity}
-% crowd weaver, aditya's book
+\subsubsection[finding crowdwork's limits]{Achieving greater complexity}\label{sec:complexity}
+\ali{maybe a subssubsubsusbsubsection on the result/implication?
+more people can make crowd tasks, but the algiorithmic manag}
 
@@
   \nobalance{}
-  \begin{appendices}
-  \section{Paragraph graveyard}
+  }
+  \endnotetext{
+  \section{Complexity notes}
     \topic{\joke{great} topic sentence here.}
@@
   \cite{10.2307/23702539,dynamo}.
-  \end{appendices}
-
 }

### 2016-09-12 drafts/labor/decomposition.tex
renamed files, moved stuff around, adding implications sections

--- 2016-09-08 drafts/labor/slicing.tex
@@
 
-\subsubsection[slicing work into smaller parts]{Slicing work smaller
-\joke{ --- or ``putting the {\tiny `micro'} in microwork''}
-}\label{sec:Slicing}
+\subsubsection[the decomposition of work]{The Decomposition of Work}\label{sec:decomposition}
 
@@
 \onlyinsubfile{
-  % \balance{}
+  \balance{}
   \printbibliography
 
-  \clearpage
-  \nobalance{}
-  \begin{appendices}
-    \section{Paragraph graveyard}
+  % \clearpage
+  % \nobalance{}
+  }
+  \endnotetext{
+    \section{Slicing notes}
     The finding here is that
@@
 
-  \end{appendices}
   }

### 2016-09-12 drafts/labor/piecework_lit.tex
renamed files, moved stuff around, adding implications sections

\documentclass[trackingWork]{subfiles}

\makeatletter
\makeatother
\begin{document}
\section{A Review of Piecework}
\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowdwork
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the subsequent sections --- and indeed, the entire argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise of piecework at a very high level,
        identifying key figures and ideas during this time; and finally
  \item look at the fall of piecework, such as it was,
        considering in particular
        the factors that may have led to piecework's eventual demise
\end{inlinelist}.

\topic{While ``piecework'' has proven difficult to concretize from the literature,
we can trace a constellation of characteristics of piece work that recur throughout the literature.}
We'll follow the history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

\topic{One of the earliest definitions of piecework,
in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
\citeauthor{hughRaynbirdTaskWork} offers what appears to be
one of the first concise definitions of piecework,
which he variously also calls ``measure work'', ``grate work'', and ``task work'':
``\dots the chief difference lies between the day--labourer,
who receives a certain some of money\dots~for his day's work,
and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''
\cite{hughRaynbirdTaskWork}.

\citeauthor{hughRaynbirdTaskWork} makes several arguments for the merits of piece work,
pointing out that\dots
``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
he knows that all depends on his own diligence and perseverance\dots~[and]
so long as he performs his work to the satisfaction of his master,
he is not under that control to which the day--labourer is always subject.''

\citeauthor{10.2307/2338394} gives a more concise definition of piecework
--- ``payments for results'' --- and further illustrates the concept with examples:
``\dots~payment is made for each hectare which is pronounced to be well ploughed~\dots~
for each living foal got from a mare;~\dots~
for each living calf got~\dots'' etc\dots
\cite{10.2307/2338394}.

\ali{
todo:
\begin{enumerate}
  \item piecework section
  \item what is it
  \item historical arc
  \item why do i care? how is it relevant
\end{enumerate}
}

\onlyinsubfile{
  % \balance{}
  \printbibliography
  \clearpage
  % \nobalance{}
  }

\end{document}

### 2016-09-12 drafts/labor/relationships.tex
renamed files, moved stuff around, adding implications sections

--- 2016-09-08 drafts/labor/flexibility.tex
@@
 
-\subsubsection[worker volatility]{Volatile quality and availability? How much can you deal with?}
-\label{sec:flexibility}
+\subsubsection[the relationships of workers]{???}
+\label{sec:relationships}
 
@@
   % \nobalance{}
-
-  \begin{appendices}
+  }
+  \endnotetext{
+    \section{Flexibility notes}
     A number of researchers have identified
@@
     \textit{as well as} longitudinally documented the success of these approaches.
-  Here, we
-
-\end{appendices}
+    Here, we \ali{\dots?}
 }
-
+\theendnotes
 \end{document}

### 2016-09-12 drafts/labor/trackingWork.tex
slicing and flexibility spin-off
complexity and some slicing stuff
more changes
renamed files, moved stuff around, adding implications sections

@@
 \documentclass[chi_draft]{sigchi}
+
+\usepackage{zref-xr}
+\usepackage{zref-xr}
+\zxrsetup{%
+  tozreflabel=false, % not needed, since `zref` is not used otherwise
+  toltxlabel=true, % classical LaTeX labels that can be referenced with \ref, \pageref
+}
+\zexternaldocument{cases.tex}% star form: only LaTeX labels are imported
 
@@
 
+\newcommand{\subsubsubsection}[1]{\textit{#1}}
+\newcommand{\pieceworkpers}{Piecework's perspective.}
+\newcommand{\crowdworkpers}{Crowdwork's perspective.}
+\newcommand{\whatchanged}{What's changed.}
+\newcommand{\implication}{Implications for crowdwork.}
+
 \newcommand{\onlyinsubfile}[1]{#1}
@@
 \newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!
+\onlyinsubfile{
+  \hypersetup{
+    citecolor  = blue,
+    linkcolor  = blue,
+  }
+}
+
+\renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
 \begin{document}
@@
 \maketitle
-\begin{abstract}
-  With growing attention toward on--demand labor
-  --- ranging from the ``sharing economy'' to information work ---
-  scholars have made connections to various frameworks 
-  and mechanisms such as worker advocacy, empowerment, and Taylorism,
-  to make sense of our observations of
-  on--demand work and the workers that power this movement.
-  We argue that the literature surrounding ``piecework''
-  informs and even predicts both the contributions
-  that have been made toward the development of on--demand labor and crowd work
-  as well as the fallout among workers and researchers
-  with regard to the disillusionment and alienation of work.
 
-  After evaluating this framing
-  through a series of case studies,
-  we look to the future
-  to identify worthwhile questions and
-  points of inquiry, such as the movement toward factories,
-  that researchers in social computing should consider
-  as we attempt to anticipate and perhaps shape the future of work.
-\end{abstract}
+% Abstract time
+\subfile{abstract.tex}
 
@@
 \subfile{introduction.tex}
+\subfile{piecework_lit.tex}
 \subfile{cases.tex}
 
+\balance{}
+\printbibliography
+
+\clearpage
+\nobalance{}
+\endnotetext{
   \section{The Bleak Future of Crowd Work}
@@
   a career in which we want our children to work.
-
-\balance{}
-\printbibliography
+}
+\theendnotes
 

### 2016-09-12 drafts/labor/cases.tex
piecework_lit should be kind of done

@@
 We look at lots of papers that use the term ``crowdsourcing'' or ``crowdwork''
-in the abstracts and titles of their papers,
-especially the papers that we cited earlier, to try
+in the abstracts and titles of their papers%
+, especially the papers that we cited earlier, to try
 to answer the questions that we posed in the introduction.
@@
 piecework answers the questions we, crowdsourcing researchers, have asked.
-We then see whether and how crowdsourcing as we know it has differed from piecework,
-and how that affects the predictions and conclusions made in the piecework literature.
+We then see whether and how crowdsourcing as we know it has differed from piecework%
+, and how that affects the predictions and conclusions made in the piecework literature.
 % look at review papers
@@
 This has largely involved
-  identifying challenges to this form of labor,
-  overcoming them through novel designs of work--flows and processes, and
+  identifying challenges to this form of labor%
+,   overcoming them through novel designs of work--flows and processes, and
   repeating the process
@@
 \begin{inlinelist}
-  \item \nameref{sec:complexity},
-  \item \nameref{sec:decomposition}, and
+  \item \nameref{sec:complexity}%
+,   \item \nameref{sec:decomposition}, and
   \item \nameref{sec:relationships}
 \end{inlinelist}.
-We'll explore these aspects of crowdsourcing,
-discussing the extents to which work can be
-decomposed,
-contextually abstracted, and
+We'll explore these aspects of crowdsourcing%
+, discussing the extents to which work can be
+decomposed%
+, contextually abstracted, and
 made more resilient to attrition of various forms.
@@
 
-\subfile{complexity}
-\subfile{slicing}
-\subfile{flexibility}
+\subfile{complexity.tex}
+\subfile{decomposition.tex}
+\subfile{relationships.tex}
 
 \endnotetext{
-Here, \citeauthor{hu1961parallel}'s work,
-saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
-parallels the approach that dominated early research into crowd work
+Here, \citeauthor{hu1961parallel}'s work%
+, saying of assembly line work that
+``it is assumed that men are of equal ability and every man can do any of the $n$ jobs''%
+, parallels the approach that dominated early research into crowd work
 --- namely, using non--expert crowds for complex work
 \cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,
-and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
-substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,
-and persists to this day as it colors
+This mindset in \citeauthor{hu1961parallel}'s analysis%
+, and indeed the study of factory and mass manufacturing labor through the \nth{20} century%
+, substantively owes its existence to scientific management
+and the rigorous decomposition of work into tasks, discussed earlier%
+, and persists to this day as it colors
 researchers' goals and objectives in the study and design of crowd work.
 
-Piecework's influence on the abstraction of work into tasks,
-described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,
-prompting work requesters to consider ways to design these now discrete tasks in ways that
+Piecework's influence on the abstraction of work into tasks%
+, described above, is more than just caused by the decomposition of work;
+work abstraction itself makes it possible for workers to come and go flexibly%
+, prompting work requesters to consider ways to design these now discrete tasks in ways that
 maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
@@
 \citeauthor{thompson1913time}'s thesis asserts that
-task work is predicated on the accurate scientific management of work,
-including the ``miscellaneous tasks''.
+task work is predicated on the accurate scientific management of work%
+, including the ``miscellaneous tasks''.
 \citeauthor{thompson1913time} argues
@@
 
-Broken down in this way, work could grow to unprecedented scales,
-but the quality of the work would remain relatively variable
+Broken down in this way, work could grow to unprecedented scales%
+, but the quality of the work would remain relatively variable
 \cite{murray1983decentralisation}.
-Textile work being a salient example,
-it took time for workers to acquire sufficient skill
+Textile work being a salient example%
+, it took time for workers to acquire sufficient skill
 to do every aspect of the work so that the garment would be accepted by the company soliciting that work
@@
 
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
-manageable routines that could be taught relatively easily,
-and whose work output could be evaluated in abstraction from the rest of the work
+A compelling solution emerged in the early \nth{20} century to break tasks down into discrete%
+, manageable routines that could be taught relatively easily%
+, and whose work output could be evaluated in abstraction from the rest of the work
 \cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,
-but a single very narrowly defined action that needed to be done on every car
+In Ford's assembly line, this meant that workers were not responsible for building a whole car%
+, but a single very narrowly defined action that needed to be done on every car
 \cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,
-``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
+By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes%
+, ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
 \cite{schoenberger1988fordism}.
 
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),
-can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,
-facilitating evaluation and parallelization
+This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos)%
+, can be seen today in crowd work and on--demand labor through the application of micro--tasks.
+\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces%
+, facilitating evaluation and parallelization
 \cite{writingMicroTasks}.
-By decomposing and recomposing tasks,
-and in particular by assigning similarly natured work to the same workers,
-workers could become ``experts'' in a small aspect of the work that they did,
-speeding their work dramatically
+By decomposing and recomposing tasks%
+, and in particular by assigning similarly natured work to the same workers%
+, workers could become ``experts'' in a small aspect of the work that they did%
+, speeding their work dramatically
 \cite{delayAndOrderLasecki}.
@@
   % \nobalance{}
-  \clearpage
+  % \clearpage
   \nobalance{}
@@
     --- though not necessarily \textit{faster} than ``macrotasks'' ---
-    yield higher quality work,
-    particularly when that work is susceptible to frequent interruptions
+    yield higher quality work%
+,     particularly when that work is susceptible to frequent interruptions
     \cite{cheng2015break}.
@@
   The metaphorical mechanics of these dynamics are still at play;
-  workers and managers continue to interact in adversarial manners,
-  despite substantive work into aligning the motivations of workers and requesters
+  workers and managers continue to interact in adversarial manners%
+,   despite substantive work into aligning the motivations of workers and requesters
 
     % \section{Case Studies}
-    The existing body of research has shed light on on--demand labor from various perspectives,
-    and revealed a number of topics that,
-    through our framing, are clearly situated together.
+    The existing body of research has shed light on on--demand labor from various perspectives%
+,     and revealed a number of topics that%
+,     through our framing, are clearly situated together.
     Those topics are, at a high level, as follows:
@@
     \item the \textbf{processes} involved in making work into tasks, or discretization;
-    \item the outcomes (and indeed the \textbf{fallout}) of that discretization,
-    both on the work itself as well as the workers;
+    \item the outcomes (and indeed the \textbf{fallout}) of that discretization%
+,     both on the work itself as well as the workers;
     and finally
@@
     \subsection{The Fallout of Crowd Work}
-    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
-    % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
+    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging%
+,     % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
     \citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
@@
     \citeauthor{uberAlgorithm}
-    find similarly that workers on gig work platforms are frustrated by the systems on which they work,
-    to say little of the policies which these systems enforce
+    find similarly that workers on gig work platforms are frustrated by the systems on which they work%
+,     to say little of the policies which these systems enforce
     \cite{uberAlgorithm}.
@@
     \citeauthor{laborEconomicsOfCrowdsourcingHorton}
-    identified problems with crowd work wages relatively early on,
-    attempting to address this imbalance from a behavioral economic perspective ---
+    identified problems with crowd work wages relatively early on%
+,     attempting to address this imbalance from a behavioral economic perspective ---
     that is, identifying and presenting a model that describes a worker's
@@
 
-    But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}'',
-    a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
+    But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}''%
+,     a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
     \cite{turkopticon}.
-    Of particular relevance here,
-    \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
+    Of particular relevance here%
+,     \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
     wage theft and
     pay rates that translate to well under minimum wage.
-    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},
-    we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
+    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton}%
+,     we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
     was \$1.38, while the mean was \$3.63
@@
     \cite{Sun20111033}.
-    This might have rung true in \citeyear{Sun20111033},
-    and certainly corroborates \citeauthor{Ross}'s findings after investigating
-    ``who are the crowdworkers'',
-    but as \citeauthor{whoareNOTtheTurkers} points out
+    This might have rung true in \citeyear{Sun20111033}%
+,     and certainly corroborates \citeauthor{Ross}'s findings after investigating
+    ``who are the crowdworkers''%
+,     but as \citeauthor{whoareNOTtheTurkers} points out
     ``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
@@
     and later
-    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
-    via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
+    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers%
+,     via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
     \cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
@@
     \cite{turner1952trade}.
-    This, \citeauthor{ebbinghaus1999institutions} argued,
-    fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
+    This, \citeauthor{ebbinghaus1999institutions} argued%
+,     fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
     (to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
@@
     We argue further that the \textit{causes} here
-    --- work decomposition,
-    work abstraction, and
+    --- work decomposition%
+,     work abstraction, and
     flexibility ---
@@
     \cite{MaliciousCrowdworkersGadiraju}.
-    While some work has cast workers as ``malicious'' or at least adversarial parties,
-    the evidence thus far suggests that
+    While some work has cast workers as ``malicious'' or at least adversarial parties%
+,     the evidence thus far suggests that
     workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
@@
     the \textit{actual} trustworthiness drops precipitously
-    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,
-    leaving only the ones who cannot or will not establish their trustworthiness.
+    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so%
+,     leaving only the ones who cannot or will not establish their trustworthiness.
 
     \subsection{Relationships Between Workers and Managers}
-    Suffice it to say that poor pay and poor work are linked,
-    and that we should not be surprised to find this relationship play out online as strongly as it does offline.
+    Suffice it to say that poor pay and poor work are linked%
+,     and that we should not be surprised to find this relationship play out online as strongly as it does offline.
     But the poor treatment of workers by managers
@@
     We discussed Fordism and Taylorism earlier in our discussions of
-    \nameref{sec:Complexity} and \nameref{sec:flexibility},
-    but here the core of these paradigmatic views
+    \nameref{sec:Complexity} and \nameref{sec:flexibility}%
+,     but here the core of these paradigmatic views
     --- the scientific management of work ---
     becomes relevant.
-    We use ``external'' here instead of ``scientific'', however,
-    to more broadly capture the disconnect between managers and workers.
-    By describing it as thus,
-    we can touch on the relationship that workers have with \textit{researchers}, as well,
-    even though that work is not strictly
+    We use ``external'' here instead of ``scientific'', however%
+,     to more broadly capture the disconnect between managers and workers.
+    By describing it as thus%
+,     we can touch on the relationship that workers have with \textit{researchers}, as well%
+,     even though that work is not strictly
     --- or just not exclusively ---
@@
     \citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
-    cognitively demanding tasks on workers,
-    informing crowdsourcing research by suggesting the use of cognitive load assessments such as
+    cognitively demanding tasks on workers%
+,     informing crowdsourcing research by suggesting the use of cognitive load assessments such as
     NASA Task Load Index surveys to evaluate workers pre and post--tasks
@@
     External management comes in other forms than scientific, as previously mentioned.
-    Researchers in particular have noticed that their relationships with on--demand workers are,
-    at the least, complex.
+    Researchers in particular have noticed that their relationships with on--demand workers are%
+,     at the least, complex.
     \citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
@@
 
-    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
-    and even physiological study of pieceworkers closely resembles the
+    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research%
+,     and even physiological study of pieceworkers closely resembles the
     research into cognitive loads and stress levels that we discussed among on--demand crowd workers
@@
     \subsubsection{Resistance}
-    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,
-    often without recourse or opportunity for feedback, let alone substantive input.
+    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems%
+,     often without recourse or opportunity for feedback, let alone substantive input.
     Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
-    many toggle their availability to avoid being dispatched to more distant locations,
-    resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
+    many toggle their availability to avoid being dispatched to more distant locations%
+,     resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
     \cite{uberAlgorithm}.
@@
 
-    Resistance against managers in piecework and factory labor settings are deeply well--explored,
-    but perhaps the most relevant case study to draw on here is to be found in
-    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',
-    where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
+    Resistance against managers in piecework and factory labor settings are deeply well--explored%
+,     but perhaps the most relevant case study to draw on here is to be found in
+    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors''%
+,     where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
     \cite{waldinger1996helots}.
-    The achievements of labor advocacy groups such as labor unions as resistant,
-    even adversarial organizations counter--balancing the management
+    The achievements of labor advocacy groups such as labor unions as resistant%
+,     even adversarial organizations counter--balancing the management
     is somewhat well--understood

### 2016-09-12 drafts/labor/complexity.tex
piecework_lit should be kind of done

@@
 The broader body of work has varied significantly in type
---- providing conversational assistants,
-interpreting medical data, and
-telling coherent and compelling stories,
-to name a few examples
+--- providing conversational assistants%
+, interpreting medical data, and
+telling coherent and compelling stories%
+, to name a few examples
 \cite{Lasecki:2013:CCC:2501988.2502057,mavandadi2012distributed,KimStoria}.
 
-\topic{This body of research has involved similar approaches to problems,
-often involving insights made in Computer Science and applied to human work--flows.}
+\topic{This body of research has involved similar approaches to problems%
+, often involving insights made in Computer Science and applied to human work--flows.}
 The crowdwork literature typically identifies target milestones in CS
-that have presented significant challenges for researchers,
-leverages some of the approaches and insights that Computer Science researchers have already made
-(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge}),
-and arranges humans as computational black boxes within those approaches and processes
+that have presented significant challenges for researchers%
+, leverages some of the approaches and insights that Computer Science researchers have already made
+(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge})%
+, and arranges humans as computational black boxes within those approaches and processes
 \cite[][and others]{crowdForgeKittur,foundry}.
@@
 \textit{Foundry}'s tools for managing and arranging expert groups into a cohort
-allow researchers to convincingly argue that expert teams can be rapidly formed,
-just like non--expert teams
+allow researchers to convincingly argue that expert teams can be rapidly formed%
+, just like non--expert teams
 \cite{foundry}.
@@
 piecework has been limited principally by the challenges of human management and oversight.}
-\citeauthor{10.2307/23702539},
-who describes piecework as
+\citeauthor{10.2307/23702539}%
+, who describes piecework as
     ``\dots~based on examination of various shop jobs,
-    which included calculation of the standard time and compensation for each task'',
-argues that piecework must be rigorously evaluated at a time that demands
+      which included calculation of the standard time and compensation for each task''%
+, argues that piecework must be rigorously evaluated at a time that demands
 \textit{other people} perform the evaluation.
@@
 This body of work identifies ways to transform a variety of duties comprising complex tasks
-and distills them into ``a narrow set of routinized duties'',
-informed in part by researchers --- acting as inspectors --- and experts
+and distills them into ``a narrow set of routinized duties''%
+, informed in part by researchers --- acting as inspectors --- and experts
 \cite[quotations from][]{10.2307/23702539}
@@
 Researchers have identified significant growth of
-the hippocampal regions of the brains in veteran drivers,
-generally understood to be responsible for spatial functions such as navigation
-\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
-      skok2000managing,Woollett1407,woollett2011acquiring}.
+the hippocampal regions of the brains in veteran drivers%
+, generally understood to be responsible for spatial functions such as navigation
+\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625%
+,       skok2000managing,Woollett1407,woollett2011acquiring}.
 Services such as Google Maps \& Waze make it possible for
@@
   \printbibliography
-
-  \clearpage
+  % \clearpage
   \nobalance{}
   }
+
   \endnotetext{
@@
     \topic{\joke{great} topic sentence here.}
-    \citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work,
-    exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
+    \citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work%
+,     exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
     \cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-    While some of this work doesn't strictly fall under ``crowdsourcing'',
-    the [scientific] management of the self as a worker
+    While some of this work doesn't strictly fall under ``crowdsourcing''%
+,     the [scientific] management of the self as a worker
     (of sorts)
@@
     \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
-    --- working for increasing amounts of time on tasks of growing complexity,
-    only to discover that requesters are not willing to pay,
-    for instance ---
+    --- working for increasing amounts of time on tasks of growing complexity%
+,     only to discover that requesters are not willing to pay%
+,     for instance ---
     but these and other dangers range an enormous landscape
@@
     have attempted to find the characteristics which
-    enable and stymie successful applications of this form of work,
-    and how to get past or around these boundaries.}
+    enable and stymie successful applications of this form of work%
+,     and how to get past or around these boundaries.}
     Whether this work has focused on the self (as in the case of ``selfsourcing'')
-    or on others (as in \textit{CrowdForge} and \textit{Foundry}),
-    crowdsourcing literature has generally studied the constituent work
+    or on others (as in \textit{CrowdForge} and \textit{Foundry})%
+,     crowdsourcing literature has generally studied the constituent work
     and attempted either to work around
@@
   (the other being the resistance of management (for different and varying reasons)).
-  In the cases of online platforms like AMT and digitally mediated platforms like Uber,
-  where in both cases workers rarely
+  In the cases of online platforms like AMT and digitally mediated platforms like Uber%
+,   where in both cases workers rarely
   if ever
   interact face--to--face and
-  opportunities for coordination and collective action are severely limited,
-  we should expect to see
+  opportunities for coordination and collective action are severely limited%
+,   we should expect to see
   the challenges stymieing collective action

### 2016-09-12 drafts/labor/decomposition.tex
piecework_lit should be kind of done

@@
 \topic{The crowdsourcing research into work decomposition
-has largely focused on minimizing the additional context necessary to do tasks,
-and making it easier to do tasks with less time.}
+has largely focused on minimizing the additional context necessary to do tasks%
+, and making it easier to do tasks with less time.}
 This first thread is
@@
 (specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
-which finds that greater work context fosters more reliably high--quality work),
-the micro task paradigm has emerged as the overwhelming favorite
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016,
-      cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
+which finds that greater work context fosters more reliably high--quality work)%
+, the micro task paradigm has emerged as the overwhelming favorite
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016%
+,       cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
 
-\topic{As the additional context necessary to complete a task diminishes,
-the marginal cost of finding and
+\topic{As the additional context necessary to complete a task diminishes%
+, the marginal cost of finding and
 \textit{doing} tasks has increasingly become the focus of research.}
-\citeauthor{taskSearch} illustrate the challenges on AMT,
-and some work has gone into ameliorating the problems specific to this work site
+\citeauthor{taskSearch} illustrate the challenges on AMT%
+, and some work has gone into ameliorating the problems specific to this work site
 (\textit{ReLauncher}), % does anyone care for project names?
@@
 (\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning})
-\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,
-      Cai:2015:WLW:2702123.2702267}.
-Yet more work looks at the general framing of tasks,
-chaining and arranging them to maximally exploit
+\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996%
+,       Cai:2015:WLW:2702123.2702267}.
+Yet more work looks at the general framing of tasks%
+, chaining and arranging them to maximally exploit
 the attention and stress threshold % ???
@@
 \cite{Cai:2016:CRI:2858036.2858237}.
-Rather than attempt to minimize the error rates in micro--tasks,
-as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested,
-we as a community have leaned \textit{into} the peril of
-low--context work,
-``embracing error'' in crowdsourcing
+Rather than attempt to minimize the error rates in micro--tasks%
+, as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested%
+, we as a community have leaned \textit{into} the peril of
+low--context work%
+, ``embracing error'' in crowdsourcing
 \cite{embracingErrorKrishna}.
@@
 systematic task decomposition
-stretch back as far as the \nth{17} century,
-when Airy employed young boys at the Greenwich Observatory who
+stretch back as far as the \nth{17} century%
+, when Airy employed young boys at the Greenwich Observatory who
 ``possessed the basic skills of mathematics, including
@@
 First, work output was quickly verifiable;
-Airy could assign variably skilled workers to compute values,
-and have other workers check their work.
+Airy could assign variably skilled workers to compute values%
+, and have other workers check their work.
 \ali{I could point out that the opportunity to check work and
@@
 The insight of breaking tasks down into smaller components didn't find its audience until
-the early \nth{20} century,
-with the rise of Fordism and scientific management (or Taylorism).
+the early \nth{20} century%
+, with the rise of Fordism and scientific management (or Taylorism).
 From scientific management, we found that
 we could measure work at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out,
-piecework most greatly benefits the instrumented measurement of workers, but certainly
+As \citeauthor{Brown01041990} points out%
+, piecework most greatly benefits the instrumented measurement of workers, but certainly
 in Ford and Taylor's time --- and certainly in Airy's time ---
 highly instrumented, automatic measurement of workers was all but impossible.
-As a result,
-the distillation of work into smaller chunks
+As a result%
+, the distillation of work into smaller chunks
 ultimately reached a limit of usefulness.
@@
     The finding here is that
-    crowdwork can be more carefully micro--managed than piecework could be,
-    and that this is a double--edged sword:
-    we can effectively give feedback to workers on everything they do,
-    but this is emboldening us to try to
+    crowdwork can be more carefully micro--managed than piecework could be%
+,     and that this is a double--edged sword:
+    we can effectively give feedback to workers on everything they do%
+,     but this is emboldening us to try to
     over--manage workers just as piecework tried to do.
@@
     \begin{enumerate}
-      \item show how this is related to the assembly line and scientific management,
-            and how piecework literature tried to measure everything,
-            but found it untenable given the extra equipment that was necessary
+      \item show how this is related to the assembly line and scientific management%
+,             and how piecework literature tried to measure everything%
+,             but found it untenable given the extra equipment that was necessary
             (but generally which didn't exist)
@@
     \cite{Cosley:2007:SUI:1216295.1216309}.
-    Much of this work and the work at the periphery of this space, then,
-    has focused on
+    Much of this work and the work at the periphery of this space, then%
+,     has focused on
     minimizing the amount of time that people need to spend doing

### 2016-09-12 drafts/labor/flexibility.tex
piecework_lit should be kind of done

@@
   \printbibliography
-  \clearpage
+  % \clearpage
   % \nobalance{}
-
-  \begin{appendices}
+  }
+  \endnotetext{
+    \section{Flexibility notes}
     A number of researchers have identified
-  worker attrition,
-  variability of worker performance,
-  and uncertainty about good versus bad--faith actors as open questions of crowdwork
+    worker attrition%
+,     variability of worker performance%
+,     and uncertainty about good versus bad--faith actors as open questions of crowdwork
     \cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
@@
     both the study and practice
-  of crowdwork, that microtasks should be designed resiliently against the variability of workers,
-  fully exploiting the abstracted nature of each piece of work
+    of crowdwork, that microtasks should be designed resiliently against the variability of workers%
+,     fully exploiting the abstracted nature of each piece of work
     \cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-  That is to say, micro--tasks should be designed such that a single worker's poor performance,
-  or a good worker's sudden departure,
-  does not significantly impact the agenda of the work as a whole.
+    That is to say, micro--tasks should be designed such that a single worker's poor performance%
+,     or a good worker's sudden departure%
+,     does not significantly impact the agenda of the work as a whole.
     While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
     in the form of higher cumulative time to complete
-  (albeit much shorter real time to complete, owing to parallelization),
-  \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+    (albeit much shorter real time to complete, owing to parallelization)%
+,     \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
     similar tasks together
@@
     \citeauthor{embracingErrorKrishna} take a different approach;
-  by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds,
-  the authors offer orders--of--magnitude improvements
+    by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds%
+,     the authors offer orders--of--magnitude improvements
     in various binary classification tasks
     \cite{embracingErrorKrishna}.
-  And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-  some researchers have designed work predicated on the expectation of it instead:
+    And rather than building tasks to \textit{tolerate} worker drop--off and attrition%
+,     some researchers have designed work predicated on the expectation of it instead:
     \citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
@@
     \begin{inlinelist}
-    \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju},
-    \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
+      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju}%
+,       \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
       \item Expecting certain levels of attrition and incorrectness and using that variability to their advantage \cite{embracingErrorKrishna}
@@
 
-  In the field of piecework,
-  the research covering this topic has both explored
+    In the field of piecework%
+,     the research covering this topic has both explored
     a breadth of tasks that might be rendered doable by piecemeal workers
     \textit{as well as} longitudinally documented the success of these approaches.
-  Here, we
-
-\end{appendices}
+    Here, we \ali{\dots?}
 }
-
+% \theendnotes
 \end{document}

### 2016-09-12 drafts/labor/piecework_lit.tex
piecework_lit should be kind of done

@@
 \documentclass[trackingWork]{subfiles}
+
+% \renewcommand{\topic}[1]{#1}
 
@@
 \topic{The HCI community has used the term ``piecework'' to describe
-myriad instantiations of on--demand labor,
-but this reference has generally been offered in passing.}
+myriad instantiations of on--demand labor%
+, but this reference has generally been offered in passing.}
 As this paper principally traces a relationship between
 the historical piecework and the contemporary crowdwork
-(or on--demand labor more generally),
-this casual familiarity with piecework may prove insufficient.
+(or on--demand labor more generally)%
+, this casual familiarity with piecework may prove insufficient.
 We'll more carefully discuss piecework in this section in order
@@
   \item define ``piecework'' as researchers in the topic understood it;
-  \item trace the rise of piecework at a very high level,
-        identifying key figures and ideas during this time; and finally
-  \item look at the fall of piecework, such as it was,
-        considering in particular
+  \item trace the rise of piecework at a very high level%
+,         identifying key figures and ideas during this time; and finally
+  \item look at the fall of piecework, such as it was%
+,         considering in particular
         the factors that may have led to piecework's eventual demise
@@
 
-\topic{While ``piecework'' has proven difficult to concretize from the literature,
-we can trace a constellation of characteristics of piece work that recur throughout the literature.}
+\subsection{What was piecework?}
+\topic{While ``piecework'' has proven difficult to concretize from the literature%
+, we can trace a constellation of characteristics of piece work that recur throughout the literature.}
 We'll follow the history of research, collecting
-descriptions,
-examples, and
+descriptions%
+, examples, and
 provided definitions of piecework, trying
@@
 
-\topic{One of the earliest definitions of piecework,
-in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
-\citeauthor{hughRaynbirdTaskWork} offers what appears to be
-one of the first concise definitions of piecework,
-which he variously also calls ``measure work'', ``grate work'', and ``task work'':
+\topic{One of the earliest definitions of piecework%
+, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
+\citeauthor{hughRaynbirdTaskWork} offers
+a concise definition of piecework
+--- which he variously also calls ``measure work'', ``grate work'', and ``task work'' ---
+by contrasting the ``task--labourer'' with the ``day--labourer'':
 ``\dots the chief difference lies between the day--labourer,
@@
 \cite{hughRaynbirdTaskWork}.
+This description offers the first rudimentary definition of piecework
+from which the practice will grow for more than a century;
+piece work, as \citeauthor{hughRaynbirdTaskWork} offers,
+``depend[s] on the quantity of work done''.
 
-\citeauthor{hughRaynbirdTaskWork} makes several arguments for the merits of piece work,
-pointing out that\dots
-``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
-he knows that all depends on his own diligence and perseverance\dots~[and]
-so long as he performs his work to the satisfaction of his master,
-he is not under that control to which the day--labourer is always subject.''
-
-\citeauthor{10.2307/2338394} gives a more concise definition of piecework
---- ``payments for results'' --- and further illustrates the concept with examples:
+\citeauthor{10.2307/2338394} gives a more illustrative definition of piecework%
+, offering examples:
 ``\dots~payment is made for each hectare which is pronounced to be well ploughed~\dots~
@@
 \cite{10.2307/2338394}.
+This framing perhaps makes the most intuitive sense;
+``payment for results'', as \citeauthor{10.2307/2338394} calls it%
+, is not only common in practice, but well--studied in labor economics as well
+\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
 
-\ali{
-todo:
-\begin{enumerate}
-  \item piecework section
-  \item what is it
-  \item historical arc
-  \item why do i care? how is it relevant
-\end{enumerate}
-}
+\topic{It's worth acknowledging that --- as \citeauthor{hart2013rise} point out ---
+``this distinction [between piece--rates and time--rates] was not completely clear--cut''
+\cite{hart2013rise}.}
+The ``Rowan premium system''%
+, which essentially paid workers a base rate for time with the opportunity for additional pay associated with output,
+was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
+muddies the waters for us later as we attempt to categorize cases of piecework
+\cite{rowan1901premium}.
+Nevertheless,
+this work offers us an intuition for what piecework is, if not a bright--line rule.
+
+\topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
+Returning to
+\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piece work%
+crop up; he points out that\dots
+``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
+he knows that all depends on his own diligence and perseverance\dots~[and]
+so long as he performs his work to the satisfaction of his master%
+, he is not under that control to which the day--labourer is always subject.''
+\citeauthor{hughRaynbirdTaskWork} (and others, as we will see)
+highlight the freedom from control that ``task--labourers'' enjoy
+\cite{hughRaynbirdTaskWork,rowan1901premium}.
+
+\topic{We see this sense of independence regardless of the time, locale, and industry.}
+\citeauthor{10.2307/3827491} offers us a look into the lives and culture of ``match girls''
+--- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
+Of particular interest was their reputation ``\dots~for generosity, independence, and protectiveness,
+but also for brashness, irregularity, low morality, and little education''
+\cite{10.2307/3827491}.
+\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia%
+, finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
+``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
+the management recognised his right to put a `substitute' or `grass' compositor in his place''
+\cite{10.2307/27508091}.
+From these accounts we should be able to identify
+a sense of independence that
+resonates across decades, industries, and locales where piecework is found.
+We'll problematize this supposed advantage as we trace the history of piecework%
+, but for now we can say that piecework affords
+independence and some sense of locus
+otherwise unknown to workers.
+
+\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
+In their reflection on the features endemic to piecework in the 1930s,
+which they describe as the ``heyday'' of piecework's prominence;
+among them were the following:
+\begin{inlinelist}
+\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
+\item workers with specific slices of skills could be more appropriately matched to suitable tasks
+\end{inlinelist}
+\cite{hart2013rise}.
+
+\topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
+suffice for our needs.}
+Those needs being that the definition be
+\begin{inlinelist}
+\item faithful to the historical cases of piecework that we see in the scholarship; and
+\item relevant and informative to potential cases of piecework today.
+\end{inlinelist}
+We offer the following:
+that \pieceworkdefinition.
+
+\subsection{What was piecework's historical arc?}
+
+\topic{Piecework's history traces back further perhaps than most would expect.}
+\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
+to calculate equations in order to better--predict the trajectories of various celestial bodies
+\cite{grier2013computers}.
+While this approach didn't become an economic powerhouse as later examples would prove%
+, Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
+determining the extent to which work can be decomposed, and
+finding the limits of complexity of that decomposed work.
+That is, Airy found that he could train youths in elementary mathematics
+to complete the majority of the calculations he would otherwise have had to solve on his own%
+, and that the greater body of work could ultimately be completed sooner
+if he arranged his work appropriately.
+
+\topic{Piecework took a circuitous path in its rise to the mainstream%
+% from the field into the home and finally into factories%
+, each time finding additional ways to leverage the advantages of piecework.}
+First applied to farm work, as
+\citeauthor{hughRaynbirdTaskWork} and others illustrate%
+, the practice remained relatively obscure until
+it was brought to the textile industry
+\cite{hughRaynbirdTaskWork}.
+At the turn of the \nth{20} century%
+, when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City%
+, \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes%
+, describing piecework comprehensively
+% ``The timeworkers' wages should be verified by the timekeeper's book.''
+\cite{riisOtherSideLives,norton1900textile}.
+Soon after,
+we saw the application of piecework systems in textile mills on the realization that
+``[pieceworkers in Italy] will work as many hours as it is possible for him to stand''
+\cite{clark1908cotton}.
+% By the turn of the \nth{20} century%
+% , \citeauthor{riisOtherSideLives} 
+% the poor working and living conditions that were pervasive in
+% an increasingly industrial United States
+% \cite{riisOtherSideLives}.
+% By the turn of the \nth{20} century,
+best practices regarding the measurement and management of
+piecework rates, and of workers in the engineering industry,
+were beginning to take shape
+\cite{burton1899commercial}.
+
+\topic{Researchers have since struggled
+to understand the mechanisms and characteristics in piecework
+which fueled its rise to popularity during this time.}
+\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
+could be found in piecework;
+the approach of paying workers for each piece of output necessitated
+the rigorous tracking, measurement, and training of workers
+for which scientific management became famous
+\cite{10.2307/23702539}.
+This argument is certainly compelling;
+it would seem to make the concurrent upswing of
+scientific management and Fordism
+through the first two--thirds of the \nth{20} century
+alongside piecework not only understandable, but predictable
+\cite{hart2013rise}.
+\citeauthor{Brown01041990} inquired from another direction, asking
+what limited the adoption of piecework in industries that otherwise gravitated toward it
+(in the case studies he examined, this mostly focused on railway engineers)
+\cite{Brown01041990}.
+
+\topic{As increasing attention revealed problems in piecework as it related to workers%
+, workers themselves began to speak out about their frustration with this new regime.}
+It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work%
+, but this led to industry organizations representing
+railway workers, mechanical engineers, and others contributing their myriad perspectives
+\cite{american1921problem,richards1904anything,riisOtherSideLives}.
+Nevertheless, piecework continued to permeate low--skilled labor.
+
+\topic{Piecework became an important contributor to the war effort in the Second World War%
+, cementing its role not only in American factories, but in industrial work around the world.}
+While piecework began to catch on at the turn of the \nth{20} century%
+, the 1930s represented a boom for piecework on an unprecedented scale%
+, especially among engineering and metalworking industries.
+As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+He attributes this to the shortage of male workers%
+, who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+
+% \topic{Necessity drove the widespread adoption of piecework and work decomposition%
+% , perhaps more even than Ford's assembly line.}
+
+\topic{Despite the intense growth of the piecework approach to renumeration%
+, this time was not without turmoil.}
+As previously discussed, a number of worker organizations weighed in on
+(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
+\cite{american1921problem,richards1904anything}.
+\citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
+\cite{10.2307/3827491}.
+
+\topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
+The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
+--- albeit critical ---
+case study in this space;
+the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
+to strike for
+an individual minimum wage, among other demands
+\cite{10.2307/2221944}.
+\citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
+\cite{10.2307/41829256}.
+The adoption of piecework time--study and other principles associated with Taylor and scientific management
+itself reliably precipitated strikes and more generally gave workers a clear enemy
+against which to rally
+\cite{jacoby1983union}.
+
+\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed just a few decades earlier.}
+\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
+In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
+We summarize some of the salient suggestions here:
+\begin{inlinelist}
+\item the emergence of more effective, more nuanced incentive models
+--- rewarding teams for complex achievements, for instance;
+\item the shifting of these industries (manufacturing, clothing, etc\dots)
+to other countries;
+\item the quality of ``multidimensional'' work becoming too difficult to evaluate.
+\end{inlinelist}
+\cite{hart2013rise}.
+% the gendered and stratified dynamics at play in the matchgirls strike
+
+% \topic{Piecework declined in popularity in the United States, but
+% its spirit thrived through the mid--\nth{20} century in
+% scientific management and Fordism.}
+% Concurrent with the fall of piecework was arguably the rise of factories as centers of work.
+
+\subsection{Why is piecework relevant to crowdwork?}
+Using the definition of piecework that we came up with earlier, we argue that
+crowdwork is fundamentally an instantiation of piecework, and
+that we can more precisely anticipate the answers to the open research questions we discussed earlier.
+We'll show that the dimensions of crowdwork that the broader HCI community has been studying
+align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
 
 \onlyinsubfile{
-  % \balance{}
   \printbibliography
-  \clearpage
-  % \nobalance{}
+  % \clearpage
   }

### 2016-09-12 drafts/labor/trackingWork.tex
piecework_lit should be kind of done

@@
 
-\usepackage{zref-xr}
-\usepackage{zref-xr}
-\zxrsetup{%
-  tozreflabel=false, % not needed, since `zref` is not used otherwise
-  toltxlabel=true, % classical LaTeX labels that can be referenced with \ref, \pageref
-}
-\zexternaldocument{cases.tex}% star form: only LaTeX labels are imported
+\usepackage{xr-hyper}
+% \zxrsetup{%
+%   tozreflabel=false, % not needed, since `zref` is not used otherwise
+%   toltxlabel=true, % classical LaTeX labels that can be referenced with \ref, \pageref
+% }
+% \zexternaldocument{cases.tex}% star form: only LaTeX labels are imported
 
@@
 \usepackage[pdflang={en-US},pdftex]{hyperref}
-
+\usepackage{mfirstuc}
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
@@
 \newcommand{\implication}{Implications for crowdwork.}
+\newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of ``payment for results''%
+, leading to and leveraging narrower skill sets required for narrowly defined tasks,
+and affording workers some amount of freedom to complete tasks
+(and consequently earn money)
+at at whatever rate and in whichever manner they wish}
 
@@
 \subfile{cases.tex}
+\subfile{research_implications.tex}
 

### 2016-09-12 drafts/labor/complexity.tex
complexity should be kind of done

@@
 
+% \renewcommand{\topic}[1]{#1}
 \subsubsection[finding crowdwork's limits]{Achieving greater complexity}\label{sec:complexity}
-\ali{maybe a subssubsubsusbsubsection on the result/implication?
-more people can make crowd tasks, but the algiorithmic manag}
-
 \subsubsubsection{\crowdworkpers}
@@
 piecework has been limited principally by the challenges of human management and oversight.}
-\citeauthor{10.2307/23702539}%
-, who describes piecework as
-    ``\dots~based on examination of various shop jobs,
-      which included calculation of the standard time and compensation for each task''%
-, argues that piecework must be rigorously evaluated at a time that demands
-\textit{other people} perform the evaluation.
-% \cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} later enumerates some of the roles required
+\citeauthor{10.2307/23702539} describes a case study in Santa Fe Railway, which
+deployed scientific management and a piecework regime in an attempt to stymie rising repair costs
+\cite{10.2307/23702539}.
+Returning to \citeauthor{hart2016rise}'s reflections on piecework's limitations,
+we recall the multidimensional problem
+--- tasks comprising of numerous, sometimes conflicting, goals
+\cite{hart2016rise}.
+It would be reasonable, then, to infer that work like this
+--- reasonably highly skilled work where quality is difficult to assess ---
+would be unsuitable for piecework.
+
+\topic{\citeauthor{hart2016rise} and \citeauthor{10.2307/23702539}%
+, without acknowledging one another%
+, seemingly corroborate one another's conclusions at different levels of observation.}
+\citeauthor{10.2307/23702539} enumerates some of the roles required
 to facilitate piecework in the early \nth{20} century:
-    ``\dots~piecework clerks, inspectors, and ``experts''\dots''
-\cite{10.2307/23702539}.
-This criterion strictly limits the extent to which piecework can grow in complexity;
-it must, for instance, be quickly evaluable by another person.
+    ``\dots~piecework clerks, inspectors, and `experts'\dots''
+\cite{hart2016rise,10.2307/23702539}.
+\citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
+to be making differing claims about the limitations of piecework, but
+we argue that \citeauthor{10.2307/23702539}
+is simply making a more concrete observation illustrating the insight that
+\citeauthor{hart2016rise} later makes.
+\citeauthor{10.2307/23702539} recognizes that
+it's necessary for a successful piecework shop to employ
+clerks,
+inspectors, and
+other experts to properly design and evaluate complex work.
+\citeauthor{hart2016rise} argues an ultimate limit to how far this can go;
+at some point, evaluating multidimensional work output for quality
+(rather than for quantity) becomes infeasible.
 
@@
 managerial overhead and the fixed cost of adopting a piecework payment regime.
+
+\topic{There are other characteristics to effective complex piecework institutions, such as
+appropriately designed management practices.}
+\citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
+``The foreman had the power to hire and fire workers and allocate workplaces,
+but then left the face--worker largely free to his own efforts so that
+often he went all day without seeing the foreman''
+\cite{10.2307/2118435}.
+The general approach adopted by these West Virginia mines was,
+as in other factories with active foremen,
+to let the foreman be the intermediary between management and the worker.
+Specifically, foremen were responsible for allocating resources and
+understanding when and how to modify work as necessary
+\cite{wray1949marginal}.
 
@@
 
-\topic{Technology more directly facilitates the subversion of expertise requirements
+\topic{Furthermore, technology more directly facilitates the subversion of expertise requirements
 by giving non--experts access to information that would otherwise be unavailable.}
@@
 
-\topic{Crowdsourcing falters when the routinization of complex work proves difficult.}
-\ali{Something about complexity being difficult to make routine when
-there are lots of little, varied, unpredictable tasks (nod to \citeauthor{Brown01041990})}
+\subsubsubsection{\implication}
+\topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
+it also signals some of the ultimate limitations of crowdwork and piecework in general.}
+While the threshold preventing task requesters from utilizing piecework
+has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
+If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
+we would benefit from finding ways to decompose varied tasks into homogeneous microtasks.
+
+We should also consider exploring the limitations that algorithmic management bring along more carefully.
+While research has touched on this subject, we've yet to make out the bigger picture of this theme
+\cite{uberAlgorithm}.
+If we can resolve the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest,
+then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciousCrowdworkersGadiraju}
+and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
+\cite{takingAHITMcInnis}.
+
+% \topic{Crowdsourcing falters when the routinization of complex work becomes too difficult.}
+% \ali{Something about complexity being difficult to make routine when
+% there are lots of little, varied, unpredictable tasks (nod to \citeauthor{Brown01041990})}
 

### 2016-09-12 drafts/labor/complexity.tex
closing the book on decomposition

@@
 (rather than for quantity) becomes infeasible.
+
+\topic{This isn't to say that complex work is outside of the realm of piecework;
+indeed, we've discussed complex applications of crowdwork already.}
+As \citeauthor{hart2013rise} described,
+the 1930s saw a flourishing of clever piecework job design out of necessity
+due to the fact that it was infeasible to provide new workers with
+the comprehensive education that was familiar to men
+\cite{hart2013rise}.
+This constraint led to much more tightly scoped work, and
+(perhaps surprisingly at the time) more efficient allocations of workers,
+who could now specialize in extremely narrowly defined roles.
+The same could be said of Airy and his \textit{computers}
+--- young boys whose preparations consisted principally of
+a relatively specific mathematics curriculum
+\cite{grier2013computers}.
 
@@
 
+Finally, and perhaps most importantly,
+we need to replicate the success of narrowly slicing education and training for expert work
+as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
+\cite{hart2013rise,grier2013computers}.
+That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work.
+To some extent, an argument can be made that
+MOOCs and other online education resources
+provide crowdworkers with the resources that they need, but 
+it remains to be seen whether that work will be appropriately valued, let alone
+properly interpreted by task solicitors
+\cite{aguaded2013mooc}.
+If we can overcome this obstacle,
+we might be able to empower crowdworkers to do complex work such as engineering and metalworking,
+rather than doom them to match girl reputations:
+``brash, irregular, immoral, and uneducated''
+\cite{10.2307/3827491}.
+
 % \topic{Crowdsourcing falters when the routinization of complex work becomes too difficult.}
@@
 
-  \endnotetext{
-  \section{Complexity notes}
-    \topic{\joke{great} topic sentence here.}
-    \citeauthor{selfsourcingTeevan2014} push the boundaries of decomposed work%
-,     exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
-    \cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
-    While some of this work doesn't strictly fall under ``crowdsourcing''%
-,     the [scientific] management of the self as a worker
-    (of sorts)
-    will prove relevant as we trace the literature surrounding piecework.
+%   \endnotetext{
+% %     \topic{We summarize the work discussed so far in this way:
+% %     researchers of crowdsourcing and on--demand labor in general
+% %     have attempted to find the characteristics which
+% %     enable and stymie successful applications of this form of work%
+% % ,     and how to get past or around these boundaries.}
+% %     Whether this work has focused on the self (as in the case of ``selfsourcing'')
+% %     or on others (as in \textit{CrowdForge} and \textit{Foundry})%
+% % ,     crowdsourcing literature has generally studied the constituent work
+% %     and attempted either to work around
+% %     or indeed with
+% %     the limitations of crowdwork to accomplish a difficult task
+% %     \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
 
-    This is also kind of off-topic, isn't it? where to take it?
+% %     Crowdforge, turkomatic, others talk about limiting factors; bring them up
 
-    \topic{A smaller but growing body of work has reflected on this
-    and brought to our attention a number of ethical questions surrounding the
-    increasing complexity of crowdwork and the hazards that increasingly arise.}
-    \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
-    --- working for increasing amounts of time on tasks of growing complexity%
-,     only to discover that requesters are not willing to pay%
-,     for instance ---
-    but these and other dangers range an enormous landscape
-    \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
+%       % Talked to someone about this section and got some feedback\dots
+%       % \begin{enumerate}
+%       %   \item Explain what ``clever arrangement of work'' in the first paragraph means
+%       %         (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
+%       %   \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
+%       %         (maybe even to justify its placement here); and
+%       %   \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
+%       %         to make the historical stuff on piecework
+%       %         directly speak to the three previous paragraphs.
+%       %   \item (Maybe?) summarize the piecework stuff
+%       %         (the last paragraph approaches that, but more could be done).
+%       % \end{enumerate}
+%       % Thoughts on this?
 
-    \topic{We summarize the work discussed so far in this way:
-    researchers of crowdsourcing and on--demand labor in general
-    have attempted to find the characteristics which
-    enable and stymie successful applications of this form of work%
-,     and how to get past or around these boundaries.}
-    Whether this work has focused on the self (as in the case of ``selfsourcing'')
-    or on others (as in \textit{CrowdForge} and \textit{Foundry})%
-,     crowdsourcing literature has generally studied the constituent work
-    and attempted either to work around
-    or indeed with
-    the limitations of crowdwork to accomplish a difficult task
-    \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
-
-    Crowdforge, turkomatic, others talk about limiting factors; bring them up
-
-      Talked to someone about this section and got some feedback\dots
-      \begin{enumerate}
-        \item Explain what ``clever arrangement of work'' in the first paragraph means
-              (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
-        \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
-              (maybe even to justify its placement here); and
-        \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
-              to make the historical stuff on piecework
-              directly speak to the three previous paragraphs.
-        \item (Maybe?) summarize the piecework stuff
-              (the last paragraph approaches that, but more could be done).
-      \end{enumerate}
-      Thoughts on this?
-
-  \citeauthor{10.2307/23702539} also argues that
-  a significant obstacle to the introduction of piecework in railroad shops was
-  the resistance of workers
-  (the other being the resistance of management (for different and varying reasons)).
-  In the cases of online platforms like AMT and digitally mediated platforms like Uber%
-,   where in both cases workers rarely
-  if ever
-  interact face--to--face and
-  opportunities for coordination and collective action are severely limited%
-,   we should expect to see
-  the challenges stymieing collective action
-  that we see in \citeauthor{dynamo} and elsewhere
-  \cite{10.2307/23702539,dynamo}.
-}
+% %   \citeauthor{10.2307/23702539} also argues that
+% %   a significant obstacle to the introduction of piecework in railroad shops was
+% %   the resistance of workers
+% %   (the other being the resistance of management (for different and varying reasons)).
+% %   In the cases of online platforms like AMT and digitally mediated platforms like Uber%
+% % ,   where in both cases workers rarely
+% %   if ever
+% %   interact face--to--face and
+% %   opportunities for coordination and collective action are severely limited%
+% % ,   we should expect to see
+% %   the challenges stymieing collective action
+% %   that we see in \citeauthor{dynamo} and elsewhere
+% %   \cite{10.2307/23702539,dynamo}.
+% }
 

### 2016-09-12 drafts/labor/decomposition.tex
complexity should be kind of done
closing the book on decomposition

@@
 \subsubsubsection{\crowdworkpers}
-% \ali{Does this belong here at all? The complexity argument just dives in.}
 \topic{The crowdsourcing research into work decomposition
@@
 ,       Cai:2015:WLW:2702123.2702267}.
+% \topic{One of the emergent properties of micro--tasks has been the relative cost of
+%     \textit{finding} worthwhile tasks.}
+%     The research community has documented and to some extent attempted to intervene in
+%     the discovery of worthwhile tasks
+%     \cite{taskSearch}.
+\citeauthor{Cosley:2007:SUI:1216295.1216309}
+attempts to address this by
+directing workers to tasks through
+``intelligent task routing''
+\cite{Cosley:2007:SUI:1216295.1216309}.
+Much of this work and the work at the periphery of this space, then%
+, has focused on
+minimizing the amount of time that people need to spend doing
+anything other than the work for which they are paid.
+
 Yet more work looks at the general framing of tasks%
@@
 \cite{embracingErrorKrishna}.
-
-Not all of the work toward optimizing crowd work--flows has gone toward
-minimizing the creative input of crowd workers;
-a thriving body of literature adopts
-practices such as pipelining to allow experts to participate in crowd work
-\cite{foundry}.
 
@@
 , and have other workers check their work.
-\ali{I could point out that the opportunity to check work and
-repeat the task is a little like find--fix--verify, but is that jumping the gun?}
 Second, tasks were discrete --- that is, independent from one another.
@@
 
-\ali{\citeauthor{marx2012economic} in here? Alienated from the context of the work?
-Critiques about piecework marginalizing workers?
-\citeauthor{american1921problem} wrote about this as well
-(albeit they were advocating from workers' perspectives, but who isn't?)}
+\subsubsubsection{\whatchanged}
+ \topic{A number of factors in crowdwork are different from piecework,
+ chief among them being the relative ease with which
+ the metaphorical ``assembly line'' can be changed.}
+ Computers make it possible to switch from one task to another
+unlike any arbitrary manufacturing factory possibly could;
+a worker could do any number of
+different \textit{types} of tasks in the span of just a few minutes
+\cite{delayAndOrderLasecki}.
+This has spurred an entire body of work investigating the effects of
+ordering,
+pacing,
+interruptions, and
+other factors in piecework that would have been
+all but impossible to measure consistently as few as 20 years ago
+\cite{cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
 
-\subsubsubsection{\whatchanged}
-\begin{inlinelist}
-  \item Computers make it possible to switch from one task to another
-  unlike any arbitrary manufacturing factory possibly could;
-  \item we've sliced works to such small sizes that the marginal costs
-  --- things like task--finding, cognitive load switching, etc\dots ---
-  have become relatively large;
-  \item instrumentation has become so advanced that
-  the curve of diminishing value on measuring and tracking workers has shifted significantly
-  (but not been obliterated);
-  \ali{some companies have suggested self--tracking through
-  programs that give workers fitbits and whatnot --- I could make the argument that
-  this is just an illustration of that, but
-  it's not really about work \textit{per se} unless
-  you think about it as the general management of workers. Thoughts?}
-\end{inlinelist}
+Further, we've sliced work to such small scales that the marginal activities
+--- things like finding work, cognitive task switching, etc\dots ---
+have become relatively large compared to the tasks themselves
+\cite{taskSearch}.
+In the historical case of piecework,
+moving metallurgical tools, mining equipment, or
+other industry materials would have been prohibitively difficult and slow;
+workers were encouraged to specialize in a single set of tasks,
+allowing pieceworkers to sequence their tasks optimally on their own
+\cite{hart2013rise}.
+
+Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
+--- one which where crowdworkers are rendered as
+``modular, protocol--defined computational services'' ---
+we may yield better results from crowdwork if we think of workers as similar to
+specialized, repurposable tools
+\cite{irani2015cultural}.
+\ali{feeling meh about this argument\dots}
+
+Finally, instrumentation has reached a sufficiently advanced and ubiquitous point that
+the dream of scientific management and Taylorism
+--- to measure every motion at every point throughout the workday and beyond ---
+is not only doable, but trivial
+\cite{waltz2012quantified}.
+One of the major challenges \citeauthor{10.2307/23702539} cites as
+preventing scientific management from being fully utilized,
+the difficulty of tracking work \& workers, no longer exists
+\cite{10.2307/23702539}.
+
+\subsubsubsection{\implication}
+Crowdwork research today is on the right track to investigate pipelining and meta--task design.
+That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
+\cite[see, for example,][]{turkopticon}.
+It's not clear how much benefit there is in the further decomposition of work,
+given that we've hit bottlenecks with the cognitive stresses of switching between tasks
+as \citeauthor{delayAndOrderLasecki} highlight
+\cite{delayAndOrderLasecki}.
 
@@
   }
-  \endnotetext{
-    \section{Slicing notes}
-    The finding here is that
-    crowdwork can be more carefully micro--managed than piecework could be%
-,     and that this is a double--edged sword:
-    we can effectively give feedback to workers on everything they do%
-,     but this is emboldening us to try to
-    over--manage workers just as piecework tried to do.
+%   \endnotetext{
+% %     \section{Slicing notes}
+% %     The finding here is that
+% %     crowdwork can be more carefully micro--managed than piecework could be%
+% % ,     and that this is a double--edged sword:
+% %     we can effectively give feedback to workers on everything they do%
+% % ,     but this is emboldening us to try to
+% %     over--manage workers just as piecework tried to do.
 
-    My goal for this section is to make two points:
-    \begin{enumerate}
-      \item show how this is related to the assembly line and scientific management%
-,             and how piecework literature tried to measure everything%
-,             but found it untenable given the extra equipment that was necessary
-            (but generally which didn't exist)
-            to track every movement and action that workers took.
-      \item show how this work was enabled by the ``verifiability'' of work output(?)
-    \end{enumerate}
+% %     My goal for this section is to make two points:
+% %     \begin{enumerate}
+% %       \item show how this is related to the assembly line and scientific management%
+% % ,             and how piecework literature tried to measure everything%
+% % ,             but found it untenable given the extra equipment that was necessary
+% %             (but generally which didn't exist)
+% %             to track every movement and action that workers took.
+% %       \item show how this work was enabled by the ``verifiability'' of work output(?)
+% %     \end{enumerate}
 
-    As a result, the prevalent mindset of designing work for crowd workers
-    --- one which treats micro task workers as
-    ``modular, protocol--defined computational services'' ---
-    has inexorably alienated workers from the greater context of their work
-    \cite{irani2015cultural}.
+%     % \ali{i don't like this section \textit{here}, but i like it in general. what do?}
 
-    \ali{i don't like this section \textit{here}, but i like it in general. what do?}
-    \topic{One of the emergent properties of micro--tasks has been the relative cost of
-    \textit{finding} worthwhile tasks.}
-    The research community has documented and to some extent attempted to intervene in
-    the discovery of worthwhile tasks
-    \cite{taskSearch}.
-    \citeauthor{Cosley:2007:SUI:1216295.1216309}
-    attempts to address this by
-    directing workers to tasks through
-    ``intelligent task routing''
-    \cite{Cosley:2007:SUI:1216295.1216309}.
-    Much of this work and the work at the periphery of this space, then%
-,     has focused on
-    minimizing the amount of time that people need to spend doing
-    anything other than the work for which they are paid.
-
-    What we take away from this and the previous set of work is that
-    the value of adopting crowdsourcing for any particular task
-    seems to be mediated by two questions:
-    \begin{inlinelist}
-      \item How long does it take to train workers to do the work in question? and
-      \item How long does it take for the worker to do the work?
-    \end{inlinelist}
-    Minimizing these criteria has become
-    the overarching motivation of the crowdsourcing work design community
-    \cite{cheng2015break,Newell:2016:OMA:2858036.2858490}
-
-  }
+%   }
 

### 2016-09-12 drafts/labor/relationships.tex
closing the book on decomposition

@@
 
-\subsubsection[the relationships of workers]{???}
+\subsubsection[the relationships of workers]{The Relationships of Workers to Work, Peers, and Others}
 \label{sec:relationships}
+\subsubsubsection{\crowdworkpers}
+\topic{A number of ethical questions surrounding the
+    increasing complexity of crowdwork and the hazards have increasingly arisen.}
+    \citeauthor{professionalCrowdworkEthics} bring some of these issues at stake
+    --- working for increasing amounts of time on tasks of growing complexity%
+,     only to discover that requesters are not willing to pay%
+,     for instance ---
+    but these and other dangers range an enormous landscape
+    \cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd,dynamo}.
 
@@
   \printbibliography
-  \clearpage
+  % \clearpage
   % \nobalance{}
@@
     A number of researchers have identified
-    worker attrition,
-    variability of worker performance,
-    and uncertainty about good versus bad--faith actors as open questions of crowdwork
+    worker attrition%
+,     variability of worker performance%
+,     and uncertainty about good versus bad--faith actors as open questions of crowdwork
     \cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
@@
     both the study and practice
-    of crowdwork, that microtasks should be designed resiliently against the variability of workers,
-    fully exploiting the abstracted nature of each piece of work
+    of crowdwork, that microtasks should be designed resiliently against the variability of workers%
+,     fully exploiting the abstracted nature of each piece of work
     \cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-    That is to say, micro--tasks should be designed such that a single worker's poor performance,
-    or a good worker's sudden departure,
-    does not significantly impact the agenda of the work as a whole.
+    That is to say, micro--tasks should be designed such that a single worker's poor performance%
+,     or a good worker's sudden departure%
+,     does not significantly impact the agenda of the work as a whole.
     While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
     in the form of higher cumulative time to complete
-    (albeit much shorter real time to complete, owing to parallelization),
-    \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+    (albeit much shorter real time to complete, owing to parallelization)%
+,     \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
     similar tasks together
@@
     \citeauthor{embracingErrorKrishna} take a different approach;
-    by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds,
-    the authors offer orders--of--magnitude improvements
+    by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds%
+,     the authors offer orders--of--magnitude improvements
     in various binary classification tasks
     \cite{embracingErrorKrishna}.
-    And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-    some researchers have designed work predicated on the expectation of it instead:
+    And rather than building tasks to \textit{tolerate} worker drop--off and attrition%
+,     some researchers have designed work predicated on the expectation of it instead:
     \citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
@@
     \begin{inlinelist}
-      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju},
-      \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
+      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju}%
+,       \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
       \item Expecting certain levels of attrition and incorrectness and using that variability to their advantage \cite{embracingErrorKrishna}
@@
 
-    In the field of piecework,
-    the research covering this topic has both explored
+    In the field of piecework%
+,     the research covering this topic has both explored
     a breadth of tasks that might be rendered doable by piecemeal workers
@@
 }
-\theendnotes
+% \theendnotes
 \end{document}

### 2016-09-13 drafts/labor/cases.tex
msb comments

@@
 \section{Major Research Questions}
+
+\msb{this needs a transition into it.
+Before you lay out our method, lay out what we're trying to do.
+What's our goal here?
+THEN the method can explain how we achieved it.
+This might involve switching the order of some of the text here}
 
@@
 We'll also point to corresponding piecework literature addressing these aspects.
+\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
 Finally, we'll discuss how these elements will serve

### 2016-09-13 drafts/labor/complexity.tex
msb comments

@@
 , often involving insights made in Computer Science and applied to human work--flows.}
-The crowdwork literature typically identifies target milestones in CS
+The crowdwork literature typically identifies target milestones in CS \msb{don't use CS, it could be either crowdsourcing or computer science}
 that have presented significant challenges for researchers%
@@
 it leverages the in--built advantages that technology and digital media afford.
-\textit{Foundry}'s tools for managing and arranging expert groups into a cohort
+For example, \textit{Foundry}'s tools for managing and arranging expert groups into a cohort
 allow researchers to convincingly argue that expert teams can be rapidly formed%
@@
 \subsubsubsection{\pieceworkpers}
+\msb{needs a transition to set the tone here. e.g., The piecework literature also examined the complexity limits possible with piecework.}
 \topic{The research into piecework makes the case that
@@
 --- reasonably highly skilled work where quality is difficult to assess ---
-would be unsuitable for piecework.
+would be unsuitable for piecework. \msb{that conclusion came way too fast. give a little detail on the railway. how did they come to that conclusion?}
 
+\msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
 \topic{\citeauthor{hart2016rise} and \citeauthor{10.2307/23702539}%
@@
 is simply making a more concrete observation illustrating the insight that
-\citeauthor{hart2016rise} later makes.
+\citeauthor{hart2016rise} later makes. \msb{what insight? I've lost the thread here.}
 \citeauthor{10.2307/23702539} recognizes that
@@
 a relatively specific mathematics curriculum
-\cite{grier2013computers}.
+\cite{grier2013computers}. \msb{so what are we to take out of this? that there is no limit? just our creativity?}
 
 \topic{Piecework researchers also make claims regarding
-the organizations that benefit from piecework in the first place.}
+the organizations that benefit from piecework in the first place.} \msb{I think a reader wouldn't understand why that claim is relevant.
+Let's invert it, something like how the researchers point out that only certain kinds of organizations can effectively make use of piecework.}
 \citeauthor{Brown01041990} discusses the factors necessary for piecework to thrive:
@@
     a variety of duties than in jobs with a narrow set of routinized duties''
-\cite{Brown01041990}.
+\cite{Brown01041990} \msb{that's confusing, is incentive pay the same as piecework? can you cut the first part of the quote and give us the noun instead of forcing us to figure it out?}.
 \citeauthor{10.2307/23702539} adds further, that
@@
 make a persuasive argument that piecework is limited in complexity by
-managerial overhead and the fixed cost of adopting a piecework payment regime.
+managerial overhead \msb{wait, wasn't that in a previous paragraph?
+that should be joined with the managerial overhead text above} and the fixed cost of adopting a piecework payment regime \msb{that seems like it's the new point here}.
 
@@
 understanding when and how to modify work as necessary
-\cite{wray1949marginal}.
+\cite{wray1949marginal}. \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
 
 \subsubsubsection{\whatchanged}
+\msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework.
+You're about to draw on those points to make your argument, so they need to be at the top of my mind here.}
 \topic{Digital media have expanded the scope of
@@
 the nature of the work is limited in complexity
-\cite{Brown01041990}.
+\cite{Brown01041990} \msb{I don't recall this point in the prior section: I remember management overhead and fixed costs of materials, I don't remember a paragraph about complexity.
+If we want to draw on that point, make it in the earlier section}.
 Given the flourishing of on--demand labor platforms such as
@@
         ``economies of scale'' at very little cost
-        to the system--designer \cite{lessig2006code,miller2011understanding}
+        to the system--designer \cite{lessig2006code,miller2011understanding} \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
 \end{inlinelist}.
@@
 within the domain of experts.}
+\msb{I don't yet buy the following argument.
+If the point is that technology makes us more expert, I disagree that the CrowdCrit/Voyant systems are using technology to do this.
+They are building the smarts into their OWN workflows, rather than giving workers EXTERNAL tools that make them smarter.
+Giving workers a calculator is an external tool; the mathematical tables project already demonstrated that you can build smarts into the workflow if you don't have one.}
 \citeauthor{yuanAlmost} builds on the work of others
@@
 \cite{uberAlgorithm,10.2307/23702539}.
+}
 
@@
 ranging topics such as police activity, congestion, construction, etc\dots
-\cite{silva2013traffic,hind2014outsmarting}.
+\cite{silva2013traffic,hind2014outsmarting}. \msb{what's the insight I should take away from this paragraph? what does this say about crowd work?}
 
 \subsubsubsection{\implication}
+\msb{This paragraph can be expanded to make a more concrete argument. What will be possible? What won't?}
 \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
@@
 has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
+\msb{is that your prediction?
+I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
 If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
-we would benefit from finding ways to decompose varied tasks into homogeneous microtasks.
+we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
 
+\msb{this doesn't seem like a concrete prediction. what would piecework say will happen if we didn't resolve the tension?}
 We should also consider exploring the limitations that algorithmic management bring along more carefully.
@@
 as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-\cite{hart2013rise,grier2013computers}.
-That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work.
+\cite{hart2013rise,grier2013computers} \msb{remind us of what those were}.
+That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work. \msb{I don't understand: why?}
 To some extent, an argument can be made that
@@
 ``brash, irregular, immoral, and uneducated''
-\cite{10.2307/3827491}.
+\cite{10.2307/3827491}. \msb{how will that reduce any of the problems except the last one?}
 

### 2016-09-13 drafts/labor/introduction.tex
renamed files, moved stuff around, adding implications sections
msb comments

@@
 the constituent parts of larger tasks.}
-This framing of work into abstract blocks
+This framing of work into \msb{modular, instead of abstract?} abstract blocks
 has allowed people to engage in work despite
@@
 that complex tasks can be accomplished by directing and managing crowds of workers
-has spurred the research and and industry communities to flock to sites of labor
+has spurred the research and industry communities to flock to sites of labor
 like Amazon's Mechanical Turk (AMT)
-to explore the limits of this distributed, fleeting workforce.
-Researchers in particular have taken to the space in earnest,
+to explore the limits of this distributed, \msb{fleeting sounds pejorative. on-demand? it sounds weird to flock to a site for a pejorative reason} fleeting workforce.
+Researchers in particular \msb{no more ``in particular'' than industry...Crowdflower is big, much bigger than research} have taken to the space in earnest,
 finding opportunities to enable new forms of work
@@
 
+\msb{I suggest merging the next two paragraphs into one.
+Right now it's not clear by the time I get to the third paragraph where this is going.
+Instead, I think the topic of the single (merged, snipped-down) paragraph would be how this stuff is growing, yay, and it's giving rise to challenges.}
 \topic{This form of work has grown considerably in size,
@@
 
-% \topic{This matters because\dots WHYYYYYYYYYYYYYYY\\}
-% Indeed, this on--demand workforce has sparked interest
-% across industries ranging
-% driving for hire (for example Uber),
-% house--cleaning (Handy),
-% and generalized services (TaskRabbit)
-% \cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-% Today, a rapidly growing transient workforce is forming,
-% itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
-% to benefit from a latent pool of previously vetted workers
-% \cite{pewSharing}.
-
 \topic{For all the growth we've observed in this labor market,
@@
 
+\msb{there's a lot of flipping between on-demand work, crowd work, etc.
+in this intro.
+Name them all up near the front, state which one you're going to use in \textit{italics}, and then be consistent throughout.}
 \topic{The extant body of work has ostensibly sought to answer one underlying question:
-What does the future hold for work and those that do it?}
+What does the future hold for \msb{on-demand/crowd} work and those that do it?}
 Researchers have offered their input on this open question along three major threads:
@@
         \begin{inlinelist}
-          \item how complex can crowd work get? and
+          \item how complex can crowd work get? \msb{how complex are the goals that crowd work can accomplish?} and
           \item how far will crowd work reach into the everyday lives of people?
+\msb{which people?
+workers?
+requesters?
+do you mean, will recruiting crowd work be the domain of experts or everyday end users?}
         \end{inlinelist}
@@
               Hahn:2016:KAB:2858036.2858364};
-  \item To what extent can we decompose work (especially into micro--tasks)?
+  \item To what extent can we decompose work (especially into micro--tasks)? \msb{How far can work be decomposed into smaller and smaller microtasks?}
         \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
               whyWouldAnyoneBrewer,takingAHITMcInnis}
+        \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
 \end{enumerate}
@@
 but has not offered a persuasive framing for holistically explaining
+\msb{the next two lines are vaguely referencing the last paragraph, but the argument would be much stronger if it explicitly tied back to those questions.
+Like that they haven't holistically explained the answers to those questions}
 the developments in worker processes that researchers have developer, or
@@
 We offer a framing for crowd work spanning the aforementioned industries
-collectively as a contemporary instantiation of ``piecework''.
+collectively as a contemporary instantiation of \msb{the historical arc of} ``piecework''.
+\msb{one sentence defining piecework}
 Piecework as a metaphor for the type of work at hand is not new.
@@
 
-More concretely, by looking at task--based or ``gig'' work as
+More concretely, by looking at task--based or ``gig'' work \msb{another example of the floating task-based, gig, crowd, on-demand work term} as
 an instantiation (or even a continuation) of piecework,
@@
         and finally,
-  \item offer predictions of what social computing researchers,
+  \item offer predictions of what social computing \msb{crowdsourcing?} researchers,
         and workers themselves,
@@
 \end{inlinelist}
+\msb{what I think would really help this paragraph is a single great example, in one or two sentences, of how this kind of framing would help.
+``For example, we draw on ...
+to explain ...''.
+Reader goes AHA I see why you're doing this now!}
 
-We'll look at a broad range of cases under a number of major themes
+We'll \msb{use contractions sparingly, this sounds too informal} look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence.
+get specific.}
 we propose as broadly describing the types of research being done in crowd work
@@
   \printbibliography
-  \nobalance{}
-  % \clearpage
-  % \begin{appendices}
-  % \section{Sleeping kittens}
-  % \end{appendices}
-
 }

### 2016-09-13 drafts/labor/piecework_lit.tex
complexity should be kind of done
msb comments

@@
 As this paper principally traces a relationship between
-the historical piecework and the contemporary crowdwork
+the historical piecework and the contemporary crowdwork \msb{consistency: crowdwork or crowd work?}
 (or on--demand labor more generally)%
@@
 ,         considering in particular
-        the factors that may have led to piecework's eventual demise
+        the factors that may have led to piecework's \msb{(and crowd work's?)} eventual demise
 \end{inlinelist}.
@@
 \topic{One of the earliest definitions of piecework%
-, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
+, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.} \msb{The circumspect bit is weird, not sure how to interpret that.}
 \citeauthor{hughRaynbirdTaskWork} offers
@@
 muddies the waters for us later as we attempt to categorize cases of piecework
-\cite{rowan1901premium}.
+\cite{rowan1901premium}. \msb{That point runs way too fast for me --- why does it muddy the waters? Either simplify, cut, or slow it down.}
 Nevertheless,
-this work offers us an intuition for what piecework is, if not a bright--line rule.
+this work offers us an intuition for what piecework is, if not a bright--line rule. \msb{I'd cut that, I think it's fine...}
 
@@
 , but for now we can say that piecework affords
-independence and some sense of locus
+independence and some sense of locus \msb{not sure what that means...a sense of place?}
 otherwise unknown to workers.
@@
 \cite{hart2013rise}.
+\msb{explain that in your own words now --- that the piecework infrastructure allows lower-skilled people to be used for skilled positions, by building the expertise into the workflow}
 
 \topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
-suffice for our needs.}
+suffice for our needs.} \msb{Again, I don't think the goal for the reader should be to arrive at a definition through a narrative, they instead need to know how to think about piecework.
+This  paragraph could probably be cut}
 Those needs being that the definition be
@@
 to calculate equations in order to better--predict the trajectories of various celestial bodies
-\cite{grier2013computers}.
+\cite{grier2013computers}. \msb{year?}
 While this approach didn't become an economic powerhouse as later examples would prove%
@@
 
+\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
+What I got out of it is that a bunch of people did piecework, but I don't know why or in fact why these are different than the sources we cite earlier.
+Can you hone it?}
 \topic{Piecework took a circuitous path in its rise to the mainstream%
 % from the field into the home and finally into factories%
-, each time finding additional ways to leverage the advantages of piecework.}
+, each time finding additional ways to leverage the advantages of piecework.} \msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?}
 First applied to farm work, as
@@
 , \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes%
-, describing piecework comprehensively
+, describing piecework comprehensively \msb{here's an example of why I'm confused. there's guidance and comprehensive description, but this text doesn't tell me what actually happened with piecework.}
 % ``The timeworkers' wages should be verified by the timekeeper's book.''
@@
 % By the turn of the \nth{20} century,
-best practices regarding the measurement and management of
+Best practices regarding the measurement and management of
 piecework rates, and of workers in the engineering industry,
@@
 (in the case studies he examined, this mostly focused on railway engineers)
-\cite{Brown01041990}.
+\cite{Brown01041990}. \msb{Seems weird to leave the question hanging. Either un-questionify it, or answer it.}
 
@@
 \subsection{Why is piecework relevant to crowdwork?}
+\msb{I assume this is forthcoming?}
 Using the definition of piecework that we came up with earlier, we argue that

### 2016-09-13 drafts/labor/abstract.tex
piecework_lit should be kind of done
lots of changes all over the place

@@
 \makeatother
-% \zexternaldocument*{complexity}
 
@@
 \begin{abstract}
-Networked computation is enabling the rise of
-crowdwork,
+The Internet is enabling the rise of
+crowd work,
 gig work, and
@@
 A large and growing body of scholarship has sought to predict
-the socio--technical outcomes of this shift, especially
-\nameref{sec:complexity} how complex and interdependent crowdwork's tasks can be,
-\nameref{sec:decomposition} how thinly crowdwork can be sliced and modularized, and
-\nameref{sec:relationships} what the collective outcomes will be for crowdworkers.
+the socio--technical outcomes of this shift, especially along three threads: %. This question has been dominated by three threads:
+\begin{inlinelist}
+\item \nameref{sec:complexity};
+\item \nameref{sec:decomposition}; and
+\item \nameref{sec:relationships}.
+\end{inlinelist}
 In this paper, we look to the historical scholarship on piecework
---- a strikingly similar trend of
+--- a similar trend of
 work decomposition,
@@
 payment that was popular at the turn of the \nth{20} century ---
-to understand how these questions might play out with modern crowdwork.
+to understand how these questions might play out with modern crowd work.
 To do so, we identify the mechanisms that limited piecework historically, and
-identify whether crowdwork faces the same mechanism limits or might differentiate itself.
-
-% We consider the aforementioned open questions in the body of research on crowdwork
-% under the following framework:
-% \begin{inlinelist}
-% \item we look to the familiar literature in human computation;
-%       then
-% \item we investigate the scholarship on piecework;
-% \item we identify what (if anything) has changed in crowdwork to differentiate it from piecework;
-%       and finally
-% \item we attempt to provide a concise prediction on these open questions,
-%       informed by piecework's insights.
-% \end{inlinelist}
-
-% With growing attention toward on--demand labor
-% --- ranging from the ``sharing economy'' to information work ---
-% scholars have made connections to various frameworks 
-% and mechanisms such as worker advocacy, empowerment, and Taylorism,
-% to make sense of our observations of
-% on--demand work and the workers that power this movement.
-% We argue that the literature surrounding ``piecework''
-% informs and even predicts both the contributions
-% that have been made toward the development of on--demand labor and crowd work
-% as well as the fallout among workers and researchers
-% with regard to the disillusionment and alienation of work.
-
-% After evaluating this framing
-% through a series of case studies,
-% we look to the future
-% to identify worthwhile questions and
-% points of inquiry, such as the movement toward factories,
-% that researchers in social computing should consider
-% as we attempt to anticipate and perhaps shape the future of work.
+identify whether crowd work faces the same mechanism limits or might differentiate itself.
+This approach to understanding crowd work provides additional theoretical framing
+to make sense of a topic with which researchers continue to struggle to understand.
 \end{abstract}
-
-% \onlyinsubfile{
-% \clearpage
-% \subfile{cases}
-% }
-
 \end{document}

### 2016-09-13 drafts/labor/cases.tex
lots of changes all over the place

--- 2016-09-12 drafts/labor/cases.tex
@@
 
-We look at lots of papers that use the term ``crowdsourcing'' or ``crowdwork''
+We look at lots of papers that use the term ``crowdsourcing'' or ``crowd work''
 in the abstracts and titles of their papers%
@@
 , and how that affects the predictions and conclusions made in the piecework literature.
-% look at review papers
-
-% what's the methodology?
-% make piecework literature much clearer about argument
-% make headers clearer to signal what i'm doing here.
 
@@
 
-%                        _ _      _ 
-%  _ __   __ _ _ __ __ _| | | ___| |
-% | '_ \ / _` | '__/ _` | | |/ _ \ |
-% | |_) | (_| | | | (_| | | |  __/ |
-% | .__/ \__,_|_|  \__,_|_|_|\___|_|
-% |_|
-
 Piecework has seen work along this dimension spanning decades;
@@
 
-\subsubsection{So how does this affect crowdwork?}
+\subsubsection{So how does this affect crowd work?}
 The work we've seen so far 
@@
 ,     and certainly corroborates \citeauthor{Ross}'s findings after investigating
-    ``who are the crowdworkers''%
+    ``who are the crowd workers''%
 ,     but as \citeauthor{whoareNOTtheTurkers} points out
@@
     ``understanding malicious behavior''
-    \cite{MaliciousCrowdworkersGadiraju}.
+    \cite{MaliciouscrowdworkersGadiraju}.
     While some work has cast workers as ``malicious'' or at least adversarial parties%

### 2016-09-13 drafts/labor/complexity.tex
lots of changes all over the place

--- 2016-09-12 drafts/labor/complexity.tex
@@
 % \renewcommand{\topic}[1]{#1}
-\subsubsection[finding crowdwork's limits]{Achieving greater complexity}\label{sec:complexity}
+\subsubsection[finding crowd work's limits]{Achieving greater complexity}\label{sec:complexity}
 \subsubsubsection{\crowdworkpers}
@@
 , often involving insights made in Computer Science and applied to human work--flows.}
-The crowdwork literature typically identifies target milestones in CS
+The crowd work literature typically identifies target milestones in CS
 that have presented significant challenges for researchers%
@@
 \topic{This isn't to say that complex work is outside of the realm of piecework;
-indeed, we've discussed complex applications of crowdwork already.}
+indeed, we've discussed complex applications of crowd work already.}
 As \citeauthor{hart2013rise} described,
@@
 \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
-it also signals some of the ultimate limitations of crowdwork and piecework in general.}
+it also signals some of the ultimate limitations of crowd work and piecework in general.}
 While the threshold preventing task requesters from utilizing piecework
@@
 If we can resolve the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest,
-then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciousCrowdworkersGadiraju}
+then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciouscrowdworkersGadiraju}
 and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
@@
 \cite{hart2013rise,grier2013computers}.
-That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work.
+That is, we need to identify new ways to train crowd workers for uniquely narrowly defined work.
 To some extent, an argument can be made that
 MOOCs and other online education resources
-provide crowdworkers with the resources that they need, but 
+provide crowd workers with the resources that they need, but 
 it remains to be seen whether that work will be appropriately valued, let alone
@@
 If we can overcome this obstacle,
-we might be able to empower crowdworkers to do complex work such as engineering and metalworking,
+we might be able to empower crowd workers to do complex work such as engineering and metalworking,
 rather than doom them to match girl reputations:
@@
 
-% \topic{Crowdsourcing falters when the routinization of complex work becomes too difficult.}
-% \ali{Something about complexity being difficult to make routine when
-% there are lots of little, varied, unpredictable tasks (nod to \citeauthor{Brown01041990})}
-
 \onlyinsubfile{
-  \balance{}
   \printbibliography
-  % \clearpage
-  \nobalance{}
   }
 
-%   \endnotetext{
-% %     \topic{We summarize the work discussed so far in this way:
-% %     researchers of crowdsourcing and on--demand labor in general
-% %     have attempted to find the characteristics which
-% %     enable and stymie successful applications of this form of work%
-% % ,     and how to get past or around these boundaries.}
-% %     Whether this work has focused on the self (as in the case of ``selfsourcing'')
-% %     or on others (as in \textit{CrowdForge} and \textit{Foundry})%
-% % ,     crowdsourcing literature has generally studied the constituent work
-% %     and attempted either to work around
-% %     or indeed with
-% %     the limitations of crowdwork to accomplish a difficult task
-% %     \cite{selfsourcingTeevan2014,selfsourcingTeevan2016,crowdForgeKittur,foundry}.
-
-% %     Crowdforge, turkomatic, others talk about limiting factors; bring them up
-
-%       % Talked to someone about this section and got some feedback\dots
-%       % \begin{enumerate}
-%       %   \item Explain what ``clever arrangement of work'' in the first paragraph means
-%       %         (should be simple; Find--Fix--Verify is the \textit{sort} of thing I mean);
-%       %   \item I should make the ``selfsourcing'' stuff clearer for people unfamiliar with it
-%       %         (maybe even to justify its placement here); and
-%       %   \item I should break up \citeauthor{10.2307/23702539} \& \citeauthor{Brown01041990}
-%       %         to make the historical stuff on piecework
-%       %         directly speak to the three previous paragraphs.
-%       %   \item (Maybe?) summarize the piecework stuff
-%       %         (the last paragraph approaches that, but more could be done).
-%       % \end{enumerate}
-%       % Thoughts on this?
-
-% %   \citeauthor{10.2307/23702539} also argues that
-% %   a significant obstacle to the introduction of piecework in railroad shops was
-% %   the resistance of workers
-% %   (the other being the resistance of management (for different and varying reasons)).
-% %   In the cases of online platforms like AMT and digitally mediated platforms like Uber%
-% % ,   where in both cases workers rarely
-% %   if ever
-% %   interact face--to--face and
-% %   opportunities for coordination and collective action are severely limited%
-% % ,   we should expect to see
-% %   the challenges stymieing collective action
-% %   that we see in \citeauthor{dynamo} and elsewhere
-% %   \cite{10.2307/23702539,dynamo}.
-% }
-
 \end{document}

### 2016-09-13 drafts/labor/decomposition.tex
lots of changes all over the place

@@
 anything other than the work for which they are paid.
+
+Earlier we discussed \citeauthor{cheng2015break}'s work
+measuring the impact that interruption has on worker performance
+\cite{cheng2015break}.
+This work illustrates a broader sentiment in
+both the study and practice
+of crowd work, that microtasks should be designed resiliently against the variability of workers%
+, fully exploiting the abstracted nature of each piece of work
+\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
+That is to say, micro--tasks should be designed such that a single worker's poor performance%
+, or a good worker's sudden departure%
+, does not significantly impact the agenda of the work as a whole.
+While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
+in the form of higher cumulative time to complete
+(albeit much shorter real time to complete, owing to parallelization)%
+, \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+similar tasks together
+\cite[respectively]{cheng2015break,delayAndOrderLasecki}.
 
@@
 \subsubsubsection{\whatchanged}
- \topic{A number of factors in crowdwork are different from piecework,
+ \topic{A number of factors in crowd work are different from piecework,
  chief among them being the relative ease with which
@@
 a worker could do any number of
-different \textit{types} of tasks in the span of just a few minutes
-\cite{delayAndOrderLasecki}.
+different \textit{types} of tasks in the span of just a few minutes,
+driven in particular by the power \citeauthor{lessig2006code} points to ---
+that system--designers can share, modify, and instantiate environments
+like sites of labor in a few lines of code
+\cite{delayAndOrderLasecki,lessig2006code}.
 This has spurred an entire body of work investigating the effects of
@@
 Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
---- one which where crowdworkers are rendered as
+--- one which where crowd workers are rendered as
 ``modular, protocol--defined computational services'' ---
-we may yield better results from crowdwork if we think of workers as similar to
+we may yield better results from crowd work if we think of workers as similar to
 specialized, repurposable tools
@@
 \subsubsubsection{\implication}
-Crowdwork research today is on the right track to investigate pipelining and meta--task design.
+crowd work research today is on the right track to investigate pipelining and meta--task design.
 That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
@@
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-
-  % \clearpage
-  % \nobalance{}
-  }
-%   \endnotetext{
-% %     \section{Slicing notes}
-% %     The finding here is that
-% %     crowdwork can be more carefully micro--managed than piecework could be%
-% % ,     and that this is a double--edged sword:
-% %     we can effectively give feedback to workers on everything they do%
-% % ,     but this is emboldening us to try to
-% %     over--manage workers just as piecework tried to do.
-
-% %     My goal for this section is to make two points:
-% %     \begin{enumerate}
-% %       \item show how this is related to the assembly line and scientific management%
-% % ,             and how piecework literature tried to measure everything%
-% % ,             but found it untenable given the extra equipment that was necessary
-% %             (but generally which didn't exist)
-% %             to track every movement and action that workers took.
-% %       \item show how this work was enabled by the ``verifiability'' of work output(?)
-% %     \end{enumerate}
-
-%     % \ali{i don't like this section \textit{here}, but i like it in general. what do?}
-
-%   }
-
 \end{document}

### 2016-09-13 drafts/labor/flexibility.tex
lots of changes all over the place

@@
 ,     variability of worker performance%
-,     and uncertainty about good versus bad--faith actors as open questions of crowdwork
-    \cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
+,     and uncertainty about good versus bad--faith actors as open questions of crowd work
+    \cite{MaliciouscrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
     \ali{We can and should discuss the distinction between presumably ``bad faith'' workers
@@
     both the study and practice
-    of crowdwork, that microtasks should be designed resiliently against the variability of workers%
+    of crowd work, that microtasks should be designed resiliently against the variability of workers%
 ,     fully exploiting the abstracted nature of each piece of work
@@
     \begin{inlinelist}
-      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju}%
+      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciouscrowdworkersGadiraju}%
 ,       \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and

### 2016-09-13 drafts/labor/piecework_lit.tex
lots of changes all over the place

--- 2016-09-12 drafts/labor/piecework_lit.tex
@@
 As this paper principally traces a relationship between
-the historical piecework and the contemporary crowdwork
+the historical piecework and the contemporary crowd work
 (or on--demand labor more generally)%
@@
 
-\subsection{Why is piecework relevant to crowdwork?}
+\subsection{Why is piecework relevant to crowd work?}
 Using the definition of piecework that we came up with earlier, we argue that
-crowdwork is fundamentally an instantiation of piecework, and
+crowd work is fundamentally an instantiation of piecework, and
 that we can more precisely anticipate the answers to the open research questions we discussed earlier.
-We'll show that the dimensions of crowdwork that the broader HCI community has been studying
+We'll show that the dimensions of crowd work that the broader HCI community has been studying
 align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.

### 2016-09-13 drafts/labor/relationships.tex
lots of changes all over the place

--- 2016-09-12 drafts/labor/relationships.tex
@@
 
-\subsubsection[the relationships of workers]{???}
+\subsubsection[the relationships of workers]{The Relationships of Workers to Work, Peers, and Others}
 \label{sec:relationships}
 
-\ali{TODO:
-\begin{enumerate}
-  \item \crowdworkpers
-  \item \pieceworkpers
-  \item \whatchanged
-\end{enumerate}
-}
+\subsubsubsection{\crowdworkpers}
+\topic{The relationships of workers with their work, with their peers, and with others
+are complex and not especially well--understood.}
+While researchers have begun to appreciate
+the sociality of crowd workers in labor markets,
+the study of these communities is made enormously more challenging by
+the limited access of online spaces 
+\cite{crowdcollab}.
+Nevertheless, a number of interesting findings have
+already emerged in the crowd work literature%
+, which we can report here.
+
+\topic{A number of ethical questions surrounding the
+increasing complexity of crowd work have arisen in recent years.}
+\citeauthor{professionalcrowdworkEthics} bring some of these issues at stake
+--- working for increasing amounts of time on tasks of growing complexity%
+, only to discover that requesters are not willing to pay%
+, for instance ---
+but these and other dangers range an enormous landscape
+\cite{crowdworkFuture,professionalcrowdworkEthics,nickerson2013crowd,dynamo}.
+
+Some research already looks at research such as investing in workers, and
+informally, we know that this happens among industry requesters
+\cite{jonBrelig,shepherdingDow}.
+AMT, meanwhile, offers requesters the ability to create tasks which are
+not just hidden from unqualified workers by default, but completely.
+Requesters have taken to using lists of worker IDs which reference
+workers who have proven their reliability,
+representing a sort of proto--organization of loosely connected workers.
+
+\subsubsubsection{\pieceworkpers}
+The rise of labor unions in the \nth{20} century seems to have been precipitated by
+egregiously unjust conditions imposed on workers in factories and elsewhere
+\cite{ebbinghaus1999institutions}.
+Incidents broadly describing this dynamic can be found in research on AMT
+\cite{turkopticon,dynamo}.
+If these are prototypical labor advocacy organizations of contemporary on--demand work,
+the next question we should look to is if
+--- and indeed \textit{how} ---
+these institutions might face challenges in the future.
+
+For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
+and identify that
+``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
+\cite{levi2009union}.
+Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
+% unable to satisfy their needs or otherwise more invested in the 
+This perception already appears to be emerging in digitally mediated peer--governed organizations,
+as \citeauthor{keegan2010egalitarians} and others have illustratively documented
+\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
+If these organizations and others are to avoid the same fate that labor unions faced,
+they should take care to study this phenomenon and attempt to avoid it.
+
+\subsubsubsection{\whatchanged}
+\topic{While online and distributed workers can be harder to find,
+other features make the study thereof rewarding for the substantively different ways
+that we can approach their study.}
+The longitudinal analysis of everyday discourse, for instance,
+is trivial in the study of online communities
+
+\subsubsubsection{\implication}
 
 \onlyinsubfile{
-  % \balance{}
 \printbibliography
-  \clearpage
-  % \nobalance{}
 }
-  \endnotetext{
-    \section{Flexibility notes}
-    A number of researchers have identified
-    worker attrition,
-    variability of worker performance,
-    and uncertainty about good versus bad--faith actors as open questions of crowdwork
-    \cite{MaliciousCrowdworkersGadiraju,Ipeirotis:2010:QMA:1837885.1837906}.
-    \ali{We can and should discuss the distinction between presumably ``bad faith'' workers
-    \& workers who are merely responding in kind to bad requesters
-    --- and the broader questions surrounding
-    the roles that requesters as well as workers should play ---
-    but let it suffice to say that
-    requesters have been trying to understand and manage what appears to them as inconsistent work.
-    Their ways of responding to that variance in work quality has largely involved
-    making the work more flexible and resilient to work
-    (although some work has gone into investigating the causes, rather than treating the symptoms)}
-
-    Earlier we discussed \citeauthor{cheng2015break}'s work
-    measuring the impact that interruption has on worker performance
-    \cite{cheng2015break}.
-    This work illustrates a broader sentiment in
-    both the study and practice
-    of crowdwork, that microtasks should be designed resiliently against the variability of workers,
-    fully exploiting the abstracted nature of each piece of work
-    \cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-    That is to say, micro--tasks should be designed such that a single worker's poor performance,
-    or a good worker's sudden departure,
-    does not significantly impact the agenda of the work as a whole.
-    While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
-    in the form of higher cumulative time to complete
-    (albeit much shorter real time to complete, owing to parallelization),
-    \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
-    similar tasks together
-    \cite[respectively]{cheng2015break,delayAndOrderLasecki}.
-
-    \citeauthor{embracingErrorKrishna} take a different approach;
-    by ``embracing error'' and forming models describing the latency of workers in classifying objects at rapid speeds,
-    the authors offer orders--of--magnitude improvements
-    in various binary classification tasks
-    \cite{embracingErrorKrishna}.
-    And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
-    some researchers have designed work predicated on the expectation of it instead:
-    \citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
-    crowd workers would never be given enough information to piece together sensitive information about
-    any single topic
-    \cite{sensitiveTasks}.
-
-    The work thus far seems to attempt to maximize the quality of work among workers through various means:
-    \begin{inlinelist}
-      \item Identifying ``bad'' workers (fraught with problems as this characterization is) \cite{MaliciousCrowdworkersGadiraju},
-      \item Designing tasks with break points to facilitate the on--boarding and off--boarding that happens anyway \cite{cheng2015break}, and
-      \item Expecting certain levels of attrition and incorrectness and using that variability to their advantage \cite{embracingErrorKrishna}
-    \end{inlinelist}.
-
-    Flexibility has been explored through the lens of Fordism, perhaps best illustrated by
-    \citeauthor{tolliday1986between}'s treatment describing
-    turnover rates rising above 300\% in the decade leading to the introduction of the assembly line in 1913.
-    Specifically, the utilization of ``\dots~`semi--special' machine tools which could be adapted
-    [and]~\dots~added flexibility through seasonal layoffs for production workers and the use of
-    piece rates~\dots~rather than a day wage system''
-    \cite{tolliday1986between}.
-
-    In the field of piecework,
-    the research covering this topic has both explored
-    a breadth of tasks that might be rendered doable by piecemeal workers
-    \textit{as well as} longitudinally documented the success of these approaches.
-    Here, we \ali{\dots?}
-}
-\theendnotes
 \end{document}

### 2016-09-13 drafts/labor/trackingWork.tex
complexity should be kind of done
lots of changes all over the place

@@
 \setlist{nolistsep}
-\def\plaintitle{Reexamining Crowd Work: A Historical Framing of On--Demand Labor as Piecework}
+\def\plaintitle{Examining Crowd Work Through The Lens of On--Demand Labor as Piecework} 
 \def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
@@
 \newcommand{\pieceworkpers}{Piecework's perspective.}
-\newcommand{\crowdworkpers}{Crowdwork's perspective.}
-\newcommand{\whatchanged}{What's changed.}
-\newcommand{\implication}{Implications for crowdwork.}
-\newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of ``payment for results''%
+\newcommand{\crowdworkpers}{Crowd work's perspective.}
+\newcommand{\whatchanged}{What's different about crowd work.}
+\newcommand{\implication}{Implications for crowd work research.}
+\newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of
+``payment for results''%
 , leading to and leveraging narrower skill sets required for narrowly defined tasks,
@@
 \endnotetext{
-  \section{The Bleak Future of Crowd Work}
-  We've traced a path from piecework itself through
-  the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
-  in tracing this process, we touched on the relationships between decomposition,
-  work \& worker abstraction,
-  flexibility, and followed through both the general fallout of crowd work in the research community
-  as well as the fallout between workers and the managers and other external parties
-  --- including researchers.
+  % \section{The Bleak Future of Crowd Work}
+  % We've traced a path from piecework itself through
+  % the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
+  % in tracing this process, we touched on the relationships between decomposition,
+  % work \& worker abstraction,
+  % flexibility, and followed through both the general fallout of crowd work in the research community
+  % as well as the fallout between workers and the managers and other external parties
+  % --- including researchers.
 
-  Throughout these case studies, we have pointed out the parallels between
-  the contemporary research in on--demand labor and
-  the much larger body of research constituting our understanding of topics such as
-  piecework, factory work, and laborer relations.
-  If we agree that this framing is useful and informative, then
-  several topics emerge as relatively open questions in the study of crowd work and on--demand labor.
-  Two of the most pressing questions are
-  \begin{inlinelist}
-  \item the beginnings of factories, and 
-  \item the decline of relevance of worker advocacy organizations.
-  \end{inlinelist}
-  We will discuss those questions here.
+  % Throughout these case studies, we have pointed out the parallels between
+  % the contemporary research in on--demand labor and
+  % the much larger body of research constituting our understanding of topics such as
+  % piecework, factory work, and laborer relations.
+  % If we agree that this framing is useful and informative, then
+  % several topics emerge as relatively open questions in the study of crowd work and on--demand labor.
+  % Two of the most pressing questions are
+  % \begin{inlinelist}
+  % \item the beginnings of factories, and 
+  % \item the decline of relevance of worker advocacy organizations.
+  % \end{inlinelist}
+  % We will discuss those questions here.
 
-  \subsection{The beginnings of factories}\label{sec:Factorization}
-  We established earlier that abstracted work and low wages tend to result in variable outcomes,
-  which presents problems for employers.
-  % So work output is variable, and that's bad. What are we doing about it?
-  Historically, this is what led to factories; by employing a cohort of known workers,
-  we can be reasonably assured that the quality of the work will be better than random.
-  Furthermore, we can invest more resources in training workers and
-  get workers to do more complex work with more context.
-  % you would pay some experts who understood what they needed to do
-  % and they would form a cohort into which you invest more.
+  % \subsection{The beginnings of factories}\label{sec:Factorization}
+  % We established earlier that abstracted work and low wages tend to result in variable outcomes,
+  % which presents problems for employers.
+  % % So work output is variable, and that's bad. What are we doing about it?
+  % Historically, this is what led to factories; by employing a cohort of known workers,
+  % we can be reasonably assured that the quality of the work will be better than random.
+  % Furthermore, we can invest more resources in training workers and
+  % get workers to do more complex work with more context.
+  % % you would pay some experts who understood what they needed to do
+  % % and they would form a cohort into which you invest more.
 
-  Some research already looks at research such as investing in workers, and
-  informally, we know that this happens among industry requesters
-  \cite{jonBrelig,shepherdingDow}.
-  AMT, meanwhile, offers requesters the ability to create tasks which are
-  not just hidden from unqualified workers by default, but completely.
-  Requesters have taken to using lists of worker IDs which reference
-  workers who have proven their reliability,
-  representing a sort of proto--organization of loosely connected workers.
-
-  This, then, suggests that the beginning of the regularization of workforces
-  --- a sort of coalescence of factories ---
-  is already happening.
-  If our framing of on--demand labor is accurately describing an underlying relationship with piecework,
-  then we should watch for the emergence and popularization of persistent teams of workers.
+  % This, then, suggests that the beginning of the regularization of workforces
+  % --- a sort of coalescence of factories ---
+  % is already happening.
+  % If our framing of on--demand labor is accurately describing an underlying relationship with piecework,
+  % then we should watch for the emergence and popularization of persistent teams of workers.
   % While this field site will likely prove exceedingly difficult to 
@@
 
-  \subsection{The decline of advocacy organizations}
-  The rise of labor unions in the \nth{20} century seems to have been precipitated by
-  severely unjust conditions imposed on workers in factories and elsewhere
-  \cite{ebbinghaus1999institutions}.
-  Incidents broadly describing this dynamic can be found in research on AMT
-  \cite{turkopticon,dynamo}.
-  If these are prototypical labor advocacy organizations of contemporary on--demand work,
-  the next question we should look to is if
-  --- and indeed \textit{how} ---
-  these institutions might face challenges in the future.
+  % \subsection{The decline of advocacy organizations}
 
-  For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
-  and identify that
-  ``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
-  \cite{levi2009union}.
-  Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
-  % unable to satisfy their needs or otherwise more invested in the 
-  This perception already appears to be emerging in digitally mediated peer--governed organizations,
-  as \citeauthor{keegan2010egalitarians} and others have illustratively documented
-  \cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
-  If these organizations and others are to avoid the same fate that labor unions faced,
-  they should take care to study this phenomenon and attempt to avoid it.
+  % \section{Implications for Design}
+  % If it's agreed that
+  % the major topics we've discussed thus far are related and
+  % --- at least to \textit{some} extent ---
+  % precipitated in the fashion we argue,
+  % then we have a rare opportunity as researchers,
+  % and as agents of change in the communities we study,
+  % to affect change on the dynamics of crowd and on--demand work
+  % as they continue to develop.
 
-  \section{Implications for Design}
-  If it's agreed that
-  the major topics we've discussed thus far are related and
-  --- at least to \textit{some} extent ---
-  precipitated in the fashion we argue,
-  then we have a rare opportunity as researchers,
-  and as agents of change in the communities we study,
-  to affect change on the dynamics of crowd and on--demand work
-  as they continue to develop.
+  % Without claiming to have easy, cut--and--dry solutions to these problems,
+  % we can nevertheless bring to attention a number of critical opportunities to
+  % learn from historical parallels in piecework and factory labor,
+  % and make informed decisions regarding whether
+  % (or indeed how)
+  % we may want to influence outcomes.
+  % The challenges we bring to attention here are as follows:
+  % \begin{inlinelist}
+  %   \item codifying investment toward collective goods into the designs of systems;
+  %   \item (re--)decentralizing the internet; and
+  %   \item enabling reputation transferral.
+  % \end{inlinelist}
 
-  Without claiming to have easy, cut--and--dry solutions to these problems,
-  we can nevertheless bring to attention a number of critical opportunities to
-  learn from historical parallels in piecework and factory labor,
-  and make informed decisions regarding whether
-  (or indeed how)
-  we may want to influence outcomes.
-  The challenges we bring to attention here are as follows:
-  \begin{inlinelist}
-    \item codifying investment toward collective goods into the designs of systems;
-    \item (re--)decentralizing the internet; and
-    \item enabling reputation transferral.
-  \end{inlinelist}
+  % \subsection{Codify the common good}
+  % As \citeauthor{lessig2006code} points out in his book,
+  % digital media give designers the opportunity to design and build into the systems
+  % policies and practices to contribute to the collective benefit of the people therein
+  % \cite{lessig2006code}.
+  % Historically, the confluence of forces \citeauthor{lessig2006code} describes
+  % would ultimately result in outcomes such as benefits for workers,
+  % funds for sick leave and vacation, and other conveniences.
+  % The transient nature of on--demand work would seem to problematize this arrangement,
+  % but we can discuss and explore the viability of building into systems the mechanisms necessary
+  % to save a portion of payment from every gig,
+  % record taxable income, or
+  % myriad other generally administrative tasks automatically.
 
-  \subsection{Codify the common good}
-  As \citeauthor{lessig2006code} points out in his book,
-  digital media give designers the opportunity to design and build into the systems
-  policies and practices to contribute to the collective benefit of the people therein
-  \cite{lessig2006code}.
-  Historically, the confluence of forces \citeauthor{lessig2006code} describes
-  would ultimately result in outcomes such as benefits for workers,
-  funds for sick leave and vacation, and other conveniences.
-  The transient nature of on--demand work would seem to problematize this arrangement,
-  but we can discuss and explore the viability of building into systems the mechanisms necessary
-  to save a portion of payment from every gig,
-  record taxable income, or
-  myriad other generally administrative tasks automatically.
+  % \subsection{Decentralize the internet --- again}
+  % Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
+  % forcing workers either to choose one or juggle participation in these markets with great difficulty.
+  % An ``API'' for on--demand labor markets could make it possible for any person
+  % or organization to instantiate their own marketplace and inter--operate with.
+  % This can be changed, and indeed must, if we are to realize the hopes of early researchers who advocated the
+  % democratizing nature and power of the internet
+  % \cite{barlow2009declaration,lanier2014owns}.
 
-  \subsection{Decentralize the internet --- again}
-  Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
-  forcing workers either to choose one or juggle participation in these markets with great difficulty.
-  An ``API'' for on--demand labor markets could make it possible for any person
-  or organization to instantiate their own marketplace and inter--operate with.
-  This can be changed, and indeed must, if we are to realize the hopes of early researchers who advocated the
-  democratizing nature and power of the internet
-  \cite{barlow2009declaration,lanier2014owns}.
+  % \subsection{Deal with reputation}
+  % Reputation systems in on--demand labor markets are fundamentally broken.
+  % To say nothing of the fact that information workers (such as those on AMT) can't transfer their reputations to qualitatively different forms of labor
+  % like driving--for--hire (e.g. Uber), even within the same industry
+  % it's currently not feasible for workers to transfer their reputations or other information from one place to another.
+  % This affects more than the reputation and trustworthiness of workers;
+  % accounting for things such as taxes, benefits, etc\dots is all but left to the individual workers, who struggle with myriad bureaucratic obstacles.
+  % We can design systems that facilitate the aggregation and, more importantly, the transferral of reputation, income, and other features of work.
 
-  \subsection{Deal with reputation}
-  Reputation systems in on--demand labor markets are fundamentally broken.
-  To say nothing of the fact that information workers (such as those on AMT) can't transfer their reputations to qualitatively different forms of labor
-  like driving--for--hire (e.g. Uber), even within the same industry
-  it's currently not feasible for workers to transfer their reputations or other information from one place to another.
-  This affects more than the reputation and trustworthiness of workers;
-  accounting for things such as taxes, benefits, etc\dots is all but left to the individual workers, who struggle with myriad bureaucratic obstacles.
-  We can design systems that facilitate the aggregation and, more importantly, the transferral of reputation, income, and other features of work.
+  % \section{Discussion}
+  % We've discussed a number of aspects of on--demand work that
+  % offer parallels with historical piecework.
+  % Perhaps more importantly, we've hopefully demonstrated that the dynamics we observe in on--demand work
+  % are interrelated and follow from one another just as necessarily as they did in
+  % the development and maturation of piecework and factory work through the \nth{20} century.
+  % This framing on on--demand work should, we hope, provide us with the necessary historical context
+  % to make better--informed design decisions about how we want ``the future of crowd work'' to look.
 
-  \section{Discussion}
-  We've discussed a number of aspects of on--demand work that
-  offer parallels with historical piecework.
-  Perhaps more importantly, we've hopefully demonstrated that the dynamics we observe in on--demand work
-  are interrelated and follow from one another just as necessarily as they did in
-  the development and maturation of piecework and factory work through the \nth{20} century.
-  This framing on on--demand work should, we hope, provide us with the necessary historical context
-  to make better--informed design decisions about how we want ``the future of crowd work'' to look.
-
-  \section{Conclusion}
-  \citeauthor{crowdworkFuture} discussed many of the challenges and problems in crowd work in \citeyear{crowdworkFuture},
-  but didn't necessarily situate the notion of crowd work in a broader context.
-  This paper attempts to fill that gap, and in doing so hopes
-  to give the research community theoretical grounding to work with and within on--demand labor more successfully.
-  But more than that, we hope to have addressed important questions to inform how we actually might make crowd work
-  a career in which we want our children to work.
+  % \section{Conclusion}
+  % \citeauthor{crowdworkFuture} discussed many of the challenges and problems in crowd work in \citeyear{crowdworkFuture},
+  % but didn't necessarily situate the notion of crowd work in a broader context.
+  % This paper attempts to fill that gap, and in doing so hopes
+  % to give the research community theoretical grounding to work with and within on--demand labor more successfully.
+  % But more than that, we hope to have addressed important questions to inform how we actually might make crowd work
+  % a career in which we want our children to work.
 }

### 2016-09-13 drafts/labor/complexity.tex

@@
 , often involving insights made in Computer Science and applied to human work--flows.}
-The crowd work literature typically identifies target milestones in CS
+
+The crowd work literature typically identifies target milestones in computer science
 that have presented significant challenges for researchers%
@@
 as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-\cite{hart2013rise,grier2013computers}.
-That is, we need to identify new ways to train crowd workers for uniquely narrowly defined work.
+\cite{hart2013rise,grier2013computers} \msb{remind us of what those were}.
+That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work. \msb{I don't understand: why?}
 To some extent, an argument can be made that

### 2016-09-13 drafts/labor/piecework_lit.tex

--- 2016-09-13 drafts/labor/piecework_lit.tex
@@
 Using the definition of piecework that we came up with earlier, we argue that
-crowdwork is fundamentally an instantiation of piecework, and
+crowd work is fundamentally an instantiation of piecework, and
 that we can more precisely anticipate the answers to the open research questions we discussed earlier.
-We'll show that the dimensions of crowdwork that the broader HCI community has been studying
+We'll show that the dimensions of crowd work that the broader HCI community has been studying
 align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.

### 2016-09-14 drafts/labor/complexity.tex
expanding on relationships

@@
 , informed in part by researchers --- acting as inspectors --- and experts
-\cite[quotations from][]{10.2307/23702539}
+\cite[quotations from][]{10.2307/23702539}.
 Where \citeauthor{10.2307/23702539} would call additionally for the identification of
@@
 \cite{uberAlgorithm,10.2307/23702539}.
-}
 

### 2016-09-14 drafts/labor/introduction.tex
lots of changes all over the place
expanding on relationships

@@
 the constituent parts of larger tasks.}
-This framing of work into \msb{modular, instead of abstract?} abstract blocks
+This framing of work into modular blocks
 has allowed people to engage in work despite
@@
 like Amazon's Mechanical Turk (AMT)
-to explore the limits of this distributed, \msb{fleeting sounds pejorative. on-demand? it sounds weird to flock to a site for a pejorative reason} fleeting workforce.
-Researchers in particular \msb{no more ``in particular'' than industry...Crowdflower is big, much bigger than research} have taken to the space in earnest,
+to explore the limits of this distributed, on--demand workforce.
+Researchers have taken to the space in earnest,
 finding opportunities to enable new forms of work
 using this population of ``Turkers''
-\note{as test subjects in a sort of at--scale laboratory for economic and other behavioral studies}
 \cite{bernsteinSoylent,redballoon,paolacci2010running}.
@@
               whyWouldAnyoneBrewer,takingAHITMcInnis}
-        \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
+        % \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
 \end{enumerate}

### 2016-09-14 drafts/labor/relationships.tex
expanding on relationships

@@
 are complex and not especially well--understood.}
-While researchers have begun to appreciate
-the sociality of crowd workers in labor markets,
-the study of these communities is made enormously more challenging by
-the limited access of online spaces 
+Researchers have begun to appreciate
+the sociality of crowd workers in labor markets
 \cite{crowdcollab}.
-Nevertheless, a number of interesting findings have
-already emerged in the crowd work literature%
-, which we can report here.
+Still, the study of these communities is made more challenging by
+the limited access to workers on these sites of work
+inherent to digital spaces made without social affordances
+\cite{miller2011understanding}.
+% Nevertheless, a number of interesting findings have
+% already emerged in the crowd work literature%
+% , which we can report here.
+We can break this general body of work into the three groups we named earlier:
+workers' relationships
+\begin{inlinelist}
+\item with \nameref{ch:workRelationship},
+\item with \textit{requesters}, and
+\item with \textit{other workers}.
+\end{inlinelist}
+We'll look at workers' relationships with work itself, which we'll discover
+gives us insight into why people engage in crowd work in the first place.
 
+\paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
 \topic{A number of ethical questions surrounding the
 increasing complexity of crowd work have arisen in recent years.}
-\citeauthor{professionalcrowdworkEthics} bring some of these issues at stake
+\citeauthor{professionalcrowdworkEthics} bring some of these issues to light
 --- working for increasing amounts of time on tasks of growing complexity%
@@
 but these and other dangers range an enormous landscape
-\cite{crowdworkFuture,professionalcrowdworkEthics,nickerson2013crowd,dynamo}.
+\cite{professionalcrowdworkEthics}.
+\citeauthor{crowdworkFuture}
+list a few of the problems they identified in \citeyear{crowdworkFuture}
+--- motivation, feedback, reputation, quality control, to name a few ---
+while others discuss challenges such as fostering collective action
+and the opportunity for learning and career advancement
+\cite{crowdworkFuture,nickerson2013crowd,dynamo}.
 
@@
 Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
-% unable to satisfy their needs or otherwise more invested in the 
 This perception already appears to be emerging in digitally mediated peer--governed organizations,

### 2016-09-14 drafts/labor/abstract.tex
abstract and title closed out

@@
 
+\onlyinsubfile{\usepackage{xr-hyper}
+\usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{relationships}
+\externaldocument{decomposition}
+}
 \makeatletter
@@
 \makeatother
-
 \begin{document}
-
+\onlyinsubfile{\plaintitle}
 \begin{abstract}
@@
 other forms of on--demand labor.
-A large and growing body of scholarship has sought to predict
+A large and growing body of scholarship has attempted to predict
 the socio--technical outcomes of this shift, especially along three threads: %. This question has been dominated by three threads:
@@
 This approach to understanding crowd work provides additional theoretical framing
-to make sense of a topic with which researchers continue to struggle to understand.
+to make sense of a topic with which researchers continue to struggle to understand,
+which we propose to use to answer broader questions surrounding the future of crowd work.
 \end{abstract}

### 2016-09-14 drafts/labor/abstract.tex
abstract revisions

@@
 \begin{inlinelist}
-\item \nameref{sec:complexity};
-\item \nameref{sec:decomposition}; and
-\item \nameref{sec:relationships}.
+\item \nameref{sec:complexity}?
+\item \nameref{sec:decomposition}? and
+\item \nameref{sec:relationships}?
 \end{inlinelist}
@@
 to understand how these questions might play out with modern crowd work.
-To do so, we identify the mechanisms that limited piecework historically, and
+To do so, we identify the mechanisms that enabled and limited piecework historically, and
 identify whether crowd work faces the same mechanism limits or might differentiate itself.
-This approach to understanding crowd work provides additional theoretical framing
-to make sense of a topic with which researchers continue to struggle to understand,
-which we propose to use to answer broader questions surrounding the future of crowd work.
+This approach introduces theoretical grounding that can help address
+challenging, persistent questions in crowd work,
+and suggests design interventions that may help
+avoid unintentionally repeating history.
 \end{abstract}

### 2016-09-14 drafts/labor/abstract.tex
pitfalls and mechanisms

@@
 To do so, we identify the mechanisms that enabled and limited piecework historically, and
-identify whether crowd work faces the same mechanism limits or might differentiate itself.
+identify whether crowd work faces the same pitfalls or might differentiate itself.
 This approach introduces theoretical grounding that can help address
@@
 and suggests design interventions that may help
-avoid unintentionally repeating history.
+avoid repeating history.
 \end{abstract}

### 2016-09-14 drafts/labor/abstract.tex
repeating history

@@
 This approach introduces theoretical grounding that can help address
-challenging, persistent questions in crowd work,
-and suggests design interventions that may help
-avoid repeating history.
+some of the most pernicious questions in crowd work,
+and suggests design interventions that learn from history
+rather than repeat it.
 \end{abstract}

### 2016-09-16 drafts/labor/introduction.tex
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)

--- 2016-09-06 drafts/labor/introduction.tex
@@
 \onlyinsubfile{
-  \hypersetup{
-    citecolor  = blue,
-    linkcolor  = JokeGreen,
+  \usepackage{xr-hyper}
+  \usepackage{hyperref}
+  \externaldocument{complexity}
+  \externaldocument{relationships}
+  \externaldocument{decomposition}
+  \externaldocument{piecework_lit}
 }
-}
+
+% \onlyinsubfile{
+%   \hypersetup{
+%     citecolor  = blue,
+%     linkcolor  = JokeGreen,
+%   }
+% }
+% \renewcommand{\topic}[1]{#1}
 
@@
 \begin{document}
-\section{Introduction}
+
+\section{Introduction}\label{sec:introduction}
+
 \topic{The past decade has seen a flourishing of on--demand work,
@@
 the constituent parts of larger tasks.}
-This framing of work into abstract blocks
+This framing of work into modular blocks
 has allowed people to engage in work despite
@@
 \cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
+We've seen on--demand work passing by a number of aliases:
+as \textit{crowd work} --- on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
+as \textit{gig work} --- typically involving platforms for one--off jobs, like driving, courier services, etc\dots;
+as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)on an ad hoc basis when they don't need it.
 The realization
 that complex tasks can be accomplished by directing and managing crowds of workers
-has spurred the research and and industry communities to flock to sites of labor
-like Amazon's Mechanical Turk (AMT)
-to explore the limits of this distributed, fleeting workforce.
-Researchers in particular have taken to the space in earnest,
+has spurred the research and industry communities to flock to sites of labor
+like AMT
+to explore the limits of this distributed, on--demand workforce.
+Researchers have taken to the space in earnest,
 finding opportunities to enable new forms of work
 using this population of ``Turkers''
-\note{as test subjects in a sort of at--scale laboratory for economic and other behavioral studies}
 \cite{bernsteinSoylent,redballoon,paolacci2010running}.
 
+\ali{this mega--paragraph may need to be reduced since it started as two, did very little/nothing to reduce it.}
 \topic{This form of work has grown considerably in size,
@@
 \cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-Today, on--demand work promises to become a \joke{yuge industry}
-\joke{[citation needed]}.
-
-% \topic{This matters because\dots WHYYYYYYYYYYYYYYY\\}
-% Indeed, this on--demand workforce has sparked interest
-% across industries ranging
-% driving for hire (for example Uber),
-% house--cleaning (Handy),
-% and generalized services (TaskRabbit)
-% \cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-% Today, a rapidly growing transient workforce is forming,
-% itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
-% to benefit from a latent pool of previously vetted workers
-% \cite{pewSharing}.
-
-\topic{For all the growth we've observed in this labor market,
-we have also seen a complicated and conflicted culture emerge among its constituent workers.}
+For all the growth we've observed in this labor market,
+we have also seen a complicated and conflicted culture emerge among its constituent workers.
 Researchers have made efforts to understand the people
@@
 \topic{The extant body of work has ostensibly sought to answer one underlying question:
-What does the future hold for work and those that do it?}
+What does the future hold for crowd work and those that do it?}
 Researchers have offered their input on this open question along three major threads:
 \begin{enumerate}
-  \item What are the limits of crowdsourcing?
+  \item \nameref{sec:complexity}?
         Specifically,
         \begin{inlinelist}
-          \item how complex can crowd work get? and
+          \item how complex are the goals that crowd work can accomplish? and
           \item how far will crowd work reach into the everyday lives of people?
+          \msb{which people? workers? requesters? do you mean, will recruiting crowd work be the domain of experts or everyday end users?}
+          \ali{I think I'm trying to ask how much of the world will do crowd work --- the whole world, or just representative slices of lots of different types of people (experts, non--experts, etc\dots)}
         \end{inlinelist}
@@
               Hahn:2016:KAB:2858036.2858364};
-  \item To what extent can we decompose work (especially into micro--tasks)?
+  \item \nameref{sec:decomposition}?
+  % To what extent can we decompose work (especially into micro--tasks)? \msb{How far can work be decomposed into smaller and smaller microtasks?}
         \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
               Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
-  \item What will work and the place of work look like for the workers?
+  \item \nameref{sec:relationships}?
         \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
               whyWouldAnyoneBrewer,takingAHITMcInnis}
+        % \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
 \end{enumerate}
@@
 gone as far as predict future developments.
+To the extent that the research community has explored the boundaries of crowd work,
+it nevertheless has not reflected on the underlying mechanisms determining those boundaries.
+Similarly, work directed toward the decomposition of tasks has begun to appreciate the limits of decomposition,
+but on the whole it hasn't considered the relationship between the decomposition of work and the complexity of work.
+% \ali{Mixed feelings about this because I think I'll need to go back into decomposition and
+% give a whole thing on \citeauthor{marx2012economic}
+% if I bring this up as a real failing of the crowd work research to date.
+% Thoughts?}
+While researchers are quickly picking up on the importance of
+the relationships between workers in these platforms,
+this research seems to be unfolding without the benefit of any of
+the inter--personal labor research that has been informing modern work for the better part of a century.
+% \ali{Overall: too long? too vague? overstating the shortcomings?}
 
 We offer a framing for crowd work spanning the aforementioned industries
-collectively as a contemporary instantiation of ``piecework''.
-Piecework as a metaphor for the type of work at hand is not new.
+collectively as a contemporary instantiation of \namerefl{sec:pieceworkArc} ---
+a work and payment structure which breaks tasks down into standalone contracts,
+wherein payment is made for \textit{work output}, rather than for \textit{time}.
+Piecework as a metaphor for crowd work is not new.
 Indeed,
@@
 
-More concretely, by looking at task--based or ``gig'' work as
+More concretely, by looking at crowd work as
 an instantiation (or even a continuation) of piecework,
@@
         and finally,
-  \item offer predictions of what social computing researchers,
+  \item offer predictions of what crowd work researchers,
         and workers themselves,
@@
 \end{inlinelist}
+For example,
+we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
+and come up with theoretically grounded mechanisms explaining the limitations researchers have struggled to overcome thus far.
+% \msb{what I think would really help this paragraph is a single great example, in one or two sentences, of how this kind of framing would help.
+``For example, we draw on ...
+to explain ...''.
+Reader goes AHA I see why you're doing this now!}
 
-We'll look at a broad range of cases under a number of major themes
-we propose as broadly describing the types of research being done in crowd work
-and more generally in what we argue is contemporary piecework.
-After validating this lens as a way of reasoning about on--demand labor,
-we'll attempt to use this perspective to suggest areas of research worth anticipating,
-and developments we should expect to see in the maturation of digitally mediated work.
+% We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
+% we propose as broadly describing the types of research being done in crowd work
+% and more generally in what we argue is contemporary piecework.
+We will look at the above questions in crowd work through the lens
+first of the body of research made by crowd work researchers,
+and then of the body of work piecework researchers have offered.
+After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
+we will identify differences between piecework as historically understood and
+digitally mediated crowd work as we experience it today,
+and how those differences influence the predictions that researchers in piecework have already made.
+% we will use this perspective to suggest areas of research worth anticipating,
+% and developments we should expect to see in the maturation of digitally mediated work.
 Finally, we will offer design implications based on this research.

### 2016-09-16 drafts/labor/trackingWork.tex
expanding on relationships
abstract and title closed out
compilation time indications
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)

@@
 \usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref,textcase}
 \usepackage[normalem]{ulem}
@@
 \usepackage[pdflang={en-US},pdftex]{hyperref}
-\usepackage{mfirstuc}
+\usepackage{mfirstuc,lastpage}
+\usepackage[en-GB]{datetime2}
+\usepackage{fancyhdr}
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
@@
 \makeatother
+
+\providecommand*{\namerefl}[1]{%
+\refused{#1}%
+\MakeTextLowercase{%
+\getrefbykeydefault{#1}{name}{??}%
+}%
+}
 
@@
 \setlist{nolistsep}
-\def\plaintitle{Examining Crowd Work Through The Lens of On--Demand Labor as Piecework} 
-\def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
+\def\plaintitle{Examining Crowd Work Through The Lens of Piecework}
+\def\plainauthor{Ali Alkhatib, Margaret Levi, Michael Scott Bernstein}
 \def\emptyauthor{}
-\def\plainkeywords{Crowdsourcing}
+\def\plainkeywords{Crowdsourcing; human computation; research vision}
 \def\plaingeneralterms{Documentation, Standardization}
@@
   colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
+  citecolor=blue, % black,
+  filecolor=blue, % black,
+  linkcolor=blue, % black,
   urlcolor=linkColor,
@@
 \newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!
-\onlyinsubfile{
-  \hypersetup{
-    citecolor  = blue,
-    linkcolor  = blue,
-  }
-}
+
+\let\internalNoteali\ali
+\let\internalNotemsb\msb
+\onlyinsubfile{\renewcommand{\ali}[1]{\onlyinsubfile{\internalNoteali{#1}}}}
+\onlyinsubfile{\renewcommand{\msb}[1]{\onlyinsubfile{\internalNotemsb{#1}}}}
+
+\DTMsetdatestyle{mmddyyyy}
+\DTMsetup{datesep=/}
+\pagestyle{fancy}
+\lfoot{Last compiled \today~at \DTMcurrenttime}
+\cfoot{}
+\rfoot{page \thepage/\pageref*{LastPage}}
+\fancyhead{}
+\pagenumbering{arabic}
+\renewcommand{\headrulewidth}{0pt}
+
+% \onlyinsubfile{
+%   \usepackage{xr-hyper}
+%   \usepackage{hyperref}
+% }
 
@@
 \author{%
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
-  \alignauthor{Leave Authors Anonymous\\
-    \affaddr{for Submission}\\
-    \affaddr{City, Country}\\
-    \email{e-mail address}}\\
+  \alignauthor{Ali Alkhatib\\
+    \affaddr{Stanford University}\\
+    \affaddr{Stanford, USA}\\
+    \email{ali.alkhatib@cs.stanford.edu}}\\
+  \alignauthor{Michael Scott Bernstein\\
+    \affaddr{Stanford University}\\
+    \affaddr{Stanford, USA}\\
+    \email{msb@cs.stanford.edu}}\\
+  \alignauthor{Margaret Levi\\
+    \affaddr{CASBS}\\
+    \affaddr{Stanford, USA}\\
+    \email{mlevi@stanford.edu}}\\
 }
-
+\author{
+\alignauthor{Anonymous for submission\\
+\affaddr{Affiliation fields filled}\\
+    \affaddr{to take representative space}\\
+    \email{of final document}}\\
+}
 \maketitle
-
-% Abstract time
 \subfile{abstract.tex}

### 2016-09-16 drafts/labor/piecework_lit.tex
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)
most notes on piecework_lit; will finish up tomorrow!

--- 2016-09-13 drafts/labor/piecework_lit.tex
@@
 \topic{The HCI community has used the term ``piecework'' to describe
-myriad instantiations of on--demand labor%
-, but this reference has generally been offered in passing.}
+myriad instantiations of on--demand labor,
+but this reference has generally been offered in passing.}
 As this paper principally traces a relationship between
 the historical piecework and the contemporary crowd work
-(or on--demand labor more generally)%
-, this casual familiarity with piecework may prove insufficient.
+(or on--demand labor more generally),
+this casual familiarity with piecework may prove insufficient.
 We'll more carefully discuss piecework in this section in order
@@
   \item define ``piecework'' as researchers in the topic understood it;
-  \item trace the rise of piecework at a very high level%
-,         identifying key figures and ideas during this time; and finally
-  \item look at the fall of piecework, such as it was%
-,         considering in particular
-        the factors that may have led to piecework's eventual demise
-\end{inlinelist}.
+  \item trace the rise of piecework at a very high level,
+        identifying key figures and ideas during this time; and finally
+  \item look at the fall of piecework, such as it was,
+        considering in particular
+        the factors that may have led to piecework's
+        \msb{(and crowd work's?)}
+        \ali{I wanted to leave the door open later to argue that
+        crowd work can change its trajectory and \textit{not have a demise}.
+        Is that too subtle here? I could have an aside like
+        ``(and help crowd work avoid the same fate)''. Thoughts?}
+        eventual demise.
+\end{inlinelist}
 
 \subsection{What was piecework?}
-\topic{While ``piecework'' has proven difficult to concretize from the literature%
-, we can trace a constellation of characteristics of piece work that recur throughout the literature.}
+\topic{While ``piecework'' has proven difficult to concretize from the literature,
+we can trace a constellation of characteristics of piece work that recur throughout the literature.}
 We'll follow the history of research, collecting
-descriptions%
-, examples, and
+descriptions,
+examples, and
 provided definitions of piecework, trying
@@
 
-\topic{One of the earliest definitions of piecework%
-, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
 \citeauthor{hughRaynbirdTaskWork} offers
@@
 \cite{hughRaynbirdTaskWork}.
-This description offers the first rudimentary definition of piecework
-from which the practice will grow for more than a century;
-piece work, as \citeauthor{hughRaynbirdTaskWork} offers,
-``depend[s] on the quantity of work done''.
-
-\citeauthor{10.2307/2338394} gives a more illustrative definition of piecework%
-, offering examples:
+\citeauthor{10.2307/2338394} gives a more illustrative definition of piecework,
+offering examples:
 ``\dots~payment is made for each hectare which is pronounced to be well ploughed~\dots~
@@
 This framing perhaps makes the most intuitive sense;
-``payment for results'', as \citeauthor{10.2307/2338394} calls it%
-, is not only common in practice, but well--studied in labor economics as well
+``payment for results'', as \citeauthor{10.2307/2338394} calls it,
+is not only common in practice, but well--studied in labor economics as well
 \cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
 
-\topic{It's worth acknowledging that --- as \citeauthor{hart2013rise} point out ---
-``this distinction [between piece--rates and time--rates] was not completely clear--cut''
+\topic{It's worth acknowledging that
+``this distinction [between piece--rates and time--rates] was not completely clear--cut'';
+indeed, we see work that adopts
+piece--rate compensation in some aspects and
+ time--rate compensation in others
 \cite{hart2013rise}.}
-The ``Rowan premium system''%
-, which essentially paid workers a base rate for time with the opportunity for additional pay associated with output,
+The ``Rowan premium system'',
+which essentially paid workers
+a base rate for time plus
+(the potential for) an additional pay dependent on output,
 was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
@@
 \cite{rowan1901premium}.
-Nevertheless,
-this work offers us an intuition for what piecework is, if not a bright--line rule.
+As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
+regardless of the worker's productive output
+\textit{as well as} an additional compensation tied to performance,
+workers under this regime were in some senses ``task--labourers'',
+and in other senses
+(more \ali{(familiar|conventional)})
+``day--labourers''.
+\ali{I want to come back to this later when we talk about Uber \& Lyft, and
+     how they increasingly offer to guarantee drivers an hourly rate
+     \cite[see][]{uberHourly,lyftHourly}.
+     That's why I'm nitpicking about this muddied system --- to avoid someone saying
+     ``it doesn't seem like piecework to me anymore if there's an hourly wage floor''}
 
@@
 he knows that all depends on his own diligence and perseverance\dots~[and]
-so long as he performs his work to the satisfaction of his master%
-, he is not under that control to which the day--labourer is always subject.''
+so long as he performs his work to the satisfaction of his master,
+he is not under that control to which the day--labourer is always subject.''
 \citeauthor{hughRaynbirdTaskWork} (and others, as we will see)
@@
 \cite{10.2307/3827491}.
-\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia%
-, finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
+\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
+finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
 ``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
@@
 resonates across decades, industries, and locales where piecework is found.
-We'll problematize this supposed advantage as we trace the history of piecework%
-, but for now we can say that piecework affords
-independence and some sense of locus
-otherwise unknown to workers.
+We'll problematize this supposed advantage as we trace the history of piecework,
+but for now we can say that piecework affords
+independence and some sense of autonomy
+new to people in the working class.
 
@@
 \cite{hart2013rise}.
+Piecework thus opened the door for people who previously couldn't participate in the labor market
+--- either for lack of training or for other reasons ---
+to do so, and to acquire job skills incrementally.
+Workers without conventional training
+--- like women, who had no such opportunities to engage in engineering and metalworking apprenticeships as men did ---
+could be trained very narrowly on a very tightly constrained task,
+demonstrate proficiency, and become experts in their own ways.
+% \msb{explain that in your own words now --- that the piecework infrastructure allows lower-skilled people to be used for skilled positions, by building the expertise into the workflow}
 
-\topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
-suffice for our needs.}
-Those needs being that the definition be
-\begin{inlinelist}
-\item faithful to the historical cases of piecework that we see in the scholarship; and
-\item relevant and informative to potential cases of piecework today.
-\end{inlinelist}
-We offer the following:
-that \pieceworkdefinition.
+% \topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
+% suffice for our needs.} \msb{Again, I don't think the goal for the reader should be to arrive at a definition through a narrative, they instead need to know how to think about piecework.
+This  paragraph could probably be cut}
+% Those needs being that the definition be
+% \begin{inlinelist}
+% \item faithful to the historical cases of piecework that we see in the scholarship; and
+% \item relevant and informative to potential cases of piecework today.
+% \end{inlinelist}
+% We offer the following:
+% that \pieceworkdefinition.
 
-\subsection{What was piecework's historical arc?}
+\ali{Without a summarizing paragraph here, I worry that people won't take essentially the topic sentences from each paragraph.
+That is, I want people to understand that piecework:
+\begin{enumerate}
+  \item paid workers for quantity of work done, rather than time done;
+  \item was sometimes a little muddily defined;
+  \item afforded workers freedom that they didn't have before; and
+  \item enabled people who didn't even engage in the formal labor force to participate.
+\end{enumerate}
+I wanted to kind of crystallize this, but if you think it's unnecessary, that's fine.}
+
+\subsection{The Historical Arc of Piecework}\label{sec:pieceworkArc}
 
@@
 \citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
-to calculate equations in order to better--predict the trajectories of various celestial bodies
+to calculate equations in order
+to better--predict the trajectories of various celestial bodies in the \nth{19} century
 \cite{grier2013computers}.
-While this approach didn't become an economic powerhouse as later examples would prove%
-, Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
+While this approach didn't become an economic powerhouse as later examples would prove,
+Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
 determining the extent to which work can be decomposed, and
@@
 That is, Airy found that he could train youths in elementary mathematics
-to complete the majority of the calculations he would otherwise have had to solve on his own%
-, and that the greater body of work could ultimately be completed sooner
+to complete the majority of the calculations he would otherwise have had to solve on his own,
+and that the greater body of work could ultimately be completed sooner
 if he arranged his work appropriately.
 
-\topic{Piecework took a circuitous path in its rise to the mainstream%
-% from the field into the home and finally into factories%
-, each time finding additional ways to leverage the advantages of piecework.}
+\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
+What I got out of it is that a bunch of people did piecework, but I don't know why or in fact why these are different than the sources we cite earlier.
+Can you hone it?}
+
+\ali{I wanted that paragraph to be about the rising popularity and application of piecework
+(especially as it approached its ``heyday'' (I like \cite{hart2013rise})),
+coming from humble beginnings as it found its footing.
+Given that intent\dots~
+
+Should I \textbf{refactor} or \textbf{rewrite}?}
+
+% \topic{Piecework took a circuitous path in its rise to the mainstream%
+% each time finding additional ways to leverage the advantages of piecework.}\msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?} \ali{Yeah, this is really weird; commenting out for now.}
 First applied to farm work, as
-\citeauthor{hughRaynbirdTaskWork} and others illustrate%
-, the practice remained relatively obscure until
+\citeauthor{hughRaynbirdTaskWork} and others illustrate,
+the practice remained relatively obscure until
 it was brought to the textile industry
 \cite{hughRaynbirdTaskWork}.
-At the turn of the \nth{20} century%
-, when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City%
-, \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes%
-, describing piecework comprehensively
-% ``The timeworkers' wages should be verified by the timekeeper's book.''
+At the turn of the \nth{20} century,
+when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+offering guidance on how best to manage pieceworkers
 \cite{riisOtherSideLives,norton1900textile}.
-Soon after,
-we saw the application of piecework systems in textile mills on the realization that
-``[pieceworkers in Italy] will work as many hours as it is possible for him to stand''
+\citeauthor{clark1908cotton}, for instance,
+relays his observations of textile mill pieceworkers and his realization that
+``When he works by the day the Italian operative wishes to leave before the whistle blows,
+but if he works by the piece he will work as many hours as it is possible for him to stand''
 \cite{clark1908cotton}.
-% By the turn of the \nth{20} century%
-% , \citeauthor{riisOtherSideLives} 
-% the poor working and living conditions that were pervasive in
-% an increasingly industrial United States
-% \cite{riisOtherSideLives}.
-% By the turn of the \nth{20} century,
 Best practices regarding the measurement and management of
@@
 what limited the adoption of piecework in industries that otherwise gravitated toward it
-(in the case studies he examined, this mostly focused on railway engineers)
+(in the case studies he examined, this mostly focused on railway engineers),
+ultimately arguing that factors such as the nature of the work design
+(specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
+are likely major contributing factors the limited use of piecework
 \cite{Brown01041990}.
 
-\topic{As increasing attention revealed problems in piecework as it related to workers%
-, workers themselves began to speak out about their frustration with this new regime.}
-It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work%
-, but this led to industry organizations representing
+\topic{As increasing attention revealed problems in piecework as it related to workers,
+workers themselves began to speak out about their frustration with this new regime.}
+It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
+but this led to industry organizations representing
 railway workers, mechanical engineers, and others contributing their myriad perspectives
@@
 
-\topic{Piecework became an important contributor to the war effort in the Second World War%
-, cementing its role not only in American factories, but in industrial work around the world.}
-While piecework began to catch on at the turn of the \nth{20} century%
-, the 1930s represented a boom for piecework on an unprecedented scale%
-, especially among engineering and metalworking industries.
+\topic{Piecework became an important contributor to the war effort in the Second World War,
+cementing its role not only in American factories, but in industrial work around the world.}
+While piecework began to catch on at the turn of the \nth{20} century,
+the 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.
 As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
@@
 as the ``heyday'' of the use of piecework.
-He attributes this to the shortage of male workers%
-, who would have gone through a conventional apprenticeship process
+He attributes this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
 affording them more comprehensive knowledge of the total scope of work.
@@
 
-\topic{Despite the intense growth of the piecework approach to renumeration%
-, this time was not without turmoil.}
+\topic{Despite the intense growth of the piecework approach to renumeration,
+this time was not without turmoil.}
 As previously discussed, a number of worker organizations weighed in on
@@
 
-\subsection{Why is piecework relevant to crowd work?}
+\subsection{Why is piecework relevant to crowdwork?}
+\msb{I assume this is forthcoming?}
 Using the definition of piecework that we came up with earlier, we argue that

### 2016-09-16 drafts/labor/introduction.tex
most notes on piecework_lit; will finish up tomorrow!
typo in intro

@@
   \item \nameref{sec:decomposition}?
-  % To what extent can we decompose work (especially into micro--tasks)? \msb{How far can work be decomposed into smaller and smaller microtasks?}
         \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
 but has not offered a persuasive framing for holistically explaining
-the developments in worker processes that researchers have developer, or
+the developments in worker processes that researchers have developed, or
 the phenomena in social environments we have observed;

### 2016-09-16 drafts/labor/introduction.tex
space issue

@@
 as \textit{gig work} --- typically involving platforms for one--off jobs, like driving, courier services, etc\dots;
-as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)on an ad hoc basis when they don't need it.
+as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)
+on an ad hoc basis when they don't need it.
 The realization

### 2016-09-16 drafts/labor/complexity.tex
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)
incremental changes on complexity - still more to fix

@@
 \documentclass[trackingWork]{subfiles}
+\onlyinsubfile{
+  \usepackage{xr-hyper}
+  \usepackage{hyperref}
+}
+
 \makeatletter
@@
 
-% \renewcommand{\topic}[1]{#1}
-\subsubsection[finding crowd work's limits]{Achieving greater complexity}\label{sec:complexity}
+\subsubsection[What are the limits of crowd work]{Identifying the Limits of Crowd Work}\label{sec:complexity}
 \subsubsubsection{\crowdworkpers}
@@
 The broader body of work has varied significantly in type
---- providing conversational assistants%
-, interpreting medical data, and
-telling coherent and compelling stories%
-, to name a few examples
+--- providing conversational assistants,
+interpreting medical data, and
+telling coherent and compelling stories,
+to name a few examples
 \cite{Lasecki:2013:CCC:2501988.2502057,mavandadi2012distributed,KimStoria}.
 
-\topic{This body of research has involved similar approaches to problems%
-, often involving insights made in Computer Science and applied to human work--flows.}
-
+\topic{This body of research has involved similar approaches to problems,
+often involving insights made in Computer Science and applied to human work--flows.}
 The crowd work literature typically identifies target milestones in computer science
-that have presented significant challenges for researchers%
-, leverages some of the approaches and insights that Computer Science researchers have already made
-(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge})%
-, and arranges humans as computational black boxes within those approaches and processes
+that have presented significant challenges for researchers,
+leverages some of the approaches and insights that Computer Science researchers have already made
+(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge}),
+and arranges humans as computational black boxes within those approaches and processes
 \cite[][and others]{crowdForgeKittur,foundry}.
@@
 For example, \textit{Foundry}'s tools for managing and arranging expert groups into a cohort
-allow researchers to convincingly argue that expert teams can be rapidly formed%
-, just like non--expert teams
+allow researchers to convincingly argue that expert teams can be rapidly formed,
+just like non--expert teams
 \cite{foundry}.
@@
 \subsubsubsection{\pieceworkpers}
-\msb{needs a transition to set the tone here. e.g., The piecework literature also examined the complexity limits possible with piecework.}
-\topic{The research into piecework makes the case that
-piecework has been limited principally by the challenges of human management and oversight.}
+\topic{Piecework researchers have found themselves trying to understand what characteristics limit piecework,
+or more precisely what has prevented piecework from becoming more prevalent.}
+The research into piecework makes the case that
+piecework has been limited principally by the challenges of human management and oversight.
 \citeauthor{10.2307/23702539} describes a case study in Santa Fe Railway, which
@@
 \cite{10.2307/23702539}.
-Returning to \citeauthor{hart2016rise}'s reflections on piecework's limitations,
-we recall the multidimensional problem
---- tasks comprising of numerous, sometimes conflicting, goals
-\cite{hart2016rise}.
-It would be reasonable, then, to infer that work like this
---- reasonably highly skilled work where quality is difficult to assess ---
-would be unsuitable for piecework. \msb{that conclusion came way too fast. give a little detail on the railway. how did they come to that conclusion?}
+\citeauthor{10.2307/23702539} reports on the hiring of Harrington Emerson,
+an ``efficiency expert'', who went on to develop a ``standard time'' for each task at the company
+informed by ``thousands of individual operations at the Topeka shops''.
+The cost of measuring workers in such excruciating detail at the turn of the \nth{20} century was undoubtedly immense,
+but this ``standard time'' value, which determined the pay that workers would earn for each task they do,
+was the only viable approach at the time to determine appropriate pay given the task
+\cite{10.2307/23702539}.
+But the repeated measurement of workers' time to complete tasks had shortcomings;
+for one thing, pay rates for rarer tasks
+were necessarily less certain than for more common tasks,
+which had the simple benefit of a larger sample size.
+\ali{Do I need ``for another thing\dots''? This paragraph is kind of huge already. Not sure how to break it up/down as it is.}
+One might conclude from \citeauthor{10.2307/23702539}'s observations that
+complex, creative work --- which is inherently heterogeneous and difficult to routinize ---
+would be unsuitable for piecework.
+% We might reflect on \citeauthor{hart2016rise}'s piecework's limitations,
+% and specifically the challenge of multidimensional work
+% --- tasks comprising of numerous, sometimes conflicting, goals
+% \cite{hart2016rise}.
 
-\msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
-\topic{\citeauthor{hart2016rise} and \citeauthor{10.2307/23702539}%
-, without acknowledging one another%
-, seemingly corroborate one another's conclusions at different levels of observation.}
+% It would be reasonable, then, to infer that work like this
+% --- reasonably highly skilled work where quality is difficult to assess ---
+% would be unsuitable for piecework.
+% \msb{that conclusion came way too fast. give a little detail on the railway. how did they come to that conclusion?}
+
+% \msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
+\topic{Determining appropriate pay rates, informed by the careful measurement of workers,
+isn't the only major challenge piecework faced; evaluation proved a limiting factor as well.}
 \citeauthor{10.2307/23702539} enumerates some of the roles required
-to facilitate piecework in the early \nth{20} century:
+to facilitate piecework in the early \nth{20} century --- among them
     ``\dots~piecework clerks, inspectors, and `experts'\dots''
-\cite{hart2016rise,10.2307/23702539}.
-\citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
-to be making differing claims about the limitations of piecework, but
-we argue that \citeauthor{10.2307/23702539}
-is simply making a more concrete observation illustrating the insight that
-\citeauthor{hart2016rise} later makes. \msb{what insight? I've lost the thread here.}
+\cite{10.2307/23702539}.
+% \citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
+% to be making differing claims about the limitations of piecework, but
+% we argue that \citeauthor{10.2307/23702539}
+% is simply making a more concrete observation illustrating the insight that
+% \citeauthor{hart2016rise} later makes.
 \citeauthor{10.2307/23702539} recognizes that
@@
 other experts to properly design and evaluate complex work.
-\citeauthor{hart2016rise} argues an ultimate limit to how far this can go;
+\citeauthor{hart2016rise} later makes a more concrete observation of this hurdle,
+as he argues an ultimate limit to how far this can go;
 at some point, evaluating multidimensional work output for quality
 (rather than for quantity) becomes infeasible.
+In his words,
+``if the quality of the output is more difficult to measure than the quantity,
+perhaps because of `difficult--to--observe' production techniques,
+then a piecework system is likely to encourage
+an over--emphasis on quantity produced and an under-emphasis on quality''
+\cite{hart2016rise}.
+This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
+and ultimately precipitated its downfall,
+especially with the increasing complexity of manufacturing work
+over the course of the \nth{20} century.
+\ali{just to be clear, this whole point is queuing up so that later I can be like
+``Oh hey computers are terrific at evaluating some stuff really quickly.
+And researchers in crowd work have done some work on workers evaluating other workers
+(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line) \cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
 
@@
 As \citeauthor{hart2013rise} described,
-the 1930s saw a flourishing of clever piecework job design out of necessity
-due to the fact that it was infeasible to provide new workers with
-the comprehensive education that was familiar to men
+the 1930s saw a flourishing of ingenious piecework design
+almost strictly out of necessity ---
+it was infeasible to provide new workers with
+the comprehensive education that was familiar to men through apprenticeships
 \cite{hart2013rise}.
@@
 a relatively specific mathematics curriculum
-\cite{grier2013computers}. \msb{so what are we to take out of this? that there is no limit? just our creativity?}
+\cite{grier2013computers}.
+Instead, the literature suggests that
+piecework is tightly limited only when the application of piecework follows
+a direct, perhaps even unimaginative,
+mapping from an time--based regime to an output--based one.
+When the work is redesigned from the ground up 
+--- as we see with mathematicians in the \nth{19} century and
+with the metalworking industry during the Second World War ---
+it seems that we don't yet know the limits of complexity with regard to piecework.
+% \msb{so what are we to take out of this? that there is no limit? just our creativity?}
+\ali{This is \textit{almost} making a prediction,
+but I wanted to take something punchy away from the stuff I cite here.}
 
-\topic{Piecework researchers also make claims regarding
-the organizations that benefit from piecework in the first place.} \msb{I think a reader wouldn't understand why that claim is relevant.
-Let's invert it, something like how the researchers point out that only certain kinds of organizations can effectively make use of piecework.}
-\citeauthor{Brown01041990} discusses the factors necessary for piecework to thrive:
-    ``\dots~incentive pay is less likely in jobs with
+\topic{Piecework researchers also argue that,
+in addition to constraints on the kind of \textit{work} that's amenable to piecework,
+only certain kinds of \textit{organizations} are amenable to piecework.}
+% \msb{I think a reader wouldn't understand why that claim is relevant.
+% Let's invert it, something like how
+% the researchers point out that
+% only certain kinds of organizations can effectively make use of piecework.}
+\citeauthor{Brown01041990} discusses the organizational factors necessary for piecework to thrive,
+arguing that piecework
+``\dots~ is less likely in jobs with
 a variety of duties than in jobs with a narrow set of routinized duties''
-\cite{Brown01041990} \msb{that's confusing, is incentive pay the same as piecework? can you cut the first part of the quote and give us the noun instead of forcing us to figure it out?}.
+\cite{Brown01041990}
+% \msb{that's confusing, is incentive pay the same as piecework?
+% can you cut the first part of the quote and
+% give us the noun instead of forcing us to figure it out?}.
 \citeauthor{10.2307/23702539} adds further, that
@@
 make a persuasive argument that piecework is limited in complexity by
-managerial overhead \msb{wait, wasn't that in a previous paragraph?
-that should be joined with the managerial overhead text above} and the fixed cost of adopting a piecework payment regime \msb{that seems like it's the new point here}.
+the capacity to endure managerial overhead while transitioning to a new system.
+\msb{wait, wasn't that in a previous paragraph?
+that should be joined with the managerial overhead text above}
+\ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
+the first is that there are limits on doing hard, ``difficult--to--observe'' work,
+and the second is that the cost propagates up to management,
+putting limits on who can run piecework systems.
+So\dots\\
+\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
+\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
+\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop}
 
@@
 This body of work identifies ways to transform a variety of duties comprising complex tasks
-and distills them into ``a narrow set of routinized duties''%
-, informed in part by researchers --- acting as inspectors --- and experts
+and distills them into ``a narrow set of routinized duties'',
+informed in part by researchers --- acting as inspectors --- and experts
 \cite[quotations from][]{10.2307/23702539}.
@@
 Researchers have identified significant growth of
-the hippocampal regions of the brains in veteran drivers%
-, generally understood to be responsible for spatial functions such as navigation
-\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625%
-,       skok2000managing,Woollett1407,woollett2011acquiring}.
+the hippocampal regions of the brains in veteran drivers,
+generally understood to be responsible for spatial functions such as navigation
+\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
+      skok2000managing,Woollett1407,woollett2011acquiring}.
 Services such as Google Maps \& Waze make it possible for
@@
 \onlyinsubfile{
+  \balance{}
   \printbibliography

### 2016-09-17 drafts/labor/introduction.tex
MSB intro edits

--- 2016-09-16 drafts/labor/introduction.tex
@@
 
-\topic{The past decade has seen a flourishing of on--demand work,
+\topic{The past decade has seen a flourishing of \textit{on--demand work},
 largely driven by the reformulation of work as
 the constituent parts of larger tasks.}
-This framing of work into modular blocks
-has allowed people to engage in work despite
-limited time,
-little to no awareness of the broader context of the work, and 
-(often) fleeting identities and associations
-\cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
-We've seen on--demand work passing by a number of aliases:
-as \textit{crowd work} --- on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
-as \textit{gig work} --- typically involving platforms for one--off jobs, like driving, courier services, etc\dots;
-as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)on an ad hoc basis when they don't need it.
-The realization
-that complex tasks can be accomplished by directing and managing crowds of workers
-has spurred the research and industry communities to flock to sites of labor
-like AMT
-to explore the limits of this distributed, on--demand workforce.
-Researchers have taken to the space in earnest,
-finding opportunities to enable new forms of work
-using this population of ``Turkers''
-\cite{bernsteinSoylent,redballoon,paolacci2010running}.
+This framing of work into de-contextualized, modular blocks
+enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+Distributed paid participants then engage in the work whenever their schedules allow,
+with little to no awareness of the broader context of the work, and 
+with (often) fleeting identities and associations
+\cite{martin2014being,uberAlgorithm}.
+In this paper, we use the term on--demand work to join a pair of related phenomena:%We've seen on--demand work passing by a number of aliases:
+(1) \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
+(2) \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
+%as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)
+%on an ad hoc basis when they don't need it.
+The realization that complex tasks can be accomplished by directing and managing these crowds of workers spurred industry to flock to sites of labor
+like AMT and Uber to explore the limits of this distributed, on--demand workforce.
+Researchers have also taken to the space in earnest,
+developing systems and designs that enable new forms of production~(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
 
-\ali{this mega--paragraph may need to be reduced since it started as two, did very little/nothing to reduce it.}
-\topic{This form of work has grown considerably in size,
-far beyond the domain of ``information work'' from which it first sprang.}
-While \citeauthor{howe2008crowdsourcing} described crowdsourcing as
-``outsourcing [work] to an undefined, generally large group of people in the form of an open call'',
-for years the instantiation of this work was limited to the utilization of
-human intelligence to process data and act on information
+\topic{As on--demand work has grown far beyond the domain of information work from which it first sprang, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who empower it.}
+Originally, \citeauthor{howe2008crowdsourcing} described crowdsourcing in general terms as
+``outsourcing [work] to an undefined, generally large group of people in the form of an open call''.
+However, for years its instantiation was limited to the utilization of
+human intelligence to process data, participate in scientific studies, and perform information work
 \cite{CrowdsourcingUserStudies,movieSummarizationWu,
       yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
-More recently, crowdsourcing of embodied work
---- driving, cleaning, for instance ---
-has become a focus of on--demand labor markets
+More recently, crowdsourcing of physically embodied work
+--- driving and cleaning, for instance ---
+has become a focus for on--demand labor markets
 \cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-For all the growth we've observed in this labor market,
-we have also seen a complicated and conflicted culture emerge among its constituent workers.
-Researchers have made efforts to understand the people
-that have gravitated toward crowdsourcing platforms
-since its emergence and popularization,
-but as the form of work has grown and changed, so too have the demographics of workers
-\cite{Ross,whoareNOTtheTurkers}.
+This growth prompted increasing efforts to understand the workers who gravitate toward these platforms~\cite{Ross,whoareNOTtheTurkers}.
 Some of this research has been motivated by the identification of the sociality of gig work,
 and the frustration and disenfranchisement that these systems embody
-\cite{turkopticon,dynamo}.
+\cite{turkopticon,martin2014being,takingAHITMcInnis}.
 Other work has focused on the \textit{outcomes} of this frustration,
 reflecting on the resistance workers express against digitally mediated labor markets
-\cite{uberAlgorithm}.
+\cite{uberAlgorithm,dynamo}.
 
-\topic{The extant body of work has ostensibly sought to answer one underlying question:
-What does the future hold for crowd work and those that do it?}
+\topic{This body of research has sought to answer one central question:
+What does the future hold for on--demand work and those that do it?}
 Researchers have offered their input on this open question along three major threads:
@@
         \begin{inlinelist}
-          \item how complex are the goals that crowd work can accomplish? and
-          \item how far will crowd work reach into the everyday lives of people?
-          \msb{which people? workers? requesters? do you mean, will recruiting crowd work be the domain of experts or everyday end users?}
-          \ali{I think I'm trying to ask how much of the world will do crowd work --- the whole world, or just representative slices of lots of different types of people (experts, non--experts, etc\dots)}
+          \item How complex are the goals that crowd work can accomplish?, and
+          \item What kinds of goals and industries may eventually utilize it?
         \end{inlinelist}
@@
 \end{enumerate}
+This research literature has largely sought to answer these questions by examining the present phenomenon.
+So far, it has not offered a framing for holistically explaining
+the developments in worker processes that researchers have developed, or
+the emergent phenomena in social environments;
+nor has any research, to our knowledge,
+gone as far as predict future developments.
 
 \subsection{Piecework as a lens to understand crowdsourcing}
-This large and growing body of research has conversed
-to varying degrees with labor scholarship,
-but has not offered a persuasive framing for holistically explaining
-the developments in worker processes that researchers have developed, or
-the phenomena in social environments we have observed;
-nor has any research, to our knowledge,
-gone as far as predict future developments.
-To the extent that the research community has explored the boundaries of crowd work,
-it nevertheless has not reflected on the underlying mechanisms determining those boundaries.
-Similarly, work directed toward the decomposition of tasks has begun to appreciate the limits of decomposition,
-but on the whole it hasn't considered the relationship between the decomposition of work and the complexity of work.
+% MSB Cutting because it seems like it's just attacking the literature unnecessarily...
+% ----
+%This large and growing body of research has conversed
+%to varying degrees with labor scholarship,
+% ...
+%To the extent that the research community has explored the boundaries of crowd work,
+%it nevertheless has not reflected on the underlying mechanisms determining those boundaries.
+%Similarly, work directed toward the decomposition of tasks has begun to appreciate the limits of decomposition,
+%but on the whole it hasn't considered the relationship between the decomposition of work and the complexity of work.
 % \ali{Mixed feelings about this because I think I'll need to go back into decomposition and
@@
 % Thoughts?}
-While researchers are quickly picking up on the importance of
-the relationships between workers in these platforms,
-this research seems to be unfolding without the benefit of any of
-the inter--personal labor research that has been informing modern work for the better part of a century.
+%While researchers are quickly picking up on the importance of
+%the relationships between workers in these platforms,
+%this research seems to be unfolding without the benefit of any of
+%the inter--personal labor research that has been informing modern work for the better part of a century.
 % \ali{Overall: too long? too vague? overstating the shortcomings?}
-
-We offer a framing for crowd work spanning the aforementioned industries
-collectively as a contemporary instantiation of \namerefl{sec:pieceworkArc} ---
+In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework}: %\namerefl{sec:pieceworkArc} ---
 a work and payment structure which breaks tasks down into standalone contracts,
@@
 the framing of on--demand labor as a re--instantiation of piecework
-gives us more material to make sense of the broader research on this new form of work
-by evaluating this work through a much more refined theoretical lens,
-informed by decades of rigorous, empirically based research.
+gives us years of historical material to make sense of the broader research on this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.
 
-More concretely, by looking at crowd work as
+More concretely, by looking at on--demand work as
 an instantiation (or even a continuation) of piecework,
@@
   \item make sense of the phenomena so far as part of a much larger series of interrelated events;
-  \item bring into focus the ongoing work among
+  \item reflect on similarities in the ongoing work among
         workers,
@@
         and finally,
-  \item offer predictions of what crowd work researchers,
+  \item to the extent that history repeats itself, offer predictions of what on--demand work researchers,
         and workers themselves,
-        should expect to see on the horizon of on--demand work.
+        should expect to see on the horizon.
 \end{inlinelist}
-For example,
-we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
-and come up with theoretically grounded mechanisms explaining the limitations researchers have struggled to overcome thus far.
+For example, we will draw on the piecework literature such as case studies of the Santa Fe Railway to understand the historical complexity limits in piecework, and leverage that understanding to suggest which modern complexity limits in crowd work~\cite{crowdworkFuture} may be fundamental and which may be overcome.
+%we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
+%and come up with theoretically grounded mechanisms explaining the limitations researchers have struggled to overcome thus far.
 % \msb{what I think would really help this paragraph is a single great example, in one or two sentences, of how this kind of framing would help.
@@
 
+We organize this paper as follows: we first review the literature on piecework to lay  groundwork and make clear the analogy to on--demand work.
+Then, we interrogate the three major research questions above from a piecework frame.
+We will
 % We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
@@
 % and more generally in what we argue is contemporary piecework.
-We will look at the above questions in crowd work through the lens
-first of the body of research made by crowd work researchers,
-and then of the body of work piecework researchers have offered.
-After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
-we will identify differences between piecework as historically understood and
-digitally mediated crowd work as we experience it today,
-and how those differences influence the predictions that researchers in piecework have already made.
+%We will look at the above questions in crowd work through the lens
+%first of the body of research made by crowd work researchers,
+%and then of the body of work piecework researchers have offered.
+%After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
+we will identify similarities and differences between piecework as historically understood and
+on--demand work as we experience it today.
+Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
 % we will use this perspective to suggest areas of research worth anticipating,
 % and developments we should expect to see in the maturation of digitally mediated work.
-Finally, we will offer design implications based on this research.
+Finally, we will offer design implications for researchers and practitioners based on our results.
 

### 2016-09-17 drafts/labor/piecework_lit.tex
MSB history edits

@@
 We'll more carefully discuss piecework in this section in order
-to inform the subsequent sections --- and indeed, the entire argument.
+to inform the the rest of the argument.
 Specifically, we will
@@
         the factors that may have led to piecework's
-        \msb{(and crowd work's?)}
-        \ali{I wanted to leave the door open later to argue that
-        crowd work can change its trajectory and \textit{not have a demise}.
-        Is that too subtle here? I could have an aside like
-        ``(and help crowd work avoid the same fate)''. Thoughts?}
-        eventual demise.
+        eventual demise \msb{did it demise? there are sweatshops around the world today}.
 \end{inlinelist}
@@
 \subsection{What was piecework?}
-\topic{While ``piecework'' has proven difficult to concretize from the literature,
-we can trace a constellation of characteristics of piece work that recur throughout the literature.}
-We'll follow the history of research, collecting
+\topic{Aligning on--demand work with piecework requires an understanding of what piecework is. While ``piecework'' has had multiple definitions over time,
+we can trace a constellation of characteristics that recur throughout the literature.}
+We will follow this history of research, collecting
 descriptions,
@@
 \topic{It's worth acknowledging that
-``this distinction [between piece--rates and time--rates] was not completely clear--cut'';
+``this distinction [between piece--rates and time--rates] was not completely clear--cut'' \msb{cite? who said that? Hart?};
 indeed, we see work that adopts
@@
 Returning to
-\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piece work%
+\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piece work \msb{inconsistent use of ``piece work'' vs. ``piecework'' in the non-quoted text. Which do you want?}%
 crop up; he points out that\dots
@@
 \topic{We see this sense of independence regardless of the time, locale, and industry.}
-\citeauthor{10.2307/3827491} offers us a look into the lives and culture of ``match girls''
+\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match girls''
 --- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
-Of particular interest was their reputation ``\dots~for generosity, independence, and protectiveness,
+Of particular interest was their independent nature, via their reputation ``\dots~for generosity, independence, and protectiveness,
 but also for brashness, irregularity, low morality, and little education''
@@
 
-\ali{Without a summarizing paragraph here, I worry that people won't take essentially the topic sentences from each paragraph.
-That is, I want people to understand that piecework:
+In summary, piecework:
 \begin{enumerate}
-  \item paid workers for quantity of work done, rather than time done;
-  \item was sometimes a little muddily defined;
-  \item afforded workers freedom that they didn't have before; and
-  \item enabled people who didn't even engage in the formal labor force to participate.
+  \item paid workers for quantity of work done, rather than time done, but occasionally mixed the two payment models;
+  \item afforded workers freedom in when and how much to work; and
+  \item structured tasks such that people who didn't have the training to engage in the traditional labor force could still participate.
 \end{enumerate}
-I wanted to kind of crystallize this, but if you think it's unnecessary, that's fine.}
 
@@
 \cite{grier2013computers}.
-While this approach didn't become an economic powerhouse as later examples would prove,
-Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
+While this approach didn't become the same economic powerhouse as later examples would,
+Airy \msb{This is the first time Airy comes up. Who are they?} and others arguably found the kernel of insight that we pursue throughout this discussion:
 determining the extent to which work can be decomposed, and
@@
 
-\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
-What I got out of it is that a bunch of people did piecework, but I don't know why or in fact why these are different than the sources we cite earlier.
-Can you hone it?}
-
-\ali{I wanted that paragraph to be about the rising popularity and application of piecework
-(especially as it approached its ``heyday'' (I like \cite{hart2013rise})),
-coming from humble beginnings as it found its footing.
-Given that intent\dots~
-
-Should I \textbf{refactor} or \textbf{rewrite}?}
-
 % \topic{Piecework took a circuitous path in its rise to the mainstream%
 % each time finding additional ways to leverage the advantages of piecework.}\msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?} \ali{Yeah, this is really weird; commenting out for now.}
-First applied to farm work, as
+Piecework then began to grow: first applied to farm work, as
 \citeauthor{hughRaynbirdTaskWork} and others illustrate,
 the practice remained relatively obscure until
-it was brought to the textile industry
+it blossomed in the textile industry
 \cite{hughRaynbirdTaskWork}.
-At the turn of the \nth{20} century,
-when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
-\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+This growth was so marked that by the turn of the \nth{20} century,
+\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
 offering guidance on how best to manage pieceworkers
@@
 \cite{clark1908cotton}.
-Best practices regarding the measurement and management of
+During this period, best practices regarding the measurement and management of
 piecework rates, and of workers in the engineering industry,
@@
 
-\topic{Researchers have since struggled
-to understand the mechanisms and characteristics in piecework
-which fueled its rise to popularity during this time.}
+\topic{Researchers sought to understand the characteristics of piecework
+that fueled its rise to popularity.}
 \citeauthor{10.2307/23702539} argued that the first sparks of scientific management
@@
 (specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
-are likely major contributing factors the limited use of piecework
+were the major contributing factors that determined the use of piecework
 \cite{Brown01041990}.
 
-\topic{As increasing attention revealed problems in piecework as it related to workers,
-workers themselves began to speak out about their frustration with this new regime.}
-It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
-but this led to industry organizations representing
-railway workers, mechanical engineers, and others contributing their myriad perspectives
-\cite{american1921problem,richards1904anything,riisOtherSideLives}.
-Nevertheless, piecework continued to permeate low--skilled labor.
-
-\topic{Piecework became an important contributor to the war effort in the Second World War,
+\topic{At this point, piecework became an important factor in the war effort for the Second World War,
 cementing its role not only in American factories, but in industrial work around the world.}
-While piecework began to catch on at the turn of the \nth{20} century,
-the 1930s represented a boom for piecework on an unprecedented scale,
+The 1930s represented a boom for piecework on an unprecedented scale,
 especially among engineering and metalworking industries.
@@
 
+\topic{As increasing attention revealed problems in piecework as it related to workers,
+workers themselves began to speak out about their frustration with this new regime.}
+It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
+but this led to industry organizations representing
+railway workers, mechanical engineers, and others contributing their perspectives
+\cite{american1921problem,richards1904anything,riisOtherSideLives}.
+%Nevertheless, piecework continued to permeate low--skilled labor.
 % \topic{Necessity drove the widespread adoption of piecework and work decomposition%
 % , perhaps more even than Ford's assembly line.}
-
-\topic{Despite the intense growth of the piecework approach to renumeration,
-this time was not without turmoil.}
-As previously discussed, a number of worker organizations weighed in on
-(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
-\cite{american1921problem,richards1904anything}.
-\citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
+%\topic{Despite the intense growth of the piecework approach to renumeration,
+%this time was not without turmoil.}
+%These worker organizations weighed in on
+%(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
+%\cite{american1921problem,richards1904anything}.
+For example, \citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
 \cite{10.2307/3827491}.
@@
 \cite{10.2307/41829256}.
-The adoption of piecework time--study and other principles associated with Taylor and scientific management
-itself reliably precipitated strikes and more generally gave workers a clear enemy
-against which to rally
+The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
 \cite{jacoby1983union}.
 
-\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed just a few decades earlier.}
+\topic{However, piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
 \citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
 In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
-We summarize some of the salient suggestions here:
+The salient suggestions include:
 \begin{inlinelist}
@@
 
-\subsection{Why is piecework relevant to crowdwork?}
-\msb{I assume this is forthcoming?}
-Using the definition of piecework that we came up with earlier, we argue that
-crowd work is fundamentally an instantiation of piecework, and
-that we can more precisely anticipate the answers to the open research questions we discussed earlier.
-We'll show that the dimensions of crowd work that the broader HCI community has been studying
-align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
+\subsection{From piecework to on--demand work}
+Crowd work and gig work are fundamentally an instantion of piecework.
+First, workers on platforms such as Mechanical Turk and Uber are generally incentivized by unit of work, even if some may be offered an hourly base salary as well.
+Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
+Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+
+Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
+In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.
 

### 2016-09-19 drafts/labor/complexity.tex
pushing whatever i have...

@@
 % \citeauthor{hart2016rise} later makes.
-\citeauthor{10.2307/23702539} recognizes that
+\citeauthor{10.2307/23702539} further recognizes that
 it's necessary for a successful piecework shop to employ
@@
 And researchers in crowd work have done some work on workers evaluating other workers
-(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line) \cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
+(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line)
+\cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
+
+\topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
+particularly because maximizing on the advantages of piecework seemed to reward
+smaller, more constrained, more narrowly trained tasks.}
+For most of the \nth{19} century,
+piecework was applied almost exclusively to farm and textile work.
+% \citeauthor{grier2013computers} describes Airy's training and employment of young men to compute astronomical observations,
+% mentioning that boys needed only to have ``the basic skills of mathematics''
+% \cite{grier2013computers}.
+% \citeauthor{hart2013rise} argue that piecework flourished in the first half of the \nth{20} century because
+% tasks were decomposed --- a discussion we will return to later ---
+% such that untrained workers could be trained relatively quickly on a very narrowly designed task
+% \cite{hart2013rise}.
+Work was simple and widely understood
+--- farm workers didn't need to be trained on how to plow fields, or birth foals;
+seamstresses knew how to sew together denim
+\cite{10.2307/2338394,riisOtherSideLives}.
+% The general argument for decomposition, we will see later,
+% hinges on the promise that
+% tasks can be \textit{made} less complex and time--consuming,
+% if they're not already so.
+% \ali{this section is here because i later talk about the complexity of work, and
+% I want to keep the whole thing about how Uber, Google Maps, etc\dots make it possible to
+% \begin{inlinelist}
+% \item train workers more quickly (honestly just transmitting instructions over the Internet is so much quicker than what Airy had), and
+% \item give workers necessary tools/information to get work done
+% (e.g. ``Waze knows you should avoid the intersection 3 blocks down the road because of an accident''
+% and (maybe) ``not great at math? WolframAlpha can confirm or just give you your answer'')
+% \end{inlinelist}.}
 
@@
 indeed, we've discussed complex applications of crowd work already.}
-As \citeauthor{hart2013rise} described,
-the 1930s saw a flourishing of ingenious piecework design
-almost strictly out of necessity ---
+% As \citeauthor{hart2013rise} described,
+While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design,
+much of it arose out of necessity ---
+% almost strictly out of necessity ---
 it was infeasible to provide new workers with
@@
 \cite{hart2013rise}.
-This constraint led to much more tightly scoped work, and
-(perhaps surprisingly at the time) more efficient allocations of workers,
-who could now specialize in extremely narrowly defined roles.
+While this constraint led to much more tightly scoped work
+% (perhaps surprisingly at the time) more efficient allocations of workers,
+who now had to specialize in extremely narrowly defined roles.
 The same could be said of Airy and his \textit{computers}
@@
 \cite{grier2013computers}.
-Instead, the literature suggests that
+Instead, we argue that the literature suggests that
 piecework is tightly limited only when the application of piecework follows
@@
 \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
-\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop}
+\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop
+
+Maybe I should just merge it with the next paragraph?}
 
@@
 understanding when and how to modify work as necessary
-\cite{wray1949marginal}. \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
+\cite{wray1949marginal}.
+% \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
+The management of pieceworkers %, \citeauthor{10.2307/2118435} and later \citeauthor{wray1949marginal} argue,
+demanded people in positions akin to foremen
+--- intermediate managers people who were
+\begin{inlinelist}
+  \item familiar with and even sympathetic to the needs of workers,
+  \item empowered by higher level management to make decisions, and
+  \item relaxed enough in day--to--day work to allow workers to go about their work
+\end{inlinelist}
+\cite{wray1949marginal,10.2307/2118435}.
 
@@
 You're about to draw on those points to make your argument, so they need to be at the top of my mind here.}
+\topic{Piecework makes a number of observations leading to the conclusion that
+piecework's complexity is fundamentally bounded by several limitations,
+chief among them the costs of managerial overhead and the transition thereto.}
+\citeauthor{Brown01041990} and \citeauthor{10.2307/23702539}'s claims that
+organizations can't adopt piecework unless
+they're sufficiently large to absorb the cost of transitioning to a piecework system;
+\citeauthor{10.2307/2118435} and \citeauthor{wray1949marginal}'s observations for
+the importance of competent, effective managerial oversight
+--- a human resource, which made the scaling cost prohibitively expensive for many
+% in their times (\citeyear{10.2307/2118435} and \citeyear{wray1949marginal}, respectively)
+\cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
+\ali{Something like this? Should I dig deeper?
+I can line these points up in a way to make the next paragraph sort of obvious or inevitably,
+depending on how actively engaged someone is while reading this\dots}
+
 \topic{Digital media have expanded the scope of
@@
 the nature of the work is limited in complexity
-\cite{Brown01041990} \msb{I don't recall this point in the prior section: I remember management overhead and fixed costs of materials, I don't remember a paragraph about complexity.
-If we want to draw on that point, make it in the earlier section}.
+\cite{Brown01041990}.
+% \msb{I don't recall this point in the prior section:
+% I remember management overhead and fixed costs of materials,
+% I don't remember a paragraph about complexity.
+% If we want to draw on that point, make it in the earlier section}
 Given the flourishing of on--demand labor platforms such as
@@
 the Internet has trivialized
-the costs and challenges of the earlier limiting factors for two reasons:
+the costs and challenges of the earlier limiting factors %  for two reasons:
+% \begin{inlinelist}
+  % \item 
+  because technology make it easier
   \begin{inlinelist}
-  \item Technology make it much easier to do complex work aided by computers; and % is there a source for this? Is this necessary to cite?
-  \item The Internet allows us to leverage the benefits of
-        ``economies of scale'' at very little cost
-        to the system--designer \cite{lessig2006code,miller2011understanding} \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
-\end{inlinelist}.
+  \item to do complex work aided by computers and
+  \item to evaluate and manage workers as they do increasingly complex work,
+        even observing their work to an otherwise unprecedented granularity.
+  \end{inlinelist}
+
+%   \item The Internet allows us to leverage the benefits of
+%         ``economies of scale'' at very little cost
+%         to the system--designer \cite{lessig2006code,miller2011understanding}.% \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
+% \end{inlinelist}
 

### 2016-09-19 drafts/labor/introduction.tex
pushing whatever i have...

@@
 the constituent parts of larger tasks.}
-This framing of work into de-contextualized, modular blocks
-enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+This framing of work into de--contextualized, modular blocks
+enables computation to hire workers at scale through open calls on the Internet
+\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
 Distributed paid participants then engage in the work whenever their schedules allow,
@@
 \cite{martin2014being,uberAlgorithm}.
-In this paper, we use the term on--demand work to join a pair of related phenomena:%We've seen on--demand work passing by a number of aliases:
-(1) \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
-(2) \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
-%as the \textit{sharing economy} --- where participants can rent out their valuables (such as a car, a home, etc\dots)
-%on an ad hoc basis when they don't need it.
+In this paper, we use the term on--demand work to join a pair of related phenomena:
+\begin{inlinelist}
+\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
+\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
+\end{inlinelist}
 The realization that complex tasks can be accomplished by directing and managing these crowds of workers spurred industry to flock to sites of labor
@@
 Researchers have also taken to the space in earnest,
-developing systems and designs that enable new forms of production~(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
+developing systems and designs that enable new forms of production
+(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
 
@@
               whyWouldAnyoneBrewer,takingAHITMcInnis}
-        % \ali{is this about collective action? yes; also, career growth? education? governance? sociality? wages?}
 \end{enumerate}

### 2016-09-19 drafts/labor/piecework_lit.tex
pushing whatever i have...

--- 2016-09-16 drafts/labor/piecework_lit.tex
@@
 \documentclass[trackingWork]{subfiles}
-
-% \renewcommand{\topic}[1]{#1}
-
 \makeatletter
@@
 \begin{document}
+
 \section{A Review of Piecework}
+
 \topic{The HCI community has used the term ``piecework'' to describe
@@
 \subsection{What was piecework?}
+
 \topic{While ``piecework'' has proven difficult to concretize from the literature,
@@
 demonstrate proficiency, and become experts in their own ways.
-% \msb{explain that in your own words now --- that the piecework infrastructure allows lower-skilled people to be used for skilled positions, by building the expertise into the workflow}
-
-% \topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
-% suffice for our needs.} \msb{Again, I don't think the goal for the reader should be to arrive at a definition through a narrative, they instead need to know how to think about piecework.
-This  paragraph could probably be cut}
-% Those needs being that the definition be
-% \begin{inlinelist}
-% \item faithful to the historical cases of piecework that we see in the scholarship; and
-% \item relevant and informative to potential cases of piecework today.
-% \end{inlinelist}
-% We offer the following:
-% that \pieceworkdefinition.
 
@@
 \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
-What I got out of it is that a bunch of people did piecework, but I don't know why or in fact why these are different than the sources we cite earlier.
+What I got out of it is that a bunch of people did piecework, but
+I don't know why or in fact why these are different than the sources we cite earlier.
 Can you hone it?}
@@
 \ali{I wanted that paragraph to be about the rising popularity and application of piecework
-(especially as it approached its ``heyday'' (I like \cite{hart2013rise})),
+(especially as it approached its ``heyday'' (a la \cite{hart2013rise})),
 coming from humble beginnings as it found its footing.
@@
 
-% \topic{Piecework took a circuitous path in its rise to the mainstream%
-% each time finding additional ways to leverage the advantages of piecework.}\msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?} \ali{Yeah, this is really weird; commenting out for now.}
-First applied to farm work, as
+\topic{First applied to farm work, as
 \citeauthor{hughRaynbirdTaskWork} and others illustrate,
@@
 it was brought to the textile industry
-\cite{hughRaynbirdTaskWork}.
+\cite{hughRaynbirdTaskWork}.}
 At the turn of the \nth{20} century,
@@
 
-% \topic{Necessity drove the widespread adoption of piecework and work decomposition%
-% , perhaps more even than Ford's assembly line.}
-
 \topic{Despite the intense growth of the piecework approach to renumeration,
@@
 \cite{hart2013rise}.
-% the gendered and stratified dynamics at play in the matchgirls strike
-
-% \topic{Piecework declined in popularity in the United States, but
-% its spirit thrived through the mid--\nth{20} century in
-% scientific management and Fordism.}
-% Concurrent with the fall of piecework was arguably the rise of factories as centers of work.
 
 \subsection{Why is piecework relevant to crowdwork?}
+
 \msb{I assume this is forthcoming?}

### 2016-09-19 drafts/labor/relationships.tex
abstract and title closed out
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)
pushing whatever i have...

@@
 
-\subsubsection[the relationships of workers]{The Relationships of Workers to Work, Peers, and Others}
+\subsubsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}
 \label{sec:relationships}
@@
 \begin{inlinelist}
-\item with \nameref{ch:workRelationship},
+\item with \textit{work}, % \nameref{ch:workRelationship},
 \item with \textit{requesters}, and
@@
 
-\paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
+\ali{
+
+\subsubsubsection{\crowdworkpers}
+\begin{enumerate}
+  \item with work: what is this to workers? is it social, as \citeauthor{Jung:2016:SFG:2851581.2892328} says?
+        it seems like no according to \citeauthor{fort2011amazon}.
+        a few have tried to understand this from a turker's view
+        (see \citeauthor{martin2014being} and bigham's blog post);
+        generally trying to understand why people would do this work (\citeauthor{whyWouldAnyoneBrewer})
+        \cite{whyWouldAnyoneBrewer,Jung:2016:SFG:2851581.2892328,
+              martin2014being,fort2011amazon}.
+  \item with requesters: most of what we know about crowd workers as they relate to requesters, or employers,
+        comes from our recognition of the frustration workers have;
+        \citeauthor{fixingChaos} talks about the chaos of turking and the unilateral power requesters have;
+        \citeauthor{turkopticon} highlight the information asymmetry between workers and requesters on AMT.
+        \citeauthor{dynamo} point out the frustration workers experience trying
+        to communicate collectively with requesters.
+        some research frames this tension as the turker's problem
+        (see \citeauthor{MaliciousCrowdworkersGadiraju}),
+        but some has taken a more even--handed view (\citeauthor{takingAHITMcInnis})
+        \cite{turkopticon,dynamo,MaliciousCrowdworkersGadiraju,takingAHITMcInnis}.
+  \item with other workers: \citeauthor{dynamo} specifically points out the ``mega--drama'' among workers;
+        \citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
+        the culture of crowd work and the study thereof.
+        \citeauthor{crowdcollab} quantifies and maps this social network of turkers
+        (this is different from the requesters topic because
+        researchers like Irani, Gray, etc\dots~aren't Turkers, but are as close to
+        ``cultural anthropologists doing ethnography'' as we see on AMT, maybe online in general)
+        \cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab}.
+\end{enumerate}
+
+\subsubsubsection{\pieceworkpers}
+\begin{enumerate}
+  \item with work: how do workers relate to work?
+        \citeauthor{riisOtherSideLives} documents this to some extent,
+        but doesn't say that much about interviews or anything like that.
+        a handful of sources suggest to know something about
+        how workers relate to the work, like \citeauthor{clark1908cotton},
+        but this distinctly doesn't bring the worker's perspective into the discussion;
+        it projects an etic view, rather than attempts to understand from an emic view.
+        then again, this was a time before Boas and Malinowski, so maybe this is natural.
+        \cite{riisOtherSideLives,clark1908cotton}.
+  \item with requesters:
+        \cite{levi2009union,ahlquist2013interest,
+              mccallum2013global,jacoby1983union}?
+  \item with other workers: this is harder; crowd work research doesn't just benefit from
+        digitally mediated technology allowing us to make
+        relationship networks like \citeauthor{crowdcollab} did;
+        we also benefit from what we consider modern Anthropology and ethnography,
+        something which fundamentally didn't exist or was still in its infancy during piecework's life.
+        Still, we can look at primary sources, like \cite{american1921problem} by
+        the president of the American Federation of Labor,
+        to give us some hint of how they related to each other
+        \cite{american1921problem,crowdcollab}
+\end{enumerate}
+
+\subsubsubsection{\whatchanged}
+
+\subsubsubsection{\implication}
+
+}
+
+% \paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
 \topic{A number of ethical questions surrounding the

### 2016-09-19 drafts/labor/piecework_lit.tex

--- 2016-09-17 drafts/labor/piecework_lit.tex
@@
 \subsection{What was piecework?}
+
+\topic{While ``piecework'' has proven difficult to concretize from the literature,
+we can trace a constellation of characteristics of piece work that recur throughout the literature.}
+% We'll follow the history of research, collecting
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is. While ``piecework'' has had multiple definitions over time,
@@
 if he arranged his work appropriately.
+
+\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
+What I got out of it is that a bunch of people did piecework, but
+I don't know why or in fact why these are different than the sources we cite earlier.
+Can you hone it?}
+
+\ali{I wanted that paragraph to be about the rising popularity and application of piecework
+(especially as it approached its ``heyday'' (a la \cite{hart2013rise})),
+coming from humble beginnings as it found its footing.
+Given that intent\dots~
+
+Should I \textbf{refactor} or \textbf{rewrite}?}
+
+\topic{First applied to farm work, as
+\citeauthor{hughRaynbirdTaskWork} and others illustrate,
+the practice remained relatively obscure until
+it was brought to the textile industry
+\cite{hughRaynbirdTaskWork}.}
+At the turn of the \nth{20} century,
+when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
 
@@
 
+\topic{As increasing attention revealed problems in piecework as it related to workers,
+workers themselves began to speak out about their frustration with this new regime.}
+It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
+but this led to industry organizations representing
+railway workers, mechanical engineers, and others contributing their myriad perspectives
+\cite{american1921problem,richards1904anything,riisOtherSideLives}.
+Nevertheless, piecework continued to permeate low--skilled labor.
+
+% \topic{Piecework became an important contributor to the war effort in the Second World War,
+
 \topic{At this point, piecework became an important factor in the war effort for the Second World War,
@@
 
+\topic{Despite the intense growth of the piecework approach to renumeration,
+this time was not without turmoil.}
+As previously discussed, a number of worker organizations weighed in on
+(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
+\cite{american1921problem,richards1904anything}.
+\citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
 \topic{As increasing attention revealed problems in piecework as it related to workers,
@@
 \cite{hart2013rise}.
-% the gendered and stratified dynamics at play in the matchgirls strike
 
-% \topic{Piecework declined in popularity in the United States, but
-% its spirit thrived through the mid--\nth{20} century in
-% scientific management and Fordism.}
-% Concurrent with the fall of piecework was arguably the rise of factories as centers of work.
+\subsection{Why is piecework relevant to crowdwork?}
 
+\msb{I assume this is forthcoming?}
+Using the definition of piecework that we came up with earlier, we argue that
+crowd work is fundamentally an instantiation of piecework, and
+that we can more precisely anticipate the answers to the open research questions we discussed earlier.
+We'll show that the dimensions of crowd work that the broader HCI community has been studying
+align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
 \subsection{From piecework to on--demand work}

### 2016-09-20 drafts/labor/relationships.tex
relationships expanded

--- 2016-09-14 drafts/labor/relationships.tex
@@
 \begin{inlinelist}
-\item with \nameref{ch:workRelationship},
+% \item with \textit{work}, % \nameref{ch:workRelationship},
 \item with \textit{requesters}, and
@@
 
-\paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
-\topic{A number of ethical questions surrounding the
-increasing complexity of crowd work have arisen in recent years.}
-\citeauthor{professionalcrowdworkEthics} bring some of these issues to light
---- working for increasing amounts of time on tasks of growing complexity%
-, only to discover that requesters are not willing to pay%
-, for instance ---
-but these and other dangers range an enormous landscape
-\cite{professionalcrowdworkEthics}.
-\citeauthor{crowdworkFuture}
-list a few of the problems they identified in \citeyear{crowdworkFuture}
---- motivation, feedback, reputation, quality control, to name a few ---
-while others discuss challenges such as fostering collective action
-and the opportunity for learning and career advancement
-\cite{crowdworkFuture,nickerson2013crowd,dynamo}.
+\topic{\ali{some topic sentence that brings together the debate where
+one side blames Turkers for being bad at Turking and
+another side blames requesters for not understanding Turkers as a culture of people.}}
+Early on, \citeauthor{turkopticon} highlighted
+the information asymmetry between workers and requesters on AMT,
+leading to the creation of \textit{Turkopticon}, a site which allows
+Turkers to rate and review requesters
+\cite{turkopticon}.
+\citeauthor{dynamo} took this critique on information asymmetry and power imbalances a step further,
+% pointing out the frustration workers experience trying
+designing \textit{Dynamo} to facilitate
+Turkers acting collectively to bring about changes to their circumstances
+--- this led to the Academic Requester Guidelines
+\cite{dynamo}.
+\citeauthor{fixingChaos} has since identified the chaos inherent to ``Turking'',
+the unbridled power requesters have, and
+the distress this causes for workers,
+calling for a solution
 
-Some research already looks at research such as investing in workers, and
-informally, we know that this happens among industry requesters
-\cite{jonBrelig,shepherdingDow}.
-AMT, meanwhile, offers requesters the ability to create tasks which are
-not just hidden from unqualified workers by default, but completely.
-Requesters have taken to using lists of worker IDs which reference
-workers who have proven their reliability,
-representing a sort of proto--organization of loosely connected workers.
+some research frames this tension as the turker's problem
+(see \citeauthor{MaliciousCrowdworkersGadiraju}),
+but some has taken a more even--handed view (\citeauthor{takingAHITMcInnis})
+\cite{turkopticon,dynamo,MaliciousCrowdworkersGadiraju,takingAHITMcInnis}.
+
+\topic{}
+\citeauthor{dynamo} specifically points out the ``mega--drama'' among workers;
+\citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
+the culture of crowd work and the study thereof.
+\citeauthor{crowdcollab} quantifies and maps this social network of turkers
+(this is different from the requesters topic because
+researchers like Irani, Gray, etc\dots~aren't Turkers, but are as close to
+``cultural anthropologists doing ethnography'' as we see on AMT, maybe online in general)
+\cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab}.
+% \end{enumerate}
 
 \subsubsubsection{\pieceworkpers}
-The rise of labor unions in the \nth{20} century seems to have been precipitated by
-egregiously unjust conditions imposed on workers in factories and elsewhere
-\cite{ebbinghaus1999institutions}.
-Incidents broadly describing this dynamic can be found in research on AMT
-\cite{turkopticon,dynamo}.
-If these are prototypical labor advocacy organizations of contemporary on--demand work,
-the next question we should look to is if
---- and indeed \textit{how} ---
-these institutions might face challenges in the future.
-
-For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
-and identify that
-``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
-\cite{levi2009union}.
-Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
-This perception already appears to be emerging in digitally mediated peer--governed organizations,
-as \citeauthor{keegan2010egalitarians} and others have illustratively documented
-\cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
-If these organizations and others are to avoid the same fate that labor unions faced,
-they should take care to study this phenomenon and attempt to avoid it.
+% \begin{enumerate}
+  how do workers relate to work?
+  %       \citeauthor{riisOtherSideLives} documents this to some extent,
+  %       but doesn't say that much about interviews or anything like that.
+  %       a handful of sources suggest to know something about
+  %       how workers relate to the work, like \citeauthor{clark1908cotton},
+  %       but this distinctly doesn't bring the worker's perspective into the discussion;
+  %       it projects an etic view, rather than attempts to understand from an emic view.
+  %       then again, this was a time before Boas and Malinowski, so maybe this is natural.
+  %       \cite{riisOtherSideLives,clark1908cotton}.
+  % \item
+          The primary avenue for workers to interact with requesters or managers has been through
+        laborer advocacy groups, which later became guilds, unions, etc\dots
+        and through that lens, there's a lot of research on
+        the relationships between workers and requesters
+        \cite{levi2009union,ahlquist2013interest,
+              mccallum2013global,jacoby1983union}.
+        One component of collectively negotiating with managers has been the process
+        of collective action, a topic which has been substantively explored
+        but is not quite yet answered
+        \cite{russell1982collective,olsonlogic}.
+  % \item
+  this is harder; crowd work research doesn't just benefit from
+        digitally mediated technology allowing us to make
+        relationship networks like \citeauthor{crowdcollab} did;
+        we also benefit from what we consider modern Anthropology and ethnography,
+        something which fundamentally didn't exist or was still in its infancy during piecework's life.
+        Still, we can look at primary sources, like \cite{american1921problem} by
+        the president of the American Federation of Labor,
+        to give us some hint of how they related to each other
+        \cite{american1921problem,crowdcollab}.
+% \end{enumerate}
 
 \subsubsubsection{\whatchanged}
-\topic{While online and distributed workers can be harder to find,
-other features make the study thereof rewarding for the substantively different ways
-that we can approach their study.}
-The longitudinal analysis of everyday discourse, for instance,
-is trivial in the study of online communities
+\topic{The differences between crowd workers and pieceworkers seem defined largely by
+the differences in the places of work.}
+Whereas it arguably became inevitable that workers would have a place to
+meet, discuss, and collaborate
+when they began sharing places of work,
+online spaces make it much harder to do so.
+Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
+without any affiliation with
+--- or even knowledge of ---
+communities of peers
+\ali{multiple citations that labor unions came out of factories here}
+\cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
+
+While the historical management of workers had to be done by hiring a foreman,
+who necessarily had an intuitive
+--- perhaps sympathetic ---
+relationship with workers,
+the foreman of the \nth{20} century has largely been replaced
+by the algorithm of the \nth{21} century
+\cite{uberAlgorithm}.
+The result of this change is that the agents managing work are now
+cold and logical, if unforgiving.
+Where a person might recognize that the ``attention check'' questions 
+proposed by \citeauthor{le2010ensuring} ensure that malicious and inattentive are stopped,
+some implementations of these approaches
+(\cite[see][]{MaliciousCrowdworkersGadiraju}
+only seem to antagonize workers
+\cite{le2010ensuring,MaliciousCrowdworkersGadiraju}
 
 \subsubsubsection{\implication}
+
+% \paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
+% \topic{A number of ethical questions surrounding the
+% increasing complexity of crowd work have arisen in recent years.}
+% \citeauthor{professionalcrowdworkEthics} bring some of these issues to light
+% --- working for increasing amounts of time on tasks of growing complexity%
+% , only to discover that requesters are not willing to pay%
+% , for instance ---
+% but these and other dangers range an enormous landscape
+% \cite{professionalcrowdworkEthics}.
+% \citeauthor{crowdworkFuture}
+% list a few of the problems they identified in \citeyear{crowdworkFuture}
+% --- motivation, feedback, reputation, quality control, to name a few ---
+% while others discuss challenges such as fostering collective action
+% and the opportunity for learning and career advancement
+% \cite{crowdworkFuture,nickerson2013crowd,dynamo}.
+
+% Some research already looks at research such as investing in workers, and
+% informally, we know that this happens among industry requesters
+% \cite{jonBrelig,shepherdingDow}.
+% AMT, meanwhile, offers requesters the ability to create tasks which are
+% not just hidden from unqualified workers by default, but completely.
+% Requesters have taken to using lists of worker IDs which reference
+% workers who have proven their reliability,
+% representing a sort of proto--organization of loosely connected workers.
+
+% \subsubsubsection{\pieceworkpers}
+% The rise of labor unions in the \nth{20} century seems to have been precipitated by
+% egregiously unjust conditions imposed on workers in factories and elsewhere
+% \cite{ebbinghaus1999institutions}.
+% Incidents broadly describing this dynamic can be found in research on AMT
+% \cite{turkopticon,dynamo}.
+% If these are prototypical labor advocacy organizations of contemporary on--demand work,
+% the next question we should look to is if
+% --- and indeed \textit{how} ---
+% these institutions might face challenges in the future.
+
+% For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
+% and identify that
+% ``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
+% \cite{levi2009union}.
+% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
+% This perception already appears to be emerging in digitally mediated peer--governed organizations,
+% as \citeauthor{keegan2010egalitarians} and others have illustratively documented
+% \cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
+% If these organizations and others are to avoid the same fate that labor unions faced,
+% they should take care to study this phenomenon and attempt to avoid it.
+
+% \subsubsubsection{\whatchanged}
+% \topic{While online and distributed workers can be harder to find,
+% other features make the study thereof rewarding for the substantively different ways
+% that we can approach their study.}
+% The longitudinal analysis of everyday discourse, for instance,
+% is trivial in the study of online communities
+
+% \subsubsubsection{\implication}
 

### 2016-09-20 drafts/labor/relationships.tex
wrapping this up to take to gates

@@
 \subsubsubsection{\crowdworkpers}
-\topic{The relationships of workers with their work, with their peers, and with others
-are complex and not especially well--understood.}
+\topic{The relationships of workers
+% with their work,
+with their peers and
+with requesters
+are nuanced and not especially well--understood.}
 Researchers have begun to appreciate
-the sociality of crowd workers in labor markets
-\cite{crowdcollab}.
-Still, the study of these communities is made more challenging by
+the sociality of crowd workers in labor markets;
+still, the study of these communities is made more challenging by
 the limited access to workers on these sites of work
 inherent to digital spaces made without social affordances
-\cite{miller2011understanding}.
-% Nevertheless, a number of interesting findings have
-% already emerged in the crowd work literature%
-% , which we can report here.
-We can break this general body of work into the three groups we named earlier:
+\cite{crowdcollab,miller2011understanding}.
+We can break this general body of work into two subgroups:
 workers' relationships
@@
 another side blames requesters for not understanding Turkers as a culture of people.}}
+Some research frames this tension as the Turker's problem
+(see, for example, \citeauthor{MaliciousCrowdworkersGadiraju}'s work,
+which frames the problem of unpredictable work as the result of ``malicious'' crowd workers),
+\cite{MaliciousCrowdworkersGadiraju,Sheng:2008:GLI:1401890.1401965,Ipeirotis:2010:QMA:1837885.1837906}.
+
 Early on, \citeauthor{turkopticon} highlighted
@@
 \citeauthor{dynamo} took this critique on information asymmetry and power imbalances a step further,
-% pointing out the frustration workers experience trying
 designing \textit{Dynamo} to facilitate
@@
 \cite{dynamo}.
-\citeauthor{fixingChaos} has since identified the chaos inherent to ``Turking'',
-the unbridled power requesters have, and
-the distress this causes for workers,
-calling for a solution
+This unbridled power that requesters have over workers and
+the resultant stress and frustration that this generates
+has been part of the undercurrent of research into
+the tense relationships between workers and requesters
+\cite{fixingChaos,dynamo}.
 
-some research frames this tension as the turker's problem
-(see \citeauthor{MaliciousCrowdworkersGadiraju}),
-but some has taken a more even--handed view (\citeauthor{takingAHITMcInnis})
-\cite{turkopticon,dynamo,MaliciousCrowdworkersGadiraju,takingAHITMcInnis}.
-
-\topic{}
-\citeauthor{dynamo} specifically points out the ``mega--drama'' among workers;
+\topic{The frustration that workers experience dealing with requesters
+seems to precipitate frustration and mistrust between crowd workers, as well.}
+\citeauthor{dynamo} describes ``mega--drama'' among workers on forums for Turkers;
 \citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
 the culture of crowd work and the study thereof.
-\citeauthor{crowdcollab} quantifies and maps this social network of turkers
-(this is different from the requesters topic because
-researchers like Irani, Gray, etc\dots~aren't Turkers, but are as close to
-``cultural anthropologists doing ethnography'' as we see on AMT, maybe online in general)
-\cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab}.
-% \end{enumerate}
+\citeauthor{crowdcollab} quantifies and maps this social network of Turkers.
+\citeauthor{takingAHITMcInnis} takes these observations and considers
+what a crowd work platform might look like if it were to be designed more inclusively
+\cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab,takingAHITMcInnis}.
+The overarching theme of the research in this space has been
+documenting the struggle of crowd workers
+and attempting to intervene in constructive ways, while walking the balancing act
+(especially in the cases of \citeauthor{irani2015cultural} and later \citeauthor{storiesIraniSilberman})
+as we think about the culture of crowd workers.
 
 \subsubsubsection{\pieceworkpers}
-% \begin{enumerate}
-  how do workers relate to work?
-  %       \citeauthor{riisOtherSideLives} documents this to some extent,
-  %       but doesn't say that much about interviews or anything like that.
-  %       a handful of sources suggest to know something about
-  %       how workers relate to the work, like \citeauthor{clark1908cotton},
-  %       but this distinctly doesn't bring the worker's perspective into the discussion;
-  %       it projects an etic view, rather than attempts to understand from an emic view.
-  %       then again, this was a time before Boas and Malinowski, so maybe this is natural.
-  %       \cite{riisOtherSideLives,clark1908cotton}.
-  % \item
-          The primary avenue for workers to interact with requesters or managers has been through
-        laborer advocacy groups, which later became guilds, unions, etc\dots
-        and through that lens, there's a lot of research on
+\topic{The questions surrounding
+the ways pieceworkers related to managers might be best answered by
+the work that has been done in the emergence and proliferation of labor unions.}
+The primary avenue for workers to interact with managers has been through
+laborer advocacy groups such as the American Federation of Labor,
+(one of the forerunners of the largest and
+most politically influential labor union in the United States).
+Looking through that lens, we find copious research on
 the relationships between workers and requesters
@@
 \cite{russell1982collective,olsonlogic}.
-  % \item
-  this is harder; crowd work research doesn't just benefit from
-        digitally mediated technology allowing us to make
-        relationship networks like \citeauthor{crowdcollab} did;
-        we also benefit from what we consider modern Anthropology and ethnography,
-        something which fundamentally didn't exist or was still in its infancy during piecework's life.
-        Still, we can look at primary sources, like \cite{american1921problem} by
-        the president of the American Federation of Labor,
+
+\topic{Answering how workers related to one another is arguably more challenging
+for a number of reasons.}
+For one thing, the research methods we typically associate with the exploratory study of cultures
+--- Anthropology, and namely participant--observation, ethnography, etc\dots ---
+didn't exist quite as we know them at the turn of the \nth{20} century, and
+wouldn't for several more decades.
+Still, we can look at primary sources, like \citetitle{american1921problem}
 to give us some hint of how they related to each other
-        \cite{american1921problem,crowdcollab}.
-% \end{enumerate}
+\cite{american1921problem}.
+
+\ali{What do I want to do/say here? I think I want to point to point out that
+\dots~I dunno. I'll review the \citetitle{american1921problem} book and maybe some other stuff}
 
@@
 only seem to antagonize workers
-\cite{le2010ensuring,MaliciousCrowdworkersGadiraju}
+\cite{le2010ensuring,MaliciousCrowdworkersGadiraju}.
 
 \subsubsubsection{\implication}
+\topic{What we've done in the field of crowd work might be able to tell us something about piecework
+just as piecework has told us so much about crowd work.}
+Crowd work research doesn't just benefit from
+digital media allowing us to make
+relationship networks like \citeauthor{crowdcollab} do;
+we benefit from the firmer theoretical basis of Anthropology that
+existed in a radically different form at the turn of the \nth{20} century,
+when piecework began to emerge.
+\citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and
+other luminaries throughout the first half of the \nth{20} century
+effectively defined Cultural Anthropology as we know it today;
+\textit{participant--observation},
+the \textit{etic} and the \textit{emic} understanding of culture, and
+\textit{reflexivity}
+didn't take even a resemblance of their contemporary forms until these works
+\cite{malinowski2002argonauts,boas1940race,mead1973coming}.
 

### 2016-09-20 drafts/labor/complexity.tex
fix ref in complexity

@@
 If we can resolve the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest,
-then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciouscrowdworkersGadiraju}
+then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciousCrowdworkersGadiraju}
 and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate

### 2016-09-21 drafts/labor/cases.tex
fixing ref in cases.tex
remove flexibility. no flexibility!

@@
 We look at lots of papers that use the term ``crowdsourcing'' or ``crowd work''
-in the abstracts and titles of their papers%
-, especially the papers that we cited earlier, to try
+in the abstracts and titles of their papers,  especially the papers that we cited earlier, to try
 to answer the questions that we posed in the introduction.
@@
 piecework answers the questions we, crowdsourcing researchers, have asked.
-We then see whether and how crowdsourcing as we know it has differed from piecework%
-, and how that affects the predictions and conclusions made in the piecework literature.
+We then see whether and how crowdsourcing as we know it has differed from piecework,  and how that affects the predictions and conclusions made in the piecework literature.
 
@@
 This has largely involved
-  identifying challenges to this form of labor%
-,   overcoming them through novel designs of work--flows and processes, and
+  identifying challenges to this form of labor,    overcoming them through novel designs of work--flows and processes, and
   repeating the process
@@
 \begin{inlinelist}
-  \item \nameref{sec:complexity}%
-,   \item \nameref{sec:decomposition}, and
+  \item \nameref{sec:complexity},    \item \nameref{sec:decomposition}, and
   \item \nameref{sec:relationships}
 \end{inlinelist}.
-We'll explore these aspects of crowdsourcing%
-, discussing the extents to which work can be
-decomposed%
-, contextually abstracted, and
+We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
+decomposed,  contextually abstracted, and
 made more resilient to attrition of various forms.
@@
 \endnotetext{
-Here, \citeauthor{hu1961parallel}'s work%
-, saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs''%
-, parallels the approach that dominated early research into crowd work
+Here, \citeauthor{hu1961parallel}'s work,  saying of assembly line work that
+``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',  parallels the approach that dominated early research into crowd work
 --- namely, using non--expert crowds for complex work
 \cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis%
-, and indeed the study of factory and mass manufacturing labor through the \nth{20} century%
-, substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier%
-, and persists to this day as it colors
+This mindset in \citeauthor{hu1961parallel}'s analysis,  and indeed the study of factory and mass manufacturing labor through the \nth{20} century,  substantively owes its existence to scientific management
+and the rigorous decomposition of work into tasks, discussed earlier,  and persists to this day as it colors
 researchers' goals and objectives in the study and design of crowd work.
 
-Piecework's influence on the abstraction of work into tasks%
-, described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly%
-, prompting work requesters to consider ways to design these now discrete tasks in ways that
+Piecework's influence on the abstraction of work into tasks,  described above, is more than just caused by the decomposition of work;
+work abstraction itself makes it possible for workers to come and go flexibly,  prompting work requesters to consider ways to design these now discrete tasks in ways that
 maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
@@
 \citeauthor{thompson1913time}'s thesis asserts that
-task work is predicated on the accurate scientific management of work%
-, including the ``miscellaneous tasks''.
+task work is predicated on the accurate scientific management of work,  including the ``miscellaneous tasks''.
 \citeauthor{thompson1913time} argues
@@
 
-Broken down in this way, work could grow to unprecedented scales%
-, but the quality of the work would remain relatively variable
+Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable
 \cite{murray1983decentralisation}.
-Textile work being a salient example%
-, it took time for workers to acquire sufficient skill
+Textile work being a salient example,  it took time for workers to acquire sufficient skill
 to do every aspect of the work so that the garment would be accepted by the company soliciting that work
@@
 
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete%
-, manageable routines that could be taught relatively easily%
-, and whose work output could be evaluated in abstraction from the rest of the work
+A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work
 \cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car%
-, but a single very narrowly defined action that needed to be done on every car
+In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car
 \cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes%
-, ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
+By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
 \cite{schoenberger1988fordism}.
 
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos)%
-, can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces%
-, facilitating evaluation and parallelization
+This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),  can be seen today in crowd work and on--demand labor through the application of micro--tasks.
+\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization
 \cite{writingMicroTasks}.
-By decomposing and recomposing tasks%
-, and in particular by assigning similarly natured work to the same workers%
-, workers could become ``experts'' in a small aspect of the work that they did%
-, speeding their work dramatically
+By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically
 \cite{delayAndOrderLasecki}.
@@
     --- though not necessarily \textit{faster} than ``macrotasks'' ---
-    yield higher quality work%
-,     particularly when that work is susceptible to frequent interruptions
+    yield higher quality work,      particularly when that work is susceptible to frequent interruptions
     \cite{cheng2015break}.
@@
   The metaphorical mechanics of these dynamics are still at play;
-  workers and managers continue to interact in adversarial manners%
-,   despite substantive work into aligning the motivations of workers and requesters
+  workers and managers continue to interact in adversarial manners,    despite substantive work into aligning the motivations of workers and requesters
 
     % \section{Case Studies}
-    The existing body of research has shed light on on--demand labor from various perspectives%
-,     and revealed a number of topics that%
-,     through our framing, are clearly situated together.
+    The existing body of research has shed light on on--demand labor from various perspectives,      and revealed a number of topics that,      through our framing, are clearly situated together.
     Those topics are, at a high level, as follows:
@@
     \item the \textbf{processes} involved in making work into tasks, or discretization;
-    \item the outcomes (and indeed the \textbf{fallout}) of that discretization%
-,     both on the work itself as well as the workers;
+    \item the outcomes (and indeed the \textbf{fallout}) of that discretization,      both on the work itself as well as the workers;
     and finally
@@
     \subsection{The Fallout of Crowd Work}
-    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging%
-,     % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
+    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,      % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
     \citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
@@
     \citeauthor{uberAlgorithm}
-    find similarly that workers on gig work platforms are frustrated by the systems on which they work%
-,     to say little of the policies which these systems enforce
+    find similarly that workers on gig work platforms are frustrated by the systems on which they work,      to say little of the policies which these systems enforce
     \cite{uberAlgorithm}.
@@
     \citeauthor{laborEconomicsOfCrowdsourcingHorton}
-    identified problems with crowd work wages relatively early on%
-,     attempting to address this imbalance from a behavioral economic perspective ---
+    identified problems with crowd work wages relatively early on,      attempting to address this imbalance from a behavioral economic perspective ---
     that is, identifying and presenting a model that describes a worker's
@@
 
-    But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}''%
-,     a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
+    But we turn to \citeauthor{turkopticon}'s discussion of ``\TO'',      a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
     \cite{turkopticon}.
-    Of particular relevance here%
-,     \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
+    Of particular relevance here,      \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
     wage theft and
     pay rates that translate to well under minimum wage.
-    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton}%
-,     we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
+    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},      we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
     was \$1.38, while the mean was \$3.63
@@
     \cite{Sun20111033}.
-    This might have rung true in \citeyear{Sun20111033}%
-,     and certainly corroborates \citeauthor{Ross}'s findings after investigating
-    ``who are the crowd workers''%
-,     but as \citeauthor{whoareNOTtheTurkers} points out
+    This might have rung true in \citeyear{Sun20111033},      and certainly corroborates \citeauthor{Ross}'s findings after investigating
+    ``who are the crowd workers'',      but as \citeauthor{whoareNOTtheTurkers} points out
     ``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
@@
     and later
-    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers%
-,     via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
+    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,      via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
     \cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
@@
     \cite{turner1952trade}.
-    This, \citeauthor{ebbinghaus1999institutions} argued%
-,     fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
+    This, \citeauthor{ebbinghaus1999institutions} argued,      fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
     (to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
@@
     We argue further that the \textit{causes} here
-    --- work decomposition%
-,     work abstraction, and
+    --- work decomposition,      work abstraction, and
     flexibility ---
@@
     ``understanding malicious behavior''
-    \cite{MaliciouscrowdworkersGadiraju}.
-    While some work has cast workers as ``malicious'' or at least adversarial parties%
-,     the evidence thus far suggests that
+    \cite{MaliciousCrowdworkersGadiraju}
+    While some work has cast workers as ``malicious'' or at least adversarial parties,      the evidence thus far suggests that
     workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
@@
     the \textit{actual} trustworthiness drops precipitously
-    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so%
-,     leaving only the ones who cannot or will not establish their trustworthiness.
+    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,      leaving only the ones who cannot or will not establish their trustworthiness.
 
     \subsection{Relationships Between Workers and Managers}
-    Suffice it to say that poor pay and poor work are linked%
-,     and that we should not be surprised to find this relationship play out online as strongly as it does offline.
+    Suffice it to say that poor pay and poor work are linked,      and that we should not be surprised to find this relationship play out online as strongly as it does offline.
     But the poor treatment of workers by managers
@@
     We discussed Fordism and Taylorism earlier in our discussions of
-    \nameref{sec:Complexity} and \nameref{sec:flexibility}%
-,     but here the core of these paradigmatic views
+    \nameref{sec:Complexity} and \nameref{sec:flexibility},      but here the core of these paradigmatic views
     --- the scientific management of work ---
     becomes relevant.
-    We use ``external'' here instead of ``scientific'', however%
-,     to more broadly capture the disconnect between managers and workers.
-    By describing it as thus%
-,     we can touch on the relationship that workers have with \textit{researchers}, as well%
-,     even though that work is not strictly
+    We use ``external'' here instead of ``scientific'', however,      to more broadly capture the disconnect between managers and workers.
+    By describing it as thus,      we can touch on the relationship that workers have with \textit{researchers}, as well,      even though that work is not strictly
     --- or just not exclusively ---
@@
     \citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
-    cognitively demanding tasks on workers%
-,     informing crowdsourcing research by suggesting the use of cognitive load assessments such as
+    cognitively demanding tasks on workers,      informing crowdsourcing research by suggesting the use of cognitive load assessments such as
     NASA Task Load Index surveys to evaluate workers pre and post--tasks
@@
     External management comes in other forms than scientific, as previously mentioned.
-    Researchers in particular have noticed that their relationships with on--demand workers are%
-,     at the least, complex.
+    Researchers in particular have noticed that their relationships with on--demand workers are,      at the least, complex.
     \citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
@@
 
-    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research%
-,     and even physiological study of pieceworkers closely resembles the
+    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,      and even physiological study of pieceworkers closely resembles the
     research into cognitive loads and stress levels that we discussed among on--demand crowd workers
@@
     \subsubsection{Resistance}
-    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems%
-,     often without recourse or opportunity for feedback, let alone substantive input.
+    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,      often without recourse or opportunity for feedback, let alone substantive input.
     Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
-    many toggle their availability to avoid being dispatched to more distant locations%
-,     resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
+    many toggle their availability to avoid being dispatched to more distant locations,      resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
     \cite{uberAlgorithm}.
@@
     Resistance has sometimes been more coordinated, as well;
-    we see this in \citeauthor{turkopticon}'s coverage on \textit{Turkopticon} as workers collectively accumulated information about requesters, and
-    in \citeauthor{dynamo}'s work on \textit{Dynamo}, which generated ``Guidelines for Academic Requesters'' written by crowd workers
+    we see this in \citeauthor{turkopticon}'s coverage on \TO
+    as workers collectively accumulated information about requesters, and
+    in \citeauthor{dynamo}'s work on \DO, which generated ``Guidelines for Academic Requesters'' written by crowd workers
     \cite{turkopticon,dynamo}.
 
-    Resistance against managers in piecework and factory labor settings are deeply well--explored%
-,     but perhaps the most relevant case study to draw on here is to be found in
-    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors''%
-,     where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
+    Resistance against managers in piecework and factory labor settings are deeply well--explored,      but perhaps the most relevant case study to draw on here is to be found in
+    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',      where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
     \cite{waldinger1996helots}.
-    The achievements of labor advocacy groups such as labor unions as resistant%
-,     even adversarial organizations counter--balancing the management
+    The achievements of labor advocacy groups such as labor unions as resistant,      even adversarial organizations counter--balancing the management
     is somewhat well--understood

### 2016-09-21 drafts/labor/complexity.tex
remove flexibility. no flexibility!

@@
 \documentclass[trackingWork]{subfiles}
-\onlyinsubfile{
-  \usepackage{xr-hyper}
-  \usepackage{hyperref}
-}
 
@@
 would be unsuitable for piecework.
-% We might reflect on \citeauthor{hart2016rise}'s piecework's limitations,
-% and specifically the challenge of multidimensional work
-% --- tasks comprising of numerous, sometimes conflicting, goals
-% \cite{hart2016rise}.
 
-% It would be reasonable, then, to infer that work like this
-% --- reasonably highly skilled work where quality is difficult to assess ---
-% would be unsuitable for piecework.
-% \msb{that conclusion came way too fast. give a little detail on the railway. how did they come to that conclusion?}
-
-% \msb{this paragraph needs a topic sentence. it's a wandering paragraph currently. is the claim here that evaluation is the limiting factor?}
 \topic{Determining appropriate pay rates, informed by the careful measurement of workers,
@@
 \cite{10.2307/23702539}.
-% \citeauthor{10.2307/23702539} and \citeauthor{hart2016rise} may seem
-% to be making differing claims about the limitations of piecework, but
-% we argue that \citeauthor{10.2307/23702539}
-% is simply making a more concrete observation illustrating the insight that
-% \citeauthor{hart2016rise} later makes.
 \citeauthor{10.2307/23702539} further recognizes that
@@
 piecework was applied almost exclusively to farm and textile work.
-% \citeauthor{grier2013computers} describes Airy's training and employment of young men to compute astronomical observations,
-% mentioning that boys needed only to have ``the basic skills of mathematics''
-% \cite{grier2013computers}.
-% \citeauthor{hart2013rise} argue that piecework flourished in the first half of the \nth{20} century because
-% tasks were decomposed --- a discussion we will return to later ---
-% such that untrained workers could be trained relatively quickly on a very narrowly designed task
-% \cite{hart2013rise}.
 Work was simple and widely understood
@@
 \cite{10.2307/2338394,riisOtherSideLives}.
-% The general argument for decomposition, we will see later,
-% hinges on the promise that
-% tasks can be \textit{made} less complex and time--consuming,
-% if they're not already so.
-% \ali{this section is here because i later talk about the complexity of work, and
-% I want to keep the whole thing about how Uber, Google Maps, etc\dots make it possible to
-% \begin{inlinelist}
-% \item train workers more quickly (honestly just transmitting instructions over the Internet is so much quicker than what Airy had), and
-% \item give workers necessary tools/information to get work done
-% (e.g. ``Waze knows you should avoid the intersection 3 blocks down the road because of an accident''
-% and (maybe) ``not great at math? WolframAlpha can confirm or just give you your answer'')
-% \end{inlinelist}.}
 
@@
 indeed, we've discussed complex applications of crowd work already.}
-% As \citeauthor{hart2013rise} described,
 While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design,
 much of it arose out of necessity ---
-% almost strictly out of necessity ---
 it was infeasible to provide new workers with
@@
 While this constraint led to much more tightly scoped work
-% (perhaps surprisingly at the time) more efficient allocations of workers,
 who now had to specialize in extremely narrowly defined roles.
@@
 it seems that we don't yet know the limits of complexity with regard to piecework.
-% \msb{so what are we to take out of this? that there is no limit? just our creativity?}
 \ali{This is \textit{almost} making a prediction,

### 2016-09-21 drafts/labor/decomposition.tex
expanding on relationships
abstract and title closed out
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)
remove flexibility. no flexibility!

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
+% \makeatletter
+% \def\blx@maxline{77}
+% \makeatother
+\onlyinsubfile{\usepackage{xr-hyper}
+\usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{relationships}
+\externaldocument{decomposition}
+}
 \begin{document}
 
-\subsubsection[the decomposition of work]{The Decomposition of Work}\label{sec:decomposition}
+\subsubsection[How far can work be decomposed into smaller and smaller microtasks]{Decomposing Work}\label{sec:decomposition}
 
@@
 \topic{The crowdsourcing research into work decomposition
-has largely focused on minimizing the additional context necessary to do tasks%
-, and making it easier to do tasks with less time.}
+has largely focused on minimizing the additional context necessary to do tasks,
+and making it easier to do tasks with less time.}
 This first thread is
@@
 (specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
-which finds that greater work context fosters more reliably high--quality work)%
-, the micro task paradigm has emerged as the overwhelming favorite
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016%
-,       cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
+which finds that greater work context fosters more reliably high--quality work),
+the micro task paradigm has emerged as the overwhelming favorite
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016,
+      cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
 
-\topic{As the additional context necessary to complete a task diminishes%
-, the marginal cost of finding and
+\topic{As the additional context necessary to complete a task diminishes,
+the marginal cost of finding and
 \textit{doing} tasks has increasingly become the focus of research.}
-\citeauthor{taskSearch} illustrate the challenges on AMT%
-, and some work has gone into ameliorating the problems specific to this work site
+\citeauthor{taskSearch} illustrate the challenges on AMT,
+and some work has gone into ameliorating the problems specific to this work site
 (\textit{ReLauncher}), % does anyone care for project names?
@@
 (\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning})
-\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996%
-,       Cai:2015:WLW:2702123.2702267}.
+\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,
+      Cai:2015:WLW:2702123.2702267}.
 % \topic{One of the emergent properties of micro--tasks has been the relative cost of
@@
 \cite{Cosley:2007:SUI:1216295.1216309}.
-Much of this work and the work at the periphery of this space, then%
-, has focused on
+Much of this work and the work at the periphery of this space, then,
+has focused on
 minimizing the amount of time that people need to spend doing
@@
 both the study and practice
-of crowd work, that microtasks should be designed resiliently against the variability of workers%
-, fully exploiting the abstracted nature of each piece of work
+of crowd work, that microtasks should be designed resiliently against the variability of workers,
+fully exploiting the abstracted nature of each piece of work
 \cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-That is to say, micro--tasks should be designed such that a single worker's poor performance%
-, or a good worker's sudden departure%
-, does not significantly impact the agenda of the work as a whole.
+That is to say, micro--tasks should be designed such that a single worker's poor performance,
+or a good worker's sudden departure,
+does not significantly impact the agenda of the work as a whole.
 While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components
 in the form of higher cumulative time to complete
-(albeit much shorter real time to complete, owing to parallelization)%
-, \citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
+(albeit much shorter real time to complete, owing to parallelization),
+\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
 similar tasks together
@@
 
-Yet more work looks at the general framing of tasks%
-, chaining and arranging them to maximally exploit
+Yet more work looks at the general framing of tasks,
+chaining and arranging them to maximally exploit
 the attention and stress threshold % ???
@@
 \cite{Cai:2016:CRI:2858036.2858237}.
-Rather than attempt to minimize the error rates in micro--tasks%
-, as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested%
-, we as a community have leaned \textit{into} the peril of
-low--context work%
-, ``embracing error'' in crowdsourcing
+Rather than attempt to minimize the error rates in micro--tasks,
+as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested,
+we as a community have leaned \textit{into} the peril of
+low--context work,
+``embracing error'' in crowdsourcing
 \cite{embracingErrorKrishna}.
@@
 systematic task decomposition
-stretch back as far as the \nth{17} century%
-, when Airy employed young boys at the Greenwich Observatory who
+stretch back as far as the \nth{19} century,
+when Airy employed young boys at the Greenwich Observatory who
 ``possessed the basic skills of mathematics, including
@@
 First, work output was quickly verifiable;
-Airy could assign variably skilled workers to compute values%
-, and have other workers check their work.
+Airy could assign variably skilled workers to compute values,
+and have other workers check their work.
 Second, tasks were discrete --- that is, independent from one another.
@@
 The insight of breaking tasks down into smaller components didn't find its audience until
-the early \nth{20} century%
-, with the rise of Fordism and scientific management (or Taylorism).
+the early \nth{20} century,
+with the rise of Fordism and scientific management (or Taylorism).
 From scientific management, we found that
 we could measure work at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out%
-, piecework most greatly benefits the instrumented measurement of workers, but certainly
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers, but certainly
 in Ford and Taylor's time --- and certainly in Airy's time ---
 highly instrumented, automatic measurement of workers was all but impossible.
-As a result%
-, the distillation of work into smaller chunks
+As a result,
+the distillation of work into smaller chunks
 ultimately reached a limit of usefulness.
+
+\topic{Piecework researchers address a number of the problems with the decomposition of work,
+and the conflicting pressures managers and workers put forth.}
+\citeauthor{bewley1999wages} in particular points out that
+the approach of paying workers by the piece is
+``\dots~not practical for workers doing many tasks, because of
+the cost of establishing the rates and because
+piecework does not compensate workers for time spent switching tasks''.
+Ultimately, \citeauthor{bewley1999wages} argues that
+``[piecework is] infeasible, because \dots
+total output is the joint product of varying groups of people''
+\cite{bewley1999wages}.
+\ali{This is becoming both
+a complexity    argument and
+a decomposition argument\dots
+I'm okay with using \citeauthor{bewley1999wages} in both, but
+I want people to read this paper and not be like
+``wow this author sure got their money's worth from those three books!''}
 

### 2016-09-21 drafts/labor/relationships.tex
remove flexibility. no flexibility!

@@
 the information asymmetry between workers and requesters on AMT,
-leading to the creation of \textit{Turkopticon}, a site which allows
+leading to the creation of \TO, a site which allows
 Turkers to rate and review requesters
@@
 \citeauthor{dynamo} took this critique on information asymmetry and power imbalances a step further,
-designing \textit{Dynamo} to facilitate
+designing \DO to facilitate
 Turkers acting collectively to bring about changes to their circumstances
@@
 
-\ali{What do I want to do/say here? I think I want to point to point out that
-\dots~I dunno. I'll review the \citetitle{american1921problem} book and maybe some other stuff}
+The ethos of American labor advocacy organizations``not only as railroad employees but also as members of the larger life of the community''
+and other notions of affiliation and recognition of shared underlying goals and hopes.
+I could point to \citeauthor{riisOtherSideLives} and be like
+``what \textit{he} did was illustrate through this relatively new medium --- photography ---
+how bad pieceworkers had it, foreshadowing
+this later urging by nascent labor advocacy groups to recognize that
+workers all shared these problems and needed to associate.''
+
+This can also foreshadow crowd sourcing efforts like
+\TO~and \DO.
 

### 2016-09-21 drafts/labor/complexity.tex
relationships.tex

@@
 a variety of duties than in jobs with a narrow set of routinized duties''
-\cite{Brown01041990}
+\cite{Brown01041990}.
+\citeauthor{SJOE:SJOE371} points out the phenomenon here as a market effect:
+``\dots~in an environment with multi--tasking,
+pay schemes based on tightly specified performance may \textit{induce workers
+to neglect tasks} that are less easy to measure [emphasis added]''
+\cite{SJOE:SJOE371}.
 % \msb{that's confusing, is incentive pay the same as piecework?
@@
 % give us the noun instead of forcing us to figure it out?}.
-\citeauthor{10.2307/23702539} adds further, that
+\citeauthor{10.2307/23702539} argues that
 successful cases of piecework owed themselves in part to the fact that

### 2016-09-21 drafts/labor/relationships.tex
relationships.tex

--- 2016-09-20 drafts/labor/relationships.tex
@@
 \citeauthor{dynamo} took this critique on information asymmetry and power imbalances a step further,
-designing \textit{Dynamo} to facilitate
+designing \DO~to facilitate
 Turkers acting collectively to bring about changes to their circumstances
@@
 
-\ali{What do I want to do/say here? I think I want to point to point out that
-\dots~I dunno. I'll review the \citetitle{american1921problem} book and maybe some other stuff}
+The driving force of American labor advocacy organizations was to get piecework railroad workers
+to identify
+``not only as railroad employees but also as members of the larger life of the community''
+\cite{american1921problem}.
+Doing this, 
+\citeauthor{ostrom1990governing} and others argued,
+would facilitate collective action and perhaps collective governance
+\cite{ostrom1990governing,russell1982collective,olsonlogic}.
+\citeauthor{riisOtherSideLives} had contributed to this sense of shared struggle and endurance
+by the time \citetitle{american1921problem} was published
+by documenting pieceworkers in their home--workplaces,
+literally bringing to light the grim circumstances in which pieceworkers lived and worked
+\cite{riisOtherSideLives}.
+
+% This can also foreshadow crowd sourcing efforts like
+% \TO~and \DO.
 
@@
 communities of peers
-\ali{multiple citations that labor unions came out of factories here}
+% \ali{multiple citations that labor unions came out of factories here}
 \cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
 
-While the historical management of workers had to be done by hiring a foreman,
-who necessarily had an intuitive
+\topic{We further find the sources of differences between crowd work and piecework
+in the nature of the relationship between workers and requesters
+--- or rather, the lack thereof.}
+While historically the management of workers had to be done through a foreman
+(who necessarily had an intuitive
 --- perhaps sympathetic ---
-relationship with workers,
+relationship with workers),
 the foreman of the \nth{20} century has largely been replaced
-by the algorithm of the \nth{21} century
+by algorithms of the \nth{21} century
 \cite{uberAlgorithm}.
@@
 some implementations of these approaches
-(\cite[see][]{MaliciousCrowdworkersGadiraju}
-only seem to antagonize workers
+only seem to antagonize workers \ali{shots fired}
 \cite{le2010ensuring,MaliciousCrowdworkersGadiraju}.
+% We've known for decades that workers prefer human management 
+\citeauthor{10.2307/2555446} told us more than 30 years ago
+--- in \citeyear{10.2307/2555446} --- that
+``\dots~when performance is difficult to evaluate,
+imperfect input measures and
+a manager's subjective judgment are preferable to
+defective (simple, observable) output measures''
+\cite{10.2307/2555446}.
+\ali{here's an idea. feel free to push back or revise or something:\\
+``Should we be surprised that
+management approaches we've known to
+frustrate pieceworkers since \citeyear{10.2307/2555446}
+would only aggravate crowd workers in many of the same ways?''}
 
@@
 \cite{malinowski2002argonauts,boas1940race,mead1973coming}.
+
+\topic{The research on piecework still offers to guide us on perhaps the most rudimentary aspects of worker management}
+\citeauthor{10.2307/2555446} drew a dichotomous line between
+``defective (simple, observable) output measures'' and
+``a manager's subjective judgment'',
+but such a dichotomy need not necessarily represent our work management styles
+\cite{10.2307/2555446}.
+We can develop tools that better inform humans, rather than
+(perhaps futilely) attempt to delegate all worker management to machines.
+This is an area we should pursue, but haven't yet.
+If the literature on piecework is to be believed,
+more considerate \textit{human} management may resolve
+many of the tensions we've discovered among among crowd workers.
 

### 2016-09-21 drafts/labor/decomposition.tex
decomposition revised

--- 2016-09-12 drafts/labor/decomposition.tex
@@
 \makeatother
+\onlyinsubfile{
+  \usepackage{xr-hyper}
+  \usepackage{hyperref}
+}
+
 \begin{document}
 
-\subsubsection[the decomposition of work]{The Decomposition of Work}\label{sec:decomposition}
+\subsection[How far can work be decomposed into smaller microtasks]{Decomposing Work}\label{sec:decomposition}
 
-\subsubsubsection{\crowdworkpers}
-\topic{The crowdsourcing research into work decomposition
-has largely focused on minimizing the additional context necessary to do tasks%
-, and making it easier to do tasks with less time.}
-This first thread is
-perhaps best described by \citeauthor{verroios2014context} as
-making crowd workers ``\dots able to act with
-global understanding when each contributor only has access to local views''
-\cite{verroios2014context}.
-With the exception of a few cases
-(specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
-which finds that greater work context fosters more reliably high--quality work)%
-, the micro task paradigm has emerged as the overwhelming favorite
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016%
-,       cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
+At its core, on--demand work has been enabled by decomposition of large goals into many small tasks.
+As such, one of the central questions in the literature is how to design these microtasks, and which kinds of tasks are amenable to decomposition.
+In this section, we place these questions in the context of piecework's Tayloristic evolution.
 
-\topic{As the additional context necessary to complete a task diminishes%
-, the marginal cost of finding and
-\textit{doing} tasks has increasingly become the focus of research.}
-\citeauthor{taskSearch} illustrate the challenges on AMT%
-, and some work has gone into ameliorating the problems specific to this work site
-(\textit{ReLauncher}), % does anyone care for project names?
-while other work designs tasks around gap time
-(\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning})
-\cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996%
-,       Cai:2015:WLW:2702123.2702267}.
+\begin{comment}
+Outline:
+Crowd work
+	- How do we take work and split it up into smaller work?
+	- How small can we go?
+	- Once we split up, what happens? Marketplace choosing
+Piecework
+	- How do we take work and split it up?
+	[ - how small can we go? ]
+	[- Nothing about how people choose. ]
+Same/different
+	[- task data science ]
+	[- Cognitive barriers, task switching, etc, dominates]
+	[ - Switching is different, marketplace model ]
+\end{comment}
+
+\subsubsection{\crowdworkpers}
+\topic{Many contributions to the design and engineering of crowd work consist of creative methods for decomposing goals.} Even when tasks such as writing and editing cannot be reliably performed by individual workers, researchers demonstrated that decompositions of these tasks into workflows can succeed~\cite{crowdForgeKittur,bernsteinSoylent,writingMicroTasks,Nebeling:2016:WCW:2858036.2858169}.
+These decompositions typically take the form of workflows, which are algorithmic sequences of tasks that manage interdependencies~\cite{Bigham2014}. 
+Workflows often utilize a first sequence of tasks to identify an area of focus (e.g., a paragraph topic~\cite{crowdForgeKittur}, an error~\cite{bernsteinSoylent}, or a concept~\cite{Yu2016a,Yu2016b} and a second sequence of tasks to execute work on that area.
+This decomposition style has been successfully applied across many areas, including food labeling~\cite{noronha2011platemate}, brainstorming~\cite{siangliulue2015toward,yu2014distributed}, and accessibility~\cite{lasecki2013chorus,lasecki2012real,Lasecki2011}.
+
+\topic{If decomposition is key to success in crowd work, the question arises: what can, and can't, be decomposed?} 
+Moreover, how thinly can work be sliced and subdivided into smaller and smaller tasks? 
+The general trend has been that smaller is better, and the microtask paradigm has emerged as the overwhelming favorite
+\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+This work illustrates a broader sentiment in both the study and practice
+of crowd work, that microtasks should be designed resiliently against the variability of workers, preventing a single errant submission from impacting the agenda of the work as a whole %, fully exploiting the abstracted nature of each piece of work
+~\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
+In this sense, finer decompositions are seen as more robust --- both to interruptions and errors~\cite{cheng2015break} --- even if they incur a fixed time cost.
+At the extreme, recent work has attempted demonstrated microtasks that take seconds~\cite{Vaish:2014:TCC:2611222.2556996,Cai:2015:WLW:2702123.2702267} or even tenths of a second~\cite{embracingErrorKrishna}.
+%Earlier we discussed \citeauthor{cheng2015break}'s work measuring the impact that interruption has on worker performance \cite{cheng2015break}.%While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components in the form of higher cumulative time to complete (albeit much shorter real time to complete, owing to parallelization)%
+However, workers perform better when similar tasks are strung together 
+\cite{delayAndOrderLasecki}, or chained and arranged to maximize  the attention threshold of workers~\cite{Cai:2016:CRI:2858036.2858237}.
+Despite this, 
+%rather than attempt to minimize the error rates in micro--tasks
+%, as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested%
+we as a community have leaned \textit{into} the peril of
+low--context work, ``embracing error'' in crowdsourcing~\cite{embracingErrorKrishna}.
+
+\topic{The general lesson has been that the more micro the task, and the more fine the decomposition, the greater the risk that workers lose context necessary to perform the work well.}
+For example, workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017}, interpret tasks in different ways~\cite{kairam2016parting}, and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
+Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher-level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.
+
+% \topic{The crowdsourcing research into work decomposition
+% has largely focused on minimizing the additional context necessary to do tasks%
+% , and making it easier to do tasks with less time.}
+% This first thread is
+% perhaps best described by \citeauthor{verroios2014context} as
+% making crowd workers ``\dots able to act with
+% global understanding when each contributor only has access to local views''
+% \cite{verroios2014context}.
+% With the exception of a few cases
+% (specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
+% which finds that greater work context fosters more reliably high--quality work)%
+% , the micro task paradigm has emerged as the overwhelming favorite
+% \cite{selfsourcingTeevan2014,selfsourcingTeevan2016%
+% ,       cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
+
+\topic{As the additional context necessary to complete a task diminishes, the invisible labor of \textit{finding} tasks~\cite{martin2014being} has arisen as a major issue.}
+\citeauthor{taskSearch} illustrate the task search challenges on AMT.%
+Workers seek out good requesters~\cite{martin2014being} and then  ``streak'' to perform many tasks of that same type~\cite{taskSearch}.
+%, and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}), % does anyone care for project names?
+%while other work designs tasks around gap time (\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning}) \cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,       Cai:2015:WLW:2702123.2702267}.
 % \topic{One of the emergent properties of micro--tasks has been the relative cost of
@@
 %     \cite{taskSearch}.
-\citeauthor{Cosley:2007:SUI:1216295.1216309}
-attempts to address this by
-directing workers to tasks through
-``intelligent task routing''
-\cite{Cosley:2007:SUI:1216295.1216309}.
-Much of this work and the work at the periphery of this space, then%
-, has focused on
-minimizing the amount of time that people need to spend doing
-anything other than the work for which they are paid.
+Researchers have reacted by designing task recommendation systems (e.g.,~\cite{Cosley:2007:SUI:1216295.1216309})
+%attempts to address this by directing workers to tasks through ``intelligent task routing'' \cite{Cosley:2007:SUI:1216295.1216309}.
+and others focused on minimizing the amount of time that people need to spend doing anything other than the work for which they are paid~\cite{callison2014crowd}.
 
-Yet more work looks at the general framing of tasks%
-, chaining and arranging them to maximally exploit
-the attention and stress threshold % ???
-of workers
-\cite{Cai:2016:CRI:2858036.2858237}.
-Rather than attempt to minimize the error rates in micro--tasks%
-, as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested%
-, we as a community have leaned \textit{into} the peril of
-low--context work%
-, ``embracing error'' in crowdsourcing
-\cite{embracingErrorKrishna}.
-
-\subsubsubsection{\pieceworkpers}
+\subsubsection{\pieceworkpers}
 \topic{The research community relating to
@@
 `Arithmetic, the use of Logarithms, and Elementary Algebra'~''
-to compute, by hand, astronomical phenomena
+to compute astronomical phenomena
 \cite{grier2013computers}.
-These workers became the first \textit{computers}.
+%These workers became the first \textit{computers}.
 
-The work Airy solicited was interesting for several reasons.
+The work that Airy solicited resonates with modern crowd work for several reasons.
 First, work output was quickly verifiable;
@@
 
-The insight of breaking tasks down into smaller components didn't find its audience until
-the early \nth{20} century%
-, with the rise of Fordism and scientific management (or Taylorism).
-From scientific management, we found that
-we could measure work at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out%
-, piecework most greatly benefits the instrumented measurement of workers, but certainly
-in Ford and Taylor's time --- and certainly in Airy's time ---
-highly instrumented, automatic measurement of workers was all but impossible.
-As a result%
-, the distillation of work into smaller chunks
-ultimately reached a limit of usefulness.
+This approach found its audience in the early \nth{20} century with the rise of Fordism and scientific management (or Taylorism).
+Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
+As \citeauthor{Brown01041990} points out, piecework most greatly benefits the instrumented measurement of workers, but certainly in Ford and Taylor's time, highly instrumented, automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+As a result, the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+\msb{This subsection is too shallow and needs a bit more. For example, can you give examples of the last point?}
 
-\subsubsubsection{\whatchanged}
- \topic{A number of factors in crowdwork are different from piecework,
- chief among them being the relative ease with which
- the metaphorical ``assembly line'' can be changed.}
- Computers make it possible to switch from one task to another
-unlike any arbitrary manufacturing factory possibly could;
-a worker could do any number of
-different \textit{types} of tasks in the span of just a few minutes
-\cite{delayAndOrderLasecki}.
-This has spurred an entire body of work investigating the effects of
+\subsubsection{\whatchanged}
+\begin{comment}
+outline
+	- measurement is more precise, so decomposition is deeper
+	- not a single position, but a marketplace
+\end{comment}
+
+\topic{Where measurement and instrumentation were limiting factors for historical piecework, computation has changed the situation so that a dream of scientific management and Taylorism
+--- to measure every motion at every point throughout the workday and beyond ---
+is not only doable, but trivial
+\cite{waltz2012quantified}.}
+Where \citeauthor{10.2307/23702539} directly implicates measurement as
+% cites as
+preventing scientific management from being fully utilized,
+%the difficulty of tracking work \& workers, no longer exists
+%\cite{10.2307/23702539}.}
+modern crowd work is measuring and modeling every click, scroll, and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
+The result is that on--demand work can articulate and track far more carefully than piecework historically could.
+
+\topic{A second shift is  
+%A number of factors in crowd work are different from piecework, chief among them being 
+the relative ease with which the metaphorical ``assembly line'' can be changed.}
+% Computers make it possible to switch from one task to another unlike any arbitrary manufacturing factory possibly could;
+%modern on--demand workers on AMT and other platforms can more readily do any number of
+%different \textit{types} of tasks in the span of just a few minutes
+Historical manufacturing equipment was not Turing-complete, and could not quickly be assembled, edited, and redeployed.\msb{cite for historical piecework equipment being a large time + money investment?}
+In contrast, today system--designers can share, modify, and instantiate environments
+like sites of labor in a few lines of code
+\cite{lessig2006code,turkitLittle}.
+This opportunity has spurred an entire body of work investigating the effects of
 ordering,
@@
 other factors in piecework that would have been
-all but impossible to measure consistently as few as 20 years ago
-\cite{cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
+all but impossible to manipulate as few as 20 years ago
+\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
 
-Further, we've sliced work to such small scales that the marginal activities
---- things like finding work, cognitive task switching, etc\dots ---
+Third, modern crowd work has sliced work to such small scales that the marginal activities
+--- things like finding work and cognitive task switching ---
 have become relatively large compared to the tasks themselves
@@
 \cite{hart2013rise}.
+The result is that crowd workers are more free agents than historically was the case.
+However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take-home earnings than before.
 
-Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
---- one which where crowdworkers are rendered as
-``modular, protocol--defined computational services'' ---
-we may yield better results from crowdwork if we think of workers as similar to
-specialized, repurposable tools
-\cite{irani2015cultural}.
-\ali{feeling meh about this argument\dots}
+% Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
+% --- one which where crowd workers are rendered as
+% ``modular, protocol--defined computational services'' ---
+% we may yield better results from crowd work if we think of workers as similar to
+% specialized, repurposable tools
+% \cite{irani2015cultural}.
+% \ali{feeling meh about this argument\dots}
 
-Finally, instrumentation has reached a sufficiently advanced and ubiquitous point that
-the dream of scientific management and Taylorism
---- to measure every motion at every point throughout the workday and beyond ---
-is not only doable, but trivial
-\cite{waltz2012quantified}.
-One of the major challenges \citeauthor{10.2307/23702539} cites as
-preventing scientific management from being fully utilized,
-the difficulty of tracking work \& workers, no longer exists
-\cite{10.2307/23702539}.
+\subsubsection{\implication}
+If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely-sliced and highly decomposed than ever before.
+Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved, or near enough that they no longer limit task decomposition.
+Now, not just tasks, but entire workers' histories~\cite{hata2017glimpse}, can be collected and analyzed in detail.
+% crowd work research today is on the right track to investigate pipelining and meta--task design.
+% That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
+% \cite[see, for example,][]{turkopticon}.
 
-\subsubsubsection{\implication}
-Crowdwork research today is on the right track to investigate pipelining and meta--task design.
-That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
-\cite[see, for example,][]{turkopticon}.
-It's not clear how much benefit there is in the further decomposition of work,
-given that we've hit bottlenecks with the cognitive stresses of switching between tasks
-as \citeauthor{delayAndOrderLasecki} highlight
-\cite{delayAndOrderLasecki}.
+However, decomposition has hit a second bottleneck: cognition. 
+Task switching costs and other cognitive costs make it difficult to work tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
+There will of course be tasks that can be decomposed without much context, and these will form the most fine-grained of microtasks.
+However, other tasks cannot be freed from context --- for example, logo design requires a deep understanding of the client and their goals.
+In part due to this limitation, 99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-
-  % \clearpage
-  % \nobalance{}
-  }
-%   \endnotetext{
-% %     \section{Slicing notes}
-% %     The finding here is that
-% %     crowdwork can be more carefully micro--managed than piecework could be%
-% % ,     and that this is a double--edged sword:
-% %     we can effectively give feedback to workers on everything they do%
-% % ,     but this is emboldening us to try to
-% %     over--manage workers just as piecework tried to do.
-
-% %     My goal for this section is to make two points:
-% %     \begin{enumerate}
-% %       \item show how this is related to the assembly line and scientific management%
-% % ,             and how piecework literature tried to measure everything%
-% % ,             but found it untenable given the extra equipment that was necessary
-% %             (but generally which didn't exist)
-% %             to track every movement and action that workers took.
-% %       \item show how this work was enabled by the ``verifiability'' of work output(?)
-% %     \end{enumerate}
-
-%     % \ali{i don't like this section \textit{here}, but i like it in general. what do?}
-
-%   }
+So, ultimately, the levels of decomposition are likely to follow the contours of context required. Low-context work will be extremely highly decomposed. High-context work will continue to be limited.
 

### 2016-09-21 drafts/labor/decomposition.tex

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
-\onlyinsubfile{
-  \usepackage{xr-hyper}
+% \makeatletter
+% \def\blx@maxline{77}
+% \makeatother
+\onlyinsubfile{\usepackage{xr-hyper}
 \usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{relationships}
+\externaldocument{decomposition}
 }
-
 \begin{document}
@@
 %while other work designs tasks around gap time (\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning}) \cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,       Cai:2015:WLW:2702123.2702267}.
+
 % \topic{One of the emergent properties of micro--tasks has been the relative cost of
@@
 \msb{This subsection is too shallow and needs a bit more. For example, can you give examples of the last point?}
+
+\topic{Piecework researchers enumerate a number of problems with the decomposition of work, and the conflicting pressures managers and workers put forth.}
+\citeauthor{bewley1999wages} in particular points out that
+the approach of paying workers by the piece is
+``\dots~not practical for workers doing many tasks, because of
+the cost of establishing the rates and because
+piecework does not compensate workers for time spent switching tasks''.
+Ultimately, \citeauthor{bewley1999wages} argues that
+``[piecework is] infeasible, because \dots
+total output is the joint product of varying groups of people''
+\cite{bewley1999wages}.
 

### 2016-09-21 drafts/labor/piecework_lit.tex
piecework labeling

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\makeatother
+\onlyinsubfile{
+  \usepackage{xr-hyper}
+  \usepackage{hyperref}
+  \externaldocument{complexity}
+  \externaldocument{relationships}
+  \externaldocument{decomposition}
+}
 \begin{document}
@@
         the factors that may have led to piecework's
-        eventual demise \msb{did it demise? there are sweatshops around the world today}.
+        eventual demise in the American and European labor markets.
+        % \msb{did it demise? there are sweatshops around the world today}.
+        % \ali{is it okay to corner it off like this (by just being about the US/EU)?}
 \end{inlinelist}
 
+While piecework has been studied from myriad perspectives through the lenses of
+political science,
+economics (both micro-- and macro--), and
+a plethora of other fields
+\ali{should I cite examples of all of these?}, for the sake of brevity
+we will discuss the history of piecework in the contexts of the three questions we brought up with crowd work:
+``\namerefl{sec:complexity}'',
+``\namerefl{sec:decomposition}'', and 
+``\namerefl{sec:relationships}''.
+
 \subsection{What was piecework?}
-
-\topic{While ``piecework'' has proven difficult to concretize from the literature,
-we can trace a constellation of characteristics of piece work that recur throughout the literature.}
-% We'll follow the history of research, collecting
-\topic{Aligning on--demand work with piecework requires an understanding of what piecework is. While ``piecework'' has had multiple definitions over time,
-we can trace a constellation of characteristics that recur throughout the literature.}
+\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
+While ``piecework'' has had multiple definitions over time,
+we can trace a constellation of characteristics that recur throughout the literature.
 We will follow this history of research, collecting
@@
 \topic{It's worth acknowledging that
-``this distinction [between piece--rates and time--rates] was not completely clear--cut'' \msb{cite? who said that? Hart?};
-indeed, we see work that adopts
+``this distinction [between piece--rates and time--rates] was not completely clear--cut''
+\cite{hart2013rise}.}
+Indeed, we see work that adopts
 piece--rate compensation in some aspects and
- time--rate compensation in others
-\cite{hart2013rise}.}
+time--rate compensation in others.
 The ``Rowan premium system'',
@@
 and in other senses
-(more \ali{(familiar|conventional)})
+(more conventional)
 ``day--labourers''.
-\ali{I want to come back to this later when we talk about Uber \& Lyft, and
-     how they increasingly offer to guarantee drivers an hourly rate
-     \cite[see][]{uberHourly,lyftHourly}.
-     That's why I'm nitpicking about this muddied system --- to avoid someone saying
-     ``it doesn't seem like piecework to me anymore if there's an hourly wage floor''}
+% \ali{I want to come back to this later when we talk about Uber \& Lyft, and
+%      how they increasingly offer to guarantee drivers an hourly rate
+%      \cite[see][]{uberHourly,lyftHourly}.
+%      That's why I'm nitpicking about this muddied system --- to avoid someone saying
+%      ``it doesn't seem like piecework to me anymore if there's an hourly wage floor''}
 
@@
 Returning to
-\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piece work \msb{inconsistent use of ``piece work'' vs. ``piecework'' in the non-quoted text. Which do you want?}%
+\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
+% \msb{inconsistent use of ``piece work'' vs. ``piecework'' in the non-quoted text. Which do you want?}
+% \ali{Okay here's the problem: I'm happy to use ``piecework'' and that's the word that exists in the dictionary, but
+% at the time that \citeauthor{hughRaynbirdTaskWork} was writing, it was ``piece work'' and I don't know if
+% this is a trivial enough thing that I can just quietly concatenate his/their quotations so it looks like they say ``piecework''.
+% I think at this point I was just calling it ``piece work'' because at this point in history people were calling it that.
+% I've just changed everything out of quotation marks to ``piecework[er[s]]'', and
+% left everything in quotation marks as ``piece work''.}
 crop up; he points out that\dots
@@
 \citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
-finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
+finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
 ``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
@@
 \end{enumerate}
+\ali{woah i thought this was a comment; is it okay as is?}
 
 \subsection{The Historical Arc of Piecework}\label{sec:pieceworkArc}
-
 \topic{Piecework's history traces back further perhaps than most would expect.}
@@
 if he arranged his work appropriately.
-
 \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
@@
 Can you hone it?}
-
 \ali{I wanted that paragraph to be about the rising popularity and application of piecework
-(especially as it approached its ``heyday'' (a la \cite{hart2013rise})),
+(especially as it approached its ``heyday'' \cite{hart2013rise}),
 coming from humble beginnings as it found its footing.
-Given that intent\dots~
-
+Given that intent\dots\\
 Should I \textbf{refactor} or \textbf{rewrite}?}
 
-\topic{First applied to farm work, as
-\citeauthor{hughRaynbirdTaskWork} and others illustrate,
+\ali{I'm thinking of dropping most of the last paragraph and merging it with the next one. Thoughts?}
+
+\ali{Complexity?}
+\topic{First largely used among farm workers, as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
 the practice remained relatively obscure until
@@
 \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+\cite{riisOtherSideLives,norton1900textile}.
 
-% \topic{Piecework took a circuitous path in its rise to the mainstream%
-% each time finding additional ways to leverage the advantages of piecework.}\msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?} \ali{Yeah, this is really weird; commenting out for now.}
-Piecework then began to grow: first applied to farm work, as
+\ali{Complexity?}
+\topic{Piecework then began to grow: first applied to farm work, as
 \citeauthor{hughRaynbirdTaskWork} and others illustrate,
@@
 it blossomed in the textile industry
-\cite{hughRaynbirdTaskWork}.
+\cite{hughRaynbirdTaskWork}.}
 This growth was so marked that by the turn of the \nth{20} century,
@@
 \cite{burton1899commercial}.
+% \ali{Everything up till now has been pointing at complexity (right? if not, then~\dots~lol)}
 
+\ali{Complexity + Decomposition?}
 \topic{Researchers sought to understand the characteristics of piecework
@@
 \cite{Brown01041990}.
+% \ali{that last paragraph is a mix of complexity and decomposition I think}
 
+\ali{??? Why is this here? Can it be moved?}
 \topic{As increasing attention revealed problems in piecework as it related to workers,
@@
 Nevertheless, piecework continued to permeate low--skilled labor.
+% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
 
-% \topic{Piecework became an important contributor to the war effort in the Second World War,
-
-\topic{At this point, piecework became an important factor in the war effort for the Second World War,
+\ali{Decomposition + Relationships?}
+\topic{Piecework became an important factor in the war effort for the Second World War,
 cementing its role not only in American factories, but in industrial work around the world.}
@@
 affording them more comprehensive knowledge of the total scope of work.
+One might reflect on the observation that ``Rosie the Riveter''
+--- an icon of mid--\nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating} ---
+was herself a pieceworker
+\cite{davies2014origins}.
 
+\ali{Relationships}
 \topic{Despite the intense growth of the piecework approach to renumeration,
@@
 \cite{american1921problem,richards1904anything}.
-\citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
-\topic{As increasing attention revealed problems in piecework as it related to workers,
-workers themselves began to speak out about their frustration with this new regime.}
-It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
-but this led to industry organizations representing
-railway workers, mechanical engineers, and others contributing their perspectives
+As mounting attention increasingly revealed problems in piecework as it related to workers,
+workers themselves began to speak out about their frustration with this new regime.
+\citeauthor{riisOtherSideLives}'s photo--documentary work
+brought light to otherwise--invisible pieceworkers in their homes;
+this and other events led to organizations representing
+railway workers, mechanical engineers, and others
+who began to mount advocacy in defense and favor of workers
 \cite{american1921problem,richards1904anything,riisOtherSideLives}.
-%Nevertheless, piecework continued to permeate low--skilled labor.
-% \topic{Necessity drove the widespread adoption of piecework and work decomposition%
-% , perhaps more even than Ford's assembly line.}
-%\topic{Despite the intense growth of the piecework approach to renumeration,
-%this time was not without turmoil.}
-%These worker organizations weighed in on
-%(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
-%\cite{american1921problem,richards1904anything}.
-For example, \citeauthor{10.2307/3827491} describes worker resistance among a largely disempowered community --- young women employed by piecework
+\citeauthor{10.2307/3827491} tells us of the match--girls strike of 1888,
+one of the earliest and most famous successful worker strikes and perhaps
+the beginning of ``militant trade unionism''
 \cite{10.2307/3827491}.
+It wasn't an exaggeration when \citeauthor{weyer1894history} later said that
+``the match--girls' victory turned a new leaf in Trade Union annals'' \cite{weyer1894history};
+in the 30 years since the match--girls strike,
+the Trade Union Movement grew from
+``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
+\cite{webb1920history}.
 
+\ali{Relationships}
 \topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
@@
 
-\topic{However, piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
+\ali{???}
+\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
 \citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
@@
 
-\subsection{Why is piecework relevant to crowdwork?}
+\subsection{Why is piecework relevant to crowd work?}
 

### 2016-09-21 drafts/labor/cases.tex
decomposition revised
moved a bunch of piecework history to its constituent sections. exciting!

@@
 \begin{document}
-\section{Major Research Questions}
+\section{Research Questions}
 
@@
 
-\endnotetext{
-Here, \citeauthor{hu1961parallel}'s work,  saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',  parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work
-\cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,  and indeed the study of factory and mass manufacturing labor through the \nth{20} century,  substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,  and persists to this day as it colors
-researchers' goals and objectives in the study and design of crowd work.
+% \endnotetext{
+% Here, \citeauthor{hu1961parallel}'s work,  saying of assembly line work that
+% ``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',  parallels the approach that dominated early research into crowd work
+% --- namely, using non--expert crowds for complex work
+% \cite{hu1961parallel}.
+% This mindset in \citeauthor{hu1961parallel}'s analysis,  and indeed the study of factory and mass manufacturing labor through the \nth{20} century,  substantively owes its existence to scientific management
+% and the rigorous decomposition of work into tasks, discussed earlier,  and persists to this day as it colors
+% researchers' goals and objectives in the study and design of crowd work.
 
-Piecework's influence on the abstraction of work into tasks,  described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,  prompting work requesters to consider ways to design these now discrete tasks in ways that
-maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
-\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
-It's to this area that we now turn our attention.
+% Piecework's influence on the abstraction of work into tasks,  described above, is more than just caused by the decomposition of work;
+% work abstraction itself makes it possible for workers to come and go flexibly,  prompting work requesters to consider ways to design these now discrete tasks in ways that
+% maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
+% \textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
+% It's to this area that we now turn our attention.
 
-Piecework has seen work along this dimension spanning decades;
-\citeauthor{thompson1913time} investigate some of the ways that
-construction can benefit from the principles of scientific management.
-\citeauthor{thompson1913time}'s thesis asserts that
-task work is predicated on the accurate scientific management of work,  including the ``miscellaneous tasks''.
-\citeauthor{thompson1913time} argues
---- as early as \citeyear{thompson1913time} ---
-that ``\dots one may be challenged to find any class of work
-involving labor either indoors or out--of--doors
-where tasks cannot be fixed by proper time--study''
-\cite{thompson1913time}.
+% Piecework has seen work along this dimension spanning decades;
+% \citeauthor{thompson1913time} investigate some of the ways that
+% construction can benefit from the principles of scientific management.
+% \citeauthor{thompson1913time}'s thesis asserts that
+% task work is predicated on the accurate scientific management of work,  including the ``miscellaneous tasks''.
+% \citeauthor{thompson1913time} argues
+% --- as early as \citeyear{thompson1913time} ---
+% that ``\dots one may be challenged to find any class of work
+% involving labor either indoors or out--of--doors
+% where tasks cannot be fixed by proper time--study''
+% \cite{thompson1913time}.
 
-Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable
-\cite{murray1983decentralisation}.
-Textile work being a salient example,  it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work
-\cite{vezina1992light}.
+% Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable
+% \cite{murray1983decentralisation}.
+% Textile work being a salient example,  it took time for workers to acquire sufficient skill
+% to do every aspect of the work so that the garment would be accepted by the company soliciting that work
+% \cite{vezina1992light}.
 
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work
-\cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car
-\cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
-\cite{schoenberger1988fordism}.
+% A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work
+% \cite{restructuringPieceworkBaker}.
+% In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car
+% \cite{towardsGlobalFordism}.
+% By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
+% \cite{schoenberger1988fordism}.
 
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),  can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization
-\cite{writingMicroTasks}.
-By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically
-\cite{delayAndOrderLasecki}.
-Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage
-\cite{rogstadius2011assessment}.
+% This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),  can be seen today in crowd work and on--demand labor through the application of micro--tasks.
+% \citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization
+% \cite{writingMicroTasks}.
+% By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically
+% \cite{delayAndOrderLasecki}.
+% Perhaps more important, however, was that
+% the breaking down of work into tasks has made it more practical to evaluate work at each stage
+% \cite{rogstadius2011assessment}.
 
-\subsubsection{So how does this affect crowd work?}
-The work we've seen so far 
-\begin{itemize}
-  \item worst case: assembling iPhones (extant)
-  \item average case: railroad workers and assembly lines
-  \item high (complexity) case: 
-\end{itemize}
+% \subsubsection{So how does this affect crowd work?}
+% The work we've seen so far 
+% \begin{itemize}
+%   \item worst case: assembling iPhones (extant)
+%   \item average case: railroad workers and assembly lines
+%   \item high (complexity) case: 
+% \end{itemize}
 
-% so CW?
-% mechanism/limitation in pw?
-% differences in cw? tech/access to intelligence?
-% predicted outcome then? -> decompose effort/expertise access
-% possible quirks/diffs?
+% % so CW?
+% % mechanism/limitation in pw?
+% % differences in cw? tech/access to intelligence?
+% % predicted outcome then? -> decompose effort/expertise access
+% % possible quirks/diffs?
 
-% what can be decomposed? what can be evaluated?
-% technology/turuing completeness gets us a win
-% predefined expertise? you can find someone almost instantly;
-% ease of doing this makes it qualitatively different due to Channel Factors (diverting a channel/river)
-% huddler paper cites some decomposition stuff workflows are good/bad at things
-% 3-way slack w/msb and mav
-% academy of management 
-% van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
-}
+% % what can be decomposed? what can be evaluated?
+% % technology/turuing completeness gets us a win
+% % predefined expertise? you can find someone almost instantly;
+% % ease of doing this makes it qualitatively different due to Channel Factors (diverting a channel/river)
+% % huddler paper cites some decomposition stuff workflows are good/bad at things
+% % 3-way slack w/msb and mav
+% % academy of management 
+% % van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
+% }
 

### 2016-09-21 drafts/labor/complexity.tex
decomposition revised
moved a bunch of piecework history to its constituent sections. exciting!

@@
 
-\subsubsection[What are the limits of crowd work]{Identifying the Limits of Crowd Work}\label{sec:complexity}
-\subsubsubsection{\crowdworkpers}
+\subsection[What are the limits of crowd work]{Identifying the Limits of Crowd Work}\label{sec:complexity}
+\subsubsection{\crowdworkpers}
 \topic{Crowdsourcing research has spent the better part of a decade attempting
@@
 
-\subsubsubsection{\pieceworkpers}
+\subsubsection{\pieceworkpers}
+\topic{First largely used among farm workers, as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
+the practice remained relatively obscure until
+it was brought to the textile industry
+\cite{hughRaynbirdTaskWork}.}
+At the turn of the \nth{20} century,
+when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+\cite{riisOtherSideLives,norton1900textile}.
+
+\topic{Piecework then began to grow: first applied to farm work, as
+\citeauthor{hughRaynbirdTaskWork} and others illustrate,
+the practice remained relatively obscure until
+it blossomed in the textile industry
+\cite{hughRaynbirdTaskWork}.}
+This growth was so marked that by the turn of the \nth{20} century,
+\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+offering guidance on how best to manage pieceworkers
+\cite{riisOtherSideLives,norton1900textile}.
+\citeauthor{clark1908cotton}, for instance,
+relays his observations of textile mill pieceworkers and his realization that
+``When he works by the day the Italian operative wishes to leave before the whistle blows,
+but if he works by the piece he will work as many hours as it is possible for him to stand''
+\cite{clark1908cotton}.
+During this period, best practices regarding the measurement and management of
+piecework rates, and of workers in the engineering industry,
+were beginning to take shape
+\cite{burton1899commercial}.
+% \ali{Everything up till now has been pointing at complexity (right? if not, then~\dots~lol)}
+
+\topic{Researchers sought to understand the characteristics of piecework
+that fueled its rise to popularity.}
+\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
+could be found in piecework;
+the approach of paying workers for each piece of output necessitated
+the rigorous tracking, measurement, and training of workers
+for which scientific management became famous
+\cite{10.2307/23702539}.
+This argument is certainly compelling;
+it would seem to make the concurrent upswing of
+scientific management and Fordism
+through the first two--thirds of the \nth{20} century
+alongside piecework not only understandable, but predictable
+\cite{hart2013rise}.
+\citeauthor{Brown01041990} inquired from another direction, asking
+what limited the adoption of piecework in industries that otherwise gravitated toward it
+(in the case studies he examined, this mostly focused on railway engineers),
+ultimately arguing that factors such as the nature of the work design
+(specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
+were the major contributing factors that determined the use of piecework
+\cite{Brown01041990}.
+% \ali{that last paragraph is a mix of complexity and decomposition I think}
+
 \topic{Piecework researchers have found themselves trying to understand what characteristics limit piecework,
@@
 
-\subsubsubsection{\whatchanged}
+\subsubsection{\whatchanged}
 \msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework.
@@
 
-\subsubsubsection{\implication}
+\subsubsection{\implication}
 \msb{This paragraph can be expanded to make a more concrete argument. What will be possible? What won't?}

### 2016-09-21 drafts/labor/decomposition.tex
moved a bunch of piecework history to its constituent sections. exciting!

@@
 \subsubsection{\pieceworkpers}
+\topic{Piecework became an important factor in the war effort for the Second World War,
+cementing its role not only in American factories, but in industrial work around the world.}
+The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.
+As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+He attributes this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+One might reflect on the observation that ``Rosie the Riveter''
+--- an icon of mid--\nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating} ---
+was herself a pieceworker
+\cite{davies2014origins}.
+
 \topic{The research community relating to

### 2016-09-21 drafts/labor/piecework_lit.tex
moved a bunch of piecework history to its constituent sections. exciting!

@@
 
-While piecework has been studied from myriad perspectives through the lenses of
-political science,
-economics (both micro-- and macro--), and
-a plethora of other fields
-\ali{should I cite examples of all of these?}, for the sake of brevity
-we will discuss the history of piecework in the contexts of the three questions we brought up with crowd work:
-``\namerefl{sec:complexity}'',
-``\namerefl{sec:decomposition}'', and 
-``\namerefl{sec:relationships}''.
+% While piecework has been studied from myriad perspectives through the lenses of
+% political science,
+% economics (both micro-- and macro--), and
+% a plethora of other fields
+% \ali{should I cite examples of all of these?}, for the sake of brevity
+% we will discuss the history of piecework in the contexts of the three questions we brought up with crowd work:
+% ``\namerefl{sec:complexity}'',
+% ``\namerefl{sec:decomposition}'', and 
+% ``\namerefl{sec:relationships}''.
 
@@
 
+\ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}
+
 \subsection{The Historical Arc of Piecework}\label{sec:pieceworkArc}
@@
 
-\ali{I'm thinking of dropping most of the last paragraph and merging it with the next one. Thoughts?}
-
-\ali{Complexity?}
-\topic{First largely used among farm workers, as
-\citeauthor{hughRaynbirdTaskWork} and others discuss,
-the practice remained relatively obscure until
-it was brought to the textile industry
-\cite{hughRaynbirdTaskWork}.}
-At the turn of the \nth{20} century,
-when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
-\citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
-\cite{riisOtherSideLives,norton1900textile}.
-
-\ali{Complexity?}
-\topic{Piecework then began to grow: first applied to farm work, as
-\citeauthor{hughRaynbirdTaskWork} and others illustrate,
-the practice remained relatively obscure until
-it blossomed in the textile industry
-\cite{hughRaynbirdTaskWork}.}
-This growth was so marked that by the turn of the \nth{20} century,
-\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
-and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
-offering guidance on how best to manage pieceworkers
-\cite{riisOtherSideLives,norton1900textile}.
-\citeauthor{clark1908cotton}, for instance,
-relays his observations of textile mill pieceworkers and his realization that
-``When he works by the day the Italian operative wishes to leave before the whistle blows,
-but if he works by the piece he will work as many hours as it is possible for him to stand''
-\cite{clark1908cotton}.
-During this period, best practices regarding the measurement and management of
-piecework rates, and of workers in the engineering industry,
-were beginning to take shape
-\cite{burton1899commercial}.
-% \ali{Everything up till now has been pointing at complexity (right? if not, then~\dots~lol)}
-
-\ali{Complexity + Decomposition?}
-\topic{Researchers sought to understand the characteristics of piecework
-that fueled its rise to popularity.}
-\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
-could be found in piecework;
-the approach of paying workers for each piece of output necessitated
-the rigorous tracking, measurement, and training of workers
-for which scientific management became famous
-\cite{10.2307/23702539}.
-This argument is certainly compelling;
-it would seem to make the concurrent upswing of
-scientific management and Fordism
-through the first two--thirds of the \nth{20} century
-alongside piecework not only understandable, but predictable
-\cite{hart2013rise}.
-\citeauthor{Brown01041990} inquired from another direction, asking
-what limited the adoption of piecework in industries that otherwise gravitated toward it
-(in the case studies he examined, this mostly focused on railway engineers),
-ultimately arguing that factors such as the nature of the work design
-(specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
-were the major contributing factors that determined the use of piecework
-\cite{Brown01041990}.
-% \ali{that last paragraph is a mix of complexity and decomposition I think}
+\ali{\sout{I'm thinking of dropping most of the last paragraph and merging it with the next one. Thoughts?}
+I moved it to \nameref{sec:complexity}}
 
@@
 % \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
-
-\ali{Decomposition + Relationships?}
-\topic{Piecework became an important factor in the war effort for the Second World War,
-cementing its role not only in American factories, but in industrial work around the world.}
-The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.
-As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-He attributes this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter''
---- an icon of mid--\nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating} ---
-was herself a pieceworker
-\cite{davies2014origins}.
-
-\ali{Relationships}
-\topic{Despite the intense growth of the piecework approach to renumeration,
-this time was not without turmoil.}
-As previously discussed, a number of worker organizations weighed in on
-(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
-\cite{american1921problem,richards1904anything}.
-As mounting attention increasingly revealed problems in piecework as it related to workers,
-workers themselves began to speak out about their frustration with this new regime.
-\citeauthor{riisOtherSideLives}'s photo--documentary work
-brought light to otherwise--invisible pieceworkers in their homes;
-this and other events led to organizations representing
-railway workers, mechanical engineers, and others
-who began to mount advocacy in defense and favor of workers
-\cite{american1921problem,richards1904anything,riisOtherSideLives}.
-\citeauthor{10.2307/3827491} tells us of the match--girls strike of 1888,
-one of the earliest and most famous successful worker strikes and perhaps
-the beginning of ``militant trade unionism''
-\cite{10.2307/3827491}.
-It wasn't an exaggeration when \citeauthor{weyer1894history} later said that
-``the match--girls' victory turned a new leaf in Trade Union annals'' \cite{weyer1894history};
-in the 30 years since the match--girls strike,
-the Trade Union Movement grew from
-``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
-\cite{webb1920history}.
-
-\ali{Relationships}
-\topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
-The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
---- albeit critical ---
-case study in this space;
-the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
-to strike for
-an individual minimum wage, among other demands
-\cite{10.2307/2221944}.
-\citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
-\cite{10.2307/41829256}.
-The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
-\cite{jacoby1983union}.
 

### 2016-09-21 drafts/labor/relationships.tex
decomposition revised
moved a bunch of piecework history to its constituent sections. exciting!

@@
 
-\subsubsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}
+\subsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}
 \label{sec:relationships}
 
-\subsubsubsection{\crowdworkpers}
+\subsubsection{\crowdworkpers}
 \topic{The relationships of workers
@@
 
-\subsubsubsection{\pieceworkpers}
+\subsubsection{\pieceworkpers}
+\topic{Despite the intense growth of the piecework approach to renumeration,
+this time was not without turmoil.}
+As previously discussed, a number of worker organizations weighed in on
+(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
+\cite{american1921problem,richards1904anything}.
+As mounting attention increasingly revealed problems in piecework as it related to workers,
+workers themselves began to speak out about their frustration with this new regime.
+\citeauthor{riisOtherSideLives}'s photo--documentary work
+brought light to otherwise--invisible pieceworkers in their homes;
+this and other events led to organizations representing
+railway workers, mechanical engineers, and others
+who began to mount advocacy in defense and favor of workers
+\cite{american1921problem,richards1904anything,riisOtherSideLives}.
+\citeauthor{10.2307/3827491} tells us of the match--girls strike of 1888,
+one of the earliest and most famous successful worker strikes and perhaps
+the beginning of ``militant trade unionism''
+\cite{10.2307/3827491}.
+It wasn't an exaggeration when \citeauthor{weyer1894history} later said that
+``the match--girls' victory turned a new leaf in Trade Union annals'' \cite{weyer1894history};
+in the 30 years since the match--girls strike,
+the Trade Union Movement grew from
+``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
+\cite{webb1920history}.
+
+\topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
+The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
+--- albeit critical ---
+case study in this space;
+the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
+to strike for
+an individual minimum wage, among other demands
+\cite{10.2307/2221944}.
+\citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
+\cite{10.2307/41829256}.
+The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
+\cite{jacoby1983union}.
+
 \topic{The questions surrounding
@@
 
-\subsubsubsection{\whatchanged}
+\subsubsection{\whatchanged}
 \topic{The differences between crowd workers and pieceworkers seem defined largely by
@@
 
-\subsubsubsection{\implication}
+\subsubsection{\implication}
 \topic{What we've done in the field of crowd work might be able to tell us something about piecework

### 2016-09-21 drafts/labor/complexity.tex
piecework history of complexity revised

--- 2016-09-21 drafts/labor/complexity.tex
@@
 
-\subsubsection[What are the limits of crowd work]{Identifying the Limits of Crowd Work}\label{sec:complexity}
-\subsubsubsection{\crowdworkpers}
-\topic{Crowdsourcing research has spent the better part of a decade attempting
-to prove the viability of crowdsourcing in increasingly complex work.}  % this is my topic sentence.
-\citeauthor{crowdForgeKittur}
-map the discussion toward this goal in their work on
-crowdsourcing complex work
-\cite{crowdForgeKittur}.
-The broader body of work has varied significantly in type
---- providing conversational assistants,
-interpreting medical data, and
-telling coherent and compelling stories,
-to name a few examples
-\cite{Lasecki:2013:CCC:2501988.2502057,mavandadi2012distributed,KimStoria}.
+\begin{comment}
+Crowdwork
+  - Kittur said let's do complex stuff
+  - This works by using CS techniques
+  - Clear that this works in focused cases
+  - More recent shift toward using experts
+\end{comment}
 
-\topic{This body of research has involved similar approaches to problems,
-often involving insights made in Computer Science and applied to human work--flows.}
-The crowd work literature typically identifies target milestones in computer science
-that have presented significant challenges for researchers,
-leverages some of the approaches and insights that Computer Science researchers have already made
-(for example, MapReduce in the case of \citeauthor{crowdForgeKittur}'s \textit{CrowdForge}),
-and arranges humans as computational black boxes within those approaches and processes
-\cite[][and others]{crowdForgeKittur,foundry}.
+\subsection[What are the complexity limits of crowd work]{Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
+
+\subsubsection{\crowdworkpers}
+
+\topic{Crowdsourcing research has spent the better part of a decade proving the viability of crowdsourcing in complex work.}  % this is my topic sentence.
+Unless crowdsourcing can demonstrate viability for meaningfully complex tasks, the argument runs, it will be incapable of ensuring a pro-social outcome for work and workers~\cite{crowdworkFuture}.
+\citeauthor{crowdForgeKittur} first opened the question of whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
+Their work demonstrated proof-of-concept crowdsourcing of a simple encyclopedia article and news summary.
+Seeking to raise the complexity ceiling~\cite{myers2000past}, researchers have since created additional proof-of-concept applications and techniques, including conversational assistants~\cite{Lasecki:2013:CCC:2501988.2502057}, medical data interpreters~\cite{Lasecki:2013:CCC:2501988.2502057}, and idea generation~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}, to name a few examples.
+
+\topic{This body of research has often invoked insights from Computer Science and applied them to human workflows.}
+The crowd work literature typically identifies target milestones that have presented significant challenges for researchers, leverages an insight from Computer Science (for example, MapReduce~\cite{crowdForgeKittur} or sequence alignment algorithms~\cite{lasecki2012real}), and arranges humans as computational black boxes within those approaches.
 This approach has proven a compelling one because
-it leverages the in--built advantages that technology and digital media afford.
-For example, \textit{Foundry}'s tools for managing and arranging expert groups into a cohort
-allow researchers to convincingly argue that expert teams can be rapidly formed,
-just like non--expert teams
-\cite{foundry}.
+it leverages the in--built advantages of scale, automation, and programmability that software affords.
+%For example, \textit{Foundry}'s tools for managing and arranging expert groups into a cohort allow researchers to convincingly argue that expert teams can be rapidly formed, just like non--expert teams \cite{foundry}.
 
-\subsubsubsection{\pieceworkpers}
-\topic{Piecework researchers have found themselves trying to understand what characteristics limit piecework,
-or more precisely what has prevented piecework from becoming more prevalent.}
-The research into piecework makes the case that
-piecework has been limited principally by the challenges of human management and oversight.
-\citeauthor{10.2307/23702539} describes a case study in Santa Fe Railway, which
-deployed scientific management and a piecework regime in an attempt to stymie rising repair costs
-\cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} reports on the hiring of Harrington Emerson,
-an ``efficiency expert'', who went on to develop a ``standard time'' for each task at the company
-informed by ``thousands of individual operations at the Topeka shops''.
-The cost of measuring workers in such excruciating detail at the turn of the \nth{20} century was undoubtedly immense,
-but this ``standard time'' value, which determined the pay that workers would earn for each task they do,
-was the only viable approach at the time to determine appropriate pay given the task
-\cite{10.2307/23702539}.
-But the repeated measurement of workers' time to complete tasks had shortcomings;
-for one thing, pay rates for rarer tasks
-were necessarily less certain than for more common tasks,
-which had the simple benefit of a larger sample size.
-\ali{Do I need ``for another thing\dots''? This paragraph is kind of huge already. Not sure how to break it up/down as it is.}
+It is now clear that this approach works with focused complex tasks, but the broader wicked problems~\cite{rittel1973dilemmas} remain unsolved. 
+As a first example, idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}, but there is as yet no general crowdsourced solution for the broader goal of invention and innovation~\cite{fuge2014analysis}.
+Second, focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,writingMicroTasks,agapie2015crowdsourcing}, but there is no general solution to create a high-quality crowd-powered author across domains.
+Third, data analysis tasks such as clustering~\cite{chilton2013cascade}, categorization~\cite{andre2014crowd}, and outlining~\cite{luther2015crowdlines} are possible, but there is no general solution for sensemaking.
+It is not yet clear what insights would be required to enable crowdsourced solutions for these broader wicked problems and other crowdsourced goals such as addressing climate change~\cite{introne2011climate}.
+
+Restricting attention to non-expert microtask workers limited the ambitions of research in the area. 
+So, \citeauthor{foundry} introduced the idea of crowdsourcing with online paid \textit{experts} from platforms such as Upwork.
+Expert crowdsourcing enables access to a much broader set of workers, for example designers and programmers.
+Researchers then apply similar Computer Science techniques but to expert ``macro-tasks''~\cite{cheng2015break,haas2015argonaut}.
+This approach has led to successful crowdsourcing of goals such as the user-centered design process~\cite{foundry}, programming tasks~\cite{latoza2014microtask,Fast2016,Chen2016}, and mentorship~\cite{suzukiAtelier}.
+However, there remains the open question of how complex the work outcomes from these approaches will be.
+
+\subsubsection{\pieceworkpers}
+
+\begin{comment}
+- Farm workers-->textile
+- Limit: human management and oversight
+- Evaluation
+- Skilled work harder
+- Only some organizations can use it
+- Management practices
+\end{comment}
+
+\topic{\citeauthor{grier2013computers} gives early accounts of a piecework strategy in Airy's creation of the British Nautical Almanac~\cite{grier2013computers}.} 
+Airy's goal was complex --- mathematical calculations to produce tables that would allow sailors to locate themselves by starlight from sea. 
+Many of his contributors did not have high-level mathematical training, so Airy broke down the task into simpler calculations and distributed them by mail, accomplishing the complex goal through piecework tasks that paid little.
+
+However, when piecework entered the American economy, it was not used for complex work.
+Initially for farm work, as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
+the practice remained relatively obscure until
+%it was brought to the textile industry
+%\cite{hughRaynbirdTaskWork}.
+%\topic{Piecework then began to grow: first applied to farm work, as
+%\citeauthor{hughRaynbirdTaskWork} and others illustrate,
+%the practice remained relatively obscure until
+it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+Complexity levels remained low at the turn of the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+%was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+However, writers of the time focused their attentive on wage~\cite{burton1899commercial} and management regimes \cite{norton1900textile}.
+%\citeauthor{clark1908cotton}, for instance,
+%relays his observations of textile mill pieceworkers and his realization that
+%``When he works by the day the Italian operative wishes to leave before the whistle blows,
+%but if he works by the piece he will work as many hours as it is possible for him to stand''
+%\cite{clark1908cotton}.
+%During this period, best practices regarding the measurement and management of
+%piecework rates, and of workers in the engineering industry,
+%were beginning to take shape
+%\cite{burton1899commercial}.
+% \ali{Everything up till now has been pointing at complexity (right? if not, then~\dots~lol)}
+The work remained low-skilled in part because it was infeasible to provide new pieceworkers with the comprehensive education that was familiar through apprenticeships \cite{hart2013rise}.
+
+Measurement limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework. 
+When \citeauthor{Brown01041990} investigated what limited the adoption of piecework in industries that otherwise gravitated toward it (e.g., railway engineers), the homogeneity of tasks %and  and the costs associated with adopting a piecework model
+arose as a major contributing factor~\cite{Brown01041990}.
+% \ali{that last paragraph is a mix of complexity and decomposition I think}
+%\topic{Piecework researchers have found themselves trying to understand what characteristics limit piecework,
+%or more precisely what has prevented piecework from becoming more prevalent.}
+%The research into piecework makes the case that piecework has been limited principally by the challenges of human management and oversight.
+\citeauthor{10.2307/23702539} concurs via a case study in Santa Fe Railway, which used ``efficiency experts'' to develop a ``standard time'' to determine pay for each task at the company informed by ``thousands of individual operations at the Topeka shops''~\cite{10.2307/23702539}.
+%The cost of measuring workers in such excruciating detail at the turn of the \nth{20} century was undoubtedly immense,
+%but this ``standard time'' value, which determined the pay that workers would earn for each task they do,
+%was the only viable approach at the time to determine appropriate pay given the task
+%\cite{10.2307/23702539}.
+%But the repeated measurement of workers' time to complete tasks had shortcomings;
+%for one thing, pay rates for rarer tasks
+%were necessarily less certain than for more common tasks,
+%which had the simple benefit of a larger sample size.
+%\ali{Do I need ``for another thing\dots''? This paragraph is kind of huge already. Not sure how to break it up/down as it is.}
 One might conclude from \citeauthor{10.2307/23702539}'s observations that
@@
 
-\topic{Determining appropriate pay rates, informed by the careful measurement of workers,
-isn't the only major challenge piecework faced; evaluation proved a limiting factor as well.}
-\citeauthor{10.2307/23702539} enumerates some of the roles required
-to facilitate piecework in the early \nth{20} century --- among them
-    ``\dots~piecework clerks, inspectors, and `experts'\dots''
+\topic{Piecework was also limited to tasks that could be clearly evaluated.}
+The roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''
 \cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} further recognizes that
-it's necessary for a successful piecework shop to employ
-clerks,
-inspectors, and
-other experts to properly design and evaluate complex work.
-\citeauthor{hart2016rise} later makes a more concrete observation of this hurdle,
-as he argues an ultimate limit to how far this can go;
-at some point, evaluating multidimensional work output for quality
-(rather than for quantity) becomes infeasible.
+%\citeauthor{10.2307/23702539} further recognizes that
+%it's necessary for a successful piecework shop to employ
+%clerks,
+%inspectors, and
+%other experts to properly design and evaluate complex work.
+\citeauthor{hart2016rise} argues for an ultimate complexity limit: at some point, evaluating multidimensional work for quality (rather than for quantity) becomes infeasible.
 In his words,
-``if the quality of the output is more difficult to measure than the quantity,
-perhaps because of `difficult--to--observe' production techniques,
+``if the quality of the output is more difficult to measure than the quantity [\ldots]
+%perhaps because of `difficult--to--observe' production techniques,
 then a piecework system is likely to encourage
@@
 \cite{hart2016rise}.
-This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
-and ultimately precipitated its downfall,
-especially with the increasing complexity of manufacturing work
-over the course of the \nth{20} century.
-\ali{just to be clear, this whole point is queuing up so that later I can be like
-``Oh hey computers are terrific at evaluating some stuff really quickly.
-And researchers in crowd work have done some work on workers evaluating other workers
-(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line)
-\cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
+%This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
+%and ultimately precipitated its downfall,
+%especially with the increasing complexity of manufacturing work
+%over the course of the \nth{20} century.
+%\ali{just to be clear, this whole point is queuing up so that later I can be like
+%``Oh hey computers are terrific at evaluating some stuff really quickly.
+%And researchers in crowd work have done some work on workers evaluating other workers
+%(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line)
+%\cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
+Complex work, which is often subjective to evaluate, falls victim to this criteria.
+
+\topic{This focus on measurement and tracking had consequences.}
+\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
+could be found in piecework:
+the approach of paying workers for each piece of output necessitated
+the rigorous tracking, measurement, and training of workers
+for which scientific management became famous
+\cite{10.2307/23702539}.
+If true, 
+%This argument is certainly compelling;
+%it would seem to make 
+the concurrent upswing of
+scientific management and Fordism
+through the first two--thirds of the \nth{20} century
+alongside piecework was not only understandable, but predictable
+\cite{hart2013rise}.
+
+\topic{Piecework researchers also argue that,
+in addition to constraints on the kind of \textit{work} that's amenable to piecework,
+only certain kinds of \textit{organizations} are amenable to piecework.}
+Researchers detail three organizational criteria.
+% \msb{I think a reader wouldn't understand why that claim is relevant.
+% Let's invert it, something like how
+% the researchers point out that
+% only certain kinds of organizations can effectively make use of piecework.}
+First, \citeauthor{Brown01041990} argues that piecework ``is less likely in jobs with
+a variety of duties than in jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
+\citeauthor{SJOE:SJOE371} points out the phenomenon here as a market effect:
+``in an environment with multi--tasking, pay schemes based on tightly specified performance may \textit{induce workers to neglect tasks} that are less easy to measure''
+(emphasis added) \cite{SJOE:SJOE371}.
+% \msb{that's confusing, is incentive pay the same as piecework?
+% can you cut the first part of the quote and
+% give us the noun instead of forcing us to figure it out?}.
+Second, complexity was limited by access to capital to create the necessary infrastructure.
+As \citeauthor{10.2307/23702539} reports, %that successful cases of piecework owed themselves in part to the fact that
+only the largest and most wealthy railroads had the resources to ``pay the overhead involved in installing work reorganization'' \cite{10.2307/23702539}.
+% Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
+% make a persuasive argument that piecework is limited in complexity by
+% the capacity to endure managerial overhead while transitioning to a new system.
+% \msb{wait, wasn't that in a previous paragraph?
+% that should be joined with the managerial overhead text above}
+% \ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
+% the first is that there are limits on doing hard, ``difficult--to--observe'' work,
+% and the second is that the cost propagates up to management,
+% putting limits on who can run piecework systems.
+% So\dots\\
+% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
+% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
+% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop
+% Maybe I should just merge it with the next paragraph?}
+Third, organizations required capable managers in charge of the pieceworkers.
+% \topic{There are other characteristics to effective complex piecework institutions, such as
+% appropriately designed management practices.}
+% \citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
+% ``The foreman had the power to hire and fire workers and allocate workplaces,
+% but then left the face--worker largely free to his own efforts so that
+% often he went all day without seeing the foreman''
+% \cite{10.2307/2118435}.
+The West Virginia mines, for example, hired foremen 
+% The general approach adopted by these West Virginia mines was,
+% as in other factories with active foremen,
+to be the intermediary between upper management and the worker~\cite{10.2307/2118435}.
+Specifically, foremen were responsible for allocating resources and
+understanding when and how to modify work as necessary
+\cite{wray1949marginal}.
+% \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
+% The management of pieceworkers %, \citeauthor{10.2307/2118435} and later \citeauthor{wray1949marginal} argue,
+% demanded people in positions akin to foremen
+% --- intermediate managers people who were
+% \begin{inlinelist}
+%   \item familiar with and even sympathetic to the needs of workers,
+%   \item empowered by higher level management to make decisions, and
+%   \item relaxed enough in day--to--day work to allow workers to go about their work
+% \end{inlinelist}
+% \cite{wray1949marginal,10.2307/2118435}.
+So, in sum, organizations historically could only take advantage of piecework if they had homogenous work to be done, access to capital to purchase the necessary equipment, and the ability to hire intermediate management.
 
@@
 
-\topic{This isn't to say that complex work is outside of the realm of piecework;
-indeed, we've discussed complex applications of crowd work already.}
-While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design,
-much of it arose out of necessity ---
-it was infeasible to provide new workers with
-the comprehensive education that was familiar to men through apprenticeships
-\cite{hart2013rise}.
-While this constraint led to much more tightly scoped work
-who now had to specialize in extremely narrowly defined roles.
-The same could be said of Airy and his \textit{computers}
---- young boys whose preparations consisted principally of
-a relatively specific mathematics curriculum
-\cite{grier2013computers}.
-Instead, we argue that the literature suggests that
-piecework is tightly limited only when the application of piecework follows
-a direct, perhaps even unimaginative,
-mapping from an time--based regime to an output--based one.
-When the work is redesigned from the ground up 
---- as we see with mathematicians in the \nth{19} century and
-with the metalworking industry during the Second World War ---
-it seems that we don't yet know the limits of complexity with regard to piecework.
-\ali{This is \textit{almost} making a prediction,
-but I wanted to take something punchy away from the stuff I cite here.}
+%\topic{This isn't to say that complex work is outside of the realm of piecework; indeed, we've discussed complex applications of crowd work already.}
+%While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design, much of it arose out of necessity --- it was infeasible to provide new workers with the comprehensive education that was familiar to men through apprenticeships \cite{hart2013rise}.
+%While this constraint led to much more tightly scoped work who now had to specialize in extremely narrowly defined roles.
+%The same could be said of Airy and his \textit{computers} --- young boys whose preparations consisted principally of a relatively specific mathematics curriculum \cite{grier2013computers}.
+% Instead, we argue that the literature suggests that piecework is tightly limited only when the application of piecework follows a direct, perhaps even unimaginative, mapping from an time--based regime to an output--based one.
+% When the work is redesigned from the ground up --- as we see with mathematicians in the \nth{19} century and with the metalworking industry during the Second World War --- it seems that we don't yet know the limits of complexity with regard to piecework.
+%\ali{This is \textit{almost} making a prediction, but I wanted to take something punchy away from the stuff I cite here.}
 
-\topic{Piecework researchers also argue that,
-in addition to constraints on the kind of \textit{work} that's amenable to piecework,
-only certain kinds of \textit{organizations} are amenable to piecework.}
-% \msb{I think a reader wouldn't understand why that claim is relevant.
-% Let's invert it, something like how
-% the researchers point out that
-% only certain kinds of organizations can effectively make use of piecework.}
-\citeauthor{Brown01041990} discusses the organizational factors necessary for piecework to thrive,
-arguing that piecework
-``\dots~ is less likely in jobs with
-a variety of duties than in jobs with a narrow set of routinized duties''
-\cite{Brown01041990}
-% \msb{that's confusing, is incentive pay the same as piecework?
-% can you cut the first part of the quote and
-% give us the noun instead of forcing us to figure it out?}.
-\citeauthor{10.2307/23702539} adds further, that
-successful cases of piecework owed themselves in part to the fact that
-    ``\dots~only [the largest and most wealthy railroads] had the resources to~\dots
-    pay the overhead involved in installing work reorganization''
-\cite{10.2307/23702539}.
-Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
-make a persuasive argument that piecework is limited in complexity by
-the capacity to endure managerial overhead while transitioning to a new system.
-\msb{wait, wasn't that in a previous paragraph?
-that should be joined with the managerial overhead text above}
-\ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
-the first is that there are limits on doing hard, ``difficult--to--observe'' work,
-and the second is that the cost propagates up to management,
-putting limits on who can run piecework systems.
-So\dots\\
-\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
-\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
-\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop
-
-Maybe I should just merge it with the next paragraph?}
-
-\topic{There are other characteristics to effective complex piecework institutions, such as
-appropriately designed management practices.}
-\citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
-``The foreman had the power to hire and fire workers and allocate workplaces,
-but then left the face--worker largely free to his own efforts so that
-often he went all day without seeing the foreman''
-\cite{10.2307/2118435}.
-The general approach adopted by these West Virginia mines was,
-as in other factories with active foremen,
-to let the foreman be the intermediary between management and the worker.
-Specifically, foremen were responsible for allocating resources and
-understanding when and how to modify work as necessary
-\cite{wray1949marginal}.
-% \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
-The management of pieceworkers %, \citeauthor{10.2307/2118435} and later \citeauthor{wray1949marginal} argue,
-demanded people in positions akin to foremen
---- intermediate managers people who were
-\begin{inlinelist}
-  \item familiar with and even sympathetic to the needs of workers,
-  \item empowered by higher level management to make decisions, and
-  \item relaxed enough in day--to--day work to allow workers to go about their work
-\end{inlinelist}
-\cite{wray1949marginal,10.2307/2118435}.
-
-\subsubsubsection{\whatchanged}
+\subsubsection{\whatchanged}
 \msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework.

### 2016-09-21 drafts/labor/decomposition.tex
piecework history of complexity revised

@@
 \subsubsection{\pieceworkpers}
+\citeauthor{Brown01041990} inquired from another direction, asking
+what limited the adoption of piecework in industries that otherwise gravitated toward it
+(in the case studies he examined, this mostly focused on railway engineers),
+ultimately arguing that factors such as the nature of the work design
+(specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
+were the major contributing factors that determined the use of piecework
+\cite{Brown01041990}.
+
 \topic{Piecework became an important factor in the war effort for the Second World War,

### 2016-09-21 drafts/labor/piecework_lit.tex
added a ref shortname
piecework checkin

@@
 
+\ali{Maybe something here about how this isn't comprehensive, but it gives you enough to understand the
+era about which we're talking, the very high--level things that were going on, etc\dots~so that
+the later discussions actually make sense?}
+
 % While piecework has been studied from myriad perspectives through the lenses of
@@
 
-\subsection{What was piecework?}
+\subsection{What is piecework?}
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
@@
 ``day--labourers''.
-% \ali{I want to come back to this later when we talk about Uber \& Lyft, and
-%      how they increasingly offer to guarantee drivers an hourly rate
-%      \cite[see][]{uberHourly,lyftHourly}.
-%      That's why I'm nitpicking about this muddied system --- to avoid someone saying
-%      ``it doesn't seem like piecework to me anymore if there's an hourly wage floor''}
 
@@
 \citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
-% \msb{inconsistent use of ``piece work'' vs. ``piecework'' in the non-quoted text. Which do you want?}
-% \ali{Okay here's the problem: I'm happy to use ``piecework'' and that's the word that exists in the dictionary, but
-% at the time that \citeauthor{hughRaynbirdTaskWork} was writing, it was ``piece work'' and I don't know if
-% this is a trivial enough thing that I can just quietly concatenate his/their quotations so it looks like they say ``piecework''.
-% I think at this point I was just calling it ``piece work'' because at this point in history people were calling it that.
-% I've just changed everything out of quotation marks to ``piecework[er[s]]'', and
-% left everything in quotation marks as ``piece work''.}
 crop up; he points out that\dots
@@
 
-\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
+\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward
+the question of the features that sprout from piecework.}
 In their reflection on the features endemic to piecework in the 1930s,
@@
 \begin{inlinelist}
-\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
-\item workers with specific slices of skills could be more appropriately matched to suitable tasks
+  \item ``female workers who generally had less training'' had to be trained in
+        narrower subsets of the general body of skills that
+        conventional (male) apprentices would undertake, and
+  \item workers with specific slices of skills could be more appropriately matched to
+        suitable tasks
 \end{inlinelist}
@@
 Workers without conventional training
---- like women, who had no such opportunities to engage in engineering and metalworking apprenticeships as men did ---
+--- like women, who had no such opportunities
+to engage in engineering and metalworking apprenticeships as men did ---
 could be trained very narrowly on a very tightly constrained task,
@@
 \begin{enumerate}
-  \item paid workers for quantity of work done, rather than time done, but occasionally mixed the two payment models;
+  \item paid workers for quantity of work done, rather than time done,
+        but occasionally mixed the two payment models;
   \item afforded workers freedom in when and how much to work; and
-  \item structured tasks such that people who didn't have the training to engage in the traditional labor force could still participate.
+  \item structured tasks such that people who didn't have the training
+        to engage in the traditional labor force could still participate.
 \end{enumerate}
-\ali{woah i thought this was a comment; is it okay as is?}
+% \ali{woah i thought this was a comment; is it okay as is?}
 
-\ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}
+% \ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}
 
-\subsection{The Historical Arc of Piecework}\label{sec:pieceworkArc}
+\subsection{A Primer of Piecework}\label{sec:pieceworkPrimer} % yay rhyming?
+\topic{In this section we will offer a brief overview of the history of piecework;
+this should not be mistaken for a comprehensive or especially rigorous background.}
+Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
+In other words, framing piecework in the contexts of
+the early days of the industrial revolution,
+through the political and economic turmoil of the early and mid--\nth{20} century,
+and into the \nth{21} century.
+While the previous section provided a \textit{definition} of piecework,
+this section attempts to shine a light on the \textit{zeitgeist} of piecework.
+
 \topic{Piecework's history traces back further perhaps than most would expect.}
@@
 \cite{grier2013computers}.
-While this approach didn't become the same economic powerhouse as later examples would,
-Airy \msb{This is the first time Airy comes up. Who are they?} and others arguably found the kernel of insight that we pursue throughout this discussion:
-determining the extent to which work can be decomposed, and
-finding the limits of complexity of that decomposed work.
-That is, Airy found that he could train youths in elementary mathematics
-to complete the majority of the calculations he would otherwise have had to solve on his own,
-and that the greater body of work could ultimately be completed sooner
-if he arranged his work appropriately.
-\msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
-What I got out of it is that a bunch of people did piecework, but
-I don't know why or in fact why these are different than the sources we cite earlier.
-Can you hone it?}
-\ali{I wanted that paragraph to be about the rising popularity and application of piecework
-(especially as it approached its ``heyday'' \cite{hart2013rise}),
-coming from humble beginnings as it found its footing.
-Given that intent\dots\\
-Should I \textbf{refactor} or \textbf{rewrite}?}
+George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
+by breaking complex calculations into constituent parts, and
+training young men to solve simple algebraic problems,
+Airy could distribute work to many more people than could otherwise complete the full calculations.
 
-\ali{\sout{I'm thinking of dropping most of the last paragraph and merging it with the next one. Thoughts?}
-I moved it to \nameref{sec:complexity}}
+% While this approach didn't become the same economic powerhouse as later examples would,
+% Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
+% determining the extent to which work can be decomposed, and
+% finding the limits of complexity of that decomposed work.
+% Airy found that he could train youths in elementary mathematics
+% to complete the majority of the calculations he would otherwise have had to solve on his own,
+% and that the greater body of work could ultimately be completed sooner
+% if he arranged his work appropriately.
+% \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
+% What I got out of it is that a bunch of people did piecework, but
+% I don't know why or in fact why these are different than the sources we cite earlier.
+% Can you hone it?}
+% \ali{I wanted that paragraph to be about the rising popularity and application of piecework
+% (especially as it approached its ``heyday'' \cite{hart2013rise}),
+% coming from humble beginnings as it found its footing.
+% Given that intent\dots\\
+% Should I \textbf{refactor} or \textbf{rewrite}?}
 
-\ali{??? Why is this here? Can it be moved?}
-\topic{As increasing attention revealed problems in piecework as it related to workers,
-workers themselves began to speak out about their frustration with this new regime.}
-It began, arguably, with \citeauthor{riisOtherSideLives}'s photo--documentary work,
-but this led to industry organizations representing
+% \ali{\sout{I'm thinking of
+% dropping most of the last paragraph and merging it with the next one. Thoughts?}
+% I moved it to \nameref{sec:complexity}}
+
+\ali{Where do I want to take this? Where are we now?
+\begin{enumerate}
+  \item Piecework started in the \nth{19} century (above). [DONE]
+  \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
+  \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
+  \item Piecework died pretty quickly (below). [DONE]
+\end{enumerate}
+Sound good?}
+
+\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
+but it found its foothold in manual labor.}
+Piecework took off on in farm work \cite{hughRaynbirdTaskWork},
+in textiles \cite{restructuringPieceworkBaker,riisOtherSideLives},
+on railroads \cite{Brown01041990}, and 
+elsewhere in manufacturing \cite{10.2307/3827491}.
+This proliferation of manual piecework led to discussion surrounding how best to manage these workers
+\cite{norton1900textile,clark1908cotton},
+which seemed to lead to frustration among workers over their poor working and living conditions
+(famously documented by
+% \ali{??? Why is this here? Can it be moved?}
+% {As increasing attention revealed problems in piecework as it related to workers,
+% workers themselves began to speak out about their frustration with this new regime.}
+\citeauthor{riisOtherSideLives})
+\cite{riisOtherSideLives}.
+This led to industry organizations representing
 railway workers, mechanical engineers, and others contributing their myriad perspectives
-\cite{american1921problem,richards1904anything,riisOtherSideLives}.
-Nevertheless, piecework continued to permeate low--skilled labor.
+\cite{american1921problem,richards1904anything}.
+% Nevertheless, piecework continued to permeate low--skilled labor.
 % \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
 
-\ali{???}
+\topic{Fordism and scientific management thrust piecework into higher gear, especially as
+mass manufacturing and
+a depleted wartime workforce forced industry to find new ways to eke out more production capacity.}
+\citeauthor{hart2013rise} point out that the Second World War,
+which called millions of Americans to military service,
+necessitated the rapid training and employment of
+a labor pool that hadn't historically been utilized in industrial labor: women
+\cite{hart2013rise}.
+% With little time and virtually none of the training that was conventionally expected of metalworkers and other industrial work
+% (e.g. years--long apprenticeships),
+% tasks had to be redesigned and workers trained for these dramatically unorthodox task designs
+\ali{I'd like to bring the Rosie the Riveter anecdote in here but it's not feeling natural\dots}
+% \cite{honey1985creating,davies2014origins}.
+
 \topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
@@
 \item the shifting of these industries (manufacturing, clothing, etc\dots)
-to other countries;
-\item the quality of ``multidimensional'' work becoming too difficult to evaluate.
+to other countries; and
+\item the quality of ``multidimensional'' work becoming too difficult to evaluate
 \end{inlinelist}

### 2016-09-21 drafts/labor/complexity.tex

@@
 \subsection[What are the complexity limits of crowd work]{Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
+A key question to the future of on--demand work is \textit{what} precisely will become part of this economy.
+Paid crowdsourcing began with simple microtasks on platforms such as Amazon Mechanical Turk, but microtasks are only helpful if they build up to a larger whole.
+So, our first question: how complex can the work outcomes from crowd work be?
 

### 2016-09-21 drafts/labor/relationships.tex

@@
 \subsubsection{\pieceworkpers}
+\citeauthor{clark1908cotton}, for instance,
+relays his observations of textile mill pieceworkers and his realization that
+``When he works by the day the Italian operative wishes to leave before the whistle blows,
+but if he works by the piece he will work as many hours as it is possible for him to stand''
+
+\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
+and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
+offering guidance on how best to manage pieceworkers
+\cite{riisOtherSideLives,norton1900textile}.
+
 \topic{Despite the intense growth of the piecework approach to renumeration,

### 2016-09-21 drafts/labor/complexity.tex
looked through complexity

@@
 
-\subsection[What are the complexity limits of crowd work]{Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
-A key question to the future of on--demand work is \textit{what} precisely will become part of this economy.
-Paid crowdsourcing began with simple microtasks on platforms such as Amazon Mechanical Turk, but microtasks are only helpful if they build up to a larger whole.
-So, our first question: how complex can the work outcomes from crowd work be?
+\subsection[What are the complexity limits of crowd work]
+{Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
+A key question to the future of on--demand work
+\ali{is this an old remnant of me being indecisive about terms?} is
+\textit{what} precisely will become part of this economy.
+Paid crowdsourcing began with simple microtasks on platforms such as
+Amazon Mechanical Turk, but
+microtasks are only helpful if they build up to a larger whole.
+So, our first question:
+how complex can the work outcomes from crowd work be?
 
@@
 
-\topic{Crowdsourcing research has spent the better part of a decade proving the viability of crowdsourcing in complex work.}  % this is my topic sentence.
-Unless crowdsourcing can demonstrate viability for meaningfully complex tasks, the argument runs, it will be incapable of ensuring a pro-social outcome for work and workers~\cite{crowdworkFuture}.
-\citeauthor{crowdForgeKittur} first opened the question of whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
-Their work demonstrated proof-of-concept crowdsourcing of a simple encyclopedia article and news summary.
-Seeking to raise the complexity ceiling~\cite{myers2000past}, researchers have since created additional proof-of-concept applications and techniques, including conversational assistants~\cite{Lasecki:2013:CCC:2501988.2502057}, medical data interpreters~\cite{Lasecki:2013:CCC:2501988.2502057}, and idea generation~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}, to name a few examples.
+\topic{Crowdsourcing research has spent the better part of a decade
+proving the viability of crowdsourcing in complex work.}
+Unless crowdsourcing can demonstrate viability for meaningfully complex tasks,
+the argument runs,
+it will be incapable of ensuring a pro--social outcome for work and workers
+\cite{crowdworkFuture}.
+\citeauthor{crowdForgeKittur} first opened the question of
+whether crowdsourcing could be used for goals that are not simple parallel tasks
+\cite{crowdForgeKittur}.
+Their work demonstrated proof--of--concept crowdsourcing of
+a simple encyclopedia article and news summary
+--- tasks which could be verified or repeated
+with reasonable expectations of similar outcomes.
+Seeking to raise the complexity ceiling
+\cite{myers2000past},
+researchers have since created
+additional proof--of--concept applications and techniques,
+including conversational assistants \cite{Lasecki:2013:CCC:2501988.2502057},
+medical data interpreters \cite{Lasecki:2013:CCC:2501988.2502057}, and
+idea generation \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+to name a few examples.
 
-\topic{This body of research has often invoked insights from Computer Science and applied them to human workflows.}
-The crowd work literature typically identifies target milestones that have presented significant challenges for researchers, leverages an insight from Computer Science (for example, MapReduce~\cite{crowdForgeKittur} or sequence alignment algorithms~\cite{lasecki2012real}), and arranges humans as computational black boxes within those approaches.
+\topic{This body of research has often invoked insights from
+Computer Science and applied them to human workflows.}
+The crowd work literature typically identifies target milestones that have
+presented significant challenges for researchers,
+leverages an insight from Computer Science
+(for example, MapReduce
+\cite{crowdForgeKittur} or
+sequence alignment algorithms
+\cite{lasecki2012real}),
+and arranges humans as computational black boxes within those approaches.
+\ali{We could quote \citeauthor{storiesIraniSilberman} and say ``human APIs'',
+but that gets a little critique--y, and we might want to save that for later.}
 This approach has proven a compelling one because
-it leverages the in--built advantages of scale, automation, and programmability that software affords.
-%For example, \textit{Foundry}'s tools for managing and arranging expert groups into a cohort allow researchers to convincingly argue that expert teams can be rapidly formed, just like non--expert teams \cite{foundry}.
+it leverages the in--built advantages of
+scale,
+automation, and
+programmability that software affords.
 
-It is now clear that this approach works with focused complex tasks, but the broader wicked problems~\cite{rittel1973dilemmas} remain unsolved. 
-As a first example, idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}, but there is as yet no general crowdsourced solution for the broader goal of invention and innovation~\cite{fuge2014analysis}.
-Second, focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,writingMicroTasks,agapie2015crowdsourcing}, but there is no general solution to create a high-quality crowd-powered author across domains.
-Third, data analysis tasks such as clustering~\cite{chilton2013cascade}, categorization~\cite{andre2014crowd}, and outlining~\cite{luther2015crowdlines} are possible, but there is no general solution for sensemaking.
-It is not yet clear what insights would be required to enable crowdsourced solutions for these broader wicked problems and other crowdsourced goals such as addressing climate change~\cite{introne2011climate}.
+It is now clear that this approach works with focused complex tasks, but
+the broader wicked problems \citeauthor{rittel1973dilemmas} brings up remain unsolved
+\cite{rittel1973dilemmas}.
+As a first example,
+idea generation shows promise
+\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+but there is as yet no general crowdsourced solution for
+the broader goal of invention and innovation
+\cite{fuge2014analysis}.
+Second,
+focused writing tasks are now feasible
+\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
+      writingMicroTasks,agapie2015crowdsourcing}, but
+there is no general solution to create
+a high--quality crowd--powered author across domains. 
+Third,
+data analysis tasks such as
+clustering \cite{chilton2013cascade},
+categorization \cite{andre2014crowd}, and
+outlining \cite{luther2015crowdlines}
+are possible, but there is no general solution for sensemaking.
+It is not yet clear what insights would be required
+to enable crowdsourced solutions for these broader wicked problems and
+other crowdsourced goals such as addressing climate change
+\cite{introne2011climate}.
 
-Restricting attention to non-expert microtask workers limited the ambitions of research in the area. 
-So, \citeauthor{foundry} introduced the idea of crowdsourcing with online paid \textit{experts} from platforms such as Upwork.
-Expert crowdsourcing enables access to a much broader set of workers, for example designers and programmers.
-Researchers then apply similar Computer Science techniques but to expert ``macro-tasks''~\cite{cheng2015break,haas2015argonaut}.
-This approach has led to successful crowdsourcing of goals such as the user-centered design process~\cite{foundry}, programming tasks~\cite{latoza2014microtask,Fast2016,Chen2016}, and mentorship~\cite{suzukiAtelier}.
-However, there remains the open question of how complex the work outcomes from these approaches will be.
+Restricting attention to non--expert microtask workers limited
+the ambitions of research in the area.
+So, \citeauthor{foundry} introduced the idea of crowdsourcing with
+online paid \textit{experts} from platforms such as Upwork.
+Expert crowdsourcing enables access to a much broader set of workers,
+for example designers and programmers.
+Researchers then apply similar Computer Science techniques but
+to expert ``macro-tasks''
+\cite{cheng2015break,haas2015argonaut}.
+This approach has led to successful crowdsourcing of goals such as
+the user--centered design process \cite{foundry},
+programming tasks \cite{latoza2014microtask,Fast2016,Chen2016}, and
+mentorship \cite{suzukiAtelier}.
+However, there remains the open question of
+how complex the work outcomes from these approaches will be.
 
@@
 
-\topic{\citeauthor{grier2013computers} gives early accounts of a piecework strategy in Airy's creation of the British Nautical Almanac~\cite{grier2013computers}.} 
-Airy's goal was complex --- mathematical calculations to produce tables that would allow sailors to locate themselves by starlight from sea. 
-Many of his contributors did not have high-level mathematical training, so Airy broke down the task into simpler calculations and distributed them by mail, accomplishing the complex goal through piecework tasks that paid little.
+\topic{\citeauthor{grier2013computers} gives early accounts of
+a piecework strategy in Airy's creation of
+the British Nautical Almanac
+\cite{grier2013computers}.}
+Airy's goal was complex
+--- mathematical calculations to produce tables that would
+allow sailors to locate themselves by starlight from sea.
+Many of his contributors did not have high--level mathematical training,
+so Airy broke down the task into simpler calculations and
+distributed them by mail,
+accomplishing the complex goal through piecework tasks that paid little.
 
-However, when piecework entered the American economy, it was not used for complex work.
+However, when piecework entered the American economy,
+it was not used for complex work.
 Initially for farm work, as
@@
 the practice remained relatively obscure until
-%it was brought to the textile industry
-%\cite{hughRaynbirdTaskWork}.
-%\topic{Piecework then began to grow: first applied to farm work, as
-%\citeauthor{hughRaynbirdTaskWork} and others illustrate,
-%the practice remained relatively obscure until
-it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
-Complexity levels remained low at the turn of the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
-%was documenting abhorrent working \& living conditions of pieceworkers in New York City,
-However, writers of the time focused their attentive on wage~\cite{burton1899commercial} and management regimes \cite{norton1900textile}.
-%\citeauthor{clark1908cotton}, for instance,
-%relays his observations of textile mill pieceworkers and his realization that
-%``When he works by the day the Italian operative wishes to leave before the whistle blows,
-%but if he works by the piece he will work as many hours as it is possible for him to stand''
-%\cite{clark1908cotton}.
-%During this period, best practices regarding the measurement and management of
-%piecework rates, and of workers in the engineering industry,
-%were beginning to take shape
-%\cite{burton1899commercial}.
-% \ali{Everything up till now has been pointing at complexity (right? if not, then~\dots~lol)}
-The work remained low-skilled in part because it was infeasible to provide new pieceworkers with the comprehensive education that was familiar through apprenticeships \cite{hart2013rise}.
+it blossomed in the textile industry
+\cite{hughRaynbirdTaskWork}.
+Complexity levels remained low at the turn of
+the \nth{20} century as piecework saturated New York City
+\cite{riisOtherSideLives}.
+However, writers of the time focused their attention on
+wage \cite{burton1899commercial} and
+management regimes \cite{norton1900textile}.
+\ali{\citeauthor{hart2013rise} would later argue of this time that\dots?
+I'm beginning to notice that I'm really gravitating toward naming authors,
+which is totally weird when I read through conventional HCI papers.}
+The work remained low--skilled in part because
+it was infeasible to provide new pieceworkers with the comprehensive education
+that apprenticeships imparted \cite{hart2013rise}.
+\ali{I feel like this sort of jumps ahead without mentioning it
+(in my head it's a jump from
+\citeauthor{norton1900textile}'s \citeyear{norton1900textile} to
+\citeauthor{hart2013rise}'s ~1940s), and
+I felt like this was a phenomenon that
+I'd like to shine in a positive(--ish?) light.}
 
-Measurement limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework. 
-When \citeauthor{Brown01041990} investigated what limited the adoption of piecework in industries that otherwise gravitated toward it (e.g., railway engineers), the homogeneity of tasks %and  and the costs associated with adopting a piecework model
-arose as a major contributing factor~\cite{Brown01041990}.
-% \ali{that last paragraph is a mix of complexity and decomposition I think}
-%\topic{Piecework researchers have found themselves trying to understand what characteristics limit piecework,
-%or more precisely what has prevented piecework from becoming more prevalent.}
-%The research into piecework makes the case that piecework has been limited principally by the challenges of human management and oversight.
-\citeauthor{10.2307/23702539} concurs via a case study in Santa Fe Railway, which used ``efficiency experts'' to develop a ``standard time'' to determine pay for each task at the company informed by ``thousands of individual operations at the Topeka shops''~\cite{10.2307/23702539}.
-%The cost of measuring workers in such excruciating detail at the turn of the \nth{20} century was undoubtedly immense,
-%but this ``standard time'' value, which determined the pay that workers would earn for each task they do,
-%was the only viable approach at the time to determine appropriate pay given the task
-%\cite{10.2307/23702539}.
-%But the repeated measurement of workers' time to complete tasks had shortcomings;
-%for one thing, pay rates for rarer tasks
-%were necessarily less certain than for more common tasks,
-%which had the simple benefit of a larger sample size.
-%\ali{Do I need ``for another thing\dots''? This paragraph is kind of huge already. Not sure how to break it up/down as it is.}
+Measurement limited the complexity of piecework:
+only tasks that could be measured and priced could be completed via piecework. 
+When \citeauthor{Brown01041990} investigated
+what limited the adoption of piecework in industries that otherwise gravitated toward it
+(e.g., railway engineers),
+the homogeneity of tasks arose as a major contributing factor
+\cite{Brown01041990}.
+\citeauthor{10.2307/23702539} concurs via a case study in Santa Fe Railway,
+which used ``efficiency experts'' to develop a ``standard time''
+to determine pay for each task at the company informed by
+``thousands of individual operations'' % removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
+\cite{10.2307/23702539}.
 One might conclude from \citeauthor{10.2307/23702539}'s observations that
-complex, creative work --- which is inherently heterogeneous and difficult to routinize ---
+complex, creative work
+--- which is inherently heterogeneous and difficult to routinize ---
 would be unsuitable for piecework.
@@
 \topic{Piecework was also limited to tasks that could be clearly evaluated.}
-The roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''
+The roles required to facilitate piecework in the early \nth{20} century included
+``piecework clerks, inspectors, and `experts'''
 \cite{10.2307/23702539}.
-%\citeauthor{10.2307/23702539} further recognizes that
-%it's necessary for a successful piecework shop to employ
-%clerks,
-%inspectors, and
-%other experts to properly design and evaluate complex work.
-\citeauthor{hart2016rise} argues for an ultimate complexity limit: at some point, evaluating multidimensional work for quality (rather than for quantity) becomes infeasible.
+\citeauthor{hart2016rise} argues for an ultimate complexity limit:
+at some point, evaluating multidimensional work for quality
+(rather than for quantity) becomes infeasible.
 In his words,
 ``if the quality of the output is more difficult to measure than the quantity [\ldots]
-%perhaps because of `difficult--to--observe' production techniques,
 then a piecework system is likely to encourage
@@
 \cite{hart2016rise}.
-%This, \citeauthor{hart2016rise} argues, may have fundamentally hamstrung piecework,
-%and ultimately precipitated its downfall,
-%especially with the increasing complexity of manufacturing work
-%over the course of the \nth{20} century.
-%\ali{just to be clear, this whole point is queuing up so that later I can be like
-%``Oh hey computers are terrific at evaluating some stuff really quickly.
-%And researchers in crowd work have done some work on workers evaluating other workers
-%(e.g. Find--Fix--Verify, arguably \textit{PeerStudio} does something along this line)
-%\cite{bernsteinSoylent,Kulkarni:2015:PRP:2724660.2724670}''}
 Complex work, which is often subjective to evaluate, falls victim to this criteria.
@@
 If true, 
-%This argument is certainly compelling;
-%it would seem to make 
 the concurrent upswing of
@@
 Researchers detail three organizational criteria.
-% \msb{I think a reader wouldn't understand why that claim is relevant.
-% Let's invert it, something like how
-% the researchers point out that
-% only certain kinds of organizations can effectively make use of piecework.}
-First, \citeauthor{Brown01041990} argues that piecework ``is less likely in jobs with
-a variety of duties than in jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
+First, \citeauthor{Brown01041990} argues that
+piecework ``is less likely in jobs with a variety of duties than in
+jobs with a narrow set of routinized duties''
+\cite{Brown01041990}.
 \citeauthor{SJOE:SJOE371} points out the phenomenon here as a market effect:
-``in an environment with multi--tasking, pay schemes based on tightly specified performance may \textit{induce workers to neglect tasks} that are less easy to measure''
+``in an environment with multi--tasking,
+pay schemes based on tightly specified performance may
+\textit{induce workers to neglect tasks} that are less easy to measure''
 (emphasis added) \cite{SJOE:SJOE371}.
-% \msb{that's confusing, is incentive pay the same as piecework?
-% can you cut the first part of the quote and
-% give us the noun instead of forcing us to figure it out?}.
-Second, complexity was limited by access to capital to create the necessary infrastructure.
-As \citeauthor{10.2307/23702539} reports, %that successful cases of piecework owed themselves in part to the fact that
-only the largest and most wealthy railroads had the resources to ``pay the overhead involved in installing work reorganization'' \cite{10.2307/23702539}.
-% Together, \citeauthor{10.2307/23702539} and \citeauthor{Brown01041990}
-% make a persuasive argument that piecework is limited in complexity by
-% the capacity to endure managerial overhead while transitioning to a new system.
-% \msb{wait, wasn't that in a previous paragraph?
-% that should be joined with the managerial overhead text above}
-% \ali{I'm trying to make this argument that \cite{10.2307/23702539} is making two claims;
-% the first is that there are limits on doing hard, ``difficult--to--observe'' work,
-% and the second is that the cost propagates up to management,
-% putting limits on who can run piecework systems.
-% So\dots\\
-% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Rewrite\\
-% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Refactor\\
-% \makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{1em}} Drop
-% Maybe I should just merge it with the next paragraph?}
+Second,
+complexity was limited by access to capital to create the necessary infrastructure.
+As \citeauthor{10.2307/23702539} reports,
+only the largest and most wealthy railroads had the resources
+to ``pay the overhead involved in installing work reorganization''
+\cite{10.2307/23702539}.
 Third, organizations required capable managers in charge of the pieceworkers.
-% \topic{There are other characteristics to effective complex piecework institutions, such as
-% appropriately designed management practices.}
-% \citeauthor{10.2307/2118435} describe the role of the foreman in West Virginia coal mines under the piecework model:
-% ``The foreman had the power to hire and fire workers and allocate workplaces,
-% but then left the face--worker largely free to his own efforts so that
-% often he went all day without seeing the foreman''
-% \cite{10.2307/2118435}.
 The West Virginia mines, for example, hired foremen 
-% The general approach adopted by these West Virginia mines was,
-% as in other factories with active foremen,
-to be the intermediary between upper management and the worker~\cite{10.2307/2118435}.
+to be the intermediary between upper management and the worker
+\cite{10.2307/2118435}.
 Specifically, foremen were responsible for allocating resources and
@@
 \cite{wray1949marginal}.
-% \msb{Currently unclear what I'm supposed to take out of this. Summarize for me. What does this say about the complexity limits of piecework?}
-% The management of pieceworkers %, \citeauthor{10.2307/2118435} and later \citeauthor{wray1949marginal} argue,
-% demanded people in positions akin to foremen
-% --- intermediate managers people who were
-% \begin{inlinelist}
-%   \item familiar with and even sympathetic to the needs of workers,
-%   \item empowered by higher level management to make decisions, and
-%   \item relaxed enough in day--to--day work to allow workers to go about their work
-% \end{inlinelist}
-% \cite{wray1949marginal,10.2307/2118435}.
-So, in sum, organizations historically could only take advantage of piecework if they had homogenous work to be done, access to capital to purchase the necessary equipment, and the ability to hire intermediate management.
+So, in sum,
+organizations historically could only take advantage of piecework if they had
+homogeneous work to be done,
+access to capital to purchase the necessary equipment, and
+the ability to hire intermediate management.
+\ali{will people understand the nuance here of ``intermediate'' meaning that
+the management act as \textit{intermediaries}?
+I know that sounds stupid, but
+I feel like the wording has a connotation of just being ``middle management'',
+which doesn't communicate that these are bidirectional negotiators.}
 
@@
 
-%\topic{This isn't to say that complex work is outside of the realm of piecework; indeed, we've discussed complex applications of crowd work already.}
-%While \citeauthor{hart2013rise} described a flourishing of ingenious piecework design, much of it arose out of necessity --- it was infeasible to provide new workers with the comprehensive education that was familiar to men through apprenticeships \cite{hart2013rise}.
-%While this constraint led to much more tightly scoped work who now had to specialize in extremely narrowly defined roles.
-%The same could be said of Airy and his \textit{computers} --- young boys whose preparations consisted principally of a relatively specific mathematics curriculum \cite{grier2013computers}.
-% Instead, we argue that the literature suggests that piecework is tightly limited only when the application of piecework follows a direct, perhaps even unimaginative, mapping from an time--based regime to an output--based one.
-% When the work is redesigned from the ground up --- as we see with mathematicians in the \nth{19} century and with the metalworking industry during the Second World War --- it seems that we don't yet know the limits of complexity with regard to piecework.
-%\ali{This is \textit{almost} making a prediction, but I wanted to take something punchy away from the stuff I cite here.}
-
 \subsubsection{\whatchanged}
-\msb{Before you get into this, summarize what I am supposed to have learned from the prior section on piecework.
-You're about to draw on those points to make your argument, so they need to be at the top of my mind here.}
 \topic{Piecework makes a number of observations leading to the conclusion that
@@
 --- a human resource, which made the scaling cost prohibitively expensive for many
-% in their times (\citeyear{10.2307/2118435} and \citeyear{wray1949marginal}, respectively)
 \cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
-\ali{Something like this? Should I dig deeper?
-I can line these points up in a way to make the next paragraph sort of obvious or inevitably,
-depending on how actively engaged someone is while reading this\dots}
 
@@
 \cite{Brown01041990}.
-% \msb{I don't recall this point in the prior section:
-% I remember management overhead and fixed costs of materials,
-% I don't remember a paragraph about complexity.
-% If we want to draw on that point, make it in the earlier section}
 Given the flourishing of on--demand labor platforms such as
@@
 the Internet has trivialized
-the costs and challenges of the earlier limiting factors %  for two reasons:
-% \begin{inlinelist}
-  % \item 
+the costs and challenges of the earlier limiting factors
 because technology make it easier
@@
 
-%   \item The Internet allows us to leverage the benefits of
-%         ``economies of scale'' at very little cost
-%         to the system--designer \cite{lessig2006code,miller2011understanding}.% \msb{I don't understand that argument. What is scaling here? The people? The tools for complex work?}
-% \end{inlinelist}
-
+\ali{maybe this should go away?
 \topic{Technology has made it possible
@@
 within the domain of experts.}
-\msb{I don't yet buy the following argument.
-If the point is that technology makes us more expert, I disagree that the CrowdCrit/Voyant systems are using technology to do this.
-They are building the smarts into their OWN workflows, rather than giving workers EXTERNAL tools that make them smarter.
-Giving workers a calculator is an external tool; the mathematical tables project already demonstrated that you can build smarts into the workflow if you don't have one.}
 \citeauthor{yuanAlmost} builds on the work of others
@@
 today algorithms manage workers as pieceworkers once did
-\cite{uberAlgorithm,10.2307/23702539}.
+\cite{uberAlgorithm,10.2307/23702539}.}
 
-\topic{Furthermore, technology more directly facilitates the subversion of expertise requirements
+\topic{Technology facilitates the subversion of expertise requirements
 by giving non--experts access to information that would otherwise be unavailable.}
@@
 ranging topics such as police activity, congestion, construction, etc\dots
-\cite{silva2013traffic,hind2014outsmarting}. \msb{what's the insight I should take away from this paragraph? what does this say about crowd work?}
+\cite{silva2013traffic,hind2014outsmarting}.
+\msb{what's the insight I should take away from this paragraph? what does this say about crowd work?}
+\ali{Maybe I'm burnt out but I feel like the topic sentence is the takeaway and
+I'm not entirely sure what else to put here.
+Maybe that digital media can make it possible for non--experts to do increasingly expert work,
+informed by ``insider'' or ``expert'' knowledge of work, the way Uber, Waze, etc\dots do?}
 
 \subsubsection{\implication}
-\msb{This paragraph can be expanded to make a more concrete argument. What will be possible? What won't?}
-\topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
-it also signals some of the ultimate limitations of crowd work and piecework in general.}
-While the threshold preventing task requesters from utilizing piecework
-has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
-\msb{is that your prediction?
-I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
-If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
-we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
+% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
+% it also signals some of the ultimate limitations of crowd work and piecework in general.}
+% While the threshold preventing task requesters from utilizing piecework
+% has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
+% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
+% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
+% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
 
-\msb{this doesn't seem like a concrete prediction. what would piecework say will happen if we didn't resolve the tension?}
-We should also consider exploring the limitations that algorithmic management bring along more carefully.
-While research has touched on this subject, we've yet to make out the bigger picture of this theme
+We should consider exploring the limitations that algorithmic management bring along more carefully.
+While crowd work research has touched on this subject, we've yet to make out the bigger picture of this theme
 \cite{uberAlgorithm}.
@@
 \cite{takingAHITMcInnis}.
+\ali{What will happen if we don't do this?
+Maybe this is where I should bring up
+the ``market for lemons'' metaphor and say that
+the market will just continue to decline in quality until
+nobody is willing to rely on it either as a source of data/information (requesters), or
+a means of sustaining oneself financially (workers).}
 
@@
 as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-\cite{hart2013rise,grier2013computers} \msb{remind us of what those were}.
-That is, we need to identify new ways to train crowdworkers for uniquely narrowly defined work. \msb{I don't understand: why?}
+of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
+\ali{If we can find better ways to train crowd workers for narrowly defined work,
+we can envision more complex applications and outcomes of crowd work,
+as in the case of metalwork breaking away from the historical requirements of lengthy,
+comprehensive apprenticeships.}
 To some extent, an argument can be made that
@@
 ``brash, irregular, immoral, and uneducated''
-\cite{10.2307/3827491}. \msb{how will that reduce any of the problems except the last one?}
+\cite{10.2307/3827491}.
+\msb{how will that reduce any of the problems except the last one?}
+\ali{I was thinking of fleshing out that the antipathy in general
+(``brash, immoral, etc\dots'') stems from these workers not being high--skill;
+we see similar descriptions used of Turkers when people talk about
+malicious and
+low--skill Turkers and getting them to ``expert--level work''.
+the historical counterexample is that
+metalworking pieceworkers didn't have these reputations,
+and we can suppose that this is because they \textit{were} trained
+(albeit unconventionally), and therefore considered educated, regular, etc\dots
+Too much of a stretch?}
 

### 2016-09-21 drafts/labor/trackingWork.tex
incremental changes on complexity - still more to fix
MSB intro edits
pushing whatever i have...
remove flexibility. no flexibility!
moved a bunch of piecework history to its constituent sections. exciting!
added a ref shortname
piecework history of complexity revised
looked through complexity

@@
 \setlist*[inlinelist,1]{%
+   % label=\alph*),
  label=\arabic*),
@@
 \setlist{nolistsep}
-\def\plaintitle{Examining Crowd Work Through The Lens of Piecework}
+\def\plaintitle{Examining Crowd Work Through The Historical Lens of Piecework}
 \def\plainauthor{Ali Alkhatib, Margaret Levi, Michael Scott Bernstein}
 \def\emptyauthor{}
-\def\plainkeywords{Crowdsourcing; human computation; research vision}
+\def\plainkeywords{Crowdsourcing; on--demand labor; gig work}
 \def\plaingeneralterms{Documentation, Standardization}
@@
 \newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}
+\newcommand{\TO}{\textit{Turkopticon}}
+\newcommand{\DO}{\textit{Dynamo}}
 
@@
 \newcommand{\subsubsubsection}[1]{\textit{#1}}
-\newcommand{\pieceworkpers}{Piecework's perspective.}
-\newcommand{\crowdworkpers}{Crowd work's perspective.}
-\newcommand{\whatchanged}{What's different about crowd work.}
-\newcommand{\implication}{Implications for crowd work research.}
+\newcommand{\pieceworkpers}{Piecework's perspective}
+\newcommand{\crowdworkpers}{Crowd work's perspective}
+\newcommand{\whatchanged}{Comparing the phenomena}
+\newcommand{\implication}{Implications for crowd work}
 \newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of
-``payment for results''%
-, leading to and leveraging narrower skill sets required for narrowly defined tasks,
+``payment for results'',
+leading to and leveraging narrower skill sets required for narrowly defined tasks,
 and affording workers some amount of freedom to complete tasks
 (and consequently earn money)
-at at whatever rate and in whichever manner they wish}
+at at whatever rate and in whichever manner they wish} % do i use this more than once? even that much?
 
@@
 
-\category{H.5.m.}{Information Interfaces and Presentation
-  (e.g. HCI)}{Miscellaneous} \category{See
-  \url{http://acm.org/about/class/1998/} for the full list of ACM
-  classifiers. This section is required.}{}{}
+\category{H.5.3.}{Information Interfaces and Presentation
+  (e.g. HCI)}{Group and Organization Interfaces}
 

### 2016-09-21 drafts/labor/complexity.tex
partway through implications

@@
 {Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
-A key question to the future of on--demand work
-\ali{is this an old remnant of me being indecisive about terms?} is
+A key question to the future of crowd work is
 \textit{what} precisely will become part of this economy.
@@
 
-\topic{This body of research has often invoked insights from
-Computer Science and applied them to human workflows.}
-The crowd work literature typically identifies target milestones that have
-presented significant challenges for researchers,
-leverages an insight from Computer Science
+\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing workflows.}
+Beginning with a goal that has
+presented significant challenges for computers,
+the researcher leverages an insight from Computer Science
 (for example, MapReduce
@@
 sequence alignment algorithms
-\cite{lasecki2012real}),
+\cite{lasecki2012real})
 and arranges humans as computational black boxes within those approaches.
-\ali{We could quote \citeauthor{storiesIraniSilberman} and say ``human APIs'',
-but that gets a little critique--y, and we might want to save that for later.}
+% \ali{We could quote \citeauthor{storiesIraniSilberman} and say ``human APIs'',
+% but that gets a little critique--y, and we might want to save that for later.}
 This approach has proven a compelling one because
@@
 It is now clear that this approach works with focused complex tasks, but
-the broader wicked problems \citeauthor{rittel1973dilemmas} brings up remain unsolved
+the broader wicked problem that each complex goal represents remains unsolved
 \cite{rittel1973dilemmas}.
@@
 there is no general solution to create
-a high--quality crowd--powered author across domains. 
+a cross-domain, high--quality crowd--powered author. 
 Third,
@@
 It is not yet clear what insights would be required
-to enable crowdsourced solutions for these broader wicked problems and
-other crowdsourced goals such as addressing climate change
-\cite{introne2011climate}.
+to enable crowdsourced solutions for these broader wicked problems.
+%  and
+% other crowdsourced goals such as addressing climate change
+% \cite{introne2011climate}.
 
-Restricting attention to non--expert microtask workers limited
-the ambitions of research in the area.
+Restricting attention to non--expert, microtask workers proved limiting.
 So, \citeauthor{foundry} introduced the idea of crowdsourcing with
@@
 for example designers and programmers.
-Researchers then apply similar Computer Science techniques but
-to expert ``macro-tasks''
-\cite{cheng2015break,haas2015argonaut}.
-This approach has led to successful crowdsourcing of goals such as
-the user--centered design process \cite{foundry},
-programming tasks \cite{latoza2014microtask,Fast2016,Chen2016}, and
+The same ideas can then be applied to expert ``macro-tasks''
+\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
+programming \cite{latoza2014microtask,Fast2016,Chen2016}, and
 mentorship \cite{suzukiAtelier}.
 However, there remains the open question of
-how complex the work outcomes from these approaches will be.
+how complex the work outcomes from expert crowds can be.
 
@@
 it was not used for complex work.
-Initially for farm work, as
+One reason for low complexity was workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
+that apprenticeships imparted \cite{hart2013rise}.
+So, initially piecework arose for farm work, and as
 \citeauthor{hughRaynbirdTaskWork} and others discuss,
@@
 wage \cite{burton1899commercial} and
-management regimes \cite{norton1900textile}.
-\ali{\citeauthor{hart2013rise} would later argue of this time that\dots?
-I'm beginning to notice that I'm really gravitating toward naming authors,
-which is totally weird when I read through conventional HCI papers.}
-The work remained low--skilled in part because
-it was infeasible to provide new pieceworkers with the comprehensive education
-that apprenticeships imparted \cite{hart2013rise}.
-\ali{I feel like this sort of jumps ahead without mentioning it
-(in my head it's a jump from
-\citeauthor{norton1900textile}'s \citeyear{norton1900textile} to
-\citeauthor{hart2013rise}'s ~1940s), and
-I felt like this was a phenomenon that
-I'd like to shine in a positive(--ish?) light.}
+management regimes \cite{norton1900textile}
+rather than training.
 
-Measurement limited the complexity of piecework:
+Measurement also limited the complexity of piecework:
 only tasks that could be measured and priced could be completed via piecework. 
@@
 (e.g., railway engineers),
-the homogeneity of tasks arose as a major contributing factor
-\cite{Brown01041990}.
-\citeauthor{10.2307/23702539} concurs via a case study in Santa Fe Railway,
+the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
+\citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
 which used ``efficiency experts'' to develop a ``standard time''
 to determine pay for each task at the company informed by
-``thousands of individual operations'' % removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
-\cite{10.2307/23702539}.
+``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
+~\cite{10.2307/23702539}.
 One might conclude from \citeauthor{10.2307/23702539}'s observations that
@@
 
-\topic{Piecework was also limited to tasks that could be clearly evaluated.}
-The roles required to facilitate piecework in the early \nth{20} century included
-``piecework clerks, inspectors, and `experts'''
-\cite{10.2307/23702539}.
-\citeauthor{hart2016rise} argues for an ultimate complexity limit:
+\topic{Piecework was limited to tasks that could be clearly evaluated.}
+For example, the roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
+\citeauthor{hart2016rise} argues that evaluation causes an ultimate complexity limit:
 at some point, evaluating multidimensional work for quality
@@
 \topic{This focus on measurement and tracking had consequences.}
-\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
+\citeauthor{10.2307/23702539} suggests that the first sparks of scientific management
 could be found in piecework:
@@
 in addition to constraints on the kind of \textit{work} that's amenable to piecework,
-only certain kinds of \textit{organizations} are amenable to piecework.}
+only certain kinds of \textit{organizations} were amenable to piecework.}
 Researchers detail three organizational criteria.
@@
 pay schemes based on tightly specified performance may
-\textit{induce workers to neglect tasks} that are less easy to measure''
-(emphasis added) \cite{SJOE:SJOE371}.
+induce workers to neglect tasks that are less easy to measure''
+\cite{SJOE:SJOE371}.
 Second,
@@
 only the largest and most wealthy railroads had the resources
-to ``pay the overhead involved in installing work reorganization''
-\cite{10.2307/23702539}.
+necessary
+% ``pay the overhead involved in installing work reorganization''
+~\cite{10.2307/23702539}.
 Third, organizations required capable managers in charge of the pieceworkers.
 The West Virginia mines, for example, hired foremen 
-to be the intermediary between upper management and the worker
-\cite{10.2307/2118435}.
-Specifically, foremen were responsible for allocating resources and
+to be the intermediary between upper management and the workers~\cite{10.2307/2118435}.
+These foremen were responsible for allocating resources and
 understanding when and how to modify work as necessary
@@
 access to capital to purchase the necessary equipment, and
-the ability to hire intermediate management.
-\ali{will people understand the nuance here of ``intermediate'' meaning that
-the management act as \textit{intermediaries}?
-I know that sounds stupid, but
-I feel like the wording has a connotation of just being ``middle management'',
-which doesn't communicate that these are bidirectional negotiators.}
+the ability to hire people who could serve as intermediaries between pieceworkers and management.
 
 \topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
-particularly because maximizing on the advantages of piecework seemed to reward
-smaller, more constrained, more narrowly trained tasks.}
+particularly because maximizing the advantages of piecework seemed to reward
+smaller, more constrained, more narrowly-trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
 For most of the \nth{19} century,
@@
 \subsubsection{\whatchanged}
-\topic{Piecework makes a number of observations leading to the conclusion that
-piecework's complexity is fundamentally bounded by several limitations,
-chief among them the costs of managerial overhead and the transition thereto.}
-\citeauthor{Brown01041990} and \citeauthor{10.2307/23702539}'s claims that
-organizations can't adopt piecework unless
-they're sufficiently large to absorb the cost of transitioning to a piecework system;
-\citeauthor{10.2307/2118435} and \citeauthor{wray1949marginal}'s observations for
-the importance of competent, effective managerial oversight
---- a human resource, which made the scaling cost prohibitively expensive for many
-\cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
+\begin{comment}
+  mangeerial overhead limits, so what's different
+  more people can now do complex work without training (more complex)
+  parts of management can be automated (more firms)
+  cheaper to create the infrastructure (more complex)
+\end{comment}
+% \topic{Piecework makes a number of observations leading to the conclusion that
+% piecework's complexity is fundamentally bounded by several limitations,
+% chief among them the costs of managerial overhead and the transition thereto.}
+% \citeauthor{Brown01041990} and \citeauthor{10.2307/23702539}'s claims that
+% organizations can't adopt piecework unless
+% they're sufficiently large to absorb the cost of transitioning to a piecework system;
+% \citeauthor{10.2307/2118435} and \citeauthor{wray1949marginal}'s observations for
+% the importance of competent, effective managerial oversight
+% --- a human resource, which made the scaling cost prohibitively expensive for many
+% \cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
 
-\topic{Digital media have expanded the scope of
-viable piecework by pushing drastically on the limits
-cited by piecework researchers.}
+% \topic{Digital media have expanded the scope of
+% viable piecework by pushing drastically on the limits
+% cited by piecework researchers.}
 The research on piecework tells us that
@@
 the costs and challenges of the earlier limiting factors
-because technology make it easier
+because technology makes it easier
 \begin{inlinelist}
-\item to do complex work aided by computers and
-\item to evaluate and manage workers as they do increasingly complex work,
-      even observing their work to an otherwise unprecedented granularity.
+\item for workers to do complex work without training,
+\item to manage workers in doing complex work, and 
+\item to create the infrastructure necessary to manage the workers.
 \end{inlinelist}
 
-\ali{maybe this should go away?
-\topic{Technology has made it possible
-for non--experts to do work that was once considered
-within the domain of experts.}
-\citeauthor{yuanAlmost} builds on the work of others
-(\textit{Voyant} and, more relevantly, \textit{CrowdCrit})
-to design workflows that yield ``expert--level feedback''
-\cite{yuanAlmost,Xu:2014:VGS:2531602.2531604,Luther:2014:CCA:2556420.2556788}.
-This body of work identifies ways to transform a variety of duties comprising complex tasks
-and distills them into ``a narrow set of routinized duties'',
-informed in part by researchers --- acting as inspectors --- and experts
-\cite[quotations from][]{10.2307/23702539}.
-Where \citeauthor{10.2307/23702539} would call additionally for the identification of
-crowdsourcing's version of ``piecework clerks'', we point out that
-today algorithms manage workers as pieceworkers once did
-\cite{uberAlgorithm,10.2307/23702539}.}
-
-\topic{Technology facilitates the subversion of expertise requirements
-by giving non--experts access to information that would otherwise be unavailable.}
-Taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge''
---- a demonstration of the driver's comprehensive familiarity.
-Researchers have identified significant growth of
-the hippocampal regions of the brains in veteran drivers,
-generally understood to be responsible for spatial functions such as navigation
+\topic{Technology increases non--experts' levels of expertise by giving access to information that would otherwise be unavailable.}
+For example, taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge''
+--- a demonstration of the driver's comprehensive familiarity with the city's roads.
+This test is so challenging that veteran drivers exhibit significantly larger
+the regions of the brain associated with spatial functions such as navigation
 \cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
       skok2000managing,Woollett1407,woollett2011acquiring}.
-Services such as Google Maps \& Waze make it possible for
+In contrast, with on--demand platforms such as Uber, services such as Google Maps \& Waze make it possible for
 people entirely unfamiliar with a city
-to know more about a city even than experts through
-the collective data generated by other users
-ranging topics such as police activity, congestion, construction, etc\dots
-\cite{silva2013traffic,hind2014outsmarting}.
-\msb{what's the insight I should take away from this paragraph? what does this say about crowd work?}
-\ali{Maybe I'm burnt out but I feel like the topic sentence is the takeaway and
-I'm not entirely sure what else to put here.
-Maybe that digital media can make it possible for non--experts to do increasingly expert work,
-informed by ``insider'' or ``expert'' knowledge of work, the way Uber, Waze, etc\dots do?}
+to operate profesionally~\cite{silva2013traffic,hind2014outsmarting}.
+Other examples include search engines enabling information retrieval, and word processors enabling spelling and grammar checking.
+By augmentating the human intellect~\cite{engelbart2001augmenting}, computing has shifted the complexity of work that is possible without training.
+
+Algorithms have automated some tasks that previously fell to management.
+Computational systems hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
+In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
+These algorithms, however, are less able than human managers to manage contingencies that were not programmed into them.
+
+% \ali{maybe this should go away?
+% \topic{Technology has made it possible
+% for non--experts to do work that was once considered
+% within the domain of experts.}
+% \citeauthor{yuanAlmost} builds on the work of others
+% (\textit{Voyant} and, more relevantly, \textit{CrowdCrit})
+% to design workflows that yield ``expert--level feedback''
+% \cite{yuanAlmost,Xu:2014:VGS:2531602.2531604,Luther:2014:CCA:2556420.2556788}.
+% This body of work identifies ways to transform a variety of duties comprising complex tasks
+% and distills them into ``a narrow set of routinized duties'',
+% informed in part by researchers --- acting as inspectors --- and experts
+% \cite[quotations from][]{10.2307/23702539}.
+% Where \citeauthor{10.2307/23702539} would call additionally for the identification of
+% crowdsourcing's version of ``piecework clerks'', we point out that
+% today algorithms manage workers as pieceworkers once did
+% \cite{uberAlgorithm,10.2307/23702539}.}
+Finally, the organizational limit on infrastructure creation is somewhat lessened. Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
+\citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
+As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use crowd work.
 
@@
 % we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
+As in expert crowdsourcing, being able to make stronger assumptions about workers' abilities will increase the complexity of crowd work outcomes.
+As workers have access to better tools, requesters will be able to make stronger assumptions in task design.
+However, these tool shifts will be tied to technological progress.
+Further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
+of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
+To some extent, an argument can be made that
+MOOCs and other online education resources
+provide crowd workers with the resources that they need, but 
+it remains to be seen whether that work will be appropriately valued, let alone
+properly interpreted by task solicitors
+\cite{aguaded2013mooc}.
+If we can overcome this obstacle,
+we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
+rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
+However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work.
+
+Algorithmic management 
+was not Turing-complete, and
 
@@
 
-Finally, and perhaps most importantly,
-we need to replicate the success of narrowly slicing education and training for expert work
-as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
-\ali{If we can find better ways to train crowd workers for narrowly defined work,
-we can envision more complex applications and outcomes of crowd work,
-as in the case of metalwork breaking away from the historical requirements of lengthy,
-comprehensive apprenticeships.}
-To some extent, an argument can be made that
-MOOCs and other online education resources
-provide crowd workers with the resources that they need, but 
-it remains to be seen whether that work will be appropriately valued, let alone
-properly interpreted by task solicitors
-\cite{aguaded2013mooc}.
-If we can overcome this obstacle,
-we might be able to empower crowd workers to do complex work such as engineering and metalworking,
-rather than doom them to match girl reputations:
-``brash, irregular, immoral, and uneducated''
-\cite{10.2307/3827491}.
-\msb{how will that reduce any of the problems except the last one?}
-\ali{I was thinking of fleshing out that the antipathy in general
-(``brash, immoral, etc\dots'') stems from these workers not being high--skill;
-we see similar descriptions used of Turkers when people talk about
-malicious and
-low--skill Turkers and getting them to ``expert--level work''.
-the historical counterexample is that
-metalworking pieceworkers didn't have these reputations,
-and we can suppose that this is because they \textit{were} trained
-(albeit unconventionally), and therefore considered educated, regular, etc\dots
-Too much of a stretch?}
-
 \onlyinsubfile{

### 2016-09-21 drafts/labor/complexity.tex
done with complexity pass

@@
 % we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
-As in expert crowdsourcing, being able to make stronger assumptions about workers' abilities will increase the complexity of crowd work outcomes.
-As workers have access to better tools, requesters will be able to make stronger assumptions in task design.
-However, these tool shifts will be tied to technological progress.
+Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes much like the shift to expert crowdsourcing.
 Further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
@@
 rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
-However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work.
+However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.
 
-Algorithmic management 
-was not Turing-complete, and
+Will the shift from human managers to Turing-complete algorithms raise the complexity ceiling? 
+By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
+%At worst, as now, they are far simpler~\cite{uberAlgorithm}.
+So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
+However, as a resource constraint, algorithms are a fixed cost and not a per-person cost like human managers.
+So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford on--demand labor.
+This shift may enable complex work that was not cost-effective before to become feasible.
 
-We should consider exploring the limitations that algorithmic management bring along more carefully.
-While crowd work research has touched on this subject, we've yet to make out the bigger picture of this theme
-\cite{uberAlgorithm}.
-If we can resolve the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest,
-then we may be able to break a toxic cycle of mistrustful requesters \cite[for example][]{MaliciousCrowdworkersGadiraju}
-and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
-\cite{takingAHITMcInnis}.
-\ali{What will happen if we don't do this?
-Maybe this is where I should bring up
-the ``market for lemons'' metaphor and say that
-the market will just continue to decline in quality until
-nobody is willing to rely on it either as a source of data/information (requesters), or
-a means of sustaining oneself financially (workers).}
+Because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re-introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
+This move, however, will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
+% and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
+% \cite{takingAHITMcInnis}.
+% \ali{What will happen if we don't do this?
+% Maybe this is where I should bring up
+% the ``market for lemons'' metaphor and say that
+% the market will just continue to decline in quality until
+% nobody is willing to rely on it either as a source of data/information (requesters), or
+% a means of sustaining oneself financially (workers).}
+
+Finally, the cost of creating piecework infrastructure has dropped. 
+Expensive manufacturing equipment has been largely replaced by computer code~\cite{lessig2006code}.
+As with lowered costs of management, lowered infrastructure costs will make crowd work accessible to a broader set of people and organizations.
+This in and of itself does not raise the complexity ceiling, but by broadening the potential market for crowd work, it may enable a new set of goals and needs take part.
 

### 2016-09-21 drafts/labor/discussion.tex
renamed research_implications.tex to discussion.tex

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Implications for Research}
\begin{enumerate}
  \item \ali{the past is no guarantee of the future; what could be unforeseen?
  \item it's easy to fall into the (dys|u)--topian camp of what piecework/crowd work will look like
  \item what should we be focusing on? the future of crowd work}
\end{enumerate}

\end{document}

### 2016-09-21 drafts/labor/complexity.tex

@@
 % we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
-Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes much like the shift to expert crowdsourcing.
-Further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
+Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
+Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
+Beyond this, further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
 of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
@@
 However, as a resource constraint, algorithms are a fixed cost and not a per-person cost like human managers.
-So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford on--demand labor.
-This shift may enable complex work that was not cost-effective before to become feasible.
-
-Because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re-introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
-This move, however, will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
+So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford crowd work.
+This shift may enable complex goals that were not cost-effective before to become feasible.
+However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re-introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
+This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
 % and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate

### 2016-09-21 drafts/labor/decomposition.tex
looked through complexity

@@
 \documentclass[trackingWork]{subfiles}
-% \makeatletter
-% \def\blx@maxline{77}
-% \makeatother
-\onlyinsubfile{\usepackage{xr-hyper}
-\usepackage{hyperref}
-\externaldocument{complexity}
-\externaldocument{relationships}
-\externaldocument{decomposition}
-}
+\makeatletter
+\def\blx@maxline{77}
+\makeatother
+% \onlyinsubfile{
+% \usepackage{xr-hyper}
+% \usepackage{hyperref}
+% \externaldocument{complexity}
+% \externaldocument{relationships}
+% % \externaldocument{decomposition}
+% }
 \begin{document}
@@
 %different \textit{types} of tasks in the span of just a few minutes
-Historical manufacturing equipment was not Turing-complete, and could not quickly be assembled, edited, and redeployed.\msb{cite for historical piecework equipment being a large time + money investment?}
+Historical manufacturing equipment could not quickly be assembled, edited, and redeployed.\msb{cite for historical piecework equipment being a large time + money investment?}
 In contrast, today system--designers can share, modify, and instantiate environments
@@
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-}
+% \onlyinsubfile{
+%   \balance{}
+%   \printbibliography
+% }
 

### 2016-09-21 drafts/labor/discussion.tex
wait is discussion being committed?

@@
 \begin{document}
-\section{Implications for Research}
+\section{Discussion}
+\topic{Having taken a comprehensive look toward crowd work through the piecework lens,
+we can't help but take a step back to consider a number of meta--issues that arose in our analysis.}
+
+\subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
+
+\subsection{Polarizing Tendencies}\label{sec:polarizationOfCrowdWork}
+
+\subsection{Taking Stock of our Research Agenda}\label{sec:whatShouldBeTheFuture}
+
 \begin{enumerate}

### 2016-09-21 drafts/labor/discussion.tex
checking in enumerated bullet points in discussion.tex and compiled pdfs

--- 2016-09-21 drafts/labor/discussion.tex
@@
 \begin{document}
-\section{Implications for Research}
+\section{Discussion}
+\topic{Having taken a comprehensive look toward crowd work through the piecework lens,
+we can't help but take a step back to consider a number of meta--issues that arose in our analysis.}
+Stated briefly, these issues are 
+\begin{inlinelist}
+  \item \namerefl{sec:perilousProblemsPredicting},
+  \item \namerefl{sec:polarizationOfCrowdWork}, and
+  \item \namerefl{sec:whatShouldBeTheFuture}.
+\end{inlinelist}
+We will attempt to grapple with these questions here based on
+what we brought up in the earlier case studies.
+
+\subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
 \begin{enumerate}
-  \item \ali{the past is no guarantee of the future; what could be unforeseen?
-  \item it's easy to fall into the (dys|u)--topian camp of what piecework/crowd work will look like
-  \item what should we be focusing on? the future of crowd work}
+  \item The past isn't a perfect indicator of the future
+  \item ``it would be wrong to conclude that
+        in the realm of digital labor
+        there is nothing new under the sun''
+        \cite{scholz2012digital}.
+  \item Digital sites of labor are so radically different
+        (for many of the reasons that \DO
+        articulated in the discussion about
+        organizing people and the challenges therein
+        \cite{dynamo}, as well as for
+        many of the rhizomatic features \citeauthor{miller2011understanding}
+        discusses
+        \cite{miller2011understanding}).
 \end{enumerate}
 
+\subsection{Polarizing Tendencies}\label{sec:polarizationOfCrowdWork}
+\begin{enumerate}
+  \item It's easy to think crowd work as heading toward some extreme
+  \item Activists describe speculative workers as having
+        ``essentially been turned into modern--day slaves''
+        \cite{activistsHuffPoLawsuit}.
+  \item Meanwhile, advocates describe it as
+        ``a project of sharing
+        aimed at providing ordinary people
+        with more economic opportunities and
+        improving their lives\dots''
+        \cite{uberPropaganda}.
+  \item There may be truth buried in both claims, but the important takeaway is that
+        the closest we can get to a single truth is somewhere in the middled,
+        buried in nuance.
+  \item There may be a Utopian world at the end of the tunnel that is our world now:
+        all work that we engage in could become speculative and risky;
+        the layers between us and our managers might increasingly become
+        ``defective (simple, observable)'' algorithms \cite{10.2307/2555446}
+        --- the same ones which already seem to frustrate
+        on--demand workers on Uber and other markets for labor
+        \cite{uberAlgorithm,dynamo,turkopticon}.
+  \item On the other hand, the future of crowd work could be eminently promising:
+        piecework's nascent years were nothing short of grim, but
+        they precipitated a century of the strongest labor advocacy the world has ever seen.
+        In India, workers across the nation engaged in
+        the largest labor strike in human history
+        % \cite{indiaStrikeIndustriall,indiaStrikeAlJazeera}
+        --- perhaps as many as 150 million
+        \cite{indiaStrikeRealNews}.
+        Arguably the Internet has made collective action on an unprecedented scale much easier.
+  \item So are we heading toward \textit{Mad Max} or\dots
+        \ali{is there a Utopian future movie?}
+        If we do nothing, probably somewhere in the middle.
+        It's possible that the difficulties of enforcing laws on multinational corporations is
+        tipping the balance of power in the direction of corporations, in which case
+        things will get worse;
+        but arguably we can affect change on the trajectory of crowd work,
+        benefiting from everything piecework scholars have learned about
+        collective action and governance
+        among workers, while also avoiding some of the perils they faced.
+\end{enumerate}
+
+\subsection{Deciding our Research Agenda}\label{sec:whatShouldBeTheFuture}
+\begin{enumerate}
+  \item We need to take a step back from the work that's been done so far and
+        think about where we are going with the crowd work research agenda.
+  \item Piecework researchers pointed out long ago that piecework is problematized by the fact that
+        ``piecework does not compensate workers for time spent switching tasks''
+        \cite{bewley1999wages} --- we've studied this phenomenon in crowd work
+        \cite{delayAndOrderLasecki,taskSearch}
+        and we should consider whether this remains a worthwhile area to explore unless
+        we're actually finding ways to minimize the many forms of costs of switching tasks.
+  \item What questions are we trying to answer?
+        If, as we assumed in
+        the previous section on \namerefl{sec:polarizationOfCrowdWork},
+        our driving motivation is to empower workers, we
+
+\end{enumerate}
+
+\endnotetext{
+\begin{enumerate}
+  \item the past is no guarantee of the future; what could be unforeseen?
+  \item it's easy to fall into the (dys|u)--topian camp of what piecework/crowd work will look like
+  \item what should we be focusing on? the future of crowd work
+\end{enumerate}
+}
+
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+}
+
 \end{document}

### 2016-09-21 drafts/labor/relationships.tex
looked through complexity
relationships crowd perspective

@@
 
+HCI and CSCW have largely framed themselves around supporting work rather than becoming an infrastructural layer enabling it.
+While all artifacts have politics, this shift into computational labor systems has directly impacted the lives and livelihood of workers.
+So, it is important to understand: what will the future look like for the workers who use these systems?
+
 \subsubsection{\crowdworkpers}
-\topic{The relationships of workers
-% with their work,
-with their peers and
-with requesters
-are nuanced and not especially well--understood.}
-Researchers have begun to appreciate
-the sociality of crowd workers in labor markets;
-still, the study of these communities is made more challenging by
-the limited access to workers on these sites of work
-inherent to digital spaces made without social affordances
-\cite{crowdcollab,miller2011understanding}.
-We can break this general body of work into two subgroups:
-workers' relationships
-\begin{inlinelist}
-% \item with \textit{work}, % \nameref{ch:workRelationship},
-\item with \textit{requesters}, and
-\item with \textit{other workers}.
-\end{inlinelist}
-We'll look at workers' relationships with work itself, which we'll discover
-gives us insight into why people engage in crowd work in the first place.
+\begin{comment}
+	- Workers do it for money
+	- Workers coordinate and communicate (gray, being a turker, team stuff). there is mega-drama
+	- Workers don't like requesters. have tried to organize
+\end{comment}
 
-\topic{\ali{some topic sentence that brings together the debate where
-one side blames Turkers for being bad at Turking and
-another side blames requesters for not understanding Turkers as a culture of people.}}
-Some research frames this tension as the Turker's problem
-(see, for example, \citeauthor{MaliciousCrowdworkersGadiraju}'s work,
-which frames the problem of unpredictable work as the result of ``malicious'' crowd workers),
-\cite{MaliciousCrowdworkersGadiraju,Sheng:2008:GLI:1401890.1401965,Ipeirotis:2010:QMA:1837885.1837906}.
+One of the initial questions that researchers asked was, who are the crowd workers and what draws them to crowd work?
+Early literature emphasized motivations like fun and spare change, but this narrative soon shifted to emphasize that many workers use platforms such as Amazon Mechanical Turk as a primary source of income~\cite{kaufmann2011more,ipeirotis2010demographics,Antin2012a}.
+Despite this, Mechanical Turk is a low-wage affair for most workers in the United States~\cite{ipeirotis2010demographics,martin2014being,gupta2014turk}.
+Thus, those who choose to opt out of the traditional labor force and spend significant time on Mechanical Turk are especially motivated by the opportunity for autonomy and skill variety~\cite{kaufmann2011more}.
+Due to valuing autonomy, it is tempting to ascribe attitudes of ``pity the workers'' to Turkers, but this frame is increasingly rejected by workers and designers as patronizing~\cite{storiesIraniSilberman}.
 
-Early on, \citeauthor{turkopticon} highlighted
-the information asymmetry between workers and requesters on AMT,
-leading to the creation of \TO, a site which allows
-Turkers to rate and review requesters
-\cite{turkopticon}.
-\citeauthor{dynamo} took this critique on information asymmetry and power imbalances a step further,
-designing \DO~to facilitate
-Turkers acting collectively to bring about changes to their circumstances
---- this led to the Academic Requester Guidelines
-\cite{dynamo}.
+% \topic{The relationships of workers
+% % with their work,
+% with their peers and
+% with requesters
+% are nuanced and not especially well--understood.}
+% We can break this general body of work into two subgroups:
+% workers' relationships
+% \begin{inlinelist}
+% % \item with \textit{work}, % \nameref{ch:workRelationship},
+% \item with \textit{requesters}, and
+% \item with \textit{other workers}.
+% \end{inlinelist}
+% We'll look at workers' relationships with work itself, which we'll discover
+% gives us insight into why people engage in crowd work in the first place.
+
+Workers' relationships with requesters are fraught.
+% \topic{\ali{some topic sentence that brings together the debate where
+% one side blames Turkers for being bad at Turking and
+% another side blames requesters for not understanding Turkers as a culture of people.}}
+Workers are often blamed for any low-quality work, regardless of whether they are responsible~\cite{martin2014being,takingAHITMcInnis}.
+Some research is extremely open about this position, blaming unpredictable work on ``malicious'' workers~\cite{MaliciousCrowdworkersGadiraju} or those with ``a lack of expertise, dedication [or] interest''~\cite{Sheng:2008:GLI:1401890.1401965}.
+% (see, for example, \citeauthor{MaliciousCrowdworkersGadiraju}'s work,
+% which frames the problem of unpredictable work as the result of ``malicious'' crowd workers),
+% \cite{MaliciousCrowdworkersGadiraju,Sheng:2008:GLI:1401890.1401965,Ipeirotis:2010:QMA:1837885.1837906}.
+Workers resent this position --- for good reason.
+\citeauthor{turkopticon} highlighted the information asymmetry between workers and requesters on AMT, leading to the creation of \TO, a site which allows Turkers to rate and review requesters~\cite{turkopticon}.
+Dynamo then took this critique on information asymmetry and power imbalances a step further, designing a platform to facilitate
+Turkers acting collectively to bring about changes to their circumstances~\cite{dynamo}.
 This unbridled power that requesters have over workers and
@@
 
-\topic{The frustration that workers experience dealing with requesters
-seems to precipitate frustration and mistrust between crowd workers, as well.}
-\citeauthor{dynamo} describes ``mega--drama'' among workers on forums for Turkers;
-\citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
-the culture of crowd work and the study thereof.
-\citeauthor{crowdcollab} quantifies and maps this social network of Turkers.
-\citeauthor{takingAHITMcInnis} takes these observations and considers
-what a crowd work platform might look like if it were to be designed more inclusively
-\cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab,takingAHITMcInnis}.
-The overarching theme of the research in this space has been
-documenting the struggle of crowd workers
-and attempting to intervene in constructive ways, while walking the balancing act
-(especially in the cases of \citeauthor{irani2015cultural} and later \citeauthor{storiesIraniSilberman})
-as we think about the culture of crowd workers.
+Researchers have also begun to appreciate the sociality of crowd workers.
+Because the platforms do not typically include social spaces, workers instead congregate off-platform in forums and mailing lists.
+There, Turkers exchange advice on high-paying work, talk about their earnings, build social connections, and discuss requesters~\cite{martin2014being}.
+Many crowd workers know each other through offline and online connections, coordinating behind-the-scenes despite the platforms encouraging independent work~\cite{crowdcollab,yin2016communication}.
+However, the frustration and mistrust that workers experience with requesters does occasionally boil over on the forums. 
+This behavior has come to be known as ``mega--drama'' amongst such workers~\cite{dynamo}.
+Still, the study of these communities is made challenging because most of these platforms do not themselves include social affordances for workers~\cite{miller2011understanding}.
+
+% \topic{The frustration that workers experience dealing with requesters
+% seems to precipitate frustration and mistrust between crowd workers, as well.}
+% \citeauthor{dynamo} describes ``mega--drama'' among workers on forums for Turkers;
+% \citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
+% the culture of crowd work and the study thereof.
+% \citeauthor{crowdcollab} quantifies and maps this social network of Turkers.
+% \citeauthor{takingAHITMcInnis} takes these observations and considers
+% what a crowd work platform might look like if it were to be designed more inclusively
+% \cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab,takingAHITMcInnis}.
+% The overarching theme of the research in this space has been
+% documenting the struggle of crowd workers
+% and attempting to intervene in constructive ways, while walking the balancing act
+% (especially in the cases of \citeauthor{irani2015cultural} and later \citeauthor{storiesIraniSilberman})
+% as we think about the culture of crowd workers.
 
@@
 
-\onlyinsubfile{
-\printbibliography
-}
+% \onlyinsubfile{
+% \balance{}
+% \printbibliography
+% }
 \end{document}

### 2016-09-21 drafts/labor/discussion.tex
turning enumerate into paragraphs

@@
 \subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
-\begin{enumerate}
-  \item The past isn't a perfect indicator of the future
-  \item ``it would be wrong to conclude that
+The past can't be a perfect predictor for the future;
+as \cite{scholz2012digital} points out,
+``it would be wrong to conclude that
 in the realm of digital labor
@@
 \cite{scholz2012digital}.
-  \item Digital sites of labor are so radically different
-        (for many of the reasons that \DO
-        articulated in the discussion about
-        organizing people and the challenges therein
-        \cite{dynamo}, as well as for
-        many of the rhizomatic features \citeauthor{miller2011understanding}
-        discusses
-        \cite{miller2011understanding}).
-\end{enumerate}
+Many of the challenges that \DO~overcame
+--- determining trustworthiness, ensuring anonymity, etc\dots
+\cite{dynamo} ---
+were relatively unique challenges precipitated by the affordances of the Internet, specifically
+that people could (and often do) contribute to communities in a one--off manner
+\cite{mcinnis2016one}.
+The Internet seemed to make this kind of loose affiliation feasible where before it wasn't
+\cite{catalyst,dynamo}.
+% articulated in the discussion about
+% organizing people and the challenges therein
+% \cite{dynamo}, as well as for
+% many of the rhizomatic features \citeauthor{miller2011understanding}
+% discusses
+% \cite{miller2011understanding}).
+
+But this does not mean that
+attempting to draw meaningfully from historical scholarship would be folly;
+as we have shown in the preceding sections,
+enough of piecework can and does inform crowdsourcing that
+we should take this as a cue to look first for
+historical framings on phenomena that we encounter,
+as this work can outline both the mechanisms that we should expect to play out
+as well as one of the (perhaps many) possible futures.
+
+In particular, the predictions that have emerged surrounding crowd work have run the spectrum
+from deep pessimism to exuberant optimism. \ali{should I cite this?}
+In the next section, we will use
+the piecework foundation which informed our case studies
+to trace out possible Dystopian and Utopian futures for crowd work.
 

### 2016-09-21 drafts/labor/discussion.tex
changes on subsections in discussion

@@
 \subsection{Polarizing Tendencies}\label{sec:polarizationOfCrowdWork}
-\begin{enumerate}
-  \item It's easy to think crowd work as heading toward some extreme
-  \item Activists describe speculative workers as having
+\topic{It would be easy
+to think about the future of crowd work and
+end up at one of two extremes.}
+On one hand, crowd work researchers
+imagine the application of crowdsourcing as
+a potentially bright future that enables the achievement of near--impossible goals
+\cite{redballoon,crowdworkFuture,vizwiz}; % \ali{and who else should I cite here?}
+on the other hand, researchers
+warn against crowd work as potentially exploitative sites of dispossession \cite{scholz2012digital},
+a site of racial discrimination \cite{edelman2015racial},
+and invisible, deeply frustrated workers \cite{turkopticon,bighamHalfWorkday}.
+
+\topic{A uniquely challenging facet of this topic of inquiry is the public attention that this domain has attracted.}
+Activists have described speculative work as having
 ``essentially been turned into modern--day slaves''
 \cite{activistsHuffPoLawsuit}.
-  \item Meanwhile, advocates describe it as
+Meanwhile, advocates describe it as
 ``a project of sharing
@@
 \cite{uberPropaganda}.
-  \item There may be truth buried in both claims, but the important takeaway is that
-        the closest we can get to a single truth is somewhere in the middled,
-        buried in nuance.
-  \item There may be a Utopian world at the end of the tunnel that is our world now:
-        all work that we engage in could become speculative and risky;
+\ali{these are salient, but with the last paragraph it feels really flimsy.
+I'd really like to get this Uber quote in, but
+not if it means this paragraph gets that much weaker.}
+
+% \topic{There is, of course, a kernel of truth in everything here.}
+\topic{There is evidence to support the claim that
+a Dystopian world awaits us at the end of the tunnel.}
+The arbitrary nature of on--demand work may permeate our lives, inducing us
+``to neglect tasks that are less easy to measure'' \cite{SJOE:SJOE371}
+rewarding us not for creativity but predictability;
+payment for this work may ultimately be determined by
+an algorithm that fundamentally doesn't understand us;
 the layers between us and our managers might increasingly become
-        ``defective (simple, observable)'' algorithms \cite{10.2307/2555446}
-        --- the same ones which already seem to frustrate
-        on--demand workers on Uber and other markets for labor
+``defective (simple, observable)'' algorithms \cite{10.2307/2555446},
+just like those which already frustrate
+on--demand workers
 \cite{uberAlgorithm,dynamo,turkopticon}.
-  \item On the other hand, the future of crowd work could be eminently promising:
-        piecework's nascent years were nothing short of grim, but
-        they precipitated a century of the strongest labor advocacy the world has ever seen.
-        In India, workers across the nation engaged in
+
+\topic{On the other hand, there is some evidence that offers hope.}
+Yes, piecework's nascent years were arguably grim, but
+they precipitated a century of the strongest labor advocacy the world has ever seen
+\cite{hart2013rise,mccallum2013global}.
+Even today, the geist that came out of the labor union revolution
+inspires collective action and worker empowerment around the world:
+in India, workers across the nation engaged in
 the largest labor strike in human history
-        % \cite{indiaStrikeIndustriall,indiaStrikeAlJazeera}
 --- perhaps as many as 150 million
 \cite{indiaStrikeRealNews}.
-        Arguably the Internet has made collective action on an unprecedented scale much easier.
-  \item So are we heading toward \textit{Mad Max} or\dots
-        \ali{is there a Utopian future movie?}
-        If we do nothing, probably somewhere in the middle.
-        It's possible that the difficulties of enforcing laws on multinational corporations is
-        tipping the balance of power in the direction of corporations, in which case
-        things will get worse;
-        but arguably we can affect change on the trajectory of crowd work,
-        benefiting from everything piecework scholars have learned about
-        collective action and governance
-        among workers, while also avoiding some of the perils they faced.
-\end{enumerate}
+If labor advocacy groups can find ways to permeate on--demand labor markets,
+as some have called for
+\cite{futureUnions},
+then the future of crowd work may follow
+the same trajectory of worker empowerment that piecework \textit{later} found.
+\ali{The Internet may make this possible!}
+% Arguably the Internet has made collective action on an unprecedented scale much easier.
+
+\topic{Perhaps the most pressing question is this:
+is the future of crowd work going to look more \textit{Dystopian} or \textit{Utopian}?}
+% The unsatisfying answer is that
+If we have learn nothing, crowd work will probably end up somewhere in the middle.
+% It's possible that the difficulties of enforcing laws on multinational corporations is
+% tipping the balance of power in the direction of corporations, in which case
+% things will get worse;
+% but arguably we can affect change on the trajectory of crowd work,
+% benefiting from everything piecework scholars have learned about
+% collective action and governance
+% among workers, while also avoiding some of the perils they faced.
 

### 2016-09-21 drafts/labor/relationships.tex

@@
 \subsubsection{\pieceworkpers}
-\citeauthor{clark1908cotton}, for instance,
-relays his observations of textile mill pieceworkers and his realization that
+Early observers believed that workers were strongly motivated by the piecework model.
+\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
 ``When he works by the day the Italian operative wishes to leave before the whistle blows,
-but if he works by the piece he will work as many hours as it is possible for him to stand''
+but if he works by the piece he will work as many hours as it is possible for him to stand.''
+Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
 
-\citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City,
-and \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes,
-offering guidance on how best to manage pieceworkers
-\cite{riisOtherSideLives,norton1900textile}.
-
-\topic{Despite the intense growth of the piecework approach to renumeration,
-this time was not without turmoil.}
-As previously discussed, a number of worker organizations weighed in on
-(or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time
-\cite{american1921problem,richards1904anything}.
-As mounting attention increasingly revealed problems in piecework as it related to workers,
-workers themselves began to speak out about their frustration with this new regime.
-\citeauthor{riisOtherSideLives}'s photo--documentary work
-brought light to otherwise--invisible pieceworkers in their homes;
-this and other events led to organizations representing
-railway workers, mechanical engineers, and others
-who began to mount advocacy in defense and favor of workers
-\cite{american1921problem,richards1904anything,riisOtherSideLives}.
-\citeauthor{10.2307/3827491} tells us of the match--girls strike of 1888,
-one of the earliest and most famous successful worker strikes and perhaps
-the beginning of ``militant trade unionism''
-\cite{10.2307/3827491}.
-It wasn't an exaggeration when \citeauthor{weyer1894history} later said that
-``the match--girls' victory turned a new leaf in Trade Union annals'' \cite{weyer1894history};
-in the 30 years since the match--girls strike,
-the Trade Union Movement grew from
-``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
-\cite{webb1920history}.
-
-\topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
-The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
---- albeit critical ---
-case study in this space;
-the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
-to strike for
-an individual minimum wage, among other demands
-\cite{10.2307/2221944}.
-\citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
-\cite{10.2307/41829256}.
+Workers' relationships with employers quickly soured.
+The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+% in the 30 years since the match--girls strike,
+% the Trade Union Movement grew from
+% ``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
+% \cite{webb1920history}.
+This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
 The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
@@
 
-\topic{The questions surrounding
-the ways pieceworkers related to managers might be best answered by
-the work that has been done in the emergence and proliferation of labor unions.}
-The primary avenue for workers to interact with managers has been through
-laborer advocacy groups such as the American Federation of Labor,
-(one of the forerunners of the largest and
-most politically influential labor union in the United States).
-Looking through that lens, we find copious research on
-the relationships between workers and requesters
-\cite{levi2009union,ahlquist2013interest,
+Soon, many worker organizations were weighing in on (or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
+As mounting attention increasingly revealed problems in piecework's treatment of workers, theworkers themselves began to speak out about their frustration with this new regime.
+Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers
+\cite{american1921problem,richards1904anything}.
+
+% \topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
+% The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
+% --- albeit critical ---
+% case study in this space;
+% the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
+% to strike for
+% an individual minimum wage, among other demands
+% \cite{10.2307/2221944}.
+% \citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
+% \cite{10.2307/41829256}.
+% The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
+% \cite{jacoby1983union}.
+
+Pieceworkers' relationships with their employers eventually developed a pattern of using 
+% \topic{The questions surrounding
+% the ways pieceworkers related to managers might be best answered by
+% the work that has been done in the emergence and proliferation of labor unions.}
+% The primary avenue for workers to interact with managers has been through
+laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,
       mccallum2013global,jacoby1983union}.
-One component of collectively negotiating with managers has been the process
-of collective action, a topic which has been substantively explored
-but is not quite yet answered
-\cite{russell1982collective,olsonlogic}.
+% (one of the forerunners of the largest and
+% most politically influential labor union in the United States).
+% Looking through that lens, we find copious research on
+% the relationships between workers and requesters
+% \cite{levi2009union,ahlquist2013interest,
+%       mccallum2013global,jacoby1983union}.
+Collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.
 
-\topic{Answering how workers related to one another is arguably more challenging
-for a number of reasons.}
+\topic{Less is known about how pieceworkers related to each other.}
 For one thing, the research methods we typically associate with the exploratory study of cultures
 --- Anthropology, and namely participant--observation, ethnography, etc\dots ---
-didn't exist quite as we know them at the turn of the \nth{20} century, and
-wouldn't for several more decades.
-Still, we can look at primary sources, like \citetitle{american1921problem}
-to give us some hint of how they related to each other
-\cite{american1921problem}.
-
-The driving force of American labor advocacy organizations was to get piecework railroad workers
-to identify
+didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
+% Still, we can look at primary sources, like \citetitle{american1921problem}
+% to give us some hint of how they related to each other
+% \cite{american1921problem}.
+Primary sources indicate that labor organizations wished for workers to identify as a collective group, 
+% The driving force of American labor advocacy organizations was to get piecework railroad workers
+% to identify
 ``not only as railroad employees but also as members of the larger life of the community''
@@
 \cite{ostrom1990governing,russell1982collective,olsonlogic}.
-\citeauthor{riisOtherSideLives} had contributed to this sense of shared struggle and endurance
-by the time \citetitle{american1921problem} was published
+\citeauthor{riisOtherSideLives} also contributed to this sense of shared struggle and endurance
+% by the time \citetitle{american1921problem} was published
 by documenting pieceworkers in their home--workplaces,
@@
 \subsubsection{\whatchanged}
+
+\begin{comment}
+\end{comment}
+
 \topic{The differences between crowd workers and pieceworkers seem defined largely by

### 2016-09-21 drafts/labor/relationships.tex
differences

@@
 \begin{comment}
+	*workers make little money but love autonomy --- workers make little money
+	workers blamed for quality --- ???
+	both cases, sociality is hard
+	*collective action hard --- collective action succeeded
+
+	- algorithms, not managers
 \end{comment}
 
-\topic{The differences between crowd workers and pieceworkers seem defined largely by
-the differences in the places of work.}
-Whereas it arguably became inevitable that workers would have a place to
-meet, discuss, and collaborate
-when they began sharing places of work,
-online spaces make it much harder to do so.
-Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
-without any affiliation with
---- or even knowledge of ---
-communities of peers
-% \ali{multiple citations that labor unions came out of factories here}
-\cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
+While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', there was generally less written about quality concerns for historical pieceworkers than there is in modern crowd work.
+Why the difference? 
+One possibility is that, through writing web scripts and applying them to many tasks, it is possible for a small number of spammers have an outsized influence.
+Historically, it was much harder for such workers to move and get new jobs --- today, they can simply accept a different task on Mechanical Turk.
+Another possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen, online workers are known by little more than an account ID.
 
-\topic{We further find the sources of differences between crowd work and piecework
-in the nature of the relationship between workers and requesters
---- or rather, the lack thereof.}
-While historically the management of workers had to be done through a foreman
+% \topic{The differences between crowd workers and pieceworkers seem defined largely by
+% the differences in the places of work.}
+% Whereas it arguably became inevitable that workers would have a place to
+% meet, discuss, and collaborate
+% when they began sharing places of work,
+% online spaces make it much harder to do so.
+% Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
+% without any affiliation with
+% --- or even knowledge of ---
+% communities of peers
+% % \ali{multiple citations that labor unions came out of factories here}
+% \cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
+
+\topic{The relationship between workers and employers has also shifted: while historically the management of workers had to be done through a foreman
 (who necessarily had an intuitive
@@
 by algorithms of the \nth{21} century
-\cite{uberAlgorithm}.
+\cite{uberAlgorithm}.}
 The result of this change is that the agents managing work are now
-cold and logical, if unforgiving.
+cold, logical, and unforgiving.
 Where a person might recognize that the ``attention check'' questions 
-proposed by \citeauthor{le2010ensuring} ensure that malicious and inattentive are stopped,
+proposed by \citeauthor{le2010ensuring} ensure that malicious and inattentive workers are stopped,
 some implementations of these approaches
-only seem to antagonize workers \ali{shots fired}
-\cite{le2010ensuring,MaliciousCrowdworkersGadiraju}.
-% We've known for decades that workers prefer human management 
-\citeauthor{10.2307/2555446} told us more than 30 years ago
---- in \citeyear{10.2307/2555446} --- that
-``\dots~when performance is difficult to evaluate,
+only seem to antagonize workers.
+More than 30 years ago, \citeauthor{10.2307/2555446} wrote:
+``When performance is difficult to evaluate,
 imperfect input measures and
@@
 \cite{10.2307/2555446}.
-\ali{here's an idea. feel free to push back or revise or something:\\
-``Should we be surprised that
-management approaches we've known to
-frustrate pieceworkers since \citeyear{10.2307/2555446}
-would only aggravate crowd workers in many of the same ways?''}
+This frustration has only grown as requesters have had to rely on automatic management mechanisms. Only a few use the equivalent of human foremen~\cite{haas2015argonaut,kulkarni2012mobileworks}.
+
+% \ali{here's an idea. feel free to push back or revise or something:\\
+% ``Should we be surprised that
+% management approaches we've known to
+% frustrate pieceworkers since \citeyear{10.2307/2555446}
+% would only aggravate crowd workers in many of the same ways?''}
+
+Relative to the mature state of collective action for pieceworkers offline, crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
+Both pieceworkers and crowd workers have struggled at times to form a collective identity necessary to organize.
+With workers joining and leaving the crowd labor force continuously, and with many part-time members, it is extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
+However, even when they can: whereas pieceworkers could physically block access to a site of production, online labor markets provide no facilities for workers to change the experience of other workers.
+This is a key limitation --- without it, workers cannot enforce a strike.
 

### 2016-09-21 drafts/labor/conclusion.tex
added a conclusion.tex file. finished with discussion (i think)

--- 2016-09-21 drafts/labor/discussion.tex
@@
 \makeatother
+
 \begin{document}
-\section{Discussion}
-\topic{Having taken a comprehensive look toward crowd work through the piecework lens,
-we can't help but take a step back to consider a number of meta--issues that arose in our analysis.}
-Stated briefly, these issues are 
-\begin{inlinelist}
-  \item \namerefl{sec:perilousProblemsPredicting},
-  \item \namerefl{sec:polarizationOfCrowdWork}, and
-  \item \namerefl{sec:whatShouldBeTheFuture}.
-\end{inlinelist}
-We will attempt to grapple with these questions here based on
-what we brought up in the earlier case studies.
-
-\subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
-\begin{enumerate}
-  \item The past isn't a perfect indicator of the future
-  \item ``it would be wrong to conclude that
-        in the realm of digital labor
-        there is nothing new under the sun''
-        \cite{scholz2012digital}.
-  \item Digital sites of labor are so radically different
-        (for many of the reasons that \DO
-        articulated in the discussion about
-        organizing people and the challenges therein
-        \cite{dynamo}, as well as for
-        many of the rhizomatic features \citeauthor{miller2011understanding}
-        discusses
-        \cite{miller2011understanding}).
-\end{enumerate}
-
-\subsection{Polarizing Tendencies}\label{sec:polarizationOfCrowdWork}
-\begin{enumerate}
-  \item It's easy to think crowd work as heading toward some extreme
-  \item Activists describe speculative workers as having
-        ``essentially been turned into modern--day slaves''
-        \cite{activistsHuffPoLawsuit}.
-  \item Meanwhile, advocates describe it as
-        ``a project of sharing
-        aimed at providing ordinary people
-        with more economic opportunities and
-        improving their lives\dots''
-        \cite{uberPropaganda}.
-  \item There may be truth buried in both claims, but the important takeaway is that
-        the closest we can get to a single truth is somewhere in the middled,
-        buried in nuance.
-  \item There may be a Utopian world at the end of the tunnel that is our world now:
-        all work that we engage in could become speculative and risky;
-        the layers between us and our managers might increasingly become
-        ``defective (simple, observable)'' algorithms \cite{10.2307/2555446}
-        --- the same ones which already seem to frustrate
-        on--demand workers on Uber and other markets for labor
-        \cite{uberAlgorithm,dynamo,turkopticon}.
-  \item On the other hand, the future of crowd work could be eminently promising:
-        piecework's nascent years were nothing short of grim, but
-        they precipitated a century of the strongest labor advocacy the world has ever seen.
-        In India, workers across the nation engaged in
-        the largest labor strike in human history
-        % \cite{indiaStrikeIndustriall,indiaStrikeAlJazeera}
-        --- perhaps as many as 150 million
-        \cite{indiaStrikeRealNews}.
-        Arguably the Internet has made collective action on an unprecedented scale much easier.
-  \item So are we heading toward \textit{Mad Max} or\dots
-        \ali{is there a Utopian future movie?}
-        If we do nothing, probably somewhere in the middle.
-        It's possible that the difficulties of enforcing laws on multinational corporations is
-        tipping the balance of power in the direction of corporations, in which case
-        things will get worse;
-        but arguably we can affect change on the trajectory of crowd work,
-        benefiting from everything piecework scholars have learned about
-        collective action and governance
-        among workers, while also avoiding some of the perils they faced.
-\end{enumerate}
-
-\subsection{Deciding our Research Agenda}\label{sec:whatShouldBeTheFuture}
-\begin{enumerate}
-  \item We need to take a step back from the work that's been done so far and
-        think about where we are going with the crowd work research agenda.
-  \item Piecework researchers pointed out long ago that piecework is problematized by the fact that
-        ``piecework does not compensate workers for time spent switching tasks''
-        \cite{bewley1999wages} --- we've studied this phenomenon in crowd work
-        \cite{delayAndOrderLasecki,taskSearch}
-        and we should consider whether this remains a worthwhile area to explore unless
-        we're actually finding ways to minimize the many forms of costs of switching tasks.
-  \item What questions are we trying to answer?
-        If, as we assumed in
-        the previous section on \namerefl{sec:polarizationOfCrowdWork},
-        our driving motivation is to empower workers, we
-
-\end{enumerate}
-
-\endnotetext{
-\begin{enumerate}
-  \item the past is no guarantee of the future; what could be unforeseen?
-  \item it's easy to fall into the (dys|u)--topian camp of what piecework/crowd work will look like
-  \item what should we be focusing on? the future of crowd work
-\end{enumerate}
-}
-
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-}
-
+\section{Conclusion}
+????
 \end{document}

### 2016-09-21 drafts/labor/relationships.tex

@@
 \subsubsection{\implication}
-\topic{What we've done in the field of crowd work might be able to tell us something about piecework
-just as piecework has told us so much about crowd work.}
-Crowd work research doesn't just benefit from
-digital media allowing us to make
-relationship networks like \citeauthor{crowdcollab} do;
-we benefit from the firmer theoretical basis of Anthropology that
-existed in a radically different form at the turn of the \nth{20} century,
-when piecework began to emerge.
+
+%	*workers make little money but love autonomy --- workers make little money
+%	workers blamed for quality --- ???
+%	both cases, sociality is hard
+%	*collective action hard --- collective action succeeded
+%	- algorithms, not managers
+
+The decentralization and anonymization of crowd work will continue to make many of its social relationships a struggle.
+While some workers get to know each other well on forums~\cite{martin2014being,crowdcollab}, many never engage in these social spaces.
+Without intervention, worker relationships and collectivism are likely to be inhibited by this decentralized design.
+One option is to build worker centralizing points into the platform, for example asking workers to vote on each others' reputation or allowing groups of workers to collectively reject a task from the platform~\cite{crowdguilds}.
+
+\topic{The history of piecework further suggests that relationships between workers and employers might be improved if employers engaged in more human management styles.}
+% \citeauthor{10.2307/2555446} drew a dichotomous line between
+% ``defective (simple, observable) output measures'' and
+% ``a manager's subjective judgment'',
+% but such a dichotomy need not necessarily represent our work management styles
+% \cite{10.2307/2555446}.
+Instead of delegating as many management tasks as possible to an algorithm, it might be possible to build dashboards and other information tools that empower modern crowd work foremen~\cite{kulkarni2012mobileworks}.
+If the literature on piecework is to be believed,
+more considerate \textit{human} management may resolve
+many of the tensions we've discovered among among crowd workers.
+
+\topic{Reciprocally, crowd work may be able to inform piecework research.}
+There exists far less literature about pieceworkers' relationships than there does today about crowd workers' relationships.
+Crowd work research benefits from both the accessibility of digital platforms, as well as the firmer theoretical basis of Anthropology than
+existed at the turn of the \nth{20} century, when piecework began to emerge.
 \citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and
@@
 \cite{malinowski2002argonauts,boas1940race,mead1973coming}.
-
-\topic{The research on piecework still offers to guide us on perhaps the most rudimentary aspects of worker management}
-\citeauthor{10.2307/2555446} drew a dichotomous line between
-``defective (simple, observable) output measures'' and
-``a manager's subjective judgment'',
-but such a dichotomy need not necessarily represent our work management styles
-\cite{10.2307/2555446}.
-We can develop tools that better inform humans, rather than
-(perhaps futilely) attempt to delegate all worker management to machines.
-This is an area we should pursue, but haven't yet.
-If the literature on piecework is to be believed,
-more considerate \textit{human} management may resolve
-many of the tensions we've discovered among among crowd workers.
+Modern crowd work may give us an opportunity to revisit open questions in piecework with a more refined lens.
 

### 2016-09-21 drafts/labor/piecework_lit.tex
piecework_lit

@@
 \end{inlinelist}
-
-\ali{Maybe something here about how this isn't comprehensive, but it gives you enough to understand the
-era about which we're talking, the very high--level things that were going on, etc\dots~so that
-the later discussions actually make sense?}
+This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.
 
@@
 by contrasting the ``task--labourer'' with the ``day--labourer'':
-``\dots the chief difference lies between the day--labourer,
+``the chief difference lies between the day--labourer,
 who receives a certain some of money\dots~for his day's work,
-and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''
-\cite{hughRaynbirdTaskWork}.
+and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''~\cite{hughRaynbirdTaskWork}.
 \citeauthor{10.2307/2338394} gives a more illustrative definition of piecework,
 offering examples:
-``\dots~payment is made for each hectare which is pronounced to be well ploughed~\dots~
-for each living foal got from a mare;~\dots~
-for each living calf got~\dots'' etc\dots
-\cite{10.2307/2338394}.
+``payment is made for each hectare which is pronounced to be well ploughed~[\dots]~
+for each living foal got from a mare;~[\dots]~
+for each living calf got~[\dots]''~\cite{10.2307/2338394}.
 This framing perhaps makes the most intuitive sense;
@@
 \cite{hart2013rise}.}
-Indeed, we see work that adopts
-piece--rate compensation in some aspects and
+Indeed, work adopted piece--rate compensation in some aspects and
 time--rate compensation in others.
@@
 \citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
-crop up; he points out that\dots
-``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion\dots
-he knows that all depends on his own diligence and perseverance\dots~[and]
+crop up; he points out that 
+``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~[\dots]~
+he knows that all depends on his own diligence and perseverance~[\dots and]~
 so long as he performs his work to the satisfaction of his master,
 he is not under that control to which the day--labourer is always subject.''
-\citeauthor{hughRaynbirdTaskWork} (and others, as we will see)
-highlight the freedom from control that ``task--labourers'' enjoy
-\cite{hughRaynbirdTaskWork,rowan1901premium}.
+\citeauthor{hughRaynbirdTaskWork} highlight the freedom from control that ``task--labourers'' enjoy~\cite{hughRaynbirdTaskWork,rowan1901premium}.
 
@@
 ``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
-the management recognised his right to put a `substitute' or `grass' compositor in his place''
-\cite{10.2307/27508091}.
-From these accounts we should be able to identify
-a sense of independence that
-resonates across decades, industries, and locales where piecework is found.
-We'll problematize this supposed advantage as we trace the history of piecework,
-but for now we can say that piecework affords
-independence and some sense of autonomy
-new to people in the working class.
+the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
+From these accounts we identify a sense of independence and autonomy that resonates across decades, industries, and locales where piecework is found.
+% We'll problematize this supposed advantage as we trace the history of piecework,
+% but for now we can say that piecework affords
+% independence and some sense of autonomy
+% new to people in the working class.
 
-\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward
-the question of the features that sprout from piecework.}
-In their reflection on the features endemic to piecework in the 1930s,
-which they describe as the ``heyday'' of piecework's prominence;
-among them were the following:
-\begin{inlinelist}
-  \item ``female workers who generally had less training'' had to be trained in
-        narrower subsets of the general body of skills that
-        conventional (male) apprentices would undertake, and
-  \item workers with specific slices of skills could be more appropriately matched to
-        suitable tasks
-\end{inlinelist}
-\cite{hart2013rise}.
-Piecework thus opened the door for people who previously couldn't participate in the labor market
---- either for lack of training or for other reasons ---
-to do so, and to acquire job skills incrementally.
+\topic{Piecework opened the door for people who previously couldn't participate in the labor market --- for example due to lack of training --- to do so, and to acquire job skills incrementally.}
+For example, women could receive training in narrow subsets of the general body of skills, enabling them to act in capacities similar to what conventional (male) apprentices would undertake~\cite{hart2013rise}.
+% offer another series of compelling insights toward
+% the question of the features that sprout from piecework.}
+% In their reflection on the features endemic to piecework in the 1930s,
+% which they describe as the ``heyday'' of piecework's prominence;
+% among them were the following:
+% \begin{inlinelist}
+%   \item ``female workers who generally had less training'' had to be trained in
+%         narrower subsets of the general body of skills that
+%         conventional (male) apprentices would undertake, and
+In addition, workers with specific slices of skills could be matched to suitable tasks.
 Workers without conventional training
@@
 
-\subsection{A Primer of Piecework}\label{sec:pieceworkPrimer} % yay rhyming?
+\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % yay rhyming?
 \topic{In this section we will offer a brief overview of the history of piecework;
-this should not be mistaken for a comprehensive or especially rigorous background.}
+this should not be mistaken for a comprehensive background.}
 Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
-In other words, framing piecework in the contexts of
-the early days of the industrial revolution,
+In other words, it will frame piecework in the contexts of
+the early days of the Industrial Revolution,
 through the political and economic turmoil of the early and mid--\nth{20} century,
@@
 
-\ali{Where do I want to take this? Where are we now?
-\begin{enumerate}
-  \item Piecework started in the \nth{19} century (above). [DONE]
-  \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
-  \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
-  \item Piecework died pretty quickly (below). [DONE]
-\end{enumerate}
-Sound good?}
+% \ali{Where do I want to take this? Where are we now?
+% \begin{enumerate}
+%   \item Piecework started in the \nth{19} century (above). [DONE]
+%   \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
+%   \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
+%   \item Piecework died pretty quickly (below). [DONE]
+% \end{enumerate}
+% Sound good?}
 
@@
 elsewhere in manufacturing \cite{10.2307/3827491}.
-This proliferation of manual piecework led to discussion surrounding how best to manage these workers
-\cite{norton1900textile,clark1908cotton},
-which seemed to lead to frustration among workers over their poor working and living conditions
-(famously documented by
-% \ali{??? Why is this here? Can it be moved?}
-% {As increasing attention revealed problems in piecework as it related to workers,
-% workers themselves began to speak out about their frustration with this new regime.}
-\citeauthor{riisOtherSideLives})
-\cite{riisOtherSideLives}.
-This led to industry organizations representing
-railway workers, mechanical engineers, and others contributing their myriad perspectives
-\cite{american1921problem,richards1904anything}.
-% Nevertheless, piecework continued to permeate low--skilled labor.
-% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
-
-\topic{Fordism and scientific management thrust piecework into higher gear, especially as
+Fordism and scientific management thrust piecework into higher gear, especially as
 mass manufacturing and
@@
 % \cite{honey1985creating,davies2014origins}.
+
+The early proliferation of manual piecework led to discussion surrounding how best to manage pieceworkers
+\cite{norton1900textile,clark1908cotton}.
+Despite this, workers' means were mostly ignored,
+leading to frustration over poor working and living conditions
+(famously documented by
+% \ali{??? Why is this here? Can it be moved?}
+% {As increasing attention revealed problems in piecework as it related to workers,
+% workers themselves began to speak out about their frustration with this new regime.}
+\citeauthor{riisOtherSideLives})
+\cite{riisOtherSideLives}.
+This led to industry organizations representing
+railway workers, mechanical engineers, and others beginning to speak out on pieceworkers' behalf~\cite{american1921problem,richards1904anything}.
+% Nevertheless, piecework continued to permeate low--skilled labor.
+% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
 
@@
 
-\subsection{Why is piecework relevant to crowd work?}
+\subsection{Piecework and crowd work}
 

### 2016-09-21 drafts/labor/discussion.tex

@@
 \makeatother
+\onlyinsubfile{
+\usepackage{xr-hyper}
+\usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{relationships}
+\externaldocument{decomposition}
+}
+
 \begin{document}
@@
 is the future of crowd work going to look more \textit{Dystopian} or \textit{Utopian}?}
-% The unsatisfying answer is that
 If we have learn nothing, crowd work will probably end up somewhere in the middle.
-% It's possible that the difficulties of enforcing laws on multinational corporations is
-% tipping the balance of power in the direction of corporations, in which case
-% things will get worse;
-% but arguably we can affect change on the trajectory of crowd work,
-% benefiting from everything piecework scholars have learned about
-% collective action and governance
-% among workers, while also avoiding some of the perils they faced.
+It's possible that the difficulties of enforcing laws on multinational corporations is
+tipping the balance of power in the direction of corporations, in which case
+things will get worse;
+but arguably we can affect change on the trajectory of crowd work,
+benefiting from everything piecework scholars have learned about
+collective action and governance
+among workers, while also avoiding some of the perils they faced.
+\ali{I really don't like this paragraph, but I felt like we needed to plant a flag on the question in the topic sentence}
 
-\subsection{Deciding our Research Agenda}\label{sec:whatShouldBeTheFuture}
-\begin{enumerate}
-  \item We need to take a step back from the work that's been done so far and
-        think about where we are going with the crowd work research agenda.
-  \item Piecework researchers pointed out long ago that piecework is problematized by the fact that
+\subsection{Determining our Research Agenda}\label{sec:whatShouldBeTheFuture}
+Taking a step back from the work that's been done so far and
+thinking about where the body of crowd work research is attempting to go may help us identify
+areas of research on which we should be pushing forward on.
+We return to the three questions that motivated this paper:
+\begin{inlinelist}
+  \item ``\namerefl{sec:complexity}?''
+  \item ``\namerefl{sec:decomposition}?''
+        and
+  \item ``\namerefl{sec:relationships}?''
+\end{inlinelist}.
+
+While we have arguably outpaced piecework with regard to the limits on the complexity of work,
+we have yet to find insurmountable limits to crowdsourcing.
+Still, we can learn from the piecework literature as it relates to
+the stymieing effect that mismanagement has on workers;
+research into the complexity limits should emphasize on finding new ways to manage workers,
+in particular using humans --- perhaps other crowd workers ---
+to act as modern foremen.
+
+Piecework researchers looking into decomposition pointed out long ago that
+piecework is saddled by a lower limit on decomposition:
 ``piecework does not compensate workers for time spent switching tasks''
-        \cite{bewley1999wages} --- we've studied this phenomenon in crowd work
-        \cite{delayAndOrderLasecki,taskSearch}
-        and we should consider whether this remains a worthwhile area to explore unless
-        we're actually finding ways to minimize the many forms of costs of switching tasks.
-  \item What questions are we trying to answer?
-        If, as we assumed in
-        the previous section on \namerefl{sec:polarizationOfCrowdWork},
-        our driving motivation is to empower workers, we
+\cite{bewley1999wages}.
+We've since studied this phenomenon in crowd work to great length both
+observationally \cite{taskSearch} and
+experimentally \cite{delayAndOrderLasecki}.
+We should consider whether this remains a worthwhile area to explore; %  unless
+% we're actually finding ways to minimize the many forms of costs of switching tasks.
+unless the work we put forth directly affects the costs of task--switching
+--- for instance, the cost of suboptimal task search, or the cognitive burden of changing tasks ---
+we may only make incremental advances in micro--task decomposition.
 
-\end{enumerate}
+Finally, we turn to the relationships of crowd workers.
+The crowd work literature here can convincingly speak back to
+the piecework scholarship perhaps more than in the other sections.
+The tools that are available to us today
+--- not just technical, but \textit{methodological} ---
+make it possible to
+discover,
+study, and
+empower \ali{I feel like Lily would take some umbrage at this because she seemed to not like the narrative of us researchers just unilaterally bequeathing power upon Turkers (well, Turkers certainly disliked it).
+Thoughts on refactoring?}
+crowd workers in ways that were unimaginable to piecework researchers.
+A professor engages in crowd work \cite{bighamHalfWorkday}
+not just because it's possible, but because our community
+(of mostly computer scientists)
+appreciates the importance of approaches such as participant--observation and ethnography as a whole
+\cite{olson2014ways}.
+% Whereas piecework didn't have the benefit of Anthropology as we know it or the myriad ways of discovering, learning about, and empowering workers
+% What questions are we trying to answer?
+% If, as we assumed in
+% the previous section on \namerefl{sec:polarizationOfCrowdWork},
+% our driving motivation is to empower workers, we
+
+% \end{enumerate}
 

### 2016-09-21 drafts/labor/piecework_lit.tex

@@
 
+Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition. 
+Platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well. 
+% Second, workers on these platforms have significant freedom in choosing which tasks to work on, and when to log on and perform work~\cite{martin2014being,whyWouldAnyoneBrewer}. 
+% Finally, computational workflows have successfully scaffolded many complex tasks into a series of small microtasks that can be performed without expertise.
+%  \subsection{Piecework and crowd work}
+% \msb{I assume this is forthcoming?}
+% Using the definition of piecework that we came up with earlier, we argue that
+% crowd work is fundamentally an instantiation of piecework, and
+% that we can more precisely anticipate the answers to the open research questions we discussed earlier.
+% We'll show that the dimensions of crowd work that the broader HCI community has been studying
+% align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
+% \subsection{From piecework to on--demand work}
+% Crowd work and gig work are fundamentally an instantion of piecework.
+% First, workers on platforms such as Mechanical Turk and Uber are generally incentivized by unit of work, even if some may be offered an hourly base salary as well.
+Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
+Finally, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+
+Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
+In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.
+
 % \ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}
@@
 
-\subsection{Piecework and crowd work}
-
-\msb{I assume this is forthcoming?}
-Using the definition of piecework that we came up with earlier, we argue that
-crowd work is fundamentally an instantiation of piecework, and
-that we can more precisely anticipate the answers to the open research questions we discussed earlier.
-We'll show that the dimensions of crowd work that the broader HCI community has been studying
-align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
-\subsection{From piecework to on--demand work}
-Crowd work and gig work are fundamentally an instantion of piecework.
-First, workers on platforms such as Mechanical Turk and Uber are generally incentivized by unit of work, even if some may be offered an hourly base salary as well.
-Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
-
-Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
-In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.
-
 \onlyinsubfile{

### 2016-09-21 drafts/labor/piecework_lit.tex
compilation fix

@@
 mass manufacturing and
-a depleted wartime workforce forced industry to find new ways to eke out more production capacity.}
+a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
 \citeauthor{hart2013rise} point out that the Second World War,

### 2016-09-21 drafts/labor/complexity.tex

@@
 there is no general solution to create
-a cross-domain, high--quality crowd--powered author. 
+a cross--domain, high--quality crowd--powered author. 
 Third,
@@
 for example designers and programmers.
-The same ideas can then be applied to expert ``macro-tasks''
+The same ideas can then be applied to expert ``macro--tasks''
 \cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
@@
 then a piecework system is likely to encourage
-an over--emphasis on quantity produced and an under-emphasis on quality''
+an over--emphasis on quantity produced and an under--emphasis on quality''
 \cite{hart2016rise}.
@@
 particularly because maximizing the advantages of piecework seemed to reward
-smaller, more constrained, more narrowly-trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
+smaller, more constrained, more narrowly--trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
 For most of the \nth{19} century,
@@
 Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
-Beyond this, further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
+Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
 of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
@@
 
-Will the shift from human managers to Turing-complete algorithms raise the complexity ceiling? 
+Will the shift from human managers to Turing--complete algorithms raise the complexity ceiling? 
 By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
@@
 So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
-However, as a resource constraint, algorithms are a fixed cost and not a per-person cost like human managers.
+However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
 So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford crowd work.
-This shift may enable complex goals that were not cost-effective before to become feasible.
-However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re-introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
+This shift may enable complex goals that were not cost--effective before to become feasible.
+However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re--introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
 This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.

### 2016-09-21 drafts/labor/decomposition.tex

@@
 % \onlyinsubfile{
-% \usepackage{xr-hyper}
+% \usepackage{xr--hyper}
 % \usepackage{hyperref}
@@
 For example, workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017}, interpret tasks in different ways~\cite{kairam2016parting}, and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
-Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher-level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.
+Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher--level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.
 
@@
 affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter''
---- an icon of mid--\nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating} ---
-was herself a pieceworker
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating},
+was a pieceworker
 \cite{davies2014origins}.
@@
 \cite{grier2013computers}.
-%These workers became the first \textit{computers}.
-
 The work that Airy solicited resonates with modern crowd work for several reasons.
@@
 Second, tasks were discrete --- that is, independent from one another.
-Finally, knowledge of the full scope of the project
---- indeed, knowledge of anything more than the problem set at hand ---
-was unnecessary.
+Finally, workers could be trained on a very narrow subset of mathematical skills to be
+sufficiently qualified to do this work.
 
@@
 %\cite{10.2307/23702539}.}
-modern crowd work is measuring and modeling every click, scroll, and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
+modern crowd work is measuring and modeling every click, scroll, and keyboard event
+\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
 The result is that on--demand work can articulate and track far more carefully than piecework historically could.
@@
 \topic{A second shift is  
-%A number of factors in crowd work are different from piecework, chief among them being 
 the relative ease with which the metaphorical ``assembly line'' can be changed.}
-% Computers make it possible to switch from one task to another unlike any arbitrary manufacturing factory possibly could;
-%modern on--demand workers on AMT and other platforms can more readily do any number of
-%different \textit{types} of tasks in the span of just a few minutes
-Historical manufacturing equipment could not quickly be assembled, edited, and redeployed.\msb{cite for historical piecework equipment being a large time + money investment?}
+Historical manufacturing equipment could not quickly be assembled, edited, and redeployed
+\cite{hu1961parallel}.
 In contrast, today system--designers can share, modify, and instantiate environments
@@
 --- things like finding work and cognitive task switching ---
-have become relatively large compared to the tasks themselves
+have become large relative to the tasks themselves
 \cite{taskSearch}.
@@
 The result is that crowd workers are more free agents than historically was the case.
-However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take-home earnings than before.
-
-% Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
-% --- one which where crowd workers are rendered as
-% ``modular, protocol--defined computational services'' ---
-% we may yield better results from crowd work if we think of workers as similar to
-% specialized, repurposable tools
-% \cite{irani2015cultural}.
-% \ali{feeling meh about this argument\dots}
+However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take--home earnings than before.
 
 \subsubsection{\implication}
-If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely-sliced and highly decomposed than ever before.
+If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
 Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved, or near enough that they no longer limit task decomposition.
 Now, not just tasks, but entire workers' histories~\cite{hata2017glimpse}, can be collected and analyzed in detail.
-% crowd work research today is on the right track to investigate pipelining and meta--task design.
-% That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
-% \cite[see, for example,][]{turkopticon}.
 
@@
 Task switching costs and other cognitive costs make it difficult to work tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
-There will of course be tasks that can be decomposed without much context, and these will form the most fine-grained of microtasks.
+There will of course be tasks that can be decomposed without much context, and these will form the most fine--grained of microtasks.
 However, other tasks cannot be freed from context --- for example, logo design requires a deep understanding of the client and their goals.
@@
 
-So, ultimately, the levels of decomposition are likely to follow the contours of context required. Low-context work will be extremely highly decomposed. High-context work will continue to be limited.
+So, ultimately, the levels of decomposition are likely to follow the contours of context required.
+Low--context work will be extremely highly decomposed.
+High--context work will continue to be limited.
 

### 2016-09-21 drafts/labor/trackingWork.tex
renamed research_implications.tex to discussion.tex
added a conclusion.tex file. finished with discussion (i think)
maybe i should also let the trackingWork PDF finish compiling
conclusion

@@
   colorlinks,
-  citecolor=blue, % black,
-  filecolor=blue, % black,
-  linkcolor=blue, % black,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=black,
   urlcolor=linkColor,
@@
 
-\DTMsetdatestyle{mmddyyyy}
-\DTMsetup{datesep=/}
-\pagestyle{fancy}
-\lfoot{Last compiled \today~at \DTMcurrenttime}
-\cfoot{}
-\rfoot{page \thepage/\pageref*{LastPage}}
-\fancyhead{}
-\pagenumbering{arabic}
+% \DTMsetdatestyle{mmddyyyy}
+% \DTMsetup{datesep=/}
+% \pagestyle{fancy}
+% \lfoot{Last compiled \today~at \DTMcurrenttime}
+% \cfoot{}
+% \rfoot{page \thepage/\pageref*{LastPage}}
+% \fancyhead{}
+% \pagenumbering{arabic}
 \renewcommand{\headrulewidth}{0pt}
@@
 
+\renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
 \renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
@@
 \subfile{cases.tex}
-\subfile{research_implications.tex}
+\subfile{discussion.tex}
+\subfile{conclusion.tex}
 
@@
 }
-\theendnotes
+% \theendnotes
 

### 2016-09-21 drafts/labor/cases.tex
added a ref shortname
citation fixes

--- 2016-09-21 drafts/labor/cases.tex
@@
 
-\subsection{What are the limits of crowdsourcing?}
+%\subsection{What are the limits of crowdsourcing?}
 \topic{Research in crowdsourcing has spent the better part of a decade
@@
 ``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',  parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work
-\cite{hu1961parallel}.
+--- namely, using non--expert crowds for complex work~\cite{hu1961parallel}.
 This mindset in \citeauthor{hu1961parallel}'s analysis,  and indeed the study of factory and mass manufacturing labor through the \nth{20} century,  substantively owes its existence to scientific management
@@
 involving labor either indoors or out--of--doors
-where tasks cannot be fixed by proper time--study''
-\cite{thompson1913time}.
+where tasks cannot be fixed by proper time--study''~\cite{thompson1913time}.
 
-Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable
-\cite{murray1983decentralisation}.
+Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable~\cite{murray1983decentralisation}.
 Textile work being a salient example,  it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work
-\cite{vezina1992light}.
+to do every aspect of the work so that the garment would be accepted by the company soliciting that work~\cite{vezina1992light}.
 
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work
-\cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car
-\cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''
-\cite{schoenberger1988fordism}.
+A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work~\cite{restructuringPieceworkBaker}.
+In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car~\cite{towardsGlobalFordism}.
+By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''~\cite{schoenberger1988fordism}.
 
 This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),  can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization
-\cite{writingMicroTasks}.
-By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically
-\cite{delayAndOrderLasecki}.
+\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization~\cite{writingMicroTasks}.
+By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically~\cite{delayAndOrderLasecki}.
 Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage
-\cite{rogstadius2011assessment}.
+the breaking down of work into tasks has made it more practical to evaluate work at each stage~\cite{rogstadius2011assessment}.
 

### 2016-09-21 drafts/labor/complexity.tex
citation fixes

@@
 the argument runs,
-it will be incapable of ensuring a pro--social outcome for work and workers
-\cite{crowdworkFuture}.
+it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
 \citeauthor{crowdForgeKittur} first opened the question of
-whether crowdsourcing could be used for goals that are not simple parallel tasks
-\cite{crowdForgeKittur}.
+whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
 Their work demonstrated proof--of--concept crowdsourcing of
@@
 with reasonable expectations of similar outcomes.
-Seeking to raise the complexity ceiling
-\cite{myers2000past},
+Seeking to raise the complexity ceiling~\cite{myers2000past},
 researchers have since created
@@
 the researcher leverages an insight from Computer Science
-(for example, MapReduce
-\cite{crowdForgeKittur} or
-sequence alignment algorithms
-\cite{lasecki2012real})
+(for example, MapReduce~\cite{crowdForgeKittur} or
+sequence alignment algorithms~\cite{lasecki2012real})
 and arranges humans as computational black boxes within those approaches.
@@
 It is now clear that this approach works with focused complex tasks, but
-the broader wicked problem that each complex goal represents remains unsolved
-\cite{rittel1973dilemmas}.
+the broader wicked problem that each complex goal represents remains unsolved~\cite{rittel1973dilemmas}.
 As a first example,
-idea generation shows promise
-\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
 but there is as yet no general crowdsourced solution for
-the broader goal of invention and innovation
-\cite{fuge2014analysis}.
+the broader goal of invention and innovation~\cite{fuge2014analysis}.
 Second,
-focused writing tasks are now feasible
-\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
+focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
       writingMicroTasks,agapie2015crowdsourcing}, but
@@
 for example designers and programmers.
-The same ideas can then be applied to expert ``macro--tasks''
-\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
+The same ideas can then be applied to expert ``macro--tasks''~\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
 programming \cite{latoza2014microtask,Fast2016,Chen2016}, and
@@
 a piecework strategy in Airy's creation of
-the British Nautical Almanac
-\cite{grier2013computers}.}
+the British Nautical Almanac~\cite{grier2013computers}.}
 Airy's goal was complex
@@
 the practice remained relatively obscure until
-it blossomed in the textile industry
-\cite{hughRaynbirdTaskWork}.
+it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
 Complexity levels remained low at the turn of
-the \nth{20} century as piecework saturated New York City
-\cite{riisOtherSideLives}.
+the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
 However, writers of the time focused their attention on
@@
 then a piecework system is likely to encourage
-an over--emphasis on quantity produced and an under--emphasis on quality''
-\cite{hart2016rise}.
+an over--emphasis on quantity produced and an under--emphasis on quality''~\cite{hart2016rise}.
 Complex work, which is often subjective to evaluate, falls victim to this criteria.
@@
 the rigorous tracking, measurement, and training of workers
-for which scientific management became famous
-\cite{10.2307/23702539}.
+for which scientific management became famous~\cite{10.2307/23702539}.
 If true, 
@@
 through the first two--thirds of the \nth{20} century
-alongside piecework was not only understandable, but predictable
-\cite{hart2013rise}.
+alongside piecework was not only understandable, but predictable~\cite{hart2013rise}.
 
@@
 piecework ``is less likely in jobs with a variety of duties than in
-jobs with a narrow set of routinized duties''
-\cite{Brown01041990}.
+jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
 \citeauthor{SJOE:SJOE371} points out the phenomenon here as a market effect:
@@
 pay schemes based on tightly specified performance may
-induce workers to neglect tasks that are less easy to measure''
-\cite{SJOE:SJOE371}.
+induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
 Second,
@@
 These foremen were responsible for allocating resources and
-understanding when and how to modify work as necessary
-\cite{wray1949marginal}.
+understanding when and how to modify work as necessary~\cite{wray1949marginal}.
 So, in sum,
@@
 --- farm workers didn't need to be trained on how to plow fields, or birth foals;
-seamstresses knew how to sew together denim
-\cite{10.2307/2338394,riisOtherSideLives}.
+seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.
 
@@
 we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity
-\cite{Brown01041990}.
+the nature of the work is limited in complexity~\cite{Brown01041990}.
 Given the flourishing of on--demand labor platforms such as
@@
 This test is so challenging that veteran drivers exhibit significantly larger
-the regions of the brain associated with spatial functions such as navigation
-\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
+the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
       skok2000managing,Woollett1407,woollett2011acquiring}.
@@
 it remains to be seen whether that work will be appropriately valued, let alone
-properly interpreted by task solicitors
-\cite{aguaded2013mooc}.
+properly interpreted by task solicitors~\cite{aguaded2013mooc}.
 If we can overcome this obstacle,

### 2016-09-21 drafts/labor/decomposition.tex
citation fixes

@@
 Moreover, how thinly can work be sliced and subdivided into smaller and smaller tasks? 
-The general trend has been that smaller is better, and the microtask paradigm has emerged as the overwhelming favorite
-\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+The general trend has been that smaller is better, and the microtask paradigm has emerged as the overwhelming favorite~\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
 This work illustrates a broader sentiment in both the study and practice
@@
 %Earlier we discussed \citeauthor{cheng2015break}'s work measuring the impact that interruption has on worker performance \cite{cheng2015break}.%While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components in the form of higher cumulative time to complete (albeit much shorter real time to complete, owing to parallelization)%
-However, workers perform better when similar tasks are strung together 
-\cite{delayAndOrderLasecki}, or chained and arranged to maximize  the attention threshold of workers~\cite{Cai:2016:CRI:2858036.2858237}.
+However, workers perform better when similar tasks are strung together ~\cite{delayAndOrderLasecki}, or chained and arranged to maximize  the attention threshold of workers~\cite{Cai:2016:CRI:2858036.2858237}.
 Despite this, 
@@
 (specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
-were the major contributing factors that determined the use of piecework
-\cite{Brown01041990}.
+were the major contributing factors that determined the use of piecework~\cite{Brown01041990}.
 
@@
 an icon of \nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating},
-was a pieceworker
-\cite{davies2014origins}.
+was a pieceworker~\cite{davies2014origins}.
 
@@
 `Arithmetic, the use of Logarithms, and Elementary Algebra'~''
-to compute astronomical phenomena
-\cite{grier2013computers}.
+to compute astronomical phenomena~\cite{grier2013computers}.
 The work that Airy solicited resonates with modern crowd work for several reasons.
@@
 ``[piecework is] infeasible, because \dots
-total output is the joint product of varying groups of people''
-\cite{bewley1999wages}.
+total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
 
@@
 --- to measure every motion at every point throughout the workday and beyond ---
-is not only doable, but trivial
-\cite{waltz2012quantified}.}
+is not only doable, but trivial~\cite{waltz2012quantified}.}
 Where \citeauthor{10.2307/23702539} directly implicates measurement as
@@
 %\cite{10.2307/23702539}.}
-modern crowd work is measuring and modeling every click, scroll, and keyboard event
-\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
+modern crowd work is measuring and modeling every click, scroll, and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
 The result is that on--demand work can articulate and track far more carefully than piecework historically could.
@@
 the relative ease with which the metaphorical ``assembly line'' can be changed.}
-Historical manufacturing equipment could not quickly be assembled, edited, and redeployed
-\cite{hu1961parallel}.
+Historical manufacturing equipment could not quickly be assembled, edited, and redeployed~\cite{hu1961parallel}.
 In contrast, today system--designers can share, modify, and instantiate environments
-like sites of labor in a few lines of code
-\cite{lessig2006code,turkitLittle}.
+like sites of labor in a few lines of code~\cite{lessig2006code,turkitLittle}.
 This opportunity has spurred an entire body of work investigating the effects of
@@
 other factors in piecework that would have been
-all but impossible to manipulate as few as 20 years ago
-\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
+all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
 
@@
 --- things like finding work and cognitive task switching ---
-have become large relative to the tasks themselves
-\cite{taskSearch}.
+have become large relative to the tasks themselves~\cite{taskSearch}.
 In the historical case of piecework,
@@
 workers were encouraged to specialize in a single set of tasks,
-allowing pieceworkers to sequence their tasks optimally on their own
-\cite{hart2013rise}.
+allowing pieceworkers to sequence their tasks optimally on their own~\cite{hart2013rise}.
 The result is that crowd workers are more free agents than historically was the case.

### 2016-09-21 drafts/labor/discussion.tex
citation fixes

@@
 in the realm of digital labor
-there is nothing new under the sun''
-\cite{scholz2012digital}.
+there is nothing new under the sun''~\cite{scholz2012digital}.
 Many of the challenges that \DO~overcame
---- determining trustworthiness, ensuring anonymity, etc\dots
-\cite{dynamo} ---
+--- determining trustworthiness, ensuring anonymity, etc\dots~\cite{dynamo} ---
 were relatively unique challenges precipitated by the affordances of the Internet, specifically
-that people could (and often do) contribute to communities in a one--off manner
-\cite{mcinnis2016one}.
-The Internet seemed to make this kind of loose affiliation feasible where before it wasn't
-\cite{catalyst,dynamo}.
+that people could (and often do) contribute to communities in a one--off manner~\cite{mcinnis2016one}.
+The Internet seemed to make this kind of loose affiliation feasible where before it wasn't~\cite{catalyst,dynamo}.
 % articulated in the discussion about
@@
 imagine the application of crowdsourcing as
-a potentially bright future that enables the achievement of near--impossible goals
-\cite{redballoon,crowdworkFuture,vizwiz}; % \ali{and who else should I cite here?}
+a potentially bright future that enables the achievement of near--impossible goals~\cite{redballoon,crowdworkFuture,vizwiz}; % \ali{and who else should I cite here?}
 on the other hand, researchers
-warn against crowd work as potentially exploitative sites of dispossession \cite{scholz2012digital},
-a site of racial discrimination \cite{edelman2015racial},
-and invisible, deeply frustrated workers \cite{turkopticon,bighamHalfWorkday}.
+warn against crowd work
+as potentially exploitative sites
+of dispossession \cite{scholz2012digital},
+of racial discrimination \cite{edelman2015racial}, and
+of invisible, deeply frustrated workers \cite{turkopticon,bighamHalfWorkday}.
 
@@
 Activists have described speculative work as having
-``essentially been turned into modern--day slaves''
-\cite{activistsHuffPoLawsuit}.
+``essentially been turned into modern--day slaves''~\cite{activistsHuffPoLawsuit}.
 Meanwhile, advocates describe it as
@@
 with more economic opportunities and
-improving their lives\dots''
-\cite{uberPropaganda}.
+improving their lives\dots''~\cite{uberPropaganda}.
 \ali{these are salient, but with the last paragraph it feels really flimsy.
@@
 just like those which already frustrate
-on--demand workers
-\cite{uberAlgorithm,dynamo,turkopticon}.
+on--demand workers~\cite{uberAlgorithm,dynamo,turkopticon}.
 
@@
 Yes, piecework's nascent years were arguably grim, but
-they precipitated a century of the strongest labor advocacy the world has ever seen
-\cite{hart2013rise,mccallum2013global}.
+they precipitated a century of the strongest labor advocacy the world has ever seen~\cite{hart2013rise,mccallum2013global}.
 Even today, the geist that came out of the labor union revolution
@@
 the largest labor strike in human history
---- perhaps as many as 150 million
-\cite{indiaStrikeRealNews}.
+--- perhaps as many as 150 million~\cite{indiaStrikeRealNews}.
 If labor advocacy groups can find ways to permeate on--demand labor markets,
-as some have called for
-\cite{futureUnions},
+as some have called for~\cite{futureUnions},
 then the future of crowd work may follow
@@
 piecework is saddled by a lower limit on decomposition:
-``piecework does not compensate workers for time spent switching tasks''
-\cite{bewley1999wages}.
+``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}.
 We've since studied this phenomenon in crowd work to great length both
@@
 (of mostly computer scientists)
-appreciates the importance of approaches such as participant--observation and ethnography as a whole
-\cite{olson2014ways}.
+appreciates the importance of approaches such as participant--observation and ethnography as a whole~\cite{olson2014ways}.
 % Whereas piecework didn't have the benefit of Anthropology as we know it or the myriad ways of discovering, learning about, and empowering workers

### 2016-09-21 drafts/labor/introduction.tex
relationships.tex
added a conclusion.tex file. finished with discussion (i think)
citation fixes

@@
 This framing of work into de--contextualized, modular blocks
-enables computation to hire workers at scale through open calls on the Internet
-\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+enables computation to hire workers at scale through open calls on the Internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
 Distributed paid participants then engage in the work whenever their schedules allow,
 with little to no awareness of the broader context of the work, and 
-with (often) fleeting identities and associations
-\cite{martin2014being,uberAlgorithm}.
+with (often) fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
 In this paper, we use the term on--demand work to join a pair of related phenomena:
@@
 \end{inlinelist}
-The realization that complex tasks can be accomplished by directing and managing these crowds of workers spurred industry to flock to sites of labor
-like AMT and Uber to explore the limits of this distributed, on--demand workforce.
+The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
+like AMT to explore the limits of this distributed, on--demand workforce.
 Researchers have also taken to the space in earnest,
@@
 \topic{As on--demand work has grown far beyond the domain of information work from which it first sprang, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who empower it.}
-Originally, \citeauthor{howe2008crowdsourcing} described crowdsourcing in general terms as
+\citeauthor{howe2008crowdsourcing} first described crowdsourcing in general terms as
 ``outsourcing [work] to an undefined, generally large group of people in the form of an open call''.
 However, for years its instantiation was limited to the utilization of
-human intelligence to process data, participate in scientific studies, and perform information work
-\cite{CrowdsourcingUserStudies,movieSummarizationWu,
+human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,
       yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
@@
 --- driving and cleaning, for instance ---
-has become a focus for on--demand labor markets
-\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-This growth prompted increasing efforts to understand the workers who gravitate toward these platforms~\cite{Ross,whoareNOTtheTurkers}.
+has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+This growth prompted efforts to understand the workers who gravitate toward these platforms~\cite{Ross,whoareNOTtheTurkers}.
 Some of this research has been motivated by the identification of the sociality of gig work,
-and the frustration and disenfranchisement that these systems embody
-\cite{turkopticon,martin2014being,takingAHITMcInnis}.
+and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
 Other work has focused on the \textit{outcomes} of this frustration,
-reflecting on the resistance workers express against digitally mediated labor markets
-\cite{uberAlgorithm,dynamo}.
+reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.
 
@@
         Specifically,
-        \begin{inlinelist}
+        \begin{inlinelist}[label=(\alph*)]
           \item How complex are the goals that crowd work can accomplish?, and
@@
 referenced crowd work as ``piecework'' briefly
-as a loose analogy to the form of work emerging at the time
-\cite{crowdworkFuture}.
+as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
 But more than this,
@@
 Then, we interrogate the three major research questions above from a piecework frame.
-We will
 % We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
@@
 %After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
-we will identify similarities and differences between piecework as historically understood and
+We will identify similarities and differences between piecework as historically understood and
 on--demand work as we experience it today.

### 2016-09-21 drafts/labor/relationships.tex
citation fixes

@@
 has been part of the undercurrent of research into
-the tense relationships between workers and requesters
-\cite{fixingChaos,dynamo}.
+the tense relationships between workers and requesters~\cite{fixingChaos,dynamo}.
 
@@
 This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
-The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
-\cite{jacoby1983union}.
+The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally~\cite{jacoby1983union}.
 
@@
 As mounting attention increasingly revealed problems in piecework's treatment of workers, theworkers themselves began to speak out about their frustration with this new regime.
-Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers
-\cite{american1921problem,richards1904anything}.
+Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers~\cite{american1921problem,richards1904anything}.
 
@@
 % to identify
-``not only as railroad employees but also as members of the larger life of the community''
-\cite{american1921problem}.
+``not only as railroad employees but also as members of the larger life of the community''~\cite{american1921problem}.
 Doing this, 
 \citeauthor{ostrom1990governing} and others argued,
-would facilitate collective action and perhaps collective governance
-\cite{ostrom1990governing,russell1982collective,olsonlogic}.
+would facilitate collective action and perhaps collective governance~\cite{ostrom1990governing,russell1982collective,olsonlogic}.
 \citeauthor{riisOtherSideLives} also contributed to this sense of shared struggle and endurance
@@
 by documenting pieceworkers in their home--workplaces,
-literally bringing to light the grim circumstances in which pieceworkers lived and worked
-\cite{riisOtherSideLives}.
+literally bringing to light the grim circumstances in which pieceworkers lived and worked~\cite{riisOtherSideLives}.
 
@@
 the foreman of the \nth{20} century has largely been replaced
-by algorithms of the \nth{21} century
-\cite{uberAlgorithm}.}
+by algorithms of the \nth{21} century~\cite{uberAlgorithm}.}
 The result of this change is that the agents managing work are now
@@
 a manager's subjective judgment are preferable to
-defective (simple, observable) output measures''
-\cite{10.2307/2555446}.
+defective (simple, observable) output measures''~\cite{10.2307/2555446}.
 This frustration has only grown as requesters have had to rely on automatic management mechanisms. Only a few use the equivalent of human foremen~\cite{haas2015argonaut,kulkarni2012mobileworks}.
@@
 \textit{reflexivity}
-didn't take even a resemblance of their contemporary forms until these works
-\cite{malinowski2002argonauts,boas1940race,mead1973coming}.
+didn't take even a resemblance of their contemporary forms until these works~\cite{malinowski2002argonauts,boas1940race,mead1973coming}.
 Modern crowd work may give us an opportunity to revisit open questions in piecework with a more refined lens.

### 2016-09-21 drafts/labor/discussion.tex
revised discussion

--- 2016-09-21 drafts/labor/discussion.tex
@@
 \section{Discussion}
-\topic{Having taken a comprehensive look toward crowd work through the piecework lens,
-we can't help but take a step back to consider a number of meta--issues that arose in our analysis.}
-Stated briefly, these issues are 
+\topic{In our analysis of crowd work via the piecework lens, three issues arose:}
 \begin{inlinelist}
@@
 \end{inlinelist}
-We will attempt to grapple with these questions here based on
-what we brought up in the earlier case studies.
+We will attempt to grapple with these questions here explicitly.
 
@@
 \cite{scholz2012digital}.
-Many of the challenges that \DO~overcame
---- determining trustworthiness, ensuring anonymity, etc\dots
-\cite{dynamo} ---
-were relatively unique challenges precipitated by the affordances of the Internet, specifically
-that people could (and often do) contribute to communities in a one--off manner
+Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern crowd work.
+For example, many of the challenges that \DO~overcame in crowd collective action, such as designing for trustworthiness and ensuring anonymity, were relatively unique challenges precipitated by the affordances of the internet.
+For example, unlike physical work environments, people can (and often do) contribute to online communities in a one--off manner
 \cite{mcinnis2016one}.
-The Internet seemed to make this kind of loose affiliation feasible where before it wasn't
-\cite{catalyst,dynamo}.
+The Internet makes this kind of loose affiliation feasible.
+While we have attempted to understand the likely overlaps and differences between history and modern day, no analysis is perfect.
 % articulated in the discussion about
@@
 attempting to draw meaningfully from historical scholarship would be folly;
-as we have shown in the preceding sections,
 enough of piecework can and does inform crowdsourcing that
-we should take this as a cue to look first for
-historical framings on phenomena that we encounter,
-as this work can outline both the mechanisms that we should expect to play out
-as well as one of the (perhaps many) possible futures.
+% we should take this as a cue to look first for
+HCI and CSCW might seek out historical framings for other phenomena of study as well.
+% as this work can outline both the mechanisms that we should expect to play out
+While we can only speculate one of (perhaps many) possible futures, history does allow us to articulate and bound which futures appear more likely.
 
 In particular, the predictions that have emerged surrounding crowd work have run the spectrum
-from deep pessimism to exuberant optimism. \ali{should I cite this?}
+from deep pessimism~\cite{fort2011amazon} to exuberant optimism~\cite{crowdworkFuture}.
 In the next section, we will use
 the piecework foundation which informed our case studies
-to trace out possible Dystopian and Utopian futures for crowd work.
+to trace out possible dystopian and utopian futures for crowd work.
 
-\subsection{Polarizing Tendencies}\label{sec:polarizationOfCrowdWork}
-\topic{It would be easy
-to think about the future of crowd work and
-end up at one of two extremes.}
+\subsection{Utopian and Dystopian Visions}\label{sec:polarizationOfCrowdWork}
+\topic{An easy narrative is to characterize the future of crowd work  at one of two extremes.}
 On one hand, crowd work researchers
 imagine the application of crowdsourcing as
-a potentially bright future that enables the achievement of near--impossible goals
-\cite{redballoon,crowdworkFuture,vizwiz}; % \ali{and who else should I cite here?}
-on the other hand, researchers
-warn against crowd work as potentially exploitative sites of dispossession \cite{scholz2012digital},
-a site of racial discrimination \cite{edelman2015racial},
-and invisible, deeply frustrated workers \cite{turkopticon,bighamHalfWorkday}.
+a potentially bright future that enables the achievement of near--impossible goals and career opportunities~\cite{redballoon,crowdworkFuture,vizwiz,suzukiAtelier}.
+% \ali{and who else should I cite here?}
+On the other hand, researchers
+warn that crowd work will create exploitative sites of dispossession~\cite{scholz2012digital},
+racial discrimination~\cite{edelman2015racial},
+and invisible, deeply frustrated workers~\cite{turkopticon,bighamHalfWorkday}.
 
@@
 with more economic opportunities and
-improving their lives\dots''
-\cite{uberPropaganda}.
-\ali{these are salient, but with the last paragraph it feels really flimsy.
-I'd really like to get this Uber quote in, but
-not if it means this paragraph gets that much weaker.}
+improving their lives''~\cite{uberPropaganda}.
 
 % \topic{There is, of course, a kernel of truth in everything here.}
-\topic{There is evidence to support the claim that
-a Dystopian world awaits us at the end of the tunnel.}
-The arbitrary nature of on--demand work may permeate our lives, inducing us
-``to neglect tasks that are less easy to measure'' \cite{SJOE:SJOE371}
+\topic{Piecework teaches us that, without appropriate norms and policy, the dystopian outcome has happened and will happen again.}
+The piecework nature of on--demand work induces us
+``to neglect tasks that are less easy to measure'' \cite{SJOE:SJOE371},
 rewarding us not for creativity but predictability;
 payment for this work may ultimately be determined by
-an algorithm that fundamentally doesn't understand us;
+an algorithm that fundamentally doesn't understand people;
 the layers between us and our managers might increasingly become
@@
 \cite{uberAlgorithm,dynamo,turkopticon}.
+However, social policy has advanced since the early 1900s, so as crowd work gains popularity a repeat of \textit{How the Other Half Lives}~\cite{riisOtherSideLives} seems less likely.
 
-\topic{On the other hand, there is some evidence that offers hope.}
-Yes, piecework's nascent years were arguably grim, but
-they precipitated a century of the strongest labor advocacy the world has ever seen
-\cite{hart2013rise,mccallum2013global}.
+\topic{On the other hand, while piecework's nascent years were grim, 
+they precipitated a century of extremely strong labor advocacy~\cite{hart2013rise,mccallum2013global}.}
 Even today, the geist that came out of the labor union revolution
 inspires collective action and worker empowerment around the world:
-in India, workers across the nation engaged in
+in India, workers across the nation recently engaged in
 the largest labor strike in human history
@@
 \cite{indiaStrikeRealNews}.
-If labor advocacy groups can find ways to permeate on--demand labor markets,
-as some have called for
+If labor advocacy groups can find ways to permeate on--demand labor markets as some have called for
 \cite{futureUnions},
@@
 the same trajectory of worker empowerment that piecework \textit{later} found.
-\ali{The Internet may make this possible!}
+
 % Arguably the Internet has made collective action on an unprecedented scale much easier.
 
-\topic{Perhaps the most pressing question is this:
-is the future of crowd work going to look more \textit{Dystopian} or \textit{Utopian}?}
-If we have learn nothing, crowd work will probably end up somewhere in the middle.
-It's possible that the difficulties of enforcing laws on multinational corporations is
-tipping the balance of power in the direction of corporations, in which case
-things will get worse;
-but arguably we can affect change on the trajectory of crowd work,
-benefiting from everything piecework scholars have learned about
-collective action and governance
-among workers, while also avoiding some of the perils they faced.
-\ali{I really don't like this paragraph, but I felt like we needed to plant a flag on the question in the topic sentence}
+\topic{The history of piecework suggests that the utopian and dystopian outcomes will \textit{both} occur, in different parts of the world and to different people.}
+When piecework plummeted in the United States, outsourcing rose --- creating major labor issues around the world.
+It is entirely possible that we will create a new brand of flexible online career in developed countries, while simultaneously fueling an unskilled decentralized labor force in developing nations.
+As designers and researchers, this prompts the question: which outcome are we attempting to promote or avoid for who?
 
-\subsection{Determining our Research Agenda}\label{sec:whatShouldBeTheFuture}
-Taking a step back from the work that's been done so far and
-thinking about where the body of crowd work research is attempting to go may help us identify
-areas of research on which we should be pushing forward on.
+% \topic{Perhaps the most pressing question is this:
+% is the future of crowd work going to look more \textit{Dystopian} or \textit{Utopian}?}
+% If we have learn nothing, crowd work will probably end up somewhere in the middle.
+% It's possible that the difficulties of enforcing laws on multinational corporations is
+% tipping the balance of power in the direction of corporations, in which case
+% things will get worse;
+% but arguably we can affect change on the trajectory of crowd work,
+% benefiting from everything piecework scholars have learned about
+% collective action and governance
+% among workers, while also avoiding some of the perils they faced.
+% \ali{I really don't like this paragraph, but I felt like we needed to plant a flag on the question in the topic sentence}
+
+\subsection{A Research Agenda}\label{sec:whatShouldBeTheFuture}
+Piecework also helps bring into focus the
+% Taking a step back from the work that's been done so far and
+% thinking about where the body of crowd work research is attempting to go may help us identify
+areas of research that might bear the most fruit.
 We return to the three questions that motivated this paper:
@@
 While we have arguably outpaced piecework with regard to the limits on the complexity of work,
-we have yet to find insurmountable limits to crowdsourcing.
-Still, we can learn from the piecework literature as it relates to
+the most complex and open-ended wicked problems~\cite{rittel1973dilemmas} remain the domain of older human collectives such as governments and organizations. 
+% we have yet to find insurmountable limits to crowdsourcing.
+In addition, we can learn from the piecework literature as it relates to
 the stymieing effect that mismanagement has on workers;
@@
 we may only make incremental advances in micro--task decomposition.
+When the cognitive cost of understanding a task and its inputs outstrips the effort required to complete the task, decomposition seems a poor choice.
 
@@
 study, and
-empower \ali{I feel like Lily would take some umbrage at this because she seemed to not like the narrative of us researchers just unilaterally bequeathing power upon Turkers (well, Turkers certainly disliked it).
+partner with 
+% \ali{I feel like Lily would take some umbrage at this because she seemed to not like the narrative of us researchers just unilaterally bequeathing power upon Turkers (well, Turkers certainly disliked it).
 Thoughts on refactoring?}
 crowd workers in ways that were unimaginable to piecework researchers.
-A professor engages in crowd work \cite{bighamHalfWorkday}
+A professor engages in crowd work~\cite{bighamHalfWorkday}
 not just because it's possible, but because our community
-(of mostly computer scientists)
-appreciates the importance of approaches such as participant--observation and ethnography as a whole
-\cite{olson2014ways}.
+% (of mostly computer scientists)
+appreciates the importance of approaches such as participant--observation and ethnography as a whole~\cite{olson2014ways}.
 % Whereas piecework didn't have the benefit of Anthropology as we know it or the myriad ways of discovering, learning about, and empowering workers

### 2016-09-21 drafts/labor/conclusion.tex

@@
 
+\onlyinsubfile{
+\usepackage{xr-hyper}
+\usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{decomposition}
+\externaldocument{relationships}
+}
+
 \begin{document}
 \section{Conclusion}
-????
+In this paper we considered three overarching questions in crowd work
+--- \begin{inlinelist}
+  \item ``\namerefl{sec:complexity}?'';
+  \item ``\namerefl{sec:decomposition}?'';
+        and
+  \item ``\namerefl{sec:relationships}?''
+\end{inlinelist} ---
+by using the scholarship on piecework to inform our predictions and
+to contextualize the answers that crowd work researchers have already uncovered.
+We subsequently take this historical framing to ask broader questions of crowd work,
+including among others what the future of crowd work holds for workers.
 \end{document}

### 2016-09-21 drafts/labor/piecework_lit.tex

@@
 ``payment for results'', as \citeauthor{10.2307/2338394} calls it,
-is not only common in practice, but well--studied in labor economics as well
-\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
+is not only common in practice, but well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
 
 \topic{It's worth acknowledging that
-``this distinction [between piece--rates and time--rates] was not completely clear--cut''
-\cite{hart2013rise}.}
+``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
 Indeed, work adopted piece--rate compensation in some aspects and
@@
 was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
-muddies the waters for us later as we attempt to categorize cases of piecework
-\cite{rowan1901premium}.
+muddies the waters for us later as we attempt to categorize cases of piecework~\cite{rowan1901premium}.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
@@
 Of particular interest was their independent nature, via their reputation ``\dots~for generosity, independence, and protectiveness,
-but also for brashness, irregularity, low morality, and little education''
-\cite{10.2307/3827491}.
+but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
 \citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
@@
 to calculate equations in order
-to better--predict the trajectories of various celestial bodies in the \nth{19} century
-\cite{grier2013computers}.
+to better--predict the trajectories of various celestial bodies in the \nth{19} century~\cite{grier2013computers}.
 George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
@@
 necessitated the rapid training and employment of
-a labor pool that hadn't historically been utilized in industrial labor: women
-\cite{hart2013rise}.
+a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
 % With little time and virtually none of the training that was conventionally expected of metalworkers and other industrial work
@@
 
-The early proliferation of manual piecework led to discussion surrounding how best to manage pieceworkers
-\cite{norton1900textile,clark1908cotton}.
+The early proliferation of manual piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
 Despite this, workers' means were mostly ignored,
@@
 % workers themselves began to speak out about their frustration with this new regime.}
-\citeauthor{riisOtherSideLives})
-\cite{riisOtherSideLives}.
+\citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
 This led to industry organizations representing
@@
 \item the quality of ``multidimensional'' work becoming too difficult to evaluate
-\end{inlinelist}
-\cite{hart2013rise}.
+\end{inlinelist}~\cite{hart2013rise}.
 

### 2016-09-21 drafts/labor/conclusion.tex
conc

--- 2016-09-21 drafts/labor/conclusion.tex
@@
 
+\onlyinsubfile{
+\usepackage{xr-hyper}
+\usepackage{hyperref}
+\externaldocument{complexity}
+\externaldocument{decomposition}
+\externaldocument{relationships}
+}
+
 \begin{document}
 \section{Conclusion}
-????
+Crowd work is not new: it is a modern instantiation of piecework.
+In this paper, we reconsidered three major research questions in crowd work using the lens of piecework: \begin{inlinelist}
+  \item ``\namerefl{sec:complexity}?'';
+  \item ``\namerefl{sec:decomposition}?'';
+        and
+  \item ``\namerefl{sec:relationships}?''
+\end{inlinelist}
+To do so, we drew on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon. 
+% predictions and to contextualize the answers that crowd work researchers have already uncovered.
+% We subsequently take this historical framing to ask broader questions of crowd work, including among others what the future of crowd work holds for workers.
+
 \end{document}

### 2016-09-21 drafts/labor/discussion.tex
internet, not Internet

@@
 The past can't be a perfect predictor for the future;
-as \cite{scholz2012digital} points out,
+as \citeauthor{scholz2012digital} points out,
 ``it would be wrong to conclude that
@@
 \cite{mcinnis2016one}.
-The Internet makes this kind of loose affiliation feasible.
+The internet makes this kind of loose affiliation feasible.
 While we have attempted to understand the likely overlaps and differences between history and modern day, no analysis is perfect.
@@
 
-% Arguably the Internet has made collective action on an unprecedented scale much easier.
+% Arguably the internet has made collective action on an unprecedented scale much easier.
 

### 2016-09-21 drafts/labor/introduction.tex
internet, not Internet

@@
 This framing of work into de--contextualized, modular blocks
-enables computation to hire workers at scale through open calls on the Internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
 Distributed paid participants then engage in the work whenever their schedules allow,

### 2016-09-21 drafts/labor/conclusion.tex
conc fightin words

@@
 \section{Conclusion}
-Crowd work is not new: it is a modern instantiation of piecework.
-In this paper, we reconsidered three major research questions in crowd work using the lens of piecework: \begin{inlinelist}
+Crowd work and on--demand work are not new: they are modern instantiations of piecework.
+In this paper, we reconsider three major research questions in crowd work using the lens of piecework: \begin{inlinelist}
   \item ``\namerefl{sec:complexity}?'';
@@
 \end{inlinelist}
-To do so, we drew on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon. 
+To do so, we draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon. 
 % predictions and to contextualize the answers that crowd work researchers have already uncovered.
 % We subsequently take this historical framing to ask broader questions of crowd work, including among others what the future of crowd work holds for workers.
+Reciprocally, we believe that modern crowd work will teach us about the broader phenomenon of piecework as well.
+If history really does repeat itself, the best we can do is be prepared.
 

### 2016-09-22 drafts/labor/introduction.tex
intro revisions

@@
 
-\topic{As on--demand work has grown far beyond the domain of information work from which it first sprang, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who empower it.}
+\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who study it.}
 \citeauthor{howe2008crowdsourcing} first described crowdsourcing in general terms as
-``outsourcing [work] to an undefined, generally large group of people in the form of an open call''.
+``outsourcing [work] to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
 However, for years its instantiation was limited to the utilization of
@@
 has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-This growth prompted efforts to understand the workers who gravitate toward these platforms~\cite{Ross,whoareNOTtheTurkers}.
+This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
 Some of this research has been motivated by the identification of the sociality of gig work,
 and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
-Other work has focused on the \textit{outcomes} of this frustration,
+Other work has focused on the outcomes of this frustration,
 reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.
 
-\topic{This body of research has sought to answer one central question:
-What does the future hold for on--demand work and those that do it?}
-Researchers have offered their input on this open question along three major threads:
+\topic{This body of research has broadly sought to answer one central question:
+What does the future hold for on--demand work and those who do it?}
+Researchers have offered insights on this question along three major threads:
 \begin{enumerate}
@@
 \end{enumerate}
-This research literature has largely sought to answer these questions by examining the present phenomenon.
+This research has largely sought to answer these questions by examining the present on--demand work phenomenon.
 So far, it has not offered a framing for holistically explaining
@@
 nor has any research, to our knowledge,
-gone as far as predict future developments.
+gone as far as directly predict future developments.
 
@@
 a work and payment structure which breaks tasks down into standalone contracts,
-wherein payment is made for \textit{work output}, rather than for \textit{time}.
+wherein payment is made for work output, rather than for time.
 Piecework as a metaphor for crowd work is not new.
@@
 \citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
-referenced crowd work as ``piecework'' briefly
+referenced crowd work as piecework briefly
 as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
@@
 the framing of on--demand labor as a re--instantiation of piecework
-gives us years of historical material to make sense of the broader research on this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.
+gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.
 
@@
   \item make sense of the phenomena so far as part of a much larger series of interrelated events;
-  \item reflect on similarities in the ongoing work among
-        workers,
-        system--designers, and
-        researchers in this space;
+  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today;
         and finally,
@@
         and workers themselves,
-        should expect to see on the horizon.
+        might expect to see on the horizon.
 \end{inlinelist}
-For example, we will draw on the piecework literature such as case studies of the Santa Fe Railway to understand the historical complexity limits in piecework, and leverage that understanding to suggest which modern complexity limits in crowd work~\cite{crowdworkFuture} may be fundamental and which may be overcome.
+For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.
 %we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
@@
 
-We organize this paper as follows: we first review the literature on piecework to lay  groundwork and make clear the analogy to on--demand work.
-Then, we interrogate the three major research questions above from a piecework frame.
+We organize this paper as follows: we first review the definition and historical arc of piecework to lay  groundwork and make clear the analogy to on--demand work (which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature).
+Then, we interrogate the three major research questions above using the lens of piecework.
 % We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
@@
 % and developments we should expect to see in the maturation of digitally mediated work.
-Finally, we will offer design implications for researchers and practitioners based on our results.
+Finally, we will offer implications for researchers and practitioners based on our results.
 

### 2016-09-22 drafts/labor/abstract.tex
intro notes resolved, some requests for feedback i also changed it so that \ali and \msb notes don't show up in the final document, but do show up in the constituent parts. if you want to undo that, comment lines 124-127 (or just 126-127)
internet, not Internet
remove spacing from abstract

@@
 \documentclass[trackingWork]{subfiles}
-
 \onlyinsubfile{\usepackage{xr-hyper}
@@
 \begin{abstract}
-The Internet is enabling the rise of
+The internet is enabling the rise of
 crowd work,
@@
 \begin{inlinelist}
-\item \nameref{sec:complexity}?
-\item \nameref{sec:decomposition}? and
+\item \nameref{sec:complexity}?,
+\item \nameref{sec:decomposition}?, and
 \item \nameref{sec:relationships}?
@@
 to understand how these questions might play out with modern crowd work.
-To do so, we identify the mechanisms that enabled and limited piecework historically, and
+We identify the mechanisms that enabled and limited piecework historically, and
 identify whether crowd work faces the same pitfalls or might differentiate itself.

### 2016-09-22 drafts/labor/piecework_lit.tex
piecework_lit

@@
 \begin{inlinelist}
-  \item define ``piecework'' as researchers in the topic understood it;
-  \item trace the rise of piecework at a very high level,
-        identifying key figures and ideas during this time; and finally
-  \item look at the fall of piecework, such as it was,
-        considering in particular
-        the factors that may have led to piecework's
-        eventual demise in the American and European labor markets.
+  \item define ``piecework'' as researchers in the topic understood it; and
+  \item trace the rise and fall of piecework at a very high level,
+        identifying key figures and ideas during this time;
+  %        and finally
+  % \item look at the fall of piecework, such as it was,
+  %       considering in particular
+  %       the factors that may have led to piecework's
+  %       eventual demise in the American and European labor markets.
         % \msb{did it demise? there are sweatshops around the world today}.
@@
 for each living foal got from a mare;~[\dots]~
-for each living calf got~[\dots]''~\cite{10.2307/2338394}.
+for each living calf got''~\cite{10.2307/2338394}.
 This framing perhaps makes the most intuitive sense;
-``payment for results'', as \citeauthor{10.2307/2338394} calls it,
+``payment for results,'' as \citeauthor{10.2307/2338394} calls it,
 is not only common in practice, but well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
@@
 Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition. 
-Platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well. 
+First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well. 
 % Second, workers on these platforms have significant freedom in choosing which tasks to work on, and when to log on and perform work~\cite{martin2014being,whyWouldAnyoneBrewer}. 
@@
 Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-Finally, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
-
+Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
 Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.

### 2016-09-22 drafts/labor/complexity.tex
internet, not Internet
complexity tweaks

@@
 medical data interpreters \cite{Lasecki:2013:CCC:2501988.2502057}, and
-idea generation \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+idea generation workflows \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
 to name a few examples.
@@
 
-It is now clear that this approach works with focused complex tasks, but
-the broader wicked problem that each complex goal represents remains unsolved~\cite{rittel1973dilemmas}.
+It is now clear that this computational workflow approach works with focused complex tasks, but
+the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.
 As a first example,
@@
 We argue that
-the Internet has trivialized
+the internet has trivialized
 the costs and challenges of the earlier limiting factors
@@
 % While the threshold preventing task requesters from utilizing piecework
-% has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
+% has dropped thanks to affordances of the internet, the ceiling on task complexity hasn't moved significantly.
 % \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}

### 2016-09-22 drafts/labor/complexity.tex
typos

@@
 For example, the roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
-\citeauthor{hart2016rise} argues that evaluation causes an ultimate complexity limit:
+\citeauthor{hart2016rise} argue that evaluation is the ultimate complexity limit:
 at some point, evaluating multidimensional work for quality
 (rather than for quantity) becomes infeasible.
-In his words,
+In their words,
 ``if the quality of the output is more difficult to measure than the quantity [\ldots]

### 2016-09-22 drafts/labor/complexity.tex
tweak

@@
 --- a demonstration of the driver's comprehensive familiarity with the city's roads.
-This test is so challenging that veteran drivers exhibit significantly larger
+This test is so challenging that veteran drivers develop significantly larger
 the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,

### 2016-09-22 drafts/labor/cases.tex
removed dumb paragraph

@@
 
-\msb{this needs a transition into it.
+% \msb{this needs a transition into it.
 Before you lay out our method, lay out what we're trying to do.
@@
 
-We look at lots of papers that use the term ``crowdsourcing'' or ``crowd work''
-in the abstracts and titles of their papers,  especially the papers that we cited earlier, to try
-to answer the questions that we posed in the introduction.
-Then we look to the piecework literature to see whether and to what extent
-piecework answers the questions we, crowdsourcing researchers, have asked.
-We then see whether and how crowdsourcing as we know it has differed from piecework,  and how that affects the predictions and conclusions made in the piecework literature.
+% We look at lots of papers that use the term ``crowdsourcing'' or ``crowd work''
+% in the abstracts and titles of their papers,  especially the papers that we cited earlier, to try
+% to answer the questions that we posed in the introduction.
+% Then we look to the piecework literature to see whether and to what extent
+% piecework answers the questions we, crowdsourcing researchers, have asked.
+% We then see whether and how crowdsourcing as we know it has differed from piecework,  and how that affects the predictions and conclusions made in the piecework literature.
 

### 2016-09-28 cleanrefs.py
cleaned refs

@@ -2,9 +2,12 @@ import re,itertools
 with open('references.bib') as references:
   refs=references.read()
 
-refObjects=re.findall("(@(?:.*?)\})\n", refs,re.MULTILINE|re.DOTALL)
+refObjects=re.findall("(@(?:.*?)\}\n\})\n", refs,re.MULTILINE|re.DOTALL)
 print("Found {} references...".format(len(refObjects)))
 
+
+# print(refObjects[0])
+# quit()
 bibliography=list()
 longest_field=0
 
@@ -20,7 +23,7 @@ for obj in fields:
   temp=set(itertools.chain(*filter(None,map(fieldName.findall,obj))))
   results=results.union(temp)
 # print(results)
-longest_field=len(max(results,key=lambda n:len(n)))
+longest_field=max(results,key=lambda n:len(n))
 
 # print(longest_field)
 # quit()
@@ -29,10 +32,16 @@ for ref in refObjects:
   for line in ref.split("\n"):
     output=list()
     result=re.findall("^\s*(\S+)\s*=\s*\{(.*)\}(?:,|\})",line)
-    output.append("  {field:{width}} = {{{entry}}},".format(field=result[0][0],width=longest_field,entry=result[0][1]) if result else line or "}")
+    # if result and result[0][0]=="year":
+    #   print(result)
+    #   quit()
+    output.append("  {field}{space} = {{{entry}}},".format(field=result[0][0],
+                                                           space=" "*(len(longest_field)+15-len(result[0][0])),
+                                                           entry=result[0][1]) if result else line or "}")
     bibliography.append(output[0])
-    # print("\n".join(output))
+    print("\n".join(output))
 
-print("\n".join(bibliography))
-# for n in bibliography:
-#   print(n)
+
+# print("\n".join(bibliography))
+for n in bibliography:
+  print(n)

### 2016-11-21 drafts/labor/rebuttal.tex
rebuttal init

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,times,hyperref,coloremoji}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\bibliography{references}
% \setlength{\parskip}{.4em}

\begin{document}
We'd like to thank the reviewers for the feedback they offered
--- their input will absolutely strengthen this paper.
We'll focus on calling to attention the potential hazards of
using historical analyses to draw ahistorical conclusions in our discussion,
and attempt to keep it

Dear reviewers,

Thanks, keep it 💯 yo

\printbibliography{}
\end{document}

### 2016-11-21 admin/IRB/performances/interview_questions/main.tex
rebuttal init

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{nameref,color}
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  % pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  % pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
% \setlength{\parskip}{.4em}
\pagenumbering{gobble}
\begin{document}

Interviews with performance artists under this protocol are designed to be
(at least somewhat)
open--ended, but given that we're still going in with a handful of
\textbf{a priori} notions of what we want to learn about,
we have a list of prepared questions that we plan to use to prompt
further discussion about whatever proves to be of interest to participants.
That is, while we have questions about performance artists'
\hyperref[sec:Background]{backgrounds},
\hyperref[sec:LearningCraft]{learning their craft},
\hyperref[sec:EngagingAudience]{engaging with an audience that has more direct avenues of communication},
\&
\hyperref[sec:DealingWithLaw]{negotiating the political and legal issues within online platforms}.

\begin{enumerate}

  \section*{Background}
            \label{sec:Background}
    \item How long have you been (streaming | YouTubing | podcasting | etc\dots)?
    \item What did you do before that?
    \item What got you into this?

  \section*{Learning the craft}
            \label{sec:LearningCraft}
    \item I imagine there are some skills that you have to learn to be a good
          (streamer, podcaster, etc\dots); can you talk about any of the skills
          you found you had to learn to be good at this?
    \item What sorts of things do you have to become good at
          that your audience might not realize (if anything)?

  \section*{Engaging the Audience}
            \label{sec:EngagingAudience}
    \item Do you see feedback or engagement from your audience?
          How?
    \item What avenues do people take to engage with you?
          Do you actively open and manage those channels, or
          do people simply find them?
    \item How do you respond to that kind of engagement?
          That is,
            do you prompt it?
            Encourage it?
            Keep it at arm's length?
    \item Do you encounter much criticism?
          (If so\dots)
    \item How do you deal with criticism?
    \item Is this feedback and engagement overall a good thing, or
          is it generally more trouble than the value it provides?

  \section*{Dealing with ``law'' and politics}
            \label{sec:DealingWithLaw}
    \item Do any policies or rules come up in your videos?
          To give an example, do you find yourself fending off DMCA complaints,
          or having to argue that your content isn't in violation of some rule?
    \item Do rules and policies ever change?
          (Can you describe how?)
    \item What influence do you (or others in your position) have
          (if any)
          in affecting policies?
    \item How do you think of yourself
          in relation to the platform you publish on?
    \item To be more specific,
          how much agency do you feel you have in
          changing policies that don't make sense
          (or are unreasonably applied)?
\end{enumerate}

% \printbibliography{}
\end{document}

### 2016-11-21 drafts/compensating_task_search/introduction.tex
rebuttal init

\documentclass[main]{subfiles}
\onlyinsubfile{
  \usepackage{xr-hyper}
  \usepackage{hyperref}
  % \externaldocument{nameHere}
}

\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\section{Introduction}\label{sec:introduction}

Motivated by~\cite{crowdcollab} and the general insight that a fair amount of work
isn't compensated because it doesn't fall within the confines of narrowly defined tasks,
we address the known problem of task search, described by~\cite{taskSearch}
and studied by a number of subsequent researchers.

Workers on Mechanical Turk already spend time helping one another
(e.g. see \href{https://www.reddit.com/r/HITsWorthTurkingFor/}{/r/HITsWorthTurkingFor},
a sub--community on Reddit to assist ``Turkers'' in finding worthwhile tasks),
but their reward for such work is relatively unclear;
some prior research into hiring referrals suggests that
people who make worthwhile connections are afforded a certain amount of
social capital~\cite[see][]{Gemkow2011}.

Still, social aspects of job referral as they've been studied in the past have focused

\onlyinsubfile{
  % \balance{}
  \printbibliography
}

\end{document}

### 2016-11-21 drafts/task_referral/introduction.tex
rebuttal init

\documentclass[main]{subfiles}
\onlyinsubfile{
  \usepackage{xr-hyper}
  \usepackage{hyperref}
  % \externaldocument{nameHere}
}

\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\section{Introduction}\label{sec:introduction}

\ali{
\begin{enumerate}
  \item What problems are really holding crowd work back?
  \item Complexity. Workers are afraid
        to do creative work and get shot down for sticking their necks out.
  \item Okay, well \cite{haas2015argonaut} addresses that.
        Is that paper not solving that problem?
        Or is something else the problem?
  \item It addresses that problem to an extent.
        Tasks are also (prohibitively) poorly designed.
  \item There's some work on that as well [task iteration paper].
        So, again, what's missing?
  \item Fundamentally, nothing. We have all the pieces to this puzzle.
        But between iterating on task design,
        setting appropriate pay rates, etc\dots
        the emergent complexity in managing crowd workers has become a full time job.
  \item So what's the fix? Is there one?
  \item The fix is to take \citeauthor{haas2015argonaut}'s work and
        extend it to the logical conclusion
        --- \textit{foremen}, middle managers who are responsible for guiding a number of workers
        (or in this case, tasks) to satisfactory completion.
  \item While the working of these broken--down tasks is called ``micro--work'',
        we call the broken--down management of workers ``micro--management''. :P
\end{enumerate}
}

On--demand workers have expressed frustration with the algorithmic management that largely dictates their work
\cite{uberAlgorithm}.
The fear of deviating from expected results causes online information workers 
--- such as workers on Amazon Mechanical Turk (AMT), or ``Turkers'' ---
to avoid creative or open--ended tasks due to the perceived riskiness of
investing a significant amount of time into a task only to be rejected.
\ali{[citation needed]}
\citeauthor{haas2015argonaut} address some of these problems by bootstrapping
human reviewers from vetted crowd workers
\cite{haas2015argonaut}.

We propose an extension of this relationship assigning intermediary management tasks to crowd workers,
building on a body of literature reflecting on the myriad roles and
ultimate importance of assembly line managers known as ``foremen''.
We propose to explore other ways that crowd workers can
guide \& inform workers as they do creative tasks,
as well as consolidate \& propagate feedback to requesters when it becomes apparent something is wrong.
More than reviewing tasks,
the administration of tasks itself becomes an eligible candidate for crowdsourcing.

\onlyinsubfile{
  % \balance{}
  \printbibliography
}

\end{document}

### 2016-11-21 qual_management/main.tex
rebuttal init

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,times,hyperref,titling,microtype}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[compact]{titlesec}
\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
  \textbf{\LARGE{\@title}}

  \@author
\end{flushleft}\egroup
}
\makeatother
\title{Externalizing Worker Qualifications}
\author{Ali Alkhatib}

\begin{document}
\maketitle{}

\section*{Background}
Labor market platforms ranging industries and styles of work
(e.g. Uber, TaskRabbit, UpWork, and Amazon Mechanical Turk)
have struggled for years with persistent and
in some cases growing
challenges relating to worker qualifications.
These issues range a wide spectrum:
in some cases, worker qualifications are non--transferable,
leading requesters to ``re--invent the wheel'' as they attempt to determine
in their own way whether a potential worker is qualified and reliable.

Problems determining a worker's qualifications start on day 1;
labor markets begin their relationships with new workers
almost entirely uncertain about the worker's competence in any type of task.
Gathering this information through qualification exams is generally
time--consuming and costly.

Challenges mount as workers' skill sets develop;
work requiring more training and skill (for example, translating or programming)
are either verified by individual \textit{requesters} (e.g. Amazon Mechanical Turk)
or are verified by the \textit{platform} itself (e.g. UpWork).
While the \textit{UpWork} model avoids needless repetitive work by
generally consolidating qualifying exams at the platform level,
these labor platforms nevertheless find themselves in the unenviable
(and often unexpected) position of
having to develop new qualifications exams to outpace would--be cheaters.

% \textit{I want to communicate something along the lines of how
% these labor markets shouldn't be expected
% to handle these problems very well, because
% they've been focusing on sparking a form of commerce that hasn't existed before,
% etc\dots
% but I feel like this sounds too much like
% ``It's no wonder you all have failed!'' and
% I feel like there's no way to dig myself out of that hole.
% Thoughts?}

\section*{A (Potential) Solution}
In offline labor markets involving skilled workers,
credentials are sometimes managed by
external, trusted organizations:
the state requires electricians
to serve in apprenticeships and
pass licensing exams;
lawyers take exams administered by the American Bar Association (ABA);
doctors take similar exams given by their own oversight organization (the NBME).

This approach may prove useful in
alleviating the burden online labor markets are increasingly taking on.
By externalizing worker qualifications, a number of benefits can emerge:
\begin{enumerate}
  \item New workers can provide some evidence of their work history to labor platforms,
        mitigating or even resolving the ``cold start'' problem of
        not knowing the trustworthiness or competence of a new user.
  \item Aggregated (even heterogeneous)
        ratings sourced from differing labor markets can provide
        a more holistic picture of a worker's areas of competence.
        Workers can use this information to identify specialization more easily,
        reducing wasted time \textit{searching for} or \textit{working on} suboptimal tasks.
        Requesters can benefit from this information by
        making more informed decisions about which worker to contract.
  \item By turning the reputation management of workers into an external entity
        (agnostic to labor platform),
        designers and developers can focus more fully on the marketplace itself.
\end{enumerate}

% Rather than invest time and effort into worker reputation systems,
% qualifications exams, and all of the overhead associated with these challenges,
% labor platforms could
% externalize worker reputations and qualifications to other entities.
% Most labor markets

\section*{Measurable Variables}

\section*{First Steps}

% \printbibliography{}
\end{document}

### 2016-11-22 drafts/labor/rebuttal.tex
rebuttal draft?

@@
 \documentclass[11pt]{article}
-\usepackage{balance,graphics,setspace,parskip,times,hyperref,coloremoji}
+\usepackage{balance,graphics,setspace,parskip,times,hyperref,nth}
 \usepackage[margin=1in]{geometry}
@@
 % \setlength{\parskip}{.4em}
-
+\pagenumbering{gobble}
 \begin{document}
@@
 --- their input will absolutely strengthen this paper.
-We'll focus on calling to attention the potential hazards of
-using historical analyses to draw ahistorical conclusions in our discussion,
-and attempt to keep it
 
-Dear reviewers,
+1AC and R5 bring up that one of the issues with historical analysis such as ours
+is that the method strips away context in pursuit of ahistoric conclusions.
+This is absolutely an excellent point, and
+a shortcoming which we struggled to articulate in our discussion section.
+We'll add more to the discussion section to highlight the potential hazards
+of applying historical analyses due to the different
+politics, economies, cultures, and other factors which
+may influence any phenomena.
 
-Thanks, keep it 💯 yo
+\textit{There's almost certainly something by Rosenberg about how
+        historical analysis is flawed but not totally unhelpful.
+        I'll quote it.}
+
+R5 also points out that we give attention to \citeauthor{grier2013computers}'s work and
+the case study of human computers,
+perhaps at the expense of the other case studies
+(that is, the cases of the matchstick girls and railroad workers).
+While we'll occasionally reference other cases
+(such as the industrial workers outside of
+the railroad industry in the early and mid--\nth{20} century),
+we'll bring the other two case studies to
+a similar level of scrutiny that the case study of Airy's human computers
+(as described by \citeauthor{grier2013computers}~\cite{grier2013computers}) is afforded.
+
+R3 noted that we decided to cluster crowd work research in a particular way
+which consolidated some research
+--- for instance, into ``professional development'' --- into
+the broader topic of ``complexity'';
+the observation is well taken, and we'll dedicate some space to reflect on
+the decisions we made with regard to clustering.
+
+R4 and R5 suggest a number of works in scientific management
+(\textbf{Hounsell}, \textbf{Roe Smith}, and \citeauthor{williamson2016}) for a more complete view on scientific management.
+We struggled with this topic in the process of writing,
+as we wanted to communicate a high--level overview of some topics only as they
+were relevant to the core argument of piecework.
+We worried that bringing \textbf{Hounsell}, \textbf{Roe Smith}, and others into
+the section mentioning scientific management would have necessitated
+a much more involved discussion than
+this venue's format would afford.
+Given this feedback, we'll
+acknowledge this influential work and point out that there's
+significantly more detail and nuance to scientific management than
+we could afford to explore, given our prerogative.
+
+% {\LARGE
+% Dear reviewers,
+
+% You da real MVPs
+
+% Your feedback was {\Huge😎} I'm all like {\Huge🙇}
+
+% But R5 I mean {\Huge😒}
+
+% Nah I'm just kidding {\Huge😂} keep it {\Huge💯}!
+% }
 

### 2016-11-22 drafts/labor/complexity.tex
rebuttal draft?

@@
 additional proof--of--concept applications and techniques,
-including conversational assistants \cite{Lasecki:2013:CCC:2501988.2502057},
-medical data interpreters \cite{Lasecki:2013:CCC:2501988.2502057}, and
+including conversational assistants \cite{lasecki2013chorus},
+medical data interpreters \cite{lasecki2013chorus}, and
 idea generation workflows \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},

### 2016-11-22 drafts/labor/rebuttal.tex
rebuttal rev 2.0

--- 2016-11-21 drafts/labor/rebuttal.tex
@@
 \bibliography{references}
-% \setlength{\parskip}{.4em}
+\newlist{inlinelist}{enumerate*}{1}
+\setlist*[inlinelist,1]{%
+   % label=\alph*),
+ label=\arabic*),
+}
+\pagenumbering{gobble}
+\begin{document}
+We'd sincerely like to thank the reviewers for the feedback they generously offered
+--- your comments have pointed out some opportunities for clarification
+and potential changes in our framing that we think will help strengthen the paper.
 
-\begin{document}
-We'd like to thank the reviewers for the feedback they offered
---- their input will absolutely strengthen this paper.
-We'll focus on calling to attention the potential hazards of
-using historical analyses to draw ahistorical conclusions in our discussion,
-and attempt to keep it
+1AC and R5 bring up that one of the potential hazards with
+historical analysis is that
+it tends to strip away context in pursuit of ahistoric conclusions.
+This is an excellent point, and
+the reference to~\citeauthor{rosenberg1994exploring}'s
+work in particular is especially beneficial to us, as his work
+(in~\citeyear{rosenberg1982inside} and later~\citeyear{rosenberg1994exploring})
+offers a concise description of the spirit of the method we adopted here,
+and articulates a limited use case for this approach.
 
-Dear reviewers,
+In our revision, we'll attempt to clarify two things:
+\begin{enumerate}
+  \item that this particular contribution offers
+        to fill some of the gaps in crowd work and suggests potential futures
+        (perhaps specifically offering
+            ``to narrow our estimates
+              and thus to concentrate resources
+              in directions that are more likely to have useful payoffs''~\cite{rosenberg1994exploring});
+  and
+  \item that the overarching method here may be
+        a powerful tool for researchers attempting
+        to make sense of (seemingly) new phenomena
+        (for example,
+        arguing ``that past history is
+                  an indispensable source of information
+                  to anyone interested in characterizing technologies''~\cite{rosenberg1982inside}).
+\end{enumerate}
 
-Thanks, keep it 💯 yo
+% To quote~\citeauthor{rosenberg1994exploring},
+% ``Science will often provide the capability to acquire information
+% \dots that we do not presently possess, but
+% \textit{science does not make the acquisition of this information costless}''
+%~\cite{rosenberg1994exploring}.
+% In that spirit, 
+% This is absolutely an excellent point, and
+% a shortcoming which we struggled to articulate in our discussion section.
+% We'll add more to the discussion section to highlight the potential hazards
+% of applying historical analyses due to the different
+% politics, economies, cultures, and other factors which
+% may influence any phenomena.
+
+% \textit{There's almost certainly something by Rosenberg about how
+%         historical analysis is flawed but not totally unhelpful.
+%         I'll quote it.}
+
+R5 also points out that we give attention to~\citeauthor{grier2013computers}'s work and
+the case study of human computers,
+perhaps at the expense of the other case studies
+(that is, the cases of the matchstick girls and railroad workers).
+While we occasionally bring to light other cases
+(such as the industrial workers during the Second World War),
+we'll attempt to bring the two major case studies to
+a similar level of attention as we afford Airy's human computers,
+% which~\citeauthor{grier2013computers} offers~\cite{grier2013computers},
+making these three case studies into
+more equal threads which recur throughout the paper.
+
+R3 noted our decision to cluster crowd work research in a particular way,
+which consolidated some research
+--- for instance, the work into ``professional development'' ---
+into the broader topic of ``complexity'';
+the concern R3 raises is well taken, and
+we'll dedicate some space to reflect on
+the decisions we made with regard to clustering research topics.
+
+R4 and R5 offer a number of works
+(e.g. by~\textbf{Hounsell}, \textbf{Roe Smith}, and~\citeauthor{williamson2016},
+separately)
+for a more comprehensive discussion of scientific management.
+We agree that these works will
+substantively add to a reader's understanding of scientific management.
+% We'll attempt to bring into discussion these works, but
+% our original concern was that a satisfactory review of scientific management
+% would be difficult to articulate concisely.
+We'll attempt to crystallize the body of work concisely, and
+point out that there's much more to be said about these topics.
+
+% We found ourselves struggling to reconcile
+% how adequately to bring up some of the leading thinkers in
+% scientific management
+% --- and their contributions ---
+% given the tight constraints.
+% We can only gesture to~\textbf{Hounsell}, \textbf{Roe Smith}, and~\citeauthor{williamson2016}
+% and point out that there's substantially more to scientific management than we
+% can ``afford'' to communicate.
+% Given this feedback, we'll attempt to carefully balance the desire
+% to satisfactorily discuss leading minds in these fields with
+% the limited space afforded to provide readers with context
+% We struggled with this topic in the process of writing,
+% as we wanted to communicate a high--level overview of some topics only as they
+% were relevant to the core argument of piecework.
+% We worried that bringing \textbf{Hounsell}, \textbf{Roe Smith}, and others into
+% the section mentioning scientific management would have necessitated
+% a much more involved discussion than
+% this venue's format would afford.
+% Given this feedback, we'll
+% acknowledge this influential work and point out that there's
+% significantly more detail and nuance to scientific management than
+% we could afford to explore, given our prerogative.
+
+% {\LARGE
+% Dear reviewers,
+
+% You da real MVPs
+
+% Your feedback was {\Huge😎} I'm all like {\Huge🙇}
+
+% But R5 I mean {\Huge😒}
+
+% Nah I'm just kidding {\Huge😂} keep it {\Huge💯}!
+% }
 

### 2016-11-22 drafts/labor/rebuttal.tex
msb comments

@@
 \begin{document}
-We'd sincerely like to thank the reviewers for the feedback they generously offered
---- your comments have pointed out some opportunities for clarification
-and potential changes in our framing that we think will help strengthen the paper.
+We sincerely thank the reviewers for the feedback they offered.
+Your comments have identified several opportunities for clarification and potential changes in our framing that we think will help strengthen the paper.
 
+\section{RISKS OF A HISTORICAL ANALYSIS}
 1AC and R5 bring up that one of the potential hazards with
@@
 offers a concise description of the spirit of the method we adopted here,
-and articulates a limited use case for this approach.
+and articulates a use case for this approach.
 
-In our revision, we'll attempt to clarify two things:
+In our revision, we will to clarify:
 \begin{enumerate}
-  \item that this particular contribution offers
-        to fill some of the gaps in crowd work and suggests potential futures
+  \item that our contribution offers to fill some of the gaps in crowd work and suggests potential futures
         (perhaps specifically offering
@@
   and
-  \item that the overarching method here may be
+  \item that our method of connecting history to modern socio-technical phenomena may be
         a powerful tool for researchers attempting
@@
 
+\section{CASE STUDY FOCUS}
 R5 also points out that we give attention to~\citeauthor{grier2013computers}'s work and
 the case study of human computers,
-perhaps at the expense of the other case studies
-(that is, the cases of the matchstick girls and railroad workers).
+perhaps at the expense of the other case studies.
+Our goal had been to equally highlight the cases of the matchstick girls and railroad workers.
 While we occasionally bring to light other cases
@@
 
-R3 noted our decision to cluster crowd work research in a particular way,
-which consolidated some research
---- for instance, the work into ``professional development'' ---
-into the broader topic of ``complexity'';
-the concern R3 raises is well taken, and
-we'll dedicate some space to reflect on
+\section{TOPIC SELECTION}
+R3 and R4 noted our decision to cluster crowd work research around three questions
+that consolidated some research
+--- for instance, the ``quantity-quality dilemma'' (R4), ``professional development'' (R3) and ``incentive structures'' (R3) --- into other broader topics.
+This critique is well taken. We will dedicate some space to reflect on
 the decisions we made with regard to clustering research topics.
 
+\section{RELATED WORK}
 R4 and R5 offer a number of works

### 2016-11-22 drafts/labor/rebuttal.tex
ethics

@@
 
-\section{RISKS OF A HISTORICAL ANALYSIS}
+\section{RISKS OF HISTORICAL ANALYSIS}
 1AC and R5 bring up that one of the potential hazards with
@@
 
+\section{ETHICS}
+R4 asks whether our analysis can shed any light onto the question of ``whether it is ethical, to make use of crowd work in HCI research''.
+There are two possible interpretations of this question: 1) whether piecework itself is ethical, or 2) whether research on crowdwork is ethical. 
+We are assuming that R4 meant the former.
+We will add this topic to the Discussion of our paper, integrating R4's suggested Williamson paper.
+Briefly, the literature on the history of labor does not position itself as stating that paid work is strictly ethical or unethical: the question, rather, is: what are the conditions under which is it exploitative?
+This literature argues that the exploitation occurs when conditions put workers in harm's way, such as in sweatshops and agricultural work with pesticides, or where the employers are underpaying or overworking by the standards of the day.
+The question then becomes whether the socio-technical infrastructure of Mechanical Turk and other marketplaces are putting workers in harm's way, underpaying, or overworking.
+Mechanical Turk itself does not directly require any amount of payment or work.
+However, its design does encourage employers to engage in such behaviors: piecework rates, for example, undervalue workers' task search time, and the task design interfaces explicitly encourage replication across multiple workers of uncertain quality rather than identifying one worker and paying them more.
+
 % We found ourselves struggling to reconcile

### 2016-11-23 drafts/labor/rebuttal.tex
added 2000 characters, making it 2000 characters over the 5k char limit...

--- 2016-11-22 drafts/labor/rebuttal.tex
@@
 
-\section{RISKS OF A HISTORICAL ANALYSIS}
+\section*{RISKS OF HISTORICAL ANALYSIS}
 1AC and R5 bring up that one of the potential hazards with
@@
 
-\section{CASE STUDY FOCUS}
+\section*{CASE STUDY FOCUS}
 R5 also points out that we give attention to~\citeauthor{grier2013computers}'s work and
@@
 perhaps at the expense of the other case studies.
-Our goal had been to equally highlight the cases of the matchstick girls and railroad workers.
+Our goal had been to equally highlight the cases of the ``match--girls'' and railroad workers.
 While we occasionally bring to light other cases
@@
 
-\section{TOPIC SELECTION}
+We can elaborate on a number of aspects of match--girls' work and
+how it relates to crowd work under the case studies of
+decomposition and worker relations in the following ways:
+
+% \subsection*{Complexity}
+
+\subsection*{Decomposition}
+We can delve into the parallels between
+Airy's computers, match--girls, farm work, and railroad work in the context of
+the decomposition of work by looking at
+the compensation models which directed attention toward
+discrete, measurable output (namely, rather than remuneration by time).
+We can further flesh this topic out by highlighting
+the overarching driving force to discretize work output and
+render payment for each individual unit of output,
+in some cases (e.g. match--girls' work manufacturing individual matches)
+more effectively than others (e.g. railroad workers renewing a coupler, or
+                                   farm workers birthing a foal).
+Hypothesizing
+(and in the case of railroad work, deriving some insight from
+the critiques outlined by~\citeauthor{american1921problem}'s grievances \cite{american1921problem})
+from these cases, we might infer that inadequate measurement approaches
+--- particularly those which leave open the opportunity for systematized gaming ---
+stymied piecework's popularity in some cases, while in others it flourished.
+
+\subsection*{Relationships}
+With regard to the relationships among workers and between workers and their managers,
+we see ample opportunity to further develop the case studies of match--girls in both
+the context of their cooperation
+(that is, how they collectively acted to initiate a workers' strike), and
+in the context of their relationships with managers
+(that is, a decidedly adversarial one, fraught with
+arbitrary and sometimes punitive payment deductions).
+Regarding railroad workers (and to a lesser extent farm workers),
+our sources gave some insight to the relationships between workers and their managers,
+but ``anthropological'' studies of the sociality of workers and
+their relationships with managers and each other was, as we conclude later,
+notably sparse relative to what we now know about crowd work and its culture.
+
+\section*{TOPIC SELECTION}
 R3 and R4 noted our decision to cluster crowd work research around three questions
@@
 
-\section{RELATED WORK}
+\section*{RELATED WORK}
 R4 and R5 offer a number of works
-(e.g. by~\textbf{Hounsell}, \textbf{Roe Smith}, and~\citeauthor{williamson2016},
-separately)
+(e.g.~\citeauthor{williamson2016})
 for a more comprehensive discussion of scientific management.
@@
 point out that there's much more to be said about these topics.
+
+\section*{ETHICS}
+R4 asks whether our analysis can shed any light onto the question of
+``whether it is ethical, to make use of crowd work in HCI research''.
+We had two interpretations of this question:
+\begin{inlinelist}
+  \item whether piecework itself is ethical, or
+  \item whether research on crowdwork is ethical. 
+\end{inlinelist}
+We'll engage here with the former.
+
+We will add this topic to the Discussion of our paper,
+integrating R4's suggested~\citeauthor{williamson2016} paper.
+Briefly, the literature on the history of labor does not
+position itself as stating that piecework is inherently ethical or unethical;
+the question, rather, is what conditions render it exploitative?
+
+This literature argues that the exploitation occurs when
+conditions put workers in harm's way, such as
+in sweatshops and agricultural work with pesticides, or where
+employers systematically underpay or overwork laborers by contemporary standards.
+
+The question then becomes
+whether the socio--technical infrastructure of Mechanical Turk
+and other marketplaces
+are putting workers in harm's way, underpaying, or overworking.
+Mechanical Turk itself does not directly require any amount of payment or work.
+However, its design does encourage employers to engage in such behaviors:
+piecework rates, for example, undervalue workers' task search time, and
+the task design interfaces explicitly encourage
+replication across multiple workers of uncertain quality rather than
+identifying one worker and paying them more.
 

### 2016-11-23 drafts/labor/rebuttal.tex
rebuttal revisions to make the case studies sections less bad

--- 2016-11-22 drafts/labor/rebuttal.tex
@@
 \usepackage[tiny,compact]{titlesec}
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
+\usepackage[citestyle=authoryear,backend=bibtex]{biblatex}
 \bibliography{references}
@@
 This is an excellent point, and
-the reference to~\citeauthor{rosenberg1994exploring}'s
+the reference to~\citeauthor{rosenberg1994exploring}'s~(\citeyear{rosenberg1994exploring}
+and later~\citeyear{rosenberg1994exploring})
 work in particular is especially beneficial to us, as his work
-(in~\citeyear{rosenberg1982inside} and later~\citeyear{rosenberg1994exploring})
 offers a concise description of the spirit of the method we adopted here,
@@
 
-In our revision, we will to clarify:
+In our revision, we will clarify that:
 \begin{enumerate}
-  \item that our contribution offers to fill some of the gaps in crowd work and suggests potential futures
+  \item our contribution offers to fill some of the gaps in crowdwork and suggests potential futures
         (perhaps specifically offering
@@
               and thus to concentrate resources
-              in directions that are more likely to have useful payoffs''~\cite{rosenberg1994exploring});
+              in directions that are more likely to have useful payoffs''~(\cite{rosenberg1994exploring}));
   and
-  \item that our method of connecting history to modern socio-technical phenomena may be
+  \item our method of connecting history to modern socio-technical phenomena may be
         a powerful tool for researchers attempting
@@
                   an indispensable source of information
-                  to anyone interested in characterizing technologies''~\cite{rosenberg1982inside}).
+                  to anyone interested in characterizing technologies''~(\cite{rosenberg1982inside})).
 \end{enumerate}
@@
 % \textit{science does not make the acquisition of this information costless}''
-%~\cite{rosenberg1994exploring}.
+%~(\cite{rosenberg1994exploring}).
 % In that spirit, 
@@
 
-\section{CASE STUDY FOCUS}
-R5 also points out that we give attention to~\citeauthor{grier2013computers}'s work and
+\section*{CASE STUDY FOCUS}
+R5 also points out that we give attention to~\citeauthor{grier2013computers}'s~(\citeyear{grier2013computers}) work and
 the case study of human computers,
 perhaps at the expense of the other case studies.
-Our goal had been to equally highlight the cases of the matchstick girls and railroad workers.
+Our goal had been to equally highlight the ``major'' case studies of the match--girls and railroad workers.
 While we occasionally bring to light other cases
 (such as the industrial workers during the Second World War),
-we'll attempt to bring the two major case studies to
-a similar level of attention as we afford Airy's human computers,
-% which~\citeauthor{grier2013computers} offers~\cite{grier2013computers},
-making these three case studies into
-more equal threads which recur throughout the paper.
+% we'll attempt to bring the major case studies to
+% a similar level of attention as we afforded the human computers,
+% % which~\citeauthor{grier2013computers} offers~(\cite{grier2013computers}),
+% making these three case studies into
+% threads of peer importance, recurring through the paper.
+% We can elaborate on a number of aspects of match--girls' work and
+% how it relates to crowdwork under the case studies of
+% decomposition and worker relations in the following ways:
+% \subsection*{Complexity}
+% \subsection*{Decomposition}
+we'll bring more focus on the case studies of
+the match--girls and the railroad workers
+to elevate them to
+\citeauthor{grier2013computers}'s detail of human computers~(\citeyear{grier2013computers}).
 
-\section{TOPIC SELECTION}
-R3 and R4 noted our decision to cluster crowd work research around three questions
+To use the match--girls as an example, we'll address
+the reorientation of payment for work from
+a time--centric approach to
+an output--centric one and how
+this discretization and decomposition of work allowed managers
+to track workers more fastidiously,
+evaluate the working styles and resultant performance,
+and encourage or discourage certain behavior much more granularly via
+precise disciplinary measures.
+This analysis will also
+allow us to make deeper conclusions about
+the implications of piecework decomposition on crowdwork.
+
+% We can delve into the parallels between
+% Airy's computers, match--girls, farm work, and railroad work in the context of
+% the decomposition of work by looking at
+% the compensation models which directed attention toward
+% discrete, measurable output (namely, rather than remuneration by time).
+% We can further flesh this topic out by highlighting
+% the overarching driving force to discretize work output and
+% render payment for each individual unit of output,
+% in some cases (e.g. match--girls' work manufacturing individual matches)
+% more effectively than others (e.g. railroad workers renewing a coupler, or
+%                                    farm workers birthing a foal).
+% Hypothesizing
+% (and in the case of railroad work, deriving some insight from
+% the critiques outlined by~\citeauthor{american1921problem}'s~(\citeyear{american1921problem}) grievances)
+% from these cases, we might infer that inadequate measurement approaches
+% --- particularly those which leave open the opportunity for systematized gaming ---
+% stymied piecework's popularity in some cases, while in others it flourished.
+
+% \subsection*{Relationships}
+
+% With regard to the relationships among workers and between workers and their managers,
+We'll also return to literature on the relationships
+among match--girls in their nascent labor movement and
+the notably adversarial relationships they had with their managers,
+whose above mentioned disciplinary methods later took on
+punitive, arbitrary qualities.
+This discussion will expose further parallels between
+the internal and external relationships
+of crowdworkers and pieceworkers.
+
+% we see ample opportunity to further develop the case studies of match--girls in both
+% the context of their cooperation
+% (that is, how they collectively acted to initiate a workers' strike), and
+% in the context of their relationships with managers
+% (that is, a decidedly adversarial one, fraught with
+% arbitrary and sometimes punitive payment deductions).
+% Regarding railroad workers (and to a lesser extent farm workers),
+% our sources gave some insight to the relationships between workers and their managers,
+% but ``anthropological'' studies of the sociality of workers and
+% their relationships with managers and each other was, as we conclude later,
+% notably sparse relative to what we now know about crowdwork and its culture.
+
+\section*{TOPIC SELECTION}
+R3 and R4 noted our decision to cluster crowdwork research around three questions
 that consolidated some research
@@
 
-\section{RELATED WORK}
+\section*{RELATED WORK}
 R4 and R5 offer a number of works
-(e.g. by~\textbf{Hounsell}, \textbf{Roe Smith}, and~\citeauthor{williamson2016},
-separately)
+(\cite[e.g.][]{williamson2016})
 for a more comprehensive discussion of scientific management.
@@
 point out that there's much more to be said about these topics.
+
+\section*{ETHICS}
+R4 asks whether our analysis can shed any light onto the question of
+``whether it is ethical, to make use of crowdwork in HCI research''.
+We had two interpretations of this question:
+\begin{inlinelist}
+  \item whether piecework itself is ethical, or
+  \item whether research on crowdwork is ethical. 
+\end{inlinelist}
+We'll engage here with the former.
+
+We will add this topic to the Discussion of our paper,
+integrating R4's suggested paper~(\cite[namely,][]{williamson2016}).
+Briefly, the literature on the history of labor does not
+position itself as stating that piecework is inherently ethical or unethical;
+the question, rather, is what conditions render it exploitative?
+
+This literature argues that the exploitation occurs when
+conditions put workers in harm's way, such as
+in sweatshops and agricultural work with pesticides, or where
+employers systematically underpay or overwork laborers by contemporary standards.
+
+The question then becomes
+whether the socio--technical infrastructure of Mechanical Turk
+and other marketplaces
+are putting workers in harm's way, underpaying, or overworking.
+Mechanical Turk itself does not directly require any amount of payment or work.
+However, its design does encourage employers to engage in such behaviors:
+piecework rates, for example, undervalue workers' task search time, and
+the task design interfaces explicitly encourage
+replication across multiple workers of uncertain quality rather than
+identifying one worker and paying them more.
 
@@
 
-\printbibliography{}
+% \printbibliography{}
 \end{document}

### 2016-11-23 drafts/labor/relationships.tex
rebuttal revisions to make the case studies sections less bad

@@
 Soon, many worker organizations were weighing in on (or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
-As mounting attention increasingly revealed problems in piecework's treatment of workers, theworkers themselves began to speak out about their frustration with this new regime.
+As mounting attention increasingly revealed problems in piecework's treatment of workers, the workers themselves began to speak out about their frustration with this new regime.
 Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers~\cite{american1921problem,richards1904anything}.

### 2016-11-23 drafts/labor/rebuttal.tex
more character shaving

@@
 \section*{RISKS OF HISTORICAL ANALYSIS}
-1AC and R5 bring up that one of the potential hazards with
+1AC \& R5 bring up that one of the potential hazards with
 historical analysis is that
@@
 This is an excellent point, and
-the reference to~\citeauthor{rosenberg1994exploring}'s~(\citeyear{rosenberg1994exploring}
-and later~\citeyear{rosenberg1994exploring})
-work in particular is especially beneficial to us, as his work
-offers a concise description of the spirit of the method we adopted here,
+the pointers to~\citeauthor{rosenberg1994exploring}'s works
+in particular are especially beneficial;
+his~\citeyear{rosenberg1982inside} and~\citeyear{rosenberg1994exploring} works
+offer concise descriptions of the spirit of the method we adopted here,
 and articulates a use case for this approach.
-
 In our revision, we will clarify that:
@@
   \item our contribution offers to fill some of the gaps in crowdwork and suggests potential futures
-        (perhaps specifically offering
+        (carefully stated, offering
             ``to narrow our estimates
@@
   and
-  \item our method of connecting history to modern socio-technical phenomena may be
-        a powerful tool for researchers attempting
+  \item our method of relating history to modern socio--technical systems may be
+        a useful tool for researchers attempting
         to make sense of (seemingly) new phenomena
-        (for example,
-        arguing ``that past history is
-                  an indispensable source of information
+        (i.e., offering ``that past history is an indispensable source of information
                           to anyone interested in characterizing technologies''~(\cite{rosenberg1982inside})).
@@
 \section*{CASE STUDY FOCUS}
-R5 also points out that we give attention to~\citeauthor{grier2013computers}'s~(\citeyear{grier2013computers}) work and
-the case study of human computers,
+R5 points out that we give attention to~\citeauthor{grier2013computers}'s~(\citeyear{grier2013computers}) work and
+the case of human computers,
 perhaps at the expense of the other case studies.
-Our goal had been to equally highlight the ``major'' case studies of the match--girls and railroad workers.
+Our goal had been to equally highlight the case studies of
+the match--girls and railroad workers.
 While we occasionally bring to light other cases
@@
 % \subsection*{Decomposition}
-we'll bring more focus on the case studies of
-the match--girls and the railroad workers
-to elevate them to
-\citeauthor{grier2013computers}'s detail of human computers~(\citeyear{grier2013computers}).
+we'll bring more focus on these ``major'' case studies
+(the match--girls and the railroad workers)
+to afford them
+the scrutiny human computers earned.
 
@@
 the reorientation of payment for work from
-a time--centric approach to
-an output--centric one and how
-this discretization and decomposition of work allowed managers
-to track workers more fastidiously,
-evaluate the working styles and resultant performance,
+a time--centric to
+output--centric, and how
+this led to discretization and decomposition of work,
+allowing managers
+to track workers more minutely,
+evaluate working styles \& resultant performance,
 and encourage or discourage certain behavior much more granularly via
@@
 \section*{TOPIC SELECTION}
-R3 and R4 noted our decision to cluster crowdwork research around three questions
+R3 \& R4 note our decision to cluster crowdwork research around three questions
 that consolidated some research
---- for instance, the ``quantity-quality dilemma'' (R4), ``professional development'' (R3) and ``incentive structures'' (R3) --- into other broader topics.
-This critique is well taken. We will dedicate some space to reflect on
+--- for instance,
+the ``quantity-quality dilemma'' (R4),
+``professional development'' (R3),
+and ``incentive structures'' (R3) --- into other broader topics.
+This critique is well taken.
+We will dedicate some space to reflect on
 the decisions we made with regard to clustering research topics.
@@
 \section*{RELATED WORK}
-R4 and R5 offer a number of works
+R4 \& R5 offer a number of works
 (\cite[e.g.][]{williamson2016})
@@
 We will add this topic to the Discussion of our paper,
-integrating R4's suggested paper~(\cite[namely,][]{williamson2016}).
+integrating the resource R4 offered~(\cite{williamson2016}).
 Briefly, the literature on the history of labor does not
-position itself as stating that piecework is inherently ethical or unethical;
-the question, rather, is what conditions render it exploitative?
+frame the question as ``whether piecework is inherently ethical or unethical'',
+instead asking what conditions render it exploitative?
 
-This literature argues that the exploitation occurs when
-conditions put workers in harm's way, such as
+This literature we brought to bear suggests that exploitation occurs when
+conditions harm workers directly or indirectly, such as
 in sweatshops and agricultural work with pesticides, or where
@@
 
-The question then becomes
-whether the socio--technical infrastructure of Mechanical Turk
+The question then is
+whether socio--technical infrastructure like Mechanical Turk (AMT)
 and other marketplaces
-are putting workers in harm's way, underpaying, or overworking.
-Mechanical Turk itself does not directly require any amount of payment or work.
-However, its design does encourage employers to engage in such behaviors:
+similarly harm, underpay, or overwork workers.
+AMT itself does not directly require any amount of payment or work,
+but its design encourages employers to engage in such behaviors:
 piecework rates, for example, undervalue workers' task search time, and
-the task design interfaces explicitly encourage
-replication across multiple workers of uncertain quality rather than
-identifying one worker and paying them more.
+task design interfaces undeniably frame
+workers as unreliable by recommending
+replication with multiple workers rather than
+trusting and paying individual workers more.
+% let alone individuating them as people.
 

### 2016-11-23 drafts/labor/rebuttal.tex
exploitative upwards inflection

@@
 \begin{enumerate}
-  \item our contribution offers to fill some of the gaps in crowdwork and suggests potential futures
-        (carefully stated, offering
+  \item Our contribution %offers to fill some of the gaps in crowdwork and 
+  suggests potential futures, offering
             ``to narrow our estimates
               and thus to concentrate resources
-              in directions that are more likely to have useful payoffs''~(\cite{rosenberg1994exploring}));
+              in directions that are more likely to have useful payoffs''~(\cite{rosenberg1994exploring});
   and
-  \item our method of relating history to modern socio--technical systems may be
+  \item Our method of relating history to modern socio--technical systems may be
         a useful tool for researchers attempting
-        to make sense of (seemingly) new phenomena
-        (i.e., offering ``that past history is an indispensable source of information
-                          to anyone interested in characterizing technologies''~(\cite{rosenberg1982inside})).
+        to make sense of ostensibly new phenomena.
+        In other words, offering ``that past history is an indispensable source of information
+                          to anyone interested in characterizing technologies''~(\cite{rosenberg1982inside}).
 \end{enumerate}
@@
 To use the match--girls as an example, we'll address
-the reorientation of payment for work from
-a time--centric to
-output--centric, and how
-this led to discretization and decomposition of work,
+the reorientation of their payment for work from
+time--centric to
+output--centric.
+This reorientation led to further discretization and decomposition,
 allowing managers
-to track workers more minutely,
-evaluate working styles \& resultant performance,
-and encourage or discourage certain behavior much more granularly via
+to track the match--girls more minutely,
+evaluate their working styles \& resultant performance,
+and encourage or discourage certain behaviors much more granularly via
 precise disciplinary measures.
@@
 the notably adversarial relationships they had with their managers,
-whose above mentioned disciplinary methods later took on
+whose disciplinary methods later took on
 punitive, arbitrary qualities.
@@
 \section*{TOPIC SELECTION}
-R3 \& R4 note our decision to cluster crowdwork research around three questions
-that consolidated some research
+R3 \& R4 note that our decision to cluster crowdwork research around three questions
+consolidated some research
 --- for instance,
@@
 frame the question as ``whether piecework is inherently ethical or unethical'',
-instead asking what conditions render it exploitative?
+instead asking what conditions render it exploitative.
 

### 2016-11-30 drafts/labor/rebuttal.tex
IRB waiver of documentation paper thing

@@
 --- for instance,
-the ``quantity-quality dilemma'' (R4),
+the ``quantity--quality dilemma'' (R4),
 ``professional development'' (R3),

### 2016-11-30 admin/IRB/performances/assent/main.tex
IRB waiver of documentation paper thing

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp,titling,nameref,color,fancyhdr,paralist}
\usepackage[T1]{fontenc}
\usepackage[margin=1in,headheight=14pt]{geometry}
\usepackage[inline]{enumitem}
\usepackage[compact]{titlesec}
\PassOptionsToPackage{hyphens}{url}\usepackage[pdflang={en-US},pdftex]{hyperref}

\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=red,
  breaklinks=true,
  hypertexnames=false
}
\usepackage[dvipsnames]{xcolor}

\newcommand{\contact}[1]{{\color{red}#1}}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
\pagenumbering{gobble}

\titlespacing*{\section}{0em}{0.5em}{-0.5em}

\begin{document}
\renewcommand{\headrulewidth}{0pt}

\newenvironment{headerfont}{\color{gray}\fontfamily{ptm}\selectfont}{}
\pagestyle{fancy}{
  \lhead{\begin{headerfont}Stanford University\end{headerfont}}
  \chead{\begin{headerfont}\textbf{\underline{Waiver of Documentation}}\end{headerfont}}
  \rhead{\textbf{Approval Date}: \underline{November 30 2016}}
}

\section*{{\LARGE Ethnographic Study of Online Performance Artists}}

\section*{What's this study about?}
We're trying to learn about
the relationships online performers have with each other and
with the
organizations that run
the platforms that they stream and
upload to.
Some of this touches on things like
dealing with inter--personal issues and
how the community deals with major events
as well as questions about what it's like to
make your income from online public performances.

\section*{What are my choices?}
You can choose not to participate in the study at all, or
you can respond to some questions but not others.
You can later ask that I not use some or
any of the information from any previous conversations
(until that data,
like an anonymized quotation, 
is published in a form that I can't change after the fact).
If you tell me you want to withdraw or
simply not participate in the first place,
I'll delete the data you gave
(if any) and that'll be the end of it.

\section*{Will anyone know I'm in the study?}
Only if you want ---
the only people in our research group who will know individual participants' names
will be people directly involved in the study.
The only reason other people will know you participated will be if
you tell them, or you tell me
to mention you when I email them.

\section*{What else should I know?}
If you're interested in participating, you should know
\begin{compactitem}
  % \item Since it would be too difficult to get a signature from participants,
  %       (to say nothing of verifying the signature),
  %       we're using a ``\textbf{waiver of documentation}'',
  %       so this is purely informational.
  \item The first conversation will probably take 30--45 minutes.
        Later conversations (if you're okay with it) would likely be shorter
        (mostly to clarify things or ask follow--up questions).
  \item I can't compensate you (or anyone, really) for their time.
  \item I plan on recording the audio of our conversation
        so I can transcribe what you said
        (so I don't misremember or misquote anything).
        Once I have a transcript, I'll delete the audio.
\end{compactitem}

\section*{Who can I talk to about the study?}
If you have any questions or concerns with the study,
you can talk to the researchers directly, or
you can talk to Stanford's Institutional Review Board (IRB),
who oversee our research and make sure participants are treated ethically.

\textbf{Talk to the researchers}:
If you have any questions or problems related to the study,
feel free to contact me, \contact{Ali Alkhatib},
by phone at \contact{408--348--2720} or
by email (\contact{\url{ali.alkhatib@cs.stanford.edu}}).

\textbf{Talk to the IRB}:
If you're not satisfied with how this study is being conducted, or
if you have any concerns, complaints, or
questions about the research or
your rights as a participant, please feel free to contact the
(IRB) to speak to someone independent of the research team at
\contact{(650)--723--2480} or
toll free at \contact{1--866--680--2906}.
You can also email \url{irb2-manager@lists.stanford.edu} or
write to:

\begin{center}\contact{Stanford IRB, Stanford University,\\
         3000 El Camino Real, Five Palo Alto Square, 4th Floor,\\
         Palo Alto, CA 94306}\end{center}

\end{document}

### 2016-11-30 admin/IRB/performances/recruitment/main.tex
IRB waiver of documentation paper thing

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\usepackage{nameref,color}
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  % pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  % pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
% \setlength{\parskip}{.4em}
\pagenumbering{gobble}
\begin{document}
Hi,

My name's Ali Alkhatib. % link to homepage
I'm a PhD student at Stanford.
I'm studying professional online performers
(like YouTubers, Twitch streamers, etc\dots)
--- basically anyone who does it as a major source of income.

Specifically, I'm looking into
the relationships people have with their platforms
--- so for instance, what it's like for
YouTubers to make their living from the videos they put on YouTube,
talking to and working with YouTube's admins,
and dealing with controversial ``meta'' topics like
sudden changes to rules on the platform.

I'm emailing because I saw your
[\textbf{online streaming/broadcasting channel}]
and I thought your perspective would be interesting.
I was wondering if you'd be willing to chat with me about anything relating to
[\textbf{streaming|YouTubing|etc\dots}]. % make it clear this'll be anonymous if desired

% What I had in mind is a chat that would probably take about 30 minutes,
% possibly with a few follow up conversations.
% I was figuring to video chat via
% FaceTime or Google Hangout or something for the first
% ``long'' conversation, and email for later follow ups.

If you're open to chatting about online performances and stuff, feel free to
\href{mailto:ali.alkhatib@cs.stanford.edu}{email me}
(\texttt{ali\allowbreak .alkhatib\allowbreak @cs\allowbreak .stanford.edu}) or
call me at my cell phone: ${+1~(408)~348-2720}$

PS: if you're curious what sort of research I generally do,
feel free to poke around at \href{https://ali-alkhatib.com}{my personal site}
(\texttt{https://\allowbreak ali-\allowbreak alkhatib\allowbreak .com}).

\end{document}

### 2016-12-02 qual_management/main.tex
among other things, a (maybe really ugly-looking) business card

@@
 \documentclass[11pt]{article}
-\usepackage{balance,graphics,setspace,parskip,times,hyperref,titling,microtype}
+\usepackage{balance,graphics,setspace,parskip,sourceserifpro,hyperref,titling,microtype}
 \usepackage[margin=1in]{geometry}
@@
 work requiring more training and skill (for example, translating or programming)
-are either verified by individual \textit{requesters} (e.g. Amazon Mechanical Turk)
-or are verified by the \textit{platform} itself (e.g. UpWork).
-While the \textit{UpWork} model avoids needless repetitive work by
+are either verified by individual \textbf{requesters} (e.g. Amazon Mechanical Turk)
+or are verified by the \textbf{platform} itself (e.g. UpWork).
+While the \textbf{UpWork} model avoids needless repetitive work by
 generally consolidating qualifying exams at the platform level,
@@
 
-% \textit{I want to communicate something along the lines of how
+% \textbf{I want to communicate something along the lines of how
 % these labor markets shouldn't be expected
@@
         Workers can use this information to identify specialization more easily,
-        reducing wasted time \textit{searching for} or \textit{working on} suboptimal tasks.
+        reducing wasted time \textbf{searching for} or \textbf{working on} suboptimal tasks.
         Requesters can benefit from this information by

### 2016-12-05 drafts/aggregator/introduction.tex
proposal (why didn't i put this up sooner?)

\documentclass[main]{subfiles}

\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\section{Introduction}\label{sec:introduction}

Online labor markets ranging a variety of industries of (largely) service work
--- e.g. on--demand drivers (Uber, Lyft, Sidecar, etc\dots) and
    information work (Amazon Mechanical Turk, Upwork, CrowdFlower, etc\dots) ---
have struggled for some time with pernicious, in some cases \textit{growing},
difficulty assessing the quality of its workforce,
especially as it applies to a given task.
In other words, whether a worker on Amazon Mechanical Turk (AMT) is likely
to complete a task ``accurately'' remains an open question,
leading researchers to address this pervasive problem in a number of ways
\cite

% Labor market platforms ranging industries and styles of work
% (e.g. Uber, TaskRabbit, UpWork, and Amazon Mechanical Turk)
% have struggled for years with persistent
% (generally growing)
% challenges relating to worker qualifications.
% These issues range a wide spectrum:
% in some cases, worker qualifications are non--transferable,
% leading requesters to ``re--invent the wheel'' as they attempt to determine
% in their own way whether a potential worker is qualified and reliable.

% Problems determining a worker's qualifications start on day 1;
% labor markets begin their relationships with new workers
% almost entirely uncertain about the worker's competence in any type of task.
% Gathering this information through qualification exams is generally
% time--consuming and costly.

% Challenges mount as workers' skill sets develop;
% work requiring more training and skill (for example, translating or programming)
% are either verified by individual \textit{requesters} (e.g. Amazon Mechanical Turk)
% or are verified by the \textit{platform} itself (e.g. UpWork).
% While the \textit{UpWork} model avoids needless repetitive work by
% generally consolidating qualifying exams at the platform level,
% these labor platforms nevertheless find themselves in the unenviable
% (and often unexpected) position of
% having to develop new qualifications exams to outpace would--be cheaters.

\onlyinsubfile{
% \balance{}
\printbibliography{}
}

\end{document}

### 2016-12-05 drafts/labor/abstract.tex
proposal (why didn't i put this up sooner?)

@@
 \documentclass[trackingWork]{subfiles}
-\onlyinsubfile{\usepackage{xr-hyper}
+\onlyinsubfile{
+  \usepackage{xr-hyper}
   \usepackage{hyperref}
@@
 A large and growing body of scholarship has attempted to predict
-the socio--technical outcomes of this shift, especially along three threads: %. This question has been dominated by three threads:
+the socio--technical outcomes of this shift, especially along three threads:
 \begin{inlinelist}

### 2016-12-05 drafts/labor/cases.tex
proposal (why didn't i put this up sooner?)

--- 2016-09-21 drafts/labor/cases.tex
@@
 
-\msb{this needs a transition into it.
-Before you lay out our method, lay out what we're trying to do.
-What's our goal here?
-THEN the method can explain how we achieved it.
-This might involve switching the order of some of the text here}
-
-We look at lots of papers that use the term ``crowdsourcing'' or ``crowd work''
-in the abstracts and titles of their papers,  especially the papers that we cited earlier, to try
-to answer the questions that we posed in the introduction.
-Then we look to the piecework literature to see whether and to what extent
-piecework answers the questions we, crowdsourcing researchers, have asked.
-We then see whether and how crowdsourcing as we know it has differed from piecework,  and how that affects the predictions and conclusions made in the piecework literature.
-
-%\subsection{What are the limits of crowdsourcing?}
 \topic{Research in crowdsourcing has spent the better part of a decade
@@
   identifying challenges to this form of labor,    overcoming them through novel designs of work--flows and processes, and
-  repeating the process
-\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
+  repeating the process~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
 The question that has emerged
@@
 \begin{inlinelist}
-  \item \nameref{sec:complexity},    \item \nameref{sec:decomposition}, and
-  \item \nameref{sec:relationships}
-\end{inlinelist}.
+  \item \nameref{sec:complexity},
+  \item \nameref{sec:decomposition}, and
+  \item \nameref{sec:relationships}.
+\end{inlinelist}
 We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
@@
 \end{itemize}
-
-% so CW?
-% mechanism/limitation in pw?
-% differences in cw? tech/access to intelligence?
-% predicted outcome then? -> decompose effort/expertise access
-% possible quirks/diffs?
-
-% what can be decomposed? what can be evaluated?
-% technology/turuing completeness gets us a win
-% predefined expertise? you can find someone almost instantly;
-% ease of doing this makes it qualitatively different due to Channel Factors (diverting a channel/river)
-% huddler paper cites some decomposition stuff workflows are good/bad at things
-% 3-way slack w/msb and mav
-% academy of management 
-% van der ven paper ~1972 limits of workflows cited in daniela's flash orgs paper
 }
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
+% \onlyinsubfile{
+%   \balance{}
+%   \printbibliography
 %   \nobalance{}
-  % \clearpage
-  \nobalance{}
-  }
-  % \begin{appendices}
-% \section{Graveyard}
+%   }
 \endnotetext{

### 2016-12-05 drafts/labor/conclusion.tex
intro revisions
proposal (why didn't i put this up sooner?)

@@
 \section{Conclusion}
-Crowd work and on--demand work are not new: they are modern instantiations of piecework.
-In this paper, we reconsider three major research questions in crowd work using the lens of piecework: \begin{inlinelist}
+Crowd work and on--demand work are not new: they are contemporary instantiations of piecework.
+In this paper, we reconsider three major research questions in crowd work using the lens of piecework:
+\begin{inlinelist}
   \item ``\namerefl{sec:complexity}?'';
-  \item ``\namerefl{sec:decomposition}?'';
-        and
+  \item ``\namerefl{sec:decomposition}?''; and
   \item ``\namerefl{sec:relationships}?''
@@
 To do so, we draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon.
-% predictions and to contextualize the answers that crowd work researchers have already uncovered.
-% We subsequently take this historical framing to ask broader questions of crowd work, including among others what the future of crowd work holds for workers.
 Reciprocally, we believe that modern crowd work will teach us about the broader phenomenon of piecework as well.

### 2016-12-05 drafts/labor/decomposition.tex
proposal (why didn't i put this up sooner?)

@@
 
-At its core, on--demand work has been enabled by decomposition of large goals into many small tasks.
-As such, one of the central questions in the literature is how to design these microtasks, and which kinds of tasks are amenable to decomposition.
-In this section, we place these questions in the context of piecework's Tayloristic evolution.
+At its core,
+on--demand work has been enabled by decomposition of large goals into many small tasks.
+As such,
+one of the central questions in the literature is how to design these microtasks,
+and which kinds of tasks are amenable to decomposition.
+In this section,
+we place these questions in the context of piecework's Tayloristic evolution.
 
@@
 \subsubsection{\crowdworkpers}
-\topic{Many contributions to the design and engineering of crowd work consist of creative methods for decomposing goals.} Even when tasks such as writing and editing cannot be reliably performed by individual workers, researchers demonstrated that decompositions of these tasks into workflows can succeed~\cite{crowdForgeKittur,bernsteinSoylent,writingMicroTasks,Nebeling:2016:WCW:2858036.2858169}.
-These decompositions typically take the form of workflows, which are algorithmic sequences of tasks that manage interdependencies~\cite{Bigham2014}. 
-Workflows often utilize a first sequence of tasks to identify an area of focus (e.g., a paragraph topic~\cite{crowdForgeKittur}, an error~\cite{bernsteinSoylent}, or a concept~\cite{Yu2016a,Yu2016b} and a second sequence of tasks to execute work on that area.
-This decomposition style has been successfully applied across many areas, including food labeling~\cite{noronha2011platemate}, brainstorming~\cite{siangliulue2015toward,yu2014distributed}, and accessibility~\cite{lasecki2013chorus,lasecki2012real,Lasecki2011}.
+\topic{Many contributions to the design and engineering of crowd work
+consist of creative methods for decomposing goals.}
+Even when tasks such as writing and editing cannot be reliably performed by individual workers,
+researchers demonstrated that
+decompositions of these tasks into workflows can succeed~\cite{crowdForgeKittur,
+                                                               bernsteinSoylent,
+                                                               writingMicroTasks,
+                                                               Nebeling:2016:WCW:2858036.2858169}. 
+These decompositions typically take the form of workflows,
+which are algorithmic sequences of tasks that manage interdependencies~\cite{Bigham2014}. 
+Workflows often utilize a first sequence of tasks to identify an area of focus (e.g.,
+a paragraph topic~\cite{crowdForgeKittur},
+an error~\cite{bernsteinSoylent},
+or a concept~\cite{Yu2016a,Yu2016b} and a second sequence of tasks to execute work on that area. 
+This decomposition style has been successfully applied across many areas,
+including food labeling~\cite{noronha2011platemate},
+brainstorming~\cite{siangliulue2015toward,yu2014distributed},
+and accessibility~\cite{lasecki2013chorus,lasecki2012real,Lasecki2011}.
 
-\topic{If decomposition is key to success in crowd work, the question arises: what can, and can't, be decomposed?} 
-Moreover, how thinly can work be sliced and subdivided into smaller and smaller tasks? 
-The general trend has been that smaller is better, and the microtask paradigm has emerged as the overwhelming favorite~\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
+\topic{If decomposition is key to success in crowd work,
+the question arises: what can,
+and can't,
+be decomposed?} 
+Moreover,
+how thinly can work be sliced and subdivided into smaller and smaller tasks? 
+The general trend has been that smaller is better,
+and the microtask paradigm has emerged as the overwhelming favorite~\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
 This work illustrates a broader sentiment in both the study and practice
-of crowd work, that microtasks should be designed resiliently against the variability of workers, preventing a single errant submission from impacting the agenda of the work as a whole %, fully exploiting the abstracted nature of each piece of work
+of crowd work,
+that microtasks should be designed resiliently against the variability of workers,
+preventing a single errant submission from impacting the agenda of the work as a whole %,
+fully exploiting the abstracted nature of each piece of work
 ~\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
-In this sense, finer decompositions are seen as more robust --- both to interruptions and errors~\cite{cheng2015break} --- even if they incur a fixed time cost.
-At the extreme, recent work has attempted demonstrated microtasks that take seconds~\cite{Vaish:2014:TCC:2611222.2556996,Cai:2015:WLW:2702123.2702267} or even tenths of a second~\cite{embracingErrorKrishna}.
-%Earlier we discussed \citeauthor{cheng2015break}'s work measuring the impact that interruption has on worker performance \cite{cheng2015break}.%While \citeauthor{cheng2015break} found costs with breaking tasks into smaller components in the form of higher cumulative time to complete (albeit much shorter real time to complete, owing to parallelization)%
-However, workers perform better when similar tasks are strung together ~\cite{delayAndOrderLasecki}, or chained and arranged to maximize  the attention threshold of workers~\cite{Cai:2016:CRI:2858036.2858237}.
+In this sense,
+finer decompositions are seen as more robust
+--- both to interruptions and errors~\cite{cheng2015break} ---
+even if they incur a fixed time cost.
+At the extreme,
+recent work has attempted demonstrated
+microtasks that take seconds~\cite{Vaish:2014:TCC:2611222.2556996,Cai:2015:WLW:2702123.2702267}
+or even fractions of a second~\cite{embracingErrorKrishna}.
+However,
+workers perform better when similar tasks are strung together~\cite{delayAndOrderLasecki},
+or
+chained and arranged to maximize the attention threshold of workers~\cite{Cai:2016:CRI:2858036.2858237}.
 Despite this,
-%rather than attempt to minimize the error rates in micro--tasks
-%, as \citeauthor{Kinnaird:2012:WTM:2389176.2389219} suggested%
 we as a community have leaned \textit{into} the peril of
@@
 
-\topic{The general lesson has been that the more micro the task, and the more fine the decomposition, the greater the risk that workers lose context necessary to perform the work well.}
-For example, workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017}, interpret tasks in different ways~\cite{kairam2016parting}, and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
-Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher--level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.
+\topic{The general lesson has been that the more micro the task,
+and the more fine the decomposition,
+the greater the risk that workers lose context necessary to perform the work well.}
+For example,
+workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017},
+interpret tasks in different ways~\cite{kairam2016parting},
+and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
+Research has sought to ameliorate this issue by
+designing workflows help workers ``act with global understanding when
+each contributor only has access to local views''~\cite{verroios2014context},
+typically by automatically or manually generating higher--level representations
+for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.
 
-% \topic{The crowdsourcing research into work decomposition
-% has largely focused on minimizing the additional context necessary to do tasks%
-% , and making it easier to do tasks with less time.}
-% This first thread is
-% perhaps best described by \citeauthor{verroios2014context} as
-% making crowd workers ``\dots able to act with
-% global understanding when each contributor only has access to local views''
-% \cite{verroios2014context}.
-% With the exception of a few cases
-% (specifically, \citeauthor{Kinnaird:2012:WTM:2389176.2389219}'s work
-% which finds that greater work context fosters more reliably high--quality work)%
-% , the micro task paradigm has emerged as the overwhelming favorite
-% \cite{selfsourcingTeevan2014,selfsourcingTeevan2016%
-% ,       cheng2015break,Kinnaird:2012:WTM:2389176.2389219}.
+\topic{As the additional context necessary to complete a task diminishes,
+the invisible labor of \textit{finding} tasks~\cite{martin2014being} has arisen as a major issue.}
+\citeauthor{taskSearch} illustrate the task search challenges on AMT.
+Workers seek out good requesters~\cite{martin2014being} and then 
+``streak'' to perform many tasks of that same type~\cite{taskSearch}.
+and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}),
 
-\topic{As the additional context necessary to complete a task diminishes, the invisible labor of \textit{finding} tasks~\cite{martin2014being} has arisen as a major issue.}
-\citeauthor{taskSearch} illustrate the task search challenges on AMT.%
-Workers seek out good requesters~\cite{martin2014being} and then  ``streak'' to perform many tasks of that same type~\cite{taskSearch}.
-%, and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}), % does anyone care for project names?
-%while other work designs tasks around gap time (\textit{Twitch Crowdsourcing} \& \textit{Wait--Learning}) \cite{taskSearch,KucherbaevReLauncher,Vaish:2014:TCC:2611222.2556996,       Cai:2015:WLW:2702123.2702267}.
-
-% \topic{One of the emergent properties of micro--tasks has been the relative cost of
-%     \textit{finding} worthwhile tasks.}
-%     The research community has documented and to some extent attempted to intervene in
-%     the discovery of worthwhile tasks
-%     \cite{taskSearch}.
-Researchers have reacted by designing task recommendation systems (e.g.,~\cite{Cosley:2007:SUI:1216295.1216309})
-%attempts to address this by directing workers to tasks through ``intelligent task routing'' \cite{Cosley:2007:SUI:1216295.1216309}.
-and others focused on minimizing the amount of time that people need to spend doing anything other than the work for which they are paid~\cite{callison2014crowd}.
+Researchers have reacted by designing task recommendation systems~\cite[e.g.][]{Cosley:2007:SUI:1216295.1216309}
+and others focused on
+minimizing the amount of time that people need to spend doing anything other than
+the work for which they are paid~\cite{callison2014crowd}.
 
 \subsubsection{\pieceworkpers}
-\citeauthor{Brown01041990} inquired from another direction, asking
+\citeauthor{Brown01041990} inquired from another direction,
+asking
 what limited the adoption of piecework in industries that otherwise gravitated toward it
-(in the case studies he examined, this mostly focused on railway engineers),
+(in the case studies he examined,
+this mostly focused on railway engineers),
 ultimately arguing that factors such as the nature of the work design
-(specifically, the homogeneity of tasks) and the costs associated with adopting a piecework model
+(specifically,
+the homogeneity of tasks) and the costs associated with adopting a piecework model
 were the major contributing factors that determined the use of piecework~\cite{Brown01041990}.
@@
 \topic{Piecework became an important factor in the war effort for the Second World War,
-cementing its role not only in American factories, but in industrial work around the world.}
+cementing its role not only in American factories,
+but in industrial work around the world.}
 The 1930s represented a boom for piecework on an unprecedented scale,
@@
 One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating},
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women \cite{honey1985creating},
 was a pieceworker~\cite{davies2014origins}.
@@
 when Airy employed young boys at the Greenwich Observatory who
-``possessed the basic skills of mathematics, including
-`Arithmetic, the use of Logarithms, and Elementary Algebra'~''
+``possessed the basic skills of mathematics,
+including
+`Arithmetic,
+the use of Logarithms,
+and Elementary Algebra'~''
 to compute astronomical phenomena~\cite{grier2013computers}.
 The work that Airy solicited resonates with modern crowd work for several reasons.
-First, work output was quickly verifiable;
+First,
+work output was quickly verifiable;
 Airy could assign variably skilled workers to compute values,
 and have other workers check their work.
-Second, tasks were discrete --- that is, independent from one another.
-Finally, workers could be trained on a very narrow subset of mathematical skills to be
+Second,
+tasks were discrete
+--- that is,
+independent from one another.
+Finally,
+workers could be trained on a very narrow subset of mathematical skills to be
 sufficiently qualified to do this work.
@@
 Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out, piecework most greatly benefits the instrumented measurement of workers, but certainly in Ford and Taylor's time, highly instrumented, automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-As a result, the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
-\msb{This subsection is too shallow and needs a bit more. For example, can you give examples of the last point?}
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers,
+but certainly in Ford and Taylor's time,
+highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+As a result,
+the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
 
-\topic{Piecework researchers enumerate a number of problems with the decomposition of work, and the conflicting pressures managers and workers put forth.}
+\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
+and the conflicting pressures managers and workers put forth.}
 \citeauthor{bewley1999wages} in particular points out that
 the approach of paying workers by the piece is
-``\dots~not practical for workers doing many tasks, because of
+``\dots~not practical for workers doing many tasks,
+because of
 the cost of establishing the rates and because
@@
 Ultimately, \citeauthor{bewley1999wages} argues that
-``[piecework is] infeasible, because \dots
+``[piecework is] infeasible,
+because \dots
 total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
@@
 outline
-	- measurement is more precise, so decomposition is deeper
-	- not a single position, but a marketplace
+	- measurement is more precise,
+  so decomposition is deeper
+	- not a single position,
+  but a marketplace
 \end{comment}
 
-\topic{Where measurement and instrumentation were limiting factors for historical piecework, computation has changed the situation so that a dream of scientific management and Taylorism
+\topic{Where measurement and instrumentation were limiting factors for historical piecework,
+computation has changed the situation so that a dream of scientific management and Taylorism
 --- to measure every motion at every point throughout the workday and beyond ---
-is not only doable, but trivial~\cite{waltz2012quantified}.}
+is not only doable,
+but trivial~\cite{waltz2012quantified}.}
 Where \citeauthor{10.2307/23702539} directly implicates measurement as
@@
 preventing scientific management from being fully utilized,
-%the difficulty of tracking work \& workers, no longer exists
+%the difficulty of tracking work \& workers,
+no longer exists
 %\cite{10.2307/23702539}.}
-modern crowd work is measuring and modeling every click, scroll, and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
+modern crowd work is measuring and modeling every click,
+scroll,
+and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
 The result is that on--demand work can articulate and track far more carefully than piecework historically could.
@@
 the relative ease with which the metaphorical ``assembly line'' can be changed.}
-Historical manufacturing equipment could not quickly be assembled, edited, and redeployed~\cite{hu1961parallel}.
-In contrast, today system--designers can share, modify, and instantiate environments
+Historical manufacturing equipment could not quickly be assembled,
+edited,
+and redeployed~\cite{hu1961parallel}.
+In contrast,
+today system--designers can share,
+modify,
+and instantiate environments
 like sites of labor in a few lines of code~\cite{lessig2006code,turkitLittle}.
@@
 pacing,
-interruptions, and
+interruptions,
+and
 other factors in piecework that would have been
-all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
+all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,
+                                                              Cai:2016:CRI:2858036.2858237,
+                                                              cheng2015break,
+                                                              measuringCrowdsourcingCheng,
+                                                              embracingErrorKrishna}.
 
-Third, modern crowd work has sliced work to such small scales that the marginal activities
+Third,
+modern crowd work has sliced work to such small scales that the marginal activities
 --- things like finding work and cognitive task switching ---
@@
 In the historical case of piecework,
-moving metallurgical tools, mining equipment, or
+moving metallurgical tools,
+mining equipment,
+or
 other industry materials would have been prohibitively difficult and slow;
@@
 The result is that crowd workers are more free agents than historically was the case.
-However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take--home earnings than before.
+However,
+because they spend significant time searching for tasks,
+the piece rate is less a good estimate of take--home earnings than before.
 
 \subsubsection{\implication}
-If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
-Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved, or near enough that they no longer limit task decomposition.
-Now, not just tasks, but entire workers' histories~\cite{hata2017glimpse}, can be collected and analyzed in detail.
+If measurement precision limited the depth of decomposition for piecework historically,
+as \citeauthor{10.2307/23702539} argues,
+then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
+Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved,
+or near enough that they no longer limit task decomposition.
+Now,
+not just tasks,
+but entire workers' histories~\cite{hata2017glimpse},
+can be collected and analyzed in detail.
 
-However, decomposition has hit a second bottleneck: cognition. 
-Task switching costs and other cognitive costs make it difficult to work tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
-There will of course be tasks that can be decomposed without much context, and these will form the most fine--grained of microtasks.
-However, other tasks cannot be freed from context --- for example, logo design requires a deep understanding of the client and their goals.
-In part due to this limitation, 99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
+However,
+decomposition has hit a second bottleneck: cognition. 
+Task switching costs and other cognitive costs make it difficult
+to work on tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
+There will of course be tasks that can be decomposed without much context,
+and these will form the most fine--grained of microtasks.
+However, other tasks cannot be freed from context
+--- for example,
+logo design requires a deep understanding of the client and their goals.
+In part due to this limitation,
+99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
 
-So, ultimately, the levels of decomposition are likely to follow the contours of context required.
+So,
+ultimately,
+the levels of decomposition are likely to follow the contours of context required.
 Low--context work will be extremely highly decomposed.

### 2016-12-05 qual_management/main.tex
proposal (why didn't i put this up sooner?)

--- 2016-11-21 qual_management/main.tex
@@
 \documentclass[11pt]{article}
-\usepackage{balance,graphics,setspace,parskip,times,hyperref,titling,microtype}
+\usepackage{balance,graphics,setspace,parskip,fourier,hyperref,titling,microtype}
 \usepackage[margin=1in]{geometry}
@@
 \bibliography{references}
-\makeatletter
-\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
-\begin{flushleft}
-  \textbf{\LARGE{\@title}}
-
-  \@author
-\end{flushleft}\egroup
-}
-\makeatother
-\title{Externalizing Worker Qualifications}
-\author{Ali Alkhatib}
 
 \begin{document}
-\maketitle{}
-
+\section*{{\LARGE Externalizing Worker Qualifications}}
 \section*{Background}
@@
 (e.g. Uber, TaskRabbit, UpWork, and Amazon Mechanical Turk)
-have struggled for years with persistent and
-in some cases growing
+have struggled for years with persistent
+(generally growing)
 challenges relating to worker qualifications.
@@
   \item By turning the reputation management of workers into an external entity
-        (agnostic to labor platform),
+        (agnostic to the labor platform),
         designers and developers can focus more fully on the marketplace itself.
@@
 \section*{Measurable Variables}
+We see two measurable variables that might be relevant to
+the goal of better, or higher--quality, crowd (and other on--demand) work.
+The first, accuracy of predicting worker acceptance rates, is fairly straightforward.
+The second, determining whether workers better self--evaluate in their selection of work
+based on more comprehensive awareness of their reputation (or credibility).
+We'll call this latter dimension ``worker self--specialization'' to allude to
+the emergent property of workers identifying their own niche expertises and
+selecting for tasks in those domains.
+
+\subsection*{Worker Accuracy}
+Determining a worker's likelihood of completing a task correctly is
+fundamentally the goal of using past worker approval rates
+(like on AMT, Upwork, and other platforms)
+as a low pass filter for workers.
+The thinking goes that past performance is an indicator towards future performance.
+
+The baseline approaches requesters have used typically falls into two paths:
+\begin{enumerate*}
+  \item Relying principally on the platform's rating system as a broad--spectrum filter, and
+  \item Implementing one's own rating or qualification system as a determiner of
+        whether or not to allow someone to do tasks of similar nature.
+\end{enumerate*}
+
+We can conduct post--facto analyses of workers' data
+(aggregated from their various work platforms) to see
+if any combination and/or weighting of their
+worker profiles from multiple platforms
+more accurately predicts their likelihood of doing good work than the ``baseline''
+(that is, informed solely by the platform on which the work was solicited).
+
+To run this analysis, we would need to solicit new information from workers
+than they've traditionally shared.
+Specifically,
+workers would need to be willing to share approval rates from other work platforms
+(such as AMT or CrowdFlower).
+
+\subsection*{Worker Self--Specialization}
+The next dimension that we could evaluate would be
+to use this data to better--inform workers about their approval trends.
+The best way to run this experiment would be
+to randomly assign some number of workers to a control group and
+some others to an experimental group, wherein these workers have the opportunity
+(after sharing their aggregated profiles with us)
+to see at a higher level on what \textit{types} of work they typically fare better.
+
+The operationalization of this dimension would almost certainly be
+the measurement of overall approval rates over time after intervention.
+To make this more concrete, we would provide workers with guidance on what
+sorts of work they typically get approved, perhaps
+helping to guide them to those kinds of work more actively than otherwise.
+
+There are problems with the experimental design described at the outset.
+For one thing, finding willing participants might be difficult, and
+randomly sampling for potential participants would make
+recruitment of a sufficiently large cohort difficult.
+Moreover, the self--selection of workers into this group
+(or rather, out)
+would almost certainly compromise the ``randomness'' of the assignment approach.
+
+Instead, it may be better simply to allow
+a limited number of interested participants
+to sign up to participate in a ``pilot program'',
+incorporating their work data from other platforms.
+While it's true that this would otherwise be ``giving up'' on randomized experimentation,
+the question of
+what types of workers would prefer this sort of work is
+its own important area of discussion that we shouldn't try to evade.
 
 \section*{First Steps}
+The first step is to solicit workers
+(either on AMT, Upwork, or another platform)
+who may be interested in participating in a study of this nature.
+We'll ask these participants to self--report their own approval rates from each platform,
+perhaps asking for an annotated history of their tasks
+(if feasible).
+
+Two methods of evaluation may be viable, depending on the homogeneity of the tasks.
+One approach would be
+to train a machine learning model on a subset of the aggregated data
+and test it against other subset of that data.
+
+The other approach would be
+to investigate the nature of the tasks that are being done,
+categorize them by similarity, and
+focus on trying to predict worker quality from the aggregated approval rates across platforms
+(for instance, looking at the worker's approval rates for translations across AMT and Upwork).
+
+% During this time, or perhaps later,
+% we would run workers through a number of tasks.
+% At first, an \textit{a priori} estimate of
+% their likelihood to return a correct response would be unnecessary,
+% as we would initially be exploring the data for indicators
+% across the aggregated data
+% that may be predictive.
+
+% We could either use a training--test approach
+% (a fairly straightforward approach in machine learning ---
+% to train on a portion of the dataset and
+% test on another (specifically, one that the learning algorithm has never seen)).
+
+A later step is
+to build a prototypical server serving the basic needs of
+an aggregated reputation service.
+That is,
+if a client queries the service for information about a worker on Upwork,
+the client will receive a reasonably comprehensive snapshot of the worker's
+history on other platforms.
 

### 2016-12-08 drafts/labor/complexity.tex
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.

@@
 
-\topic{\citeauthor{grier2013computers} gives early accounts of
-a piecework strategy in Airy's creation of
-the British Nautical Almanac~\cite{grier2013computers}.}
-Airy's goal was complex
---- mathematical calculations to produce tables that would
-allow sailors to locate themselves by starlight from sea.
-Many of his contributors did not have high--level mathematical training,
-so Airy broke down the task into simpler calculations and
-distributed them by mail,
-accomplishing the complex goal through piecework tasks that paid little.
+% \topic{\citeauthor{grier2013computers} gives early accounts of
+% a piecework strategy in Airy's creation of
+% the British Nautical Almanac~\cite{grier2013computers}.}
+% Airy's goal was complex
+% --- mathematical calculations to produce tables that would
+% allow sailors to locate themselves by starlight from sea.
+% Many of his contributors did not have high--level mathematical training,
+% so Airy broke down the task into simpler calculations and
+% distributed them by mail,
+% accomplishing the complex goal through piecework tasks that paid little.
 
-However, when piecework entered the American economy,
-it was not used for complex work.
-One reason for low complexity was workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
-that apprenticeships imparted \cite{hart2013rise}.
-So, initially piecework arose for farm work, and as
-\citeauthor{hughRaynbirdTaskWork} and others discuss,
-the practice remained relatively obscure until
-it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
-Complexity levels remained low at the turn of
-the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
-However, writers of the time focused their attention on
-wage \cite{burton1899commercial} and
-management regimes \cite{norton1900textile}
-rather than training.
+% However, when piecework entered the American economy,
+% it was not used for complex work.
+% One reason for low complexity was workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
+% that apprenticeships imparted \cite{hart2013rise}.
+% So, initially piecework arose for farm work, and as
+% \citeauthor{hughRaynbirdTaskWork} and others discuss,
+% the practice remained relatively obscure until
+% it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+% Complexity levels remained low at the turn of
+% the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+% However, writers of the time focused their attention on
+% wage \cite{burton1899commercial} and
+% management regimes \cite{norton1900textile}
+% rather than training.
 

### 2016-12-08 drafts/labor/decomposition.tex
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.

@@
 
-\topic{Piecework became an important factor in the war effort for the Second World War,
-cementing its role not only in American factories,
-but in industrial work around the world.}
-The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.
-As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-He attributes this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women \cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
+% \topic{Piecework became an important factor in the war effort for the Second World War,
+% cementing its role not only in American factories,
+% but in industrial work around the world.}
+% The 1930s represented a boom for piecework on an unprecedented scale,
+% especially among engineering and metalworking industries.
+% As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
+% --- and more broadly the first half of the \nth{20} century ---
+% as the ``heyday'' of the use of piecework.
+% He attributes this to the shortage of male workers,
+% who would have gone through a conventional apprenticeship process
+% affording them more comprehensive knowledge of the total scope of work.
+% One might reflect on the observation that ``Rosie the Riveter'',
+% an icon of \nth{20} century America who
+% represented empowerment and opportunity for women \cite{honey1985creating},
+% was a pieceworker~\cite{davies2014origins}.
 
-\topic{The research community relating to
-piecework and labor
-has been wrestling with the decomposition of work for centuries.}
-The beginnings of
-systematic task decomposition
-stretch back as far as the \nth{19} century,
-when Airy employed young boys at the Greenwich Observatory who
-``possessed the basic skills of mathematics,
-including
-`Arithmetic,
-the use of Logarithms,
-and Elementary Algebra'~''
-to compute astronomical phenomena~\cite{grier2013computers}.
-The work that Airy solicited resonates with modern crowd work for several reasons.
-First,
-work output was quickly verifiable;
-Airy could assign variably skilled workers to compute values,
-and have other workers check their work.
-Second,
-tasks were discrete
---- that is,
-independent from one another.
-Finally,
-workers could be trained on a very narrow subset of mathematical skills to be
-sufficiently qualified to do this work.
+% \topic{The research community relating to
+% piecework and labor
+% has been wrestling with the decomposition of work for centuries.}
+% The beginnings of
+% systematic task decomposition
+% stretch back as far as the \nth{19} century,
+% when Airy employed young boys at the Greenwich Observatory who
+% ``possessed the basic skills of mathematics,
+% including
+% `Arithmetic,
+% the use of Logarithms,
+% and Elementary Algebra'~''
+% to compute astronomical phenomena~\cite{grier2013computers}.
+% The work that Airy solicited resonates with modern crowd work for several reasons.
+% First,
+% work output was quickly verifiable;
+% Airy could assign variably skilled workers to compute values,
+% and have other workers check their work.
+% Second,
+% tasks were discrete
+% --- that is,
+% independent from one another.
+% Finally,
+% workers could be trained on a very narrow subset of mathematical skills to be
+% sufficiently qualified to do this work.
 
-This approach found its audience in the early \nth{20} century with the rise of Fordism and scientific management (or Taylorism).
-Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out,
-piecework most greatly benefits the instrumented measurement of workers,
-but certainly in Ford and Taylor's time,
-highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-As a result,
-the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+% This approach found its audience in the early \nth{20} century with the rise of Fordism and scientific management (or Taylorism).
+% Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
+% As \citeauthor{Brown01041990} points out,
+% piecework most greatly benefits the instrumented measurement of workers,
+% but certainly in Ford and Taylor's time,
+% highly instrumented,
+% automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+% As a result,
+% the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
 
-\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
-and the conflicting pressures managers and workers put forth.}
-\citeauthor{bewley1999wages} in particular points out that
-the approach of paying workers by the piece is
-``\dots~not practical for workers doing many tasks,
-because of
-the cost of establishing the rates and because
-piecework does not compensate workers for time spent switching tasks''.
-Ultimately, \citeauthor{bewley1999wages} argues that
-``[piecework is] infeasible,
-because \dots
-total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
+% \topic{Piecework researchers enumerate a number of problems with the decomposition of work,
+% and the conflicting pressures managers and workers put forth.}
+% \citeauthor{bewley1999wages} in particular points out that
+% the approach of paying workers by the piece is
+% ``\dots~not practical for workers doing many tasks,
+% because of
+% the cost of establishing the rates and because
+% piecework does not compensate workers for time spent switching tasks''.
+% Ultimately, \citeauthor{bewley1999wages} argues that
+% ``[piecework is] infeasible,
+% because \dots
+% total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
 

### 2016-12-08 drafts/labor/piecework_lit.tex
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.

@@
 \textit{as well as} an additional compensation tied to performance,
-workers under this regime were in some senses ``task--labourers'',
-and in other senses
-(more conventional)
-``day--labourers''.
+workers under this regime were
+in some senses ``task--labourers'', and
+in other senses (more conventional) ``day--labourers''.
 
@@
 
-\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % yay rhyming?
+\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
 \topic{In this section we will offer a brief overview of the history of piecework;
@@
 
+\subsection{Case studies in piecework}
+\topic{We'll try to discuss the context and general details of three instructive cases of piecework,
+       which will substantively inform the three questions we investigate later
+       (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
+The three major threads we'll explore will be
+\begin{inlinelist}
+  \item Airy's employment of \textit{human computers},
+  \item the ``matchstick'' girls, and
+  \item railroad workers across the United States.
+\end{inlinelist}
+
+\subsubsection{Airy's Computers}
+\topic{\citeauthor{grier2013computers} gives early accounts of
+a piecework strategy in Airy's creation of
+the British Nautical Almanac~\cite{grier2013computers}.}
+Airy's goal was complex
+--- mathematical calculations to produce tables that would
+allow sailors to locate themselves by starlight from sea.
+Many of his contributors did not have high--level mathematical training,
+so Airy broke down the task into simpler calculations and
+distributed them by mail,
+accomplishing the complex goal through piecework tasks that paid little.
+
+However, when piecework entered the American economy,
+it was not used for complex work.
+One reason for low complexity was workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
+that apprenticeships imparted \cite{hart2013rise}.
+So, initially piecework arose for farm work, and as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
+the practice remained relatively obscure until
+it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+Complexity levels remained low at the turn of
+the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+However, writers of the time focused their attention on
+wage \cite{burton1899commercial} and
+management regimes \cite{norton1900textile}
+rather than training.
+
+\topic{Piecework became an important factor in the war effort for the Second World War,
+cementing its role not only in American factories,
+but in industrial work around the world.}
+The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.
+As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+He attributes this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women \cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
+
+\topic{The research community relating to
+piecework and labor
+has been wrestling with the decomposition of work for centuries.}
+The beginnings of
+systematic task decomposition
+stretch back as far as the \nth{19} century,
+when Airy employed young boys at the Greenwich Observatory who
+``possessed the basic skills of mathematics,
+including
+`Arithmetic,
+the use of Logarithms,
+and Elementary Algebra'~''
+to compute astronomical phenomena~\cite{grier2013computers}.
+The work that Airy solicited resonates with modern crowd work for several reasons.
+First,
+work output was quickly verifiable;
+Airy could assign variably skilled workers to compute values,
+and have other workers check their work.
+Second,
+tasks were discrete
+--- that is,
+independent from one another.
+Finally,
+workers could be trained on a very narrow subset of mathematical skills to be
+sufficiently qualified to do this work.
+
+This approach found its audience in the early \nth{20} century with the rise of Fordism and scientific management (or Taylorism).
+Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers,
+but certainly in Ford and Taylor's time,
+highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+As a result,
+the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+
+\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
+and the conflicting pressures managers and workers put forth.}
+\citeauthor{bewley1999wages} in particular points out that
+the approach of paying workers by the piece is
+``\dots~not practical for workers doing many tasks,
+because of
+the cost of establishing the rates and because
+piecework does not compensate workers for time spent switching tasks''.
+Ultimately, \citeauthor{bewley1999wages} argues that
+``[piecework is] infeasible,
+because \dots
+total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
+
+Workers' relationships with employers quickly soured.
+The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
+The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally~\cite{jacoby1983union}.
+
+\subsubsection{Human Computers}
+
+\subsubsection{The Matchstick Girls}
+\begin{itemize}
+  \item foo
+  \item foo
+  \item foo
+  \item foo
+\end{itemize}
+
+\subsubsection{Railroad Workers}
+\begin{itemize}
+  \item foo
+  \item foo
+  \item foo
+  \item foo
+\end{itemize}
+
 \onlyinsubfile{

### 2016-12-08 drafts/labor/relationships.tex
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.

@@
 
-Workers' relationships with employers quickly soured.
-The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
-As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
-% in the 30 years since the match--girls strike,
-% the Trade Union Movement grew from
-% ``20 per cent of adult male manual--working wage--earners [to] over 60 per cent''
-% \cite{webb1920history}.
-This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
-The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally~\cite{jacoby1983union}.
+% Workers' relationships with employers quickly soured.
+% The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+% As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+% This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
+% The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally~\cite{jacoby1983union}.
 
@@
 
-% \topic{While many workers participated in piecework, worker sentiment toward the practice was --- by all accounts --- mostly negative.}
-% The match girls strikes which \citeauthor{10.2307/3827491} describes were just one early
-% --- albeit critical ---
-% case study in this space;
-% the national coal strike of 1912 led to an overwhelming vote among federated coal miner pieceworkers
-% to strike for
-% an individual minimum wage, among other demands
-% \cite{10.2307/2221944}.
-% \citeauthor{10.2307/41829256} documents a series of efforts among women in the garment industries in Philadelphia to negotiate collective bargaining rights and recognition of their own labor union
-% \cite{10.2307/41829256}.
-% The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally
-% \cite{jacoby1983union}.
-
 Pieceworkers' relationships with their employers eventually developed a pattern of using 
-% \topic{The questions surrounding
-% the ways pieceworkers related to managers might be best answered by
-% the work that has been done in the emergence and proliferation of labor unions.}
-% The primary avenue for workers to interact with managers has been through
 laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,
       mccallum2013global,jacoby1983union}.
-% (one of the forerunners of the largest and
-% most politically influential labor union in the United States).
-% Looking through that lens, we find copious research on
-% the relationships between workers and requesters
-% \cite{levi2009union,ahlquist2013interest,
-%       mccallum2013global,jacoby1983union}.
 Collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.
@@
 didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
-% Still, we can look at primary sources, like \citetitle{american1921problem}
-% to give us some hint of how they related to each other
-% \cite{american1921problem}.
 Primary sources indicate that labor organizations wished for workers to identify as a collective group, 
-% The driving force of American labor advocacy organizations was to get piecework railroad workers
-% to identify
 ``not only as railroad employees but also as members of the larger life of the community''~\cite{american1921problem}.
@@
 \citeauthor{riisOtherSideLives} also contributed to this sense of shared struggle and endurance
-% by the time \citetitle{american1921problem} was published
 by documenting pieceworkers in their home--workplaces,
 literally bringing to light the grim circumstances in which pieceworkers lived and worked~\cite{riisOtherSideLives}.
-
-% This can also foreshadow crowd sourcing efforts like
-% \TO~and \DO.
 

### 2016-12-08 drafts/labor/trackingWork.tex
intro revisions
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.

@@
 \usepackage{xr-hyper}
-% \zxrsetup{%
-%   tozreflabel=false, % not needed, since `zref` is not used otherwise
-%   toltxlabel=true, % classical LaTeX labels that can be referenced with \ref, \pageref
-% }
-% \zexternaldocument{cases.tex}% star form: only LaTeX labels are imported
-
 \usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
@@
 % Use \plainauthor for final version.
-%  pdfauthor={\plainauthor},
-  pdfauthor={\emptyauthor},
+  pdfauthor={\plainauthor},
+  % pdfauthor={\emptyauthor},
   pdfkeywords={\plainkeywords},
@@
 
-% create a shortcut to typeset table headings
-% \newcommand\tabhead[1]{\small\textbf{#1}}
 \bibliography{references}
-% \addbibresource{references}
 
@@
 
-% \DTMsetdatestyle{mmddyyyy}
-% \DTMsetup{datesep=/}
-% \pagestyle{fancy}
-% \lfoot{Last compiled \today~at \DTMcurrenttime}
-% \cfoot{}
-% \rfoot{page \thepage/\pageref*{LastPage}}
-% \fancyhead{}
-% \pagenumbering{arabic}
 \renewcommand{\headrulewidth}{0pt}
 
-% \onlyinsubfile{
-%   \usepackage{xr-hyper}
-%   \usepackage{hyperref}
-% }
-
-\renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
+% \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
 \renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
@@
 }
-\author{
-\alignauthor{Anonymous for submission\\
-\affaddr{Affiliation fields filled}\\
-    \affaddr{to take representative space}\\
-    \email{of final document}}\\
-}
+% \author{
+% \alignauthor{Anonymized for Review\\
+% \affaddr{~}\\
+%     \affaddr{~}\\
+%     \email{~}}\\
+% }
 \maketitle
@@
 \endnotetext{
-  % \section{The Bleak Future of Crowd Work}
-  % We've traced a path from piecework itself through
-  % the processes that describe the design and implementation of piece work and crowd work as part of the same thread;
-  % in tracing this process, we touched on the relationships between decomposition,
-  % work \& worker abstraction,
-  % flexibility, and followed through both the general fallout of crowd work in the research community
-  % as well as the fallout between workers and the managers and other external parties
-  % --- including researchers.
-
-  % Throughout these case studies, we have pointed out the parallels between
-  % the contemporary research in on--demand labor and
-  % the much larger body of research constituting our understanding of topics such as
-  % piecework, factory work, and laborer relations.
-  % If we agree that this framing is useful and informative, then
-  % several topics emerge as relatively open questions in the study of crowd work and on--demand labor.
-  % Two of the most pressing questions are
-  % \begin{inlinelist}
-  % \item the beginnings of factories, and 
-  % \item the decline of relevance of worker advocacy organizations.
-  % \end{inlinelist}
-  % We will discuss those questions here.
-
-  % \subsection{The beginnings of factories}\label{sec:Factorization}
-  % We established earlier that abstracted work and low wages tend to result in variable outcomes,
-  % which presents problems for employers.
-  % % So work output is variable, and that's bad. What are we doing about it?
-  % Historically, this is what led to factories; by employing a cohort of known workers,
-  % we can be reasonably assured that the quality of the work will be better than random.
-  % Furthermore, we can invest more resources in training workers and
-  % get workers to do more complex work with more context.
-  % % you would pay some experts who understood what they needed to do
-  % % and they would form a cohort into which you invest more.
-
-  % This, then, suggests that the beginning of the regularization of workforces
-  % --- a sort of coalescence of factories ---
-  % is already happening.
-  % If our framing of on--demand labor is accurately describing an underlying relationship with piecework,
-  % then we should watch for the emergence and popularization of persistent teams of workers.
-  % While this field site will likely prove exceedingly difficult to 
-  % Social computing researchers should seek to understand this process more
-  % fully, as we a
-
-  % \subsection{The decline of advocacy organizations}
-
-  % \section{Implications for Design}
-  % If it's agreed that
-  % the major topics we've discussed thus far are related and
-  % --- at least to \textit{some} extent ---
-  % precipitated in the fashion we argue,
-  % then we have a rare opportunity as researchers,
-  % and as agents of change in the communities we study,
-  % to affect change on the dynamics of crowd and on--demand work
-  % as they continue to develop.
-
-  % Without claiming to have easy, cut--and--dry solutions to these problems,
-  % we can nevertheless bring to attention a number of critical opportunities to
-  % learn from historical parallels in piecework and factory labor,
-  % and make informed decisions regarding whether
-  % (or indeed how)
-  % we may want to influence outcomes.
-  % The challenges we bring to attention here are as follows:
-  % \begin{inlinelist}
-  %   \item codifying investment toward collective goods into the designs of systems;
-  %   \item (re--)decentralizing the internet; and
-  %   \item enabling reputation transferral.
-  % \end{inlinelist}
-
-  % \subsection{Codify the common good}
-  % As \citeauthor{lessig2006code} points out in his book,
-  % digital media give designers the opportunity to design and build into the systems
-  % policies and practices to contribute to the collective benefit of the people therein
-  % \cite{lessig2006code}.
-  % Historically, the confluence of forces \citeauthor{lessig2006code} describes
-  % would ultimately result in outcomes such as benefits for workers,
-  % funds for sick leave and vacation, and other conveniences.
-  % The transient nature of on--demand work would seem to problematize this arrangement,
-  % but we can discuss and explore the viability of building into systems the mechanisms necessary
-  % to save a portion of payment from every gig,
-  % record taxable income, or
-  % myriad other generally administrative tasks automatically.
-
-  % \subsection{Decentralize the internet --- again}
-  % Digitally mediated on--demand labor markets have historically been insular and incompatible with one another,
-  % forcing workers either to choose one or juggle participation in these markets with great difficulty.
-  % An ``API'' for on--demand labor markets could make it possible for any person
-  % or organization to instantiate their own marketplace and inter--operate with.
-  % This can be changed, and indeed must, if we are to realize the hopes of early researchers who advocated the
-  % democratizing nature and power of the internet
-  % \cite{barlow2009declaration,lanier2014owns}.
-
-  % \subsection{Deal with reputation}
-  % Reputation systems in on--demand labor markets are fundamentally broken.
-  % To say nothing of the fact that information workers (such as those on AMT) can't transfer their reputations to qualitatively different forms of labor
-  % like driving--for--hire (e.g. Uber), even within the same industry
-  % it's currently not feasible for workers to transfer their reputations or other information from one place to another.
-  % This affects more than the reputation and trustworthiness of workers;
-  % accounting for things such as taxes, benefits, etc\dots is all but left to the individual workers, who struggle with myriad bureaucratic obstacles.
-  % We can design systems that facilitate the aggregation and, more importantly, the transferral of reputation, income, and other features of work.
-
-  % \section{Discussion}
-  % We've discussed a number of aspects of on--demand work that
-  % offer parallels with historical piecework.
-  % Perhaps more importantly, we've hopefully demonstrated that the dynamics we observe in on--demand work
-  % are interrelated and follow from one another just as necessarily as they did in
-  % the development and maturation of piecework and factory work through the \nth{20} century.
-  % This framing on on--demand work should, we hope, provide us with the necessary historical context
-  % to make better--informed design decisions about how we want ``the future of crowd work'' to look.
-
-  % \section{Conclusion}
-  % \citeauthor{crowdworkFuture} discussed many of the challenges and problems in crowd work in \citeyear{crowdworkFuture},
-  % but didn't necessarily situate the notion of crowd work in a broader context.
-  % This paper attempts to fill that gap, and in doing so hopes
-  % to give the research community theoretical grounding to work with and within on--demand labor more successfully.
-  % But more than that, we hope to have addressed important questions to inform how we actually might make crowd work
-  % a career in which we want our children to work.
 }
-% \theendnotes
 
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2016-12-13 drafts/labor/cases.tex
verrrrry rough draft of moving the lit review stuff around (more info below) moving the contextual stuff from , , and  to , with the ultimate goal of putting most of the general-purpose historical background in the lit review section as a roadmap/overview of the three main threads that we use to evaluate these questions.
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

@@
 
-\endnotetext{
-Here, \citeauthor{hu1961parallel}'s work,  saying of assembly line work that
-``it is assumed that men are of equal ability and every man can do any of the $n$ jobs'',  parallels the approach that dominated early research into crowd work
---- namely, using non--expert crowds for complex work~\cite{hu1961parallel}.
-This mindset in \citeauthor{hu1961parallel}'s analysis,  and indeed the study of factory and mass manufacturing labor through the \nth{20} century,  substantively owes its existence to scientific management
-and the rigorous decomposition of work into tasks, discussed earlier,  and persists to this day as it colors
-researchers' goals and objectives in the study and design of crowd work.
-
-Piecework's influence on the abstraction of work into tasks,  described above, is more than just caused by the decomposition of work;
-work abstraction itself makes it possible for workers to come and go flexibly,  prompting work requesters to consider ways to design these now discrete tasks in ways that
-maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
-\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
-It's to this area that we now turn our attention.
-
-Piecework has seen work along this dimension spanning decades;
-\citeauthor{thompson1913time} investigate some of the ways that
-construction can benefit from the principles of scientific management.
-\citeauthor{thompson1913time}'s thesis asserts that
-task work is predicated on the accurate scientific management of work,  including the ``miscellaneous tasks''.
-\citeauthor{thompson1913time} argues
---- as early as \citeyear{thompson1913time} ---
-that ``\dots one may be challenged to find any class of work
-involving labor either indoors or out--of--doors
-where tasks cannot be fixed by proper time--study''~\cite{thompson1913time}.
-
-Broken down in this way, work could grow to unprecedented scales,  but the quality of the work would remain relatively variable~\cite{murray1983decentralisation}.
-Textile work being a salient example,  it took time for workers to acquire sufficient skill
-to do every aspect of the work so that the garment would be accepted by the company soliciting that work~\cite{vezina1992light}.
-
-A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,  manageable routines that could be taught relatively easily,  and whose work output could be evaluated in abstraction from the rest of the work~\cite{restructuringPieceworkBaker}.
-In Ford's assembly line, this meant that workers were not responsible for building a whole car,  but a single very narrowly defined action that needed to be done on every car~\cite{towardsGlobalFordism}.
-By the mid--\nth{20} century, \citeauthor{schoenberger1988fordism} writes,  ``\dots~the intensification of the labor process is argued to have hit mental, physical, and social limits.''~\cite{schoenberger1988fordism}.
-
-This approach, ``Fordism'' (and its better--known contemporary ``Taylorism'' of similar ethos),  can be seen today in crowd work and on--demand labor through the application of micro--tasks.
-\citeauthor{writingMicroTasks} highlight some of the advantages of breaking work into pieces,  facilitating evaluation and parallelization~\cite{writingMicroTasks}.
-By decomposing and recomposing tasks,  and in particular by assigning similarly natured work to the same workers,  workers could become ``experts'' in a small aspect of the work that they did,  speeding their work dramatically~\cite{delayAndOrderLasecki}.
-Perhaps more important, however, was that
-the breaking down of work into tasks has made it more practical to evaluate work at each stage~\cite{rogstadius2011assessment}.
-
-\subsubsection{So how does this affect crowd work?}
-The work we've seen so far 
-\begin{itemize}
-  \item worst case: assembling iPhones (extant)
-  \item average case: railroad workers and assembly lines
-  \item high (complexity) case: 
-\end{itemize}
-}
-
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-%   \nobalance{}
-%   }
-\endnotetext{
-\section{Cases notes}
-    \citeauthor{cheng2015break} found that microtasks
-    --- though not necessarily \textit{faster} than ``macrotasks'' ---
-    yield higher quality work,      particularly when that work is susceptible to frequent interruptions
-    \cite{cheng2015break}.
-  % }
-  % \endnotetext{
-  \subsection{What forms of work design and worker management are viable?}
-  \begin{itemize}
-    \item researchers have looked at how to increase worker productivity
-    (e.g. finding the maximal speed at which gig workers can be expected to work before making errors)
-    \cite{measuringCrowdsourcingCheng}.
-    \item we've also seen people ``embrace error''
-    \cite{embracingErrorKrishna}.
-    \item still other research has looked into ways to sandbox workers from the context of their work 
-    \item but scholarship looking into the design and management of work and workers isn't new;
-    lots of research into getting pieceworkers to do work more quickly
-    \cite{seymour1954manual}.
-    \item Researchers have even asked the age old question of \textit{what motivates} pieceworkers
-    (echoing similar research on Wikipedia and Mechanical Turk)
-    \cite{roy1953work,Nov:2007:MW:1297797.1297798,kaufmann2011more}
-  \end{itemize}
-
-  \subsection{What will work and the place of work look like for workers?}
-  The metaphorical mechanics of these dynamics are still at play;
-  workers and managers continue to interact in adversarial manners,    despite substantive work into aligning the motivations of workers and requesters
-
-    % \section{Case Studies}
-    The existing body of research has shed light on on--demand labor from various perspectives,      and revealed a number of topics that,      through our framing, are clearly situated together.
-    Those topics are, at a high level, as follows:
-    \begin{enumerate}
-    \item the \textbf{processes} involved in making work into tasks, or discretization;
-    \item the outcomes (and indeed the \textbf{fallout}) of that discretization,      both on the work itself as well as the workers;
-    and finally
-    \item the \textbf{relationships} between workers and requesters of the work
-    --- both \textit{cooperative} and \textit{adversarial} cases.
-    \end{enumerate}
-
-    \subsection{The Fallout of Crowd Work}
-    % If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,      % the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
-    \citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
-    (see also \citeauthor{dynamo}'s work
-    continuing in the spirit of this observation to generate collective action to improve worker conditions)
-    \cite{turkopticon,dynamo}.
-    \citeauthor{uberAlgorithm}
-    find similarly that workers on gig work platforms are frustrated by the systems on which they work,      to say little of the policies which these systems enforce
-    \cite{uberAlgorithm}.
-
-    We discussed the benefits of flexibility
-    (both in the sense of having arbitrary workers perform tasks and
-    in the sense that we can design tasks to be more resilient to poor work)
-    in the previous section.
-    It's from that point in the literature that we turn our attention to
-    the perhaps unintended effects of crowd work
-    and the affordances for transience that we build into this mode of work.
-    We'll address two major areas of work under this subject:
-    \begin{inlinelist}
-    \item \nameref{sec:lowPay}; and
-    \item \nameref{sec:varQualWork}.
-    \end{inlinelist}
-
-    \subsubsection{Low Pay}
-
-    \citeauthor{laborEconomicsOfCrowdsourcingHorton}
-    identified problems with crowd work wages relatively early on,      attempting to address this imbalance from a behavioral economic perspective ---
-    that is, identifying and presenting a model that describes a worker's
-    ``\textit{reservation wage}''
-    \cite{laborEconomicsOfCrowdsourcingHorton}.
-    This work has largely informed much of the research into and practice of estimating crowd work compensation
-    \cite{incentivesShaw,paolacci2010running}.
-
-    But we turn to \citeauthor{turkopticon}'s discussion of ``\TO'',      a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
-    \cite{turkopticon}.
-    Of particular relevance here,      \citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
-    wage theft and
-    pay rates that translate to well under minimum wage.
-    Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},      we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
-    was \$1.38, while the mean was \$3.63
-    \cite{laborEconomicsOfCrowdsourcingHorton}.
-
-    Understanding workers' motivations given these conditions has thus become a goal for some researchers
-    \cite{whyWouldAnyoneBrewer}.
-    \citeauthor{Sun20111033} conclude that
-    ``\dots~solvers participate in online tasks
-    not only for money
-    but also for enjoyment
-    or the sense of self--worth''
-    \cite{Sun20111033}.
-    This might have rung true in \citeyear{Sun20111033},      and certainly corroborates \citeauthor{Ross}'s findings after investigating
-    ``who are the crowd workers'',      but as \citeauthor{whoareNOTtheTurkers} points out
-    ``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
-    \cite{Ross,whoareNOTtheTurkers}.
-
-    Now, \citeauthor{turkopticon}
-    and later
-    \citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,      via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s contributions in this space
-    \cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.
-
-    On--demand workers were not the first to be exploited along the dimension of low pay rates.
-    Frustration over low (and declining) pay was one of the chief grievances among then nascent
-    British labor unions in the early \nth{20} century
-    \cite{turner1952trade}.
-    This, \citeauthor{ebbinghaus1999institutions} argued,      fueled the rocketing union membership rates through the mid--\nth{20} century until 1980
-    (to which we'll return when we discuss \citeauthor{levi2009union}'s reexamination of labor unions)
-    \cite{ebbinghaus1999institutions,levi2009union}.
-    This realization has similarly fueled a body of research into
-    the various incentive structures available to piecework employers
-    \cite{roy1953work}.
-
-    The parallels between the complaints of low pay among crowd workers and other on--demand workers
-    and the pieceworkers and later factory workers in the \nth{20} century
-    are inescapable.
-    We argue further that the \textit{causes} here
-    --- work decomposition,      work abstraction, and
-    flexibility ---
-    lead inexorably to low and declining pay for workers.
-    Moreover, we point out that low pay leads to other negative outcomes both
-    in on--demand work
-    as well as
-    in piecework and on assembly lines.
-
-    \subsubsection{Variable quality work}
-    Researchers have struggled with what we might generously call work of ``variable quality''
-    along two dimensions.
-    The first, to use the characterization of one of these contributions, we can call
-    ``understanding malicious behavior''
-    \cite{MaliciousCrowdworkersGadiraju}
-    While some work has cast workers as ``malicious'' or at least adversarial parties,      the evidence thus far suggests that
-    workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
-    (a topic of discussion to which we'll return later)
-    \cite{uberAlgorithm}.
-    The second dimension of research in this space generally attempts
-    to eke out the highest quality work possible from workers
-    given the apparent difficulty in predicting work outcomes
-    \cite{embracingErrorKrishna}.
-
-    The effect low wages have had on piece work and factory workers is well--known;
-    \citeauthor{gantt1913work} discuss this exact mechanism in his book on
-    % \citetitle{gantt1913work}, pointing out that
-    ``\dots~where there is no union,
-         the class wage is practically gauged by the wages the poor workman will accept,
-         and the good workman soon becomes discouraged and \textit{sets his pace by that of his less efficient neighbor},
-         with the result that the general tone of the shop is lowered'' (emphasis added)
-    \cite{gantt1913work}.
-
-    This research is similar to, but subtly different from, the notion of the ``market for `lemons'''
-    which \citeauthor{fort2011amazon} discuss;
-    specifically, \citeauthor{akerlof1970market}'s writing of a ``market for `lemons'''
-    describes a marketplace where the quality of the product or service is unknown to the buyer
-    \cite{fort2011amazon,akerlof1970market}.
-    The effect of this \textit{perceived} uncertainty is that
-    the \textit{actual} trustworthiness drops precipitously
-    as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,      leaving only the ones who cannot or will not establish their trustworthiness.
-
-    \subsection{Relationships Between Workers and Managers}
-    Suffice it to say that poor pay and poor work are linked,      and that we should not be surprised to find this relationship play out online as strongly as it does offline.
-    But the poor treatment of workers by managers
-    --- both human and algorithmic ---
-    do more than affect the economic relationships between workers and employers.
-    Here, then, we turn to examine this facet of on--demand work
-    and how these dynamics strikingly replicate the relationships
-    researchers in labor advocacy encountered in the study of piecework and factory work.
-
-    This topic can be condensed into two major areas:
-    \begin{inlinelist}
-    \item external (scientific) management, and the evaluation of workers as functional modules; and
-    % \item the alienation of workers under this form of evaluation; and
-    \item the consequential resistance workers express due to their perceived alienation and distance from managing forces.
-    \end{inlinelist}
-
-    \subsubsection{External Management}
-    We discussed Fordism and Taylorism earlier in our discussions of
-    \nameref{sec:Complexity} and \nameref{sec:flexibility},      but here the core of these paradigmatic views
-    --- the scientific management of work ---
-    becomes relevant.
-    We use ``external'' here instead of ``scientific'', however,      to more broadly capture the disconnect between managers and workers.
-    By describing it as thus,      we can touch on the relationship that workers have with \textit{researchers}, as well,      even though that work is not strictly
-    --- or just not exclusively ---
-    of the same nature as the management and experience as when interacting with requesters.
-
-    First, intuitively, the variable--quality work we discussed previously has led to
-    a large and growing body of research attempting to evaluate workers' performance and error rates
-    across numerous dimensions;
-    for example, \citeauthor{measuringCrowdsourcingCheng} explore the error rates of workers by
-    operating on a sliding scale giving workers varying amounts of time to accomplish micro--tasks
-    \cite{measuringCrowdsourcingCheng}.
-    \citeauthor{storiesIraniSilberman} describe the treatment of workers
-    as sorts of ``human APIs'' that can, importantly, be rigorously evaluated
-    \cite{storiesIraniSilberman}.
-    \citeauthor{gevins2003neurophysiological} began to explore the neurophysiological effects of
-    cognitively demanding tasks on workers,      informing crowdsourcing research by suggesting the use of cognitive load assessments such as
-    NASA Task Load Index surveys to evaluate workers pre and post--tasks
-    \cite{embracingErrorKrishna,measuringCrowdsourcingCheng}.
-
-    External management comes in other forms than scientific, as previously mentioned.
-    Researchers in particular have noticed that their relationships with on--demand workers are,      at the least, complex.
-    \citeauthor{storiesIraniSilberman} point out that their relationships with Turkers are highly complex;
-    specifically, their interactions with field sites in which they work
-    as designers and mediators of change influence the relationships they have with Turkers
-    \cite{storiesIraniSilberman}.
-
-    The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,      and even physiological study of pieceworkers closely resembles the
-    research into cognitive loads and stress levels that we discussed among on--demand crowd workers
-    \cite{hu1961parallel,pieceworkBiologicalHarm}.
-    Even the complicated relationships between observers and workers themselves are not necessarily new;
-    \citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
-    to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
-    \cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.
-
-    Similarly, \citeauthor{pollard1963factory}'s words on the punishment factory workers faced ---
-    for example, that ``unsatisfactory work was punished \dots~by fines or by dismissal'' ---
-    seems especially relevant given the fears we now know to be ubiquitous on platforms such as AMT, Uber, and other on--demand markets
-    \cite{pollard1963factory,uberAlgorithm,dynamo,turkopticon,takingAHITMcInnis}.
-
-    \subsubsection{Resistance}
-    It shouldn't surprise us, then, that workers have resisted the management imposed on them both by other people and their systems,      often without recourse or opportunity for feedback, let alone substantive input.
-    Indeed, \citeauthor{uberAlgorithm} discover of Uber drivers that 
-    many toggle their availability to avoid being dispatched to more distant locations,      resisting the intent of the designers of the systems and their ``algorithmic and data--driven management''
-    \cite{uberAlgorithm}.
-
-    Resistance has sometimes been more coordinated, as well;
-    we see this in \citeauthor{turkopticon}'s coverage on \TO
-    as workers collectively accumulated information about requesters, and
-    in \citeauthor{dynamo}'s work on \DO, which generated ``Guidelines for Academic Requesters'' written by crowd workers
-    \cite{turkopticon,dynamo}.
-
-    Resistance against managers in piecework and factory labor settings are deeply well--explored,      but perhaps the most relevant case study to draw on here is to be found in
-    \citeauthor{waldinger1996helots}'s case study of ``Justice for Janitors'',      where marginalized workers managed to raise awareness for their plight and secure support for badly needed reforms
-    \cite{waldinger1996helots}.
-    The achievements of labor advocacy groups such as labor unions as resistant,      even adversarial organizations counter--balancing the management
-    is somewhat well--understood
-    \cite{russell1982collective,craig1992behavior}.
-    We argue that these threads of resistance against management in various forms
-    are in fact one.
-  }
-  % \theendnotes
-  % \end{appendices}
-
 \end{document}

### 2016-12-13 drafts/labor/complexity.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

--- 2016-11-22 drafts/labor/complexity.tex
@@
 
+\onlyinsubfile{\newpage}
 \subsubsection{\crowdworkpers}
@@
 additional proof--of--concept applications and techniques,
-including conversational assistants \cite{lasecki2013chorus},
-medical data interpreters \cite{lasecki2013chorus}, and
-idea generation workflows \cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+including conversational assistants~\cite{lasecki2013chorus},
+medical data interpreters~\cite{lasecki2013chorus}, and
+idea generation workflows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
 to name a few examples.
@@
 and arranges humans as computational black boxes within those approaches.
-% \ali{We could quote \citeauthor{storiesIraniSilberman} and say ``human APIs'',
-% but that gets a little critique--y, and we might want to save that for later.}
 This approach has proven a compelling one because
@@
 Second,
-focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
-      writingMicroTasks,agapie2015crowdsourcing}, but
+focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,writingMicroTasks,agapie2015crowdsourcing}, but
 there is no general solution to create
@@
 data analysis tasks such as
-clustering \cite{chilton2013cascade},
-categorization \cite{andre2014crowd}, and
-outlining \cite{luther2015crowdlines}
+clustering~\cite{chilton2013cascade},
+categorization~\cite{andre2014crowd}, and
+outlining~\cite{luther2015crowdlines}
 are possible, but there is no general solution for sensemaking.
@@
 to enable crowdsourced solutions for these broader wicked problems.
-%  and
-% other crowdsourced goals such as addressing climate change
-% \cite{introne2011climate}.
 
@@
 for example designers and programmers.
-The same ideas can then be applied to expert ``macro--tasks''~\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
-programming \cite{latoza2014microtask,Fast2016,Chen2016}, and
-mentorship \cite{suzukiAtelier}.
+The same ideas can then be applied to expert ``macro--tasks''~\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design~\cite{foundry},
+programming~\cite{latoza2014microtask,Fast2016,Chen2016}, and
+mentorship~\cite{suzukiAtelier}.
 However, there remains the open question of
@@
 
+\onlyinsubfile{\newpage}
 \subsubsection{\pieceworkpers}
@@
 
-\topic{\citeauthor{grier2013computers} gives early accounts of
-a piecework strategy in Airy's creation of
-the British Nautical Almanac~\cite{grier2013computers}.}
-Airy's goal was complex
---- mathematical calculations to produce tables that would
-allow sailors to locate themselves by starlight from sea.
-Many of his contributors did not have high--level mathematical training,
-so Airy broke down the task into simpler calculations and
-distributed them by mail,
-accomplishing the complex goal through piecework tasks that paid little.
-
-However, when piecework entered the American economy,
-it was not used for complex work.
-One reason for low complexity was workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
-that apprenticeships imparted \cite{hart2013rise}.
-So, initially piecework arose for farm work, and as
-\citeauthor{hughRaynbirdTaskWork} and others discuss,
-the practice remained relatively obscure until
-it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
-Complexity levels remained low at the turn of
-the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
-However, writers of the time focused their attention on
-wage \cite{burton1899commercial} and
-management regimes \cite{norton1900textile}
-rather than training.
-
+\ali{I'm not sure that this really stays close to the three case studies that emerged as major threads (now in the piecework lit review).
+Should I refactor/rewrite this to make it more relevant?}
 Measurement also limited the complexity of piecework:
@@
 Researchers detail three organizational criteria.
-First, \citeauthor{Brown01041990} argues that
+\begin{Numberlist}
+\item \citeauthor{Brown01041990} argues that
 piecework ``is less likely in jobs with a variety of duties than in
@@
 induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
-Second,
-complexity was limited by access to capital to create the necessary infrastructure.
+\item complexity was limited by access to capital to create the necessary infrastructure.
 As \citeauthor{10.2307/23702539} reports,
@@
 ~\cite{10.2307/23702539}.
-Third, organizations required capable managers in charge of the pieceworkers.
+\item organizations required capable managers in charge of the pieceworkers.
 The West Virginia mines, for example, hired foremen 
@@
 understanding when and how to modify work as necessary~\cite{wray1949marginal}.
+\end{Numberlist}
 So, in sum,
@@
 
+\onlyinsubfile{\newpage}
 \subsubsection{\whatchanged}
@@
 \end{comment}
-% \topic{Piecework makes a number of observations leading to the conclusion that
-% piecework's complexity is fundamentally bounded by several limitations,
-% chief among them the costs of managerial overhead and the transition thereto.}
-% \citeauthor{Brown01041990} and \citeauthor{10.2307/23702539}'s claims that
-% organizations can't adopt piecework unless
-% they're sufficiently large to absorb the cost of transitioning to a piecework system;
-% \citeauthor{10.2307/2118435} and \citeauthor{wray1949marginal}'s observations for
-% the importance of competent, effective managerial oversight
-% --- a human resource, which made the scaling cost prohibitively expensive for many
-% \cite{10.2307/2118435,wray1949marginal,10.2307/23702539,Brown01041990}.
 
-% \topic{Digital media have expanded the scope of
-% viable piecework by pushing drastically on the limits
-% cited by piecework researchers.}
 The research on piecework tells us that
@@
 
-% \ali{maybe this should go away?
-% \topic{Technology has made it possible
-% for non--experts to do work that was once considered
-% within the domain of experts.}
-% \citeauthor{yuanAlmost} builds on the work of others
-% (\textit{Voyant} and, more relevantly, \textit{CrowdCrit})
-% to design workflows that yield ``expert--level feedback''
-% \cite{yuanAlmost,Xu:2014:VGS:2531602.2531604,Luther:2014:CCA:2556420.2556788}.
-% This body of work identifies ways to transform a variety of duties comprising complex tasks
-% and distills them into ``a narrow set of routinized duties'',
-% informed in part by researchers --- acting as inspectors --- and experts
-% \cite[quotations from][]{10.2307/23702539}.
-% Where \citeauthor{10.2307/23702539} would call additionally for the identification of
-% crowdsourcing's version of ``piecework clerks'', we point out that
-% today algorithms manage workers as pieceworkers once did
-% \cite{uberAlgorithm,10.2307/23702539}.}
 Finally, the organizational limit on infrastructure creation is somewhat lessened. Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
@@
 \subsubsection{\implication}
-% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
-% it also signals some of the ultimate limitations of crowd work and piecework in general.}
-% While the threshold preventing task requesters from utilizing piecework
-% has dropped thanks to affordances of the internet, the ceiling on task complexity hasn't moved significantly.
-% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
-% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
-% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
 Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
@@
 Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
+of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
 To some extent, an argument can be made that
@@
 By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
-%At worst, as now, they are far simpler~\cite{uberAlgorithm}.
 So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
@@
 However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re--introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
-This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
-% and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
-% \cite{takingAHITMcInnis}.
-% \ali{What will happen if we don't do this?
-% Maybe this is where I should bring up
-% the ``market for lemons'' metaphor and say that
-% the market will just continue to decline in quality until
-% nobody is willing to rely on it either as a source of data/information (requesters), or
-% a means of sustaining oneself financially (workers).}
+This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters~\cite{MaliciousCrowdworkersGadiraju}.
 

### 2016-12-13 drafts/labor/decomposition.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

--- 2016-12-05 drafts/labor/decomposition.tex
@@
 
-\topic{Piecework became an important factor in the war effort for the Second World War,
-cementing its role not only in American factories,
-but in industrial work around the world.}
-The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.
-As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-He attributes this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women \cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
-
+\ali{This can be trimmed to just reference back to the piecework review}
 \topic{The research community relating to
 piecework and labor
-has been wrestling with the decomposition of work for centuries.}
+has been wrestling with the decomposition of work for the better part of a century.}
 The beginnings of
@@
 The work that Airy solicited resonates with modern crowd work for several reasons.
-First,
-work output was quickly verifiable;
+\begin{Numberlist}
+\item work output was quickly verifiable;
 Airy could assign variably skilled workers to compute values,
 and have other workers check their work.
-Second,
-tasks were discrete
+\item tasks were discrete
 --- that is,
 independent from one another.
-Finally,
-workers could be trained on a very narrow subset of mathematical skills to be
+\item workers could be trained on a very narrow subset of mathematical skills to be
 sufficiently qualified to do this work.
-
-This approach found its audience in the early \nth{20} century with the rise of Fordism and scientific management (or Taylorism).
-Scientific management suggested that it was possible to measure work  at unprecedented resolution and precision.
-As \citeauthor{Brown01041990} points out,
-piecework most greatly benefits the instrumented measurement of workers,
-but certainly in Ford and Taylor's time,
-highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-As a result,
-the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+\end{Numberlist}
 
@@
 Where \citeauthor{10.2307/23702539} directly implicates measurement as
-% cites as
 preventing scientific management from being fully utilized,
-%the difficulty of tracking work \& workers,
 no longer exists
-%\cite{10.2307/23702539}.}
 modern crowd work is measuring and modeling every click,
@@
 other factors in piecework that would have been
-all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,
-                                                              Cai:2016:CRI:2858036.2858237,
-                                                              cheng2015break,
-                                                              measuringCrowdsourcingCheng,
-                                                              embracingErrorKrishna}.
+all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
 
@@
 
+\onlyinsubfile{\newpage}
 \subsubsection{\implication}
@@
 Task switching costs and other cognitive costs make it difficult
-to work on tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
+to work on tasks so far decontextualized from their original intention~\cite{delayAndOrderLasecki}.
 There will of course be tasks that can be decomposed without much context,
@@
 
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-% }
+\onlyinsubfile{
+  \balance{}
+  \printbibliography
+}
 

### 2016-12-13 drafts/labor/introduction.tex
cleaned refs
rebuttal init
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

@@
 \section{Introduction}\label{sec:introduction}
-
 \topic{The past decade has seen a flourishing of \textit{on--demand work},
@@
 \item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
-\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
+\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.~
 \end{inlinelist}
@@
 
-\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who study it.}
-\citeauthor{howe2008crowdsourcing} first described crowdsourcing in general terms as
-``outsourcing [work] to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
+\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complex, conflicted culture amongst both the workers who enable it and the researchers who study it.}
+\citeauthor{howe2008crowdsourcing} first described crowdsourcing broadly as
+``outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
 However, for years its instantiation was limited to the utilization of
@@
 \subsection{Piecework as a lens to understand crowdsourcing}
-% MSB Cutting because it seems like it's just attacking the literature unnecessarily...
-% ----
-%This large and growing body of research has conversed
-%to varying degrees with labor scholarship,
-% ...
-%To the extent that the research community has explored the boundaries of crowd work,
-%it nevertheless has not reflected on the underlying mechanisms determining those boundaries.
-%Similarly, work directed toward the decomposition of tasks has begun to appreciate the limits of decomposition,
-%but on the whole it hasn't considered the relationship between the decomposition of work and the complexity of work.
-% \ali{Mixed feelings about this because I think I'll need to go back into decomposition and
-% give a whole thing on \citeauthor{marx2012economic}
-% if I bring this up as a real failing of the crowd work research to date.
-% Thoughts?}
-%While researchers are quickly picking up on the importance of
-%the relationships between workers in these platforms,
-%this research seems to be unfolding without the benefit of any of
-%the inter--personal labor research that has been informing modern work for the better part of a century.
-% \ali{Overall: too long? too vague? overstating the shortcomings?}
 In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework}: %\namerefl{sec:pieceworkArc} ---
@@
 For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.
-%we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
-%and come up with theoretically grounded mechanisms explaining the limitations researchers have struggled to overcome thus far.
-% \msb{what I think would really help this paragraph is a single great example, in one or two sentences, of how this kind of framing would help.
-``For example, we draw on ...
-to explain ...''.
-Reader goes AHA I see why you're doing this now!}
 
@@
 Then, we interrogate the three major research questions above using the lens of piecework.
-% We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
-% we propose as broadly describing the types of research being done in crowd work
-% and more generally in what we argue is contemporary piecework.
-%We will look at the above questions in crowd work through the lens
-%first of the body of research made by crowd work researchers,
-%and then of the body of work piecework researchers have offered.
-%After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
 We will identify similarities and differences between piecework as historically understood and
@@
 Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
-% we will use this perspective to suggest areas of research worth anticipating,
-% and developments we should expect to see in the maturation of digitally mediated work.
 Finally, we will offer implications for researchers and practitioners based on our results.

### 2016-12-13 drafts/labor/piecework_lit.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

--- 2016-09-22 drafts/labor/piecework_lit.tex
@@
         identifying key figures and ideas during this time;
-  %        and finally
-  % \item look at the fall of piecework, such as it was,
-  %       considering in particular
-  %       the factors that may have led to piecework's
-  %       eventual demise in the American and European labor markets.
-        % \msb{did it demise? there are sweatshops around the world today}.
-        % \ali{is it okay to corner it off like this (by just being about the US/EU)?}
 \end{inlinelist}
 This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.
-
-% While piecework has been studied from myriad perspectives through the lenses of
-% political science,
-% economics (both micro-- and macro--), and
-% a plethora of other fields
-% \ali{should I cite examples of all of these?}, for the sake of brevity
-% we will discuss the history of piecework in the contexts of the three questions we brought up with crowd work:
-% ``\namerefl{sec:complexity}'',
-% ``\namerefl{sec:decomposition}'', and 
-% ``\namerefl{sec:relationships}''.
 
@@
 \topic{We see this sense of independence regardless of the time, locale, and industry.}
-\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match girls''
+\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
 --- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
@@
 From these accounts we identify a sense of independence and autonomy that resonates across decades, industries, and locales where piecework is found.
-% We'll problematize this supposed advantage as we trace the history of piecework,
-% but for now we can say that piecework affords
-% independence and some sense of autonomy
-% new to people in the working class.
 
@@
 For example, women could receive training in narrow subsets of the general body of skills, enabling them to act in capacities similar to what conventional (male) apprentices would undertake~\cite{hart2013rise}.
-% offer another series of compelling insights toward
-% the question of the features that sprout from piecework.}
-% In their reflection on the features endemic to piecework in the 1930s,
-% which they describe as the ``heyday'' of piecework's prominence;
-% among them were the following:
-% \begin{inlinelist}
-%   \item ``female workers who generally had less training'' had to be trained in
-%         narrower subsets of the general body of skills that
-%         conventional (male) apprentices would undertake, and
 In addition, workers with specific slices of skills could be matched to suitable tasks.
@@
 In summary, piecework:
-\begin{enumerate}
+\begin{inlinelist}
   \item paid workers for quantity of work done, rather than time done,
@@
         to engage in the traditional labor force could still participate.
-\end{enumerate}
-% \ali{woah i thought this was a comment; is it okay as is?}
+\end{inlinelist}
 
 Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
-First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well. 
-% Second, workers on these platforms have significant freedom in choosing which tasks to work on, and when to log on and perform work~\cite{martin2014being,whyWouldAnyoneBrewer}. 
-% Finally, computational workflows have successfully scaffolded many complex tasks into a series of small microtasks that can be performed without expertise.
-%  \subsection{Piecework and crowd work}
-% \msb{I assume this is forthcoming?}
-% Using the definition of piecework that we came up with earlier, we argue that
-% crowd work is fundamentally an instantiation of piecework, and
-% that we can more precisely anticipate the answers to the open research questions we discussed earlier.
-% We'll show that the dimensions of crowd work that the broader HCI community has been studying
-% align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
-% \subsection{From piecework to on--demand work}
-% Crowd work and gig work are fundamentally an instantion of piecework.
-% First, workers on platforms such as Mechanical Turk and Uber are generally incentivized by unit of work, even if some may be offered an hourly base salary as well.
-Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+\begin{Numberlist}
+\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
+\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
+\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+\end{Numberlist}
 Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
@@
 
-% \ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}
-
-\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % yay rhyming?
+\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
 \topic{In this section we will offer a brief overview of the history of piecework;
@@
 
-% While this approach didn't become the same economic powerhouse as later examples would,
-% Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
-% determining the extent to which work can be decomposed, and
-% finding the limits of complexity of that decomposed work.
-% Airy found that he could train youths in elementary mathematics
-% to complete the majority of the calculations he would otherwise have had to solve on his own,
-% and that the greater body of work could ultimately be completed sooner
-% if he arranged his work appropriately.
-% \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
-% What I got out of it is that a bunch of people did piecework, but
-% I don't know why or in fact why these are different than the sources we cite earlier.
-% Can you hone it?}
-% \ali{I wanted that paragraph to be about the rising popularity and application of piecework
-% (especially as it approached its ``heyday'' \cite{hart2013rise}),
-% coming from humble beginnings as it found its footing.
-% Given that intent\dots\\
-% Should I \textbf{refactor} or \textbf{rewrite}?}
-
-% \ali{\sout{I'm thinking of
-% dropping most of the last paragraph and merging it with the next one. Thoughts?}
-% I moved it to \nameref{sec:complexity}}
-
-% \ali{Where do I want to take this? Where are we now?
-% \begin{enumerate}
-%   \item Piecework started in the \nth{19} century (above). [DONE]
-%   \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
-%   \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
-%   \item Piecework died pretty quickly (below). [DONE]
-% \end{enumerate}
-% Sound good?}
-
 \topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
 but it found its foothold in manual labor.}
-Piecework took off on in farm work \cite{hughRaynbirdTaskWork},
-in textiles \cite{restructuringPieceworkBaker,riisOtherSideLives},
-on railroads \cite{Brown01041990}, and 
-elsewhere in manufacturing \cite{10.2307/3827491}.
+Piecework took off on in farm work~\cite{hughRaynbirdTaskWork},
+in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
+on railroads~\cite{Brown01041990}, and 
+elsewhere in manufacturing~\cite{10.2307/3827491}.
 Fordism and scientific management thrust piecework into higher gear, especially as
@@
 a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
-% With little time and virtually none of the training that was conventionally expected of metalworkers and other industrial work
-% (e.g. years--long apprenticeships),
-% tasks had to be redesigned and workers trained for these dramatically unorthodox task designs
-\ali{I'd like to bring the Rosie the Riveter anecdote in here but it's not feeling natural\dots}
-% \cite{honey1985creating,davies2014origins}.
 
-The early proliferation of manual piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
-Despite this, workers' means were mostly ignored,
+The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
+Despite this, workers' needs were mostly ignored,
 leading to frustration over poor working and living conditions
 (famously documented by
-% \ali{??? Why is this here? Can it be moved?}
-% {As increasing attention revealed problems in piecework as it related to workers,
-% workers themselves began to speak out about their frustration with this new regime.}
 \citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
-This led to industry organizations representing
-railway workers, mechanical engineers, and others beginning to speak out on pieceworkers' behalf~\cite{american1921problem,richards1904anything}.
-% Nevertheless, piecework continued to permeate low--skilled labor.
-% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}
+Discontent reached a crescendo when industry organizations representing
+railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.
 
@@
 
+\subsection{Case studies in piecework}
+We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
+       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
+\begin{numberlist}
+  \item Airy's employment of \textit{human computers},
+  \item domestic and farm work (in particular, the ``match--girls''), and
+  \item railroad workers across the United States.
+\end{numberlist}
+
+\ali{I'm coming out of this with a sort of structure that \textit{loosely} describes piecework chronologically.
+Strictly speaking, I'm talking about each of these topics
+(let's call them
+\textbf{human computation},
+\textbf{domestic and farm work}, and
+\textbf{industrial work}),
+and some of these things overlap with each other to some extent, but
+for the most part
+Airy and the human computers came in the mid--\nth{19} century,
+farm work and the ``domestic'' work (like making matches) came around at the turn of the \nth{20} century, and
+the industrial work (railroad workers, the WWII war effort, and the rise of the labor union movement) unfolded from the early \nth{20} century onward.}
+
+\subsubsection{Airy's Computers}
+\topic{Some of the first systematic cases of what we would recognize as crowd work
+can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
+% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
+% piecework, found in George Airy's creation of
+% the British Nautical Almanac.}
+Airy needed to compute tables that would
+allow sailors to locate themselves by starlight from sea.
+This work ostensibly called for educated people who comprehensively understood mathematics.
+Airy realized that he could break the tasks down and delegate the constituent parts
+to ``human computers'' who
+``\dots~possessed the basic skills of mathematics,
+including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
+As a result, many of Airy's computers had relatively rudimentary educations
+compared to the background of education that typically worked in the calculation of solar tables.
+Airy distributed tasks by mail,
+allowing work to be completed by a somewhat geographically distributed workforce,
+and paid for each piece of work completed.
+
+\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
+For one thing,
+Airy had set a policy of firing his computers once they reached age 23.
+This practice ensured two outcomes that arguably disfavored workers.
+\begin{Numberlist}
+\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
+      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
+\item it limited workers' ability to organize
+      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
+\end{Numberlist}
+% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
+% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
+
+\subsubsection{Domestic and Farmhand Labor}
+\topic{When piecework entered the American economy, it was not used for complex work.}
+One reason for low complexity was workers' skills:
+it was infeasible to provide new pieceworkers with the comprehensive education
+that apprenticeships imparted~\cite{hart2013rise}.
+So, initially piecework in the United States arose for farm work, and as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
+the practice remained relatively obscure until
+it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+The complexity of the work remained low at the turn of
+the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+However, writers of the time focused their attention on
+wage~\cite{burton1899commercial} and
+management regimes~\cite{norton1900textile}
+rather than training.
+Mass manufacturing,
+such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
+flourished under piecework systems in densely populated cities.
+
+\topic{Workers' relationships with employers quickly soured.}
+The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
+and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+As \citeauthor{weyer1894history} described,
+``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
+the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+
+\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
+In 1912, national coal miners effected an individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
+Women in the garment industry in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
+It was in the midst of this time that \citeauthor{taylor1914principles} published
+the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
+% and it was this framing on work that
+It was this framing on work \& worker management that gave workers a concrete adversary
+--- if not Taylor himself, then \textit{Taylorism}  ---
+against which to rally~\cite{jacoby1983union,parker1920casual}.
+
+\subsubsection{Railroad and other Industrial Workers}
+\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
+piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
+Scientific management argued that it was possible to measure work at
+unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers,
+% but certainly in Ford and Taylor's time,
+but in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
+as rudimentary guides on per--task compensation.
+This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
+railroad maintenance, automobile manufacturing, and later weapons manufacturing.
+\ali{???? better word for ``weapons manufacturing''?}
+% Instead, as a result,
+% the distillation of work into smaller units ultimately
+% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+
+% \topic{With the formalization of scientific management,
+% piecework became an important component of the mass manufacturing
+% in American factories.}
+% \topic{}
+
+\topic{The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.}
+\citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+They attribute this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
+
 \onlyinsubfile{
   \printbibliography
-  % \clearpage
   }

### 2016-12-13 drafts/labor/relationships.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

@@
 
-% \topic{The relationships of workers
-% % with their work,
-% with their peers and
-% with requesters
-% are nuanced and not especially well--understood.}
-% We can break this general body of work into two subgroups:
-% workers' relationships
-% \begin{inlinelist}
-% % \item with \textit{work}, % \nameref{ch:workRelationship},
-% \item with \textit{requesters}, and
-% \item with \textit{other workers}.
-% \end{inlinelist}
-% We'll look at workers' relationships with work itself, which we'll discover
-% gives us insight into why people engage in crowd work in the first place.
-
 Workers' relationships with requesters are fraught.
-% \topic{\ali{some topic sentence that brings together the debate where
-% one side blames Turkers for being bad at Turking and
-% another side blames requesters for not understanding Turkers as a culture of people.}}
 Workers are often blamed for any low-quality work, regardless of whether they are responsible~\cite{martin2014being,takingAHITMcInnis}.
 Some research is extremely open about this position, blaming unpredictable work on ``malicious'' workers~\cite{MaliciousCrowdworkersGadiraju} or those with ``a lack of expertise, dedication [or] interest''~\cite{Sheng:2008:GLI:1401890.1401965}.
-% (see, for example, \citeauthor{MaliciousCrowdworkersGadiraju}'s work,
-% which frames the problem of unpredictable work as the result of ``malicious'' crowd workers),
-% \cite{MaliciousCrowdworkersGadiraju,Sheng:2008:GLI:1401890.1401965,Ipeirotis:2010:QMA:1837885.1837906}.
 Workers resent this position --- for good reason.
@@
 
-% \topic{The frustration that workers experience dealing with requesters
-% seems to precipitate frustration and mistrust between crowd workers, as well.}
-% \citeauthor{dynamo} describes ``mega--drama'' among workers on forums for Turkers;
-% \citeauthor{irani2015cultural} and \citeauthor{storiesIraniSilberman} discuss
-% the culture of crowd work and the study thereof.
-% \citeauthor{crowdcollab} quantifies and maps this social network of Turkers.
-% \citeauthor{takingAHITMcInnis} takes these observations and considers
-% what a crowd work platform might look like if it were to be designed more inclusively
-% \cite{dynamo,irani2015cultural,storiesIraniSilberman,crowdcollab,takingAHITMcInnis}.
-% The overarching theme of the research in this space has been
-% documenting the struggle of crowd workers
-% and attempting to intervene in constructive ways, while walking the balancing act
-% (especially in the cases of \citeauthor{irani2015cultural} and later \citeauthor{storiesIraniSilberman})
-% as we think about the culture of crowd workers.
-
+\onlyinsubfile{\newpage}
 \subsubsection{\pieceworkpers}
@@
 Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
-
-% Workers' relationships with employers quickly soured.
-% The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
-% As \citeauthor{weyer1894history} described, ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike, the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
-% This strike was followed by others, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
-% The adoption of piecework's time--studies and other Taylorist and scientific management approaches reliably precipitated strikes and more generally gave workers a clear enemy against which to rally~\cite{jacoby1983union}.
 
@@
 	*collective action hard --- collective action succeeded
-
 	- algorithms, not managers
@@
 Another possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen, online workers are known by little more than an account ID.
-
-% \topic{The differences between crowd workers and pieceworkers seem defined largely by
-% the differences in the places of work.}
-% Whereas it arguably became inevitable that workers would have a place to
-% meet, discuss, and collaborate
-% when they began sharing places of work,
-% online spaces make it much harder to do so.
-% Crowd workers can ``lurk'' and do tasks, or just do the occasional one--off task,
-% without any affiliation with
-% --- or even knowledge of ---
-% communities of peers
-% % \ali{multiple citations that labor unions came out of factories here}
-% \cite{miller2011understanding,mcinnis2016one,earl2011digitally}.
 
@@
 
-% \ali{here's an idea. feel free to push back or revise or something:\\
-% ``Should we be surprised that
-% management approaches we've known to
-% frustrate pieceworkers since \citeyear{10.2307/2555446}
-% would only aggravate crowd workers in many of the same ways?''}
-
 Relative to the mature state of collective action for pieceworkers offline, crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
@@
 
+\onlyinsubfile{\newpage}
 \subsubsection{\implication}
-
-%	*workers make little money but love autonomy --- workers make little money
-%	workers blamed for quality --- ???
-%	both cases, sociality is hard
-%	*collective action hard --- collective action succeeded
-%	- algorithms, not managers
 
@@
 \topic{The history of piecework further suggests that relationships between workers and employers might be improved if employers engaged in more human management styles.}
-% \citeauthor{10.2307/2555446} drew a dichotomous line between
-% ``defective (simple, observable) output measures'' and
-% ``a manager's subjective judgment'',
-% but such a dichotomy need not necessarily represent our work management styles
-% \cite{10.2307/2555446}.
 Instead of delegating as many management tasks as possible to an algorithm, it might be possible to build dashboards and other information tools that empower modern crowd work foremen~\cite{kulkarni2012mobileworks}.
@@
 
-% \paragraph[relationships to work]{Relationships to work}\label{ch:workRelationship}
-% \topic{A number of ethical questions surrounding the
-% increasing complexity of crowd work have arisen in recent years.}
-% \citeauthor{professionalcrowdworkEthics} bring some of these issues to light
-% --- working for increasing amounts of time on tasks of growing complexity%
-% , only to discover that requesters are not willing to pay%
-% , for instance ---
-% but these and other dangers range an enormous landscape
-% \cite{professionalcrowdworkEthics}.
-% \citeauthor{crowdworkFuture}
-% list a few of the problems they identified in \citeyear{crowdworkFuture}
-% --- motivation, feedback, reputation, quality control, to name a few ---
-% while others discuss challenges such as fostering collective action
-% and the opportunity for learning and career advancement
-% \cite{crowdworkFuture,nickerson2013crowd,dynamo}.
-
-% Some research already looks at research such as investing in workers, and
-% informally, we know that this happens among industry requesters
-% \cite{jonBrelig,shepherdingDow}.
-% AMT, meanwhile, offers requesters the ability to create tasks which are
-% not just hidden from unqualified workers by default, but completely.
-% Requesters have taken to using lists of worker IDs which reference
-% workers who have proven their reliability,
-% representing a sort of proto--organization of loosely connected workers.
-
-% \subsubsubsection{\pieceworkpers}
-% The rise of labor unions in the \nth{20} century seems to have been precipitated by
-% egregiously unjust conditions imposed on workers in factories and elsewhere
-% \cite{ebbinghaus1999institutions}.
-% Incidents broadly describing this dynamic can be found in research on AMT
-% \cite{turkopticon,dynamo}.
-% If these are prototypical labor advocacy organizations of contemporary on--demand work,
-% the next question we should look to is if
-% --- and indeed \textit{how} ---
-% these institutions might face challenges in the future.
-
-% For insight on this, we return to \citeyear{levi2009union}'s study of labor unions,
-% and identify that
-% ``Scholars who evaluate union governance by procedural criteria generally find that oligarchy tends to arise and persist even when democratic procedures are in place''
-% \cite{levi2009union}.
-% Indeed, \citeauthor{levi2009union} writes about the general perception that labor unions were either
-% This perception already appears to be emerging in digitally mediated peer--governed organizations,
-% as \citeauthor{keegan2010egalitarians} and others have illustratively documented
-% \cite{beschastnikh2008wikipedian,keegan2010egalitarians}.
-% If these organizations and others are to avoid the same fate that labor unions faced,
-% they should take care to study this phenomenon and attempt to avoid it.
-
-% \subsubsubsection{\whatchanged}
-% \topic{While online and distributed workers can be harder to find,
-% other features make the study thereof rewarding for the substantively different ways
-% that we can approach their study.}
-% The longitudinal analysis of everyday discourse, for instance,
-% is trivial in the study of online communities
-
-% \subsubsubsection{\implication}
-
-% \onlyinsubfile{
-% \balance{}
-% \printbibliography
-% }
+\onlyinsubfile{
+\balance{}
+\printbibliography
+}
 \end{document}

### 2016-12-15 drafts/labor/piecework_lit.tex
notes from msb added
moved some things around and refactored the earlier parts of the piecework literature review
i didn't bother to change the narrative of the sections yet, in part because i don't really know how to rewrite it from the chronological ordering, especially going from the industrial work->farmwork sections, since taylorism was sort of a directional bridge. maybe i could write it so it's more like 'where did taylorism come from though? for that, we turn to farmwork...'?

@@
 Specifically, we will
-\begin{inlinelist}
-  \item define ``piecework'' as researchers in the topic understood it; and
+\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
+  \item define ``piecework'' as researchers in the topic understood it;
   \item trace the rise and fall of piecework at a very high level,
@@
 In summary, piecework:
-\begin{inlinelist}
+\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
   \item paid workers for quantity of work done, rather than time done,
         but occasionally mixed the two payment models;
-  \item afforded workers freedom in when and how much to work; and
+  \item afforded workers freedom in when and how much to work;
   \item structured tasks such that people who didn't have the training
@@
 \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
-\topic{In this section we will offer a brief overview of the history of piecework;
-this should not be mistaken for a comprehensive background.}
-Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
-In other words, it will frame piecework in the contexts of
-the early days of the Industrial Revolution,
+\topic{In this section we'll discuss the history of piecework at a high level
+to contextualize the observations and conclusions we later make,
+rather than provide a comprehensive historical account.}
+% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
+In other words, it will frame piecework,
+in the early days of the Industrial Revolution \&
 through the political and economic turmoil of the early and mid--\nth{20} century,
@@
 While the previous section provided a \textit{definition} of piecework,
-this section attempts to shine a light on the \textit{zeitgeist} of piecework.
+this section attempts to shine a light on the \textit{zeitgeist} of piecework;
+as close to a \textit{thick description} of piecework
+as we can afford~\cite{geertz1973interpretation}.
 
@@
 (famously documented by
-\citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
+\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
 Discontent reached a crescendo when industry organizations representing
@@
 The salient suggestions include:
-\begin{inlinelist}
+\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
 \item the emergence of more effective, more nuanced incentive models
@@
 \item the shifting of these industries (manufacturing, clothing, etc\dots)
-to other countries; and
+to other countries;
 \item the quality of ``multidimensional'' work becoming too difficult to evaluate
@@
        % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
-\begin{numberlist}
-  \item Airy's employment of \textit{human computers},
-  \item domestic and farm work (in particular, the ``match--girls''), and
-  \item railroad workers across the United States.
+\begin{numberlist}[itemjoin*={~and~},itemjoin={~}]
+  \item railroad workers across the United States;
+  \item Airy's employment of \textit{human computers};
+  \item domestic and farm work (in particular, the ``match--girls'').
+
 \end{numberlist}
@@
 
-\subsubsection{Airy's Computers}
-\topic{Some of the first systematic cases of what we would recognize as crowd work
-can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
-% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
-% piecework, found in George Airy's creation of
-% the British Nautical Almanac.}
-Airy needed to compute tables that would
-allow sailors to locate themselves by starlight from sea.
-This work ostensibly called for educated people who comprehensively understood mathematics.
-Airy realized that he could break the tasks down and delegate the constituent parts
-to ``human computers'' who
-``\dots~possessed the basic skills of mathematics,
-including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
-As a result, many of Airy's computers had relatively rudimentary educations
-compared to the background of education that typically worked in the calculation of solar tables.
-Airy distributed tasks by mail,
-allowing work to be completed by a somewhat geographically distributed workforce,
-and paid for each piece of work completed.
+% \ali{maybe go backwards on this}
 
-\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
-For one thing,
-Airy had set a policy of firing his computers once they reached age 23.
-This practice ensured two outcomes that arguably disfavored workers.
-\begin{Numberlist}
-\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
-      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
-\item it limited workers' ability to organize
-      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
-\end{Numberlist}
 % \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
 % Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
+
+\subsubsection{Railroad and other Industrial Workers}
+\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
+piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
+Scientific management argued that it was possible to measure work at
+unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
+As \citeauthor{Brown01041990} points out,
+piecework most greatly benefits the instrumented measurement of workers,
+but in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
+as rudimentary guides on per--task compensation.
+This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
+railroad maintenance, automobile manufacturing, and later weapons manufacturing.
+\ali{???? better word for ``weapons manufacturing''?}
+% Instead, as a result,
+% the distillation of work into smaller units ultimately
+% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+
+% \topic{With the formalization of scientific management,
+% piecework became an important component of the mass manufacturing
+% in American factories.}
+% \topic{}
+
+\topic{The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.}
+\citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+They attribute this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
 
@@
 
-\subsubsection{Railroad and other Industrial Workers}
-\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
-piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
-Scientific management argued that it was possible to measure work at
-unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
-As \citeauthor{Brown01041990} points out,
-piecework most greatly benefits the instrumented measurement of workers,
-% but certainly in Ford and Taylor's time,
-but in \citeauthor{taylor1914principles}'s time highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
-as rudimentary guides on per--task compensation.
-This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
-railroad maintenance, automobile manufacturing, and later weapons manufacturing.
-\ali{???? better word for ``weapons manufacturing''?}
-% Instead, as a result,
-% the distillation of work into smaller units ultimately
-% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+\subsubsection{Airy's Computers}
+\topic{Some of the first systematic cases of what we would recognize as crowd work
+can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
+% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
+% piecework, found in George Airy's creation of
+% the British Nautical Almanac.}
+Airy needed to compute tables that would
+allow sailors to locate themselves by starlight from sea.
+This work ostensibly called for educated people who comprehensively understood mathematics.
+Airy realized that he could break the tasks down and delegate the constituent parts
+to ``human computers'' who
+``\dots~possessed the basic skills of mathematics,
+including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
+As a result, many of Airy's computers had relatively rudimentary educations
+compared to the background of education that typically worked in the calculation of solar tables.
+Airy distributed tasks by mail,
+allowing work to be completed by a somewhat geographically distributed workforce,
+and paid for each piece of work completed.
+\ali{also talk about the challenge/competition Airy was in with other astronomers,
+and the approach of verifying worker results}
 
-% \topic{With the formalization of scientific management,
-% piecework became an important component of the mass manufacturing
-% in American factories.}
-% \topic{}
-
-\topic{The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.}
-\citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-They attribute this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women~\cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
+\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
+For one thing,
+Airy had set a policy of firing his computers once they reached age 23.
+\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
+\begin{Numberlist}
+\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
+      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
+\item it limited workers' ability to organize
+      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
+\end{Numberlist}}
 

### 2016-12-17 drafts/labor/cases.tex
piecework_lit drafting

@@
 
-\topic{Research in crowdsourcing has spent the better part of a decade
-exploring how to grow the limits of crowdsourcing and
-find the boundaries of crowd work and microtasks.}
-This has largely involved
-  identifying challenges to this form of labor,    overcoming them through novel designs of work--flows and processes, and
-  repeating the process~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
-The question that has emerged
-among these researchers and
-through the work that they have produced then
-has been driving at \textit{whether} there are limits to crowdsourcing
-(and, if so, what factors determine those limits).
-Through this lens, we can point to
-a number of contributions to the field that have extended the boundaries of crowd work.
+{Research in crowdsourcing has spent the better part of a decade
+exploring how to grow the limits of crowdsourcing,
+finding the boundaries of crowd work and microtasks.}
+This has largely involved iteratively
+  identifying challenges to increasing complexity and
+  overcoming them through novel designs of work--flows and processes~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
+The question that has emerged then
+has been \textit{whether} there are limits to crowdsourcing
+--- and if so, what factors determine them.
+Through this lens,
+a number of contributions to the field have pushed the boundaries of crowd work.
 
@@
 manipulating and extending along three dimensions:
-\begin{inlinelist}
-  \item \nameref{sec:complexity},
-  \item \nameref{sec:decomposition}, and
-  \item \nameref{sec:relationships}.
-\end{inlinelist}
+\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
+  \item \namerefl{sec:complexity}
+  \item \namerefl{sec:decomposition}
+  \item \namerefl{sec:relationships}.
+\end{Numberlist}
 We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be

### 2016-12-17 drafts/labor/introduction.tex
piecework_lit drafting

@@
 \section{Introduction}\label{sec:introduction}
-\topic{The past decade has seen a flourishing of \textit{on--demand work},
+{The past decade has seen a flourishing of \textit{on--demand work},
 largely driven by the reformulation of work as
-the constituent parts of larger tasks.}
-This framing of work into de--contextualized, modular blocks
-enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
-Distributed paid participants then engage in the work whenever their schedules allow,
+constituent parts of larger tasks.}
+This framing of work into modular blocks
+has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+Distributed paid participants then engage in work whenever their schedules allow,
 with little to no awareness of the broader context of the work, and
-with (often) fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
+often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
 In this paper, we use the term on--demand work to join a pair of related phenomena:
-\begin{inlinelist}
-\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
-\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.~
-\end{inlinelist}
+\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
+\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work
+\item \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
+\end{numberlist}
 The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
@@
 developing systems and designs that enable new forms of production
-(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
+\cite[e.g.][]{bernsteinSoylent,vizwiz,paolacci2010running}.
 
@@
 What does the future hold for on--demand work and those who do it?}
-Researchers have offered insights on this question along three major threads:
-\begin{enumerate}
+Researchers have offered insights on this question along three major threads.
+\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
   \item \nameref{sec:complexity}?
@@
               Nebeling:2016:WCW:2858036.2858169,
-              Hahn:2016:KAB:2858036.2858364};
-  \item \nameref{sec:decomposition}?
+              Hahn:2016:KAB:2858036.2858364}
+  \item \nameref{sec:decomposition}
         \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
               Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
-              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
+              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
   \item \nameref{sec:relationships}?
         \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
-              whyWouldAnyoneBrewer,takingAHITMcInnis}
-\end{enumerate}
-This research has largely sought to answer these questions by examining the present on--demand work phenomenon.
+              whyWouldAnyoneBrewer,takingAHITMcInnis}.
+\end{Numberlist}
+
+This research has largely sought to answer these questions by examining extant on--demand work phenomena.
 So far, it has not offered a framing for holistically explaining
@@
 the emergent phenomena in social environments;
-nor has any research, to our knowledge,
-gone as far as directly predict future developments.
+nor has any research
+% , to our knowledge,
+gone so far as to directly predict future developments.
 
@@
 an instantiation (or even a continuation) of piecework,
-and by looking for patterns of behavior that the corresponding literature predicts
-on this basis, we can
-\begin{inlinelist}
-  \item make sense of the phenomena so far as part of a much larger series of interrelated events;
-  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today;
-        and finally,
-  \item to the extent that history repeats itself, offer predictions of what on--demand work researchers,
+and by looking for patterns that the corresponding literature predicts
+on this basis, we can:
+\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
+  \item make sense of past events as part of a much larger series of an interrelated phenomenon
+  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today
+  \item to some extent, offer predictions of what on--demand work researchers,
         and workers themselves,
         might expect to see on the horizon.
-\end{inlinelist}
-For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.
+\end{numberlist}
+For example, we will draw on the piecework literature studying task decomposition,
+which was historically limited by technological limits in measurement and instrumentation, and
+leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.
 

### 2016-12-17 drafts/labor/piecework_lit.tex
piecework_lit drafting

@@
 
+\onlyinsubfile{\clearpage}
 \section{A Review of Piecework}
@@
 Specifically, we will
-\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
-  \item define ``piecework'' as researchers in the topic understood it;
+\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
+  \item define ``piecework'' as researchers in the topic understood it
   \item trace the rise and fall of piecework at a very high level,
-        identifying key figures and ideas during this time;
+        identifying key figures and ideas during this time
 \end{inlinelist}
@@
 
+% \onlyinsubfile{\clearpage}
 \subsection{What is piecework?}
@@
 In summary, piecework:
-\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
+\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
   \item paid workers for quantity of work done, rather than time done,
-        but occasionally mixed the two payment models;
-  \item afforded workers freedom in when and how much to work;
+        but occasionally mixed the two payment models
+  \item afforded workers freedom in when and how much to work
   \item structured tasks such that people who didn't have the training
@@
 \item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
 \end{Numberlist}
@@
 
+% \onlyinsubfile{\clearpage}
 \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
@@
 The salient suggestions include:
-\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
+\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
 \item the emergence of more effective, more nuanced incentive models
---- rewarding teams for complex achievements, for instance;
+--- rewarding teams for complex achievements, for instance
 \item the shifting of these industries (manufacturing, clothing, etc\dots)
-to other countries;
+to other countries
 \item the quality of ``multidimensional'' work becoming too difficult to evaluate
@@
 
+% \onlyinsubfile{\clearpage}
 \subsection{Case studies in piecework}
@@
        % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
-\begin{numberlist}[itemjoin*={~and~},itemjoin={~}]
-  \item railroad workers across the United States;
-  \item Airy's employment of \textit{human computers};
+\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
+  \item railroad and other industrial workers across the United States
+  \item Airy's employment of \textit{human computers}
   \item domestic and farm work (in particular, the ``match--girls'').
-
 \end{numberlist}
+We'll approach this by looking at familiar cases of decomposed, distributed work (industrial workers),
+and make our way backwards through the \nth{20} and then \nth{19} century,
+tracing the origins of the ideas that
 
@@
 % Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
+\onlyinsubfile{\clearpage}
+\subsubsection{Railroad and other Industrial Workers}
+{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
+It was here that
+scientific management, Fordism, and Taylorism
+dramatically influenced how workers were managed and
+the ways in which workers were perceived and envisioned paradigmatically;
+it's here we'll start our overview of piecework's history through case studies.
 
-\subsubsection{Railroad and other Industrial Workers}
-\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
-piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
-Scientific management argued that it was possible to measure work at
-unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
-As \citeauthor{Brown01041990} points out,
-piecework most greatly benefits the instrumented measurement of workers,
-but in \citeauthor{taylor1914principles}'s time highly instrumented,
+\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
+found its way into the war effort during World War II.}
+With the vast majority of men drafted into service,
+factories found themselves turning to
+a predominantly female workforce that had neither
+the formal training nor
+the years of apprenticeship experience
+that conventional workers would have had.
+Rather than attempting to train this new labor force in every aspect of industrial work,
+these women were trained for individual tasks
+% --- such as riveting component to another ---
+and assigned to that task. %, decomposed from a more comprehensive task.
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
+
+% This approach would turn out to yield two benefits:
+% \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
+% \item these workers could be productive with significantly less ``on--boarding'' cost
+% \item workers' performance could be tracked \& evaluated, and better methods could be disseminated to workers.
+% \end{numberlist}
+
+{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's eponymously named \textit{Fordism}),
+piecework in the early and mid--\nth{20} century surged, especially in industrial work.}
+Scientific management promised that the careful measurement of workers would yield
+higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
+While \citeauthor{Brown01041990} points out that
+piecework dramatically advanced the instrumented measurement of workers,
+in \citeauthor{taylor1914principles}'s time highly instrumented,
 automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
-as rudimentary guides on per--task compensation.
-This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
-railroad maintenance, automobile manufacturing, and later weapons manufacturing.
-\ali{???? better word for ``weapons manufacturing''?}
+Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
+using completion times to inform per--task compensation.
+% Under this regime, piecework became commonplace in railroad maintenance and a number 
+% This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
+% railroad maintenance, automobile manufacturing, and later weapons manufacturing.
+% \ali{???? better word for ``weapons manufacturing''?}
 % Instead, as a result,
@@
 affording them more comprehensive knowledge of the total scope of work.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women~\cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
 
+\onlyinsubfile{\clearpage}
 \subsubsection{Domestic and Farmhand Labor}
@@
 \topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
-In 1912, national coal miners effected an individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
-Women in the garment industry in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
+Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
+Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
 It was in the midst of this time that \citeauthor{taylor1914principles} published
 the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
-% and it was this framing on work that
 It was this framing on work \& worker management that gave workers a concrete adversary
---- if not Taylor himself, then \textit{Taylorism}  ---
+--- if not Taylor, then \textit{Taylorism}  ---
 against which to rally~\cite{jacoby1983union,parker1920casual}.
 
+\onlyinsubfile{\clearpage}
 \subsubsection{Airy's Computers}
 \topic{Some of the first systematic cases of what we would recognize as crowd work
-can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
-% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
-% piecework, found in George Airy's creation of
-% the British Nautical Almanac.}
+can be found in the study of astronomy.}
+In the \nth{19} century, the calculation of celestial bodies had become
+a competitive field with 
 Airy needed to compute tables that would
@@
 and paid for each piece of work completed.
-\ali{also talk about the challenge/competition Airy was in with other astronomers,
-and the approach of verifying worker results}
+% \ali{also talk about the challenge/competition Airy was in with other astronomers,
+% and the approach of verifying worker results}
+% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
+% For one thing,
+Airy also instituted a policy of firing his computers once they reached age 23.
 
-\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
-For one thing,
-Airy had set a policy of firing his computers once they reached age 23.
 \ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
@@
 \onlyinsubfile{
-  \printbibliography
+  % \printbibliography
   }

### 2016-12-22 drafts/labor/abstract.tex
intro and history

@@
 \begin{abstract}
-The internet is enabling the rise of
+The internet is empowering the rise of
 crowd work,
@@
 A large and growing body of scholarship has attempted to predict
-the socio--technical outcomes of this shift, especially along three threads:
+the socio--technical outcomes of this shift, especially addressing three questions:
 \begin{inlinelist}

### 2016-12-22 drafts/labor/introduction.tex
intro and history

@@
 \section{Introduction}\label{sec:introduction}
+\msb{Please take a full pass through for consistent terminology.
+In the intro, we align on on--demand work and call it the union of crowd work and gig work.
+But then, in the title and the research questions, everything is crowd work, crowd work, crowd work.
+I suggest either using crowd work in the introduction and explaining that you're using that to mean both traditional crowd work as well as modern gig work, or using on--demand work in the title and in the research questions.
+I see the point ``(which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature)'', but it feels like a cop out.
+Which one do we want to use, really?
+Let's pick one and be consistent throughout.}
+
 {The past decade has seen a flourishing of \textit{on--demand work},
@@
 In this paper, we use the term on--demand work to join a pair of related phenomena:
+\msb{this list is not rendering correctly in the document. check out all the whitespace before `second'. I think it has to do with the use of ~and~.}
 \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
@@
 The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
-like AMT to explore the limits of this distributed, on--demand workforce.
+such as AMT to explore the limits of this distributed, on--demand workforce.
 Researchers have also taken to the space in earnest,
 developing systems and designs that enable new forms of production
+\msb{The e.g., should be in parentheses, not brackets. (e.g., [1, 2, 3])}
 \cite[e.g.][]{bernsteinSoylent,vizwiz,paolacci2010running}.
@@
 Researchers have offered insights on this question along three major threads.
+\msb{The questions are rendering in the PDF as capitalized in the middle of a sentence.
+It's cool, but for future papers please proofread the rendered PDF before sending me to look at the camready draft}
+\msb{Some of these are missing question marks at the end of the sentence too}
 \begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
@@
 referenced crowd work as piecework briefly
-as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
+as a loose analogy~\cite{crowdworkFuture}.
 But more than this,
 the framing of on--demand labor as a re--instantiation of piecework
-gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.
+gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirical research.
 
@@
 on this basis, we can:
+\msb{again weird whitespace within the sentence}
 \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
-  \item make sense of past events as part of a much larger series of an interrelated phenomenon
+  \item make sense of past events as part of a much larger series of interrelated phenomena
   \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today
@@
 leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.
+\msb{can you make this more concrete?
+what specific suggestion will we make?
+rather than just stating that we will make one...}
 
@@
 on--demand work as we experience it today.
-Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
-Finally, we will offer implications for researchers and practitioners based on our results.
+Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework, and offer implications for researchers and practitioners.
 

### 2016-12-22 drafts/labor/piecework_lit.tex
intro and history

@@
 this casual familiarity with piecework may prove insufficient.
-We'll more carefully discuss piecework in this section in order
+We will more carefully discuss piecework in this section in order
 to inform the the rest of the argument.
@@
 \end{inlinelist}
+\msb{missing ending punctuation}
 This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.
@@
 examples, and
-provided definitions of piecework, trying
-to trace the outline of a working understanding of
-\textit{what piecework is}.
+provided definitions of piecework, tracing a working understanding.
 
@@
 offering examples:
+\msb{the spacing is not even on the two sides of the ~[\dots]~ in the PDF}
 ``payment is made for each hectare which is pronounced to be well ploughed~[\dots]~
@@
 time--rate compensation in others.
-The ``Rowan premium system'',
+The Rowan premium system,
 which essentially paid workers
@@
 (the potential for) an additional pay dependent on output,
-was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
-muddies the waters for us later as we attempt to categorize cases of piecework~\cite{rowan1901premium}.
+was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms%, which
+%muddies the waters for us later as we attempt to categorize cases of piecework
+~\cite{rowan1901premium}.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
@@
 crop up; he points out that 
+\msb{more uneven spacing in the quotes. will stop pointing it out here, please fix throughout}
 ``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~[\dots]~
@@
 Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. 
+\msb{This transition seems misplaced with the Piecework primer and Case Studies subsections up next:}
 In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.
@@
 \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
-\topic{In this section we'll discuss the history of piecework at a high level
-to contextualize the observations and conclusions we later make,
-rather than provide a comprehensive historical account.}
+\msb{This section is so brief that it feels like a bit of a cheat.
+I think we need to frame it differently.
+Maybe that it should survey the different ``eras'' of piecework?
+And each paragraph is laying out a major historical era that we will return to later?
+Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
+\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.}
 % Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
-In other words, it will frame piecework,
+This section will frame piecework,
 in the early days of the Industrial Revolution \&
@@
 as we can afford~\cite{geertz1973interpretation}.
+\msb{is this really thick? as a reader it flies by really quickly and super high level...}
 
@@
 but it found its foothold in manual labor.}
-Piecework took off on in farm work~\cite{hughRaynbirdTaskWork},
+Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
 in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
@@
 a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
+\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
+\msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
 \citeauthor{hart2013rise} point out that the Second World War,
@@
 
+\msb{The topic sentence is about management, but the content is about discontent. Which is it?}
 The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
-Despite this, workers' needs were mostly ignored,
+In this conversation, workers' needs were mostly ignored,
 leading to frustration over poor working and living conditions
@@
 
+\msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
 \topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
+\msb{How quickly did it plummet? By when was it gone? Where is it now?}
+\msb{This sentence is empty calorie:}
 \citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
+\msb{Very few calorie sentence:}
 In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
@@
 --- rewarding teams for complex achievements, for instance
-\item the shifting of these industries (manufacturing, clothing, etc\dots)
+\item the shifting of these industries (e.g., manufacturing, clothing)
 to other countries
@@
 \subsection{Case studies in piecework}
-We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
+Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
        % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
 \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
-  \item railroad and other industrial workers across the United States
+  \item railroad and other industrial workers
   \item Airy's employment of \textit{human computers}
@@
 \end{numberlist}
-We'll approach this by looking at familiar cases of decomposed, distributed work (industrial workers),
-and make our way backwards through the \nth{20} and then \nth{19} century,
+Here, we introduce the basic historical details of these three cases.
+We begin with the most familiar case (industrial workers),
+working backwards through the \nth{20} and then \nth{19} century,
 tracing the origins of the ideas that 
+\msb{incomplete sentence}
 
@@
 the ways in which workers were perceived and envisioned paradigmatically;
-it's here we'll start our overview of piecework's history through case studies.
+it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}
 
+\msb{I'm feeling very disoriented in this paragraph. The subsection title is about railroads, but we're all fast-forwarded to auto manufacturing?}
 \topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
@@
 
+\msb{confusing timing; now we're going backwards?}
 {With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
@@
 
+\msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
+\msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
 \topic{The 1930s represented a boom for piecework on an unprecedented scale,
@@
 affording them more comprehensive knowledge of the total scope of work.
+
+\msb{This subsection feels like it never delivered on what it promised: a case study of one group of pieceworkers.
+I thought the goal was to introduce the railroad workers --- who they were, what they did, what their conditions were like.}
 
@@
 it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+\msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
 The complexity of the work remained low at the turn of
 the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+\msb{how is this sentence relevant to the case study?:}
 However, writers of the time focused their attention on
@@
 \topic{Workers' relationships with employers quickly soured.}
+\msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
 The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
@@
 \topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
+\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
 Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
 Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
+\msb{Why is Taylor relevant to this case study?}
 It was in the midst of this time that \citeauthor{taylor1914principles} published
@@
 \subsubsection{Airy's Computers}
+\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
 \topic{Some of the first systematic cases of what we would recognize as crowd work

### 2016-12-22 drafts/labor/trackingWork.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess
notes from msb added
piecework_lit drafting
intro and history

@@
 
+\newlist{numberlist}{enumerate*}{1}
+\setlist*[numberlist,1]{%
+   % label=\alph*),
+ label={\nthwords*,},
+}
+
+\newlist{Numberlist}{enumerate*}{1}
+\setlist*[Numberlist,1]{%
+   % label=\alph*),
+ label={\Nthwords*,},
+}
 \newlist{inlinelist}{enumerate*}{1}
@@
 \setlist{nolistsep}
-\def\plaintitle{Examining Crowd Work Through The Historical Lens of Piecework}
-\def\plainauthor{Ali Alkhatib, Margaret Levi, Michael Scott Bernstein}
+\def\plaintitle{Examining Crowd Work through The Historical Lens of Piecework}
+\def\plainauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi}
 \def\emptyauthor{}
@@
 
-\numberofauthors{3}
+\numberofauthors{1}
 \author{%
-  \alignauthor{Ali Alkhatib\\
+  \alignauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi\\
+    \affaddr{Computer Science Department and CASBS}\\
     \affaddr{Stanford University}\\
-    \affaddr{Stanford, USA}\\
-    \email{ali.alkhatib@cs.stanford.edu}}\\
-  \alignauthor{Michael Scott Bernstein\\
-    \affaddr{Stanford University}\\
-    \affaddr{Stanford, USA}\\
-    \email{msb@cs.stanford.edu}}\\
-  \alignauthor{Margaret Levi\\
-    \affaddr{CASBS}\\
-    \affaddr{Stanford, USA}\\
-    \email{mlevi@stanford.edu}}\\
+    \email{\{ali.alkhatib, msb\}@cs.stanford.edu, mlevi@stanford.edu}}\\
 }
@@
 \keywords{\plainkeywords}
+\ali{maybe put a table with high level summaries}
 \subfile{introduction.tex}

### 2016-12-31 drafts/labor/introduction.tex
piecework changes to intro and lit review

--- 2016-12-17 drafts/labor/introduction.tex
@@
 However, for years its instantiation was limited to the utilization of
-human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,
-      yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
+human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
 More recently, crowdsourcing of physically embodied work
@@
 This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
-Some of this research has been motivated by the identification of the sociality of gig work,
+Some of this research has been motivated by the identification of the sociality of gig work~\cite{crowdcollab},
 and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
@@
 What does the future hold for on--demand work and those who do it?}
-Researchers have offered insights on this question along three major threads.
-\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
+Researchers have offered insights on this question along three major threads:
+\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
   \item \nameref{sec:complexity}?
@@
               whyWouldAnyoneBrewer,takingAHITMcInnis}.
-\end{Numberlist}
+\end{numberlist}
 
@@
 nor has any research
-% , to our knowledge,
 gone so far as to directly predict future developments.
@@
 \subsection{Piecework as a lens to understand crowdsourcing}
-In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework}: %\namerefl{sec:pieceworkArc} ---
+In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
 a work and payment structure which breaks tasks down into standalone contracts,
@@
 an instantiation (or even a continuation) of piecework,
-and by looking for patterns that the corresponding literature predicts
-on this basis, we can:
-\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
+and by interrogating patterns that the corresponding literature predicts
+on this basis, we can
+\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
   \item make sense of past events as part of a much larger series of an interrelated phenomenon
-  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today
+  \item reflect on differences in the factors that impacted piecework historically and their impact on--demand work today
   \item to some extent, offer predictions of what on--demand work researchers,
@@
 \end{numberlist}
-For example, we will draw on the piecework literature studying task decomposition,
-which was historically limited by technological limits in measurement and instrumentation, and
+For example, we will draw on the piecework literature regarding task decomposition,
+which was historically limited by shortcomings in measurement and instrumentation, and
 leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.
 
-We organize this paper as follows: we first review the definition and historical arc of piecework to lay  groundwork and make clear the analogy to on--demand work (which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature).
-Then, we interrogate the three major research questions above using the lens of piecework.
+We organize this paper as follows:
+\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
+  \item we review the definition and historical arc of piecework
+  to lay groundwork and make clear the analogy to on--demand work
+  (which we'll call \textit{crowd work} hereon
+  for consistency with the existing literature)
+  \item we interrogate the three major research questions above using the lens of piecework. 
   We will identify similarities and differences between piecework as historically understood and
-on--demand work as we experience it today.
-Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
-Finally, we will offer implications for researchers and practitioners based on our results.
+  on--demand work as we know it today
+  \item we will make predictions of future developments based on how those similarities and differences influenced piecework
+  \item we will offer implications for researchers and practitioners based on our results.
+\end{numberlist}
 
-\onlyinsubfile{
-  \balance{}
-  \printbibliography
-}
+% \onlyinsubfile{
+%   \balance{}
+%   \printbibliography
+% }
 

### 2016-12-31 drafts/labor/piecework_lit.tex
piecework changes to intro and lit review

--- 2016-12-17 drafts/labor/piecework_lit.tex
@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \section{A Review of Piecework}
@@
 Specifically, we will
-\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
-  \item define ``piecework'' as researchers in the topic understood it
-  \item trace the rise and fall of piecework at a very high level,
-        identifying key figures and ideas during this time
-\end{inlinelist}
+\begin{numberlist}[itemjoin*={, and },label={\nthwords*}]
+  \item define ``piecework'' as researchers in its field understand it
+  \item trace the rise and fall of piecework at a high level,
+        identifying key figures and ideas during this time.
+\end{numberlist}
 This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.
@@
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
-While ``piecework'' has had multiple definitions over time,
+While it has had several definitions over time,
 we can trace a constellation of characteristics that recur throughout the literature.
-We will follow this history of research, collecting
+We'll follow this research, collecting
 descriptions,
 examples, and
-provided definitions of piecework, trying
-to trace the outline of a working understanding of
+definitions of piecework,
+tracing an outline of % a working understanding of
 \textit{what piecework is}.
 
-\citeauthor{hughRaynbirdTaskWork} offers
+\citeauthor{hughRaynbirdTaskWork} provides
 a concise definition of piecework
---- which he variously also calls ``measure work'', ``grate work'', and ``task work'' ---
-by contrasting the ``task--labourer'' with the ``day--labourer'':
+(which he also calls ``measure work'', ``grate work'', and ``task work'')
+by contrasting with the ``day--labourer'':
 ``the chief difference lies between the day--labourer,
-who receives a certain some of money\dots~for his day's work,
-and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''~\cite{hughRaynbirdTaskWork}.
-\citeauthor{10.2307/2338394} gives a more illustrative definition of piecework,
-offering examples:
-``payment is made for each hectare which is pronounced to be well ploughed~[\dots]~
-for each living foal got from a mare;~[\dots]~
-for each living calf got''~\cite{10.2307/2338394}.
-This framing perhaps makes the most intuitive sense;
-``payment for results,'' as \citeauthor{10.2307/2338394} calls it,
-is not only common in practice, but well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
+who receives a certain some of money~\dots~for his day's work,
+and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
+\citeauthor{10.2307/2338394} defines it through examples:
+``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
+This framing offers a more intuitive sense;
+``payment for results,'' as he calls it,
+is not only common in practice, but
+well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
 
@@
 ``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
-Indeed, work adopted piece--rate compensation in some aspects and
+Indeed, employers adopted piece--rate compensation in some aspects and
 time--rate compensation in others.
-The ``Rowan premium system'',
+The Rowan premium system,
 which essentially paid workers
-a base rate for time plus
-(the potential for) an additional pay dependent on output,
-was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
-muddies the waters for us later as we attempt to categorize cases of piecework~\cite{rowan1901premium}.
+a base rate for time plus, potentially,
+an additional pay dependent on output~\cite{rowan1901premium};
+this was just one of several alternatives to categorical time-- and piece--rate renumeration paradigms.
+% which will complicate our categorization of cases of piecework.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
 regardless of the worker's productive output
-\textit{as well as} an additional compensation tied to performance,
+\textit{as well as} additional compensation tied to performance,
 workers under this regime were
-in some senses ``task--labourers'', and
-in other senses (more conventional) ``day--labourers''.
+in some senses ``task--labourers'', but
+in other senses ``day--labourers''.
 
 \topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
-Returning to
-\citeauthor{hughRaynbirdTaskWork}, several arguments for the merits of piecework
-crop up; he points out that 
-``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~[\dots]~
-he knows that all depends on his own diligence and perseverance~[\dots and]~
-so long as he performs his work to the satisfaction of his master,
-he is not under that control to which the day--labourer is always subject.''
-\citeauthor{hughRaynbirdTaskWork} highlight the freedom from control that ``task--labourers'' enjoy~\cite{hughRaynbirdTaskWork,rowan1901premium}.
+\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework, % do you like what i did there?
+pointing out that
+``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~\dots~he knows that all depends on his own diligence and perseverance~\dots~[and] so long as he performs his work to the satisfaction of his master, he is not under that control to which the day--labourer is always subject''.
+The argument that ``task--labourers'' enjoy freedom from control crops up in \citeauthor{hughRaynbirdTaskWork}'s and later \citeauthor{rowan1901premium}'s works~\cite{hughRaynbirdTaskWork,rowan1901premium}.
 
-\topic{We see this sense of independence regardless of the time, locale, and industry.}
+\topic{We see this sense of independence in myriad times, locales, and industries.}
 \citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
---- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
-Of particular interest was their independent nature, via their reputation ``\dots~for generosity, independence, and protectiveness,
+--- young pieceworkers, mostly women, who assembled matchsticks in the late \nth{19} century.
+Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
 but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
-\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
+\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
 finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
-``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
+``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night,
 the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
-From these accounts we identify a sense of independence and autonomy that resonates across decades, industries, and locales where piecework is found.
+We therefore draw a \textit{sense} of independence and autonomy inherent to piecework.
+%that resonates across decades, industries, and locales where piecework is found.
 
-\topic{Piecework opened the door for people who previously couldn't participate in the labor market --- for example due to lack of training --- to do so, and to acquire job skills incrementally.}
-For example, women could receive training in narrow subsets of the general body of skills, enabling them to act in capacities similar to what conventional (male) apprentices would undertake~\cite{hart2013rise}.
-In addition, workers with specific slices of skills could be matched to suitable tasks.
-Workers without conventional training
---- like women, who had no such opportunities
-to engage in engineering and metalworking apprenticeships as men did ---
-could be trained very narrowly on a very tightly constrained task,
-demonstrate proficiency, and become experts in their own ways.
+\topic{Piecework opened the door for people who previously couldn't participate in the labor market to do so, and to acquire job skills incrementally.}
+During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (i.e. male) workers~\cite{hart2013rise}.
+Workers with specific skill subsets could be matched to suitable tasks.
+Women previously had virtually no opportunities
+to engage in engineering and metalworking apprenticeships as men did;
+now, they
+could be trained quickly on narrowly scoped tasks,
+demonstrate proficiency, and become experts.
 
@@
 % Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{Railroad and other Industrial Workers}
@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{Domestic and Farmhand Labor}
@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{Airy's Computers}
@@
 \onlyinsubfile{
-  % \printbibliography
+  \printbibliography
   }

### 2017-01-03 drafts/labor/cases.tex

@@
   identifying challenges to increasing complexity and
-  overcoming them through novel designs of work--flows and processes~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
+  overcoming them through novel designs of work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
 The question that has emerged then
@@
 manipulating and extending along three dimensions:
-\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
-  \item \namerefl{sec:complexity}
-  \item \namerefl{sec:decomposition}
-  \item \namerefl{sec:relationships}.
-\end{Numberlist}
+First, \namerefl{sec:complexity}.
+Second, \namerefl{sec:decomposition}.
+And third, \namerefl{sec:relationships}.
 We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
@@
 
+\onlyinsubfile{
+  \printbibliography{}
+}
+
 \end{document}

### 2017-01-03 drafts/labor/introduction.tex

@@
 \section{Introduction}\label{sec:introduction}
-{The past decade has seen a flourishing of \textit{on--demand work},
+\msb{Please take a full pass through for consistent terminology.
+In the intro, we align on on--demand work and call it the union of crowd work and gig work.
+But then, in the title and the research questions, everything is crowd work, crowd work, crowd work.
+I suggest either using crowd work in the introduction and explaining that you're using that to mean both traditional crowd work as well as modern gig work, or using on--demand work in the title and in the research questions.
+I see the point ``(which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature)'', but it feels like a cop out.
+Which one do we want to use, really?
+Let's pick one and be consistent throughout.}
+
+\ali{\textit{On--demand work} is more ``correct'', but I'm worried that crowdsourcing researchers will skim it and feel there's nothing relevant to them. Is that giving them too little credit?}
+
+\topic{The past decade has seen a flourishing of \textit{on--demand work},
 largely driven by the reformulation of work as
 constituent parts of larger tasks.}
-This framing of work into modular blocks
+This framing of work into modular
 has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
@@
 In this paper, we use the term on--demand work to join a pair of related phenomena:
-\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
-\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work
-\item \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
-\end{numberlist}
+first, \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
+and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
 The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
-like AMT to explore the limits of this distributed, on--demand workforce.
+such as AMT to explore the limits of this distributed, on--demand workforce.
 Researchers have also taken to the space in earnest,
 developing systems and designs that enable new forms of production
-\cite[e.g.][]{bernsteinSoylent,vizwiz,paolacci2010running}.
+(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
 
@@
 Researchers have offered insights on this question along three major threads:
-\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
-  \item \nameref{sec:complexity}?
+% \begin{numberlist}[itemjoin*={.~And~},itemjoin={.~}]
+  First, \nameref{sec:complexity}?
         Specifically,
@@
               Nebeling:2016:WCW:2858036.2858169,
-              Hahn:2016:KAB:2858036.2858364}
-  \item \nameref{sec:decomposition}
+              Hahn:2016:KAB:2858036.2858364}.
+  Second, \nameref{sec:decomposition}
         \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
               Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
-              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
-  \item \nameref{sec:relationships}?
+              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}.
+  And third, \nameref{sec:relationships}?
         \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
               whyWouldAnyoneBrewer,takingAHITMcInnis}.
-\end{numberlist}
+% \end{numberlist}
 
@@
 on this basis, we can
-\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
-  \item make sense of past events as part of a much larger series of an interrelated phenomenon
-  \item reflect on differences in the factors that impacted piecework historically and their impact on--demand work today
-  \item to some extent, offer predictions of what on--demand work researchers,
+% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
+first, make sense of past events as part of a much larger series of an interrelated phenomenon;
+second, reflect on differences in the factors that impacted piecework historically and their impact on--demand work today;
+and third, to some extent, offer predictions of what on--demand work researchers,
 and workers themselves,
 might expect to see on the horizon.
-\end{numberlist}
+% \end{numberlist}
 For example, we will draw on the piecework literature regarding task decomposition,
 which was historically limited by shortcomings in measurement and instrumentation, and
-leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.
+leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
+--- namely, by enabling precise tracking and measurement of workers via algorithms and software.
 
 We organize this paper as follows:
-\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
-  \item we review the definition and historical arc of piecework
+% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
+first, we review the definition and historical arc of piecework
 to lay groundwork and make clear the analogy to on--demand work
 (which we'll call \textit{crowd work} hereon
-  for consistency with the existing literature)
-  \item we interrogate the three major research questions above using the lens of piecework. 
+for consistency with the existing literature);
+second, we interrogate the three major research questions above using the lens of piecework. 
 We will identify similarities and differences between piecework as historically understood and
-  on--demand work as we know it today
-  \item we will make predictions of future developments based on how those similarities and differences influenced piecework
-  \item we will offer implications for researchers and practitioners based on our results.
-\end{numberlist}
+on--demand work as we know it today;
+third, we will make predictions of future developments based on how those similarities and differences influenced piecework;
+finally, we will offer implications for researchers and practitioners based on our results.
+% \end{numberlist}
 

### 2017-01-03 drafts/labor/piecework_lit.tex

@@
 % \onlyinsubfile{\clearpage}
-\subsection{What is piecework?}
+\subsection{What is piecework?}\label{sec:whatIsPiecework}
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
@@
 \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
-\topic{In this section we'll discuss the history of piecework at a high level
-to contextualize the observations and conclusions we later make,
-rather than provide a comprehensive historical account.}
+\msb{This section is so brief that it feels like a bit of a cheat.
+I think we need to frame it differently.
+Maybe that it should survey the different ``eras'' of piecework?
+And each paragraph is laying out a major historical era that we will return to later?
+Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
+\ali{Do you mean that we'd have (sub)subsections under this?
+I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead.
+Not exactly sure how that would work (i.e.
+would I interleave the history or just append it, basically removing the header).}
+\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
 % Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
-In other words, it will frame piecework,
+This section will frame piecework,
 in the early days of the Industrial Revolution \&
@@
 and into the \nth{21} century.
-While the previous section provided a \textit{definition} of piecework,
-this section attempts to shine a light on the \textit{zeitgeist} of piecework;
-as close to a \textit{thick description} of piecework
-as we can afford~\cite{geertz1973interpretation}.
+% While the previous section provided a \textit{definition} of piecework,
+% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
+% as close to a \textit{thick description} of piecework
+% as we can afford~\cite{geertz1973interpretation}.
+% \msb{is this really thick? as a reader it flies by really quickly and super high level...}
 
@@
 a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
-\citeauthor{hart2013rise} point out that the Second World War,
-which called millions of Americans to military service,
-necessitated the rapid training and employment of
-a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
+\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
+\ali{I'm not sure; I'll look for some data.}
+% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
+% \citeauthor{hart2013rise} point out that the Second World War,
+% which called millions of Americans to military service,
+% necessitated the rapid training and employment of
+% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
 
-The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
-Despite this, workers' needs were mostly ignored,
-leading to frustration over poor working and living conditions
+% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
+\topic{The early growth of piecework led to discussion surrounding how best to manage pieceworkers,
+generally in an adversarial approach, consequently
+frustrating workers~\cite{norton1900textile,clark1908cotton}.}
+This frustrations of poor working and living conditions
 (famously documented by
@@
 \subsection{Case studies in piecework}
-We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
+Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
        % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
-\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
-  \item railroad and other industrial workers across the United States
-  \item Airy's employment of \textit{human computers}
-  \item domestic and farm work (in particular, the ``match--girls'').
-\end{numberlist}
-We'll approach this by looking at familiar cases of decomposed, distributed work (industrial workers),
-and make our way backwards through the \nth{20} and then \nth{19} century,
-tracing the origins of the ideas that
+first, railroad and other industrial workers;
+second, Airy's employment of \textit{human computers};
+and third, domestic and farm work (in particular, the ``match--girls'').
+Here, we introduce the basic historical details of these three cases.
+We begin with the most familiar case (industrial workers),
+working backwards through the \nth{20} and then \nth{19} century,
+tracing the origins of the ideas that \dots
+% \msb{incomplete sentence}
 

### 2017-01-04 drafts/labor/introduction.tex
some changes to piecework lit's beginning thinking more carefully about how i want to frame and organize the whole section. in slack i told msb that i'd probably merge the definition and primer stuff together; they don't really make sense apart anyway, and both feel a little too thin

@@
 often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
-In this paper, we use the term on--demand work to join a pair of related phenomena:
+In this paper, we use ``on--demand work'' to capture a pair of related phenomena:
 first, \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
@@
 Researchers have offered insights on this question along three major threads:
-% \begin{numberlist}[itemjoin*={.~And~},itemjoin={.~}]
 First, \nameref{sec:complexity}?
@@
 on this basis, we can
-% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
 first, make sense of past events as part of a much larger series of an interrelated phenomenon;
@@
 We organize this paper as follows:
-% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
 first, we review the definition and historical arc of piecework

### 2017-01-04 drafts/labor/piecework_lit.tex
some changes to piecework lit
i'm thinking more carefully about how i want to frame and organize the whole section.
in slack i told msb that i'd probably merge the definition and primer stuff together;
they don't really make sense apart anyway, and both feel a little too thin

@@
 \topic{The HCI community has used the term ``piecework'' to describe
-myriad instantiations of on--demand labor,
-but this reference has generally been offered in passing.}
-As this paper principally traces a relationship between
-the historical piecework and the contemporary crowd work
-(or on--demand labor more generally),
-this casual familiarity with piecework may prove insufficient.
-We will more carefully discuss piecework in this section in order
-to inform the the rest of the argument.
-Specifically, we will
-\begin{numberlist}[itemjoin*={, and },label={\nthwords*}]
-  \item define ``piecework'' as researchers in its field understand it
-  \item trace the rise and fall of piecework at a high level,
+myriad instantiations of on--demand labor, but
+researchers have generally made this allusion in passing.}
+Since this paper traces a much stronger parallel between
+(historical) piecework and (contemporary) crowd work
+(and, more generally, on--demand labor),
+a more comprehensive background on piecework will be useful.
+We will more carefully discuss piecework in this section
+to help make our observations and arguments with better familiarity with the topic.
+Specifically,
+first, we'll define ``piecework'' as researchers in its field understand it;
+and second, we'll trace the rise and fall of piecework at a high level,
 identifying key figures and ideas during this time.
-\end{numberlist}
-This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.
+This section is not intended to be comprehensive:
+instead, it sets up the scaffolding necessary for
+our later investigations of crowd work's three questions:
+complexity limits,
+task decomposition, and
+worker relationships.
 
-% \onlyinsubfile{\clearpage}
-\subsection{What is piecework?}\label{sec:whatIsPiecework}
+\subsection{What is piecework? A Primer and Timeline}\label{sec:whatIsPiecework}
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
@@
 In summary, piecework:
-\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
+\begin{inlinelist}
   \item paid workers for quantity of work done, rather than time done,
@@
 Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
-\begin{Numberlist}
-\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
-\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
-\end{Numberlist}
+First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
+Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
+Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
 Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. 
@@
 The salient suggestions include:
-\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
+\begin{inlinelist}
 \item the emergence of more effective, more nuanced incentive models
@@
 % This approach would turn out to yield two benefits:
-% \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
 % \item these workers could be productive with significantly less ``on--boarding'' cost
 % \item workers' performance could be tracked \& evaluated, and better methods could be disseminated to workers.
-% \end{numberlist}
 
@@
 % For one thing,
-Airy also instituted a policy of firing his computers once they reached age 23.
+Airy also instituted a policy of firing his computers once they reached age 23
+\ali{this has a ``gosh isn't that interesting?'' feeling to it.
+should i just leave it out until later when there's something to say about it
+(namely, that the mode of work was initially designed to stymie professional growth)?}.
 
 \ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
-\begin{Numberlist}
-\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
+First, it eliminated any potential to advance professionally, as
+workers' careers in this area ended relatively early in their careers,
 and without formal education in mathematics they struggled to find work for which their experience was meaningful.
-\item it limited workers' ability to organize
+Second, it limited workers' ability to organize
 by ensuring that workers hardly spent sufficient time to successfully rally their peers.
-\end{Numberlist}}
+}
 

### 2017-01-06 drafts/labor/introduction.tex
tweaks in introduction

@@
 
+\msb{I think this is a question of which community you want to be influencing.
+Do you want people who study crowdsourcing to be the primary readers+citers of this paper?
+Or people who study Uber and algorithms at work?
+One way to decide this is to name five people who you really hope will read and like this paper, and note which community they are a part of.}
+
 \topic{The past decade has seen a flourishing of \textit{on--demand work},
@@
 constituent parts of larger tasks.}
-This framing of work into modular
+This framing of work into modular \msb{missing word?}
 has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
 Distributed paid participants then engage in work whenever their schedules allow,
-with little to no awareness of the broader context of the work, and
+often with little to no awareness of the broader context of the work, and
 often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
@@
 \topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complex, conflicted culture amongst both the workers who enable it and the researchers who study it.}
+\msb{This paragraph's topic sentence is about the conflicted culture, but the content seems to be about the evolution of the phenomenon. What's the goal here?}
 \citeauthor{howe2008crowdsourcing} first described crowdsourcing broadly as
@@
 leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
---- namely, by enabling precise tracking and measurement of workers via algorithms and software.
+--- namely, by enabling precise tracking and measurement of workers via algorithms and software. \msb{leading to... (let's close the loop)}
 

### 2017-01-06 drafts/labor/piecework_lit.tex
lit rev revs interim

@@
 \section{A Review of Piecework}
+
+\ali{I'm interleaving the paragraphs, so this paragraph needs to be refactored a bit to reflect that instead of suggesting that the road map will look like two distinct threads.
+I'll probably try to take a shot at it in a bit, but for now don't take the wording too seriously.}
 
@@
 \subsection{What is piecework? A Primer and Timeline}\label{sec:whatIsPiecework}
+
+\ali{
+So I'm thinking that this should flow like this:
+``We'll try to understand what piecework was by tracing the development of and discourse around piecework.''
+This'll give us the ability to switch back and forth between the timeline of the practice of piecework
+--- that is, the ``piecework primer'' stuff down below ---
+and the contemporary/concurrent research going on around the same time
+(i.e. this stuff immediately following).
+
+I'm going to interleave the paragraphs, but
+if you think that's a flawed approach, then I'll bet you won't like what you'll see below that much.%
+}
+
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
-While it has had several definitions over time,
+While it has had several definitions over the years,
 we can trace a constellation of characteristics that recur throughout the literature.
@@
 examples, and
-definitions of piecework,
+definitions,
 tracing an outline of % a working understanding of
-\textit{what piecework is}.
+\textit{what piecework is}
+alongside piecework's contemporary developments in practice.
+\ali{This is worded a little awkwardly;
+I want to say that we're looking at
+the discourse about piecework at the same time that we look at
+how piecework is unfolding. (wait, is that how I should word it?)}
 
-\citeauthor{hughRaynbirdTaskWork} provides
+\topic{Piecework's history traces back further perhaps than most would expect.}
+\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
+to calculate equations in order
+to better--predict the trajectories of various celestial bodies in the 1830s~\cite{grier2013computers}.
+George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
+by breaking complex calculations into constituent parts, and
+training young men to solve simple algebraic problems,
+Airy could distribute work to many more people than could otherwise complete the full calculations.
+
+\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
+but it found its foothold in manual labor.}
+Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
+in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
+on railroads~\cite{Brown01041990}, and 
+elsewhere in manufacturing~\cite{10.2307/3827491}.
+Fordism and scientific management thrust piecework into higher gear, especially as
+mass manufacturing and
+a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
+\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
+\ali{I'm not sure; I'll look for some data.}
+
+% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
+\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
 a concise definition of piecework
-(which he also calls ``measure work'', ``grate work'', and ``task work'')
-by contrasting with the ``day--labourer'':
+in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
+(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
+particularly driven toward encapsulating the manual labor of farmwork.}
+\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'':
 ``the chief difference lies between the day--labourer,
@@
 \citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
-finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
+finding similar notions of independence and autonomy among piecework newspaper compositors:
 ``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night,
 the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
-We therefore draw a \textit{sense} of independence and autonomy inherent to piecework.
+This sense of independence and autonomy appears to be an inherent component of piecework.
 %that resonates across decades, industries, and locales where piecework is found.
+
+\ali{I think now that this is moved to this area there's a good opportunity to frame this more as
+``now that workers could choose their own schedule, style, etc\dots,
+newfound interest in how to manage these workers emerged''.
+Thoughts?}
+\topic{The early growth of piecework led to discussion on how best to manage pieceworkers,
+generally regarding workers antagonistically~\cite{norton1900textile,clark1908cotton}.}
+This was a far cry from the earlier rhetoric on piecework, which promised that
+piece workers would gladly work as diligently and as hard as possible because
+incentive--based pay would reward workers directly for hard work~[].
+
+This frustrations of poor working and living conditions
+(famously documented by
+\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
+Discontent reached a crescendo when industry organizations representing
+railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.
 
@@
 
+% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
+\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
+Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
+by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
+\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
+--- where more than 50\% worked under incentive wage plans ---
+were principally in clothes--making (e.g. hosiery, footwear, and garments).
+% \msb{How quickly did it plummet? By when was it gone? Where is it now?}
+% \msb{This sentence is empty calorie:}
+\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
+
+\ali{let's try cutting it\dots
+\msb{Very few calorie sentence:}
+In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
+The salient suggestions include:
+\begin{inlinelist}
+\item the emergence of more effective, more nuanced incentive models
+--- rewarding teams for complex achievements, for instance
+\item the shifting of these industries (e.g., manufacturing, clothing)
+to other countries
+\item the quality of ``multidimensional'' work becoming too difficult to evaluate~\cite{hart2013rise}.
+\end{inlinelist}}
+
+\ali{This feels like the Australia of paragraphs: one sentence constitutes the whole paragraph.
+Should I add some analysis here?}
 In summary, piecework:
@@
 % \onlyinsubfile{\clearpage}
-\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
-\msb{This section is so brief that it feels like a bit of a cheat.
+% \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
+% \msb{This section is so brief that it feels like a bit of a cheat.
 I think we need to frame it differently.
@@
 Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
-\ali{Do you mean that we'd have (sub)subsections under this?
+% \ali{Do you mean that we'd have (sub)subsections under this?
 I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead.
@@
 would I interleave the history or just append it, basically removing the header).}
+\ali{there's no need for this paragraph:
+
 \topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
@@
 through the political and economic turmoil of the early and mid--\nth{20} century,
-and into the \nth{21} century.
+and into the \nth{21} century.}
 % While the previous section provided a \textit{definition} of piecework,
@@
 
-\topic{Piecework's history traces back further perhaps than most would expect.}
-\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
-to calculate equations in order
-to better--predict the trajectories of various celestial bodies in the \nth{19} century~\cite{grier2013computers}.
-George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
-by breaking complex calculations into constituent parts, and
-training young men to solve simple algebraic problems,
-Airy could distribute work to many more people than could otherwise complete the full calculations.
-
-\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
-but it found its foothold in manual labor.}
-Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
-in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
-on railroads~\cite{Brown01041990}, and 
-elsewhere in manufacturing~\cite{10.2307/3827491}.
-Fordism and scientific management thrust piecework into higher gear, especially as
-mass manufacturing and
-a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
-\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
-\ali{I'm not sure; I'll look for some data.}
 % \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
@@
 % \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
-\topic{The early growth of piecework led to discussion surrounding how best to manage pieceworkers,
-generally in an adversarial approach, consequently
-frustrating workers~\cite{norton1900textile,clark1908cotton}.}
-This frustrations of poor working and living conditions
-(famously documented by
-\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
-Discontent reached a crescendo when industry organizations representing
-railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.
-
-\msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
-\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
-\msb{How quickly did it plummet? By when was it gone? Where is it now?}
-\msb{This sentence is empty calorie:}
-\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
-\msb{Very few calorie sentence:}
-In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
-The salient suggestions include:
-\begin{inlinelist}
-\item the emergence of more effective, more nuanced incentive models
---- rewarding teams for complex achievements, for instance
-\item the shifting of these industries (e.g., manufacturing, clothing)
-to other countries
-\item the quality of ``multidimensional'' work becoming too difficult to evaluate
-\end{inlinelist}~\cite{hart2013rise}.
 

### 2017-01-06 drafts/labor/piecework_lit.tex
lit review not finished but ready for more feedback; switching to case studies to hack stuff up

@@
 \section{A Review of Piecework}
-
-\ali{I'm interleaving the paragraphs, so this paragraph needs to be refactored a bit to reflect that instead of suggesting that the road map will look like two distinct threads.
-I'll probably try to take a shot at it in a bit, but for now don't take the wording too seriously.}
 
@@
 
-\ali{
-So I'm thinking that this should flow like this:
-``We'll try to understand what piecework was by tracing the development of and discourse around piecework.''
-This'll give us the ability to switch back and forth between the timeline of the practice of piecework
---- that is, the ``piecework primer'' stuff down below ---
-and the contemporary/concurrent research going on around the same time
-(i.e. this stuff immediately following).
-
-I'm going to interleave the paragraphs, but
-if you think that's a flawed approach, then I'll bet you won't like what you'll see below that much.%
-}
-
 \topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
@@
 alongside piecework's contemporary developments in practice.
-\ali{This is worded a little awkwardly;
-I want to say that we're looking at
-the discourse about piecework at the same time that we look at
-how piecework is unfolding. (wait, is that how I should word it?)}
 
@@
 \msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
-\ali{I'm not sure; I'll look for some data.}
+\ali{I'm not sure; I'll look for some data.
+
+update: hard to find anything saying ``piecework represented eleventy billion dollars of the U.S. economy'', which is what I'd like to say,
+but I can say something like ``at its height, $k\%$ of garment workers were pieceworkers'' or something.
+that seems a bit wishy--washy though; I'd like to say something like
+``$k\%$ of the \textit{entire U.S. workforce} was working under some sort of piecework regime''.}
 
@@
 particularly driven toward encapsulating the manual labor of farmwork.}
-\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'':
+\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
 ``the chief difference lies between the day--labourer,
@@
 
-This frustrations of poor working and living conditions
-(famously documented by
-\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
-Discontent reached a crescendo when industry organizations representing
-railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.
-
 \topic{Piecework opened the door for people who previously couldn't participate in the labor market to do so, and to acquire job skills incrementally.}
@@
 were principally in clothes--making (e.g. hosiery, footwear, and garments).
-% \msb{How quickly did it plummet? By when was it gone? Where is it now?}
-% \msb{This sentence is empty calorie:}
 \citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
 
-\ali{let's try cutting it\dots
-\msb{Very few calorie sentence:}
-In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
-The salient suggestions include:
-\begin{inlinelist}
-\item the emergence of more effective, more nuanced incentive models
---- rewarding teams for complex achievements, for instance
-\item the shifting of these industries (e.g., manufacturing, clothing)
-to other countries
-\item the quality of ``multidimensional'' work becoming too difficult to evaluate~\cite{hart2013rise}.
-\end{inlinelist}}
+\ali{``, ultimately arguing that\dots''?}
 
-\ali{This feels like the Australia of paragraphs: one sentence constitutes the whole paragraph.
-Should I add some analysis here?}
 In summary, piecework:
@@
 Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
-\msb{This transition seems misplaced with the Piecework primer and Case Studies subsections up next:}
-In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.
 
-% \onlyinsubfile{\clearpage}
-% \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
-% \msb{This section is so brief that it feels like a bit of a cheat.
-I think we need to frame it differently.
-Maybe that it should survey the different ``eras'' of piecework?
-And each paragraph is laying out a major historical era that we will return to later?
-Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
-% \ali{Do you mean that we'd have (sub)subsections under this?
-I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead.
-Not exactly sure how that would work (i.e.
-would I interleave the history or just append it, basically removing the header).}
-\ali{there's no need for this paragraph:
-
-\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
-% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
-This section will frame piecework,
-in the early days of the Industrial Revolution \&
-through the political and economic turmoil of the early and mid--\nth{20} century,
-and into the \nth{21} century.}
-% While the previous section provided a \textit{definition} of piecework,
-% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
-% as close to a \textit{thick description} of piecework
-% as we can afford~\cite{geertz1973interpretation}.
-% \msb{is this really thick? as a reader it flies by really quickly and super high level...}
-
-% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
-% \citeauthor{hart2013rise} point out that the Second World War,
-% which called millions of Americans to military service,
-% necessitated the rapid training and employment of
-% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
-
-% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
-
-% \onlyinsubfile{\clearpage}
 \subsection{Case studies in piecework}
-Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
-       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
-first, railroad and other industrial workers;
-second, Airy's employment of \textit{human computers};
-and third, domestic and farm work (in particular, the ``match--girls'').
-Here, we introduce the basic historical details of these three cases.
-We begin with the most familiar case (industrial workers),
-working backwards through the \nth{20} and then \nth{19} century,
-tracing the origins of the ideas that \dots
-% \msb{incomplete sentence}
+\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
+railroad and other industrial workers;
+Airy's employment of \textit{human computers};
+and domestic and farm work (in particular, the ``match--girls'').}
+We'll introduce the facts of these three cases,
+beginning with the most familiar case of piecework
+--- industrial work, such as assembly line manufacturing ---
+working backwards through the \nth{20} and \nth{19} century.
+In doing so, we'll trace the history of piecework
+while also framing the later analysis of the major research threads we named earlier
+(questions toward complexity, decomposition, and relationships, respectively).
 
@@
 
-% \ali{maybe go backwards on this}
-
-% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
-% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
-% \onlyinsubfile{\clearpage}
 \subsubsection{Railroad and other Industrial Workers}
-{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
+\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
 It was here that
@@
 \msb{Why is Taylor relevant to this case study?}
+\ali{I'm just clumsily interleaving the contemporary literature with the events.
+I'd like to say ``This is the geist of work in their time'';
+is this a bad idea?}
 It was in the midst of this time that \citeauthor{taylor1914principles} published
@@
 and paid for each piece of work completed.
-% \ali{also talk about the challenge/competition Airy was in with other astronomers,
-% and the approach of verifying worker results}
-% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
-% For one thing,
 Airy also instituted a policy of firing his computers once they reached age 23
@@
 should i just leave it out until later when there's something to say about it
-(namely, that the mode of work was initially designed to stymie professional growth)?}.
+(namely, that the mode of work was initially designed to stymie professional growth)?}%
+.
 
-\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
+\ali{This should go later someplace, but not clear where:
+
+This practice ensured two outcomes that arguably disfavored workers.
 First, it eliminated any potential to advance professionally, as
@@
 Second, it limited workers' ability to organize
-by ensuring that workers hardly spent sufficient time to successfully rally their peers.
-}
+by ensuring that workers hardly spent sufficient time to successfully rally their peers.}
 

### 2017-01-06 drafts/labor/complexity.tex
notes in the case studies
i just wrote comments indicating the assumed and introduced facts in each of the questions (complexity, decomp, and relationships).
i think a bit of a word cloud is forming in my head, so this should give me something to structure the case studies in the lit review

@@
 
-\onlyinsubfile{\newpage}
 \subsubsection{\crowdworkpers}
@@
 medical data interpreters~\cite{lasecki2013chorus}, and
-idea generation workflows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
 to name a few examples.
@@
 
-It is now clear that this computational workflow approach works with focused complex tasks, but
-the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.
+\topic{It is now clear that this computational workflow approach works with focused complex tasks, but
+the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.}
 As a first example,
@@
 outlining~\cite{luther2015crowdlines}
-are possible, but there is no general solution for sensemaking.
+are possible, but there is no general solution for sense--making.
 It is not yet clear what insights would be required
@@
 
-Restricting attention to non--expert, microtask workers proved limiting.
+\topic{Restricting attention to non--expert, microtask workers proved limiting.}
 So, \citeauthor{foundry} introduced the idea of crowdsourcing with
@@
 
-\onlyinsubfile{\newpage}
+\clearpage
 \subsubsection{\pieceworkpers}
@@
 
-\ali{I'm not sure that this really stays close to the three case studies that emerged as major threads (now in the piecework lit review).
-Should I refactor/rewrite this to make it more relevant?}
-Measurement also limited the complexity of piecework:
-only tasks that could be measured and priced could be completed via piecework. 
+\begin{comment}
+notes: what info do i assume the reader has seen already?
+- Graves: railway companies used ``efficiency experts'' to study how long tasks should take
+- Hart: evaluation limits complexity (we can affect that with peer evaluation!)
+- Graves: sparks of Scientific Management in Piecework
+- organization types are important determinants of piecework viability: lots of types of tasks? bad
+  - Hart (I think?): variability in workers is fine though
+- Foreman is important
+- 19th century: piecework was mostly cottage industry with untrained or informally trained workers
+  (unlike industrial metal workers during WWII)
+\end{comment}
+
+{\topic{Only tasks that could be measured and priced could be completed via piecework
+--- a fact which fundamentally limited the potential complexity of piecework.}
 When \citeauthor{Brown01041990} investigated
@@
 --- farm workers didn't need to be trained on how to plow fields, or birth foals;
-seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.
+seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.}
 
-\onlyinsubfile{\newpage}
+\clearpage
 \subsubsection{\whatchanged}
@@
 
-The research on piecework tells us that
+\topic{The research on piecework tells us that
 we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity~\cite{Brown01041990}.
+the nature of the work is limited in complexity~\cite{Brown01041990}.}
 Given the flourishing of on--demand labor platforms such as
@@
 This test is so challenging that veteran drivers develop significantly larger
-the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,
-      skok2000managing,Woollett1407,woollett2011acquiring}.
+the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,skok2000managing,Woollett1407,woollett2011acquiring}.
 In contrast, with on--demand platforms such as Uber, services such as Google Maps \& Waze make it possible for
 people entirely unfamiliar with a city
-to operate profesionally~\cite{silva2013traffic,hind2014outsmarting}.
-Other examples include search engines enabling information retrieval, and word processors enabling spelling and grammar checking.
-By augmentating the human intellect~\cite{engelbart2001augmenting}, computing has shifted the complexity of work that is possible without training.
+to operate professionally~\cite{silva2013traffic,hind2014outsmarting}.
+Other examples include search engines enabling information retrieval, and
+word processors enabling spelling and grammar checking.
+By augmenting the human intellect~\cite{engelbart2001augmenting},
+computing has shifted the complexity of work that is possible without training.
 
-Algorithms have automated some tasks that previously fell to management.
+\topic{Algorithms have automated some tasks that previously fell to management.}
 Computational systems hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
@@
 
-Finally, the organizational limit on infrastructure creation is somewhat lessened. Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
+\topic{Finally, the organizational limit on infrastructure creation is somewhat lessened.}
+Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
 \citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
@@
 
+\clearpage
 \subsubsection{\implication}
@@
 This shift may enable complex goals that were not cost--effective before to become feasible.
-However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re--introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
+However, because algorithms remain far from replicating all of the foreman's responsibilities,
+most likely is a middle ground in which crowd work re--introduces
+the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
 This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters~\cite{MaliciousCrowdworkersGadiraju}.

### 2017-01-06 drafts/labor/decomposition.tex
notes in the case studies
i just wrote comments indicating the assumed and introduced facts in each of the questions (complexity, decomp, and relationships).
i think a bit of a word cloud is forming in my head, so this should give me something to structure the case studies in the lit review

@@
 In this section,
-we place these questions in the context of piecework's Tayloristic evolution.
+we place these questions in the context of piecework's Taylorist evolution.
 
@@
 
-\onlyinsubfile{\newpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{\crowdworkpers}
@@
 
-\onlyinsubfile{\newpage}
+\onlyinsubfile{\clearpage}
 \subsubsection{\pieceworkpers}
+
+\begin{comment}
+notes: what info do i assume the reader has seen already?
+- Brown: Task variability matters
+- Airy and his human computers were great:
+  - quickly verifiable
+  - independent tasks (could be checked without the whole product)
+  - narrowly trainable
+\end{comment}
+
 \citeauthor{Brown01041990} inquired from another direction,
@@
 
-\ali{This can be trimmed to just reference back to the piecework review}
 \topic{The research community relating to
@@
 ``[piecework is] infeasible,
-because \dots
-total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
+because~\dots~total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
 
-\onlyinsubfile{\newpage}
+\onlyinsubfile{\clearpage}
 \subsubsection{\whatchanged}
@@
 
-\onlyinsubfile{\newpage}
+\onlyinsubfile{\clearpage}
 \subsubsection{\implication}

### 2017-01-06 drafts/labor/relationships.tex
notes in the case studies
i just wrote comments indicating the assumed and introduced facts in each of the questions (complexity, decomp, and relationships).
i think a bit of a word cloud is forming in my head, so this should give me something to structure the case studies in the lit review

@@
 
-\subsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}
-\label{sec:relationships}
+\subsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}\label{sec:relationships}
 
-HCI and CSCW have largely framed themselves around supporting work rather than becoming an infrastructural layer enabling it.
+HCI and CSCW have framed themselves around supporting work rather than becoming an infrastructural layer enabling it.
 While all artifacts have politics, this shift into computational labor systems has directly impacted the lives and livelihood of workers.
@@
 
-\onlyinsubfile{\newpage}
+\onlyinsubfile{\clearpage}
 \subsubsection{\pieceworkpers}
+
+\begin{comment}
+notes: what info do i assume the reader has seen already?
+- Clark: pieceworkers work harder, more diligently, etc...
+- Riis saw terrible conditions, documented and communicated it to the world
+- Worker advocacy groups arose to speak out against piecework
+\end{comment}
+
 Early observers believed that workers were strongly motivated by the piecework model.
@@
 For one thing, the research methods we typically associate with the exploratory study of cultures
---- Anthropology, and namely participant--observation, ethnography, etc\dots ---
+--- Anthropology, and namely participant--observation, ethnography, etc\dots~---
 didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
@@
 
-\onlyinsubfile{\newpage}
+\onlyinsubfile{\clearpage}
 \subsubsection{\whatchanged}
@@
 
-\onlyinsubfile{
-\balance{}
-\printbibliography
-}
+% \onlyinsubfile{
+% \balance{}
+% \printbibliography
+% }
 \end{document}

### 2017-01-06 drafts/labor/piecework_lit.tex
slight comments on piecework review

@@
 (where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
-particularly driven toward encapsulating the manual labor of farmwork.}
+particularly driven toward encapsulating the manual labor of farmwork.} \msb{This is the first time the reader has seen ``Raynbird'', and they don't know which essay you're talking about.
+They might be thinking, ``What is this essay?
+Is it famous or something?
+Am I supposed to know about it?''}
 \citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
@@
 ``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
-This framing offers a more intuitive sense;
+This framing offers a more intuitive sense \msb{is this a comparative? more intuitive than what?};
 ``payment for results,'' as he calls it,
@@
 \topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
-Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
+Between 1938 and 1942, the proportion of metal workers who were pieceworkers had climbed steeply 11\% to 60\%~\cite{hart2005piecework};
 by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
-\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
+\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions of piecework
 --- where more than 50\% worked under incentive wage plans ---
@@
 
+\msb{If you want to make the argument in the last paragraph of this subsection that piecework assumed no professional skills, it's worth calling that out slightly more visibly throughout the subsection.}
 Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.

### 2017-01-06 drafts/labor/complexity.tex
lit review except for airy's computers done

@@
 
-\clearpage
+% \clearpage
 \subsubsection{\pieceworkpers}
@@
 --- a fact which fundamentally limited the potential complexity of piecework.}
-When \citeauthor{Brown01041990} investigated
-what limited the adoption of piecework in industries that otherwise gravitated toward it
-(e.g., railway engineers),
-the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
-\citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
-which used ``efficiency experts'' to develop a ``standard time''
-to determine pay for each task at the company informed by
-``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
-~\cite{10.2307/23702539}.
-One might conclude from \citeauthor{10.2307/23702539}'s observations that
+% When \citeauthor{Brown01041990} investigated
+% what limited the adoption of piecework in industries that otherwise gravitated toward it
+% (e.g., railway engineers),
+% the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
+% \citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
+% which used ``efficiency experts'' to develop a ``standard time''
+% to determine pay for each task at the company informed by
+% ``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
+% ~\cite{10.2307/23702539}.
+Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
+one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
 complex, creative work
 --- which is inherently heterogeneous and difficult to routinize ---
-would be unsuitable for piecework.
+would be unsuitable for piecework~\cite{10.2307/23702539}.
+\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
+``efficiency experts'' can effectively gauge how long known tasks should take, but
+would find themselves overwhelmed if they attempted to assess creative work like design,
+which can take an arbitrary number of iterations before proceeding to a subsequent step
+\ali{should i cite some design research here?}.
 
@@
 
-\clearpage
+% \clearpage
 \subsubsection{\whatchanged}
@@
 
-\clearpage
+% \clearpage
 \subsubsection{\implication}

### 2017-01-06 drafts/labor/piecework_lit.tex
lit review except for airy's computers done

--- 2017-01-06 drafts/labor/piecework_lit.tex
@@
 definitions,
-tracing an outline of % a working understanding of
+tracing an outline of
 \textit{what piecework is}
@@
 
-% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
-\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
+\topic{By \citeyear{hughRaynbirdTaskWork} we find
 a concise definition of piecework
-in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
+in \citeauthor{hughRaynbirdTaskWork}'s essay
 (where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
@@
 this was just one of several alternatives to categorical time-- and piece--rate renumeration paradigms.
-% which will complicate our categorization of cases of piecework.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
@@
 \topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
-\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework, % do you like what i did there?
+\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework,
 pointing out that
@@
 
-% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
 \topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
@@
 Airy's employment of \textit{human computers};
-and domestic and farm work (in particular, the ``match--girls'').}
-We'll introduce the facts of these three cases,
-beginning with the most familiar case of piecework
---- industrial work, such as assembly line manufacturing ---
-working backwards through the \nth{20} and \nth{19} century.
-In doing so, we'll trace the history of piecework
+and domestic and farm work (specifically, the ``match--girls'').}
+In introducing these cases at a high level,
+we'll trace the history of piecework
 while also framing the later analysis of the major research threads we named earlier
@@
 
-\ali{I'm coming out of this with a sort of structure that \textit{loosely} describes piecework chronologically.
-Strictly speaking, I'm talking about each of these topics
-(let's call them
-\textbf{human computation},
-\textbf{domestic and farm work}, and
-\textbf{industrial work}),
-and some of these things overlap with each other to some extent, but
-for the most part
-Airy and the human computers came in the mid--\nth{19} century,
-farm work and the ``domestic'' work (like making matches) came around at the turn of the \nth{20} century, and
-the industrial work (railroad workers, the WWII war effort, and the rise of the labor union movement) unfolded from the early \nth{20} century onward.}
+\begin{comment}
+What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)
 
-\subsubsection{Railroad and other Industrial Workers}
+- Graves: railway companies used ``efficiency experts'' to study how long tasks should take
+- Hart: evaluation limits complexity (we can affect that with peer evaluation!)
+- Graves: sparks of Scientific Management in Piecework
+- organization types are important determinants of piecework viability: lots of types of tasks? bad
+  - Hart (I think?): variability in *worker* quality is fine
+- Foreman is important
+- Worker advocacy groups arose to speak out against piecework
+
+\end{comment}
+
+\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
 \topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
-It was here that
-scientific management, Fordism, and Taylorism
-dramatically influenced how workers were managed and
-the ways in which workers were perceived and envisioned paradigmatically;
-it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}
+In the railroad industry and especially on the assembly line
+we see many of the mechanisms of work and worker management that later made their way into crowd work.
+Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
+We'll expose these facets of piecework very superficially in the context of industrial work
+to better speak to these research questions in crowd work later.
 
-\msb{I'm feeling very disoriented in this paragraph. The subsection title is about railroads, but we're all fast-forwarded to auto manufacturing?}
+\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
+what followed was a flourishing in the management practices with pieceworkers,
+as railway companies struggled to find effective ways
+to motivate and evaluate this skilled workforce of engineers.}
+\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
+finding that they employed ``efficiency experts'' to develop a ``standard time''
+to determine pay for each task at the company informed by
+``thousands of individual operations''; % ~\cite{10.2307/23702539}.
+\citeauthor{10.2307/23702539} goes on to list
+some of the roles required to facilitate piecework
+in the early \nth{20} century
+--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
+
+\topic{This oversight, while controversial
+(especially among workers~\cite{american1921problem}),
+paved the way for piecework to grow substantially.}
+With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's eponymously named \textit{Fordism}),
+piecework in the early and mid--\nth{20} century surged, especially in industrial work.
+Scientific management promised that the careful measurement of workers would yield
+higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
+While \citeauthor{Brown01041990} points out that
+piecework dramatically advanced the instrumented measurement of workers,
+in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
+using completion times to inform per--task compensation,
+similarly to efficiency experts in the Santa Fe Railway, but
+substantially more precise.
+
 \topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
@@
 
-% This approach would turn out to yield two benefits:
-% \item these workers could be productive with significantly less ``on--boarding'' cost
-% \item workers' performance could be tracked \& evaluated, and better methods could be disseminated to workers.
+% \msb{confusing timing; now we're going backwards?}
 
-\msb{confusing timing; now we're going backwards?}
-{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism}),
-piecework in the early and mid--\nth{20} century surged, especially in industrial work.}
-Scientific management promised that the careful measurement of workers would yield
-higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
-While \citeauthor{Brown01041990} points out that
-piecework dramatically advanced the instrumented measurement of workers,
-in \citeauthor{taylor1914principles}'s time highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
-using completion times to inform per--task compensation.
-% Under this regime, piecework became commonplace in railroad maintenance and a number 
-% This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
-% railroad maintenance, automobile manufacturing, and later weapons manufacturing.
-% \ali{???? better word for ``weapons manufacturing''?}
-% Instead, as a result,
-% the distillation of work into smaller units ultimately
-% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
-
-% \topic{With the formalization of scientific management,
-% piecework became an important component of the mass manufacturing
-% in American factories.}
-% \topic{}
-
-\msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
-\msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
+% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
+% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
 \topic{The 1930s represented a boom for piecework on an unprecedented scale,
@@
 
-\msb{This subsection feels like it never delivered on what it promised: a case study of one group of pieceworkers.
-I thought the goal was to introduce the railroad workers --- who they were, what they did, what their conditions were like.}
+\ali{some things I didn't cover:
+\begin{itemize}
+  \item research finding that task variability is bad for piecework, but worker variability is good
+  \item talk about the foreman
+  \item worker advocacy groups
+\end{itemize}
+should I be making a serious effort to get that stuff in here? this section is kind of long already
+(and it'll definitely be much longer than the other case studies).}
 
@@
 \subsubsection{Domestic and Farmhand Labor}
-\topic{When piecework entered the American economy, it was not used for complex work.}
-One reason for low complexity was workers' skills:
-it was infeasible to provide new pieceworkers with the comprehensive education
-that apprenticeships imparted~\cite{hart2013rise}.
-So, initially piecework in the United States arose for farm work, and as
-\citeauthor{hughRaynbirdTaskWork} and others discuss,
-the practice remained relatively obscure until
-it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
-\msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
-The complexity of the work remained low at the turn of
-the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
-\msb{how is this sentence relevant to the case study?:}
-However, writers of the time focused their attention on
-wage~\cite{burton1899commercial} and
-management regimes~\cite{norton1900textile}
-rather than training.
-Mass manufacturing,
-such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
-flourished under piecework systems in densely populated cities.
 
-\topic{Workers' relationships with employers quickly soured.}
-\msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
-The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
+\begin{comment}
+What did I pull from the threads that are related to domestic/farmwork?
+
+- Graves: sparks of Scientific Management in Piecework, especially starting here
+- 19th century: piecework was mostly cottage industry with untrained or informally trained workers
+  (unlike industrial metal workers during WWII)
+- Brown: Task variability matters
+- Clark: pieceworkers work harder, more diligently, etc...
+- Riis saw terrible conditions, documented and communicated it to the world
+
+\end{comment}
+
+\topic{The application of piecework to farm work in the late \nth{19} century and
+later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
+proved to be a formative period for piecework as we would come to know it.}
+Here, then, we'll discuss the implementation of piecework regimes in farms,
+in homes where workers assembled clothing and were paid by the \textit{piece} of output, and
+finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
+
+\topic{Farm and field work emerged early on as a rich field for the use of piecework,
+perhaps in part due to the relatively low skillset necessary to complete work.}
+We discussed earlier that farm work applied the idea of piecework to the farm by
+paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
+\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
+the only factor important in piecework was
+whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
+Piecework's advantage,
+\citeauthor{hughRaynbirdTaskWork} argues,
+was this hands--off management of workers.
+
+\topic{Piecework was beginning to find its way into homes near the end of the \nth{19} century.}
+Textile manufacturers found that they could deliver fabric to people at their homes,
+paying them to sew together clothing.
+The manufacturers would later return to retrieve the finished garments,
+paying these workers for each \textit{piece} of clothing made.
+They could, in principle, assemble as much or as little clothing as they wanted;
+the reality was more grim, as
+\citeauthor{riisOtherSideLives} documented in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
+He found that
+workers endured bleak living conditions and
+worked long hours attempting to scrape together a living.
+
+\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
+as well as the early years of worker collective organization and action.}
+At the end of the \nth{19} century,
+manufacturers employed women to assemble matchsticks in factories.
+These match--makers were some of the first workers in mass manufacturing
+to successfully rally for political causes,
+first in the form of a march on parliament in 1871 to protest a proposed tax, and 
+later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.
+
+\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
+much deeper resentment had been boiling over for years.}
+Match--girls were already frustrated with
+the arbitrariness of management,
+poor working conditions, and
+having to work with phosphoric materials (which was proving hazardous).
+\ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}
+
+\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
+This was one of the earliest and most famous successful worker strikes,
 and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
@@
 
-\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
-\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
-Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
-Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
-\msb{Why is Taylor relevant to this case study?}
-\ali{I'm just clumsily interleaving the contemporary literature with the events.
-I'd like to say ``This is the geist of work in their time'';
-is this a bad idea?}
-It was in the midst of this time that \citeauthor{taylor1914principles} published
-the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
-It was this framing on work \& worker management that gave workers a concrete adversary
---- if not Taylor, then \textit{Taylorism}  ---
-against which to rally~\cite{jacoby1983union,parker1920casual}.
+\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history,
+we should shine a light onto the reputation that had formed for these workers.}
+Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
+according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
+\citeauthor{10.2307/3827491} noted that match girls
+``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
+But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
+These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
+Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
+making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.
+
+\ali{I want to discuss a few things here:
+\begin{itemize}
+  \item there were protests about something sort of like surge pricing in crop--picking, but i don't know how to make that a paragraph (maybe merge it with match--girls?)
+\end{itemize}}
+
+% \topic{When piecework entered the American economy, it was not used for complex work.}
+% One reason for low complexity was workers' skills:
+% it was infeasible to provide new pieceworkers with the comprehensive education
+% that apprenticeships imparted~\cite{hart2013rise}.
+% So, initially piecework in the United States arose for farm work, and as
+% \citeauthor{hughRaynbirdTaskWork} and others discuss,
+% the practice remained relatively obscure until
+% it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+% \msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
+% The complexity of the work remained low at the turn of
+% the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+% \msb{how is this sentence relevant to the case study?:}
+% However, writers of the time focused their attention on
+% wage~\cite{burton1899commercial} and
+% management regimes~\cite{norton1900textile}
+% rather than training.
+% Mass manufacturing,
+% such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
+% flourished under piecework systems in densely populated cities.
+
+% \topic{Workers' relationships with employers quickly soured.}
+% \msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
+% The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
+% and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+% As \citeauthor{weyer1894history} described,
+% ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
+% the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+
+% \topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
+% \msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
+% Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
+% Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
+% \msb{Why is Taylor relevant to this case study?}
+% \ali{I'm just clumsily interleaving the contemporary literature with the events.
+% I'd like to say ``This is the geist of work in their time'';
+% is this a bad idea?}
+% It was in the midst of this time that \citeauthor{taylor1914principles} published
+% the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
+% It was this framing on work \& worker management that gave workers a concrete adversary
+% --- if not Taylor, then \textit{Taylorism}  ---
+% against which to rally~\cite{jacoby1983union,parker1920casual}.
 
@@
 \subsubsection{Airy's Computers}
+
+\begin{comment}
+What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)
+
+- Airy and his human computers were great:
+  - quickly verifiable
+  - independent tasks (could be checked without the whole product)
+  - narrowly trainable
+
+\end{comment}
+
 \msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}

### 2017-01-06 drafts/labor/piecework_lit.tex

@@
 a concise definition of piecework
-in \citeauthor{hughRaynbirdTaskWork}'s essay
-(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
+in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework
+(where he also calls it ``measure work'', ``grate work'', and ``task work''),
 particularly driven toward encapsulating the manual labor of farmwork.}
+\ali{I just made the line ``Raynbird's essay on piecework''; is that enough to make it sound less important?}
 \citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
@@
 ``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
-This framing offers a more intuitive sense;
+This framing offers an intuitive sense of piecework;
 ``payment for results,'' as he calls it,
@@
 \end{inlinelist}
+
+\msb{If you want to make the argument in the last paragraph of this subsection that piecework assumed no professional skills, it's worth calling that out slightly more visibly throughout the subsection.}\ali{I don't want to suggest that, actually, at least not ultimately; in the beginning, sure, but by WWII it definitely required certain specialized training (arguably in Airy's case too)}
 
@@
 
-% \topic{When piecework entered the American economy, it was not used for complex work.}
-% One reason for low complexity was workers' skills:
-% it was infeasible to provide new pieceworkers with the comprehensive education
-% that apprenticeships imparted~\cite{hart2013rise}.
-% So, initially piecework in the United States arose for farm work, and as
-% \citeauthor{hughRaynbirdTaskWork} and others discuss,
-% the practice remained relatively obscure until
-% it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
-% \msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
-% The complexity of the work remained low at the turn of
-% the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
-% \msb{how is this sentence relevant to the case study?:}
-% However, writers of the time focused their attention on
-% wage~\cite{burton1899commercial} and
-% management regimes~\cite{norton1900textile}
-% rather than training.
-% Mass manufacturing,
-% such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
-% flourished under piecework systems in densely populated cities.
-
-% \topic{Workers' relationships with employers quickly soured.}
-% \msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
-% The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
-% and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
-% As \citeauthor{weyer1894history} described,
-% ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
-% the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
-
-% \topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
-% \msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
-% Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
-% Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
-% \msb{Why is Taylor relevant to this case study?}
-% \ali{I'm just clumsily interleaving the contemporary literature with the events.
-% I'd like to say ``This is the geist of work in their time'';
-% is this a bad idea?}
-% It was in the midst of this time that \citeauthor{taylor1914principles} published
-% the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
-% It was this framing on work \& worker management that gave workers a concrete adversary
-% --- if not Taylor, then \textit{Taylorism}  ---
-% against which to rally~\cite{jacoby1983union,parker1920casual}.
-
 % \onlyinsubfile{\clearpage}

### 2017-01-07 drafts/labor/piecework_lit.tex
piecework review refactor

@@
 \subsection{Case studies in piecework}
-\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
-railroad and other industrial workers;
-Airy's employment of \textit{human computers};
-and domestic and farm work (specifically, the ``match--girls'').}
+\topic{Throughout the rest of the paper, we will return to four case studies to frame our analyses:
+Airy's employment of human computers;
+domestic and farm work;
+the ``match--girls'' strike;
+and industrial and assembly-line work.}
 In introducing these cases at a high level,
 we'll trace the history of piecework
-while also framing the later analysis of the major research threads we named earlier
-(questions toward complexity, decomposition, and relationships, respectively).
-
-\begin{comment}
-What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)
-
-- Graves: railway companies used ``efficiency experts'' to study how long tasks should take
-- Hart: evaluation limits complexity (we can affect that with peer evaluation!)
-- Graves: sparks of Scientific Management in Piecework
-- organization types are important determinants of piecework viability: lots of types of tasks? bad
-  - Hart (I think?): variability in *worker* quality is fine
-- Foreman is important
-- Worker advocacy groups arose to speak out against piecework
-
-\end{comment}
-
-\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
-\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
-In the railroad industry and especially on the assembly line
-we see many of the mechanisms of work and worker management that later made their way into crowd work.
-Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
-We'll expose these facets of piecework very superficially in the context of industrial work
-to better speak to these research questions in crowd work later.
-
-\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
-what followed was a flourishing in the management practices with pieceworkers,
-as railway companies struggled to find effective ways
-to motivate and evaluate this skilled workforce of engineers.}
-\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
-finding that they employed ``efficiency experts'' to develop a ``standard time''
-to determine pay for each task at the company informed by
-``thousands of individual operations''; % ~\cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} goes on to list
-some of the roles required to facilitate piecework
-in the early \nth{20} century
---- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
-
-\topic{This oversight, while controversial
-(especially among workers~\cite{american1921problem}),
-paved the way for piecework to grow substantially.}
-With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism}),
-piecework in the early and mid--\nth{20} century surged, especially in industrial work.
-Scientific management promised that the careful measurement of workers would yield
-higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
-While \citeauthor{Brown01041990} points out that
-piecework dramatically advanced the instrumented measurement of workers,
-in \citeauthor{taylor1914principles}'s time highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
-using completion times to inform per--task compensation,
-similarly to efficiency experts in the Santa Fe Railway, but
-substantially more precise.
-
-\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
-found its way into the war effort during World War II.}
-With the vast majority of men drafted into service,
-factories found themselves turning to
-a predominantly female workforce that had neither
-the formal training nor
-the years of apprenticeship experience
-that conventional workers would have had.
-Rather than attempting to train this new labor force in every aspect of industrial work,
-these women were trained for individual tasks
-% --- such as riveting component to another ---
-and assigned to that task. %, decomposed from a more comprehensive task.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women~\cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
-
-% \msb{confusing timing; now we're going backwards?}
-
-% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
-% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
-\topic{The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.}
-\citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-They attribute this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
-
-\ali{some things I didn't cover:
-\begin{itemize}
-  \item research finding that task variability is bad for piecework, but worker variability is good
-  \item talk about the foreman
-  \item worker advocacy groups
-\end{itemize}
-should I be making a serious effort to get that stuff in here? this section is kind of long already
-(and it'll definitely be much longer than the other case studies).}
-
-% \onlyinsubfile{\clearpage}
-\subsubsection{Domestic and Farmhand Labor}
-
-\begin{comment}
-What did I pull from the threads that are related to domestic/farmwork?
-
-- Graves: sparks of Scientific Management in Piecework, especially starting here
-- 19th century: piecework was mostly cottage industry with untrained or informally trained workers
-  (unlike industrial metal workers during WWII)
-- Brown: Task variability matters
-- Clark: pieceworkers work harder, more diligently, etc...
-- Riis saw terrible conditions, documented and communicated it to the world
-
-\end{comment}
-
-\topic{The application of piecework to farm work in the late \nth{19} century and
-later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
-proved to be a formative period for piecework as we would come to know it.}
-Here, then, we'll discuss the implementation of piecework regimes in farms,
-in homes where workers assembled clothing and were paid by the \textit{piece} of output, and
-finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
-
-\topic{Farm and field work emerged early on as a rich field for the use of piecework,
-perhaps in part due to the relatively low skillset necessary to complete work.}
-We discussed earlier that farm work applied the idea of piecework to the farm by
-paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
-\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
-the only factor important in piecework was
-whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
-Piecework's advantage,
-\citeauthor{hughRaynbirdTaskWork} argues,
-was this hands--off management of workers.
-
-\topic{Piecework was beginning to find its way into homes near the end of the \nth{19} century.}
-Textile manufacturers found that they could deliver fabric to people at their homes,
-paying them to sew together clothing.
-The manufacturers would later return to retrieve the finished garments,
-paying these workers for each \textit{piece} of clothing made.
-They could, in principle, assemble as much or as little clothing as they wanted;
-the reality was more grim, as
-\citeauthor{riisOtherSideLives} documented in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
-He found that
-workers endured bleak living conditions and
-worked long hours attempting to scrape together a living.
-
-\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
-as well as the early years of worker collective organization and action.}
-At the end of the \nth{19} century,
-manufacturers employed women to assemble matchsticks in factories.
-These match--makers were some of the first workers in mass manufacturing
-to successfully rally for political causes,
-first in the form of a march on parliament in 1871 to protest a proposed tax, and 
-later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.
-
-\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
-much deeper resentment had been boiling over for years.}
-Match--girls were already frustrated with
-the arbitrariness of management,
-poor working conditions, and
-having to work with phosphoric materials (which was proving hazardous).
-\ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}
-
-\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
-This was one of the earliest and most famous successful worker strikes,
-and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
-As \citeauthor{weyer1894history} described,
-``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
-the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
-
-\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history,
-we should shine a light onto the reputation that had formed for these workers.}
-Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
-according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
-\citeauthor{10.2307/3827491} noted that match girls
-``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
-But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
-These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
-Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
-making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.
-
-\ali{I want to discuss a few things here:
-\begin{itemize}
-  \item there were protests about something sort of like surge pricing in crop--picking, but i don't know how to make that a paragraph (maybe merge it with match--girls?)
-\end{itemize}}
+while also framing the later analysis of the major research threads we named earlier: complexity, decomposition, and relationships.
 
@@
 and paid for each piece of work completed.
-Airy also instituted a policy of firing his computers once they reached age 23
-\ali{this has a ``gosh isn't that interesting?'' feeling to it.
-should i just leave it out until later when there's something to say about it
-(namely, that the mode of work was initially designed to stymie professional growth)?}%
-.
+Airy also instituted a policy of firing his computers once they reached age 23.
+%\ali{this has a ``gosh isn't that interesting?'' feeling to it.
+%should i just leave it out until later when there's something to say about it
+%(namely, that the mode of work was initially designed to stymie professional growth)?}%
 
-\ali{This should go later someplace, but not clear where:
+The human computers captured several aspects of task decomposition that would become common.
+First, the work was designed such that it could be done independently and without collaboration.
+This enabled geographically-distributed workers to complete the tasks.
+Second, the work was designed so that intermediate results could be quickly verifiable as correct: Airy would have two workers each do the calculation, and a another person compare their answers.
+Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.
+%\ali{This should go later someplace, but not clear where:
 
@@
 Second, it limited workers' ability to organize
-by ensuring that workers hardly spent sufficient time to successfully rally their peers.}
+by ensuring that workers hardly spent sufficient time to successfully rally their peers.
+
+% \onlyinsubfile{\clearpage}
+\subsubsection{Domestic and Farmhand Labor}
+
+\begin{comment}
+What did I pull from the threads that are related to domestic/farmwork?
+
+- Graves: sparks of Scientific Management in Piecework, especially starting here
+- 19th century: piecework was mostly cottage industry with untrained or informally trained workers
+  (unlike industrial metal workers during WWII)
+- Brown: Task variability matters
+- Clark: pieceworkers work harder, more diligently, etc...
+- Riis saw terrible conditions, documented and communicated it to the world
+
+\end{comment}
+
+\topic{The application of piecework to farm work in the late \nth{19} century and
+later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
+proved to be a formative period for piecework as we would come to know it.}
+Piecework regimes in farms and in homes engaged workers in assembling clothing, where they were paid by the \textit{piece} of output. %, and
+%finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
+%\topic{Farm and field work emerged early on as a rich field for the use of piecework,
+%perhaps in part due to the relatively low skillset necessary to complete work.}
+Textile manufacturers found that they could deliver fabric to people at their homes,
+paying them to sew together clothing.
+The manufacturers would later return to retrieve the finished garments,
+paying these workers for each \textit{piece} of clothing made. 
+Farm work applied the idea of piecework by
+paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
+\msb{cutting this next part, it seems irrelevant to the paragraph:}
+%\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
+%the only factor important in piecework was
+%whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
+%Piecework's advantage,
+%\citeauthor{hughRaynbirdTaskWork} argues,
+%was this hands--off management of workers.
+
+Workers could, in principle, assemble as much or as little clothing as they wanted;
+the reality was more grim, as
+\citeauthor{riisOtherSideLives} documented in ``\citetitle{riisOtherSideLives}'' in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
+He found that
+workers endured bleak living conditions and
+worked long hours attempting to scrape together a living.
+
+\subsubsection{The Match--Girls' Strike}
+
+%\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
+%as well as the early years of worker collective organization and action.}
+\topic{Match--makers were some of the first workers in mass manufacturing
+to successfully rally for political causes.}
+At the end of the \nth{19} century,
+manufacturers employed women to assemble matchsticks in factories.
+These women rallied first in the form of a march on parliament in 1871 to protest a proposed tax, and 
+later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.
+
+\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
+much deeper resentment had been boiling over for years.}
+Match--girls were already frustrated with
+the arbitrariness of management,
+poor working conditions, and
+having to work with phosphoric materials (which was proving hazardous).
+\ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}
+
+\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
+This was one of the earliest and most famous successful worker strikes,
+and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
+As \citeauthor{weyer1894history} described,
+``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
+the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+
+%\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history, we should shine a light onto the reputation that had formed for these workers.}
+Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
+according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
+\citeauthor{10.2307/3827491} noted that match girls
+``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
+But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
+These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
+Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
+making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.
+
+\ali{I want to discuss a few things here:
+\begin{itemize}
+  \item there were protests about something sort of like surge pricing in crop--picking, but i don't know how to make that a paragraph (maybe merge it with match--girls?)
+\end{itemize}}
+
+\subsubsection{Industrial Workers}
+\topic{Piecework might be most familiar to the HCI researcher in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
+%In the railroad industry and especially on the assembly line we see many of the mechanisms of work and worker management that later made their way into crowd work.
+%Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
+%We'll expose these facets of piecework very superficially in the context of industrial work to better speak to these research questions in crowd work later.
+\topic{Before the factory assembly line arose, railway companies adopted piecework regimes in the early \nth{20} century.
+\msb{What was pieceworky about this?
+They were paid per meter of railroad that they laid down?}
+What followed was a flourishing of piecework management practices,
+as railway companies worked to find effective ways
+to motivate and evaluate this skilled workforce of engineers.}
+\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
+finding that they employed ``efficiency experts'' to develop a ``standard time''
+to determine pay for each task at the company informed by
+``thousands of individual operations''; % ~\cite{10.2307/23702539}.
+\citeauthor{10.2307/23702539} goes on to list
+some of the roles required to facilitate piecework
+in the early \nth{20} century
+--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
+
+\topic{This oversight, while controversial
+(especially among workers~\cite{american1921problem}),
+paved the way for piecework to grow substantially.}
+With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's eponymously named \textit{Fordism}),
+piecework in the early and mid--\nth{20} century surged, especially in industrial work.
+Scientific management promised that the careful measurement of workers would yield
+higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
+While \citeauthor{Brown01041990} points out that
+piecework dramatically advanced the instrumented measurement of workers,
+in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
+using completion times to inform per--task compensation,
+similarly to efficiency experts in the Santa Fe Railway, but
+substantially more precise.
+
+\topic{The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.}
+\citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+They attribute this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
+
+\topic{Piecework %through the \nth{20} century centered around automobile and other mass manufacturing, but
+found its way into the war effort during World War II.}
+With the vast majority of men drafted into service,
+factories found themselves turning to
+a predominantly female workforce that had neither
+the formal training nor
+the years of apprenticeship experience
+that conventional workers would have had.
+Rather than attempting to train this new labor force in every aspect of industrial work,
+these women were trained for individual tasks
+% --- such as riveting component to another ---
+and assigned to that task. %, decomposed from a more comprehensive task.
+``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
+
+% \msb{confusing timing; now we're going backwards?}
+
+% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
+% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
+
+\ali{some things I didn't cover:
+\begin{itemize}
+  \item research finding that task variability is bad for piecework, but worker variability is good
+  \item talk about the foreman
+  \item worker advocacy groups
+\end{itemize}
+should I be making a serious effort to get that stuff in here? this section is kind of long already
+(and it'll definitely be much longer than the other case studies).}
+
+\begin{comment}
+What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)
+
+- Graves: railway companies used ``efficiency experts'' to study how long tasks should take
+- Hart: evaluation limits complexity (we can affect that with peer evaluation!)
+- Graves: sparks of Scientific Management in Piecework
+- organization types are important determinants of piecework viability: lots of types of tasks? bad
+  - Hart (I think?): variability in *worker* quality is fine
+- Foreman is important
+- Worker advocacy groups arose to speak out against piecework
+
+\end{comment}
 

### 2017-01-08 drafts/labor/cases.tex
rough pw perspective

@@
 This has largely involved iteratively
-  identifying challenges to increasing complexity and
+identifying barriers on complexity and
 overcoming them through novel designs of work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
-The question that has emerged then
-has been \textit{whether} there are limits to crowdsourcing
---- and if so, what factors determine them.
-Through this lens,
-a number of contributions to the field have pushed the boundaries of crowd work.
+The question then
+has become \textit{whether} there are limits to crowdsourcing,
+and if so, what factors determine them.
+To this question, a number of contributions to the field have pressed for answers.
 
-The exploration of crowdsourcing's potential and limits has principally looked at
+The exploration of on--demand work's potential and limits has principally looked at
 manipulating and extending along three dimensions:
@@
 We'll also point to corresponding piecework literature addressing these aspects.
-\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
+% \msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
 Finally, we'll discuss how these elements will serve
@@
 
+\notinsubfile{
   \subfile{complexity.tex}
@@
   \subfile{relationships.tex}
-
+}
 \onlyinsubfile{

### 2017-01-08 drafts/labor/complexity.tex
rough pw perspective

@@
 
-\subsection[What are the complexity limits of crowd work]
-{Identifying the Complexity Limits of Crowd Work}\label{sec:complexity}
-A key question to the future of crowd work is
+\subsection[What are the complexity limits of on--demand work]
+{Identifying the Complexity Limits of On--Demand Work}\label{sec:complexity}
+A key question to the future of on--demand work is
 \textit{what} precisely will become part of this economy.
@@
 So, our first question:
-how complex can the work outcomes from crowd work be?
+how complex can the work outcomes from on--demand work be?
 
@@
 proving the viability of crowdsourcing in complex work.}
-Unless crowdsourcing can demonstrate viability for meaningfully complex tasks,
+Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
 the argument runs,
@@
 
-\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing workflows.}
+\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing work--flows.}
 Beginning with a goal that has
@@
 \citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
-As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use crowd work.
+As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use on--demand work.
 
@@
 \subsubsection{\implication}
-Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
+Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
 Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
@@
 If we can overcome this obstacle,
-we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
+we might be able to empower more of these workers to do complex work such as engineering and metalworking,
 rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
-However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.
+However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to on--demand work unless it makes common expertise more broadly available.
 
@@
 However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
-So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford crowd work.
+So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to benefit from soliciting on--demand work.
 This shift may enable complex goals that were not cost--effective before to become feasible.
 However, because algorithms remain far from replicating all of the foreman's responsibilities,
-most likely is a middle ground in which crowd work re--introduces
+most likely is a middle ground in which on--demand work re--introduces
 the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
@@
 Expensive manufacturing equipment has been largely replaced by computer code~\cite{lessig2006code}.
-As with lowered costs of management, lowered infrastructure costs will make crowd work accessible to a broader set of people and organizations.
-This in and of itself does not raise the complexity ceiling, but by broadening the potential market for crowd work, it may enable a new set of goals and needs take part.
+As with lowered costs of management, lowered infrastructure costs will make on--demand work accessible to a broader set of people and organizations.
+This in and of itself does not raise the complexity ceiling, but by broadening the potential market for on--demand work, it may enable a new set of goals and needs take part.
 

### 2017-01-08 drafts/labor/conclusion.tex
rough pw perspective

@@
 \section{Conclusion}
-Crowd work and on--demand work are not new: they are contemporary instantiations of piecework.
-In this paper, we reconsider three major research questions in crowd work using the lens of piecework:
+On--demand work is not new, but a contemporary instantiations of piecework.
+In this paper, we reconsider three major research questions in on--demand work using the lens of piecework:
 \begin{inlinelist}
@@
 To do so, we draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon.
-Reciprocally, we believe that modern crowd work will teach us about the broader phenomenon of piecework as well.
+Reciprocally, we believe that modern on--demand work will teach us about the broader phenomenon of piecework as well.
 If history really does repeat itself, the best we can do is be prepared.

### 2017-01-08 drafts/labor/decomposition.tex
rough pw perspective

@@
 
-\topic{If decomposition is key to success in crowd work,
+\topic{If decomposition is key to success in on--demand work,
 the question arises: what can,
@@
 The work that Airy solicited resonates with modern crowd work for several reasons.
-\begin{Numberlist}
-\item work output was quickly verifiable;
+First, work output was quickly verifiable;
 Airy could assign variably skilled workers to compute values,
 and have other workers check their work.
-\item tasks were discrete
+Second, tasks were discrete
 --- that is,
 independent from one another.
-\item workers could be trained on a very narrow subset of mathematical skills to be
+Third, workers could be trained on a very narrow subset of mathematical skills to be
 sufficiently qualified to do this work.
-\end{Numberlist}
 
@@
 allowing pieceworkers to sequence their tasks optimally on their own~\cite{hart2013rise}.
-The result is that crowd workers are more free agents than historically was the case.
+The result is that on--demand workers are more free agents than historically was the case.
 However,

### 2017-01-08 drafts/labor/discussion.tex
rough pw perspective

@@
 \section{Discussion}
-\topic{In our analysis of crowd work via the piecework lens, three issues arose:}
+\topic{In our analysis of on--demand work via the piecework lens, three issues arose:}
 \begin{inlinelist}
@@
 \cite{scholz2012digital}.
-Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern crowd work.
+Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
 For example, many of the challenges that \DO~overcame in crowd collective action, such as designing for trustworthiness and ensuring anonymity, were relatively unique challenges precipitated by the affordances of the internet.
@@
 attempting to draw meaningfully from historical scholarship would be folly;
-enough of piecework can and does inform crowdsourcing that
-% we should take this as a cue to look first for
+enough of piecework can and does inform on--demand work that
 HCI and CSCW might seek out historical framings for other phenomena of study as well.
-% as this work can outline both the mechanisms that we should expect to play out
 While we can only speculate one of (perhaps many) possible futures, history does allow us to articulate and bound which futures appear more likely.
@@
 the piecework foundation which informed our case studies
-to trace out possible dystopian and utopian futures for crowd work.
+to trace out possible dystopian and utopian futures for on--demand labor.
 
 \subsection{Utopian and Dystopian Visions}\label{sec:polarizationOfCrowdWork}
-\topic{An easy narrative is to characterize the future of crowd work  at one of two extremes.}
+\topic{An easy narrative is to characterize the future of on--demand labor at one of two extremes.}
 On one hand, crowd work researchers
@@
 a potentially bright future that enables the achievement of near--impossible goals and career opportunities~\cite{redballoon,crowdworkFuture,vizwiz,suzukiAtelier}.
-% \ali{and who else should I cite here?}
 On the other hand, researchers
-warn that crowd work will create exploitative sites of dispossession~\cite{scholz2012digital},
+warn that on--demand labor will create exploitative sites of dispossession~\cite{scholz2012digital},
 racial discrimination~\cite{edelman2015racial},
@@
 \cite{uberAlgorithm,dynamo,turkopticon}.
-However, social policy has advanced since the early 1900s, so as crowd work gains popularity a repeat of \textit{How the Other Half Lives}~\cite{riisOtherSideLives} seems less likely.
+However, social policy has advanced since the early 1900s, so as on--demand work gains popularity a repeat of \textit{How the Other Half Lives}~\cite{riisOtherSideLives} seems less likely.
 
@@
 \cite{futureUnions},
-then the future of crowd work may follow
+then the future of on--demand labor may follow
 the same trajectory of worker empowerment that piecework \textit{later} found.
@@
 
-% \topic{Perhaps the most pressing question is this:
-% is the future of crowd work going to look more \textit{Dystopian} or \textit{Utopian}?}
-% If we have learn nothing, crowd work will probably end up somewhere in the middle.
-% It's possible that the difficulties of enforcing laws on multinational corporations is
-% tipping the balance of power in the direction of corporations, in which case
-% things will get worse;
-% but arguably we can affect change on the trajectory of crowd work,
-% benefiting from everything piecework scholars have learned about
-% collective action and governance
-% among workers, while also avoiding some of the perils they faced.
-% \ali{I really don't like this paragraph, but I felt like we needed to plant a flag on the question in the topic sentence}
-
 \subsection{A Research Agenda}\label{sec:whatShouldBeTheFuture}
 Piecework also helps bring into focus the
-% Taking a step back from the work that's been done so far and
-% thinking about where the body of crowd work research is attempting to go may help us identify
 areas of research that might bear the most fruit.
@@
 the most complex and open-ended wicked problems~\cite{rittel1973dilemmas} remain the domain of older human collectives such as governments and organizations. 
-% we have yet to find insurmountable limits to crowdsourcing.
 In addition, we can learn from the piecework literature as it relates to
@@
 partner with 
-% \ali{I feel like Lily would take some umbrage at this because she seemed to not like the narrative of us researchers just unilaterally bequeathing power upon Turkers (well, Turkers certainly disliked it).
-Thoughts on refactoring?}
 crowd workers in ways that were unimaginable to piecework researchers.
@@
 not just because it's possible, but because our community
-% (of mostly computer scientists)
 appreciates the importance of approaches such as participant--observation and ethnography as a whole~\cite{olson2014ways}.
-% Whereas piecework didn't have the benefit of Anthropology as we know it or the myriad ways of discovering, learning about, and empowering workers
-% What questions are we trying to answer?
-% If, as we assumed in
-% the previous section on \namerefl{sec:polarizationOfCrowdWork},
-% our driving motivation is to empower workers, we
-
-% \end{enumerate}
-
-\endnotetext{
-\begin{enumerate}
-  \item the past is no guarantee of the future; what could be unforeseen?
-  \item it's easy to fall into the (dys|u)--topian camp of what piecework/crowd work will look like
-  \item what should we be focusing on? the future of crowd work
-\end{enumerate}
-}
 

### 2017-01-08 drafts/labor/introduction.tex
rough pw perspective

@@
 \makeatother
+
 \begin{document}
@@
 
-\topic{The past decade has seen a flourishing of \textit{on--demand work},
-largely driven by the reformulation of work as
+\topic{The past decade has seen a flourishing of \textit{on--demand work}
+driven by the reformulation of work as
 constituent parts of larger tasks.}
-This framing of work into modular \msb{missing word?}
+This framing of work into modular components
 has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
-Distributed paid participants then engage in work whenever their schedules allow,
+Distributed workers then engage in work whenever their schedules allow,
 often with little to no awareness of the broader context of the work, and
@@
 
-\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complex, conflicted culture amongst both the workers who enable it and the researchers who study it.}
-\msb{This paragraph's topic sentence is about the conflicted culture, but the content seems to be about the evolution of the phenomenon. What's the goal here?}
-\citeauthor{howe2008crowdsourcing} first described crowdsourcing broadly as
+\topic{As on--demand work has grown far beyond information work, it has given rise to scholarship exploring myriad facets of on--demand work.}
+\citeauthor{howe2008crowdsourcing} first described crowdsourcing as
 ``outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
@@
 In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
-a work and payment structure which breaks tasks down into standalone contracts,
-wherein payment is made for work output, rather than for time.
-Piecework as a metaphor for crowd work is not new.
-Indeed,
+a work and payment structure which breaks tasks down into discrete jobs,
+wherein payment is made for work \textit{output}, rather than for \textit{time}.
+Piecework use as a lens on on--demand work is not new.
 \citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
@@
 might expect to see on the horizon.
-% \end{numberlist}
 For example, we will draw on the piecework literature regarding task decomposition,
@@
 leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
---- namely, by enabling precise tracking and measurement of workers via algorithms and software. \msb{leading to... (let's close the loop)}
+--- namely, by enabling precise tracking and measurement of workers via algorithms and software.
+By doing so, on--demand labor will find ways to decompose tasks to unprecedented levels.
 
 We organize this paper as follows:
-first, we review the definition and historical arc of piecework
-to lay groundwork and make clear the analogy to on--demand work
+first, we review the definition and history of piecework
+to make clear the analogy to on--demand work
 (which we'll call \textit{crowd work} hereon
@@
 finally, we will offer implications for researchers and practitioners based on our results.
-% \end{numberlist}
 
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-% }
+\onlyinsubfile{
+  \clearpage
+  \balance{}
+  \printbibliography
+}
 

### 2017-01-08 drafts/labor/piecework_lit.tex
rough pw perspective

--- 2017-01-06 drafts/labor/piecework_lit.tex
@@
 Since this paper traces a much stronger parallel between
-(historical) piecework and (contemporary) crowd work
+(historical) piecework and (contemporary) on--demand work
 (and, more generally, on--demand labor),
@@
 instead, it sets up the scaffolding necessary for
-our later investigations of crowd work's three questions:
+our later investigations of on--demand work's three questions:
 complexity limits,
@@
 a concise definition of piecework
-in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework
-(where he also calls it ``measure work'', ``grate work'', and ``task work''),
+in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
+% (where he also calls it ``measure work'', ``grate work'', and ``task work''),
 particularly driven toward encapsulating the manual labor of farmwork.}
-\ali{I just made the line ``Raynbird's essay on piecework''; is that enough to make it sound less important?}
 \citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
@@
 ``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
-Indeed, employers adopted piece--rate compensation in some aspects and
-time--rate compensation in others.
-The Rowan premium system,
-which essentially paid workers
-a base rate for time plus, potentially,
-an additional pay dependent on output~\cite{rowan1901premium};
-this was just one of several alternatives to categorical time-- and piece--rate renumeration paradigms.
+Employers implemented piece--rates in some aspects and
+time--rates in others.
+The Rowan premium system, for example,
+essentially paid workers
+a base rate for time plus
+additional pay dependent on output~\cite{rowan1901premium}.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
@@
 \textit{as well as} additional compensation tied to performance,
-workers under this regime were
+workers were
 in some senses ``task--labourers'', but
 in other senses ``day--labourers''.
+This was just one of several alternatives to strict time-- and piece--rate renumeration paradigms.
 
@@
 
-\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
-Between 1938 and 1942, the proportion of metal workers who were pieceworkers had climbed steeply 11\% to 60\%~\cite{hart2005piecework};
-by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
-\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions of piecework
+\topic{Piecework's popularity in the United States and Europe fell almost as quickly as it had climbed.}
+Between 1938 and 1942, the proportion of metal workers under piecework systems had climbed steeply from 11\% to 60\%~\cite{hart2005piecework}.
+By 1961, that proportion dropped to 8\%~\cite{carlson1982time}.
+\citeauthor{carlson1982time} details that, from 1973 to 1980, the holdouts of piecework
 --- where more than 50\% worked under incentive wage plans ---
@@
 
-\ali{``, ultimately arguing that\dots''?}
-
 In summary, piecework:
 \begin{inlinelist}
-  \item paid workers for quantity of work done, rather than time done,
-        but occasionally mixed the two payment models
-  \item afforded workers freedom in when and how much to work
-  \item structured tasks such that people who didn't have the training
-        to engage in the traditional labor force could still participate.
+  \item paid workers for \textit{quantity} of work done, rather than \textit{time} done,
+        but occasionally mixed the two payment models;
+  \item afforded workers a sense of freedom and independence; and
+  \item structured tasks in such a way as to facilitate more narrowly scoped training and education.
 \end{inlinelist}
@@
 
-Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
+Viewing on--demand work as a modern instantiation of piecework is relatively straightforward by this definition.
 First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
@@
 
+% \onlyinsubfile{\clearpage}
 \subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
-\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
+\topic{Piecework might be most familiar to the HCI researcher at the assembly line, which largely defined manufacturing through the \nth{20} century.}
 In the railroad industry and especially on the assembly line
@@
 We'll expose these facets of piecework very superficially in the context of industrial work
-to better speak to these research questions in crowd work later.
+to better speak to these research questions in on--demand labor later.
 
@@
 - Riis saw terrible conditions, documented and communicated it to the world
+
+\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
+``When he works by the day the Italian operative wishes to leave before the whistle blows,
+but if he works by the piece he will work as many hours as it is possible for him to stand''~\cite{clark1908cotton}.
 

### 2017-01-08 drafts/labor/relationships.tex
rough pw perspective

@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{\pieceworkpers}
@@
 
-Early observers believed that workers were strongly motivated by the piecework model.
-\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
-``When he works by the day the Italian operative wishes to leave before the whistle blows,
-but if he works by the piece he will work as many hours as it is possible for him to stand.''
-Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
+\topic{We discussed earlier that observers believed that workers were strongly motivated by the piecework model~\cite{clark1908cotton},
+but the emergent trend contrasted with this early rhetoric,
+when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.}
+Workers, \citeauthor{roy1954efficiency} found, formed even more starkly antagonistic relationships with their managers.
+% Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
 
-Soon, many worker organizations were weighing in on (or, more precisely, against) piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
-As mounting attention increasingly revealed problems in piecework's treatment of workers, the workers themselves began to speak out about their frustration with this new regime.
-Organizations representing railway workers, mechanical engineers, and others began to mount advocacy in defense of workers~\cite{american1921problem,richards1904anything}.
-
+\topic{As managers became increasingly onerous in their management of pieceworkers,
+workers began resisting piecework regimes and the methods that came along with it.}
+Soon, many worker organizations were weighing in on
+(or, more precisely, against)
+piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
+As mounting attention increasingly revealed problems in piecework's treatment of workers,
+workers themselves began to speak out about their frustration with this new regime.
+Organizations representing
+railway workers,
+mechanical engineers, and
+others began to mount advocacy in defense of workers~\cite{american1921problem,richards1904anything}.
 Pieceworkers' relationships with their employers eventually developed a pattern of using 
-laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,
-      mccallum2013global,jacoby1983union}.
+laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,mccallum2013global,jacoby1983union}.
 Collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.
 
-\topic{Less is known about how pieceworkers related to each other.}
-For one thing, the research methods we typically associate with the exploratory study of cultures
---- Anthropology, and namely participant--observation, ethnography, etc\dots~---
-didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
-Primary sources indicate that labor organizations wished for workers to identify as a collective group, 
+\topic{There's a noticeable dearth of information on the inter--personal relationships among pieceworkers
+beyond the match--girls at the end of the \nth{19} century.}
+% For one thing, the research methods we typically associate with the exploratory study of cultures
+% --- Anthropology, and namely participant--observation, ethnography, etc\dots~---
+% didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
+Nevertheless, we can offer some observations:
+primary sources indicate that labor organizations wished for workers to identify as a collective group, 
 ``not only as railroad employees but also as members of the larger life of the community''~\cite{american1921problem}.
 Doing this, 
-\citeauthor{ostrom1990governing} and others argued,
+\citeauthor{ostrom1990governing} and others later argued,
 would facilitate collective action and perhaps collective governance~\cite{ostrom1990governing,russell1982collective,olsonlogic}.
@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{\whatchanged}
@@
 
-While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', there was generally less written about quality concerns for historical pieceworkers than there is in modern crowd work.
+While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', there was generally less written about quality concerns for historical pieceworkers than there is in modern on--demand work.
 Why the difference? 
@@
 
-Relative to the mature state of collective action for pieceworkers offline, crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
-Both pieceworkers and crowd workers have struggled at times to form a collective identity necessary to organize.
+Relative to the mature state of collective action for pieceworkers offline,
+crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
+Both pieceworkers and on--demand workers have struggled at times to form a collective identity necessary to organize.
 With workers joining and leaving the crowd labor force continuously, and with many part-time members, it is extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
@@
 
-\onlyinsubfile{\clearpage}
+% \onlyinsubfile{\clearpage}
 \subsubsection{\implication}
 
-The decentralization and anonymization of crowd work will continue to make many of its social relationships a struggle.
+The decentralization and anonymization of on--demand work, especially online crowd work, will continue to make many of its social relationships a struggle.
 While some workers get to know each other well on forums~\cite{martin2014being,crowdcollab}, many never engage in these social spaces.
@@
 
-\topic{The history of piecework further suggests that relationships between workers and employers might be improved if employers engaged in more human management styles.}
-Instead of delegating as many management tasks as possible to an algorithm, it might be possible to build dashboards and other information tools that empower modern crowd work foremen~\cite{kulkarni2012mobileworks}.
+\topic{The history of piecework further suggests that
+relationships between workers and employers might be improved
+if employers engaged in more human management styles.}
+Instead of delegating as many management tasks as possible to an algorithm,
+it might be possible to build dashboards and
+other information tools that empower modern crowd work foremen~\cite{kulkarni2012mobileworks}.
 If the literature on piecework is to be believed,
@@
 \topic{Reciprocally, crowd work may be able to inform piecework research.}
-There exists far less literature about pieceworkers' relationships than there does today about crowd workers' relationships.
-Crowd work research benefits from both the accessibility of digital platforms, as well as the firmer theoretical basis of Anthropology than
+There exists far less literature about pieceworkers' relationships than there does today about on--demand workers' relationships.
+Research into on--demand labor benefits from both the accessibility of digital platforms as well as the firmer theoretical basis of Anthropology than
 existed at the turn of the \nth{20} century, when piecework began to emerge.
@@
 didn't take even a resemblance of their contemporary forms until these works~\cite{malinowski2002argonauts,boas1940race,mead1973coming}.
-Modern crowd work may give us an opportunity to revisit open questions in piecework with a more refined lens.
+On--demand labor today may give us an opportunity to revisit open questions in piecework with a more refined lens.
 

### 2017-01-08 drafts/labor/piecework_lit.tex

--- 2017-01-07 drafts/labor/piecework_lit.tex
@@
 \end{comment}
+
+% \onlyinsubfile{\clearpage}
+\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
+\topic{Piecework might be most familiar to the HCI researcher at the assembly line, which largely defined manufacturing through the \nth{20} century.}
+In the railroad industry and especially on the assembly line
+we see many of the mechanisms of work and worker management that later made their way into crowd work.
+Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
+We'll expose these facets of piecework very superficially in the context of industrial work
+to better speak to these research questions in on--demand labor later.
+
+\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
+what followed was a flourishing in the management practices with pieceworkers,
+as railway companies struggled to find effective ways
+to motivate and evaluate this skilled workforce of engineers.}
+\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
+finding that they employed ``efficiency experts'' to develop a ``standard time''
+to determine pay for each task at the company informed by
+``thousands of individual operations''; % ~\cite{10.2307/23702539}.
+\citeauthor{10.2307/23702539} goes on to list
+some of the roles required to facilitate piecework
+in the early \nth{20} century
+--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
+
+\topic{This oversight, while controversial
+(especially among workers~\cite{american1921problem}),
+paved the way for piecework to grow substantially.}
+With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's eponymously named \textit{Fordism}),
+piecework in the early and mid--\nth{20} century surged, especially in industrial work.
+Scientific management promised that the careful measurement of workers would yield
+higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
+While \citeauthor{Brown01041990} points out that
+piecework dramatically advanced the instrumented measurement of workers,
+in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
+using completion times to inform per--task compensation,
+similarly to efficiency experts in the Santa Fe Railway, but
+substantially more precise.
+
+\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
+found its way into the war effort during World War II.}
+With the vast majority of men drafted into service,
+factories found themselves turning to
+a predominantly female workforce that had neither
+the formal training nor
+the years of apprenticeship experience
+that conventional workers would have had.
+Rather than attempting to train this new labor force in every aspect of industrial work,
+these women were trained for individual tasks
+% --- such as riveting component to another ---
+and assigned to that task. %, decomposed from a more comprehensive task.
+One might reflect on the observation that ``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+was a pieceworker~\cite{davies2014origins}.
+
+% \msb{confusing timing; now we're going backwards?}
+
+% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
+% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
+\topic{The 1930s represented a boom for piecework on an unprecedented scale,
+especially among engineering and metalworking industries.}
+\citeauthor{hart2013rise} characterize the 1930s
+--- and more broadly the first half of the \nth{20} century ---
+as the ``heyday'' of the use of piecework.
+They attribute this to the shortage of male workers,
+who would have gone through a conventional apprenticeship process
+affording them more comprehensive knowledge of the total scope of work.
 

### 2017-01-08 drafts/labor/relationships.tex
tweaks on relationships

@@
 
-\topic{We discussed earlier that observers believed that workers were strongly motivated by the piecework model~\cite{clark1908cotton},
-but the emergent trend contrasted with this early rhetoric,
+\topic{Early observers believed that workers were strongly motivated by the autonomy of working in the piecework model}.
+\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
+``When he works by the day the Italian operative wishes to leave before the whistle blows,
+but if he works by the piece he will work as many hours as it is possible for him to stand.''
+However, the emergent trend contrasted with this early rhetoric, as
 when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.}
@@
 
-\topic{As managers became increasingly onerous in their management of pieceworkers,
-workers began resisting piecework regimes and the methods that came along with it.}
+\topic{Soon workers began resisting piecework regimes and the methods that came along with it.}
+The match--girls engaged in their famous strike of 1888, particularly pushing to abolish the fines that were taken out of their wages.
+Soon others followed suit, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}
+
 Soon, many worker organizations were weighing in on
@@
 laborer advocacy groups~\cite{levi2009union,ahlquist2013interest,mccallum2013global,jacoby1983union}.
-Collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.
+Following the template of the match--girls, collective action grew to become a central component of negotiating with managers~\cite{russell1982collective,olsonlogic}.
 

### 2017-01-08 drafts/labor/decomposition.tex
decomposition revision

@@
 
-\citeauthor{Brown01041990} inquired from another direction,
-asking
-what limited the adoption of piecework in industries that otherwise gravitated toward it
-(in the case studies he examined,
-this mostly focused on railway engineers),
-ultimately arguing that factors such as the nature of the work design
-(specifically,
-the homogeneity of tasks) and the costs associated with adopting a piecework model
-were the major contributing factors that determined the use of piecework~\cite{Brown01041990}.
-
 \topic{The research community relating to
@@
 has been wrestling with the decomposition of work for the better part of a century.}
-The beginnings of
-systematic task decomposition 
-stretch back as far as the \nth{19} century,
-when Airy employed young boys at the Greenwich Observatory who
-``possessed the basic skills of mathematics,
-including
-`Arithmetic,
-the use of Logarithms,
-and Elementary Algebra'~''
-to compute astronomical phenomena~\cite{grier2013computers}.
-The work that Airy solicited resonates with modern crowd work for several reasons.
-First, work output was quickly verifiable;
-Airy could assign variably skilled workers to compute values,
-and have other workers check their work.
-Second, tasks were discrete
---- that is,
-independent from one another.
-Third, workers could be trained on a very narrow subset of mathematical skills to be
-sufficiently qualified to do this work.
+Decomposition was arguably the main innovation of Airy's human computers.
+Rather than hire expert computers, Airy identified ways to break down astrological calculations into steps that could be completed with only a basic knowledge of mathematics.
+Future piecework efforts followed a similar model.
+
+\topic{Decomposability is core to piecework: one cannot pay by the piece unless the pieces are small and discretizable.}
+Focusing on our case study of railway engineers, \citeauthor{Brown01041990} asked 
+what enabled and limited the adoption of piecework.
+He argued that the main determinants were task homogeneity and the fixed costs of the machinery and training~\cite{Brown01041990}.
+In other words, if tasks could be clearly decomposed and defined, and if the firm could afford the training and machinery for the pieceworkers, piecework tended to be adopted.
+
+Decomposition ultimately led to quantification and scientific management.
+With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+(and Henry Ford's eponymously named \textit{Fordism}),
+piecework in the early and mid--\nth{20} century surged, especially in industrial work.
+Scientific management promised that the careful measurement of workers would yield
+higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
+While \citeauthor{Brown01041990} points out that
+piecework dramatically advanced the instrumented measurement of workers,
+in \citeauthor{taylor1914principles}'s time highly instrumented,
+automatic measurement of workers was all but impossible~\cite{Brown01041990}.
+Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
+using completion times to inform per--task compensation,
+similarly to the efficiency experts hired in the Santa Fe Railway, but
+substantially more precise.
+As a result, the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
 
@@
 piecework does not compensate workers for time spent switching tasks''.
-Ultimately, \citeauthor{bewley1999wages} argues that
-``[piecework is] infeasible,
-because~\dots~total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
+%Ultimately, \citeauthor{bewley1999wages} argues that
+%``[piecework is] infeasible,
+%because~\dots~total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
+To \citeauthor{bewley1999wages}, the main limiter was the ability to define and measure the work.
 

### 2017-01-08 drafts/labor/complexity.tex
two attempts at pw perspective in complexity

@@
 
-{\topic{Only tasks that could be measured and priced could be completed via piecework
---- a fact which fundamentally limited the potential complexity of piecework.}
-% When \citeauthor{Brown01041990} investigated
-% what limited the adoption of piecework in industries that otherwise gravitated toward it
-% (e.g., railway engineers),
-% the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
-% \citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
-% which used ``efficiency experts'' to develop a ``standard time''
-% to determine pay for each task at the company informed by
-% ``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
-% ~\cite{10.2307/23702539}.
+\ali{What do we know?
+
+Airy's computers:
+\begin{itemize}
+  \item we can do some pretty complex stuff if we break tasks down into quickly verifiable parts
+  \item there's no need to train people to be experts for this work; non--experts can do the job just as well.
+\end{itemize}
+
+% Domestic and farm work
+% \begin{itemize}
+%   \item ???
+% \end{itemize}
+
+% match--girls
+% \begin{itemize}
+%   \item ???
+% \end{itemize}
+
+Industrial workers
+\begin{itemize}
+  \item revolutionary: we can train people on a narrow slice of the work and have them just do that work
+\end{itemize}
+}
+
+\topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
+Airy's human computers and among industrial workers, especially during World War II.}
+The majority of piecework's perspective, on this topic, will come from these areas.
+
+\topic{Airy's work opened the door to greater task complexity by decomposition.}
+Airy's computers had relatively limited education in mathematics, but
+they were nevertheless collectively able to accomplish complex goals by
+approaching distinct, subdivided, quickly verifiable tasks in chunks~\cite{grier2013computers}.
+\ali{do i need to say more?}
+
+\topic{It seemed, too, that only tasks that could be measured and priced could be completed via piecework
+--- a conclusion which would fundamentally limit the potential complexity of piecework.}
 Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
@@
 would find themselves overwhelmed if they attempted to assess creative work like design,
-which can take an arbitrary number of iterations before proceeding to a subsequent step
-\ali{should i cite some design research here?}.
+which can take an arbitrary number of iterations before proceeding to a subsequent step.
 
-\topic{Piecework was limited to tasks that could be clearly evaluated.}
-For example, the roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
-\citeauthor{hart2016rise} argue that evaluation is the ultimate complexity limit:
+\topic{The next innovation in piecework's drive toward complexity came on
+the assembly line, specifically during the war effort.}
+At a time when manufacturing firms needed more able--bodied workers, but
+all of the conventionally trained metalworkers had joined the armed forces,
+factories turned to women~\cite{honey1985creating,davies2014origins}.
+This need to train workers rapidly to do very narrowly defined tasks resulted in
+a \textit{recomposition} of work, wherein
+workers were only prepared for the tasks they would do;
+riveters would only be trained on how to do rivet and
+to pass the output to the next person on the assembly line,
+who could evaluate and send the work back or proceed with the next step.
+
+\ali{OR\dots}
+
+\topic{Only tasks that could be measured and priced could be completed via piecework
+--- a fact which fundamentally limited the potential complexity of piecework.}
+Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
+one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
+complex, creative work
+--- which is inherently heterogeneous and difficult to routinize ---
+would be unsuitable for piecework~\cite{10.2307/23702539}.
+\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
+``efficiency experts'' can effectively gauge how long known tasks should take, but
+would find themselves overwhelmed if they attempted to assess creative work like design,
+which can take an arbitrary number of iterations before proceeding to a subsequent step.
+
+\topic{Piecework was limited to tasks that could be quickly and accurately evaluated.}
+For example, the roles required to facilitate piecework
+in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
+\citeauthor{hart2016rise} argue that evaluation limited piecework's complexity:
 at some point, evaluating multidimensional work for quality
 (rather than for quantity) becomes infeasible.
-In their words,
+In his words,
 ``if the quality of the output is more difficult to measure than the quantity [\ldots]
 then a piecework system is likely to encourage
-an over--emphasis on quantity produced and an under--emphasis on quality''~\cite{hart2016rise}.
-Complex work, which is often subjective to evaluate, falls victim to this criteria.
+an over--emphasis on quantity~\dots~and an under--emphasis on quality''~\cite{hart2016rise}.
 
-\topic{This focus on measurement and tracking had consequences.}
-\citeauthor{10.2307/23702539} suggests that the first sparks of scientific management
-could be found in piecework:
-the approach of paying workers for each piece of output necessitated
-the rigorous tracking, measurement, and training of workers
-for which scientific management became famous~\cite{10.2307/23702539}.
-If true, 
-the concurrent upswing of
-scientific management and Fordism
-through the first two--thirds of the \nth{20} century
-alongside piecework was not only understandable, but predictable~\cite{hart2013rise}.
+% \topic{This focus on measurement and tracking had consequences.}
+% \citeauthor{10.2307/23702539} suggests that the first sparks of scientific management
+% could be found in piecework:
+% the approach of paying workers for each piece of output necessitated
+% the rigorous tracking, measurement, and training of workers
+% for which scientific management became famous~\cite{10.2307/23702539}.
+% If true, 
+% the concurrent upswing of
+% scientific management and Fordism
+% through the first two--thirds of the \nth{20} century
+% alongside piecework was not only understandable, but predictable~\cite{hart2013rise}.
 
-\topic{Piecework researchers also argue that,
-in addition to constraints on the kind of \textit{work} that's amenable to piecework,
-only certain kinds of \textit{organizations} were amenable to piecework.}
-Researchers detail three organizational criteria.
-\begin{Numberlist}
-\item \citeauthor{Brown01041990} argues that
-piecework ``is less likely in jobs with a variety of duties than in
-jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
-\citeauthor{SJOE:SJOE371} points out the phenomenon here as a market effect:
-``in an environment with multi--tasking,
-pay schemes based on tightly specified performance may
-induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
-\item complexity was limited by access to capital to create the necessary infrastructure.
-As \citeauthor{10.2307/23702539} reports,
-only the largest and most wealthy railroads had the resources
-necessary
-% ``pay the overhead involved in installing work reorganization''
-~\cite{10.2307/23702539}.
-\item organizations required capable managers in charge of the pieceworkers.
-The West Virginia mines, for example, hired foremen 
-to be the intermediary between upper management and the workers~\cite{10.2307/2118435}.
-These foremen were responsible for allocating resources and
-understanding when and how to modify work as necessary~\cite{wray1949marginal}.
-\end{Numberlist}
-So, in sum,
-organizations historically could only take advantage of piecework if they had
-homogeneous work to be done,
-access to capital to purchase the necessary equipment, and
-the ability to hire people who could serve as intermediaries between pieceworkers and management.
+\topic{Piecework researchers were also finding that certain types of \textit{organizations},
+as well as certain types of \textit{work},
+were more amenable to piecework regimes.}
+\citeauthor{Brown01041990} wrote that piecework was
+``less likely in jobs with a variety of duties than in
+jobs with a narrow set of routinized duties''~\cite{Brown01041990};
+\citeauthor{SJOE:SJOE371} hypothesized this as an effect of market forces
+--- suggesting that
+``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
+Airy seemed to recognize this,
+decomposing complex mathematical problems into easily verified components, % ~\cite{grier2013computers},
+but applying this model to tasks like railway engineering may have been prohibitively difficult.
 
-\topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
-particularly because maximizing the advantages of piecework seemed to reward
-smaller, more constrained, more narrowly--trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
-For most of the \nth{19} century,
-piecework was applied almost exclusively to farm and textile work.
-Work was simple and widely understood
---- farm workers didn't need to be trained on how to plow fields, or birth foals;
-seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.}
+\topic{But it was more than the inherent qualities of work and the organization of the firm that mattered;
+management practices themselves had to be appropriately geared to working with pieceworkers.}
+Competent foremen were integral both for allocating resources and
+for liaising between workers and upper management~\cite{wray1949marginal,10.2307/2118435}.
+% The West Virginia mines, as \citeauthor{10.2307/2118435} writes,
+% ~\cite{10.2307/2118435}.
+
+% \topic{Piecework researchers also argued that,
+% in addition to constraints on the kind of \textit{work} that's amenable to piecework,
+% only certain kinds of \textit{organizations} were amenable to piecework.}
+% Researchers detail three organizational criteria.
+% First, \citeauthor{Brown01041990} found that
+% piecework was ``less likely in jobs with a variety of duties than in
+% jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
+% \citeauthor{SJOE:SJOE371} pointed out this phenomenon as a market effect:
+% ``in an environment with multi--tasking,
+% pay schemes based on tightly specified performance may
+% induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
+% Second, complexity was limited by access to capital to create the necessary infrastructure.
+% As \citeauthor{10.2307/23702539} reported,
+% only the largest and most wealthy railroads had the resources
+% necessary to change existing systems to piecework regimes~\cite{10.2307/23702539}.
+% Third, organizations required capable managers in charge of the pieceworkers.
+% The West Virginia mines, for example, hired foremen 
+% to be the intermediary between upper management and the workers~\cite{10.2307/2118435}.
+% These foremen were responsible for allocating resources and
+% understanding when and how to modify work as necessary~\cite{wray1949marginal}.
+% So, in sum,
+% organizations historically could only take advantage of piecework if they had
+% homogeneous work to be done,
+% access to capital to purchase the necessary equipment, and
+% the ability to hire people who could serve as intermediaries between pieceworkers and management.}
+
+% \topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
+% particularly because maximizing the advantages of piecework seemed to reward
+% smaller, more constrained, more narrowly--trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
+% For most of the \nth{19} century,
+% piecework was applied almost exclusively to farm and textile work.
+% Work was simple and widely understood
+% --- farm workers didn't need to be trained on how to plow fields, or birth foals;
+% seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.
 

### 2017-01-08 drafts/labor/piecework_lit.tex
decomposition revision
two attempts at pw perspective in complexity

@@
 \end{comment}
-
-% \onlyinsubfile{\clearpage}
-\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
-\topic{Piecework might be most familiar to the HCI researcher at the assembly line, which largely defined manufacturing through the \nth{20} century.}
-In the railroad industry and especially on the assembly line
-we see many of the mechanisms of work and worker management that later made their way into crowd work.
-Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
-We'll expose these facets of piecework very superficially in the context of industrial work
-to better speak to these research questions in on--demand labor later.
-
-\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
-what followed was a flourishing in the management practices with pieceworkers,
-as railway companies struggled to find effective ways
-to motivate and evaluate this skilled workforce of engineers.}
-\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
-finding that they employed ``efficiency experts'' to develop a ``standard time''
-to determine pay for each task at the company informed by
-``thousands of individual operations''; % ~\cite{10.2307/23702539}.
-\citeauthor{10.2307/23702539} goes on to list
-some of the roles required to facilitate piecework
-in the early \nth{20} century
---- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
-
-\topic{This oversight, while controversial
-(especially among workers~\cite{american1921problem}),
-paved the way for piecework to grow substantially.}
-With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism}),
-piecework in the early and mid--\nth{20} century surged, especially in industrial work.
-Scientific management promised that the careful measurement of workers would yield
-higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
-While \citeauthor{Brown01041990} points out that
-piecework dramatically advanced the instrumented measurement of workers,
-in \citeauthor{taylor1914principles}'s time highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
-using completion times to inform per--task compensation,
-similarly to efficiency experts in the Santa Fe Railway, but
-substantially more precise.
-
-\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
-found its way into the war effort during World War II.}
-With the vast majority of men drafted into service,
-factories found themselves turning to
-a predominantly female workforce that had neither
-the formal training nor
-the years of apprenticeship experience
-that conventional workers would have had.
-Rather than attempting to train this new labor force in every aspect of industrial work,
-these women were trained for individual tasks
-% --- such as riveting component to another ---
-and assigned to that task. %, decomposed from a more comprehensive task.
-One might reflect on the observation that ``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women~\cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
-
-% \msb{confusing timing; now we're going backwards?}
-
-% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
-% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
-\topic{The 1930s represented a boom for piecework on an unprecedented scale,
-especially among engineering and metalworking industries.}
-\citeauthor{hart2013rise} characterize the 1930s
---- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
-They attribute this to the shortage of male workers,
-who would have gone through a conventional apprenticeship process
-affording them more comprehensive knowledge of the total scope of work.
 

### 2017-01-08 drafts/labor/decomposition.tex

@@
 
-\topic{The research community relating to
-piecework and labor
-has been wrestling with the decomposition of work for the better part of a century.}
-Decomposition was arguably the main innovation of Airy's human computers.
+\topic{Four major stages characterize decomposition in the history of piecework.}
+The first stage was decomposition of an expert task such that it could be done by non-experts. 
+This was arguably the main innovation of Airy's human computers.
 Rather than hire expert computers, Airy identified ways to break down astrological calculations into steps that could be completed with only a basic knowledge of mathematics.
-Future piecework efforts followed a similar model.
+%\topic{Decomposability is core to piecework: one cannot pay by the piece unless the pieces are small and discretizable.}
+Likewise, \citeauthor{Brown01041990} argued that piecework arose in industries where there was significant task homogeneity and low fixed costs of machinery and training~\cite{Brown01041990}.
+%In other words, if tasks could be clearly decomposed and defined, and if the firm could afford the training and machinery for the pieceworkers, piecework tended to be adopted.
 
-\topic{Decomposability is core to piecework: one cannot pay by the piece unless the pieces are small and discretizable.}
-Focusing on our case study of railway engineers, \citeauthor{Brown01041990} asked 
-what enabled and limited the adoption of piecework.
-He argued that the main determinants were task homogeneity and the fixed costs of the machinery and training~\cite{Brown01041990}.
-In other words, if tasks could be clearly decomposed and defined, and if the firm could afford the training and machinery for the pieceworkers, piecework tended to be adopted.
+After decomposing tasks for amateurs, the second major stage was to apply the same methods to domain experts. 
+Unlike Airy's human computers, railway engineers had significant expertise~\cite{Brown01041990}.
+As \citeauthor{Brown01041990} noted, however, it was still possible to discretize and measure their work.
+Thus, experts such as railway engineers became pieceworkers as well.
 
-Decomposition ultimately led to quantification and scientific management.
+Third, decomposition led to quantification and scientific management.
+What can be modularized can be measured, and what can be measured can be optimized.
 With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
@@
 substantially more precise.
-As a result, the distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
+The distillation of work into smaller units ultimately bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.
 
-\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
-and the conflicting pressures managers and workers put forth.}
-\citeauthor{bewley1999wages} in particular points out that
-the approach of paying workers by the piece is
-``\dots~not practical for workers doing many tasks,
-because of
-the cost of establishing the rates and because
-piecework does not compensate workers for time spent switching tasks''.
+The fourth and final stage was narrow expertise training.
+As World War II raged and there was a dearth of skilled workers, managers trained women just enough to be able to complete their tasks~\cite{hart2013rise}.
+Over time, these women could gain proficiency and gain broader expertise.
+
+%\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
+%and the conflicting pressures managers and workers put forth.}
+%\citeauthor{bewley1999wages} in particular points out that
+%the approach of paying workers by the piece is
+%``\dots~not practical for workers doing many tasks,
+%because of
+%the cost of establishing the rates and because
+%piecework does not compensate workers for time spent switching tasks''.
 %Ultimately, \citeauthor{bewley1999wages} argues that
@@
 %because~\dots~total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
-To \citeauthor{bewley1999wages}, the main limiter was the ability to define and measure the work.
+%To \citeauthor{bewley1999wages}, the main limiter was the ability to define and measure the work.
 
@@
 
-Third,
-modern crowd work has sliced work to such small scales that the marginal activities
---- things like finding work and cognitive task switching ---
-have become large relative to the tasks themselves~\cite{taskSearch}.
-In the historical case of piecework,
-moving metallurgical tools,
-mining equipment,
-or
-other industry materials would have been prohibitively difficult and slow;
-workers were encouraged to specialize in a single set of tasks,
-allowing pieceworkers to sequence their tasks optimally on their own~\cite{hart2013rise}.
-The result is that on--demand workers are more free agents than historically was the case.
-However,
-because they spend significant time searching for tasks,
-the piece rate is less a good estimate of take--home earnings than before.
+%Third,
+%modern crowd work has sliced work to such small scales that the marginal activities
+%--- things like finding work and cognitive task switching ---
+%have become large relative to the tasks themselves~\cite{taskSearch}.
+%In the historical case of piecework,
+%moving metallurgical tools,
+%mining equipment,
+%or
+%other industry materials would have been prohibitively difficult and slow;
+%workers were encouraged to specialize in a single set of tasks,
+%allowing pieceworkers to sequence their tasks optimally on their own~\cite{hart2013rise}.
+%The result is that on--demand workers are more free agents than historically was the case.
+%However,
+%because they spend significant time searching for tasks,
+%the piece rate is less a good estimate of take--home earnings than before.
 
@@
 \subsubsection{\implication}
-If measurement precision limited the depth of decomposition for piecework historically,
-as \citeauthor{10.2307/23702539} argues,
-then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
-Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved,
-or near enough that they no longer limit task decomposition.
-Now,
-not just tasks,
-but entire workers' histories~\cite{hata2017glimpse},
-can be collected and analyzed in detail.
+If decomposition in piecework progressed in four stages, we have seen three of them in on--demand work so far.
+First, as with piecework, on--demand work began by decomposing tasks so that anyone could complete them, as with data labeling on Amazon Mechanical Turk.
+Second, we began to modularize and measure existing expertise (e.g., software engineering, design) so that it could be brought into crowdsourcing systems~\cite{foundry,Chen2016}.
+Third, we used measurement to mathematically optimize workers' behaviors so that we could make the systems more efficient~\cite{weld2010decision}.
 
-However,
-decomposition has hit a second bottleneck: cognition. 
-Task switching costs and other cognitive costs make it difficult
+The fourth stage, then, appears likely to occur: narrow training of workers for expert tasks. 
+There is demand for skilled workers in many crowdsourcing tasks, and systems to help train workers~\cite{suzukiAtelier}. 
+We might expect to see the rise of systems that scaffold workers into extremely narrow areas of expertise.
+
+Finally, improved measurement and lowered costs of production have made it feasible to apply piecework methods to many domains where it may not have historically been possible.
+The limit is no longer measurement precision, but human cognition. 
+%as \citeauthor{10.2307/23702539} argues,
+%then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
+%Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved,
+%or near enough that they no longer limit task decomposition.
+%Now,
+%not just tasks,
+%but entire workers' histories~\cite{hata2017glimpse},
+%can be collected and analyzed in detail.
+Task switching and other cognitive costs make it difficult
 to work on tasks so far decontextualized from their original intention~\cite{delayAndOrderLasecki}.
@@
 logo design requires a deep understanding of the client and their goals.
-In part due to this limitation,
-99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
+%In part due to this limitation,
+%99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
 
-So,
-ultimately,
-the levels of decomposition are likely to follow the contours of context required.
-Low--context work will be extremely highly decomposed.
-High--context work will continue to be limited.
+%So,
+%ultimately,
+%the levels of decomposition are likely to follow the contours of context required.
+%Low--context work will be extremely highly decomposed.
+%High--context work will continue to be limited.
 

### 2017-01-08 drafts/labor/piecework_lit.tex

@@
 in the early \nth{20} century
---- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
-
-\topic{This oversight, while controversial
+--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}. \topic{This oversight, while controversial
 (especially among workers~\cite{american1921problem}),
 paved the way for piecework to grow substantially.}
-With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism}),
-piecework in the early and mid--\nth{20} century surged, especially in industrial work.
-Scientific management promised that the careful measurement of workers would yield
-higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
-While \citeauthor{Brown01041990} points out that
-piecework dramatically advanced the instrumented measurement of workers,
-in \citeauthor{taylor1914principles}'s time highly instrumented,
-automatic measurement of workers was all but impossible~\cite{Brown01041990}.
-Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
-using completion times to inform per--task compensation,
-similarly to efficiency experts in the Santa Fe Railway, but
-substantially more precise.
 

### 2017-01-08 drafts/labor/relationships.tex

@@
 However, the emergent trend contrasted with this early rhetoric, as
-when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.}
+when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.
 Workers, \citeauthor{roy1954efficiency} found, formed even more starkly antagonistic relationships with their managers.

### 2017-01-08 drafts/labor/complexity.tex
complexity

@@
 
+\msb{ summarizing the later conclusion so we can support it here:
+complexity is determined by whether we can encode it into workflows
+complexity is determined by whether we can modularize and measure it
+complexity is determined by how much management overhead there is
+complexity is determined by the infrastructure that can support it
+}
+
 \topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
-Airy's human computers and among industrial workers, especially during World War II.}
+Airy's human computers and among industrial workers.}
 The majority of piecework's perspective, on this topic, will come from these areas.
 
-\topic{Airy's work opened the door to greater task complexity by decomposition.}
-Airy's computers had relatively limited education in mathematics, but
-they were nevertheless collectively able to accomplish complex goals by
-approaching distinct, subdivided, quickly verifiable tasks in chunks~\cite{grier2013computers}.
-\ali{do i need to say more?}
+\topic{Airy's work on astronomical charts opened the door to greater task complexity by encoding the intelligence into the process rather than the people.}
+Airy's computers had relatively limited education in mathematics, but by combining individual simple mathematical operations, Airy was able to create a complex composite outcome~\cite{grier2013computers}.
+Likewise, in Ford's factories, no individual could build the entire car, but the process could emergently produce one.
 
-\topic{It seemed, too, that only tasks that could be measured and priced could be completed via piecework
---- a conclusion which would fundamentally limit the potential complexity of piecework.}
-Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
-one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
-complex, creative work
+However, when piecework intially entered the American economy,
+it was not used for complex work.
+Without having designed processes to combine work intelligently, piecework managers were restricted to available workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
+that apprenticeships imparted \cite{hart2013rise}.
+So, initially piecework arose for farm work, and as
+\citeauthor{hughRaynbirdTaskWork} and others discuss,
+the practice remained relatively obscure until
+it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
+Complexity levels remained low at the turn of
+the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+
+\topic{Measurement also limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework.}
+Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers.
+They identified task homogeneity and measurement as key requirements for piecework.
+However, complex, creative work
 --- which is inherently heterogeneous and difficult to routinize ---
-would be unsuitable for piecework~\cite{10.2307/23702539}.
+would then be unsuitable~\cite{10.2307/23702539}.
 \citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
 ``efficiency experts'' can effectively gauge how long known tasks should take, but
-would find themselves overwhelmed if they attempted to assess creative work like design,
+would find themselves overwhelmed if they attempted to assess creative tasks like scientific research,
 which can take an arbitrary number of iterations before proceeding to a subsequent step.
 
-\topic{The next innovation in piecework's drive toward complexity came on
-the assembly line, specifically during the war effort.}
-At a time when manufacturing firms needed more able--bodied workers, but
-all of the conventionally trained metalworkers had joined the armed forces,
-factories turned to women~\cite{honey1985creating,davies2014origins}.
-This need to train workers rapidly to do very narrowly defined tasks resulted in
-a \textit{recomposition} of work, wherein
-workers were only prepared for the tasks they would do;
-riveters would only be trained on how to do rivet and
-to pass the output to the next person on the assembly line,
-who could evaluate and send the work back or proceed with the next step.
+%\topic{The next innovation in piecework's drive toward complexity came on
+%the assembly line, specifically during the war effort.}
+%At a time when manufacturing firms needed more able--bodied workers, but
+%all of the conventionally trained metalworkers had joined the armed forces,
+%factories turned to women~\cite{honey1985creating,davies2014origins}.
+%This need to train workers rapidly to do very narrowly defined tasks resulted in
+%a \textit{recomposition} of work, wherein
+%workers were only prepared for the tasks they would do;
+%riveters would only be trained on how to do rivet and
+%to pass the output to the next person on the assembly line,
+%who could evaluate and send the work back or proceed with the next step.
 
-\ali{OR\dots}
+%\ali{OR\dots
 
-\topic{Only tasks that could be measured and priced could be completed via piecework
---- a fact which fundamentally limited the potential complexity of piecework.}
-Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
-one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
-complex, creative work
---- which is inherently heterogeneous and difficult to routinize ---
-would be unsuitable for piecework~\cite{10.2307/23702539}.
-\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
-``efficiency experts'' can effectively gauge how long known tasks should take, but
-would find themselves overwhelmed if they attempted to assess creative work like design,
-which can take an arbitrary number of iterations before proceeding to a subsequent step.
+%\topic{Only tasks that could be measured and priced could be completed via piecework
+%--- a fact which fundamentally limited the potential complexity of piecework.}
+%Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
+%one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
+%complex, creative work
+%--- which is inherently heterogeneous and difficult to routinize ---
+%would be unsuitable for piecework~\cite{10.2307/23702539}.
+%\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
+%``efficiency experts'' can effectively gauge how long known tasks should take, but
+%would find themselves overwhelmed if they attempted to assess creative work like design,
+%which can take an arbitrary number of iterations before proceeding to a subsequent step.
 
-\topic{Piecework was limited to tasks that could be quickly and accurately evaluated.}
-For example, the roles required to facilitate piecework
-in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
+\topic{Moreover, piecework was limited to tasks that could be quickly and accurately evaluated.}
+%For example, the roles required to facilitate piecework
+%in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
 \citeauthor{hart2016rise} argue that evaluation limited piecework's complexity:
@@
 an over--emphasis on quantity~\dots~and an under--emphasis on quality''~\cite{hart2016rise}.
+Complex work, which is often subjective to evaluate, falls victim to this criteria.
+%``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
 
@@
 
-\topic{Piecework researchers were also finding that certain types of \textit{organizations},
-as well as certain types of \textit{work},
-were more amenable to piecework regimes.}
-\citeauthor{Brown01041990} wrote that piecework was
-``less likely in jobs with a variety of duties than in
-jobs with a narrow set of routinized duties''~\cite{Brown01041990};
-\citeauthor{SJOE:SJOE371} hypothesized this as an effect of market forces
---- suggesting that
-``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
-Airy seemed to recognize this,
-decomposing complex mathematical problems into easily verified components, % ~\cite{grier2013computers},
-but applying this model to tasks like railway engineering may have been prohibitively difficult.
+%\topic{Piecework researchers were also finding that certain types of \textit{organizations},
+%as well as certain types of \textit{work},
+%were more amenable to piecework regimes.}
+%\citeauthor{Brown01041990} wrote that piecework was
+%``less likely in jobs with a variety of duties than in
+%jobs with a narrow set of routinized duties''~\cite{Brown01041990};
+%\citeauthor{SJOE:SJOE371} hypothesized this as an effect of market forces
+%--- suggesting that
+%``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
+%Airy seemed to recognize this,
+%decomposing complex mathematical problems into easily verified components, % ~\cite{grier2013computers},
+%but applying this model to tasks like railway engineering may have been prohibitively difficult.
 
-\topic{But it was more than the inherent qualities of work and the organization of the firm that mattered;
-management practices themselves had to be appropriately geared to working with pieceworkers.}
-Competent foremen were integral both for allocating resources and
-for liaising between workers and upper management~\cite{wray1949marginal,10.2307/2118435}.
+%\topic{But it was more than the inherent qualities of work and the organization of the firm that mattered;
+%management practices themselves had to be appropriately geared to working with pieceworkers.}
+%Competent foremen were integral both for allocating resources and
+%for liaising between workers and upper management~\cite{wray1949marginal,10.2307/2118435}.
 % The West Virginia mines, as \citeauthor{10.2307/2118435} writes,
@@
 
+%}
 % \clearpage
@@
 we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity~\cite{Brown01041990}.}
+the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
 Given the flourishing of on--demand labor platforms such as
@@
 what --- if anything --- has changed.
-We argue that
-the internet has trivialized
-the costs and challenges of the earlier limiting factors
-because technology makes it easier
-\begin{inlinelist}
-  \item for workers to do complex work without training,
-  \item to manage workers in doing complex work, and 
-  \item to create the infrastructure necessary to manage the workers.
-\end{inlinelist}
+
+%We argue that
+%the internet has trivialized
+%the costs and challenges of the earlier limiting factors
+%because technology makes it easier
+%\begin{inlinelist}
+%  \item for workers to do complex work without training,
+%  \item to manage workers in doing complex work, and 
+%  \item to create the infrastructure necessary to manage the workers.
+%\end{inlinelist}
 
@@
 \topic{Algorithms have automated some tasks that previously fell to management.}
-Computational systems hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
-In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
-These algorithms, however, are less able than human managers to manage contingencies that were not programmed into them.
+Computational systems now %hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and 
+act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
+%In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
+These algorithms are less able than human managers to evaluate subjective work.
+So, assessment remains a bottleneck.
 
-\topic{Finally, the organizational limit on infrastructure creation is somewhat lessened.}
-Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
-\citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
-As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use on--demand work.
+%\topic{Finally, the organizational limit on infrastructure creation is somewhat lessened.}
+%Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
+%\citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
+%As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use on--demand work.
 
@@
 \subsubsection{\implication}
-Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
+Algorithms are undoubtably capable of shepherding more complex work than the linear processes available to Airy and Ford.
+However, as work becomes more complex, it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
+So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, it seems likely that this increase is not enough to achieve truly complex work.
+
+However, technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
 Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
@@
 If we can overcome this obstacle,
-we might be able to empower more of these workers to do complex work such as engineering and metalworking,
-rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
+we might be able to empower more of these workers to do complex work such as engineering,
+rather than doom them to ``uneducated'' match--girl reputations~\cite{10.2307/3827491}.
 However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to on--demand work unless it makes common expertise more broadly available.
 
-Will the shift from human managers to Turing--complete algorithms raise the complexity ceiling? 
-By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
-So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
-However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
-So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to benefit from soliciting on--demand work.
-This shift may enable complex goals that were not cost--effective before to become feasible.
-However, because algorithms remain far from replicating all of the foreman's responsibilities,
-most likely is a middle ground in which on--demand work re--introduces
-the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
-This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters~\cite{MaliciousCrowdworkersGadiraju}.
+Evaluation remains as difficult for crowd work as it did for the efficiency experts. 
+Reputation systems for crowdsourcing platforms remain notoriously inflated~\cite{Horton2015a}.
+Creating communities of practice for peer assessment~\cite{crowdguilds} may offer a path forward, but many aspects of assessment remain subjective: whether a logo made for a client is fantastic or terrible may depend on taste.
 
-Finally, the cost of creating piecework infrastructure has dropped. 
-Expensive manufacturing equipment has been largely replaced by computer code~\cite{lessig2006code}.
-As with lowered costs of management, lowered infrastructure costs will make on--demand work accessible to a broader set of people and organizations.
-This in and of itself does not raise the complexity ceiling, but by broadening the potential market for on--demand work, it may enable a new set of goals and needs take part.
+So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes. 
+It will be far more broadly distributed---reaching many more tasks and areas of expertise by virtue of the internet---but not necessarily far more complex than what piecework previously achieved.
+
+%Will the shift from human managers to Turing--complete algorithms raise the complexity ceiling? 
+%By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
+%So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
+%However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
+%So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to benefit from soliciting on--demand work.
+%This shift may enable complex goals that were not cost--effective before to become feasible.
+%However, because algorithms remain far from replicating all of the foreman's responsibilities,
+%most likely is a middle ground in which on--demand work re--introduces
+%the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
+%This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters~\cite{MaliciousCrowdworkersGadiraju}.
+
+%Finally, the cost of creating piecework infrastructure has dropped. 
+%Expensive manufacturing equipment has been largely replaced by computer code~\cite{lessig2006code}.
+%As with lowered costs of management, lowered infrastructure costs will make on--demand work accessible to a broader set of people and organizations.
+%This in and of itself does not raise the complexity ceiling, but by broadening the potential market for on--demand work, it may enable a new set of goals and needs take part.
 

### 2017-01-08 drafts/labor/complexity.tex
complexity negativity

@@
 
-So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes. 
-It will be far more broadly distributed---reaching many more tasks and areas of expertise by virtue of the internet---but not necessarily far more complex than what piecework previously achieved.
+So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than it has today.
+Improvements in workflows, measurement and evaluation have already been made, and it is not clear that the remaining challenges are readily solveable.
+On--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet---but not necessarily far more complex than what piecework previously achieved.
 

### 2017-01-08 drafts/labor/introduction.tex
abstract and intro

@@
 
-\topic{The past decade has seen a flourishing of \textit{on--demand work}
-driven by the reformulation of work as
-constituent parts of larger tasks.}
-This framing of work into modular components
-has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
-Distributed workers then engage in work whenever their schedules allow,
+\topic{The past decade has seen a flourishing of computationally-mediated labor.}
+The framing of work into modular components
+has enabled computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+Distributed workers engage in work whenever their schedules allow,
 often with little to no awareness of the broader context of the work, and
 often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
-In this paper, we use ``on--demand work'' to capture a pair of related phenomena:
-first, \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
-and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
-The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
+
+For years, such labor was limited to information work such as data annotation and surveys~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
+Howver, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+In this paper we will use the term \textit{on--demand work} to capture this pair of related phenomena:
+first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
+and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
+%\topic{As on--demand work has grown, it has given rise to scholarship exploring the domains it occurs in and the people who engage in it.}
+%When \citeauthor{howe2008crowdsourcing} coined crowdsourcing, he defined it in relationship to paid labor:
+%``taking a function once performed by employees and outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
+
+The realization that complex goals can be accomplished by directing  crowds of workers has spurred the technology industry to flock to sites of labor
 such as AMT to explore the limits of this distributed, on--demand workforce.
 Researchers have also taken to the space in earnest,
-developing systems and designs that enable new forms of production
-(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}).
-
-\topic{As on--demand work has grown far beyond information work, it has given rise to scholarship exploring myriad facets of on--demand work.}
-\citeauthor{howe2008crowdsourcing} first described crowdsourcing as
-``outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
-However, for years its instantiation was limited to the utilization of
-human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
-More recently, crowdsourcing of physically embodied work
---- driving and cleaning, for instance ---
-has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
-Some of this research has been motivated by the identification of the sociality of gig work~\cite{crowdcollab},
-and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
-Other work has focused on the outcomes of this frustration,
-reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.
+developing systems that enable new forms of production
+(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}) and pursuing social scientific inquiry into the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
+This research has called out identified the sociality of gig work~\cite{crowdcollab},
+as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
+Others have focused on the outcomes of this frustration,
+reflecting on the resistance that workers express against digitally-mediated labor markets~\cite{uberAlgorithm,dynamo}.
 
@@
 Researchers have offered insights on this question along three major threads:
-First, \nameref{sec:complexity}?
-      Specifically,
-      \begin{inlinelist}[label=(\alph*)]
-        \item How complex are the goals that crowd work can accomplish?, and
-        \item What kinds of goals and industries may eventually utilize it?
-      \end{inlinelist}
+First, \namerefl{sec:complexity}~\cite{crowdForgeKittur}?
+      Specifically, how complex are the goals that crowd work can accomplish, and what kinds of goals and industries may eventually utilize it?
       \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
@@
             Hahn:2016:KAB:2858036.2858364}.
-Second, \nameref{sec:decomposition}
+Second, \namerefl{sec:decomposition}
       \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
@@
             Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
-            Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}.
-And third, \nameref{sec:relationships}?
+            Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
+And third, \namerefl{sec:relationships}
       \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
-            whyWouldAnyoneBrewer,takingAHITMcInnis}.
+            whyWouldAnyoneBrewer,takingAHITMcInnis}?
 % \end{numberlist}
@@
 an instantiation (or even a continuation) of piecework,
-and by interrogating patterns that the corresponding literature predicts
-on this basis, we can
-first, make sense of past events as part of a much larger series of an interrelated phenomenon;
-second, reflect on differences in the factors that impacted piecework historically and their impact on--demand work today;
-and third, to some extent, offer predictions of what on--demand work researchers,
+and by interrogating patterns that the historical literature identifies, we can achieve three goals.
+First, we can make sense of past events as part of a much larger series of an interrelated phenomenon.
+Second, we can reflect on differences in the features that impacted piecework historically and on--demand work today.
+And third, to some extent, we can use these differences to offer some predictions of what on--demand work researchers,
 and workers themselves,
@@
 --- namely, by enabling precise tracking and measurement of workers via algorithms and software.
-By doing so, on--demand labor will find ways to decompose tasks to unprecedented levels.
 
@@
 first, we review the definition and history of piecework
-to make clear the analogy to on--demand work
-(which we'll call \textit{crowd work} hereon
-for consistency with the existing literature);
+to make clear the analogy to on--demand work;
 second, we interrogate the three major research questions above using the lens of piecework. 
-We will identify similarities and differences between piecework as historically understood and
-on--demand work as we know it today;
-third, we will make predictions of future developments based on how those similarities and differences influenced piecework;
-finally, we will offer implications for researchers and practitioners based on our results.
+For each question, we will survey piecework's historical perspective as well as on--demand work's current perspective, call out similarities and differences, and then hazard predictions on what piecework's history might suggest about the answers to these persistent questions for on--demand work.
 

### 2017-01-08 drafts/labor/trackingWork.tex
piecework changes to intro and lit review
lit rev revs interim
notes in the case studies
i just wrote comments indicating the assumed and introduced facts in each of the questions (complexity, decomp, and relationships).
i think a bit of a word cloud is forming in my head, so this should give me something to structure the case studies in the lit review
rough pw perspective
tweaking keywords

@@
 \def\emptyauthor{}
-\def\plainkeywords{Crowdsourcing; on--demand labor; gig work}
-\def\plaingeneralterms{Documentation, Standardization}
+\def\plainkeywords{Crowd work; gig work; on--demand work; piecework}
+\def\plaingeneralterms{Group and Organization Interfaces}
 
@@
 \definecolor{PineGreen}{HTML}{008800}
-\definecolor{BrickRed}{HTML}{FF0000}
+\definecolor{Red}{HTML}{FF0000}
+\definecolor{Blue}{HTML}{0000FF}
 \definecolor{JokeGreen}{HTML}{00C953}
 \newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
-\newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}
+\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
 \newcommand{\TO}{\textit{Turkopticon}}
@@
 \newcommand{\note}[1]{\footnote{#1}}
-\newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
+\newcommand{\topic}[1]{{\color{Blue}#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
 \newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!
@@
 \renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
+
 \begin{document}
@@
 
-\clearpage
-\nobalance{}
-\endnotetext{
-}
-
 \end{document}

### 2017-01-08 drafts/labor/piecework_lit.tex
piecework lit update

@@
 Since this paper traces a much stronger parallel between
-(historical) piecework and (contemporary) on--demand work
-(and, more generally, on--demand labor),
+(historical) piecework and (contemporary) on--demand work,
 a more comprehensive background on piecework will be useful.
-We will more carefully discuss piecework in this section
-to help make our observations and arguments with better familiarity with the topic.
+%We will more carefully discuss piecework in this section
+%to help make our observations and arguments with better familiarity with the topic.
 Specifically,
@@
 
-\subsection{What is piecework? A Primer and Timeline}\label{sec:whatIsPiecework}
+\subsection{What is piecework?: A primer and timeline}\label{sec:whatIsPiecework}
 
@@
 definitions,
-tracing an outline of
-\textit{what piecework is}
-alongside piecework's contemporary developments in practice.
+to develop a working understanding.
 
@@
 the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
-This sense of independence and autonomy appears to be an inherent component of piecework.
+This sense of independence and autonomy appears to be a common component of piecework.
 %that resonates across decades, industries, and locales where piecework is found.
@@
 Thoughts?}
-\topic{The early growth of piecework led to discussion on how best to manage pieceworkers,
-generally regarding workers antagonistically~\cite{norton1900textile,clark1908cotton}.}
-This was a far cry from the earlier rhetoric on piecework, which promised that
-piece workers would gladly work as diligently and as hard as possible because
-incentive--based pay would reward workers directly for hard work~[].
+\topic{Since workers could now choose their own schedule and style, there arose a discussion on how best to manage pieceworkers.}
+This conversation generally regarded workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
+piece workers would gladly work as diligently and as hard as possible because incentive--based pay rewards hard work~\cite{clark1908cotton}.
 
 \topic{Piecework opened the door for people who previously couldn't participate in the labor market to do so, and to acquire job skills incrementally.}
-During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (i.e. male) workers~\cite{hart2013rise}.
-Workers with specific skill subsets could be matched to suitable tasks.
+During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (male) workers~\cite{hart2013rise}.
+%Workers with specific skill subsets could be matched to suitable tasks.
 Women previously had virtually no opportunities
@@
 were principally in clothes--making (e.g. hosiery, footwear, and garments).
-\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
+\citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
+The salient suggestions include:
+\begin{inlinelist}
+\item the emergence of more effective, more nuanced incentive models
+--- rewarding teams for complex achievements, for instance;
+\item the shifting of piecework industries such as manufacturing and textiles to other countries; and
+\item the quality of ``multidimensional'' work becoming too difficult to evaluate
+\end{inlinelist}~\cite{hart2013rise}.
 
@@
 Airy's employment of human computers;
-domestic and farm work;
+domestic and farm workers;
 the ``match--girls'' strike;
-and industrial and assembly-line work.}
+and industrial and assembly-line workers.}
 In introducing these cases at a high level,
@@
 % \onlyinsubfile{\clearpage}
-\subsubsection{Airy's Computers}
+\subsubsection{Airy's computers}
 
@@
 \msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
-\topic{Some of the first systematic cases of what we would recognize as crowd work
-can be found in the study of astronomy.}
-In the \nth{19} century, the calculation of celestial bodies had become
-a competitive field with 
-Airy needed to compute tables that would
-allow sailors to locate themselves by starlight from sea.
+\topic{In the \nth{19} century, the calculation of celestial bodies had become a competitive field, and Airy needed to compute tables that would allow sailors to locate themselves by starlight from sea.}
 This work ostensibly called for educated people who comprehensively understood mathematics.
 Airy realized that he could break the tasks down and delegate the constituent parts
-to ``human computers'' who
-``\dots~possessed the basic skills of mathematics,
+to human computers, or people who could compute basic functions.
+These human computers ``\dots~possessed the basic skills of mathematics,
 including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
@@
 First, the work was designed such that it could be done independently and without collaboration. 
-This enabled geographically-distributed workers to complete the tasks.
-Second, the work was designed so that intermediate results could be quickly verifiable as correct: Airy would have two workers each do the calculation, and a another person compare their answers.
+%This enabled geographically-distributed workers to complete the tasks. 
+Second, the work was designed so that intermediate results could be quickly verified: Airy would have two workers each do the calculation, and another person compare their answers.
 Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.
@@
 
-This practice ensured two outcomes that arguably disfavored workers.
+This practice ensured two outcomes that disfavored workers.
 First, it eliminated any potential to advance professionally, as
-workers' careers in this area ended relatively early in their careers,
+workers' careers ended relatively early in their careers,
 and without formal education in mathematics they struggled to find work for which their experience was meaningful.
 Second, it limited workers' ability to organize
-by ensuring that workers hardly spent sufficient time to successfully rally their peers.
+by ensuring that workers were barely in communication with each other.% hardly spent sufficient time to successfully rally their peers.
 
 % \onlyinsubfile{\clearpage}
-\subsubsection{Domestic and Farmhand Labor}
+\subsubsection{Domestic and farmhand labor}
 
@@
 proved to be a formative period for piecework as we would come to know it.}
-Piecework regimes in farms and in homes engaged workers in assembling clothing, where they were paid by the \textit{piece} of output. %, and
+Piecework regimes in farms and in homes engaged workers in assembling clothing. %, and
 %finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
@@
 %perhaps in part due to the relatively low skillset necessary to complete work.}
-Textile manufacturers found that they could deliver fabric to people at their homes,
-paying them to sew together clothing.
+Textile manufacturers found that they could deliver fabric to people at their homes, asking them to sew together clothing.
 The manufacturers would later return to retrieve the finished garments,
-paying these workers for each \textit{piece} of clothing made. 
+paying these workers for each piece of clothing completed. 
 Farm work applied the idea of piecework by
-paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
+paying workers for tasks like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
 \msb{cutting this next part, it seems irrelevant to the paragraph:}
@@
 
-\subsubsection{The Match--Girls' Strike}
+\subsubsection{The match--girls' strike}
 
@@
 \topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
-much deeper resentment had been boiling over for years.}
+much deeper resentment had been simmering for years.}
 Match--girls were already frustrated with
@@
 poor working conditions, and
-having to work with phosphoric materials (which was proving hazardous).
+having to work with hazardous phosphoric materials.
 \ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}
@@
 %\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history, we should shine a light onto the reputation that had formed for these workers.}
-Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
+Match--girls were the only group in \citeyear{booth1903life} to have formed a trade union,
 according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
-\citeauthor{10.2307/3827491} noted that match girls
+\citeauthor{10.2307/3827491} noted that match--girls
 ``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
@@
 
-\subsubsection{Industrial Workers}
-\topic{Piecework might be most familiar to the HCI researcher in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
+\subsubsection{Industrial workers}
+\topic{Piecework might be most familiar in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
 %In the railroad industry and especially on the assembly line we see many of the mechanisms of work and worker management that later made their way into crowd work.
@@
 %We'll expose these facets of piecework very superficially in the context of industrial work to better speak to these research questions in crowd work later.
-\topic{Before the factory assembly line arose, railway companies adopted piecework regimes in the early \nth{20} century.
+\topic{Before the factory assembly line arose, however, railway companies adopted piecework regimes in the early \nth{20} century.
 \msb{What was pieceworky about this?
@@
 \topic{Piecework %through the \nth{20} century centered around automobile and other mass manufacturing, but
-found its way into the war effort during World War II.}
+found its way into the war effort during World War~II.}
 With the vast majority of men drafted into service,

### 2017-01-08 drafts/labor/complexity.tex
complexity section revisions

@@
 \subsection[What are the complexity limits of on--demand work]
-{Identifying the Complexity Limits of On--Demand Work}\label{sec:complexity}
+{Complexity Limits of On--Demand Work}\label{sec:complexity}
 A key question to the future of on--demand work is
@@
 
-\topic{Crowdsourcing research has spent the better part of a decade
-proving the viability of crowdsourcing in complex work.}
-Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
-the argument runs,
-it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
-\citeauthor{crowdForgeKittur} first opened the question of
-whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
+%\topic{Crowdsourcing research has spent the better part of a decade
+%proving the viability of crowdsourcing in complex work.}
+%Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
+%the argument runs,
+%it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
+\topic{\citeauthor{crowdForgeKittur} first opened the question of
+whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.}
 Their work demonstrated proof--of--concept crowdsourcing of
-a simple encyclopedia article and news summary
+an encyclopedia article and news summary
 --- tasks which could be verified or repeated
 with reasonable expectations of similar outcomes.
-Seeking to raise the complexity ceiling~\cite{myers2000past},
+Seeking to raise the complexity ceiling,
 researchers have since created
@@
 medical data interpreters~\cite{lasecki2013chorus}, and
-idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
-to name a few examples.
+idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
 
 \topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing work--flows.}
-Beginning with a goal that has
-presented significant challenges for computers,
-the researcher leverages an insight from Computer Science
-(for example, MapReduce~\cite{crowdForgeKittur} or
-sequence alignment algorithms~\cite{lasecki2012real})
-and arranges humans as computational black boxes within those approaches.
+The system designer leverages techniques such as MapReduce~\cite{crowdForgeKittur} and
+sequence alignment algorithms~\cite{lasecki2012real}, arranging humans as computational black boxes within.
 This approach has proven a compelling one because
@@
 
-\topic{It is now clear that this computational workflow approach works with focused complex tasks, but
+\topic{It is now clear that this computational workflow approach works with some classes of complex tasks, but
 the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.}
@@
 Airy's human computers and among industrial workers.}
-The majority of piecework's perspective, on this topic, will come from these areas.
+%The majority of piecework's perspective, on this topic, will come from these areas.
 
@@
 it was not used for complex work.
-Without having designed processes to combine work intelligently, piecework managers were restricted to available workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
+Without having designed complex work processes, piecework managers were restricted to available workers' skills such as sewing: it was infeasible to provide new pieceworkers with the comprehensive education
 that apprenticeships imparted \cite{hart2013rise}.
@@
 --- which is inherently heterogeneous and difficult to routinize ---
-would then be unsuitable~\cite{10.2307/23702539}.
-\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
-``efficiency experts'' can effectively gauge how long known tasks should take, but
+was unsuitable~\cite{10.2307/23702539}.
+\citeauthor{Brown01041990}'s description of ``efficiency experts'' would corroborate this;
+efficiency experts can effectively gauge how long known tasks should take, but
 would find themselves overwhelmed if they attempted to assess creative tasks like scientific research,
@@
 the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
-Given the flourishing of on--demand labor platforms such as
-Uber, AMT, and others, we ask ourselves
-what --- if anything --- has changed.
+%Given the flourishing of on--demand labor platforms such as
+%Uber, AMT, and others, we ask ourselves
+%what --- if anything --- has changed.
+However, has computation shifted piecework's previous limits of expertise, measurement, and evaluation?
 
@@
 
-\topic{Technology increases non--experts' levels of expertise by giving access to information that would otherwise be unavailable.}
-For example, taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge''
---- a demonstration of the driver's comprehensive familiarity with the city's roads.
+\topic{In some ways, yes: technology increases non--experts' levels of expertise by giving access to information that would otherwise be unavailable.}
+For example, taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge'': a demonstration of the driver's comprehensive familiarity with the city's roads.
 This test is so challenging that veteran drivers develop significantly larger
 the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,skok2000managing,Woollett1407,woollett2011acquiring}.
-In contrast, with on--demand platforms such as Uber, services such as Google Maps \& Waze make it possible for
+In contrast, with on--demand platforms such as Uber, services such as Google Maps and Waze make it possible for
 people entirely unfamiliar with a city
@@
 %In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
-These algorithms are less able than human managers to evaluate subjective work.
-So, assessment remains a bottleneck.
+However, these algorithms are less able than human managers to evaluate subjective work.
+%So, assessment remains a bottleneck.
 
@@
 However, as work becomes more complex, it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
-So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, it seems likely that this increase is not enough to achieve truly complex work.
+So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, there is a fundamental limit to how complex such work can become.
 
-However, technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
+Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
 Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
@@
 Reputation systems for crowdsourcing platforms remain notoriously inflated~\cite{Horton2015a}.
-Creating communities of practice for peer assessment~\cite{crowdguilds} may offer a path forward, but many aspects of assessment remain subjective: whether a logo made for a client is fantastic or terrible may depend on taste.
+Ultimately, many aspects of assessment remain subjective: whether a logo made for a client is fantastic or terrible may depend on taste.
 
-So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than it has today.
+So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than piecework did.
 Improvements in workflows, measurement and evaluation have already been made, and it is not clear that the remaining challenges are readily solveable.
-On--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet---but not necessarily far more complex than what piecework previously achieved.
+However, on--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet.
 

### 2017-01-08 drafts/labor/discussion.tex
revisions of the discussion and conclusion

@@
 \section{Discussion}
-\topic{In our analysis of on--demand work via the piecework lens, three issues arose:}
+\topic{In our analysis of on--demand work via the piecework lens, three issues arise:}
 \begin{inlinelist}
@@
 Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
-For example, many of the challenges that \DO~overcame in crowd collective action, such as designing for trustworthiness and ensuring anonymity, were relatively unique challenges precipitated by the affordances of the internet.
+%For example, many of the challenges that \DO~overcame in crowd collective action, such as designing for trustworthiness and ensuring anonymity, were relatively unique challenges precipitated by the affordances of the internet.
 For example, unlike physical work environments, people can (and often do) contribute to online communities in a one--off manner
@@
 
-In particular, the predictions that have emerged surrounding crowd work have run the spectrum
-from deep pessimism~\cite{fort2011amazon} to exuberant optimism~\cite{crowdworkFuture}.
-In the next section, we will use
-the piecework foundation which informed our case studies
-to trace out possible dystopian and utopian futures for on--demand labor.
+%In particular, the predictions that have emerged in the literature surrounding the future of crowd work have run the spectrum
+%from deep pessimism~\cite{fort2011amazon} to exuberant optimism~\cite{crowdworkFuture}.
+%In the next section, we will use
+%the piecework foundation which informed our case studies
+%to trace out possible dystopian and utopian futures for on--demand labor.
+
+\msb{Shouldn't we bring up the discussion from the rebuttal about history/future connections here?}
 
@@
 \begin{inlinelist}
-  \item ``\namerefl{sec:complexity}?''
-  \item ``\namerefl{sec:decomposition}?''
-        and
-  \item ``\namerefl{sec:relationships}?''
-\end{inlinelist}.
+  \item \nameref{sec:complexity}?
+  \item \nameref{sec:decomposition}?
+  \item \nameref{sec:relationships}?
+\end{inlinelist}
 

### 2017-01-09 drafts/labor/discussion.tex
table in discussion is maybe done

--- 2017-01-08 drafts/labor/discussion.tex
@@
 \begin{document}
+
+\begin{table*}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{l X X X}
+    \toprule
+    & \textit{Observations in piecework} & \textit{Mechanism} & \textit{Implications} \\
+    \midrule
+    Complexity &
+    \small{Substantial growth in task complexity by reconstituting large tasks into smaller components. Later, innovations in specialized training enabled expert pieceworkers.} &
+    \small{Smaller components could be more easily verified, but this became a limiting factor as well; tasks which are inherently difficult to evaluate are difficult to decompose.} &
+    \small{We've already seen a push on complexity; we may be at the tail end of complexity, or we may yet see revolutionary advances in task complexity through recomposition.} \\ \hline
+
+    Decomposition &
+    \small{Work tasks could be decomposed and measured, but measurement and tracking tools at the time were rudimentary, frustrating workers in myriad industries.} &
+    \small{Scientific Management and Taylorism informed and drove decomposition by measuring and facilitating the optimization of smaller and smaller tasks.} &
+    \small{ % Piecework was limited by technology; on--demand work is enabled by it. 
+    The research has reached an ``atomic'' level of decomposition, exhausting the cognitive limits of the human brain, past which point task--switching becomes a substantial cost.} \\ \hline
+
+    Relationships &
+    \small{Firms antagonized and exploited workers, leading workers to support one another independently, ultimately resulting in strong advocacy groups counterbalancing firms.} &
+    \small{The features of piecework (independence and transience) were both the fulcrum managers used to exploit workers as well as the focal point around which workers bonded.} &
+    \small{On--demand workers will recognize a common pattern of exploitation, coordinate their collective power to form worker advocacy groups.} \\ 
+    \bottomrule
+  \end{tabularx}
+  \caption{First pass at a summary table of findings. Note that all of the cells have the same number of lines. Pretty neat, huh?}
+\end{table*}
+
 \section{Discussion}

### 2017-01-09 drafts/labor/trackingWork.tex
table in discussion is maybe done

@@
 \usepackage[en-GB]{datetime2}
-\usepackage{fancyhdr}
+\usepackage{fancyhdr,tabularx}
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
@@
 \setlist{nolistsep}
-\def\plaintitle{Examining Crowd Work through The Historical Lens of Piecework}
+\def\plaintitle{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
+% \def\plaintitle{Crowd Work and Gig Work [as] Seen Through the Historical Lens of Piecework}
 \def\plainauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi}
 \def\emptyauthor{}
-\def\plainkeywords{Crowd work; gig work; on--demand work; piecework}
-\def\plaingeneralterms{Group and Organization Interfaces}
+\def\plainkeywords{Crowdsourcing; on--demand labor; gig work}
+\def\plaingeneralterms{Documentation, Standardization}
 
@@
 \newcommand{\subsubsubsection}[1]{\textit{#1}}
-\newcommand{\pieceworkpers}{Piecework's perspective}
-\newcommand{\crowdworkpers}{Crowd work's perspective}
+\newcommand{\pieceworkpers}{The perspective of piecework}
+\newcommand{\crowdworkpers}{The perspective of on--demand work}
 \newcommand{\whatchanged}{Comparing the phenomena}
-\newcommand{\implication}{Implications for crowd work}
+\newcommand{\implication}{Implications for on--demand work}
 \newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of

### 2017-01-09 drafts/labor/abstract.tex

@@
 payment that was popular at the turn of the \nth{20} century ---
-to understand how these questions might play out with modern crowd work.
+to understand how these questions might play out with modern on--demand work.
 We identify the mechanisms that enabled and limited piecework historically, and
-identify whether crowd work faces the same pitfalls or might differentiate itself.
+identify whether on--demand work faces the same pitfalls or might differentiate itself.
 This approach introduces theoretical grounding that can help address

### 2017-01-09 drafts/labor/cases.tex

@@
 {Research in crowdsourcing has spent the better part of a decade
-exploring how to grow the limits of crowdsourcing,
-finding the boundaries of crowd work and microtasks.}
+exploring how to grow the limits of crowdsourcing.}
 This has largely involved iteratively
-identifying barriers on complexity and
-overcoming them through novel designs of work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
-The question then
-has become \textit{whether} there are limits to crowdsourcing,
+identifying barriers to high-quality, complex work, then
+overcoming them through novel designs of systems, work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
+The question 
+has become \textit{whether} there are limits to on--demand work,
 and if so, what factors determine them.
@@
 
-The exploration of on--demand work's potential and limits has principally looked at
-manipulating and extending along three dimensions:
-First, \namerefl{sec:complexity}.
-Second, \namerefl{sec:decomposition}.
-And third, \namerefl{sec:relationships}.
-We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
-decomposed,  contextually abstracted, and
-made more resilient to attrition of various forms.
-We'll also point to corresponding piecework literature addressing these aspects.
-% \msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
-Finally, we'll discuss how these elements will serve
-to constrain the upper and lower bounds of crowdsourcing as it relates
-to the question of the furthest limits of crowdsourcing.
+The exploration of on--demand work's potential and limits has principally 
+interrogated three dimensions:
+First, \namerefl{sec:complexity}?
+Second, \namerefl{sec:decomposition}?
+And third, \namerefl{sec:relationships}?
+We'll explore these aspects of on--demand work by connecting to corresponding piecework literature
+and comparing its lessons to the current state of on--demand work.
 

### 2017-01-09 drafts/labor/decomposition.tex

@@
 As such,
-one of the central questions in the literature is how to design these microtasks,
-and which kinds of tasks are amenable to decomposition.
+one of the central questions in the literature is how finely-sliced these microtasks can become, and which kinds of tasks are amenable to decomposition.
 In this section,
@@
 an error~\cite{bernsteinSoylent},
-or a concept~\cite{Yu2016a,Yu2016b} and a second sequence of tasks to execute work on that area. 
+or a concept~\cite{Yu2016a,Yu2016b}) and a second sequence of tasks to execute work on that area. 
 This decomposition style has been successfully applied across many areas,
@@
 Research has sought to ameliorate this issue by
-designing workflows help workers ``act with global understanding when
+designing workflows to help workers ``act with global understanding when
 each contributor only has access to local views''~\cite{verroios2014context},
@@
 ``streak'' to perform many tasks of that same type~\cite{taskSearch}.
-and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}),
-
+%and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}),
 Researchers have reacted by designing task recommendation systems~\cite[e.g.][]{Cosley:2007:SUI:1216295.1216309}
-and others focused on
-minimizing the amount of time that people need to spend doing anything other than
+and minimizing the amount of time that people need to spend doing anything other than
 the work for which they are paid~\cite{callison2014crowd}.
@@
 The fourth and final stage was narrow expertise training. 
-As World War II raged and there was a dearth of skilled workers, managers trained women just enough to be able to complete their tasks~\cite{hart2013rise}.
+Even after work is decomposed and measured, there may not be enough workers available to do it. 
+So, as World War II raged and there was a dearth of skilled workers, managers trained women just enough to be able to complete their tasks~\cite{hart2013rise}. 
 Over time, these women could gain proficiency and gain broader expertise.
@@
 preventing scientific management from being fully utilized,
-no longer exists
 modern crowd work is measuring and modeling every click,
@@
 \topic{A second shift is  
-the relative ease with which the metaphorical ``assembly line'' can be changed.}
+the relative ease with which the metaphorical ``assembly line'' can be experimented with and measured.}
 Historical manufacturing equipment could not quickly be assembled,
@@
 and
-other factors in piecework that would have been
-all but impossible to manipulate as few as 20 years ago~\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
+other factors accelerating scientific management that would have been
+all but impossible as few as 20 years ago~\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.
 
@@
 First, as with piecework, on--demand work began by decomposing tasks so that anyone could complete them, as with data labeling on Amazon Mechanical Turk.
-Second, we began to modularize and measure existing expertise (e.g., software engineering, design) so that it could be brought into crowdsourcing systems~\cite{foundry,Chen2016}.
+Second, we began to modularize and measure external expertise (e.g., software engineering, design) so that it could be brought into crowdsourcing systems~\cite{foundry,Chen2016}.
 Third, we used measurement to mathematically optimize workers' behaviors so that we could make the systems more efficient~\cite{weld2010decision}.

### 2017-01-09 drafts/labor/relationships.tex

@@
 
-HCI and CSCW have framed themselves around supporting work rather than becoming an infrastructural layer enabling it.
-While all artifacts have politics, this shift into computational labor systems has directly impacted the lives and livelihood of workers.
+HCI and CSCW have historically framed themselves around supporting work.
+While all artifacts have politics, the recent shift into computational labor systems has directly impacted the lives and livelihood of workers in new ways.
 So, it is important to understand: what will the future look like for the workers who use these systems?
@@
 
-One of the initial questions that researchers asked was, who are the crowd workers and what draws them to crowd work?
+Who are the crowd workers and what draws them to crowd work?
 Early literature emphasized motivations like fun and spare change, but this narrative soon shifted to emphasize that many workers use platforms such as Amazon Mechanical Turk as a primary source of income~\cite{kaufmann2011more,ipeirotis2010demographics,Antin2012a}.
@@
 Thus, those who choose to opt out of the traditional labor force and spend significant time on Mechanical Turk are especially motivated by the opportunity for autonomy and skill variety~\cite{kaufmann2011more}.
-Due to valuing autonomy, it is tempting to ascribe attitudes of ``pity the workers'' to Turkers, but this frame is increasingly rejected by workers and designers as patronizing~\cite{storiesIraniSilberman}.
+While some carry attitudes of ``pity the workers'' toward Turkers, this frame is increasingly rejected by workers and designers as patronizing~\cite{storiesIraniSilberman}.
 
 Workers' relationships with requesters are fraught.
+The unbridled power that requesters have over workers, and
+the resultant frustration that this generates,
+has motivated research into
+the tense relationships between workers and requesters~\cite{fixingChaos,dynamo}.
 Workers are often blamed for any low-quality work, regardless of whether they are responsible~\cite{martin2014being,takingAHITMcInnis}.
@@
 Turkers acting collectively to bring about changes to their circumstances~\cite{dynamo}.
-This unbridled power that requesters have over workers and
-the resultant stress and frustration that this generates
-has been part of the undercurrent of research into
-the tense relationships between workers and requesters~\cite{fixingChaos,dynamo}.
 
@@
 However, the frustration and mistrust that workers experience with requesters does occasionally boil over on the forums. 
-This behavior has come to be known as ``mega--drama'' amongst such workers~\cite{dynamo}.
-Still, the study of these communities is made challenging because most of these platforms do not themselves include social affordances for workers~\cite{miller2011understanding}.
+%This behavior has come to be known as ``mega--drama'' amongst such workers~\cite{dynamo}.
+%Still, the study of these communities is made challenging because most of these platforms do not themselves include social affordances for workers~\cite{miller2011understanding}.
 
@@
 when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.
-Workers, \citeauthor{roy1954efficiency} found, formed even more starkly antagonistic relationships with their managers.
+Piece workers, \citeauthor{roy1954efficiency} found, would form even more starkly antagonistic relationships with their managers.
 % Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
 
-\topic{Soon workers began resisting piecework regimes and the methods that came along with it.}
+\topic{Soon workers began resisting piecework regimes.}
 The match--girls engaged in their famous strike of 1888, particularly pushing to abolish the fines that were taken out of their wages.
-Soon others followed suit, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}
+Soon others followed suit, including women in the garment industry in Philadelphia who established collective bargaining rights~\cite{10.2307/41829256} and national coal miners who effected an individual minimum wage in 1912~\cite{10.2307/2221944}.
 
-Soon, many worker organizations were weighing in on
+Many worker organizations began weighing in on
 (or, more precisely, against)
@@
 
-\topic{There's a noticeable dearth of information on the inter--personal relationships among pieceworkers
+\topic{Relative to the modern on--demand workers, there is a noticeable dearth of information on the inter--personal relationships among pieceworkers
 beyond the match--girls at the end of the \nth{19} century.}
@@
 
-While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', there was generally less written about quality concerns for historical pieceworkers than there is in modern on--demand work.
+%While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', 
+There was generally less written about work quality concerns for historical pieceworkers than there is in modern on--demand work. 
 Why the difference? 
 One possibility is that, through writing web scripts and applying them to many tasks, it is possible for a small number of spammers have an outsized influence.
-Historically, it was much harder for such workers to move and get new jobs --- today, they can simply accept a different task on Mechanical Turk.
-Another possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen, online workers are known by little more than an account ID.
+Another possibility: it was much harder for piece workers to move to a new location and get a new job --- today, workers can simply accept a different task on Mechanical Turk.
+A third possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen and known to them, online workers are known by little more than an account ID.
 
@@
 crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
-Both pieceworkers and on--demand workers have struggled at times to form a collective identity necessary to organize.
-With workers joining and leaving the crowd labor force continuously, and with many part-time members, it is extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
-However, even when they can: whereas pieceworkers could physically block access to a site of production, online labor markets provide no facilities for workers to change the experience of other workers.
+%Both pieceworkers and on--demand workers have struggled at times to form a collective identity necessary to organize.
+With workers joining and leaving the on--demand labor force continuously, and with many part-time members, it is extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
+However, even when they can, enforcement remains a challenge: whereas pieceworkers could physically block access to a site of production, online labor markets provide no facilities for workers to change the experience of other workers.
 This is a key limitation --- without it, workers cannot enforce a strike.
@@
 more considerate \textit{human} management may resolve
-many of the tensions we've discovered among among crowd workers.
+many of the tensions.
 
-\topic{Reciprocally, crowd work may be able to inform piecework research.}
-There exists far less literature about pieceworkers' relationships than there does today about on--demand workers' relationships.
-Research into on--demand labor benefits from both the accessibility of digital platforms as well as the firmer theoretical basis of Anthropology than
-existed at the turn of the \nth{20} century, when piecework began to emerge.
+\topic{Reciprocally, crowd work may be able to inform piecework research in this domain.}
+There exists far less literature about piece workers' relationships than there does today about on--demand workers' relationships.
+Two reasons stand out: first, modern platforms are visible to researchers in ways that the sites of piece work labor were not.
+Second, Anthropology stands on a firmer theoretical and methodological basis than it did at the turn of the \nth{20} century.
 \citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and

### 2017-01-09 drafts/labor/introduction.tex
table v2

@@
 
+\begin{table*}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{l X X X}
+    \toprule
+    & \textit{Observations in piecework} & \textit{Mechanism} & \textit{Implications for On--demand Work} \\
+    \midrule
+    Complexity &
+    \small{Growth from simple tasks such as sewing to more complex composite outcomes on the assembly line floor.} &
+    \small{Complexity was limited by what could be easily measured and evaluated for payment by the piece.} &
+    \small{Measurement and verification remain challenges and will limit complexity unless solved.} \\ \hline
+
+    Decomposition &
+    \small{Work began sliced such that non-experts could perfom each piece, but over time was sliced such that non-overlapping expertise was required for each step.} &
+    \small{Scientific Management and Taylorism informed and drove decomposition by measuring and facilitating the optimization of smaller tasks.} &
+    \small{After scientific management matured, piecework began  specialized training to create experts in narrow tasks. A similar shift seems feasible with on--demand work.} \\ \hline
+    %The research has reached an ``atomic'' level of decomposition, exhausting the cognitive limits of the human brain, past which point task--switching becomes a substantial cost.} \\ \hline
+
+    Workers &
+    \small{Firms antagonized and exploited workers, leading workers to support one another independently, ultimately resulting in strong advocacy groups counterbalancing firms.} &
+    \small{The features of piecework (independence and transience) were both the fulcrum managers used to exploit workers as well as the focal point around which workers bonded.} &
+    \small{While worker frustrations are similar, the decentralized nature of on--demand work will limit collective action until there exist platforms to coordinate and exert pressure.} \\ 
+    \bottomrule
+  \end{tabularx}
+  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely-sliced tasks can become, and what the workplace looks like for workers.
+We connect piecework's history (left) to the mechansims that impacted it (center) to derive predictions for modern on--demand work (right).
+}
+\end{table*}
+
 \subsection{Piecework as a lens to understand crowdsourcing}

### 2017-01-09 drafts/labor/complexity.tex
deleting stuff

@@
 \subsubsection{\crowdworkpers}
-
-%\topic{Crowdsourcing research has spent the better part of a decade
-%proving the viability of crowdsourcing in complex work.}
-%Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
-%the argument runs,
-%it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
 \topic{\citeauthor{crowdForgeKittur} first opened the question of
@@
 
-% \clearpage
 \subsubsection{\pieceworkpers}
@@
 
-\ali{What do we know?
-
-Airy's computers:
-\begin{itemize}
-  \item we can do some pretty complex stuff if we break tasks down into quickly verifiable parts
-  \item there's no need to train people to be experts for this work; non--experts can do the job just as well.
-\end{itemize}
-
-% Domestic and farm work
-% \begin{itemize}
-%   \item ???
-% \end{itemize}
-
-% match--girls
-% \begin{itemize}
-%   \item ???
-% \end{itemize}
-
-Industrial workers
-\begin{itemize}
-  \item revolutionary: we can train people on a narrow slice of the work and have them just do that work
-\end{itemize}
-}
-
-\msb{ summarizing the later conclusion so we can support it here:
-complexity is determined by whether we can encode it into workflows
-complexity is determined by whether we can modularize and measure it
-complexity is determined by how much management overhead there is
-complexity is determined by the infrastructure that can support it
-}
-
 \topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
 Airy's human computers and among industrial workers.}
-%The majority of piecework's perspective, on this topic, will come from these areas.
 
@@
 
-%\topic{The next innovation in piecework's drive toward complexity came on
-%the assembly line, specifically during the war effort.}
-%At a time when manufacturing firms needed more able--bodied workers, but
-%all of the conventionally trained metalworkers had joined the armed forces,
-%factories turned to women~\cite{honey1985creating,davies2014origins}.
-%This need to train workers rapidly to do very narrowly defined tasks resulted in
-%a \textit{recomposition} of work, wherein
-%workers were only prepared for the tasks they would do;
-%riveters would only be trained on how to do rivet and
-%to pass the output to the next person on the assembly line,
-%who could evaluate and send the work back or proceed with the next step.
-
-%\ali{OR\dots
-
-%\topic{Only tasks that could be measured and priced could be completed via piecework
-%--- a fact which fundamentally limited the potential complexity of piecework.}
-%Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
-%one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
-%complex, creative work
-%--- which is inherently heterogeneous and difficult to routinize ---
-%would be unsuitable for piecework~\cite{10.2307/23702539}.
-%\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
-%``efficiency experts'' can effectively gauge how long known tasks should take, but
-%would find themselves overwhelmed if they attempted to assess creative work like design,
-%which can take an arbitrary number of iterations before proceeding to a subsequent step.
-
 \topic{Moreover, piecework was limited to tasks that could be quickly and accurately evaluated.}
-%For example, the roles required to facilitate piecework
-%in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
 \citeauthor{hart2016rise} argue that evaluation limited piecework's complexity:
@@
 Complex work, which is often subjective to evaluate, falls victim to this criteria.
-%``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
 
-% \topic{This focus on measurement and tracking had consequences.}
-% \citeauthor{10.2307/23702539} suggests that the first sparks of scientific management
-% could be found in piecework:
-% the approach of paying workers for each piece of output necessitated
-% the rigorous tracking, measurement, and training of workers
-% for which scientific management became famous~\cite{10.2307/23702539}.
-% If true, 
-% the concurrent upswing of
-% scientific management and Fordism
-% through the first two--thirds of the \nth{20} century
-% alongside piecework was not only understandable, but predictable~\cite{hart2013rise}.
-
-%\topic{Piecework researchers were also finding that certain types of \textit{organizations},
-%as well as certain types of \textit{work},
-%were more amenable to piecework regimes.}
-%\citeauthor{Brown01041990} wrote that piecework was
-%``less likely in jobs with a variety of duties than in
-%jobs with a narrow set of routinized duties''~\cite{Brown01041990};
-%\citeauthor{SJOE:SJOE371} hypothesized this as an effect of market forces
-%--- suggesting that
-%``pay schemes based on~\dots~performance may induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
-%Airy seemed to recognize this,
-%decomposing complex mathematical problems into easily verified components, % ~\cite{grier2013computers},
-%but applying this model to tasks like railway engineering may have been prohibitively difficult.
-
-%\topic{But it was more than the inherent qualities of work and the organization of the firm that mattered;
-%management practices themselves had to be appropriately geared to working with pieceworkers.}
-%Competent foremen were integral both for allocating resources and
-%for liaising between workers and upper management~\cite{wray1949marginal,10.2307/2118435}.
-% The West Virginia mines, as \citeauthor{10.2307/2118435} writes,
-% ~\cite{10.2307/2118435}.
-
-% \topic{Piecework researchers also argued that,
-% in addition to constraints on the kind of \textit{work} that's amenable to piecework,
-% only certain kinds of \textit{organizations} were amenable to piecework.}
-% Researchers detail three organizational criteria.
-% First, \citeauthor{Brown01041990} found that
-% piecework was ``less likely in jobs with a variety of duties than in
-% jobs with a narrow set of routinized duties''~\cite{Brown01041990}.
-% \citeauthor{SJOE:SJOE371} pointed out this phenomenon as a market effect:
-% ``in an environment with multi--tasking,
-% pay schemes based on tightly specified performance may
-% induce workers to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371}.
-% Second, complexity was limited by access to capital to create the necessary infrastructure.
-% As \citeauthor{10.2307/23702539} reported,
-% only the largest and most wealthy railroads had the resources
-% necessary to change existing systems to piecework regimes~\cite{10.2307/23702539}.
-% Third, organizations required capable managers in charge of the pieceworkers.
-% The West Virginia mines, for example, hired foremen 
-% to be the intermediary between upper management and the workers~\cite{10.2307/2118435}.
-% These foremen were responsible for allocating resources and
-% understanding when and how to modify work as necessary~\cite{wray1949marginal}.
-% So, in sum,
-% organizations historically could only take advantage of piecework if they had
-% homogeneous work to be done,
-% access to capital to purchase the necessary equipment, and
-% the ability to hire people who could serve as intermediaries between pieceworkers and management.}
-
-% \topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
-% particularly because maximizing the advantages of piecework seemed to reward
-% smaller, more constrained, more narrowly--trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
-% For most of the \nth{19} century,
-% piecework was applied almost exclusively to farm and textile work.
-% Work was simple and widely understood
-% --- farm workers didn't need to be trained on how to plow fields, or birth foals;
-% seamstresses knew how to sew together denim~\cite{10.2307/2338394,riisOtherSideLives}.
-
-%}
-% \clearpage
 \subsubsection{\whatchanged}
@@
 the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
-%Given the flourishing of on--demand labor platforms such as
-%Uber, AMT, and others, we ask ourselves
-%what --- if anything --- has changed.
 However, has computation shifted piecework's previous limits of expertise, measurement, and evaluation?
-
-%We argue that
-%the internet has trivialized
-%the costs and challenges of the earlier limiting factors
-%because technology makes it easier
-%\begin{inlinelist}
-%  \item for workers to do complex work without training,
-%  \item to manage workers in doing complex work, and 
-%  \item to create the infrastructure necessary to manage the workers.
-%\end{inlinelist}
 
@@
 \topic{Algorithms have automated some tasks that previously fell to management.}
-Computational systems now %hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and 
-act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
-%In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
+Computational systems now act as ``piecework clerks''~\cite{10.2307/23702539} to
+inspect,
+modify and
+combine work~\cite{turkopticon,takingAHITMcInnis}.
 However, these algorithms are less able than human managers to evaluate subjective work.
-%So, assessment remains a bottleneck.
 
-%\topic{Finally, the organizational limit on infrastructure creation is somewhat lessened.}
-%Writing web scripts takes fewer people and fewer hours than creating physical equipment for piecework.
-%\citeauthor{turkitLittle}'s vision was that any user with basic programming skills could tap into on--demand human intelligence.
-%As better toolkits lower this threshold~\cite{myers2000past} and computational thinking diffuses, a broader population will be able to use on--demand work.
-
-% \clearpage
 \subsubsection{\implication}
@@
 Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
-of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
-To some extent, an argument can be made that
+of human computation~\cite{grier2013computers} and
+drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
+An argument might be made that
 MOOCs and other online education resources
@@
 
-%Will the shift from human managers to Turing--complete algorithms raise the complexity ceiling? 
-%By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
-%So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
-%However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
-%So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to benefit from soliciting on--demand work.
-%This shift may enable complex goals that were not cost--effective before to become feasible.
-%However, because algorithms remain far from replicating all of the foreman's responsibilities,
-%most likely is a middle ground in which on--demand work re--introduces
-%the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
-%This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters~\cite{MaliciousCrowdworkersGadiraju}.
-
-%Finally, the cost of creating piecework infrastructure has dropped. 
-%Expensive manufacturing equipment has been largely replaced by computer code~\cite{lessig2006code}.
-%As with lowered costs of management, lowered infrastructure costs will make on--demand work accessible to a broader set of people and organizations.
-%This in and of itself does not raise the complexity ceiling, but by broadening the potential market for on--demand work, it may enable a new set of goals and needs take part.
-
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-% }
-
 \end{document}

### 2017-01-09 drafts/labor/discussion.tex
deleting stuff

--- 2017-01-08 drafts/labor/discussion.tex
@@
 \begin{document}
+
 \section{Discussion}
@@
 Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
-%For example, many of the challenges that \DO~overcame in crowd collective action, such as designing for trustworthiness and ensuring anonymity, were relatively unique challenges precipitated by the affordances of the internet.
 For example, unlike physical work environments, people can (and often do) contribute to online communities in a one--off manner
@@
 While we have attempted to understand the likely overlaps and differences between history and modern day, no analysis is perfect.
-% articulated in the discussion about
-% organizing people and the challenges therein
-% \cite{dynamo}, as well as for
-% many of the rhizomatic features \citeauthor{miller2011understanding}
-% discusses
-% \cite{miller2011understanding}).
 
@@
 While we can only speculate one of (perhaps many) possible futures, history does allow us to articulate and bound which futures appear more likely.
-
-%In particular, the predictions that have emerged in the literature surrounding the future of crowd work have run the spectrum
-%from deep pessimism~\cite{fort2011amazon} to exuberant optimism~\cite{crowdworkFuture}.
-%In the next section, we will use
-%the piecework foundation which informed our case studies
-%to trace out possible dystopian and utopian futures for on--demand labor.
 
@@
 
-% \topic{There is, of course, a kernel of truth in everything here.}
 \topic{Piecework teaches us that, without appropriate norms and policy, the dystopian outcome has happened and will happen again.}
@@
 the same trajectory of worker empowerment that piecework \textit{later} found.
-
-% Arguably the internet has made collective action on an unprecedented scale much easier.
 
@@
 experimentally \cite{delayAndOrderLasecki}.
-We should consider whether this remains a worthwhile area to explore; %  unless
-% we're actually finding ways to minimize the many forms of costs of switching tasks.
+We should consider whether this remains a worthwhile area to explore;
 unless the work we put forth directly affects the costs of task--switching

### 2017-01-09 drafts/labor/introduction.tex
deleting stuff

@@
 
-% \onlyinsubfile{
-%   \hypersetup{
-%     citecolor  = blue,
-%     linkcolor  = JokeGreen,
-%   }
-% }
-% \renewcommand{\topic}[1]{#1}
-
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
-
 \begin{document}
@@
 \section{Introduction}\label{sec:introduction}
-\msb{Please take a full pass through for consistent terminology.
-In the intro, we align on on--demand work and call it the union of crowd work and gig work.
-But then, in the title and the research questions, everything is crowd work, crowd work, crowd work.
-I suggest either using crowd work in the introduction and explaining that you're using that to mean both traditional crowd work as well as modern gig work, or using on--demand work in the title and in the research questions.
-I see the point ``(which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature)'', but it feels like a cop out.
-Which one do we want to use, really?
-Let's pick one and be consistent throughout.}
-
-\ali{\textit{On--demand work} is more ``correct'', but I'm worried that crowdsourcing researchers will skim it and feel there's nothing relevant to them. Is that giving them too little credit?}
-
-\msb{I think this is a question of which community you want to be influencing.
-Do you want people who study crowdsourcing to be the primary readers+citers of this paper?
-Or people who study Uber and algorithms at work?
-One way to decide this is to name five people who you really hope will read and like this paper, and note which community they are a part of.}
 
@@
 and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
-%\topic{As on--demand work has grown, it has given rise to scholarship exploring the domains it occurs in and the people who engage in it.}
-%When \citeauthor{howe2008crowdsourcing} coined crowdsourcing, he defined it in relationship to paid labor:
-%``taking a function once performed by employees and outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
 
@@
             whyWouldAnyoneBrewer,takingAHITMcInnis}?
-% \end{numberlist}
 
@@
     \midrule
-    Complexity &
+    {Complexity} &
     \small{Growth from simple tasks such as sewing to more complex composite outcomes on the assembly line floor.} &
@@
 
-    Decomposition &
+    {Decomposition} &
     \small{Work began sliced such that non-experts could perfom each piece, but over time was sliced such that non-overlapping expertise was required for each step.} &
@@
     \small{After scientific management matured, piecework began  specialized training to create experts in narrow tasks. A similar shift seems feasible with on--demand work.} \\ \hline
-    %The research has reached an ``atomic'' level of decomposition, exhausting the cognitive limits of the human brain, past which point task--switching becomes a substantial cost.} \\ \hline
 
-    Workers &
+    {Workers} &
     \small{Firms antagonized and exploited workers, leading workers to support one another independently, ultimately resulting in strong advocacy groups counterbalancing firms.} &
@@
 
-\onlyinsubfile{
-  \clearpage
-  \balance{}
-  \printbibliography
-}
-
 \end{document}

### 2017-01-09 drafts/labor/piecework_lit.tex
deleting stuff

@@
 
-% \onlyinsubfile{\clearpage}
 \section{A Review of Piecework}
@@
 a more comprehensive background on piecework will be useful.
-%We will more carefully discuss piecework in this section
-%to help make our observations and arguments with better familiarity with the topic.
 Specifically,
@@
 a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
-\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
-\ali{I'm not sure; I'll look for some data.
-
-update: hard to find anything saying ``piecework represented eleventy billion dollars of the U.S. economy'', which is what I'd like to say,
-but I can say something like ``at its height, $k\%$ of garment workers were pieceworkers'' or something.
-that seems a bit wishy--washy though; I'd like to say something like
-``$k\%$ of the \textit{entire U.S. workforce} was working under some sort of piecework regime''.}
 
@@
 in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
-% (where he also calls it ``measure work'', ``grate work'', and ``task work''),
 particularly driven toward encapsulating the manual labor of farmwork.}
@@
 This sense of independence and autonomy appears to be a common component of piecework.
-%that resonates across decades, industries, and locales where piecework is found.
 
-\ali{I think now that this is moved to this area there's a good opportunity to frame this more as
-``now that workers could choose their own schedule, style, etc\dots,
-newfound interest in how to manage these workers emerged''.
-Thoughts?}
 \topic{Since workers could now choose their own schedule and style, there arose a discussion on how best to manage pieceworkers.}
 This conversation generally regarded workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
-piece workers would gladly work as diligently and as hard as possible because incentive--based pay rewards hard work~\cite{clark1908cotton}.
+piece workers would gladly work as diligently and as hard as possible because
+incentive--based pay rewards hard work~\cite{clark1908cotton}.
+\ali{\citeauthor{norton1900textile} talks about guarding against collusion among workers, hence the instinct to include it, but
+with \citeauthor{clark1908cotton}'s paper coming out in \citeyear{clark1908cotton} and \citeauthor{norton1900textile} in \citeyear{norton1900textile}, I guess the timing looks weird.}
 
@@
 During World War II, women received training in narrow subsets of more comprehensive jobs, enabling work in capacities similar to conventional (male) workers~\cite{hart2013rise}.
-%Workers with specific skill subsets could be matched to suitable tasks.
 Women previously had virtually no opportunities
@@
 
-\msb{If you want to make the argument in the last paragraph of this subsection that piecework assumed no professional skills, it's worth calling that out slightly more visibly throughout the subsection.}\ali{I don't want to suggest that, actually, at least not ultimately; in the beginning, sure, but by WWII it definitely required certain specialized training (arguably in Airy's case too)}
-
 Viewing on--demand work as a modern instantiation of piecework is relatively straightforward by this definition.
@@
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{Airy's computers}
@@
 
-\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
 \topic{In the \nth{19} century, the calculation of celestial bodies had become a competitive field, and Airy needed to compute tables that would allow sailors to locate themselves by starlight from sea.}
@@
 Airy also instituted a policy of firing his computers once they reached age 23.
-%\ali{this has a ``gosh isn't that interesting?'' feeling to it.
-%should i just leave it out until later when there's something to say about it
-%(namely, that the mode of work was initially designed to stymie professional growth)?}%
 
@@
 First, the work was designed such that it could be done independently and without collaboration. 
-%This enabled geographically-distributed workers to complete the tasks. 
 Second, the work was designed so that intermediate results could be quickly verified: Airy would have two workers each do the calculation, and another person compare their answers.
 Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.
-%\ali{This should go later someplace, but not clear where:
 
@@
 Second, it limited workers' ability to organize
-by ensuring that workers were barely in communication with each other.% hardly spent sufficient time to successfully rally their peers.
+by ensuring that workers were barely in communication with each other.
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{Domestic and farmhand labor}
@@
 Piecework regimes in farms and in homes engaged workers in assembling clothing. %, and
-%finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
-%\topic{Farm and field work emerged early on as a rich field for the use of piecework,
-%perhaps in part due to the relatively low skillset necessary to complete work.}
 Textile manufacturers found that they could deliver fabric to people at their homes, asking them to sew together clothing.
@@
 paying workers for tasks like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
-\msb{cutting this next part, it seems irrelevant to the paragraph:}
-%\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
-%the only factor important in piecework was
-%whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
-%Piecework's advantage,
-%\citeauthor{hughRaynbirdTaskWork} argues,
-%was this hands--off management of workers.
 
@@
 
-%\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
-%as well as the early years of worker collective organization and action.}
 \topic{Match--makers were some of the first workers in mass manufacturing
@@
 having to work with hazardous phosphoric materials.
-\ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}
 
@@
 
-%\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history, we should shine a light onto the reputation that had formed for these workers.}
 Match--girls were the only group in \citeyear{booth1903life} to have formed a trade union,
@@
 
-\ali{I want to discuss a few things here:
-\begin{itemize}
-  \item there were protests about something sort of like surge pricing in crop--picking, but i don't know how to make that a paragraph (maybe merge it with match--girls?)
-\end{itemize}}
-
 \subsubsection{Industrial workers}
 \topic{Piecework might be most familiar in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
-%In the railroad industry and especially on the assembly line we see many of the mechanisms of work and worker management that later made their way into crowd work.
-%Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
-%We'll expose these facets of piecework very superficially in the context of industrial work to better speak to these research questions in crowd work later.
-\topic{Before the factory assembly line arose, however, railway companies adopted piecework regimes in the early \nth{20} century.
-\msb{What was pieceworky about this?
-They were paid per meter of railroad that they laid down?}
+Before the factory assembly line arose, however, railway companies adopted piecework regimes in the early \nth{20} century.
 What followed was a flourishing of piecework management practices,
 as railway companies worked to find effective ways
-to motivate and evaluate this skilled workforce of engineers.}
+to motivate and evaluate this skilled workforce of engineers.
 \citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
@@
 in the early \nth{20} century
---- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}. \topic{This oversight, while controversial
+--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.
+This oversight, while controversial
 (especially among workers~\cite{american1921problem}),
-paved the way for piecework to grow substantially.}
+paved the way for piecework to grow substantially.
 
@@
 
-\topic{Piecework %through the \nth{20} century centered around automobile and other mass manufacturing, but
-found its way into the war effort during World War~II.}
+\topic{Piecework found its way into the war effort during World War~II.}
 With the vast majority of men drafted into service,
@@
 these women were trained for individual tasks
-% --- such as riveting component to another ---
-and assigned to that task. %, decomposed from a more comprehensive task.
+and assigned to that task.
 ``Rosie the Riveter'',
@@
 was a pieceworker~\cite{davies2014origins}.
-
-% \msb{confusing timing; now we're going backwards?}
-
-% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
-% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
-
-\ali{some things I didn't cover:
-\begin{itemize}
-  \item research finding that task variability is bad for piecework, but worker variability is good
-  \item talk about the foreman
-  \item worker advocacy groups
-\end{itemize}
-should I be making a serious effort to get that stuff in here? this section is kind of long already
-(and it'll definitely be much longer than the other case studies).}
 
@@
 
-\onlyinsubfile{
-  \printbibliography
-  }
-
 \end{document}

### 2017-01-09 drafts/labor/discussion.tex
added rosenberg refs in discussion

@@
 \subsection{The Hazards of Predicting the Future}\label{sec:perilousProblemsPredicting}
-The past can't be a perfect predictor for the future;
-as \citeauthor{scholz2012digital} points out,
-``it would be wrong to conclude that
-in the realm of digital labor
-there is nothing new under the sun''
-\cite{scholz2012digital}.
+The past isn't a perfect predictor for the future;
+as \citeauthor{scholz2012digital} cautions,
+``it would be wrong to conclude that in the realm of digital labor there is nothing new under the sun''~\cite{scholz2012digital}.
 Our analysis is limited by the differences, foreseen and unforeseen, between historical piecework and modern on--demand work.
-For example, unlike physical work environments, people can (and often do) contribute to online communities in a one--off manner
-\cite{mcinnis2016one}.
-The internet makes this kind of loose affiliation feasible.
-While we have attempted to understand the likely overlaps and differences between history and modern day, no analysis is perfect.
+For example, unlike physical work environments, people can (and often do) make one--off contributions to online communities~\cite{mcinnis2016one}.
+While we have attempted to identify all of the likely parallels and divergences between piecework and on--demand work,
+we can't claim to have accounted for everything.
 
-But this does not mean that
-attempting to draw meaningfully from historical scholarship would be folly;
+But this does not mean that attempting to draw meaningfully from historical scholarship would be folly;
 enough of piecework can and does inform on--demand work that
@@
 
-\msb{Shouldn't we bring up the discussion from the rebuttal about history/future connections here?}
+\citeauthor{rosenberg1994exploring} and others have contributed substantially to the practice in part by clearly limiting the extent of their claims
+--- only offering, for instance,
+``to narrow our estimates and thus to concentrate resources in directions that are more likely to have useful payoffs''~\cite{rosenberg1994exploring}.
+Using this approach,
+our method of relating history to modern socio--technical systems may be
+a useful tool for researchers attempting
+to make sense of ostensibly new phenomena.
+In other words, offering ``that past history is an indispensable source of information
+to anyone interested in characterizing technologies''~\cite{rosenberg1982inside}.
+
+% \msb{Shouldn't we bring up the discussion from the rebuttal about history/future connections here?}
 
@@
 
-\topic{A uniquely challenging facet of this topic of inquiry is the public attention that this domain has attracted.}
+\topic{A uniquely challenging facet of this domain is the public attention that it has garnered.}
 Activists have described speculative work as having
 ``essentially been turned into modern--day slaves''~\cite{activistsHuffPoLawsuit}.
-Meanwhile, advocates describe it as
-``a project of sharing
-aimed at providing ordinary people
-with more economic opportunities and
-improving their lives''~\cite{uberPropaganda}.
+Meanwhile, advocates have described it as
+``a project of sharing aimed at providing ordinary people with more economic opportunities and improving their lives''~\cite{uberPropaganda}.
 
-\topic{Piecework teaches us that, without appropriate norms and policy, the dystopian outcome has happened and will happen again.}
+\topic{Piecework teaches us that, without appropriate norms and policies, the dystopian outcome has happened and may happen again.}
 The piecework nature of on--demand work induces us
-``to neglect tasks that are less easy to measure'' \cite{SJOE:SJOE371},
+``to neglect tasks that are less easy to measure''~\cite{SJOE:SJOE371},
 rewarding us not for creativity but predictability;
 payment for this work may ultimately be determined by
-an algorithm that fundamentally doesn't understand people;
-the layers between us and our managers might increasingly become
-``defective (simple, observable)'' algorithms \cite{10.2307/2555446},
+algorithms that fundamentally don't understand people;
+the layers between us and our managers might eventually become
+``defective (simple, observable)'' algorithms~\cite{10.2307/2555446},
 just like those which already frustrate
-on--demand workers
-\cite{uberAlgorithm,dynamo,turkopticon}.
-However, social policy has advanced since the early 1900s, so as on--demand work gains popularity a repeat of \textit{How the Other Half Lives}~\cite{riisOtherSideLives} seems less likely.
+on--demand workers~\cite{uberAlgorithm,dynamo,turkopticon}.
+However, social policy has advanced since the early 1900s, so as on--demand work grows, a repeat of \textit{How the Other Half Lives}~\cite{riisOtherSideLives} seems less likely.
 
 \topic{On the other hand, while piecework's nascent years were grim, 
-they precipitated a century of extremely strong labor advocacy~\cite{hart2013rise,mccallum2013global}.}
-Even today, the geist that came out of the labor union revolution
+they precipitated a century of some of the most potent labor advocacy organizations in modern history~\cite{hart2013rise,mccallum2013global}.}
+Even today, the \textit{geist} of the labor union revolution
 inspires collective action and worker empowerment around the world:
-in India, workers across the nation recently engaged in
+Recently, in India, workers across the nation engaged in
 the largest labor strike in human history
---- perhaps as many as 150 million
 \cite{indiaStrikeRealNews}.
-If labor advocacy groups can find ways to permeate on--demand labor markets as some have called for
+If labor advocacy groups can find ways to effect change in on--demand work as some have called for
 \cite{futureUnions},
 then the future of on--demand labor may follow
-the same trajectory of worker empowerment that piecework \textit{later} found.
+the same trajectory of worker empowerment that piecework saw.
 

### 2017-01-10 drafts/labor/abstract.tex
chinmay's suggestions integrated

@@
 }
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
+
 \begin{document}
@@
 \begin{abstract}
+\ali{I want to do a substantial revision of this.
+We haven't touched it much throughout the whole paper's revision process, and
+now I'm not so sure that this abstract is
+as clear or focused on precisely what the paper contributes.}
+
 The internet is empowering the rise of
@@
 This approach introduces theoretical grounding that can help address
-some of the most pernicious questions in crowd work,
+some of the most persistent questions in crowd work,
 and suggests design interventions that learn from history

### 2017-01-10 drafts/labor/cases.tex
chinmay's suggestions integrated

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
+
 \begin{document}
@@
 
-{Research in crowdsourcing has spent the better part of a decade
+\topic{Research in crowdsourcing has spent the better part of a decade
 exploring how to grow the limits of crowdsourcing.}
@@
 }
+
 \onlyinsubfile{

### 2017-01-10 drafts/labor/complexity.tex
chinmay's suggestions integrated

@@
 \topic{It is now clear that this computational workflow approach works with some classes of complex tasks, but
-the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.}
+the broader wicked problems largely remain unsolved.}
 As a first example,
@@
 \subsubsection{\pieceworkpers}
-
 \begin{comment}

### 2017-01-10 drafts/labor/conclusion.tex
revisions of the discussion and conclusion
table in discussion is maybe done
chinmay's suggestions integrated

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
 
@@
 \section{Conclusion}
-On--demand work is not new, but a contemporary instantiations of piecework.
+\topic{On--demand work is not new, but a contemporary instantiations of piecework.}
 In this paper, we reconsider three major research questions in on--demand work using the lens of piecework:
 \begin{inlinelist}
-  \item ``\namerefl{sec:complexity}?'';
-  \item ``\namerefl{sec:decomposition}?''; and
-  \item ``\namerefl{sec:relationships}?''
+  \item \nameref{sec:complexity}?
+  \item \nameref{sec:decomposition}?
+  \item \nameref{sec:relationships}?
 \end{inlinelist}

### 2017-01-10 drafts/labor/decomposition.tex
apparently a typo I forgot to check in
deleting stuff
chinmay's suggestions integrated

@@
 \documentclass[trackingWork]{subfiles}
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
-% \onlyinsubfile{
-% \usepackage{xr--hyper}
-% \usepackage{hyperref}
-% \externaldocument{complexity}
-% \externaldocument{relationships}
-% % \externaldocument{decomposition}
-% }
 \begin{document}
@@
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{\crowdworkpers}
@@
 the invisible labor of \textit{finding} tasks~\cite{martin2014being} has arisen as a major issue.}
-\citeauthor{taskSearch} illustrate the task search challenges on AMT.
+\citeauthor{taskSearch} illustrate the task search challenges on AMT~\cite{taskSearch}.
 Workers seek out good requesters~\cite{martin2014being} and then 
 ``streak'' to perform many tasks of that same type~\cite{taskSearch}.
-%and some work has gone into ameliorating the problems specific to this work site (\textit{ReLauncher}),
-Researchers have reacted by designing task recommendation systems~\cite[e.g.][]{Cosley:2007:SUI:1216295.1216309}
+Researchers have reacted by designing task recommendation systems (e.g.~\cite{Cosley:2007:SUI:1216295.1216309})
 and minimizing the amount of time that people need to spend doing anything other than
@@
 
-\onlyinsubfile{\clearpage}
 \subsubsection{\pieceworkpers}
@@
 Rather than hire expert computers, Airy identified ways to break down astrological calculations into steps that could be completed with only a basic knowledge of mathematics.
-%\topic{Decomposability is core to piecework: one cannot pay by the piece unless the pieces are small and discretizable.}
 Likewise, \citeauthor{Brown01041990} argued that piecework arose in industries where there was significant task homogeneity and low fixed costs of machinery and training~\cite{Brown01041990}.
-%In other words, if tasks could be clearly decomposed and defined, and if the firm could afford the training and machinery for the pieceworkers, piecework tended to be adopted.
 
@@
 With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism}),
+(and Henry Ford's eponymously named \textit{Fordism})~,
 piecework in the early and mid--\nth{20} century surged, especially in industrial work.
@@
 
-%\topic{Piecework researchers enumerate a number of problems with the decomposition of work,
-%and the conflicting pressures managers and workers put forth.}
-%\citeauthor{bewley1999wages} in particular points out that
-%the approach of paying workers by the piece is
-%``\dots~not practical for workers doing many tasks,
-%because of
-%the cost of establishing the rates and because
-%piecework does not compensate workers for time spent switching tasks''.
-%Ultimately, \citeauthor{bewley1999wages} argues that
-%``[piecework is] infeasible,
-%because~\dots~total output is the joint product of varying groups of people''~\cite{bewley1999wages}.
-%To \citeauthor{bewley1999wages}, the main limiter was the ability to define and measure the work.
-
-\onlyinsubfile{\clearpage}
 \subsubsection{\whatchanged}
@@
 Where \citeauthor{10.2307/23702539} directly implicates measurement as
-preventing scientific management from being fully utilized,
+preventing scientific management from being fully utilized~\cite{10.2307/23702539},
 modern crowd work is measuring and modeling every click,
@@
 
-%Third,
-%modern crowd work has sliced work to such small scales that the marginal activities
-%--- things like finding work and cognitive task switching ---
-%have become large relative to the tasks themselves~\cite{taskSearch}.
-%In the historical case of piecework,
-%moving metallurgical tools,
-%mining equipment,
-%or
-%other industry materials would have been prohibitively difficult and slow;
-%workers were encouraged to specialize in a single set of tasks,
-%allowing pieceworkers to sequence their tasks optimally on their own~\cite{hart2013rise}.
-%The result is that on--demand workers are more free agents than historically was the case.
-%However,
-%because they spend significant time searching for tasks,
-%the piece rate is less a good estimate of take--home earnings than before.
-
-\onlyinsubfile{\clearpage}
 \subsubsection{\implication}
@@
 
-The fourth stage, then, appears likely to occur: narrow training of workers for expert tasks. 
+The fourth stage, then, appears likely to occur: narrow training of workers for these decomposed expert tasks. 
 There is demand for skilled workers in many crowdsourcing tasks, and systems to help train workers~\cite{suzukiAtelier}. 
@@
 The limit is no longer measurement precision, but human cognition. 
-%as \citeauthor{10.2307/23702539} argues,
-%then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
-%Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved,
-%or near enough that they no longer limit task decomposition.
-%Now,
-%not just tasks,
-%but entire workers' histories~\cite{hata2017glimpse},
-%can be collected and analyzed in detail.
 Task switching and other cognitive costs make it difficult
@@
 logo design requires a deep understanding of the client and their goals.
-%In part due to this limitation,
-%99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.
-
-%So,
-%ultimately,
-%the levels of decomposition are likely to follow the contours of context required.
-%Low--context work will be extremely highly decomposed.
-%High--context work will continue to be limited.
-
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-% }
 

### 2017-01-10 drafts/labor/discussion.tex
chinmay's suggestions integrated

@@
 to anyone interested in characterizing technologies''~\cite{rosenberg1982inside}.
-
-% \msb{Shouldn't we bring up the discussion from the rebuttal about history/future connections here?}
 
@@
 piecework is saddled by a lower limit on decomposition:
-``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}.
+``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}. % Chinmay wanted me to specifically call out that this is from chapter 6. Is that necessary...?
+\ali{Chinmay: <suggestion to look at \cite{stigler1962information} and friction> --- not sure if worth getting into.}
 We've since studied this phenomenon in crowd work to great length both
@@
 
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
-% }
+\ali{
+Discussions from Slack:
+
+Chinmay: I think the most provocative claim is ``The history of piecework suggests that the utopian and dystopian outcomes will both occur, in different parts of the world and to different people.''\\
+seems contrary to the idea that marketplaces are global\\
+I would also like some more thought into the Research Agenda section\\
+
+Ali: so the dystopian and utopian claim is partly influenced by the fact that piecework seemed to affect some industries differently than others\\
+
+Chinmay: ah ok\\
+i think you could claim the same about online piecework\\
+differnt industries\\
+
+Ali: garment-makers were pretty thoroughly marginalized, but industrial metalwork led to this whole WWII phenomenon with expert training\\
+mmmm\\
+but now it's not just about the industry, but about the design of the labor market\\
+we can influence the ``architecture'' in a way that we couldn't before\\
+this is all lessig and stuff\\
+
+Chinmay: oh\\
+i understand\\
+hmm something there\\
+might poke more :)
+
+\dotfill
+
+Chinmay> i wish you had a little more criticality in taylorism but I'm a downer :)
+
+Ali> actually a reviewer wanted more on taylorism but i was afraid that we just didn't have the space to really give anyone into taylorism what they want
+
+Chinmay> essentially that would be a great place to talk about how those three questions you mentioned interact
+
+Ali> oh
+that might be worth fitting in
+i mean we inexplicably got it down to 9.5 pages or so
+or 9.75
+we were over for like 99% of the drafting and then inexplicably something just imploded and we got like half a page
+so yeah i'll try and write something giving taylorism more space and using that to talk about the confluence of these three questions
+}
 

### 2017-01-10 drafts/labor/piecework_lit.tex
chinmay's suggestions integrated

@@
 elsewhere in manufacturing~\cite{10.2307/3827491}.
-Fordism and scientific management thrust piecework into higher gear, especially as
+At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
+--- a mechanical engineer with a fascination in work efficiency ---
+began studying and formalizing the decomposition, tracking, and management of tasks.
+In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
+concretizing an idea that had nebulously been forming,
+and which he had been working out himself, for years~\cite{taylor1914principles}.
+Scientific Management (and Fordism) thrust piecework into higher gear, especially as
 mass manufacturing and
 a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
+
+\ali{Somewhere here(?) I want to talk about Gibreth's works~\cite{gilbreth1911motion,gilbreth1911motion}. Should I save it to be the saving grace in the discussion?}
 
@@
 particularly driven toward encapsulating the manual labor of farmwork.}
-\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
+He does this by contrasting with the ``day--labourer'' ---
 ``the chief difference lies between the day--labourer,
@@
 \citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
---- young pieceworkers, mostly women, who assembled matchsticks in the late \nth{19} century.
+--- teenage women who assembled matchsticks in the late \nth{19} century in London.
+% added that these women were in London since Chinmay critiqued that we're glossing over the labor advocacy movement that grew in Europe, which revealed that we completely left out the fact that this all happened in England.
+\ali{Chinmay: but still not convinced; this is a post hoc ergo propter hoc argument}
 Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
@@
 finding similar notions of independence and autonomy among piecework newspaper compositors:
-``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night,
-the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
+``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night, the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
 This sense of independence and autonomy appears to be a common component of piecework.
@@
 incentive--based pay rewards hard work~\cite{clark1908cotton}.
-\ali{\citeauthor{norton1900textile} talks about guarding against collusion among workers, hence the instinct to include it, but
-with \citeauthor{clark1908cotton}'s paper coming out in \citeyear{clark1908cotton} and \citeauthor{norton1900textile} in \citeyear{norton1900textile}, I guess the timing looks weird.}
 
@@
 but if he works by the piece he will work as many hours as it is possible for him to stand''~\cite{clark1908cotton}.
-
 \end{comment}
@@
 the reality was more grim, as
-\citeauthor{riisOtherSideLives} documented in ``\citetitle{riisOtherSideLives}'' in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
+\citeauthor{riisOtherSideLives} documented in \textit{How the Other Half Lives} in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
 He found that

### 2017-01-10 drafts/labor/relationships.tex
chinmay's suggestions integrated

@@
 HCI and CSCW have historically framed themselves around supporting work.
-While all artifacts have politics, the recent shift into computational labor systems has directly impacted the lives and livelihood of workers in new ways.
-So, it is important to understand: what will the future look like for the workers who use these systems?
+While all artifacts have politics,
+the recent shift into computational labor systems has directly impacted
+the lives and livelihood of workers in new ways.
+So, it is important to understand:
+what will the future look like for the workers who use these systems?
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{\crowdworkpers}
@@
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{\pieceworkpers}
@@
 Piece workers, \citeauthor{roy1954efficiency} found, would form even more starkly antagonistic relationships with their managers.
-% Workers' situations were quite dire: \citeauthor{riisOtherSideLives}  documented abhorrent working and living conditions of pieceworkers in New York City~\cite{riisOtherSideLives}.
 
@@
 beyond the match--girls at the end of the \nth{19} century.}
-% For one thing, the research methods we typically associate with the exploratory study of cultures
-% --- Anthropology, and namely participant--observation, ethnography, etc\dots~---
-% didn't exist quite as we know them at the turn of the \nth{20} century, and wouldn't for several more decades.
 Nevertheless, we can offer some observations:
@@
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{\whatchanged}
@@
 
-%While historical pieceworkers could be looked down on, as the match-stick girls were characterized by ``brashness, irregularity, low morality, and little education'', 
 There was generally less written about work quality concerns for historical pieceworkers than there is in modern on--demand work. 
@@
 Where a person might recognize that the ``attention check'' questions
-proposed by \citeauthor{le2010ensuring} ensure that malicious and inattentive workers are stopped,
+proposed by \citeauthor{le2010ensuring} and others ensure that
+malicious and inattentive workers are stopped~\cite{le2010ensuring,AAAIW113995},
 some implementations of these approaches
-only seem to antagonize workers.
+only seem to antagonize workers~\cite{takingAHITMcInnis}.
 More than 30 years ago, \citeauthor{10.2307/2555446} wrote:
@@
 
-% \onlyinsubfile{\clearpage}
 \subsubsection{\implication}
-
 The decentralization and anonymization of on--demand work, especially online crowd work, will continue to make many of its social relationships a struggle.
@@
 Second, Anthropology stands on a firmer theoretical and methodological basis than it did at the turn of the \nth{20} century.
-\citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and
+Malinowski, Boas, Mead, and
+% \citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and % this was naming Boas twice, so.
 other luminaries throughout the first half of the \nth{20} century

### 2017-01-11 drafts/labor/cases.tex
many minor changes through complexity

@@
 \topic{Research in crowdsourcing has spent the better part of a decade
-exploring how to grow the limits of crowdsourcing.}
+exploring how to grow its limits.}
 This has largely involved iteratively
-identifying barriers to high-quality, complex work, then
-overcoming them through novel designs of systems, work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
+identifying barriers to high--quality, complex work, then
+overcoming them through novel designs of
+systems, work--flows, and processes
+(e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
 The question 
@@
 
-The exploration of on--demand work's potential and limits has principally 
-interrogated three dimensions:
+The exploration of on--demand labor's potential and limits has principally 
+navigated three dimensions:
 First, \namerefl{sec:complexity}?
@@
 And third, \namerefl{sec:relationships}?
-We'll explore these aspects of on--demand work by connecting to corresponding piecework literature
-and comparing its lessons to the current state of on--demand work.
+We'll explore these aspects of on--demand labor by connecting to corresponding piecework literature
+and comparing its lessons to the current state of on--demand labor.
 

### 2017-01-11 drafts/labor/complexity.tex
many minor changes through complexity

@@
 
-\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing work--flows.}
+\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new work--flows.}
 The system designer leverages techniques such as MapReduce~\cite{crowdForgeKittur} and
@@
 This approach has proven a compelling one because
-it leverages the in--built advantages of
+it leverages the inherent advantages of
 scale,
@@
 \topic{Airy's work on astronomical charts opened the door to greater task complexity by encoding the intelligence into the process rather than the people.}
-Airy's computers had relatively limited education in mathematics, but by combining individual simple mathematical operations, Airy was able to create a complex composite outcome~\cite{grier2013computers}.
+Airy's computers had relatively limited education in mathematics, but by combining simple mathematical operations, Airy was able to create a complex composite outcome~\cite{grier2013computers}.
 Likewise, in Ford's factories, no individual could build the entire car, but the process could emergently produce one.
@@
 Complexity levels remained low at the turn of
-the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
+the \nth{20} century as piecework saturated densely populated urban areas
+such as London and New York City~\cite{riisOtherSideLives}.
 
@@
 an over--emphasis on quantity~\dots~and an under--emphasis on quality''~\cite{hart2016rise}.
-Complex work, which is often subjective to evaluate, falls victim to this criteria.
+Complex work, which is often subjective to evaluate, falls victim to this pitfall.
 
@@
 \topic{The research on piecework tells us that
-we should expect piecework to thrive in industries where
-the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
-However, has computation shifted piecework's previous limits of expertise, measurement, and evaluation?
+we should expect it to thrive in industries where
+the nature of the work is limited in complexity~\cite{Brown01041990}, and
+become less common as work becomes more complex.}
+Has computation shifted piecework's previous limits of
+expertise, measurement, and evaluation?
 
@@
 By augmenting the human intellect~\cite{engelbart2001augmenting},
-computing has shifted the complexity of work that is possible without training.
+computing has shifted the complexity of work that is possible with minimal or no training.
 
@@
 combine work~\cite{turkopticon,takingAHITMcInnis}.
-However, these algorithms are less able than human managers to evaluate subjective work.
+However, these algorithms are less able than human managers to evaluate subjective work, and
+their ability to exercise discretion.
 
 \subsubsection{\implication}
-Algorithms are undoubtably capable of shepherding more complex work than the linear processes available to Airy and Ford.
-However, as work becomes more complex, it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
+Algorithms are undoubtedly capable of shepherding more complex work than the linear processes available to Airy and Ford.
+However, as work becomes more complex,
+it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
 So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, there is a fundamental limit to how complex such work can become.

### 2017-01-11 drafts/labor/introduction.tex
many minor changes through complexity

@@
 
-\topic{The past decade has seen a flourishing of computationally-mediated labor.}
+\topic{The past decade has seen a flourishing of computationally--mediated labor.}
 The framing of work into modular components
@@
 For years, such labor was limited to information work such as data annotation and surveys~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
-Howver, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-In this paper we will use the term \textit{on--demand work} to capture this pair of related phenomena:
+However, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
+In this paper we will use the term \textit{on--demand labor}, as \citeauthor to capture this pair of related phenomena:
 first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
-and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
+and second, \textit{gig work}~\cite{friedman2014workers}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
+\ali{I'm a little worried about people offering contradicting terms, but the argument I \textit{want} to make is that
+these terms (``gig work'' and ``crowd work'') were sort of born out of marketing and business interests (just like ``sharing economy''),
+and as a result they're somewhat fuzzy and imprecise (for example, see~\cite{de2015crowdsourcing}); what we're trying to do is
+place these terms where they're most accurate and tightly scoped, so that their meaningfulness is retained;
+then we use  ``on--demand work'' to capture this work more holistically,
+much more carefully selecting the terminology for academic discourse.
 
-The realization that complex goals can be accomplished by directing  crowds of workers has spurred the technology industry to flock to sites of labor
+Too off--course?}
+
+The realization that complex goals can be accomplished by directing crowds of workers has spurred industry to flock to sites of labor
 such as AMT to explore the limits of this distributed, on--demand workforce.
@@
 as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
-Others have focused on the outcomes of this frustration,
-reflecting on the resistance that workers express against digitally-mediated labor markets~\cite{uberAlgorithm,dynamo}.
+Others have focused on the \textit{outcomes} of this frustration,
+reflecting on the resistance that workers express against digitally--mediated labor markets~\cite{uberAlgorithm,dynamo}.
 
-\topic{This body of research has broadly sought to answer one central question:
-What does the future hold for on--demand work and those who do it?}
+\topic{This body of research has broadly worked toward the answer to one central question:
+\textit{What does the future hold for on--demand work and those who do it?}}
 Researchers have offered insights on this question along three major threads:
-First, \namerefl{sec:complexity}~\cite{crowdForgeKittur}?
-      Specifically, how complex are the goals that crowd work can accomplish, and what kinds of goals and industries may eventually utilize it?
-      \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
-            Nebeling:2016:WCW:2858036.2858169,
-            Hahn:2016:KAB:2858036.2858364}.
-Second, \namerefl{sec:decomposition}
-      \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
-            LykourentzouPersonalityMatters,KucherbaevReLauncher,
-            Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
-            Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
-And third, \namerefl{sec:relationships}
-      \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
-            whyWouldAnyoneBrewer,takingAHITMcInnis}?
+First, \namerefl{sec:complexity} ---
+specifically, how complex are the goals that crowd work can accomplish, and
+what kinds of goals and industries may eventually utilize it~\cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364,crowdForgeKittur}?
+Second, \namerefl{sec:decomposition}~\cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
+And third, \namerefl{sec:relationships}~\cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}?
 
 This research has largely sought to answer these questions by examining extant on--demand work phenomena.
-So far, it has not offered a framing for holistically explaining
+So far, it has not offered an ontology to describe or understand
 the developments in worker processes that researchers have developed, or
@@
 nor has any research
-gone so far as to directly predict future developments.
+gone so far as to anticipate future developments.
 
@@
     {Decomposition} &
-    \small{Work began sliced such that non-experts could perfom each piece, but over time was sliced such that non-overlapping expertise was required for each step.} &
+    \small{Work began sliced such that non--experts could perform each piece, but over time was sliced such that non--overlapping expertise was required for each step.} &
     \small{Scientific Management and Taylorism informed and drove decomposition by measuring and facilitating the optimization of smaller tasks.} &
@@
   \end{tabularx}
-  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely-sliced tasks can become, and what the workplace looks like for workers.
-We connect piecework's history (left) to the mechansims that impacted it (center) to derive predictions for modern on--demand work (right).
+  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely--sliced tasks can become, and what the workplace looks like for workers.
+We connect piecework's history (left) to the mechanisms that impacted it (center) to derive predictions for modern on--demand work (right).
 }
@@
 a work and payment structure which breaks tasks down into discrete jobs,
-wherein payment is made for work \textit{output}, rather than for \textit{time}.
+wherein payment is made for \textit{output}, rather than for \textit{time}.
 Piecework use as a lens on on--demand work is not new.
-\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
-referenced crowd work as piecework briefly
+In \citeyear{crowdworkFuture}, for example,
+\citeauthor{crowdworkFuture} referenced crowd work as piecework briefly
 as a loose analogy~\cite{crowdworkFuture}.
 But more than this,
-the framing of on--demand labor as a re--instantiation of piecework
-gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirical research.
+the framing of on--demand labor as a reinstantiation of piecework % google said reinstantion is a word
+gives us years of historical material to make sense of this new form of work, and
+allows us to reflect on--demand work through a theoretical lens, informed by years of rigorous, empirical research.
 
-More concretely, by looking at on--demand work as
+More concretely, by looking at on--demand labor as
 an instantiation (or even a continuation) of piecework,
 and by interrogating patterns that the historical literature identifies, we can achieve three goals.
-First, we can make sense of past events as part of a much larger series of an interrelated phenomenon.
+First, we can make sense of past events as part of a much larger series of an interrelated phenomena.
 Second, we can reflect on differences in the features that impacted piecework historically and on--demand work today.
-And third, to some extent, we can use these differences to offer some predictions of what on--demand work researchers,
-and workers themselves,
+And third, to some extent, we can use these differences to offer some predictions of what on--demand labor researchers,
+and the workers,
 might expect to see on the horizon.
@@
 leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
---- namely, by enabling precise tracking and measurement of workers via algorithms and software.
+--- namely, by enabling precise tracking and measurement via algorithms and software.
 
@@
 to make clear the analogy to on--demand work;
-second, we interrogate the three major research questions above using the lens of piecework. 
-For each question, we will survey piecework's historical perspective as well as on--demand work's current perspective, call out similarities and differences, and then hazard predictions on what piecework's history might suggest about the answers to these persistent questions for on--demand work.
+and second, we interrogate the three major research questions above using the lens of piecework. 
+For each question, we will
+contrast the perspective the piecework scholarship offers with on--demand labor's body of research,
+identify similarities and differences, and then
+offer predictions for on--demand work.
+
+% \ali{
+% The pervasiveness of rating and reviewing mechanisms is a chief issue in the analysis ofthe gig-economy, one that spans practically all the forms of platform-work.
+De Stefano also makes the argument for analyzing jointly the two main types of work in this realm, "crowdwork" and "work-on-demand via app." In crowdwork, workers complete job tasks through online platforms, such as Amazon Mechanical Turk, Crowdflower, and Clickworker, which connect an indefinite number of organizations and individuals through the internet.
+In "work-on-demand via app," traditional working activities like transport, cleaning, and running errands are channeled through apps managed by firms like Uber, Taskrabbit, and Handy, that also variously intervene in setting quality standards of service and in the selection and management of the workforce.}
 

### 2017-01-11 drafts/labor/piecework_lit.tex
many minor changes through complexity

@@
 researchers have generally made this allusion in passing.}
-Since this paper traces a much stronger parallel between
+Since we trace a much stronger parallel between
 (historical) piecework and (contemporary) on--demand work,
@@
 definitions,
-to develop a working understanding.
+to develop a sense of piecework.
 
@@
 to calculate equations in order
-to better--predict the trajectories of various celestial bodies in the 1830s~\cite{grier2013computers}.
-George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
+to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
+George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
 by breaking complex calculations into constituent parts, and
@@
 but it found its foothold in manual labor.}
-Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
+Piecework took hold in farm work~\cite{hughRaynbirdTaskWork}, % hold foothold
 in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
@@
 particularly driven toward encapsulating the manual labor of farmwork.}
-He does this by contrasting with the ``day--labourer'' ---
+He does this by contrasting two paradigms ---
 ``the chief difference lies between the day--labourer,
@@
 a base rate for time plus
-additional pay dependent on output~\cite{rowan1901premium}.
+additional pay depending on output~\cite{rowan1901premium}.
 As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
@@
 ``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night, the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
-This sense of independence and autonomy appears to be a common component of piecework.
+This sense of independence and autonomy appears to be a common thread of piecework.
 
 \topic{Since workers could now choose their own schedule and style, there arose a discussion on how best to manage pieceworkers.}
-This conversation generally regarded workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
-piece workers would gladly work as diligently and as hard as possible because
+This conversation came to regard workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
+piece workers would gladly work as diligently for as long as possible because
 incentive--based pay rewards hard work~\cite{clark1908cotton}.
@@
 were principally in clothes--making (e.g. hosiery, footwear, and garments).
-\citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
+\citeauthor{hart2013rise} offer a number of explanations for the sudden demise of piecework.
+% \ali{Should I add ``(especially in the United States but more broadly in the developed world)'' or something?
+Since a lot of developing nations still have sweatshops and piecework to some extent or another?}
 The salient suggestions include:
@@
 \item the shifting of piecework industries such as manufacturing and textiles to other countries; and
-\item the quality of ``multidimensional'' work becoming too difficult to evaluate
-\end{inlinelist}~\cite{hart2013rise}.
+\item the quality of ``multidimensional'' work, which was too difficult to evaluate~\cite{hart2013rise}.
+\end{inlinelist}
 
@@
 Viewing on--demand work as a modern instantiation of piecework is relatively straightforward by this definition.
-First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
+First, platforms such as Amazon Mechanical Turk (AMT), Uber, Upwork, and TaskRabbit pay by the task, though some mix systems in similar ways to the Rowan system's combination of piece rate and time rate pay.
 Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
@@
 we'll trace the history of piecework
-while also framing the later analysis of the major research threads we named earlier: complexity, decomposition, and relationships.
+while also framing the later analysis of the leading research threads we named earlier:
+complexity, decomposition, and relationships.
 
@@
 First, it eliminated any potential to advance professionally, as
-workers' careers ended relatively early in their careers,
+workers' careers ended relatively quickly,
 and without formal education in mathematics they struggled to find work for which their experience was meaningful.
 Second, it limited workers' ability to organize
-by ensuring that workers were barely in communication with each other.
+by ensuring that workers were in relatively little communication with each other.
 
@@
 
-Match--girls were the only group in \citeyear{booth1903life} to have formed a trade union,
-according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
+Match--girls were some of the earliest to have formed a trade union,
+according to \citeauthor{booth1903life}'s account in \citeyear{booth1903life}. % ~\cite{booth1903life}.
 \citeauthor{10.2307/3827491} noted that match--girls
@@
 But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
-These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
+These were workers who treasured their independence, but also fiercely protected one another. % , contributing to the common good.
 Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
@@
 \topic{Piecework might be most familiar in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
-Before the factory assembly line arose, however, railway companies adopted piecework regimes in the early \nth{20} century.
-What followed was a flourishing of piecework management practices,
+Before the factory assembly line arose, however, railway companies adopted piecework regimes at the turn of the \nth{20} century.
+What followed was a flourishing of management practices,
 as railway companies worked to find effective ways
@@
 to determine pay for each task at the company informed by
-``thousands of individual operations''; % ~\cite{10.2307/23702539}.
+``thousands of individual operations'';
 \citeauthor{10.2307/23702539} goes on to list
@@
 --- and more broadly the first half of the \nth{20} century ---
-as the ``heyday'' of the use of piecework.
+as the ``heyday'' of piecework.
 They attribute this to the shortage of male workers,
@@
 factories found themselves turning to
-a predominantly female workforce that had neither
+a mostly female workforce that had neither
 the formal training nor
@@
 these women were trained for individual tasks
-and assigned to that task.
+and correspondingly assigned to that or a similar task.
 ``Rosie the Riveter'',
@@
 represented empowerment and opportunity for women~\cite{honey1985creating},
-was a pieceworker~\cite{davies2014origins}.
+would have been a pieceworker~\cite{davies2014origins}. \ali{was? would have been?}
 

### 2017-01-11 drafts/labor/discussion.tex
michael readthrough

@@
 For example, unlike physical work environments, people can (and often do) make one--off contributions to online communities~\cite{mcinnis2016one}.
-While we have attempted to identify all of the likely parallels and divergences between piecework and on--demand work,
+While we have attempted to identify some likely parallels and divergences between piecework and on--demand work,
 we can't claim to have accounted for everything.
@@
 Even today, the \textit{geist} of the labor union revolution
-inspires collective action and worker empowerment around the world:
+inspires collective action and worker empowerment around the world.
 Recently, in India, workers across the nation engaged in

### 2017-01-11 drafts/labor/introduction.tex
michael readthrough

@@
 \topic{The past decade has seen a flourishing of computationally--mediated labor.}
-The framing of work into modular components
-has enabled computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
-Distributed workers engage in work whenever their schedules allow,
+A framing of work into modular, pre-defined components
+enables computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
+In this regime, distributed workers engage in work whenever their schedules allow,
 often with little to no awareness of the broader context of the work, and
@@
 However, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
-In this paper we will use the term \textit{on--demand labor}, as \citeauthor to capture this pair of related phenomena:
+In this paper we will use the term \textit{on--demand labor}, to capture this pair of related phenomena:
 first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
@@
 (e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}) and pursuing social scientific inquiry into the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
-This research has called out identified the sociality of gig work~\cite{crowdcollab},
+This research has identified the sociality of gig work~\cite{crowdcollab},
 as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
@@
   \end{tabularx}
-  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely--sliced tasks can become, and what the workplace looks like for workers.
-We connect piecework's history (left) to the mechanisms that impacted it (center) to derive predictions for modern on--demand work (right).
+  \label{tab:overview}
+  \caption{Piecework and on--demand work have both wrestled with questions of how complex work can get, how finely--sliced tasks can become, and what the workplace will look like for workers.
+We connect piecework's history (left) to the mechanisms that determined its outcomes to these three questions (center) in order to derive predictions for modern on--demand work (right).
 }
@@
 
-\subsection{Piecework as a lens to understand crowdsourcing}
+\subsection{Piecework as a lens to understand on--demand work}
 In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
@@
 wherein payment is made for \textit{output}, rather than for \textit{time}.
-Piecework use as a lens on on--demand work is not new.
-In \citeyear{crowdworkFuture}, for example,
+We are not the first to relate on--demand work to piecework: in \citeyear{crowdworkFuture}, for example,
 \citeauthor{crowdworkFuture} referenced crowd work as piecework briefly
 as a loose analogy~\cite{crowdworkFuture}.
+Our goal in this paper is to inspect the relationship much more closely.
 But more than this,
-the framing of on--demand labor as a reinstantiation of piecework % google said reinstantion is a word
-gives us years of historical material to make sense of this new form of work, and
-allows us to reflect on--demand work through a theoretical lens, informed by years of rigorous, empirical research.
+the framing of on--demand labor as a reinstantiation of piecework 
+gives us years of historical material to help us make sense of this new form of work, and
+allows us to reflect on--demand work through a theoretical lens that is informed by years of rigorous, empirical research.
 
-More concretely, by looking at on--demand labor as
-an instantiation (or even a continuation) of piecework,
-and by interrogating patterns that the historical literature identifies, we can achieve three goals.
-First, we can make sense of past events as part of a much larger series of an interrelated phenomena.
-Second, we can reflect on differences in the features that impacted piecework historically and on--demand work today.
-And third, to some extent, we can use these differences to offer some predictions of what on--demand labor researchers,
-and the workers,
+More concretely, by interrogating on--demand labor as
+an instantiation or even a continuation of piecework, we can make sense of past events as part of a much larger series of an interrelated phenomena (Table~\ref{tab:overview}).
+We can reflect on differences in the features that impacted piecework historically and on--demand work today.
+And, to some extent, we can use these differences to offer some predictions of what on--demand work researchers
+and workers themselves
 might expect to see on the horizon.
@@
 to make clear the analogy to on--demand work;
-and second, we interrogate the three major research questions above using the lens of piecework. 
+and second, we examine the three major research questions above using the lens of piecework. 
 For each question, we will

### 2017-01-11 drafts/labor/piecework_lit.tex
michael readthrough

@@
 to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
-George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
+In the first half of the \nth{19} century, George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
 by breaking complex calculations into constituent parts, and
@@
 on railroads~\cite{Brown01041990}, and 
-elsewhere in manufacturing~\cite{10.2307/3827491}.
-At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
---- a mechanical engineer with a fascination in work efficiency ---
-began studying and formalizing the decomposition, tracking, and management of tasks.
-In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
-concretizing an idea that had nebulously been forming,
-and which he had been working out himself, for years~\cite{taylor1914principles}.
-Scientific Management (and Fordism) thrust piecework into higher gear, especially as
-mass manufacturing and
-a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
-
-\ali{Somewhere here(?) I want to talk about Gibreth's works~\cite{gilbreth1911motion,gilbreth1911motion}. Should I save it to be the saving grace in the discussion?}
-
+elsewhere in manufacturing~\cite{10.2307/3827491} by the mid-19th century.
 \topic{By \citeyear{hughRaynbirdTaskWork} we find
@@
 
+At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
+--- a mechanical engineer with a fascination in work efficiency ---
+began studying and formalizing the decomposition, tracking, and management of tasks.
+In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
+concretizing an idea that had nebulously been forming,
+and which he had been working out himself, for years~\cite{taylor1914principles}.
+Scientific Management (and Fordism) thrust piecework into higher gear, especially as
+mass manufacturing and
+a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
+
+\ali{Somewhere here(?) I want to talk about Gibreth's works~\cite{gilbreth1911motion,gilbreth1911motion}. Should I save it to be the saving grace in the discussion?}
+
 \topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
@@
 \topic{We see this sense of independence in myriad times, locales, and industries.}
-\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
---- teenage women who assembled matchsticks in the late \nth{19} century in London.
+\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls'', teenage women who assembled matchsticks in the late \nth{19} century in London.
 % added that these women were in London since Chinmay critiqued that we're glossing over the labor advocacy movement that grew in Europe, which revealed that we completely left out the fact that this all happened in England.
@@
 demonstrate proficiency, and become experts.
+``Rosie the Riveter'',
+an icon of \nth{20} century America who
+represented empowerment and opportunity for women~\cite{honey1985creating},
+would have been a pieceworker~\cite{davies2014origins}.
 
@@
 Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
-Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
-Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.
+Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
+Given this alignment, many of the same historical properties of piecework will apply to on--demand work as well.
 
@@
 and correspondingly assigned to that or a similar task.
-``Rosie the Riveter'',
-an icon of \nth{20} century America who
-represented empowerment and opportunity for women~\cite{honey1985creating},
-would have been a pieceworker~\cite{davies2014origins}. \ali{was? would have been?}
 

### 2017-01-11 drafts/labor/introduction.tex
cleaned up intro a bit

--- 2017-01-11 drafts/labor/introduction.tex
@@
 specifically, how complex are the goals that crowd work can accomplish, and
-what kinds of goals and industries may eventually utilize it~\cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364,crowdForgeKittur}?
-Second, \namerefl{sec:decomposition}~\cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
-And third, \namerefl{sec:relationships}~\cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}?
+what kinds of goals and industries may eventually utilize it~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}?
+Second, \namerefl{sec:decomposition}~\cite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
+And third, \namerefl{sec:relationships}~\cite{turkopticon,storiesIraniSilberman,crowdcollab,takingAHITMcInnis}?
 
@@
 
+\onlyinsubfile{
+  \printbibliography
+}
+
 % \ali{

### 2017-01-11 drafts/labor/complexity.tex

@@
 medical data interpreters~\cite{lasecki2013chorus}, and
-idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
+idea generation workflows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
 
-\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new work--flows.}
+\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new workflows.}
 The system designer leverages techniques such as MapReduce~\cite{crowdForgeKittur} and
@@
 Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers.
-They identified task homogeneity and measurement as key requirements for piecework.
+They identified task homogeneity and measurement as key requirements for piecework to be successful.
 However, complex, creative work
@@
 was unsuitable~\cite{10.2307/23702539}.
-\citeauthor{Brown01041990}'s description of ``efficiency experts'' would corroborate this;
+\citeauthor{Brown01041990}'s description of ``efficiency experts'' would corroborate this:
 efficiency experts can effectively gauge how long known tasks should take, but

### 2017-01-11 drafts/labor/decomposition.tex

@@
 At the extreme,
-recent work has attempted demonstrated
+recent work has demonstrated
 microtasks that take seconds~\cite{Vaish:2014:TCC:2611222.2556996,Cai:2015:WLW:2702123.2702267}
@@
 With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
-(and Henry Ford's eponymously named \textit{Fordism})~,
+(and Henry Ford's eponymously named \textit{Fordism}),
 piecework in the early and mid--\nth{20} century surged, especially in industrial work.
@@
 The fourth and final stage was narrow expertise training. 
-Even after work is decomposed and measured, there may not be enough workers available to do it. 
+Even after work is decomposed and measured, there were not enough qualified workers available to do it. 
 So, as World War II raged and there was a dearth of skilled workers, managers trained women just enough to be able to complete their tasks~\cite{hart2013rise}. 

### 2017-01-11 drafts/labor/relationships.tex

@@
 
-\subsection[What will work and the place of work look like for workers]{The Relationships of Workers to Work, Peers, and Others}\label{sec:relationships}
+\subsection[What will work and the place of work look like for workers]{Workers' Relationships to their Work}\label{sec:relationships}
 
@@
 One possibility is that, through writing web scripts and applying them to many tasks, it is possible for a small number of spammers have an outsized influence.
-Another possibility: it was much harder for piece workers to move to a new location and get a new job --- today, workers can simply accept a different task on Mechanical Turk.
+Another possibility is that historical pieceworkers faced much more risk in shirking: it was much harder for piece workers to move to a new location and find a new job.
+Today, Mechanical Turk workers can simply accept a different task.
 A third possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen and known to them, online workers are known by little more than an account ID.
@@
 relationship with workers),
-the foreman of the \nth{20} century has largely been replaced
-by algorithms of the \nth{21} century~\cite{uberAlgorithm}.}
+the foremen of the \nth{20} century has largely been replaced
+by the algorithms of the \nth{21} century~\cite{uberAlgorithm}.}
 The result of this change is that the agents managing work are now

### 2017-01-11 drafts/labor/discussion.tex
release candidate 2

--- 2017-01-09 drafts/labor/discussion.tex
@@
 piecework is saddled by a lower limit on decomposition:
-``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}.
+``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}. % Chinmay wanted me to specifically call out that this is from chapter 6. Is that necessary...?
+% \ali{Chinmay: <suggestion to look at \cite{stigler1962information} and friction> --- not sure if worth getting into.}
 We've since studied this phenomenon in crowd work to great length both
@@
 make it possible to
-discover,
-study, and
-partner with 
+discover, study, and partner with 
 crowd workers in ways that were unimaginable to piecework researchers.
@@
 
-% \onlyinsubfile{
-%   \balance{}
-%   \printbibliography
+We should also take a moment to explore opportunity to discuss the ethics of on--demand labor,
+as \citeauthor{williamson2016} does~\cite{williamson2016}.
+% with a set of tools and perspectives that took shape only after piecework's nascent years.
+The literature on the history of piecework does not
+frame the question as \textit{whether piecework is inherently ethical or unethical},
+instead asking what conditions render it exploitative.
+The literature we have brought to bear suggests that exploitation occurs when
+conditions harm workers directly or indirectly, such as
+in sweatshops and agricultural work with pesticides, or where
+employers \textit{systematically} underpay or overwork laborers by contemporary standards.
+
+The question then is
+whether given socio--technical infrastructures % like Mechanical Turk (AMT)
+% and other marketplaces
+similarly harm, underpay, or overwork workers.
+As an example, Amazon Mechanical Turk itself does not directly require any rate of payment for work,
+but its design encourages employers to engage in exploitative behavior:
+the use of piecework remuneration, for example, workers' task search time poorly;
+task design interfaces undeniably frame
+workers as unreliable by recommending
+replication with multiple workers rather than
+trusting and paying individual workers more.
+
+Piecework also has lessons for many other research questions in crowdsourcing.
+For example, future work could more deeply explore the evolution of
+scientific management as it relates to
+crowdsourcing optimization methods,
+quality control methods between the two eras, and
+a further analysis of incentive structures.
+
+% This is
+
+% \ali{
+% Discussions from Slack:
+
+% Chinmay: I think the most provocative claim is ``The history of piecework suggests that the utopian and dystopian outcomes will both occur, in different parts of the world and to different people.''\\
+% seems contrary to the idea that marketplaces are global\\
+% I would also like some more thought into the Research Agenda section\\
+
+% Ali: so the dystopian and utopian claim is partly influenced by the fact that piecework seemed to affect some industries differently than others\\
+
+% Chinmay: ah ok\\
+% i think you could claim the same about online piecework\\
+% differnt industries\\
+
+% Ali: garment-makers were pretty thoroughly marginalized, but industrial metalwork led to this whole WWII phenomenon with expert training\\
+% mmmm\\
+% but now it's not just about the industry, but about the design of the labor market\\
+% we can influence the ``architecture'' in a way that we couldn't before\\
+% this is all lessig and stuff\\
+
+% Chinmay: oh\\
+% i understand\\
+% hmm something there\\
+% might poke more :)
+
+% \dotfill
+
+% Chinmay> i wish you had a little more criticality in taylorism but I'm a downer :)
+
+% Ali> actually a reviewer wanted more on taylorism but i was afraid that we just didn't have the space to really give anyone into taylorism what they want
+
+% Chinmay> essentially that would be a great place to talk about how those three questions you mentioned interact
+
+% Ali> oh
+% that might be worth fitting in
+% i mean we inexplicably got it down to 9.5 pages or so
+% or 9.75
+% we were over for like 99\% of the drafting and then inexplicably something just imploded and we got like half a page
+% so yeah i'll try and write something giving taylorism more space and using that to talk about the confluence of these three questions
 % }

### 2017-01-11 drafts/labor/piecework_lit.tex
wordsmithing during readthrough; biggest change was merging match--girls strike paragraphs into one

@@
 
-\topic{Piecework's history traces back further perhaps than most would expect.}
-\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
+\topic{Piecework's history traces back further than most would likely expect.}
+\citeauthor{grier2013computers} describes the process astronomers adopted of hiring teenage men
 to calculate equations in order
@@
 
-\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
+\topic{Piecework began in the intellectual domain of astronomical calculations and projections,
 but it found its foothold in manual labor.}
@@
 on railroads~\cite{Brown01041990}, and 
-elsewhere in manufacturing~\cite{10.2307/3827491} by the mid-19th century.
+elsewhere in manufacturing~\cite{10.2307/3827491} by the mid--\nth{19} century.
 \topic{By \citeyear{hughRaynbirdTaskWork} we find
@@
 particularly driven toward encapsulating the manual labor of farmwork.}
-He does this by contrasting two paradigms ---
+He does this by contrasting two paradigms:
 ``the chief difference lies between the day--labourer,
@@
 and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
-\citeauthor{10.2307/2338394} defines it through examples:
+\citeauthor{10.2307/2338394} offers a number of illustrative examples:
 ``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
-This framing offers an intuitive sense of piecework;
+This framing gives us an intuitive sense of piecework;
 ``payment for results,'' as he calls it,
 is not only common in practice, but
-well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
+well--studied in labor economics~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.
 
@@
 ``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
-Employers implemented piece--rates in some aspects and
+Employers adopted piece--rates in some aspects and
 time--rates in others.
@@
 in other senses ``day--labourers''.
-This was just one of several alternatives to strict time-- and piece--rate renumeration paradigms.
+This was just one of several alternatives to strict time-- and piece--rate remuneration paradigms.
 
-At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
---- a mechanical engineer with a fascination in work efficiency ---
-began studying and formalizing the decomposition, tracking, and management of tasks.
+In the late years of the \nth{19} century, \citeauthor{taylor1914principles}
+--- a mechanical engineer with an interest in work efficiency ---
+began studying and formalizing the decomposition, tracking, and management of tasks~\cite{taylor1896piece}.
 In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
@@
 
-\topic{Since workers could now choose their own schedule and style, there arose a discussion on how best to manage pieceworkers.}
+\topic{Since workers could now choose their own schedule and style, a discussion arose on how best to manage pieceworkers.}
 This conversation came to regard workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
-piece workers would gladly work as diligently for as long as possible because
-incentive--based pay rewards hard work~\cite{clark1908cotton}.
+pieceworkers would gladly work diligently and for as long as possible, as
+incentive--based pay rewarded exactly, and thus aligned the goals of both managers and workers~\cite{clark1908cotton}.
 
@@
 Between 1938 and 1942, the proportion of metal workers under piecework systems had climbed steeply from 11\% to 60\%~\cite{hart2005piecework}.
-By 1961, that proportion dropped to 8\%~\cite{carlson1982time}.
-\citeauthor{carlson1982time} details that, from 1973 to 1980, the holdouts of piecework
+By 1961, \citeauthor{carlson1982time} finds, the proportion dropped to 8\%~\cite{carlson1982time}.
+He notes that, from 1973 to 1980, the holdouts of piecework
 --- where more than 50\% worked under incentive wage plans ---
@@
 \subsection{Case studies in piecework}
-\topic{Throughout the rest of the paper, we will return to four case studies to frame our analyses:
-Airy's employment of human computers;
+\topic{Throughout the paper, we will return to four case studies to frame our analyses:
+Airy's use of human computers;
 domestic and farm workers;
 the ``match--girls'' strike;
-and industrial and assembly-line workers.}
+and industrial and assembly--line workers.}
 In introducing these cases at a high level,
@@
 As a result, many of Airy's computers had relatively rudimentary educations
-compared to the background of education that typically worked in the calculation of solar tables.
+compared to those that typically worked in the calculation of solar tables.
 Airy distributed tasks by mail,
@@
 and paid for each piece of work completed.
-Airy also instituted a policy of firing his computers once they reached age 23.
 
@@
 Second, the work was designed so that intermediate results could be quickly verified: Airy would have two workers each do the calculation, and another person compare their answers.
-Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.
+Third, Airy identified ways to decompose the large task into narrowly--trainable subtasks.
 
+Some of Airy's policies were more controversial, for example
+firing computers once they reached age 23.
 This practice ensured two outcomes that disfavored workers.
-First, it eliminated any potential to advance professionally, as
-workers' careers ended relatively quickly,
-and without formal education in mathematics they struggled to find work for which their experience was meaningful.
-Second, it limited workers' ability to organize
-by ensuring that workers were in relatively little communication with each other.
+First, it drastically reduced professional advancement, as
+workers' careers ended quickly,
+and without conventional backgrounds in mathematics
+they later struggled to find work for which their experience was meaningful.
+And second, it limited workers' ability to organize
+by ensuring that workers were in little communication with each other,
+and that they had almost no opportunity to recognize their circumstances and
+to coordinate.
 
@@
 At the end of the \nth{19} century,
-manufacturers employed women to assemble matchsticks in factories.
+manufacturers had begun to employ teenage women to assemble matchsticks in factories.
 These women rallied first in the form of a march on parliament in 1871 to protest a proposed tax, and 
 later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.
-
-\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
+\topic{This later strike was sparked by a worker's arbitrary docking of pay, but
 much deeper resentment had been simmering for years.}
@@
 poor working conditions, and
-having to work with hazardous phosphoric materials.
+having to work with hazardous materials such as white phosphorus, the improper handling of which
+caused serious, painful, disfiguring medical conditions in the bones and ultimately death.
 
-\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
+\topic{Regardless of what prompted it, the lasting impact of the match--girls strike of 1888 was profound.}
 This was one of the earliest and most famous successful worker strikes,
@@
 As \citeauthor{weyer1894history} described,
-``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
-the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.
+``the match--girls' victory turned a new leaf in Trade Union annals'': in the 30 years after the match--girls strike,
+the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%~\cite{weyer1894history}.
 
@@
 These were workers who treasured their independence, but also fiercely protected one another. % , contributing to the common good.
-Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
+``Brashness'' may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
 making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.
@@
 the formal training nor
-the years of apprenticeship experience
-that conventional workers would have had.
+years of experience that men would have had from apprenticeships.
+% that conventional workers would have had.
 Rather than attempting to train this new labor force in every aspect of industrial work,

### 2017-01-11 drafts/labor/complexity.tex
small wording changes on final read-through

@@
 \subsubsection{\crowdworkpers}
-\topic{\citeauthor{crowdForgeKittur} first opened the question of
-whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.}
-Their work demonstrated proof--of--concept crowdsourcing of
-an encyclopedia article and news summary
+\topic{\citeauthor{crowdForgeKittur} were among the first to ask
+whether crowdsourcing could be used for more than parallelizing tasks~\cite{crowdForgeKittur}.}
+Their work showed that it could, with proof--of--concept crowdsourcing of
+encyclopedia articles and news summaries
 --- tasks which could be verified or repeated
-with reasonable expectations of similar outcomes.
+with reasonable expectations of similar results.
 Seeking to raise the complexity ceiling,
 researchers have since created
-additional proof--of--concept applications and techniques,
+yet more applications and techniques,
 including conversational assistants~\cite{lasecki2013chorus},
 medical data interpreters~\cite{lasecki2013chorus}, and
-idea generation workflows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
+idea generation~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
 

### 2017-01-11 drafts/labor/introduction.tex
small wording changes on final read-through

@@
 \topic{The past decade has seen a flourishing of computationally--mediated labor.}
-A framing of work into modular, pre-defined components
+A framing of work into modular, pre--defined components
 enables computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
@@
 
-The realization that complex goals can be accomplished by directing crowds of workers has spurred industry to flock to sites of labor
-such as AMT to explore the limits of this distributed, on--demand workforce.
+The realization that complex goals can be accomplished by directing crowds of workers has spurred firms to explore sites of labor
+such as AMT to find the limits of this distributed, on--demand workforce.
 Researchers have also taken to the space in earnest,
@@
 This research has identified the sociality of gig work~\cite{crowdcollab},
-as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
-Others have focused on the \textit{outcomes} of this frustration,
+as well as the frustration and disenfranchisement that these systems effect~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
+Others have focused on the responses to this frustration,
 reflecting on the resistance that workers express against digitally--mediated labor markets~\cite{uberAlgorithm,dynamo}.
@@
 specifically, how complex are the goals that crowd work can accomplish, and
-what kinds of goals and industries may eventually utilize it~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}?
+what kinds of industries may eventually utilize it~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}?
 Second, \namerefl{sec:decomposition}~\cite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
@@
     \small{Growth from simple tasks such as sewing to more complex composite outcomes on the assembly line floor.} &
-    \small{Complexity was limited by what could be easily measured and evaluated for payment by the piece.} &
-    \small{Measurement and verification remain challenges and will limit complexity unless solved.} \\ \hline
+    \small{Complexity was limited to tasks that could be easily measured and evaluated for payment by the piece.} &
+    \small{Measurement and verification will remain persistent challenges that will limit complexity unless solved.} \\ \hline
 
@@
 gives us years of historical material to help us make sense of this new form of work, and
-allows us to reflect on--demand work through a theoretical lens that is informed by years of rigorous, empirical research.
+allows us to study on--demand work through a theoretical lens that is informed by years of rigorous, empirical research.
 
-More concretely, by interrogating on--demand labor as
-an instantiation or even a continuation of piecework, we can make sense of past events as part of a much larger series of an interrelated phenomena (Table~\ref{tab:overview}).
+More concretely, by positioning on--demand labor as
+an instantiation or even a continuation of piecework, we can make sense of past events as part of a much larger series of interrelated phenomena (Table~\ref{tab:overview}).
 We can reflect on differences in the features that impacted piecework historically and on--demand work today.
@@
 might expect to see on the horizon.
-For example, we will draw on the piecework literature regarding task decomposition,
+For example, we will draw on piecework's scholarship on task decomposition,
 which was historically limited by shortcomings in measurement and instrumentation, and
-leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
---- namely, by enabling precise tracking and measurement via algorithms and software.
+leverage that insight to suggest how modern technology affects this mechanism in on--demand work
+--- namely, enabling precise tracking and measurement via algorithms and software.
 

### 2017-01-11 drafts/labor/trackingWork.tex
cleaned up intro a bit
small wording changes on final read-through

@@
-\documentclass[chi_draft]{sigchi}
+\documentclass{sigchi}
 
@@
 \newcommand{\implication}{Implications for on--demand work}
-\newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of
+\newcommand{\pieceworkdefinition}{piecework is the paradigm of remuneration that is largely made compelling for its use of
 ``payment for results'',
@@
 \begin{document}
+\CopyrightYear{2017} 
+\setcopyright{acmlicensed}
+\conferenceinfo{CHI 2017,}{May 06 - 11, 2017, Denver, CO, USA}
+\isbn{978-1-4503-4655-9/17/05}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/3025453.3025974}
 \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
@@
 
-\balance{}
-\printbibliography
+% \balance{}
+% \printbibliography
 

### 2017-01-11 drafts/labor/conclusion.tex
release candidate 3

@@
 \section{Conclusion}
-\topic{On--demand work is not new, but a contemporary instantiations of piecework.}
-In this paper, we reconsider three major research questions in on--demand work using the lens of piecework:
+\topic{On--demand work is not new, but a contemporary instantiation of piecework.}
+In this paper, we reconsider three major research questions in on--demand work using piecework as a lens:
 \begin{inlinelist}
@@
 \end{inlinelist}
-To do so, we draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon.
-Reciprocally, we believe that modern on--demand work will teach us about the broader phenomenon of piecework as well.
+We draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change soon.
+Reciprocally, we believe that modern on--demand work can teach us about the broader phenomenon of piecework as well.
 If history really does repeat itself, the best we can do is be prepared.
 
+\section{Acknowledgements}
+This work was supported by the Stanford Cyber Initiative and National Science Foundation award IIS-1351131.
+
 \end{document}

### 2017-01-11 drafts/labor/discussion.tex
release candidate 3

--- 2017-01-11 drafts/labor/discussion.tex
@@
 enough of piecework can and does inform on--demand work that
-HCI and CSCW might seek out historical framings for other phenomena of study as well.
+HCI and CSCW researchers might seek out historical framings for other phenomena of study as well.
 While we can only speculate one of (perhaps many) possible futures, history does allow us to articulate and bound which futures appear more likely.
@@
 warn that on--demand labor will create exploitative sites of dispossession~\cite{scholz2012digital},
-racial discrimination~\cite{edelman2015racial},
+discrimination~\cite{edelman2015racial},
 and invisible, deeply frustrated workers~\cite{turkopticon,bighamHalfWorkday}.
@@
 \topic{On the other hand, while piecework's nascent years were grim, 
-they precipitated a century of some of the most potent labor advocacy organizations in modern history~\cite{hart2013rise,mccallum2013global}.}
+what followed was a century of some of the most potent labor advocacy organizations in modern history~\cite{hart2013rise,mccallum2013global}.}
 Even today, the \textit{geist} of the labor union revolution
@@
 Recently, in India, workers across the nation engaged in
-the largest labor strike in human history
-\cite{indiaStrikeRealNews}.
-If labor advocacy groups can find ways to effect change in on--demand work as some have called for
-\cite{futureUnions},
+the largest labor strike in human history~\cite{indiaStrikeRealNews}.
+If labor advocacy groups can find ways to effect change in on--demand work as some have called for~\cite{futureUnions},
 then the future of on--demand labor may follow
@@
 
-\topic{The history of piecework suggests that the utopian and dystopian outcomes will \textit{both} occur, in different parts of the world and to different people.}
-When piecework plummeted in the United States, outsourcing rose --- creating major labor issues around the world.
-It is entirely possible that we will create a new brand of flexible online career in developed countries, while simultaneously fueling an unskilled decentralized labor force in developing nations.
-As designers and researchers, this prompts the question: which outcome are we attempting to promote or avoid for who?
+\topic{The history of piecework suggests that the utopian and dystopian outcomes will \textit{both} occur,
+in different parts of the world and to different groups of people.}
+When piecework largely disappeared in the United States, outsourcing appeared --- creating major labor issues around the world.
+It'ss entirely possible that we will create a new brand of flexible online career in developed countries,
+while simultaneously fueling an unskilled decentralized labor force in developing nations.
+As designers and researchers, this prompts the question: which outcome are we attempting to promote or avoid for whom?
 
@@
 While we have arguably outpaced piecework with regard to the limits on the complexity of work,
-the most complex and open-ended wicked problems~\cite{rittel1973dilemmas} remain the domain of older human collectives such as governments and organizations. 
+the most complex and open--ended wicked problems~\cite{rittel1973dilemmas} remain the domain of older human collectives such as governments and organizations. 
 In addition, we can learn from the piecework literature as it relates to
 the stymieing effect that mismanagement has on workers;
-research into the complexity limits should emphasize on finding new ways to manage workers,
-in particular using humans --- perhaps other crowd workers ---
-to act as modern foremen.
+research into complexity limits should focus on finding new ways to manage workers,
+in particular using humans (perhaps other crowd workers)
+to act as modern ``foremen''.
 
@@
 piecework is saddled by a lower limit on decomposition:
-``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}. % Chinmay wanted me to specifically call out that this is from chapter 6. Is that necessary...?
-\ali{Chinmay: <suggestion to look at \cite{stigler1962information} and friction> --- not sure if worth getting into.}
-We've since studied this phenomenon in crowd work to great length both
+as \citeauthor{bewley1999wages} mentions,
+``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}. 
+We've since studied this phenomenon in crowd work both
 observationally \cite{taskSearch} and
@@
 we may only make incremental advances in micro--task decomposition.
-When the cognitive cost of understanding a task and its inputs outstrips the effort required to complete the task, decomposition seems a poor choice.
+When the cognitive cost of understanding a task and its inputs outstrips the effort required to complete it, decomposition seems a poor choice.
 
@@
 make it possible to
-discover,
-study, and
-partner with 
+discover, study, and partner with 
 crowd workers in ways that were unimaginable to piecework researchers.
-A professor engages in crowd work~\cite{bighamHalfWorkday}
+\citeauthor{bighamHalfWorkday} engages in crowd work~\cite{bighamHalfWorkday}
 not just because it's possible, but because our community
@@
 
-\ali{
-Discussions from Slack:
+We should also take a moment to explore the opportunity to discuss the ethics of on--demand labor,
+as \citeauthor{williamson2016} does~\cite{williamson2016}.
+The literature on the history of piecework does not
+frame the question as \textit{whether piecework is inherently ethical or unethical},
+instead asking \textit{what conditions render it exploitative}.
+The literature we have brought to bear suggests that exploitation occurs when
+conditions harm workers directly or indirectly, such as
+in sweatshops and agricultural work with pesticides, or where
+employers systematically underpay or overwork laborers by contemporary standards.
 
-Chinmay: I think the most provocative claim is ``The history of piecework suggests that the utopian and dystopian outcomes will both occur, in different parts of the world and to different people.''\\
-seems contrary to the idea that marketplaces are global\\
-I would also like some more thought into the Research Agenda section\\
+The question then is
+whether given socio--technical infrastructures 
+systematically harm, underpay, or overwork workers.
+For example, Amazon Mechanical Turk does not directly require any rate of payment for work,
+but its design encourages employers to engage in exploitative behavior:
+piece--rate pay, for example, does not value workers' task search time;
+additionally, task design interfaces undeniably frame
+workers as unreliable by recommending
+replication with multiple workers rather than
+trusting and paying individual workers more.
 
-Ali: so the dystopian and utopian claim is partly influenced by the fact that piecework seemed to affect some industries differently than others\\
-
-Chinmay: ah ok\\
-i think you could claim the same about online piecework\\
-differnt industries\\
-
-Ali: garment-makers were pretty thoroughly marginalized, but industrial metalwork led to this whole WWII phenomenon with expert training\\
-mmmm\\
-but now it's not just about the industry, but about the design of the labor market\\
-we can influence the ``architecture'' in a way that we couldn't before\\
-this is all lessig and stuff\\
-
-Chinmay: oh\\
-i understand\\
-hmm something there\\
-might poke more :)
-
-\dotfill
-
-Chinmay> i wish you had a little more criticality in taylorism but I'm a downer :)
-
-Ali> actually a reviewer wanted more on taylorism but i was afraid that we just didn't have the space to really give anyone into taylorism what they want
-
-Chinmay> essentially that would be a great place to talk about how those three questions you mentioned interact
-
-Ali> oh
-that might be worth fitting in
-i mean we inexplicably got it down to 9.5 pages or so
-or 9.75
-we were over for like 99% of the drafting and then inexplicably something just imploded and we got like half a page
-so yeah i'll try and write something giving taylorism more space and using that to talk about the confluence of these three questions
-}
+Piecework also has lessons for a number of other research questions in crowdsourcing.
+For example, future work could more deeply explore the evolution of
+scientific management as it relates to
+crowdsourcing optimization;
+quality control approaches between the two eras; and
+further analysis of incentive structures.
 

### 2017-01-11 drafts/labor/relationships.tex
release candidate 3

@@
 the recent shift into computational labor systems has directly impacted
-the lives and livelihood of workers in new ways.
-So, it is important to understand:
-what will the future look like for the workers who use these systems?
+the lives and livelihoods of workers in new ways.
+So, it's imperative to ask:
+What will the future look like for the workers who use these systems?
 
@@
 	- Workers do it for money
-	- Workers coordinate and communicate (gray, being a turker, team stuff). there is mega-drama
+	- Workers coordinate and communicate (gray, being a turker, team stuff). there is mega--drama
 	- Workers don't like requesters. have tried to organize
@@
 Early literature emphasized motivations like fun and spare change, but this narrative soon shifted to emphasize that many workers use platforms such as Amazon Mechanical Turk as a primary source of income~\cite{kaufmann2011more,ipeirotis2010demographics,Antin2012a}.
-Despite this, Mechanical Turk is a low-wage affair for most workers in the United States~\cite{ipeirotis2010demographics,martin2014being,gupta2014turk}.
-Thus, those who choose to opt out of the traditional labor force and spend significant time on Mechanical Turk are especially motivated by the opportunity for autonomy and skill variety~\cite{kaufmann2011more}.
-While some carry attitudes of ``pity the workers'' toward Turkers, this frame is increasingly rejected by workers and designers as patronizing~\cite{storiesIraniSilberman}.
+Despite this, Mechanical Turk is a disappointingly low--wage worksite for most people in the United States~\cite{ipeirotis2010demographics,martin2014being,gupta2014turk}.
+Thus, those who choose to opt out of the traditional labor force and spend significant time on Mechanical Turk are especially motivated by the opportunity for autonomy and transience between tasks~\cite{kaufmann2011more}.
+While some describe Turkers as powerless victims or even unaware of what's going on,
+this framing is increasingly being rejected by workers and designers as
+``cast[ing] Turkers as dopes in the system.''~\cite{storiesIraniSilberman}.
 
@@
 the tense relationships between workers and requesters~\cite{fixingChaos,dynamo}.
-Workers are often blamed for any low-quality work, regardless of whether they are responsible~\cite{martin2014being,takingAHITMcInnis}.
+Workers are often blamed for any low--quality work, regardless of whether they are responsible~\cite{martin2014being,takingAHITMcInnis}.
 Some research is extremely open about this position, blaming unpredictable work on ``malicious'' workers~\cite{MaliciousCrowdworkersGadiraju} or those with ``a lack of expertise, dedication [or] interest''~\cite{Sheng:2008:GLI:1401890.1401965}.
-Workers resent this position --- for good reason.
-\citeauthor{turkopticon} highlighted the information asymmetry between workers and requesters on AMT, leading to the creation of \TO, a site which allows Turkers to rate and review requesters~\cite{turkopticon}.
-Dynamo then took this critique on information asymmetry and power imbalances a step further, designing a platform to facilitate
-Turkers acting collectively to bring about changes to their circumstances~\cite{dynamo}.
+Workers resent this position, and for good reason.
+\citeauthor{turkopticon} highlighted the information asymmetry between workers and requesters on AMT,
+which led to the creation of Turkopticon, a site which allows Turkers to rate and review requesters~\cite{turkopticon}.
+Dynamo then took this critique on information asymmetry and power imbalances further, designing a platform to facilitate
+collective action among Turkers to changes to their circumstances~\cite{dynamo}.
 
 Researchers have also begun to appreciate the sociality of crowd workers.
-Because the platforms do not typically include social spaces, workers instead congregate off-platform in forums and mailing lists.
-There, Turkers exchange advice on high-paying work, talk about their earnings, build social connections, and discuss requesters~\cite{martin2014being}.
-Many crowd workers know each other through offline and online connections, coordinating behind-the-scenes despite the platforms encouraging independent work~\cite{crowdcollab,yin2016communication}.
+Because the platforms do not typically include social spaces, workers instead congregate off--platform in forums and mailing lists.
+There, Turkers exchange advice on high--paying work, talk about their earnings, build social connections, and discuss requesters~\cite{martin2014being}.
+Many crowd workers know each other through offline and online connections, coordinating behind--the--scenes despite the platforms encouraging independent work~\cite{crowdcollab,yin2016communication}.
 However, the frustration and mistrust that workers experience with requesters does occasionally boil over on the forums. 
@@
 ``When he works by the day the Italian operative wishes to leave before the whistle blows,
-but if he works by the piece he will work as many hours as it is possible for him to stand.''
+but if he works by the piece he will work as many hours as it is possible for him to stand''~\cite{clark1908cotton}.
 However, the emergent trend contrasted with this early rhetoric, as
 when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.
-Piece workers, \citeauthor{roy1954efficiency} found, would form even more starkly antagonistic relationships with their managers.
+Piece workers, \citeauthor{roy1954efficiency} found, would form acrimonious relationships with their managers.
 
-\topic{Soon workers began resisting piecework regimes.}
+\topic{Soon, workers began resisting piecework regimes.}
 The match--girls engaged in their famous strike of 1888, particularly pushing to abolish the fines that were taken out of their wages.
@@
 
-Many worker organizations began weighing in on
-(or, more precisely, against)
+Many worker organizations began weighing in against
+% (or, more precisely, against)
 piecework and the myriad oversights it made in valuing workers' time~\cite{american1921problem,richards1904anything}.
@@
 
-\topic{Relative to the modern on--demand workers, there is a noticeable dearth of information on the inter--personal relationships among pieceworkers
+\topic{Relative to the modern on--demand workers, there is a noticeable dearth of information on the interpersonal relationships among pieceworkers
 beyond the match--girls at the end of the \nth{19} century.}
@@
 Why the difference? 
-One possibility is that, through writing web scripts and applying them to many tasks, it is possible for a small number of spammers have an outsized influence.
-Another possibility is that historical pieceworkers faced much more risk in shirking: it was much harder for piece workers to move to a new location and find a new job.
-Today, Mechanical Turk workers can simply accept a different task.
-A third possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and where pieceworkers could be directly observed by foremen and known to them, online workers are known by little more than an account ID.
+One possibility is that, by writing web scripts and applying them to many tasks,
+a small number of spammers have an outsized influence on the perception of bad actors.
+Another possibility is that historical pieceworkers faced much more risk in shirking:
+it was much harder for pieceworkers to move to a new location and find a new job.
+Today, Mechanical Turk workers can work for a dozen or more different groups in the span of a day.
+A third possibility: online anonymity breeds distrust~\cite{friedman2000trust}, and
+where pieceworkers could be directly observed by foremen and known to them,
+online workers are known by little more than an inscrutible alphanumeric string, like A2XJMS2J2FMVXK.
 
-\topic{The relationship between workers and employers has also shifted: while historically the management of workers had to be done through a foreman
-(who necessarily had an intuitive
---- perhaps sympathetic ---
-relationship with workers),
-the foremen of the \nth{20} century has largely been replaced
-by the algorithms of the \nth{21} century~\cite{uberAlgorithm}.}
-The result of this change is that the agents managing work are now
+\topic{The relationship between workers and employers has also shifted: while historically the management of workers had to be done through a foreman,
+% (who necessarily had an intuitive
+% --- perhaps sympathetic ---
+% relationship with workers),
+foremen of the \nth{20} century have largely been replaced
+by algorithms of the \nth{21} century~\cite{uberAlgorithm}.}
+% The result of this change is that
+Consequently,
+the agents managing work are now
 cold, logical, and unforgiving.
-Where a person might recognize that the ``attention check'' questions
+While a person might recognize that the ``attention check'' questions
 proposed by \citeauthor{le2010ensuring} and others ensure that
@@
 only seem to antagonize workers~\cite{takingAHITMcInnis}.
-More than 30 years ago, \citeauthor{10.2307/2555446} wrote:
-``When performance is difficult to evaluate,
+As \citeauthor{10.2307/2555446} wrote in \citeyear{10.2307/2555446},
+``when performance is difficult to evaluate,
 imperfect input measures and
@@
 defective (simple, observable) output measures''~\cite{10.2307/2555446}.
-This frustration has only grown as requesters have had to rely on automatic management mechanisms. Only a few use the equivalent of human foremen~\cite{haas2015argonaut,kulkarni2012mobileworks}.
+This frustration has only grown as requesters have had to rely on automatic management mechanisms.
+Only a few use the equivalent of human foremen~\cite{haas2015argonaut,kulkarni2012mobileworks}.
 
-Relative to the mature state of collective action for pieceworkers offline,
-crowd workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
-%Both pieceworkers and on--demand workers have struggled at times to form a collective identity necessary to organize.
-With workers joining and leaving the on--demand labor force continuously, and with many part-time members, it is extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
-However, even when they can, enforcement remains a challenge: whereas pieceworkers could physically block access to a site of production, online labor markets provide no facilities for workers to change the experience of other workers.
+Relative to the history of collective action for pieceworkers,
+on--demand workers have struggled to make their voices heard~\cite{dynamo,storiesIraniSilberman,turkopticon}.
+With workers constantly drifting through these platforms, and with many part--time members,
+it's extremely difficult to corral the group to make a collective decision~\cite{dynamo}.
+Even when they can, enforcement remains a challenge:
+while pieceworkers could physically block access to a site of production and convince other workers to join them,
+online labor markets provide no facilities for workers to change the experience of other workers.
 This is a key limitation --- without it, workers cannot enforce a strike.
@@
 Malinowski, Boas, Mead, and
-% \citeauthor{malinowski2002argonauts,boas1940race,mead1973coming} and % this was naming Boas twice, so.
 other luminaries throughout the first half of the \nth{20} century

### 2017-01-12 chi/piecework/complexity.tex
finalized, moved to /chi, etc

--- 2017-01-11 drafts/labor/complexity.tex
@@
 \topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new workflows.}
-The system designer leverages techniques such as MapReduce~\cite{crowdForgeKittur} and
-sequence alignment algorithms~\cite{lasecki2012real}, arranging humans as computational black boxes within.
+System designers leverage techniques such as MapReduce~\cite{crowdForgeKittur} and
+sequence alignment algorithms~\cite{lasecki2012real}, arranging humans as computational black boxes.
 This approach has proven a compelling one because
@@
 
-However, when piecework intially entered the American economy,
+But when piecework intially entered the American economy,
 it was not used for complex work.
@@
 that apprenticeships imparted \cite{hart2013rise}.
-So, initially piecework arose for farm work, and as
+So, initially piecework arose for farmwork, and as
 \citeauthor{hughRaynbirdTaskWork} and others discuss,
@@
 Computational systems now act as ``piecework clerks''~\cite{10.2307/23702539} to
-inspect,
-modify and
-combine work~\cite{turkopticon,takingAHITMcInnis}.
-However, these algorithms are less able than human managers to evaluate subjective work, and
-their ability to exercise discretion.
+inspect and modify work~\cite{turkopticon,takingAHITMcInnis}.
+However, these algorithms are less competent than humans at evaluating subjective work, as well as
+in their ability to exercise discretion, causing new problems for workers and managers.
 
@@
 So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than piecework did.
-Improvements in workflows, measurement and evaluation have already been made, and it is not clear that the remaining challenges are readily solveable.
-However, on--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet.
+Improvements in workflows, measurement, and evaluation have already been made, and it'ss not immediately clear that the remaining challenges are readily solvable.
+However, on--demand work will be far more broadly distributed than piecework historically was
+--- reaching many more tasks and areas of expertise by virtue of the internet.
 

### 2017-01-12 chi/piecework/decomposition.tex
finalized, moved to /chi, etc

--- 2017-01-11 drafts/labor/decomposition.tex
@@
 As such,
-one of the central questions in the literature is how finely-sliced these microtasks can become, and which kinds of tasks are amenable to decomposition.
+one of the central questions in the literature is how finely--sliced these microtasks can become, and which kinds of tasks are amenable to decomposition.
 In this section,
@@
 Even when tasks such as writing and editing cannot be reliably performed by individual workers,
-researchers demonstrated that
-decompositions of these tasks into workflows can succeed~\cite{crowdForgeKittur,
+researchers have demonstrated that
+the decomposition of these tasks into workflows can succeed~\cite{crowdForgeKittur,
 bernsteinSoylent,
@@
 These decompositions typically take the form of workflows,
-which are algorithmic sequences of tasks that manage interdependencies~\cite{Bigham2014}. 
+instantiated as algorithmically managed sequences of tasks that resolve interdependencies~\cite{Bigham2014}. 
 Workflows often utilize a first sequence of tasks to identify an area of focus (e.g.,
@@
 be decomposed?} 
-Moreover,
-how thinly can work be sliced and subdivided into smaller and smaller tasks? 
+More pointedly,
+how thinly \textit{should} work be sliced and subdivided into smaller and smaller tasks? 
 The general trend has been that smaller is better,
@@
 \topic{Four major stages characterize decomposition in the history of piecework.}
-The first stage was decomposition of an expert task such that it could be done by non-experts. 
+The first stage was decomposition of an expert task such that it could be done by non--experts. 
 This was arguably the main innovation of Airy's human computers.
 Rather than hire expert computers, Airy identified ways to break down astrological calculations into steps that could be completed with only a basic knowledge of mathematics.
-Likewise, \citeauthor{Brown01041990} argued that piecework arose in industries where there was significant task homogeneity and low fixed costs of machinery and training~\cite{Brown01041990}.
+Likewise, \citeauthor{Brown01041990} argued that piecework arose in industries with homogeneous tasks and low fixed costs of machinery and training~\cite{Brown01041990}.
 
@@
 There is demand for skilled workers in many crowdsourcing tasks, and systems to help train workers~\cite{suzukiAtelier}. 
-We might expect to see the rise of systems that scaffold workers into extremely narrow areas of expertise.
+We might expect to see the rise of systems that scaffold workers into extremely narrow areas of expertise,
+for instance using online courses as proof of expertise in a specific domain necessary for a microtask.
 

### 2017-01-12 chi/piecework/complexity.tex
fixing final paper name and other stuff
actual forreal final version

@@
-\documentclass[trackingWork]{subfiles}
+\documentclass[pn4226]{subfiles}
 
@@
 \subsubsection{\crowdworkpers}
-\topic{\citeauthor{crowdForgeKittur} were among the first to ask
+\topic{Kittur et al. were among the first to ask
 whether crowdsourcing could be used for more than parallelizing tasks~\cite{crowdForgeKittur}.}
@@
 \topic{Restricting attention to non--expert, microtask workers proved limiting.}
-So, \citeauthor{foundry} introduced the idea of crowdsourcing with
+So, Retelny et al. introduced the idea of crowdsourcing with
 online paid \textit{experts} from platforms such as Upwork.
@@
 that apprenticeships imparted \cite{hart2013rise}.
-So, initially piecework arose for farmwork, and as
-\citeauthor{hughRaynbirdTaskWork} and others discuss,
+So, initially piecework arose for farm work, and as
+
+Raynbird and others discuss,
 the practice remained relatively obscure until
@@
 \topic{Measurement also limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework.}
-Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers.
+Earlier we discussed Graves's and later Brown's analysis of railway workers.
 They identified task homogeneity and measurement as key requirements for piecework to be successful.
@@
 was unsuitable~\cite{10.2307/23702539}.
-\citeauthor{Brown01041990}'s description of ``efficiency experts'' would corroborate this:
+
+Brown's description of ``efficiency experts'' would corroborate this:
 efficiency experts can effectively gauge how long known tasks should take, but
@@
 \topic{Moreover, piecework was limited to tasks that could be quickly and accurately evaluated.}
-\citeauthor{hart2016rise} argue that evaluation limited piecework's complexity:
+
+Hart argues that evaluation limited piecework's complexity:
 at some point, evaluating multidimensional work for quality
@@
 Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
-Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
+Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as Hart and Roberts and later Grier described in their piecework examples
 of human computation~\cite{grier2013computers} and

### 2017-01-12 chi/piecework/decomposition.tex
fixing final paper name and other stuff
actual forreal final version

@@
-\documentclass[trackingWork]{subfiles}
+\documentclass[pn4226]{subfiles}
 \begin{document}
@@
 the invisible labor of \textit{finding} tasks~\cite{martin2014being} has arisen as a major issue.}
-\citeauthor{taskSearch} illustrate the task search challenges on AMT~\cite{taskSearch}.
+Chilton et al. illustrate the task search challenges on AMT~\cite{taskSearch}.
 Workers seek out good requesters~\cite{martin2014being} and then 
@@
 Rather than hire expert computers, Airy identified ways to break down astrological calculations into steps that could be completed with only a basic knowledge of mathematics.
-Likewise, \citeauthor{Brown01041990} argued that piecework arose in industries with homogeneous tasks and low fixed costs of machinery and training~\cite{Brown01041990}.
+Likewise, Brown argued that piecework arose in industries with homogeneous tasks and low fixed costs of machinery and training~\cite{Brown01041990}.
 
@@
 Unlike Airy's human computers, railway engineers had significant expertise~\cite{Brown01041990}.
-As \citeauthor{Brown01041990} noted, however, it was still possible to discretize and measure their work.
+As Brown noted, however, it was still possible to discretize and measure their work.
 Thus, experts such as railway engineers became pieceworkers as well.
@@
 What can be modularized can be measured, and what can be measured can be optimized.
-With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
+With Taylor's formalization of scientific management in \textit{Taylorism}
 (and Henry Ford's eponymously named \textit{Fordism}),
@@
 higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
-While \citeauthor{Brown01041990} points out that
+While Brown points out that
 piecework dramatically advanced the instrumented measurement of workers,
-in \citeauthor{taylor1914principles}'s time highly instrumented,
+in Taylor's time highly instrumented,
 automatic measurement of workers was all but impossible~\cite{Brown01041990}.
@@
 but trivial~\cite{waltz2012quantified}.}
-Where \citeauthor{10.2307/23702539} directly implicates measurement as
+Where Graves directly implicates measurement as
 preventing scientific management from being fully utilized~\cite{10.2307/23702539},

### 2017-01-12 chi/piecework/discussion.tex
fixing final paper name and other stuff
actual forreal final version

--- 2017-01-11 drafts/labor/discussion.tex
@@
-\documentclass[trackingWork]{subfiles}
+\documentclass[pn4226]{subfiles}
 \makeatletter
@@
 The past isn't a perfect predictor for the future;
-as \citeauthor{scholz2012digital} cautions,
+as Scholz cautions,
 ``it would be wrong to conclude that in the realm of digital labor there is nothing new under the sun''~\cite{scholz2012digital}.
@@
 
-\citeauthor{rosenberg1994exploring} and others have contributed substantially to the practice in part by clearly limiting the extent of their claims
+Rosenberg and others have contributed substantially to the practice in part by clearly limiting the extent of their claims
 --- only offering, for instance,
@@
 piecework is saddled by a lower limit on decomposition:
-as \citeauthor{bewley1999wages} mentions,
+as Bewley mentions,
 ``piecework does not compensate workers for time spent switching tasks''~\cite{bewley1999wages}. 
@@
 crowd workers in ways that were unimaginable to piecework researchers.
-\citeauthor{bighamHalfWorkday} engages in crowd work~\cite{bighamHalfWorkday}
+Bigham engages in crowd work~\cite{bighamHalfWorkday}
 not just because it's possible, but because our community
@@
 We should also take a moment to explore the opportunity to discuss the ethics of on--demand labor,
-as \citeauthor{williamson2016} does~\cite{williamson2016}.
+as Williamson does~\cite{williamson2016}.
 The literature on the history of piecework does not

### 2017-01-12 chi/piecework/introduction.tex
fixing final paper name and other stuff
actual forreal final version

--- 2017-01-11 drafts/labor/introduction.tex
@@
-\documentclass[trackingWork]{subfiles}
+\documentclass[pn4226]{subfiles}
 \onlyinsubfile{
@@
 first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
-and second, \textit{gig work}~\cite{friedman2014workers}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
-\ali{I'm a little worried about people offering contradicting terms, but the argument I \textit{want} to make is that
-these terms (``gig work'' and ``crowd work'') were sort of born out of marketing and business interests (just like ``sharing economy''),
-and as a result they're somewhat fuzzy and imprecise (for example, see~\cite{de2015crowdsourcing}); what we're trying to do is
-place these terms where they're most accurate and tightly scoped, so that their meaningfulness is retained;
-then we use  ``on--demand work'' to capture this work more holistically,
-much more carefully selecting the terminology for academic discourse.
-
-Too off--course?}
+and second, \textit{gig work}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
 
@@
 We are not the first to relate on--demand work to piecework: in \citeyear{crowdworkFuture}, for example,
-\citeauthor{crowdworkFuture} referenced crowd work as piecework briefly
+Kittur et al. referenced crowd work as piecework briefly
 as a loose analogy~\cite{crowdworkFuture}.
@@
 
-% \ali{
-% The pervasiveness of rating and reviewing mechanisms is a chief issue in the analysis ofthe gig-economy, one that spans practically all the forms of platform-work.
-De Stefano also makes the argument for analyzing jointly the two main types of work in this realm, "crowdwork" and "work-on-demand via app." In crowdwork, workers complete job tasks through online platforms, such as Amazon Mechanical Turk, Crowdflower, and Clickworker, which connect an indefinite number of organizations and individuals through the internet.
-In "work-on-demand via app," traditional working activities like transport, cleaning, and running errands are channeled through apps managed by firms like Uber, Taskrabbit, and Handy, that also variously intervene in setting quality standards of service and in the selection and management of the workforce.}
-
 \end{document}

### 2017-01-12 chi/piecework/piecework_lit.tex
fixing final paper name and other stuff
actual forreal final version

--- 2017-01-11 drafts/labor/piecework_lit.tex
@@
 \topic{Piecework's history traces back further than most would likely expect.}
-\citeauthor{grier2013computers} describes the process astronomers adopted of hiring teenage men
+Grier describes the process astronomers adopted of hiring teenage men
 to calculate equations in order
@@
 a concise definition of piecework
-in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
-particularly driven toward encapsulating the manual labor of farmwork.}
+in Raynbird's essay on piecework,
+particularly driven toward encapsulating the manual labor of farm work.}
 He does this by contrasting two paradigms:
@@
 and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
-\citeauthor{10.2307/2338394} offers a number of illustrative examples:
+Chadwick offers a number of illustrative examples:
 ``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
@@
 additional pay depending on output~\cite{rowan1901premium}.
-As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
+As Rowan's premium system guaranteed an hourly rate
 regardless of the worker's productive output
@@
 
-In the late years of the \nth{19} century, \citeauthor{taylor1914principles}
+In the late years of the \nth{19} century, Taylor
 --- a mechanical engineer with an interest in work efficiency ---
 began studying and formalizing the decomposition, tracking, and management of tasks~\cite{taylor1896piece}.
-In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
+In \citeyear{taylor1914principles} he published \textit{The principles of scientific management},
 concretizing an idea that had nebulously been forming,
@@
 
-\ali{Somewhere here(?) I want to talk about Gibreth's works~\cite{gilbreth1911motion,gilbreth1911motion}. Should I save it to be the saving grace in the discussion?}
-
 \topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
-\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework,
+Raynbird argues for the merits of piecework,
 pointing out that
 ``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~\dots~he knows that all depends on his own diligence and perseverance~\dots~[and] so long as he performs his work to the satisfaction of his master, he is not under that control to which the day--labourer is always subject''.
-The argument that ``task--labourers'' enjoy freedom from control crops up in \citeauthor{hughRaynbirdTaskWork}'s and later \citeauthor{rowan1901premium}'s works~\cite{hughRaynbirdTaskWork,rowan1901premium}.
+The argument that ``task--labourers'' enjoy freedom from control crops up in Raynbird's and later Rowan's works~\cite{hughRaynbirdTaskWork,rowan1901premium}.
 
 \topic{We see this sense of independence in myriad times, locales, and industries.}
-\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls'', teenage women who assembled matchsticks in the late \nth{19} century in London.
+Satre offers a look into the lives and culture of ``match--girls'', teenage women who assembled matchsticks in the late \nth{19} century in London.
 % added that these women were in London since Chinmay critiqued that we're glossing over the labor advocacy movement that grew in Europe, which revealed that we completely left out the fact that this all happened in England.
-\ali{Chinmay: but still not convinced; this is a post hoc ergo propter hoc argument}
+% \ali{Chinmay: but still not convinced; this is a post hoc ergo propter hoc argument}
 Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
 but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
-\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
+Hagan and Fisher document piecework from 1850 through 1930 in Australia,
 finding similar notions of independence and autonomy among piecework newspaper compositors:
@@
 Between 1938 and 1942, the proportion of metal workers under piecework systems had climbed steeply from 11\% to 60\%~\cite{hart2005piecework}.
-By 1961, \citeauthor{carlson1982time} finds, the proportion dropped to 8\%~\cite{carlson1982time}.
+By 1961, Carlson finds, the proportion dropped to 8\%~\cite{carlson1982time}.
 He notes that, from 1973 to 1980, the holdouts of piecework
@@
 were principally in clothes--making (e.g. hosiery, footwear, and garments).
-\citeauthor{hart2013rise} offer a number of explanations for the sudden demise of piecework.
+Hart and Roberts offer a number of explanations for the sudden demise of piecework.
 % \ali{Should I add ``(especially in the United States but more broadly in the developed world)'' or something?
@@
 
-\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
+\ali{\citeauthor{clark1908cotton}} observed textile mill pieceworkers and reported,
 ``When he works by the day the Italian operative wishes to leave before the whistle blows,
@@
 the reality was more grim, as
-\citeauthor{riisOtherSideLives} documented in \textit{How the Other Half Lives} in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
+Riis documented in \textit{How the Other Half Lives} in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
 He found that
@@
 and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
-As \citeauthor{weyer1894history} described,
+As Webb and Webb described,
 ``the match--girls' victory turned a new leaf in Trade Union annals'': in the 30 years after the match--girls strike,
@@
 Match--girls were some of the earliest to have formed a trade union,
-according to \citeauthor{booth1903life}'s account in \citeyear{booth1903life}. % ~\cite{booth1903life}.
-\citeauthor{10.2307/3827491} noted that match--girls
+according to Booth's account in \citeyear{booth1903life}. % ~\cite{booth1903life}.
+Satre noted that match--girls
 ``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
-But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
+But they were also, as Satre confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
 These were workers who treasured their independence, but also fiercely protected one another. % , contributing to the common good.
@@
 to motivate and evaluate this skilled workforce of engineers.
-\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
+Graves takes up a case study of the Santa Fe Railway,
 finding that they employed ``efficiency experts'' to develop a ``standard time''
@@
 ``thousands of individual operations'';
-\citeauthor{10.2307/23702539} goes on to list
+Graves goes on to list
 some of the roles required to facilitate piecework
@@
 especially among engineering and metalworking industries.}
-\citeauthor{hart2013rise} characterize the 1930s
+Hart and Roberts characterize the 1930s
 --- and more broadly the first half of the \nth{20} century ---

### 2017-01-12 chi/piecework/relationships.tex
fixing final paper name and other stuff
actual forreal final version

--- 2017-01-11 drafts/labor/relationships.tex
@@
 Workers resent this position, and for good reason.
-\citeauthor{turkopticon} highlighted the information asymmetry between workers and requesters on AMT,
+Irani and Silberman highlighted the information asymmetry between workers and requesters on AMT,
 which led to the creation of Turkopticon, a site which allows Turkers to rate and review requesters~\cite{turkopticon}.
@@
 \topic{Early observers believed that workers were strongly motivated by the autonomy of working in the piecework model}.
-\citeauthor{clark1908cotton} observed textile mill pieceworkers and reported,
+Clark observed textile mill pieceworkers and reported,
 ``When he works by the day the Italian operative wishes to leave before the whistle blows,
@@
 when workers began instituting ``The Fix'', deliberately slow work to game efficiency experts~\cite{roy1954efficiency}.
-Piece workers, \citeauthor{roy1954efficiency} found, would form acrimonious relationships with their managers.
+Piece workers, Roy found, would form acrimonious relationships with their managers.
 
@@
 Doing this, 
-\citeauthor{ostrom1990governing} and others later argued,
+Ostrom and others later argued,
 would facilitate collective action and perhaps collective governance~\cite{ostrom1990governing,russell1982collective,olsonlogic}.
-\citeauthor{riisOtherSideLives} also contributed to this sense of shared struggle and endurance
+Riis also contributed to this sense of shared struggle and endurance
 by documenting pieceworkers in their home--workplaces,
@@
 While a person might recognize that the ``attention check'' questions
-proposed by \citeauthor{le2010ensuring} and others ensure that
+proposed by Le et al. and others ensure that
 malicious and inattentive workers are stopped~\cite{le2010ensuring,AAAIW113995},
@@
 only seem to antagonize workers~\cite{takingAHITMcInnis}.
-As \citeauthor{10.2307/2555446} wrote in \citeyear{10.2307/2555446},
+As Anderson and Schmittlein wrote in \citeyear{10.2307/2555446},
 ``when performance is difficult to evaluate,

### 2017-01-14 chi/piecework/complexity.tex
fixed a few bugs: - fixed a citation on page 4 that handled citeyear incorrectly (and other citeyear and citeauthor issues) - fixed a very weird sudden line break on page 5 (h/t Ranjay for spotting it during a read through) - fixed a smaller spelling error (it'ss -> it's) on page 6 (also Ranjay)

@@
 So, initially piecework arose for farm work, and as
-
 Raynbird and others discuss,
@@
 So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than piecework did.
-Improvements in workflows, measurement, and evaluation have already been made, and it'ss not immediately clear that the remaining challenges are readily solvable.
+Improvements in workflows, measurement, and evaluation have already been made, and it's not immediately clear that the remaining challenges are readily solvable.
 However, on--demand work will be far more broadly distributed than piecework historically was

### 2017-01-14 chi/piecework/relationships.tex
fixed a few bugs: - fixed a citation on page 4 that handled citeyear incorrectly (and other citeyear and citeauthor issues) - fixed a very weird sudden line break on page 5 (h/t Ranjay for spotting it during a read through) - fixed a smaller spelling error (it'ss -> it's) on page 6 (also Ranjay)

@@
 only seem to antagonize workers~\cite{takingAHITMcInnis}.
-As Anderson and Schmittlein wrote in \citeyear{10.2307/2555446},
+As Anderson and Schmittlein wrote in 1984,
 ``when performance is difficult to evaluate,

### 2017-01-14 drafts/aggregator/introduction.tex
fixed a few bugs: - fixed a citation on page 4 that handled citeyear incorrectly (and other citeyear and citeauthor issues) - fixed a very weird sudden line break on page 5 (h/t Ranjay for spotting it during a read through) - fixed a smaller spelling error (it'ss -> it's) on page 6 (also Ranjay)

@@
 
-Online labor markets ranging a variety of industries of (largely) service work
---- e.g. on--demand drivers (Uber, Lyft, Sidecar, etc\dots) and
-    information work (Amazon Mechanical Turk, Upwork, CrowdFlower, etc\dots) ---
-have struggled for some time with pernicious, in some cases \textit{growing},
-difficulty assessing the quality of its workforce,
-especially as it applies to a given task.
-In other words, whether a worker on Amazon Mechanical Turk (AMT) is likely
-to complete a task ``accurately'' remains an open question,
-leading researchers to address this pervasive problem in a number of ways
-\cite
+On--demand work has so far been a compelling way to arrange work and imagine workers;
+the fleeting nature of work seems to appeal greatly to workers, and
+the looser expectations of responsibility toward workers appeals to employers and clients.
+But in the exchange we've made for more fleeting relationships,
+we've instantiated --- or at least added fuel to ---
+another class of challenges and problems.
+Among them remains the challenge of ensuring that workers
+--- first crowd workers, and later gig workers --- both
+\begin{inlinelist}
+\item know the skills necessary to do the immediate task and
+\item will continue to produce work at the desired level of quality.
+\end{inlinelist}
+By far, two approaches have taken hold:
+First, qualification exams (e.g. Amazon Mechanical Turk and Upwork),
+       which ask workers to take a test of some sort to demonstrate that they both
+       understand the task in question \textit{have the ability} to do the task.
+Second, relying on outside certification bodies (e.g. Uber, \ali{do we know of others that match this profile?}),
+        which assumes that outside bodies
+        --- in Uber's case, local DMVs and similar bodies ---
+        have sufficiently evaluated a person's ability to do the tasks necessary.
+
+Other approaches exist, but have not (yet) garnered the interest of on--demand firms in general.
+
+% Research has looked into this question from a handful of perspectives;
+% some have tackled the qualifications problem as a pure assessment problem.
+% others have looked to
+% [\textit{Krishna et al. on long--term workers}]
+
+\ali{I want citations on \textbf{worker qualifications} and on \textbf{worker effort and management}.
+I think Ranjay's paper on looking forward etc\dots~at CSCW 2017 would be right here.
+When do bibtex things come out?}
+
+% Online labor markets ranging a variety of industries of (largely) service work
+% --- e.g. on--demand drivers (Uber, Lyft, Sidecar, etc\dots) and
+%     information work (Amazon Mechanical Turk, Upwork, CrowdFlower, etc\dots) ---
+% have struggled for some time with pernicious, in some cases \textit{growing},
+% difficulty assessing the quality of its workforce,
+% especially as it applies to a given task.
+% In other words, whether a worker on Amazon Mechanical Turk (AMT) is likely
+% to complete a task ``accurately'' remains an open question,
+% leading researchers to address this pervasive problem in a number of ways
 
@@
 
-\onlyinsubfile{
-% \balance{}
-\printbibliography{}
-}
-
 \end{document}

### 2017-01-20 drafts/aggregator/introduction.tex
very minimal introduction (literally just numbered list of maybe-topic sentences)

@@
 
+\ali{
+  Components of a good intro (anything else to add?):
+  \begin{itemize}
+    \item state of the world
+    \item what's the problem?
+    \item why this problem is such a big deal(?)
+    \item what's making this problem so persistent?
+    \item what's the insight we bring?
+    \item what we built/did
+    \item what did we find?
+  \end{itemize}
+}
+
+\begin{enumerate}
+  % state of the world
+  \item on--demand work has grown significantly
+  \item task diversity is growing
+  \item new platforms continue to emerge
+
+  % what's wrong?
+  \item workers are struggling to signal quality:
+  \begin{enumerate}
+     \item new workers --- cold start problem
+     \item old workers --- too much noise
+  \end{enumerate}
+  \item requesters have problems as well:
+  \begin{enumerate}
+    \onlyinsubfile{
+      \item did trump really quote Bane in his inaugural?
+      \item omg like really
+    }
+
+    % why is this problem a big deal?
+    \item requesters don't know whom to trust and have to bootstrap their own trustworthiness evaluations
+          (e.g. qualifications exams, tentatively and very cautiously trusting coarse heuristic data)
+    \item requesters tend to undervalue workers --- at least at first --- leading to a ``market for lemons''
+  \end{enumerate}
+
+  % Why is this so hard?
+  \item a few problems frustrate the design of interventions
+  \begin{enumerate}
+    \item platforms have little incentive to facilitate external agents of intervention
+    \item the stakeholders that platforms are more interested in seem to be those with the money (i.e. requesters)
+          \ali{maybe too argumentative?}
+    \item on--demand workers gravitate toward these platforms in large part because of the transience and ephemeral nature of the work
+          (therefore, working against that will meet with resistance (and possibly attrition) from the very workers who constitute the platform today)
+  \end{enumerate}
+
+  % What's our insight?
+  \item in ``conventional'' work, the problem of signaling a worker's quality was solved by using past worker performance as a proxy for future performance
+  \item on--demand work platforms have tried to do this with approval and rating systems, but they're generally too vague to be useful
+        % (on AMT, requesters limit work to those with approval ratings higher than 95\% or even 99\%, or workers with thousands or tens of thousands of tasks completed, or both)
+  \item conventional work r\'{e}sum\'{e}s usually allow the worker to emphasize and omit information that's not relevant --- \textbf{what if we facilitated that?}
+
+  % So what did we do?
+  \item we built a system that collected data from workers' profiles on online crowdsourcing platforms
+  (first AMT, then others), and parsed that data to present more meaningful analytic data on things like
+  \begin{enumerate}
+    \item the nature of the task; objective (i.e. factually based and evaluable) versus subjective (e.g. survey \ali{maybe also object labeling?})
+    \item the type of requester; academic versus industry
+    \item \textit{others}
+  \end{enumerate}
+  \item we also built a proof--of--concept qualifications management system abstracted from both our worker profile system and the work platform itself
+  (to explore the potential to abstract worker qualifications and credentials from the system itself, allowing other parties to specialize in this task)
+
+  % What did we find/how did it go?
+  \item we evaluated this system from two perspectives:
+  \begin{enumerate}
+    \item did requesters who used this system find that the worker data better--informed their qualifications management and worker solicitation?
+          in other words, was the output of the work more reliably accurate when they relied on the analytic data that we provided versus the coarse approval/rejection rating?
+    \item did \textit{workers} benefit from having high level data providing some insight into their work trends?
+did this lead them to better work more quickly?
+did they take a more active role in the management of their reputation / ``worker profile'' / r\'{e}sum\'{e}?
+  \end{enumerate}
+  \item we also explored whether the externalized worker qualification system worked effectively, and considered the potential advantages and disadvantages of a set of decoupled systems rather than an integrated one (market, qualifications assessment system, payment platform, etc\dots).
+\end{enumerate}
+
+\ali{
 On--demand work has so far been a compelling way to arrange work and imagine workers;
@@
 
-Other approaches exist, but have not (yet) garnered the interest of on--demand firms in general.
+Other approaches exist, but have not (yet) garnered the interest of on--demand firms in general.}
 

### 2017-02-06 admin/CURIS2017/curis.tex
CURIS proposal

\documentclass[12pt]{article}
\usepackage{balance,times,graphics,setspace,microtype}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage{hyperref}

\setlength{\parskip}{.4em}
\setlength{\parindent}{0cm}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
\pagenumbering{gobble}

\begin{document}
\section{Project Name}

Making a ``R\'{e}sum\'{e} for Gig Workers''

\section{Description}

Gig workers are increasingly
piecing together earnings from various work platforms to make ends meet.
But insight into their work performance and history
are virtually nonexistent.
Requesters
--- the people that hire workers ---
increasingly struggle with existing qualifications and credentials services,
which lack sufficient detail about worker quality, driving them to undervalue workers.
Requesters don't get the information they need to make confident, informed decisions, and
as a result they're reluctant to trust gig workers to pay for creative, complex work.

This summer we'll be building a \textit{gig worker r\'{e}sum\'{e}} service,
an online system that aggregates
workers' histories, provides analytics on their performance over time, and helps requesters find high--quality workers.
You'll play a major part in designing and implementing this system,
involving web scraping, some statistical analysis, data visualization, and potentially more.
You'll also have the opportunity to study how people use the system,
using a mix of qualitative and quantitative analysis,
to inform our next steps.

\section{Recommended Background}\label{sec:RecommendedBackground}

Ideally, you have experience in
Human--Computer Interaction
(think CS 147, 247).
Much of the day--to--day work will involve programming for the web
(think CS 142),
so regardless of your research goals you should be able to code
(and be reasonably competent with existing libraries, frameworks, and APIs).

It might be worth pointing out that
I don't expect you to be an expert in these things right now.
You should be competent enough to start working on day 1, but
this is a chance to hone your skills and maybe learn something completely new.

If your background doesn't fit with anything above,
please tell us the most ambitious thing you've undertaken,
academic or not!
(Hard work can overcome lack of preparation!)

\section{Number of students allowed}

2

\section{Prerequisite/Preparation}

If you feel unprepared in any of the areas listed under the
``\nameref{sec:RecommendedBackground}'' section,
then your preparation between now and the start of the program will be to
try to fill in gaps.
If you want, we can discuss this further and start to prepare for your time through spring quarter.

\bibliographystyle{acm}
\bibliography{../references.bib}
\end{document}

### 2017-02-06 presentations/hasso_talk/hasso_talk.tex
hasso presentation

% !TEX program = xelatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{beamer}
\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth}
% \usepackage[sfdefault]{FiraSans} %% option 'sfdefault' activates Fira Sans as the default text font
% \usepackage[T1]{fontenc}
% \usepackage[T1]{fontenc}
\usepackage[export]{adjustbox}
% \usepackage[ddmmyyyy]{datetime}
 \usepackage[en-US]{datetime2}

\beamertemplatenavigationsymbolsempty

\mode<presentation>{
  \usetheme{metropolis}
  \setbeamercolor{institute in head/foot}{fg=stanfordRedText}
  \setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=stanfordRed}
}

\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}

\author{\textbf{Ali Alkhatib}, Margaret Levi, Michael Bernstein\\
\texttt{ \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib} }}

\institute[Stanford]{Stanford University}

\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

% Notes
% Hi everyone. My name's Ali, and I want to tell you about the work on ``Cooperative Labor Markets'' that I've been doing for the past several months with a number of collaborators.

\begin{frame}{A New Form of Work: On--Demand Labor}
  \begin{itemize}[<+- | alert@+>]
    \item Crowdsourcing
    \begin{figure}
    \includegraphics[scale=0.3]{figures/amt.png}
    \includegraphics[scale=0.1]{figures/upwork.png}
    \end{figure}
    \item Gig work
    \begin{figure}
    \includegraphics[scale=0.1]{figures/uber.png}~~~~~
    \includegraphics[scale=0.15]{figures/handy.png}
    \end{figure}
    % \item (Not \textbf{so} obviously) labor unions might be problematic:
    %   \\ --- logistics are complicated (solvable)
    %   \\ --- cultural mismatch
    %     \cite{dynamo}
  \end{itemize}
\end{frame}

\begin{frame}{Growing Number of Questions}
\begin{itemize}[<+- | alert@+>]
  \item What are the complexity limits of on--demand work?
  \item How far can work be decomposed into smaller microtasks?
  \item What will work and the place of work look like for workers?
\end{itemize}

\end{frame}

\section{Piecework as a lens to understand on--demand work}

% \section{What is Piecework?}
\begin{frame}{A (\textit{quick}) Primer on Piecework}
  \begin{itemize}[<+- | alert@+>]
    \item Payment made for \textit{output}, rather than for \textit{time}
  \end{itemize}
\end{frame}

\begin{frame}{An Example: Pieceworkers at the turn of the \nth{20} century}
\begin{figure}
    \includegraphics[scale=0.27]{figures/pieceworkers.jpg}  
\end{figure}
\end{frame}

\begin{frame}{One Question: The Limits on Task Decomposition}
  \begin{itemize}[<+- | alert@+>]
  \item Piecework: Lack of instrumentation limited task decomposition
  \item Today? Tools and technology can track every move workers make
  \item We've hit upon what seems to be an atomic limit on decomposition ---
        cognitive switching between tasks is becoming the main barrier
  \end{itemize}
\end{frame}

\begin{frame}{Future Directions}
  \begin{itemize}[<+- | alert@+>]
  \item If on--demand labor follows the same trajectory as piecework, what can we expect?
  \item If we can design the entire platforms of labor in ways we couldn't imagine a century ago, what can we do to affect different outcomes?
  \end{itemize}
\end{frame}

% \begin{frame}{Thank You}
%   \author{\textbf{Ali Alkhatib}, Margaret Levi, Michael Bernstein\\
% \texttt{ \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
%          \href{http://twitter.com/_alialkhatib}{@\_alialkhatib} }}
% \end{frame}
% Notes
% Now, one path to take is to attempt to reform labor unions so that they better reflect the needs of workers, and particularly the threats that they face. That's perfectly valid, but we've seen this as an opportunity to think more diversely about design solutions that mitigate the adversarial nature of these markets from the outset.
% This thought has led us to worker cooperatives.

% \begin{frame}{Worker Cooperatives}
%   \begin{itemize}[<+- | alert@+>]
%     \item Worker cooperatives are reasonably well--explored
%       \cite{craig1992behavior,mellor1988worker}
%     \item \textbf{Collective governance} is a challenge
%       \cite{russell1982collective,ostrom1990governing,polletta2002freedom}
%   \end{itemize}
% \end{frame}

% Notes
% Worker cooperatives are actually fairly well explored and some of their advantages are clear, as are potential shortcomings and challenges to watch out for.
% Namely, managing collectively governed systems like these is challenging, and the nature of the challenge seems to change as the organization grows, so the solutions that worked once no longer work down the road.

% \begin{frame}{What is piecework?}

%   \begin{figure}
%     \includegraphics[scale=0.12]{figures/figure.png}  
%   \end{figure}
% \end{frame}

% Notes
% Working together with the NDWA, we're building an online worker cooperative for house cleaners that provides workers with useful tools to manage their work, giving them a starting point to discuss how they want this system to run, making decisions collectively.

% \section{Lots of open questions remain}

% % Notes
% % There are still many questions that remain unanswered, and even questions that have yet to be asked, but I think this approach is the right one to begin the ``anthropological investigation'' of collective governance.

\begin{frame}
  \frametitle{Contact}
    name: \href{https://ali-alkhatib.com}{Ali Alkhatib} \\
    % human: find me walking around this week \\
    email: \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} \\
    twitter: \href{https://twitter.com/_alialkhatib}{@\_alialkhatib} \\
\end{frame}

% % Notes
% % I'm over time [just guessing, but really], but if you have any questions, or criticism, or suggestions, I would really like to hear from you. My contact information is all there, so please approach me so we can talk in further depth about these topics - Thanks!

% \begin{frame}[allowframebreaks]{References}
%   \bibliography{references}
%   \bibliographystyle{abbrv}
% \end{frame}

\end{document}

### 2017-03-06 proposals/SIGFproposal_2016/ali_alkhatib_sigf.tex
sigf17

% !TEX program = xelatex
\documentclass{article}
\usepackage{balance,graphics,setspace,hyperref,etoolbox,refcount}
\usepackage{graphicx,sidecap,fancyhdr,titling,blindtext,fontspec}
\usepackage{xltxtra,indentfirst,xunicode,xesearch}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[square,numbers]{natbib}
\bibliographystyle{newapa}

\titlespacing*{\section}
{0pt}{4pt}{0pt}

\titleformat*{\subsection}{\it}

\setmainfont[
Ligatures=TeX,
Extension=.otf,
UprightFont= *-regular,
BoldFont=*-bold,
ItalicFont=*-italic,
BoldItalicFont=*-bolditalic]{texgyretermes}

\setmonofont{Source Code Pro}

\newcounter{countitems}
\newcounter{nextitemizecount}
\newcommand{\setupcountitems}{%
  \stepcounter{nextitemizecount}%
  \setcounter{countitems}{0}%
  \preto\item{\stepcounter{countitems}}%
}
\makeatletter
\newcommand{\computecountitems}{%
  \edef\@currentlabel{\number\c@countitems}%
  \label{countitems@\number\numexpr\value{nextitemizecount}-1\relax}%
}
\newcommand{\nextitemizecount}{%
  \getrefnumber{countitems@\number\c@nextitemizecount}%
}
\makeatother
\AtBeginEnvironment{inlinelist}{\setupcountitems}
\AtEndEnvironment{inlinelist}{\computecountitems}
\AtBeginEnvironment{enumerate}{\setupcountitems}
\AtEndEnvironment{enumerate}{\computecountitems}
\AtBeginEnvironment{itemize}{\setupcountitems}
\AtEndEnvironment{itemize}{\computecountitems}

\pagestyle{fancy}

\newcounter{words}
\newenvironment{counted}{%
  \setcounter{words}{0}
  \SearchList!{wordcount}{\stepcounter{words}}
    {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
    n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
  \UndoBoundary{'}
  \SearchOrder{p;}}{%
  \StopSearching}

\lhead{} % controls the left corner of the header
% \chead{} % controls the center of the header
\rhead{Ali Alkhatib} % controls the right corner of the header
\lfoot{} % controls the left corner of the footer
\cfoot{} % controls the center of the footer
\rfoot{Page~\thepage} % controls the right corner of the footer
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}

% \thispagestyle{empty}% Removes the header from the first page. Change plain to empty to remove the numbering entirely.

\setlength{\parskip}{.2em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}

\setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
\setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
\newcommand{\sectitle}[1]{\textit{#1}}
\title{Designing for Collective Governance in the Gig Economy}
\author{Ali Alkhatib}
\date{Submitted \today}

\begin{document}
\begin{titlingpage}
    \maketitle
    % \begin{counted}
    \begin{abstract}
Digitally mediated marketplaces have led to a resurgence of ``gig'' work,
wherein workers are connected by the Internet rather than geography.
Without shared physical spaces,
disenfranchised workers struggle to coordinate and take collective action. I will engage in fieldwork with digitally managed workers in information work,
domestic services,
and on--demand driving to understand how workers resist algorithmic systems and act collectively. I will then design and launch a new cooperatively--governed platform for work,
in collaboration with the National Domestic Workers Alliance (NDWA),
to develop frameworks and tools to facilitate collective action and governance.
    \end{abstract}
    % \end{counted}
% \footnotesize{word count: \thewords~(100 max)}
\end{titlingpage}

\section*{Background}

Today, with the growth of digitally mediated systems to manage workers and decontextualized pieces of work,
\textit{gig} work
--- characterized by the discrete jobs (or \textit{gigs}) which workers agree to do  ---
has increasingly become the focus of public and academic interest
\citep{martin2014being,uberAlgorithm,hall2015analysis}.
But gig work platforms have achieved some level of notoriety as settings of disempowerment
fueled by information asymmetry and opaque algorithmic management;
while researchers have made efforts to empower workers,
perhaps most notably on Amazon Mechanical Turk (AMT),
these efforts have thus far only ameliorated,
rather than resolved,
tension between
the various stakeholders of these systems
\citep{turkopticon,dynamo}.
This persisting frustration among \textit{Turkers} (the colloquial name for workers on AMT) parallels the frustration
workers in the \textit{gig} economy
--- namely, on--demand services such as Uber and Handy
\citep{dillahunt2016does,uberRiots,fedsUber,hall2015analysis}.

With gig work increasingly representing the labor paradigm under which people work,
and with this mode of labor being exported to other parts of the world,
it is not only important that we attempt to understand the culture of those engaged in the discretization of work,
but that we attempt to participate in this process with design interventions
to improve the circumstances of workers by empowering them.

Working with the National Domestic Workers Alliance (NDWA),
I began to explore methods of empowering workers and ensuring their benefits, safety, and well--being.
Historically, labor unions have advocated on behalf of workers,
typically representing an industry or company's workforce.
In the case of gig work,
the geographic diffusion associated with digitally mediated systems
makes collective organization difficult.
More importantly, however,
the findings we uncovered among Turkers during our work on Dynamo
suggests that gig workers are reluctant regarding what they perceive labor unions to represent.
As one Turker described the sentiment I found more generally,
``\dots~I would not feel comfortable taking part [in a labor union].
It runs against my grain because I am an individualist.
[\dots] I consider myself self--employed\dots~not working for anyone in particular''
\citep{dynamo}.

This cultural disconnect between workers and the perceived qualities of labor unions presented us with an opportunity
to explore more creative ways of dealing with the adversarial relationships between workers and managers.
In this search, we found that \textit{worker cooperatives}
--- which have struggled to reach large scales in the past in part due to the mounting struggle associated with growing collectively governed groups ---
may in fact be feasible given the unique features of the Internet, 
described by Miller and others,
which make it possible for people to collaborate rapidly and cheaply,
obviating many of the concerns that typically stymie efforts to organize,
such as geography
\citep{miller2011understanding}.

I propose to conduct ethnographic research on gig work in digitally mediated platforms,
through the historical lens of \textit{piece work},
to better understand the characteristics that influence and affect the culture of gig workers today.
The ethnographic research we conduct will yield an \textit{emic} understanding of
the values gig workers hold and their sense of self as they relate to the work they do.
With this ethnographic data,
I will then design and build a worker cooperative in collaboration with gig workers and the NDWA.
This system will reflect on the ethnographic findings we will have made from the first study,
as well as test guidelines toward enabling and fostering collective governance strategies.

While researchers have contributed to our understanding of worker cooperatives and participatory democracy,
we nevertheless find that worker cooperatives struggle to grow to the scales that firm--owned companies do.
Hardin, Olson, and others offer general framing of collective governance,
especially as it relates to the instances of collective action I explored in our work on Dynamo
\citep{russell1982collective,olsonlogic}.
Ostrom in particular provides substantive guidance on the mechanisms necessary to facilitate collective governance
\citep{ostrom1990governing}.

Ultimately, while we know how to design and engineer systems for collaboration,
and in particular we can generate instances of collective action where the target action is finite and focused,
we do not yet know how to design online systems for the sustained collective governance that we see only occasionally
(e.g. Wikipedia)
\citep{catalyst,dynamo,foundry}.
The core questions in this research, exploring the mechanisms of collective governance on the Internet, are social ones.

\section*{Research plan}
As stated previously,
a critical component of the research proposal consists of
understanding the culture of gig workers.
Indeed, this proposal consists of two distinct research endeavors:
\begin{inlinelist}
\item Ethnographic study of gig workers, to understand the ways that workers do and indeed may support one another, and
\item design interventions consisting of the creation of a cooperative labor market designed collaboratively with workers.
\end{inlinelist}

\subsection*{Study 1: Ethnographic study of gig workers}
The first study will consist of ethnographic fieldwork studying workers in modern piece work labor markets 
--- e.g., Mechanical Turk, Uber, and Handy ---
to understand how social coping strategies emerge among workers.
How do workers air their grievances?
How do they come together to push for change?
How have they adapted to the working conditions of these digital hiring halls?

I will perform one year of fieldwork with digital piece workers.
To understand their breadth of experience, I will sample roughly fifty workers from information work
(for example,
Mechanical Turk and CrowdFlower),
and another fifty workers in the ``sharing economy''
(including Uber,
Lyft,
and AirBnB),
and fifty in gig work
(namely Handy and TaskRabbit).
Given existing relationships and reputations as credible researchers in communities of Turkers, I can recruit participants through known existing channels.
In the cases of platforms such as Uber,
AirBnB,
and Handy, I can recruit workers for nominally legitimate work
(for example,
requesting that a driver drives for 15 minutes to a previously selected destination)
and invite them to participate in our interview.

\subsection*{Study 2: Design and engineering of a cooperative labor market}
The second component of this research will involve
the engineering and launch a cooperatively owned ``gig" marketplace for domestic workers
in collaboration with the National Domestic Workers Alliance
(NDWA).
We will contribute the first digitally mediated gig labor market collaboratively designed with
workers and labor advocacy groups.
% My prior work established a collective action platform for crowd workers
% \citep{dynamo}.
Informed by my work on Dynamo and the ethnographic fieldwork in the first study,
I will then seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
The NDWA's Fair Care Labs are providing access to interested domestic workers
(house cleaners), and we are designing and implementing the platform with a pilot group.

The goal of this research is
to understand how to design for a set of distributed workers to achieve consensus on issues such as
management, service fees, and arbitration.
Through the creation of this labor market and platform, we will explore and test theories behind
enabling collective governance, consensus--building, and participatory democracy.
This study will yield guidelines for system--designers to spark and foster these values.

Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction.

\section*{Interdisciplinarity}
For this research to succeed, the methodological approaches of Political Science
and the engineering skills of Computer Science must not only be in conversation, but in concert.
The challenge in designing for collective governance,
as Hardin (1982) differentiates it from collective action,
is such that collective governance
``takes on the character of a nearly anthropological investigation of minute interrelationships''
\citep{russell1982collective}.
In order to design a system that empowers workers,
a researchers must first attempt to understand a wide range of workers' circumstances:
The needs of workers;
the contexts in which they work;
their relationships with one another, with other groups, \& with institutions such as governments;
and more generally the paradigmatic views of gig workers.
Only then can one reasonably hope to design a system consistent with the views and broader culture of gig work.

Lee et al. and Gray et al. have identified a number of ways that workers subvert and circumvent the intents of system--designers,
both in digital workplaces and where work is simply mediated digitally
\citep{uberAlgorithm,crowdcollab}.
These patterns of behavior elude algorithmic tracking and measurement because they deliberately avoid the \textit{a priori} assumptions made by the designers of systems,
who attempt to structure these sites of work in ways to create incentives for preferred behavior.
Identifying and understanding the details of this behavior thus begins with a qualitative,
ethnographic endeavor.

\section*{Anticipated Contributions \& Implications}
This research is important given the recent trends toward the discretization of work.
With increasing numbers of people joining the gig economy,
and lacking sufficient work toward digitally mediated forms of collective action and governance,
disenfranchised and marginalized laborers in particular stand to lose significantly.
Indeed, advances made by labor advocacy groups in the past century have already eroded
due to the classification of many of these workers as ``contractors'' rather than ``employees'',
signaling the larger trend of the reversal of benefits and protections for which workers have fought.

More broadly, this project's will speak to the critique that the Internet,
far from the democratizing force that Barlow predicted it would be,
has become an infrastructure facilitating the corporate centralization of power
\citep{barlow2009declaration,jones2011does,EboCybertopia}.
The success of this project will therefore not only
bring a more nuanced understanding of the shifting climate and culture of labor,
but may
--- in a small way ---
deliver on the promise Barlow made that
cyberspace and the Internet would be a tool for democratizing access to information, and with that,
empowering more than the tech elite who engineered it.
% but will add to the increasingly minority claim that the Internet should be a democratizing force,
% rather than a centralizing one;
% \citep{barlow2009declaration}.

% To ensure that system--designers have reasonable access to our findings,
% our research targets will be top--tier venues in Human--Computer Interaction
% which support open access publication paradigms
% (e.g., ACM CHI, ACM CSCW).
% A second set of outlets will be through the NDWA's relationships with labor advocates and the press.

\pagebreak

% \bibliographystyle{acm}
\bibliography{../../references.bib}
\end{document}

### 2017-03-06 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
sigf17

\documentclass{article}
\usepackage{xr-hyper}
\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase}
\usepackage[normalem]{ulem}
\usepackage{url}
% \usepackage{natbib}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[T1]{fontenc}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{mfirstuc,lastpage}
\usepackage[en-GB]{datetime2}
\usepackage{fancyhdr,tabularx}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only

\usepackage{mfirstuc,lastpage}
\bibliography{references}
\usepackage[margin=1in]{geometry}
\pagestyle{fancy}

\lhead{} % controls the left corner of the header
% \chead{} % controls the center of the header
\rhead{Ali Alkhatib} % controls the right corner of the header
\lfoot{} % controls the left corner of the footer
\cfoot{} % controls the center of the footer
\rfoot{Page~\thepage} % controls the right corner of the footer
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}

\thispagestyle{empty}% Removes the header from the first page. Change plain to empty to remove the numbering entirely.

% \setlength{\parskip}{.2em}

% \newlist{inlinelist}{enumerate*}{1}
% \setlist*[inlinelist,1]{
%   label=\arabic*),
% }

% \setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
% \setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
% \newcommand{\sectitle}[1]{\textit{#1}}
\title{Designing for Complex Gig Work and Careers}
\author{Ali Alkhatib}
\date{Submitted \today}

% take half 
% motivating work: historical analysis of crowd 
% gig work instead of on--demand work
%
%
%
%
\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
\definecolor{Gray}{HTML}{BABABA}
% \newcommand{\blah}[1]{{\color{Gray}\textit{#1}}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

\begin{document}
  % \maketitle
  \begin{abstract}
On--demand labor has become a new instantiation of piecework.
Without shared factories and water coolers,
how do these workers coordinate,
build solidarity,
and take collective action?
We will engage in fieldwork with pieceworkers in data entry,
domestic services,
and on--demand driving to understand how they counter algorithmic systems and engage in collective action.
Our goal is to both understand the factors that drive collective behavior in supposedly distributed labor platforms,
and to build infrastructures that enable the growth of worker collective action in digitally mediated on--demand labor.
Our results shed light on the potential for digital piecework cooperativism,
as well as the policies necessary for it to succeed. % ~108 words
  \end{abstract}

\section*{Motivating Work: Historical Analysis of Piecework}
% \topic{The past decade has seen a flourishing of computationally--mediated labor.}
% A framing of work into modular, pre--defined components
% enables computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
% In this regime, distributed workers engage in work whenever their schedules allow,
% often with little to no awareness of the broader context of the work, and
% often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.

\topic{Piecework has historically disempowered workers and empowered employers.}
Subdividing work and paying per task drives employees to work faster and longer,
affording them little control over working conditions.
For most of its history, piecework offered no benefits, occupational health and safety, or social insurance.
When piecework combined with job routinization, workers became interchangeable and deindividuated.
This combination of piecework and routinization produced the dominant payment system during the industrial era,
starting with agriculture and home production but quickly moving into factories.
However, by the turn of the 20th century in the United States,
campaigns for workers' rights yielded regulations on conditions, and by the late 20th century,
piecework was associated primarily with migrant labor and sweatshops.
This contracted piecework economy has remained relatively stable since.

\topic{Today, with the growth of on--demand labor, piecework is reemerging in new forms.}
Networked computational infrastructure is now mediating each unit of work and worker.
Uber's taxi drivers are paid per ride and assigned jobs by an algorithm~\cite{uberAlgorithm,hall2015analysis}.
Information workers, such as those on Amazon Mechanical Turk and Upwork, are paid per task, performing vast quantities of data and administrative work~\cite{martin2014being}.
Increasingly numerous workers make money through one--off tasks like housecleaning and food delivery, all mediated by online platforms.
Current legal and policy frameworks classify these workers as independent contractors, precluding them from accessing employee benefits and job security associated with employment.
Complicating matters, today's workers often despise unfair employers but prize the flexibility and autonomy afforded by their platforms~\cite{martin2014being}.
Many view themselves as masters of their own fate, free to enter and leave the workforce at will and without repercussion.
Despite this, they still seek collective representation, as Uber drivers have done repeatedly to combat falling rates.

\topic{How do workers react to, and create collective counterbalances to, the digital piecework economy?}
On the one hand, studies of online communities predict that workers online will face greater challenges, because
trust is more difficult to build in digitally mediated environments~\cite{successfulOnlineCommunities,kollock2005managing,cook2005cooperation}.
Without factory floors, water coolers, and shared neighborhoods~\cite{waterCooler},
how do workers coordinate and build solidarity?
On the other hand,
online forums may make it easier for workers to share information and strategies,
and successful collaborative efforts such as Wikipedia suggest that the online environment may support forms of cooperation that could not have succeeded as readily offline.

\topic{As yet,
we have limited knowledge of how online pieceworkers could adapt to contemporary labor markets.}
We do not know how workers are engaging in coordinated and cooperative actions,
nor what form they should take in an era of digital work.
Are their efforts similar to historical behaviors for offline pieceworkers,
such as unions and worker cooperatives?
Or do the affordances of the online world result in different outcomes?
Furthermore,
could technological affordances grant pieceworkers the ability to organize more effectively,
and even to run their own marketplaces?

We propose paired grounded ethnographic fieldwork and system development to 1) understand the forces shaping informal and formal worker reactions to the online piecework economy,
and 2) create a technical and design foundation for collective action,
ownership and governance in online piecework.

First, we will analyze worker behavior in the current sociotechnical ecosystem through a year of fieldwork on major Mechanical Turk
(data entry)
platforms and forums,
and on gig economy platforms such as Uber
(driving)
and Handy
(domestic work).
Through this fieldwork, we will investigate the techniques that workers use to manage their work environments and job opportunities.
We will seek to understand the challenges they face when balancing
an identity of being self--employed with an identity of being a pieceworker for a privately--owned labor platform, and
the techniques that workers and employers use to circumvent each other's restrictions, barriers, and behaviors.

Second, we will create and launch a cooperatively owned ``gig" marketplace for domestic workers in collaboration with the National Domestic Workers Alliance
(NDWA).
Our own prior work established a collective action platform for crowd workers~\cite{dynamo}.
We now seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
The NDWA's Fair Care Labs are providing access to interested domestic workers
(house cleaners), and we are designing and implementing the platform with them as a pilot group.
The goal is to understand how to design for a set of distributed workers to achieve consensus on issues such as management, service fees, and arbitration.

Our team is uniquely qualified to pursue this opportunity.
Levi has years of experience studying labor, collectives, and unions.
She is pairing with Bernstein, a computer scientist who studies online labor platforms and collective action.
Bernstein, alongside PhD students Salehi and Alkhatib, created the first platform for collective action efforts amongst digital pieceworkers, focused on the Amazon Mechanical Turk platform.
This effort has galvanized hundreds of workers and resulted in outcomes including public media campaigns and the creation of ethical work guidelines for Mechanical Turk employers.

This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.
What affordances and protections do workers need in order to make their voices heard and to have collective power?
What protections should platforms be required to provide to their workers?
What support is necessary for people who work on multiple platforms, frequently moving between different employers?
Are unions the answer, or some other form of worker organization?
% Today, with the growth of digitally mediated systems to manage workers and decontextualized pieces of work,
% \textit{gig} work
% --- characterized by the discrete jobs (or \textit{gigs}) which workers agree to do  ---
% has increasingly become the focus of public and academic interest
% \cite{martin2014being,uberAlgorithm,hall2015analysis}.
% But gig work platforms have achieved some level of notoriety as settings of disempowerment
% fueled by information asymmetry and opaque algorithmic management;
% while researchers have made efforts to empower workers,
% perhaps most notably on Amazon Mechanical Turk (AMT),
% these efforts have thus far only ameliorated,
% rather than resolved,
% tension between
% the various stakeholders of these systems
% \cite{turkopticon,dynamo}.
% This persisting frustration among \textit{Turkers} (the colloquial name for workers on AMT) parallels the frustration
% workers in the \textit{gig} economy
% --- namely, on--demand services such as Uber and Handy
% \cite{dillahunt2016does,uberRiots,fedsUber,hall2015analysis}.

% With gig work increasingly representing the labor paradigm under which people work,
% and with this mode of labor being exported to other parts of the world,
% it is not only important that we attempt to understand the culture of those engaged in the discretization of work,
% but that we attempt to participate in this process with design interventions
% to improve the circumstances of workers by empowering them.

% Working with the National Domestic Workers Alliance (NDWA),
% I began to explore methods of empowering workers and ensuring their benefits, safety, and well--being.
% Historically, labor unions have advocated on behalf of workers,
% typically representing an industry or company's workforce.
% In the case of gig work,
% the geographic diffusion associated with digitally mediated systems
% makes collective organization difficult.
% More importantly, however,
% the findings we uncovered among Turkers during our work on Dynamo
% suggests that gig workers are reluctant regarding what they perceive labor unions to represent.
% As one Turker described the sentiment I found more generally,
% ``\dots~I would not feel comfortable taking part [in a labor union].
% It runs against my grain because I am an individualist.
% [\dots] I consider myself self--employed\dots~not working for anyone in particular''
% \cite{dynamo}.

% This cultural disconnect between workers and the perceived qualities of labor unions presented us with an opportunity
% to explore more creative ways of dealing with the adversarial relationships between workers and managers.
% In this search, we found that \textit{worker cooperatives}
% --- which have struggled to reach large scales in the past in part due to the mounting struggle associated with growing collectively governed groups ---
% may in fact be feasible given the unique features of the Internet, 
% described by Miller and others,
% which make it possible for people to collaborate rapidly and cheaply,
% obviating many of the concerns that typically stymie efforts to organize,
% such as geography
% \cite{miller2011understanding}.

% I propose to conduct ethnographic research on gig work in digitally mediated platforms,
% through the historical lens of \textit{piece work},
% to better understand the characteristics that influence and affect the culture of gig workers today.
% The ethnographic research we conduct will yield an \textit{emic} understanding of
% the values gig workers hold and their sense of self as they relate to the work they do.
% With this ethnographic data,
% I will then design and build a worker cooperative in collaboration with gig workers and the NDWA.
% This system will reflect on the ethnographic findings we will have made from the first study,
% as well as test guidelines toward enabling and fostering collective governance strategies.

% While researchers have contributed to our understanding of worker cooperatives and participatory democracy,
% we nevertheless find that worker cooperatives struggle to grow to the scales that firm--owned companies do.
% Hardin, Olson, and others offer general framing of collective governance,
% especially as it relates to the instances of collective action I explored in our work on Dynamo
% \cite{russell1982collective,olsonlogic}.
% Ostrom in particular provides substantive guidance on the mechanisms necessary to facilitate collective governance
% \cite{ostrom1990governing}.

% Ultimately, while we know how to design and engineer systems for collaboration,
% and in particular we can generate instances of collective action where the target action is finite and focused,
% we do not yet know how to design online systems for the sustained collective governance that we see only occasionally
% (e.g. Wikipedia)
% \cite{catalyst,dynamo,foundry}.
% The core questions in this research, exploring the mechanisms of collective governance on the Internet, are social ones.

\section*{Research plan}
\begin{blah}
We intend to create a platform that can become a template for online work that better represents workers' needs.
However, solutions that exist for traditional, non-gig workers
--- especially labor unions ---
face resistance from workers in the gig economy~\cite{martin2014being,dynamo}.
This distaste may be due to a general decline in public opinion of unions.
However, the traditional mechanics of a labor union are also a poor match: challenging issues include managing the benefits and proportional representation of workers whose contracts may last only minutes or hours, enforcing collective decisions on a decentralized network of workers, and establishing trust between workers who may never meet face--to--face.
In our early fieldwork, for example, a worker on Amazon Mechanical Turk stated:
``If by `union' you mean a `labor union', I would not feel comfortable taking part.
It runs against my grain because I am an individualist.
[\dots] I consider myself self--employed\dots\  not working for anyone in particular.''

Given these constraints, what alternative model would give workers more power in an online gig--economy platform while still enabling a market to emerge and function? To explore one such alternative, we will collaborate with the National Domestic Workers Alliance's innovation arm, the Fair Care Labs, to create tools for online pieceworkers to collectively manage and govern their own labor markets.
While the literature in social computing design and crowdsourcing already features many tools for successful decentralized collaborations (e.g., Wikipedia), and collective action movements (e.g., change.org), templates and guidelines for sustained collective action and governance online in high--stakes situations that affect workers' livelihoods are rare.
Such efforts tend to swell quickly, but soon find themselves embroiled as they struggle to satisfy myriad stakeholders .
Our goal is to introduce design mechanisms to enable collectives of gig workers to form, debate policy, and implement their decisions over a long period of time.

Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction. The web and mobile application, Alia, enables the basic affordances of ``gig'' work for domestic work
--- rapid scheduling of any available worker in the network (Figure 1).
\end{blah}

% As stated previously,
% a critical component of the research proposal consists of
% understanding the culture of gig workers.
% Indeed, this proposal consists of two distinct research endeavors:
% \begin{itemize}
% \item Ethnographic study of gig workers, to understand the ways that workers do and indeed may support one another, and
% \item design interventions consisting of the creation of a cooperative labor market designed collaboratively with workers.
% \end{itemize}

% \subsection*{Study 1: Ethnographic study of gig workers}
% The first study will consist of ethnographic fieldwork studying workers in modern piece work labor markets 
% --- e.g., Mechanical Turk, Uber, and Handy ---
% to understand how social coping strategies emerge among workers.
% How do workers air their grievances?
% How do they come together to push for change?
% How have they adapted to the working conditions of these digital hiring halls?

% I will perform one year of fieldwork with digital piece workers.
% To understand their breadth of experience, I will sample roughly fifty workers from information work
% (for example,
% Mechanical Turk and CrowdFlower),
% and another fifty workers in the ``sharing economy''
% (including Uber,
% Lyft,
% and AirBnB),
% and fifty in gig work
% (namely Handy and TaskRabbit).
% Given existing relationships and reputations as credible researchers in communities of Turkers, I can recruit participants through known existing channels.
% In the cases of platforms such as Uber,
% AirBnB,
% and Handy, I can recruit workers for nominally legitimate work
% (for example,
% requesting that a driver drives for 15 minutes to a previously selected destination)
% and invite them to participate in our interview.

% \subsection*{Study 2: Design and engineering of a cooperative labor market}
% The second component of this research will involve
% the engineering and launch a cooperatively owned ``gig" marketplace for domestic workers
% in collaboration with the National Domestic Workers Alliance
% (NDWA).
% We will contribute the first digitally mediated gig labor market collaboratively designed with
% workers and labor advocacy groups.
% % My prior work established a collective action platform for crowd workers
% % \cite{dynamo}.
% Informed by my work on Dynamo and the ethnographic fieldwork in the first study,
% I will then seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
% The NDWA's Fair Care Labs are providing access to interested domestic workers
% (house cleaners), and we are designing and implementing the platform with a pilot group.

% The goal of this research is
% to understand how to design for a set of distributed workers to achieve consensus on issues such as
% management, service fees, and arbitration.
% Through the creation of this labor market and platform, we will explore and test theories behind
% enabling collective governance, consensus--building, and participatory democracy.
% This study will yield guidelines for system--designers to spark and foster these values.

% Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
% This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
% However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
% While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
% In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction.

\section*{Interdisciplinarity}
For this research to succeed, the methodological approaches of Political Science
and the engineering skills of Computer Science must not only be in conversation, but in concert.
The challenge in designing for collective governance,
as Hardin (1982) differentiates it from collective action,
is such that collective governance
``takes on the character of a nearly anthropological investigation of minute interrelationships''~\cite{russell1982collective}.
In order to design a system that empowers workers,
a researchers must first attempt to understand a wide range of workers' circumstances:
The needs of workers;
the contexts in which they work;
their relationships with one another, with other groups, \& with institutions such as governments;
and more generally the paradigmatic views of gig workers.
Only then can one reasonably hope to design a system consistent with the views and broader culture of gig work.

Lee et al. and Gray et al. have identified a number of ways that workers subvert and circumvent the intents of system--designers,
both in digital workplaces and where work is simply mediated digitally~\cite{uberAlgorithm,crowdcollab}.
These patterns of behavior elude algorithmic tracking and measurement because they deliberately avoid the \textit{a priori} assumptions made by the designers of systems,
who attempt to structure these sites of work in ways to create incentives for preferred behavior.
Identifying and understanding the details of this behavior thus begins with a qualitative,
ethnographic endeavor.

\section*{Anticipated Contributions \& Implications}
This research is important given the recent trends toward the discretization of work.
With increasing numbers of people joining the gig economy,
and lacking sufficient work toward digitally mediated forms of collective action and governance,
disenfranchised and marginalized laborers in particular stand to lose significantly.
Indeed, advances made by labor advocacy groups in the past century have already eroded
due to the classification of many of these workers as ``contractors'' rather than ``employees'',
signaling the larger trend of the reversal of benefits and protections for which workers have fought.

More broadly, this project's will speak to the critique that the Internet,
far from the democratizing force that Barlow predicted it would be,
has become an infrastructure facilitating the corporate centralization of power~\cite{barlow2009declaration,jones2011does,EboCybertopia}.
The success of this project will therefore not only
bring a more nuanced understanding of the shifting climate and culture of labor,
but may
--- in a small way ---
deliver on the promise Barlow made that
cyberspace and the Internet would be a tool for democratizing access to information, and with that,
empowering more than the tech elite who engineered it.
% but will add to the increasingly minority claim that the Internet should be a democratizing force,
% rather than a centralizing one;
% \cite{barlow2009declaration}.

% To ensure that system--designers have reasonable access to our findings,
% our research targets will be top--tier venues in Human--Computer Interaction
% which support open access publication paradigms
% (e.g., ACM CHI, ACM CSCW).
% A second set of outlets will be through the NDWA's relationships with labor advocates and the press.

\pagebreak
\printbibliography{}
% \begin{document}
% \bibliographystyle{acm}
% \bibliography{references.bib}
\end{document}

### 2017-03-06 proposals/SIGFproposal_2017/old_cyber.tex
sigf17

--- 2016-05-16 proposals/CyberProposal/paper.tex
@@
 Levi,
-Johari and Valentine have an ongoing funded Cyber project titled ``Cyber Work: The Future of Networked Labor".
+Johari and Valentine have an ongoing funded Cyber project titled ``Cyber Work: The Future of Networked Labor''.
 This project outlined major goals:
@@
 
-Second, we will create and launch a cooperatively owned ``gig" marketplace for domestic workers in collaboration with the National Domestic Workers Alliance
+Second, we will create and launch a cooperatively owned ``gig'' marketplace for domestic workers in collaboration with the National Domestic Workers Alliance
 (NDWA).
@@
 % In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction.
-% The web and mobile application, Alia, enables the basic affordances of ``gig" work for domestic work 
+% The web and mobile application, Alia, enables the basic affordances of ``gig'' work for domestic work 
 % --- rapid scheduling of any available worker in the network
@@
 \bibliographystyle{apalike}
-\bibliography{../references.bib}
+\bibliography{references.bib}
 \end{document}

### 2017-03-06 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
major changes to background section

@@
 \usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase,comment}
 \usepackage[normalem]{ulem}
@@
 \definecolor{Blue}{HTML}{0000FF}
-\newcommand{\topic}[1]{{\color{Blue}#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
+\newcommand{\topic}[1]{{\color{Blue}#1}}
 \definecolor{Gray}{HTML}{BABABA}
-% \newcommand{\blah}[1]{{\color{Gray}\textit{#1}}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
+\newcommand{\ugh}[1]{{\color{Gray}\textit{#1}}}
 \newenvironment{blah}{\par\color{Gray}\itshape}{\par}
@@
 \section*{Motivating Work: Historical Analysis of Piecework}
-% \topic{The past decade has seen a flourishing of computationally--mediated labor.}
-% A framing of work into modular, pre--defined components
-% enables computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
-% In this regime, distributed workers engage in work whenever their schedules allow,
-% often with little to no awareness of the broader context of the work, and
-% often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
-
 \topic{Piecework has historically disempowered workers and empowered employers.}
@@
 starting with agriculture and home production but quickly moving into factories.
-However, by the turn of the 20th century in the United States,
-campaigns for workers' rights yielded regulations on conditions, and by the late 20th century,
+However, by the turn of the \nth{20} century in the United States,
+campaigns for workers' rights yielded regulations on conditions, and by the late \nth{20} century,
 piecework was associated primarily with migrant labor and sweatshops.
@@
 Uber's taxi drivers are paid per ride and assigned jobs by an algorithm~\cite{uberAlgorithm,hall2015analysis}.
-Information workers, such as those on Amazon Mechanical Turk and Upwork, are paid per task, performing vast quantities of data and administrative work~\cite{martin2014being}.
+Information workers, such as those on Amazon Mechanical Turk and Upwork, are paid per task,
+performing vast quantities of data and administrative work~\cite{martin2014being}.
 Increasingly numerous workers make money through one--off tasks like housecleaning and food delivery, all mediated by online platforms.
+Much of this work has been limited in complexity to easily verifiable work, stymieing gig workers' career opportunities
+in many of the same ways that historically limited information workers' careers~\cite{grier2013computers}.
+
+\begin{comment}
 Current legal and policy frameworks classify these workers as independent contractors, precluding them from accessing employee benefits and job security associated with employment.
@@
 Despite this, they still seek collective representation, as Uber drivers have done repeatedly to combat falling rates.
+% that bit's a little/a lot too much about collective representation. not really the right tact, right?
+\end{comment}
 
+\begin{comment}
+% this feels a little collective action--y;
+% should i cut it and reframe this to be more about individuals making their work more complex/having a better trajectory?
 \topic{How do workers react to, and create collective counterbalances to, the digital piecework economy?}
@@
 and successful collaborative efforts such as Wikipedia suggest that the online environment may support forms of cooperation that could not have succeeded as readily offline.
+\end{comment}
 
+\begin{comment}
 \topic{As yet,
-we have limited knowledge of how online pieceworkers could adapt to contemporary labor markets.}
-We do not know how workers are engaging in coordinated and cooperative actions,
-nor what form they should take in an era of digital work.
-Are their efforts similar to historical behaviors for offline pieceworkers,
-such as unions and worker cooperatives?
-Or do the affordances of the online world result in different outcomes?
-Furthermore,
-could technological affordances grant pieceworkers the ability to organize more effectively,
-and even to run their own marketplaces?
+Will their fates follow those of historical offline pieceworkers,
+or will the affordances of the online world result in different outcomes?
+In our most recent work~\cite{pieceworkCrowdworkGigwork}, we argue that
+the trajectory of historical pieceworkers can substantively inform the trajectory of gig work today,
+but that we must take an active role in the design and development of these platforms
+--- that is, taking advantage of the potential to influence the entire architectures as \citeauthor{lessig2006code} points out~\cite{lessig2006code}.
+\end{comment}
+
+\topic{In our examination of crowd work and gig work as a reinstantiation of piecework~\cite{pieceworkCrowdworkGigwork}, we identify a number of parallels that relate to this question.}
+First, that the first ``human computers'' were geographically distributed and only retained for several years to stifle career development opportunities~\cite{grier2013computers};
+second, that effective uses of historical piecework were generally limited to easily measured work due to a general mistrust~\cite{american1921problem,richards1904anything};
+third, this mistrust led to such a starkly adversarial relationship between workers and managers that
+it may have precipitated the wave of labor advocacy which defined the first half of the \nth{20} century~\cite{hart2013rise,hart2016rise}.
+What remains then are serious questions about the future of gig work as informed by piecework's history.
+We have limited knowledge of how gig workers today are adapting to digitally mediated, constantly changing labor markets.
+Similarly, we know very little about gig work's potential to support complex work, let alone how to turn that potential into reality.
+
+\begin{comment}
 
@@
 This effort has galvanized hundreds of workers and resulted in outcomes including public media campaigns and the creation of ethical work guidelines for Mechanical Turk employers.
+\end{comment}
 
-This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.
-What affordances and protections do workers need in order to make their voices heard and to have collective power?
-What protections should platforms be required to provide to their workers?
-What support is necessary for people who work on multiple platforms, frequently moving between different employers?
-Are unions the answer, or some other form of worker organization?
-% Today, with the growth of digitally mediated systems to manage workers and decontextualized pieces of work,
-% \textit{gig} work
-% --- characterized by the discrete jobs (or \textit{gigs}) which workers agree to do  ---
-% has increasingly become the focus of public and academic interest
-% \cite{martin2014being,uberAlgorithm,hall2015analysis}.
-% But gig work platforms have achieved some level of notoriety as settings of disempowerment
-% fueled by information asymmetry and opaque algorithmic management;
-% while researchers have made efforts to empower workers,
-% perhaps most notably on Amazon Mechanical Turk (AMT),
-% these efforts have thus far only ameliorated,
-% rather than resolved,
-% tension between
-% the various stakeholders of these systems
-% \cite{turkopticon,dynamo}.
-% This persisting frustration among \textit{Turkers} (the colloquial name for workers on AMT) parallels the frustration
-% workers in the \textit{gig} economy
-% --- namely, on--demand services such as Uber and Handy
-% \cite{dillahunt2016does,uberRiots,fedsUber,hall2015analysis}.
+\topic{This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.}
+Will gig work represent empowered, independent professionals who freely control their destinies, or 
+will gig work become a geographically distributed ``digital sweatshop''~\cite{dawnDigitalSweatshopCushing}?
+Whereas researchers had relatively little control over the markets that emerged throughout the \nth{20} century,
+researchers of digital sites of work have the ability not only to influence, but to architect these settings~\cite{lessig2006code}.
+It can be argued that with a few lines of code, we can effect outsize change on the future of gig work
+--- a form of work that continues to grow by the year~\cite{pewSharing,pewSharing24percent}.
 
-% With gig work increasingly representing the labor paradigm under which people work,
-% and with this mode of labor being exported to other parts of the world,
-% it is not only important that we attempt to understand the culture of those engaged in the discretization of work,
-% but that we attempt to participate in this process with design interventions
-% to improve the circumstances of workers by empowering them.
-
-% Working with the National Domestic Workers Alliance (NDWA),
-% I began to explore methods of empowering workers and ensuring their benefits, safety, and well--being.
-% Historically, labor unions have advocated on behalf of workers,
-% typically representing an industry or company's workforce.
-% In the case of gig work,
-% the geographic diffusion associated with digitally mediated systems
-% makes collective organization difficult.
-% More importantly, however,
-% the findings we uncovered among Turkers during our work on Dynamo
-% suggests that gig workers are reluctant regarding what they perceive labor unions to represent.
-% As one Turker described the sentiment I found more generally,
-% ``\dots~I would not feel comfortable taking part [in a labor union].
-% It runs against my grain because I am an individualist.
-% [\dots] I consider myself self--employed\dots~not working for anyone in particular''
-% \cite{dynamo}.
-
-% This cultural disconnect between workers and the perceived qualities of labor unions presented us with an opportunity
-% to explore more creative ways of dealing with the adversarial relationships between workers and managers.
-% In this search, we found that \textit{worker cooperatives}
-% --- which have struggled to reach large scales in the past in part due to the mounting struggle associated with growing collectively governed groups ---
-% may in fact be feasible given the unique features of the Internet, 
-% described by Miller and others,
-% which make it possible for people to collaborate rapidly and cheaply,
-% obviating many of the concerns that typically stymie efforts to organize,
-% such as geography
-% \cite{miller2011understanding}.
-
-% I propose to conduct ethnographic research on gig work in digitally mediated platforms,
-% through the historical lens of \textit{piece work},
-% to better understand the characteristics that influence and affect the culture of gig workers today.
-% The ethnographic research we conduct will yield an \textit{emic} understanding of
-% the values gig workers hold and their sense of self as they relate to the work they do.
-% With this ethnographic data,
-% I will then design and build a worker cooperative in collaboration with gig workers and the NDWA.
-% This system will reflect on the ethnographic findings we will have made from the first study,
-% as well as test guidelines toward enabling and fostering collective governance strategies.
-
-% While researchers have contributed to our understanding of worker cooperatives and participatory democracy,
-% we nevertheless find that worker cooperatives struggle to grow to the scales that firm--owned companies do.
-% Hardin, Olson, and others offer general framing of collective governance,
-% especially as it relates to the instances of collective action I explored in our work on Dynamo
-% \cite{russell1982collective,olsonlogic}.
-% Ostrom in particular provides substantive guidance on the mechanisms necessary to facilitate collective governance
-% \cite{ostrom1990governing}.
-
-% Ultimately, while we know how to design and engineer systems for collaboration,
-% and in particular we can generate instances of collective action where the target action is finite and focused,
-% we do not yet know how to design online systems for the sustained collective governance that we see only occasionally
-% (e.g. Wikipedia)
-% \cite{catalyst,dynamo,foundry}.
-% The core questions in this research, exploring the mechanisms of collective governance on the Internet, are social ones.
+% What affordances and protections do workers need in order to make their voices heard and to have collective power?
+% What protections should platforms be required to provide to their workers?
+% What support is necessary for people who work on multiple platforms, frequently moving between different employers?
+% Are unions the answer, or some other form of worker organization?
 
@@
 
-% As stated previously,
-% a critical component of the research proposal consists of
-% understanding the culture of gig workers.
-% Indeed, this proposal consists of two distinct research endeavors:
-% \begin{itemize}
-% \item Ethnographic study of gig workers, to understand the ways that workers do and indeed may support one another, and
-% \item design interventions consisting of the creation of a cooperative labor market designed collaboratively with workers.
-% \end{itemize}
-
-% \subsection*{Study 1: Ethnographic study of gig workers}
-% The first study will consist of ethnographic fieldwork studying workers in modern piece work labor markets 
-% --- e.g., Mechanical Turk, Uber, and Handy ---
-% to understand how social coping strategies emerge among workers.
-% How do workers air their grievances?
-% How do they come together to push for change?
-% How have they adapted to the working conditions of these digital hiring halls?
-
-% I will perform one year of fieldwork with digital piece workers.
-% To understand their breadth of experience, I will sample roughly fifty workers from information work
-% (for example,
-% Mechanical Turk and CrowdFlower),
-% and another fifty workers in the ``sharing economy''
-% (including Uber,
-% Lyft,
-% and AirBnB),
-% and fifty in gig work
-% (namely Handy and TaskRabbit).
-% Given existing relationships and reputations as credible researchers in communities of Turkers, I can recruit participants through known existing channels.
-% In the cases of platforms such as Uber,
-% AirBnB,
-% and Handy, I can recruit workers for nominally legitimate work
-% (for example,
-% requesting that a driver drives for 15 minutes to a previously selected destination)
-% and invite them to participate in our interview.
-
-% \subsection*{Study 2: Design and engineering of a cooperative labor market}
-% The second component of this research will involve
-% the engineering and launch a cooperatively owned ``gig" marketplace for domestic workers
-% in collaboration with the National Domestic Workers Alliance
-% (NDWA).
-% We will contribute the first digitally mediated gig labor market collaboratively designed with
-% workers and labor advocacy groups.
-% % My prior work established a collective action platform for crowd workers
-% % \cite{dynamo}.
-% Informed by my work on Dynamo and the ethnographic fieldwork in the first study,
-% I will then seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
-% The NDWA's Fair Care Labs are providing access to interested domestic workers
-% (house cleaners), and we are designing and implementing the platform with a pilot group.
-
-% The goal of this research is
-% to understand how to design for a set of distributed workers to achieve consensus on issues such as
-% management, service fees, and arbitration.
-% Through the creation of this labor market and platform, we will explore and test theories behind
-% enabling collective governance, consensus--building, and participatory democracy.
-% This study will yield guidelines for system--designers to spark and foster these values.
-
-% Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
-% This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
-% However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
-% While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
-% In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction.
-
 \section*{Interdisciplinarity}
@@
 The challenge in designing for collective governance,
-as Hardin (1982) differentiates it from collective action,
+as~\citeauthor{russell1982collective} (\citeyear{russell1982collective}) differentiates it from collective action,
 is such that collective governance
@@
 
-Lee et al. and Gray et al. have identified a number of ways that workers subvert and circumvent the intents of system--designers,
+\citeauthor{uberAlgorithm} and \citeauthor{crowdcollab} have identified a number of ways that workers subvert and circumvent the intents of system--designers,
 both in digital workplaces and where work is simply mediated digitally~\cite{uberAlgorithm,crowdcollab}.
@@
 
-More broadly, this project's will speak to the critique that the Internet,
+More broadly, this project's will speak to the critique that the internet,
 far from the democratizing force that Barlow predicted it would be,

### 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
some changes to interdisciplinarity and contributions

@@
 \section*{Motivating Work: Historical Analysis of Piecework}
-\topic{Piecework has historically disempowered workers and empowered employers.}
-Subdividing work and paying per task drives employees to work faster and longer,
+\topic{Piecework has historically disempowered workers and empowered employers, and
+many the same dynamics of disenfranchisement and frustration are poised to return today.}
+In our most recent work~\cite{pieceworkCrowdworkGigwork}, we found that a number of aspects of piecework
+--- a form of work payment that began more than a century ago ---
+bears striking resemblance to the contemporary labor we call ``gig work''.
+By subdividing work and paying for each task, employees are driven to work faster and longer,
 affording them little control over working conditions.
-For most of its history, piecework offered no benefits, occupational health and safety, or social insurance.
+And for most of its history, piecework yielded no benefits, occupational health and safety, or social insurance.
 When piecework combined with job routinization, workers became interchangeable and deindividuated.
@@
 \section*{Research plan}
-\begin{blah}
+\topic{We intend to build a system that tracks gig workers' professional histories across myriad work platforms,
+aggregating \& consolidating that data, and
+allowing workers to curate their professional identities, instantiated by a ``digital r\'{e}sum\'{e}''.}
+A number of open questions stand to be informed by this endeavor:
+first, do tools and systems like these help various stakeholders (such as workers, or requesters) in measurable ways?
+And second, what can we learn about collective governance from a system that is designed and operated collectively with workers themselves?
+
+\topic{At a high level, we can study the usefulness of this system using standard experimental approaches.}
+First, through quantitative methods (for example, comparing workers' earnings over time); and
+second, through qualitative  methods (for example, through interviews to determine whether workers find and exploit specialty niches as a result of this tool).
+We can also explore employers perspectives in tangible ways;
+for example, given concerns about the quality of work~\cite{Ipeirotis:2010:QMA:1837885.1837906,le2010ensuring,Law:2017:CTR:2998181.2998197},
+we will investigate whether a system that better communicates worker expertise allows employers to rely more confidently on crowdwork.
+
+\topic{More broadly, we can think of this as an opportunity to engage in participatory design and democracy, and
+to learn best practices through experimentation.}
+We can take this opportunity to study collective governance in online settings
+--- extending, as \citeauthor{russell1982collective} describes it,
+from the ``on--shot'' cases of collective action that we instantiated in our earlier work~\cite{dynamo} to ``ongoing'' collective action~\cite{russell1982collective}.
+While considerable effort has gone toward collective governance
+(e.g.~\cite{ostrom1990governing,mccallum2013global,ahlquist2013interest,olsonlogic,polletta2002freedom}),
+many of these mechanisms rely on collocation to some extent or another;
+the best practices of collective governance for communities that never meet
+(e.g moderators and administrators on Wikipedia)
+remains unclear.
+
+\begin{comment}
+\topic{We're pursuing the design of a system that offers a very narrow slice of workers' needs based on
+what we've learned so far from gig workers, and from the historical timeline that we saw in piecework.}
+Informed by our observations from fieldwork while organizing and motivating collective action~\cite{dynamo}
+and from other research~\cite{martin2014being},
+we've found that gig workers are highly resistant to traditional labor advocacy and empowerment tools, such as labor unions.
+In part, this may be because the conventional systems of affiliations and labor unions are a poor match;
+to say nothing of responding to the varied needs of people who participate in the gig economy to differing degrees,
+what we consider conventional labor advocacy structures is anathema to gig workers:
+``If you mean a `labor union', I would not feel comfortable taking part.
+It runs against my grain because I am an individualist.
+[\dots]~I consider myself self--employed\dots~not working for anyone in particular''.
+Looking through piecework's history, we found that similar sentiments existed in the early years of piecework~\cite{pieceworkCrowdworkGigwork}.
+While this sentiment shifted over the following decades, it's worth considering that this shift occurred over a number of years, and
+only as people became increasingly familiar with the conclusion that forms of gig 
+\end{comment}
+
+\begin{comment}
 We intend to create a platform that can become a template for online work that better represents workers' needs.
-However, solutions that exist for traditional, non-gig workers
+However, solutions that exist for traditional, non--gig workers
 --- especially labor unions ---
@@
 --- rapid scheduling of any available worker in the network (Figure 1).
-\end{blah}
+\end{comment}
 
 \section*{Interdisciplinarity}
-For this research to succeed, the methodological approaches of Political Science
-and the engineering skills of Computer Science must not only be in conversation, but in concert.
-The challenge in designing for collective governance,
-as~\citeauthor{russell1982collective} (\citeyear{russell1982collective}) differentiates it from collective action,
-is such that collective governance
-``takes on the character of a nearly anthropological investigation of minute interrelationships''~\cite{russell1982collective}.
+\topic{This research will call on methods and skills originally from and generally found in disparate fields.}
+As we showed with our paper examining the relationship between piecework and gig work~\cite{pieceworkCrowdworkGigwork},
+the lens of a social scientist can inform conversations about digitally mediated work in ways that many computer scientists are not otherwise able to offer.
+At the same time, the perspectives of computer scientists
+--- and the challenges digitally mediated spaces bring with them ---
+offer to reinvigorate deeply theoretically grounded discussion and offer field sites with which to experiment and test these theories.
+
+\begin{blah}
 In order to design a system that empowers workers,
@@
 Only then can one reasonably hope to design a system consistent with the views and broader culture of gig work.
+\end{blah}
 
+\begin{blah}
 \citeauthor{uberAlgorithm} and \citeauthor{crowdcollab} have identified a number of ways that workers subvert and circumvent the intents of system--designers,
@@
 ethnographic endeavor.
+\end{blah}
 
 \section*{Anticipated Contributions \& Implications}
-This research is important given the recent trends toward the discretization of work.
-With increasing numbers of people joining the gig economy,
-and lacking sufficient work toward digitally mediated forms of collective action and governance,
-disenfranchised and marginalized laborers in particular stand to lose significantly.
-Indeed, advances made by labor advocacy groups in the past century have already eroded
-due to the classification of many of these workers as ``contractors'' rather than ``employees'',
-signaling the larger trend of the reversal of benefits and protections for which workers have fought.
+This research is of significant importance given the recent trends in the ``gig economy''.
+With more and more people joining the workforce in this capacity~\cite{pewSharing,pewSharing24percent},
+concerns about ``the future of work''~\cite{waterCooler,crowdworkFuture} continue to grow.
+Without compelling career trajectories in this space,
+the optimistic prediction seems to be that we will have little say in the outcome of this shift in work.
+Our hope is that, concretely, the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
+More broadly, we expect to offer two major contributions to the Human--Computer Interaction community:
+first, an existence proof that a gig work market for complex, specialized workers is possible; and
+second, insights into how best to scaffold and foster ongoing collective governance.
+% and lacking sufficient work toward digitally mediated forms of collective action and governance,
+% disenfranchised and marginalized laborers in particular stand to lose significantly.
+% Indeed, advances made by labor advocacy groups in the past century have already eroded
+% due to the classification of many of these workers as ``contractors'' rather than ``employees'',
+% signaling the larger trend of the reversal of benefits and protections for which workers have fought.
 
-More broadly, this project's will speak to the critique that the internet,
+This project's will speak to the critique that the internet,
 far from the democratizing force that Barlow predicted it would be,

### 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
abstract revision --- 98 words!

@@
   \begin{abstract}
-On--demand labor has become a new instantiation of piecework.
-Without shared factories and water coolers,
-how do these workers coordinate,
-build solidarity,
-and take collective action?
-We will engage in fieldwork with pieceworkers in data entry,
-domestic services,
-and on--demand driving to understand how they counter algorithmic systems and engage in collective action.
-Our goal is to both understand the factors that drive collective behavior in supposedly distributed labor platforms,
-and to build infrastructures that enable the growth of worker collective action in digitally mediated on--demand labor.
-Our results shed light on the potential for digital piecework cooperativism,
-as well as the policies necessary for it to succeed. % ~108 words
+Gig work has become a new instantiation of piecework.
+Further, gig work platforms allow employers to geographically disperse workers and silo workers' information,
+making organization and coordination all but impossible,
+and stymieing work specialization and expertise.
+Our research will draw on our past work, using historical analysis of piecework to inform this contemporary phenomenon,
+to design and develop a ``digital hiring hall''.
+Our research will
+shed light on the potential complexity limits of gig work,
+point to possible futures of labor advocacy for gig workers,
+and offer some insights into the best practices of principally online collective governance.
+
+% The results of this research will make it easier for gig workers to identify and focus on specializations,
+% fostering rewarding careers in gig work and giving workers spaces to organize and coordinate.
+
+% Our goal is to both understand the factors that drive collective behavior in supposedly distributed labor platforms,
+% and to build infrastructures that enable the growth of worker collective action in digitally mediated on--demand labor.
+% Our results shed light on the potential for digital piecework cooperativism,
+% as well as the policies necessary for it to succeed. % ~108 words
   \end{abstract}

### 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
statement revisions based on earlier feedback

@@
 
+% Michael Bernstein [5:42 PM] 
+% feedback:
+% [x] first paragraph: "gig work” needs examples (Uber, Mechanical Turk) so people know what you’re talking about.
+I know you get there in the second paragraph, but I think a reader will tune out the first paragraph if they don’t know concretely what you’re talking about.
+% [x] research plan: I’d like to see a stronger transition. Ideally, a question that the historical analysis left open that you are now poised to answer
+% [x] research plan: make clearer that your proposal is to transition this historical framing into a novel system that you are building (I don’t think that readers will automatically know this)
+% [ ] research plan: whoa, you jump to experimental analysis before you even describe what it is!
+I need 1-2 paragraphs describing the system that you’re building.
+Walk through what it does, what a worker will do with it.
+% [x] research plan: can cut the third paragraph for space if needed
+% [ ] interdisciplinarity: call out the obvious: that you are a trained social scientist now in a computer science department
+
 \begin{document}
@@
 and offer some insights into the best practices of principally online collective governance.
-
-% The results of this research will make it easier for gig workers to identify and focus on specializations,
-% fostering rewarding careers in gig work and giving workers spaces to organize and coordinate.
-
-% Our goal is to both understand the factors that drive collective behavior in supposedly distributed labor platforms,
-% and to build infrastructures that enable the growth of worker collective action in digitally mediated on--demand labor.
-% Our results shed light on the potential for digital piecework cooperativism,
-% as well as the policies necessary for it to succeed. % ~108 words
   \end{abstract}
@@
 --- a form of work payment that began more than a century ago ---
-bears striking resemblance to the contemporary labor we call ``gig work''.
+bears striking resemblance to the sort of work we have begun to see emerge in Uber, Lyft, TaskRabbit, and more
+--- collectively called ``gig work''.
 By subdividing work and paying for each task, employees are driven to work faster and longer,
@@
 
-\begin{comment}
-Current legal and policy frameworks classify these workers as independent contractors, precluding them from accessing employee benefits and job security associated with employment.
-Complicating matters, today's workers often despise unfair employers but prize the flexibility and autonomy afforded by their platforms~\cite{martin2014being}.
-Many view themselves as masters of their own fate, free to enter and leave the workforce at will and without repercussion.
-Despite this, they still seek collective representation, as Uber drivers have done repeatedly to combat falling rates.
-% that bit's a little/a lot too much about collective representation. not really the right tact, right?
-\end{comment}
-
-\begin{comment}
-% this feels a little collective action--y;
-% should i cut it and reframe this to be more about individuals making their work more complex/having a better trajectory?
-\topic{How do workers react to, and create collective counterbalances to, the digital piecework economy?}
-On the one hand, studies of online communities predict that workers online will face greater challenges, because
-trust is more difficult to build in digitally mediated environments~\cite{successfulOnlineCommunities,kollock2005managing,cook2005cooperation}.
-Without factory floors, water coolers, and shared neighborhoods~\cite{waterCooler},
-how do workers coordinate and build solidarity?
-On the other hand,
-online forums may make it easier for workers to share information and strategies,
-and successful collaborative efforts such as Wikipedia suggest that the online environment may support forms of cooperation that could not have succeeded as readily offline.
-\end{comment}
-
-\begin{comment}
-\topic{As yet,
-Will their fates follow those of historical offline pieceworkers,
-or will the affordances of the online world result in different outcomes?
-In our most recent work~\cite{pieceworkCrowdworkGigwork}, we argue that
-the trajectory of historical pieceworkers can substantively inform the trajectory of gig work today,
-but that we must take an active role in the design and development of these platforms
---- that is, taking advantage of the potential to influence the entire architectures as \citeauthor{lessig2006code} points out~\cite{lessig2006code}.
-\end{comment}
-
 \topic{In our examination of crowd work and gig work as a reinstantiation of piecework~\cite{pieceworkCrowdworkGigwork}, we identify a number of parallels that relate to this question.}
@@
 
-\begin{comment}
-
-We propose paired grounded ethnographic fieldwork and system development to 1) understand the forces shaping informal and formal worker reactions to the online piecework economy,
-and 2) create a technical and design foundation for collective action,
-ownership and governance in online piecework.
-
-First, we will analyze worker behavior in the current sociotechnical ecosystem through a year of fieldwork on major Mechanical Turk
-(data entry)
-platforms and forums,
-and on gig economy platforms such as Uber
-(driving)
-and Handy
-(domestic work).
-Through this fieldwork, we will investigate the techniques that workers use to manage their work environments and job opportunities.
-We will seek to understand the challenges they face when balancing
-an identity of being self--employed with an identity of being a pieceworker for a privately--owned labor platform, and
-the techniques that workers and employers use to circumvent each other's restrictions, barriers, and behaviors.
-
-Second, we will create and launch a cooperatively owned ``gig" marketplace for domestic workers in collaboration with the National Domestic Workers Alliance
-(NDWA).
-Our own prior work established a collective action platform for crowd workers~\cite{dynamo}.
-We now seek to augment this exploration of union--style collective by creating a cooperative that can scale up and set its own standards in the marketplace.
-The NDWA's Fair Care Labs are providing access to interested domestic workers
-(house cleaners), and we are designing and implementing the platform with them as a pilot group.
-The goal is to understand how to design for a set of distributed workers to achieve consensus on issues such as management, service fees, and arbitration.
-
-Our team is uniquely qualified to pursue this opportunity.
-Levi has years of experience studying labor, collectives, and unions.
-She is pairing with Bernstein, a computer scientist who studies online labor platforms and collective action.
-Bernstein, alongside PhD students Salehi and Alkhatib, created the first platform for collective action efforts amongst digital pieceworkers, focused on the Amazon Mechanical Turk platform.
-This effort has galvanized hundreds of workers and resulted in outcomes including public media campaigns and the creation of ethical work guidelines for Mechanical Turk employers.
-\end{comment}
-
 \topic{This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.}
@@
 
-% What affordances and protections do workers need in order to make their voices heard and to have collective power?
-% What protections should platforms be required to provide to their workers?
-% What support is necessary for people who work on multiple platforms, frequently moving between different employers?
-% Are unions the answer, or some other form of worker organization?
+\section*{Research plan}
+\topic{While our research considered a number of possible futures for gig work in general and gig workers in particular,
+several important questions remain unresolved.}
+For example, will gig work remain an undesirable career path,
+largely taken up by people who can't or won't commit to conventional work arrangements;
+or can gig work someday come to represent highly skilled, sought after work?
+While we offered a framework for answering questions such as this one, the reality is that
+any researcher or team exploring this area
+--- especially on the basis of the parallel we drew ---
+would fundamentally need to be conversant with several deep bodies of research spanning myriad fields.
 
-\section*{Research plan}
-\topic{We intend to build a system that tracks gig workers' professional histories across myriad work platforms,
+Informed by our research into gig work, we propose to build one of the potential futures that we discussed.
+\topic{We will build a system that tracks gig workers' professional histories across myriad work platforms,
 aggregating \& consolidating that data, and
-allowing workers to curate their professional identities, instantiated by a ``digital r\'{e}sum\'{e}''.}
-A number of open questions stand to be informed by this endeavor:
-first, do tools and systems like these help various stakeholders (such as workers, or requesters) in measurable ways?
-And second, what can we learn about collective governance from a system that is designed and operated collectively with workers themselves?
+allowing workers to curate their professional identities, instantiated by a ``digital hiring hall''.}
+% A number of open questions stand to be informed by this endeavor:
+% first, do tools and systems like these help various stakeholders (such as workers, or requesters) in measurable ways?
+% And second, what can we learn about collective governance from a system that is designed and operated collectively with workers themselves?
+
+\begin{blah}
+% [ ] research plan: whoa, you jump to experimental analysis before you even describe what it is!
+I need 1-2 paragraphs describing the system that you’re building.
+Walk through what it does, what a worker will do with it.
+% \topic{1--2 paragraphs describing the system. What does it do? what will workers do with it?}
+\topic{The system we intend to build will serve two roles --- one for workers and one for employers.} % that the worker can curate and that requesters and platforms at large can access to make hiring decisions algorithmically.
+Gig workers will share their work histories from platforms such as Amazon Mechanical Turk with our repository,
+which will allow us to provide the worker with information upon which they can reflect
+(e.g. detailed analysis of their work trends, approval rates under various circumstances, etc\dots)
+as well as information in relation to other workers
+(e.g. offering a percentile ranking overall for approval rates, or within specific types of tasks).
+In doing so, we hope to find that workers will identify cues and gravitate toward qualities in work
+in which they identify competitive advantage (for example, )
+
+\topic{The second aspect }
+and second, we will incorporate their data into responses from platforms and employers looking for workers qualified to do complex work.
+
+By disconnecting workers' data from the platforms themselves, our repository will act as a sort of ``digital hiring hall'',
+connecting employers with workers whose varied work histories provide them with the expertise necessary to do given tasks.
+\end{blah}
 
@@
 
+\begin{comment}
 \topic{More broadly, we can think of this as an opportunity to engage in participatory design and democracy, and
@@
 --- extending, as \citeauthor{russell1982collective} describes it,
-from the ``on--shot'' cases of collective action that we instantiated in our earlier work~\cite{dynamo} to ``ongoing'' collective action~\cite{russell1982collective}.
+from the ``one--shot'' cases of collective action that we instantiated in our earlier work~\cite{dynamo}
+      to ``ongoing'' collective action~\cite{russell1982collective}.
 While considerable effort has gone toward collective governance
@@
 remains unclear.
+\end{comment}
 
@@
 
-\begin{blah}
+\begin{comment}
 In order to design a system that empowers workers,
@@
 Only then can one reasonably hope to design a system consistent with the views and broader culture of gig work.
-\end{blah}
+\end{comment}
 
-\begin{blah}
+\begin{comment}
 \citeauthor{uberAlgorithm} and \citeauthor{crowdcollab} have identified a number of ways that workers subvert and circumvent the intents of system--designers,
@@
 ethnographic endeavor.
-\end{blah}
+\end{comment}
 

### 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
cleared out the bulleted list of suggestions

@@
 % [x] research plan: make clearer that your proposal is to transition this historical framing into a novel system that you are building (I don’t think that readers will automatically know this)
-% [ ] research plan: whoa, you jump to experimental analysis before you even describe what it is!
+% [x] research plan: whoa, you jump to experimental analysis before you even describe what it is!
 I need 1-2 paragraphs describing the system that you’re building.
@@
 % [x] research plan: can cut the third paragraph for space if needed
-% [ ] interdisciplinarity: call out the obvious: that you are a trained social scientist now in a computer science department
+% [x] interdisciplinarity: call out the obvious: that you are a trained social scientist now in a computer science department
 
@@
 
-\begin{blah}
+% \begin{blah}
 % [ ] research plan: whoa, you jump to experimental analysis before you even describe what it is!
@@
 In doing so, we hope to find that workers will identify cues and gravitate toward qualities in work
-in which they identify competitive advantage (for example, )
+in which they identify competitive advantage.
 
-\topic{The second aspect }
-and second, we will incorporate their data into responses from platforms and employers looking for workers qualified to do complex work.
+\topic{The second role that this repository will help employers identify and source the best workers for highly skilled work.}
+By aggregating the data from work platforms and
+offering more in--depth analysis of workers' histories, we hope to find that
+we can direct employers to gig workers whose overall experiences match the employer's given needs.
+For example, if an employer or a platform queries our platform for
+workers who are proficient at translating between two specific languages,
+we can yield a list of workers who would be suitable for the task.
 
+\topic{This approach is similar in some senses to the role that hiring halls played through the \nth{20} century.}
 By disconnecting workers' data from the platforms themselves, our repository will act as a sort of ``digital hiring hall'',
 connecting employers with workers whose varied work histories provide them with the expertise necessary to do given tasks.
-\end{blah}
+By engaging workers in the management of this repository, we will advance the metaphor in much the same way that
+labor unions maintained --- and continue to operate --- hiring halls of their workers.
+
+% \end{blah}
 
@@
 offer to reinvigorate deeply theoretically grounded discussion and offer field sites with which to experiment and test these theories.
+As a Computer Science Ph.D. student originally trained as an Anthropologist, I believe
+I'm the best candidate to combine these bodies of literature to make this project succeed.
 

### 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
cleared out the bulleted list of suggestions from 9:40ish

@@
 % }
+\makeatletter         
+\def\@maketitle{   % custom maketitle 
+{\Large \scshape \@title} ~~~\@author ~ on  \@date \par 
+\smallskip \hrule \smallskip }
 
-% \setlist[itemize]{itemsep=0pt,topsep=0pt,parsep=0pt}
-% \setenumerate{itemsep=0pt,topsep=0pt,parsep=0pt}
-% \newcommand{\sectitle}[1]{\textit{#1}}
-\title{Designing for Complex Gig Work and Careers}
+\title{Designing Digital Hiring Halls}
 \author{Ali Alkhatib}
-\date{Submitted \today}
+\date{\today}
 
@@
 \newcommand{\topic}[1]{{\color{Blue}#1}}
+\renewcommand{\topic}[1]{{#1}}
 \definecolor{Gray}{HTML}{BABABA}
@@
 
-% Michael Bernstein [5:42 PM] 
-% feedback:
-% [x] first paragraph: "gig work” needs examples (Uber, Mechanical Turk) so people know what you’re talking about.
-I know you get there in the second paragraph, but I think a reader will tune out the first paragraph if they don’t know concretely what you’re talking about.
-% [x] research plan: I’d like to see a stronger transition. Ideally, a question that the historical analysis left open that you are now poised to answer
-% [x] research plan: make clearer that your proposal is to transition this historical framing into a novel system that you are building (I don’t think that readers will automatically know this)
-% [x] research plan: whoa, you jump to experimental analysis before you even describe what it is!
-I need 1-2 paragraphs describing the system that you’re building.
-Walk through what it does, what a worker will do with it.
-% [x] research plan: can cut the third paragraph for space if needed
-% [x] interdisciplinarity: call out the obvious: that you are a trained social scientist now in a computer science department
+% Michael Bernstein [9:40 PM]
+% [x] add a title
+% [x] “new instantiation” —> “modern instantiation”
+% [x] why “further” in a second sentence of an abstract?
+% [x] "design and develop a 'digital hiring hall’”— specify for gig workers?
+% [x] need a sentence in the abstract describing what your system actually does (e.g., create a portable resume across services by scraping work histories, letting workers use that history to demonstrate abilities to a new platform)
+% [x] final sentence of abstract has three broadbrush implications, choose one that it maximizes
+% [x] intro to Research Plan section needs a more focused question, specifically the one that you’re solving.
+"will gig work remain an undesirable career path” is too broad for our project to answer.
+likewise, "we propose to build one of the potential futures that we discussed” is too vague.
+Instead, start with a focused question more like (and this isn’t refined, but:) "How can workers demonstrate their skill?” And explain in one sentence this was an issue historically, and it was only partially solved b/c they couldn’t X.
+But now you can build technology to X.
+And that’s what you’re going to do.
+% [x] "who are proficient at translating between two specific languages”: ground this.
+e.g., they have lots of previously recorded work by a translation service on Mechanical Turk to vouch for their quality
+% [x] Interdisciplinarity: first sentence is weaksauce — get specific
+% [x] Contributions: “little say” — why is this the _optimistic_ prediction? wouldn’t the optimist say that we _can_ steer it?
 
 \begin{document}
-  % \maketitle
+  \maketitle
   \begin{abstract}
-Gig work has become a new instantiation of piecework.
-Further, gig work platforms allow employers to geographically disperse workers and silo workers' information,
+    Gig work has become a modern instantiation of piecework.
+    Their platforms allow employers to
+    disperse workers geographically and silo workers' information,
     making organization and coordination all but impossible,
     and stymieing work specialization and expertise.
-Our research will draw on our past work, using historical analysis of piecework to inform this contemporary phenomenon,
-to design and develop a ``digital hiring hall''.
+    Our research will draw on our past work,
+    using historical analysis of piecework to inform this contemporary phenomenon,
+    to design and develop a ``digital hiring hall''
+    --- a platform--independent repository for workers to store their body of work, curate their professional image,
+    and better signal expertise to employers.
     Our research will
-shed light on the potential complexity limits of gig work,
-point to possible futures of labor advocacy for gig workers,
-and offer some insights into the best practices of principally online collective governance.
+    begin to instantiate one of the potential futures of labor advocacy for gig workers.
   \end{abstract}
@@
 
-\topic{In our examination of crowd work and gig work as a reinstantiation of piecework~\cite{pieceworkCrowdworkGigwork}, we identify a number of parallels that relate to this question.}
-First, that the first ``human computers'' were geographically distributed and only retained for several years to stifle career development opportunities~\cite{grier2013computers};
-second, that effective uses of historical piecework were generally limited to easily measured work due to a general mistrust~\cite{american1921problem,richards1904anything};
+\topic{In our examination of crowd work and gig work as
+a reinstantiation of piecework~\cite{pieceworkCrowdworkGigwork},
+we identify a number of parallels that relate to this question.}
+First, that the first ``human computers'' were geographically distributed and
+only retained for several years to stifle career development opportunities~\cite{grier2013computers};
+second, that effective uses of historical piecework were generally limited to
+easily measured work due to a general mistrust~\cite{american1921problem,richards1904anything};
 third, this mistrust led to such a starkly adversarial relationship between workers and managers that
-it may have precipitated the wave of labor advocacy which defined the first half of the \nth{20} century~\cite{hart2013rise,hart2016rise}.
+it may have precipitated the wave of labor advocacy which defined
+the first half of the \nth{20} century~\cite{hart2013rise,hart2016rise}.
 What remains then are serious questions about the future of gig work as informed by piecework's history.
@@
 
-\section*{Research plan}
+% [x] intro to Research Plan section needs a more focused question, specifically the one that you’re solving.
+"will gig work remain an undesirable career path” is too broad for our project to answer.
+likewise, "we propose to build one of the potential futures that we discussed” is too vague.
+Instead, start with a focused question more like (and this isn’t refined, but:) "How can workers demonstrate their skill?” And explain in one sentence this was an issue historically, and it was only partially solved b/c they couldn’t X.
+But now you can build technology to X.
+And that’s what you’re going to do.
+\section*{Research Plan}
 \topic{While our research considered a number of possible futures for gig work in general and gig workers in particular,
 several important questions remain unresolved.}
-For example, will gig work remain an undesirable career path,
-largely taken up by people who can't or won't commit to conventional work arrangements;
-or can gig work someday come to represent highly skilled, sought after work?
+For example, we pointed to the general question that researchers had been asking for some time ---
+whether gig work will become a site of complex, rewarding work
+--- or if it will remain a setting for low quality tasks with little upward mobility.
 While we offered a framework for answering questions such as this one, the reality is that
@@
 
-Informed by our research into gig work, we propose to build one of the potential futures that we discussed.
+We concluded that gig work's potential will be determined by
+how easily skilled workers can secure good work through these platforms.
+In order to do that, workers need to be able to communicate rich information about their qualifications to employers,
+which existing markets have so far struggled to do.
 \topic{We will build a system that tracks gig workers' professional histories across myriad work platforms,
-aggregating \& consolidating that data, and
-allowing workers to curate their professional identities, instantiated by a ``digital hiring hall''.}
-% A number of open questions stand to be informed by this endeavor:
-% first, do tools and systems like these help various stakeholders (such as workers, or requesters) in measurable ways?
-% And second, what can we learn about collective governance from a system that is designed and operated collectively with workers themselves?
+aggregating, consolidating, \& analyzing that data, and
+allowing workers to curate their professional identities, instantiated by a ``digital hiring hall'' for gig workers.}
 
-% \begin{blah}
-% [ ] research plan: whoa, you jump to experimental analysis before you even describe what it is!
-I need 1-2 paragraphs describing the system that you’re building.
-Walk through what it does, what a worker will do with it.
-% \topic{1--2 paragraphs describing the system. What does it do? what will workers do with it?}
-\topic{The system we intend to build will serve two roles --- one for workers and one for employers.} % that the worker can curate and that requesters and platforms at large can access to make hiring decisions algorithmically.
+\topic{The system we intend to build will serve two roles --- one for workers and one for employers.}
 Gig workers will share their work histories from platforms such as Amazon Mechanical Turk with our repository,
@@
 workers who are proficient at translating between two specific languages,
-we can yield a list of workers who would be suitable for the task.
+we can yield a list of workers who would be suitable for the task, determined by
+the workers' comprehensive histories on Mechanical Turk and, potentially, elsewhere.
 
@@
 labor unions maintained --- and continue to operate --- hiring halls of their workers.
-
-% \end{blah}
 
@@
 
-\begin{comment}
-\topic{More broadly, we can think of this as an opportunity to engage in participatory design and democracy, and
-to learn best practices through experimentation.}
-We can take this opportunity to study collective governance in online settings
---- extending, as \citeauthor{russell1982collective} describes it,
-from the ``one--shot'' cases of collective action that we instantiated in our earlier work~\cite{dynamo}
-      to ``ongoing'' collective action~\cite{russell1982collective}.
-While considerable effort has gone toward collective governance
-(e.g.~\cite{ostrom1990governing,mccallum2013global,ahlquist2013interest,olsonlogic,polletta2002freedom}),
-many of these mechanisms rely on collocation to some extent or another;
-the best practices of collective governance for communities that never meet
-(e.g moderators and administrators on Wikipedia)
-remains unclear.
-\end{comment}
-
-\begin{comment}
-\topic{We're pursuing the design of a system that offers a very narrow slice of workers' needs based on
-what we've learned so far from gig workers, and from the historical timeline that we saw in piecework.}
-Informed by our observations from fieldwork while organizing and motivating collective action~\cite{dynamo}
-and from other research~\cite{martin2014being},
-we've found that gig workers are highly resistant to traditional labor advocacy and empowerment tools, such as labor unions.
-In part, this may be because the conventional systems of affiliations and labor unions are a poor match;
-to say nothing of responding to the varied needs of people who participate in the gig economy to differing degrees,
-what we consider conventional labor advocacy structures is anathema to gig workers:
-``If you mean a `labor union', I would not feel comfortable taking part.
-It runs against my grain because I am an individualist.
-[\dots]~I consider myself self--employed\dots~not working for anyone in particular''.
-Looking through piecework's history, we found that similar sentiments existed in the early years of piecework~\cite{pieceworkCrowdworkGigwork}.
-While this sentiment shifted over the following decades, it's worth considering that this shift occurred over a number of years, and
-only as people became increasingly familiar with the conclusion that forms of gig 
-\end{comment}
-
-\begin{comment}
-We intend to create a platform that can become a template for online work that better represents workers' needs.
-However, solutions that exist for traditional, non--gig workers
---- especially labor unions ---
-face resistance from workers in the gig economy~\cite{martin2014being,dynamo}.
-This distaste may be due to a general decline in public opinion of unions.
-However, the traditional mechanics of a labor union are also a poor match: challenging issues include managing the benefits and proportional representation of workers whose contracts may last only minutes or hours, enforcing collective decisions on a decentralized network of workers, and establishing trust between workers who may never meet face--to--face.
-In our early fieldwork, for example, a worker on Amazon Mechanical Turk stated:
-``If by `union' you mean a `labor union', I would not feel comfortable taking part.
-It runs against my grain because I am an individualist.
-[\dots] I consider myself self--employed\dots\  not working for anyone in particular.''
-
-Given these constraints, what alternative model would give workers more power in an online gig--economy platform while still enabling a market to emerge and function?
-To explore one such alternative, we will collaborate with the National Domestic Workers Alliance's innovation arm, the Fair Care Labs, to create tools for online pieceworkers to collectively manage and govern their own labor markets.
-While the literature in social computing design and crowdsourcing already features many tools for successful decentralized collaborations (e.g., Wikipedia), and collective action movements (e.g., change.org), templates and guidelines for sustained collective action and governance online in high--stakes situations that affect workers' livelihoods are rare.
-Such efforts tend to swell quickly, but soon find themselves embroiled as they struggle to satisfy myriad stakeholders .
-Our goal is to introduce design mechanisms to enable collectives of gig workers to form, debate policy, and implement their decisions over a long period of time.
-
-Together with the Fair Care Labs, we are designing and launching a cooperative gig labor market for domestic workers (house cleaners).
-This cooperative labor market, Alia, is designed for domestic workers in the California Bay Area.
-However, our interest (and Fair Care Labs') is to generalize its insights to other workers and industries.
-While the driving domain is already crowded with services such as Lyft and Uber, domestic work is a more ideal area to launch a research prototype.
-In addition, the NDWA Fair Care Labs' support and network will help us gain rapid traction. The web and mobile application, Alia, enables the basic affordances of ``gig'' work for domestic work
---- rapid scheduling of any available worker in the network (Figure 1).
-\end{comment}
-
 \section*{Interdisciplinarity}
-\topic{This research will call on methods and skills originally from and generally found in disparate fields.}
+% \topic{This research will call on methods and skills originally from and generally found in disparate fields.}
+% \topic{The insights and expertise necessary to conduct this research successfully will call on a variety of skills not typically seen together, such as
+% participant observation, participatory design techniques, and machine learning approaches.}
 As we showed with our paper examining the relationship between piecework and gig work~\cite{pieceworkCrowdworkGigwork},
@@
 
-\begin{comment}
-In order to design a system that empowers workers,
-a researchers must first attempt to understand a wide range of workers' circumstances:
-The needs of workers;
-the contexts in which they work;
-their relationships with one another, with other groups, \& with institutions such as governments;
-and more generally the paradigmatic views of gig workers.
-Only then can one reasonably hope to design a system consistent with the views and broader culture of gig work.
-\end{comment}
-
-\begin{comment}
-\citeauthor{uberAlgorithm} and \citeauthor{crowdcollab} have identified a number of ways that workers subvert and circumvent the intents of system--designers,
-both in digital workplaces and where work is simply mediated digitally~\cite{uberAlgorithm,crowdcollab}.
-These patterns of behavior elude algorithmic tracking and measurement because they deliberately avoid the \textit{a priori} assumptions made by the designers of systems,
-who attempt to structure these sites of work in ways to create incentives for preferred behavior.
-Identifying and understanding the details of this behavior thus begins with a qualitative,
-ethnographic endeavor.
-\end{comment}
-
 \section*{Anticipated Contributions \& Implications}
@@
 concerns about ``the future of work''~\cite{waterCooler,crowdworkFuture} continue to grow.
-Without compelling career trajectories in this space,
-the optimistic prediction seems to be that we will have little say in the outcome of this shift in work.
-Our hope is that, concretely, the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
+Our hope is that the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
 More broadly, we expect to offer two major contributions to the Human--Computer Interaction community:
@@
 second, insights into how best to scaffold and foster ongoing collective governance.
-% and lacking sufficient work toward digitally mediated forms of collective action and governance,
-% disenfranchised and marginalized laborers in particular stand to lose significantly.
-% Indeed, advances made by labor advocacy groups in the past century have already eroded
-% due to the classification of many of these workers as ``contractors'' rather than ``employees'',
-% signaling the larger trend of the reversal of benefits and protections for which workers have fought.
 
@@
 empowering more than the tech elite who engineered it.
-% but will add to the increasingly minority claim that the Internet should be a democratizing force,
-% rather than a centralizing one;
-% \cite{barlow2009declaration}.
-
-% To ensure that system--designers have reasonable access to our findings,
-% our research targets will be top--tier venues in Human--Computer Interaction
-% which support open access publication paradigms
-% (e.g., ACM CHI, ACM CSCW).
-% A second set of outlets will be through the NDWA's relationships with labor advocates and the press.
 
@@
 \printbibliography{}
-% \begin{document}
-% \bibliographystyle{acm}
-% \bibliography{references.bib}
 \end{document}

### 2017-03-07 drafts/aggregator/introduction.tex
various drafts and stuff idk

--- 2016-12-05 drafts/aggregator/introduction.tex
@@
 \documentclass[main]{subfiles}
-
-\makeatletter
-\def\blx@maxline{77}
-\makeatother
 \begin{document}
-
 \section{Introduction}\label{sec:introduction}
 
-Online labor markets ranging a variety of industries of (largely) service work
---- e.g. on--demand drivers (Uber, Lyft, Sidecar, etc\dots) and
-    information work (Amazon Mechanical Turk, Upwork, CrowdFlower, etc\dots) ---
-have struggled for some time with pernicious, in some cases \textit{growing},
-difficulty assessing the quality of its workforce,
-especially as it applies to a given task.
-In other words, whether a worker on Amazon Mechanical Turk (AMT) is likely
-to complete a task ``accurately'' remains an open question,
-leading researchers to address this pervasive problem in a number of ways
-\cite
+% \ali{
+%   Components of a good intro (anything else to add?):
+%   \begin{itemize}
+%     \item state of the world
+%     \item what's the problem?
+%     \item why this problem is such a big deal(?)
+%     \item what's making this problem so persistent?
+%     \item what's the insight we bring?
+%     \item what we built/did
+%     \item what did we find?
+%   \end{itemize}
+% }
 
-% Labor market platforms ranging industries and styles of work
-% (e.g. Uber, TaskRabbit, UpWork, and Amazon Mechanical Turk)
-% have struggled for years with persistent
-% (generally growing)
-% challenges relating to worker qualifications.
-% These issues range a wide spectrum:
-% in some cases, worker qualifications are non--transferable,
-% leading requesters to ``re--invent the wheel'' as they attempt to determine
-% in their own way whether a potential worker is qualified and reliable.
+% \topic{Crowdsourcing platforms have principally been limited to low--quality work, with
+% relatively little trust in either the quality or intent of workers~\cite{Ipeirotis:2010:QMA:1837885.1837906,MaliciousCrowdworkersGadiraju}.}
+% Workers who are new to a platform represent an ``unknown quantity'' to the platform,
+% which typically substantially limits the worker's eligibility to claim tasks (for instance, in the case of Amazon Mechanical Turk);
+% even when a worker is known to the platform, requesters themselves don't know whether a worker's reputation
+% --- for instance, a 99\% approval rating ---
+% represents a worker's competence in the sorts of tasks for which the requester is soliciting.
 
-% Problems determining a worker's qualifications start on day 1;
-% labor markets begin their relationships with new workers
-% almost entirely uncertain about the worker's competence in any type of task.
-% Gathering this information through qualification exams is generally
-% time--consuming and costly.
+% \topic{This problem has grown over time, and
+% various stopgap solutions to this fundamental absence of trust have emerged.}
+% A significant amount of effort has notably gone toward
+% designing workflows and processes that empower non--expert workers to satisfying complete work tasks.
+% Examples of this can be found in domains such as
+% creative writing~\cite{Kim2017,writingMicroTasks},
+% \ali{and, you know, other things here}, and more.
+% These approaches, with few exceptions, make the same assumption: that 
+% crowd workers are approximately interchangeable, and that
+% workers don't particularly specialize aside from a single--dimensional measure of approval rates
+% (e.g. limiting access to tasks to those with higher than 99.5\% approval ratings).
 
-% Challenges mount as workers' skill sets develop;
-% work requiring more training and skill (for example, translating or programming)
-% are either verified by individual \textit{requesters} (e.g. Amazon Mechanical Turk)
-% or are verified by the \textit{platform} itself (e.g. UpWork).
-% While the \textit{UpWork} model avoids needless repetitive work by
-% generally consolidating qualifying exams at the platform level,
-% these labor platforms nevertheless find themselves in the unenviable
-% (and often unexpected) position of
-% having to develop new qualifications exams to outpace would--be cheaters.
+% Many of these approaches assume that workers are cut from the same cloth;
+% that is, most workers are generally similarly qualified and thus interchangeable,
+% or rather to be more precise that
+% a worker's qualification is generally a function involving a small number of dimensions.
 
-\onlyinsubfile{
-% \balance{}
-\printbibliography{}
+% Our research begins by questioning this assumption.
+
+% \topic{On--demand labor has a problem with reputation.}
+
+\topic{One persistent source of frustration in crowdsourcing has been how to make more complex work viable.}
+A great deal of research has gone into various processes and work--flows that would facilitate complex outcomes, but
+much of this work presupposes that workers are generally unskilled.
+\citeauthor{crowdworkFuture} asked a number of years ago whether crowd work would ever become more than it was at the time ---
+``largely a dead--end job''~\cite{crowdworkFuture} ---
+and thus far the answer appears, reluctantly, to be``no''.
+
+\ali{Notes to do:
+\begin{enumerate}
+  \item challenges for workers
+  \begin{enumerate}
+    \item new to the platform
+    \item new to any unfamiliar requester
+  \end{enumerate}
+  \item challenges for requesters
+  \begin{enumerate}
+    \item complex qualifications are an unknown challenge (e.g. ``someone who's a good image tagger who also knows Tamil''?)
+    \item confidence in existing simple reputation systems is already poor;
+          at best, approval rating of past tasks is not very sensitive
+          (i.e. lots of workers with 99\% approval rates --- says nothing about specific approval rates of tagging or translating or whatever)
+  \end{enumerate}
+\end{enumerate}
 }
 
+\deadkittens{
+\topic{In the past decade, as markets for transient on--demand labor have grown in popularity,
+a number of challenges associated with these modes of work have emerged and grown as well.}
+In particular
+it remains to be seen how, or indeed \textit{if}, expert workers will participate in on--demand labor markets.
+While crowdsourcing and the gig economy have made it possible for millions of people to
+participate in the workforce in ways that were previously unavailable to them~\cite[][\ali{todo: add others}]{Zyskowski:2015:ACU:2675133.2675158},
+the longer--term trajectories of these people's careers are still as uncertain as they were when \citeauthor{crowdworkFuture} called crowd work
+``\dots largely a dead--end job''~\cite{crowdworkFuture}.
+So we find ourselves more than 5 years later answering the same question:
+will these people have the opportunity to engage in high--skill work in time?
+% as \citeauthor{crowdworkFuture} hypothesize~\cite{crowdworkFuture}?
+While research has found myriad ways
+to arrange and manage non--expert crowd workers to produce expert quality work,
+these achievements nevertheless do little to affirm that
+the future of work will afford for skilled workers, instead suggesting that
+the work experts do today may someday be done by a handful of non--experts.
+
+\topic{Lack of trust and reputation has held back high--skilled work.}
+Much of the work into complex crowdsourcing applications has operated on the premise that
+workers are roughly similar; they're all non--experts from a nebulous pool that can be sourced arbitrarily.
+Even where researchers recognize worker persistence,
+workers are often perceived and treated as untrustworthy~\cite{sensitiveTasks}.
+As a result, crowdsourcing innovations typically focus in the domain of workflows and processes
+\ali{--- find--fix--verify, and other examples? }~\cite{bernsteinSoylent,crowdForgeKittur}.
+And, on some level, this is in line with the ethos of on--demand work itself:
+as researchers have found,
+many crowd workers appreciate
+the fleeting nature of their relationship with
+the platform and with the people for whom they work
+~\ali{insert something here about how they like not being obligated to do work at certain times, etc\dots~(I'm pretty sure i have that in a citation or notes somewhere)}
+
+\topic{The reputation problem is closely related to a ``cold start'' problem.}
+Workers who are new to a platform can't signal to the platform that they're any good.
+Even when they're established in the platform,
+it can be difficult for requesters to discern high--quality workers;
+rating inflation
+(exhibited both in AMT as well as on Uber, etc\dots)
+complicates simple approval rating--based judgments, and
+in general these platforms only offer aggregate, coarse data on a worker's quality
+--- for example, their approval rate overall,
+or over the past several months,
+rather than within certain subsets of tasks.
+
+\topic{A historical analysis of on--demand work can inform potential solutions.}
+\ali{When we look at \citeauthor{pieceworkCrowdworkGigwork}'s work, we see that historical analysis can be really useful.
+We look to the solutions that emerged from nascent piecework and we identify \textit{hiring halls}, which not only served as places to source workers, but also managed the worker's credentials and qualifications.
+If a worker consistently did poor work, he might be kicked out of the group that ran the hiring hall, ensuring a certain level of quality.
+More importantly, if the worker}
+
+\topic{Aggregated places for people to store and reference their r\'{e}sum\'{e}s}
+
+\topic{What we built}
+
+\topic{What we found quantitatively}
+
+\topic{What we found qualitatively}
+}
+% \pagebreak
+
+  \textbf{crowd work has a reputation problem};
+        on--demand work has grown significantly;
+        task diversity is growing;
+        new platforms continue to emerge;
+        trust still a problem
+
+  \textbf{this is a cold start problem};
+     workers new to the platform are stuck slogging it;
+     workers new to a requester stuck too.
+
+  \textbf{existing platforms aren't great at this stuff};
+  Mechanical Turk 
+    % \begin{enumerate}
+    %   \item masters, upwork tests
+    %   \item hard to upkeep, long tail doom
+    % \end{enumerate}
+    % \item a few problems frustrate the design of interventions
+    % \begin{enumerate}
+    %   \item platforms have little incentive to facilitate external agents of intervention
+    %   \item \ali{does upwork outsource qualifications}
+    %   \item the stakeholders that platforms are more interested in seem to be those with the money (i.e. requesters)
+    %         \ali{maybe too argumentative?}
+    %   \item on--demand workers gravitate toward these platforms in large part because of the transience and ephemeral nature of the work
+    %         (therefore, working against that will meet with resistance (and possibly attrition) from the very workers who constitute the platform today)
+    % \end{enumerate}
+
+    \textbf{clever solution: r\'{e}sum\'{e}s for on--demand workers};
+    % \ali{imagine you have a resume with just your name and the total number of hours you've worked in your life}
+    % requesters don't know whom to trust and have to bootstrap their own trustworthiness evaluations
+    % platforms are stuck too (e.g. qualifications exams, credentials etc\dots)
+
+  % \end{enumerate}
+  % why is this problem a big deal?
+% \item requesters tend to undervalue workers --- at least at first --- leading to a ``market for lemons''
+
+  % Why is this so hard?
+
+  % What's our insight?
+  in ``conventional'' work, the problem of signaling a worker's quality was solved by using past worker performance as a proxy for future performance
+
+  on--demand work platforms have tried to do this with approval and rating systems, but they're generally too vague to be useful
+        % (on AMT, requesters limit work to those with approval ratings higher than 95\% or even 99\%, or workers with thousands or tens of thousands of tasks completed, or both)
+
+  conventional work r\'{e}sum\'{e}s usually allow the worker to emphasize and omit information that's not relevant --- \textbf{what if we facilitated that?}
+
+  % So what did we do?
+  we built a system that collected data from workers' profiles on online crowdsourcing platforms
+  (first AMT, then others), and parsed that data to present more meaningful analytic data on things like
+  \begin{enumerate}
+    \item the nature of the task; objective (i.e. factually based and evaluable) versus subjective (e.g. survey \ali{maybe also object labeling?})
+    \item the type of requester; academic versus industry
+    \item \textit{others}
+  \end{enumerate}
+  we also built a proof--of--concept qualifications management system abstracted from both our worker profile system and the work platform itself
+  (to explore the potential to abstract worker qualifications and credentials from the system itself, allowing other parties to specialize in this task)
+
+  % What did we find/how did it go?
+  we evaluated this system from two perspectives:
+  \begin{enumerate}
+    \item did requesters who used this system find that the worker data better--informed their qualifications management and worker solicitation?
+          in other words, was the output of the work more reliably accurate when they relied on the analytic data that we provided versus the coarse approval/rejection rating?
+    \item did \textit{workers} benefit from having high level data providing some insight into their work trends?
+did this lead them to better work more quickly?
+did they take a more active role in the management of their reputation / ``worker profile'' / r\'{e}sum\'{e}?
+  \end{enumerate}
+  we also explored whether the externalized worker qualification system worked effectively, and considered the potential advantages and disadvantages of a set of decoupled systems rather than an integrated one (market, qualifications assessment system, payment platform, etc\dots).
+% \end{enumerate}
+
+\section{dead kittens}
+\ali{this is getting a little argumentative, yeah?
+I want this to get in the direction of
+\textit{``Things aren't going the way they need to in order for crowd work to be good for people.''}}
+\topic{Today's on--demand labor markets don't seem to be driving toward a bright future for workers.}
+At one end of the spectrum,
+workers are treated as interchangeable cogs whose purposes are as
+backdrops upon which to project rubrics, workflows, and processes~\cite{yuanAlmost}.
+In this space, workers' lack of expertise is precisely what makes this form of labor compelling ---
+as an opportunity to show that, when arranged and managed correctly,
+``\dots non--experts can achieve better coverage and latency than a professional\dots''~\cite{lasecki2012real}.
+At the other end of the spectrum,
+workers in on--demand labor markets are seen as the walking dead ---
+human laborers who will ``satisfice'' until artificially intelligent agents can take over~\ali{[cite Kalanick being Kalanick]}.
+
 \end{document}

### 2017-03-07 drafts/streaming_busking/main.tex
various drafts and stuff idk

--- 2017-01-11 drafts/labor/trackingWork.tex
@@
-\documentclass{sigchi}
+\documentclass[chi_draft]{sigchi}
 
@@
 \usepackage[normalem]{ulem}
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
+% \usepackage{url}
+% \usepackage{natbib}
+% \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
 \usepackage[T1]{fontenc}
@@
 
-% \bibliographystyle{SIGCHI-Reference-Format}
 \renewcommand*{\notesname}{Graveyard of old paragraphs}
@@
 
-\newlist{numberlist}{enumerate*}{1}
-\setlist*[numberlist,1]{%
-   % label=\alph*),
- label={\nthwords*,},
-}
-
-\newlist{Numberlist}{enumerate*}{1}
-\setlist*[Numberlist,1]{%
-   % label=\alph*),
- label={\Nthwords*,},
-}
 \newlist{inlinelist}{enumerate*}{1}
@@
 \setlist{nolistsep}
-\def\plaintitle{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
-% \def\plaintitle{Crowd Work and Gig Work [as] Seen Through the Historical Lens of Piecework}
-\def\plainauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi}
+\def\plaintitle{Online Performance Art, and the Politics of Their Platforms}
+\def\plainauthor{Ali Alkhatib, Michael Scott Bernstein}
 \def\emptyauthor{}
-\def\plainkeywords{Crowd work; gig work; on--demand work; piecework}
-\def\plaingeneralterms{Group and Organization Interfaces}
+\def\plainkeywords{TODO}
+\def\plaingeneralterms{}
 
@@
 
-\bibliography{references}
-
 \newcommand{\subsubsubsection}[1]{\textit{#1}}
@@
 
-% \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
-\renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
-
 \begin{document}
-\CopyrightYear{2017} 
-\setcopyright{acmlicensed}
-\conferenceinfo{CHI 2017,}{May 06 - 11, 2017, Denver, CO, USA}
-\isbn{978-1-4503-4655-9/17/05}\acmPrice{\$15.00}
-\doi{http://dx.doi.org/10.1145/3025453.3025974}
 \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
@@
 
-\numberofauthors{1}
+\numberofauthors{2}
 \author{%
-  \alignauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi\\
-    \affaddr{Computer Science Department and CASBS}\\
+  \alignauthor{Ali Alkhatib\\
     \affaddr{Stanford University}\\
-    \email{\{ali.alkhatib, msb\}@cs.stanford.edu, mlevi@stanford.edu}}\\
+    \affaddr{Stanford, USA}\\
+    \email{ali.alkhatib@cs.stanford.edu}}\\
+  \alignauthor{Michael Scott Bernstein\\
+    \affaddr{Stanford University}\\
+    \affaddr{Stanford, USA}\\
+    \email{msb@cs.stanford.edu}}\\
 }
-% \author{
-% \alignauthor{Anonymized for Review\\
-% \affaddr{~}\\
-%     \affaddr{~}\\
-%     \email{~}}\\
-% }
+\author{
+\alignauthor{Anonymized for Review\\
+\affaddr{~}\\
+    \affaddr{~}\\
+    \email{~}}\\
+}
 \maketitle
@@
 \keywords{\plainkeywords}
-\ali{maybe put a table with high level summaries}
-\subfile{introduction.tex}
-\subfile{piecework_lit.tex}
-\subfile{cases.tex}
-\subfile{discussion.tex}
-\subfile{conclusion.tex}
+\subfile{introduction}
 
-% \balance{}
-% \printbibliography
+\bibliographystyle{SIGCHI-Reference-Format}
+\bibliography{references}
 

### 2017-03-08 proposals/Brown_Magic_Grant_2017/main.tex
magic grant init

--- 2017-03-07 proposals/SIGFproposal_2017/ali_alkhatib_sigf.tex
@@
-\documentclass{article}
+\documentclass[10pt]{article}
 \usepackage{xr-hyper}
-\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase,comment}
+\usepackage{todonotes,balance,graphics,color,alphalph,subfiles,appendix,endnotes,times}
+\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase,comment,titlesec}
 \usepackage[normalem]{ulem}
@@
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-\usepackage[T1]{fontenc}
+% \usepackage[T1]{fontenc}
 \usepackage[pdflang={en-US},pdftex]{hyperref}
 \usepackage{mfirstuc,lastpage}
-\usepackage[en-GB]{datetime2}
+\usepackage[en-US]{datetime2}
 \usepackage{fancyhdr,tabularx}
@@
 
-\usepackage{mfirstuc,lastpage}
+% \usepackage{mfirstuc,lastpage}
 \bibliography{references}
@@
 \pagestyle{fancy}
+\titlespacing*{\section}{0.0ex}{2ex plus .3ex}{1ex plus .3ex}
 
@@
 \makeatletter
-\def\@maketitle{   % custom maketitle 
-{\Large \scshape \@title} ~~~\@author ~ on  \@date \par 
-\smallskip \hrule \smallskip }
+
+\def\maketitle{%
+\par{\centering {\scshape \huge \textbf{\@title}}%
+\par{\@author}
+\par{submitted \today}%
+\par}}
+% \smallskip \hrule \smallskip }
 

### 2017-03-16 proposals/Brown_Magic_Grant_2017/main.tex
magic proposal

@@
 \usepackage{mfirstuc,lastpage}
-\usepackage[en-US]{datetime2}
 \usepackage{fancyhdr,tabularx}
@@
 \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+\usepackage[en-US]{datetime2}
+\DTMsettimestyle{default}
+\DTMsetup{showseconds=true}
 
@@
 \rhead{Ali Alkhatib} % controls the right corner of the header
-\lfoot{} % controls the left corner of the footer
+\lfoot{last updated} % controls the left corner of the footer
 \cfoot{} % controls the center of the footer
@@
 
-\thispagestyle{empty}% Removes the header from the first page. Change plain to empty to remove the numbering entirely.
+\thispagestyle{empty}
 
@@
 \par{\@author}
-\par{submitted \today}%
+\par{DRAFT \DTMcurrenttime~\today}%
+% \par{DRAFT \DTMcurrenttime~on~\today}%
 \par}}
@@
 
-\title{Designing Digital Hiring Halls}
+\definecolor{PineGreen}{HTML}{008800}
+\definecolor{Red}{HTML}{FF0000}
+\definecolor{Blue}{HTML}{0000FF}
+\definecolor{JokeGreen}{HTML}{00C953}
+\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
+\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
+
+\title{Ethnographic Study of Online Entertainers}
 \author{Ali Alkhatib}
@@
 
-% take half 
-% motivating work: historical analysis of crowd 
-% gig work instead of on--demand work
-%
-%
-%
-%
 \definecolor{Blue}{HTML}{0000FF}
 \newcommand{\topic}[1]{{\color{Blue}#1}}
-\renewcommand{\topic}[1]{{#1}}
+% \renewcommand{\topic}[1]{{#1}}
 \definecolor{Gray}{HTML}{BABABA}
@@
 
-% Michael Bernstein [9:40 PM]
-% [x] add a title
-% [x] “new instantiation” —> “modern instantiation”
-% [x] why “further” in a second sentence of an abstract?
-% [x] "design and develop a 'digital hiring hall’”— specify for gig workers?
-% [x] need a sentence in the abstract describing what your system actually does (e.g., create a portable resume across services by scraping work histories, letting workers use that history to demonstrate abilities to a new platform)
-% [x] final sentence of abstract has three broadbrush implications, choose one that it maximizes
-% [x] intro to Research Plan section needs a more focused question, specifically the one that you’re solving.
-"will gig work remain an undesirable career path” is too broad for our project to answer.
-likewise, "we propose to build one of the potential futures that we discussed” is too vague.
-Instead, start with a focused question more like (and this isn’t refined, but:) "How can workers demonstrate their skill?” And explain in one sentence this was an issue historically, and it was only partially solved b/c they couldn’t X.
-But now you can build technology to X.
-And that’s what you’re going to do.
-% [x] "who are proficient at translating between two specific languages”: ground this.
-e.g., they have lots of previously recorded work by a translation service on Mechanical Turk to vouch for their quality
-% [x] Interdisciplinarity: first sentence is weaksauce — get specific
-% [x] Contributions: “little say” — why is this the _optimistic_ prediction? wouldn’t the optimist say that we _can_ steer it?
-
 \begin{document}
@@
   \begin{abstract}
-    Gig work has become a modern instantiation of piecework.
-    Their platforms allow employers to
-    disperse workers geographically and silo workers' information,
-    making organization and coordination all but impossible,
-    and stymieing work specialization and expertise.
-    Our research will draw on our past work,
-    using historical analysis of piecework to inform this contemporary phenomenon,
-    to design and develop a ``digital hiring hall''
-    --- a platform--independent repository for workers to store their body of work, curate their professional image,
-    and better signal expertise to employers.
-    Our research will
-    begin to instantiate one of the potential futures of labor advocacy for gig workers.
+  TODO? ahhhh
   \end{abstract}
 
-\section*{Motivating Work: Historical Analysis of Piecework}
-\topic{Piecework has historically disempowered workers and empowered employers, and
-many the same dynamics of disenfranchisement and frustration are poised to return today.}
-In our most recent work~\cite{pieceworkCrowdworkGigwork}, we found that a number of aspects of piecework
---- a form of work payment that began more than a century ago ---
-bears striking resemblance to the sort of work we have begun to see emerge in Uber, Lyft, TaskRabbit, and more
---- collectively called ``gig work''.
-By subdividing work and paying for each task, employees are driven to work faster and longer,
-affording them little control over working conditions.
-And for most of its history, piecework yielded no benefits, occupational health and safety, or social insurance.
-When piecework combined with job routinization, workers became interchangeable and deindividuated.
-This combination of piecework and routinization produced the dominant payment system during the industrial era,
-starting with agriculture and home production but quickly moving into factories.
-However, by the turn of the \nth{20} century in the United States,
-campaigns for workers' rights yielded regulations on conditions, and by the late \nth{20} century,
-piecework was associated primarily with migrant labor and sweatshops.
-This contracted piecework economy has remained relatively stable since.
+\section*{Situation}
+\topic{In the past decade
+online video sharing % \ali{, such as YouTube and Vimeo,}
+and streaming sites % \ali{, like Twitch and others,}
+have increasingly taken on characteristics of
+Deweyan publics~\cite{dewey2012public,disalvo2009design}.}
+The ``YouTubers'' and ``streamers'' who use these platforms have come to rely both on
+the technical infrastructure these platforms provide \textit{as well as}
+the sociality of the venue itself.
+These sites have in some sense democratized media and entertainment for online performers by
+making the avenues to large audiences substantially more accessible.
+In these transformations into publics, these sites
+have grown into bustling hubs for artists to engage with audiences and practice their trade.
 
-\topic{Today, with the growth of on--demand labor, piecework is reemerging in new forms.}
-Networked computational infrastructure is now mediating each unit of work and worker.
-Uber's taxi drivers are paid per ride and assigned jobs by an algorithm~\cite{uberAlgorithm,hall2015analysis}.
-Information workers, such as those on Amazon Mechanical Turk and Upwork, are paid per task,
-performing vast quantities of data and administrative work~\cite{martin2014being}.
-Increasingly numerous workers make money through one--off tasks like housecleaning and food delivery, all mediated by online platforms.
-Much of this work has been limited in complexity to easily verifiable work, stymieing gig workers' career opportunities
-in many of the same ways that historically limited information workers' careers~\cite{grier2013computers}.
+\section*{Problem}
+\topic{Recently, tensions between online performers and the platforms on which they work have shaken people's
+assumptions regarding the nature of the space they're using.}
+Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
+YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
+and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
+\ali{now i want to say something like
+``these are all just symptoms of the fact that YouTube and these other platforms aren't the public spaces they were originally perceived to be;
+in fact these are carefully controlled spaces, and we need to better understand this stuff''}
 
-\topic{In our examination of crowd work and gig work as
-a reinstantiation of piecework~\cite{pieceworkCrowdworkGigwork},
-we identify a number of parallels that relate to this question.}
-First, that the first ``human computers'' were geographically distributed and
-only retained for several years to stifle career development opportunities~\cite{grier2013computers};
-second, that effective uses of historical piecework were generally limited to
-easily measured work due to a general mistrust~\cite{american1921problem,richards1904anything};
-third, this mistrust led to such a starkly adversarial relationship between workers and managers that
-it may have precipitated the wave of labor advocacy which defined
-the first half of the \nth{20} century~\cite{hart2013rise,hart2016rise}.
-What remains then are serious questions about the future of gig work as informed by piecework's history.
-We have limited knowledge of how gig workers today are adapting to digitally mediated, constantly changing labor markets.
-Similarly, we know very little about gig work's potential to support complex work, let alone how to turn that potential into reality.
+\section*{Insight}
+\topic{Studying contemporary developments through an established lens can be a helpful way of
+making sense of the phenomena we've seen so far, and
+provide us with some grounding to make predictions and inform future design decisions.}
+We've done this before, studying crowdsourcing and gig work as
+an instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork};
+this new framing of an existing domain with historical context not only
+provided us with guidance toward some of the major threads of research
+in the field of crowdsourcing, but
+also informed some of the potential futures of crowdsourcing and
+the even faster--growing gig work labor markets.
 
-\topic{This work carries policy implications for the emerging, and as yet largely unregulated, gig economy.}
-Will gig work represent empowered, independent professionals who freely control their destinies, or 
-will gig work become a geographically distributed ``digital sweatshop''~\cite{dawnDigitalSweatshopCushing}?
-Whereas researchers had relatively little control over the markets that emerged throughout the \nth{20} century,
-researchers of digital sites of work have the ability not only to influence, but to architect these settings~\cite{lessig2006code}.
-It can be argued that with a few lines of code, we can effect outsize change on the future of gig work
---- a form of work that continues to grow by the year~\cite{pewSharing,pewSharing24percent}.
+To some extent, we can liken online performance art on YouTube, Twitch, and other platforms to street performances.
+Online performers
+--- streamers and YouTubers in particular ---
+tend not to have the formal training of many conventional performers and artists,
+instead growing in proficiency and becoming more skilled principally through their practice;
+in much the same way,
+street performers are 
+%
+Online performance artists have the ability
+to engage directly with their audiences and tailor their performances to the desires of the crowd
+--- YouTubers call upon their audiences to like, dislike, and submit comments to the video, while
+    streamers literally change their performances on the fly to suit the whims of those watching and chatting;
+meanwhile, the most successful street performers typically engage participants,
+drawing them into the performance itself, even bantering with passersby.
+%
+Finally, online performance has increasingly become a site of political engagement and activism,
+sometimes subverting authority
+--- in these cases online moderators and algorithms ---
+in the process of or for the purpose of making a point;
+the history of street performance as a channel for political action and advocacy has been deeply studied
+(see, for example, \cite{harrison1990drawing,cohen1998radical}).
+\section*{Incomplete!}
+\section*{Proposal}
 
-% [x] intro to Research Plan section needs a more focused question, specifically the one that you’re solving.
-"will gig work remain an undesirable career path” is too broad for our project to answer.
-likewise, "we propose to build one of the potential futures that we discussed” is too vague.
-Instead, start with a focused question more like (and this isn’t refined, but:) "How can workers demonstrate their skill?” And explain in one sentence this was an issue historically, and it was only partially solved b/c they couldn’t X.
-But now you can build technology to X.
-And that’s what you’re going to do.
+% \section*{Motivating Work: Historical Analysis of Contemporary Phenomena}
+
+% Online performers share an uncanny number of similarities with street performers
+
+% \ali{generally speaking, each major bullet point = roughly 1 paragraph}
+\begin{itemize}
+  \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
+        a growing industry for entertainers.
+        Further, it has ``democratized'' entertainment, providing people with a public space
+        (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
+  \item Our research into crowdsourcing and gig work has yielded
+        a compelling framework for these phenomena as an instantiation of piecework.
+        We've found that this approach
+        --- that is, drawing parallels with historical phenomena ---
+        can generally give researchers of sociotechnical systems substantive framing of a research space.
+  \item To some extent, this can be likened to \textbf{street performance} inasmuch as
+  \begin{itemize}
+    \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
+    \item entertainers could get more direct feedback from their audience than other avenues could afford them
+    \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
+  \end{itemize}
+  \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
+        \ali{need a clearer \textbf{research question}?}
+  \item When we look at this dimension of online performance art, the metaphor begins to break down:
+  \begin{itemize}
+    \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
+    \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
+    \item communication between the performer and the audience is carefully mediated
+          (and literally moderated) by the platform itself
+    \item online performances are (generally) more permanent than street performances
+          \ali{technology has thrown a wrench into this lately, with
+              cameras allowing us to record street performances and
+              tools making it impossible (or at least difficult) to save online streams, but
+              \textit{let's not worry about all that}.}
+  \end{itemize}
+  \item So while there are compelling parallels between online performances and the street art of busking,
+        this has only taken us partway in answering important questions about a rapidly growing field.
+\end{itemize}
+
 \section*{Research Plan}
-\topic{While our research considered a number of possible futures for gig work in general and gig workers in particular,
-several important questions remain unresolved.}
-For example, we pointed to the general question that researchers had been asking for some time ---
-whether gig work will become a site of complex, rewarding work
---- or if it will remain a setting for low quality tasks with little upward mobility.
-While we offered a framework for answering questions such as this one, the reality is that
-any researcher or team exploring this area
---- especially on the basis of the parallel we drew ---
-would fundamentally need to be conversant with several deep bodies of research spanning myriad fields.
+% \begin{itemize}
+%   \item We propose to direct our research energy toward answering one question: 
+% \end{itemize}
 
-We concluded that gig work's potential will be determined by
-how easily skilled workers can secure good work through these platforms.
-In order to do that, workers need to be able to communicate rich information about their qualifications to employers,
-which existing markets have so far struggled to do.
-\topic{We will build a system that tracks gig workers' professional histories across myriad work platforms,
-aggregating, consolidating, \& analyzing that data, and
-allowing workers to curate their professional identities, instantiated by a ``digital hiring hall'' for gig workers.}
+% We propose to conduct ethnographic research of online performance artists who principally earn their income either from Twitch or YouTube (or both).
 
-\topic{The system we intend to build will serve two roles --- one for workers and one for employers.}
-Gig workers will share their work histories from platforms such as Amazon Mechanical Turk with our repository,
-which will allow us to provide the worker with information upon which they can reflect
-(e.g. detailed analysis of their work trends, approval rates under various circumstances, etc\dots)
-as well as information in relation to other workers
-(e.g. offering a percentile ranking overall for approval rates, or within specific types of tasks).
-In doing so, we hope to find that workers will identify cues and gravitate toward qualities in work
-in which they identify competitive advantage.
+% \topic{While our research considered a number of possible futures for gig work in general and gig workers in particular,
+% several important questions remain unresolved.}
+% For example, we pointed to the general question that researchers had been asking for some time ---
+% whether gig work will become a site of complex, rewarding work
+% --- or if it will remain a setting for low quality tasks with little upward mobility.
+% While we offered a framework for answering questions such as this one, the reality is that
+% any researcher or team exploring this area
+% --- especially on the basis of the parallel we drew ---
+% would fundamentally need to be conversant with several deep bodies of research spanning myriad fields.
 
-\topic{The second role that this repository will help employers identify and source the best workers for highly skilled work.}
-By aggregating the data from work platforms and
-offering more in--depth analysis of workers' histories, we hope to find that
-we can direct employers to gig workers whose overall experiences match the employer's given needs.
-For example, if an employer or a platform queries our platform for
-workers who are proficient at translating between two specific languages,
-we can yield a list of workers who would be suitable for the task, determined by
-the workers' comprehensive histories on Mechanical Turk and, potentially, elsewhere.
+% We concluded that gig work's potential will be determined by
+% how easily skilled workers can secure good work through these platforms.
+% In order to do that, workers need to be able to communicate rich information about their qualifications to employers,
+% which existing markets have so far struggled to do.
+% \topic{We will build a system that tracks gig workers' professional histories across myriad work platforms,
+% aggregating, consolidating, \& analyzing that data, and
+% allowing workers to curate their professional identities, instantiated by a ``digital hiring hall'' for gig workers.}
 
-\topic{This approach is similar in some senses to the role that hiring halls played through the \nth{20} century.}
-By disconnecting workers' data from the platforms themselves, our repository will act as a sort of ``digital hiring hall'',
-connecting employers with workers whose varied work histories provide them with the expertise necessary to do given tasks.
-By engaging workers in the management of this repository, we will advance the metaphor in much the same way that
-labor unions maintained --- and continue to operate --- hiring halls of their workers.
+% \topic{The system we intend to build will serve two roles --- one for workers and one for employers.}
+% Gig workers will share their work histories from platforms such as Amazon Mechanical Turk with our repository,
+% which will allow us to provide the worker with information upon which they can reflect
+% (e.g. detailed analysis of their work trends, approval rates under various circumstances, etc\dots)
+% as well as information in relation to other workers
+% (e.g. offering a percentile ranking overall for approval rates, or within specific types of tasks).
+% In doing so, we hope to find that workers will identify cues and gravitate toward qualities in work
+% in which they identify competitive advantage.
 
-\topic{At a high level, we can study the usefulness of this system using standard experimental approaches.}
-First, through quantitative methods (for example, comparing workers' earnings over time); and
-second, through qualitative  methods (for example, through interviews to determine whether workers find and exploit specialty niches as a result of this tool).
-We can also explore employers perspectives in tangible ways;
-for example, given concerns about the quality of work~\cite{Ipeirotis:2010:QMA:1837885.1837906,le2010ensuring,Law:2017:CTR:2998181.2998197},
-we will investigate whether a system that better communicates worker expertise allows employers to rely more confidently on crowdwork.
+% \topic{The second role that this repository will help employers identify and source the best workers for highly skilled work.}
+% By aggregating the data from work platforms and
+% offering more in--depth analysis of workers' histories, we hope to find that
+% we can direct employers to gig workers whose overall experiences match the employer's given needs.
+% For example, if an employer or a platform queries our platform for
+% workers who are proficient at translating between two specific languages,
+% we can yield a list of workers who would be suitable for the task, determined by
+% the workers' comprehensive histories on Mechanical Turk and, potentially, elsewhere.
 
-\section*{Interdisciplinarity}
-% \topic{This research will call on methods and skills originally from and generally found in disparate fields.}
-% \topic{The insights and expertise necessary to conduct this research successfully will call on a variety of skills not typically seen together, such as
-% participant observation, participatory design techniques, and machine learning approaches.}
-As we showed with our paper examining the relationship between piecework and gig work~\cite{pieceworkCrowdworkGigwork},
-the lens of a social scientist can inform conversations about digitally mediated work in ways that many computer scientists are not otherwise able to offer.
-At the same time, the perspectives of computer scientists
---- and the challenges digitally mediated spaces bring with them ---
-offer to reinvigorate deeply theoretically grounded discussion and offer field sites with which to experiment and test these theories.
-As a Computer Science Ph.D. student originally trained as an Anthropologist, I believe
-I'm the best candidate to combine these bodies of literature to make this project succeed.
+% \topic{This approach is similar in some senses to the role that hiring halls played through the \nth{20} century.}
+% By disconnecting workers' data from the platforms themselves, our repository will act as a sort of ``digital hiring hall'',
+% connecting employers with workers whose varied work histories provide them with the expertise necessary to do given tasks.
+% By engaging workers in the management of this repository, we will advance the metaphor in much the same way that
+% labor unions maintained --- and continue to operate --- hiring halls of their workers.
 
-\section*{Anticipated Contributions \& Implications}
-This research is of significant importance given the recent trends in the ``gig economy''.
-With more and more people joining the workforce in this capacity~\cite{pewSharing,pewSharing24percent},
-concerns about ``the future of work''~\cite{waterCooler,crowdworkFuture} continue to grow.
-Our hope is that the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
-More broadly, we expect to offer two major contributions to the Human--Computer Interaction community:
-first, an existence proof that a gig work market for complex, specialized workers is possible; and
-second, insights into how best to scaffold and foster ongoing collective governance.
+% \topic{At a high level, we can study the usefulness of this system using standard experimental approaches.}
+% First, through quantitative methods (for example, comparing workers' earnings over time); and
+% second, through qualitative  methods (for example, through interviews to determine whether workers find and exploit specialty niches as a result of this tool).
+% We can also explore employers perspectives in tangible ways;
+% for example, given concerns about the quality of work~\cite{Ipeirotis:2010:QMA:1837885.1837906,le2010ensuring,Law:2017:CTR:2998181.2998197},
+% we will investigate whether a system that better communicates worker expertise allows employers to rely more confidently on crowdwork.
 
-This project's will speak to the critique that the internet,
-far from the democratizing force that Barlow predicted it would be,
-has become an infrastructure facilitating the corporate centralization of power~\cite{barlow2009declaration,jones2011does,EboCybertopia}.
-The success of this project will therefore not only
-bring a more nuanced understanding of the shifting climate and culture of labor,
-but may
---- in a small way ---
-deliver on the promise Barlow made that
-cyberspace and the Internet would be a tool for democratizing access to information, and with that,
-empowering more than the tech elite who engineered it.
+% \section*{Interdisciplinarity}
+% % \topic{This research will call on methods and skills originally from and generally found in disparate fields.}
+% % \topic{The insights and expertise necessary to conduct this research successfully will call on a variety of skills not typically seen together, such as
+% % participant observation, participatory design techniques, and machine learning approaches.}
+% As we showed with our paper examining the relationship between piecework and gig work~\cite{pieceworkCrowdworkGigwork},
+% the lens of a social scientist can inform conversations about digitally mediated work in ways that many computer scientists are not otherwise able to offer.
+% At the same time, the perspectives of computer scientists
+% --- and the challenges digitally mediated spaces bring with them ---
+% offer to reinvigorate deeply theoretically grounded discussion and offer field sites with which to experiment and test these theories.
+% As a Computer Science Ph.D. student originally trained as an Anthropologist, I believe
+% I'm the best candidate to combine these bodies of literature to make this project succeed.
+
+% \section*{Anticipated Contributions \& Implications}
+% This research is of significant importance given the recent trends in the ``gig economy''.
+% With more and more people joining the workforce in this capacity~\cite{pewSharing,pewSharing24percent},
+% concerns about ``the future of work''~\cite{waterCooler,crowdworkFuture} continue to grow.
+% Our hope is that the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
+% More broadly, we expect to offer two major contributions to the Human--Computer Interaction community:
+% first, an existence proof that a gig work market for complex, specialized workers is possible; and
+% second, insights into how best to scaffold and foster ongoing collective governance.
+
+% This project's will speak to the critique that the internet,
+% far from the democratizing force that Barlow predicted it would be,
+% has become an infrastructure facilitating the corporate centralization of power~\cite{barlow2009declaration,jones2011does,EboCybertopia}.
+% The success of this project will therefore not only
+% bring a more nuanced understanding of the shifting climate and culture of labor,
+% but may
+% --- in a small way ---
+% deliver on the promise Barlow made that
+% cyberspace and the Internet would be a tool for democratizing access to information, and with that,
+% empowering more than the tech elite who engineered it.
 

### 2017-03-16 proposals/Brown_Magic_Grant_2017/main.tex
relatively significant revisions to problem and insight

@@
 % \renewcommand{\topic}[1]{{#1}}
-\definecolor{Gray}{HTML}{BABABA}
-\newcommand{\ugh}[1]{{\color{Gray}\textit{#1}}}
+\definecolor{Gray}{HTML}{A0A0A0}
+\newcommand{\ugh}[1]{{\color{Gray}#1}}
 \newenvironment{blah}{\par\color{Gray}\itshape}{\par}
@@
 \section*{Situation}
-\topic{In the past decade
-online video sharing % \ali{, such as YouTube and Vimeo,}
+\topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
 and streaming sites % \ali{, like Twitch and others,}
-have increasingly taken on characteristics of
-Deweyan publics~\cite{dewey2012public,disalvo2009design}.}
-The ``YouTubers'' and ``streamers'' who use these platforms have come to rely both on
-the technical infrastructure these platforms provide \textit{as well as}
-the sociality of the venue itself.
-These sites have in some sense democratized media and entertainment for online performers by
+have garnered the attention of audiences around the world
+in the span of little more than a decade.} % taken on characteristics of
+So widely used are these platforms
+that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
+content producers
+--- ``YouTubers'' and ``streamers'' ---
+have begun to make their livings on these sites.
+Indeed, these performers have come to rely on
+the technical infrastructure these platforms provide as well as
+the access to audiences that these platforms provide as well.
+In doing so, these sites have in a real sense democratized media and entertainment industries by
 making the avenues to large audiences substantially more accessible.
-In these transformations into publics, these sites
-have grown into bustling hubs for artists to engage with audiences and practice their trade.
+And thus, in these transformations into publics, what started as
+a recreational hobby
+has become a bustling hub for professional artists to
+engage with audiences,
+practice their craft, and 
+hone their skills~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}.
 
 \section*{Problem}
-\topic{Recently, tensions between online performers and the platforms on which they work have shaken people's
+\topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
 assumptions regarding the nature of the space they're using.}
@@
 and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
-\ali{now i want to say something like
-``these are all just symptoms of the fact that YouTube and these other platforms aren't the public spaces they were originally perceived to be;
-in fact these are carefully controlled spaces, and we need to better understand this stuff''}
+\ugh{These events fundamentally speak to the contentious politics of an emerging form of expression and art,
+and specifically to the uncertainty regarding the status of these artists who tend to subvert not just
+the norms of their chosen arts, but also
+orthodox labels of performers in the first place.}
 
 \section*{Insight}
-\topic{Studying contemporary developments through an established lens can be a helpful way of
-making sense of the phenomena we've seen so far, and
-provide us with some grounding to make predictions and inform future design decisions.}
-We've done this before, studying crowdsourcing and gig work as
-an instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork};
-this new framing of an existing domain with historical context not only
+\topic{Studying contemporary phenomena through an established framing
+can be a helpful way of
+making sense of what we've seen so far, as well as
+providing us with some grounding to make predictions.}
+We've done this before, studying crowd work and gig work as
+a modern instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork}.
+This positioning of an existing subject in historical context not only
 provided us with guidance toward some of the major threads of research
-in the field of crowdsourcing, but
+in crowdsourcing, but
 also informed some of the potential futures of crowdsourcing and
-the even faster--growing gig work labor markets.
+the ever growing gig work labor markets.
 
-To some extent, we can liken online performance art on YouTube, Twitch, and other platforms to street performances.
+\topic{We can liken some aspects of online performance art on YouTube and Twitch
+to street performance and busking,
+allowing us to relate ostensibly new phenomena to robust scholarship.}
+As is the case with street performance,
+online performance isn't restricted to classically trained professionals;
+instead, performers (both online and in the streets) practice more experimental 
+drawing on their candid interaction with their audiences,
+which can change from one performance to the next.
+
+\begin{blah}
 Online performers
---- streamers and YouTubers in particular ---
 tend not to have the formal training of many conventional performers and artists,
@@
 (see, for example, \cite{harrison1990drawing,cohen1998radical}).
+\end{blah}
+
 \section*{Incomplete!}

### 2017-03-16 proposals/Brown_Magic_Grant_2017/main.tex
framing rough draft done?

@@
   \begin{abstract}
-  TODO? ahhhh
+  \ali{something along the lines of: online performers are becoming a big deal, and
+       these performers are possibly the forerunners of a new type of expression and art.
+       The frustration and turmoil they're experiencing
+       --- both interpersonally and with the environment on which they ``work'' ---
+       demands to be better understood if we hope to make informed design decisions that empower these people.
+       We propose to conduct ethnographic fieldwork of this domain,
+       initially informed by a historical framing, but
+       ultimately studying how digitally mediated publics such as these affect
+       the politics of this group of performers.}
   \end{abstract}
 
-\section*{Situation}
+% \section*{Situation}
 \topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
@@
 
-\section*{Problem}
+% \section*{Problem}
 \topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
@@
 and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
-\ugh{These events fundamentally speak to the contentious politics of an emerging form of expression and art,
-and specifically to the uncertainty regarding the status of these artists who tend to subvert not just
-the norms of their chosen arts, but also
-orthodox labels of performers in the first place.}
+These events highlight the politics of an emerging form of expression and art,
+and specifically the sometimes controversial nature of the experimentation that performers engage in.
 
-\section*{Insight}
+% \section*{Insight}
 \topic{Studying contemporary phenomena through an established framing
@@
 
-\begin{blah}
-Online performers
-tend not to have the formal training of many conventional performers and artists,
-instead growing in proficiency and becoming more skilled principally through their practice;
-in much the same way,
-street performers are 
-%
-Online performance artists have the ability
-to engage directly with their audiences and tailor their performances to the desires of the crowd
---- YouTubers call upon their audiences to like, dislike, and submit comments to the video, while
-    streamers literally change their performances on the fly to suit the whims of those watching and chatting;
-meanwhile, the most successful street performers typically engage participants,
-drawing them into the performance itself, even bantering with passersby.
-%
-Finally, online performance has increasingly become a site of political engagement and activism,
-sometimes subverting authority
---- in these cases online moderators and algorithms ---
-in the process of or for the purpose of making a point;
-the history of street performance as a channel for political action and advocacy has been deeply studied
-(see, for example, \cite{harrison1990drawing,cohen1998radical}).
-\end{blah}
+\topic{This connection has purchase, but it has flaws, mostly owing to the unique nature of ``networked publics''~\cite{boyd2007youth}.}
+For one thing, most video sharing websites keep videos
+% --- the modern performances ---
+in perpetuity
+--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
+The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
+YouTube comments linger for years after they're made.
 
-\section*{Incomplete!}
-\section*{Proposal}
-
-% \section*{Motivating Work: Historical Analysis of Contemporary Phenomena}
-
-% Online performers share an uncanny number of similarities with street performers
+% \section*{TODO: Problem with this model}
+% \section*{TODO: Proposal}
+% \section*{Incomplete!}
+% \section*{Proposal}
 
 % \ali{generally speaking, each major bullet point = roughly 1 paragraph}
-\begin{itemize}
-  \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
-        a growing industry for entertainers.
-        Further, it has ``democratized'' entertainment, providing people with a public space
-        (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
-  \item Our research into crowdsourcing and gig work has yielded
-        a compelling framework for these phenomena as an instantiation of piecework.
-        We've found that this approach
-        --- that is, drawing parallels with historical phenomena ---
-        can generally give researchers of sociotechnical systems substantive framing of a research space.
-  \item To some extent, this can be likened to \textbf{street performance} inasmuch as
-  \begin{itemize}
-    \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
-    \item entertainers could get more direct feedback from their audience than other avenues could afford them
-    \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
-  \end{itemize}
-  \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
-        \ali{need a clearer \textbf{research question}?}
-  \item When we look at this dimension of online performance art, the metaphor begins to break down:
-  \begin{itemize}
-    \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
-    \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
-    \item communication between the performer and the audience is carefully mediated
-          (and literally moderated) by the platform itself
-    \item online performances are (generally) more permanent than street performances
-          \ali{technology has thrown a wrench into this lately, with
-              cameras allowing us to record street performances and
-              tools making it impossible (or at least difficult) to save online streams, but
-              \textit{let's not worry about all that}.}
-  \end{itemize}
-  \item So while there are compelling parallels between online performances and the street art of busking,
-        this has only taken us partway in answering important questions about a rapidly growing field.
-\end{itemize}
+% \begin{itemize}
+%   \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
+%         a growing industry for entertainers.
+%         Further, it has ``democratized'' entertainment, providing people with a public space
+%         (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
+%   \item Our research into crowdsourcing and gig work has yielded
+%         a compelling framework for these phenomena as an instantiation of piecework.
+%         We've found that this approach
+%         --- that is, drawing parallels with historical phenomena ---
+%         can generally give researchers of sociotechnical systems substantive framing of a research space.
+%   \item To some extent, this can be likened to \textbf{street performance} inasmuch as
+%   \begin{itemize}
+%     \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
+%     \item entertainers could get more direct feedback from their audience than other avenues could afford them
+%     \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
+%   \end{itemize}
+%   \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
+%         \ali{need a clearer \textbf{research question}?}
+%   \item When we look at this dimension of online performance art, the metaphor begins to break down:
+%   \begin{itemize}
+%     \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
+%     \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
+%     \item communication between the performer and the audience is carefully mediated
+%           (and literally moderated) by the platform itself
+%     \item online performances are (generally) more permanent than street performances
+%           \ali{technology has thrown a wrench into this lately, with
+%               cameras allowing us to record street performances and
+%               tools making it impossible (or at least difficult) to save online streams, but
+%               \textit{let's not worry about all that}.}
+%   \end{itemize}
+%   \item So while there are compelling parallels between online performances and the street art of busking,
+%         this has only taken us partway in answering important questions about a rapidly growing field.
+% \end{itemize}
 
-\section*{Research Plan}
+\section*{Research Proposal}
+
+\section*{Broader Impact}
+\topic{Formally recognized art forms have assimilated emergent genres and types of performance art in the past, and
+we should expect that they will continue to do so.}
+Jazz was born out of what largely ``\dots was recognized~\dots~as being but one species of the genus, ragtime played by ear by fakers\dots''~\cite{10.2307/779456}.
+Even further back in history, \nth{17} century street performance in Japan blended ``\dots every imaginable genus and species of popular song and dance\dots'' and
+kabuki, in its early years,
+``would hardly have been recognized or condoned [by traditional performers]''~\cite{groemer2016street}.
+
+\ali{something like: if this continues, then we should infer that
+     the future of this type of performance art will bear heavily on
+     the direction and trajectory of digitally mediated entertainment and performance}
+
+\section*{Demonstrate viability}
+
+\section*{Expected outcomes}
+
 % \begin{itemize}

### 2017-03-17 proposals/Brown_Magic_Grant_2017/main.tex
oh right forgot to state the obvious
updates to framing based on conversation with msb and some of ethan's feedback also added some bullet points to the research plan

@@
 
+\topic{One issue frustrating this potential research area is that
+the platforms have changed so dramatically in as few as 5 to 10 years that
+it can seem as though researchers are trying to wade through shifting sands.}
+Scholarship on content creators loses bearing rapidly when economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
+the tension between these entertainers experimenting with content remixing and
+the holders of the copyright of the source material
+(for example, see~\cite{Hilderbrand48})
+shifts dramatically when
+the platform's copyright enforcement mechanisms advance
+not just incrementally but substantially~\cite{kim2012institutionalization}.
+
 % \section*{Insight}
-\topic{Studying contemporary phenomena through an established framing
+\topic{We recognize that \textit{historical analysis}
+% Studying contemporary phenomena through an established framing
 can be a helpful way of
-making sense of what we've seen so far, as well as
-providing us with some grounding to make predictions.}
-We've done this before, studying crowd work and gig work as
-a modern instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork}.
-This positioning of an existing subject in historical context not only
-provided us with guidance toward some of the major threads of research
-in crowdsourcing, but
-also informed some of the potential futures of crowdsourcing and
-the ever growing gig work labor markets.
+making sense of contemporary phenomena.}
+As we found in a recent paper framing crowd work and gig work as a modern instantiation of \textit{piecework},
+scholarship from a parallel or similar domain can
+provide us some grounding to make sense of what we've seen, as well as inform predictions~\cite{pieceworkCrowdworkGigwork}.
+% This positioning of an existing subject in historical context not only
+% provided us with guidance toward some of the major threads of research
+% in crowdsourcing, but
+% also informed some of the potential futures of crowdsourcing and
+% the ever growing gig work labor markets.
 
-\topic{We can liken some aspects of online performance art on YouTube and Twitch
+\topic{To that end, we can liken some aspects of online performance art on YouTube and Twitch
 to street performance and busking,
@@
 YouTube comments linger for years after they're made.
+And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
+the near--perfect ability of these platforms to redirect revenue and even people's focus instantaneously.
 
@@
 
-\section*{Research Proposal}
+\section*{Research Plan}
+\ali{Basic plan:
+\begin{itemize}
+  \item contact YouTube channels that have more than 1M subscribers
+        (and have mentioned YouTube policies at some point
+          (suggesting maybe passing interest in discussing meta topics)?)
+  \item Reach out by any contact methods they provide (if none, then skip)
+  \item Semi--structured interview guiding towards:
+  \begin{itemize}
+    \item interaction with platform \& its designers
+    \item engagement with audience
+    \item coordination with other YouTubers (if any)
+  \end{itemize}
+  \item tentative questions include:
+  \begin{itemize}
+    \item platform questions
+    \begin{enumerate}
+      \item when did you switch to (streaming/YouTubing) as primary career?
+      \item what's YouTube's relationship with you?
+      \item if you were telling a new YouTuber how to get started,
+            what would you say they should know about YouTube
+            (the system, the company, etc\dots)
+            that's not immediately obvious?
+    \end{enumerate}
+    \item audience questions
+    \begin{enumerate}
+      \item how do you interact with audiences?
+      \item what sorts of comments do you get?
+      \item do you reply?
+      \item do you change your future content based on feedback in the comments?
+    \end{enumerate}
+    \item peer questions
+    \begin{enumerate}
+      \item what's your sense of the community as a whole?
+      \item how many other YouTubers do you know
+            (let's say you've had a chat with them in the last month and
+             will probably do so again in the next month)?
+      \item are you all cooperative? competitive? what's the culture like?
+    \end{enumerate}
+  \end{itemize}
+\end{itemize}
+}
 
-\section*{Broader Impact}
-\topic{Formally recognized art forms have assimilated emergent genres and types of performance art in the past, and
-we should expect that they will continue to do so.}
-Jazz was born out of what largely ``\dots was recognized~\dots~as being but one species of the genus, ragtime played by ear by fakers\dots''~\cite{10.2307/779456}.
-Even further back in history, \nth{17} century street performance in Japan blended ``\dots every imaginable genus and species of popular song and dance\dots'' and
-kabuki, in its early years,
-``would hardly have been recognized or condoned [by traditional performers]''~\cite{groemer2016street}.
+% \section*{Broader Impact}
+% \topic{Formally recognized art forms have assimilated emergent genres and types of performance art in the past, and
+% we should expect that they will continue to do so.}
+% Jazz was born out of what largely ``\dots was recognized~\dots~as being but one species of the genus, ragtime played by ear by fakers\dots''~\cite{10.2307/779456}.
+% Even further back in history, \nth{17} century street performance in Japan blended ``\dots every imaginable genus and species of popular song and dance\dots'' and
+% kabuki, in its early years,
+% ``would hardly have been recognized or condoned [by traditional performers]''~\cite{groemer2016street}.
 
-\ali{something like: if this continues, then we should infer that
-     the future of this type of performance art will bear heavily on
-     the direction and trajectory of digitally mediated entertainment and performance}
+% \ali{something like: if this continues, then we should infer that
+%      the future of this type of performance art will bear heavily on
+%      the direction and trajectory of digitally mediated entertainment and performance}
 
-\section*{Demonstrate viability}
+% \section*{Demonstrate viability}
 
-\section*{Expected outcomes}
+% \section*{Expected outcomes}
 

### 2017-03-17 proposals/Brown_Magic_Grant_2017/main.tex
proposal coming close to 'complete' status

@@
 
+\usepackage{caption}
+\captionsetup[table]{labelformat=empty}
+
 % \usepackage{mfirstuc,lastpage}
@@
 % \chead{} % controls the center of the header
-\rhead{Ali Alkhatib} % controls the right corner of the header
+\rhead{Alkhatib \& Nagy} % controls the right corner of the header
 \lfoot{last updated} % controls the left corner of the footer
@@
 
+\AtBeginBibliography{\small}
 \thispagestyle{empty}
@@
 
-% \newlist{inlinelist}{enumerate*}{1}
-% \setlist*[inlinelist,1]{
-%   label=\arabic*),
-% }
+\newlist{inlinelist}{enumerate*}{1}
+\setlist*[inlinelist,1]{
+  label=\arabic*),
+}
 \makeatletter
@@
 \def\maketitle{%
-\par{\centering {\scshape \huge \textbf{\@title}}%
+\par{\centering {\scshape \LARGE \textbf{\@title}}%
 \par{\@author}
@@
 \definecolor{Blue}{HTML}{0000FF}
+\definecolor{Black}{HTML}{000000}
 \definecolor{JokeGreen}{HTML}{00C953}
@@
 
-\title{Ethnographic Study of Online Entertainers}
-\author{Ali Alkhatib}
+\title{An Ethnography of Online Performance Artists}
+\author{Ali Alkhatib \& Jeffrey Nagy}
 \date{\today}
+\urlstyle{same}
+\hypersetup{
+  colorlinks,
+  citecolor=Black,
+  urlcolor=Blue}
 
@@
 \newcommand{\topic}[1]{{\color{Blue}#1}}
-% \renewcommand{\topic}[1]{{#1}}
+\renewcommand{\topic}[1]{{#1}}
 \definecolor{Gray}{HTML}{A0A0A0}
@@
 it can seem as though researchers are trying to wade through shifting sands.}
-Scholarship on content creators loses bearing rapidly when economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
+Scholarship on content creators loses bearing rapidly when
+economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
 the tension between these entertainers experimenting with content remixing and
@@
 not just incrementally but substantially~\cite{kim2012institutionalization}.
+And the popularity of candidate field sites can vary chaotically
+(see, for example, Vine reaching as many as 200 million active users and
+abrupt shutdown less than a year later~\cite{vineDecline}).
 
@@
 provide us some grounding to make sense of what we've seen, as well as inform predictions~\cite{pieceworkCrowdworkGigwork}.
-% This positioning of an existing subject in historical context not only
-% provided us with guidance toward some of the major threads of research
-% in crowdsourcing, but
-% also informed some of the potential futures of crowdsourcing and
-% the ever growing gig work labor markets.
 
@@
 
-\topic{This connection has purchase, but it has flaws, mostly owing to the unique nature of ``networked publics''~\cite{boyd2007youth}.}
+\topic{This connection has purchase, but it has gaps, mostly owing to the unique nature of the internet~\cite{miller2011understanding}.}
 For one thing, most video sharing websites keep videos
-% --- the modern performances ---
 in perpetuity
@@
 The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
-YouTube comments linger for years after they're made.
+YouTube comments linger for years after they're made, forming in some sense ``networked publics''~\cite{boyd2007youth}.
 And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
 the near--perfect ability of these platforms to redirect revenue and even people's focus instantaneously.
+
+\topic{We propose to fill in these gaps in our framing of online performance as a modern instantiation of street performance.}
+How does the \textit{permanence} of digitally archived videos affect people's willingness to experiment with new formats and styles,
+and \textit{how can we enable and encourage riskier, potentially more rewarding experimentation}?
+How do networked publics influence whether and how performers engage with their audiences,
+and \textit{how can we design publics that engage the audience with each other and with the performer}?
+How do laws, and the algorithmically instantiated policies, on video sharing and streaming platforms spur or stifle creative remixing?
 
@@
 \section*{Research Plan}
-\ali{Basic plan:
-\begin{itemize}
-  \item contact YouTube channels that have more than 1M subscribers
-        (and have mentioned YouTube policies at some point
-          (suggesting maybe passing interest in discussing meta topics)?)
-  \item Reach out by any contact methods they provide (if none, then skip)
-  \item Semi--structured interview guiding towards:
-  \begin{itemize}
-    \item interaction with platform \& its designers
-    \item engagement with audience
-    \item coordination with other YouTubers (if any)
-  \end{itemize}
-  \item tentative questions include:
-  \begin{itemize}
-    \item platform questions
-    \begin{enumerate}
-      \item when did you switch to (streaming/YouTubing) as primary career?
-      \item what's YouTube's relationship with you?
-      \item if you were telling a new YouTuber how to get started,
-            what would you say they should know about YouTube
-            (the system, the company, etc\dots)
-            that's not immediately obvious?
-    \end{enumerate}
-    \item audience questions
-    \begin{enumerate}
-      \item how do you interact with audiences?
-      \item what sorts of comments do you get?
-      \item do you reply?
-      \item do you change your future content based on feedback in the comments?
-    \end{enumerate}
-    \item peer questions
-    \begin{enumerate}
-      \item what's your sense of the community as a whole?
-      \item how many other YouTubers do you know
-            (let's say you've had a chat with them in the last month and
-             will probably do so again in the next month)?
-      \item are you all cooperative? competitive? what's the culture like?
-    \end{enumerate}
-  \end{itemize}
-\end{itemize}
-}
+\topic{We propose to carry out qualitative research with major online content producers
+on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.}
+Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
+open--ended exploration of the field site as we find parallels and divergences in the domain
+of online performance art from street and other emergent formats of performance and entertainment.
+% With that being said, we can generally frame our agenda in five components
+% ---
+% \begin{inlinelist}
+%   \item Surveying the landscape of online performance platforms,
+%         identifying performers who may be willing to serve as ``informants'', and
+%         recruiting interviewing participants;
+%   \item Investigating the relationships and tensions performers have with \textit{audiences};
+%   \item Exploring the relationships performers have with \textit{other performers};
+%   \item Studying the relationships performers have with \textit{the platforms themselves}; and finally
+%   \item Designing interventions to facilitate creative experimentation in this nascent format,
+%         informed largely by each of the earlier facets of online performance.
+% \end{inlinelist}
+A rough timeline of these phases can be described as such:
+
+\begin{tabular}{p{2cm}|p{13cm}}
+  \textbf{Term} & \textbf{Plan} \\
+\hline
+Q1  & Identify potential participants and begin recruiting participants;
+      begin preliminary interviews and fieldwork.
+      Research specifically focusing on performers' \textit{relationships with audiences}
+      and noting areas of frustration for performers and intervention opportunities \\
+\hline
+Q2  & Qualitative fieldwork toward understanding performers' \textit{relationships with other performers},
+      with some attention toward the potential for performer affiliations \\
+\hline
+Q3  & Fieldwork looking at performers' \textit{relationships with their respective platforms}
+      (for example: YouTube, Twitch) \\
+\hline
+Q4  & Exploring and iterating on \textit{design interventions} with YouTubers and streamers
+      to advance values discussed earlier
+      (audience engagement, online performer affiliation, etc\dots)  \\
+\end{tabular}
+
+\topic{Our interview process will }
+We will use semi--structured interviews
+to remain open to participants guiding our research agenda toward
+whatever they consider to be compelling topics.
+With that being said
+
+% \ali{Basic plan:
+% \begin{itemize}
+%   \item contact YouTube channels that have more than 1M subscribers
+%         (and have mentioned YouTube policies at some point
+%           (suggesting maybe passing interest in discussing meta topics)?)
+%   \item Reach out by any contact methods they provide (if none, then skip)
+%   \item Semi--structured interview guiding towards:
+%   \begin{itemize}
+%     \item interaction with platform \& its designers
+%     \item engagement with audience
+%     \item coordination with other YouTubers (if any)
+%   \end{itemize}
+%   \item tentative questions include:
+%   \begin{itemize}
+%     \item platform questions
+%     \begin{enumerate}
+%       \item when did you switch to (streaming/YouTubing) as primary career?
+%       \item what's YouTube's relationship with you?
+%       \item if you were telling a new YouTuber how to get started,
+%             what would you say they should know about YouTube
+%             (the system, the company, etc\dots)
+%             that's not immediately obvious?
+%     \end{enumerate}
+%     \item audience questions
+%     \begin{enumerate}
+%       \item how do you interact with audiences?
+%       \item what sorts of comments do you get?
+%       \item do you reply?
+%       \item do you change your future content based on feedback in the comments?
+%     \end{enumerate}
+%     \item peer questions
+%     \begin{enumerate}
+%       \item what's your sense of the community as a whole?
+%       \item how many other YouTubers do you know
+%             (let's say you've had a chat with them in the last month and
+%              will probably do so again in the next month)?
+%       \item are you all cooperative? competitive? what's the culture like?
+%     \end{enumerate}
+%   \end{itemize}
+% \end{itemize}
+% }
 
@@
 
-\pagebreak
+% \pagebreak
+% \renewcommand\bibname{{\small References}}
+\setlength\bibitemsep{0pt}
+
+\renewcommand*{\bibfont}{\tiny}
+
 \printbibliography{}

### 2017-03-17 proposals/Brown_Magic_Grant_2017/main.tex
abstract changes (hah, get it?)

@@
 % \chead{} % controls the center of the header
-\rhead{Alkhatib \& Nagy} % controls the right corner of the header
-\lfoot{last updated} % controls the left corner of the footer
+\rhead{Ali Alkhatib \& Jeff Nagy} % controls the right corner of the header
+\lfoot{} % controls the left corner of the footer
+\lfoot{last updated \DTMcurrenttime~\today} % controls the left corner of the footer
 \cfoot{} % controls the center of the footer
-\rfoot{Page~\thepage} % controls the right corner of the footer
+\rfoot{} % controls the right corner of the footer
+% \rfoot{Page~\thepage} % controls the right corner of the footer
 \renewcommand{\headrulewidth}{0.0pt}
@@
 \thispagestyle{empty}
-
-% \setlength{\parskip}{.2em}
 
@@
 \par{DRAFT \DTMcurrenttime~\today}%
-% \par{DRAFT \DTMcurrenttime~on~\today}%
 \par}}
@@
   \begin{abstract}
-  \ali{something along the lines of: online performers are becoming a big deal, and
-       these performers are possibly the forerunners of a new type of expression and art.
+  Online video sharing and streaming platforms have grown quickly in the past decade
+  to support an emerging community of unconventional entertainers.
+  This eclectic group of artists has collectively been
+  exploring boundaries and experimenting with styles and expectations,
+  critically influencing both our cultural expectations of performers and the norms of performance itself.
+  But these dramatic shifts have coincided with turmoil among this new class of entertainers and artists.
   The frustration and turmoil they're experiencing
-       --- both interpersonally and with the environment on which they ``work'' ---
-       demands to be better understood if we hope to make informed design decisions that empower these people.
-       We propose to conduct ethnographic fieldwork of this domain,
+  --- both with one another and with the very platform on which they ``work'' ---
+  calls for a better understanding of the culture and position of this nascent culture.
+  We propose to conduct ethnographic fieldwork in this domain,
   initially informed by a historical framing, but
   ultimately studying how digitally mediated publics such as these affect
-       the politics of this group of performers.}
+  the politics of this group of performers.
   \end{abstract}
@@
 % \section*{Insight}
-\topic{We recognize that \textit{historical analysis}
+\topic{\textit{Historical analysis}
 % Studying contemporary phenomena through an established framing
 can be a helpful way of
-making sense of contemporary phenomena.}
+making sense of contemporary phenomena and informing early fieldwork.}
 As we found in a recent paper framing crowd work and gig work as a modern instantiation of \textit{piecework},
 scholarship from a parallel or similar domain can
-provide us some grounding to make sense of what we've seen, as well as inform predictions~\cite{pieceworkCrowdworkGigwork}.
+provide us some grounding to make sense of what we've seen and offer a framework for making reasonable predictions~\cite{pieceworkCrowdworkGigwork}.
+We believe that a similar approach can be used here to help make sense of online streaming and video performances.
 

### 2017-03-17 proposals/Brown_Magic_Grant_2017/main.tex
spent a while working on the abstract

@@
 
-\AtBeginBibliography{\small}
+% \AtBeginBibliography{\small}
 \thispagestyle{empty}
@@
 
-\def\maketitle{%
-\par{\centering {\scshape \LARGE \textbf{\@title}}%
-\par{\@author}
-\par{DRAFT \DTMcurrenttime~\today}%
-\par}}
+% \def\maketitle{%
+% \par{{\scshape \LARGE \textbf{\@title}}%
+% \par{\@author}
+% \par{DRAFT \DTMcurrenttime~\today}%
+% \par}}
 % \smallskip \hrule \smallskip }
+\makeatletter
+\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
+\begin{flushleft}
+  {\scshape \Large \textbf{\@title}}
 
+  \@author
+\end{flushleft}\egroup
+}
+\makeatother
 \definecolor{PineGreen}{HTML}{008800}
@@
 \title{An Ethnography of Online Performance Artists}
-\author{Ali Alkhatib \& Jeffrey Nagy}
+\author{Ali Alkhatib \& Jeffrey Nagy
+--- DRAFT \DTMcurrenttime~\today}
 \date{\today}
@@
 \newcommand{\topic}[1]{{\color{Blue}#1}}
-\renewcommand{\topic}[1]{{#1}}
+% \renewcommand{\topic}[1]{{#1}}
 \definecolor{Gray}{HTML}{A0A0A0}
@@
 
+\renewenvironment{abstract}{%
+\hfill\begin{minipage}{0.95\textwidth}
+% \rule{\textwidth}{1pt}
+\itshape
+}
+{
+% \noindent\rule{\textwidth}{1pt}
+\bigskip
+\end{minipage}}
+
 \begin{document}
@@
   \begin{abstract}
-  Online video sharing and streaming platforms have grown quickly in the past decade
-  to support an emerging community of unconventional entertainers.
-  This eclectic group of artists has collectively been
-  exploring boundaries and experimenting with styles and expectations,
-  critically influencing both our cultural expectations of performers and the norms of performance itself.
-  But these dramatic shifts have coincided with turmoil among this new class of entertainers and artists.
-  The frustration and turmoil they're experiencing
-  --- both with one another and with the very platform on which they ``work'' ---
-  calls for a better understanding of the culture and position of this nascent culture.
+  Online video sharing and streaming platforms have flourished into a bustling industry in the past decade.
+  % to support an emerging community of unconventional entertainers.
+  The artists that have emerged on these platforms since then have
+  explored boundaries and experimented with styles and formats,
+  influencing both audiences' expectations of performers and of the performances themselves.
+  But these shifts have come with turmoil among this new group of artists,
+  calling for better understanding of the culture and position of this new, growing group.
   We propose to conduct ethnographic fieldwork in this domain,
@@
 % \section*{Situation}
-\topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
-and streaming sites % \ali{, like Twitch and others,}
-have garnered the attention of audiences around the world
+\topic{Video sharing sites like YouTube and Twitch
+have captured the attention of audiences of billions of people around the world
 in the span of little more than a decade.} % taken on characteristics of
-So widely used are these platforms
-that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
+% So widely used are these platforms
+% that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
 content producers
@@
 Q1  & Identify potential participants and begin recruiting participants;
-      begin preliminary interviews and fieldwork.
+      begin interviews and fieldwork.
       Research specifically focusing on performers' \textit{relationships with audiences}
@@
 
-\topic{Our interview process will }
-We will use semi--structured interviews
-to remain open to participants guiding our research agenda toward
-whatever they consider to be compelling topics.
-With that being said
+\topic{We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
+--- in other words, people who have gone ``full time''.}
+Our initial interviews will be semi--structured in nature, as we'll want to be open to the possibility that 
+% We will use semi--structured interviews
+% to remain open to participants guiding our research agenda toward
+% whatever they consider to be compelling topics.
 
@@
 % \renewcommand\bibname{{\small References}}
-\setlength\bibitemsep{0pt}
+% \setlength\bibitemsep{0pt}
 
-\renewcommand*{\bibfont}{\tiny}
-
+% \renewcommand*{\bibfont}{\tiny}
+\pagebreak
 \printbibliography{}

### 2017-03-18 proposals/Brown_Magic_Grant_2017/main.tex
one or two steps from a ready-to-submit draft

@@
 \usepackage{xr-hyper}
-\usepackage{todonotes,balance,graphics,color,alphalph,subfiles,appendix,endnotes,times}
-\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,nth,cleveref,textcase,comment,titlesec}
+\usepackage{balance,times}
+\usepackage{xcolor,microtype,enumitem,nth,,titlesec}
 \usepackage[normalem]{ulem}
-\usepackage{url}
-% \usepackage{natbib}
+\usepackage{url,graphicx}
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-% \usepackage[T1]{fontenc}
 \usepackage[pdflang={en-US},pdftex]{hyperref}
-\usepackage{mfirstuc,lastpage}
-\usepackage{fancyhdr,tabularx}
+\usepackage{wrapfig,fancyhdr}
 \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
@@
 
-\usepackage{caption}
-\captionsetup[table]{labelformat=empty}
-
-% \usepackage{mfirstuc,lastpage}
+% \usepackage{caption,setspace}
 \bibliography{references}
 \usepackage[margin=1in]{geometry}
-\pagestyle{fancy}
+% \pagestyle{fancy}
 \titlespacing*{\section}{0.0ex}{2ex plus .3ex}{1ex plus .3ex}
@@
 }
-\makeatletter
 
-% \def\maketitle{%
-% \par{{\scshape \LARGE \textbf{\@title}}%
-% \par{\@author}
-% \par{DRAFT \DTMcurrenttime~\today}%
-% \par}}
-% \smallskip \hrule \smallskip }
 \makeatletter
@@
 \begin{flushleft}
-  {\scshape \Large \textbf{\@title}}
+  {\scshape \LARGE \textbf{\@title}}
 
@@
 \makeatother
+
 \definecolor{PineGreen}{HTML}{008800}
@@
 
-\title{An Ethnography of Online Performance Artists}
+\title{Ethnography of Online Performance Artists}
 \author{Ali Alkhatib \& Jeffrey Nagy
@@
 \hfill\begin{minipage}{0.95\textwidth}
-% \rule{\textwidth}{1pt}
 \itshape
@@
 {
-% \noindent\rule{\textwidth}{1pt}
 \bigskip
@@
   \begin{abstract}
-  Online video sharing and streaming platforms have flourished into a bustling industry in the past decade.
-  % to support an emerging community of unconventional entertainers.
-  The artists that have emerged on these platforms since then have
+  Abstract: Online video sharing and streaming platforms have flourished over the past decade.
+  The artists on these platforms have
   explored boundaries and experimented with styles and formats,
@@
 
-% \section*{Situation}
-\topic{Video sharing sites like YouTube and Twitch
-have captured the attention of audiences of billions of people around the world
-in the span of little more than a decade.} % taken on characteristics of
-% So widely used are these platforms
-% that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
-content producers
---- ``YouTubers'' and ``streamers'' ---
+\begin{wrapfigure}{r}{0.5\textwidth}
+\centering
+\includegraphics[width=0.5\textwidth]{figures/cgpgrey.png}
+\caption{\label{fig:Magic}A popular YouTuber discussing election results and various methods of determining election winners.}
+\end{wrapfigure}
+Video sharing sites like YouTube and Twitch
+have captured the attention of billions of people around the world
+in the span of little more than a decade.
+Content producers
+--- known informally as ``YouTubers'' and ``streamers'' ---
 have begun to make their livings on these sites.
@@
 
-% \section*{Problem}
-\topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
-assumptions regarding the nature of the space they're using.}
-Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
+But recently, tensions between these online performers and the platforms on which they work have shaken people's
+assumptions regarding the nature of the space they're using.
+Performers on now--defunct Vine called (unsuccessfully) to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
 YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
@@
 
-\topic{One issue frustrating this potential research area is that
+One issue frustrating this potential research area is that
 the platforms have changed so dramatically in as few as 5 to 10 years that
-it can seem as though researchers are trying to wade through shifting sands.}
+it can seem as though researchers are trying to wade through shifting sands.
 Scholarship on content creators loses bearing rapidly when
@@
 % \section*{Insight}
-\topic{\textit{Historical analysis}
+\textit{Historical analysis}
 % Studying contemporary phenomena through an established framing
 can be a helpful way of
-making sense of contemporary phenomena and informing early fieldwork.}
+making sense of contemporary phenomena and informing early fieldwork.
 As we found in a recent paper framing crowd work and gig work as a modern instantiation of \textit{piecework},
@@
 
-\topic{To that end, we can liken some aspects of online performance art on YouTube and Twitch
+To that end, we can liken some aspects of online performance art on YouTube and Twitch
 to street performance and busking,
-allowing us to relate ostensibly new phenomena to robust scholarship.}
+allowing us to relate ostensibly new phenomena to robust scholarship.
 As is the case with street performance,
@@
 
-\topic{This connection has purchase, but it has gaps, mostly owing to the unique nature of the internet~\cite{miller2011understanding}.}
+This connection has purchase, but it has gaps, mostly owing to the unique nature of the internet~\cite{miller2011understanding}.
 For one thing, most video sharing websites keep videos
@@
 
-\topic{We propose to fill in these gaps in our framing of online performance as a modern instantiation of street performance.}
+We propose to fill in these gaps in our framing of online performance as a modern instantiation of street performance.
 How does the \textit{permanence} of digitally archived videos affect people's willingness to experiment with new formats and styles,
@@
 
-% \section*{TODO: Problem with this model}
-% \section*{TODO: Proposal}
-% \section*{Incomplete!}
-% \section*{Proposal}
-
-% \ali{generally speaking, each major bullet point = roughly 1 paragraph}
-% \begin{itemize}
-%   \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
-%         a growing industry for entertainers.
-%         Further, it has ``democratized'' entertainment, providing people with a public space
-%         (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
-%   \item Our research into crowdsourcing and gig work has yielded
-%         a compelling framework for these phenomena as an instantiation of piecework.
-%         We've found that this approach
-%         --- that is, drawing parallels with historical phenomena ---
-%         can generally give researchers of sociotechnical systems substantive framing of a research space.
-%   \item To some extent, this can be likened to \textbf{street performance} inasmuch as
-%   \begin{itemize}
-%     \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
-%     \item entertainers could get more direct feedback from their audience than other avenues could afford them
-%     \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
-%   \end{itemize}
-%   \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
-%         \ali{need a clearer \textbf{research question}?}
-%   \item When we look at this dimension of online performance art, the metaphor begins to break down:
-%   \begin{itemize}
-%     \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
-%     \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
-%     \item communication between the performer and the audience is carefully mediated
-%           (and literally moderated) by the platform itself
-%     \item online performances are (generally) more permanent than street performances
-%           \ali{technology has thrown a wrench into this lately, with
-%               cameras allowing us to record street performances and
-%               tools making it impossible (or at least difficult) to save online streams, but
-%               \textit{let's not worry about all that}.}
-%   \end{itemize}
-%   \item So while there are compelling parallels between online performances and the street art of busking,
-%         this has only taken us partway in answering important questions about a rapidly growing field.
-% \end{itemize}
-
 \section*{Research Plan}
-\topic{We propose to carry out qualitative research with major online content producers
-on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.}
+We propose to carry out qualitative research with major online content producers
+on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.
 Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
@@
 of online performance art from street and other emergent formats of performance and entertainment.
-% With that being said, we can generally frame our agenda in five components
-% ---
-% \begin{inlinelist}
-%   \item Surveying the landscape of online performance platforms,
-%         identifying performers who may be willing to serve as ``informants'', and
-%         recruiting interviewing participants;
-%   \item Investigating the relationships and tensions performers have with \textit{audiences};
-%   \item Exploring the relationships performers have with \textit{other performers};
-%   \item Studying the relationships performers have with \textit{the platforms themselves}; and finally
-%   \item Designing interventions to facilitate creative experimentation in this nascent format,
-%         informed largely by each of the earlier facets of online performance.
-% \end{inlinelist}
-A rough timeline of these phases can be described as such:
+A timeline of these phases is outlined here:
 
-\begin{tabular}{p{2cm}|p{13cm}}
-  \textbf{Term} & \textbf{Plan} \\
+\vspace*{5pt}
+
+\noindent\begin{tabular}{p{.5in}|p{5.5in}}
+  \textbf{Term} & \textbf{Primary Research Focus} \\
 \hline
-Q1  & Identify potential participants and begin recruiting participants;
-      begin interviews and fieldwork.
-      Research specifically focusing on performers' \textit{relationships with audiences}
-      and noting areas of frustration for performers and intervention opportunities \\
+Q1  & Identifying potential participants and recruiting participants;
+      preliminary interviewing and other surveying fieldwork. \\
 \hline
-Q2  & Qualitative fieldwork toward understanding performers' \textit{relationships with other performers},
-      with some attention toward the potential for performer affiliations \\
+Q2  & Performers' \textit{relationships with audiences},
+      noting areas of frustration for performers and intervention opportunities, and
+      investigating parallel or similar situations that affect street performers. \\
 \hline
-Q3  & Fieldwork looking at performers' \textit{relationships with their respective platforms}
-      (for example: YouTube, Twitch) \\
+Q3  & Performers' \textit{relationships with other performers},
+      with an eye toward themes such as affiliations or associations of performers representing performers,
+      as well as informal supporting relationships. \\
 \hline
-Q4  & Exploring and iterating on \textit{design interventions} with YouTubers and streamers
-      to advance values discussed earlier
-      (audience engagement, online performer affiliation, etc\dots)  \\
+Q4  & Performers' \textit{relationships with their respective platforms},
+      with particular attention to the politics of using an \textit{ostensibly} ``public'' resource,
+      and the circumstances that make salient that these platforms are in fact \textit{privately} owned and operated.
 \end{tabular}
 
-\topic{We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
---- in other words, people who have gone ``full time''.}
-Our initial interviews will be semi--structured in nature, as we'll want to be open to the possibility that 
-% We will use semi--structured interviews
-% to remain open to participants guiding our research agenda toward
-% whatever they consider to be compelling topics.
+\vspace*{5pt}
 
-% \ali{Basic plan:
-% \begin{itemize}
-%   \item contact YouTube channels that have more than 1M subscribers
-%         (and have mentioned YouTube policies at some point
-%           (suggesting maybe passing interest in discussing meta topics)?)
-%   \item Reach out by any contact methods they provide (if none, then skip)
-%   \item Semi--structured interview guiding towards:
-%   \begin{itemize}
-%     \item interaction with platform \& its designers
-%     \item engagement with audience
-%     \item coordination with other YouTubers (if any)
-%   \end{itemize}
-%   \item tentative questions include:
-%   \begin{itemize}
-%     \item platform questions
-%     \begin{enumerate}
-%       \item when did you switch to (streaming/YouTubing) as primary career?
-%       \item what's YouTube's relationship with you?
-%       \item if you were telling a new YouTuber how to get started,
-%             what would you say they should know about YouTube
-%             (the system, the company, etc\dots)
-%             that's not immediately obvious?
-%     \end{enumerate}
-%     \item audience questions
-%     \begin{enumerate}
-%       \item how do you interact with audiences?
-%       \item what sorts of comments do you get?
-%       \item do you reply?
-%       \item do you change your future content based on feedback in the comments?
-%     \end{enumerate}
-%     \item peer questions
-%     \begin{enumerate}
-%       \item what's your sense of the community as a whole?
-%       \item how many other YouTubers do you know
-%             (let's say you've had a chat with them in the last month and
-%              will probably do so again in the next month)?
-%       \item are you all cooperative? competitive? what's the culture like?
-%     \end{enumerate}
-%   \end{itemize}
-% \end{itemize}
-% }
+We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
+--- in other words, people who have gone ``full time''.
+We'll aim for approximately two--thirds of our participants to be principally active on YouTube and one--third on Twitch, as
+the former is significantly larger (and the community seemingly more diverse), suggesting
+a need to get more varied perspectives from participants.
 
-% \section*{Broader Impact}
-% \topic{Formally recognized art forms have assimilated emergent genres and types of performance art in the past, and
-% we should expect that they will continue to do so.}
-% Jazz was born out of what largely ``\dots was recognized~\dots~as being but one species of the genus, ragtime played by ear by fakers\dots''~\cite{10.2307/779456}.
-% Even further back in history, \nth{17} century street performance in Japan blended ``\dots every imaginable genus and species of popular song and dance\dots'' and
-% kabuki, in its early years,
-% ``would hardly have been recognized or condoned [by traditional performers]''~\cite{groemer2016street}.
+Our interviews will be semi--structured as
+we want to be open to the potential for our participants to take us in different directions than originally planned;
+nevertheless, we have a bank of interview questions that will
+guide the interview along the three major themes we identified earlier:
+\begin{inlinelist}
+  \item questions about the participant's \textit{relationship with their audiences}
+        (e.g. ``do you have personal relationships with other [streamers or YouTubers]?'',
+              ``do you think of other [streamers or YouTubers] as `the competition', or potential collaborators, or as something else (or a mix)?'');
+  \item questions about the participant's \textit{relationship with other performers}
+        (e.g. ``what sorts of comments and feedback do you get?'',
+              ``do you respond to comments? change your format or style?''); and
+  \item questions about the participant's perceived \textit{relationship with the platform}
+        (e.g. ``when did you go `full time'?'',
+              ``what sorts of things get you in trouble on [your platform]?'').
+\end{inlinelist}
 
-% \ali{something like: if this continues, then we should infer that
-%      the future of this type of performance art will bear heavily on
-%      the direction and trajectory of digitally mediated entertainment and performance}
+% Each stage of research will focus on a specific thematic question for two reasons:
+% first, to incrementally inform our subsequent, potentially more contentious, research questions; and
+% second, to provide us with focused topics which we can scrutinize with the aforementioned historical lens of street performance.
 
-% \section*{Demonstrate viability}
-
-% \section*{Expected outcomes}
-
-% \begin{itemize}
-%   \item We propose to direct our research energy toward answering one question: 
-% \end{itemize}
-
-% We propose to conduct ethnographic research of online performance artists who principally earn their income either from Twitch or YouTube (or both).
-
-% \topic{While our research considered a number of possible futures for gig work in general and gig workers in particular,
-% several important questions remain unresolved.}
-% For example, we pointed to the general question that researchers had been asking for some time ---
-% whether gig work will become a site of complex, rewarding work
-% --- or if it will remain a setting for low quality tasks with little upward mobility.
-% While we offered a framework for answering questions such as this one, the reality is that
-% any researcher or team exploring this area
-% --- especially on the basis of the parallel we drew ---
-% would fundamentally need to be conversant with several deep bodies of research spanning myriad fields.
-
-% We concluded that gig work's potential will be determined by
-% how easily skilled workers can secure good work through these platforms.
-% In order to do that, workers need to be able to communicate rich information about their qualifications to employers,
-% which existing markets have so far struggled to do.
-% \topic{We will build a system that tracks gig workers' professional histories across myriad work platforms,
-% aggregating, consolidating, \& analyzing that data, and
-% allowing workers to curate their professional identities, instantiated by a ``digital hiring hall'' for gig workers.}
-
-% \topic{The system we intend to build will serve two roles --- one for workers and one for employers.}
-% Gig workers will share their work histories from platforms such as Amazon Mechanical Turk with our repository,
-% which will allow us to provide the worker with information upon which they can reflect
-% (e.g. detailed analysis of their work trends, approval rates under various circumstances, etc\dots)
-% as well as information in relation to other workers
-% (e.g. offering a percentile ranking overall for approval rates, or within specific types of tasks).
-% In doing so, we hope to find that workers will identify cues and gravitate toward qualities in work
-% in which they identify competitive advantage.
-
-% \topic{The second role that this repository will help employers identify and source the best workers for highly skilled work.}
-% By aggregating the data from work platforms and
-% offering more in--depth analysis of workers' histories, we hope to find that
-% we can direct employers to gig workers whose overall experiences match the employer's given needs.
-% For example, if an employer or a platform queries our platform for
-% workers who are proficient at translating between two specific languages,
-% we can yield a list of workers who would be suitable for the task, determined by
-% the workers' comprehensive histories on Mechanical Turk and, potentially, elsewhere.
-
-% \topic{This approach is similar in some senses to the role that hiring halls played through the \nth{20} century.}
-% By disconnecting workers' data from the platforms themselves, our repository will act as a sort of ``digital hiring hall'',
-% connecting employers with workers whose varied work histories provide them with the expertise necessary to do given tasks.
-% By engaging workers in the management of this repository, we will advance the metaphor in much the same way that
-% labor unions maintained --- and continue to operate --- hiring halls of their workers.
-
-% \topic{At a high level, we can study the usefulness of this system using standard experimental approaches.}
-% First, through quantitative methods (for example, comparing workers' earnings over time); and
-% second, through qualitative  methods (for example, through interviews to determine whether workers find and exploit specialty niches as a result of this tool).
-% We can also explore employers perspectives in tangible ways;
-% for example, given concerns about the quality of work~\cite{Ipeirotis:2010:QMA:1837885.1837906,le2010ensuring,Law:2017:CTR:2998181.2998197},
-% we will investigate whether a system that better communicates worker expertise allows employers to rely more confidently on crowdwork.
-
-% \section*{Interdisciplinarity}
-% % \topic{This research will call on methods and skills originally from and generally found in disparate fields.}
-% % \topic{The insights and expertise necessary to conduct this research successfully will call on a variety of skills not typically seen together, such as
-% % participant observation, participatory design techniques, and machine learning approaches.}
-% As we showed with our paper examining the relationship between piecework and gig work~\cite{pieceworkCrowdworkGigwork},
-% the lens of a social scientist can inform conversations about digitally mediated work in ways that many computer scientists are not otherwise able to offer.
-% At the same time, the perspectives of computer scientists
-% --- and the challenges digitally mediated spaces bring with them ---
-% offer to reinvigorate deeply theoretically grounded discussion and offer field sites with which to experiment and test these theories.
-% As a Computer Science Ph.D. student originally trained as an Anthropologist, I believe
-% I'm the best candidate to combine these bodies of literature to make this project succeed.
-
-% \section*{Anticipated Contributions \& Implications}
-% This research is of significant importance given the recent trends in the ``gig economy''.
-% With more and more people joining the workforce in this capacity~\cite{pewSharing,pewSharing24percent},
-% concerns about ``the future of work''~\cite{waterCooler,crowdworkFuture} continue to grow.
-% Our hope is that the artifact produced by this research raises the limit on the complexity of work that can reliably be crowdsourced.
-% More broadly, we expect to offer two major contributions to the Human--Computer Interaction community:
-% first, an existence proof that a gig work market for complex, specialized workers is possible; and
-% second, insights into how best to scaffold and foster ongoing collective governance.
-
-% This project's will speak to the critique that the internet,
-% far from the democratizing force that Barlow predicted it would be,
-% has become an infrastructure facilitating the corporate centralization of power~\cite{barlow2009declaration,jones2011does,EboCybertopia}.
-% The success of this project will therefore not only
-% bring a more nuanced understanding of the shifting climate and culture of labor,
-% but may
-% --- in a small way ---
-% deliver on the promise Barlow made that
-% cyberspace and the Internet would be a tool for democratizing access to information, and with that,
-% empowering more than the tech elite who engineered it.
-
-% \pagebreak
-% \renewcommand\bibname{{\small References}}
-% \setlength\bibitemsep{0pt}
-
-% \renewcommand*{\bibfont}{\tiny}
 \pagebreak

### 2017-03-18 proposals/Brown_Magic_Grant_2017/main.tex
maybe almost done (except for a conclusion para

@@
 \usepackage{url,graphicx}
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
+\usepackage[citestyle=authoryear-icomp,backend=bibtex]{biblatex}
 \usepackage[pdflang={en-US},pdftex]{hyperref}
@@
   \begin{abstract}
-  Abstract: Online video sharing and streaming platforms have flourished over the past decade.
+  Online video sharing and streaming platforms have flourished over the past decade.
   The artists on these platforms have
@@
 practice their craft, and 
-hone their skills~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}.
+hone their skills~(\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}).
 
@@
 assumptions regarding the nature of the space they're using.
-Performers on now--defunct Vine called (unsuccessfully) to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
-YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
-and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
+Performers on now--defunct Vine called (unsuccessfully) to be better compensated~(\cite{vineWantsMoney,vineInsiderMeeting});
+YouTubers meanwhile struggle to defend their unconventional performance work~(\cite{h3h3Lawsuit})
+and individually and collectively field accusations of misconduct of various forms~(\cite{youtubeDramaResponses}).
 These events highlight the politics of an emerging form of expression and art,
@@
 Scholarship on content creators loses bearing rapidly when
-economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
+economic factors motivate aspiring professional entertainers to join the community~(\cite{Hamilton:2014:STF:2611105.2557048});
 the tension between these entertainers experimenting with content remixing and
 the holders of the copyright of the source material
-(for example, see~\cite{Hilderbrand48})
+(for example, see~(\cite{Hilderbrand48}))
 shifts dramatically when
 the platform's copyright enforcement mechanisms advance
-not just incrementally but substantially~\cite{kim2012institutionalization}.
+not just incrementally but substantially~(\cite{kim2012institutionalization}).
 And the popularity of candidate field sites can vary chaotically
 (see, for example, Vine reaching as many as 200 million active users and
-abrupt shutdown less than a year later~\cite{vineDecline}).
+abrupt shutdown less than a year later~(\cite{vineDecline})).
 
@@
 scholarship from a parallel or similar domain can
-provide us some grounding to make sense of what we've seen and offer a framework for making reasonable predictions~\cite{pieceworkCrowdworkGigwork}.
+provide us some grounding to make sense of what we've seen and offer a framework for making reasonable predictions~(\cite{pieceworkCrowdworkGigwork}).
 We believe that a similar approach can be used here to help make sense of online streaming and video performances.
@@
 
-This connection has purchase, but it has gaps, mostly owing to the unique nature of the internet~\cite{miller2011understanding}.
+This connection has purchase, but it also has gaps, mostly owing to the unique nature of the internet~(\cite{miller2011understanding}).
 For one thing, most video sharing websites keep videos
@@
 The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
-YouTube comments linger for years after they're made, forming in some sense ``networked publics''~\cite{boyd2007youth}.
+YouTube comments linger for years after they're made, forming in some sense ``networked publics''~(\cite{boyd2007youth}).
 And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
-the near--perfect ability of these platforms to redirect revenue and even people's focus instantaneously.
+the near--perfect ability of these platforms to redirect revenue and even people's focus in an instant.
 
@@
 We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
---- in other words, people who have gone ``full time''.
+--- in other words, people who have gone ``full--time''.
 We'll aim for approximately two--thirds of our participants to be principally active on YouTube and one--third on Twitch, as
@@
   \item questions about the participant's perceived \textit{relationship with the platform}
-        (e.g. ``when did you go `full time'?'',
+        (e.g. ``when did you go `full--time'?'',
               ``what sorts of things get you in trouble on [your platform]?'').

### 2017-03-18 proposals/Brown_Magic_Grant_2017/main.tex
fixes based on jeff's notes

@@
 \author{Ali Alkhatib \& Jeffrey Nagy
---- DRAFT \DTMcurrenttime~\today}
+}
 \date{\today}
@@
 the technical infrastructure these platforms provide as well as
-the access to audiences that these platforms provide as well.
+the access to broad audiences and the promise of exposure.
 In doing so, these sites have in a real sense democratized media and entertainment industries by
 making the avenues to large audiences substantially more accessible.
-And thus, in these transformations into publics, what started as
+And thus, in these transformations into Deweyan--style publics~(\cite{dewey2012public,disalvo2009design}), what started as
 a recreational hobby
@@
 
-But recently, tensions between these online performers and the platforms on which they work have shaken people's
-assumptions regarding the nature of the space they're using.
+But recently, tensions between these online performers and the platforms on which they work have shaken
+assumptions performers held regarding the nature of the space they're using,
+revealing breakdowns in the expectations of these systems~(\cite{doi:10.1177/1461444809342738}).
 Performers on now--defunct Vine called (unsuccessfully) to be better compensated~(\cite{vineWantsMoney,vineInsiderMeeting});
@@
 
-One issue frustrating this potential research area is that
-the platforms have changed so dramatically in as few as 5 to 10 years that
-it can seem as though researchers are trying to wade through shifting sands.
+% One issue frustrating this potential research area is that
+One issue research in this area must overcome is that
+platforms have changed so dramatically in as few as 5 to 10 years that
+it can seem as though researchers are trying to navigate through what seem at times like shifting sands.
 Scholarship on content creators loses bearing rapidly when
-economic factors motivate aspiring professional entertainers to join the community~(\cite{Hamilton:2014:STF:2611105.2557048});
+economic factors motivate performers to turn a recreation or a hobby
+--- such as video game streaming~(\cite{Hamilton:2014:STF:2611105.2557048}) --- into a career;
 the tension between these entertainers experimenting with content remixing and
@@
 online performance isn't restricted to classically trained professionals;
-instead, performers (both online and in the streets) practice more experimental 
-drawing on their candid interaction with their audiences,
+instead, performers (both online and in the streets) practice more experimentally,
+drawing on candid interactions with their audiences,
 which can change from one performance to the next.
 
-This connection has purchase, but it also has gaps, mostly owing to the unique nature of the internet~(\cite{miller2011understanding}).
+This connection has purchase, but it also has gaps,
+mostly owing to the unique nature of the internet~(\cite{miller2011understanding}).
 For one thing, most video sharing websites keep videos
@@
 --- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
-The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
+The public that congregates around a performer, too, has a similar permanence that offline busking doesn't;
 YouTube comments linger for years after they're made, forming in some sense ``networked publics''~(\cite{boyd2007youth}).
-And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
-the near--perfect ability of these platforms to redirect revenue and even people's focus in an instant.
+We must also account for the different nature of algorithmic enforcement of rules and policies, and
+the near--perfect ability of these platforms to redirect revenue and
+even an individual audience member's focus in an instant.
 
@@
 We propose to carry out qualitative research with major online content producers
-on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.
+on video \textit{streaming} sites such as Twitch and Periscope, and video \textit{sharing} sites like YouTube.
 Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
@@
 
-We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
+We will recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
 --- in other words, people who have gone ``full--time''.
-We'll aim for approximately two--thirds of our participants to be principally active on YouTube and one--third on Twitch, as
+We will attempt to recruit approximately two--thirds of our participants from YouTube and one--third from Twitch, as
 the former is significantly larger (and the community seemingly more diverse), suggesting
@@
 
-% Each stage of research will focus on a specific thematic question for two reasons:
-% first, to incrementally inform our subsequent, potentially more contentious, research questions; and
-% second, to provide us with focused topics which we can scrutinize with the aforementioned historical lens of street performance.
-
 \pagebreak

### 2017-03-26 chi/piecework/presentation/presentation.tex
cleaning up diffs and some roadmapping for chi presentation

% !TEX program = lualatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{beamer}

\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor}
\usepackage[citestyle=author,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}
\bibliography{references}
\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\renewcommand\textbullet{\ensuremath{\bullet}}
\definecolor{stanfordRed}{HTML}{8C1515}
% \mode<presentation>{
  % \usetheme{metropolis}
  \usecolortheme{seagull}

\definecolor{Orange}{HTML}{E98125}
\definecolor{Blue}{HTML}{0645AD}
\hypersetup{colorlinks,linkcolor=,urlcolor=Blue}

% Theme colors are derived from these two elements
\setbeamercolor{alerted text}{fg=Orange}
\setsansfont[BoldFont={Source Sans Pro Semibold},
              Numbers={OldStyle}]{Source Sans Pro}
\setmonofont{Source Code Pro}
% \setbeamercolor{frametitle}{bg=stanfordRed}

  % \setbeamercolor*{palette tertiary}{use=structure,fg=stanfordRed,bg=stanfordRed}
% }
% \metroset{set}
\setbeamercolor{institute in head/foot}{fg=stanfordRed}
% \setbeamercolor{progress bar}{fg=stanfordRed}
% \setbeamercolor{progress bar}{bg=gr}
% \setsansfont{Source Sans Pro}

\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
\metroset{titleformat=smallcaps}

\author{\textbf{Ali Alkhatib},
                Michael Bernstein,
                Margaret Levi\\
\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}} }}

\institute[Stanford]{Stanford University}
\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

% PAPER ROADMAP
% Open problems in crowdsourcing: complexity, decomposition, relationships
% Piecework!
% Wait what's piecework again
% Some cool case studies: 
%     airy's computers
%     domestic/farmhand labor
%     matchstick girls
%     industrial workers
% Okay let's get into this
% Complexity limits
%     what did crowdsourcing say?
%     what did piecework say?
%     comparisons?
%     implications?
% Decomposition limits
%     what did crowdsourcing say?
%     what did piecework say?
%     comparisons?
%     implications?
% Worker relationships
%     what did crowdsourcing say?
%     what did piecework say?
%     comparisons?
%     implications?
% Discussion
%     this is a very tricky thing to do, predicting the future
%     anyway let's predict the future: utopian/dystopian visions
%     Research agendas! go go gadget contextual analysis

\section{Introduction}

\begin{frame}{Open problems in crowdsourcing}
\begin{itemize}[<+- | alert@+>]
  \item Complexity~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
  \item Decomposition
  \item Relationships
\end{itemize}
\end{frame}

\begin{frame}{Emerging field of on--demand work}
\begin{itemize}[<+- | alert@+>]
  \item what's that about??
\end{itemize}
\end{frame}

\begin{frame}{Contact}

    % \faTwitter
    name: {Ali Alkhatib} \\
    email: \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} \\
    twitter: \href{https://twitter.com/_alialkhatib}{@\_alialkhatib} \\
\end{frame}

% \bibliographystyle{SIGCHI-Reference-Format}
\printbibliography{}
\end{document}

### 2017-03-27 chi/piecework/presentation/presentation.tex
fleshed out sections and frames

@@
 \usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor}
-\usepackage[citestyle=author,backend=bibtex]{biblatex}
+\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
 \usepackage[export]{adjustbox}
@@
 \hypersetup{colorlinks,linkcolor=,urlcolor=Blue}
+
+\definecolor{PineGreen}{HTML}{008800}
+\definecolor{Red}{HTML}{FF0000}
+\definecolor{Blue}{HTML}{0000FF}
+\definecolor{JokeGreen}{HTML}{00C953}
+\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
+\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
 
@@
 
-% PAPER ROADMAP
-% Open problems in crowdsourcing: complexity, decomposition, relationships
-% Piecework!
-% Wait what's piecework again
-% Some cool case studies: 
-%     airy's computers
-%     domestic/farmhand labor
-%     matchstick girls
-%     industrial workers
-% Okay let's get into this
-% Complexity limits
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Decomposition limits
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Worker relationships
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Discussion
-%     this is a very tricky thing to do, predicting the future
-%     anyway let's predict the future: utopian/dystopian visions
-%     Research agendas! go go gadget contextual analysis
+% ## Intro [`1 minute`] [`1`]
+% - Open problems in crowdsourcing: complexity, decomposition, relationships [`?`]
+% - Piecework! [`?`]
+% ## Wait what's piecework again [`1 minute`] [`2`]
+% ## Some case studies: [`3 minutes`] [`5`]
+% - airy's computers [`1`]
+% - domestic/farmhand labor [`0.5`]
+% - matchstick girls [`1`]
+% - industrial workers [`0.5`]
+% ## Complexity limits [`3 minutes`] [`8`]
+% - what did crowdsourcing say? [`0.5`]
+% - what did piecework say? [`1`]
+% - comparisons? [`1`]
+% - implications? [`0.5`]
+% ## Decomposition limits [`3 minutes`] [`11`]
+% - what did crowdsourcing say? [`0.5`]
+% - what did piecework say? [`1`]
+% - comparisons? [`1`]
+% - implications? [`0.5`]
+% ## Worker relationships [`3 minutes`] [`14`]
+% - what did crowdsourcing say? [`0.5`]
+% - what did piecework say? [`1`]
+% - comparisons? [`1`]
+% - implications? [`0.5`]
+% ## Discussion [`1 minute`] [`15`]
+% - this is a very tricky thing to do, predicting the future [`?`]
+% - anyway let's predict the future: utopian/dystopian visions [`?`]
+% - Research agendas! go go gadget contextual analysis [`?`]
 
@@
 
-\begin{frame}{Open problems in crowdsourcing}
+\begin{frame}{Open problems in crowdsourcing
+                                                                                [\texttt{1 minute}]
+}
 \begin{itemize}[<+- | alert@+>]
-  \item Complexity~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
+  \item Complexity~\cite{suzukiAtelier,KimStoria,yuanAlmost,
+                         YuEncouragingOutside,
+                         Nebeling:2016:WCW:2858036.2858169,
+                         Hahn:2016:KAB:2858036.2858364}
   \item Decomposition
@@
 
-\begin{frame}{Emerging field of on--demand work}
+\section{Piecework Primer}
+\begin{frame}{Piecework Review
+                                                                                [\texttt{1 minute}]
+}
+something
+\end{frame}
+
+\section{Case Studies}
+\begin{frame}{Airy's Computers
+                                                                                [\texttt{1 minute}]
+    }
+\begin{itemize}
+  \item images of hand--drawn tables
+  \item maybe some illustration of the mathematical work that was being done
+  \item draw to attention that you could break this into pieces
+        (an example of breaking a formula down?)
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Domestic \& Farmhand Labor
+                                                                                [\texttt{0.5 minutes}]
+}
+\end{frame}
+
+\begin{frame}{The Match Girls
+                                                                                [\texttt{1 minute}]
+}
+\end{frame}
+
+\begin{frame}{Industrial Workers
+                                                                                [\texttt{0.5 minutes}]
+}
+\end{frame}
+
+\section{Research Threads}
+\begin{frame}{Complexity Limits
+                                                                                [\texttt{3 minutes}]
+}
 \begin{itemize}[<+- | alert@+>]
-  \item what's that about??
+  \item Crowdwork's perspective
+  \item Piecework's perspective
+  \item Comparisons
+  \item Implications
 \end{itemize}
+\end{frame}
+
+\begin{frame}{Decomposition Limits
+                                                                                [\texttt{3 minutes}]
+}
+\begin{itemize}[<+- | alert@+>]
+  \item Crowdwork's perspective
+  \item Piecework's perspective
+  \item Comparisons
+  \item Implications
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Worker Relationships
+                                                                                [\texttt{3 minutes}]
+}
+\begin{itemize}[<+- | alert@+>]
+  \item Crowdwork's perspective
+  \item Piecework's perspective
+  \item Comparisons
+  \item Implications
+\end{itemize}
+\end{frame}
+
+\section{Discussion}
+\begin{frame}{Discussion
+                                                                                [\texttt{1 minute}]
+}
 \end{frame}
@@
 
-    % \faTwitter
     name: {Ali Alkhatib} \\

### 2017-03-28 drafts/task_referral/introduction.tex
removed reaveal.js stuff

@@
 On--demand workers have expressed frustration with the algorithmic management that largely dictates their work
-\cite{uberAlgorithm}.
+\cite{uberAlgorithm}. \msb{blah blah blah}
 The fear of deviating from expected results causes online information workers 

### 2017-03-31 chi/piecework/complexity.tex
keynote checkin

@@
 \subsubsection{\crowdworkpers}
-\topic{Kittur et al. were among the first to ask
-whether crowdsourcing could be used for more than parallelizing tasks~\cite{crowdForgeKittur}.}
+\topic{Kittur et al. were among the first to ask whether crowdsourcing could be used for more than parallelizing tasks~\cite{crowdForgeKittur}.}
 Their work showed that it could, with proof--of--concept crowdsourcing of
@@
 medical data interpreters~\cite{lasecki2013chorus}, and
-idea generation~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.
+idea generation~\cite{Yu2016a,Yu2016b}.
 
@@
 
-\topic{It is now clear that this computational workflow approach works with some classes of complex tasks, but
-the broader wicked problems largely remain unsolved.}
+\topic{It is now clear that this computational workflow approach works with some classes of complex tasks, but the broader wicked problems largely remain unsolved.}
 As a first example,
-idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
+idea generation shows promise~\cite{Yu2016a,Yu2016b},
 but there is as yet no general crowdsourced solution for
@@
 
-\topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
-Airy's human computers and among industrial workers.}
+Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier: Airy's human computers and among industrial workers.
 
@@
 \topic{Moreover, piecework was limited to tasks that could be quickly and accurately evaluated.}
-
 Hart argues that evaluation limited piecework's complexity:
@@
 
-\topic{The research on piecework tells us that
-we should expect it to thrive in industries where
-the nature of the work is limited in complexity~\cite{Brown01041990}, and
-become less common as work becomes more complex.}
+\topic{The research on piecework tells us that we should expect it to thrive in industries where the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
 Has computation shifted piecework's previous limits of

### 2017-04-27 chi/piecework/presentation/presentation.tex
gave up and went with keynote
notes and rough roadmapping is in the document
latex chi presentation draft (in progress)

--- 2017-03-26 chi/piecework/presentation/presentation.tex
@@
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor}
-\usepackage[citestyle=author,backend=bibtex]{biblatex}
+\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,enumitem}
+\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
 \usepackage[export]{adjustbox}
@@
 
+\hypersetup{
+  colorlinks = true,
+  % linkcolor = blue,
+  citecolor = blue
+}
+
+\def\labelitemi{--}
+
+% \let\oldcite=\textcite
+% \renewcommand{\textcite}[1]{\textcolor[rgb]{.2,.2,.9}{[\oldcite{#1}]}}
+
 \begin{document}
@@
 \end{frame}
+% \renewcommand\multicitedelim{\addsemicolon\space}
 
-% PAPER ROADMAP
-% Open problems in crowdsourcing: complexity, decomposition, relationships
-% Piecework!
-% Wait what's piecework again
-% Some cool case studies: 
-%     airy's computers
-%     domestic/farmhand labor
-%     matchstick girls
-%     industrial workers
-% Okay let's get into this
-% Complexity limits
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Decomposition limits
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Worker relationships
-%     what did crowdsourcing say?
-%     what did piecework say?
-%     comparisons?
-%     implications?
-% Discussion
-%     this is a very tricky thing to do, predicting the future
-%     anyway let's predict the future: utopian/dystopian visions
-%     Research agendas! go go gadget contextual analysis
+% \setbeamercovered{transparent}
 
-\section{Introduction}
+\section*{Introduction}
 
 \begin{frame}{Open problems in crowdsourcing}
-\begin{itemize}[<+- | alert@+>]
-  \item Complexity~\cite{suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
-  \item Decomposition
-  \item Relationships
+  \begin{itemize}%[<+- | alert@+>]
+    \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
+                           Nebeling:2016:WCW:2858036.2858169,
+                           Hahn:2016:KAB:2858036.2858364}
+    \item \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
+                              Law:2016:CKC:2858036.2858144,
+                              Chang:2016:ACC:2858036.2858411,
+                              Newell:2016:OMA:2858036.2858490}
+    \item \textbf{Relationships}~\textcite{turkopticon,storiesIraniSilberman,crowdcollab,
+                              takingAHITMcInnis}
   \end{itemize}
@@
 
-\begin{frame}{Emerging field of on--demand work}
-\begin{itemize}[<+- | alert@+>]
-  \item what's that about??
+\begin{frame}[standout]{whataatatat}
+    What is the future of work?
+
+\end{frame}
+
+\begin{frame}{Introduction}
+  We hope to provide:
+      \begin{itemize}
+        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
+        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
       \end{itemize}
@@
 
+\begin{frame}{A case for comparative historical analysis}
+Historical analysis is nothing new
+    \begin{itemize}
+      \item \textcite{Wyche2006,bodker1993historical}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{A brief glossary}
+    \begin{itemize}
+      \item Crowd work: digitally mediated \textbf{information work}
+      --- for example, work done on Amazon Mechanical Turk~\cite{crowdworkFuture}
+      \item Gig work: digitally mediated --- but often \textbf{physically embodied} --- one--off jobs,
+      such as
+      \textit{driving},
+      \textit{courier services},
+      and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Complexity}
+  What kinds of problems do we mean when we talk about complexity?
+  \begin{itemize}
+    \item Can crowds improve existing works?~\cite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986}
+    \item Can crowds critique designs?~\cite{yuanAlmost}
+    \item Can crowds create things from whole cloth?~\cite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{What does the crowdsourcing literature say?}
+    something pretty insightful I bet!
+\end{frame}
+
+\begin{frame}{What does the piecework literature say?}
+    something even more insightful, I'm sure!
+\end{frame}
+
+\begin{frame}[t]\frametitle{Let's compare this stuff}
+\framesubtitle{something else}
+  \begin{columns}
+  \begin{column}{0.5\textwidth}
+  % \section*{Piecework}
+     some text here some text here some text here some text here some text here
+
+  \end{column}
+
+  \begin{column}{0.5\textwidth}
+  % \section*{CrowdWork}
+     some text here some text here some text here some text here some text here
+
+  \end{column}
+  \end{columns}
+\end{frame}
+
+% \section{Piecework Primer}
+% \begin{frame}{Piecework Review
+%                                                                                 [\texttt{1 minute}]
+% }
+% % What is piecework?
+% % \begin{figure}
+
+% % \end{figure}
+% \end{frame}
+
 \begin{frame}{Contact}
 
-    % \faTwitter
     name: {Ali Alkhatib} \\

### 2017-05-01 chi/piecework/presentation/presentation.tex
weekend updates

@@
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,enumitem}
+\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg}
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
@@
 
+\newenvironment{mystepwiseitemize}{\begin{itemize}[<+-| alert@+>]}{\end{itemize}}
 % Theme colors are derived from these two elements
@@
 \setbeamercolor{institute in head/foot}{fg=stanfordRed}
+\setbeamercovered{transparent}
+\setbeamercovered{again covered={\opaqueness<1->{15}}}
 % \setbeamercolor{progress bar}{fg=stanfordRed}
@@
 \date{\today}
+\setbeamertemplate{itemize items}{--}
 
@@
 
-\def\labelitemi{--}
+% \def\labelitemi{--}
 
@@
 
-\begin{frame}{Open problems in crowdsourcing}
-  \begin{itemize}%[<+- | alert@+>]
-    \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
+% \begin{frame}[t]{Open problems in crowdsourcing}
+%   \begin{itemize}[<+>]
+%     \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
+%                            Nebeling:2016:WCW:2858036.2858169,
+%                            Hahn:2016:KAB:2858036.2858364}
+%   \end{itemize}
+% \end{frame}
+
+% \begin{frame}[t]{Open problems in crowdsourcing}
+%   \begin{itemize}[<+>]
+%     \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
+%                            Nebeling:2016:WCW:2858036.2858169,
+%                            Hahn:2016:KAB:2858036.2858364}
+%     \item \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
+%                               Law:2016:CKC:2858036.2858144,
+%                               Chang:2016:ACC:2858036.2858411,
+%                               Newell:2016:OMA:2858036.2858490}
+%   \end{itemize}
+% \end{frame}
+
+\begin{frame}[t]{Open problems in crowdsourcing}
+\begin{columns}
+\begin{column}{0.6\textwidth}
+  \begin{itemize}%[<+>]
+    \item<1,4> \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
                            Nebeling:2016:WCW:2858036.2858169,
                            Hahn:2016:KAB:2858036.2858364}
-    \item \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
+    \item<2> \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
                               Law:2016:CKC:2858036.2858144,
@@
                               Newell:2016:OMA:2858036.2858490}
-    \item \textbf{Relationships}~\textcite{turkopticon,storiesIraniSilberman,crowdcollab,
+    \item<3> \textbf{Relationships}~\textcite{turkopticon,storiesIraniSilberman,crowdcollab,
                               takingAHITMcInnis}
+
   \end{itemize}
+
+\end{column}
+\begin{column}{0.4\textwidth}
+test
+\end{column}
+\end{columns}
+%   \begin{figure}[htbp]
+%   \centering
+
+% \end{figure}
+
+  % \begin{figure}
+  % \includesvg{test}
+  % \caption{svg image}
+  % \includegraphics<1>[width=.5\textwidth]{figures/handy.png}
+  % \includegraphics<2>[width=.5\textwidth]{figures/amt.png}
+  % \includegraphics<3>[width=.5\textwidth]{figures/uber.png}
+  % \end{figure}
 \end{frame}
@@
 \begin{frame}{What does the crowdsourcing literature say?}
-    something pretty insightful I bet!
+\begin{columns}[t]
+
+  \begin{column}[t]{0.5\textwidth}
+    \begin{itemize}
+      \item Build complexity into the process
+      \begin{itemize}
+        \item Apply CS methods to people~(\textcite{crowdForgeKittur})
+      \end{itemize}
+    \end{itemize}
+  \end{column}
+
+  \begin{column}{0.5\textwidth}
+    \begin{figure}
+    \includegraphics[width=\textwidth]{figures/mapReduce.png}
+    \end{figure}
+  \end{column}
+
+\end{columns}
 \end{frame}
@@
 
-\begin{frame}[t]\frametitle{Let's compare this stuff}
-\framesubtitle{something else}
-  \begin{columns}
-  \begin{column}{0.5\textwidth}
-  % \section*{Piecework}
-     some text here some text here some text here some text here some text here
+\begin{frame}{Comparerereer}
+% \framesubtitle{something else}
+%   \begin{columns}
+%   \begin{column}{0.5\textwidth}
+%   % \section*{Piecework}
+%      some text here some text here some text here some text here some text here
 
-  \end{column}
+%   \end{column}
 
-  \begin{column}{0.5\textwidth}
-  % \section*{CrowdWork}
-     some text here some text here some text here some text here some text here
+%   \begin{column}{0.5\textwidth}
+%   % \section*{CrowdWork}
+%      some text here some text here some text here some text here some text here
 
-  \end{column}
-  \end{columns}
+%   \end{column}
+%   \end{columns}
 \end{frame}

### 2017-05-02 chi/piecework/presentation/presentation.tex
svgs
monday updates; finished intro?

@@
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg}
+\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles}
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
@@
 
+\newcommand{\onlyinsubfile}[1]{#1}
 \begin{document}
-
+\renewcommand{\onlyinsubfile}[1]{}
 \begin{frame}
@@
 
-\section*{Introduction}
+% \subfile{intro.tex}
 
-% \begin{frame}[t]{Open problems in crowdsourcing}
-%   \begin{itemize}[<+>]
-%     \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
-%                            Nebeling:2016:WCW:2858036.2858169,
-%                            Hahn:2016:KAB:2858036.2858364}
-%   \end{itemize}
-% \end{frame}
-
-% \begin{frame}[t]{Open problems in crowdsourcing}
-%   \begin{itemize}[<+>]
-%     \item \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
-%                            Nebeling:2016:WCW:2858036.2858169,
-%                            Hahn:2016:KAB:2858036.2858364}
-%     \item \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
-%                               Law:2016:CKC:2858036.2858144,
-%                               Chang:2016:ACC:2858036.2858411,
-%                               Newell:2016:OMA:2858036.2858490}
-%   \end{itemize}
-% \end{frame}
-
-\begin{frame}[t]{Open problems in crowdsourcing}
-\begin{columns}
-\begin{column}{0.6\textwidth}
-  \begin{itemize}%[<+>]
-    \item<1,4> \textbf{Complexity}~\textcite{suzukiAtelier,KimStoria,yuanAlmost,
-                           Nebeling:2016:WCW:2858036.2858169,
-                           Hahn:2016:KAB:2858036.2858364}
-    \item<2> \textbf{Decomposition}~\textcite{sensitiveTasks,LykourentzouPersonalityMatters,
-                              Law:2016:CKC:2858036.2858144,
-                              Chang:2016:ACC:2858036.2858411,
-                              Newell:2016:OMA:2858036.2858490}
-    \item<3> \textbf{Relationships}~\textcite{turkopticon,storiesIraniSilberman,crowdcollab,
-                              takingAHITMcInnis}
-
-  \end{itemize}
-
-\end{column}
-\begin{column}{0.4\textwidth}
-test
-\end{column}
-\end{columns}
-%   \begin{figure}[htbp]
-%   \centering
-
-% \end{figure}
-
-  % \begin{figure}
-  % \includesvg{test}
-  % \caption{svg image}
-  % \includegraphics<1>[width=.5\textwidth]{figures/handy.png}
-  % \includegraphics<2>[width=.5\textwidth]{figures/amt.png}
-  % \includegraphics<3>[width=.5\textwidth]{figures/uber.png}
-  % \end{figure}
-\end{frame}
-
-\begin{frame}[standout]{whataatatat}
+\begin{frame}[standout]
     What is the future of work?

### 2017-05-03 chi/piecework/presentation/complexity.tex
adding to complexity

--- 2017-05-02 chi/piecework/presentation/presentation.tex
@@
-% !TEX program = lualatex
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beamer Presentation
-% LaTeX Template
-% Version 1.0 (10/11/12)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[presentation]{subfiles}
 
-\documentclass{beamer}
+\begin{document}
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles}
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-\usepackage[export]{adjustbox}
-\usepackage[en-US]{datetime2}
-\bibliography{references}
-\usetheme{metropolis}
-\beamertemplatenavigationsymbolsempty
-\renewcommand\textbullet{\ensuremath{\bullet}}
-\definecolor{stanfordRed}{HTML}{8C1515}
-% \mode<presentation>{
-  % \usetheme{metropolis}
-  \usecolortheme{seagull}
-
-\definecolor{Orange}{HTML}{E98125}
-\definecolor{Blue}{HTML}{0645AD}
-\hypersetup{colorlinks,linkcolor=,urlcolor=Blue}
-
-\definecolor{PineGreen}{HTML}{008800}
-\definecolor{Red}{HTML}{FF0000}
-\definecolor{Blue}{HTML}{0000FF}
-\definecolor{JokeGreen}{HTML}{00C953}
-\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
-\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
-
-\newenvironment{mystepwiseitemize}{\begin{itemize}[<+-| alert@+>]}{\end{itemize}}
-% Theme colors are derived from these two elements
-\setbeamercolor{alerted text}{fg=Orange}
-\setsansfont[BoldFont={Source Sans Pro Semibold},
-              Numbers={OldStyle}]{Source Sans Pro}
-\setmonofont{Source Code Pro}
-% \setbeamercolor{frametitle}{bg=stanfordRed}
-
-  % \setbeamercolor*{palette tertiary}{use=structure,fg=stanfordRed,bg=stanfordRed}
-% }
-% \metroset{set}
-\setbeamercolor{institute in head/foot}{fg=stanfordRed}
-\setbeamercovered{transparent}
-\setbeamercovered{again covered={\opaqueness<1->{15}}}
-% \setbeamercolor{progress bar}{fg=stanfordRed}
-% \setbeamercolor{progress bar}{bg=gr}
-% \setsansfont{Source Sans Pro}
-
-\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
-\metroset{titleformat=smallcaps}
-
-\author{\textbf{Ali Alkhatib},
-                Michael Bernstein,
-                Margaret Levi\\
-\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
-         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}} }}
-
-\institute[Stanford]{Stanford University}
-\date{\today}
-\setbeamertemplate{itemize items}{--}
-
-\hypersetup{
-  colorlinks = true,
-  % linkcolor = blue,
-  citecolor = blue
-}
-
-% \def\labelitemi{--}
-
-% \let\oldcite=\textcite
-% \renewcommand{\textcite}[1]{\textcolor[rgb]{.2,.2,.9}{[\oldcite{#1}]}}
-
-\newcommand{\onlyinsubfile}[1]{#1}
-\begin{document}
-\renewcommand{\onlyinsubfile}[1]{}
-\begin{frame}
-\titlepage
-\end{frame}
-% \renewcommand\multicitedelim{\addsemicolon\space}
-
-% \setbeamercovered{transparent}
-
-% \subfile{intro.tex}
-
-\begin{frame}[standout]
-    What is the future of work?
-
-\end{frame}
-
-\begin{frame}{Introduction}
-  We hope to provide:
-      \begin{itemize}
-        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
-        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
-      \end{itemize}
-\end{frame}
-
-\begin{frame}{A case for comparative historical analysis}
-Historical analysis is nothing new
-    \begin{itemize}
-      \item \textcite{Wyche2006,bodker1993historical}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{A brief glossary}
-    \begin{itemize}
-      \item Crowd work: digitally mediated \textbf{information work}
-      --- for example, work done on Amazon Mechanical Turk~\cite{crowdworkFuture}
-      \item Gig work: digitally mediated --- but often \textbf{physically embodied} --- one--off jobs,
-      such as
-      \textit{driving},
-      \textit{courier services},
-      and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
-    \end{itemize}
-\end{frame}
+\section{Complexity}
 
@@
   What kinds of problems do we mean when we talk about complexity?
-  \begin{itemize}
+  \begin{mystepwiseitemize}
     \item Can crowds improve existing works?~\cite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986}
@@
     \item Can crowds create things from whole cloth?~\cite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
-  \end{itemize}
+  \end{mystepwiseitemize}
 \end{frame}
@@
 \begin{frame}{What does the crowdsourcing literature say?}
-\begin{columns}[t]
-
-  \begin{column}[t]{0.5\textwidth}
+\begin{columns}
+  \begin{column}{0.5\textwidth}
     \begin{itemize}
       \item Build complexity into the process
-      \begin{itemize}
+      \begin{mystepwiseitemize}
         \item Apply CS methods to people~(\textcite{crowdForgeKittur})
-      \end{itemize}
+        \item see other stuff
+      \end{mystepwiseitemize}
     \end{itemize}
@@
     \begin{figure}
-    \includegraphics[width=\textwidth]{figures/mapReduce.png}
+    % \visible<2>{
+    \includegraphics[width=0.7\textwidth]{figures/mapReduce.png}
+    % }
+    % \visible<1>{
+
+    \includegraphics[width=0.7\textwidth]{figures/complexity/cw_literature/scribe.png}
+    % }
     \end{figure}
@@
 
-\begin{frame}{Comparerereer}
+\begin{frame}{Comparisons}
 % \framesubtitle{something else}
@@
 \end{frame}
-
-% \section{Piecework Primer}
-% \begin{frame}{Piecework Review
-%                                                                                 [\texttt{1 minute}]
-% }
-% % What is piecework?
-% % \begin{figure}
-
-% % \end{figure}
-% \end{frame}
-
-\begin{frame}{Contact}
-
-    name: {Ali Alkhatib} \\
-    email: \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} \\
-    twitter: \href{https://twitter.com/_alialkhatib}{@\_alialkhatib} \\
-\end{frame}
-
-% \bibliographystyle{SIGCHI-Reference-Format}
-\printbibliography{}
 \end{document}

### 2017-05-03 chi/piecework/presentation/presentation.tex
fixed some intro stuff
adding to complexity

@@
 \end{frame}
-% \renewcommand\multicitedelim{\addsemicolon\space}
-
 % \setbeamercovered{transparent}
 
-% \subfile{intro.tex}
+\subfile{intro.tex}
 
@@
     What is the future of work?
+\end{frame}
+
+\begin{frame}{What is the future of work?}
 
@@
 \begin{frame}{A case for comparative historical analysis}
-Historical analysis is nothing new
 \begin{itemize}
-      \item \textcite{Wyche2006,bodker1993historical}
+  \item Historical analysis isn't new
+  \begin{itemize}
+    \item In general (see~\textcite{rosenberg1994exploring,rosenberg1982inside})
+    \item In HCI (see~\textcite{Wyche2006,bodker1993historical})
+  \end{itemize}
+  % \item Still can be useful.
 \end{itemize}
@@
 
-\begin{frame}{Complexity}
-  What kinds of problems do we mean when we talk about complexity?
-  \begin{itemize}
-    \item Can crowds improve existing works?~\cite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986}
-    \item Can crowds critique designs?~\cite{yuanAlmost}
-    \item Can crowds create things from whole cloth?~\cite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{What does the crowdsourcing literature say?}
-\begin{columns}[t]
-
-  \begin{column}[t]{0.5\textwidth}
-    \begin{itemize}
-      \item Build complexity into the process
-      \begin{itemize}
-        \item Apply CS methods to people~(\textcite{crowdForgeKittur})
-      \end{itemize}
-    \end{itemize}
-  \end{column}
-
-  \begin{column}{0.5\textwidth}
-    \begin{figure}
-    \includegraphics[width=\textwidth]{figures/mapReduce.png}
-    \end{figure}
-  \end{column}
-
-\end{columns}
-\end{frame}
-
-\begin{frame}{What does the piecework literature say?}
-    something even more insightful, I'm sure!
-\end{frame}
-
-\begin{frame}{Comparerereer}
-% \framesubtitle{something else}
-%   \begin{columns}
-%   \begin{column}{0.5\textwidth}
-%   % \section*{Piecework}
-%      some text here some text here some text here some text here some text here
-
-%   \end{column}
-
-%   \begin{column}{0.5\textwidth}
-%   % \section*{CrowdWork}
-%      some text here some text here some text here some text here some text here
-
-%   \end{column}
-%   \end{columns}
-\end{frame}
-
-% \section{Piecework Primer}
-% \begin{frame}{Piecework Review
-%                                                                                 [\texttt{1 minute}]
-% }
-% % What is piecework?
-% % \begin{figure}
-
-% % \end{figure}
-% \end{frame}
-
-\begin{frame}{Contact}
-
-    name: {Ali Alkhatib} \\
-    email: \href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} \\
-    twitter: \href{https://twitter.com/_alialkhatib}{@\_alialkhatib} \\
-\end{frame}
+\subfile{complexity}
+% \subfile{decomposition}
+% \subfile{relationships}
+% \subfile{contact}
 

### 2017-05-03 chi/piecework/presentation/complexity.tex
complexity lit review done?

@@
 \documentclass[presentation]{subfiles}
+
+\onlyinsubfile{
+\bibliography{references}
+  % \beamertemplategridbackground[1mm]
+}
 
@@
   What kinds of problems do we mean when we talk about complexity?
-  \begin{mystepwiseitemize}
-    \item Can crowds improve existing works?~\cite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986}
-    \item Can crowds critique designs?~\cite{yuanAlmost}
-    \item Can crowds create things from whole cloth?~\cite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
-  \end{mystepwiseitemize}
+  \begin{itemize}
+    \item<1-> Can crowds help you write something?
+
+    \scriptsize{
+      \textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}
+    }
+    \item<2-> Can crowds critique novel designs?
+
+    \scriptsize{
+      \textcite{yuanAlmost,fuge2014analysis}
+    }
+    \item<3-> Can crowds create things from whole cloth?
+
+    \scriptsize{
+      \textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
+    }
+  \end{itemize}
 \end{frame}
@@
 \begin{columns}
-  \begin{column}{0.5\textwidth}
+  \begin{column}{0.6\textwidth}
     \begin{itemize}
@@
       \begin{mystepwiseitemize}
-        \item Apply CS methods to people~(\textcite{crowdForgeKittur})
-        \item see other stuff
+        \item Apply CS methods to people
+
+        \scriptsize{
+          \textcite{crowdForgeKittur}
+        }
+        \item Humans as computational units
+
+        \scriptsize{
+          \textcite{Lasecki:2014:LSR:2661334.2661352}
+        }
+        \item Crowdsourcing workflows as function state machines
+
+        \scriptsize{\textcite{latoza2014microtask}
+        }
       \end{mystepwiseitemize}
@@
 
-  \begin{column}{0.5\textwidth}
+  \begin{column}{0.4\textwidth}
     \begin{figure}
-    % \visible<2>{
-    \includegraphics[width=0.7\textwidth]{figures/mapReduce.png}
-    % }
-    % \visible<1>{
+    \only<1>{
+    \includegraphics[width=\textwidth]{figures/complexity/cw_literature/mapReduce.png}
+    }
 
-    \includegraphics[width=0.7\textwidth]{figures/complexity/cw_literature/scribe.png}
-    % }
+    \only<2>{
+      \includegraphics[width=\textwidth]{figures/complexity/cw_literature/scribe.png}
+    }
+    \only<3>{
+      \begin{adjustbox}{max totalsize={\textwidth}{\textheight},center}
+      \begin{tikzpicture}[->,>=stealth',auto, node distance=4cm,
+                          thick,transform shape]
+        \node[state,label=above:{Described}]                       (A)                    {};
+        \node[state,label=left:{Described written}]                (B) [below of=A]       {};
+        \node[state,label=left:{Run Tests}]                        (D) [below of=B]       {};
+        \node[state,label=below:{Described written buggy}]         (C) [below right of=D] {};
+        \node[state,label=below:{Described written buggy}]         (E) [below left of=D]  {};
+
+        \path[every node/.style={sloped,anchor=south}]
+              (A) edge              node {Write description} (B)
+              (B) edge [loop right] node {Edit code} (B)
+                  edge              node {Edit code} (D)
+              (D) edge              node {} (C)
+                  edge              node {} (E)
+              (C) edge [bend right] node {Debug} (B)
+                  edge [bend right] node {Debug} (D)
+              (E) edge [bend right] node {Edit Code} (D)
+                  edge [bend left]  node {Edit Code} (B);
+        \end{tikzpicture}
+        \end{adjustbox}
+    }
     \end{figure}
@@
 \begin{frame}{What does the piecework literature say?}
-    something even more insightful, I'm sure!
+    % \begin{itemize}
+      % \item
+      George Airy (astronomer) used a very similar approach~\cite{grier2013computers}
+      % \begin{itemize}
+      %   \item Break work up so that individual tasks are simple
+
+      %   Make the process \textit{itself} compile the final product.
+      % \end{itemize}
+    % \end{itemize}
+    % \vspace{0mm}
+    \begin{columns}
+    \begin{column}{0.6\textwidth}
+      \begin{figure}
+        \includegraphics[width=\textwidth]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.4\textwidth}
+      \begin{itemize}
+        \item Employed computers
+        \item 13--20 years old
+        \item Overworked
+        \item Underpaid
+        \item Could be fired at will
+      \end{itemize}
+    \end{column}
+    \end{columns}
+
+\end{frame}
+
+\begin{frame}{George Airy --- whiz kid}
+    Airy built complexity into the process, assigning \textit{human computers} 
+    to compute, verify, and correct the right ascension and declination of stars.
+
+    \begin{figure}
+    \includegraphics[width=\textwidth]{figures/complexity/pw_literature/airy.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Humble origins}
+  First implementations of piecework:
+  \visible<2->{farms}\visible<3->{, textiles}\visible<4->{, and other low--skill labor.}
+  \begin{columns}
+
+  \visible<2->{
+  \begin{column}{0.3\textwidth}
+  \begin{figure}
+  \includegraphics[width=\textwidth]{figures/idk.png}
+  \end{figure}
+  \end{column}
+  }
+
+  \visible<3->{
+  \begin{column}{0.3\textwidth}
+  \begin{figure}
+  \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
+  \end{figure}
+  \end{column}
+  }
+
+  \visible<4->{
+  \begin{column}{0.3\textwidth}
+  \begin{figure}
+  \includegraphics[width=\textwidth]{figures/photo/Matchgirl_strikers.PNG}
+  \end{figure}
+  \end{column}
+  }
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Ford}
+    Fordism
+
 \end{frame}
@@
 \end{frame}
+
+\onlyinsubfile{
+  \printbibliography{}
+}
 \end{document}

### 2017-05-03 chi/piecework/presentation/introduction.tex
complexity lit review done?

\documentclass[presentation]{subfiles}

\onlyinsubfile{
  % \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
  % \bibliography{~/Library/texmf/bibtex/bib/local/references}
  % \beamertemplategridbackground[1mm]
}

\begin{document}
% \section*{Introduction}

\begin{frame}{Open problems in crowdsourcing}
\begin{columns}
\begin{column}[T]{0.5\textwidth}
  \begin{itemize}%[<+>]
    \item<2,4> \textbf{\large Complexity}

    \scriptsize{
    \textcite{suzukiAtelier,KimStoria,yuanAlmost,Yu2016b,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
    }
    \vspace{8mm}

    \item<3> \textbf{\large Decomposition}

    \scriptsize{
    \textcite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
    }
    \vspace{8mm}

    \item<5> \textbf{\large Relationships}

    \scriptsize{
      \textcite{turkopticon,storiesIraniSilberman,crowdcollab,takingAHITMcInnis,dynamo,uberAlgorithm}
    }

  \end{itemize}

\end{column}

\begin{column}[T]{0.25\textwidth}
  \begin{figure}
    \centering
    \vspace{-5mm}

    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}
    }\visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
                  \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
    }

    \vspace{5mm}
    \includegraphics[width=\textwidth]{figures/complexity/hexRepeat.png}
    \vspace{5mm}

    \visible<3->{
    \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
    }

    \visible<5->{
      \vspace{10mm}
      \includegraphics[width=\textwidth]{figures/idk.png}
    }
  \end{figure}

\end{column}

\begin{column}[T]{0.25\textwidth}
\centering
\vspace{5mm}

\visible<2->{
  \small{Complexity}
  \vspace{10mm}
}

\small{Tasks}

\visible<3->{
  \vspace{10mm}
  \small{Decomposition}
}

\end{column}
\end{columns}
\end{frame}

\begin{frame}[standout]
    What is the future of work?
\end{frame}

\begin{frame}{What is the future of work?}
    How will \textbf{technology} affect the \textbf{complexity} of the work that on--demand workers do?

    What are the \textbf{limits} of complexity in on--demand work?

    How can we \textbf{reach} those limits?
\end{frame}

\notinsubfile{
  \subfile{timeline.tex} % this takes a while to compile...
}

\begin{frame}{Introduction}
  We hope to provide:
      \begin{itemize}
        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
      \end{itemize}
\end{frame}

\end{document}

### 2017-05-03 chi/piecework/presentation/presentation.tex
complexity lit review done?

@@
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles}
+\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
+\usetikzlibrary{arrows,automata}
 \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
@@
   \usecolortheme{seagull}
+
+\usepackage{pgfcalendar}
+\usepackage{datenumber,xparse,fp}
+\usetikzlibrary{arrows.meta,backgrounds,fixedpointarithmetic}
+\makeatletter
+\ExplSyntaxOn
+\tl_new:N \l_chronos_dateformat_tl
+\tl_set:Nn \l_chronos_dateformat_tl { Y/m/d }
+% YY yn lle YYYY
+\cs_new_protected_nopar:Npn \chronos_year_shorten:n #1
+{
+  \chronos_year_shorten_aux:w #1 \q_stop % expl3 manuaal, 46
+}
+\cs_new_protected_nopar:Npn \chronos_year_shorten_aux:w #1 #2 #3 #4 \q_stop
+{
+  #3 #4
+}
+\cs_generate_variant:Nn \chronos_year_shorten:n { V , c }
+% dangos dyddiadau
+\cs_new_protected_nopar:Npn \chronos_show_date:n #1
+{
+  \tl_map_inline:Nn \l_chronos_dateformat_tl
+  {
+    \str_case:nnF { ##1 }
+    {
+      { a } { \pgfcalendarweekdayshortname{\thechronos@weekday} }
+      { A } { \pgfcalendarweekdayname{\thechronos@weekday} }
+      { b } { \pgfcalendarmonthshortname{\csname chronos@#1month\endcsname} }
+      { B } { \pgfcalendarmonthname{\csname chronos@#1month\endcsname} }
+      { d } { \csname chronos@#1day\endcsname }
+      { m } { \csname chronos@#1month\endcsname }
+      { y } { \chronos_year_shorten:c { chronos@#1year } }
+      { Y } { \csname chronos@#1year\endcsname }
+      { @ } { ~ }
+    }
+    {
+      ##1
+    }
+  }
+}
+% user interface
+\NewDocumentCommand \chronossetdateformat { m }
+{
+  \tl_set:Nn \l_chronos_dateformat_tl { #1 }
+  \tl_replace_all:Nnn \l_chronos_dateformat_tl { ~ } { @ }
+}
+% for pgf/tikz convenience
+\NewDocumentCommand \chronos@showdate { m }
+{
+  \pgfcalendarjuliantoweekday{\csname thechronos@#1date\endcsname}{\c@chronos@weekday}%
+  \chronos_show_date:n { #1 }
+}
+\ExplSyntaxOff
+\newcounter{chronos@date}
+\newcounter{chronos@startdate}
+\newcounter{chronos@enddate}
+\newcounter{chronos@startyear}
+\newcounter{chronos@endyear}
+\newcounter{chronos@yeardate}
+\newcounter{chronos@thingdate}
+\newcounter{chronos@otherthingdate}
+\newcounter{chronos@weekday}
+\newif\ifchronos@marks
+\chronos@marksfalse
+\newif\ifchronos@timeline@showyears
+\chronos@timeline@showyearstrue
+\newif\ifchronos@eventyearsonline
+\chronos@eventyearsonlinefalse
+\pgfkeys{/pgf/number format,
+  int detect,
+  set thousands separator={},
+}
+\tikzset{%
+  /handlers/.chronos too/.code={%
+    \edef\chronos@tempc{\pgfkeyscurrentpath}%
+    \edef\chronos@tempd{#1}%
+    \tikzset{%
+      \pgfkeyscurrentpath @too/.code={%
+        \tikzset{%
+          /chronos/\chronos@tempd/.append style={##1},
+        }%
+      },
+      \chronos@tempc/.forward to=\chronos@tempc @too,
+    }%
+  },
+  chronos/.code={% https://tex.stackexchange.com/a/159856/ - Claudio Fiandrino
+    \tikzset{%
+      align=center,
+      anchor=mid,
+      fixed point arithmetic,
+      /chronos/.cd,
+      #1,
+    }%
+    \pgfmathsetmacro\chronos@unit{(\chronos@width-20pt)/(\thechronos@enddate-\thechronos@startdate)}%
+    \draw [/chronos/timeline@line, line width=\chronos@height] (-10pt,0) coordinate (chronos pre) -- +(\chronos@width,0) coordinate (chronos post);
+    \coordinate (chronos base) at (0,-.5*\chronos@height);
+    \coordinate (chronos top) at (0,.5*\chronos@height);
+    \coordinate (chronos foot) at (0,{-.5*\chronos@height-\chronos@borderheight});
+    \coordinate (chronos head) at (0,{.5*\chronos@height+\chronos@borderheight});
+    \coordinate (chronos start) at (0,0);
+    \coordinate (chronos end) at ([xshift=-10pt]chronos post);
+    \ifdim\chronos@borderheight>0pt
+      \fill [bottom color=chronos@borderinner, top color=chronos@borderouter] (chronos pre |- chronos head) rectangle (chronos post |- chronos top);
+      \fill [top color=chronos@borderinner, bottom color=chronos@borderouter] (chronos pre |- chronos base) rectangle (chronos post |- chronos foot);
+    \fi
+    \ifchronos@timeline@showyears
+      \pgfmathsetcounter{chronos@startyear}{\chronos@startyear}%
+      \pgfmathsetcounter{chronos@endyear}{\chronos@endyear}%
+      \def\tempa{none}%
+      \ifx\chronos@stepfrom\tempa
+        \def\tempa{01}%
+        \ifx\chronos@startmonth\tempa
+          \ifx\chronos@startday\tempa
+            \else\stepcounter{chronos@startyear}%
+          \fi
+          \else\stepcounter{chronos@startyear}%
+        \fi
+      \else
+        \pgfmathsetcounter{chronos@startyear}{\chronos@stepfrom}%
+      \fi
+      \def\tempa{12}%
+      \def\tempb{31}%
+      \ifx\chronos@endmonth\tempa
+        \ifx\chronos@endday\tempb
+          \stepcounter{chronos@endyear}%
+        \fi
+      \fi
+      \pgfmathsetmacro\chronos@nextstep{int(\thechronos@startyear+\chronos@stepyears)}%
+      \foreach \i in {\thechronos@startyear,\chronos@nextstep,...,\thechronos@endyear} {%
+        \pgfcalendardatetojulian{{\i}-01-01}{\c@chronos@yeardate}%
+        \pgfmathsetmacro\chronos@tempa{(\thechronos@yeardate-\thechronos@startdate)*\chronos@unit}%
+        \def\tempa{on line}%
+        \ifx\tempa\chronos@timelineyears
+          \node (chronos@year@\i) [/chronos/.cd, timeline@years, timeline year on line] at (\chronos@tempa pt,0) {\i};
+          \ifchronos@marks
+             \path [/chronos/timeline mark on line] (chronos@year@\i.south) -- (chronos@year@\i |- chronos base);
+             \path [/chronos/timeline mark on line] (chronos@year@\i.north) --  (chronos@year@\i |- chronos top);
+          \fi
+        \else
+          \node (chronos@year@\i) [/chronos/.cd, timeline@years, timeline year off line] at (\chronos@tempa pt,0) {\i};
+          \ifchronos@marks
+            \path [shorten <=.5*\chronos@height, /chronos/timeline mark off line] (\chronos@tempa pt, 0) -- (chronos@year@\i);
+          \fi
+        \fi
+      }%
+    \fi
+    \ifchronos@eventyearsonline
+    \tikzset{%
+      /chronos/timeline years=on line,
+    }%
+    \fi
+  },
+  chronos set date/.code args={#1:#2:#3:#4}{%
+    \pgfcalendardatetojulian{{#1}-#2-#3}{\c@chronos@date}%
+    \setcounter{chronos@#4date}{\thechronos@date}%
+    \expandafter\def\csname chronos@#4year\endcsname{#1}%
+    \expandafter\def\csname chronos@#4month\endcsname{#2}%
+    \expandafter\def\csname chronos@#4day\endcsname{#3}%
+  },
+  chronos date/.style args={#1-#2-#3}{%
+    /tikz/chronos set date/.expanded={#1:#2:#3:thing}%
+  },
+  chronos period date/.style args={#1-#2-#3}{%
+    /tikz/chronos set date/.expanded={#1:#2:#3:otherthing}%
+  },
+  /chronos/.search also={/tikz},
+  /chronos/.cd,
+  step years/.store in=\chronos@stepyears,
+  step from year/.store in=\chronos@stepfrom,
+  start date/.style args={#1-#2-#3}{%
+    /tikz/chronos set date/.expanded={#1:#2:#3:start},
+  },
+  end date/.style args={#1-#2-#3}{%
+    /tikz/chronos set date/.expanded={#1:#2:#3:end},
+  },
+  timeline width/.store in=\chronos@width,
+  timeline height/.store in=\chronos@height,
+  width/.forward to=/chronos/timeline width,
+  height/.forward to=/chronos/timeline height,
+  timeline border height/.store in=\chronos@borderheight,
+  timeline border inner colour/.code={\colorlet{chronos@borderinner}{#1}},
+  timeline border outer colour/.code={\colorlet{chronos@borderouter}{#1}},
+  timeline years set/.store in=\chronos@timelineyears,
+  timeline years/.is choice,
+  timeline years/.forward to=/chronos/timeline years set,
+  timeline years/above/.code={%
+    \tikzset{%
+      /chronos/timeline@years/.style={above, anchor=south, yshift=.5*\chronos@height},
+    }%
+  },
+  timeline years/below/.code={%
+    \tikzset{%
+      /chronos/timeline@years/.style={below, anchor=north, yshift=-.5*\chronos@height},
+    }%
+  },
+  timeline years/on line/.code={%
+    \tikzset{%
+      /chronos/timeline@years/.style={anchor=center},
+    }%
+  },
+  only years/.code={%
+    \chronossetdateformat{Y}%
+    \tikzset{%
+      /tikz/chronos date/.style={%
+        /tikz/chronos set date/.expanded={##1:01:01:thing}%
+      },
+      /tikz/chronos period date/.style={%
+        /tikz/chronos set date/.expanded={##1:01:01:otherthing}%
+      },
+    }%
+  },
+  only text/.code={%
+    \tikzset{/chronos/only years}%
+    \chronossetdateformat{}%
+  },
+  background/.code={%
+    \colorlet{chronos@background}{#1}%
+  },
+  timeline marks/.is if=chronos@marks,
+  timeline show years/.is if=chronos@timeline@showyears,
+  timeline no years/.code={%
+    \tikzset{%
+      /chronos/timeline show years=false,
+    }%
+  },
+  lines/.style={draw, {Triangle[width=0pt 3,reversed,length=0pt 1.5]}-{Triangle[width=0pt 3,reversed,length=0pt 1.5]}, shorten <={.5*\chronos@height}},
+  events/.style={fill=chronos@background, fill opacity=.95, text opacity=1, draw opacity=1, rounded corners, align=center, font=\footnotesize},
+  period/.style={draw},
+  period event line/.style={/chronos/lines},
+  period event/.style={/chronos/events},
+  event line/.style={/chronos/lines},
+  event/.style={/chronos/events},
+  event years on line/.is if=chronos@eventyearsonline,
+  event year on line/.style={/chronos/timeline@years, /chronos/timeline year on line},
+  special date/.store in=\chronos@specialdate,
+  timeline@line/.style={draw},
+  timeline line/.chronos too=timeline@line,
+  timeline year off line/.style={fill=chronos@background, text opacity=1, fill opacity=.95},
+  timeline mark off line/.style={draw, {Triangle[width=0pt 3,reversed,length=0pt 1.5]}-, thin, shorten >=-2pt},
+  timeline year on line/.style={text=chronos@background, inner sep=1pt},
+  timeline mark on line/.style={draw=chronos@background, shorten >=1.5pt},
+  timeline mark too/.code={%
+    \tikzset{%
+      /chronos/.cd,
+      timeline mark on line/.append style={#1},
+      timeline mark off line/.append style={#1},
+    }%
+  },
+  timeline year too/.code={%
+    \tikzset{%
+      /chronos/.cd,
+      timeline year on line/.append style={#1},
+      timeline year off line/.append style={#1},
+    }%
+  },
+  timeline mark/.forward to=/chronos/timeline mark too,
+  timeline year/.forward to=/chronos/timeline year too,
+  start date=1001-10-01,
+  end date=1003-06-14,
+  timeline width=100mm,
+  timeline height=1pt,
+  timeline border height=0pt,
+  chronos date=1850-01-01,
+  chronos period date=1851-01-01,
+  step years=1,
+  timeline years=above,
+  background=white,
+  timeline border inner colour=black,
+  timeline border outer colour=chronos@background,
+  step from year=none,
+  special date=none,
+}
+\NewDocumentCommand \chronosevent { O {} m O {} +m D () { -10pt } }
+{%
+  \tikzset{%
+    chronos date/.expanded={#2},
+  }5
+  \pgfmathsetmacro\chronos@offset{(#5 < 0pt) ? (#5-.5*\chronos@height-.5*\chronos@borderheight) : (#5+.5*\chronos@height+.5*\chronos@borderheight)}%
+  \pgfmathsetmacro\chronos@anchor{(#5 < 0pt) ? "north" : "south"}%
+  \scoped[on background layer]{\path [postaction={/chronos/event line, #1}] ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- +(0,\chronos@offset pt) node [anchor=\chronos@anchor, /chronos/event, #3] {\chronos@showdate{thing}\\#4};}
+  \ifchronos@eventyearsonline
+    \edef\tempa{none}%
+    \edef\tempb{\chronos@specialdate}%
+    \ifx\tempa\tempb\else\let\chronos@thingyear\chronos@specialdate\tikzset{/chronos/special date=none}\fi
+    \node [/chronos/.cd, event year on line] at ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) {\chronos@thingyear};
+  \fi
+}
+\NewDocumentCommand \chronosspecialdate { m }
+{%
+  \tikzset{%
+    /chronos/special date=#1,
+  }%
+}
+\NewDocumentCommand \chronosperiod { O {} m O {} m O {} +m D () { -10pt } }
+{%
+  \tikzset{%
+    chronos date/.expanded={#2}, chronos period date/.expanded={#4}
+  }%
+  \pgfmathsetmacro\chronos@offset{(#7 < 0pt) ? (#7-.5*\chronos@height-.5*\chronos@borderheight) : (#7+.5*\chronos@height+.5*\chronos@borderheight)}%
+  \pgfmathsetmacro\chronos@anchor{(#7 < 0pt) ? "north" : "south"}%
+  \path [postaction={line width=\chronos@height, /chronos/period, #1}] ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- ({(\thechronos@otherthingdate-\thechronos@startdate)*\chronos@unit pt},0);
+  \scoped[on background layer]{\path [postaction={/chronos/period event line, #3}] ({(.5*\thechronos@otherthingdate+.5*\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- +(0,\chronos@offset pt) node [anchor=\chronos@anchor, /chronos/period event, #5] {\chronos@showdate{thing}--\chronos@showdate{otherthing}\\#6};}
+}
+\makeatother
+
+\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}
 
@@
 \newcommand{\onlyinsubfile}[1]{#1}
+\newcommand{\notinsubfile}[1]{}
 \begin{document}
 \renewcommand{\onlyinsubfile}[1]{}
+\renewcommand{\notinsubfile}[1]{#1}
+
 \begin{frame}
@@
 
-\subfile{intro.tex}
-
-\begin{frame}[standout]
-    What is the future of work?
-\end{frame}
-
-\begin{frame}{What is the future of work?}
-
-\end{frame}
-
-\begin{frame}{Introduction}
-  We hope to provide:
-      \begin{itemize}
-        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
-        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
-      \end{itemize}
-\end{frame}
+\subfile{introduction.tex}
 
@@
   \begin{itemize}
-    \item In general (see~\textcite{rosenberg1994exploring,rosenberg1982inside})
-    \item In HCI (see~\textcite{Wyche2006,bodker1993historical})
+    \item In general
+
+    \scriptsize{
+      \textcite{rosenberg1994exploring,rosenberg1982inside}
+    }
+    \item In HCI
+
+    \scriptsize{
+      \textcite{Wyche2006,bodker1993historical}
+    }
   \end{itemize}

### 2017-05-03 chi/piecework/presentation/complexity.tex
complexity stuff

@@
 
-\begin{frame}{Humble origins}
-  First implementations of piecework:
-  \visible<2->{farms}\visible<3->{, textiles}\visible<4->{, and other low--skill labor.}
+\begin{frame}{Cottage Industry}
+  First appearances of piecework:
+\visible<2->{farms}%
+\visible<3->{, textiles}%
+\visible<4->{, and matchsticks.}
   \begin{columns}
@@
   \visible<2->{
-  \begin{column}{0.3\textwidth}
+  \begin{column}{0.33\textwidth}
   \begin{figure}
@@
   \visible<3->{
-  \begin{column}{0.3\textwidth}
+  \begin{column}{0.33\textwidth}
   \begin{figure}
@@
   \visible<4->{
-  \begin{column}{0.3\textwidth}
+  \begin{column}{0.33\textwidth}
   \begin{figure}
@@
 
+\begin{frame}\frametitle{Planes, trains, and automobiles}
+             \framesubtitle{{\dots~not in that order}}
+
+  \begin{columns}
+  % \vspace{1mm}
+  \only<1,4->{
+  \begin{column}{0.33\textwidth}
+  \centering
+  Trains
+  \begin{figure}
+    \begin{overlayarea}{\textwidth}{4cm}
+    \begin{minipage}[c][4cm]{\textwidth}
+      \includegraphics[width=\textwidth]{figures/photo/train-engineer.jpg}
+    \end{minipage}
+    \end{overlayarea}
+    % \caption{i wanna fly the train}
+    \end{figure}
+\end{column}
+
+\only<1>{
+    \begin{column}{0.66\textwidth}
+      \begin{itemize}
+        \item Train engineers were among the first skilled workers to experience piece--rate payment systems
+        \item ``Efficiency experts'' occasionally came by to evaluate how long it took to do various jobs
+        \item This led to slow--walking the job, to offset the fact that
+              efficiency experts weren't accounting for liminal work.
+      \end{itemize}
+    \end{column}
+  }
+}
+
+  \only<2,4->{
+    \only<2>{
+      \begin{column}{0.33\textwidth}
+      \begin{itemize}
+        \item Ford, Taylor, and others moved quickly
+              monitor and measure professionals.
+      \end{itemize}
+      \end{column}
+    }
+
+  \begin{column}{0.33\textwidth}
+    \centering
+      Automobiles
+      \begin{figure}
+        \begin{overlayarea}{\textwidth}{4cm}
+        \begin{minipage}[c][4cm]{\textwidth}
+          \includegraphics[width=\textwidth]{figures/idk.png}
+        \end{minipage}
+        \end{overlayarea}
+        % \caption{something something}
+      \end{figure}
+  \end{column}
+
+  \only<2>{
+    \begin{column}{0.33\textwidth}
+    \begin{itemize}
+      \item \textit{Manufacturing} proved amenable
+            to assembly line processes.
+    \end{itemize}
+    \end{column}
+  }
+}
+
+  \only<3->{
+    \only<3>{
+      \begin{column}{0.66\textwidth}
+      \end{column}
+    }
+  \begin{column}{0.33\textwidth}
+    \centering
+      Planes
+      \begin{figure}
+        \begin{overlayarea}{\textwidth}{4cm}
+        \begin{minipage}[c][4cm]{\textwidth}
+            \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
+      \end{minipage}
+      \end{overlayarea}
+      % \caption{i wanna fly the plane}
+      \end{figure}
+  \end{column}
+  }
+
+  \end{columns}
+
+\end{frame}
+
 \begin{frame}{Ford}
-    Fordism
+    % Fordism
 

### 2017-05-04 chi/piecework/presentation/complexity.tex
checkin -- near finished

@@
     \begin{minipage}[c][4cm]{\textwidth}
-      \includegraphics[width=\textwidth]{figures/photo/train-engineer.jpg}
+      \includegraphics[width=\textwidth]{figures/photo/train_workers.jpg}
     \end{minipage}
@@
       \begin{itemize}
-        \item Train engineers were among the first skilled workers to experience piece--rate payment systems
-        \item ``Efficiency experts'' occasionally came by to evaluate how long it took to do various jobs
-        \item This led to slow--walking the job, to offset the fact that
-              efficiency experts weren't accounting for liminal work.
+        \item ``Efficiency experts'' measured how long it would take to do various jobs
+        \item These measurements would be used to assign values for each specific task
+        \item Train engineers performed work more slowly when inspectors were around
       \end{itemize}
@@
       \begin{itemize}
-        \item Ford, Taylor, and others moved quickly
-              monitor and measure professionals.
+        \item Fordism, Taylorism, and Scientific Management in full force
       \end{itemize}
@@
         \begin{minipage}[c][4cm]{\textwidth}
-          \includegraphics[width=\textwidth]{figures/idk.png}
+          \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
         \end{minipage}
@@
 
-\begin{frame}{Ford}
-    % Fordism
-
+\begin{frame}{Comparisons}
+\begin{itemize}
+  \item Limited range of tasks > arbitrary changes (building planes is easier than fixing trains)
+  \item Has technology changed this? Yes
+  \begin{itemize}
+    \item Technology makes complex tasks relatively trivial
+    \item Measuring workers is easier than ever
+  \end{itemize}
+\end{itemize}
 \end{frame}
 
-\begin{frame}{Comparisons}
-% \framesubtitle{something else}
+\begin{frame}{Complexity}{Cab Drivers}
 % \begin{columns}
-%   \begin{column}{0.5\textwidth}
-%   % \section*{Piecework}
-%      some text here some text here some text here some text here some text here
-
+  % \begin{column}{0.20\textwidth}
+    % \begin{itemize}
+    %   \item 
+      % Cab drivers
+    % \end{itemize}
   % \end{column}
 
-%   \begin{column}{0.5\textwidth}
-%   % \section*{CrowdWork}
-%      some text here some text here some text here some text here some text here
+  % \begin{column}{0.80\textwidth}
+  \begin{figure}
+  \includegraphics<1>[width=\textwidth]{figures/photo/knowledge-student.jpg}
+  \includegraphics<2>[width=\textwidth]{figures/photo/gps-map.jpg}
 
+  \end{figure}
   % \end{column}
 % \end{columns}
+\end{frame}
+
+\begin{frame}{Complexity}{Algorithmic Measurement}
+    notes
+    \begin{itemize}
+      \item I'm thinking of pointing to UpWork's screen recording tool as a way to measure workers
+      \item also maybe google analytics and other ways of tracking web--based workers
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Implications}
+  \begin{itemize}
+    \item We make stronger assumptions about workers' abilities thanks to technology
+    \item Evaluation remains difficult, but we're trying to find stopgap solutions through decomposition
+    \item We're still not solving the problems of inherently subjectively judged work
+  \end{itemize}
 \end{frame}

### 2017-05-05 chi/piecework/presentation/complexity.tex
some formatting fixes in the cottage industry section
timing of slides in cottage industry was shifted by one; fixed
presentation draft

@@
 \onlyinsubfile{
+\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
 \bibliography{references}
@@
       \item Build complexity into the process
-      \begin{mystepwiseitemize}
-        \item Apply CS methods to people
+      \begin{itemize}
+        \item<1> Apply CS methods to people
 
@@
         }
-        \item Humans as computational units
+        \item<2> Humans as computational units
 
@@
         }
-        \item Crowdsourcing workflows as function state machines
+        \item<3> Crowdsourcing workflows as function state machines
 
@@
         }
-      \end{mystepwiseitemize}
+      \end{itemize}
     \end{itemize}
@@
 \begin{frame}{Cottage Industry}
-  First appearances of piecework:
-\visible<2->{farms}%
-\visible<3->{, textiles}%
-\visible<4->{, and matchsticks.}
+  % First appearances of piecework:
+% \visible<2->{farms}%
+% \visible<3->{, textiles}%
+% \visible<4->{, and matchsticks.}
   \begin{columns}
-
-  \visible<2->{
+  \visible<1->{
     \begin{column}{0.33\textwidth}
+    \centering
+    Farms
     \begin{figure}
@@
 
-  \visible<3->{
+  \visible<2->{
   \begin{column}{0.33\textwidth}
+  \centering
+    Textiles
   \begin{figure}
@@
 
-  \visible<4->{
+  \visible<3->{
   \begin{column}{0.33\textwidth}
+  \centering
+    Matchsticks
   \begin{figure}
@@
       \begin{itemize}
-        \item ``Efficiency experts'' measured how long it would take to do various jobs
-        \item These measurements would be used to assign values for each specific task
-        \item Train engineers performed work more slowly when inspectors were around
+        \item ``Efficiency experts'' measured how long it would take to do various jobs~\cite{doi:10.2307/1883615}
+        \item These measurements would be used to assign values for each specific task~\cite{american1921problem}
+        \item Train engineers instituted ``The Fix'' to correct perceived unfairness~\cite{roy1954efficiency}
       \end{itemize}
@@
 \begin{itemize}
-  \item Limited range of tasks > arbitrary changes (building planes is easier than fixing trains)
-  \item Has technology changed this? Yes
+  \item Limited array of tasks versus arbitrarily complex work
+  \begin{itemize}
+    \item \textit{Building} planes versus \textit{fixing} trains
+  \end{itemize}
+  \item \normalsize{Has technology changed this?}
     \begin{itemize}

### 2017-05-05 chi/piecework/presentation/introduction.tex
presentation draft

@@
 
-\begin{frame}{Open problems in crowdsourcing}
+% \frame<1-3>[frame1]{
+\begin{frame}<-2>[label=returnable]{A brief glossary}
+      \begin{itemize}
+        \item<1> \textit{Crowd work}: digitally mediated \textbf{information work}
+        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)~\cite{crowdworkFuture}
+        \item<2> \textit{Gig work}: digitally mediated (but often \textbf{physically embodied}) one--off jobs,
+        such as
+        \textit{driving},
+        \textit{courier services},
+        and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
+        \item<3> \textit{Piecework}: Payment for \textit{output} rather than for \textit{time}
+      \end{itemize}
+\end{frame}
+
+\begin{frame}{Open problems in on--demand work}
 \begin{columns}
@@
   \end{itemize}
-
 \end{column}
@@
 
-    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}
-    }\visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
+    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
+    \visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
                   \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
@@
 
+% \begin{frame}[label=ndinpi]{Non-determinism in the $\pi$-calculus (or ``global'' choice)}
+%   % \only<2>{\frametitle{Example: a p\^{a}tisserie}}
+%   \centering
+%   \only<1>{
+%   something 1
+%   }\only<2>{
+%   something 2
+%   }
+% \end{frame}
+
+\againframe<3->{returnable}
+
+% \begin{frame}{A brief glossary}
+%     \begin{itemize}
+%       \item \textit{Crowd work}: digitally mediated \textbf{information work}
+%       (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)~\cite{crowdworkFuture}
+%       \item \textit{Gig work}: digitally mediated (but often \textbf{physically embodied}) one--off jobs,
+%       such as
+%       \textit{driving},
+%       \textit{courier services},
+%       and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
+%       \item<2> \textit{Piecework}: Payment for \textit{output} rather than for \textit{time}
+%     \end{itemize}
+% \end{frame}
+
+% \againframe<2>{ndinpi}
+
 \notinsubfile{
@@
 
+\begin{frame}{A case for comparative historical analysis}
+\begin{itemize}
+  \item Historical analysis isn't new
+  \begin{itemize}
+    \item In general
+
+    \scriptsize{
+      \textcite{rosenberg1994exploring,rosenberg1982inside}
+    }
+    \item In HCI
+
+    \scriptsize{
+      \textcite{Wyche2006,bodker1993historical}
+    }
+  \end{itemize}
+  \item Still, it's an underutilized method
+  \begin{itemize}
+    \item Provide some basic framing for ostensibly new phenomena
+    \item Theoretically ground ourselves
+    \item Flesh out \textit{differences} and their implications
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\onlyinsubfile{
+  \printbibliography{}
+}
+
 \end{document}

### 2017-05-05 chi/piecework/presentation/presentation.tex
complexity stuff
presentation draft

@@
 \usepackage[en-US]{datetime2}
+
 \bibliography{references}
@@
   \usecolortheme{seagull}
+
+% \setbeamercolor{framesubtitle}{fg=mDarkTeal}
+% \addtobeamert emplate{frametitle}{}{%
+%   \ifx\insertframesubtitle\@empty\else%
+%   \usebeamerfont{framesubtitle}%
+%   \usebeamercolor[fg]{framesubtitle}%
+%   \small{\insertframesubtitle}%
+%   \fi%
+% }
+
+\setbeamertemplate{frametitle}
+{\vskip-3pt
+  \leavevmode
+  \hbox{%
+  \begin{beamercolorbox}[wd=\paperwidth,ht=5.8ex,dp=1ex]{frametitle}%
+    \raggedright\hspace*{2em}\large{\insertframetitle}\\
+    \hspace*{4em}\small{\insertframesubtitle}\hspace*{2em}
+  \end{beamercolorbox}
+  }%
+}
+\setbeamertemplate{caption}{\raggedright\footnotesize{\insertcaption}\par}
 
@@
 \setbeamertemplate{itemize items}{--}
+% \setbeamertemplate{itemize subitems}{--}
 
@@
 \subfile{introduction.tex}
-
-\begin{frame}{A case for comparative historical analysis}
-\begin{itemize}
-  \item Historical analysis isn't new
-  \begin{itemize}
-    \item In general
-
-    \scriptsize{
-      \textcite{rosenberg1994exploring,rosenberg1982inside}
-    }
-    \item In HCI
-
-    \scriptsize{
-      \textcite{Wyche2006,bodker1993historical}
-    }
-  \end{itemize}
-  % \item Still can be useful.
-\end{itemize}
-\end{frame}
-
-\begin{frame}{A brief glossary}
-    \begin{itemize}
-      \item Crowd work: digitally mediated \textbf{information work}
-      --- for example, work done on Amazon Mechanical Turk~\cite{crowdworkFuture}
-      \item Gig work: digitally mediated --- but often \textbf{physically embodied} --- one--off jobs,
-      such as
-      \textit{driving},
-      \textit{courier services},
-      and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
-    \end{itemize}
-\end{frame}
-
 \subfile{complexity}

### 2017-05-05 chi/piecework/presentation/introduction.tex
nightly updates

@@
       \begin{itemize}
-        \item<1> \textit{Crowd work}: digitally mediated \textbf{information work}
+        \item<1-> \textit{Crowd work}: digitally mediated \textbf{information work}
         (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)~\cite{crowdworkFuture}
-        \item<2> \textit{Gig work}: digitally mediated (but often \textbf{physically embodied}) one--off jobs,
+        \item<2-> \textit{Gig work}: digitally mediated (but often \textbf{physically embodied}) one--off jobs,
         such as
@@
         and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
-        \item<3> \textit{Piecework}: Payment for \textit{output} rather than for \textit{time}
+        \item<3-> \textit{Piecework}: Payment for \textit{output} rather than for \textit{time}
       \end{itemize}
@@
     \vspace{5mm}
-    \includegraphics[width=\textwidth]{figures/complexity/hexRepeat.png}
+    \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
     \vspace{5mm}

### 2017-05-06 chi/piecework/presentation/complexity.tex
design changes

@@
   \begin{itemize}
-    \item<1-> Can crowds help you write something?
-
+    \item<1-> Can crowds help you write something?\\
     \scriptsize{
@@
     }
-    \item<2-> Can crowds critique novel designs?
-
+    \item<2-> Can crowds critique novel designs?\\
     \scriptsize{
@@
     }
-    \item<3-> Can crowds create things from whole cloth?
-
+    \item<3-> Can crowds create things from whole cloth?\\
     \scriptsize{
@@
       \begin{itemize}
-        \item<1> Apply CS methods to people
-
+        \item<1> Apply CS methods to people\\
         \scriptsize{
@@
         }
-        \item<2> Humans as computational units
-
+        \item<2> Humans as computational units\\
         \scriptsize{
@@
         }
-        \item<3> Crowdsourcing workflows as function state machines
-
+        \item<3> Crowdsourcing workflows as function state machines\\
         \scriptsize{\textcite{latoza2014microtask}
@@
       % \begin{itemize}
-      %   \item Break work up so that individual tasks are simple
-
+      %   \item Break work up so that individual tasks are simple\\
       %   Make the process \textit{itself} compile the final product.

### 2017-05-06 chi/piecework/presentation/introduction.tex
design changes

--- 2017-05-03 chi/piecework/presentation/introduction.tex
@@
 
-\begin{frame}{Open problems in crowdsourcing}
+% \frame<1-3>[frame1]{
+\begin{frame}<-2>[label=returnable]{A brief glossary}
+      \begin{itemize}
+        \item<1-> \alert{Crowd work}: digitally mediated \emph{information work}
+        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)~\cite{crowdworkFuture}
+        \item<2-> \alert{Gig work}: digitally mediated (often \emph{physically embodied}) one--off jobs,
+        such as
+        {driving},
+        {courier services},
+        and {administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
+        \item<3-> \alert{Piecework}: Payment for \emph{output} rather than for \emph{time}
+      \end{itemize}
+\end{frame}
+
+\begin{frame}{Open problems in on--demand work}
 \begin{columns}
@@
   \begin{itemize}%[<+>]
-    \item<2,4> \textbf{\large Complexity}
-
+    \item<2,4> \textbf{\large Complexity}\\
     \scriptsize{
@@
 
-    \item<3> \textbf{\large Decomposition}
-
+    \item<3> \textbf{\large Decomposition}\\
     \scriptsize{
@@
 
-    \item<5> \textbf{\large Relationships}
-
+    \item<5> \textbf{\large Relationships}\\
     \scriptsize{
@@
 \begin{frame}{What is the future of work?}
-    How will \textbf{technology} affect the \textbf{complexity} of the work that on--demand workers do?
+    How will \alert{technology} affect the complexity of the work that on--demand workers do?
 
-    What are the \textbf{limits} of complexity in on--demand work?
+    What are the \alert{limits} of complexity in on--demand work?
 
-    How can we \textbf{reach} those limits?
+    How can we \alert{reach} those limits?
 \end{frame}
+
+\begin{frame}{Thesis}
+    These questions have all been asked before.
+
+    History can help us answer them today.
+
+    We'll reach into the history of \alert{piecework}
+    --- of human computers, match stick makers, and metalworkers ---
+    and show how the \alert{history} of their work can
+    inform answers to questions about the \alert{future} of digital work.
+\end{frame}
+
+\againframe<2->{returnable}
 
@@
       \begin{itemize}
-        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
-        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
+        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\alert{on--demand work}'') as a resurgence of \alert{piecework}.
+        \item A method for making sense of contemporary phenomena through \alert{historical analysis}.
       \end{itemize}
@@
 
+\begin{frame}{A case for comparative historical analysis}
+\begin{itemize}
+  \item Historical analysis isn't new
+  \begin{itemize}
+    \item In general\\
+    \scriptsize{
+      \textcite{rosenberg1994exploring,rosenberg1982inside}
+    }
+    \item In HCI\\
+    \scriptsize{
+      \textcite{Wyche2006,bodker1993historical}
+    }
+  \end{itemize}
+  \item Still, it's an underutilized method
+  \begin{itemize}
+    \item Provide some basic framing for ostensibly new phenomena
+    \item Theoretically ground ourselves
+    \item Flesh out \emph{differences} and their implications
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\onlyinsubfile{
+  \printbibliography{}
+}
+
 \end{document}

### 2017-05-06 chi/piecework/presentation/complexity.tex
various minor updates and notes from msb

@@
 \begin{frame}{Complexity}
-  What kinds of problems do we mean when we talk about complexity?
+  What do we mean when we talk about complexity?
   \begin{itemize}
@@
 
-\begin{frame}{What does the crowdsourcing literature say?}
+\begin{frame}{What Does Crowdsourcing Say?}
 \begin{columns}
@@
 
-\begin{frame}{What does the piecework literature say?}
+\begin{frame}{What Does Piecework Say?}
     % \begin{itemize}
       % \item
-      George Airy (astronomer) used a very similar approach~\cite{grier2013computers}
+      George Airy (astronomer) used a very similar approach\\
+      \scriptsize{\textcite{grier2013computers}}
       % \begin{itemize}
@@
 
-\begin{frame}{George Airy --- whiz kid}
+\begin{frame}{George Airy}
     Airy built complexity into the process, assigning \textit{human computers} 
@@
 
-\begin{frame}\frametitle{Planes, trains, and automobiles}
-             \framesubtitle{{\dots~not in that order}}
-
+\begin{frame}{Planes, Trains, and Automobiles}
   \begin{columns}
@@
       \begin{itemize}
-        \item ``Efficiency experts'' measured how long it would take to do various jobs~\cite{doi:10.2307/1883615}
-        \item These measurements would be used to assign values for each specific task~\cite{american1921problem}
-        \item Train engineers instituted ``The Fix'' to correct perceived unfairness~\cite{roy1954efficiency}
+        \item ``Efficiency experts'' measured how long it would take to do various jobs\\
+        \scriptsize{\textcite{doi:10.2307/1883615}}
+        \item These measurements would be used to assign values for each specific task\\
+        \scriptsize{\textcite{american1921problem}}
+        \item Train engineers instituted ``The Fix'' to correct perceived unfairness\\
+        \scriptsize{\textcite{roy1954efficiency}}
       \end{itemize}
@@
         \end{overlayarea}
-        % \caption{something something}
       \end{figure}
@@
       \begin{column}{0.66\textwidth}
+      \begin{itemize}
+        \item Men drafted during World War II
+        \item Factories turned to a new workforce who had neither conventional training nor experience
+        \item Training for individual tasks and correspondingly assigning workers 
+      \end{itemize}
       \end{column}
@@
       \end{overlayarea}
-      % \caption{i wanna fly the plane}
       \end{figure}
@@
 
-\begin{frame}{Complexity}{Cab Drivers}
+\begin{frame}{Complexity} % {Cab Drivers}
 % \begin{columns}
@@
   \includegraphics<1>[width=\textwidth]{figures/photo/knowledge-student.jpg}
-  \includegraphics<2>[width=\textwidth]{figures/photo/gps-map.jpg}
-
+  \only<2>{
+  \begin{columns}
+  \begin{column}{0.2\textwidth}
+  \includegraphics[width=\textwidth]{figures/photo/knowledge-student.jpg}
+  \end{column}
+  \begin{column}{0.8\textwidth}
+  \includegraphics[width=\textwidth]{figures/complexity/brain_scans.png}
+  \end{column}
+  \end{columns}
+  }
+  \includegraphics<3>[width=\textwidth]{figures/photo/gps-map.jpg}
   \end{figure}
@@
 
-\begin{frame}{Complexity}{Algorithmic Measurement}
+\begin{frame}{Complexity} % {Algorithmic Measurement}
     notes

### 2017-05-06 chi/piecework/presentation/introduction.tex
various minor updates and notes from msb

@@
 % \frame<1-3>[frame1]{
-\begin{frame}<-2>[label=returnable]{A brief glossary}
+\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
       \begin{itemize}
         \item<1-> \alert{Crowd work}: digitally mediated \emph{information work}
-        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)~\cite{crowdworkFuture}
+        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)\\
+          \scriptsize{\textcite{crowdworkFuture}}
         \item<2-> \alert{Gig work}: digitally mediated (often \emph{physically embodied}) one--off jobs,
@@
         {courier services},
-        and {administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
-        \item<3-> \alert{Piecework}: Payment for \emph{output} rather than for \emph{time}
+        and {administrative support}\\
+          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}
+        \item<3-> \alert{On--demand work}: crowd work and gig work, collectively
+        \item<5-> \alert{Piecework}: Payment for \emph{output} rather than for \emph{time}
       \end{itemize}
@@
 
-\begin{frame}{Open problems in on--demand work}
+\begin{frame}{Ongoing Threads in Crowdsourcing Research}
 \begin{columns}
 \begin{column}[T]{0.5\textwidth}
-  \begin{itemize}%[<+>]
+  \begin{enumerate}%[<+>]
     \item<2,4> \textbf{\large Complexity}\\
@@
     }
-    \vspace{8mm}
+    \vspace{5mm}
 
@@
 
-  \end{itemize}
+  \end{enumerate}
 \end{column}
@@
     \centering
-    \vspace{-5mm}
+    % \vspace{-5mm}
+    \vspace{0mm}
 
@@
 
-    \vspace{5mm}
+    \vspace{4mm}
     \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
@@
     \visible<5->{
-      \vspace{10mm}
-      \includegraphics[width=\textwidth]{figures/idk.png}
+      \vspace{5mm}
+      \includegraphics[width=\textwidth]{figures/networkTransparency.png}
     }
@@
   \small{Complexity}
-  \vspace{10mm}
+  \vspace{13mm}
 }
@@
 \visible<3->{
-  \vspace{10mm}
+  \vspace{11mm}
   \small{Decomposition}
@@
 \begin{frame}[standout]
-    What is the future of work?
+    What will be the future of work?
 \end{frame}
 
-\begin{frame}{What is the future of work?}
+\begin{frame}{What will be the future of work?}
     How will \alert{technology} affect the complexity of the work that on--demand workers do?
@@
 
-\againframe<2->{returnable}
+\againframe<4->{returnable}
 
@@
       \begin{itemize}
-        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\alert{on--demand work}'') as a resurgence of \alert{piecework}.
+        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}.
         \item A method for making sense of contemporary phenomena through \alert{historical analysis}.
@@
 
-\begin{frame}{A case for comparative historical analysis}
+\begin{frame}{Comparative Historical Analysis}
 \begin{itemize}
@@
   \begin{itemize}
-    \item Provide some basic framing for ostensibly new phenomena
-    \item Theoretically ground ourselves
+    \item Provide some basic framing for \emph{ostensibly} new phenomena
+    \item \emph{Explicate} our theoretical grounding
     \item Flesh out \emph{differences} and their implications

### 2017-05-06 chi/piecework/presentation/presentation.tex
design changes
various minor updates and notes from msb

--- 2017-05-03 chi/piecework/presentation/presentation.tex
@@
-% !TEX program = lualatex
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beamer Presentation
-% LaTeX Template
-% Version 1.0 (10/11/12)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass{beamer}
 
-\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles}
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
+\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
+\usepackage{pgfcalendar}
+\usepackage{datenumber,xparse,fp}
+
+\usetikzlibrary{arrows,automata}
+\usepackage[citestyle=authoryear-comp,backend=bibtex]{biblatex}
 \usepackage[export]{adjustbox}
 \usepackage[en-US]{datetime2}
+
 \bibliography{references}
@@
 \beamertemplatenavigationsymbolsempty
-\renewcommand\textbullet{\ensuremath{\bullet}}
 \definecolor{stanfordRed}{HTML}{8C1515}
-% \mode<presentation>{
-  % \usetheme{metropolis}
-  \usecolortheme{seagull}
+\definecolor{metroBackground}{RGB}{250,250,250}
+
+% Recent Notes from msb
+% [ ] I suggest that you come out with a one sentence summary of your main thesis after 2 and before 3.
+“What I’m going to argue in this talk is that on-demand work is a modern instantiation of a much older phenomenon, piecework [put the word on the slide], and that we’re going to be able to use the history of piecework over the past two hundred years to shed light on several unanswered questions about the unfolding phenomenon of on-demand work.”
+% [ ] This way people know what you’re winding up for in this talk
+% [ ] don’t get rid of slide 5. I’m just saying put one framing sentence up front here.
+% [ ] 3: these are not the canonical cites for complexity. kittur crowdforge is most famous
+% [ ] 3: put the task hexagons at the top, not in between. the hierarchy I see what you’re trying to do, but it’s not really working
+% [ ] When you introduce the piecework term, give several examples (with pictures) that listeners will be familiar with.
+% [ ] You never stopped to equate piecework to on demand work. Why does piecework = on demand work? (in particular, because on demand work ALSO pays people per task)
+% [ ] 3: your complexity example is being built out of triangles, not hexagons
+% [ ] Get rid of the citation numbers, and replace it with years Alkhatib 2017, not Alkhatib [1]
+% [ ] 3: I think this isn’t focused correctly.
+I’d get rid of this slide and ONLY motivate the complexity question.
+Then mention briefly that in the paper you also talk about decomposition and worker relationships, but you need 10 pages and 166 references rather than a 15 minute talk to do all that.
+Return to these other themes at the end of the talk and mention some short take aways from each one.
+% [ ] 4: doesn’t make sense to ask “how will we reach those limits?”
+% [ ] LOL at the phone ringing
+% [ ] 7: the top timeline is events, the bottom is citations. change the bottom one to be events
+% [ ] 8: this is WAY too late to still be doing an intro. You’re five minutes into the talk. Intro should be done by 2min.
+% [ ] 12: the details here are irrelevant to complexity.
+why does it matter that they could be fired at will, overworked, underpaid?
+shouldn’t we be focusing on how piecework enabled the complexity?
+If it’s not important (as you say), then don’t include it.
+Parsimony, yay!
+% [ ] 12: another example of a citation I can’t follow by watching slides
+% [ ] 14: ditch the shrugging guy, he’s really weird
+% [ ] 15: trains — this has nothing to do with complexity…? cut?
+% [ ] 15: fordism — again, explain what this has to do with complexity? what does fordism have to do with complex outcomes? spell it out. what should we be learning from this?
+% [ ] 16: technology makes complex tasks relatively trivial — this is an overly broad claim. makes SOME?
+% [ ] 17: speed up The Knowledge/Gmaps example. I think it can go twice as fast and still carry as much weight
+
+\setbeamertemplate{frametitle}
+{\vskip-3pt
+  \leavevmode
+  \hbox{%
+  \begin{beamercolorbox}[wd=\paperwidth,ht=5.8ex,dp=1ex]{frametitle}%
+    \raggedright\hspace*{1em}\large{\insertframetitle}
+  \end{beamercolorbox}
+  }%
+}
+\setbeamertemplate{caption}{\raggedright\footnotesize{\insertcaption}\par}
+
+\addtobeamertemplate{frametitle}{}{%
+    \begin{textblock*}{28mm}(\paperwidth-40mm,-12mm)
+      \includegraphics[height=10mm]{figures/origami-1.png}
+      \includegraphics[height=10mm]{figures/origami-2.png}
+      \includegraphics[height=8mm]{figures/origami-white.png}
+    \end{textblock*}}
+
+\input{timeline_code}
+
+\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}
 
@@
 \setmonofont{Source Code Pro}
-% \setbeamercolor{frametitle}{bg=stanfordRed}
-
-  % \setbeamercolor*{palette tertiary}{use=structure,fg=stanfordRed,bg=stanfordRed}
-% }
-% \metroset{set}
 \setbeamercolor{institute in head/foot}{fg=stanfordRed}
@@
 \setbeamercovered{again covered={\opaqueness<1->{15}}}
-% \setbeamercolor{progress bar}{fg=stanfordRed}
-% \setbeamercolor{progress bar}{bg=gr}
-% \setsansfont{Source Sans Pro}
 
 \title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
-\metroset{titleformat=smallcaps}
+% \metroset{titleformat=smallcaps}
 
-\author{\textbf{Ali Alkhatib},
-                Michael Bernstein,
-                Margaret Levi\\
+\author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
 \texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
@@
 \date{\today}
-\setbeamertemplate{itemize items}{--}
 
@@
 
-% \def\labelitemi{--}
-
-% \let\oldcite=\textcite
-% \renewcommand{\textcite}[1]{\textcolor[rgb]{.2,.2,.9}{[\oldcite{#1}]}}
-
 \newcommand{\onlyinsubfile}[1]{#1}
+\newcommand{\notinsubfile}[1]{}
 \begin{document}
 \renewcommand{\onlyinsubfile}[1]{}
+\renewcommand{\notinsubfile}[1]{#1}
+
 \begin{frame}
@@
 
-\subfile{intro.tex}
+\subfile{introduction}
+\subfile{complexity}
+\subfile{closing}
+\subfile{contact}
 
-\begin{frame}[standout]
-    What is the future of work?
-\end{frame}
+% \begin{frame}[allowframebreaks]{References}
+% \printbibliography{}
+% \end{frame}
 
-\begin{frame}{What is the future of work?}
-
-\end{frame}
-
-\begin{frame}{Introduction}
-  We hope to provide:
-      \begin{itemize}
-        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
-        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
-      \end{itemize}
-\end{frame}
-
-\begin{frame}{A case for comparative historical analysis}
-\begin{itemize}
-  \item Historical analysis isn't new
-  \begin{itemize}
-    \item In general (see~\textcite{rosenberg1994exploring,rosenberg1982inside})
-    \item In HCI (see~\textcite{Wyche2006,bodker1993historical})
-  \end{itemize}
-  % \item Still can be useful.
-\end{itemize}
-\end{frame}
-
-\begin{frame}{A brief glossary}
-    \begin{itemize}
-      \item Crowd work: digitally mediated \textbf{information work}
-      --- for example, work done on Amazon Mechanical Turk~\cite{crowdworkFuture}
-      \item Gig work: digitally mediated --- but often \textbf{physically embodied} --- one--off jobs,
-      such as
-      \textit{driving},
-      \textit{courier services},
-      and \textit{administrative support}~\cite{friedman2014workers,Parigi:2016:GE:3026779.3013496}
-    \end{itemize}
-\end{frame}
-
-\subfile{complexity}
-% \subfile{decomposition}
-% \subfile{relationships}
-% \subfile{contact}
-
-% \bibliographystyle{SIGCHI-Reference-Format}
-\printbibliography{}
 \end{document}

### 2017-05-09 chi/piecework/presentation/complexity.tex
checkin presentation

@@
 \onlyinsubfile{
-\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-\bibliography{references}
-  % \beamertemplategridbackground[1mm]
 }
-
 \begin{document}
@@
   What do we mean when we talk about complexity?
+  % \begin{minipage}[t][25mm][t]{\textwidth}
+  %   \begin{columns}
+  %     \begin{column}[T]{0.8\textwidth}
         \begin{itemize}
-    \item<1-> Can crowds help you write something?\\
-    \scriptsize{
-      \textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}
-    }
+          \item<1->
+          Can crowds help you write something?\\
+          \scriptsize{\textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}}\normalsize{}
           \item<2-> Can crowds critique novel designs?\\
-    \scriptsize{
-      \textcite{yuanAlmost,fuge2014analysis}
-    }
+          \scriptsize{\textcite{yuanAlmost,fuge2014analysis}}\normalsize{}
           \item<3-> Can crowds create things from whole cloth?\\
-    \scriptsize{
-      \textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
-    }
+          \scriptsize{\textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
         \end{itemize}
+      % \end{column}
+      % \begin{column}[T]{0.2\textwidth}
+      %   \begin{figure}
+    % \centering
+    % \vspace{0mm}
+
+    % \visible<1->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
+    % \visible<3->{
+    % \includegraphics<1->[width=.5\textwidth]{figures/complexity/paper_resized.png}
+    %               % \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
+    % % }
+
+    % \includegraphics<2->[width=.5\textwidth]{figures/complexity/chair_resized.png}
+
+    % \includegraphics[width=.15\textwidth]{figures/paper-unfolded.png}$\rightarrow$
+    % \includegraphics[width=.15\textwidth]{figures/rhino.png}
+    % \includegraphics[width=.15\textwidth]{figures/fish.png}
+    % \includegraphics[width=.15\textwidth]{figures/origami-1b.png}
+
+    % \vspace{4mm}
+    % \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
+    % \vspace{5mm}
+
+    % \visible<2->{
+    % \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
+    % }
+
+    % \visible<4->{
+    %   \vspace{5mm}
+    %   \includegraphics[width=\textwidth]{figures/networkTransparency.png}
+    % }
+  % \end{figure}
+  %     \end{column}
+  %   \end{columns}
+  %   \end{minipage}
 \end{frame}
@@
         \item<1> Apply CS methods to people\\
-        \scriptsize{
-          \textcite{crowdForgeKittur}
-        }
+        \scriptsize{\textcite{crowdForgeKittur}}\normalsize{}
         \item<2> Humans as computational units\\
-        \scriptsize{
-          \textcite{Lasecki:2014:LSR:2661334.2661352}
-        }
+        \scriptsize{\textcite{Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
         \item<3> Crowdsourcing workflows as function state machines\\
@@
 \begin{frame}{What Does Piecework Say?}
-    % \begin{itemize}
-      % \item
       George Airy (astronomer) used a very similar approach\\
-      \scriptsize{\textcite{grier2013computers}}
-      % \begin{itemize}
-      %   \item Break work up so that individual tasks are simple\\
-      %   Make the process \textit{itself} compile the final product.
-      % \end{itemize}
-    % \end{itemize}
-    % \vspace{0mm}
+      \scriptsize{\textcite{grier2013computers}}\normalsize{}
     \begin{columns}
@@
         \item 13--20 years old
-        \item Overworked
-        \item Underpaid
-        \item Could be fired at will
+        \item no particularly strong background in mathematics
+        \item A basic understanding of logarithms, algebra, etc\dots
       \end{itemize}
@@
 \begin{frame}{George Airy}
-    Airy built complexity into the process, assigning \textit{human computers} 
-    to compute, verify, and correct the right ascension and declination of stars.
+    Airy built complexity into the process, assigning \emph{human computers} 
+    to calculate \& verify the \emph{right ascension} and \emph{declination} of stars.
 
@@
 
-\begin{frame}{Cottage Industry}
-  % First appearances of piecework:
-% \visible<2->{farms}%
-% \visible<3->{, textiles}%
-% \visible<4->{, and matchsticks.}
-  \begin{columns}
-  \visible<1->{
-    \begin{column}{0.33\textwidth}
-    \centering
-    Farms
-    \begin{figure}
-    \includegraphics[width=\textwidth]{figures/idk.png}
-    \end{figure}
-    \end{column}
-    }
-
-  \visible<2->{
-  \begin{column}{0.33\textwidth}
-  \centering
-    Textiles
-  \begin{figure}
-  \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
-  \end{figure}
-  \end{column}
-  }
-
-  \visible<3->{
-  \begin{column}{0.33\textwidth}
-  \centering
-    Matchsticks
-  \begin{figure}
-  \includegraphics[width=\textwidth]{figures/photo/Matchgirl_strikers.PNG}
-  \end{figure}
-  \end{column}
-  }
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Planes, Trains, and Automobiles}
-  \begin{columns}
-  % \vspace{1mm}
-  \only<1,4->{
-  \begin{column}{0.33\textwidth}
-  \centering
-  Trains
-  \begin{figure}
-    \begin{overlayarea}{\textwidth}{4cm}
-    \begin{minipage}[c][4cm]{\textwidth}
-      \includegraphics[width=\textwidth]{figures/photo/train_workers.jpg}
-    \end{minipage}
-    \end{overlayarea}
-    % \caption{i wanna fly the train}
-    \end{figure}
-\end{column}
-
-\only<1>{
-    \begin{column}{0.66\textwidth}
-      \begin{itemize}
-        \item ``Efficiency experts'' measured how long it would take to do various jobs\\
-        \scriptsize{\textcite{doi:10.2307/1883615}}
-        \item These measurements would be used to assign values for each specific task\\
-        \scriptsize{\textcite{american1921problem}}
-        \item Train engineers instituted ``The Fix'' to correct perceived unfairness\\
-        \scriptsize{\textcite{roy1954efficiency}}
-      \end{itemize}
-    \end{column}
-  }
-}
-
-  \only<2,4->{
-    \only<2>{
-      \begin{column}{0.33\textwidth}
-      \begin{itemize}
-        \item Fordism, Taylorism, and Scientific Management in full force
-      \end{itemize}
-      \end{column}
-    }
-
-  \begin{column}{0.33\textwidth}
-    \centering
-      Automobiles
-      \begin{figure}
-        \begin{overlayarea}{\textwidth}{4cm}
-        \begin{minipage}[c][4cm]{\textwidth}
-          \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
-        \end{minipage}
-        \end{overlayarea}
-      \end{figure}
-  \end{column}
-
-  \only<2>{
-    \begin{column}{0.33\textwidth}
-    \begin{itemize}
-      \item \textit{Manufacturing} proved amenable
-            to assembly line processes.
-    \end{itemize}
-    \end{column}
-  }
-}
-
-  \only<3->{
-    \only<3>{
-      \begin{column}{0.66\textwidth}
-      \begin{itemize}
-        \item Men drafted during World War II
-        \item Factories turned to a new workforce who had neither conventional training nor experience
-        \item Training for individual tasks and correspondingly assigning workers 
-      \end{itemize}
-      \end{column}
-    }
-  \begin{column}{0.33\textwidth}
-    \centering
-      Planes
-      \begin{figure}
-        \begin{overlayarea}{\textwidth}{4cm}
-        \begin{minipage}[c][4cm]{\textwidth}
-            \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
-      \end{minipage}
-      \end{overlayarea}
-      \end{figure}
-  \end{column}
-  }
-
-  \end{columns}
-
-\end{frame}
+\subfile{piecework-simple}
+\subfile{piecework-complex}
 
@@
   \begin{itemize}
-    \item \textit{Building} planes versus \textit{fixing} trains
+    \item \emph{Building} planes versus \emph{fixing} trains
   \end{itemize}
@@
     \begin{itemize}
-      \item Technology makes complex tasks relatively trivial
+      \item Technology makes \emph{some} complex tasks relatively trivial
       \item Measuring workers is easier than ever
@@
 \begin{frame}{Complexity} % {Cab Drivers}
-% \begin{columns}
-  % \begin{column}{0.20\textwidth}
-    % \begin{itemize}
-    %   \item 
-      % Cab drivers
-    % \end{itemize}
-  % \end{column}
-
-  % \begin{column}{0.80\textwidth}
   \begin{figure}
@@
   \end{figure}
-  % \end{column}
-% \end{columns}
 \end{frame}

### 2017-05-09 chi/piecework/presentation/introduction.tex
checkin presentation

@@
 \documentclass[presentation]{subfiles}
+\onlyinsubfile{}
+\setbeamersize{description width=30mm}
+\begin{document}
+\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
+      \begin{description}
+        \item<1-> [Crowd work] Digitally mediated \alert{information work}
+        (e.g. Amazon Mechanical Turk, UpWork)\\
+          \scriptsize{\textcite{crowdworkFuture}}\normalsize{}
+        \item<2-> [Gig work] Digitally mediated (often \alert{physically embodied}) one--off jobs,
+        such as
+        \emph{driving for hire},
+        \emph{courier services},
+        and \emph{administrative support}\\
+          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}\normalsize{}
+        \item<3-> [On--demand~work] Crowd work and gig work, collectively
+        \item<5-> [Piecework] Payment for \emph{output} rather than for \emph{time}
+      \end{description}
+\end{frame}
 
-\onlyinsubfile{
-  % \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
-  % \bibliography{~/Library/texmf/bibtex/bib/local/references}
-  % \beamertemplategridbackground[1mm]
+\begin{frame}[standout]
+On--demand work is a modern instantiation of a much older phenomenon
+--- \alert{piecework}.
+
+{\normalsize The historical arc of piecework
+can shed light on persistent questions
+in this ongoing phenomenon of on--demand work.}
+\end{frame}
+
+\againframe<4->{returnable}
+
+\begin{frame}[t]{What is piecework?}
+\centering
+  \emph{Payment for \textbf{output} rather than for \textbf{time}}
+\vspace*{2mm}
+\begin{columns}
+  \begin{column}{0.33\textwidth}
+    \centering
+    Textiles
+    \begin{figure}
+    \begin{overlayarea}{\textwidth}{33mm}
+    \begin{minipage}[c][33mm]{\textwidth}
+      \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
+    \end{minipage}
+    \end{overlayarea}
+    \end{figure}
+  \end{column}
+  \begin{column}{0.33\textwidth}
+    \centering
+    Automobiles
+    \begin{figure}
+    \begin{overlayarea}{\textwidth}{33mm}
+    \begin{minipage}[c][33mm]{\textwidth}
+      \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
+    \end{minipage}
+    \end{overlayarea}
+    \end{figure}
+  \end{column}
+  \begin{column}{0.33\textwidth}
+    \centering
+    Metalwork
+    \begin{figure}
+    \begin{overlayarea}{\textwidth}{33mm}
+    \begin{minipage}[c][33mm]{\textwidth}
+      \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
+    \end{minipage}
+    \end{overlayarea}
+    \end{figure}
+  \end{column}
+\end{columns}
+
+\visible<2->{
+  \begin{columns}
+  \centering
+    \begin{column}{0.33\textwidth}
+    \centering
+      Crowd work
+    \begin{figure}
+      \begin{overlayarea}{\textwidth}{15mm}
+      \begin{minipage}[c][15mm]{\textwidth}
+        \includegraphics[width=\textwidth]{figures/amt.png}
+      \end{minipage}
+      \end{overlayarea}
+    \end{figure}
+    \end{column}
+    \begin{column}{0.33\textwidth}
+    \centering
+      Gig Work
+    \begin{figure}
+    \begin{overlayarea}{\textwidth}{15mm}
+      \begin{minipage}[c][15mm]{\textwidth}
+        \includegraphics[width=\textwidth]{figures/uber.png}
+      \end{minipage}
+      \end{overlayarea}
+    \end{figure}
+    \end{column}
+  \end{columns}
 }
 
-\begin{document}
-% \section*{Introduction}
-
-% \frame<1-3>[frame1]{
-\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
-      \begin{itemize}
-        \item<1-> \alert{Crowd work}: digitally mediated \emph{information work}
-        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)\\
-          \scriptsize{\textcite{crowdworkFuture}}
-        \item<2-> \alert{Gig work}: digitally mediated (often \emph{physically embodied}) one--off jobs,
-        such as
-        {driving},
-        {courier services},
-        and {administrative support}\\
-          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}
-        \item<3-> \alert{On--demand work}: crowd work and gig work, collectively
-        \item<5-> \alert{Piecework}: Payment for \emph{output} rather than for \emph{time}
-      \end{itemize}
 \end{frame}
 
-\begin{frame}{Ongoing Threads in Crowdsourcing Research}
-\begin{columns}
-\begin{column}[T]{0.5\textwidth}
-  \begin{enumerate}%[<+>]
-    \item<2,4> \textbf{\large Complexity}\\
-    \scriptsize{
-    \textcite{suzukiAtelier,KimStoria,yuanAlmost,Yu2016b,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
-    }
-    \vspace{5mm}
+% \begin{frame}{Why Piecework?}
+% What makes on--demand work another instantiation of piecework?
 
-    \item<3> \textbf{\large Decomposition}\\
-    \scriptsize{
-    \textcite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
-    }
-    \vspace{5mm}
-
-    \item<5> \textbf{\large Relationships}\\
-    \scriptsize{
-      \textcite{turkopticon,storiesIraniSilberman,crowdcollab,takingAHITMcInnis,dynamo,uberAlgorithm}
-    }
-
-  \end{enumerate}
-\end{column}
-
-\begin{column}[T]{0.25\textwidth}
-  \begin{figure}
-    \centering
-    % \vspace{-5mm}
-    \vspace{0mm}
-
-    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
-    \visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
-                  \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
-    }
-
-    \vspace{4mm}
-    \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
-    \vspace{5mm}
-
-    \visible<3->{
-    \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
-    }
-
-    \visible<5->{
-      \vspace{5mm}
-      \includegraphics[width=\textwidth]{figures/networkTransparency.png}
-    }
-  \end{figure}
-
-\end{column}
-
-\begin{column}[T]{0.25\textwidth}
-\centering
-\vspace{5mm}
-
-\visible<2->{
-  \small{Complexity}
-  \vspace{13mm}
-}
-
-\small{Tasks}
-
-\visible<3->{
-  \vspace{11mm}
-  \small{Decomposition}
-}
-
-\end{column}
-\end{columns}
-\end{frame}
+% \ali{piecework }
+% \end{frame}
 
@@
     What are the \alert{limits} of complexity in on--demand work?
-
-    How can we \alert{reach} those limits?
 \end{frame}
@@
 \begin{frame}{Thesis}
-    These questions have all been asked before.
+    This question --- and others like it --- has been asked before.
 
@@
 
-\againframe<4->{returnable}
-
 \notinsubfile{
-  \subfile{timeline.tex} % this takes a while to compile...
+  \subfile{timeline.tex}
 }
@@
       \begin{itemize}
-        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}.
-        \item A method for making sense of contemporary phenomena through \alert{historical analysis}.
+        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}
+        \item A method for making sense of contemporary phenomena through \alert{historical analysis}
       \end{itemize}
@@
     \item In general\\
-    \scriptsize{
-      \textcite{rosenberg1994exploring,rosenberg1982inside}
-    }
+    \scriptsize{\textcite{rosenberg1994exploring,rosenberg1982inside}}\normalsize{}
     \item In HCI\\
-    \scriptsize{
-      \textcite{Wyche2006,bodker1993historical}
-    }
+    \scriptsize{\textcite{Wyche2006,bodker1993historical}}\normalsize{}
   \end{itemize}

### 2017-05-10 chi/piecework/presentation/complexity.tex
checkin with most of the bolded/critical fixes resolved (see main page)

@@
 
-\section{Complexity}
+\subfile{threads}
+% \againframe<2>{threads}
 
 \begin{frame}{Complexity}
-  What do we mean when we talk about complexity?
-  % \begin{minipage}[t][25mm][t]{\textwidth}
-  %   \begin{columns}
-  %     \begin{column}[T]{0.8\textwidth}
+  \begin{columns}
+    \begin{column}{.6\textwidth}
+      \visible<-4>{What do we mean when we say \alert{complexity}?}
+    \end{column}
+    \begin{column}{.4\textwidth}
+      \only<5->{\hfill \dots~In some cases. Sometimes.}
+    \end{column}
+  \end{columns}
+  \begin{columns}
+  \begin{column}{.55\textwidth}
   \begin{itemize}
-          \item<1->
-          Can crowds help you write something?\\
-          \scriptsize{\textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}}\normalsize{}
-          \item<2-> Can crowds critique novel designs?\\
-          \scriptsize{\textcite{yuanAlmost,fuge2014analysis}}\normalsize{}
-          \item<3-> Can crowds create things from whole cloth?\\
-          \scriptsize{\textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
+    \item<2-> Can crowds help you \alert{revise} work?\par
+    \scriptsize{\textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}\par}\normalsize{}
+    \item<3-> Can crowds \alert{critique} designs?\par
+    \scriptsize{\textcite{yuanAlmost,fuge2014analysis}\par}\normalsize{}
+    \item<4-> Can crowds create artifacts \alert{de novo}?\par
+    \scriptsize{\textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}\par}\normalsize{}
   \end{itemize}
-      % \end{column}
-      % \begin{column}[T]{0.2\textwidth}
-      %   \begin{figure}
-    % \centering
-    % \vspace{0mm}
-
-    % \visible<1->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
-    % \visible<3->{
-    % \includegraphics<1->[width=.5\textwidth]{figures/complexity/paper_resized.png}
-    %               % \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
-    % % }
-
-    % \includegraphics<2->[width=.5\textwidth]{figures/complexity/chair_resized.png}
-
-    % \includegraphics[width=.15\textwidth]{figures/paper-unfolded.png}$\rightarrow$
-    % \includegraphics[width=.15\textwidth]{figures/rhino.png}
-    % \includegraphics[width=.15\textwidth]{figures/fish.png}
-    % \includegraphics[width=.15\textwidth]{figures/origami-1b.png}
-
-    % \vspace{4mm}
-    % \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
-    % \vspace{5mm}
-
-    % \visible<2->{
-    % \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
-    % }
-
-    % \visible<4->{
-    %   \vspace{5mm}
-    %   \includegraphics[width=\textwidth]{figures/networkTransparency.png}
-    % }
-  % \end{figure}
-  %     \end{column}
-  %   \end{columns}
-  %   \end{minipage}
+  \end{column}
+  \begin{column}{.45\textwidth}
+      \begin{itemize}
+        \item[$\Rightarrow$]<6-> \hfill With variable success\par
+        \visible<0>{\scriptsize{\textcite{bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169}\par}\normalsize{}}
+        \item[$\Rightarrow$]<7-> \hfill With careful guidance\par
+        \visible<0>{\scriptsize{\textcite{yuanAlmost,fuge2014analysis}\par}\normalsize{}}
+        \item[$\Rightarrow$]<8-> \hfill Within narrow specifications\par
+        \visible<0>{\scriptsize{\textcite{Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}\par}\normalsize{}}
+      \end{itemize}
+  \end{column}
+  \end{columns}
 \end{frame}
 
-\begin{frame}{What Does Crowdsourcing Say?}
+\begin{frame}{What Does On--Demand Work Say?}
   \begin{columns}
     \begin{column}{0.6\textwidth}
+      Build complexity into the process
       \begin{itemize}
-      \item Build complexity into the process
-      \begin{itemize}
-        \item<1> Apply CS methods to people\\
-        \scriptsize{\textcite{crowdForgeKittur}}\normalsize{}
-        \item<2> Humans as computational units\\
-        \scriptsize{\textcite{Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
-        \item<3> Crowdsourcing workflows as function state machines\\
-        \scriptsize{\textcite{latoza2014microtask}
-        }
-      \end{itemize}
+        \item<+-> Apply CS methods to people\par
+                 \scriptsize{\textcite{crowdForgeKittur}\par}\normalsize{}
+        \item<+-> Humans as computational units\par
+                 \scriptsize{\textcite{Lasecki:2014:LSR:2661334.2661352}\par}\normalsize{}
+        \item<+-> Crowdsourcing workflows as function state machines\par
+                 \scriptsize{\textcite{latoza2014microtask}\par}\normalsize{}
       \end{itemize}
@@
     \begin{column}{0.4\textwidth}
-    \begin{figure}
       \only<1>{
-    \includegraphics[width=\textwidth]{figures/complexity/cw_literature/mapReduce.png}
+      \includegraphics[max width=\textwidth,max height=\textheight]{figures/complexity/cw_literature/mapReduce.png}
       }
@@
       \only<2>{
-      \includegraphics[width=\textwidth]{figures/complexity/cw_literature/scribe.png}
+        \includegraphics[max width=\textwidth,max height=\textheight]{figures/complexity/cw_literature/scribe.png}
       }
@@
       }
-    \end{figure}
     \end{column}
-
   \end{columns}
@@
 \begin{frame}{What Does Piecework Say?}
-      George Airy (astronomer) used a very similar approach\\
-      \scriptsize{\textcite{grier2013computers}}\normalsize{}
+  What we'll find
+  \begin{itemize}
+    \item<+-> Building complexity into the processes
+    \item<+-> Incremental advances until managers \emph{tracked} and \emph{standardized} workers and work
+    \item<+-> Insights into task specialization
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[t]{What Does Piecework Say?}
+      George Airy (astronomer) used a very similar approach\par
+      \scriptsize{\textcite{grier2013computers}\par}\normalsize{}
     \begin{columns}
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
+        \includegraphics[max width=\textwidth, max height=.7\textheight,keepaspectratio]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
       \end{figure}
@@
 \begin{frame}{Comparisons}
+
 \begin{itemize}
-  \item Limited array of tasks versus arbitrarily complex work
+    \item<+-> Building complexity into the processes
+    % \item<+-> Incremental advances until managers \emph{tracked} and \emph{standardized} workers and work
+    % \item<+-> Insights into task specialization
+  % \end{itemize}
+% \begin{itemize}
+  \item Challenges dealing with flexibility
   \begin{itemize}
@@
   \end{itemize}
-  \item \normalsize{Has technology changed this?}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Implications for On--Demand Work}
+  Has technology shifted on--demand work?
     \begin{itemize}
@@
     \end{itemize}
-\end{itemize}
 \end{frame}
 
-\begin{frame}{Complexity} % {Cab Drivers}
+\begin{frame}{Enhanced Cognition} % {Cab Drivers}
   \begin{figure}
@@
 
-\begin{frame}{Complexity} % {Algorithmic Measurement}
-    notes
-    \begin{itemize}
-      \item I'm thinking of pointing to UpWork's screen recording tool as a way to measure workers
-      \item also maybe google analytics and other ways of tracking web--based workers
-    \end{itemize}
+\begin{frame}{Tracking Work and Workers} % {Algorithmic Measurement}
+    % notes
+    % \begin{itemize}
+      \ali{Upwork's screen recording tool as a way to measure workers}
+      % \item also maybe google analytics and other ways of tracking web--based workers
+    % \end{itemize}
 \end{frame}
 
-\begin{frame}{Implications}
+\begin{frame}{Takeaways}
   \begin{itemize}
@@
 
-\onlyinsubfile{
-  \printbibliography{}
-}
+% \onlyinsubfile{
+%   \printbibliography{}
+% }
 \end{document}

### 2017-05-10 chi/piecework/presentation/introduction.tex
checkin with most of the bolded/critical fixes resolved (see main page)

@@
 \onlyinsubfile{}
-\setbeamersize{description width=30mm}
+\setbeamersize{description width=8em}
 \begin{document}
-\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
+\begin{frame}<-3>[label=definitions]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
   \begin{description}
-        \item<1-> [Crowd work] Digitally mediated \alert{information work}
-        (e.g. Amazon Mechanical Turk, UpWork)\\
-          \scriptsize{\textcite{crowdworkFuture}}\normalsize{}
+    \item<1-> [Crowd work] Digitally mediated \alert{information work}, like 
+      \emph{image tagging}, \emph{audio transcription}, and \emph{data processing}\par
+      \scriptsize{\textcite{crowdworkFuture}\par}\normalsize{}
     \item<2-> [Gig work] Digitally mediated (often \alert{physically embodied}) one--off jobs,
@@
       \emph{courier services},
-        and \emph{administrative support}\\
-          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}\normalsize{}
-        \item<3-> [On--demand~work] Crowd work and gig work, collectively
-        \item<5-> [Piecework] Payment for \emph{output} rather than for \emph{time}
+      and \emph{administrative support}\par
+      \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}\par}\normalsize{}
+    \item<3-> [On--demand work] Crowd work and gig work, collectively
+  \end{description}
+  \visible<5->{
+     \begin{center}
+        {\color{mLightBrown}
+          {\begin{tikzpicture}
+            \node  (C) at (0,0) {};
+            \node (D) at (9,0) {};
+            \path (C) to [ornament=86] (D);
+          \end{tikzpicture}}}%
+    \end{center}
+   }
+  \begin{description}
+    \item<6-> [\textrm{Piecework}] \textrm{Payment for \emph{output} rather than for \emph{time}}
   \end{description}
@@
 
-\begin{frame}[standout]
+\begin{frame}[standout,label=takeaway]{}
   On--demand work is a modern instantiation of a much older phenomenon
@@
 
-\againframe<4->{returnable}
+\againframe<4->{definitions}
 
-\begin{frame}[t]{What is piecework?}
-\centering
-  \emph{Payment for \textbf{output} rather than for \textbf{time}}
-\vspace*{2mm}
-\begin{columns}
+\begin{frame}[t]{Payment for \emph{output} rather than for \emph{time}}
+  \begin{columns}[T]
     \begin{column}{0.33\textwidth}
       \centering
-    Textiles
-    \begin{figure}
-    \begin{overlayarea}{\textwidth}{33mm}
-    \begin{minipage}[c][33mm]{\textwidth}
-      \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
-    \end{minipage}
-    \end{overlayarea}
-    \end{figure}
+      Textiles\\
+      \includegraphics[max width=\linewidth,max height=.3\textheight,keepaspectratio]{figures/pieceworkers.jpg}
     \end{column}
@@
       \centering
-    Automobiles
-    \begin{figure}
-    \begin{overlayarea}{\textwidth}{33mm}
-    \begin{minipage}[c][33mm]{\textwidth}
-      \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
-    \end{minipage}
-    \end{overlayarea}
-    \end{figure}
+      Automobiles\\
+      \includegraphics[max width=\linewidth,max height=.3\textheight,keepaspectratio]{figures/photo/ford_assembly_line.jpg}
     \end{column}
@@
       \centering
-    Metalwork
-    \begin{figure}
-    \begin{overlayarea}{\textwidth}{33mm}
-    \begin{minipage}[c][33mm]{\textwidth}
-      \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
-    \end{minipage}
-    \end{overlayarea}
-    \end{figure}
+      Metalwork\\
+      \includegraphics[max width=\linewidth,max height=.3\textheight,keepaspectratio]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
     \end{column}
   \end{columns}
+  \vspace*{7mm}
 
   \visible<2->{
-  \begin{columns}
-  \centering
-    \begin{column}{0.33\textwidth}
+    \begin{columns}[b]
+      \begin{column}[t]{0.45\textwidth}
         \centering
         Crowd work
-    \begin{figure}
-      \begin{overlayarea}{\textwidth}{15mm}
-      \begin{minipage}[c][15mm]{\textwidth}
-        \includegraphics[width=\textwidth]{figures/amt.png}
-      \end{minipage}
-      \end{overlayarea}
-    \end{figure}
+        \begin{columns}
+          \begin{column}[b]{.5\textwidth}
+            \includegraphics[max width=\linewidth,max height=\textheight,keepaspectratio]{figures/amt.png}
           \end{column}
-    \begin{column}{0.33\textwidth}
+          \begin{column}[b]{.5\textwidth}
+            \includegraphics[max width=\linewidth,max height=\textheight,keepaspectratio]{figures/upwork.png}
+          \end{column}
+        \end{columns}
+      \end{column}
+
+      \begin{column}[t]{0.45\textwidth}
         \centering
         Gig Work
-    \begin{figure}
-    \begin{overlayarea}{\textwidth}{15mm}
-      \begin{minipage}[c][15mm]{\textwidth}
-        \includegraphics[width=\textwidth]{figures/uber.png}
-      \end{minipage}
-      \end{overlayarea}
-    \end{figure}
+        \begin{columns}
+          \begin{column}{.5\textwidth}
+            \includegraphics[max width=\linewidth,max height=\textheight,keepaspectratio]{figures/uber.png}
+          \end{column}
+          \begin{column}{.5\textwidth}
+            \includegraphics[max width=\linewidth,max height=\textheight,keepaspectratio]{figures/postmates.png}
+          \end{column}
+        \end{columns}
       \end{column}
@@
   }
-
 \end{frame}
-
-% \begin{frame}{Why Piecework?}
-% What makes on--demand work another instantiation of piecework?
-
-% \ali{piecework }
-% \end{frame}
 
@@
 \begin{frame}{What will be the future of work?}
-    How will \alert{technology} affect the complexity of the work that on--demand workers do?
-
-    What are the \alert{limits} of complexity in on--demand work?
+  \begin{columns}
+    \begin{column}{0.6\textwidth}
+      \begin{itemize}
+        \item<1-> How will \alert{technology} affect the complexity of the work that on--demand workers do?
+        \item<2-> What are the \alert{limits} of complexity in on--demand work?
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.4\textwidth}
+    \only<3>{
+      \centering
+      The answers to these questions may predict the \emph{reach} of on--demand work
+    }
+    \end{column}
+    \end{columns}
 \end{frame}
@@
 
-\notinsubfile{
-  \subfile{timeline.tex}
-}
-
 \begin{frame}{Introduction}
-  We hope to provide:
+  We hope you come away with:
       \begin{itemize}
-        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}
-        \item A method for making sense of contemporary phenomena through \alert{historical analysis}
+        \item An \alert{ontological lens} for making sense of on--demand work as a resurgence of piecework
+        \item A renewed interest in the use of \alert{historical analysis} to make sense of contemporary phenomena
       \end{itemize}
@@
 \begin{frame}{Comparative Historical Analysis}
-\begin{itemize}
-  \item Historical analysis isn't new
-  \begin{itemize}
-    \item In general\\
-    \scriptsize{\textcite{rosenberg1994exploring,rosenberg1982inside}}\normalsize{}
-    \item In HCI\\
-    \scriptsize{\textcite{Wyche2006,bodker1993historical}}\normalsize{}
-  \end{itemize}
-  \item Still, it's an underutilized method
+  HCI researchers have used historical analysis in the past\par
+  \scriptsize{\textcite{Wyche2006,bodker1993historical}\par}\normalsize{}
+
+  \dots~But we haven't applied this method to make sense of on--demand work,
+        which is a missed opportunity to\dots
   \begin{itemize}
@@
   \end{itemize}
-\end{itemize}
+% \end{itemize}
 \end{frame}
 
-\onlyinsubfile{
-  \printbibliography{}
+\notinsubfile{
+  \subfile{timeline.tex}
 }

### 2017-05-10 chi/piecework/presentation/presentation.tex
checkin presentation
checkin with most of the bolded/critical fixes resolved (see main page)

@@
-\documentclass{beamer}
+\documentclass[aspectratio=169,12pt]{beamer} % or 14 or 17 or 20
 
-\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
-\usepackage{pgfcalendar}
+\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer}
+\usepackage{pgfcalendar,pgfornament,svg,subfiles,chronosys,cancel,xcolor,color,nth}
+\usepackage[outline]{contour}
 \usepackage{datenumber,xparse,fp}
-
 \usetikzlibrary{arrows,automata}
@@
 \usepackage[en-US]{datetime2}
-
 \bibliography{references}
 \usetheme{metropolis}
+% \linespread{1}
+% \usecolortheme{metropolis}
 \beamertemplatenavigationsymbolsempty
@@
 
-% Recent Notes from msb
-% [ ] I suggest that you come out with a one sentence summary of your main thesis after 2 and before 3.
-“What I’m going to argue in this talk is that on-demand work is a modern instantiation of a much older phenomenon, piecework [put the word on the slide], and that we’re going to be able to use the history of piecework over the past two hundred years to shed light on several unanswered questions about the unfolding phenomenon of on-demand work.”
-% [ ] This way people know what you’re winding up for in this talk
-% [ ] don’t get rid of slide 5. I’m just saying put one framing sentence up front here.
-% [ ] 3: these are not the canonical cites for complexity. kittur crowdforge is most famous
-% [ ] 3: put the task hexagons at the top, not in between. the hierarchy I see what you’re trying to do, but it’s not really working
-% [ ] When you introduce the piecework term, give several examples (with pictures) that listeners will be familiar with.
-% [ ] You never stopped to equate piecework to on demand work. Why does piecework = on demand work? (in particular, because on demand work ALSO pays people per task)
-% [ ] 3: your complexity example is being built out of triangles, not hexagons
-% [ ] Get rid of the citation numbers, and replace it with years Alkhatib 2017, not Alkhatib [1]
-% [ ] 3: I think this isn’t focused correctly.
-I’d get rid of this slide and ONLY motivate the complexity question.
-Then mention briefly that in the paper you also talk about decomposition and worker relationships, but you need 10 pages and 166 references rather than a 15 minute talk to do all that.
-Return to these other themes at the end of the talk and mention some short take aways from each one.
-% [ ] 4: doesn’t make sense to ask “how will we reach those limits?”
-% [ ] LOL at the phone ringing
-% [ ] 7: the top timeline is events, the bottom is citations. change the bottom one to be events
-% [ ] 8: this is WAY too late to still be doing an intro. You’re five minutes into the talk. Intro should be done by 2min.
-% [ ] 12: the details here are irrelevant to complexity.
-why does it matter that they could be fired at will, overworked, underpaid?
-shouldn’t we be focusing on how piecework enabled the complexity?
-If it’s not important (as you say), then don’t include it.
-Parsimony, yay!
-% [ ] 12: another example of a citation I can’t follow by watching slides
-% [ ] 14: ditch the shrugging guy, he’s really weird
-% [ ] 15: trains — this has nothing to do with complexity…? cut?
-% [ ] 15: fordism — again, explain what this has to do with complexity? what does fordism have to do with complex outcomes? spell it out. what should we be learning from this?
-% [ ] 16: technology makes complex tasks relatively trivial — this is an overly broad claim. makes SOME?
-% [ ] 17: speed up The Knowledge/Gmaps example. I think it can go twice as fast and still carry as much weight
+% notes from last practice talk (5/9 morning):
+% [x] spacing in the definitions — piecework should be more separated from the other definitions to make clearer that it's different
+%      tried something different. now it's Source Serif Pro. Thoughts?
+% [x] space the crowdwork and gigwork images (uber and AMT) further down to separate them from the piecework images
+% [x] move the timeline to after the historical analysis slide
+% [x] return to the standout slide at the end as a way of wrapping up
+% [x] on that note; instead of the slide that just says "complexity", maybe take the slide with the three panels (comp, decomp, relationships) and put that in its place, and then have a transition where i gray out decomp and relationships.
+then at the end be like "there's a lot more in the paper, like decomposition and relationships, but we don't have time to get deeply into them.
+i encourage you to read the paper if you found this talk provocative [or something], as we follow a similar style for the other discussions"
+% [x] there's a flickering thing that goes on in the pdf viewer i used that is super obnoxious.
+some other pdf viewers are less bad about it.
+i'm gonna trial a few viewers since i just have to go to fullscreen and click next
+
+% [ ] slide 8: title="introduction" not good.
+i think i need to move all of this stuff around so this slide is earlier and i can arguably say that the timeline (slide 7) is getting into the body of the talk (all of the events correspond to advances in *complexity*, not just crowd/piecework)
+% [ ] slide 10: i can give specific examples; "soylent _revised_ academic papers; crowd research _generates new papers_" is kind of salient 
+% [x] also slide 10: "from whole cloth" is both idiomatic and possibly the wrong idiom, so i'm just going to change it to something like "completely anew" or something 
+% [ ] generally: need to figure out whether to keep the origami birds.
+they look nice but the reference to breaking tasks down into triangles and those triangles making it possible to make more complex, nuanced things is totally lost.
+it feels more like an inside joke that the audience won't get
+% [x] reference/explain the chart on slide 13, because i don't really explicitly reference it. it's just an image showing several people doing work, contributing to and verifying each other's work 
+% [x] slide 14: farm workers need an image (mentioned before) 
+% [ ] slide 16 felt super confused and foggy. maybe it was just too abrupt. i'll try to practice my transition to it 
+% [ ] slide 17: i'm dropping figure B and maybe the tiny image on the left. sound okay? 
+% [ ] slide 18: needs real content and not just internal notes 
+% [x] slide 20: i actually left out the relationships panel. also no visualizations. thoughts? 
+% [x] slide 21: it's not twofold if i have 3 things -_- 
+% [ ] slide 22: "thanks and questions?"
+
+% more of MSB's notes
+% !!!!!!
+[x] spend a little more time motivating why the complexity question is important.
+If we knew the answer to this, we would be able to predict what kind of worker displacements might happen, build appropriate infrastructure, create strong worker advocate groups, etc.
+% !!!!!!
+[x] 11: I think this needs to be bounded.
+you need to argue that in fact NOT ALL complex stuff has been achieved.
+e.g., the three examples you gave, while we’ve made progress, are basically still open questions.
+and are they fundamentally impossible, or just facing a lack of engineering imagination?
+that’s what we turn to piecework to answer.
+<— important transition, it frames the question
+% !!!!!!
+[x] I need to hear the argument that you’re going to make about what piecework has to say about complexity, before you dive into the “ hat does piecework say?" history.
+otherwise it feels like a wandering tour rather than an argument.
+% !!!!!!
+[ ] Slide 20: instead of just handwaving that we talk about it, summarize the big takeaways from your Table 1.
+e.g., "Decomposition: After scientific management matured, piecework began specialized training to create experts in narrow tasks.
+A similar shift seems feasible with on–demand work.” and “Relationships: While worker frustrations are similar, the decentralized nature of on–demand work will limit collective action until there exist platforms to coordinate and exert pressure.”
+% [x] Don’t introduce your name and title; the session chair will do that for you.
+Make sure to show up to the room in the break before your session (at least 15min early) to test your laptop and introduce yourself to the session chair
+% [x] 2: Upwork, not UpWork
+% [x] 2: top bullet is services, bottom is categories. make them consistent
+% [x] Intro is way better now. congrats!
+% [ ] On final slide, thank Stanford Cyber Initiative for funding
+% [ ] First half of Slide 8 seems redundant. Cut? And just focus on the second half.
+% [ ] 10: wrap this together — essentially you need to say “so this reduces to a question of what can, and what can’t be crowdsourced.
+What are the limits, if any?” and put something like that on the slide
+% [x] 11: inconsistent language use in titles, for example “crowdsorucing” instead of “on-demand work” in the title? unless this is intentional
+% [ ] (by the time you get to automobiles, I don’t know why you’re telling me what you’re telling me)
+% [ ] 13: not clear how ascention and descention of stars is “in the process”.
+I suggest pointing out instead that he reduced it to simple addition etc calculations that people passed between each other
+% [ ] I don’t know what “marginal complexity” means. Either explain it or change the term
+% [ ] 15: change “values” to “payment”…clearer
+% [ ] 16 or after: make clear that technology hasn’t fundamentally changed the relationship.
+we can still do the equivalent of building planes, and still can’t do the equivalent of fixing planes (things that require breadth of knowledge and flexibility)
+% [x] 17: don’t spend time laying out the dual hypotheses.
+just say that clearly this is either causing changes in the brain, or selecting for a small subset of people (in one sentence) — either way not a complex task that just anyone can do
+% [ ] 21: be clear on the slide bullet three that this is about History as a lens on contemporary phenomena. right now it’s not clear
 
@@
 \setbeamercolor{alerted text}{fg=Orange}
-\setsansfont[BoldFont={Source Sans Pro Semibold},
+\setsansfont[BoldFont={Source Sans Pro Bold},
               Numbers={OldStyle}]{Source Sans Pro}
+\setmainfont[BoldFont={Source Serif Pro Semibold},
+              Numbers={OldStyle}]{Source Serif Pro}
 \setmonofont{Source Code Pro}
@@
 % \metroset{titleformat=smallcaps}
-
+\contourlength{0.1pt}
+\contournumber{10}
 \author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
 \texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
-         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}} }}
+         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}\normalsize{} }}
 
@@
 \subfile{closing}
+\againframe[standout]{takeaway}
 \subfile{contact}
 
-% \begin{frame}[allowframebreaks]{References}
-% \printbibliography{}
-% \end{frame}
+\begin{frame}[allowframebreaks]{References}
+  \printbibliography{}
+\end{frame}
 

### 2017-05-10 chi/piecework/presentation/complexity.tex
checkin wednesday morning

@@
   \begin{columns}
-    \begin{column}{0.6\textwidth}
+    \begin{column}{0.5\textwidth}
       Build complexity into the process
@@
 
-    \begin{column}{0.4\textwidth}
+    \begin{column}{0.5\textwidth}
       \only<1>{
-      \includegraphics[max width=\textwidth,max height=\textheight]{figures/complexity/cw_literature/mapReduce.png}
+      \includegraphics[max width=\textwidth,max height=.9\textheight]{figures/complexity/cw_literature/mapReduce.png}
       }
@@
       \only<2>{
-        \includegraphics[max width=\textwidth,max height=\textheight]{figures/complexity/cw_literature/scribe.png}
+        \includegraphics[max width=\textwidth,max height=.7\textheight]{figures/complexity/cw_literature/scribe.png}
       }
       \only<3>{
-        \begin{adjustbox}{max totalsize={\textwidth}{\textheight},center}
+        \begin{adjustbox}{max totalsize={\textwidth}{.9\textheight},center}
         \begin{tikzpicture}[->,>=stealth',auto, node distance=4cm,
@@
   \only<2>{
-  \begin{columns}
-  \begin{column}{0.2\textwidth}
-  \includegraphics[width=\textwidth]{figures/photo/knowledge-student.jpg}
-  \end{column}
-  \begin{column}{0.8\textwidth}
-  \includegraphics[width=\textwidth]{figures/complexity/brain_scans.png}
-  \end{column}
-  \end{columns}
+  % \begin{columns}
+  % \begin{column}{0.2\textwidth}
+  % \includegraphics[width=\textwidth]{figures/photo/knowledge-student.jpg}
+  % \end{column}
+  % \begin{column}{0.8\textwidth}
+  \includegraphics[width=\textwidth]{figures/complexity/brain_scansA.png}
+  % \end{column}
+  % \end{columns}
   }
@@
 \begin{frame}{Tracking Work and Workers} % {Algorithmic Measurement}
+\begin{columns}
+\begin{column}{.5\textwidth}
+\includegraphics[max width=\textwidth,max height=.8\textheight,keepaspectratio]{figures/screenshot_odeskteamapp.jpg}
+\end{column}
+\begin{column}{.5\textwidth}
+% \begin{itemize}
+  Upwork has turned to logging workers' keystrokes and taking screenshots automatically every 10 minutes
+% \end{itemize}
+% \includegraphics[max width=\textwidth,max height=.8\textheight,keepaspectratio]{figures/time-tracking-like-odesk-screenshot-monitor1.png}
+\end{column}
+\end{columns}
+
     % notes
     % \begin{itemize}
-      \ali{Upwork's screen recording tool as a way to measure workers}
+      % \ali{Upwork's screen recording tool as a way to measure workers}
       % \item also maybe google analytics and other ways of tracking web--based workers

### 2017-05-10 chi/piecework/presentation/complexity.tex
final talk!?

@@
     Airy built complexity into the process, assigning \emph{human computers} 
-    to calculate \& verify the \emph{right ascension} and \emph{declination} of stars.
+    to calculate astronomical movements.
+    % to calculate \& verify the \emph{right ascension} and \emph{declination} of stars.
 
@@
   Has technology shifted on--demand work?
+
+  In some ways, yes
         \begin{itemize}
@@
         \end{itemize}
+  In other ways, no --- we still don't have good end--to--end processes for arbitrarily complex work
+
+  \emph{We can make a routine out of building planes, but not out of fixing trains}
 \end{frame}

### 2017-05-10 chi/piecework/presentation/introduction.tex
final talk!?

@@
 
-\begin{frame}{Introduction}
-  We hope you come away with:
-      \begin{itemize}
-        \item An \alert{ontological lens} for making sense of on--demand work as a resurgence of piecework
-        \item A renewed interest in the use of \alert{historical analysis} to make sense of contemporary phenomena
-      \end{itemize}
-\end{frame}
+% \begin{frame}{Introduction}
+  % To take away:
+      % \begin{itemize}
+        % \item An \alert{ontological lens} for making sense of on--demand work as a resurgence of piecework
+        % \item A renewed interest in the use of \alert{historical analysis} to make sense of contemporary phenomena
+      % \end{itemize}
+% \end{frame}
 
 \begin{frame}{Comparative Historical Analysis}
-  HCI researchers have used historical analysis in the past\par
+  HCI researchers have used \alert{historical analysis} to understand social systems before\par
   \scriptsize{\textcite{Wyche2006,bodker1993historical}\par}\normalsize{}

### 2017-05-10 chi/piecework/presentation/presentation.tex
checkin wednesday morning
final talk!?
minor fixes

@@
-\documentclass[aspectratio=169,12pt]{beamer} % or 14 or 17 or 20
+\documentclass[aspectratio=169,12pt,noslidenumbers]{beamer} % or 14 or 17 or 20
 
@@
 % [x] also slide 10: "from whole cloth" is both idiomatic and possibly the wrong idiom, so i'm just going to change it to something like "completely anew" or something 
-% [ ] generally: need to figure out whether to keep the origami birds.
+% [?] generally: need to figure out whether to keep the origami birds.
 they look nice but the reference to breaking tasks down into triangles and those triangles making it possible to make more complex, nuanced things is totally lost.
@@
 % !!!!!!
-[ ] Slide 20: instead of just handwaving that we talk about it, summarize the big takeaways from your Table 1.
+[x] Slide 20: instead of just handwaving that we talk about it, summarize the big takeaways from your Table 1.
 e.g., "Decomposition: After scientific management matured, piecework began specialized training to create experts in narrow tasks.
@@
 % [x] Intro is way better now. congrats!
-% [ ] On final slide, thank Stanford Cyber Initiative for funding
-% [ ] First half of Slide 8 seems redundant. Cut? And just focus on the second half.
-% [ ] 10: wrap this together — essentially you need to say “so this reduces to a question of what can, and what can’t be crowdsourced.
+% [x] On final slide, thank Stanford Cyber Initiative for funding
+% [x] First half of Slide 8 seems redundant. Cut? And just focus on the second half.
+% [x] 10: wrap this together — essentially you need to say “so this reduces to a question of what can, and what can’t be crowdsourced.
 What are the limits, if any?” and put something like that on the slide
 % [x] 11: inconsistent language use in titles, for example “crowdsorucing” instead of “on-demand work” in the title? unless this is intentional
-% [ ] (by the time you get to automobiles, I don’t know why you’re telling me what you’re telling me)
-% [ ] 13: not clear how ascention and descention of stars is “in the process”.
+% [x] (by the time you get to automobiles, I don’t know why you’re telling me what you’re telling me)
+% [x] 13: not clear how ascention and descention of stars is “in the process”.
 I suggest pointing out instead that he reduced it to simple addition etc calculations that people passed between each other
-% [ ] I don’t know what “marginal complexity” means. Either explain it or change the term
-% [ ] 15: change “values” to “payment”…clearer
-% [ ] 16 or after: make clear that technology hasn’t fundamentally changed the relationship.
+% [x] I don’t know what “marginal complexity” means. Either explain it or change the term
+% [x] 15: change “values” to “payment”…clearer
+% [x] 16 or after: make clear that technology hasn’t fundamentally changed the relationship.
 we can still do the equivalent of building planes, and still can’t do the equivalent of fixing planes (things that require breadth of knowledge and flexibility)
@@
 \contournumber{10}
-\author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
-\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
-         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}\normalsize{} }}
+\author{{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
+\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
+         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}
 
@@
 
-\begin{frame}[allowframebreaks]{References}
-  \printbibliography{}
-\end{frame}
+% \begin{frame}[allowframebreaks]{References}
+%   \printbibliography{}
+% \end{frame}
 

### 2017-05-12 chi/piecework/presentation/presentation.tex
fixing references
fixed talk date to actual given date!

--- 2017-05-06 chi/piecework/presentation/presentation.tex
@@
-\documentclass{beamer}
+\documentclass[aspectratio=169,12pt,noslidenumbers]{beamer} % or 14 or 17 or 20
 
-\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
-\usepackage{pgfcalendar}
+\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer}
+\usepackage{pgfcalendar,pgfornament,svg,subfiles,chronosys,cancel,xcolor,color,nth}
+\usepackage[outline]{contour}
 \usepackage{datenumber,xparse,fp}
-
 \usetikzlibrary{arrows,automata}
@@
 \usepackage[en-US]{datetime2}
-
 \bibliography{references}
@@
 \definecolor{metroBackground}{RGB}{250,250,250}
-
-% Recent Notes from msb
-% [ ] I suggest that you come out with a one sentence summary of your main thesis after 2 and before 3.
-“What I’m going to argue in this talk is that on-demand work is a modern instantiation of a much older phenomenon, piecework [put the word on the slide], and that we’re going to be able to use the history of piecework over the past two hundred years to shed light on several unanswered questions about the unfolding phenomenon of on-demand work.”
-% [ ] This way people know what you’re winding up for in this talk
-% [ ] don’t get rid of slide 5. I’m just saying put one framing sentence up front here.
-% [ ] 3: these are not the canonical cites for complexity. kittur crowdforge is most famous
-% [ ] 3: put the task hexagons at the top, not in between. the hierarchy I see what you’re trying to do, but it’s not really working
-% [ ] When you introduce the piecework term, give several examples (with pictures) that listeners will be familiar with.
-% [ ] You never stopped to equate piecework to on demand work. Why does piecework = on demand work? (in particular, because on demand work ALSO pays people per task)
-% [ ] 3: your complexity example is being built out of triangles, not hexagons
-% [ ] Get rid of the citation numbers, and replace it with years Alkhatib 2017, not Alkhatib [1]
-% [ ] 3: I think this isn’t focused correctly.
-I’d get rid of this slide and ONLY motivate the complexity question.
-Then mention briefly that in the paper you also talk about decomposition and worker relationships, but you need 10 pages and 166 references rather than a 15 minute talk to do all that.
-Return to these other themes at the end of the talk and mention some short take aways from each one.
-% [ ] 4: doesn’t make sense to ask “how will we reach those limits?”
-% [ ] LOL at the phone ringing
-% [ ] 7: the top timeline is events, the bottom is citations. change the bottom one to be events
-% [ ] 8: this is WAY too late to still be doing an intro. You’re five minutes into the talk. Intro should be done by 2min.
-% [ ] 12: the details here are irrelevant to complexity.
-why does it matter that they could be fired at will, overworked, underpaid?
-shouldn’t we be focusing on how piecework enabled the complexity?
-If it’s not important (as you say), then don’t include it.
-Parsimony, yay!
-% [ ] 12: another example of a citation I can’t follow by watching slides
-% [ ] 14: ditch the shrugging guy, he’s really weird
-% [ ] 15: trains — this has nothing to do with complexity…? cut?
-% [ ] 15: fordism — again, explain what this has to do with complexity? what does fordism have to do with complex outcomes? spell it out. what should we be learning from this?
-% [ ] 16: technology makes complex tasks relatively trivial — this is an overly broad claim. makes SOME?
-% [ ] 17: speed up The Knowledge/Gmaps example. I think it can go twice as fast and still carry as much weight
 
@@
 % \metroset{titleformat=smallcaps}
-
-\author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
-\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
-         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}} }}
+\contourlength{0.1pt}
+\contournumber{10}
+\author{{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
+\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
+         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}
 
 \institute[Stanford]{Stanford University}
-\date{\today}
+% \newdate{talkDate}{10}{05}{2017}
+\date{May 10, 2017}
+% \date{\talkDate}
 

### 2017-07-09 admin/IRB/performances/consent_waiver/main.tex
added acmart (and other stuff?)

--- 2016-11-30 admin/IRB/performances/assent/main.tex
@@
 \documentclass[11pt]{article}
-\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp,titling,nameref,color,fancyhdr,paralist}
+\usepackage{balance,graphics,setspace,parskip,textcomp,titling,nameref,color,fancyhdr,paralist}
 \usepackage[T1]{fontenc}
@@
 \usepackage[compact]{titlesec}
-\PassOptionsToPackage{hyphens}{url}\usepackage[pdflang={en-US},pdftex]{hyperref}
+% \PassOptionsToPackage{hyphens}{url}\usepackage[pdflang={en-US},pdftex]{hyperref}
 
 \definecolor{linkColor}{RGB}{6,125,233}
-\hypersetup{
-  bookmarksnumbered,
-  pdfstartview={FitH},
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=blue,
-  urlcolor=red,
-  breaklinks=true,
-  hypertexnames=false
-}
+% \hypersetup{
+%   bookmarksnumbered,
+%   pdfstartview={FitH},
+%   colorlinks,
+%   citecolor=black,
+%   filecolor=black,
+%   linkcolor=blue,
+%   urlcolor=red,
+%   breaklinks=true,
+%   hypertexnames=false
+% }
 \usepackage[dvipsnames]{xcolor}

### 2017-07-09 admin/IRB/vitae/consent/main.tex
added acmart (and other stuff?)

--- 2016-11-30 admin/IRB/performances/assent/main.tex
@@
-\documentclass[11pt]{article}
+\documentclass[10pt]{article}
 \usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp,titling,nameref,color,fancyhdr,paralist}
@@
 \usepackage[compact]{titlesec}
-\PassOptionsToPackage{hyphens}{url}\usepackage[pdflang={en-US},pdftex]{hyperref}
 
+\setsansfont[BoldFont={Source Sans Pro Bold},
+              Numbers={OldStyle}]{Source Sans Pro}
+\setmainfont[BoldFont={Source Serif Pro Semibold},
+              Numbers={OldStyle}]{Source Serif Pro}
+\setmonofont{Source Code Pro}
 \definecolor{linkColor}{RGB}{6,125,233}
-\hypersetup{
-  bookmarksnumbered,
-  pdfstartview={FitH},
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=blue,
-  urlcolor=red,
-  breaklinks=true,
-  hypertexnames=false
-}
+
 \usepackage[dvipsnames]{xcolor}
@@
   \chead{\begin{headerfont}\textbf{\underline{Waiver of Documentation}}\end{headerfont}}
-  \rhead{\textbf{Approval Date}: \underline{November 30 2016}}
+  \rhead{\textbf{Approval Date}: \underline{Tentative}}
 }
 
-\section*{{\LARGE Ethnographic Study of Online Performance Artists}}
+\section*{{\LARGE Managing crowd and gig workers' histories and qualifications}}
 
@@
 We're trying to learn about
-the relationships online performers have with each other and
-with the
-organizations that run
-the platforms that they stream and
-upload to.
-Some of this touches on things like
-dealing with inter--personal issues and
-how the community deals with major events
-as well as questions about what it's like to
-make your income from online public performances.
+whether independently--run platforms for tracking workers' histories
+and qualifications
+can better inform whether workers are likely
+to be qualified and suited for complex gig work.
+Some of this research will involve
+the management of workers \textbf{by} other workers.
+
+In this study, we'll be asking workers managed by this system about
+their experiences using the platform (``vitae''), as well as
+interviewing people who manage these qualification groups.
+% the relationships online performers have with each other and
+% with the
+% organizations that run
+% the platforms that they stream and
+% upload to.
+% Some of this touches on things like
+% dealing with inter--personal issues and
+% how the community deals with major events
+% as well as questions about what it's like to
+% make your income from online public performances.
 
@@
 \begin{compactitem}
+  \item Participation is voluntary; if you wish not to participate, it's perfectly all right.
+  \item Participants will not be compensated for their interviewing time.
+  \item If at any point you wish to withdraw from the study,
+        we will accommodate such requests as quickly as reasonably possible.
   % \item Since it would be too difficult to get a signature from participants,
@@
   %       so this is purely informational.
-  \item The first conversation will probably take 30--45 minutes.
-        Later conversations (if you're okay with it) would likely be shorter
-        (mostly to clarify things or ask follow--up questions).
-  \item I can't compensate you (or anyone, really) for their time.
-  \item I plan on recording the audio of our conversation
-        so I can transcribe what you said
-        (so I don't misremember or misquote anything).
-        Once I have a transcript, I'll delete the audio.
+  % \item The first conversation will probably take 30--45 minutes.
+  %       Later conversations (if you're okay with it) would likely be shorter
+  %       (mostly to clarify things or ask follow--up questions).
+  % \item I can't compensate you (or anyone, really) for their time.
+  % \item I plan on recording the audio of our conversation
+  %       so I can transcribe what you said
+  %       (so I don't misremember or misquote anything).
+  %       Once I have a transcript, I'll delete the audio.
 \end{compactitem}

### 2017-07-18 proposals/mediaX/proposal.tex
renamed to proposal.tex/pdf/whatever

\documentclass[10pt]{article}
\usepackage{xcolor,enumitem,nth,titlesec,xr-hyper,balance,wrapfig,fancyhdr,fontspec,url,graphicx}
\usepackage[normalem]{ulem}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only
\usepackage[T1]{fontenc}

\usepackage[en-US]{datetime2}
% \DTMsettimestyle{default}
% \DTMsetup{showseconds=true}

\setmainfont{SourceSerifPro}[
 ItalicFont = SourceSerifPro,
 ItalicFeatures={FakeSlant=0.2}]

\bibliography{references}
\usepackage[margin=1in]{geometry}
\pagestyle{fancy}
% \titlespacing*{\section}{0.0ex}{2ex plus .3ex}{1ex plus .3ex}

\lhead{} % controls the left corner of the header
% \chead{} % controls the center of the header
\rhead{} % controls the right corner of the header
\lfoot{last updated \DTMcurrenttime~\today} % controls the left corner of the footer
\cfoot{} % controls the center of the footer
\rfoot{} % controls the right corner of the footer
\rfoot{Page~\thepage} % controls the right corner of the footer
% \renewcommand{\headrulewidth}{0.0pt}
% \renewcommand{\footrulewidth}{0.0pt}

% \AtBeginBibliography{\small}
\thispagestyle{empty}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
  {\scshape \Large \textbf{\@title}}

  \vspace{1ex}

  \@author
\end{flushleft}\egroup
}
\makeatother

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{Black}{HTML}{000000}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\title{Vitae: Aggregated Work Histories for On--Demand Workers}
\author{Ali Alkhatib \& Michael Bernstein}
\date{\today}

\urlstyle{same}

\hypersetup{
  colorlinks,
  citecolor=Black,
  urlcolor=Blue}

\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}}
\renewcommand{\topic}[1]{{#1}}
\definecolor{Gray}{HTML}{A0A0A0}
\newcommand{\ugh}[1]{{\color{Gray}#1}}
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

\renewenvironment{abstract}{%
\hfill\begin{minipage}{0.95\textwidth}
\itshape
}
{
\bigskip
\end{minipage}}

\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!

\begin{document}
  \maketitle

\begin{abstract}
todo
\end{abstract}

% \section*{Outline}
% \begin{enumerate}
%   \item Frame the problem
%   \item What are the intuitive solutions
%   \item Why is that not working?
%   \item What's our insight?
%   \item What are we building
%   \item Why is this important
%   \item What do we need
% \end{enumerate}

\section*{State of the world}
\begin{enumerate}[noitemsep]
  \item Research is showing that both the \textit{breadth} and \textit{depth} (or \textit{complexity})
        of on--demand labor will continue to grow
  \item More people will be doing more kinds of work, transiently moving between myriad platforms
  \item People's educations are becoming increasingly decomposed;
  online education platforms make it possible for people to learn and acquire new skills piecemeal,
  and demonstrate proficiency on ad hoc bases
\end{enumerate}

\section*{Trouble!}
\begin{enumerate}[noitemsep]
  \item The people who enter the workforce in this capacity are increasingly finding that
        this form of work doesn't facilitate the building of more complex, more rewarding work
  \item tracking and \textit{bringing together} the histories of these workers is difficult
  \begin{enumerate}[noitemsep]
    \item it's hard to trust individual workers
    \item platforms don't make it any easier to \textit{amalgamate} data
  \end{enumerate}
\end{enumerate}

\section*{Intuitive solutions (and why they fail)}
\begin{enumerate}[noitemsep]
  \item One solution seems to be to maintain a list of known good workers --- ``whitelisting'' good workers
  \begin{enumerate}[noitemsep]
    \item This generally suffices for large--scale requesters; they form a cohort of known good workers, and
          can (in general) afford to ``explore'' and discover new workers
    \item This solution doesn't work for smaller scale requesters:
    \begin{enumerate}[noitemsep]
      \item Existing requesters don't have much (read: any) incentive to share rosters of good workers
      \item Platforms themselves don't necessarily facilitate sharing in the first place
      \item Small scale requesters don't have sufficient traffic to make it worth it to invest in qualifications exams
      \item Workers who already spend relatively large proportions of their time \textbf{finding tasks}
            don't want to take qualifying exams to become eligible for one--off tasks
    \end{enumerate}
  \end{enumerate}
\end{enumerate}

\section*{Key insight \joke{or ``Wao Such Insight! Much \dots~Insight''}}
\begin{enumerate}[noitemsep]
  \item In our most recent paper, we highlighted some of the common themes between
        the historical piecework and the contemporary ``gig'' or
        on--demand work~\cite{pieceworkCrowdworkGigwork}.
  \item We argued that, while significant differences are to be found throughout the comparisons of
        contemporary cases of on--demand work and historical counterparts,
        these parallel cases can still \textit{inform} how we observe the evolution of work going forward
        and, indeed, how we can shape and design that future.
  \item In that light, we look to the trajectory of piecework in the \nth{20} century for inspiration and find\dots
  \item what about \textbf{hiring halls}? What's important about hiring halls?
  \begin{enumerate}[noitemsep]
    \item Hiring halls were (/are) communities of workers with various skills who band together
          to help dispatch qualified workers for various jobs.
    \item These hiring halls had \textit{de facto} collective bargaining powers.
    \item This was an important step toward collective action and collective governance
          (other topics about which we've written~\cite{dynamo}).
  \end{enumerate}
\end{enumerate}

\section*{What we're building}
\begin{enumerate}[noitemsep]
  \item Building a system that aggregates workers' histories into a single platform
  \item Same with education
  \item Through this, workers can demonstrate their proficiency by accumulating work histories from various platforms
\end{enumerate}

\section*{Why is this hard (or at least not trivial)}
\begin{enumerate}
  \item \ali{All of the reasons we've ever said that the internet complicates offline things}
  \item also, even offline communities of workers are hard to organize;
        we think we can make some of it easier through technology, but \textit{how} is still an open question
\end{enumerate}

\printbibliography{}
\end{document}

### 2017-07-19 proposals/mediaX/proposal.tex
notes from meeting earlier

@@
 
+\ali{Don't be a hundred duck sized horses; be a horse sized duck}
+
+\ali{chekhov's gun; don't put something in the story if you won't advance it to the end (or very far)}
+
 \section*{State of the world}
@@
         of on--demand labor will continue to grow
-  \item More people will be doing more kinds of work, transiently moving between myriad platforms
+  \item More people will be doing more kinds of work, transiently moving between myriad platforms \ali{more quickly than before}
+  \ali{relationships are shrinking to the span of a few minutes, and people may only interact once}
   \item People's educations are becoming increasingly decomposed;
@@
 \begin{enumerate}[noitemsep]
-  \item The people who enter the workforce in this capacity are increasingly finding that
+  \item \ali{silo'ing} The people who enter the workforce in this capacity are increasingly finding that
         this form of work doesn't facilitate the building of more complex, more rewarding work
@@
 \begin{enumerate}[noitemsep]
+  \item \ali{feel free to jump around \msb{jump, jump, jump around}, like pointing to preference of freelancers; just say that this is a wider phenomenon}
+  \item \ali{find some econ literature or something that finds that people prefer workers they know (i.e. gravitate toward whitelisting)}
   \item One solution seems to be to maintain a list of known good workers --- ``whitelisting'' good workers
@@
 \begin{enumerate}[noitemsep]
+\item \ali{don't be too meta; just cite the \textit{idea} of a hiring hall}
   \item In our most recent paper, we highlighted some of the common themes between
@@
   \end{enumerate}
+  \item \ali{aggregating workers histories to make it possible to communicate skills}
+  \item \ali{hiring halls had to be \textbf{bottlenecks}; we can be less bottlenecky}
 \end{enumerate}
@@
 \begin{enumerate}[noitemsep]
+\item \ali{be more concrete about what the system will do}
+\item \ali{maybe make a nice--looking figure}
   \item Building a system that aggregates workers' histories into a single platform
@@
 \begin{enumerate}
+
   \item \ali{All of the reasons we've ever said that the internet complicates offline things}
@@
 
+\ali{evaluation plan?}
+
+\ali{so what/impact of outcome}
+
 \printbibliography{}
